### 2021-06-28 paper/mcd2.tex
Skeleton for paper.

\documentclass[letterpaper,final,12pt,reqno]{amsart}

\usepackage[total={6.3in,9.2in},top=1.1in,left=1.1in]{geometry}

\usepackage{times,bm,bbm,empheq,fancyvrb,graphicx}
\usepackage[dvipsnames]{xcolor}
\usepackage{longtable}
\usepackage{booktabs}

\usepackage[within=section]{newfloat}

\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing}

\usepackage[kw]{pseudo}
\pseudoset{left-margin=15mm,topsep=5mm,idfont=\texttt}

% hyperref should be the last package we load
\usepackage[pdftex,
colorlinks=true,
plainpages=false, % only if colorlinks=true
linkcolor=blue,   % ...
citecolor=Red,    % ...
urlcolor=black    % ...
]{hyperref}

\renewcommand{\baselinestretch}{1.05}

\allowdisplaybreaks[1]  % allow display breaks in align environments, if they avoid major underfulls

\newtheoremstyle{claim}% name
  {5pt}% space above
  {5pt}% space below
  {\itshape}% body font
  {}% indent amount
  {\itshape}% theorem head font
  {.}% punctuation after theorem head
  {.5em}% space after theorem head
  {\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}% theorem head spec
\theoremstyle{claim}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

\newcommand{\eps}{\epsilon}
\newcommand{\RR}{\mathbb{R}}

\newcommand{\grad}{\nabla}
\newcommand{\Div}{\nabla\cdot}
\newcommand{\trace}{\operatorname{tr}}

\newcommand{\hbn}{\hat{\mathbf{n}}}

\newcommand{\bb}{\mathbf{b}}
\newcommand{\be}{\mathbf{e}}
\newcommand{\bbf}{\mathbf{f}}
\newcommand{\bg}{\mathbf{g}}
\newcommand{\bn}{\mathbf{n}}
\newcommand{\br}{\mathbf{r}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\bv}{\mathbf{v}}
\newcommand{\bw}{\mathbf{w}}
\newcommand{\bx}{\mathbf{x}}

\newcommand{\bF}{\mathbf{F}}
\newcommand{\bV}{\mathbf{V}}
\newcommand{\bX}{\mathbf{X}}

\newcommand{\bxi}{\bm{\xi}}

\newcommand{\bzero}{\bm{0}}

\newcommand{\rhoi}{\rho_{\text{i}}}

\newcommand{\ip}[2]{\left(#1,#2\right)}

\newcommand{\mR}{R^{\bm{\oplus}}}
\newcommand{\iR}{R^{\bullet}}

\newcommand{\nn}{{\text{n}}}
\newcommand{\pp}{{\text{p}}}
\newcommand{\qq}{{\text{q}}}
\newcommand{\rr}{{\text{r}}}

\newcommand{\bus}{\bu|_s}

% norm |||x|||
\newcommand{\vertiii}[1]{{\left\vert\kern-0.25ex\left\vert\kern-0.25ex\left\vert #1 \right\vert\kern-0.25ex\right\vert\kern-0.25ex\right\vert}}

% numbering
\setcounter{tocdepth}{3}
\makeatletter
\def\l@subsection{\@tocline{2}{0pt}{4pc}{5pc}{}}
\makeatother

\numberwithin{equation}{section}
\numberwithin{figure}{section}
\numberwithin{table}{section}
\numberwithin{theorem}{section}

\DeclareFloatingEnvironment[name=Pseudocode]{pcode}

\begin{document}
\title[On MCD methods for nonlinear or nonlocal VIs]{On multilevel constraint decomposition methods for nonlinear or nonlocal variational inequalities}

\author{Ed Bueler}

\date{\today}

\begin{abstract} FIXME
\end{abstract}

\maketitle

%\tableofcontents

\thispagestyle{empty}
%\bigskip

\section{Introduction} \label{sec:intro}

FIXME the goal is to generalize and improve the method of \cite{Tai2003}

\section{$p$-coercive variational inequalities} \label{sec:vi}

FIXME

\small

\bigskip
\bibliography{mcd2}
\bibliographystyle{siam}

\end{document}

### 2021-06-29 paper/mcd2.tex
Start writing some stuff.

@@
 
-\usepackage{times,bm,bbm,empheq,fancyvrb,graphicx}
+\usepackage{times,bm,bbm,empheq,fancyvrb,graphicx,amsthm}
 \usepackage[dvipsnames]{xcolor}
@@
 
-\newtheoremstyle{claim}% name
+\newtheoremstyle{cstyle}% name
   {5pt}% space above
@@
   {\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}% theorem head spec
-\theoremstyle{claim}
+\theoremstyle{cstyle}
+
 \newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
+\newtheorem{lemma}[theorem]{Lemma}
+
+\newtheoremstyle{dstyle}% name
+  {5pt}% space above
+  {5pt}% space below
+  {}%{\itshape}% body font
+  {}% indent amount
+  {\itshape}% theorem head font
+  {.}% punctuation after theorem head
+  {.5em}% space after theorem head
+  {\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}% theorem head spec
+\theoremstyle{dstyle}
+
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{example}[theorem]{Example}
+
+% numbering
+\numberwithin{equation}{section}
+\numberwithin{figure}{section}
+\numberwithin{table}{section}
+\numberwithin{theorem}{section}
 
@@
 \newcommand{\bx}{\mathbf{x}}
-
 \newcommand{\bF}{\mathbf{F}}
@@
 \newcommand{\bX}{\mathbf{X}}
+\newcommand{\bxi}{\bm{\xi}}
+\newcommand{\bzero}{\bm{0}}
 
-\newcommand{\bxi}{\bm{\xi}}
-
-\newcommand{\bzero}{\bm{0}}
+\newcommand{\cK}{\mathcal{K}}
+\newcommand{\cV}{\mathcal{V}}
 
@@
 
-\newcommand{\ip}[2]{\left(#1,#2\right)}
+\newcommand{\ip}[2]{\left<#1,#2\right>}
 
@@
 
-\newcommand{\bus}{\bu|_s}
-
-% norm |||x|||
-\newcommand{\vertiii}[1]{{\left\vert\kern-0.25ex\left\vert\kern-0.25ex\left\vert #1 \right\vert\kern-0.25ex\right\vert\kern-0.25ex\right\vert}}
-
-% numbering
-\setcounter{tocdepth}{3}
-\makeatletter
-\def\l@subsection{\@tocline{2}{0pt}{4pc}{5pc}{}}
-\makeatother
-
-\numberwithin{equation}{section}
-\numberwithin{figure}{section}
-\numberwithin{table}{section}
-\numberwithin{theorem}{section}
-
 \DeclareFloatingEnvironment[name=Pseudocode]{pcode}
@@
 \begin{document}
-\title[On MCD methods for nonlinear or nonlocal VIs]{On multilevel constraint decomposition methods for nonlinear or nonlocal variational inequalities}
+\title[On MCD methods for nonlinear and nonlocal VIs]{On multilevel constraint decomposition methods \\ for nonlinear and nonlocal variational inequalities}
 
@@
 
+Suppose $\cV$ is a reflexive Banach space with norm $\|\cdot\|$, $\cV'$ its topological dual space, and $\ip{z}{v} = z(v)$ the dual pairing of $z \in \cV'$ and $v\in \cV$.
+Note that $\|z\|_{\cV'} = \sup_{\|v\|=1} |\ip{z}{v}|$ defines the norm on $\cV'$, which is also a Banach space.
+Let $\cK \subset \cV$ be a nonempty closed and convex subset of $\cV$.
+The following definitions and theorem are standard \cite{KinderlehrerStampacchia1980}.
+
+\begin{definition}  A map $f:\cK \to \cV'$ is \emph{monotone} if
+\begin{equation}
+\ip{f(u)-f(v)}{u-v} \ge 0 \qquad \text{for all } u,v \in \cK, \label{eq:monotone}
+\end{equation}
+\emph{strictly monotone} if equality in \eqref{eq:monotone} implies $u=v$, and \emph{coercive} if there exists $w \in \cK$ so that
+\begin{equation}
+\frac{\ip{f(u)-f(w)}{u-w}}{\|u-w\|} \to +\infty \qquad \text{as } \|u\|\to +\infty.
+\end{equation}
+\end{definition}
+
+\begin{definition} For $u\in \cK$, the statement that
+\begin{equation}
+\ip{f(u)}{v-u} \ge 0 \qquad \text{for all } v\in \cK \label{eq:vi}
+\end{equation}
+is called a \emph{variational inequality} (VI).
+\end{definition}
+
+Speaking informally, and conceptualizing the dual pairing as an inner product, such a VI says that the angle between $f(u)$ and any arbitrary vector $v-u$ pointing into $\cK$ from $u$, is at most $90^\circ$, thus $f(u)$ points straight into $\cK$.
+Note that if $u$ is in the interior of $\cK$ then \eqref{eq:vi} implies $f(u)=0$, so in this sense the VI generalizes a system of equations.
+
+It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution, and furthermore that that solution is unique if $f$ is strictly monotone \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}.
+The condition of continuity in Theorem \ref{thm:viwellposed} can be weakened to continuity on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}.
+However, the continuity of $f$ will apply in our examples.
+
+As in the calculus of variations \cite{Evans2010}, the condition of coercivity in Theorem \ref{thm:viwellposed} permits a (weak) compactness argument even for unbounded sets $\cK$.
+However, the VIs solved in this paper satisfy a stronger form of coercivity which will be critical in proving convergence of MCD methods; note that \cite{Tai2003} uses ``coercive'' from ``$2$-coercive'' in the following definition.
+
+\begin{definition}  $f:\cK \to \cV'$ is \emph{$p$-coercive} if there exists $\kappa>0$ such that
+\begin{equation}
+\ip{f(u)-f(v)}{u-v} \ge \kappa \|u-v\|^p \qquad \text{for all } u,v \in \cK. \label{eq:pcoercive}
+\end{equation}
+\end{definition}
+
+It is easy to see that if $f$ is $p$-coercive for $p>1$ then it is monotone, strictly monotone, and coercive.
+The following result is thus also well-known, but it is important to state as it shows we are addressing well-posed problems.
+
+\begin{theorem}  \label{thm:viwellposed}  If $f:\cK \to \cV'$ is continuous and $p$-coercive for $p>1$ then there exists a unique $u\in \cK$ so that VI \eqref{eq:vi} holds.
+\end{theorem}
+
+\begin{example}  Inequality-constrained optimization of convex and differentiable scalar functionals generate VI problems.
+In fact, suppose $F:V \to \RR$ is convex and (G\^ateau) differentiable.
+Then VI \eqref{eq:vi} holds for $f=F'$ if and only if
+\begin{equation}
+u = \operatorname{arg-min}_{v\in\cK} F(v). \label{eq:minimization}
+\end{equation}
+Tai \cite{Tai2003} addresses MCD convergence for problem \eqref{eq:minimization} under the assumption that $f=F'$ is $2$-coercive.
+\end{example}
+
+\begin{example}  However, not all VIs \eqref{eq:vi} correspond to optimization problems.
+For example, consider the following nonlinear problem of porous-medium type.
+Suppose $\Omega \subset \RR^d$ is bounded and open with polygonal boundary.
+Let $\cV = H_0^1(\Omega)$ and consider $\cK = \{v\ge 0\} \subset \cV$.
+Let $g\in \cV'$.
+For $u$ in $\cK$ let
+\begin{equation}
+\ip{f(u)}{v} = \int_\Omega (1+e^{-u}) \grad u \cdot \grad v\,dx - \ip{g}{v}. \label{eq:nongradientexample}
+\end{equation}
+It is easy to see that $|\ip{f(u)}{v}| \le (c \|u\|_{\cV} + \|g\|_{\cV'}) \|v\|_{\cV}$, thus that $f:\cK \to \cV'$ is continuous.  FIXME IS IT $2$-coercive?  FIXME $f\ne F'$
+\end{example}
+
+\section{Multilevel constraint decomposition algorithms} \label{sec:algorithms}
+
+FIXME state the decomposition, the basis algorithm, the FAS version which has $O(m)$ residual evaluation complexity on each level
+
+\section{Convergence of the basic algorithm} \label{sec:convergence}
+
+FIXME
+
+\section{Results} \label{sec:results}
+
 FIXME

### 2021-06-30 paper/mcd2.tex
Fiddle around with text and add a reference.

@@
 
-Suppose $\cV$ is a reflexive Banach space with norm $\|\cdot\|$, $\cV'$ its topological dual space, and $\ip{z}{v} = z(v)$ the dual pairing of $z \in \cV'$ and $v\in \cV$.
+Suppose $\cV$ is a real, reflexive Banach space with norm $\|\cdot\|$ and topological dual space $\cV'$.
+Denote the dual pairing of $z \in \cV'$ and $v\in \cV$ by $\ip{z}{v} = z(v)$.
 Note that $\|z\|_{\cV'} = \sup_{\|v\|=1} |\ip{z}{v}|$ defines the norm on $\cV'$, which is also a Banach space.
-Let $\cK \subset \cV$ be a nonempty closed and convex subset of $\cV$.
-The following definitions and theorem are standard \cite{KinderlehrerStampacchia1980}.
+
+Let $\cK \subset \cV$ be a nonempty closed and convex subset of $\cV$.  We consider the following \emph{variational inequality} (VI) for $u\in \cK$:
+\begin{equation}
+\ip{f(u)}{v-u} \ge 0 \qquad \text{for all } v\in \cK. \label{eq:vi}
+\end{equation}
+Informally, conceptualizing the dual pairing as an inner product, \eqref{eq:vi} would say that the angle between $f(u)$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
+That is, $f(u)$ points straight into $\cK$ if $u$ solves \eqref{eq:vi}, and if $u$ is in the interior of $\cK$ then \eqref{eq:vi} further implies $f(u)=0$.
+Thus such a VI generalizes a system of equations.
+
+The following definitions \cite{KinderlehrerStampacchia1980} are standard.
 
@@
 
-\begin{definition} For $u\in \cK$, the statement that
+It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, and that the solution is unique if $f$ is strictly monotone.
+As in the calculus of variations \cite{Evans2010}, coercivity permits a (weak) compactness argument even for unbounded sets $\cK$.
+The condition of continuity can be weakened to continuity on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but the stronger condition will apply in our examples.
+
+Monotone VI \eqref{eq:vi} generalizes the problem of minimizing convex objective functionals.
+In fact, suppose $F:\cV \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cV \to \cV'$.
+Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
+Furthermore, if $F$ is convex and $\cK$ is convex and closed then VI \eqref{eq:vi} holds for $f=F'$ and $u\in \cK$ if and only if
 \begin{equation}
-\ip{f(u)}{v-u} \ge 0 \qquad \text{for all } v\in \cK \label{eq:vi}
+u = \operatorname{arg-min}_{v\in\cK} F(v) \label{eq:minimization}
 \end{equation}
-is called a \emph{variational inequality} (VI).
-\end{definition}
+\cite[Proposition II.2.1]{EkelandTemam1976}.  Note that \cite{Tai2003} addresses MCD convergence for problem \eqref{eq:minimization}.
 
-Speaking informally, and conceptualizing the dual pairing as an inner product, such a VI says that the angle between $f(u)$ and any arbitrary vector $v-u$ pointing into $\cK$ from $u$, is at most $90^\circ$, thus $f(u)$ points straight into $\cK$.
-Note that if $u$ is in the interior of $\cK$ then \eqref{eq:vi} implies $f(u)=0$, so in this sense the VI generalizes a system of equations.
+However, not all VIs correspond to optimization problems.  For example, consider the following problem of porous-medium type.
 
-It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution, and furthermore that that solution is unique if $f$ is strictly monotone \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}.
-The condition of continuity in Theorem \ref{thm:viwellposed} can be weakened to continuity on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}.
-However, the continuity of $f$ will apply in our examples.
+\begin{example}  Suppose $\Omega \subset \RR^d$ is bounded and open with polygonal boundary.
+Let $\cV = H_0^1(\Omega)$ and consider $\cK = \{v\ge 0\} \subset \cV$.
+For $g\in \cV'$ and $u$ in $\cK$ let
+\begin{equation}
+\ip{f(u)}{v} = \int_\Omega (1+e^{-u}) \grad u \cdot \grad v\,dx - \ip{g}{v}. \label{eq:nongradientexample}
+\end{equation}
+It is easy to see that $|\ip{f(u)}{v}| \le (c \|u\|_{\cV} + \|g\|_{\cV'}) \|v\|_{\cV}$, thus that $f:\cK \to \cV'$ is continuous.  FIXME IS IT $2$-coercive?  FIXME $f\ne F'$
+\end{example}
 
-As in the calculus of variations \cite{Evans2010}, the condition of coercivity in Theorem \ref{thm:viwellposed} permits a (weak) compactness argument even for unbounded sets $\cK$.
-However, the VIs solved in this paper satisfy a stronger form of coercivity which will be critical in proving convergence of MCD methods; note that \cite{Tai2003} uses ``coercive'' from ``$2$-coercive'' in the following definition.
+However, the VIs solved in this paper satisfy a stronger form of coercivity which will be important in proving convergence of MCD methods.
 
@@
 \end{equation}
+(Note \cite{Tai2003} uses ``coercive'' for $2$-coercive.)
 \end{definition}
 
-It is easy to see that if $f$ is $p$-coercive for $p>1$ then it is monotone, strictly monotone, and coercive.
-The following result is thus also well-known, but it is important to state as it shows we are addressing well-posed problems.
+It is easy to see that if $f$ is $p$-coercive for $p>1$ then it is monotone, strictly monotone, and coercive, and thus the following result, which states that we are addressing well-posed problems, is immediate.
 
@@
 \end{theorem}
-
-\begin{example}  Inequality-constrained optimization of convex and differentiable scalar functionals generate VI problems.
-In fact, suppose $F:V \to \RR$ is convex and (G\^ateau) differentiable.
-Then VI \eqref{eq:vi} holds for $f=F'$ if and only if
-\begin{equation}
-u = \operatorname{arg-min}_{v\in\cK} F(v). \label{eq:minimization}
-\end{equation}
-Tai \cite{Tai2003} addresses MCD convergence for problem \eqref{eq:minimization} under the assumption that $f=F'$ is $2$-coercive.
-\end{example}
-
-\begin{example}  However, not all VIs \eqref{eq:vi} correspond to optimization problems.
-For example, consider the following nonlinear problem of porous-medium type.
-Suppose $\Omega \subset \RR^d$ is bounded and open with polygonal boundary.
-Let $\cV = H_0^1(\Omega)$ and consider $\cK = \{v\ge 0\} \subset \cV$.
-Let $g\in \cV'$.
-For $u$ in $\cK$ let
-\begin{equation}
-\ip{f(u)}{v} = \int_\Omega (1+e^{-u}) \grad u \cdot \grad v\,dx - \ip{g}{v}. \label{eq:nongradientexample}
-\end{equation}
-It is easy to see that $|\ip{f(u)}{v}| \le (c \|u\|_{\cV} + \|g\|_{\cV'}) \|v\|_{\cV}$, thus that $f:\cK \to \cV'$ is continuous.  FIXME IS IT $2$-coercive?  FIXME $f\ne F'$
-\end{example}
 

### 2021-07-01 paper/mcd2.tex
Make a bit more progress.

@@
 
-\section{$p$-coercive variational inequalities} \label{sec:vi}
+\section{Coercive variational inequalities and constraint decomposition} \label{sec:vi}
 
@@
 
-However, not all VIs correspond to optimization problems.  For example, consider the following problem of porous-medium type.
+However, not all VIs correspond to optimization problems.  For example, consider the following steady advection-diffusion problem.
 
@@
 Let $\cV = H_0^1(\Omega)$ and consider $\cK = \{v\ge 0\} \subset \cV$.
-For $g\in \cV'$ and $u$ in $\cK$ let
+For $g\in \cV'$, $\bX \in L^\infty(\Omega)^d$, and $u$ in $\cK$ let
 \begin{equation}
-\ip{f(u)}{v} = \int_\Omega (1+e^{-u}) \grad u \cdot \grad v\,dx - \ip{g}{v}. \label{eq:nongradientexample}
+\ip{f(u)}{v} = \int_\Omega \grad u \cdot \grad v\,dx - (\bX \cdot \grad u) v - \ip{g}{v}. \label{eq:nongradientexample}
 \end{equation}
-It is easy to see that $|\ip{f(u)}{v}| \le (c \|u\|_{\cV} + \|g\|_{\cV'}) \|v\|_{\cV}$, thus that $f:\cK \to \cV'$ is continuous.  FIXME IS IT $2$-coercive?  FIXME $f\ne F'$
+FIXME POINT OUT THAT IS STEADY ADVECTION-DIFFUSION VI; IF $\bX=0$ THEN IT IS OPTIMIZATION BUT OTHERWISE NOT; ALWAYS CONTINUOUS; IF $\bX$ IS SMALL THEN $2$-coercive; It is easy to see that $|\ip{f(u)}{v}| \le (c \|u\|_{\cV} + \|g\|_{\cV'}) \|v\|_{\cV}$, thus that $f:\cK \to \cV'$ is continuous.
 \end{example}
+
+Other examples of VI problems which are not of optimization type are found in ice sheet applications \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids problems \cite{Bueler2021conservation}.
 
@@
 
-\section{Multilevel constraint decomposition algorithms} \label{sec:algorithms}
+The multilevel algorithms we consider later are based on both subspace \cite{Xu1992} and constraint \cite{Tai2003} decompositions, as follows.
+Suppose $\cV_i \subset \cV$ are $m<\infty$ subspaces so that
+\begin{equation}
+\cV = \sum_{i=1}^m \cV_i \label{eq:subspacedecomp}
+\end{equation}
+in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$.  Suppose further that $\cK_i \subset \cV$ are nonempty, closed, and convex subsets such that
+\begin{equation}
+\cK_i \subset \cV_i \qquad \text{and} \qquad \cK = \sum_{i=1}^m \cK_i. \label{eq:constraintdecomp}
+\end{equation}
+The sum in \eqref{eq:constraintdecomp} is required to hold in two senses: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.
+(The latter sense is automatic in \eqref{eq:subspacedecomp} because $\cV_i$ are subspaces.)  Neither decomposition is unique in general.
+Note that $\cK_i \not\subset \cK$ in most applications.
 
-FIXME state the decomposition, the basis algorithm, the FAS version which has $O(m)$ residual evaluation complexity on each level
+In practice the space $\cV$ will be a finite-dimensional finite element (FE) space because we will solve VI \eqref{eq:vi} over a mesh.  We FIXME triangulation etc. Though our
+
+\begin{example}  FIXME $\cV_i$ are 1d spaces on a single level; $\psi,\cK,\cK_i$ for obstacle problem; note $\cK_i \not\subset \cK$ when $\psi>0$
+\end{example}
+
+FIXME state the basic algorithm
+
+\section{Multilevel constraint decomposition algorithms} \label{sec:multilevel}
+
+FIXME state essentially Algorithm 4.7 \cite{GraeserKornhuber2009} but with $\text{V}(\nu_1,\nu_2)$ cycles which works for linear; observe that up-smoothing is more efficient; state the FAS version which has $O(m)$ residual evaluation complexity on each level
 

### 2021-07-02 paper/mcd2.tex
A bit more progress.

@@
 
+\newcommand{\supp}{\operatorname{supp}}
+
 \DeclareFloatingEnvironment[name=Pseudocode]{pcode}
@@
 
-Suppose $\cV$ is a real, reflexive Banach space with norm $\|\cdot\|$ and topological dual space $\cV'$.
-Denote the dual pairing of $z \in \cV'$ and $v\in \cV$ by $\ip{z}{v} = z(v)$.
-Note that $\|z\|_{\cV'} = \sup_{\|v\|=1} |\ip{z}{v}|$ defines the norm on $\cV'$, which is also a Banach space.
+Suppose $\cV$ is a real, reflexive Banach space with norm $\|v\|$ for $v\in\cV$ and topological dual space $\cV'$.
+Denote the dual pairing of $\phi \in \cV'$ and $v$ by $\ip{\phi}{v} = \phi(v)$.
+Note that $\|\phi\|_{\cV'} = \sup_{\|v\|=1} |\ip{\phi}{v}|$ defines the norm on $\cV'$, which is also a Banach space.
 
-Let $\cK \subset \cV$ be a nonempty closed and convex subset of $\cV$.  We consider the following \emph{variational inequality} (VI) for $u\in \cK$:
+Let $\cK \subset \cV$ be a nonempty closed and convex subset of $\cV$, the \emph{constraint set}.
+For continuous $f:\cK \to \cV'$ we consider the following \emph{variational inequality} (VI) for $u\in \cK$:
 \begin{equation}
@@
 \end{equation}
-Informally, conceptualizing the dual pairing as an inner product, \eqref{eq:vi} would say that the angle between $f(u)$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
-That is, $f(u)$ points straight into $\cK$ if $u$ solves \eqref{eq:vi}, and if $u$ is in the interior of $\cK$ then \eqref{eq:vi} further implies $f(u)=0$.
-Thus such a VI generalizes a system of equations.
+Informally, if we conceptualize the dual pairing as an inner product then \eqref{eq:vi} says that the angle between $f(u)$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$, that is, $f(u)$ points straight into $\cK$ if $u$ solves \eqref{eq:vi}.
+Furthermore, if $u$ is in the interior of $\cK$ then \eqref{eq:vi} implies $f(u)=0$, so VI \eqref{eq:vi} generalizes such an equation (system of equations).
 
-The following definitions \cite{KinderlehrerStampacchia1980} are standard.
+The following definitions are standard \cite{KinderlehrerStampacchia1980}.
 
@@
 \begin{equation}
-\frac{\ip{f(u)-f(w)}{u-w}}{\|u-w\|} \to +\infty \qquad \text{as } \|u\|\to +\infty.
+\frac{\ip{f(u)-f(w)}{u-w}}{\|u-w\|} \to +\infty \qquad \text{as } \|u\|\to +\infty. \label{eq:coercive}
 \end{equation}
@@
 
-It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, and that the solution is unique if $f$ is strictly monotone.
-As in the calculus of variations \cite{Evans2010}, coercivity permits a (weak) compactness argument even for unbounded sets $\cK$.
+It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, and also that the solution is unique if $f$ is strictly monotone.
+As is standard in the calculus of variations \cite{Evans2010}, coercivity permits a (weak) compactness argument even for unbounded sets $\cK$.
 The condition of continuity can be weakened to continuity on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but the stronger condition will apply in our examples.
 
-Monotone VI \eqref{eq:vi} generalizes the problem of minimizing convex objective functionals.
-In fact, suppose $F:\cV \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cV \to \cV'$.
+When $f$ is monotone, VI \eqref{eq:vi} generalizes the problem of minimizing convex functions.
+Suppose $F:\cV \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cV \to \cV'$.
 Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
-Furthermore, if $F$ is convex and $\cK$ is convex and closed then VI \eqref{eq:vi} holds for $f=F'$ and $u\in \cK$ if and only if
+Furthermore, if $F$ is convex and $\cK$ is closed and convex then \eqref{eq:vi} holds for $f=F'$ if and only if
 \begin{equation}
@@
 \end{equation}
-\cite[Proposition II.2.1]{EkelandTemam1976}.  Note that \cite{Tai2003} addresses MCD convergence for problem \eqref{eq:minimization}.
+\cite[Proposition II.2.1]{EkelandTemam1976}.  Note that reference \cite{Tai2003} addresses problem \eqref{eq:minimization}.
 
-However, not all VIs correspond to optimization problems.  For example, consider the following steady advection-diffusion problem.
+However, not all VIs correspond to optimization.
+For example, consider the following steady advection-diffusion problem over $H_0^1(\Omega)$.
+(We denote Sobolev spaces by $W^{k,p}(\Omega)$ if $k\ge 0$ and $p\ge 1$ \cite{Evans2010}, and $W^{k,2}=H^k$ in particular.)
 
-\begin{example}  Suppose $\Omega \subset \RR^d$ is bounded and open with polygonal boundary.
+\begin{example}  \label{ex:advectiondiffusion}  Suppose $\Omega \subset \RR^d$ is bounded and open with polygonal boundary.
 Let $\cV = H_0^1(\Omega)$ and consider $\cK = \{v\ge 0\} \subset \cV$.
-For $g\in \cV'$, $\bX \in L^\infty(\Omega)^d$, and $u$ in $\cK$ let
+For $g\in L^2(\Omega)$, $\bX \in L^\infty(\Omega)^d$, and $u$ in $\cK$ let
 \begin{equation}
-\ip{f(u)}{v} = \int_\Omega \grad u \cdot \grad v\,dx - (\bX \cdot \grad u) v - \ip{g}{v}. \label{eq:nongradientexample}
+\ip{f(u)}{v} = \int_\Omega \grad u \cdot \grad v\,dx - (\bX \cdot \grad u) v - gv\,dx. \label{eq:nongradientexample}
+\end{equation}
+FIXME OR DO I WANT
+\begin{equation}
+\ip{f(u)}{v} = \int_\Omega \grad u \cdot \grad v\,dx - u (\bX \cdot \grad v) - gv\,dx. \label{eq:ALTnongradientexample}
 \end{equation}
@@
 
-Other examples of VI problems which are not of optimization type are found in ice sheet applications \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids problems \cite{Bueler2021conservation}.
+Other examples of VI problems which are not of optimization type are found in ice sheet models \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids problems \cite{Bueler2021conservation}.
 
-However, the VIs solved in this paper satisfy a stronger form of coercivity which will be important in proving convergence of MCD methods.
+The VIs solved in this paper satisfy a stronger form of coercivity than \eqref{eq:coercive}.  This stronger assumption will be used in proving convergence of MCD methods.
 
-\begin{definition}  $f:\cK \to \cV'$ is \emph{$p$-coercive} if there exists $\kappa>0$ such that
+\begin{definition}  Let $p>1$.  The map $f:\cK \to \cV'$ is \emph{$p$-coercive} if there exists $\kappa>0$ such that
 \begin{equation}
@@
 
-It is easy to see that if $f$ is $p$-coercive for $p>1$ then it is monotone, strictly monotone, and coercive, and thus the following result, which states that we are addressing well-posed problems, is immediate.
+It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive, and thus the following result, which states that the VI problems considered in this paper are well-posed, is immediate.
 
-\begin{theorem}  \label{thm:viwellposed}  If $f:\cK \to \cV'$ is continuous and $p$-coercive for $p>1$ then there exists a unique $u\in \cK$ so that VI \eqref{eq:vi} holds.
+\begin{theorem}  \label{thm:viwellposed}  If $f:\cK \to \cV'$ is continuous and $p$-coercive (for $p>1$) then there exists a unique $u\in \cK$ so that VI \eqref{eq:vi} holds.
 \end{theorem}
 
-The multilevel algorithms we consider later are based on both subspace \cite{Xu1992} and constraint \cite{Tai2003} decompositions, as follows.
-Suppose $\cV_i \subset \cV$ are $m<\infty$ subspaces so that
+Our multilevel algorithms (next section) are based on both subspace and constraint decompositions, as follows.  Suppose there are $m<\infty$ subspaces $\cV_i \subset \cV$ so that
 \begin{equation}
@@
 \end{equation}
-in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$.  Suppose further that $\cK_i \subset \cV$ are nonempty, closed, and convex subsets such that
+holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$ \cite{Xu1992}.
+Suppose further that $\cK_i \subset \cV$ are nonempty, closed, and convex subsets such that
 \begin{equation}
@@
 \end{equation}
-The sum in \eqref{eq:constraintdecomp} is required to hold in two senses: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.
-(The latter sense is automatic in \eqref{eq:subspacedecomp} because $\cV_i$ are subspaces.)  Neither decomposition is unique in general.
-Note that $\cK_i \not\subset \cK$ in most applications.
+The constraint decomposition \cite{Tai2003} sum in \eqref{eq:constraintdecomp} is required to hold in two senses: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.
+(The latter sense is automatic in \eqref{eq:subspacedecomp} because $\cV_i$ are subspaces.)  Note that neither decomposition is unique (in general), and that $\cK_i \not\subset \cK$ in many applications.
+
+Following \cite{Tai2003}, for each $\cK_i$ we also assume a (generally) nonlinear restriction operator $R_i : \cK \to \cK_i$ such that
+\begin{equation}
+v = \sum_{i=1}^m R_i v \qquad \text{if } v \in \cK.  \label{eq:constraintrestrictionsum}
+\end{equation}
+An \emph{constraint decomposition} of the constraint set $\cK$ is a choice of $\cV_i,\cK_i,R_i$ satisfying the above hypotheses.
+
+In the next section we will introduce FE spaces and describe practical algorithms, but, as noted in the Introduction, constraint decompositions may be defined at the level of the continuum problem.
+The following examples illustrate constraint decompositions of \emph{obstacle problems} \cite{GraeserKornhuber2009}; Example \ref{ex:domaindecomposition} is an overlapping domain decomposition and \ref{ex:frequencydecomposition} is a nonoverlapping frequency decomposition.
+The latter example can be compared to the multilevel FE decomposition in the next section.
+
+\begin{example}  \label{ex:domaindecomposition}  Consider a bounded domain $\Omega \subset \RR^d$ with polygonal boundary, let $\cV = H_0^k(\Omega)$, $k\ge 1$, and suppose the obstacle $\psi \in H^k(\Omega)$ satisfies $\psi|_{\partial \Omega} \le 0$.
+Let $\cK = \{v \ge \psi\} \subset \cV$.
+Suppose further that $\{\phi_i\}_{i=1}^m$ is a smooth partition of unity on $\Omega$, i.e.~so that $0 \le \phi_i\le 1$ and $\sum_i \phi_i = 1$, and let $\Omega_i = \supp \phi_i$ (open support).
+Let $\cV_i = \{w \in \cV:w|_{\Omega \setminus \Omega_i} =0 \}$, $\cK_i = \{v \in V: v \ge \phi_i \psi\}$, and $R_i(v) = \max\{v,\phi_i \psi\}$.
+Then \eqref{eq:subspacedecomp}, \eqref{eq:constraintdecomp}, and \eqref{eq:constraintrestrictionsum} all hold.
+\end{example}
+
+\begin{example}  \label{ex:frequencydecomposition}  For simplicity suppose $\Omega = (0,a)^d \subset \RR^d$ for $a>0$, a cube, and let $\cV = H_{\text{per}}^k(\Omega)$, $k\ge 1$, be the periodic functions.
+Suppose $\psi \in H^k(\Omega)$ and let $\cK = \{v \ge \psi\} \subset \cV$.
+Without getting into any detailed notation of frequency representation, but noting that the frequencies are discrete, suppose $\cV_i \subset \cV$ are $m<\infty$ subspaces from an (nonoverlapping) partition by frequency, thus satisfying \eqref{eq:subspacedecomp}.
+Suppose $P_i:\cV \to \cV_i$ are the corresponding orthogonal projections, satisfying $I = \sum_i P_i$.
+Let $\cK_i = \{v \ge P_i \psi\} \subset \cV$ and $R_i(v) = \max\{v,P_i \psi\}$.
+Then \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} hold.
+\end{example}
+
+In Example \ref{ex:domaindecomposition}, note that $\cK_i \not\subset \cK$ in many cases, for example if $\psi$ is positive over portions of $\Omega$ where the decomposition into overlapping subdomains $\Omega_i$ is nontrivial.
+A similar comment applies in Example \ref{ex:frequencydecomposition}.
+
+FIXME state the basic ``sequential constraint decomposition'' and ``parallel constraint decomposition'' algorithms
 
@@
 \end{example}
-
-FIXME state the basic algorithm
 

### 2021-07-03 paper/mcd2.tex
Add introductory text.  (Note some claims are just hope at this point.)

@@
 
-\usepackage{times,bm,bbm,empheq,fancyvrb,graphicx,amsthm}
+\usepackage{times,bm,bbm,empheq,fancyvrb,graphicx,amsthm,amssymb}
 \usepackage[dvipsnames]{xcolor}
@@
 \begin{document}
-\title[On MCD methods for nonlinear and nonlocal VIs]{On multilevel constraint decomposition methods \\ for nonlinear and nonlocal variational inequalities}
+\title[On MCD methods for nonlinear and nonlocal VIs]{On multilevel constraint decomposition methods \\ for nonlinear variational inequalities}
 
@@
 
-FIXME the goal is to generalize and improve the method of \cite{Tai2003}
+The goal of this paper is to generalize the constraint decomposition (CD) method of X.-C.~Tai \cite{Tai2003}.
+We extend this solution method to nonlinear variational inequality (VI) problems for which it has not been previously considered, we address important finite element (FE) implementation details, and we demonstrate the resulting performance.
+
+The convergence of the CD method for several types of decompositions was proven in \cite{Tai2003} in the case of coercive VIs which arise from minimization of a convex functional over a convex set.
+In its multilevel form the CD method has been shown to have optimal complexity for elliptic, linear obstacle problems \cite[subsection 5.4]{Tai2003}; see also Theorem 4.6 and Algorithm 4.7 in \cite{GraeserKornhuber2009}.
+However, especially for the classical obstacle problem, i.e.~for a Laplacian functional, certain other methods are known to improve performance relative to the multilevel CD method \cite{GraeserKornhuber2009}.
+On the other hand, these improved multilevel methods either track the active set or modify the nodal basis functions, and in this sense they are discrete algorithms.
+By contrast the CD methods apply at the level of the continuum problem (see below), so the method is more fundamental than the truncated monotone multigrid method \cite{GraeserKornhuber2009,Kornhuber1994}, for example, in that sense.
+
+We extend the original algorithms and theory presented in \cite{Tai2003} in three particular directions:
+\renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
+\begin{enumerate}
+\item We do not assume that the continuum VI problem arises from optimization of a scalar objective, and, as shown by examples in Section \ref{sec:vi}, this represents the removal of a symmetry assumption in \cite{Tai2003}.
+Nonetheless we can prove convergence at essentially the same rate (Section \ref{sec:convergence}).
+% HOPE
+\item In the multilevel case we make the observation that, if implemented appropriately, V(0,1) ``up-slash'' cycles are intrinsically more efficient than the V(1,0) ``down-slash'' cycles (Section \ref{sec:multilevel}).
+This observation seems to be new for multilevel CD methods; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}.
+This preference for up-smoothing is, apparently, special to CD methods and does not arise in corresponding unconstrained subspace decomposition \cite{Xu1992} approaches.
+\item In the multilevel and nonlinear case, we make the algorithms more practical by addressing efficient storage of intermediate quantities.
+In particular we show results from a full approximation storage (FAS; \cite{Brandt1977}) version of the multilevel CD algorithm.
+\end{enumerate}
+
+As an additional difference from the presentation in \cite{Tai2003}, but an essentially formal one, we observe that the iterates from CD methods are always admissible, and thus the residual functional need only be defined on the admissible subset.
+Methods like CD, for which all iterates are admissible, have an advantage because they allow direct solutions of certain VI problems, such as the fluid-layer dynamics problems in \cite{Bueler2021conservation,JouvetBueler2012}, for which non-admissible methods such as semi-smooth methods \cite{BensonMunson2006} may require unnatural modifications of the residual form.
+
+In one example we consider a case where the residual functional is nonlocal, that is, where it is not a partial differential operator at all.
+Instead its values instead arise from the solution of Stokes problems for a layer of fluid on a substrate which forms the obstacle \cite{SayagWorster2013}, so the corresponding FE operator discretization is not sparse.
+In this case we cannot prove coercivity but we can demonstrate near-optimal complexity in practice.
+% HOPE
 
@@
 
-Let $\cK \subset \cV$ be a nonempty closed and convex subset of $\cV$, the \emph{constraint set}.
-For continuous $f:\cK \to \cV'$ we consider the following \emph{variational inequality} (VI) for $u\in \cK$:
+Let $\cK \subset \cV$ be a nonempty closed and convex subset of $\cV$, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.
+For a continuous \emph{residual functional} $f:\cK \to \cV'$ and \emph{source} $g\in \cV'$ we consider the following \emph{variational inequality} (VI) for $u\in \cK$:
 \begin{equation}
-\ip{f(u)}{v-u} \ge 0 \qquad \text{for all } v\in \cK. \label{eq:vi}
+\ip{f(u)}{v-u} \ge g(v-u) \qquad \text{for all } v\in \cK. \label{eq:vi}
 \end{equation}
-Informally, if we conceptualize the dual pairing as an inner product then \eqref{eq:vi} says that the angle between $f(u)$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$, that is, $f(u)$ points straight into $\cK$ if $u$ solves \eqref{eq:vi}.
-Furthermore, if $u$ is in the interior of $\cK$ then \eqref{eq:vi} implies $f(u)=0$, so VI \eqref{eq:vi} generalizes such an equation (system of equations).
+Because $f$ is a general, nonlinear map the use of a right-hand source function is not strictly needed---by redefining $f$ we may take $g=0$---but its inclusion will be helpful in describing algorithms in Section \ref{sec:multilevel}.
+
+Informally, if we conceptualize the dual pairing as an inner product, and if $g=0$, then \eqref{eq:vi} says that the angle between $f(u)$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$, that is, $f(u)$ points directly into $\cK$ if $u$ solves \eqref{eq:vi}.
+Furthermore, if $u$ is in the interior of $\cK$ then \eqref{eq:vi} implies $f(u)=g$; this VI generalizes the concept of a nonlinear system of equations $f(u)=g$ over $\cV$ to the constrained case where $\cK \subsetneq \cV$.
 
@@
 It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, and also that the solution is unique if $f$ is strictly monotone.
-As is standard in the calculus of variations \cite{Evans2010}, coercivity permits a (weak) compactness argument even for unbounded sets $\cK$.
-The condition of continuity can be weakened to continuity on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but the stronger condition will apply in our examples.
+The condition of continuity can be weakened to only apply on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but the stronger condition will apply in our examples.
+As in the calculus of variations \cite{Evans2010}, coercivity permits a weak compactness argument for unbounded sets $\cK$.
 
-When $f$ is monotone, VI \eqref{eq:vi} generalizes the problem of minimizing convex functions.
-Suppose $F:\cV \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cV \to \cV'$.
-Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
-Furthermore, if $F$ is convex and $\cK$ is closed and convex then \eqref{eq:vi} holds for $f=F'$ if and only if
+When $f$ is monotone, VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
+In fact, suppose $F:\cV \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cV \to \cV'$.
+Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}, and if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ and $g=0$ if and only if
 \begin{equation}
@@
 \end{equation}
-\cite[Proposition II.2.1]{EkelandTemam1976}.  Note that reference \cite{Tai2003} addresses problem \eqref{eq:minimization}.
+\cite[Proposition II.2.1]{EkelandTemam1976}.  Note that Tai \cite{Tai2003} addresses optimization problem \eqref{eq:minimization}.
 
@@
 Let $\cV = H_0^1(\Omega)$ and consider $\cK = \{v\ge 0\} \subset \cV$.
-For $g\in L^2(\Omega)$, $\bX \in L^\infty(\Omega)^d$, and $u$ in $\cK$ let
+For $\bX \in L^\infty(\Omega)^d$ such that $\Div \bX=0$, and $u$ in $\cK$ let
 \begin{equation}
-\ip{f(u)}{v} = \int_\Omega \grad u \cdot \grad v\,dx - (\bX \cdot \grad u) v - gv\,dx. \label{eq:nongradientexample}
+\ip{f(u)}{v} = \int_\Omega \grad u \cdot \grad v\,dx - (\bX \cdot \grad u) v\,dx. \label{eq:nongradientexample}
 \end{equation}
-FIXME OR DO I WANT
-\begin{equation}
-\ip{f(u)}{v} = \int_\Omega \grad u \cdot \grad v\,dx - u (\bX \cdot \grad v) - gv\,dx. \label{eq:ALTnongradientexample}
-\end{equation}
-FIXME POINT OUT THAT IS STEADY ADVECTION-DIFFUSION VI; IF $\bX=0$ THEN IT IS OPTIMIZATION BUT OTHERWISE NOT; ALWAYS CONTINUOUS; IF $\bX$ IS SMALL THEN $2$-coercive; It is easy to see that $|\ip{f(u)}{v}| \le (c \|u\|_{\cV} + \|g\|_{\cV'}) \|v\|_{\cV}$, thus that $f:\cK \to \cV'$ is continuous.
+FIXME any $g\in\cV'$; POINT OUT THAT IS STEADY ADVECTION-DIFFUSION VI; MAKE ARGUMENT FOLLOWING \cite{Elmanetal2014}; see also \cite{Kirby2010}; IF $\bX=0$ THEN IT IS OPTIMIZATION BUT OTHERWISE NOT; $2$-coercive if outflow on Neumann; It is easy to see that $|\ip{f(u)}{v}| \le c \|u\|_{\cV} \|v\|_{\cV}$, thus that $f:\cK \to \cV'$ is continuous.
 \end{example}

### 2021-07-04 paper/mcd2.tex
Put in the basic algorithms.

@@
 \newcommand{\supp}{\operatorname{supp}}
+\newcommand{\Span}{\operatorname{span}}
 
@@
 Suppose $\cV$ is a real, reflexive Banach space with norm $\|v\|$ for $v\in\cV$ and topological dual space $\cV'$.
-Denote the dual pairing of $\phi \in \cV'$ and $v$ by $\ip{\phi}{v} = \phi(v)$.
+Denote the dual pairing of $\phi \in \cV'$ and $v\in\cV$ by $\ip{\phi}{v} = \phi(v)$.
 Note that $\|\phi\|_{\cV'} = \sup_{\|v\|=1} |\ip{\phi}{v}|$ defines the norm on $\cV'$, which is also a Banach space.
@@
 \begin{equation}
-\ip{f(u)}{v-u} \ge g(v-u) \qquad \text{for all } v\in \cK. \label{eq:vi}
+\ip{f(u)}{v-u} \ge \ip{g}{v-u} \qquad \text{for all } v\in \cK. \label{eq:vi}
 \end{equation}
-Because $f$ is a general, nonlinear map the use of a right-hand source function is not strictly needed---by redefining $f$ we may take $g=0$---but its inclusion will be helpful in describing algorithms in Section \ref{sec:multilevel}.
+Because $f$ is a general, and generally nonlinear, map the inclusion of $g$ is not strictly needed---by redefining $f$ we may take $g=0$---but its inclusion will be helpful in describing algorithms in Section \ref{sec:multilevel}.
 
-Informally, if we conceptualize the dual pairing as an inner product, and if $g=0$, then \eqref{eq:vi} says that the angle between $f(u)$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$, that is, $f(u)$ points directly into $\cK$ if $u$ solves \eqref{eq:vi}.
-Furthermore, if $u$ is in the interior of $\cK$ then \eqref{eq:vi} implies $f(u)=g$; this VI generalizes the concept of a nonlinear system of equations $f(u)=g$ over $\cV$ to the constrained case where $\cK \subsetneq \cV$.
+VI \eqref{eq:vi} generalizes the nonlinear system of equations $f(u)=g$ over a vector space $\cV$ to the constrained case where $\cK \subsetneq \cV$ is a convex subset.
+Informally, if we conceptualize the dual pairing as an inner product, then \eqref{eq:vi} says that the angle between $f(u)-g$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
+That is, \eqref{eq:vi} says that $f(u)-g$ points directly into $\cK$ if $u\in\partial\cK$, while if $u$ is in the interior of $\cK$ then \eqref{eq:vi} implies $f(u)=g$.
 
@@
 The condition of continuity can be weakened to only apply on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but the stronger condition will apply in our examples.
-As in the calculus of variations \cite{Evans2010}, coercivity permits a weak compactness argument for unbounded sets $\cK$.
+As in the calculus of variations \cite{Evans2010}, coercivity permits a compactness argument even for unbounded sets $\cK$; recall that the bounded, closed subsets of a reflexive Banach space are weakly compact.
 
@@
 \begin{equation}
-\cV = \sum_{i=1}^m \cV_i \label{eq:subspacedecomp}
+\cV = \sum_{i=0}^{m-1} \cV_i \label{eq:subspacedecomp}
 \end{equation}
@@
 \begin{equation}
-\cK_i \subset \cV_i \qquad \text{and} \qquad \cK = \sum_{i=1}^m \cK_i. \label{eq:constraintdecomp}
+\cK_i \subset \cV_i \qquad \text{and} \qquad \cK = \sum_{i=0}^{m-1} \cK_i. \label{eq:constraintdecomp}
 \end{equation}
@@
 
-Following \cite{Tai2003}, for each $\cK_i$ we also assume a (generally) nonlinear restriction operator $R_i : \cK \to \cK_i$ such that
+Following \cite{Tai2003}, for each $\cK_i$ we also assume a (generally) nonlinear restriction operator $R_i : \cK \to \cK_i$ such that % SHOULD WE HAVE  R_i:\cV \to \cK_i?
 \begin{equation}
-v = \sum_{i=1}^m R_i v \qquad \text{if } v \in \cK.  \label{eq:constraintrestrictionsum}
+v = \sum_{i=0}^{m-1} R_i v \qquad \text{if } v \in \cK.  \label{eq:constraintrestrictionsum}
 \end{equation}
@@
 Let $\cK = \{v \ge \psi\} \subset \cV$.
-Suppose further that $\{\phi_i\}_{i=1}^m$ is a smooth partition of unity on $\Omega$, i.e.~so that $0 \le \phi_i\le 1$ and $\sum_i \phi_i = 1$, and let $\Omega_i = \supp \phi_i$ (open support).
+Suppose further that $\{\phi_i\}_{i=0}^{m-1}$ is a smooth partition of unity on $\Omega$, i.e.~so that $0 \le \phi_i\le 1$ and $\sum_i \phi_i = 1$, and let $\Omega_i = \supp \phi_i$ (open support).
 Let $\cV_i = \{w \in \cV:w|_{\Omega \setminus \Omega_i} =0 \}$, $\cK_i = \{v \in V: v \ge \phi_i \psi\}$, and $R_i(v) = \max\{v,\phi_i \psi\}$.
@@
 
-FIXME state the basic ``sequential constraint decomposition'' and ``parallel constraint decomposition'' algorithms
+\begin{pcode}[ht]
+\begin{pseudo*}
+\pr{cd-parallel}(f:\cK \to \cV', g\in \cV', \{\cK_i\}, \{R_i:\cK \to \cK_i\}, 0<\alpha\le 1, u\in\cK)\text{:} \\+
+    for $i \in \{0,\dots,m-1\}$: \qquad \qquad \ct{parallel for} \\+
+        $\hat w_i\in \cK_i$: \quad $\ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge g(v_i-\hat w_i) \quad \forall v_i\in \cK_i$ \\-
+    return $w=(1-\alpha) u + \alpha \sum_i \hat w_i$
+\end{pseudo*}
+\caption{The basic parallel CD algorithm updates an iterate $u$ by solving VIs over each constraint set $\cK_i$ in any order, giving $w\in \cK$.}
+\label{pc:cdparallel}
+\end{pcode}
 
-In practice the space $\cV$ will be a finite-dimensional finite element (FE) space because we will solve VI \eqref{eq:vi} over a mesh.  We FIXME triangulation etc. Though our
+\begin{pcode}[ht]
+\begin{pseudo*}
+\pr{cd-sequential}(f:\cK \to \cV', g\in \cV', \{\cK_i\}, \{R_i:\cK \to \cK_i\}, 0<\alpha\le 1, u\in\cK)\text{:} \\+
+    for $i = 0,\dots,m-1$: \qquad \qquad \ct{in order} \\+
+        $\hat w_i\in \cK_i$: \quad $\displaystyle \ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\Big)}{v_i-\hat w_i} \ge g(v_i-\hat w_i) \quad \forall v_i\in \cK_i$ \\
+        $w_i = (1-\alpha) R_i u + \alpha \hat w_i$ \\-
+    return $w= \sum_i w_i = (1-\alpha) u + \alpha \sum_i \hat w_i$
+\end{pseudo*}
+\caption{Each VI incorporates the previous solutions in this basic sequential CD algorithm; compare Pseudocode \ref{pc:cdparallel}.}
+\label{pc:cdsequential}
+\end{pcode}
 
-\begin{example}  FIXME $\cV_i$ are 1d spaces on a single level; $\psi,\cK,\cK_i$ for obstacle problem; note $\cK_i \not\subset \cK$ when $\psi>0$
+FIXME
+
+\begin{example}  FIXME if $\cV_i=\Span\{\phi_i\}$ for $\phi_i\in\cV$ are 1d spaces and if $\cK = \{v \ge \psi\} \subset \cV$ (obstacle problem), and if $R_i : \cK \to \cK_i$ for obstacle problem; note $\cK_i \not\subset \cK$ when $\psi>0$
 \end{example}
+
+In practice the space $\cV$ will be a finite element (FE) space because we will solve VI \eqref{eq:vi} over a mesh, and thus $\cV$ and each $\cV_i$ will be finite-dimensional.
 

### 2021-07-04 paper/mcd2.tex
More text and better structure.

@@
 \begin{document}
-\title[On MCD methods for nonlinear and nonlocal VIs]{On multilevel constraint decomposition methods \\ for nonlinear variational inequalities}
+\title[On MCD methods for nonlinear and nonlocal VIs]{On multilevel constraint decomposition methods \\ for nonlinear and nonlocal variational inequalities}
 
@@
 \item In the multilevel and nonlinear case, we make the algorithms more practical by addressing efficient storage of intermediate quantities.
-In particular we show results from a full approximation storage (FAS; \cite{Brandt1977}) version of the multilevel CD algorithm.
+In particular we show results from a full approximation storage (FAS; \cite{Brandt1977}) version of the multilevel CD algorithm (Section \ref{sec:resultslocal}).
 \end{enumerate}
@@
 
-In one example we consider a case where the residual functional is nonlocal, that is, where it is not a partial differential operator at all.
+In one example we consider a case where the residual functional is nonlocal, that is, where it is not a partial differential operator at all (Section \ref{sec:resultsnonlocal}).
 Instead its values instead arise from the solution of Stokes problems for a layer of fluid on a substrate which forms the obstacle \cite{SayagWorster2013}, so the corresponding FE operator discretization is not sparse.
@@
 
-When $f$ is monotone, VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
-In fact, suppose $F:\cV \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cV \to \cV'$.
-Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}, and if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ and $g=0$ if and only if
-\begin{equation}
-u = \operatorname{arg-min}_{v\in\cK} F(v) \label{eq:minimization}
-\end{equation}
-\cite[Proposition II.2.1]{EkelandTemam1976}.  Note that Tai \cite{Tai2003} addresses optimization problem \eqref{eq:minimization}.
-
-However, not all VIs correspond to optimization.
-For example, consider the following steady advection-diffusion problem over $H_0^1(\Omega)$.
-(We denote Sobolev spaces by $W^{k,p}(\Omega)$ if $k\ge 0$ and $p\ge 1$ \cite{Evans2010}, and $W^{k,2}=H^k$ in particular.)
-
-\begin{example}  \label{ex:advectiondiffusion}  Suppose $\Omega \subset \RR^d$ is bounded and open with polygonal boundary.
-Let $\cV = H_0^1(\Omega)$ and consider $\cK = \{v\ge 0\} \subset \cV$.
-For $\bX \in L^\infty(\Omega)^d$ such that $\Div \bX=0$, and $u$ in $\cK$ let
-\begin{equation}
-\ip{f(u)}{v} = \int_\Omega \grad u \cdot \grad v\,dx - (\bX \cdot \grad u) v\,dx. \label{eq:nongradientexample}
-\end{equation}
-FIXME any $g\in\cV'$; POINT OUT THAT IS STEADY ADVECTION-DIFFUSION VI; MAKE ARGUMENT FOLLOWING \cite{Elmanetal2014}; see also \cite{Kirby2010}; IF $\bX=0$ THEN IT IS OPTIMIZATION BUT OTHERWISE NOT; $2$-coercive if outflow on Neumann; It is easy to see that $|\ip{f(u)}{v}| \le c \|u\|_{\cV} \|v\|_{\cV}$, thus that $f:\cK \to \cV'$ is continuous.
-\end{example}
-
-Other examples of VI problems which are not of optimization type are found in ice sheet models \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids problems \cite{Bueler2021conservation}.
-
-The VIs solved in this paper satisfy a stronger form of coercivity than \eqref{eq:coercive}.  This stronger assumption will be used in proving convergence of MCD methods.
+The VIs solved in this paper satisfy a stronger form of coercivity than \eqref{eq:coercive}.\footnote{An exception is the problem in Section \ref{sec:resultsnonlocal}; there the functional $f$ is not known to be coercive.}
 
@@
 
-It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive, and thus the following result, which states that the VI problems considered in this paper are well-posed, is immediate.
+It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive, and thus the following result holds.
+It states that the $p$-coercive VI problems considered in this paper are well-posed.
 
-\begin{theorem}  \label{thm:viwellposed}  If $f:\cK \to \cV'$ is continuous and $p$-coercive (for $p>1$) then there exists a unique $u\in \cK$ so that VI \eqref{eq:vi} holds.
+\begin{theorem}  \label{thm:viwellposed}  If $f:\cK \to \cV'$ is continuous and $p$-coercive (for $p>1$) then there exists a unique $u\in \cK$ solving VI \eqref{eq:vi}.
 \end{theorem}
+
+When $f$ is monotone, VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
+In fact, suppose $F:\cV \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cV \to \cV'$.
+Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}, and if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ and $g=0$ if and only if
+\begin{equation}
+u = \operatorname{arg-min}_{v\in\cK} F(v) \label{eq:minimization}
+\end{equation}
+\cite[Proposition II.2.1]{EkelandTemam1976}.
+
+Note that some numerical solvers for VIs, including the constraint decomposition methods of Tai \cite{Tai2003} (below), address optimization problem \eqref{eq:minimization} subject to the assumption that $F'$ is 2-coercive.
+However, not all VIs correspond to optimization, and furthermore some optimization problems involve $p$-coercive residuals.
+(The latter possibility is mentioned by \cite{Tai2003}; see also \cite{TaiXu2002}.)  We provide three examples.
+Note that Sobolev spaces are denoted $W^{k,p}(\Omega)$ \cite{Evans2010}, with $W^{k,2}=H^k$ in particular.
+
+\begin{example}  \label{ex:advectiondiffusion}  Consider a steady advection-diffusion problem over a bounded domain $\Omega \subset \RR^d$, with polygonal boundary, and suppose the boundary is partitioned $\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$ with $\partial_D\Omega$ of positive measure.
+Let $\cV = H_0^1(\Omega)$ be the space of functions with value zero on $\partial_D\Omega$; for simplicity suppose zero Dirichlet and Neumann data.
+Consider a velocity field $\bX \in L^\infty(\Omega)^d$ such that $\Div \bX=0$, and define the functional
+\begin{equation}
+\ip{f(u)}{v} = \int_\Omega \eps \grad u \cdot \grad v - (\bX \cdot \grad u) v\,dx \label{eq:advectiondiffusion}
+\end{equation}
+for $\eps>0$.
+Finally, suppose a nonnegativity constraint $\cK = \{v\ge 0\} \subset \cV$ on the solution \cite[for example]{ChangNakshatrala2017}, and consider VI \eqref{eq:vi} for any $g\in\cV'$.
+It is easy to see that $|\ip{f(u)}{v}| \le (\eps + \|\bX\|_\infty) \|u\|_{\cV} \|v\|_{\cV}$, thus that $f:\cK \to \cV'$ is continuous.
+FIXME  MAKE ARGUMENT FOLLOWING \cite{Elmanetal2014}; see also \cite{Kirby2010}; IF $\bX=0$ THEN IT IS OPTIMIZATION BUT OTHERWISE NOT; $2$-coercive if outflow on Neumann
+\end{example}
+
+\begin{example}  \label{ex:porousgeneralization}  FIXME
+\begin{equation}
+\ip{f(u)}{v} = \int_\Omega \phi(u) \grad u \cdot \grad v\,dx \label{eq:porousgeneralization}
+\end{equation}
+\end{example}
+
+\begin{example}  \label{ex:plaplacian}  FIXME
+\begin{equation}
+\ip{f(u)}{v} = \int_\Omega |\grad u|^{p-2} \grad u \cdot \grad v\,dx \label{eq:plaplacian}
+\end{equation}
+\end{example}
+
+Examples \ref{ex:advectiondiffusion} and \ref{ex:porousgeneralization} are VI problems which do not arise from optimization.
+Further examples of such VIs are in ice sheet models \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids \cite{Bueler2021conservation}.
+Example \ref{ex:plaplacian} is an optimization-type problem of $p$-coercive type.
+Numerical solver performance for Examples \ref{ex:advectiondiffusion}--\ref{ex:plaplacian} will be considered in section \ref{sec:resultslocal}.
+% HOPE
 
@@
 \end{equation}
-The constraint decomposition \cite{Tai2003} sum in \eqref{eq:constraintdecomp} is required to hold in two senses: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.
+The sum in \eqref{eq:constraintdecomp} is required to hold in two senses: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.
 (The latter sense is automatic in \eqref{eq:subspacedecomp} because $\cV_i$ are subspaces.)  Note that neither decomposition is unique (in general), and that $\cK_i \not\subset \cK$ in many applications.
@@
 \end{equation}
-An \emph{constraint decomposition} of the constraint set $\cK$ is a choice of $\cV_i,\cK_i,R_i$ satisfying the above hypotheses.
+A \emph{constraint decomposition} (CD) of the set $\cK$ is a choice of $\cV_i,\cK_i,R_i$ satisfying the above hypotheses.
 
-In the next section we will introduce FE spaces and describe practical algorithms, but, as noted in the Introduction, constraint decompositions may be defined at the level of the continuum problem.
-The following examples illustrate constraint decompositions of \emph{obstacle problems} \cite{GraeserKornhuber2009}; Example \ref{ex:domaindecomposition} is an overlapping domain decomposition and \ref{ex:frequencydecomposition} is a nonoverlapping frequency decomposition.
-The latter example can be compared to the multilevel FE decomposition in the next section.
+In the Section \ref{sec:multilevel} we will introduce discretizations and describe practical algorithms, but CDs may be defined at the level of the continuum problem.
+The following examples illustrate CDs of obstacle problems \cite{GraeserKornhuber2009}.
+Example \ref{ex:domaindecomposition} is an overlapping domain decomposition and Example \ref{ex:frequencydecomposition} is a nonoverlapping frequency decomposition.
+The latter can be compared to the multilevel FE decomposition in Section \ref{sec:multilevel}.
 
@@
 
-\begin{pcode}[ht]
+We now state the basic CD algorithms as iterations which solve VI problems over each set $\cK_i$.
+(By contrast, the algorithms in \cite{Tai2003} involve iterative optimization over the $\cK_i$.)  There is a parallel (additive) form and a successive (multiplicative) form; these generalize the Jacobi and Gauss-Seidel iterations \cite{Greenbaum1997}, respectively.
+Each algorithm is stated as a function which computes a new iterate $w\in\cK$ from a previous iterate $y \in \cK$, the result being an improved approximation of the solution $u\in \cK$ of VI problem \eqref{eq:vi}.
+
 \begin{pseudo*}
-\pr{cd-parallel}(f:\cK \to \cV', g\in \cV', \{\cK_i\}, \{R_i:\cK \to \cK_i\}, 0<\alpha\le 1, u\in\cK)\text{:} \\+
+\pr{cd-parallel}(\{\cK_i\}, \{R_i:\cK \to \cK_i\}, 0<\alpha\le 1, y\in\cK)\text{:} \\+
     for $i \in \{0,\dots,m-1\}$: \qquad \qquad \ct{parallel for} \\+
-        $\hat w_i\in \cK_i$: \quad $\ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge g(v_i-\hat w_i) \quad \forall v_i\in \cK_i$ \\-
-    return $w=(1-\alpha) u + \alpha \sum_i \hat w_i$
+        $\hat w_i\in \cK_i$: \quad $\ip{f(y - R_i y + \hat w_i)}{v_i-\hat w_i} \ge g(v_i-\hat w_i) \quad \forall v_i\in \cK_i$ \\-
+    return $w=(1-\alpha) y + \alpha \sum_i \hat w_i$
 \end{pseudo*}
-\caption{The basic parallel CD algorithm updates an iterate $u$ by solving VIs over each constraint set $\cK_i$ in any order, giving $w\in \cK$.}
-\label{pc:cdparallel}
-\end{pcode}
 
-\begin{pcode}[ht]
 \begin{pseudo*}
-\pr{cd-sequential}(f:\cK \to \cV', g\in \cV', \{\cK_i\}, \{R_i:\cK \to \cK_i\}, 0<\alpha\le 1, u\in\cK)\text{:} \\+
+\pr{cd-successive}(\{\cK_i\}, \{R_i:\cK \to \cK_i\}, 0<\alpha\le 1, y\in\cK)\text{:} \\+
     for $i = 0,\dots,m-1$: \qquad \qquad \ct{in order} \\+
-        $\hat w_i\in \cK_i$: \quad $\displaystyle \ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\Big)}{v_i-\hat w_i} \ge g(v_i-\hat w_i) \quad \forall v_i\in \cK_i$ \\
-        $w_i = (1-\alpha) R_i u + \alpha \hat w_i$ \\-
-    return $w= \sum_i w_i = (1-\alpha) u + \alpha \sum_i \hat w_i$
+        $\hat w_i\in \cK_i$: \quad $\displaystyle \ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j y\Big)}{v_i-\hat w_i} \ge g(v_i-\hat w_i) \quad \forall v_i\in \cK_i$ \\
+        $w_i = (1-\alpha) R_i y + \alpha \hat w_i$ \\-
+    return $w = (1-\alpha) y + \alpha \sum_i \hat w_i$
 \end{pseudo*}
-\caption{Each VI incorporates the previous solutions in this basic sequential CD algorithm; compare Pseudocode \ref{pc:cdparallel}.}
-\label{pc:cdsequential}
-\end{pcode}
 
-FIXME
+Note that we assume a damping parameter $0<\alpha\le 1$.
+The convergence results in the next section require substantial damping in the parallel algorithm, i.e.~$\alpha \le 1/m$, but the successive version can be shown to converge without damping ($\alpha=1$).
+In Section \ref{sec:resultslocal} we demonstrate practical convergence for a larger range of $\alpha$ than suggested by the theory.
+% HOPE
 
-\begin{example}  FIXME if $\cV_i=\Span\{\phi_i\}$ for $\phi_i\in\cV$ are 1d spaces and if $\cK = \{v \ge \psi\} \subset \cV$ (obstacle problem), and if $R_i : \cK \to \cK_i$ for obstacle problem; note $\cK_i \not\subset \cK$ when $\psi>0$
-\end{example}
+Observe that the expression $v_i - \hat w_i$ for $v_i,\hat w_i \in \cK_i$ appears in each algorithm, but, on the other hand, the reader may confirm that the argument to the residual functional $f$ is always an element of $\cK$.
+To clarify this situation, note that by \eqref{eq:constraintrestrictionsum} one may write $v_i - \hat w_i$ as a difference of admissible vectors (i.e.~from $\cK$), namely
+\begin{align*}
+[y - R_i y + v_i] - [y - R_i y + \hat w_i] = \left[\sum_{j\ne i} R_j y + v_i\right] - \left[\sum_{j\ne i} R_j y + \hat w_i\right] &= v_i - \hat w_i, \label{eq:admissibledifference} \\
+\left[\sum_{j<i} w_j + v_i + \sum_{j>i} R_j y\right] - \left[\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j y\right] &= v_i - \hat w_i,  \notag
+\end{align*}
+for the two versions, respectively.  In other words, these algorithms require the approximate solution of VI \eqref{eq:vi} by looking among perturbations from the sets $\cK_i$.
 
-In practice the space $\cV$ will be a finite element (FE) space because we will solve VI \eqref{eq:vi} over a mesh, and thus $\cV$ and each $\cV_i$ will be finite-dimensional.
-
-\section{Multilevel constraint decomposition algorithms} \label{sec:multilevel}
-
-FIXME state essentially Algorithm 4.7 \cite{GraeserKornhuber2009} but with $\text{V}(\nu_1,\nu_2)$ cycles which works for linear; observe that up-smoothing is more efficient; state the FAS version which has $O(m)$ residual evaluation complexity on each level
+FIXME key conceptual difference between the nonlinear and linear case; rewrite VI in \pr{cd-parallel} in the linear case; non-basic algorithm needed for $O(m)$ residual evaluation complexity in the nonlinear case (Section \ref{sec:multilevel})
 
@@
 
-\section{Results} \label{sec:results}
+\section{Finite elements and multilevel constraint decomposition} \label{sec:multilevel}
+
+Until now the algorithms can be interpreted as infinite-dimensional, e.g.~supposing constraint sets as defined in Examples \ref{ex:domaindecomposition} or \ref{ex:frequencydecomposition}.
+However, in practice we will solve VI \eqref{eq:vi} over a mesh and the space $\cV$ will be a finite element (FE) space.
+Thus $\cV$, and each subspace $\cV_i$, will be finite-dimensional.
+
+FIXME triangulation
+
+\begin{example}  FIXME if $\cV_i=\Span\{\phi_i\}$ for $\phi_i\in\cV$ are 1d spaces and if $\cK = \{v \ge \psi\} \subset \cV$ (obstacle problem), and if $R_i : \cK \to \cK_i$ for obstacle problem; note $\cK_i \not\subset \cK$ when $\psi>0$
+\end{example}
+
+FIXME state essentially Algorithm 4.7 \cite{GraeserKornhuber2009} but with $\text{V}(\nu_1,\nu_2)$ cycles which works for linear; observe that up-smoothing is more efficient; state the FAS version which has $O(m)$ residual evaluation complexity on each level
+
+\section{Results for local variational inequalities} \label{sec:resultslocal}
+
+FIXME
+
+\section{Results for a nonlocal variational inequality} \label{sec:resultsnonlocal}
 
@@
 \small
-
-\bigskip
 \bibliography{mcd2}

### 2021-07-05 paper/mcd2.tex
Attempt to clarify the issues with the basic algorithms.

@@
 In its multilevel form the CD method has been shown to have optimal complexity for elliptic, linear obstacle problems \cite[subsection 5.4]{Tai2003}; see also Theorem 4.6 and Algorithm 4.7 in \cite{GraeserKornhuber2009}.
-However, especially for the classical obstacle problem, i.e.~for a Laplacian functional, certain other methods are known to improve performance relative to the multilevel CD method \cite{GraeserKornhuber2009}.
-On the other hand, these improved multilevel methods either track the active set or modify the nodal basis functions, and in this sense they are discrete algorithms.
-By contrast the CD methods apply at the level of the continuum problem (see below), so the method is more fundamental than the truncated monotone multigrid method \cite{GraeserKornhuber2009,Kornhuber1994}, for example, in that sense.
 
@@
 
-In one example we consider a case where the residual functional is nonlocal, that is, where it is not a partial differential operator at all (Section \ref{sec:resultsnonlocal}).
-Instead its values instead arise from the solution of Stokes problems for a layer of fluid on a substrate which forms the obstacle \cite{SayagWorster2013}, so the corresponding FE operator discretization is not sparse.
-In this case we cannot prove coercivity but we can demonstrate near-optimal complexity in practice.
+Especially for the classical obstacle problem, i.e.~for a Laplacian functional, certain other methods are known to improve performance relative to the multilevel CD method \cite{GraeserKornhuber2009}.
+These improved multilevel methods either track the active set or modify the nodal basis functions, and in this sense they are discrete algorithms.
+By contrast the CD methods apply at the level of the continuum problem (see below); CD methods are more fundamental than the truncated monotone multigrid method \cite{GraeserKornhuber2009,Kornhuber1994}, for example, in that sense.
+
+In one example at the end of this paper (Section \ref{sec:resultsnonlocal}) we consider a nonlocal residual functional, that is, one which is not a partial differential operator.
+One evaluation of this functional requires the solution of a Stokes problem for a layer of fluid \nocite{SayagWorster2013} on a substrate (which forms the obstacle), so the corresponding FE operator discretization is not sparse.
+In this case we cannot prove coercivity but we nonetheless succeed in demonstrating near-optimal complexity in practice.
 % HOPE
@@
 (By contrast, the algorithms in \cite{Tai2003} involve iterative optimization over the $\cK_i$.)  There is a parallel (additive) form and a successive (multiplicative) form; these generalize the Jacobi and Gauss-Seidel iterations \cite{Greenbaum1997}, respectively.
-Each algorithm is stated as a function which computes a new iterate $w\in\cK$ from a previous iterate $y \in \cK$, the result being an improved approximation of the solution $u\in \cK$ of VI problem \eqref{eq:vi}.
+Each algorithm computes a new iterate $w\in\cK$, an improved approximation of the solution $u\in \cK$ of VI problem \eqref{eq:vi} relative to the previous iterate $y \in \cK$.
+A damping parameter $0<\alpha\le 1$ is included.
 
@@
 \pr{cd-parallel}(\{\cK_i\}, \{R_i:\cK \to \cK_i\}, 0<\alpha\le 1, y\in\cK)\text{:} \\+
-    for $i \in \{0,\dots,m-1\}$: \qquad \qquad \ct{parallel for} \\+
-        $\hat w_i\in \cK_i$: \quad $\ip{f(y - R_i y + \hat w_i)}{v_i-\hat w_i} \ge g(v_i-\hat w_i) \quad \forall v_i\in \cK_i$ \\-
+    for $i \in \{0,\dots,m-1\}$: \qquad \qquad \ct{parallel} \\+
+        $\hat w_i\in \cK_i$: \quad $\boxed{\ip{f(y - R_i y + \hat w_i)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\-
     return $w=(1-\alpha) y + \alpha \sum_i \hat w_i$
@@
     for $i = 0,\dots,m-1$: \qquad \qquad \ct{in order} \\+
-        $\hat w_i\in \cK_i$: \quad $\displaystyle \ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j y\Big)}{v_i-\hat w_i} \ge g(v_i-\hat w_i) \quad \forall v_i\in \cK_i$ \\
+        $\hat w_i\in \cK_i$: \quad $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j y\Big)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\
         $w_i = (1-\alpha) R_i y + \alpha \hat w_i$ \\-
@@
 
-Note that we assume a damping parameter $0<\alpha\le 1$.
+Let $e_i = \hat w_i - R_i y \in \cV_i$ be the $i$th-subspace change (in either algorithm), and note we may then write $w = y + \alpha \sum_i e_i$ when needed.
+Also note for \pr{cd-parallel} that $y - R_i y + \hat w_i = y+e_i$.
+
+The reader may confirm that, inside each boxed VI, the argument of $f$ is an element of $\cK$.
+On the other hand, the expression $v_i - \hat w_i$ appears.
+To clarify this situation, note that by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write $v_i - \hat w_i$ as a difference of admissible vectors (i.e.~from $\cK$), namely
+\begin{align*}
+[y - R_i y + v_i] - [y - R_i y + \hat w_i] &= v_i - \hat w_i, \label{eq:admissibledifference} \\
+\left[\sum_{j<i} w_j + v_i + \sum_{j>i} R_j y\right] - \left[\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j y\right] &= v_i - \hat w_i,  \notag
+\end{align*}
+for the two versions, respectively.  In other words, these algorithms require the solution of VI \eqref{eq:vi} over (parallel or successive) perturbations from the sets $\cK_i$.
+
 The convergence results in the next section require substantial damping in the parallel algorithm, i.e.~$\alpha \le 1/m$, but the successive version can be shown to converge without damping ($\alpha=1$).
@@
 
-Observe that the expression $v_i - \hat w_i$ for $v_i,\hat w_i \in \cK_i$ appears in each algorithm, but, on the other hand, the reader may confirm that the argument to the residual functional $f$ is always an element of $\cK$.
-To clarify this situation, note that by \eqref{eq:constraintrestrictionsum} one may write $v_i - \hat w_i$ as a difference of admissible vectors (i.e.~from $\cK$), namely
-\begin{align*}
-[y - R_i y + v_i] - [y - R_i y + \hat w_i] = \left[\sum_{j\ne i} R_j y + v_i\right] - \left[\sum_{j\ne i} R_j y + \hat w_i\right] &= v_i - \hat w_i, \label{eq:admissibledifference} \\
-\left[\sum_{j<i} w_j + v_i + \sum_{j>i} R_j y\right] - \left[\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j y\right] &= v_i - \hat w_i,  \notag
-\end{align*}
-for the two versions, respectively.  In other words, these algorithms require the approximate solution of VI \eqref{eq:vi} by looking among perturbations from the sets $\cK_i$.
+In Section \ref{sec:multilevel} we will implement these basic algorithms for multilevel CDs.
+However, this implementation requires additional concepts in the case that $f$ is either: \emph{(i)} defined only on $\cK$, \emph{(ii)} nonlinear, or \emph{(iii)} nonlocal.
+(In fact we will extend the above methods by applying the full approximation storage idea of Brandt \cite{Brandt1977}; see Section \ref{sec:multilevel}.)  The issue seems not to have been addressed by references which apply CD algorithms only to the classical obstacle problem \cite[in particular]{GraeserKornhuber2009,Tai2003}.
 
-FIXME key conceptual difference between the nonlinear and linear case; rewrite VI in \pr{cd-parallel} in the linear case; non-basic algorithm needed for $O(m)$ residual evaluation complexity in the nonlinear case (Section \ref{sec:multilevel})
+To explain our concern, suppose $f:\cV \to \cV'$ is linear and local (in a sense which is explained momentarily).
+Only consider \pr{cd-parallel}, for simplicity, and suppose for concreteness that $\cV = H_0^1(\Omega)$ (for some domain $\Omega \subset \RR^d$).
+Note that for $z_i \in \cV_i$ we have $\ip{f(y - R_i y + \hat w_i)}{z_i} = \ip{f(y)}{z_i} + \ip{f(\hat w_i - R_i y)}{z_i}$, so the VI over $\cK_i$ can be written as
+\begin{equation}
+\ip{f(\hat w_i - R_i y)}{v_i-\hat w_i} \ge \tilde g(v_i-\hat w_i)  \qquad\qquad \text{\emph{($f$ linear)}} \label{eq:linearlocalcdparallel}
+\end{equation}
+for all $v_i\in \cK_i$, where $\tilde g(z) = g(z) - \ip{f(y)}{z}$.
+Note $e_i = \hat w_i - R_i y \in \cV_i$ but generally $e_i \notin \cK$, so \eqref{eq:linearlocalcdparallel} only make sense if $f$ is defined over the whole vector space.
+Now each VI problem in \pr{cd-parallel} can then be solved using a stored residual $\ip{f(y)}{\cdot}$ which is put in a fixed source term.
+Furthermore, by $f$ being ``local'' we mean that if a basis $\{\phi_i\}$ of $\cV_i$ can be constructed with each support small in $\Omega$, then $\ip{f(\phi_i)}{z}$ can be computed by an integral over the support of $\phi_i$.
+Then the solution of each VI problem \label{eq:linearlocalcdparallel} can be computed by local means over the union of the supports of the basis.
+Thus the resulting solver for the everywhere-defined, linear, and local $f$ has implementation efficiencies that are unavailable otherwise.
+In conclusion, additional ideas, like those in the algorithms in Section \ref{sec:multilevel}, are needed for practical application of CD methods to problems which are not close to the classical obstacle problem.
 

### 2021-07-05 paper/mcd2.tex
Edits on the first three pages.

@@
 
-We extend the original algorithms and theory presented in \cite{Tai2003} in three particular directions:
+We extend the original algorithms, theory, and examples presented in \cite{Tai2003} in four particular directions:
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
 \begin{enumerate}
-\item We do not assume that the continuum VI problem arises from optimization of a scalar objective, and, as shown by examples in Section \ref{sec:vi}, this represents the removal of a symmetry assumption in \cite{Tai2003}.
-Nonetheless we can prove convergence at essentially the same rate (Section \ref{sec:convergence}).
-% HOPE
-\item In the multilevel case we make the observation that, if implemented appropriately, V(0,1) ``up-slash'' cycles are intrinsically more efficient than the V(1,0) ``down-slash'' cycles (Section \ref{sec:multilevel}).
+\item We do not assume that the continuum VI problem arises from optimization of a scalar objective, an assumption in \cite{Tai2003}.
+Nonetheless we can prove convergence at the same rate (Section \ref{sec:convergence}); % HOPE
+examples are explored in Sections \ref{sec:vi}, \ref{sec:resultslocal}, and \ref{sec:resultsnonlocal}.
+\item We observe that the iterates from these CD methods are always admissible, and thus the residual functional need only be defined on the admissible subset; Sections \ref{sec:resultslocal} and \ref{sec:resultsnonlocal} include examples.
+Admissible-iterate methods allow direct solutions of certain VI problems, such as fluid-layer dynamics problems (Section \ref{sec:resultsnonlocal}; see also \cite{Bueler2021conservation,JouvetBueler2012}) for which non-admissible methods, such as semi-smooth methods \cite{BensonMunson2006}, may require unnatural modifications of the residual form.
+\item In the multilevel case we make the observation that, if implemented appropriately, V(0,$\nu$) ``up-slash'' cycles are intrinsically more efficient than the V($\nu$,0) ``down-slash'' cycles (Section \ref{sec:multilevel}).
 This observation seems to be new for multilevel CD methods; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}.
-This preference for up-smoothing is, apparently, special to CD methods and does not arise in corresponding unconstrained subspace decomposition \cite{Xu1992} approaches.
-\item In the multilevel and nonlinear case, we make the algorithms more practical by addressing efficient storage of intermediate quantities.
-In particular we show results from a full approximation storage (FAS; \cite{Brandt1977}) version of the multilevel CD algorithm (Section \ref{sec:resultslocal}).
+Strong preference for up-smoothing is, apparently, special to multilevel CD methods and does not arise in corresponding unconstrained (e.g.~\cite{Xu1992}) problems.
+\item In the multilevel and nonlinear case, we make the algorithms practical by addressing efficient storage of intermediate quantities.
+In particular we show results from a full approximation storage (FAS; \cite{Brandt1977}) version of the multilevel CD algorithm (Section \ref{sec:resultslocal}), avoiding global (Newton) linearization; compare \cite{GraeserKornhuber2009}.
 \end{enumerate}
-
-As an additional difference from the presentation in \cite{Tai2003}, but an essentially formal one, we observe that the iterates from CD methods are always admissible, and thus the residual functional need only be defined on the admissible subset.
-Methods like CD, for which all iterates are admissible, have an advantage because they allow direct solutions of certain VI problems, such as the fluid-layer dynamics problems in \cite{Bueler2021conservation,JouvetBueler2012}, for which non-admissible methods such as semi-smooth methods \cite{BensonMunson2006} may require unnatural modifications of the residual form.
 
@@
 These improved multilevel methods either track the active set or modify the nodal basis functions, and in this sense they are discrete algorithms.
-By contrast the CD methods apply at the level of the continuum problem (see below); CD methods are more fundamental than the truncated monotone multigrid method \cite{GraeserKornhuber2009,Kornhuber1994}, for example, in that sense.
+By contrast the CD methods apply at the level of the continuum problem (see below); CD methods are more fundamental than the truncated monotone multigrid method \cite{Kornhuber1994}, for example, in that sense, and we can demonstrate comparable rates of convergence (Section \ref{sec:resultslocal}).
+% HOPE
+Furthermore, truncated monotone approaches for nonlinear problems have, so far, applied a global linearization strategy \cite{GraeserKornhuber2009}, while we apply FAS (Section \ref{sec:multilevel}).
+In any case, acceleration of our nonlinear multilevel CD methods by such active-set and/or basis-level manipulations is a topic for further research.
 
 In one example at the end of this paper (Section \ref{sec:resultsnonlocal}) we consider a nonlocal residual functional, that is, one which is not a partial differential operator.
-One evaluation of this functional requires the solution of a Stokes problem for a layer of fluid \nocite{SayagWorster2013} on a substrate (which forms the obstacle), so the corresponding FE operator discretization is not sparse.
-In this case we cannot prove coercivity but we nonetheless succeed in demonstrating near-optimal complexity in practice.
+Each evaluation of this functional requires the solution of a Stokes problem for a layer of fluid \nocite{SayagWorster2013} on a substrate (which forms the obstacle), and thus the corresponding FE operator discretization is also not sparse.
+In this case we cannot prove coercivity but we nonetheless succeed in demonstrating near-optimal complexity of the Section \ref{sec:multilevel} multilevel CD algorithm in practice.
 % HOPE
 
-\section{Coercive variational inequalities and constraint decomposition} \label{sec:vi}
+\section{Coercive variational inequalities} \label{sec:vi}
 
-Suppose $\cV$ is a real, reflexive Banach space with norm $\|v\|$ for $v\in\cV$ and topological dual space $\cV'$.
-Denote the dual pairing of $\phi \in \cV'$ and $v\in\cV$ by $\ip{\phi}{v} = \phi(v)$.
-Note that $\|\phi\|_{\cV'} = \sup_{\|v\|=1} |\ip{\phi}{v}|$ defines the norm on $\cV'$, which is also a Banach space.
+Suppose $\cV$ is a real, reflexive Banach space with norm $\|\cdot\|$ and topological dual space $\cV'$.
+Denote the dual pairing of $\phi \in \cV'$ and $v\in\cV$ by $\ip{\phi}{v} = \phi(v)$, and note that $\|\phi\|_{\cV'} = \sup_{\|v\|=1} |\ip{\phi}{v}|$ defines a Banach space norm on $\cV'$.
 
-Let $\cK \subset \cV$ be a nonempty closed and convex subset of $\cV$, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.
-For a continuous \emph{residual functional} $f:\cK \to \cV'$ and \emph{source} $g\in \cV'$ we consider the following \emph{variational inequality} (VI) for $u\in \cK$:
+Let $\cK \subset \cV$ be a nonempty closed and convex subset, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.
+For a continuous \emph{residual functional} $f:\cK \to \cV'$ and \emph{source} $g\in \cV'$ we consider the following \emph{variational inequality} (VI) for the (exact) solution $u^*\in \cK$:
 \begin{equation}
-\ip{f(u)}{v-u} \ge \ip{g}{v-u} \qquad \text{for all } v\in \cK. \label{eq:vi}
+\ip{f(u^*)}{v-u^*} \ge \ip{g}{v-u^*} \qquad \text{for all } v\in \cK. \label{eq:vi}
 \end{equation}
-Because $f$ is a general, and generally nonlinear, map the inclusion of $g$ is not strictly needed---by redefining $f$ we may take $g=0$---but its inclusion will be helpful in describing algorithms in Section \ref{sec:multilevel}.
+Because $f$ is a (generally) nonlinear map, $g$ is not strictly needed, but its presence is helpful to the algorithms of Section \ref{sec:multilevel}.  (By redefining $f$ we may take $g=0$.)
 
-VI \eqref{eq:vi} generalizes the nonlinear system of equations $f(u)=g$ over a vector space $\cV$ to the constrained case where $\cK \subsetneq \cV$ is a convex subset.
-Informally, if we conceptualize the dual pairing as an inner product, then \eqref{eq:vi} says that the angle between $f(u)-g$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
-That is, \eqref{eq:vi} says that $f(u)-g$ points directly into $\cK$ if $u\in\partial\cK$, while if $u$ is in the interior of $\cK$ then \eqref{eq:vi} implies $f(u)=g$.
+VI \eqref{eq:vi} generalizes the nonlinear system of equations $f(u^*)=g$ from a vector space $\cV$ to the constrained case (over $\cK$).
+Informally, if we conceptualize the dual pairing as an inner product, then \eqref{eq:vi} says that the angle between $f(u^*)-g$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
+That is, \eqref{eq:vi} says that $f(u^*)-g$ points directly into $\cK$ if $u^* \in\partial\cK$, while if $u^* \in \cK^\circ$ (interior) then \eqref{eq:vi} implies $f(u^*)=g$.
 
@@
 
-It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, and also that the solution is unique if $f$ is strictly monotone.
+It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, and also that the solution $u^* \in \cK$ is unique if $f$ is strictly monotone.
 The condition of continuity can be weakened to only apply on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but the stronger condition will apply in our examples.
@@
 
-The VIs solved in this paper satisfy a stronger form of coercivity than \eqref{eq:coercive}.\footnote{An exception is the problem in Section \ref{sec:resultsnonlocal}; there the functional $f$ is not known to be coercive.}
+The VIs solved in this paper satisfy a stronger form of coercivity than \eqref{eq:coercive}.\footnote{An exception is the problem in Section \ref{sec:resultsnonlocal}; there $f$ is not known to be monotone or coercive.}
 
@@
 
-It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive, and thus the following result holds.
-It states that the $p$-coercive VI problems considered in this paper are well-posed.
+It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive, and thus the following result, which states that the $p$-coercive VI problems considered in this paper are well-posed, holds.
 
@@
 \begin{equation}
-u = \operatorname{arg-min}_{v\in\cK} F(v) \label{eq:minimization}
+u^* = \operatorname{arg-min}_{v\in\cK} F(v) \label{eq:minimization}
 \end{equation}
@@
 
-Note that some numerical solvers for VIs, including the constraint decomposition methods of Tai \cite{Tai2003} (below), address optimization problem \eqref{eq:minimization} subject to the assumption that $F'$ is 2-coercive.
-However, not all VIs correspond to optimization, and furthermore some optimization problems involve $p$-coercive residuals.
-(The latter possibility is mentioned by \cite{Tai2003}; see also \cite{TaiXu2002}.)  We provide three examples.
-Note that Sobolev spaces are denoted $W^{k,p}(\Omega)$ \cite{Evans2010}, with $W^{k,2}=H^k$ in particular.
+Before giving examples, note that Sobolev spaces are denoted $W^{k,p}(\Omega)$ \cite{Evans2010}, with $W^{k,2}=H^k$ in particular.
+
+\begin{example}  \label{ex:plaplacian}  FIXME
+\begin{equation}
+\ip{f(u)}{v} = \int_\Omega |\grad u|^{p-2} \grad u \cdot \grad v\,dx \label{eq:plaplacian}
+\end{equation}
+\end{example}
+
+Some numerical solvers for VIs, including the CD methods of Tai \cite{Tai2003}, address optimization problem \eqref{eq:minimization}.
+However, not all VIs correspond to optimization, and we give two such examples next.
 
@@
 
-\begin{example}  \label{ex:plaplacian}  FIXME
-\begin{equation}
-\ip{f(u)}{v} = \int_\Omega |\grad u|^{p-2} \grad u \cdot \grad v\,dx \label{eq:plaplacian}
-\end{equation}
-\end{example}
-
 Examples \ref{ex:advectiondiffusion} and \ref{ex:porousgeneralization} are VI problems which do not arise from optimization.
-Further examples of such VIs are in ice sheet models \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids \cite{Bueler2021conservation}.
-Example \ref{ex:plaplacian} is an optimization-type problem of $p$-coercive type.
+Further examples are in ice sheet models \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids \cite{Bueler2021conservation}.
 Numerical solver performance for Examples \ref{ex:advectiondiffusion}--\ref{ex:plaplacian} will be considered in section \ref{sec:resultslocal}.
 % HOPE
+
+\section{Constraint decomposition; the basic algorithms} \label{sec:cd}
 

### 2021-07-06 paper/mcd2.tex
Edits on CD.

@@
 
-Our multilevel algorithms (next section) are based on both subspace and constraint decompositions, as follows.  Suppose there are $m<\infty$ subspaces $\cV_i \subset \cV$ so that
+The multilevel algorithms in Section \ref{sec:multilevel} are based on the more general notion of a solution by constraint decomposition, as follows.
+Suppose there are $m<\infty$ subspaces $\cV_i \subset \cV$ so that
 \begin{equation}
@@
 \end{equation}
-holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$ \cite{Xu1992}.
+holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$; this is called a \emph{subspace decomposition} \cite{Xu1992}.
 Suppose further that $\cK_i \subset \cV$ are nonempty, closed, and convex subsets such that
@@
 The sum in \eqref{eq:constraintdecomp} is required to hold in two senses: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.
-(The latter sense is automatic in \eqref{eq:subspacedecomp} because $\cV_i$ are subspaces.)  Note that neither decomposition is unique (in general), and that $\cK_i \not\subset \cK$ in many applications.
+(The latter sense is automatic in \eqref{eq:subspacedecomp} because $\cV_i$ are subspaces.)  Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is unique (in general), and that $\cK_i \not\subset \cK$ in many applications.
 
-Following \cite{Tai2003}, for each $\cK_i$ we also assume a (generally) nonlinear restriction operator $R_i : \cK \to \cK_i$ such that % SHOULD WE HAVE  R_i:\cV \to \cK_i?
+Following \cite{Tai2003}, for each $\cK_i$ we also assume a bounded, (generally) nonlinear restriction operator $R_i : \cK \to \cK_i$ such that
 \begin{equation}
@@
 \end{equation}
-A \emph{constraint decomposition} (CD) of the set $\cK$ is a choice of $\cV_i,\cK_i,R_i$ satisfying the above hypotheses.
+A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,R_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}.
 
-In the Section \ref{sec:multilevel} we will introduce discretizations and describe practical algorithms, but CDs may be defined at the level of the continuum problem.
-The following examples illustrate CDs of obstacle problems \cite{GraeserKornhuber2009}.
+In the Section \ref{sec:multilevel} we will introduce discretizations and describe practical algorithms, but the CDs in this section can apply at the level of the continuum problem.
+The following two examples illustrate this for obstacle problems \cite{GraeserKornhuber2009}.
 Example \ref{ex:domaindecomposition} is an overlapping domain decomposition and Example \ref{ex:frequencydecomposition} is a nonoverlapping frequency decomposition.
-The latter can be compared to the multilevel FE decomposition in Section \ref{sec:multilevel}.
+(The multilevel FE CD in Section \ref{sec:multilevel} approximately implements a frequency decomposition.)
 
-\begin{example}  \label{ex:domaindecomposition}  Consider a bounded domain $\Omega \subset \RR^d$ with polygonal boundary, let $\cV = H_0^k(\Omega)$, $k\ge 1$, and suppose the obstacle $\psi \in H^k(\Omega)$ satisfies $\psi|_{\partial \Omega} \le 0$.
+\begin{example}  \label{ex:domaindecomposition}  Consider a bounded domain $\Omega \subset \RR^d$, let $\cV = W_0^{k,p}(\Omega)$ for $k\ge 0$ and $p\ge 1$, and suppose the obstacle $\psi \in W^{k,p}(\Omega)$ satisfies $\psi|_{\partial \Omega} \le 0$.
 Let $\cK = \{v \ge \psi\} \subset \cV$.
-Suppose further that $\{\phi_i\}_{i=0}^{m-1}$ is a smooth partition of unity on $\Omega$, i.e.~so that $0 \le \phi_i\le 1$ and $\sum_i \phi_i = 1$, and let $\Omega_i = \supp \phi_i$ (open support).
-Let $\cV_i = \{w \in \cV:w|_{\Omega \setminus \Omega_i} =0 \}$, $\cK_i = \{v \in V: v \ge \phi_i \psi\}$, and $R_i(v) = \max\{v,\phi_i \psi\}$.
+Suppose further that $\{\phi_i\}_{i=0}^{m-1}$ is a smooth partition of unity on $\Omega$, i.e.~so that $0 \le \phi_i\le 1$ and $\sum_i \phi_i = 1$, and let $\Omega_i$ be the support of $\phi_i$.
+Let $\cV_i = \{w \in \cV:w|_{\Omega \setminus \Omega_i} =0 \}$, $\cK_i = \{v \in \cV_i: v \ge \phi_i \psi\}$, and $R_i(v) = \phi_i v$.
 Then \eqref{eq:subspacedecomp}, \eqref{eq:constraintdecomp}, and \eqref{eq:constraintrestrictionsum} all hold.
@@
 
-\begin{example}  \label{ex:frequencydecomposition}  For simplicity suppose $\Omega = (0,a)^d \subset \RR^d$ for $a>0$, a cube, and let $\cV = H_{\text{per}}^k(\Omega)$, $k\ge 1$, be the periodic functions.
+%Note that the boundedness of $R_i(v) = \max\{v,\phi_i \psi\}$ depends on the relatively-low regularity of $\cV = H_0^1(\Omega)$.
+That is, if $u,v \in H^{1,p}(\Omega)$ for $1\le p \le \infty$ then $\max\{u,v\} \in H^{1,p}(\Omega)$.
+(To prove this note $\max\{u,0\} \in H^{1,p}(\Omega)$ is shown by \cite[Theorem II.A.1]{KinderlehrerStampacchia1980}; then combine with facts $\max\{a,b\} = \frac{1}{2} (a + b + |a-b|)$ and $|a| = \max\{a,0\} + \max\{-a,0\}$ for $a\in\RR$.)  However, Sobolev spaces $H^{k,p}(\Omega)$ are not generally closed under the maximum operation for $k\ge 2$.
+
+\begin{example}  \label{ex:frequencydecomposition}  For simplicity suppose $\Omega = (0,a)^d \subset \RR^d$ for $a>0$, a cube, and let $\cV = H_{\text{per}}^k(\Omega)$, $k\ge 0$, be the periodic functions.
 Suppose $\psi \in H^k(\Omega)$ and let $\cK = \{v \ge \psi\} \subset \cV$.
-Without getting into any detailed notation of frequency representation, but noting that the frequencies are discrete, suppose $\cV_i \subset \cV$ are $m<\infty$ subspaces from an (nonoverlapping) partition by frequency, thus satisfying \eqref{eq:subspacedecomp}.
+Without giving any detailed notation of frequency representation, but noting that the frequencies are discrete, suppose $\cV_i \subset \cV$ are $m<\infty$ subspaces from an (nonoverlapping) partition by frequency, thus satisfying \eqref{eq:subspacedecomp}.
 Suppose $P_i:\cV \to \cV_i$ are the corresponding orthogonal projections, satisfying $I = \sum_i P_i$.
-Let $\cK_i = \{v \ge P_i \psi\} \subset \cV$ and $R_i(v) = \max\{v,P_i \psi\}$.
+Let $\cK_i = \{v \ge P_i \psi\} \subset \cV_i$ and $R_i = P_i$.
 Then \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} hold.
@@
 We now state the basic CD algorithms as iterations which solve VI problems over each set $\cK_i$.
-(By contrast, the algorithms in \cite{Tai2003} involve iterative optimization over the $\cK_i$.)  There is a parallel (additive) form and a successive (multiplicative) form; these generalize the Jacobi and Gauss-Seidel iterations \cite{Greenbaum1997}, respectively.
+(By contrast, the algorithms in \cite{Tai2003} solve optimization problems over the $\cK_i$.)  There is a parallel (additive) form and a successive (multiplicative) form; these generalize the Jacobi and Gauss-Seidel iterations \cite{Greenbaum1997}, respectively.
 Each algorithm computes a new iterate $w\in\cK$, an improved approximation of the solution $u\in \cK$ of VI problem \eqref{eq:vi} relative to the previous iterate $y \in \cK$.

### 2021-07-06 paper/mcd2.tex
Complete a round of edits.

@@
 \begin{document}
-\title[On MCD methods for nonlinear and nonlocal VIs]{On multilevel constraint decomposition methods \\ for nonlinear and nonlocal variational inequalities}
+\title[On multilevel constraint decomposition methods]{On multilevel constraint decomposition methods \\ for nonlinear and nonlocal variational inequalities}
 
@@
 
-\section{Constraint decomposition; the basic algorithms} \label{sec:cd}
+\section{Constraint decomposition, the basic algorithm} \label{sec:cd}
 
@@
 
-We now state the basic CD algorithms as iterations which solve VI problems over each set $\cK_i$.
-(By contrast, the algorithms in \cite{Tai2003} solve optimization problems over the $\cK_i$.)  There is a parallel (additive) form and a successive (multiplicative) form; these generalize the Jacobi and Gauss-Seidel iterations \cite{Greenbaum1997}, respectively.
-Each algorithm computes a new iterate $w\in\cK$, an improved approximation of the solution $u\in \cK$ of VI problem \eqref{eq:vi} relative to the previous iterate $y \in \cK$.
-A damping parameter $0<\alpha\le 1$ is included.
+We now state the basic CD algorithm as an iteration which solves VI problems over each set $\cK_i$.
+(By contrast, algorithms in \cite{Tai2003} solve \emph{optimization} problems.)  There are parallel (additive) and successive (multiplicative) versions; these generalize the Jacobi and Gauss-Seidel iterations \cite{Greenbaum1997}, respectively.
+Each starts from a current iterate $u \in \cK$ and computes a new iterate $w\in\cK$, an improved approximation of the solution $u^* \in \cK$ of VI problem \eqref{eq:vi}, according to a damping parameter $0<\alpha\le 1$.
+If \pr{parallel} $=$ \text{\textbf{true}} then the \text{\textbf{for}} loop can be computed in any order.
 
 \begin{pseudo*}
-\pr{cd-parallel}(\{\cK_i\}, \{R_i:\cK \to \cK_i\}, 0<\alpha\le 1, y\in\cK)\text{:} \\+
-    for $i \in \{0,\dots,m-1\}$: \qquad \qquad \ct{parallel} \\+
-        $\hat w_i\in \cK_i$: \quad $\boxed{\ip{f(y - R_i y + \hat w_i)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\-
-    return $w=(1-\alpha) y + \alpha \sum_i \hat w_i$
+\pr{cd}(0<\alpha\le 1, u\in\cK)\text{:} \\+
+    if \pr{parallel}: \\+
+        for $i \in \{0,\dots,m-1\}$: \\+
+            $\hat w_i\in \cK_i$: \\+
+                 $\boxed{\ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\---
+    else: \\+
+        for $i = 0,\dots,m-1$: \\+
+            $\hat w_i\in \cK_i$: \\+
+                $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\Big)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\-
+            $w_i = (1-\alpha) R_i u + \alpha \hat w_i$ \\--
+    $\hat w = \sum_i \hat w_i$ \\
+    return $w=(1-\alpha) u + \alpha \hat w$
 \end{pseudo*}
-
-\begin{pseudo*}
-\pr{cd-successive}(\{\cK_i\}, \{R_i:\cK \to \cK_i\}, 0<\alpha\le 1, y\in\cK)\text{:} \\+
-    for $i = 0,\dots,m-1$: \qquad \qquad \ct{in order} \\+
-        $\hat w_i\in \cK_i$: \quad $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j y\Big)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\
-        $w_i = (1-\alpha) R_i y + \alpha \hat w_i$ \\-
-    return $w = (1-\alpha) y + \alpha \sum_i \hat w_i$
-\end{pseudo*}
-
-Let $e_i = \hat w_i - R_i y \in \cV_i$ be the $i$th-subspace change (in either algorithm), and note we may then write $w = y + \alpha \sum_i e_i$ when needed.
-Also note for \pr{cd-parallel} that $y - R_i y + \hat w_i = y+e_i$.
 
@@
 \begin{align*}
-[y - R_i y + v_i] - [y - R_i y + \hat w_i] &= v_i - \hat w_i, \label{eq:admissibledifference} \\
-\left[\sum_{j<i} w_j + v_i + \sum_{j>i} R_j y\right] - \left[\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j y\right] &= v_i - \hat w_i,  \notag
+[u - R_i u + v_i] - [u - R_i u + \hat w_i] &= v_i - \hat w_i, \label{eq:admissibledifference} \\
+\left[\sum_{j<i} w_j + v_i + \sum_{j>i} R_j u\right] - \left[\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\right] &= v_i - \hat w_i,  \notag
 \end{align*}
-for the two versions, respectively.  In other words, these algorithms require the solution of VI \eqref{eq:vi} over (parallel or successive) perturbations from the sets $\cK_i$.
+for the two versions, respectively.  In other words, we solve VI \eqref{eq:vi} over (parallel or successive) perturbations from the sets $\cK_i$.  In fact, if
+\begin{equation}
+e_i = \hat w_i - R_i u \in \cV_i \label{eq:ithupdate}
+\end{equation}
+is the update in $\cK_i$ then we may then write $w = u + \alpha \sum_i e_i$.
+In particular, note that $u - R_i u + \hat w_i = u + e_i$; the \pr{parallel} version computes the updates $e_i$ independently.
 
-The convergence results in the next section require substantial damping in the parallel algorithm, i.e.~$\alpha \le 1/m$, but the successive version can be shown to converge without damping ($\alpha=1$).
+The convergence results in the next section require substantial damping in the parallel algorithm, i.e.~$\alpha \le 1/m$, as in \cite{Tai2003} as well, but the successive version can be shown to converge without damping ($\alpha=1$).
 In Section \ref{sec:resultslocal} we demonstrate practical convergence for a larger range of $\alpha$ than suggested by the theory.
@@
 
-In Section \ref{sec:multilevel} we will implement these basic algorithms for multilevel CDs.
-However, this implementation requires additional concepts in the case that $f$ is either: \emph{(i)} defined only on $\cK$, \emph{(ii)} nonlinear, or \emph{(iii)} nonlocal.
-(In fact we will extend the above methods by applying the full approximation storage idea of Brandt \cite{Brandt1977}; see Section \ref{sec:multilevel}.)  The issue seems not to have been addressed by references which apply CD algorithms only to the classical obstacle problem \cite[in particular]{GraeserKornhuber2009,Tai2003}.
+In Section \ref{sec:multilevel} we will convert the basic algorithm into a multilevel approach for an obstacle problem by choosing appropriate $\cK_i$ based on a FE triangulation.
+However, in order that the algorithm has a practical and efficient implementation, we will require additional concepts when $f$ has any of the following properties: \emph{(i)} $f$ is defined only on $\cK$, \emph{(ii)} $f$ is nonlinear, or \emph{(iii)} $f$ is nonlocal.
+The issue of practical implementation for such $f$ seems not to have been addressed by references which apply CD algorithms only to the classical obstacle problem \cite[in particular]{GraeserKornhuber2009,Tai2003}.
+Consider the following example.
 
-To explain our concern, suppose $f:\cV \to \cV'$ is linear and local (in a sense which is explained momentarily).
-Only consider \pr{cd-parallel}, for simplicity, and suppose for concreteness that $\cV = H_0^1(\Omega)$ (for some domain $\Omega \subset \RR^d$).
-Note that for $z_i \in \cV_i$ we have $\ip{f(y - R_i y + \hat w_i)}{z_i} = \ip{f(y)}{z_i} + \ip{f(\hat w_i - R_i y)}{z_i}$, so the VI over $\cK_i$ can be written as
+\begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear and defined on all of $\cV$.
+Furthermore suppose $f$ is local in the sense that a basis $\{\phi_i\}$ of $\cV_i$ exists with each support small in $\Omega$, and that $\ip{f(\phi_i)}{z}$ can be computed by an integral over the support of $\phi_i$.
+(When $\cV$ is an FE space then this is the usual case for evaluating PDE weak forms over a basis of hat functions; see Section \ref{sec:multilevel}.)  Considering only the parallel algorithm, for simplicity, the VI over $\cK_i$ can be written as
 \begin{equation}
-\ip{f(\hat w_i - R_i y)}{v_i-\hat w_i} \ge \tilde g(v_i-\hat w_i)  \qquad\qquad \text{\emph{($f$ linear)}} \label{eq:linearlocalcdparallel}
+\ip{f(\hat w_i - R_i u)}{v_i-\hat w_i} \ge \tilde g(v_i-\hat w_i) \label{eq:linearlocalvi}
 \end{equation}
-for all $v_i\in \cK_i$, where $\tilde g(z) = g(z) - \ip{f(y)}{z}$.
-Note $e_i = \hat w_i - R_i y \in \cV_i$ but generally $e_i \notin \cK$, so \eqref{eq:linearlocalcdparallel} only make sense if $f$ is defined over the whole vector space.
-Now each VI problem in \pr{cd-parallel} can then be solved using a stored residual $\ip{f(y)}{\cdot}$ which is put in a fixed source term.
-Furthermore, by $f$ being ``local'' we mean that if a basis $\{\phi_i\}$ of $\cV_i$ can be constructed with each support small in $\Omega$, then $\ip{f(\phi_i)}{z}$ can be computed by an integral over the support of $\phi_i$.
-Then the solution of each VI problem \label{eq:linearlocalcdparallel} can be computed by local means over the union of the supports of the basis.
-Thus the resulting solver for the everywhere-defined, linear, and local $f$ has implementation efficiencies that are unavailable otherwise.
-In conclusion, additional ideas, like those in the algorithms in Section \ref{sec:multilevel}, are needed for practical application of CD methods to problems which are not close to the classical obstacle problem.
+for all $v_i\in \cK_i$, where
+\begin{equation}
+\tilde g(z) = g(z) - \ip{f(u)}{z}. \label{eq:linearlocalrhs}
+\end{equation}
+Noting $e_i = \hat w_i - R_i u \in \cV_i \notin \cK$, in general, VI \eqref{eq:linearlocalvi} only makes sense because $f$ is defined over $\cV_i$.
+Thus each subset problem \eqref{eq:linearlocalvi} can be solved using a stored residual $\ip{f(y)}{\cdot}$ in an updated source term.
+In other words, because $f$ has all these nice properties, the solution of each VI problem can be computed by local, thus inexpensive, means.
+\end{example}
+
+A solver which would work for Example \ref{ex:fnice} has implementation efficiencies that are unavailable in general.
+Additional ideas, like those in Section \ref{sec:multilevel}, are needed for practical application of CD methods to problems which are not close to the classical obstacle problem.
+In particular, we will extend the basic algorithm to nonlinear $f:\cK\to \cV'$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}, and in Section \ref{sec:resultsnonlocal} we will apply the algorithm in a case where residual evaluation is substantially less simple than evaluation of a PDE weak form.
 

### 2021-07-06 paper/mcd2.tex
Get a start on section 4.

@@
 \newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{assumptions}[theorem]{Assumptions}
 
@@
 
-FIXME
+To prove the convergence of the basic algorithm we make two assumptions regarding the CD and the residual functional, as follows.
+
+\begin{assumptions}  \emph{Lipschitz restrictions}: There exists $C_1>0$ so that
+\begin{equation}
+\left(\sum_{i=0}^{m-1} \|R_i u - R_i v\|^2\right)^{1/2} \le C_1 \|u-v\| \label{as:lipschitzrestrictions}
+\end{equation}
+for all $u,v\in\cK$.  \emph{Totally Lipschitz residual}: There exists $C_2>0$ so that
+\begin{equation}
+FIXME \label{as:totallylipschitzresidual}
+\end{equation}
+for all 
+\end{assumptions}
+
+Though the assumptions are technical in certain senses, the reader should note some essential features:
+\begin{itemize}
+\item The assumptions are identical to inequalities (7), (8) in \cite{Tai2003} where they are used them to prove the convergence of the optimization version of the basic CD algorithm.
+See also their unconstrained versions, inequalities (13), (14) in \cite{TaiXu2002}.
+\item Assumption \eqref{as:lipschitzrestrictions} addresses only the constraint decomposition while, essentially, \eqref{as:totallylipschitzresidual} addresses only a subspace decomposition property of the residual functional.
+For example, if $f$ is defined by a PDE weak form, as in the classical obstacle problem, then \eqref{as:totallylipschitzresidual} is verified over the subpaces $\cV_i$, and the subsets $\cK_i$ are not relevant.
+\item Tai \cite{Tai2003} gives values of the constants $C_1$ in \eqref{as:lipschitzrestrictions} for obstacle problem constraint decompositions using $P_1$ FE spaces over a shape-regular and quasi-uniform triangulation.
+In particular, constants $C_1$ are know for domain decompositions (with or without an additional coarse mesh) and standard multilevel heirarchies.
+% ASSUMING $f$ IS LIPSCHITZ WILL MAKE THE EXISTENCE OF C_2 CLEAR; SEE (14) IN \cite{TaiXu2002}
+\end{itemize}
 

### 2021-07-07 paper/mcd2.tex
Bits of thinking.

@@
 
-\begin{assumptions}  \emph{Lipschitz restrictions}: There exists $C_1>0$ so that
+\begin{assumptions}  \emph{Lipschitz restrictions}: There exists a constant $C_1>0$ so that
 \begin{equation}
@@
 \end{equation}
-for all 
+for all FIXME
 \end{assumptions}
+Note that the constants $C_1,C_2$ generally depend on $m$.
+% IF PROOF BELOW SUCCEEDS, SO ALGORITHM IS NORM DECREASING, THEN MAY BE ABLE TO ARGUE THAT IT IS O.K.
+FOR CONSTANTS TO DEPEND ON $u,v,w_{ij}$ etc.
+AS LONG AS THEY ARE UNIFORMLY BOUNDED ON CERTAIN BOUNDED SETS
 

### 2021-07-07 paper/mcd2.tex
Actually applying the CD method when the residual is non-local is a bridge too far.

@@
 \begin{document}
-\title[On multilevel constraint decomposition methods]{On multilevel constraint decomposition methods \\ for nonlinear and nonlocal variational inequalities}
+\title[On multilevel constraint decomposition methods]{On multilevel constraint decomposition methods \\ for nonlinear variational inequalities}
 
@@
 \begin{enumerate}
-\item We do not assume that the continuum VI problem arises from optimization of a scalar objective, an assumption in \cite{Tai2003}.
+\item In contrast to \cite{Tai2003}, we do not assume that the continuum VI problem arises from optimization of a scalar objective.
 Nonetheless we can prove convergence at the same rate (Section \ref{sec:convergence}); % HOPE
-examples are explored in Sections \ref{sec:vi}, \ref{sec:resultslocal}, and \ref{sec:resultsnonlocal}.
-\item We observe that the iterates from these CD methods are always admissible, and thus the residual functional need only be defined on the admissible subset; Sections \ref{sec:resultslocal} and \ref{sec:resultsnonlocal} include examples.
-Admissible-iterate methods allow direct solutions of certain VI problems, such as fluid-layer dynamics problems (Section \ref{sec:resultsnonlocal}; see also \cite{Bueler2021conservation,JouvetBueler2012}) for which non-admissible methods, such as semi-smooth methods \cite{BensonMunson2006}, may require unnatural modifications of the residual form.
+examples are explored in Sections \ref{sec:vi} and \ref{sec:results}.
+\item We observe that the iterates from the CD method are always admissible, and thus the residual functional need only be defined on the admissible subset; Section \ref{sec:results} includes an example.
+Admissible-iterate methods allow direct solutions of certain VI problems, such as fluid-layer dynamics problems \cite{Bueler2021conservation,JouvetBueler2012} for which non-admissible methods, such as semi-smooth methods \cite{BensonMunson2006}, may require unnatural modifications of the residual form.
 \item In the multilevel case we make the observation that, if implemented appropriately, V(0,$\nu$) ``up-slash'' cycles are intrinsically more efficient than the V($\nu$,0) ``down-slash'' cycles (Section \ref{sec:multilevel}).
 This observation seems to be new for multilevel CD methods; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}.
-Strong preference for up-smoothing is, apparently, special to multilevel CD methods and does not arise in corresponding unconstrained (e.g.~\cite{Xu1992}) problems.
-\item In the multilevel and nonlinear case, we make the algorithms practical by addressing efficient storage of intermediate quantities.
-In particular we show results from a full approximation storage (FAS; \cite{Brandt1977}) version of the multilevel CD algorithm (Section \ref{sec:resultslocal}), avoiding global (Newton) linearization; compare \cite{GraeserKornhuber2009}.
+Strong preference for up-smoothing is, apparently, special to multilevel CD methods and does not arise in corresponding unconstrained problems.
+\item In the multilevel and nonlinear case, we make the CD algorithm practical by addressing efficient storage of intermediate quantities.
+In particular, in Section \ref{sec:results} we show results from a full approximation storage (FAS; see \cite{Brandt1977}) implementation.
+Thereby we avoid global (Newton) linearization when solving nonlinear VI problems; compare \cite{GraeserKornhuber2009}.
 \end{enumerate}
@@
 These improved multilevel methods either track the active set or modify the nodal basis functions, and in this sense they are discrete algorithms.
-By contrast the CD methods apply at the level of the continuum problem (see below); CD methods are more fundamental than the truncated monotone multigrid method \cite{Kornhuber1994}, for example, in that sense, and we can demonstrate comparable rates of convergence (Section \ref{sec:resultslocal}).
+By contrast the CD method applies at the level of the continuum problem (see below); it is more fundamental than the truncated monotone multigrid method \cite{Kornhuber1994}, for example, in that sense, and we can demonstrate comparable rates of convergence (Section \ref{sec:results}).
 % HOPE
 Furthermore, truncated monotone approaches for nonlinear problems have, so far, applied a global linearization strategy \cite{GraeserKornhuber2009}, while we apply FAS (Section \ref{sec:multilevel}).
-In any case, acceleration of our nonlinear multilevel CD methods by such active-set and/or basis-level manipulations is a topic for further research.
+In any case, acceleration of our nonlinear multilevel CD method by such active-set and/or basis-level manipulations is a topic for further research.
 
-In one example at the end of this paper (Section \ref{sec:resultsnonlocal}) we consider a nonlocal residual functional, that is, one which is not a partial differential operator.
+% A BRIDGE TOO FAR:  In one example at the end of this paper (Section \ref{sec:resultsnonlocal}) we consider a nonlocal residual functional, that is, one which is not a partial differential operator.
 Each evaluation of this functional requires the solution of a Stokes problem for a layer of fluid \nocite{SayagWorster2013} on a substrate (which forms the obstacle), and thus the corresponding FE operator discretization is also not sparse.
 In this case we cannot prove coercivity but we nonetheless succeed in demonstrating near-optimal complexity of the Section \ref{sec:multilevel} multilevel CD algorithm in practice.
-% HOPE
 
@@
 
-The VIs solved in this paper satisfy a stronger form of coercivity than \eqref{eq:coercive}.\footnote{An exception is the problem in Section \ref{sec:resultsnonlocal}; there $f$ is not known to be monotone or coercive.}
+The VIs solved in this paper satisfy a stronger form of coercivity than \eqref{eq:coercive}.
 
@@
 Further examples are in ice sheet models \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids \cite{Bueler2021conservation}.
-Numerical solver performance for Examples \ref{ex:advectiondiffusion}--\ref{ex:plaplacian} will be considered in section \ref{sec:resultslocal}.
+Numerical solver performance for Examples \ref{ex:advectiondiffusion}--\ref{ex:plaplacian} will be considered in section \ref{sec:results}.
 % HOPE
@@
 The convergence results in the next section require substantial damping in the parallel algorithm, i.e.~$\alpha \le 1/m$, as in \cite{Tai2003} as well, but the successive version can be shown to converge without damping ($\alpha=1$).
-In Section \ref{sec:resultslocal} we demonstrate practical convergence for a larger range of $\alpha$ than suggested by the theory.
+In Section \ref{sec:results} we demonstrate practical convergence for a larger range of $\alpha$ than suggested by the theory.
 % HOPE
@@
 In Section \ref{sec:multilevel} we will convert the basic algorithm into a multilevel approach for an obstacle problem by choosing appropriate $\cK_i$ based on a FE triangulation.
-However, in order that the algorithm has a practical and efficient implementation, we will require additional concepts when $f$ has any of the following properties: \emph{(i)} $f$ is defined only on $\cK$, \emph{(ii)} $f$ is nonlinear, or \emph{(iii)} $f$ is nonlocal.
+However, in order that the algorithm has a practical and efficient implementation, we will require additional concepts when $f$ has any of the following properties: \emph{(i)} $f$ is defined only on $\cK$, or \emph{(ii)} $f$ is nonlinear, or \emph{(iii)} $f$ is non-local.
 The issue of practical implementation for such $f$ seems not to have been addressed by references which apply CD algorithms only to the classical obstacle problem \cite[in particular]{GraeserKornhuber2009,Tai2003}.
@@
 \end{equation}
-for all $v_i\in \cK_i$, where
-\begin{equation}
-\tilde g(z) = g(z) - \ip{f(u)}{z}. \label{eq:linearlocalrhs}
-\end{equation}
+for all $v_i\in \cK_i$, where $\tilde g(z) = g(z) - \ip{f(u)}{z}$.
 Noting $e_i = \hat w_i - R_i u \in \cV_i \notin \cK$, in general, VI \eqref{eq:linearlocalvi} only makes sense because $f$ is defined over $\cV_i$.
-Thus each subset problem \eqref{eq:linearlocalvi} can be solved using a stored residual $\ip{f(y)}{\cdot}$ in an updated source term.
-In other words, because $f$ has all these nice properties, the solution of each VI problem can be computed by local, thus inexpensive, means.
+Thus each subset problem \eqref{eq:linearlocalvi} can be solved using a stored residual $\ip{f(y)}{\cdot}$ (in the updated source term) and incrementally by computations over the basis supports.
+In other words, because $f$ has all these nice properties, the solution of each VI problem can be computed by inexpensive means.
 \end{example}
@@
 Additional ideas, like those in Section \ref{sec:multilevel}, are needed for practical application of CD methods to problems which are not close to the classical obstacle problem.
-In particular, we will extend the basic algorithm to nonlinear $f:\cK\to \cV'$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}, and in Section \ref{sec:resultsnonlocal} we will apply the algorithm in a case where residual evaluation is substantially less simple than evaluation of a PDE weak form.
+In particular, we will extend the basic algorithm to nonlinear $f:\cK\to \cV'$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
+However, extending the algorithm to non-local residual functionals is a topic for future research.
 
@@
 
-\section{Results for local variational inequalities} \label{sec:resultslocal}
+\section{Results for local variational inequalities} \label{sec:results}
 
@@
 
-\section{Results for a nonlocal variational inequality} \label{sec:resultsnonlocal}
-
-FIXME
+% A BRIDGE TOO FAR:  \section{Results for a nonlocal variational inequality} \label{sec:resultsnonlocal}
 

### 2021-07-07 paper/mcd2.tex
State the C_2 assumption.

@@
 \newtheorem{assumptions}[theorem]{Assumptions}
+
+\newtheoremstyle{cstyle*}% name
+  {5pt}% space above
+  {5pt}% space below
+  {\itshape}% body font
+  {}% indent amount
+  {\itshape}% theorem head font
+  {.}% punctuation after theorem head
+  {.5em}% space after theorem head
+  {\thmname{#1}}% theorem head spec
+\theoremstyle{cstyle*}
+\newtheorem{assumptions*}{Assumptions}
 
@@
 
-To prove the convergence of the basic algorithm we make two assumptions regarding the CD and the residual functional, as follows.
+To prove the convergence of the basic algorithm we make two assumptions as follows.
 
-\begin{assumptions}  \emph{Lipschitz restrictions}: There exists a constant $C_1>0$ so that
+\begin{assumptions*}  \emph{Lipschitz restrictions}: There exists a constant $C_1>0$ so that
 \begin{equation}
@@
 \begin{equation}
-FIXME \label{as:totallylipschitzresidual}
+\sum_{i=0}^{m-1} \sum_{j=0}^{m-1} \left|\ip{f(w_{ij} + v_i) - f(w_{ij})}{y_i}\right| \le C_2 \left(\sum_{i=0}^{m-1} \|v_i\|^2\right)^{1/2} \left(\sum_{i=0}^{m-1} \|y_i\|^2\right)^{1/2} \label{as:totallylipschitzresidual}
 \end{equation}
-for all FIXME
-\end{assumptions}
-Note that the constants $C_1,C_2$ generally depend on $m$.
+for all $w_{ij} \in \cK$ and $v_i,y_i \in \cV_i$ such that $w_{ij} + v_i \in \cK$ (for all $i,j$).  Note that the constants $C_1,C_2$ generally depend on $m$.
+\end{assumptions*}
 % IF PROOF BELOW SUCCEEDS, SO ALGORITHM IS NORM DECREASING, THEN MAY BE ABLE TO ARGUE THAT IT IS O.K.
@@
 \begin{itemize}
-\item The assumptions are identical to inequalities (7), (8) in \cite{Tai2003} where they are used them to prove the convergence of the optimization version of the basic CD algorithm.
+\item Assumption \eqref{as:lipschitzrestrictions} addresses only the constraint decomposition.
+Furthermore, it is often the case that \eqref{as:totallylipschitzresidual} addresses only a subspace decomposition property of the residual functional.
+(For example, in the classical obstacle problem \eqref{as:totallylipschitzresidual} is verified over the (sub)spaces $\cV$ and $\cV_i$, and the subsets $\cK$ and $\cK_i$ are not involved.)
+\item The assumptions are identical to inequalities (7), (8) in \cite{Tai2003} where they are used to prove convergence of the optimization version(s) of the basic CD algorithm.
 See also their unconstrained versions, inequalities (13), (14) in \cite{TaiXu2002}.
-\item Assumption \eqref{as:lipschitzrestrictions} addresses only the constraint decomposition while, essentially, \eqref{as:totallylipschitzresidual} addresses only a subspace decomposition property of the residual functional.
-For example, if $f$ is defined by a PDE weak form, as in the classical obstacle problem, then \eqref{as:totallylipschitzresidual} is verified over the subpaces $\cV_i$, and the subsets $\cK_i$ are not relevant.
 \item Tai \cite{Tai2003} gives values of the constants $C_1$ in \eqref{as:lipschitzrestrictions} for obstacle problem constraint decompositions using $P_1$ FE spaces over a shape-regular and quasi-uniform triangulation.
-In particular, constants $C_1$ are know for domain decompositions (with or without an additional coarse mesh) and standard multilevel heirarchies.
+In particular, constants $C_1$ are know for overlapping domain decompositions (with or without an additional coarse mesh) and standard multilevel heirarchies.
 % ASSUMING $f$ IS LIPSCHITZ WILL MAKE THE EXISTENCE OF C_2 CLEAR; SEE (14) IN \cite{TaiXu2002}
 \end{itemize}
+
+FIXME
 

### 2021-07-07 paper/mcd2.tex
Write up of the main coercive f examples.

@@
 When $f$ is monotone, VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
-In fact, suppose $F:\cV \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cV \to \cV'$.
-Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}, and if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ and $g=0$ if and only if
+In fact, suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cK \to \cV'$.
+Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
+Furthermore, Proposition II.2.1 in the same reference shows that if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ (and $g=0$) if and only if
 \begin{equation}
-u^* = \operatorname{arg-min}_{v\in\cK} F(v) \label{eq:minimization}
+u^* = \operatorname{arg-min}_{v\in\cK} F(v). \label{eq:minimization}
 \end{equation}
-\cite[Proposition II.2.1]{EkelandTemam1976}.
 
-Before giving examples, note that Sobolev spaces are denoted $W^{k,p}(\Omega)$ \cite{Evans2010}, with $W^{k,2}=H^k$ in particular.
+From now on $\Omega \subset \RR^d$ denotes a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.
+Note that Sobolev spaces \cite{Evans2010} are denoted here by $W^{k,p}(\Omega)$, with Hilbert-space cases denoted $W^{k,2}=H^k$.
 
-\begin{example}  \label{ex:plaplacian}  FIXME
+Many numerical solvers for VIs, including the CD methods of Tai \cite{Tai2003}, address optimization problem \eqref{eq:minimization} under the hypothesis that $F'$ is coercive.
+The following example includes the classical obstacle problem \cite{GraeserKornhuber2009} and the obstacle problem for the $p$-Laplacian \cite{ChoeLewis1991}.
+
+\begin{example}  \label{ex:plaplacian}  Suppose $a\in L^\infty(\Omega)$ such that $a(x)\ge a_0$ a.e.~for some constant $a_0>0$, and $p\ge 2$.
+For $u,v \in \cV = W^{1,p}_0(\Omega)$ define $f:\cV \to \cV'$ by
 \begin{equation}
-\ip{f(u)}{v} = \int_\Omega |\grad u|^{p-2} \grad u \cdot \grad v\,dx \label{eq:plaplacian}
+\ip{f(u)}{v} = \int_\Omega a(x) |\grad u|^{p-2} \grad u \cdot \grad v\,dx. \label{eq:plaplacian}
 \end{equation}
+Now, if $x,y\in\RR^d$ then $(|x|^{p-2} x - |y|^{p-2} y)\cdot (x-y) \ge 2^{2-p} |x-y|^p$ \cite[see Appendix A and references therein]{Bueler2021conservation}.
+It follows from the Poincar\'e inequality that
+    $$\ip{f(u) - f(v)}{u-v} \ge 2^{2-p} a_0 \|\grad u - \grad v\|_p^p \ge 2^{2-p} a_0 C \|u-v\|^p$$
+for some $C>0$, and thus $f$ is $p$-coercive.  (The same $f$ is coercive if $1<p<2$ but the proof is somewhat different \cite[Theorem 4.4]{Bueler2021conservation}.)  Also, for $g\in\cV'$ define
+    $$F(u) = \int_\Omega \frac{a(x)}{p} |\grad u|^p\,dx - g(u).$$
+Then $F'(u) = f(u) - g$, $F$ is a convex functional (since $f$ is coercive), and, for any closed and convex $\cK\subset \cV$, VI problem \eqref{eq:vi} for $f$ and $g$ is equivalent to optimization problem \eqref{eq:minimization} for $F$.
 \end{example}
 
-Some numerical solvers for VIs, including the CD methods of Tai \cite{Tai2003}, address optimization problem \eqref{eq:minimization}.
-However, not all VIs correspond to optimization, and we give two such examples next.
+However, not all VIs correspond to optimization.
+We give two such examples next, a linear advection-diffusion problem and then a nonlinear porous-medium-type problem, both of which are important in applications.
+Each example is preceded by a lemma.
 
-\begin{example}  \label{ex:advectiondiffusion}  Consider a steady advection-diffusion problem over a bounded domain $\Omega \subset \RR^d$, with polygonal boundary, and suppose the boundary is partitioned $\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$ with $\partial_D\Omega$ of positive measure.
-Let $\cV = H_0^1(\Omega)$ be the space of functions with value zero on $\partial_D\Omega$; for simplicity suppose zero Dirichlet and Neumann data.
-Consider a velocity field $\bX \in L^\infty(\Omega)^d$ such that $\Div \bX=0$, and define the functional
+\begin{lemma}  \label{lem:advectionskew}  \cite{Elmanetal2014}\,  Suppose $\bX :\Omega \to \RR^d$ is a smooth and bounded vector field on $\Omega$ with zero divergence ($\Div \bX=0$).
+For $u,v \in H^1(\Omega)$ let $s(u,v) = \int_\Omega (\bX \cdot \grad u) v\,dx$.
+Then $s(v,v) = \frac{1}{2} \int_{\partial \Omega} v^2 \bX\cdot \bn\,dx$ where $\bn$ is the outward normal on $\partial \Omega$.
+\end{lemma}
+
+\begin{proof}
+Integration by parts gives $s(u,v) = - s(v,u) + \int_{\partial \Omega} uv \bX\cdot \bn\,dx$, so the result follows.
+\end{proof}
+
+\begin{example}  \label{ex:advectiondiffusion}  Suppose $\partial\Omega$ is partitioned into Dirichlet and Neumann portions, i.e.~$\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$, with $\partial_D\Omega$ of positive measure.
+Let $\cV = H_0^1(\Omega)$ be the space of functions with value zero on $\partial_D\Omega$.
+(For simplicity we consider zero boundary data.)  Choose a smooth and bounded velocity field $\bX$ on $\Omega$ such that $\Div \bX=0$, but additionally assume outward flow on the Neumann boundary, i.e.~$\bX \cdot \bn \ge 0$ at points of $\partial_N\Omega$.
+For $u,v \in \cV = H_0^1(\Omega)$ and $\eps>0$ define
 \begin{equation}
-\ip{f(u)}{v} = \int_\Omega \eps \grad u \cdot \grad v - (\bX \cdot \grad u) v\,dx \label{eq:advectiondiffusion}
+\ip{f(u)}{v} = \int_\Omega \eps \grad u \cdot \grad v - (\bX \cdot \grad u) v\,dx. \label{eq:advectiondiffusion}
 \end{equation}
-for $\eps>0$.
-Finally, suppose a nonnegativity constraint $\cK = \{v\ge 0\} \subset \cV$ on the solution \cite[for example]{ChangNakshatrala2017}, and consider VI \eqref{eq:vi} for any $g\in\cV'$.
-It is easy to see that $|\ip{f(u)}{v}| \le (\eps + \|\bX\|_\infty) \|u\|_{\cV} \|v\|_{\cV}$, thus that $f:\cK \to \cV'$ is continuous.
-FIXME  MAKE ARGUMENT FOLLOWING \cite{Elmanetal2014}; see also \cite{Kirby2010}; IF $\bX=0$ THEN IT IS OPTIMIZATION BUT OTHERWISE NOT; $2$-coercive if outflow on Neumann
+Consider VI \eqref{eq:vi} for any closed and convex $\cK \subset \cV$ and any $g\in\cV'$.
+It is easy to see that $|\ip{f(u)}{v}| \le (\eps + \|\bX\|_\infty) \|u\| \|v\|$, thus that $f:\cK \to \cV'$ is continuous.
+On the other hand, by Lemma \ref{lem:advectionskew}, which says that $s(u,v)$ is skew-symmetric up to a nonnegative addition, and by the outward flow assumption and the Poincar\'e inequality,
+\begin{align*}
+\ip{f(u)-f(v)}{u-v} &= \eps \int_\Omega |\grad u - \grad v|^2\,dx + s(u-v,u-v) \\
+                    &= \eps \int_\Omega |\grad u - \grad v|^2\,dx + \frac{1}{2} \int_{\partial_N\Omega} (u-v)^2 \bX\cdot\bn \ge \eps C \|u-v\|^2.
+\end{align*}
+Thus $f$ is 2-coercive, and so VI problem \eqref{eq:vi} is well-posed for any closed and convex $\cK \subset \cV$ and $g\in \cV'$.
+
+IF $\bX=0$ THEN IT IS OPTIMIZATION BUT OTHERWISE NOT
 \end{example}
+
+In the above example we have assumed zero boundary data for simplicity.  Reference \cite{ChangNakshatrala2017} considers the above problem subject to a nonnegativity constraint ($\cK = \{v\ge 0\}$).
 

### 2021-07-08 paper/mcd2.tex
Stuck on coercivity of porous media form.

@@
 However, not all VIs correspond to optimization.
-We give two such examples next, a linear advection-diffusion problem and then a nonlinear porous-medium-type problem, both of which are important in applications.
-Each example is preceded by a lemma.
+We give two such examples next, first a linear advection-diffusion problem and then a nonlinear porous-medium-type problem.
+Each of these examples, which are important in applications, is preceded by a lemma.
 
@@
 Thus $f$ is 2-coercive, and so VI problem \eqref{eq:vi} is well-posed for any closed and convex $\cK \subset \cV$ and $g\in \cV'$.
-
-IF $\bX=0$ THEN IT IS OPTIMIZATION BUT OTHERWISE NOT
 \end{example}
 
-In the above example we have assumed zero boundary data for simplicity.  Reference \cite{ChangNakshatrala2017} considers the above problem subject to a nonnegativity constraint ($\cK = \{v\ge 0\}$).
+However, if $\bX \ne 0$ then \eqref{eq:vi} for the functional \eqref{eq:advectiondiffusion} is not a (constrained) minimization problem.
+Indeed, the Jacobian of $f$ is the bilinear form $J_f(u,v) = \int_\Omega \eps \grad u \cdot \grad v\,dx - s(u,v)$.
+For nonzero advection this is not symmetric, thus it cannot be the Hessian of an objective $F$.
+(For $u,v$ which are zero on all of $\partial \Omega$, $J_f(u,v) - J_f(v,u) = -2 s(u,v)$, and we may then construct $u,v$ so this result is nonzero anywhere $\bX$ is nonzero.)
 
-\begin{example}  \label{ex:porousgeneralization}  FIXME
+References \cite{Bueler2021conservation,ChangNakshatrala2017} consider such advection-diffusion VI problems over $\cK = \{v\ge 0\}$.
+
+\begin{example}  \label{ex:porousgeneralization}  FIXME. I THINK THIS IS NOT GOING TO WORK.  COERCIVITY IS SUPER FRAGILE W.R.T. THIS CHANGE?
 \begin{equation}

### 2021-07-09 paper/mcd2.tex
Put ideas on non-monotonicity in appendix.

@@
 
-\begin{example}  \label{ex:porousgeneralization}  FIXME. I THINK THIS IS NOT GOING TO WORK.  COERCIVITY IS SUPER FRAGILE W.R.T. THIS CHANGE?
+\begin{example}  \label{ex:porous}  FIXME. I THINK THIS IS NOT GOING TO WORK.  COERCIVITY IS SUPER FRAGILE W.R.T. THIS CHANGE?
 \begin{equation}
-\ip{f(u)}{v} = \int_\Omega \phi(u) \grad u \cdot \grad v\,dx \label{eq:porousgeneralization}
+\ip{f(u)}{v} = \int_\Omega \phi(u) \grad u \cdot \grad v\,dx \label{eq:porous}
 \end{equation}
@@
 
-Examples \ref{ex:advectiondiffusion} and \ref{ex:porousgeneralization} are VI problems which do not arise from optimization.
+Examples \ref{ex:advectiondiffusion} and \ref{ex:porous} are VI problems which do not arise from optimization.
 Further examples are in ice sheet models \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids \cite{Bueler2021conservation}.
@@
 
+\normalsize
+\appendix
+
+\section{Porous-medium problems are never monotone}
+
+Based on the examples in Section \ref{sec:vi}, the property of $p$-coercivity would seem to be a mild generalization of uniform ellipticity (including its quasi-linear meaning \cite[Section 8.3]{Evans2010}).
+Indeed, weak-form operators with certain lower-order terms (Example \ref{ex:advectiondiffusion}) or degenerate quasi-linear type (Example \ref{ex:plaplacian}) can be monotone and coercive.
+However, even the weaker property of monotonicity is ``fragile'' in a pointwise sense, as explained next, while ellipticity is more robust.
+In particular, we show that no nontrivial porous-medium type functional is monotone, nor $p$-coercive.
+We have not found any such explanation in the literature, but the observation of non-monotonicity in \cite{RocknerWang2008} is related.
+
+Consider the expression $(Ax - By)\cdot (x-y)$ where $A,B \in \RR^{d\times d}$ are symmetric matrices and $x,y\in \RR^d$.
+This expression is the quadratic form for a certain symmetric block matrix.
+In fact, regarding $x,y$ as column vectors, we observe that
+\begin{equation}
+(Ax - By)\cdot (x-y) = \begin{bmatrix} x^\top & y^\top \end{bmatrix} \begin{bmatrix}
+                        A & -\frac{1}{2}(A+B) \\
+                        -\frac{1}{2}(A+B) & B \end{bmatrix} \begin{bmatrix} x \\ y \end{bmatrix}. \label{eq:blockmatrix}
+\end{equation}
+Let $M_{A,B}$ denote the symmetric $2d \times 2d$ matrix in \eqref{eq:blockmatrix}.
+Clearly the expression is only bounded below if $M_{A,B}$ is nonnegative definite.
+Stated as the following Lemma, the important point here is that exact equality $A=B$ is required for nonnegativity of the expression; knowing $A$ is close to $B$ is not helpful.
+
+\begin{lemma}  Suppose $A,B \in \RR^{d\times d}$ are symmetric matrices which commute.
+Then $(Ax - By)\cdot (x-y)\ge 0$ for all $x,y\in \RR^d$ if and only if $A=B$ is nonnegative definite.
+\end{lemma}
+
+\newcommand{\sbvec}[2]{\left[\begin{smallmatrix} #1 \\ #2 \end{smallmatrix}\right]}
+\newcommand{\sbmat}[4]{\left[\begin{smallmatrix} #1 & #2 \\ #3 & #4 \end{smallmatrix}\right]}
+
+\begin{proof}
+If $A=B$ is nonnegative definite then nonnegativity of the expression follows immediately.
+Conversely, since $A$ is symmetric there exists an orthogonal basis $\{x_i\}$ of eigenvectors, with $Ax_i = \lambda_i x$.
+By commutativity these are also eigenvectors of $B$, with $Bx_i = \mu_i x_i$.
+Let $\big\{\sbvec{x_1}{x_1}$, $\sbvec{x_1}{-x_1}$, $\dots$, $\sbvec{x_d}{x_d}$, $\sbvec{x_d}{-x_d}\big\}$ be an ordered basis of $\RR^{2d}$.
+It is easy to confirm that in this basis $M_{A,B}$ is block diagonal with $2\times 2$ diagonal blocks $\sbmat{0}{-\frac{1}{2}(\lambda_i+\mu_i)}{-\frac{1}{2}(\lambda_i+\mu_i)}{\lambda_i+\mu_i}$.
+Thus the eigenvalues of $M_{A,B}$ come in pairs $\frac{1}{2}(\lambda_i + \mu_i) \pm \sqrt{\frac{1}{2}(\lambda_i^2 + \mu_i^2)}$ for $i=1,\dots,d$.
+By the strict concavity of the square root, $\sqrt{\frac{1}{2}(\lambda_i^2 + \mu_i^2)} > \frac{1}{2}(|\lambda_i| + |\mu_i|)$ if $\lambda_i\ne \mu_i$.
+Thus if any eigenvalues of $A$ differ from those of $B$, i.e.~$\lambda_i\ne \mu_i$ for any $i$, then there exists a negative eigenvalue of $M_{A,B}$.
+\end{proof}
+
+It is not clear if the hypotheses of symmetry and commutativity can be removed.
+However, when the matrices are scalar multiples of the identity then the Lemma asserts that $(ax-by)\cdot(x-y) \ge 0$ for all $x,y\in\RR^d$ if and only if $a=b\ge 0$.
+This scalar case suffices for the following result.
+
+\begin{theorem}
+Suppose $\phi:[0,\infty) \to (0,\infty)$ is continuous.
+For $\Omega \subset \RR^d$ open and nonempty, $\cV = H_0^1(\Omega)$, and $g\in \cV'$, consider the following functional on $\cK = \{v\ge 0\} \subset \cV$:
+\begin{equation}
+\ip{f(u)}{v} = \int_\Omega \phi(u) \grad u\cdot \grad v\,dx - g(v).  \label{eq:porousagain}
+\end{equation}
+If $\phi$ is not constant then $f$ is not monotone.
+\end{theorem}
+
+\begin{proof}
+Suppose $\phi(\alpha)\ne \phi(\beta)$ for $\alpha > 0$ and $\beta \ge 0$.
+(That is, suppose $\phi$ is not constant.)  For $x\in\Omega$, construct $u\in \cK$ smooth so that $u(x)=\alpha$ and $\xi = \grad u(x) \ne 0$.
+Using the Lemma, choose $\eta \in \RR^d$ so that $(\phi(\alpha) \xi - \phi(\beta) \eta) \cdot (\xi - \eta) < 0$.
+FIXME NOT SURE IF THIS IS GOING TO WORK
+\end{proof}
+
+The main idea is that even the slightest variation in the coefficient $\phi(u)$ violates monotonicity.
+A strong and tight ellipticity hypothesis, e.g.~the existence of $c_0>0$ and $\eps>0$ so that $c_0 \le \phi(u) \le (1+\eps)c_0$, will not imply monotonicity.
+Contrast the nontrivial coefficient dependence on $\grad u$ in the $p$-coercive $p$-Laplacian Example \ref{ex:plaplacian} (with $p>2)$.
+
+No nontrivial functional \eqref{eq:porousagain} is monotone, nor is it ever $p$-coercive for any $p>1$.
 \end{document}

### 2021-07-13 paper/mcd2.tex
Use float package to manage pseudocode formatting better.
A cartoon of constraint decomposition.

@@
 
-\usepackage[within=section]{newfloat}
-
 \usepackage{tikz}
@@
 \pseudoset{left-margin=15mm,topsep=5mm,idfont=\texttt}
+
+\usepackage{float}
 
@@
 
-\DeclareFloatingEnvironment[name=Pseudocode]{pcode}
-
 \begin{document}
@@
 %\bigskip
+
+\newfloat{pseudofloat}{t}{xyz}[section]
+\floatname{pseudofloat}{Algorithm}
 
@@
 Let $\cK \subset \cV$ be a nonempty closed and convex subset, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.
-For a continuous \emph{residual functional} $f:\cK \to \cV'$ and \emph{source} $g\in \cV'$ we consider the following \emph{variational inequality} (VI) for the (exact) solution $u^*\in \cK$:
+For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and \emph{source} $g\in \cV'$ we consider the following \emph{variational inequality} (VI) for the (exact) solution $u^*\in \cK$:
 \begin{equation}
@@
 
+\begin{figure}
+\includegraphics[width=0.55\textwidth]{genfigs/cartoon.pdf}
+\caption{A constraint decomposition (CD) in a one-sided obstacle problem with $\mathcal{K}=\{v\ge \psi\}$ looks something like this.}
+\label{fig:cartoon}
+\end{figure}
+
 In the Section \ref{sec:multilevel} we will introduce discretizations and describe practical algorithms, but the CDs in this section can apply at the level of the continuum problem.
@@
 
-We now state the basic CD algorithm as an iteration which solves VI problems over each set $\cK_i$.
+We now state the basic CD algorithm as an iteration which solves VI problems over each set $\cK_i$ (Algorithm \ref{alg:basiccd}).
 (By contrast, algorithms in \cite{Tai2003} solve \emph{optimization} problems.)  There are parallel (additive) and successive (multiplicative) versions; these generalize the Jacobi and Gauss-Seidel iterations \cite{Greenbaum1997}, respectively.
@@
 
+\begin{pseudofloat}
 \begin{pseudo*}
@@
 \end{pseudo*}
+\caption{The basic constraint decomposition (CD) algorithm for VI problem \eqref{eq:vi}.}
+\label{alg:basiccd}
+\end{pseudofloat}
 

### 2021-07-13 paper/genfigs/cartoon.py

@@ -1,0 +1,52 @@
+#!/usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.patches import Arc
+from writeout import writeout
+
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["mathtext.fontset"] = "cm"
+
+fsize = 20.0
+
+def makeaxes(xmin, xmax, ymin, ymax):
+    x = np.linspace(xmin, xmax, 2)
+    y = np.linspace(ymin, ymax, 2)
+    zero = np.zeros(np.shape(x))
+    plt.plot(x, zero, 'k', lw = 1.0)
+    plt.plot(zero, y, 'k', lw = 1.0)
+
+plt.figure(figsize=(8, 6))
+makeaxes(-0.25, 1.25, -0.1, 0.85)
+plt.text(-0.05, -0.07, r'$0$', fontsize=fsize)
+
+psi1 = 0.2
+psi2 = 0.35
+plt.fill([psi1, 1.2, 1.2, psi1, psi1],
+         [psi2, psi2, 0.8, 0.8, psi2],
+         'k', alpha = 0.2)
+plt.text(0.9, 0.6, r'$\mathcal{K}$', fontsize=fsize)
+
+plt.text(1.28, -0.02, r'$\mathcal{V}_1$', fontsize=fsize)
+plt.plot(psi1, 0.0, 'k.', ms = 12.0)
+plt.plot([psi1, 1.2], [0.0, 0.0], 'k', lw = 2.5)
+plt.text(psi1 - 0.07, - 0.085, r'$\psi(x_1)$', fontsize=fsize)
+plt.text(0.7, - 0.09, r'$\mathcal{K}_1$', fontsize=fsize)
+plt.gca().annotate('', xy=(0.7, psi2 - 0.3), xytext=(0.7, psi2 - 0.05),
+                   arrowprops=dict(arrowstyle='->, head_length=0.8, head_width=0.5'))
+plt.text(0.72, 0.15, r'$R_1$', fontsize=fsize)
+
+plt.text(-0.02, 0.89, r'$\mathcal{V}_2$', fontsize=fsize)
+plt.plot(0.0, psi2, 'k.', ms = 12.0)
+plt.plot([0.0, 0.0], [psi2, 0.8], 'k', lw = 2.5)
+plt.text(- 0.2, psi2 - 0.02, r'$\psi(x_2)$', fontsize=fsize)
[…]

### 2021-07-13 paper/mcd2.tex
Edits section 1.

@@
 The goal of this paper is to generalize the constraint decomposition (CD) method of X.-C.~Tai \cite{Tai2003}.
-We extend this solution method to nonlinear variational inequality (VI) problems for which it has not been previously considered, we address important finite element (FE) implementation details, and we demonstrate the resulting performance.
+The convergence of this method, for several types of decompositions, was proven by Tai for coercive variational inequality (VI) problems which arise from minimization of a convex functional over a convex set.
+In its multilevel form the method has been shown to have optimal complexity for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
 
-The convergence of the CD method for several types of decompositions was proven in \cite{Tai2003} in the case of coercive VIs which arise from minimization of a convex functional over a convex set.
-In its multilevel form the CD method has been shown to have optimal complexity for elliptic, linear obstacle problems \cite[subsection 5.4]{Tai2003}; see also Theorem 4.6 and Algorithm 4.7 in \cite{GraeserKornhuber2009}.
-
-We extend the original algorithms, theory, and examples presented in \cite{Tai2003} in four particular directions:
+We extend the CD method to nonlinear VI problems for which it has not been previously considered, including a new proof of convergence which removes the assumption that the problem arises from (constrained) optimization.
+Furthermore we address new finite element (FE) implementation concerns in nonlinear cases, and we demonstrate the resulting performance.
+The original algorithms and theory presented in \cite{Tai2003} are extended in four particular directions:
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
 \begin{enumerate}
-\item In contrast to \cite{Tai2003}, we do not assume that the continuum VI problem arises from optimization of a scalar objective.
-Nonetheless we can prove convergence at the same rate (Section \ref{sec:convergence}); % HOPE
+\item We do not assume that the continuum VI problem arises from optimization of a scalar objective.
+Nonetheless we can prove convergence in norm at the same rate as the original method converges in energy (Section \ref{sec:convergence}); % HOPE
 examples are explored in Sections \ref{sec:vi} and \ref{sec:results}.
-\item We observe that the iterates from the CD method are always admissible, and thus the residual functional need only be defined on the admissible subset; Section \ref{sec:results} includes an example.
-Admissible-iterate methods allow direct solutions of certain VI problems, such as fluid-layer dynamics problems \cite{Bueler2021conservation,JouvetBueler2012} for which non-admissible methods, such as semi-smooth methods \cite{BensonMunson2006}, may require unnatural modifications of the residual form.
-\item In the multilevel case we make the observation that, if implemented appropriately, V(0,$\nu$) ``up-slash'' cycles are intrinsically more efficient than the V($\nu$,0) ``down-slash'' cycles (Section \ref{sec:multilevel}).
-This observation seems to be new for multilevel CD methods; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}.
-Strong preference for up-smoothing is, apparently, special to multilevel CD methods and does not arise in corresponding unconstrained problems.
-\item In the multilevel and nonlinear case, we make the CD algorithm practical by addressing efficient storage of intermediate quantities.
-In particular, in Section \ref{sec:results} we show results from a full approximation storage (FAS; see \cite{Brandt1977}) implementation.
-Thereby we avoid global (Newton) linearization when solving nonlinear VI problems; compare \cite{GraeserKornhuber2009}.
+
+\item We make the observation that, if implemented appropriately, multilevel V(0,$\nu$) ``up-slash'' cycles are intrinsically more efficient than the V($\nu$,0) ``down-slash'' cycles (Section \ref{sec:multilevel}).
+This observation seems to be new and should replace the default use of down-slash or even balanced (e.g.~V(1,1)) cycles; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}.
+A strong preference for up-smoothing is, apparently, special to multilevel CD methods and does not arise in corresponding unconstrained problems.
+
+\item In the nonlinear case, we make the multilevel CD algorithm more practical by addressing efficient storage of intermediate quantities.
+In particular, in Section \ref{sec:results} we show results from a full approximation storage (FAS; see \cite{Brandt1977}) implementation which avoids global (Newton) linearization; compare \cite{GraeserKornhuber2009}.
+
+\item For problems of porous-media type the nonlinear operator $f$ is not coercive (Appendix A), so the existing convergence theory does not apply.
+However, a full-cycle implementation of the multilevel CD method is demonstrated numerically.
+The operator is approximated by a coercive operator, nonlinear in general, for the duration of the V-cycle, by ``freezing' the solution-dependent coefficient.
+The resulting scheme is highly-effective for doubly-nonlinear diffusion operators (Section \ref{sec:results}).
 \end{enumerate}
 
-Especially for the classical obstacle problem, i.e.~for a Laplacian functional, certain other methods are known to improve performance relative to the multilevel CD method \cite{GraeserKornhuber2009}.
-These improved multilevel methods either track the active set or modify the nodal basis functions, and in this sense they are discrete algorithms.
-By contrast the CD method applies at the level of the continuum problem (see below); it is more fundamental than the truncated monotone multigrid method \cite{Kornhuber1994}, for example, in that sense, and we can demonstrate comparable rates of convergence (Section \ref{sec:results}).
+Note that the iterates from the CD method are always admissible, and thus the operator need only be defined on the admissible subset; Section \ref{sec:results} includes an example.
+Admissible-iterate methods should allow direct solutions of certain VI problems, such as fluid-layer dynamics problems \cite{Bueler2021conservation,JouvetBueler2012}, for which non-admissible methods, such as semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
+The full-cycle scheme in \emph{(iv)} above, which applies to non-coercive VIs, is helpful in these problems.
+
+Especially for the classical obstacle problem with a Laplacian operator, certain other multilevel methods are known to improve performance relative to the multilevel CD method \cite{GraeserKornhuber2009}.
+These improved methods either track the active set in the discretization or modify the nodal basis functions, and in this sense they are discrete algorithms.
+By contrast the CD method applies at the level of the continuum problem (Sections \ref{sec:cd} and \ref{sec:convergence}), and thus it is a more fundamental method than the truncated monotone multigrid method \cite{Kornhuber1994}, for example, and in any case we can demonstrate comparable rates of convergence (Section \ref{sec:results}).
 % HOPE
-Furthermore, truncated monotone approaches for nonlinear problems have, so far, applied a global linearization strategy \cite{GraeserKornhuber2009}, while we apply FAS (Section \ref{sec:multilevel}).
-In any case, acceleration of our nonlinear multilevel CD method by such active-set and/or basis-level manipulations is a topic for further research.
+Truncated monotone approaches for nonlinear problems have, so far, applied a global linearization strategy \cite{GraeserKornhuber2009}, while we apply FAS (Section \ref{sec:multilevel}).
+Acceleration of our nonlinear multilevel CD method via active-set and/or basis-level manipulations is a topic for further research.
 

### 2021-07-14 paper/mcd2.tex
Done with section 2 edits.

@@
 Suppose $\cV$ is a real, reflexive Banach space with norm $\|\cdot\|$ and topological dual space $\cV'$.
-Denote the dual pairing of $\phi \in \cV'$ and $v\in\cV$ by $\ip{\phi}{v} = \phi(v)$, and note that $\|\phi\|_{\cV'} = \sup_{\|v\|=1} |\ip{\phi}{v}|$ defines a Banach space norm on $\cV'$.
+Denote the dual pairing of $\phi \in \cV'$ and $v\in\cV$ by $\ip{\phi}{v} = \phi(v)$, and note that $\|\phi\|_{\cV'} = \sup_{\|v\|=1} |\ip{\phi}{v}|$ defines a (Banach space) norm on $\cV'$.
 
@@
 VI \eqref{eq:vi} generalizes the nonlinear system of equations $f(u^*)=g$ from a vector space $\cV$ to the constrained case (over $\cK$).
-Informally, if we conceptualize the dual pairing as an inner product, then \eqref{eq:vi} says that the angle between $f(u^*)-g$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
-That is, \eqref{eq:vi} says that $f(u^*)-g$ points directly into $\cK$ if $u^* \in\partial\cK$, while if $u^* \in \cK^\circ$ (interior) then \eqref{eq:vi} implies $f(u^*)=g$.
+Informally, if we conceptualize the dual pairing as an inner product then \eqref{eq:vi} says that the angle between $f(u^*)-g$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
+That is, \eqref{eq:vi} says that $f(u^*)-g$ points directly into $\cK$ if $u^* \in\partial\cK$.
+Furthermore, if $u^* \in \cK^\circ$ (interior) then \eqref{eq:vi} implies $f(u^*)=g$.
 
@@
 \end{equation}
+We say VI \eqref{eq:vi} is (\emph{strictly}) \emph{monotone} if $f$ is (strictly) monotone, and likewise for \emph{coercive}.
 \end{definition}
 
-It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, and also that the solution $u^* \in \cK$ is unique if $f$ is strictly monotone.
-The condition of continuity can be weakened to only apply on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but the stronger condition will apply in our examples.
-As in the calculus of variations \cite{Evans2010}, coercivity permits a compactness argument even for unbounded sets $\cK$; recall that the bounded, closed subsets of a reflexive Banach space are weakly compact.
+It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, and also that the solution $u^* \in \cK$ is unique when $f$ is strictly monotone.
+As in the calculus of variations \cite{Evans2010}, coercivity permits a compactness argument for unbounded sets $\cK$; recall that the bounded, closed subsets of a reflexive Banach space are weakly compact.
+Note that the condition of continuity can be weakened to only apply on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but the stronger condition will apply in our examples.
 
-The VIs solved in this paper satisfy a stronger form of coercivity than \eqref{eq:coercive}.
+The coercive VIs solved in this paper satisfy a stronger inequality than \eqref{eq:coercive}.
 
@@
 
-It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive, and thus the following result, which states that the $p$-coercive VI problems considered in this paper are well-posed, holds.
+It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive, and thus the following result holds.
+It states that the $p$-coercive VI problems considered in this paper are well-posed.
 
@@
 Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
-Furthermore, Proposition II.2.1 in the same reference shows that if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ (and $g=0$) if and only if
+Furthermore, if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ and $g=0$ if and only if
 \begin{equation}
-u^* = \operatorname{arg-min}_{v\in\cK} F(v). \label{eq:minimization}
+u^* = \operatorname{arg-min}_{v\in\cK} F(v) \label{eq:minimization}
 \end{equation}
+\cite[Proposition II.2.1]{EkelandTemam1976}.
 
 From now on $\Omega \subset \RR^d$ denotes a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.
-Note that Sobolev spaces \cite{Evans2010} are denoted here by $W^{k,p}(\Omega)$, with Hilbert-space cases denoted $W^{k,2}=H^k$.
+Note that Sobolev spaces \cite{Evans2010} are denoted here by $W^{k,p}(\Omega)$, for $k$ an integer and $1\le p \le \infty$, with $W^{k,2}=H^k$ a Hilbert space.
 
-Many numerical solvers for VIs, including the CD methods of Tai \cite{Tai2003}, address optimization problem \eqref{eq:minimization} under the hypothesis that $F'$ is coercive.
-The following example includes the classical obstacle problem \cite{GraeserKornhuber2009} and the obstacle problem for the $p$-Laplacian \cite{ChoeLewis1991}.
+The CD methods of Tai \cite{Tai2003} address optimization problem \eqref{eq:minimization} under the hypothesis that $F'$ is coercive.
+The following example includes the classical obstacle problem for the linear Laplacian \cite{GraeserKornhuber2009} and the obstacle problem for the $p$-Laplacian \cite{ChoeLewis1991}.
 
@@
 Now, if $x,y\in\RR^d$ then $(|x|^{p-2} x - |y|^{p-2} y)\cdot (x-y) \ge 2^{2-p} |x-y|^p$ \cite[see Appendix A and references therein]{Bueler2021conservation}.
-It follows from the Poincar\'e inequality that
+Thus it follows from the Poincar\'e inequality that
     $$\ip{f(u) - f(v)}{u-v} \ge 2^{2-p} a_0 \|\grad u - \grad v\|_p^p \ge 2^{2-p} a_0 C \|u-v\|^p$$
-for some $C>0$, and thus $f$ is $p$-coercive.  (The same $f$ is coercive if $1<p<2$ but the proof is somewhat different \cite[Theorem 4.4]{Bueler2021conservation}.)  Also, for $g\in\cV'$ define
+for some $C>0$, and thus $f$ is $p$-coercive.
+(The map in \eqref{eq:plaplacian} is coercive if $1<p<2$ but the proof is somewhat different \cite[Theorem 4.4]{Bueler2021conservation}; in Section \ref{sec:results} we need only the $p\ge 2$ case.)  For $g\in\cV'$ define
     $$F(u) = \int_\Omega \frac{a(x)}{p} |\grad u|^p\,dx - g(u).$$
-Then $F'(u) = f(u) - g$, $F$ is a convex functional (since $f$ is coercive), and, for any closed and convex $\cK\subset \cV$, VI problem \eqref{eq:vi} for $f$ and $g$ is equivalent to optimization problem \eqref{eq:minimization} for $F$.
+Then $F'(u) = f(u) - g$, $F$ is a convex functional (since $f$ is coercive), and, for any closed and convex $\cK\subset \cV$, VI problem \eqref{eq:vi} for is equivalent to optimization problem \eqref{eq:minimization}.
 \end{example}
 
-However, not all VIs correspond to optimization.
-We give two such examples next, first a linear advection-diffusion problem and then a nonlinear porous-medium-type problem.
-Each of these examples, which are important in applications, is preceded by a lemma.
+However, not all VIs arise from optimization problems.
+We give two such examples next, first a linear advection-diffusion problem, one which is coercive, and then a nonlinear porous-medium-type problem, which is not.
+Each of these examples is important in applications.
+The first is preceded by a lemma.
 
 \begin{lemma}  \label{lem:advectionskew}  \cite{Elmanetal2014}\,  Suppose $\bX :\Omega \to \RR^d$ is a smooth and bounded vector field on $\Omega$ with zero divergence ($\Div \bX=0$).
-For $u,v \in H^1(\Omega)$ let $s(u,v) = \int_\Omega (\bX \cdot \grad u) v\,dx$.
-Then $s(v,v) = \frac{1}{2} \int_{\partial \Omega} v^2 \bX\cdot \bn\,dx$ where $\bn$ is the outward normal on $\partial \Omega$.
+For $u,v \in H^1(\Omega)$ let $b(u,v) = \int_\Omega (\bX \cdot \grad u) v\,dx$.
+Then $b(u,u) = \frac{1}{2} \int_{\partial \Omega} u^2 \bX\cdot \bn\,dx$ where $\bn$ is the outward normal on $\partial \Omega$.
 \end{lemma}
@@
 \begin{proof}
-Integration by parts gives $s(u,v) = - s(v,u) + \int_{\partial \Omega} uv \bX\cdot \bn\,dx$, so the result follows.
+Integration by parts gives $b(u,v) = - b(v,u) + \int_{\partial \Omega} uv \bX\cdot \bn\,dx$, so the result follows.
 \end{proof}
@@
 Let $\cV = H_0^1(\Omega)$ be the space of functions with value zero on $\partial_D\Omega$.
-(For simplicity we consider zero boundary data.)  Choose a smooth and bounded velocity field $\bX$ on $\Omega$ such that $\Div \bX=0$, but additionally assume outward flow on the Neumann boundary, i.e.~$\bX \cdot \bn \ge 0$ at points of $\partial_N\Omega$.
+(For simplicity we consider zero boundary data.)  Consider a smooth and bounded velocity field $\bX$ on $\Omega$ such that $\Div \bX=0$, but additionally assume that the flow is outward on the Neumann boundary, i.e.~$\bX \cdot \bn \ge 0$ at points of $\partial_N\Omega$.
 For $u,v \in \cV = H_0^1(\Omega)$ and $\eps>0$ define
@@
 Consider VI \eqref{eq:vi} for any closed and convex $\cK \subset \cV$ and any $g\in\cV'$.
-It is easy to see that $|\ip{f(u)}{v}| \le (\eps + \|\bX\|_\infty) \|u\| \|v\|$, thus that $f:\cK \to \cV'$ is continuous.
-On the other hand, by Lemma \ref{lem:advectionskew}, which says that $s(u,v)$ is skew-symmetric up to a nonnegative addition, and by the outward flow assumption and the Poincar\'e inequality,
+It is easy to see that $|\ip{f(u)}{v}| \le (\eps + \|\bX\|_\infty) \|u\| \|v\|$, thus that $f:\cK \to \cV'$ is continuous (and Lipschitz).
+On the other hand, Lemma \ref{lem:advectionskew} says that $s(u,v)$ is skew-symmetric up to a nonnegative term.
+However, by the outward flow assumption and the Poincar\'e inequality,
 \begin{align*}
-\ip{f(u)-f(v)}{u-v} &= \eps \int_\Omega |\grad u - \grad v|^2\,dx + s(u-v,u-v) \\
+\ip{f(u)-f(v)}{u-v} &= \eps \int_\Omega |\grad u - \grad v|^2\,dx + b(u-v,u-v) \\
                     &= \eps \int_\Omega |\grad u - \grad v|^2\,dx + \frac{1}{2} \int_{\partial_N\Omega} (u-v)^2 \bX\cdot\bn \ge \eps C \|u-v\|^2.
@@
 
-However, if $\bX \ne 0$ then \eqref{eq:vi} for the functional \eqref{eq:advectiondiffusion} is not a (constrained) minimization problem.
-Indeed, the Jacobian of $f$ is the bilinear form $J_f(u,v) = \int_\Omega \eps \grad u \cdot \grad v\,dx - s(u,v)$.
-For nonzero advection this is not symmetric, thus it cannot be the Hessian of an objective $F$.
-(For $u,v$ which are zero on all of $\partial \Omega$, $J_f(u,v) - J_f(v,u) = -2 s(u,v)$, and we may then construct $u,v$ so this result is nonzero anywhere $\bX$ is nonzero.)
-
+If $\bX \ne 0$ then VI \eqref{eq:vi} for $f$ in \eqref{eq:advectiondiffusion} is not a (constrained) minimization problem.
+Indeed, the Jacobian of $f$, namely the bilinear form $J_f(u,v) = \int_\Omega \eps \grad u \cdot \grad v\,dx - b(u,v)$, is not symmetric in that case  (For $u,v$ which are zero on all of $\partial \Omega$, $J_f(u,v) - J_f(v,u) = -2 b(u,v)$, and we may then construct $u,v$ so this result is nonzero anywhere $\bX$ is nonzero.)  Thus $f$ in \eqref{eq:advectiondiffusion} cannot be the Hessian of a scalar objective $F$.
 References \cite{Bueler2021conservation,ChangNakshatrala2017} consider such advection-diffusion VI problems over $\cK = \{v\ge 0\}$.
 
-\begin{example}  \label{ex:porous}  FIXME. I THINK THIS IS NOT GOING TO WORK.  COERCIVITY IS SUPER FRAGILE W.R.T. THIS CHANGE?
+\begin{example}  \label{ex:porous}  FIXME
 \begin{equation}
@@
 
-Examples \ref{ex:advectiondiffusion} and \ref{ex:porous} are VI problems which do not arise from optimization.
-Further examples are in ice sheet models \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids \cite{Bueler2021conservation}.
-Numerical solver performance for Examples \ref{ex:advectiondiffusion}--\ref{ex:plaplacian} will be considered in section \ref{sec:results}.
-% HOPE
+Numerical solver performance for Examples \ref{ex:plaplacian}, \ref{ex:advectiondiffusion}, and \ref{ex:porous} will be considered in Section \ref{sec:results}.
+Further examples of nonlinear VI problems appear in ice sheet models \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids \cite{Bueler2021conservation}.
 
@@
 
-\begin{figure}
+\begin{figure}[h]
 \includegraphics[width=0.55\textwidth]{genfigs/cartoon.pdf}
-\caption{A constraint decomposition (CD) in a one-sided obstacle problem with $\mathcal{K}=\{v\ge \psi\}$ looks something like this.}
+\caption{A constraint decomposition (CD) in a one-sided obstacle problem ($\mathcal{K}=\{v\ge \psi\}$) might like this if $\mathcal{V}$ were 2-dimensional.}
 \label{fig:cartoon}

### 2021-07-14 paper/genfigs/cartoon.py

@@ -18,7 +18,7 @@ def makeaxes(xmin, xmax, ymin, ymax):
     plt.plot(zero, y, 'k', lw = 1.0)
 
 plt.figure(figsize=(8, 6))
-makeaxes(-0.25, 1.25, -0.1, 0.85)
+makeaxes(-0.25, 1.2, -0.1, 0.8)
 plt.text(-0.05, -0.07, r'$0$', fontsize=fsize)
 
 psi1 = 0.2

### 2021-07-14 paper/mcd2.tex
Insert a couple of lemmas.

@@
 
+The following easy, but important, lemma follows directly from $p$-coercivity.
+Along the way we define a new quantity $E(v,u^*)$ which is bounded below by a function of the norm error in $v$.
+We will use this quantity in the proofs in Section \ref{sec:convergence}.
+
+\begin{lemma} \label{lem:normlike}  Suppose $f:\mathcal{K} \to \mathcal{V}'$ is $p$-coercive and $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.  For $v \in \mathcal{K}$,
+\begin{equation}
+  E(v,u^*) = \ip{f(v)}{v-u^*} - g(v-u^*) \ge \kappa \|v-u^*\|^p.  \label{eq:normlike}
+\end{equation}
+\end{lemma}
+
+\begin{proof}
+\begin{align*}
+E(v,u^*) &= \ip{f(v)}{v-u^*} - \ip{f(u^*)}{v-u^*} + \ip{f(u^*)}{v-u^*} - g(v-u^*) \\
+   &\ge \kappa \|v-u^*\|^p + 0.  \qedhere
+\end{align*}
+\end{proof}
+
 When $f$ is monotone, VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
@@
 
-From now on $\Omega \subset \RR^d$ denotes a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.
-Note that Sobolev spaces \cite{Evans2010} are denoted here by $W^{k,p}(\Omega)$, for $k$ an integer and $1\le p \le \infty$, with $W^{k,2}=H^k$ a Hilbert space.
-
 The CD methods of Tai \cite{Tai2003} address optimization problem \eqref{eq:minimization} under the hypothesis that $F'$ is coercive.
 The following example includes the classical obstacle problem for the linear Laplacian \cite{GraeserKornhuber2009} and the obstacle problem for the $p$-Laplacian \cite{ChoeLewis1991}.
+Note that from now on $\Omega \subset \RR^d$ denotes a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.
+Also, Sobolev spaces \cite{Evans2010} are denoted here by $W^{k,p}(\Omega)$, for $k$ an integer and $1\le p \le \infty$, with $W^{k,2}=H^k$ a Hilbert space.
 
@@
 Each starts from a current iterate $u \in \cK$ and computes a new iterate $w\in\cK$, an improved approximation of the solution $u^* \in \cK$ of VI problem \eqref{eq:vi}, according to a damping parameter $0<\alpha\le 1$.
-If \pr{parallel} $=$ \text{\textbf{true}} then the \text{\textbf{for}} loop can be computed in any order.
+For the parallel version the \text{\textbf{for}} loop can be computed in any order.
 
-\begin{pseudofloat}
+\begin{pseudofloat}[H]
 \begin{pseudo*}
@@
 \end{align*}
-for the two versions, respectively.  In other words, we solve VI \eqref{eq:vi} over (parallel or successive) perturbations from the sets $\cK_i$.  In fact, if
+for the two versions, respectively.  In other words, we solve VI \eqref{eq:vi} over (parallel or successive) perturbations from the sets $\cK_i$.  In fact, if we define
 \begin{equation}
@@
 \end{equation}
-is the update in $\cK_i$ then we may then write $w = u + \alpha \sum_i e_i$.
-In particular, note that $u - R_i u + \hat w_i = u + e_i$; the \pr{parallel} version computes the updates $e_i$ independently.
+as the update in $\cK_i$ then we may then write $w = u + \alpha \sum_i e_i$.
+In particular, note that $u - R_i u + \hat w_i = u + e_i$; the parallel version computes the updates $e_i$ independently.
+Also note that for one iteration of the parallel version of the basic CD algorithm, from \eqref{eq:constraintrestrictionsum} and \eqref{eq:ithupdate} observe that
+\begin{equation}
+\hat w = u + \sum_{i=0}^{m-1} e_i = u^* + \sum_{i=0}^{m-1} \hat w_i - R_i u^*.  \label{eq:ithupdatefacts}
+\end{equation}
 
@@
 
-\begin{assumptions*}  \emph{Lipschitz restrictions}: There exists a constant $C_1>0$ so that
+\begin{assumptions*}  \emph{Totally Lipschitz restrictions}: There exists a constant $C_1>0$ so that
 \begin{equation}
@@
 \begin{equation}
-\sum_{i=0}^{m-1} \sum_{j=0}^{m-1} \left|\ip{f(w_{ij} + v_i) - f(w_{ij})}{y_i}\right| \le C_2 \left(\sum_{i=0}^{m-1} \|v_i\|^2\right)^{1/2} \left(\sum_{i=0}^{m-1} \|y_i\|^2\right)^{1/2} \label{as:totallylipschitzresidual}
+\sum_{i=0}^{m-1} \sum_{j=0}^{m-1} \left|\ip{f(w_{ij} + v_i) - f(w_{ij})}{y_i}\right| \le C_2 \left(\sum_{i=0}^{m-1} \|v_i\|^2\right)^{1/2} \left(\sum_{i=0}^{m-1} \|y_i\|^2\right)^{1/2} \label{as:lipschitzresidual}
 \end{equation}
@@
 
-FIXME
+From the assumptions and the above definitions we have the following core lemma from which a convergence proof for the parallel algorithm is built.
+
+\begin{lemma} \label{lem:core}  Suppose $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.
+Suppose the constraint decomposition $\mathcal{V}_i,\mathcal{K}_i,R_i$ satisfies \eqref{as:lipschitzrestrictions}, and suppose $f$ satisfies \eqref{as:lipschitzresidual} as well.
+Suppose $\hat w$ is computed in the parallel version of Algorithm \ref{alg:basiccd}.
+Then
+\begin{equation}
+   \ip{f(\hat w)}{\hat w - u^*} \le C_2 \sum_{i=0}^{m-1} \|e_i\|^2 + C_1 C_2 \left(\sum_{i=0}^{m-1} \|e_i\|^2\right)^{1/2} \|u-u^*\|. \label{eq:core}
+\end{equation}
+\end{lemma}
+
+\begin{proof}
+FIXME: (25) in \cite{Tai2003}
+\end{proof}
+
+Note that the coercivity of $f$ is not actually used in proving Lemma \ref{lem:core}.
 

### 2021-07-15 paper/mcd2.tex
Edits and new material in sections 3 and 4.

@@
 
-The multilevel algorithms in Section \ref{sec:multilevel} are based on the more general notion of a solution by constraint decomposition, as follows.
-Suppose there are $m<\infty$ subspaces $\cV_i \subset \cV$ so that
+Suppose there are $m<\infty$ subspaces $\cV_i \subset \cV$ so that the sum
 \begin{equation}
@@
 holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$; this is called a \emph{subspace decomposition} \cite{Xu1992}.
-Suppose further that $\cK_i \subset \cV$ are nonempty, closed, and convex subsets such that
+Suppose further that $\cK_i \subset \cV_i$ are nonempty, closed, and convex subsets such that
 \begin{equation}
-\cK_i \subset \cV_i \qquad \text{and} \qquad \cK = \sum_{i=0}^{m-1} \cK_i. \label{eq:constraintdecomp}
+\cK = \sum_{i=0}^{m-1} \cK_i. \label{eq:constraintdecomp}
 \end{equation}
 The sum in \eqref{eq:constraintdecomp} is required to hold in two senses: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.
-(The latter sense is automatic in \eqref{eq:subspacedecomp} because $\cV_i$ are subspaces.)  Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is unique (in general), and that $\cK_i \not\subset \cK$ in many applications.
-
-Following \cite{Tai2003}, for each $\cK_i$ we also assume a bounded, (generally) nonlinear restriction operator $R_i : \cK \to \cK_i$ such that
+(The latter sense for \eqref{eq:subspacedecomp} is automatic because the $\cV_i$ are subspaces.)  Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique in any sense.
+Also, $\cK_i \not\subset \cK$ in many applications; see the cartoon in Figure \ref{fig:cartoon}.
+Finally, for each $\cK_i$ we assume that there are bounded, (generally) nonlinear restriction operators $R_i : \cK \to \cK_i$ such that
 \begin{equation}
@@
 \end{equation}
-A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,R_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}.
+A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,R_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} \cite{Tai2003}.
 
@@
 \includegraphics[width=0.55\textwidth]{genfigs/cartoon.pdf}
-\caption{A constraint decomposition (CD) in a one-sided obstacle problem ($\mathcal{K}=\{v\ge \psi\}$) might like this if $\mathcal{V}$ were 2-dimensional.}
+\caption{A constraint decomposition (CD) in a one-sided obstacle problem ($\mathcal{K}=\{v\ge \psi\}$) might look like this if $\mathcal{V}$ were 2-dimensional.}
 \label{fig:cartoon}
@@
 
-In the Section \ref{sec:multilevel} we will introduce discretizations and describe practical algorithms, but the CDs in this section can apply at the level of the continuum problem.
-The following two examples illustrate this for obstacle problems \cite{GraeserKornhuber2009}.
-Example \ref{ex:domaindecomposition} is an overlapping domain decomposition and Example \ref{ex:frequencydecomposition} is a nonoverlapping frequency decomposition.
-(The multilevel FE CD in Section \ref{sec:multilevel} approximately implements a frequency decomposition.)
+In Section \ref{sec:multilevel} we will introduce discretizations and describe practical algorithms, but the fundamental CD concept applies even at the level of the continuum problem.
+The following two examples illustrate this for obstacle problems \cite{GraeserKornhuber2009}, the first being an overlapping domain decomposition and the second a disjoint frequency decomposition.
+The multilevel FE CD in Section \ref{sec:multilevel} approximates such an exact frequency decomposition.
 
@@
 Let $\cK = \{v \ge \psi\} \subset \cV$.
-Suppose further that $\{\phi_i\}_{i=0}^{m-1}$ is a smooth partition of unity on $\Omega$, i.e.~so that $0 \le \phi_i\le 1$ and $\sum_i \phi_i = 1$, and let $\Omega_i$ be the support of $\phi_i$.
+Suppose further that $\{\phi_i\}_{i=0}^{m-1}$ is a smooth partition of unity on $\Omega$, satisfying $0 \le \phi_i\le 1$ and $\sum_i \phi_i = 1$, and let $\Omega_i$ be the support of $\phi_i$.
 Let $\cV_i = \{w \in \cV:w|_{\Omega \setminus \Omega_i} =0 \}$, $\cK_i = \{v \in \cV_i: v \ge \phi_i \psi\}$, and $R_i(v) = \phi_i v$.
@@
 
-%Note that the boundedness of $R_i(v) = \max\{v,\phi_i \psi\}$ depends on the relatively-low regularity of $\cV = H_0^1(\Omega)$.
-That is, if $u,v \in H^{1,p}(\Omega)$ for $1\le p \le \infty$ then $\max\{u,v\} \in H^{1,p}(\Omega)$.
-(To prove this note $\max\{u,0\} \in H^{1,p}(\Omega)$ is shown by \cite[Theorem II.A.1]{KinderlehrerStampacchia1980}; then combine with facts $\max\{a,b\} = \frac{1}{2} (a + b + |a-b|)$ and $|a| = \max\{a,0\} + \max\{-a,0\}$ for $a\in\RR$.)  However, Sobolev spaces $H^{k,p}(\Omega)$ are not generally closed under the maximum operation for $k\ge 2$.
-
-\begin{example}  \label{ex:frequencydecomposition}  For simplicity suppose $\Omega = (0,a)^d \subset \RR^d$ for $a>0$, a cube, and let $\cV = H_{\text{per}}^k(\Omega)$, $k\ge 0$, be the periodic functions.
-Suppose $\psi \in H^k(\Omega)$ and let $\cK = \{v \ge \psi\} \subset \cV$.
-Without giving any detailed notation of frequency representation, but noting that the frequencies are discrete, suppose $\cV_i \subset \cV$ are $m<\infty$ subspaces from an (nonoverlapping) partition by frequency, thus satisfying \eqref{eq:subspacedecomp}.
+\begin{example}  \label{ex:frequencydecomposition}  For simplicity suppose $\Omega = (0,a)^d \subset \RR^d$ is a cube, and let $\cV = H_{\text{per}}^k(\Omega)$, $k\ge 0$, be the periodic functions.
+Suppose $\psi \in H_{\text{per}}^k(\Omega)$ and let $\cK = \{v \ge \psi\} \subset \cV$.
+Without using any detailed notation for Fourier representation, but noting that the frequencies are discrete, suppose $\{\cV_i\}$ are $m<\infty$ subspaces of $\cV$ defined by an (nonoverlapping) partition by frequency, thus satisfying \eqref{eq:subspacedecomp} as an orthogonal decomposition.
 Suppose $P_i:\cV \to \cV_i$ are the corresponding orthogonal projections, satisfying $I = \sum_i P_i$.
 Let $\cK_i = \{v \ge P_i \psi\} \subset \cV_i$ and $R_i = P_i$.
-Then \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} hold.
+Then \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} also hold.
 \end{example}
 
-In Example \ref{ex:domaindecomposition}, note that $\cK_i \not\subset \cK$ in many cases, for example if $\psi$ is positive over portions of $\Omega$ where the decomposition into overlapping subdomains $\Omega_i$ is nontrivial.
-A similar comment applies in Example \ref{ex:frequencydecomposition}.
+Note that $\cK_i \not\subset \cK$ in most cases.
+In Example \ref{ex:domaindecomposition}, if $\psi$ is positive over portions of $\Omega$ where the decomposition into overlapping subdomains $\Omega_i$ is nontrivial, then $\cK_i \not\subset \cK$, and similarly for Example \ref{ex:frequencydecomposition}.
+Instead, the important inclusion is $\cK_i \subset \cV_i$.
 
-We now state the basic CD algorithm as an iteration which solves VI problems over each set $\cK_i$ (Algorithm \ref{alg:basiccd}).
-(By contrast, algorithms in \cite{Tai2003} solve \emph{optimization} problems.)  There are parallel (additive) and successive (multiplicative) versions; these generalize the Jacobi and Gauss-Seidel iterations \cite{Greenbaum1997}, respectively.
-Each starts from a current iterate $u \in \cK$ and computes a new iterate $w\in\cK$, an improved approximation of the solution $u^* \in \cK$ of VI problem \eqref{eq:vi}, according to a damping parameter $0<\alpha\le 1$.
-For the parallel version the \text{\textbf{for}} loop can be computed in any order.
+Algorithm \ref{alg:basiccd} below states the basic CD algorithm as an iteration which solves smaller VI problems over each set $\cK_i$.
+(By contrast, the algorithms in \cite{Tai2003} solve optimization problems over each $\cK_i$.)  The algorithm starts from a current iterate $u \in \cK$ and computes a new iterate $w\in\cK$, an improved approximation of the solution $u^* \in \cK$ of \eqref{eq:vi}, using a damping parameter $0<\alpha\le 1$.
+There are parallel (additive) and successive (multiplicative) versions of the algorithm, generalizing the Jacobi and Gauss-Seidel iterations \cite{Greenbaum1997}, respectively.
+For the parallel version the \textbf{for} loop can be computed in any order.
 
@@
 \begin{pseudo*}
-\pr{cd}(0<\alpha\le 1, u\in\cK)\text{:} \\+
+\pr{cditeration}(u\in\cK)\text{:} \\+
     if \pr{parallel}: \\+
@@
                 $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\Big)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\-
-            $w_i = (1-\alpha) R_i u + \alpha \hat w_i$ \\--
-    $\hat w = \sum_i \hat w_i$ \\
-    return $w=(1-\alpha) u + \alpha \hat w$
+            $w_i = (1-\alpha) R_i u + \alpha \hat w_i\in\cK_i$ \\--
+    $\hat w = \sum_i \hat w_i\in\cK$ \\
+    return $w=(1-\alpha) u + \alpha \hat w\in\cK$
 \end{pseudo*}
@@
 
-The reader may confirm that, inside each boxed VI, the argument of $f$ is an element of $\cK$.
-On the other hand, the expression $v_i - \hat w_i$ appears.
-To clarify this situation, note that by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write $v_i - \hat w_i$ as a difference of admissible vectors (i.e.~from $\cK$), namely
+The reader may confirm that, inside each boxed VI in Algorithm \ref{alg:basiccd}, the argument of $f$ is an element of $\cK$.
+While the $\cK_i$ difference $v_i - \hat w_i$ appears, note that by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write it as a difference of admissible vectors (i.e.~from $\cK$), namely
 \begin{align*}
@@
 \end{align*}
-for the two versions, respectively.  In other words, we solve VI \eqref{eq:vi} over (parallel or successive) perturbations from the sets $\cK_i$.  In fact, if we define
+for the two versions, respectively.  In other words, we solve VI \eqref{eq:vi} over (parallel or successive) perturbations from the sets $\cK_i$.
+
+In fact, if we define
 \begin{equation}
@@
 \end{equation}
-as the update in $\cK_i$ then we may then write $w = u + \alpha \sum_i e_i$.
-In particular, note that $u - R_i u + \hat w_i = u + e_i$; the parallel version computes the updates $e_i$ independently.
-Also note that for one iteration of the parallel version of the basic CD algorithm, from \eqref{eq:constraintrestrictionsum} and \eqref{eq:ithupdate} observe that
-\begin{equation}
-\hat w = u + \sum_{i=0}^{m-1} e_i = u^* + \sum_{i=0}^{m-1} \hat w_i - R_i u^*.  \label{eq:ithupdatefacts}
-\end{equation}
+as the $i$th update from a pass through a \textbf{for} loop in Algorithm \ref{alg:basiccd} then $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.
+Easy identities $u - R_i u + \hat w_i = u + e_i$ and $\hat w = u^* + \sum_i \hat w_i - R_i u^*$ will also be used.
 
-The convergence results in the next section require substantial damping in the parallel algorithm, i.e.~$\alpha \le 1/m$, as in \cite{Tai2003} as well, but the successive version can be shown to converge without damping ($\alpha=1$).
+The convergence results in the next section require substantial damping when applying the parallel algorithm, i.e.~$\alpha \le 1/m$ \cite{Tai2003}, but the successive version can be shown to converge without damping ($\alpha=1$).
 In Section \ref{sec:results} we demonstrate practical convergence for a larger range of $\alpha$ than suggested by the theory.
@@
 
-In Section \ref{sec:multilevel} we will convert the basic algorithm into a multilevel approach for an obstacle problem by choosing appropriate $\cK_i$ based on a FE triangulation.
-However, in order that the algorithm has a practical and efficient implementation, we will require additional concepts when $f$ has any of the following properties: \emph{(i)} $f$ is defined only on $\cK$, or \emph{(ii)} $f$ is nonlinear, or \emph{(iii)} $f$ is non-local.
-The issue of practical implementation for such $f$ seems not to have been addressed by references which apply CD algorithms only to the classical obstacle problem \cite[in particular]{GraeserKornhuber2009,Tai2003}.
-Consider the following example.
+In order for Algorithm \ref{alg:basiccd} to have a practical and efficient FE implementation (Section \ref{sec:multilevel}), we need additional notions when $f$ has any of the following properties: \emph{(i)} $f$ is defined only on $\cK$, or \emph{(ii)} $f$ is nonlinear, or \emph{(iii)} $f$ is non-local.
+The issue of practical implementation for any such $f$ seems not to have been addressed in the CD literature; in particular  \cite{GraeserKornhuber2009,Tai2003} only apply the method to the classical obstacle problem.
+The following example explains the simplifications available in that easier case.
 
 \begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear and defined on all of $\cV$.
-Furthermore suppose $f$ is local in the sense that a basis $\{\phi_i\}$ of $\cV_i$ exists with each support small in $\Omega$, and that $\ip{f(\phi_i)}{z}$ can be computed by an integral over the support of $\phi_i$.
-(When $\cV$ is an FE space then this is the usual case for evaluating PDE weak forms over a basis of hat functions; see Section \ref{sec:multilevel}.)  Considering only the parallel algorithm, for simplicity, the VI over $\cK_i$ can be written as
+Furthermore suppose $f$ is local in the sense that a basis $\{\phi_i\}$ of $\cV_i$ exists with each support small in $\Omega$ in the sense that $\ip{f(\phi_i)}{z}$ can be computed by an integral over the support of $\phi_i$.
+(When $\cV$ is an FE space then this is the usual case for evaluating PDE weak forms over a basis of hat functions---see Section \ref{sec:multilevel}---but integral operators do not generally have this property.)  Considering only the parallel algorithm for simplicity, the VI over $\cK_i$ in Algorithm \ref{alg:basiccd} can be written as
 \begin{equation}
@@
 \end{equation}
-for all $v_i\in \cK_i$, where $\tilde g(z) = g(z) - \ip{f(u)}{z}$.
+where $\tilde g(z) = g(z) - \ip{f(u)}{z}$.
 Noting $e_i = \hat w_i - R_i u \in \cV_i \notin \cK$, in general, VI \eqref{eq:linearlocalvi} only makes sense because $f$ is defined over $\cV_i$.
-Thus each subset problem \eqref{eq:linearlocalvi} can be solved using a stored residual $\ip{f(y)}{\cdot}$ (in the updated source term) and incrementally by computations over the basis supports.
-In other words, because $f$ has all these nice properties, the solution of each VI problem can be computed by inexpensive means.
+In this case each subset problem \eqref{eq:linearlocalvi} can be solved using a stored residual $\ip{f(u)}{\cdot}$ included into the updated source term.
+The VI \eqref{eq:vi} is then approximately solved in any incremental manner, by computations over the basis supports, by one application of \pr{cditeration}.
 \end{example}
 
-A solver which would work for Example \ref{ex:fnice} has implementation efficiencies that are unavailable in general.
-Additional ideas, like those in Section \ref{sec:multilevel}, are needed for practical application of CD methods to problems which are not close to the classical obstacle problem.
-In particular, we will extend the basic algorithm to nonlinear $f:\cK\to \cV'$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
+In other words, because $f$ has all these nice properties, the decomposed VI problems in Example \ref{ex:fnice} can be computed by inexpensive means.
+Note that the solution method for VI \eqref{eq:linearlocalvi} is not the concern here; we are observing that the data of problem \eqref{eq:linearlocalvi} is small in practice.
+
+A solver which would work for Example \ref{ex:fnice} can thus have implementation efficiencies which are unavailable in general.
+Additional ideas, like those in Section \ref{sec:multilevel}, are therefore needed for practical application of CD methods to more difficult problems.
+In particular, we will extend the basic algorithm to nonlinear $f$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}, in which case the decomposed problems only evaluate $f$ over (admissible) elements of $\mathcal{K}$.
 However, extending the algorithm to non-local residual functionals is a topic for future research.
@@
 
-To prove the convergence of the basic algorithm we make two assumptions as follows.
+To prove the convergence of the basic algorithm we make two assumptions, as follows, which are inequalities (7) and (8) in \cite{Tai2003}.
+These inequalities might be described as ``totally Lipschitz'' requirements for the maps $R_i$ and $f$, respectively, over the constraint decomposition.
 
-\begin{assumptions*}  \emph{Totally Lipschitz restrictions}: There exists a constant $C_1>0$ so that
+\begin{assumptions*}  There exists a constant $C_1>0$ so that
 \begin{equation}
@@
 \end{equation}
-for all $u,v\in\cK$.  \emph{Totally Lipschitz residual}: There exists $C_2>0$ so that
+for all $u,v\in\cK$.  Furthermore, there exists $C_2>0$ so that
 \begin{equation}
@@
 \end{equation}
-for all $w_{ij} \in \cK$ and $v_i,y_i \in \cV_i$ such that $w_{ij} + v_i \in \cK$ (for all $i,j$).  Note that the constants $C_1,C_2$ generally depend on $m$.
+for all $w_{ij} \in \cK$ and $v_i,y_i \in \cV_i$ such that $w_{ij} + v_i \in \cK$ (for all $i,j$).
 \end{assumptions*}
-% IF PROOF BELOW SUCCEEDS, SO ALGORITHM IS NORM DECREASING, THEN MAY BE ABLE TO ARGUE THAT IT IS O.K.
-FOR CONSTANTS TO DEPEND ON $u,v,w_{ij}$ etc.
-AS LONG AS THEY ARE UNIFORMLY BOUNDED ON CERTAIN BOUNDED SETS
 
-Though the assumptions are technical in certain senses, the reader should note some essential features:
+These technical assumptions have some essential features:
 \begin{itemize}
-\item Assumption \eqref{as:lipschitzrestrictions} addresses only the constraint decomposition.
-Furthermore, it is often the case that \eqref{as:totallylipschitzresidual} addresses only a subspace decomposition property of the residual functional.
-(For example, in the classical obstacle problem \eqref{as:totallylipschitzresidual} is verified over the (sub)spaces $\cV$ and $\cV_i$, and the subsets $\cK$ and $\cK_i$ are not involved.)
-\item The assumptions are identical to inequalities (7), (8) in \cite{Tai2003} where they are used to prove convergence of the optimization version(s) of the basic CD algorithm.
-See also their unconstrained versions, inequalities (13), (14) in \cite{TaiXu2002}.
-\item Tai \cite{Tai2003} gives values of the constants $C_1$ in \eqref{as:lipschitzrestrictions} for obstacle problem constraint decompositions using $P_1$ FE spaces over a shape-regular and quasi-uniform triangulation.
-In particular, constants $C_1$ are know for overlapping domain decompositions (with or without an additional coarse mesh) and standard multilevel heirarchies.
+\item Assumption \eqref{as:lipschitzrestrictions} addresses only the constraint decomposition, and not the map $f$.
+\item It is sometimes the case that \eqref{as:lipschitzresidual} addresses only a subspace decomposition property of the operator $f$, and not the CD sets $\cK_i$ or maps $R_i$.
+For example, in the classical obstacle problem \eqref{as:lipschitzresidual} is verified over all $w_{ij} \in \cV$.
+\item The unconstrained versions of \eqref{as:lipschitzrestrictions} and \eqref{as:lipschitzresidual} are (13), (14) in \cite{TaiXu2002}, respectively.
+\item Tai \cite{Tai2003} gives values of the constants $C_1$ in \eqref{as:lipschitzrestrictions} for obstacle problem constraint decompositions using $P_1$ FE spaces over shape-regular and quasi-uniform triangulations, which is the only case we will demonstrate (Section \ref{sec:results}).
+In particular, constants $C_1$ are know for overlapping domain decompositions (with or without an additional coarse mesh) and standard multilevel hierarchies.
+\item Constants $C_1,C_2$ generally depend on $m$.
+However, as noted in \cite{Tai2003}, and regarding the convergence proof below, if repeated application of Algorithm \ref{alg:basiccd} is known to generate a bounded sequence of iterates in $\mathcal{K}$ then constants $C_1,C_2$ are actually permitted to depend on $u,v,w_{ij},v_i,y_i$ as well.
 % ASSUMING $f$ IS LIPSCHITZ WILL MAKE THE EXISTENCE OF C_2 CLEAR; SEE (14) IN \cite{TaiXu2002}
@@
 
-From the assumptions and the above definitions we have the following core lemma from which a convergence proof for the parallel algorithm is built.
+From the assumptions and the above definitions we have the following core lemma for the parallel version.  Our proof follows \cite{Tai2003}.
 
 \begin{lemma} \label{lem:core}  Suppose $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.
-Suppose the constraint decomposition $\mathcal{V}_i,\mathcal{K}_i,R_i$ satisfies \eqref{as:lipschitzrestrictions}, and suppose $f$ satisfies \eqref{as:lipschitzresidual} as well.
-Suppose $\hat w$ is computed in the parallel version of Algorithm \ref{alg:basiccd}.
+Suppose $\mathcal{V}_i$, $\mathcal{K}_i$, $R_i$, and $f$ satisfy \eqref{as:lipschitzrestrictions} and \eqref{as:lipschitzresidual}.
+Suppose $\hat w$ is computed from $u$ by the parallel version of Algorithm \ref{alg:basiccd}.
+Recall $e_i = \hat w_i - R_i u$.
 Then
@@
 \begin{proof}
-FIXME: (25) in \cite{Tai2003}
+For $i\in \{0,1,\dots,m-1\}$ use increasing, but wrapped, indices to define the following sequence $\phi_{ij}=\phi_{i,j} \in \cK$:
+\begin{align*}
+\phi_{i,0} &= u + e_i, \\
+\phi_{i,1} &= u + e_i + e_{i+1}, \\
+  &\vdots \\
+\phi_{i,m-1} &= u + e_i + e_{i+1} + \dots + e_{m-1} + e_0 + \dots + e_{i-1}.
+\end{align*}
+(Each $\phi_{ij} \in \cK$ by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum}.)  Noting that the boxed VI in the parallel version of Algorithm \ref{alg:basiccd} can be written
+
+    FIXME
 \end{proof}
 
-Note that the coercivity of $f$ is not actually used in proving Lemma \ref{lem:core}.
+Note that the coercivity of $f$ is not used in proving Lemma \ref{lem:core}.
 

### 2021-07-16 paper/mcd2.tex
Put down a lemma, following one part of the Tai convergence argument.

@@
 
-The following easy, but important, lemma follows directly from $p$-coercivity.
-Along the way we define a new quantity $E(v,u^*)$ which is bounded below by a function of the norm error in $v$.
-We will use this quantity in the proofs in Section \ref{sec:convergence}.
+When $f$ is monotone VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
+Suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cK \to \cV'$.
+Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
+Furthermore, \cite[Proposition II.2.1]{EkelandTemam1976} shows that if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ and $g=0$ if and only if
+\begin{equation}
+u^* = \operatorname{arg-min}_{v\in\cK} F(v). \label{eq:minimization}
+\end{equation}
+The CD methods of Tai \cite{Tai2003} address problem \eqref{eq:minimization} under the hypothesis that $F'$ is coercive.
 
-\begin{lemma} \label{lem:normlike}  Suppose $f:\mathcal{K} \to \mathcal{V}'$ is $p$-coercive and $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.  For $v \in \mathcal{K}$,
-\begin{equation}
-  E(v,u^*) = \ip{f(v)}{v-u^*} - g(v-u^*) \ge \kappa \|v-u^*\|^p.  \label{eq:normlike}
-\end{equation}
-\end{lemma}
-
-\begin{proof}
-\begin{align*}
-E(v,u^*) &= \ip{f(v)}{v-u^*} - \ip{f(u^*)}{v-u^*} + \ip{f(u^*)}{v-u^*} - g(v-u^*) \\
-   &\ge \kappa \|v-u^*\|^p + 0.  \qedhere
-\end{align*}
-\end{proof}
-
-When $f$ is monotone, VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
-In fact, suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cK \to \cV'$.
-Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
-Furthermore, if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ and $g=0$ if and only if
-\begin{equation}
-u^* = \operatorname{arg-min}_{v\in\cK} F(v) \label{eq:minimization}
-\end{equation}
-\cite[Proposition II.2.1]{EkelandTemam1976}.
-
-The CD methods of Tai \cite{Tai2003} address optimization problem \eqref{eq:minimization} under the hypothesis that $F'$ is coercive.
 The following example includes the classical obstacle problem for the linear Laplacian \cite{GraeserKornhuber2009} and the obstacle problem for the $p$-Laplacian \cite{ChoeLewis1991}.
 Note that from now on $\Omega \subset \RR^d$ denotes a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.
-Also, Sobolev spaces \cite{Evans2010} are denoted here by $W^{k,p}(\Omega)$, for $k$ an integer and $1\le p \le \infty$, with $W^{k,2}=H^k$ a Hilbert space.
+Also, Sobolev spaces \cite{Evans2010} are denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$, with $W^{k,2}$ denoted by $H^k$.
 
@@
 for some $C>0$, and thus $f$ is $p$-coercive.
-(The map in \eqref{eq:plaplacian} is coercive if $1<p<2$ but the proof is somewhat different \cite[Theorem 4.4]{Bueler2021conservation}; in Section \ref{sec:results} we need only the $p\ge 2$ case.)  For $g\in\cV'$ define
+(The map in \eqref{eq:plaplacian} is coercive if $1<p<2$, but the proof is somewhat different \cite[Theorem 4.4]{Bueler2021conservation}, and in Section \ref{sec:results} we need only the $p\ge 2$ case.)  For $g\in\cV'$ define
     $$F(u) = \int_\Omega \frac{a(x)}{p} |\grad u|^p\,dx - g(u).$$
@@
 
-However, not all VIs arise from optimization problems.
-We give two such examples next, first a linear advection-diffusion problem, one which is coercive, and then a nonlinear porous-medium-type problem, which is not.
+However, not all VI problems arise from optimization.
+We give two such examples next, first a coercive and linear advection-diffusion problem, and then a nonlinear porous-medium-type problem.
 Each of these examples is important in applications.
@@
 If $\bX \ne 0$ then VI \eqref{eq:vi} for $f$ in \eqref{eq:advectiondiffusion} is not a (constrained) minimization problem.
-Indeed, the Jacobian of $f$, namely the bilinear form $J_f(u,v) = \int_\Omega \eps \grad u \cdot \grad v\,dx - b(u,v)$, is not symmetric in that case  (For $u,v$ which are zero on all of $\partial \Omega$, $J_f(u,v) - J_f(v,u) = -2 b(u,v)$, and we may then construct $u,v$ so this result is nonzero anywhere $\bX$ is nonzero.)  Thus $f$ in \eqref{eq:advectiondiffusion} cannot be the Hessian of a scalar objective $F$.
+Indeed, the Jacobian of $f$, namely the bilinear form $J_f(u,v) = \int_\Omega \eps \grad u \cdot \grad v\,dx - b(u,v)$, is not symmetric in that case.\footnote{Assume $\bX \ne 0$ is continuous for simplicity.
+For $u,v$ which are zero on $\partial \Omega$, $J_f(u,v) - J_f(v,u) = -2 b(u,v)$.
+There exist $u,v$ so $b(u,v)$ is nonzero; construct $u,v$ locally near some point where $\bX$ is nonzero.}  Thus $f$ in \eqref{eq:advectiondiffusion} cannot be the Hessian of a scalar objective $F$.
 References \cite{Bueler2021conservation,ChangNakshatrala2017} consider such advection-diffusion VI problems over $\cK = \{v\ge 0\}$.
@@
 
-Suppose there are $m<\infty$ subspaces $\cV_i \subset \cV$ so that the sum
+Suppose there are $m<\infty$ closed subspaces $\cV_i \subset \cV$ so that the sum
 \begin{equation}
@@
 \end{equation}
-The sum in \eqref{eq:constraintdecomp} is required to hold in two senses: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.
-(The latter sense for \eqref{eq:subspacedecomp} is automatic because the $\cV_i$ are subspaces.)  Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique in any sense.
+The sum in \eqref{eq:constraintdecomp} is required to hold in two senses: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.\footnote{Sense \emph{(ii)} for \eqref{eq:subspacedecomp} is automatic because the $\cV_i$ are subspaces.}
+
+Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique.
 Also, $\cK_i \not\subset \cK$ in many applications; see the cartoon in Figure \ref{fig:cartoon}.
-Finally, for each $\cK_i$ we assume that there are bounded, (generally) nonlinear restriction operators $R_i : \cK \to \cK_i$ such that
+
+Finally, for each $\cK_i$ we assume that there are bounded, (generally) nonlinear restriction operators $R_i : \cK \to \cK_i$ such that if $v \in \cK$ then
 \begin{equation}
-v = \sum_{i=0}^{m-1} R_i v \qquad \text{if } v \in \cK.  \label{eq:constraintrestrictionsum}
+v = \sum_{i=0}^{m-1} R_i v.  \label{eq:constraintrestrictionsum}
 \end{equation}
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 \includegraphics[width=0.55\textwidth]{genfigs/cartoon.pdf}
@@
 In Section \ref{sec:multilevel} we will introduce discretizations and describe practical algorithms, but the fundamental CD concept applies even at the level of the continuum problem.
-The following two examples illustrate this for obstacle problems \cite{GraeserKornhuber2009}, the first being an overlapping domain decomposition and the second a disjoint frequency decomposition.
-The multilevel FE CD in Section \ref{sec:multilevel} approximates such an exact frequency decomposition.
+The following two examples illustrate this for obstacle problems \cite{GraeserKornhuber2009}.
+First we consider an overlapping domain decomposition.
 
@@
 \end{example}
+
+Our second example is a disjoint frequency decomposition.  A multilevel FE CD, e.g.~the one proposed in Section \ref{sec:multilevel}, approximates such a frequency decomposition.
 
@@
 The reader may confirm that, inside each boxed VI in Algorithm \ref{alg:basiccd}, the argument of $f$ is an element of $\cK$.
-While the $\cK_i$ difference $v_i - \hat w_i$ appears, note that by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write it as a difference of admissible vectors (i.e.~from $\cK$), namely
+While the difference $v_i - \hat w_i \in \cV_i$ appears, note that by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write it as a difference of admissible vectors (i.e.~from $\cK$), namely
 \begin{align*}
@@
 \end{align*}
-for the two versions, respectively.  In other words, we solve VI \eqref{eq:vi} over (parallel or successive) perturbations from the sets $\cK_i$.
+for the two versions, respectively.
 
-In fact, if we define
+In other words, we solve VI \eqref{eq:vi} over perturbations from the sets $\cK_i$.  Define
 \begin{equation}
@@
 \end{equation}
-as the $i$th update from a pass through a \textbf{for} loop in Algorithm \ref{alg:basiccd} then $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.
-Easy identities $u - R_i u + \hat w_i = u + e_i$ and $\hat w = u^* + \sum_i \hat w_i - R_i u^*$ will also be used.
+as the $i$th update from a pass through either \textbf{for} loop in Algorithm \ref{alg:basiccd}.
+Then $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.
+Easy identities $u - R_i u + \hat w_i = u + e_i$ and $\hat w = u^* + \sum_i \hat w_i - R_i u^*$ also hold.
+We will use these identities in Section \ref{sec:convergence}.
 
-The convergence results in the next section require substantial damping when applying the parallel algorithm, i.e.~$\alpha \le 1/m$ \cite{Tai2003}, but the successive version can be shown to converge without damping ($\alpha=1$).
+The convergence results in Section \ref{sec:convergence} require substantial damping when applying the parallel algorithm, i.e.~$\alpha \le 1/m$, the same as in \cite{Tai2003}, but the successive version can be shown to converge without damping ($\alpha=1$).
 In Section \ref{sec:results} we demonstrate practical convergence for a larger range of $\alpha$ than suggested by the theory.
@@
 In order for Algorithm \ref{alg:basiccd} to have a practical and efficient FE implementation (Section \ref{sec:multilevel}), we need additional notions when $f$ has any of the following properties: \emph{(i)} $f$ is defined only on $\cK$, or \emph{(ii)} $f$ is nonlinear, or \emph{(iii)} $f$ is non-local.
-The issue of practical implementation for any such $f$ seems not to have been addressed in the CD literature; in particular  \cite{GraeserKornhuber2009,Tai2003} only apply the method to the classical obstacle problem.
-The following example explains the simplifications available in that easier case.
+The issue of practical implementation for any such $f$ seems not to have been addressed in the CD literature, which is why we call Algorithm \ref{alg:basiccd} the ``basic'' algorithm.
+In particular  \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem.
+The following example explains the simplifications available in that easy case.
 
 \begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear and defined on all of $\cV$.
-Furthermore suppose $f$ is local in the sense that a basis $\{\phi_i\}$ of $\cV_i$ exists with each support small in $\Omega$ in the sense that $\ip{f(\phi_i)}{z}$ can be computed by an integral over the support of $\phi_i$.
-(When $\cV$ is an FE space then this is the usual case for evaluating PDE weak forms over a basis of hat functions---see Section \ref{sec:multilevel}---but integral operators do not generally have this property.)  Considering only the parallel algorithm for simplicity, the VI over $\cK_i$ in Algorithm \ref{alg:basiccd} can be written as
+Furthermore suppose $f$ is local in the sense that a basis $\{\phi_i\}$ of $\cV_i$ exists, with each support small in $\Omega$, such that $\ip{f(\phi_i)}{z}$ can be computed by an integral over the support of $\phi_i$.
+(When $\cV$ is an FE space then this is the usual case for evaluating PDE weak forms over a basis of hat functions; see Section \ref{sec:multilevel}.
+Constrast integral operators where $\ip{f(\phi_i)}{z}$ requires an integral over $\Omega$ even if the support of $\phi_i$ is small.)  Considering only the parallel version of Algorithm \ref{alg:basiccd} for simplicity, the boxed VI over $\cK_i$ can be written as
 \begin{equation}
-\ip{f(\hat w_i - R_i u)}{v_i-\hat w_i} \ge \tilde g(v_i-\hat w_i) \label{eq:linearlocalvi}
+\ip{f(e_i)}{v_i-\hat w_i} \ge \tilde g(v_i-\hat w_i) \label{eq:linearlocalvi}
 \end{equation}
 where $\tilde g(z) = g(z) - \ip{f(u)}{z}$.
-Noting $e_i = \hat w_i - R_i u \in \cV_i \notin \cK$, in general, VI \eqref{eq:linearlocalvi} only makes sense because $f$ is defined over $\cV_i$.
-In this case each subset problem \eqref{eq:linearlocalvi} can be solved using a stored residual $\ip{f(u)}{\cdot}$ included into the updated source term.
-The VI \eqref{eq:vi} is then approximately solved in any incremental manner, by computations over the basis supports, by one application of \pr{cditeration}.
+Noting $e_i = \hat w_i - R_i u \in \cV_i \notin \cK$, in general, VIs \eqref{eq:linearlocalvi} will only make sense because $f$ is defined over $\cV$.
+Each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $\ip{f(u)}{\cdot}$ (already included into the source term).
+The VI \eqref{eq:vi} is then approximately solved in any incremental and efficient manner, by computations over the basis supports, during one application of \pr{cditeration}.
 \end{example}
 
-In other words, because $f$ has all these nice properties, the decomposed VI problems in Example \ref{ex:fnice} can be computed by inexpensive means.
-Note that the solution method for VI \eqref{eq:linearlocalvi} is not the concern here; we are observing that the data of problem \eqref{eq:linearlocalvi} is small in practice.
+In other words, when $f$ has all the nice properties proposed in Example \ref{ex:fnice} then the decomposed VI problems can be computed from inexpensive data.
+Note that the actual solution method for VI \eqref{eq:linearlocalvi} is not the concern here; we are observing that the data of problem \eqref{eq:linearlocalvi}, and the cost of residual evaluation, is small in practice.
+A solver which would work for Example \ref{ex:fnice} can thus have implementation efficiencies which are unavailable in general.
 
-A solver which would work for Example \ref{ex:fnice} can thus have implementation efficiencies which are unavailable in general.
-Additional ideas, like those in Section \ref{sec:multilevel}, are therefore needed for practical application of CD methods to more difficult problems.
-In particular, we will extend the basic algorithm to nonlinear $f$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}, in which case the decomposed problems only evaluate $f$ over (admissible) elements of $\mathcal{K}$.
+Additional ideas, like those in Section \ref{sec:multilevel}, are needed for practical application of CD methods to more difficult problems.
+In particular, we will extend the basic Algorithm \ref{alg:basiccd} to nonlinear $f$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}, in which case the decomposed problems only evaluate $f$ over (admissible) elements of $\mathcal{K}$, and we propose restrictions which allow inexpensive approximations of the decomposed problems.
 However, extending the algorithm to non-local residual functionals is a topic for future research.
@@
 
-To prove the convergence of the basic algorithm we make two assumptions, as follows, which are inequalities (7) and (8) in \cite{Tai2003}.
-These inequalities might be described as ``totally Lipschitz'' requirements for the maps $R_i$ and $f$, respectively, over the constraint decomposition.
+We now seek a proof of the convergence of Algorithm \ref{alg:basiccd}.
+This will be possible if we restrict to $2$-coercive operators and make certain assumptions as in \cite{Tai2003}.
+First we define a new quantity related to VI problem \eqref{eq:vi}.
+
+\begin{definition} Suppose $f:\cK \to \cV'$ and $g \in \cV'$.  For $u,v \in \cK$ let
+\begin{equation}
+  E(v,u) = \ip{f(v)}{v-u} - g(v-u).  \label{eq:normlikedefn}
+\end{equation}
+\end{definition}
+
+If $u^*$ solves \eqref{eq:vi} then $E(v,u^*)$ is somewhat like an ``merit function,'' something used in the context of solving nonlinear equations \cite{NocedalWright2006} to replace an objective (scalar) functional.
+(Note that $E(v,u^*)$ is \emph{not} a residual for \eqref{eq:vi}.)  The next lemma, which follows directly from $p$-coercivity, shows $E(v,u^*)$ is bounded below.
+
+\begin{lemma} \label{lem:normlike}  Suppose $f:\mathcal{K} \to \mathcal{V}'$ is $p$-coercive and $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.  For $v \in \mathcal{K}$,
+\begin{equation}
+  E(v,u^*) \ge \kappa \|v-u^*\|^p.  \label{eq:normlikebound}
+\end{equation}
+\end{lemma}
+
+\begin{proof}
+\begin{align*}
+E(v,u^*) &= \ip{f(v)}{v-u^*} - \ip{f(u^*)}{v-u^*} + \ip{f(u^*)}{v-u^*} - g(v-u^*) \\
+   &\ge \kappa \|v-u^*\|^p + 0.  \qedhere
+\end{align*}
+\end{proof}
+
+Next we make two assumptions which are essentially the same as (7) and (8) in \cite{Tai2003}.
 
@@
 \end{equation}
-for all $u,v\in\cK$.  Furthermore, there exists $C_2>0$ so that
+for all $u,v\in\cK$.  Furthermore, there exists $C_2>0$ so that for all $v_i \in \cV_i$ and $y_j \in \cV_j$,
 \begin{equation}
-\sum_{i=0}^{m-1} \sum_{j=0}^{m-1} \left|\ip{f(w_{ij} + v_i) - f(w_{ij})}{y_i}\right| \le C_2 \left(\sum_{i=0}^{m-1} \|v_i\|^2\right)^{1/2} \left(\sum_{i=0}^{m-1} \|y_i\|^2\right)^{1/2} \label{as:lipschitzresidual}
+\sum_{i=0}^{m-1} \sum_{j=0}^{m-1} \left|\ip{f(w_{ij} + v_i) - f(w_{ij})}{y_j}\right| \le C_2 \left(\sum_{i=0}^{m-1} \|v_i\|^2\right)^{1/2} \left(\sum_{j=0}^{m-1} \|y_j\|^2\right)^{1/2} \label{as:lipschitzresidual}
 \end{equation}
-for all $w_{ij} \in \cK$ and $v_i,y_i \in \cV_i$ such that $w_{ij} + v_i \in \cK$ (for all $i,j$).
+for all $w_{ij} \in \cK$ such that $w_{ij} + v_i \in \cK$ for all $i$ and $j$.
 \end{assumptions*}
 
-These technical assumptions have some essential features:
+These inequalities might be described as ``totally Lipschitz'' requirements for the maps $R_i$ and $f$, respectively.  We make the following observations:
 \begin{itemize}
-\item Assumption \eqref{as:lipschitzrestrictions} addresses only the constraint decomposition, and not the map $f$.
-\item It is sometimes the case that \eqref{as:lipschitzresidual} addresses only a subspace decomposition property of the operator $f$, and not the CD sets $\cK_i$ or maps $R_i$.
-For example, in the classical obstacle problem \eqref{as:lipschitzresidual} is verified over all $w_{ij} \in \cV$.
+\item Assumption \eqref{as:lipschitzrestrictions} addresses only the CD, and not the map $f$.
+\item It is sometimes the case that \eqref{as:lipschitzresidual} addresses only a subspace decomposition property of $f$ (and not the CD itself).
+For example, in the classical obstacle problem one may verify \eqref{as:lipschitzresidual} over all $w_{ij} \in \cV$.
 \item The unconstrained versions of \eqref{as:lipschitzrestrictions} and \eqref{as:lipschitzresidual} are (13), (14) in \cite{TaiXu2002}, respectively.
-\item Tai \cite{Tai2003} gives values of the constants $C_1$ in \eqref{as:lipschitzrestrictions} for obstacle problem constraint decompositions using $P_1$ FE spaces over shape-regular and quasi-uniform triangulations, which is the only case we will demonstrate (Section \ref{sec:results}).
-In particular, constants $C_1$ are know for overlapping domain decompositions (with or without an additional coarse mesh) and standard multilevel hierarchies.
-\item Constants $C_1,C_2$ generally depend on $m$.
-However, as noted in \cite{Tai2003}, and regarding the convergence proof below, if repeated application of Algorithm \ref{alg:basiccd} is known to generate a bounded sequence of iterates in $\mathcal{K}$ then constants $C_1,C_2$ are actually permitted to depend on $u,v,w_{ij},v_i,y_i$ as well.
-% ASSUMING $f$ IS LIPSCHITZ WILL MAKE THE EXISTENCE OF C_2 CLEAR; SEE (14) IN \cite{TaiXu2002}
+\item If we assume $f$ itself is Lipschitz then the existence of $C_2$ is clear \cite{TaiXu2002}.
+However, our bound for the convergence rate for Algorithm \ref{alg:basiccd} is improved when $C_1,C_2$ are made smaller.
+\item Tai \cite{Tai2003} gives values of the constants $C_1$ in \eqref{as:lipschitzrestrictions} for obstacle problem constraint decompositions using $P_1$ FE spaces over shape-regular and quasi-uniform triangulations.
+In particular, constants $C_1$ are known for overlapping domain decompositions (with or without an additional coarse mesh) and standard multilevel hierarchies.
+Thus the cases we will demonstrate in Section \ref{sec:results} are already covered.
+\item The constants $C_1,C_2$ generally depend on $m$.
+However, regarding the convergence proof below, \cite{Tai2003} observes that if repeated application of Algorithm \ref{alg:basiccd} generates a bounded sequence of iterates in $\mathcal{K}$ then constants $C_1,C_2$ would be permitted to depend on $u,v,w_{ij},v_i,y_i$ as well.
 \end{itemize}
 
-From the assumptions and the above definitions we have the following core lemma for the parallel version.  Our proof follows \cite{Tai2003}.
+From the above definitions and assumptions we have the following estimate for the parallel version of Algorithm \ref{alg:basiccd}.  Our proof follows \cite{Tai2003}.
 
-\begin{lemma} \label{lem:core}  Suppose $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.
+\begin{lemma} \label{lem:core}  Suppose $f$ is $2$-coercive and $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.
+Recall $E(v,u^*)$ is defined in \eqref{eq:normlikedefn}.
 Suppose $\mathcal{V}_i$, $\mathcal{K}_i$, $R_i$, and $f$ satisfy \eqref{as:lipschitzrestrictions} and \eqref{as:lipschitzresidual}.
-Suppose $\hat w$ is computed from $u$ by the parallel version of Algorithm \ref{alg:basiccd}.
-Recall $e_i = \hat w_i - R_i u$.
+Suppose $\hat w$ is computed from $u$ by the parallel version of Algorithm \ref{alg:basiccd}, and recall $e_i = \hat w_i - R_i u$.
 Then
 \begin{equation}
-   \ip{f(\hat w)}{\hat w - u^*} \le C_2 \sum_{i=0}^{m-1} \|e_i\|^2 + C_1 C_2 \left(\sum_{i=0}^{m-1} \|e_i\|^2\right)^{1/2} \|u-u^*\|. \label{eq:core}
+   E(\hat w,u^*) \le C_2 \sum_{i=0}^{m-1} \|e_i\|^2 + \kappa^{-1} C_1 C_2 \left(\sum_{i=0}^{m-1} \|e_i\|^2\right)^{1/2} E(u,u^*)^{1/2}. \label{eq:core}
 \end{equation}
@@
 
-\begin{proof}
-For $i\in \{0,1,\dots,m-1\}$ use increasing, but wrapped, indices to define the following sequence $\phi_{ij}=\phi_{i,j} \in \cK$:
-\begin{align*}
-\phi_{i,0} &= u + e_i, \\
-\phi_{i,1} &= u + e_i + e_{i+1}, \\
-  &\vdots \\
-\phi_{i,m-1} &= u + e_i + e_{i+1} + \dots + e_{m-1} + e_0 + \dots + e_{i-1}.
-\end{align*}
-(Each $\phi_{ij} \in \cK$ by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum}.)  Noting that the boxed VI in the parallel version of Algorithm \ref{alg:basiccd} can be written
-
-    FIXME
+\begin{proof}  Recalling that $\hat w = u^* + \sum_i \hat w_i - R_i u^*$, by the (parallel version) boxed VI in Algorithm \ref{alg:basiccd} with $v_i = R_i u^*$ we get
+\begin{align}
+E(\hat w,u^*) &= \sum_{i=0}^{m-1} \ip{f(\hat w)}{\hat w_i - R_i u^*} - g(\hat w_i - R_i u^*) \label{eq:startcore} \\
+    &\le \sum_{i=0}^{m-1} \ip{f(\hat w)}{\hat w_i - R_i u^*} + \ip{f(u + e_i)}{R_i u^* - \hat w_i} \notag \\
+    &= \sum_{i=0}^{m-1} \ip{f(\hat w) - f(u + e_i)}{\hat w_i - R_i u^*}. \notag
+\end{align}
+For $i\in \{0,1,\dots,m-1\}$ use wrapped indices to define elements $\phi_{i,j}$:
+\begin{align}
+\phi_{i,0} &= u + e_i, \label{eq:gridcore} \\
+\phi_{i,1} &= u + e_i + e_{i+1}, \notag \\
+  &\vdots \notag \\
+\phi_{i,m-1} &= u + e_i + e_{i+1} + \dots + e_{m-1} + e_0 + \dots + e_{i-1}, \notag
+\end{align}
+and $\phi_{i,j} \in \cK$ by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum}.
+Note that $\phi_{i,m-1} = u + \sum_i e_i = \hat w$.
+From \eqref{eq:startcore}, apply a telescoping sum using the elements in \eqref{eq:gridcore}:
+\begin{align}
+E(\hat w,u^*) &\le \sum_{i=0}^{m-1} \sum_{j=1}^{m-1} \ip{f(\phi_{i,j}) - f(\phi_{i,j-1})}{\hat w_i - R_i u^*} \label{eq:nextcore} \\
+  &\le \sum_{i=0}^{m-1} \sum_{j=0}^{m-1} \left|\ip{f(\phi_{i,j}) - f(\phi_{i,j-1})}{\hat w_i - R_i u^*}\right|. \notag
+\end{align}
+By an easy renumbering of the indices $i,j$ we may apply assumption \eqref{as:lipschitzresidual}, then the triangle inequality, and then assumption \eqref{as:lipschitzrestrictions}; we denote $\left(\sum_{i=0}^{m-1} \|e_i\|^2\right)^{1/2}$ by $Z$:
+\begin{align}
+E(\hat w,u^*) &\le C_2 Z \left(\sum_{i=0}^{m-1} \|\hat w_i - R_i u^*\|^2\right)^{1/2} = C_2 Z \left(\sum_{i=0}^{m-1} \|\hat w_i - R_i u + R_i u - R_i u^*\|^2\right)^{1/2} \label{eq:nextnextcore} \\
+  &\le C_2 Z \left(Z + \left(\sum_{i=0}^{m-1} \|R_i u - R_i u^*\|^2\right)^{1/2}\right) \le C_2 Z \left(Z + C_1 \|u-u^*\|\right). \notag
+\end{align}
+Finally apply \eqref{eq:normlikebound} to give \eqref{eq:core}.
 \end{proof}
 
-Note that the coercivity of $f$ is not used in proving Lemma \ref{lem:core}.
+FIXME NOW THE ENTIRE BATTLE IS TO GET A CONVEXITY RESULT $E(w,u) \ge C \sum_{j=0}^{m-1} \|e_j\|^2$ for some reasonable $C$
 
@@
 
-Until now the algorithms can be interpreted as infinite-dimensional, e.g.~supposing constraint sets as defined in Examples \ref{ex:domaindecomposition} or \ref{ex:frequencydecomposition}.
-However, in practice we will solve VI \eqref{eq:vi} over a mesh and the space $\cV$ will be a finite element (FE) space.
-Thus $\cV$, and each subspace $\cV_i$, will be finite-dimensional.
+In practice we will solve VI \eqref{eq:vi} over a finite-dimensional space $\cV$ based on a choice of a mesh over $\Omega$ and a finite element (FE) space.
 

### 2022-01-17 paper/mcd2.tex
Edits.  This is a good under-construction paper.

@@
 
-The goal of this paper is to generalize the constraint decomposition (CD) method of X.-C.~Tai \cite{Tai2003}.
+The goal of this paper is to generalize the constraint decomposition (CD) method of X.-C.~Tai \cite{Tai2003} to nonlinear VI problems to which it has not been previously applied.
 The convergence of this method, for several types of decompositions, was proven by Tai for coercive variational inequality (VI) problems which arise from minimization of a convex functional over a convex set.
 In its multilevel form the method has been shown to have optimal complexity for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
+We construct a new proof of convergence which removes the assumption that the problem arises from (constrained) optimization.
+Furthermore we address its finite element (FE) implementation in nonlinear cases, and we demonstrate the resulting performance.
 
-We extend the CD method to nonlinear VI problems for which it has not been previously considered, including a new proof of convergence which removes the assumption that the problem arises from (constrained) optimization.
-Furthermore we address new finite element (FE) implementation concerns in nonlinear cases, and we demonstrate the resulting performance.
-The original algorithms and theory presented in \cite{Tai2003} are extended in four particular directions:
+The algorithms and theory presented in \cite{Tai2003} are extended in four particular directions:
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
@@
 \item We do not assume that the continuum VI problem arises from optimization of a scalar objective.
-Nonetheless we can prove convergence in norm at the same rate as the original method converges in energy (Section \ref{sec:convergence}); % HOPE
+Nonetheless we can prove convergence in $H^1$ norm at the same rate as the original method converges in energy (Section \ref{sec:convergence}); % HOPE
 examples are explored in Sections \ref{sec:vi} and \ref{sec:results}.
 
-\item We make the observation that, if implemented appropriately, multilevel V(0,$\nu$) ``up-slash'' cycles are intrinsically more efficient than the V($\nu$,0) ``down-slash'' cycles (Section \ref{sec:multilevel}).
-This observation seems to be new and should replace the default use of down-slash or even balanced (e.g.~V(1,1)) cycles; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}.
+\item We make the observation that, if implemented appropriately, multilevel ``up-smoothing'' is intrinsically more efficient than ``down-smoothing'' (Section \ref{sec:multilevel}).
+This observation seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}.
 A strong preference for up-smoothing is, apparently, special to multilevel CD methods and does not arise in corresponding unconstrained problems.
 
-\item In the nonlinear case, we make the multilevel CD algorithm more practical by addressing efficient storage of intermediate quantities.
-In particular, in Section \ref{sec:results} we show results from a full approximation storage (FAS; see \cite{Brandt1977}) implementation which avoids global (Newton) linearization; compare \cite{GraeserKornhuber2009}.
+\item We make the multilevel CD algorithm more practical by addressing efficient storage of intermediate quantities.
+In Section \ref{sec:results} we show results from a full approximation storage (FAS; see \cite{Brandt1977}) implementation for nonlinear operators, one which avoids global (Newton) linearization; compare \cite{GraeserKornhuber2009}.
 
-\item For problems of porous-media type the nonlinear operator $f$ is not coercive (Appendix A), so the existing convergence theory does not apply.
+\item For problems of porous-media type the nonlinear operator $f$ is not coercive (Appendix A), so the convergence theory in \emph{(i)} does not apply.
 However, a full-cycle implementation of the multilevel CD method is demonstrated numerically.
-The operator is approximated by a coercive operator, nonlinear in general, for the duration of the V-cycle, by ``freezing' the solution-dependent coefficient.
+By ``freezing' the solution-dependent coefficient, the operator is approximated by a coercive operator, nonlinear in general, for the duration of the V-cycle.
 The resulting scheme is highly-effective for doubly-nonlinear diffusion operators (Section \ref{sec:results}).
@@
 
-Note that the iterates from the CD method are always admissible, and thus the operator need only be defined on the admissible subset; Section \ref{sec:results} includes an example.
-Admissible-iterate methods should allow direct solutions of certain VI problems, such as fluid-layer dynamics problems \cite{Bueler2021conservation,JouvetBueler2012}, for which non-admissible methods, such as semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
-The full-cycle scheme in \emph{(iv)} above, which applies to non-coercive VIs, is helpful in these problems.
+The iterates from the CD method are always admissible, and thus the operator need only be defined for admissible states; Section \ref{sec:results} includes a nontrivial example.
+Admissible-iterate methods should permit direct solutions of certain VI problems, such as fluid-layer dynamics problems \cite{Bueler2021conservation,JouvetBueler2012}, for which non-admissible methods, such as semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
+The full-cycle scheme in \emph{(iv)} above, which applies to non-coercive VIs, is also designed for these problems, which are characterized by the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation.
 
-Especially for the classical obstacle problem with a Laplacian operator, certain other multilevel methods are known to improve performance relative to the multilevel CD method \cite{GraeserKornhuber2009}.
+Especially for the classical obstacle problem with a Laplacian operator, certain multilevel techniques are known to improve performance relative to the multilevel CD method \cite{GraeserKornhuber2009}.
 These improved methods either track the active set in the discretization or modify the nodal basis functions, and in this sense they are discrete algorithms.
-By contrast the CD method applies at the level of the continuum problem (Sections \ref{sec:cd} and \ref{sec:convergence}), and thus it is a more fundamental method than the truncated monotone multigrid method \cite{Kornhuber1994}, for example, and in any case we can demonstrate comparable rates of convergence (Section \ref{sec:results}).
-% HOPE
-Truncated monotone approaches for nonlinear problems have, so far, applied a global linearization strategy \cite{GraeserKornhuber2009}, while we apply FAS (Section \ref{sec:multilevel}).
-Acceleration of our nonlinear multilevel CD method via active-set and/or basis-level manipulations is a topic for further research.
+By contrast the CD method applies at the level of the continuum problem (Sections \ref{sec:cd} and \ref{sec:convergence}), and thus it is a more fundamental method than the truncated monotone multigrid method \cite{Kornhuber1994}, for example.
+In any case, acceleration of our nonlinear multilevel CD method via active-set and/or basis-level manipulations represents a potential extension of the method here, though as a topic for future research.
 
@@
 Let $\cK \subset \cV$ be a nonempty closed and convex subset, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.
-For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and \emph{source} $g\in \cV'$ we consider the following \emph{variational inequality} (VI) for the (exact) solution $u^*\in \cK$:
+For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and \emph{source} $g\in \cV'$ we consider the following \emph{variational inequality} (VI) for the (exact) solution $u^*\in \cK$, if it exists:
 \begin{equation}
@@
 \end{equation}
-Because $f$ is a (generally) nonlinear map, $g$ is not strictly needed, but its presence is helpful to the algorithms of Section \ref{sec:multilevel}.  (By redefining $f$ we may take $g=0$.)
+Because $f$ is a (generally) nonlinear map, $g$ is not strictly needed when posing this VI, but its presence is helpful to the algorithms of Section \ref{sec:multilevel}.
+(Note that by redefining $f$ we may take $g=0$.)
 
@@
 Informally, if we conceptualize the dual pairing as an inner product then \eqref{eq:vi} says that the angle between $f(u^*)-g$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
-That is, \eqref{eq:vi} says that $f(u^*)-g$ points directly into $\cK$ if $u^* \in\partial\cK$.
-Furthermore, if $u^* \in \cK^\circ$ (interior) then \eqref{eq:vi} implies $f(u^*)=g$.
-
-The following definitions are standard \cite{KinderlehrerStampacchia1980}.
+That is, \eqref{eq:vi} says that $f(u^*)-g$ may not be zero but it points directly into $\cK$.
+In particular, if $u^* \in \cK^\circ$ (interior) then \eqref{eq:vi} implies $f(u^*)=g$.
+The following definitions are also standard \cite{KinderlehrerStampacchia1980}.
 
@@
 \end{equation}
-We say VI \eqref{eq:vi} is (\emph{strictly}) \emph{monotone} if $f$ is (strictly) monotone, and likewise for \emph{coercive}.
-\end{definition}
+We say VI \eqref{eq:vi} is \emph{monotone} if $f$ is monotone, and likewise for strictly monotone and coercive. \end{definition}
 
@@
 As in the calculus of variations \cite{Evans2010}, coercivity permits a compactness argument for unbounded sets $\cK$; recall that the bounded, closed subsets of a reflexive Banach space are weakly compact.
-Note that the condition of continuity can be weakened to only apply on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but the stronger condition will apply in our examples.
+The condition of continuity can be weakened to only apply on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but the stronger condition will apply in our examples.
 
-The coercive VIs solved in this paper satisfy a stronger inequality than \eqref{eq:coercive}.
+The coercive VIs solved in this paper satisfy a stronger inequality than \eqref{eq:coercive}, and thus they are well-posed.
 
@@
 \end{equation}
-(Note \cite{Tai2003} uses ``coercive'' for $2$-coercive.)
 \end{definition}
 
-It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive, and thus the following result holds.
-It states that the $p$-coercive VI problems considered in this paper are well-posed.
+Note \cite{Tai2003} uses ``coercive'' for $2$-coercive.  It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive, and thus the following result holds.
 
@@
 
-When $f$ is monotone VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
+When $f$ is monotone, VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
 Suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cK \to \cV'$.
 Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
-Furthermore, \cite[Proposition II.2.1]{EkelandTemam1976} shows that if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ and $g=0$ if and only if
+Furthermore, Proposition II.2.1 in \cite{EkelandTemam1976} shows that if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ and $g=0$ if and only if
 \begin{equation}
@@
 
+From now on $\Omega \subset \RR^d$ denotes a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.
+Also, Sobolev spaces \cite{Evans2010} are denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$, with $W^{k,2}$ denoted by $H^k$.
+
 The following example includes the classical obstacle problem for the linear Laplacian \cite{GraeserKornhuber2009} and the obstacle problem for the $p$-Laplacian \cite{ChoeLewis1991}.
-Note that from now on $\Omega \subset \RR^d$ denotes a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.
-Also, Sobolev spaces \cite{Evans2010} are denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$, with $W^{k,2}$ denoted by $H^k$.
 
@@
 for some $C>0$, and thus $f$ is $p$-coercive.
-(The map in \eqref{eq:plaplacian} is coercive if $1<p<2$, but the proof is somewhat different \cite[Theorem 4.4]{Bueler2021conservation}, and in Section \ref{sec:results} we need only the $p\ge 2$ case.)  For $g\in\cV'$ define
+(The map in \eqref{eq:plaplacian} is coercive if $1<p<2$, but the proof is somewhat different \cite[Theorem 4.4]{Bueler2021conservation}.
+In Section \ref{sec:results} we need only the $p\ge 2$ case.)  For $g\in\cV'$ define
     $$F(u) = \int_\Omega \frac{a(x)}{p} |\grad u|^p\,dx - g(u).$$

### 2022-07-09 paper/mcd2.tex
I do not think I can make the appendix claim (porous-medium never coercive) work

@@
 
-\small
 \bibliography{mcd2}
@@
 
-\normalsize
-\appendix
-
-\section{Porous-medium problems are never monotone}
-
-Based on the examples in Section \ref{sec:vi}, the property of $p$-coercivity would seem to be a mild generalization of uniform ellipticity (including its quasi-linear meaning \cite[Section 8.3]{Evans2010}).
-Indeed, weak-form operators with certain lower-order terms (Example \ref{ex:advectiondiffusion}) or degenerate quasi-linear type (Example \ref{ex:plaplacian}) can be monotone and coercive.
-However, even the weaker property of monotonicity is ``fragile'' in a pointwise sense, as explained next, while ellipticity is more robust.
-In particular, we show that no nontrivial porous-medium type functional is monotone, nor $p$-coercive.
-We have not found any such explanation in the literature, but the observation of non-monotonicity in \cite{RocknerWang2008} is related.
-
-Consider the expression $(Ax - By)\cdot (x-y)$ where $A,B \in \RR^{d\times d}$ are symmetric matrices and $x,y\in \RR^d$.
-This expression is the quadratic form for a certain symmetric block matrix.
-In fact, regarding $x,y$ as column vectors, we observe that
-\begin{equation}
-(Ax - By)\cdot (x-y) = \begin{bmatrix} x^\top & y^\top \end{bmatrix} \begin{bmatrix}
-                        A & -\frac{1}{2}(A+B) \\
-                        -\frac{1}{2}(A+B) & B \end{bmatrix} \begin{bmatrix} x \\ y \end{bmatrix}. \label{eq:blockmatrix}
-\end{equation}
-Let $M_{A,B}$ denote the symmetric $2d \times 2d$ matrix in \eqref{eq:blockmatrix}.
-Clearly the expression is only bounded below if $M_{A,B}$ is nonnegative definite.
-Stated as the following Lemma, the important point here is that exact equality $A=B$ is required for nonnegativity of the expression; knowing $A$ is close to $B$ is not helpful.
-
-\begin{lemma}  Suppose $A,B \in \RR^{d\times d}$ are symmetric matrices which commute.
-Then $(Ax - By)\cdot (x-y)\ge 0$ for all $x,y\in \RR^d$ if and only if $A=B$ is nonnegative definite.
-\end{lemma}
-
-\newcommand{\sbvec}[2]{\left[\begin{smallmatrix} #1 \\ #2 \end{smallmatrix}\right]}
-\newcommand{\sbmat}[4]{\left[\begin{smallmatrix} #1 & #2 \\ #3 & #4 \end{smallmatrix}\right]}
-
-\begin{proof}
-If $A=B$ is nonnegative definite then nonnegativity of the expression follows immediately.
-Conversely, since $A$ is symmetric there exists an orthogonal basis $\{x_i\}$ of eigenvectors, with $Ax_i = \lambda_i x$.
-By commutativity these are also eigenvectors of $B$, with $Bx_i = \mu_i x_i$.
-Let $\big\{\sbvec{x_1}{x_1}$, $\sbvec{x_1}{-x_1}$, $\dots$, $\sbvec{x_d}{x_d}$, $\sbvec{x_d}{-x_d}\big\}$ be an ordered basis of $\RR^{2d}$.
-It is easy to confirm that in this basis $M_{A,B}$ is block diagonal with $2\times 2$ diagonal blocks $\sbmat{0}{-\frac{1}{2}(\lambda_i+\mu_i)}{-\frac{1}{2}(\lambda_i+\mu_i)}{\lambda_i+\mu_i}$.
-Thus the eigenvalues of $M_{A,B}$ come in pairs $\frac{1}{2}(\lambda_i + \mu_i) \pm \sqrt{\frac{1}{2}(\lambda_i^2 + \mu_i^2)}$ for $i=1,\dots,d$.
-By the strict concavity of the square root, $\sqrt{\frac{1}{2}(\lambda_i^2 + \mu_i^2)} > \frac{1}{2}(|\lambda_i| + |\mu_i|)$ if $\lambda_i\ne \mu_i$.
-Thus if any eigenvalues of $A$ differ from those of $B$, i.e.~$\lambda_i\ne \mu_i$ for any $i$, then there exists a negative eigenvalue of $M_{A,B}$.
-\end{proof}
-
-It is not clear if the hypotheses of symmetry and commutativity can be removed.
-However, when the matrices are scalar multiples of the identity then the Lemma asserts that $(ax-by)\cdot(x-y) \ge 0$ for all $x,y\in\RR^d$ if and only if $a=b\ge 0$.
-This scalar case suffices for the following result.
-
-\begin{theorem}
-Suppose $\phi:[0,\infty) \to (0,\infty)$ is continuous.
-For $\Omega \subset \RR^d$ open and nonempty, $\cV = H_0^1(\Omega)$, and $g\in \cV'$, consider the following functional on $\cK = \{v\ge 0\} \subset \cV$:
-\begin{equation}
-\ip{f(u)}{v} = \int_\Omega \phi(u) \grad u\cdot \grad v\,dx - g(v).  \label{eq:porousagain}
-\end{equation}
-If $\phi$ is not constant then $f$ is not monotone.
-\end{theorem}
-
-\begin{proof}
-Suppose $\phi(\alpha)\ne \phi(\beta)$ for $\alpha > 0$ and $\beta \ge 0$.
-(That is, suppose $\phi$ is not constant.)  For $x\in\Omega$, construct $u\in \cK$ smooth so that $u(x)=\alpha$ and $\xi = \grad u(x) \ne 0$.
-Using the Lemma, choose $\eta \in \RR^d$ so that $(\phi(\alpha) \xi - \phi(\beta) \eta) \cdot (\xi - \eta) < 0$.
-FIXME NOT SURE IF THIS IS GOING TO WORK
-\end{proof}
-
-The main idea is that even the slightest variation in the coefficient $\phi(u)$ violates monotonicity.
-A strong and tight ellipticity hypothesis, e.g.~the existence of $c_0>0$ and $\eps>0$ so that $c_0 \le \phi(u) \le (1+\eps)c_0$, will not imply monotonicity.
-Contrast the nontrivial coefficient dependence on $\grad u$ in the $p$-coercive $p$-Laplacian Example \ref{ex:plaplacian} (with $p>2)$.
-
-No nontrivial functional \eqref{eq:porousagain} is monotone, nor is it ever $p$-coercive for any $p>1$.
 \end{document}

### 2022-07-09 paper/old/porous-not-coercive.tex
I do not think I can make the appendix claim (porous-medium never coercive) work

\documentclass[letterpaper,final,12pt,reqno]{amsart}

\usepackage[total={6.3in,9.2in},top=1.1in,left=1.1in]{geometry}

\usepackage{times,bm,bbm,empheq,fancyvrb,graphicx,amsthm,amssymb}
\usepackage[dvipsnames]{xcolor}
\usepackage{longtable}
\usepackage{booktabs}

\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing}

\usepackage[kw]{pseudo}
\pseudoset{left-margin=15mm,topsep=5mm,idfont=\texttt}

\usepackage{float}

% hyperref should be the last package we load
\usepackage[pdftex,
colorlinks=true,
plainpages=false, % only if colorlinks=true
linkcolor=blue,   % ...
citecolor=Red,    % ...
urlcolor=black    % ...
]{hyperref}

\renewcommand{\baselinestretch}{1.05}

\allowdisplaybreaks[1]  % allow display breaks in align environments, if they avoid major underfulls

\newtheoremstyle{cstyle}% name
  {5pt}% space above
  {5pt}% space below
  {\itshape}% body font
  {}% indent amount
  {\itshape}% theorem head font
  {.}% punctuation after theorem head
  {.5em}% space after theorem head
  {\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}% theorem head spec
\theoremstyle{cstyle}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{assumptions}[theorem]{Assumptions}

\newtheoremstyle{cstyle*}% name
  {5pt}% space above
  {5pt}% space below
  {\itshape}% body font
  {}% indent amount
  {\itshape}% theorem head font
  {.}% punctuation after theorem head
  {.5em}% space after theorem head
  {\thmname{#1}}% theorem head spec
\theoremstyle{cstyle*}
\newtheorem{assumptions*}{Assumptions}

\newtheoremstyle{dstyle}% name
  {5pt}% space above
  {5pt}% space below
  {}%{\itshape}% body font
  {}% indent amount
  {\itshape}% theorem head font
  {.}% punctuation after theorem head
  {.5em}% space after theorem head
  {\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}% theorem head spec
\theoremstyle{dstyle}

\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

% numbering
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\numberwithin{table}{section}
\numberwithin{theorem}{section}

\newcommand{\eps}{\epsilon}
\newcommand{\RR}{\mathbb{R}}

\newcommand{\grad}{\nabla}
\newcommand{\Div}{\nabla\cdot}
\newcommand{\trace}{\operatorname{tr}}

\newcommand{\hbn}{\hat{\mathbf{n}}}

\newcommand{\bb}{\mathbf{b}}
\newcommand{\be}{\mathbf{e}}
\newcommand{\bbf}{\mathbf{f}}
\newcommand{\bg}{\mathbf{g}}
\newcommand{\bn}{\mathbf{n}}
\newcommand{\br}{\mathbf{r}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\bv}{\mathbf{v}}
\newcommand{\bw}{\mathbf{w}}
\newcommand{\bx}{\mathbf{x}}
\newcommand{\bF}{\mathbf{F}}
\newcommand{\bV}{\mathbf{V}}
\newcommand{\bX}{\mathbf{X}}
\newcommand{\bxi}{\bm{\xi}}
\newcommand{\bzero}{\bm{0}}

\newcommand{\cK}{\mathcal{K}}
\newcommand{\cV}{\mathcal{V}}

\newcommand{\rhoi}{\rho_{\text{i}}}

\newcommand{\ip}[2]{\left<#1,#2\right>}

\newcommand{\mR}{R^{\bm{\oplus}}}
\newcommand{\iR}{R^{\bullet}}

\newcommand{\nn}{{\text{n}}}
\newcommand{\pp}{{\text{p}}}
\newcommand{\qq}{{\text{q}}}
\newcommand{\rr}{{\text{r}}}

\newcommand{\supp}{\operatorname{supp}}
\newcommand{\Span}{\operatorname{span}}

\begin{document}
\appendix

\begin{quote}
\emph{The Lemmas in this former appendix are correct, but I do not actually see how to prove the claimed Theorem.}
\end{quote}

\section{Porous-medium problems are never monotone}

The property of $p$-coercivity would seem to be a mild generalization of uniform ellipticity, if we include its quasi-linear meaning \cite[Section 8.3]{Evans2010}, and weak-form operators with degenerate quasi-linear type (e.g.~$p$-Laplacian) or certain lower-order terms (e.g.~advection-diffusion) can be monotone and coercive.  However, even the weaker property of monotonicity is apparently ``fragile'', as explained next, while ellipticity is more robust.  We attempt to show that no nontrivial porous-medium type functional is monotone, nor $p$-coercive.  We have not found any such explanation in the literature, but the observation of non-monotonicity in \cite{RocknerWang2008} is related.

\begin{lemma}  Suppose $a,b$ are real numbers.  Then $(ax-by)\cdot(x-y) \ge 0$ for all $x,y\in\RR^d$ if and only if $a=b\ge 0$.
\end{lemma}

\begin{proof}
If $a=b\ge 0$ then $(ax-by)\cdot(x-y) = a\|x-y\|^2 \ge 0$.

Conversely, if $a< 0$ then choose $y=0$ and $x\ne 0$ to get $(ax-by)\cdot(x-y) = a\|x\|^2<0$, a negative value from the form.  A similar strategy gives a negative value if $b<0$.  So suppose $b > 0$, $a\ge 0$, and $a\ne b$.  Consider $y=mx$ for $m$ real and define the convex quadratic scalar function $f(m) = (a-bm)(1-m)$ so that $(ax-by)\cdot(x-y) = f(m) \|x\|^2$.  We minimize $f$, giving $f(m^*) = -b(a-b)^2/4 < 0$ for $m^*=(a+b)/(2b)$.  Thus there exists $x,y$ so that $(ax-by)\cdot(x-y)<0$.  A similar argument applies if $b=0$ and $a>0$.
\end{proof}

A version of the same idea applies if $a,b$ are replaced by matrices.  Consider the expression $(Ax - By)\cdot (x-y)$ where $A,B \in \RR^{d\times d}$ are symmetric matrices and $x,y\in \RR^d$.  This expression is the quadratic form for a certain symmetric block matrix.  Regarding $x,y$ as column vectors:
\begin{equation}
(Ax - By)\cdot (x-y) = \begin{bmatrix} x^\top & y^\top \end{bmatrix} M_{A,B} \begin{bmatrix} x \\ y \end{bmatrix} \label{eq:blockmatrix}
\end{equation}
where
	$$M_{A,B} = \begin{bmatrix} A & -\frac{1}{2}(A+B) \\
                                -\frac{1}{2}(A+B) & B \end{bmatrix}$$
is a symmetric $2d \times 2d$ matrix.  Clearly \eqref{eq:blockmatrix} is only bounded below if $M_{A,B}$ is nonnegative definite.  Stated as the following Lemma, the important point here is that exact equality $A=B$ is required for nonnegativity of the expression; knowing $A$ is close to $B$ is not helpful.

\begin{lemma}  Suppose $A,B \in \RR^{d\times d}$ are symmetric matrices which commute.  Then $(Ax - By)\cdot (x-y)\ge 0$ for all $x,y\in \RR^d$ if and only if $A=B$ and $A$ is nonnegative definite. \end{lemma}

\newcommand{\sbvec}[2]{\left[\begin{smallmatrix} #1 \\ #2 \end{smallmatrix}\right]}
\newcommand{\sbmat}[4]{\left[\begin{matrix} #1 & #2 \\ #3 & #4 \end{matrix}\right]}

\begin{proof}
If $A=B$ is nonnegative definite then $(Ax - By)\cdot (x-y) = z^\top A z\ge 0$ where $z=x-y$.

Conversely, since $A$ is symmetric there exists an orthogonal basis $\{x_i\}$ of eigenvectors, with $Ax_i = \lambda_i x$.  By commutativity these are also eigenvectors of $B$, with $Bx_i = \mu_i x_i$.  Note $\lambda_i,\mu_i$ are real.

Let $\big\{\sbvec{x_1}{x_1}$, $\sbvec{x_1}{-x_1}$, $\dots$, $\sbvec{x_d}{x_d}$, $\sbvec{x_d}{-x_d}\big\}$ be an ordered basis of $\RR^{2d}$.  In this basis $M_{A,B}$ is block diagonal with symmetric $2\times 2$ diagonal blocks
   $$\sbmat{0}{\frac{1}{2}(\lambda_i-\mu_i)}{\frac{1}{2}(\lambda_i-\mu_i)}{\lambda_i+\mu_i}.$$
Thus the eigenvalues of $M_{A,B}$ come in pairs $\frac{1}{2}(\lambda_i + \mu_i) \pm \sqrt{\frac{1}{2}(\lambda_i^2 + \mu_i^2)}$ for $i=1,\dots,d$.  By the strict concavity of the square root, $\sqrt{\frac{1}{2}(\lambda_i^2 + \mu_i^2)} > \frac{1}{2}(|\lambda_i| + |\mu_i|)$ if $\lambda_i\ne \mu_i$.  Thus if any eigenvalues of $A$ differ from those of $B$, i.e.~$\lambda_i\ne \mu_i$ for any $i$, then there exists a negative eigenvalue of $M_{A,B}$.
\end{proof}

It is not clear if the hypotheses of symmetry and commutativity can be removed.

I DO NOT KNOW HOW TO PROVE THE FOLLOWING CLAIM:

\begin{theorem}
Suppose $\phi:[0,\infty) \to (0,\infty)$ is continuous.  For $\Omega \subset \RR^d$ open and nonempty, $\cV = H_0^1(\Omega)$, and $g\in \cV'$, consider the following functional on $\cK = \{v\ge 0\} \subset \cV$:
\begin{equation}
\ip{f(u)}{v} = \int_\Omega \phi(u) \grad u\cdot \grad v\,dx - \ip{g}{v}.  \label{eq:porousagain}
\end{equation}
If $\phi$ is not constant then $f$ is not monotone.
\end{theorem}

\begin{proof}
Suppose $\phi(\alpha)\ne \phi(\beta)$ for $\alpha > 0$ and $\beta \ge 0$.  (That is, suppose $\phi$ is not constant.)  For $x\in\Omega$, construct $u\in \cK$ smooth so that $u(x)=\alpha$ and $\xi = \grad u(x) \ne 0$.  Using the first Lemma, choose $\eta \in \RR^d$ so that $(\phi(\alpha) \xi - \phi(\beta) \eta) \cdot (\xi - \eta) < 0$.  FIXME NOT SURE IF THIS IS GOING TO WORK
\end{proof}

The main idea is that even the slightest variation in the coefficient $\phi(u)$ violates monotonicity.  A strong and tight ellipticity hypothesis, e.g.~the existence of $c_0>0$ and $\eps>0$ so that $c_0 \le \phi(u) \le (1+\eps)c_0$, will not imply monotonicity.  Contrast the nontrivial coefficient dependence on $\grad u$ in the $p$-coercive $p$-Laplacian example.

The above Theorem, if true, shows that no nontrivial functional \eqref{eq:porousagain} is monotone, nor is it ever $p$-coercive for any $p>1$.

\bibliography{../mcd2}
\bibliographystyle{siam}

\end{document}

### 2022-07-09 paper/mcd2.tex
edits on first three pages

@@
 
-The goal of this paper is to generalize the constraint decomposition (CD) method of X.-C.~Tai \cite{Tai2003} to nonlinear VI problems to which it has not been previously applied.
+The goal of this paper is to extend the constraint decomposition (CD) method of X.-C.~Tai \cite{Tai2003} to nonlinear VI problems to which it has not been applied.
 The convergence of this method, for several types of decompositions, was proven by Tai for coercive variational inequality (VI) problems which arise from minimization of a convex functional over a convex set.
 In its multilevel form the method has been shown to have optimal complexity for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
-We construct a new proof of convergence which removes the assumption that the problem arises from (constrained) optimization.
-Furthermore we address its finite element (FE) implementation in nonlinear cases, and we demonstrate the resulting performance.
 
-The algorithms and theory presented in \cite{Tai2003} are extended in four particular directions:
+We construct a new proof of convergence which removes the assumption that the problem arises from optimization.
+Furthermore we address finite element (FE) implementation in nonlinear cases, and we demonstrate the resulting performance.
+In fact, the theory presented in \cite{Tai2003} is extended in four particular directions:
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
 \begin{enumerate}
-\item We do not assume that the continuum VI problem arises from optimization of a scalar objective.
-Nonetheless we can prove convergence in $H^1$ norm at the same rate as the original method converges in energy (Section \ref{sec:convergence}); % HOPE
-examples are explored in Sections \ref{sec:vi} and \ref{sec:results}.
+\item We do not assume that the continuum VI problem arises from optimization of a scalar objective; examples are explored in Sections \ref{sec:vi} and \ref{sec:results}.
+We prove convergence in $H^1$ norm at the same rate as the original method converges in energy (Section \ref{sec:convergence}).
+% THIS IS THE HOPE
 
@@
 
-\item We make the multilevel CD algorithm more practical by addressing efficient storage of intermediate quantities.
+\item We make the multilevel CD algorithm more practical by addressing storage of intermediate quantities.
 In Section \ref{sec:results} we show results from a full approximation storage (FAS; see \cite{Brandt1977}) implementation for nonlinear operators, one which avoids global (Newton) linearization; compare \cite{GraeserKornhuber2009}.
 
-\item For problems of porous-media type the nonlinear operator $f$ is not coercive (Appendix A), so the convergence theory in \emph{(i)} does not apply.
+\item For problems of porous-media type the nonlinear operator $f$ is not known to be coercive, so the convergence theory in \emph{(i)} does not apply.
 However, a full-cycle implementation of the multilevel CD method is demonstrated numerically.
@@
 
-The iterates from the CD method are always admissible, and thus the operator need only be defined for admissible states; Section \ref{sec:results} includes a nontrivial example.
-Admissible-iterate methods should permit direct solutions of certain VI problems, such as fluid-layer dynamics problems \cite{Bueler2021conservation,JouvetBueler2012}, for which non-admissible methods, such as semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
-The full-cycle scheme in \emph{(iv)} above, which applies to non-coercive VIs, is also designed for these problems, which are characterized by the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation.
+The iterates from a CD method are always admissible, and thus the operator need only be defined for admissible states; Section \ref{sec:results} includes a nontrivial example.
+Admissible-iterate methods should permit direct solutions of certain VI problems, such as fluid-layer dynamics problems \cite{Bueler2021conservation,JouvetBueler2012}, for which non-admissible methods, such as semi-smooth methods \cite{BensonMunson2006}, at least require unnatural modifications of the operator formula.
+The full-cycle scheme in \emph{(iv)} above is also designed for these problems, which are characterized by the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation.
 
-Especially for the classical obstacle problem with a Laplacian operator, certain multilevel techniques are known to improve performance relative to the multilevel CD method \cite{GraeserKornhuber2009}.
+For the classical obstacle problem with a Laplacian operator, certain multilevel techniques are known to improve performance relative to the multilevel CD method \cite{GraeserKornhuber2009}.
 These improved methods either track the active set in the discretization or modify the nodal basis functions, and in this sense they are discrete algorithms.
-By contrast the CD method applies at the level of the continuum problem (Sections \ref{sec:cd} and \ref{sec:convergence}), and thus it is a more fundamental method than the truncated monotone multigrid method \cite{Kornhuber1994}, for example.
-In any case, acceleration of our nonlinear multilevel CD method via active-set and/or basis-level manipulations represents a potential extension of the method here, though as a topic for future research.
+By contrast the CD method applies at the level of the continuum problem (Sections \ref{sec:cd} and \ref{sec:convergence}); compare the truncated monotone multigrid method \cite{Kornhuber1994}, for example.
+Acceleration of our nonlinear multilevel CD method via active-set and/or basis-level manipulations represents a potential extension of the method here, and is a topic for future research.
 
@@
 Let $\cK \subset \cV$ be a nonempty closed and convex subset, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.
-For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and \emph{source} $g\in \cV'$ we consider the following \emph{variational inequality} (VI) for the (exact) solution $u^*\in \cK$, if it exists:
+For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and \emph{source functional} $g\in \cV'$ we consider the following \emph{variational inequality} (VI) for the (exact) solution $u^*\in \cK$, if it exists:
 \begin{equation}
@@
 \end{equation}
-Because $f$ is a (generally) nonlinear map, $g$ is not strictly needed when posing this VI, but its presence is helpful to the algorithms of Section \ref{sec:multilevel}.
-(Note that by redefining $f$ we may take $g=0$.)
+Because $f$ is (generally) nonlinear, $g$ is not strictly needed for posing this VI, and by redefining $f$ we may take $g=0$, but the presence of $g$ is helpful to algorithms in Section \ref{sec:multilevel}.
 
-VI \eqref{eq:vi} generalizes the nonlinear system of equations $f(u^*)=g$ from a vector space $\cV$ to the constrained case (over $\cK$).
+VI \eqref{eq:vi} generalizes the nonlinear system of equations $f(u^*)=g$ to the constraint set $\cK$.
 Informally, if we conceptualize the dual pairing as an inner product then \eqref{eq:vi} says that the angle between $f(u^*)-g$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
 That is, \eqref{eq:vi} says that $f(u^*)-g$ may not be zero but it points directly into $\cK$.
-In particular, if $u^* \in \cK^\circ$ (interior) then \eqref{eq:vi} implies $f(u^*)=g$.
-The following definitions are also standard \cite{KinderlehrerStampacchia1980}.
+In particular, if $u^*$ is in the interior $\cK^\circ$ then \eqref{eq:vi} implies $f(u^*)=g$.
 
-\begin{definition} A map $f:\cK \to \cV'$ is \emph{monotone} if
+\begin{definition} The following definitions are standard \cite{KinderlehrerStampacchia1980}.  A map $f:\cK \to \cV'$ is \emph{monotone} if
 \begin{equation}
@@
 
-Note \cite{Tai2003} uses ``coercive'' for $2$-coercive.  It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive, and thus the following result holds.
+Note that Tai \cite{Tai2003} uses ``coercive'' for what we wil call $2$-coercive.
+It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive, and thus the following result holds.
 
-\begin{theorem}  \label{thm:viwellposed}  If $f:\cK \to \cV'$ is continuous and $p$-coercive (for $p>1$) then there exists a unique $u\in \cK$ solving VI \eqref{eq:vi}.
+\begin{theorem}  \label{thm:viwellposed}  If $f:\cK \to \cV'$ is continuous and $p$-coercive then there exists a unique $u^*\in \cK$ solving VI \eqref{eq:vi}.
 \end{theorem}
 
-When $f$ is monotone, VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
+When $f$ is monotone, \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
 Suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cK \to \cV'$.
@@
 From now on $\Omega \subset \RR^d$ denotes a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.
-Also, Sobolev spaces \cite{Evans2010} are denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$, with $W^{k,2}$ denoted by $H^k$.
-
-The following example includes the classical obstacle problem for the linear Laplacian \cite{GraeserKornhuber2009} and the obstacle problem for the $p$-Laplacian \cite{ChoeLewis1991}.
+Sobolev spaces \cite{Evans2010} are denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$, with $W^{k,2}$ denoted $H^k$.
+The following example includes the classical obstacle problem for the linear Laplacian \cite{GraeserKornhuber2009} and the $p$-Laplacian for $p\ge 2$ \cite{ChoeLewis1991}.
 
@@
     $$\ip{f(u) - f(v)}{u-v} \ge 2^{2-p} a_0 \|\grad u - \grad v\|_p^p \ge 2^{2-p} a_0 C \|u-v\|^p$$
-for some $C>0$, and thus $f$ is $p$-coercive.
-(The map in \eqref{eq:plaplacian} is coercive if $1<p<2$, but the proof is somewhat different \cite[Theorem 4.4]{Bueler2021conservation}.
-In Section \ref{sec:results} we need only the $p\ge 2$ case.)  For $g\in\cV'$ define
-    $$F(u) = \int_\Omega \frac{a(x)}{p} |\grad u|^p\,dx - g(u).$$
-Then $F'(u) = f(u) - g$, $F$ is a convex functional (since $f$ is coercive), and, for any closed and convex $\cK\subset \cV$, VI problem \eqref{eq:vi} for is equivalent to optimization problem \eqref{eq:minimization}.
-\end{example}
+for some $C>0$, and thus $f$ is $p$-coercive.  On the other hand, for $g\in\cV'$ define
+    $$F(v) = \int_\Omega \frac{a(x)}{p} |\grad v|^p\,dx - \ip{g}{v}.$$
+Then $F'(v) = f(v) - g$, $F$ is a convex functional (since $f$ is coercive), and, for any closed and convex $\cK\subset \cV$, VI problem \eqref{eq:vi} for is equivalent to optimization problem \eqref{eq:minimization}.\end{example}
 
-However, not all VI problems arise from optimization.
+The map in \eqref{eq:plaplacian} is coercive if $1<p<2$, but the proof is somewhat different \cite[Theorem 4.4]{Bueler2021conservation}.  In Section \ref{sec:results} we need only the $p\ge 2$ case.
+
+Not all VI problems arise as in Example \ref{ex:plaplacian}, from optimization.
 We give two such examples next, first a coercive and linear advection-diffusion problem, and then a nonlinear porous-medium-type problem.

### 2022-07-15 paper/mcd2.tex
edits through page 6

@@
 Not all VI problems arise as in Example \ref{ex:plaplacian}, from optimization.
-We give two such examples next, first a coercive and linear advection-diffusion problem, and then a nonlinear porous-medium-type problem.
-Each of these examples is important in applications.
+We give two such examples next, first a coercive and linear advection-diffusion problem, and then a nonlinear porous-medium-type problem; each is important in applications.
 The first is preceded by a lemma.
 
-\begin{lemma}  \label{lem:advectionskew}  \cite{Elmanetal2014}\,  Suppose $\bX :\Omega \to \RR^d$ is a smooth and bounded vector field on $\Omega$ with zero divergence ($\Div \bX=0$).
+\begin{lemma}  \label{lem:advectionskew}  \cite{Elmanetal2014}\,  Suppose $\bX :\Omega \to \RR^d$ is a bounded and boundedly-differentiable vector field on $\Omega$ with zero divergence ($\Div \bX=0$).
 For $u,v \in H^1(\Omega)$ let $b(u,v) = \int_\Omega (\bX \cdot \grad u) v\,dx$.
@@
 \begin{example}  \label{ex:advectiondiffusion}  Suppose $\partial\Omega$ is partitioned into Dirichlet and Neumann portions, i.e.~$\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$, with $\partial_D\Omega$ of positive measure.
-Let $\cV = H_0^1(\Omega)$ be the space of functions with value zero on $\partial_D\Omega$.
-(For simplicity we consider zero boundary data.)  Consider a smooth and bounded velocity field $\bX$ on $\Omega$ such that $\Div \bX=0$, but additionally assume that the flow is outward on the Neumann boundary, i.e.~$\bX \cdot \bn \ge 0$ at points of $\partial_N\Omega$.
+Let $\cV = H_0^1(\Omega)$ be the space of functions which are zero along $\partial_D\Omega$.
+Consider a divergence-free velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, but additionally assume that the flow is outward on the Neumann boundary, $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.
 For $u,v \in \cV = H_0^1(\Omega)$ and $\eps>0$ define
 \begin{equation}
-\ip{f(u)}{v} = \int_\Omega \eps \grad u \cdot \grad v - (\bX \cdot \grad u) v\,dx. \label{eq:advectiondiffusion}
+\ip{f(u)}{v} = \eps \left(\grad u, \grad v\right)_{L^2(\Omega)} - b(u,v). \label{eq:advectiondiffusion}
 \end{equation}
-Consider VI \eqref{eq:vi} for any closed and convex $\cK \subset \cV$ and any $g\in\cV'$.
-It is easy to see that $|\ip{f(u)}{v}| \le (\eps + \|\bX\|_\infty) \|u\| \|v\|$, thus that $f:\cK \to \cV'$ is continuous (and Lipschitz).
-On the other hand, Lemma \ref{lem:advectionskew} says that $s(u,v)$ is skew-symmetric up to a nonnegative term.
-However, by the outward flow assumption and the Poincar\'e inequality,
+Consider VI \eqref{eq:vi} for any closed and convex $\cK \subset \cV$ and $g\in\cV'$.
+It is easy to see that $|\ip{f(u)}{v}| \le (\eps + \|\bX\|_\infty) \|u\| \|v\|$, thus that $f:\cK \to \cV'$ is continuous.
+Lemma \ref{lem:advectionskew} says that the bilinear form $s(u,v)$ is skew-symmetric up to a nonnegative term.
+By the outward flow assumption and the Poincar\'e inequality,
 \begin{align*}
@@
 \end{align*}
-Thus $f$ is 2-coercive, and so VI problem \eqref{eq:vi} is well-posed for any closed and convex $\cK \subset \cV$ and $g\in \cV'$.
+Thus $f$ is 2-coercive, and so VI problem \eqref{eq:vi} is well-posed.
 \end{example}
 
-If $\bX \ne 0$ then VI \eqref{eq:vi} for $f$ in \eqref{eq:advectiondiffusion} is not a (constrained) minimization problem.
-Indeed, the Jacobian of $f$, namely the bilinear form $J_f(u,v) = \int_\Omega \eps \grad u \cdot \grad v\,dx - b(u,v)$, is not symmetric in that case.\footnote{Assume $\bX \ne 0$ is continuous for simplicity.
-For $u,v$ which are zero on $\partial \Omega$, $J_f(u,v) - J_f(v,u) = -2 b(u,v)$.
-There exist $u,v$ so $b(u,v)$ is nonzero; construct $u,v$ locally near some point where $\bX$ is nonzero.}  Thus $f$ in \eqref{eq:advectiondiffusion} cannot be the Hessian of a scalar objective $F$.
-References \cite{Bueler2021conservation,ChangNakshatrala2017} consider such advection-diffusion VI problems over $\cK = \{v\ge 0\}$.
+References \cite{Bueler2021conservation,ChangNakshatrala2017} consider advection-diffusion VI problems like Example \ref{ex:advectiondiffusion}, specifically over the set $\cK = \{v\ge 0\}$.
+If $\bX \ne 0$ then VI \eqref{eq:vi} for $f$ in \eqref{eq:advectiondiffusion} does not correspond to a minimization problem.
+Indeed, $\ip{f(u)}{v}$ is not symmetric in that case,\footnote{Assume $\bX \ne 0$ is continuous for simplicity.
+For $u,v$ which are zero on $\partial \Omega$, note $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.
+By constructing $u,v$ locally near some point where $\bX$ is nonzero.
+one may show $b(u,v)\ne 0$.} so $f$ cannot be the gradient of a scalar objective.
 
-\begin{example}  \label{ex:porous}  FIXME
+\begin{example}  \label{ex:porous}  Suppose $\phi:[0,\infty) \to [\phi_0,\infty)$ is continuous for some constant $\phi_0>0$.
+Then the following operator, of porous medium type, is not known to be monotone (or coercive) when $\phi$ is not constant:
 \begin{equation}
-\ip{f(u)}{v} = \int_\Omega \phi(u) \grad u \cdot \grad v\,dx \label{eq:porous}
+\ip{f(u)}{v} = \int_\Omega \phi(u) \grad u \cdot \grad v\,dx. \label{eq:porous}
 \end{equation}
+As with Example \ref{ex:advectiondiffusion}, this form does not have the symmetry necessary to be the gradient of a scalar objective.
 \end{example}
@@
 Numerical solver performance for Examples \ref{ex:plaplacian}, \ref{ex:advectiondiffusion}, and \ref{ex:porous} will be considered in Section \ref{sec:results}.
-Further examples of nonlinear VI problems appear in ice sheet models \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids \cite{Bueler2021conservation}.
+Further examples of nonlinear VI problems appear in ice sheet models and other geophysical fluids \cite{Bueler2021conservation,Calvoetal2002,JouvetBueler2012}.
 
@@
 \end{equation}
-holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$; this is called a \emph{subspace decomposition} \cite{Xu1992}.
+holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$.
+Equation \eqref{eq:subspacedecomp} is called a \emph{subspace decomposition} \cite{Xu1992}.
 Suppose further that $\cK_i \subset \cV_i$ are nonempty, closed, and convex subsets such that
@@
 \end{equation}
-The sum in \eqref{eq:constraintdecomp} is required to hold in two senses: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.\footnote{Sense \emph{(ii)} for \eqref{eq:subspacedecomp} is automatic because the $\cV_i$ are subspaces.}
-
-Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique.
+The sum in \eqref{eq:constraintdecomp} is required to hold in two senses \cite{TaiTseng2002}: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.
+(Sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp} because the $\cV_i$ are subspaces.)  Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique.
 Also, $\cK_i \not\subset \cK$ in many applications; see the cartoon in Figure \ref{fig:cartoon}.
-
-Finally, for each $\cK_i$ we assume that there are bounded, (generally) nonlinear restriction operators $R_i : \cK \to \cK_i$ such that if $v \in \cK$ then
-\begin{equation}
-v = \sum_{i=0}^{m-1} R_i v.  \label{eq:constraintrestrictionsum}
-\end{equation}
-A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,R_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} \cite{Tai2003}.
 
@@
 \includegraphics[width=0.55\textwidth]{genfigs/cartoon.pdf}
-\caption{A constraint decomposition (CD) in a one-sided obstacle problem ($\mathcal{K}=\{v\ge \psi\}$) might look like this if $\mathcal{V}$ were 2-dimensional.}
+\caption{Suppose $\mathcal{V}$ is the space of real functions on a two-point set $\Omega=\{x_1,x_2\}$.
+A constraint decomposition (CD) for a one-sided obstacle problem with $\mathcal{K}=\{v\ge \psi\}$ looks like this.}
 \label{fig:cartoon}
@@
 
-In Section \ref{sec:multilevel} we will introduce discretizations and describe practical algorithms, but the fundamental CD concept applies even at the level of the continuum problem.
+For each $\cK_i$ we will also assume that there are bounded, generally-nonlinear restriction operators $R_i : \cK \to \cK_i$ such that if $v \in \cK$ then
+\begin{equation}
+v = \sum_{i=0}^{m-1} R_i v;  \label{eq:constraintrestrictionsum}
+\end{equation}
+see Figure \ref{fig:cartoon}.
+Clearly \eqref{eq:constraintrestrictionsum} implies sense \emph{(i)} for \eqref{eq:constraintdecomp}.
+A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,R_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} \cite{Tai2003}.
+
+In Section \ref{sec:multilevel} we will introduce discretizations and practical algorithms, but we observe here that the CD concept applies at the level of the continuum problem.
 The following two examples illustrate this for obstacle problems \cite{GraeserKornhuber2009}.
-First we consider an overlapping domain decomposition.
 
-\begin{example}  \label{ex:domaindecomposition}  Consider a bounded domain $\Omega \subset \RR^d$, let $\cV = W_0^{k,p}(\Omega)$ for $k\ge 0$ and $p\ge 1$, and suppose the obstacle $\psi \in W^{k,p}(\Omega)$ satisfies $\psi|_{\partial \Omega} \le 0$.
+\begin{example}  \label{ex:domaindecomposition}  Consider an overlapping domain decomposition as follows.
+For a bounded domain $\Omega \subset \RR^d$, let $\cV = W_0^{k,p}(\Omega)$ for $k\ge 0$ and $p\ge 1$, and suppose the obstacle $\psi \in W^{k,p}(\Omega)$ satisfies $\psi|_{\partial \Omega} \le 0$.
 Let $\cK = \{v \ge \psi\} \subset \cV$.

### 2022-07-17 paper/mcd2.tex
by splitting algs and commenting, I am really clear on them

@@
 
-Note that $\cK_i \not\subset \cK$ in most cases.
-In Example \ref{ex:domaindecomposition}, if $\psi$ is positive over portions of $\Omega$ where the decomposition into overlapping subdomains $\Omega_i$ is nontrivial, then $\cK_i \not\subset \cK$, and similarly for Example \ref{ex:frequencydecomposition}.
-Instead, the important inclusion is $\cK_i \subset \cV_i$.
+Note that $\cK_i \not\subset \cK$ in many cases.
+Specifically in Example \ref{ex:domaindecomposition}, if $\psi$ is positive over portions of $\Omega$ where the decomposition into overlapping subdomains $\Omega_i$ is nontrivial, then $\cK_i \not\subset \cK$, and similarly for Example \ref{ex:frequencydecomposition}.
+The important inclusion is $\cK_i \subset \cV_i$.
 
-Algorithm \ref{alg:basiccd} below states the basic CD algorithm as an iteration which solves smaller VI problems over each set $\cK_i$.
-(By contrast, the algorithms in \cite{Tai2003} solve optimization problems over each $\cK_i$.)  The algorithm starts from a current iterate $u \in \cK$ and computes a new iterate $w\in\cK$, an improved approximation of the solution $u^* \in \cK$ of \eqref{eq:vi}, using a damping parameter $0<\alpha\le 1$.
-There are parallel (additive) and successive (multiplicative) versions of the algorithm, generalizing the Jacobi and Gauss-Seidel iterations \cite{Greenbaum1997}, respectively.
-For the parallel version the \textbf{for} loop can be computed in any order.
+Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below state the basic CD method as iterations which improve a current iterate $u \in \cK$ by solving ``smaller'' VI problems over each subset $\cK_i$.
+The output $w\in\cK$ should be closer to the solution $u^* \in \cK$ of \eqref{eq:vi}.
 
@@
 \begin{pseudo*}
-\pr{cditeration}(u\in\cK)\text{:} \\+
-    if \pr{parallel}: \\+
+\pr{cd-add}(u\in\cK)\text{:} \\+
     for $i \in \{0,\dots,m-1\}$: \\+
-            $\hat w_i\in \cK_i$: \\+
-                 $\boxed{\ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\---
-    else: \\+
-        for $i = 0,\dots,m-1$: \\+
-            $\hat w_i\in \cK_i$: \\+
-                $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\Big)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)} \quad \forall v_i\in \cK_i$ \\-
-            $w_i = (1-\alpha) R_i u + \alpha \hat w_i\in\cK_i$ \\--
+        \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$ \rm{we have} \\+
+            $\boxed{{\large \strut} \ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)}$ \\--
     $\hat w = \sum_i \hat w_i\in\cK$ \\
@@
 \end{pseudo*}
-\caption{The basic constraint decomposition (CD) algorithm for VI problem \eqref{eq:vi}.}
-\label{alg:basiccd}
+\caption{Additive constraint decomposition for VI problem \eqref{eq:vi}.}
+\label{alg:basiccd-add}
 \end{pseudofloat}
 
-The reader may confirm that, inside each boxed VI in Algorithm \ref{alg:basiccd}, the argument of $f$ is an element of $\cK$.
-While the difference $v_i - \hat w_i \in \cV_i$ appears, note that by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write it as a difference of admissible vectors (i.e.~from $\cK$), namely
+The \textbf{for} loop in Algorithm \ref{alg:basiccd-add}, the additive or parallel CD method, can be computed in any order.
+Note that $u-R_iu+\hat w_i$ involves removing the part of $u$ which lies in $\mathcal{K}_i$ before adding-back the improved subset solution $\hat w_i$.
+
+The following multiplicative (successive) version orders the sets $\mathcal{K}_i$, and each subset solution updates the global iterate.
+While Algorithm \ref{alg:basiccd-add} generalizes a Jacobi iteration, this one is like Gauss-Seidel.
+The sum $\sum_{j>i} R_j u$ should be read as ``keep the parts of the current iterate which we have not yet improved.''
+
+\begin{pseudofloat}[H]
+\begin{pseudo*}
+\pr{cd-mult}(u\in\cK)\text{:} \\+
+    for $i = 0,\dots,m-1$: \\+
+        \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$ \rm{we have} \\+
+            $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\Big)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)}$ \\-
+            $w_i = (1-\alpha) R_i u + \alpha \hat w_i\in\cK_i$ \\-
+    $\hat w = \sum_i \hat w_i\in\cK$ \\
+    return $w=(1-\alpha) u + \alpha \hat w\in\cK$
+\end{pseudo*}
+\caption{Multiplicative constraint decomposition for VI problem \eqref{eq:vi}.}
+\label{alg:basiccd-mult}
+\end{pseudofloat}
+
+Note the use of a damping parameter $0<\alpha\le 1$.
+In Algorithm \ref{alg:basiccd-add} its role is obvious, as we keep part of the current iterate if $\alpha<1$, but in Algorithm \ref{alg:basiccd-mult} we only update the iterate ``so far'', namely $w_i$, using the same damping that will be applied in generating the final output $w$.
+
+In each boxed VI above the argument of $f$ is an element of $\cK$, as the reader should confirm.
+Likewise, by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write the test element $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors, from $\cK$, namely
 \begin{align*}
@@
 \end{align*}
-for the two versions, respectively.
+for the two Algorithms respectively.
 
-In other words, we solve VI \eqref{eq:vi} over perturbations from the sets $\cK_i$.  Define
+Define
 \begin{equation}
@@
 \end{equation}
-as the $i$th update from a pass through either \textbf{for} loop in Algorithm \ref{alg:basiccd}.
-Then $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.
-Easy identities $u - R_i u + \hat w_i = u + e_i$ and $\hat w = u^* + \sum_i \hat w_i - R_i u^*$ also hold.
+as the $i$th subset update in either Algorithm.
+Then, as the reader may check, $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.
+The easy identity $\hat w = u^* + \sum_i \hat w_i - R_i u^*$ also holds.
 We will use these identities in Section \ref{sec:convergence}.
 
-The convergence results in Section \ref{sec:convergence} require substantial damping when applying the parallel algorithm, i.e.~$\alpha \le 1/m$, the same as in \cite{Tai2003}, but the successive version can be shown to converge without damping ($\alpha=1$).
-In Section \ref{sec:results} we demonstrate practical convergence for a larger range of $\alpha$ than suggested by the theory.
-% HOPE
+% FOLLOWING CONTAINS ASPIRATIONS
+The convergence results in Section \ref{sec:convergence} require substantial damping when applying additive Algorithm \ref{alg:basiccd-add}, namely we require $\alpha \le 1/m$ as does \cite{Tai2003}.
+Algorithm \ref{alg:basiccd-mult} can be shown to converge without damping ($\alpha=1$).
+In Section \ref{sec:results} we demonstrate practical convergence for larger ranges of $\alpha$ than suggested by the theory.
 
-In order for Algorithm \ref{alg:basiccd} to have a practical and efficient FE implementation (Section \ref{sec:multilevel}), we need additional notions when $f$ has any of the following properties: \emph{(i)} $f$ is defined only on $\cK$, or \emph{(ii)} $f$ is nonlinear, or \emph{(iii)} $f$ is non-local.
-The issue of practical implementation for any such $f$ seems not to have been addressed in the CD literature, which is why we call Algorithm \ref{alg:basiccd} the ``basic'' algorithm.
-In particular  \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem.
-The following example explains the simplifications available in that easy case.
+We refer to Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} as the \emph{basic CD method}.
+To construct efficient FE implementations (Section \ref{sec:multilevel}) we will need additional notions.
+While the above Algorithms are indeed meaningful even when $f$ is nonlinear, non-local, or defined only on $\cK$, practical implementation for such general problems seems not to have been addressed in the literature.
+In particular, references \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem.
+The following example explains the implementation-relevant simplifications available in that case.
 
 \begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear and defined on all of $\cV$.
-Furthermore suppose $f$ is local in the sense that a basis $\{\phi_i\}$ of $\cV_i$ exists, with each support small in $\Omega$, such that $\ip{f(\phi_i)}{z}$ can be computed by an integral over the support of $\phi_i$.
-(When $\cV$ is an FE space then this is the usual case for evaluating PDE weak forms over a basis of hat functions; see Section \ref{sec:multilevel}.
-Constrast integral operators where $\ip{f(\phi_i)}{z}$ requires an integral over $\Omega$ even if the support of $\phi_i$ is small.)  Considering only the parallel version of Algorithm \ref{alg:basiccd} for simplicity, the boxed VI over $\cK_i$ can be written as
+Furthermore suppose $f$ is local in the sense that a basis $\{\phi_i\}$ of $\cV_i$ exists, with $\supp \phi_i$ compactly-contained in $\Omega$, such that for any $z\in\mathcal{V}$ the value $\ip{f(\phi_i)}{z}$ can be computed by an integral over the support of $\phi_i$.
+Considering only additive Algorithm \ref{alg:basiccd-add} for simplicity, by linearity the boxed VI over $\cK_i$ can be written as
 \begin{equation}
@@
 where $\tilde g(z) = g(z) - \ip{f(u)}{z}$.
-Noting $e_i = \hat w_i - R_i u \in \cV_i \notin \cK$, in general, VIs \eqref{eq:linearlocalvi} will only make sense because $f$ is defined over $\cV$.
-Each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $\ip{f(u)}{\cdot}$ (already included into the source term).
-The VI \eqref{eq:vi} is then approximately solved in any incremental and efficient manner, by computations over the basis supports, during one application of \pr{cditeration}.
+Note $e_i = \hat w_i - R_i u \in \cV_i \notin \cK$, in general, but VIs \eqref{eq:linearlocalvi} make sense because $f$ is defined over $\cV$.
+Each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $\ip{f(u)}{\cdot}$, namely as included here into the source term.
+The VI \eqref{eq:vi} is then approximately solved in any incremental and efficient manner, by computations over the basis supports.
 \end{example}
 
-In other words, when $f$ has all the nice properties proposed in Example \ref{ex:fnice} then the decomposed VI problems can be computed from inexpensive data.
-Note that the actual solution method for VI \eqref{eq:linearlocalvi} is not the concern here; we are observing that the data of problem \eqref{eq:linearlocalvi}, and the cost of residual evaluation, is small in practice.
-A solver which would work for Example \ref{ex:fnice} can thus have implementation efficiencies which are unavailable in general.
+The assumption that $\ip{f(\phi_i)}{z}$ can be evaluated over the support of $\phi_i$ is satisfied when $f$ is a linear differential operator, $\cV$ is an FE space, and $\phi_i$ are hat functions; see Section \ref{sec:multilevel}.
+Constrast non-local integral operators where $\ip{f(\phi_i)}{z}$ requires an integral over $\Omega$ even if $\phi_i$ is is a hat function with small support.
 
-Additional ideas, like those in Section \ref{sec:multilevel}, are needed for practical application of CD methods to more difficult problems.
-In particular, we will extend the basic Algorithm \ref{alg:basiccd} to nonlinear $f$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}, in which case the decomposed problems only evaluate $f$ over (admissible) elements of $\mathcal{K}$, and we propose restrictions which allow inexpensive approximations of the decomposed problems.
-However, extending the algorithm to non-local residual functionals is a topic for future research.
+When $f$ has all the nice properties listed in Example \ref{ex:fnice} then a solver can take advantage of implementation efficiencies which are unavailable in general.
+The actual solution process for VI \eqref{eq:linearlocalvi} is not the concern here; we are observing that the data of problem \eqref{eq:linearlocalvi}, and the cost of residual evaluation, are smaller than in the general case.
+For more difficult problems additional ideas, illustrated in Section \ref{sec:multilevel}, are needed for practical implementation.
+We will extend the basic method to nonlinear $f$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}, and the decomposed problems will only evaluate $f$ over (admissible) elements of $\mathcal{K}$.
+However, extending the algorithm to non-local functionals $f$ is a topic for future research.
 

### 2022-07-18 paper/mcd2.tex
edits in convergence lemma

@@
         \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$ \rm{we have} \\+
-            $\boxed{{\large \strut} \ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)}$ \\--
+            $\boxed{{\large \strut} \ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{g}{v_i-\hat w_i}}$ \\--
     $\hat w = \sum_i \hat w_i\in\cK$ \\
@@
         \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$ \rm{we have} \\+
-            $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\Big)}{v_i-\hat w_i} \ge g(v_i-\hat w_i)}$ \\-
+            $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\Big)}{v_i-\hat w_i} \ge \ip{g}{v_i-\hat w_i}}$ \\-
             $w_i = (1-\alpha) R_i u + \alpha \hat w_i\in\cK_i$ \\-
@@
 \begin{equation}
-\ip{f(e_i)}{v_i-\hat w_i} \ge \tilde g(v_i-\hat w_i) \label{eq:linearlocalvi}
+\ip{f(e_i)}{v_i-\hat w_i} \ge \ip{\tilde g}{v_i-\hat w_i} \label{eq:linearlocalvi}
 \end{equation}
-where $\tilde g(z) = g(z) - \ip{f(u)}{z}$.
+where $\tilde g = g - f(u)$.
 Note $e_i = \hat w_i - R_i u \in \cV_i \notin \cK$, in general, but VIs \eqref{eq:linearlocalvi} make sense because $f$ is defined over $\cV$.
@@
 
-We now seek a proof of the convergence of Algorithm \ref{alg:basiccd}.
-This will be possible if we restrict to $2$-coercive operators and make certain assumptions as in \cite{Tai2003}.
+We now prove the convergence of additive CD Algorithm \ref{alg:basiccd-add}. % ASPIRATION
+This will be possible if we restrict to $2$-coercive operators, require $\alpha \le 1/m$, and make certain other assumptions as in \cite{Tai2003}.
+However, our proof applies to VI problems which are not optimizations, and for which $f$ only applies to admissible elements, and thus extends the theory in \cite{Tai2003}.
 First we define a new quantity related to VI problem \eqref{eq:vi}.
@@
 \begin{equation}
-  E(v,u) = \ip{f(v)}{v-u} - g(v-u).  \label{eq:normlikedefn}
+  E(v,u) = \ip{f(v)}{v-u} - \ip{g}{v-u}.  \label{eq:normlikedefn}
 \end{equation}
@@
 
-If $u^*$ solves \eqref{eq:vi} then $E(v,u^*)$ is somewhat like an ``merit function,'' something used in the context of solving nonlinear equations \cite{NocedalWright2006} to replace an objective (scalar) functional.
-(Note that $E(v,u^*)$ is \emph{not} a residual for \eqref{eq:vi}.)  The next lemma, which follows directly from $p$-coercivity, shows $E(v,u^*)$ is bounded below.
+If $u^*$ solves \eqref{eq:vi} then $E(v,u^*)$ is somewhat like an ``merit function'' for an iterate $v$, as used in solving nonlinear equations \cite{NocedalWright2006}, replacing the scalar objective which is assumed in \cite{Tai2003}.
+The next lemma shows $E(v,u^*)$ is bounded below by a norm.
 
@@
 \begin{align*}
-E(v,u^*) &= \ip{f(v)}{v-u^*} - \ip{f(u^*)}{v-u^*} + \ip{f(u^*)}{v-u^*} - g(v-u^*) \\
-   &\ge \kappa \|v-u^*\|^p + 0.  \qedhere
+E(v,u^*) &= \ip{f(v)}{v-u^*} - \ip{f(u^*)}{v-u^*} + \ip{f(u^*)}{v-u^*} - \ip{g}{v-u^*} \\
+   &\ge \ip{f(v)-f(u^*)}{v-u^*} + 0 \ge \kappa \|v-u^*\|^p.  \qedhere
 \end{align*}
@@
 \end{equation}
-for all $u,v\in\cK$.  Furthermore, there exists $C_2>0$ so that for all $v_i \in \cV_i$ and $y_j \in \cV_j$,
+for all $u,v\in\cK$.  Furthermore, there exists a constant $C_2>0$ so that for all $v_i \in \cV_i$ and $y_j \in \cV_j$, and all $w_{ij} \in \cK$ such that $w_{ij} + v_i \in \cK$, it holds that
 \begin{equation}
-\sum_{i=0}^{m-1} \sum_{j=0}^{m-1} \left|\ip{f(w_{ij} + v_i) - f(w_{ij})}{y_j}\right| \le C_2 \left(\sum_{i=0}^{m-1} \|v_i\|^2\right)^{1/2} \left(\sum_{j=0}^{m-1} \|y_j\|^2\right)^{1/2} \label{as:lipschitzresidual}
+\sum_{i=0}^{m-1} \sum_{j=0}^{m-1} \left|\ip{f(w_{ij} + v_i) - f(w_{ij})}{y_j}\right| \le C_2 \left(\sum_{i=0}^{m-1} \|v_i\|^2\right)^{1/2} \left(\sum_{j=0}^{m-1} \|y_j\|^2\right)^{1/2}.
+\label{as:lipschitzresidual}
 \end{equation}
-for all $w_{ij} \in \cK$ such that $w_{ij} + v_i \in \cK$ for all $i$ and $j$.
 \end{assumptions*}
@@
 \begin{itemize}
-\item Assumption \eqref{as:lipschitzrestrictions} addresses only the CD, and not the map $f$.
-\item It is sometimes the case that \eqref{as:lipschitzresidual} addresses only a subspace decomposition property of $f$ (and not the CD itself).
-For example, in the classical obstacle problem one may verify \eqref{as:lipschitzresidual} over all $w_{ij} \in \cV$.
+\item Assumption \eqref{as:lipschitzrestrictions} addresses only the CD, and not the VI problem $f$.
+Tai \cite{Tai2003} gives a value of $C_1$ for obstacle problem constraint decompositions using $P_1$ FE spaces over shape-regular and quasi-uniform triangulations.
+In particular, $C_1$ is known for overlapping domain decompositions, with or without an additional coarse mesh, and standard multilevel hierarchies.
+The cases we will demonstrate in Section \ref{sec:results} are covered.
+\item If $f$ is defined on all of $\mathcal{V}$ then one might be able to verify \eqref{as:lipschitzresidual} over a subspace decomposition, not the CD.
+For example, in the classical obstacle problem one may verify \eqref{as:lipschitzresidual} over arbitrary $w_{ij} \in \cV$ \cite{Tai2003}.
+\item If we assume $f$ itself is Lipschitz then the existence of $C_2$ is clear \cite{TaiXu2002}.
+However, our convergence rate bound is improved when $C_2$ is smaller, so the smallest $C_2$ for which \eqref{as:lipschitzresidual} holds is used.
+\item The constants $C_1,C_2$ generally depend on $m$.
+Furthermore, regarding the convergence proof below, Tai \cite{Tai2003} observes that if the iterates from Algorithm \ref{alg:basiccd-add} are bounded in $\mathcal{K}$ then constants $C_1,C_2$ are permitted to depend continuously on $u,v,w_{ij},v_i,y_i$.
 \item The unconstrained versions of \eqref{as:lipschitzrestrictions} and \eqref{as:lipschitzresidual} are (13), (14) in \cite{TaiXu2002}, respectively.
-\item If we assume $f$ itself is Lipschitz then the existence of $C_2$ is clear \cite{TaiXu2002}.
-However, our bound for the convergence rate for Algorithm \ref{alg:basiccd} is improved when $C_1,C_2$ are made smaller.
-\item Tai \cite{Tai2003} gives values of the constants $C_1$ in \eqref{as:lipschitzrestrictions} for obstacle problem constraint decompositions using $P_1$ FE spaces over shape-regular and quasi-uniform triangulations.
-In particular, constants $C_1$ are known for overlapping domain decompositions (with or without an additional coarse mesh) and standard multilevel hierarchies.
-Thus the cases we will demonstrate in Section \ref{sec:results} are already covered.
-\item The constants $C_1,C_2$ generally depend on $m$.
-However, regarding the convergence proof below, \cite{Tai2003} observes that if repeated application of Algorithm \ref{alg:basiccd} generates a bounded sequence of iterates in $\mathcal{K}$ then constants $C_1,C_2$ would be permitted to depend on $u,v,w_{ij},v_i,y_i$ as well.
 \end{itemize}
 
-From the above definitions and assumptions we have the following estimate for the parallel version of Algorithm \ref{alg:basiccd}.  Our proof follows \cite{Tai2003}.
+From the above definitions and assumptions we have the following estimate for Algorithm \ref{alg:basiccd-add}.  Our proof follows \cite{Tai2003}.
 
 \begin{lemma} \label{lem:core}  Suppose $f$ is $2$-coercive and $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.
-Recall $E(v,u^*)$ is defined in \eqref{eq:normlikedefn}.
-Suppose $\mathcal{V}_i$, $\mathcal{K}_i$, $R_i$, and $f$ satisfy \eqref{as:lipschitzrestrictions} and \eqref{as:lipschitzresidual}.
-Suppose $\hat w$ is computed from $u$ by the parallel version of Algorithm \ref{alg:basiccd}, and recall $e_i = \hat w_i - R_i u$.
-Then
+Assume $\mathcal{V}_i$, $\mathcal{K}_i$, $R_i$ form a CD satisfying \eqref{as:lipschitzrestrictions}, and that \eqref{as:lipschitzresidual} is satisfied.
+If $\hat w$ is computed from $u \in \mathcal{K}$ by Algorithm \ref{alg:basiccd-add} then
 \begin{equation}
@@
 
-\begin{proof}  Recalling that $\hat w = u^* + \sum_i \hat w_i - R_i u^*$, by the (parallel version) boxed VI in Algorithm \ref{alg:basiccd} with $v_i = R_i u^*$ we get
+\begin{proof}  Recall that $\hat w = u^* + \sum_i \hat w_i - R_i u^*$.  Expand $E(\hat w,u^*)$ and substitute $v_i = R_i u^*$ into the boxed VI in Algorithm \ref{alg:basiccd-add} to find
 \begin{align}
-E(\hat w,u^*) &= \sum_{i=0}^{m-1} \ip{f(\hat w)}{\hat w_i - R_i u^*} - g(\hat w_i - R_i u^*) \label{eq:startcore} \\
+E(\hat w,u^*) &= \sum_{i=0}^{m-1} \ip{f(\hat w)}{\hat w_i - R_i u^*} - \ip{g}{\hat w_i - R_i u^*} \label{eq:startcore} \\
     &\le \sum_{i=0}^{m-1} \ip{f(\hat w)}{\hat w_i - R_i u^*} + \ip{f(u + e_i)}{R_i u^* - \hat w_i} \notag \\
@@
 \end{align}
-For $i\in \{0,1,\dots,m-1\}$ use wrapped indices to define elements $\phi_{i,j}$:
+
+For $i\in \{0,1,\dots,m-1\}$ we now use wrapped indices to define progressive states $\phi_{i,j}$:
 \begin{align}
@@
   &\vdots \notag \\
-\phi_{i,m-1} &= u + e_i + e_{i+1} + \dots + e_{m-1} + e_0 + \dots + e_{i-1}, \notag
+\phi_{i,m-1} &= u + e_i + e_{i+1} + \dots + e_{m-1} + e_0 + \dots + e_{i-1}. \notag
 \end{align}
-and $\phi_{i,j} \in \cK$ by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum}.
-Note that $\phi_{i,m-1} = u + \sum_i e_i = \hat w$.
-From \eqref{eq:startcore}, apply a telescoping sum using the elements in \eqref{eq:gridcore}:
+Observe that $\phi_{i,j} \in \cK$ by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum}, and that $\phi_{i,m-1} = u + \sum_i e_i = \hat w$, thus that \eqref{eq:startcore} can be written
+\begin{equation}
+E(\hat w,u^*) = \sum_{i=0}^{m-1} \ip{f(\phi_{i,m-1}) - f(\phi_{i,0})}{\hat w_i - R_i u^*}. \label{eq:startcoreclean}
+\end{equation}
+From \eqref{eq:startcoreclean}, telescope into a sum over $j=1,\dots,m-1$, and then apply the triangle inequality:
 \begin{align}
 E(\hat w,u^*) &\le \sum_{i=0}^{m-1} \sum_{j=1}^{m-1} \ip{f(\phi_{i,j}) - f(\phi_{i,j-1})}{\hat w_i - R_i u^*} \label{eq:nextcore} \\
-  &\le \sum_{i=0}^{m-1} \sum_{j=0}^{m-1} \left|\ip{f(\phi_{i,j}) - f(\phi_{i,j-1})}{\hat w_i - R_i u^*}\right|. \notag
+  &\le \sum_{i=0}^{m-1} \sum_{j=1}^{m-1} \left|\ip{f(\phi_{i,j}) - f(\phi_{i,j-1})}{\hat w_i - R_i u^*}\right|. \notag
 \end{align}
-By an easy renumbering of the indices $i,j$ we may apply assumption \eqref{as:lipschitzresidual}, then the triangle inequality, and then assumption \eqref{as:lipschitzrestrictions}; we denote $\left(\sum_{i=0}^{m-1} \|e_i\|^2\right)^{1/2}$ by $Z$:
+
+Let $Z=\left(\sum_{i=0}^{m-1} \|e_i\|^2\right)^{1/2}$.  FIXME: MAKE THE RENUMBERING EXPLICIT
+
+renumbering the indices $i,j$ we may apply assumption \eqref{as:lipschitzresidual}, then the triangle inequality, and then assumption \eqref{as:lipschitzrestrictions}:
 \begin{align}

### 2022-07-24 py/bratu1.py
fix bug: coarse solver needs to solve nonlinear problem

@@ -44,8 +44,8 @@ smootherparams = {"snes_rtol": 1.0,  # always succeed after one newton step
                   "ksp_max_it": 1,
                   "pc_type": "sor",
                   "pc_sor_forward": None}
-coarseparams = {"snes_type": "ksponly",
-                #"ksp_converged_reason": None,
+coarseparams = {"snes_converged_reason": None,
+                "ksp_converged_reason": None,
                 "ksp_type": "preonly",
                 "pc_type": "lu"}
 

### 2022-07-24 py/bratu1.py
comments and whitespace

@@ -1,6 +1,8 @@
 """Solve Bratu equation in 1D by FAS multigrid:
    -u'' - lambda e^u = f(x)
-with Dirichlet conditions u(0)=u(1)=0 and f(x) = 9 pi^2 sin(3 pi x) − lambda exp(sin(3 pi x)).  The exact solution of this problem is u(x) = sin(3 pi x).
+with Dirichlet conditions u(0)=u(1)=0 and
+   f(x) = 9 pi^2 sin(3 pi x) − lambda exp(sin(3 pi x)).
+The exact solution of this problem is u(x) = sin(3 pi x).
 
 The goal is to solve the same problem as in
     Bueler (2021) "The full approximation storage multigrid scheme:
@@ -15,10 +17,11 @@ where v is a test function and
     ell[v] = f * v * dx
 
 Solver is FAS V-cycles.  Going downward the V-cycle uses solve() for a
-down-smoother, then forms the FAS correction equation, then goes down a level.
+down-smoother, forms the FAS correction equation, and decrements the level.
 The coarsest problem is solved by LU.  Going upward the V-cycle prolongs and
 adds the computed correction and then uses solve() as an up-smoother.  The
-smoothers are a fixed number Newton iterations with a fixed number of GS sweeps.
+smoother is a fixed number of Newton iterations with a fixed number of GS sweeps.
+(FIXME: backward GS on up?)
 
 The FAS correction equation is
     Fc(uc)[vc] = ellc[vc]
@@ -26,7 +29,10 @@ where Fc has same formula as F and
     ellc[vc] = R'(resf)[vc] + Fc(R(uf))[vc]
 and
     resf[v] = ellf[v] - Ff(uf)[v]
-and R' is canonical restriction and R is injection.
+and R' is canonical restriction and R is injection.  After solution the fine
+level iterate is updated
+    u <- u + P(uc - R(uf))
+where P is prolongation.
 """
 
 from firedrake import *
@@ -53,7 +59,7 @@ coarseparams = {"snes_converged_reason": None,
 cmesh = UnitIntervalMesh(2)
 hierarchy = MeshHierarchy(cmesh, levels-1)
 
-# set up function spaces and some functions on each level
+# set up function spaces and problem on each level
 VV = [FunctionSpace(hierarchy[i], "CG", 1) for i in range(levels)]
[…]

### 2022-07-24 py/bratu1.py
use sensible options prefixes for solvers

@@ -43,8 +43,8 @@ lam = 1.0
 smootherparams = {"snes_rtol": 1.0,  # always succeed after one newton step
                   #"snes_view": None,
                   "snes_max_linear_solve_fail": 200, # don't error when KSP reports DIVERGED_ITS
-                  #"snes_converged_reason": None,
-                  #"ksp_converged_reason": None,
+                  "snes_converged_reason": None,
+                  "ksp_converged_reason": None,
                   #"ksp_monitor": None,
                   "ksp_type": "richardson",
                   "ksp_max_it": 1,
@@ -104,7 +104,7 @@ for j in range(vcycles):
     # down-smoothing and creation of FAS correction equation
     for i in range(levels-1,0,-1):   # levels-1, ..., 1
         # smoother application
-        solve(res[i] == 0, u[i], bcs=bcs[i],
+        solve(res[i] == 0, u[i], bcs=bcs[i], options_prefix = 'down%d' % i,
               solver_parameters=smootherparams)
         # construct FAS correction problem with right-hand side
         #     ell[i-1] = R'(res(u[i]))) + F[i-1](R(u[i]))
@@ -134,8 +134,8 @@ for j in range(vcycles):
                               # but restrict() seems better?
 
     # coarse-mesh problem and direct solution
-    solve(res[0] == 0, u[0], bcs=bcs[0],
-          solver_parameters=coarseparams)
+    solve(res[0] == 0, u[0], bcs=bcs[0], options_prefix = 'coarse',
+          solver_parameters=coarseparams, )
     print('%s|coarse cor|=%.6f' % ((levels-1)*'  ', norm(u[0])))
 
     # prolongation and up-smoothing
@@ -144,7 +144,7 @@ for j in range(vcycles):
         prolong(assemble(u[i-1] - Ru[i-1]), Pcor)  # assemble() needed because
                                                    # subtraction is UFL
         u[i] += Pcor
-        solve(res[i] == 0, u[i], bcs=bcs[i],
+        solve(res[i] == 0, u[i], bcs=bcs[i], options_prefix = 'up%d' % i,
               solver_parameters=smootherparams)
 
     # report on result of cycle

### 2022-07-24 py/bratu1.py
commit to redefinition of residual (res = F - ell not res = ell - F) because it allows CG iters as smoother

@@ -95,7 +95,7 @@ def error(u):
 
 # fine-mesh problem
 ell[fine] = f * v[fine] * dx
-res[fine] = ell[fine] - F[fine]
+res[fine] = F[fine] - ell[fine]
 print('initial             |residual|=%.6f  |error|=%.6f' \
       % (norm(assemble(res[fine])), error(u[fine])))
 
@@ -128,7 +128,7 @@ for j in range(vcycles):
             print(Rres.dat.data)
 
         ell[i-1] = Rres * v[i-1] * dx + FRu[i-1]
-        res[i-1] = ell[i-1] - F[i-1]
+        res[i-1] = F[i-1] - ell[i-1]
         inject(u[i], u[i-1])  # for initial value on next level
                               # equivalent to setting initial error iterate to zero
                               # but restrict() seems better?

### 2022-07-24 py/bratu1.py
split smoother parameters into down and up, but keep them same for now

@@ -50,6 +50,10 @@ smootherparams = {"snes_rtol": 1.0,  # always succeed after one newton step
                   "ksp_max_it": 1,
                   "pc_type": "sor",
                   "pc_sor_forward": None}
+downparams = smootherparams.copy()
+#downparams["pc_sor_forward"] = None
+upparams = smootherparams.copy()
+#upparams["pc_sor_backward"] = None
 coarseparams = {"snes_converged_reason": None,
                 "ksp_converged_reason": None,
                 "ksp_type": "preonly",
@@ -105,7 +109,7 @@ for j in range(vcycles):
     for i in range(levels-1,0,-1):   # levels-1, ..., 1
         # smoother application
         solve(res[i] == 0, u[i], bcs=bcs[i], options_prefix = 'down%d' % i,
-              solver_parameters=smootherparams)
+              solver_parameters=downparams)
         # construct FAS correction problem with right-hand side
         #     ell[i-1] = R'(res(u[i]))) + F[i-1](R(u[i]))
         # where R' is canonical restriction and R is injection
@@ -145,7 +149,7 @@ for j in range(vcycles):
                                                    # subtraction is UFL
         u[i] += Pcor
         solve(res[i] == 0, u[i], bcs=bcs[i], options_prefix = 'up%d' % i,
-              solver_parameters=smootherparams)
+              solver_parameters=upparams)
 
     # report on result of cycle
     print('vcycle %d (%d levels) |residual|=%.6f  |error|=%.6f' \

### 2022-07-24 py/bratu1.py
switch to a more credible smoother ... now I am irritated NGS is not available?

@@ -40,22 +40,25 @@ from firedrake import *
 levels = 2
 vcycles = 1
 lam = 1.0
-smootherparams = {"snes_rtol": 1.0,  # always succeed after one newton step
+smootherparams = {"snes_rtol": 0.1,  # reduce residual a bit
                   #"snes_view": None,
+                  #"snes_max_it": 2,  FIXME: does not work because Firedrake regards DIVERGED_MAX_IT as failure
                   "snes_max_linear_solve_fail": 200, # don't error when KSP reports DIVERGED_ITS
                   "snes_converged_reason": None,
                   "ksp_converged_reason": None,
                   #"ksp_monitor": None,
-                  "ksp_type": "richardson",
-                  "ksp_max_it": 1,
-                  "pc_type": "sor",
-                  "pc_sor_forward": None}
+                  "ksp_type": "cg",
+                  #"ksp_max_it": 10,
+                  "ksp_rtol": 0.01,
+                  #"pc_type": "ilu"}
+                  "pc_type": "sor"}
+                  #"pc_type": "lu"}
 downparams = smootherparams.copy()
 #downparams["pc_sor_forward"] = None
 upparams = smootherparams.copy()
 #upparams["pc_sor_backward"] = None
 coarseparams = {"snes_converged_reason": None,
-                "ksp_converged_reason": None,
+                #"ksp_converged_reason": None,
                 "ksp_type": "preonly",
                 "pc_type": "lu"}
 

### 2022-07-24 py/bratu1.py
document and leave in acceptable condition

@@ -19,9 +19,12 @@ where v is a test function and
 Solver is FAS V-cycles.  Going downward the V-cycle uses solve() for a
 down-smoother, forms the FAS correction equation, and decrements the level.
 The coarsest problem is solved by LU.  Going upward the V-cycle prolongs and
-adds the computed correction and then uses solve() as an up-smoother.  The
-smoother is a fixed number of Newton iterations with a fixed number of GS sweeps.
-(FIXME: backward GS on up?)
+adds the computed correction and then uses solve() as an up-smoother.
+
+The smoother is a single Newton iteration.  (Usually.  In fact we set
+-snes_rtol 0.1.  Note -snes_max_it is broken for composed solvers.)  Each
+Newton step linear system is, for now, solved with CG+SOR iteration with
+-ksp_rtol 0.01.
 
 The FAS correction equation is
     Fc(uc)[vc] = ellc[vc]
@@ -37,7 +40,7 @@ where P is prolongation.
 
 from firedrake import *
 
-levels = 2
+levels = 4
 vcycles = 1
 lam = 1.0
 smootherparams = {"snes_rtol": 0.1,  # reduce residual a bit
@@ -50,9 +53,8 @@ smootherparams = {"snes_rtol": 0.1,  # reduce residual a bit
                   "ksp_type": "cg",
                   #"ksp_max_it": 10,
                   "ksp_rtol": 0.01,
-                  #"pc_type": "ilu"}
+                  #"pc_type": "ilu"}  # same as LU because tridiagonal
                   "pc_type": "sor"}
-                  #"pc_type": "lu"}
 downparams = smootherparams.copy()
 #downparams["pc_sor_forward"] = None
 upparams = smootherparams.copy()

### 2022-07-25 py/bratu1.py
fix doc string

@@ -31,7 +31,7 @@ The FAS correction equation is
 where Fc has same formula as F and
     ellc[vc] = R'(resf)[vc] + Fc(R(uf))[vc]
 and
-    resf[v] = ellf[v] - Ff(uf)[v]
+    resf[v] = Ff(uf)[v] - ellf[v]
 and R' is canonical restriction and R is injection.  After solution the fine
 level iterate is updated
     u <- u + P(uc - R(uf))

### 2022-07-25 py/poisson0.py
convert plain MG code to 2 CG+ICC iters for smoother

@@ -2,7 +2,8 @@
 
 from firedrake import *
 
-levels = 2
+levels = 4
+vcycles = 1
 cmesh = UnitSquareMesh(2, 2)  # as small as practical for nontriviality
 hierarchy = MeshHierarchy(cmesh, levels-1)
 mesh = hierarchy[-1]
@@ -36,16 +37,19 @@ print('initial                   |residual|=%.6f  |error|=%.6f' \
 params = {"snes_type": "ksponly",
           "snes_max_linear_solve_fail": 2, # don't error when KSP reports DIVERGED_ITS
           #"snes_view": None,
-          "ksp_converged_reason": None,
-          "ksp_monitor_true_residual": None,
-          "ksp_type": "richardson",
-          "ksp_max_it": 1,
-          "mg_levels_ksp_type": "richardson",
-          "mg_levels_ksp_max_it": 1,
-          "mg_levels_pc_type": "sor",
-          "pc_type": "mg"}
+          #"ksp_converged_reason": None,
+          #"ksp_monitor_true_residual": None,
+          "ksp_type": "richardson", # classical V-cycles
+          "ksp_max_it": vcycles,
+          "pc_type": "mg",
+          "mg_levels_ksp_type": "cg",
+          "mg_levels_ksp_max_it": 2,
+          "mg_levels_ksp_converged_reason": None,
+          "mg_levels_pc_type": "icc",
+          "mg_coarse_ksp_type": "preonly",
+          "mg_coarse_pc_type": "lu"}
 u = run_solve(params)
-print('MG 2 V-cycles (%d levels)  |residual|=%.6f  |error|=%.6f' \
-      % (levels, norm(assemble(-F)), error(u)))
+print('MG V-cycles (%d cycles, %d levels)  |residual|=%.6f  |error|=%.6f' \
+      % (vcycles, levels, norm(assemble(-F)), error(u)))
 u.rename("u")
 File("u-poisson0.pvd").write(u)

### 2022-07-25 py/poisson.py
bring FAS Poisson into compliance with Bratu 1D choices ... seems to work

@@ -8,32 +8,41 @@ where v is a test function and
 and boundary conditions are homogeneous Dirichlet.
 
 Solver is FAS V-cycles.  Going downward the V-cycle uses solve() for a
-down-smoother, then forms the FAS correction equation, then goes down a level.
+down-smoother, forms the FAS correction equation, and decrements the level.
 The coarsest problem is solved by LU.  Going upward the V-cycle prolongs and
-adds the computed correction and then uses solve() as an up-smoother.  The
-smoothers are a fixed number of SSOR iterations.
+adds the computed correction and then uses solve() as an up-smoother.
+
+The smoother is a KSP-only solve, with two CG+ICC iterations.
 
 The FAS correction equation is
     Fc(uc)[vc] = ellc[vc]
 where Fc has same formula as F and
     ellc[vc] = R'(resf)[vc] + Fc(R(uf))[vc]
 and
-    resf[v] = ellf[v] - Ff(uf)[v]
-and R' is canonical restriction and R is injection.
+    resf[v] = Ff(uf)[v] - ellf[v]
+and R' is canonical restriction and R is injection.  After solution the fine
+level iterate is updated
+    u <- u + P(uc - R(uf))
+where P is prolongation.
 """
 
 from firedrake import *
 
-levels = 2
+levels = 4
 vcycles = 1
 smootherparams = {"snes_type": "ksponly",
                   "snes_max_linear_solve_fail": 2, # don't error when KSP reports DIVERGED_ITS
-                  #"ksp_converged_reason": None,
-                  "ksp_type": "richardson",
-                  "ksp_max_it": 1,
-                  "pc_type": "sor"}
+                  "ksp_converged_reason": None,
+                  "ksp_type": "cg",
+                  #"ksp_rtol": 0.01,
+                  "ksp_max_it": 2,
+                  "pc_type": "icc"}
                   #"ksp_type": "preonly",
                   #"pc_type": "lu"}
+downparams = smootherparams.copy()
+#downparams["pc_sor_forward"] = None
+upparams = smootherparams.copy()
+#upparams["pc_sor_backward"] = None
 coarseparams = {"snes_type": "ksponly",
                 #"ksp_converged_reason": None,
                 "ksp_type": "preonly",
[…]

### 2022-07-25 py/poisson.py
i am really frustrated ... FAS not working

@@ -28,15 +28,15 @@ where P is prolongation.
 
 from firedrake import *
 
-levels = 4
+levels = 2
 vcycles = 1
 smootherparams = {"snes_type": "ksponly",
                   "snes_max_linear_solve_fail": 2, # don't error when KSP reports DIVERGED_ITS
                   "ksp_converged_reason": None,
-                  "ksp_type": "cg",
+                  "ksp_type": "richardson",
                   #"ksp_rtol": 0.01,
-                  "ksp_max_it": 2,
-                  "pc_type": "icc"}
+                  "ksp_max_it": 1,
+                  "pc_type": "sor"}
                   #"ksp_type": "preonly",
                   #"pc_type": "lu"}
 downparams = smootherparams.copy()

### 2022-07-25 py/poisson0.py

@@ -2,7 +2,7 @@
 
 from firedrake import *
 
-levels = 4
+levels = 2
 vcycles = 1
 cmesh = UnitSquareMesh(2, 2)  # as small as practical for nontriviality
 hierarchy = MeshHierarchy(cmesh, levels-1)
@@ -14,11 +14,10 @@ v = TestFunction(Vf)
 
 x, y = SpatialCoordinate(mesh)
 f = -0.5*pi*pi*(4*cos(pi*x) - 5*cos(pi*x*0.5) + 2)*sin(pi*y)
+exact = sin(pi*x)*tan(pi*x*0.25)*sin(pi*y)
 
 F = dot(grad(u), grad(v)) * dx - f*v*dx
 bcs = DirichletBC(Vf, zero(), (1, 2, 3, 4))
-
-exact = sin(pi*x)*tan(pi*x*0.25)*sin(pi*y)
 
 Vc = FunctionSpace(cmesh, "CG", 1)
 
@@ -42,10 +41,10 @@ params = {"snes_type": "ksponly",
           "ksp_type": "richardson", # classical V-cycles
           "ksp_max_it": vcycles,
           "pc_type": "mg",
-          "mg_levels_ksp_type": "cg",
-          "mg_levels_ksp_max_it": 2,
+          "mg_levels_ksp_type": "richardson",
+          "mg_levels_ksp_max_it": 1,
           "mg_levels_ksp_converged_reason": None,
-          "mg_levels_pc_type": "icc",
+          "mg_levels_pc_type": "sor",
           "mg_coarse_ksp_type": "preonly",
           "mg_coarse_pc_type": "lu"}
 u = run_solve(params)

### 2022-07-26 paper/mcd2.tex
sweep of superficial edits and planning

@@
 
-We construct a new proof of convergence which removes the assumption that the problem arises from optimization.
-Furthermore we address finite element (FE) implementation in nonlinear cases, and we demonstrate the resulting performance.
-In fact, the theory presented in \cite{Tai2003} is extended in four particular directions:
+The theory presented in \cite{Tai2003} is extended in three particular directions:
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
 \begin{enumerate}
-\item We do not assume that the continuum VI problem arises from optimization of a scalar objective; examples are explored in Sections \ref{sec:vi} and \ref{sec:results}.
-We prove convergence in $H^1$ norm at the same rate as the original method converges in energy (Section \ref{sec:convergence}).
+\item We do not assume that the continuum VI problem arises from optimization of a scalar objective---examples are explored in Sections \ref{sec:vi} and \ref{sec:results}---but we prove convergence in $H^1$ norm at the same rate as the original method converges in energy (Section \ref{sec:convergence}).
 % THIS IS THE HOPE
 
-\item We make the observation that, if implemented appropriately, multilevel ``up-smoothing'' is intrinsically more efficient than ``down-smoothing'' (Section \ref{sec:multilevel}).
+\item We make finite element (FE) implementations of the multilevel CD algorithm more practical by addressing storage of intermediate quantities.
+Specifically, in Section \ref{sec:results} we show results from a full approximation storage (FAS; see \cite{Brandt1977,Bruneetal2015}) implementation for nonlinear operators, one which avoids global (Newton) linearization; compare \cite{GraeserKornhuber2009}.
+
+\item We observe that multilevel ``up-smoothing'' is intrinsically more efficient than ``down-smoothing'' (Section \ref{sec:multilevel}).
 This observation seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}.
-A strong preference for up-smoothing is, apparently, special to multilevel CD methods and does not arise in corresponding unconstrained problems.
+A strong preference for up-smoothing is special to multilevel CD methods and does not arise in unconstrained problems.
+\end{enumerate}
 
-\item We make the multilevel CD algorithm more practical by addressing storage of intermediate quantities.
-In Section \ref{sec:results} we show results from a full approximation storage (FAS; see \cite{Brandt1977}) implementation for nonlinear operators, one which avoids global (Newton) linearization; compare \cite{GraeserKornhuber2009}.
-
-\item For problems of porous-media type the nonlinear operator $f$ is not known to be coercive, so the convergence theory in \emph{(i)} does not apply.
+For problems of porous-media type the nonlinear operator $f$ is not known to be coercive, so the convergence theory in \emph{(i)} does not apply.
 However, a full-cycle implementation of the multilevel CD method is demonstrated numerically.
@@
 The resulting scheme is highly-effective for doubly-nonlinear diffusion operators (Section \ref{sec:results}).
-\end{enumerate}
-
-The iterates from a CD method are always admissible, and thus the operator need only be defined for admissible states; Section \ref{sec:results} includes a nontrivial example.
-Admissible-iterate methods should permit direct solutions of certain VI problems, such as fluid-layer dynamics problems \cite{Bueler2021conservation,JouvetBueler2012}, for which non-admissible methods, such as semi-smooth methods \cite{BensonMunson2006}, at least require unnatural modifications of the operator formula.
-The full-cycle scheme in \emph{(iv)} above is also designed for these problems, which are characterized by the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation.
 
@@
 Acceleration of our nonlinear multilevel CD method via active-set and/or basis-level manipulations represents a potential extension of the method here, and is a topic for future research.
+
+The iterates from a CD method are always admissible, and thus the operator need only be defined for admissible states; Section \ref{sec:results} includes a nontrivial example.
+Admissible-iterate methods should permit direct solutions of certain VI problems, such as fluid-layer dynamics problems \cite{Bueler2021conservation,JouvetBueler2012}, for which non-admissible methods, such as semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
+The full-cycle scheme in \emph{(ii)} above is also designed for these problems, which are characterized by the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation.
 
@@
 Denote the dual pairing of $\phi \in \cV'$ and $v\in\cV$ by $\ip{\phi}{v} = \phi(v)$, and note that $\|\phi\|_{\cV'} = \sup_{\|v\|=1} |\ip{\phi}{v}|$ defines a (Banach space) norm on $\cV'$.
-
 Let $\cK \subset \cV$ be a nonempty closed and convex subset, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.
@@
 \end{equation}
-Because $f$ is (generally) nonlinear, $g$ is not strictly needed for posing this VI, and by redefining $f$ we may take $g=0$, but the presence of $g$ is helpful to algorithms in Section \ref{sec:multilevel}.
+Because $f$ is (generally) nonlinear, $g$ is not strictly needed for posing this VI,  but the presence of $g$ is helpful to algorithms in Section \ref{sec:multilevel}.
+(By redefining $f$ we may take $g=0$.)
 
@@
 That is, \eqref{eq:vi} says that $f(u^*)-g$ may not be zero but it points directly into $\cK$.
-In particular, if $u^*$ is in the interior $\cK^\circ$ then \eqref{eq:vi} implies $f(u^*)=g$.
+However, if $u^*$ is in the interior $\cK^\circ$ then \eqref{eq:vi} implies $f(u^*)=g$.
 
@@
 
-Note that Tai \cite{Tai2003} uses ``coercive'' for what we wil call $2$-coercive.
-It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive, and thus the following result holds.
+It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive.
+Note that Tai \cite{Tai2003} uses ``coercive'' for what we call $2$-coercive.
+The following result holds.
 
@@
 When $f$ is monotone, \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
-Suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cK \to \cV'$.
+In fact, suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cK \to \cV'$.
 Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
@@
 \end{equation}
-The CD methods of Tai \cite{Tai2003} address problem \eqref{eq:minimization} under the hypothesis that $F'$ is coercive.
+The CD methods of Tai \cite{Tai2003} address problem \eqref{eq:minimization} under the hypothesis that $F'$ is 2-coercive.
 
@@
     $$F(v) = \int_\Omega \frac{a(x)}{p} |\grad v|^p\,dx - \ip{g}{v}.$$
-Then $F'(v) = f(v) - g$, $F$ is a convex functional (since $f$ is coercive), and, for any closed and convex $\cK\subset \cV$, VI problem \eqref{eq:vi} for is equivalent to optimization problem \eqref{eq:minimization}.\end{example}
+Then $F'(v) = f(v) - g$ and $F$ is a convex functional (since $f$ is coercive).
+For any closed and convex $\cK\subset \cV$, VI problem \eqref{eq:vi} for is equivalent to optimization problem \eqref{eq:minimization}.\end{example}
 
-The map in \eqref{eq:plaplacian} is coercive if $1<p<2$, but the proof is somewhat different \cite[Theorem 4.4]{Bueler2021conservation}.  In Section \ref{sec:results} we need only the $p\ge 2$ case.
+The map in \eqref{eq:plaplacian} is also coercive if $1<p<2$, but the proof is somewhat different \cite[Theorem 4.4]{Bueler2021conservation}.
+In Section \ref{sec:results} we need only the $p\ge 2$ case.
 
-Not all VI problems arise as in Example \ref{ex:plaplacian}, from optimization.
+However, not all VI problems arise from optimization.
 We give two such examples next, first a coercive and linear advection-diffusion problem, and then a nonlinear porous-medium-type problem; each is important in applications.
@@
 
-\section{Constraint decomposition, the basic algorithm} \label{sec:cd}
+\section{Constraint decomposition method} \label{sec:cd}
 
@@
 
-Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below state the basic CD method as iterations which improve a current iterate $u \in \cK$ by solving ``smaller'' VI problems over each subset $\cK_i$.
-The output $w\in\cK$ should be closer to the solution $u^* \in \cK$ of \eqref{eq:vi}.
+Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below state the basic CD method as functions which improve a current iterate $u \in \cK$, by solving ``smaller'' VI problems over each subset $\cK_i$, so that the output $w\in\cK$ should be closer to the solution $u^* \in \cK$ of \eqref{eq:vi}.
+While Algorithm \ref{alg:basiccd-add} generalizes a Jacobi iteration, Algorithm \ref{alg:basiccd-mult} generalizes Gauss-Seidel.
 
@@
 Note that $u-R_iu+\hat w_i$ involves removing the part of $u$ which lies in $\mathcal{K}_i$ before adding-back the improved subset solution $\hat w_i$.
-
-The following multiplicative (successive) version orders the sets $\mathcal{K}_i$, and each subset solution updates the global iterate.
-While Algorithm \ref{alg:basiccd-add} generalizes a Jacobi iteration, this one is like Gauss-Seidel.
+By contrast, the following multiplicative (successive) Algorithm \ref{alg:basiccd-mult} orders the sets $\mathcal{K}_i$, and each subset solution updates the global iterate before proceeding to the next subset.
 The sum $\sum_{j>i} R_j u$ should be read as ``keep the parts of the current iterate which we have not yet improved.''
@@
 
-Note the use of a damping parameter $0<\alpha\le 1$.
-In Algorithm \ref{alg:basiccd-add} its role is obvious, as we keep part of the current iterate if $\alpha<1$, but in Algorithm \ref{alg:basiccd-mult} we only update the iterate ``so far'', namely $w_i$, using the same damping that will be applied in generating the final output $w$.
+Note the use of a damping parameter $0<\alpha\le 1$.  In Algorithm \ref{alg:basiccd-mult} we must update the in-progress iterate $w_i$ using the same damping that is applied to the final output $w$.
 
 In each boxed VI above the argument of $f$ is an element of $\cK$, as the reader should confirm.
-Likewise, by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write the test element $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors, from $\cK$, namely
+On the other hand, by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may also write the test vector $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors from $\cK$, namely, for the two Algorithms respectively:
 \begin{align*}
 [u - R_i u + v_i] - [u - R_i u + \hat w_i] &= v_i - \hat w_i, \label{eq:admissibledifference} \\
-\left[\sum_{j<i} w_j + v_i + \sum_{j>i} R_j u\right] - \left[\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\right] &= v_i - \hat w_i,  \notag
+\left[\sum_{j<i} w_j + v_i + \sum_{j>i} R_j u\right] - \left[\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\right] &= v_i - \hat w_i.  \notag
 \end{align*}
-for the two Algorithms respectively.
 
-Define
+Now define
 \begin{equation}
-e_i = \hat w_i - R_i u \in \cV_i \label{eq:ithupdate}
+e_i = \hat w_i - R_i u \in \cV_i, \label{eq:ithupdate}
 \end{equation}
-as the $i$th subset update in either Algorithm.
-Then, as the reader may check, $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.
-The easy identity $\hat w = u^* + \sum_i \hat w_i - R_i u^*$ also holds.
+the $i$th subset increment in either Algorithm.
+As the reader may check, $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$, and $\hat w = u^* + \sum_i \hat w_i - R_i u^*$ also holds.
 We will use these identities in Section \ref{sec:convergence}.
@@
 % FOLLOWING CONTAINS ASPIRATIONS
-The convergence results in Section \ref{sec:convergence} require substantial damping when applying additive Algorithm \ref{alg:basiccd-add}, namely we require $\alpha \le 1/m$ as does \cite{Tai2003}.
-Algorithm \ref{alg:basiccd-mult} can be shown to converge without damping ($\alpha=1$).
+The convergence results in Section \ref{sec:convergence} require substantial damping when applying additive Algorithm \ref{alg:basiccd-add}, namely we require $\alpha \le 1/m$ as in \cite{Tai2003}.
+By constrast, Algorithm \ref{alg:basiccd-mult} can be shown to converge without damping ($\alpha\le 1$).
 In Section \ref{sec:results} we demonstrate practical convergence for larger ranges of $\alpha$ than suggested by the theory.
@@
 Note $e_i = \hat w_i - R_i u \in \cV_i \notin \cK$, in general, but VIs \eqref{eq:linearlocalvi} make sense because $f$ is defined over $\cV$.
-Each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $\ip{f(u)}{\cdot}$, namely as included here into the source term.
+Each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $\ip{f(u)}{\cdot}$, namely the quantity included here into the source term.
 The VI \eqref{eq:vi} is then approximately solved in any incremental and efficient manner, by computations over the basis supports.
@@
 
-The assumption that $\ip{f(\phi_i)}{z}$ can be evaluated over the support of $\phi_i$ is satisfied when $f$ is a linear differential operator, $\cV$ is an FE space, and $\phi_i$ are hat functions; see Section \ref{sec:multilevel}.
-Constrast non-local integral operators where $\ip{f(\phi_i)}{z}$ requires an integral over $\Omega$ even if $\phi_i$ is is a hat function with small support.
+The assumption that $\ip{f(\phi_i)}{z}$ can be evaluated over the support of $\phi_i$ is satisfied when $f$ is a differential operator, $\cV$ is an FE space, and $\phi_i$ are hat functions; see Section \ref{sec:multilevel}.
+Constrast non-local integral operators where $\ip{f(\phi_i)}{z}$ requires an integral over $\Omega$ even if $\phi_i$ is a hat function with small support.
 
@@
 For more difficult problems additional ideas, illustrated in Section \ref{sec:multilevel}, are needed for practical implementation.
-We will extend the basic method to nonlinear $f$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}, and the decomposed problems will only evaluate $f$ over (admissible) elements of $\mathcal{K}$.
+We will extend the basic method to nonlinear $f$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
 However, extending the algorithm to non-local functionals $f$ is a topic for future research.
@@
 \end{align}
-Finally apply \eqref{eq:normlikebound} to give \eqref{eq:core}.
+Finally apply Lemma \ref{lem:normlike} to give \eqref{eq:core}.
 \end{proof}
@@
 
-\section{Finite elements and multilevel constraint decomposition} \label{sec:multilevel}
+\section{Finite elements implementation} \label{sec:fe}
 
@@
 
-FIXME state essentially Algorithm 4.7 \cite{GraeserKornhuber2009} but with $\text{V}(\nu_1,\nu_2)$ cycles which works for linear; observe that up-smoothing is more efficient; state the FAS version which has $O(m)$ residual evaluation complexity on each level
+\section{Nonlinear multilevel constraint decomposition} \label{sec:multilevel}
+
+FIXME state the FAS version which has $O(m)$ residual evaluation complexity on each level
+
+FIXME observe that up-smoothing is more efficient
 
@@
 
-FIXME
+FIXME show results for 2D Firedrake implementation of NMCD for $p$-Laplacian obstacle problem (Example \ref{ex:plaplacian}), advection-diffusion obstacle problem (Example \ref{ex:advectiondiffusion}), and porous-media (Example \ref{ex:porous})
 
@@
 
+\appendix
+\section{Reductions of the nonlinear multilevel constraint decompostion algorithm}
+
+FIXME by removing obstacle reduce to ordinary FAS as in \cite{Bruneetal2015}
+
+FIXME by assuming $N$ is linear, reduce to Algorithm 4.7 in \cite{GraeserKornhuber2009} but with $\text{V}(\nu_1,\nu_2)$ cycles which works for linear
+
 \end{document}

### 2022-07-26 paper/mcd2.tex
start laying out the NMCD

@@
 
-The goal of this paper is to extend the constraint decomposition (CD) method of X.-C.~Tai \cite{Tai2003} to nonlinear VI problems to which it has not been applied.
-The convergence of this method, for several types of decompositions, was proven by Tai for coercive variational inequality (VI) problems which arise from minimization of a convex functional over a convex set.
+The goal of this paper is to extend the constraint decomposition (CD) method of X.-C.~Tai \cite{Tai2003} to nonlinear variational inequality (VI) problems, such as obstacle problems, to which it has not been applied.
+The convergence of this method, for several types of decompositions, was proven by Tai for coercive VI problems which arise from minimization of a convex functional over a convex set.
 In its multilevel form the method has been shown to have optimal complexity for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
@@
 
-\section{Finite elements implementation} \label{sec:fe}
+\section{Multilevel finite elements} \label{sec:fe}
 
-In practice we will solve VI \eqref{eq:vi} over a finite-dimensional space $\cV$ based on a choice of a mesh over $\Omega$ and a finite element (FE) space.
+In practice we will approximately solve VI \eqref{eq:vi} for obstacle problems over a finite-dimensional space based on a mesh over $\Omega$ and a finite element (FE) space.
 
-FIXME triangulation
+FIXME coarse triangulation, refine to level $J$, canonical restriction $R$, canonical prolongation $P$, injection $\iR$, monotone restriction $\mR$
 
-\begin{example}  FIXME if $\cV_i=\Span\{\phi_i\}$ for $\phi_i\in\cV$ are 1d spaces and if $\cK = \{v \ge \psi\} \subset \cV$ (obstacle problem), and if $R_i : \cK \to \cK_i$ for obstacle problem; note $\cK_i \not\subset \cK$ when $\psi>0$
-\end{example}
+Suppose our fine-level FE subspace is $\mathcal{V}^J \subset \mathcal{V}$, based on $J$ levels of mesh refinement as above.
+Note that the functional $g$ in \eqref{eq:vi} is defined and continuous over $\mathcal{V}^J$, that is, $g\in (\mathcal{V}^J)'$.
+If $\psi^J \in \mathcal{V}^J$ denotes the fine-level \emph{obstacle}, define
+\begin{equation}
+\mathcal{K}^J = \left\{v \in \mathcal{V}^J\,:\,v \ge \psi^J\right\}, \label{eq:fineconstraintset}
+\end{equation}
+the fine-level \emph{constraint set} of \emph{admissible} functions.
+We now assume $f$ is defined over $\mathcal{K}^J$, i.e.~$f:\mathcal{K}^J \to (\mathcal{V}^J)'$, and that it is continuous and $p$-coercive in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$.
+We seek the fine-level FE solution $u^J \in \mathcal{K}^J$ of the finite-dimensional VI
+\begin{equation}
+\ip{f(u^J)}{v-u^J} \ge \ip{g}{v-u^J} \qquad \text{for all } v\in \cK^J. \label{eq:fe:vi}
+\end{equation}
+
+As a technical matter, VI \eqref{eq:fe:vi} is not necessarily conforming for \eqref{eq:vi} even though we assume $\mathcal{V}^J \subset \mathcal{V}$ at the vector space level.
+Indeed, we do not specifically require $\mathcal{K}^J \subset \mathcal{K}$, and only the fine-level FE obstacle $\psi^J$ needs to be concretely defined here, while $\mathcal{K}$ remains abstract.
+This situation applies in certain geophysical situations where the obstacle is given by meshed data, the continuum limit of which is unknown \cite{Bueler2016}.
+
+The algorithm in the next section is designed to solve \eqref{eq:fe:vi}.
 
@@
 
-FIXME state the FAS version which has $O(m)$ residual evaluation complexity on each level
+The starting point of a \emph{multilevel constraint decomposition} \cite{GraeserKornhuber2009,Tai2003} method is to create constraint subsets, the sets denoted $\cK_i$ in sections \ref{sec:vi} and \ref{sec:convergence}, based on monotone restrictions of the constraint.
+In fact the algorithm in the current section is a multigrid V-cycle with two systems of constraint subsets, one for the downward part of the cycle and one for the upward part.
+
+To make this precise, suppose $w^J \in \cK^J$ is an admissible fine-level iterate which approximates the solution $u^J$ in some sense.
+From the iterate we define the \emph{defect constraint} \cite{GraeserKornhuber2009}
+\begin{equation}
+\chi^J = \psi^J - w^J,  \label{eq:fe:defectconstraintfcn}
+\end{equation}
+a nonpositive function.  For $j=J$ down to $j=1$ apply monotone restriction $\mR$ to define
+\begin{equation}
+\chi^{j-1} = \mR \chi^j, \label{eq:fe:chilevels}
+\end{equation}
+the \emph{level defect constraints}.  A four-level example is shown in Figure \ref{fig:chiphilevels}.  Denoting $\chi^{-1}=0$, for $j=0,\dots,J$ we also define the differences
+\begin{equation}
+\phi^j = \chi^j - \chi^{j-1}. \label{eq:fe:philevels}
+\end{equation}
+Note $\phi^0=\chi^0$ and that a telescoping sum applies, namely $\sum_{j=0}^J \phi^j = \chi^J$.
+Figure \ref{fig:chiphilevels} shows the functions $\phi^j$ as the gaps between the level defect constraints $\chi^j$.
+
+%REGENERATE using https://github.com/bueler/mg-glaciers:
+%$ cd mg-glaciers/py/
+%$ [modify decomposition() inside visualize.py to print \psi]
+%$ ./obstacle.py -J 5 -jcoarse 1 -random -randommodes 8 -diagnostics -o defect.pdf
+%fine level 5 (m=63): using 20 V(1,0) cycles -> 38.750 WU
+\begin{figure}
+\includegraphics[width=0.55\textwidth]{fixfigs/chiphilevels.pdf}
+\caption{Defect constraint $\chi^J = \psi^J - w^J$ is decomposed using monotone restriction.}
+\label{fig:chiphilevels}
+\end{figure}
+
+FIXME
+    $$\mathcal{D}^J = \{v\ge \chi^J\} \subset \mathcal{V}^J.$$
+(Recall that a perturbed iterate is admissible, $w^J + z^J \ge b^J$, if and only if the perturbation $z^J$ is in $\mathcal{D}^J$.)  Let $N^J$ be an FE approximation of the nonlinear operator $N_\eps$ in \eqref{eq:siaN}, but note that the detailed construction of $N^J$ occurs in the next subsection.
+Let $\ell^J[v] = \ip{a}{v}$.
+Our goal is to solve the fine-level FE version of VI \eqref{eq:siaVI} for an admissible perturbation $z^J$ in $\mathcal{D}^J$,
+\begin{equation}
+N^J(w^J+z^J)[v-z^J] \ge \ell^J[v-z^J] \label{eq:siaVIFEfinelevel}
+\end{equation}
+for all $v$ in $\mathcal{D}^J$.  Here $s^J=w^J+z^J$ is the desired surface elevation result, so the argument $v - z^J = (w^J+v)-s^J$ is a difference of admissible surface elevations.
+
+FIXME in Appendix we show how NMCD reduces to FAS if the obstacle is removed and to MCD if $f$ is assumed to be linear
 
@@
 
-FIXME show results for 2D Firedrake implementation of NMCD for $p$-Laplacian obstacle problem (Example \ref{ex:plaplacian}), advection-diffusion obstacle problem (Example \ref{ex:advectiondiffusion}), and porous-media (Example \ref{ex:porous})
+FIXME show results for 2D Firedrake implementation of NMCD for $p$-Laplacian obstacle problem (Example \ref{ex:plaplacian}), advection-diffusion obstacle problem (Example \ref{ex:advectiondiffusion}), and porous-media obstacle problem (Example \ref{ex:porous})
 

### 2022-07-26 paper/mcd2.tex
at least I can describe the algorithm I want as clearly as possible

@@
 \usepackage{booktabs}
+
+\usepackage{tabto}
+\TabPositions{1.5cm}
 
@@
 Let $\cK \subset \cV$ be a nonempty closed and convex subset, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.
-For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and \emph{source functional} $g\in \cV'$ we consider the following \emph{variational inequality} (VI) for the (exact) solution $u^*\in \cK$, if it exists:
+
+For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and a linear \emph{source functional} $\ell\in \cV'$ we consider the following \emph{variational inequality} (VI) for the (exact) solution $u^*\in \cK$, if it exists:
 \begin{equation}
-\ip{f(u^*)}{v-u^*} \ge \ip{g}{v-u^*} \qquad \text{for all } v\in \cK. \label{eq:vi}
+\ip{f(u^*)}{v-u^*} \ge \ip{\ell}{v-u^*} \qquad \text{for all } v\in \cK. \label{eq:vi}
 \end{equation}
-Because $f$ is (generally) nonlinear, $g$ is not strictly needed for posing this VI,  but the presence of $g$ is helpful to algorithms in Section \ref{sec:multilevel}.
-(By redefining $f$ we may take $g=0$.)
+Because $f$ is (generally) nonlinear, the source term $\ell$ is not strictly needed for posing this VI,\footnote{By redefining $f$ we may take $\ell=0$.} but its presence is helpful to algorithms in Section \ref{sec:multilevel}.
 
-VI \eqref{eq:vi} generalizes the nonlinear system of equations $f(u^*)=g$ to the constraint set $\cK$.
-Informally, if we conceptualize the dual pairing as an inner product then \eqref{eq:vi} says that the angle between $f(u^*)-g$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
-That is, \eqref{eq:vi} says that $f(u^*)-g$ may not be zero but it points directly into $\cK$.
-However, if $u^*$ is in the interior $\cK^\circ$ then \eqref{eq:vi} implies $f(u^*)=g$.
+VI \eqref{eq:vi} generalizes the nonlinear system of equations $f(u^*)=\ell$ to the constraint set $\cK$.
+Informally, if we conceptualize the dual pairing as an inner product then \eqref{eq:vi} says that the angle between $f(u^*)-\ell$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
+That is, \eqref{eq:vi} says that $f(u^*)-\ell$ may not be zero but it points directly into $\cK$.
+However, if $u^*$ is in the interior $\cK^\circ$ then \eqref{eq:vi} implies $f(u^*)=\ell$.
 
@@
 Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
-Furthermore, Proposition II.2.1 in \cite{EkelandTemam1976} shows that if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ and $g=0$ if and only if
+Furthermore, Proposition II.2.1 in \cite{EkelandTemam1976} shows that if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ and $\ell=0$ if and only if
 \begin{equation}
@@
 \begin{pseudo*}
-\pr{cd-add}(u\in\cK)\text{:} \\+
+\pr{cd-add}(f,\ell,u,\cK,\cK_i,R_i)\text{:} \\+
     for $i \in \{0,\dots,m-1\}$: \\+
         \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$ \rm{we have} \\+
-            $\boxed{{\large \strut} \ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{g}{v_i-\hat w_i}}$ \\--
+            $\boxed{{\large \strut} \ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}}$ \\--
     $\hat w = \sum_i \hat w_i\in\cK$ \\
@@
 \begin{pseudo*}
-\pr{cd-mult}(u\in\cK)\text{:} \\+
+\pr{cd-mult}(f,\ell,u,\cK,\cK_i,R_i)\text{:} \\+
     for $i = 0,\dots,m-1$: \\+
@@
 Note that the functional $g$ in \eqref{eq:vi} is defined and continuous over $\mathcal{V}^J$, that is, $g\in (\mathcal{V}^J)'$.
-If $\psi^J \in \mathcal{V}^J$ denotes the fine-level \emph{obstacle}, define
+If $\gamma^J \in \mathcal{V}^J$ denotes the fine-level \emph{obstacle} or \emph{constraint}, define
 \begin{equation}
-\mathcal{K}^J = \left\{v \in \mathcal{V}^J\,:\,v \ge \psi^J\right\}, \label{eq:fineconstraintset}
+\mathcal{K}^J = \left\{v \in \mathcal{V}^J\,:\,v \ge \gamma^J\right\}, \label{eq:fineconstraintset}
 \end{equation}
@@
 As a technical matter, VI \eqref{eq:fe:vi} is not necessarily conforming for \eqref{eq:vi} even though we assume $\mathcal{V}^J \subset \mathcal{V}$ at the vector space level.
-Indeed, we do not specifically require $\mathcal{K}^J \subset \mathcal{K}$, and only the fine-level FE obstacle $\psi^J$ needs to be concretely defined here, while $\mathcal{K}$ remains abstract.
+Indeed, we do not specifically require $\mathcal{K}^J \subset \mathcal{K}$, and only the fine-level FE obstacle $\gamma^J$ needs to be concretely defined here, while $\mathcal{K}$ remains abstract.
 This situation applies in certain geophysical situations where the obstacle is given by meshed data, the continuum limit of which is unknown \cite{Bueler2016}.
@@
 \begin{equation}
-\chi^J = \psi^J - w^J,  \label{eq:fe:defectconstraintfcn}
+\chi^J = \gamma^J - w^J,  \label{eq:fe:defectconstraintfcn}
 \end{equation}
-a nonpositive function.  For $j=J$ down to $j=1$ apply monotone restriction $\mR$ to define
+a nonpositive function.
+The idea of the defect constraint is that a perturbed iterate is admissible, e.g.~$w^J + z^J \ge \psi^J$, if and only if the perturbation $z^J$ exceeds the defect constraint, $z^J \ge \psi^j$.
+Now for $j=J$ down to $j=1$ apply monotone restriction $\mR$ to define
 \begin{equation}
@@
 \end{equation}
-the \emph{level defect constraints}.  A four-level example is shown in Figure \ref{fig:chiphilevels}.  Denoting $\chi^{-1}=0$, for $j=0,\dots,J$ we also define the differences
+called the \emph{level defect constraints}.  A four-level example is shown in Figure \ref{fig:chiphilevels}.
+
+%REGENERATE using https://github.com/bueler/mg-glaciers:
+%$ cd mg-glaciers/py/
+%$ [modify decomposition() inside visualize.py to turn off axes and print \gamma for obstacle]
+%$ ./obstacle.py -J 5 -jcoarse 1 -random -randommodes 8 -diagnostics -o defect.pdf
+%fine level 5 (m=63): using 20 V(1,0) cycles -> 38.750 WU
+\begin{figure}
+\includegraphics[width=0.55\textwidth]{fixfigs/chiphilevels.pdf}
+\caption{Defect constraint $\chi^J = \gamma^J - w^J$ is decomposed using monotone restriction.}
+\label{fig:chiphilevels}
+\end{figure}
+
+Denoting $\chi^{-1}=0$, for $j=0,\dots,J$ we also define the differences
 \begin{equation}
@@
 
-%REGENERATE using https://github.com/bueler/mg-glaciers:
-%$ cd mg-glaciers/py/
-%$ [modify decomposition() inside visualize.py to print \psi]
-%$ ./obstacle.py -J 5 -jcoarse 1 -random -randommodes 8 -diagnostics -o defect.pdf
-%fine level 5 (m=63): using 20 V(1,0) cycles -> 38.750 WU
-\begin{figure}
-\includegraphics[width=0.55\textwidth]{fixfigs/chiphilevels.pdf}
-\caption{Defect constraint $\chi^J = \psi^J - w^J$ is decomposed using monotone restriction.}
-\label{fig:chiphilevels}
-\end{figure}
+For the downward part of the multigrid V-cycle we will use the $\phi^j$ as constraints, and for the upward part the $\chi^j$.  That is, we define constraint subsets
+\begin{equation}
+\mathcal{D}^j = \{v\ge \phi^j\}, \qquad \mathcal{U}^j = \{v\ge \chi^j\}. \label{eq:fe:constraintsets}
+\end{equation}
+For $j=0,\dots,J$ these are subsets of the FE spaces $\mathcal{V}^j$.
+The upward sets are larger, and in fact the decomposition $\mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i$ applies for each $j$ in the same sense as \eqref{eq:constraintdecomp}.
 
-FIXME
-    $$\mathcal{D}^J = \{v\ge \chi^J\} \subset \mathcal{V}^J.$$
-(Recall that a perturbed iterate is admissible, $w^J + z^J \ge b^J$, if and only if the perturbation $z^J$ is in $\mathcal{D}^J$.)  Let $N^J$ be an FE approximation of the nonlinear operator $N_\eps$ in \eqref{eq:siaN}, but note that the detailed construction of $N^J$ occurs in the next subsection.
-Let $\ell^J[v] = \ip{a}{v}$.
-Our goal is to solve the fine-level FE version of VI \eqref{eq:siaVI} for an admissible perturbation $z^J$ in $\mathcal{D}^J$,
-\begin{equation}
-N^J(w^J+z^J)[v-z^J] \ge \ell^J[v-z^J] \label{eq:siaVIFEfinelevel}
-\end{equation}
-for all $v$ in $\mathcal{D}^J$.  Here $s^J=w^J+z^J$ is the desired surface elevation result, so the argument $v - z^J = (w^J+v)-s^J$ is a difference of admissible surface elevations.
+FIXME explain $g$ and $y $ and $z$ etc.; explain smoother semantics (in-place on final argument)
+
+\pseudoset{
+ctfont=\textsl,%
+ct-left=\qquad\qquad(,%
+ct-right=),%
+}
+
+\begin{pseudofloat}[H]
+\begin{pseudo*} \label{ps:nmcd-vcycle}
+\pr{nmcd-vcycle}(f,\ell,w^J,\gamma^J)\text{:} \\+
+    $\ell^J = \ell$ \ct{notational} \\
+    $\chi^J = w^J - \gamma^J$ \\
+    $g^J = w^J$ \ct{notational} \\
+    for $j=J$ downto $j=1$ \\+
+      $\chi^{j-1} = \mR \chi^j$ \\
+      $\phi^j = \chi^j - P\chi^{j-1}$ \\
+      $y^j = 0$ \\
+      $\text{\pr{smoother}}^{\text{\id{down}}}(f,\ell^j,\phi^j,g^j;y^j)$  \ct{smoothing in $\mathcal{D}^j$}\\
+      $g^{j-1} = \iR(g^j + y^j)$ \\
+      $\ell^{j-1} = f(g^{j-1}) - R \left(f(g^j+y^j) - \ell^j\right)$ \\-
+    $y^0 = 0$ \\
+    $\text{\pr{solver}}^{\text{\id{coarse}}}(f,\ell^0,\chi^0,g^0;y^0)$ \ct{solving in $\mathcal{D}^0=\mathcal{U}^0$} \\
+    $z^0 = y^0$ \\
+    for $j=1$ to $j=J$ \\+
+      $z^j = P z^{j-1} + y^{j}$ \\
+      $\text{\pr{smoother}}^{\text{\id{up}}}(f,\ell^j,\chi^j,g^j;z^j)$  \ct{smoothing in $\mathcal{U}^j$} \\-
+    return $w^J+z^J$
+\end{pseudo*}
+\caption{Nonlinear multilevel constraint decomposition V-cycle for VI problem \eqref{eq:fe:vi}.}
+\label{alg:nmcd}
+\end{pseudofloat}
 
@@
 \appendix
-\section{Reductions of the nonlinear multilevel constraint decompostion algorithm}
+\section{Reductions of the NMCD algorithm}
 
-FIXME by removing obstacle reduce to ordinary FAS as in \cite{Bruneetal2015}
+FIXME by removing obstacle, reduce to ordinary FAS as in \cite{Bruneetal2015}
 

### 2022-07-27 paper/mcd2.tex
work on constructing NMCD

@@
 
-FIXME coarse triangulation, refine to level $J$, canonical restriction $R$, canonical prolongation $P$, injection $\iR$, monotone restriction $\mR$
+FIXME coarse triangulation, refine to level $J$
+
+FIXME nesting
+\begin{equation}
+\mathcal{V}^0 \subset \mathcal{V}^1 \subset \dots \subset \mathcal{V}^J \subset \mathcal{V} \label{eq:fe:nestedspaces}
+\end{equation}
 
@@
 
-The algorithm in the next section is designed to solve \eqref{eq:fe:vi}.
+The algorithm in the next section is designed to solve \eqref{eq:fe:vi}, but as it is multilevel it requires transfer operators between meshes.
+In fact we require four operators, the first three of which are standard.
+
+FIXME canonical (dual) restriction $R$, canonical prolongation $P$, injection $\iR$, monotone restriction $\mR$
 
@@
 
-The starting point of a \emph{multilevel constraint decomposition} \cite{GraeserKornhuber2009,Tai2003} method is to create constraint subsets, the sets denoted $\cK_i$ in sections \ref{sec:vi} and \ref{sec:convergence}, based on monotone restrictions of the constraint.
-In fact the algorithm in the current section is a multigrid V-cycle with two systems of constraint subsets, one for the downward part of the cycle and one for the upward part.
+The starting point of a \emph{multilevel constraint decomposition} (MCD) \cite{GraeserKornhuber2009,Tai2003} method is to define constraint subsets, the sets generically denoted $\cK_i$ in sections \ref{sec:vi} and \ref{sec:convergence}, based on monotone restrictions of the constraint.
+However, our algorithm is a multigrid V-cycle with two types of constraint subsets, for the downward and upward parts of the cycle.
 
 To make this precise, suppose $w^J \in \cK^J$ is an admissible fine-level iterate which approximates the solution $u^J$ in some sense.
-From the iterate we define the \emph{defect constraint} \cite{GraeserKornhuber2009}
+From this we define the \emph{defect constraint} \cite{GraeserKornhuber2009}
 \begin{equation}
@@
 a nonpositive function.
-The idea of the defect constraint is that a perturbed iterate is admissible, e.g.~$w^J + z^J \ge \psi^J$, if and only if the perturbation $z^J$ exceeds the defect constraint, $z^J \ge \psi^j$.
-Now for $j=J$ down to $j=1$ apply monotone restriction $\mR$ to define
+The motivation of the defect constraint is that a corrected iterate is admissible, $w^J + z^J \ge \psi^J$, if and only if the perturbation $z^J$ exceeds the defect constraint, $z^J \ge \psi^j$.
+Next, for $j=J$ down to $j=1$ apply monotone restriction $\mR$ to define
 \begin{equation}
@@
 \end{equation}
-called the \emph{level defect constraints}.  A four-level example is shown in Figure \ref{fig:chiphilevels}.
+the \emph{level defect constraints}.  A $J=4$ example is shown in Figure \ref{fig:chiphilevels}.  Denoting $\chi^{-1}=0$, for $j=0,\dots,J$ we also define the differences
+\begin{equation}
+\phi^j = \chi^j - \chi^{j-1}. \label{eq:fe:philevels}
+\end{equation}
+(The functions $\phi^j$ are the gaps between the $\chi^j$ in Figure \ref{fig:chiphilevels}.)  Note $\phi^0=\chi^0$, and that the telescoping sum $\sum_{j=0}^J \phi^j = \chi^J$ applies.
 
@@
 
-Denoting $\chi^{-1}=0$, for $j=0,\dots,J$ we also define the differences
+For the downward part of the multigrid V-cycle we will use the $\phi^j$ as constraints, and for the upward part the $\chi^j$.  That is,
 \begin{equation}
-\phi^j = \chi^j - \chi^{j-1}. \label{eq:fe:philevels}
+\mathcal{D}^j = \{v\ge \phi^j\}, \qquad \mathcal{U}^j = \{v\ge \chi^j\} \label{eq:fe:constraintsets}
 \end{equation}
-Note $\phi^0=\chi^0$ and that a telescoping sum applies, namely $\sum_{j=0}^J \phi^j = \chi^J$.
-Figure \ref{fig:chiphilevels} shows the functions $\phi^j$ as the gaps between the level defect constraints $\chi^j$.
+define \emph{downward} and \emph{upward constraint subsets} of the FE subspaces $\mathcal{V}^j$, respectively.
+The upward sets are larger, and decompositions $\mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i$ apply in the same sense as \eqref{eq:constraintdecomp}.
 
-For the downward part of the multigrid V-cycle we will use the $\phi^j$ as constraints, and for the upward part the $\chi^j$.  That is, we define constraint subsets
+Returning to the heart of the MCD idea, our plan is to modify the admissible fine-level iterate $w^J\in \mathcal{K}^J$ by adding corrections from each coarser subspace $\mathcal{V}^j$.
+By nesting \eqref{eq:fe:nestedspaces} these corrections do indeed live in the fine-level subspace $\mathcal{V}^J$, but the corrected iterate needs to be admissible too.
+The accumulated corrections, over the whole multigrid V-cycle, must compute a new iterate in $\mathcal{K}^J$, so our corrections will come from the constraint subsets $\mathcal{D}^j$ and $\mathcal{U}^j$.
+
+In detail, suppose $y^{j+1} \in \mathcal{D}^{j+1}, \dots, y^J \in \mathcal{D}^J$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:nmcdvcycle}).
+The corrected iterate at this stage is admissible ($w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$) because the sum of corrections is in the largest defect constraint set ($y^J + \dots + y^{j+1} \in \mathcal{U}^J = \{v\ge \chi^J\}$).
+
+\begin{figure}[h]
+\begin{center}
+\input{tikz/nmcdvcycle.tex}
+\end{center}
+\caption{Going down, the NMCD V-cycle computes corrections $y_j \in \mathcal{D}^j$.  The upward corrections $z_j$ can expand into larger constraint subsets $\mathcal{U}^j$.}
+\label{fig:nmcdvcycle}
+\end{figure}
+
+FIXME do I need $f^j$ to be defined?  yes?  even though $f(w^J+y^J + \dots + y^{j+1} +y^j)$ is the same as $f^j(w^J+y^J + \dots + y^{j+1} +y^j)$, the latter can use less storage?
+
+FIXME Now we compute the next-coarser correction $y^j \in \mathcal{D}^j$ by the downward VI at the $j$th level,
 \begin{equation}
-\mathcal{D}^j = \{v\ge \phi^j\}, \qquad \mathcal{U}^j = \{v\ge \chi^j\}. \label{eq:fe:constraintsets}
+\ip{f(w^J+y^J + \dots + y^{j+1} +y^j)}{v-y^j} \ge \ip{\ell^{j+1}}{v-y^j}, \label{eq:fe:levelvi}
 \end{equation}
-For $j=0,\dots,J$ these are subsets of the FE spaces $\mathcal{V}^j$.
-The upward sets are larger, and in fact the decomposition $\mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i$ applies for each $j$ in the same sense as \eqref{eq:constraintdecomp}.
+for all $v \in \mathcal{D}^j$.  The source functional $\ell^j$ remains to be computed; see below.  Note that here and below we guarantee that the input to $f$ is admissible (in $\mathcal{K}^J$).
 
-FIXME explain $g$ and $y $ and $z$ etc.; explain smoother semantics (in-place on final argument)
+FIXME explain $g$; we define $g^j$ as the latest $j$th-level approximation to the solution $s$:
+\begin{equation}
+g^j = \begin{cases} w^J, & j=J \\
+                    \iR(g^{j+1} + y^{j+1}), & j < J.
+      \end{cases}  \label{eq:fe:defineg}
+\end{equation}
+That is, $g^j$ is our estimate of the solution on the $j$th level, before we add a correction which solves the $j$th-level VI.
+
+FIXME Subtract $N^{j+1}(w^J+z^{j+1})$ from both sides of \eqref{eq:fe:levelvi} gives
+\begin{align}
+&N^{j+1}(w^J+z^{j+1}+y)[v-y] \label{eq:siaVIFEprecoarsetwo} \\
+&\quad - N^{j+1}(w^J+z^{j+1})[v-y] \ge - \tilde F^{j+1}[v-y], \notag
+\end{align}
+where we define the resulting residual as
+\begin{equation}
+\tilde F^{j+1}[v] = N^{j+1}(w^J+z^{j+1})[v] - \ell^{j+1}[v].
+\end{equation}
+
+FIXME explain smoother semantics (in-place on final argument; reduces CP residual)
 

### 2022-07-27 paper/mcd2.tex
work done

@@
 
-In practice we will approximately solve VI \eqref{eq:vi} for obstacle problems over a finite-dimensional space based on a mesh over $\Omega$ and a finite element (FE) space.
+We will approximately solve VI \eqref{eq:vi} for certain nonlinear obstacle problems, over two-dimensional computational domains $\Omega \subset \RR^2$ and Banach space $\mathcal{V}=W_0^{1,p}(\Omega)$, using finite element (FE) spaces over triangular meshes (for concreteness).
+The targeted mesh will be the finest level in a hierarchy constructed by refinement of a coarsest mesh.
 
-FIXME coarse triangulation, refine to level $J$
+The mesh hierarchy and FE spaces are constructed in a standard manner \cite{Elmanetal2014}.
+Suppose $\Omega$ is polygonal and triangulated by $\mathcal{T}^0$, the \emph{coarsest mesh}, a finite set of non-overlapping triangles with union $\overline{\Omega}$ and characteristic size (cell diameter) $h_0>0$.
+Starting with $j=1$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ by edge bisection, thus each triangle becomes four similar triangles in $\mathcal{T}^j$ with characteristic size $h_j = 2^{-j} h_0$.
+For $J\ge 0$, the \emph{depth} of the hierarchy, the final triangulation $\mathcal{T}^J$ is the \emph{finest mesh}.
+(In the single-level $J=0$ case all multilevel efficiency will be lost.)  We define $m_j$ to be the number of interior nodes in $\mathcal{T}^j$, the number of \emph{degrees of freedom} in the $j$th level, and we denote $h=h_J=2^{-J} h_0$ and $m=m_J$ for the finest level.
 
-FIXME nesting
+Let $\mathcal{V}^j$ be the continuous, piecewise-polynomial FE space $P_k$, over $\mathcal{T}^j$, with zero boundary values, a subspace of $\mathcal{V}=W_0^{1,p}(\Omega)$ \cite{Elmanetal2014}.
+Our choices imply the following nesting:
 \begin{equation}
-\mathcal{V}^0 \subset \mathcal{V}^1 \subset \dots \subset \mathcal{V}^J \subset \mathcal{V} \label{eq:fe:nestedspaces}
+\mathcal{V}^0 \subset \mathcal{V}^1 \subset \dots \subset \mathcal{V}^J \subset \mathcal{V}.  \label{eq:fe:nestedspaces}
 \end{equation}
-
-Suppose our fine-level FE subspace is $\mathcal{V}^J \subset \mathcal{V}$, based on $J$ levels of mesh refinement as above.
-Note that the functional $g$ in \eqref{eq:vi} is defined and continuous over $\mathcal{V}^J$, that is, $g\in (\mathcal{V}^J)'$.
-If $\gamma^J \in \mathcal{V}^J$ denotes the fine-level \emph{obstacle} or \emph{constraint}, define
+Suppose $\gamma^J \in \mathcal{V}^J$ denotes the \emph{obstacle} and define
 \begin{equation}
@@
 \end{equation}
-the fine-level \emph{constraint set} of \emph{admissible} functions.
-We now assume $f$ is defined over $\mathcal{K}^J$, i.e.~$f:\mathcal{K}^J \to (\mathcal{V}^J)'$, and that it is continuous and $p$-coercive in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$.
-We seek the fine-level FE solution $u^J \in \mathcal{K}^J$ of the finite-dimensional VI
+the finest-level \emph{constraint set} of \emph{admissible} FE functions.
+
+One may assume that $f$ in \eqref{eq:vi} is defined over $\mathcal{K}^J$, i.e.~$f:\mathcal{K}^J \to \mathcal{V}'$.
+However, especially with nontrivial nonlinear functionals, we assume instead that the finest-level FE functional $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ only approximates $f$ because of quadrature and other approximations.
+We assume that $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$ in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$.
+The functional $\ell$ in \eqref{eq:vi} is defined and continuous over $\mathcal{V}^J$, that is, $\ell\in (\mathcal{V}^J)'$, but we allow the possibility of quadrature etc.~and therefore denote $\ell^J$ for the implemented source term.
+
+With these understandings we seek the finest-level FE solution $u^J \in \mathcal{K}^J$ of the finite-dimensional, nonlinear VI
 \begin{equation}
-\ip{f(u^J)}{v-u^J} \ge \ip{g}{v-u^J} \qquad \text{for all } v\in \cK^J. \label{eq:fe:vi}
+\ip{f^J(u^J)}{v-u^J} \ge \ip{\ell^J}{v-u^J} \qquad \text{for all } v\in \cK^J. \label{eq:fe:vi}
 \end{equation}
-
-As a technical matter, VI \eqref{eq:fe:vi} is not necessarily conforming for \eqref{eq:vi} even though we assume $\mathcal{V}^J \subset \mathcal{V}$ at the vector space level.
-Indeed, we do not specifically require $\mathcal{K}^J \subset \mathcal{K}$, and only the fine-level FE obstacle $\gamma^J$ needs to be concretely defined here, while $\mathcal{K}$ remains abstract.
+While \eqref{eq:fe:vi} certainly approximates \eqref{eq:vi}, even ignoring quadrature it may not technically be conforming for \eqref{eq:vi}, despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$.
+Indeed, we do not require $\mathcal{K}^J \subset \mathcal{K}$, and only the fine-level FE obstacle $\gamma^J$ needs to be concretely defined here, while $\mathcal{K}$ remains abstract.
 This situation applies in certain geophysical situations where the obstacle is given by meshed data, the continuum limit of which is unknown \cite{Bueler2016}.
 
-The algorithm in the next section is designed to solve \eqref{eq:fe:vi}, but as it is multilevel it requires transfer operators between meshes.
-In fact we require four operators, the first three of which are standard.
+The multilevel algorithm in the next section, designed to solve \eqref{eq:fe:vi}, requires transfer operators between meshes.
+In fact we require four such operators, the first three of which are standard, linear maps used for nonlinear multigrid methods \cite{Trottenbergetal2001}.
 
-FIXME canonical (dual) restriction $R$, canonical prolongation $P$, injection $\iR$, monotone restriction $\mR$
+Canonical prolongation $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ simply acts as the identity on coarser-mesh functions in $\mathcal{V}^{j-1}$; it is the vector space injection $\mathcal{V}^{j-1} \hookrightarrow \mathcal{V}^j$.
+Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ also acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $(R\sigma)[z] = \sigma[z]$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
+While these canonical transfers imply no loss of information, in exact arithmetic, they do involve a change of FE representation.
+The practical formula for canonical restriction is the ``full-weighting'' average \cite{Trottenbergetal2001}, the transpose of canonical prolongation.
+
+To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ we first denote by $x_q^j$, for $q=0,\dots,m_j-1$, the \emph{interior nodes} of the $j$th triangulation $\mathcal{T}^j$.
+Observe that for any given values at the (interior) nodes of $\mathcal{T}^j$ there is a unique function in $\mathcal{V}^j$ with those values, and also observe that, by construction, the nodes of $\mathcal{T}^{j-1}$ are also nodes of $\mathcal{T}^j$.
+Thus for $z\in\mathcal{V}^j$ we may define the injection $\iR z$ as the element of $\mathcal{V}^{j-1}$ which has the same nodal values as $z$ at the nodes of $\mathcal{T}^{j-1}$.
+
+Finally, monotone restriction $\mR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ is a nonlinear map which replaces a function by the 
+\begin{equation}
+  \mR z = \sum_{p=1}^{m_{j-1}} \max \{z_q \,:\, \psi_p^{j-1}(x_q^j) > 0\}\, \psi_p^{j-1} \label{eq:fe:monotonerestriction}
+\end{equation}
+Recalling that functions in $\mathcal{V}^j$ have zero boundary values, note that $\mR z \ge z$ on $\Omega$ if $z\le 0$.
+
+FIXME I NEED TO BE CAREFUL WITH $P_k$ for $k>1$
+
+Table \ref{tab:transfers} will help the reader keep track of transfer operator notation.
+
+\begin{table}
+\begin{tabular}{lll}
+\emph{name}  & \emph{action}  & \emph{linear?} \\ \hline
+canonical prolongation        & $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ & \,\checkmark \\
+canonical (dual) restriction  & $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ & \,\checkmark \\
+(nodal) injection restriction & $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ & \,\checkmark \\
+monotone restriction          & $\mR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ &  
+\end{tabular}
+
+\medskip
+\caption{Transfer operators used in this paper.}
+\label{tab:transfers}
+\end{table}
 
@@
 
-The starting point of a \emph{multilevel constraint decomposition} (MCD) \cite{GraeserKornhuber2009,Tai2003} method is to define constraint subsets, the sets generically denoted $\cK_i$ in sections \ref{sec:vi} and \ref{sec:convergence}, based on monotone restrictions of the constraint.
-However, our algorithm is a multigrid V-cycle with two types of constraint subsets, for the downward and upward parts of the cycle.
+The starting point of a \emph{multilevel constraint decomposition} (MCD) \cite{GraeserKornhuber2009,Tai2003} method is to define constraint subsets, the sets generically denoted $\cK_i$ in sections \ref{sec:vi} and \ref{sec:convergence}, based on monotone restrictions of the constraint, as defined momentarily.
+Our algorithm is a multilevel V-cycle with two types of constraint subsets, distinct for the downward and upward parts of the cycle.
 
-To make this precise, suppose $w^J \in \cK^J$ is an admissible fine-level iterate which approximates the solution $u^J$ in some sense.
-From this we define the \emph{defect constraint} \cite{GraeserKornhuber2009}
+To make this precise, suppose $w^J \in \cK^J$ is an admissible fine-level iterate which approximates the solution $u^J$.
+From this iterate we define the \emph{defect constraint} \cite{GraeserKornhuber2009}
 \begin{equation}
@@
 a nonpositive function.
-The motivation of the defect constraint is that a corrected iterate is admissible, $w^J + z^J \ge \psi^J$, if and only if the perturbation $z^J$ exceeds the defect constraint, $z^J \ge \psi^j$.
-Next, for $j=J$ down to $j=1$ apply monotone restriction $\mR$ to define
+The motivation of the defect constraint is that a corrected iterate is admissible, $w^J + z^J \ge \gamma^J$, if and only if the perturbation $z^J$ exceeds the defect constraint, $z^J \ge \chi^J$.
+Next, for $j=J,J-1,\dots,1$ apply monotone restriction $\mR$, equation \eqref{eq:fe:monotonerestriction}, to define
 \begin{equation}

### 2022-07-28 paper/mcd2.tex
getting pretty clear to me

--- 2022-07-27 paper/mcd2.tex
@@
 
-In practice we will approximately solve VI \eqref{eq:vi} for obstacle problems over a finite-dimensional space based on a mesh over $\Omega$ and a finite element (FE) space.
+We will approximately solve VI \eqref{eq:vi} for certain nonlinear obstacle problems, over two-dimensional computational domains $\Omega \subset \RR^2$ and Banach space $\mathcal{V}=W_0^{1,p}(\Omega)$, using piecewise-linear finite element (FE) spaces over triangular meshes.
+The targeted mesh will be the finest level in a hierarchy of meshes constructed by refinement of a coarsest mesh.
 
-FIXME coarse triangulation, refine to level $J$
+The mesh hierarchy and FE spaces are constructed in a standard manner \cite{Elmanetal2014}.
+Suppose $\Omega$ is polygonal and triangulated by $\mathcal{T}^0$, the \emph{coarsest mesh}, a finite set of non-overlapping triangles with union $\overline{\Omega}$ and characteristic size (cell diameter) $h_0>0$.
+Starting with $j=1$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ by edge bisection, thus each triangle becomes four similar triangles in $\mathcal{T}^j$ with characteristic size $h_j = 2^{-j} h_0$.
+Refinement proceeds to level $J\ge 0$, called the \emph{depth} of the hierarchy, giving $\mathcal{T}^J$ as the \emph{finest mesh}.
+(In the single-level $J=0$ case all multilevel efficiency will be lost.)  Let $m_j$ be the number of interior nodes in $\mathcal{T}^j$, the \emph{degrees of freedom} at the $j$th level, and denote $h=h_J=2^{-J} h_0$ and $m=m_J$ for the finest level.
 
-FIXME nesting
+Let $\mathcal{V}^j$ be the continuous, piecewise-polynomial FE space $P_1$,\footnote{The FE method herein apparently extends to $Q_1$ spaces over quadrilateral meshes \cite{Elmanetal2014}.
+However, for $k>1$ the $P_k$ space seems not to be suitable because the monotone restriction operator, \eqref{eq:fe:monotonerestriction} below, depends on element-wise monotonicity.
+That is, for $P_1$ or $Q_1$ if the nodal values of an FE function (non-strictly) increase then the FE function also increases.
+As shown by Figure 1.7 of \cite{Elmanetal2014}, for example, such monotonicity is not true for $P_2$.} over $\mathcal{T}^j$, with zero boundary values.
+Note $\mathcal{V}^j$ is a subspace of $\mathcal{V}=W_0^{1,p}(\Omega)$ for any $p\ge 1$ \cite{Elmanetal2014}.
+Our choices imply the following nesting:
 \begin{equation}
-\mathcal{V}^0 \subset \mathcal{V}^1 \subset \dots \subset \mathcal{V}^J \subset \mathcal{V} \label{eq:fe:nestedspaces}
+\mathcal{V}^0 \subset \mathcal{V}^1 \subset \dots \subset \mathcal{V}^J \subset \mathcal{V}.  \label{eq:fe:nestedspaces}
 \end{equation}
-
-Suppose our fine-level FE subspace is $\mathcal{V}^J \subset \mathcal{V}$, based on $J$ levels of mesh refinement as above.
-Note that the functional $g$ in \eqref{eq:vi} is defined and continuous over $\mathcal{V}^J$, that is, $g\in (\mathcal{V}^J)'$.
-If $\gamma^J \in \mathcal{V}^J$ denotes the fine-level \emph{obstacle} or \emph{constraint}, define
+Suppose $\gamma^J \in \mathcal{V}^J$ denotes the \emph{obstacle} and define
 \begin{equation}
@@
 \end{equation}
-the fine-level \emph{constraint set} of \emph{admissible} functions.
-We now assume $f$ is defined over $\mathcal{K}^J$, i.e.~$f:\mathcal{K}^J \to (\mathcal{V}^J)'$, and that it is continuous and $p$-coercive in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$.
-We seek the fine-level FE solution $u^J \in \mathcal{K}^J$ of the finite-dimensional VI
+the finest-level \emph{constraint set} of \emph{admissible} FE functions.
+
+One might assume that $f$ in \eqref{eq:vi} is defined over $\mathcal{K}^J$, i.e.~$f:\mathcal{K}^J \to \mathcal{V}'$.
+However, especially for nontrivial nonlinear functionals, the finest-level, discretized FE functional $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ is likely to only approximate $f$ because of quadrature, or other implementation details.
+The functional $\ell$ in \eqref{eq:vi} is defined and continuous over $\mathcal{V}^J$, thus $\ell\in (\mathcal{V}^J)'$, and we assume for simplicity that it can be computed exactly, but we denote $\ell^J \in (\mathcal{V}^J)'$ to indicate that it must be represented in the FE sense.
+
+With these understandings we seek the finest-level FE solution $u^J \in \mathcal{K}^J$ of this finite-dimensional, nonlinear VI
 \begin{equation}
-\ip{f(u^J)}{v-u^J} \ge \ip{g}{v-u^J} \qquad \text{for all } v\in \cK^J. \label{eq:fe:vi}
+\ip{f^J(u^J)}{v-u^J} \ge \ip{\ell^J}{v-u^J} \qquad \text{for all } v\in \cK^J. \label{eq:fe:vi}
 \end{equation}
+As for \eqref{eq:vi}, if $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, or $f^J$ is continuous, coercive and strictly monotone, then \eqref{eq:fe:vi} is well-posed \cite{KinderlehrerStampacchia1980}.
 
-As a technical matter, VI \eqref{eq:fe:vi} is not necessarily conforming for \eqref{eq:vi} even though we assume $\mathcal{V}^J \subset \mathcal{V}$ at the vector space level.
-Indeed, we do not specifically require $\mathcal{K}^J \subset \mathcal{K}$, and only the fine-level FE obstacle $\gamma^J$ needs to be concretely defined here, while $\mathcal{K}$ remains abstract.
-This situation applies in certain geophysical situations where the obstacle is given by meshed data, the continuum limit of which is unknown \cite{Bueler2016}.
+While \eqref{eq:fe:vi} approximates \eqref{eq:vi}, even if quadrature and other implementation details are ignored it may not technically be conforming for \eqref{eq:vi}.
+That is, despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require $\mathcal{K}^J \subset \mathcal{K}$.
+Only the fine-level FE obstacle $\gamma^J$ needs to be defined, while $\mathcal{K}$ remains abstract.
+In fact, if a continuum obstacle exists at all then admissibility with respect to its FE interpolant $\gamma^J$ might not imply continuum admissiblity.
+These caveats apply in geophysical situations where the obstacle is given by meshed elevation data, the continuum limit of which is undefined \cite{Bueler2016}.
 
-The algorithm in the next section is designed to solve \eqref{eq:fe:vi}, but as it is multilevel it requires transfer operators between meshes.
-In fact we require four operators, the first three of which are standard.
+The multilevel algorithm in the next section, designed to solve \eqref{eq:fe:vi}, requires transfer operators between meshes.
+We use four such operators, three of which are the standard and linear maps used for nonlinear multigrid methods \cite{Trottenbergetal2001}.
+Table \ref{tab:transfers} will help the reader keep track.
 
-FIXME canonical (dual) restriction $R$, canonical prolongation $P$, injection $\iR$, monotone restriction $\mR$
+\begin{table}
+\begin{tabular}{lll}
+\emph{name}  & \emph{action}  & \emph{linear?} \\ \hline
+canonical prolongation        & $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ & \,\checkmark \\
+canonical (dual) restriction  & $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ & \,\checkmark \\
+(nodal) injection restriction & $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ & \,\checkmark \\
+monotone restriction          & $\mR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ &  
+\end{tabular}
+
+\medskip
+\caption{Transfer operators used in this paper.}
+\label{tab:transfers}
+\end{table}
+
+Canonical prolongation $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ acts as the identity on coarser-mesh functions in $\mathcal{V}^{j-1}$; it is the linear vector space injection $\mathcal{V}^{j-1} \hookrightarrow \mathcal{V}^j$.
+Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ also acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $(R\sigma)[z] = \sigma[z]$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
+While these canonical transfers do not lose information, they typically involve a change of FE representation.
+For example, the practical formula for canonical restriction, derived by writing coarser-mesh hat functions in terms of finer-mesh ones, is a ``full-weighting'' average \cite{Trottenbergetal2001}.
+It is the transpose of canonical prolongation.
+
+To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, observe that for any given values at the (interior) nodes of $\mathcal{T}^j$ there is a unique function in $\mathcal{V}^j$ with those values.
+Also observe that, by our refinement construction, the nodes of $\mathcal{T}^{j-1}$ are also nodes of $\mathcal{T}^j$.
+For $z\in\mathcal{V}^j$ we define the injection $\iR z$ as the element of $\mathcal{V}^{j-1}$ which has the same nodal values as $z$ at the nodes of $\mathcal{T}^{j-1}$.
+Note that injection clearly preserves ordering, that is, if $w \le z$ then $\iR w \le \iR z$.
+
+Finally, monotone restriction $\mR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ is a nonlinear map which maximizes nodal values over the supports of coarse-mesh hat functions.
+To make the definition precise, first denote the \emph{interior nodes} of triangulation $\mathcal{T}^j$ by $x_q^j$ for $q=0,\dots,m_j-1$, and then the corresponding hat function \cite{Elmanetal2014} by $\psi_q^j$.
+Monotone restriction of $z\in \mathcal{V}^j$ is
+\begin{equation}
+  \mR z = \sum_{p=1}^{m_{j-1}} \max \{z_q \,:\, \psi_p^{j-1}(x_q^j) > 0\}\, \psi_p^{j-1}. \label{eq:fe:monotonerestriction}
+\end{equation}
+Recalling that the piecewise-linear functions in $\mathcal{V}^j$ have zero boundary values, note that if $z\le 0$ then $0\ge \mR z \ge z$ on $\Omega$.
 
@@
 The starting point of a \emph{multilevel constraint decomposition} (MCD) \cite{GraeserKornhuber2009,Tai2003} method is to define constraint subsets, the sets generically denoted $\cK_i$ in sections \ref{sec:vi} and \ref{sec:convergence}, based on monotone restrictions of the constraint.
-However, our algorithm is a multigrid V-cycle with two types of constraint subsets, for the downward and upward parts of the cycle.
+In fact our algorithm is a multilevel V-cycle with two types of constraint subsets, distinctly for the downward and upward parts of the cycle.
 
-To make this precise, suppose $w^J \in \cK^J$ is an admissible fine-level iterate which approximates the solution $u^J$ in some sense.
-From this we define the \emph{defect constraint} \cite{GraeserKornhuber2009}
+To make this precise, suppose $w^J \in \cK^J$ is an admissible fine-level iterate which approximates the solution $u^J$.
+From this iterate we define the \emph{defect constraint} \cite{GraeserKornhuber2009}
 \begin{equation}
@@
 a nonpositive function.
-The motivation of the defect constraint is that a corrected iterate is admissible, $w^J + z^J \ge \psi^J$, if and only if the perturbation $z^J$ exceeds the defect constraint, $z^J \ge \psi^j$.
-Next, for $j=J$ down to $j=1$ apply monotone restriction $\mR$ to define
+The motivation of the defect constraint is that a corrected iterate is admissible, $w^J + z^J \ge \gamma^J$, if and only if the perturbation $z^J$ exceeds the defect constraint, $z^J \ge \chi^J$.
+Next, for $j=J,J-1,\dots,1$ apply monotone restriction $\mR$ to define the \emph{level defect constraints}
 \begin{equation}
-\chi^{j-1} = \mR \chi^j, \label{eq:fe:chilevels}
+\chi^{j-1} = \mR \chi^j. \label{eq:fe:chilevels}
 \end{equation}
-the \emph{level defect constraints}.  A $J=4$ example is shown in Figure \ref{fig:chiphilevels}.  Denoting $\chi^{-1}=0$, for $j=0,\dots,J$ we also define the differences
+As observed after equation \eqref{eq:fe:monotonerestriction}, $0\ge \chi^{j-1} \ge \chi^j$.  A $J=4$ example is shown in Figure \ref{fig:chiphilevels}.
+
+For $j=0,\dots,J$ we also denote the differences between the level defect constraints by
 \begin{equation}
@@
 \end{equation}
-(The functions $\phi^j$ are the gaps between the $\chi^j$ in Figure \ref{fig:chiphilevels}.)  Note $\phi^0=\chi^0$, and that the telescoping sum $\sum_{j=0}^J \phi^j = \chi^J$ applies.
+(This uses the convenience definition $\chi^{-1}=0$.)  These are the gaps between the $\chi^j$ shown in Figure \ref{fig:chiphilevels}.
+Note $\phi^0=\chi^0$, and that the telescoping sum $\sum_{j=0}^J \phi^j = \chi^J$ applies.
 
@@
 
-For the downward part of the multigrid V-cycle we will use the $\phi^j$ as constraints, and for the upward part the $\chi^j$.  That is,
+For the downward part of the multigrid V-cycle we will use the $\phi^j$ as constraints, and for the upward part the $\chi^j$.  That is, let
 \begin{equation}
@@
 
-Returning to the heart of the MCD idea, our plan is to modify the admissible fine-level iterate $w^J\in \mathcal{K}^J$ by adding corrections from each coarser subspace $\mathcal{V}^j$.
-By nesting \eqref{eq:fe:nestedspaces} these corrections do indeed live in the fine-level subspace $\mathcal{V}^J$, but the corrected iterate needs to be admissible too.
-The accumulated corrections, over the whole multigrid V-cycle, must compute a new iterate in $\mathcal{K}^J$, so our corrections will come from the constraint subsets $\mathcal{D}^j$ and $\mathcal{U}^j$.
+Returning to the heart of the MCD idea, our plan is to correct the admissible fine-level iterate $w^J\in \mathcal{K}^J$ by adding modifications from each coarser subspace $\mathcal{V}^j$.
+By nesting \eqref{eq:fe:nestedspaces} these corrections live in the fine-level subspace $\mathcal{V}^J$, but the corrected iterate needs to be admissible as well.
+The accumulated corrections, from the constraint subsets $\mathcal{D}^j$ and $\mathcal{U}^j$, over the whole multigrid V-cycle, compute a new iterate in $\mathcal{K}^J$.
 
 In detail, suppose $y^{j+1} \in \mathcal{D}^{j+1}, \dots, y^J \in \mathcal{D}^J$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:nmcdvcycle}).
-The corrected iterate at this stage is admissible ($w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$) because the sum of corrections is in the largest defect constraint set ($y^J + \dots + y^{j+1} \in \mathcal{U}^J = \{v\ge \chi^J\}$).
+The so-far corrected iterate is admissible, $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$, because the sum of corrections is in the (finest) defect constraint set, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J = \{v\ge \chi^J\}$.
 
@@
 
-FIXME do I need $f^j$ to be defined?  yes?  even though $f(w^J+y^J + \dots + y^{j+1} +y^j)$ is the same as $f^j(w^J+y^J + \dots + y^{j+1} +y^j)$, the latter can use less storage?
+Now, when computing the next-coarser correction we will need to use the coarser-mesh representation, and, as in nonlinear FAS multigrid schemes \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, we must coarsen the iterate as well as the residual and the correction.\footnote{Here we diverge from algorithms by Tai  \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009}.
+If $f$ is linear, as in the implemented examples in those works, there is no need to coarsen the solution.
+(In classical linear multigrid only the residual and the correction must be coarsened.)  We follow the FAS philosophy for nonlinear functionals.}  We do this using injection \cite[section 5.3]{Trottenbergetal2001}.
+Define
+\begin{equation}
+g^j = \begin{cases} w^J, & j=J \\
+                    \iR(g^{j+1} + y^{j+1}), & j < J.
+      \end{cases}  \label{eq:fe:defineg}
+\end{equation}
+That is, $g^j$ is a $j$th-level mesh approximation to the solution $u^J$, before we add the correction which solves the coarsened VI problem at the $j$th level.\footnote{If we were to inject the finest-mesh obstacle downward to coarser meshes, i.e.~$\gamma^{j-1} = \iR \gamma^j$, then, because $\iR$ preserves inequalities and $y^j \in \mathcal{D}^j$  it would follow that $g^j \ge \gamma^j$.
+In this sense $g^j$ is an admissible function on the $j$th-level mesh.
+However, our use of the defect constraints $\chi^j$, following \cite{GraeserKornhuber2009}, avoids the need to generate these $\gamma^j$; their definition in this footnote can be ignored.}
+
+FIXME do I need $f^j$ to be defined?  yes.  Tai assumes $f(w^J+y^J + \dots + y^{j+1} +y^j)$ is defined, which is true, but $f^j(g^j)$ uses the $j$th level representation
 
@@
 for all $v \in \mathcal{D}^j$.  The source functional $\ell^j$ remains to be computed; see below.  Note that here and below we guarantee that the input to $f$ is admissible (in $\mathcal{K}^J$).
-
-FIXME explain $g$; we define $g^j$ as the latest $j$th-level approximation to the solution $s$:
-\begin{equation}
-g^j = \begin{cases} w^J, & j=J \\
-                    \iR(g^{j+1} + y^{j+1}), & j < J.
-      \end{cases}  \label{eq:fe:defineg}
-\end{equation}
-That is, $g^j$ is our estimate of the solution on the $j$th level, before we add a correction which solves the $j$th-level VI.
 
@@
 \begin{pseudo*} \label{ps:nmcd-vcycle}
-\pr{nmcd-vcycle}(f,\ell,w^J,\gamma^J)\text{:} \\+
-    $\ell^J = \ell$ \ct{notational} \\
+\pr{nmcd-vcycle}(\ell^J,w^J,\gamma^J)\text{:} \\+
     $\chi^J = w^J - \gamma^J$ \\
-    $g^J = w^J$ \ct{notational} \\
+    $g^J = w^J$ \ct{notational; no copy needed} \\
     for $j=J$ downto $j=1$ \\+
@@
       $y^j = 0$ \\
-      $\text{\pr{smoother}}^{\text{\id{down}}}(f,\ell^j,\phi^j,g^j;y^j)$  \ct{smoothing in $\mathcal{D}^j$}\\
+      $\text{\pr{smoother}}^{\text{\id{down}}}(f^j,\ell^j,\phi^j,g^j;y^j)$  \ct{smoothing in $\mathcal{D}^j$}\\
       $g^{j-1} = \iR(g^j + y^j)$ \\
-      $\ell^{j-1} = f(g^{j-1}) - R \left(f(g^j+y^j) - \ell^j\right)$ \\-
+      $\ell^{j-1} = f^{j-1}(g^{j-1}) - R \left(f^j(g^j+y^j) - \ell^j\right)$ \\-
     $y^0 = 0$ \\
-    $\text{\pr{solver}}^{\text{\id{coarse}}}(f,\ell^0,\chi^0,g^0;y^0)$ \ct{solving in $\mathcal{D}^0=\mathcal{U}^0$} \\
+    $\text{\pr{solver}}^{\text{\id{coarse}}}(f^0,\ell^0,\chi^0,g^0;y^0)$ \ct{solving in $\mathcal{D}^0=\mathcal{U}^0$} \\
     $z^0 = y^0$ \\
@@
       $z^j = P z^{j-1} + y^{j}$ \\
-      $\text{\pr{smoother}}^{\text{\id{up}}}(f,\ell^j,\chi^j,g^j;z^j)$  \ct{smoothing in $\mathcal{U}^j$} \\-
+      $\text{\pr{smoother}}^{\text{\id{up}}}(f^j,\ell^j,\chi^j,g^j;z^j)$  \ct{smoothing in $\mathcal{U}^j$} \\-
     return $w^J+z^J$
 \end{pseudo*}
-\caption{Nonlinear multilevel constraint decomposition V-cycle for VI problem \eqref{eq:fe:vi}.}
+\caption{Nonlinear multilevel constraint decomposition V-cycle for the finest-level VI problem \eqref{eq:fe:vi}, using mesh-level discretizations $f^j$ of the original nonlinear $f$ in problem \eqref{eq:vi}.}
 \label{alg:nmcd}

### 2022-07-29 paper/mcd2.tex
progress

@@
 \usepackage[kw]{pseudo}
-\pseudoset{left-margin=15mm,topsep=5mm,idfont=\texttt}
+\pseudoset{%
+left-margin=15mm,%
+topsep=5mm,%
+label=\footnotesize\arabic*,%
+idfont=\texttt,%
+ctfont=\textsl,%
+ct-left=\qquad\qquad(,%
+ct-right=),%
+}
 
@@
         \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$ \rm{we have} \\+
-            $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\Big)}{v_i-\hat w_i} \ge \ip{g}{v_i-\hat w_i}}$ \\-
+            $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}}$ \\-
             $w_i = (1-\alpha) R_i u + \alpha \hat w_i\in\cK_i$ \\-
@@
 
-The starting point of a \emph{multilevel constraint decomposition} (MCD) \cite{GraeserKornhuber2009,Tai2003} method is to define constraint subsets, the sets generically denoted $\cK_i$ in sections \ref{sec:vi} and \ref{sec:convergence}, based on monotone restrictions of the constraint.
-In fact our algorithm is a multilevel V-cycle with two types of constraint subsets, distinctly for the downward and upward parts of the cycle.
+The starting point of a \emph{multilevel constraint decomposition} (MCD) \cite{GraeserKornhuber2009,Tai2003} method is to define constraint subsets, the sets generically denoted $\cK_i$ in sections \ref{sec:vi} and \ref{sec:convergence}, based on monotone restrictions of a defect constraint (see below).
+Our multilevel V-cycle algorithm will define two types of constraint subsets in this manner, distinctly for the downward and upward parts of the cycle.
+We also diverge from the MCD algorithms by Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear \emph{full approximation scheme} (FAS) philosophy \cite{BrandtLivne2011}.
+When $f$ is linear, as in the implemented examples in \cite{GraeserKornhuber2009,Tai2003}, there is no need to coarsen the solution itself, as in classical linear multigrid only the residual and correction are coarsened, but in the FAS approach here we coarsen the solution iterate using the injection restriction.
 
-To make this precise, suppose $w^J \in \cK^J$ is an admissible fine-level iterate which approximates the solution $u^J$.
+To make this precise, suppose $w^J \in \cK^J$ is an admissible fine-level iterate which approximates the solution $u^J$ of VI \eqref{eq:fe:vi}.
 From this iterate we define the \emph{defect constraint} \cite{GraeserKornhuber2009}
@@
 a nonpositive function.
-The motivation of the defect constraint is that a corrected iterate is admissible, $w^J + z^J \ge \gamma^J$, if and only if the perturbation $z^J$ exceeds the defect constraint, $z^J \ge \chi^J$.
-Next, for $j=J,J-1,\dots,1$ apply monotone restriction $\mR$ to define the \emph{level defect constraints}
+The motivation here is that a corrected iterate is admissible, $w^J + z^J \ge \gamma^J$, if and only if the perturbation $z^J$ exceeds the defect constraint, $z^J \ge \chi^J$.
+Next, for $j=J,J-1,\dots,1$ we apply monotone restriction $\mR$ to define the \emph{level defect constraints}:
 \begin{equation}
@@
 
-For $j=0,\dots,J$ we also denote the differences between the level defect constraints by
+We also denote the differences between the level defect constraints by
 \begin{equation}
@@
 \end{equation}
-(This uses the convenience definition $\chi^{-1}=0$.)  These are the gaps between the $\chi^j$ shown in Figure \ref{fig:chiphilevels}.
-Note $\phi^0=\chi^0$, and that the telescoping sum $\sum_{j=0}^J \phi^j = \chi^J$ applies.
+(This uses the convenience definition $\chi^{-1}=0$, thus $\phi^0=\chi^0$ in particular.)  These are the gaps between the $\chi^j$ shown in Figure \ref{fig:chiphilevels}.
+Note the telescoping sum $\sum_{j=0}^J \phi^j = \chi^J$ decomposition.
 
@@
 \begin{equation}
-\mathcal{D}^j = \{v\ge \phi^j\}, \qquad \mathcal{U}^j = \{v\ge \chi^j\} \label{eq:fe:constraintsets}
+\mathcal{D}^j = \{v \in \mathcal{V}^j \,:\, v\ge \phi^j\}, \qquad \mathcal{U}^j = \{v \in \mathcal{V}^j \,:\, v\ge \chi^j\} \label{eq:fe:constraintsets}
 \end{equation}
 define \emph{downward} and \emph{upward constraint subsets} of the FE subspaces $\mathcal{V}^j$, respectively.
-The upward sets are larger, and decompositions $\mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i$ apply in the same sense as \eqref{eq:constraintdecomp}.
+The upward sets are larger, $\mathcal{D}^j \subset \mathcal{U}^j$, and set decompositions $\mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i$ apply in the same sense as \eqref{eq:constraintdecomp}.
 
-Returning to the heart of the MCD idea, our plan is to correct the admissible fine-level iterate $w^J\in \mathcal{K}^J$ by adding modifications from each coarser subspace $\mathcal{V}^j$.
+Returning to the central MCD idea, we will correct the finest-level iterate $w^J\in \mathcal{K}^J$ by adding modifications from each coarser subspace $\mathcal{V}^j$.
 By nesting \eqref{eq:fe:nestedspaces} these corrections live in the fine-level subspace $\mathcal{V}^J$, but the corrected iterate needs to be admissible as well.
@@
 In detail, suppose $y^{j+1} \in \mathcal{D}^{j+1}, \dots, y^J \in \mathcal{D}^J$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:nmcdvcycle}).
-The so-far corrected iterate is admissible, $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$, because the sum of corrections is in the (finest) defect constraint set, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J = \{v\ge \chi^J\}$.
+The so-far corrected iterate is admissible, namely $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$, because the sum of corrections is in the (finest) defect constraint set, i.e.~$y^J + \dots + y^{j+1} \in \mathcal{U}^J = \{v\ge \chi^J\}$.
 
@@
 
-Now, when computing the next-coarser correction we will need to use the coarser-mesh representation, and, as in nonlinear FAS multigrid schemes \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, we must coarsen the iterate as well as the residual and the correction.\footnote{Here we diverge from algorithms by Tai  \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009}.
-If $f$ is linear, as in the implemented examples in those works, there is no need to coarsen the solution.
-(In classical linear multigrid only the residual and the correction must be coarsened.)  We follow the FAS philosophy for nonlinear functionals.}  We do this using injection \cite[section 5.3]{Trottenbergetal2001}.
+When computing the next-coarser correction one must use the coarser-mesh representation of the problem.
+As in nonlinear FAS multigrid schemes \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, this requires that we coarsen the iterate as well as the residual and the correction, and here injection is used \cite[section 5.3]{Trottenbergetal2001}.
 Define
@@
 \end{equation}
-That is, $g^j$ is a $j$th-level mesh approximation to the solution $u^J$, before we add the correction which solves the coarsened VI problem at the $j$th level.\footnote{If we were to inject the finest-mesh obstacle downward to coarser meshes, i.e.~$\gamma^{j-1} = \iR \gamma^j$, then, because $\iR$ preserves inequalities and $y^j \in \mathcal{D}^j$  it would follow that $g^j \ge \gamma^j$.
-In this sense $g^j$ is an admissible function on the $j$th-level mesh.
-However, our use of the defect constraints $\chi^j$, following \cite{GraeserKornhuber2009}, avoids the need to generate these $\gamma^j$; their definition in this footnote can be ignored.}
+Function $g^j$ is the $j$th-level mesh approximation to the solution of \eqref{eq:fe:vi}, before smoothing and correction at that level.\footnote{If we were to inject the finest-mesh obstacle downward to coarser meshes, $\gamma^{j-1} = \iR \gamma^j$, then, because $\iR$ preserves inequalities and $y^j \in \mathcal{D}^j$, it would follow that $g^j \ge \gamma^j$.
+In this sense $g^j$ is admissible on the $j$th-level mesh.
+However, following \cite{GraeserKornhuber2009}, our use of defect constraints $\chi^j$ avoids the need to generate the $\gamma^j$.}
 
-FIXME do I need $f^j$ to be defined?  yes.  Tai assumes $f(w^J+y^J + \dots + y^{j+1} +y^j)$ is defined, which is true, but $f^j(g^j)$ uses the $j$th level representation
+Down-smoothing makes progress toward solving the following VI for $y^j \in \mathcal{D}^j$:
+\begin{equation}
+\ip{f^j(g^j + y^j)}{v-y^j} \ge \ip{\ell^j}{v-y^j} \qquad \text{for all } v\in \mathcal{D}^j. \label{eq:fe:downvi}
+\end{equation}
+Here $f^j$ is a discretization of $f$ in \eqref{eq:vi},\footnote{Again we diverge from the algorithm by Tai \cite{Tai2003}, who essentially assumes that ``$f(g^j + y^j)$'' is used here.
+If $f$ is linear then efficient implementation via that formula is possible, but otherwise one must apply $f^j$.} using quadrature etc.~as addressed for $f^J$ in \eqref{eq:fe:vi}.
 
-FIXME Now we compute the next-coarser correction $y^j \in \mathcal{D}^j$ by the downward VI at the $j$th level,
+For $j<J$ the source function $\ell^j$ is determined by an FAS-type formula, as follows.
+First note that the $j+1$ level version of \eqref{eq:fe:downvi} is not solved exactly by the correction $y^{j+1}$.
+Fixing $y^{j+1} \in \mathcal{D}^{j+1}$, we would want to find a further correction $y$ so that $g^{j+1}+y^{j+1}+y$ is still admissible, and so that
 \begin{equation}
-\ip{f(w^J+y^J + \dots + y^{j+1} +y^j)}{v-y^j} \ge \ip{\ell^{j+1}}{v-y^j}, \label{eq:fe:levelvi}
+\ip{f^{j+1}(g^{j+1}+y^{j+1}+y)}{v-y^{j+1}-y} \ge \ip{\ell^{j+1}}{v-y^{j+1}-y} \qquad \text{\emph{(notional)}} \label{eq:fe:downvinotional}
 \end{equation}
-for all $v \in \mathcal{D}^j$.  The source functional $\ell^j$ remains to be computed; see below.  Note that here and below we guarantee that the input to $f$ is admissible (in $\mathcal{K}^J$).
+held exactly.
+However, in a multilevel method we actually compute this next correction $y$ on the next-coarser level $j$.
+To do this we modify the notional finer-level VI \eqref{eq:fe:downvinotional} in five steps which mimic the standard argument for the FAS correction equations for PDEs \cite{BrandtLivne2011,Trottenbergetal2001}:
+\begin{enumerate}
+\item group $v-y^{j+1}$ as a coarser-level test function $v\in \mathcal{D}^j$,
+\item subtract the known $f^{j+1}(g^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides,
+\item replace the residual $\ell^{j+1}-f^{j+1}(g^{j+1}+y^{j+1})$ by its restriction,
+\item replace $g^{j+1}+y^{j+1}$ where it appears by its restriction,
+\item and replace $f^{j+1}$ by its coarser rediscretization $f^j$ where it appears on the left.
+\end{enumerate}
+Steps \emph{(iii)}--\emph{(v)} are justified, as usual, by the assumption that the residual and the correction should be smooth as a result of the progress made by the smoother on the $j+1$ level.
+These steps now yield the precise $j$th-level VI problem, in cluttered form but the same as \eqref{eq:fe:downvi}:
+\begin{align}
+&\ip{f^j\left(\iR(g^{j+1}+y^{j+1})+y^j\right)}{v-y^j} - \ip{f^j\left(\iR(g^{j+1}+y^{j+1})\right)}{v-y^j} \label{eq:fe:downvicluttered} \\
+&\qquad \ge \ip{R\left(\ell^{j+1}-f^{j+1}(g^{j+1}+y^{j+1})\right)}{v-y^j} \notag
+\end{align}
 
-FIXME Subtract $N^{j+1}(w^J+z^{j+1})$ from both sides of \eqref{eq:fe:levelvi} gives
-\begin{align}
-&N^{j+1}(w^J+z^{j+1}+y)[v-y] \label{eq:siaVIFEprecoarsetwo} \\
-&\quad - N^{j+1}(w^J+z^{j+1})[v-y] \ge - \tilde F^{j+1}[v-y], \notag
-\end{align}
-where we define the resulting residual as
+Thus we define $\ell^j \in (\mathcal{V}^j)'$ in \eqref{eq:fe:downvi} by the following recursive definition, starting from $\ell^J$ already defined for the finest-level problem \eqref{eq:fe:vi}, and incorporating definition \eqref{eq:fe:defineg}:
 \begin{equation}
-\tilde F^{j+1}[v] = N^{j+1}(w^J+z^{j+1})[v] - \ell^{j+1}[v].
+\ell^j = f^j(g^j) + R\left(\ell^{j+1}-f^{j+1}(g^{j+1}+y^{j+1})\right). \label{eq:fe:levelsource}
 \end{equation}
+This is identical, even if the notation differs, to the standard FAS formula for a PDE source term, for example equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.
 
-FIXME explain smoother semantics (in-place on final argument; reduces CP residual)
+At each level in the downward part of the V-cycle (Figure \ref{fig:nmcdvcycle}) we apply a smoother, a partial solver of \eqref{eq:fe:downvi} which should rapidly reduce the high-frequencies in the error, to generate a correction $y^j \in \mathcal{D}^j$.
+At the coarsest $j=0$ level we solve \eqref{eq:fe:downvi}, accurately if possible, yielding a correction $y^0$ in the coarsest defect constraint set $\mathcal{D}^0=\mathcal{U}^0 = \{v \ge \chi^0\}$.
+Note that the coarsest-level solver may be the same as the other smoothers.
 
-\pseudoset{
-ctfont=\textsl,%
-ct-left=\qquad\qquad(,%
-ct-right=),%
-}
+In the upward part of the V-cycle the smoother starts from an initial iterate $z_0^j$ which has accumulated the coarser-level corrections.  Define $z_0^0 = y^0$ and
+\begin{equation}
+z_0^j = P z_0^{j-1} + y^j  \label{eq:fe:upwardaccumulation}
+\end{equation}
+where $P$ is canonical prolongation; stated simply, $z_0^j = y^j + \dots + y^0$.
+By the decompositions $\mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i$, noted earlier, $z_0^j \in \mathcal{U}^j$.
+This initial $z_0^j$ is smoothed to generate a result $z^j \in \mathcal{U}^j$ which approximately solves the same VI as the downward version \eqref{eq:fe:downvi}, except now we can solver over a larger set:
+\begin{equation}
+\ip{f^j(g^j + z^j)}{v-z^j} \ge \ip{\ell^j}{v-z^j} \qquad \text{for all } v\in \mathcal{U}^j. \label{eq:fe:upvi}
+\end{equation}
+The reason we can solve over $\mathcal{U}^j$, unlike during the downward phase, is that there is no forthcoming coarser correction which could violate admissibility.
 
 \begin{pseudofloat}[H]
-\begin{pseudo*} \label{ps:nmcd-vcycle}
-\pr{nmcd-vcycle}(\ell^J,w^J,\gamma^J)\text{:} \\+
+\begin{pseudo} \label{ps:nmcd-vcycle}
+\pr{nmcd-vcycle}(\ell^J,\gamma^J,w^J)\text{:} \\+
     $\chi^J = w^J - \gamma^J$ \\
-    $g^J = w^J$ \ct{notational; no copy needed} \\
+    $g^J = w^J$ \\
     for $j=J$ downto $j=1$ \\+
@@
       $y^j = 0$ \\
-      $\text{\pr{smoother}}^{\text{\id{down}}}(f^j,\ell^j,\phi^j,g^j;y^j)$  \ct{smoothing in $\mathcal{D}^j$}\\
+      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\phi^j,g^j;y^j)$  \ct{smoothing of \eqref{eq:fe:downvi} in $\mathcal{D}^j$}\\
       $g^{j-1} = \iR(g^j + y^j)$ \\
-      $\ell^{j-1} = f^{j-1}(g^{j-1}) - R \left(f^j(g^j+y^j) - \ell^j\right)$ \\-
+      $\ell^{j-1} = f^{j-1}(g^{j-1}) + R \left(\ell^j - f^j(g^j+y^j)\right)$ \\-
     $y^0 = 0$ \\
-    $\text{\pr{solver}}^{\text{\id{coarse}}}(f^0,\ell^0,\chi^0,g^0;y^0)$ \ct{solving in $\mathcal{D}^0=\mathcal{U}^0$} \\
+    $\text{\pr{solve}}(\ell^0,\chi^0,g^0;y^0)$ \ct{solving of \eqref{eq:fe:downvi} in $\mathcal{D}^0=\mathcal{U}^0$} \\
     $z^0 = y^0$ \\
@@
       $z^j = P z^{j-1} + y^{j}$ \\
-      $\text{\pr{smoother}}^{\text{\id{up}}}(f^j,\ell^j,\chi^j,g^j;z^j)$  \ct{smoothing in $\mathcal{U}^j$} \\-
+      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\chi^j,g^j;z^j)$  \ct{smoothing of \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
     return $w^J+z^J$
-\end{pseudo*}
+\end{pseudo}
 \caption{Nonlinear multilevel constraint decomposition V-cycle for the finest-level VI problem \eqref{eq:fe:vi}, using mesh-level discretizations $f^j$ of the original nonlinear $f$ in problem \eqref{eq:vi}.}
@@
 
+During one application of \pr{nmcd-vcycle}, storage must be allocated for $\chi^j$, $\phi^j$, $g^j$, $y^j$, and $\ell^j$ on each level.
+However, in algorithm lines 13 and 15 no allocation is strictly required because the $y^j$ space can be reused for $z^j$ as the coarser corrections are prolonged and accumulated.
+Strictly-speaking there another such optimization: in line 3 $g^J$ need not be a copy of $w^J$.
+The total memory usage of \pr{nmcd-vcycle}, excluding storage usage in the smoothers/solver but adding storage for the original obstacle $\gamma^J$, is about $8 m^J$ real numbers\footnote{That is: $8\approx 5(4/3) + 1$.} under the standard estimate that a 2D, uniform-mesh V-cycle uses $4/3$ the memory of the finest level alone \cite[Section 2.4]{Trottenbergetal2001}, whereas a single-level method needs only $3 m^J$ for vectors $\ell^J,w^J,\gamma^J$ themselves.
+
+Regarding notation and semantics for the smoothers, note they may be iterated \id{down} and \id{up} times, respectively.
+Both the smoothers and the coarsest-level solver are supposed to do in-place modifications of their final arguments, and of course they are in-exact solvers.
+They must not modify their first three arguments.
+
+Furthermore, repeated application of \pr{nmcd-vcycle}, or of the smoothers and solver at the levels, will not reduce the classical residual $f^J(u^J) - \ell^J$ to zero everywhere.
+Recall that a finite-dimensional VI problem is equivalent to a \emph{nonlinear complementarity problem} (NCP) \cite{FacchineiPang2003}.
+For example, corresponding to \eqref{eq:fe:vi} one could write
+\begin{equation}
+u^J \ge \gamma^J, \qquad f^J(u^J) \ge \ell^J, \qquad (u^J - \gamma^J)\left(f^J(u^J) - \ell^J\right) = 0.  \label{eq:fe:ncp}
+\end{equation}
+Thus, at convergence we only expect $f^J(u^J) = \ell^J$ at those locations where the constraint is inactive, where $u^J > \gamma^J$.
+Therefore, recalling that a $j$th-level hat function based at node $x_p^j$ is denoted $\psi_p^j$, for an iterate $w^j$ define the \emph{downward nodal residual} by
+\begin{equation}
+(\hat r_{\text{down}}^j(w))_p = \begin{cases} \ip{f^j(w^j)-\ell^j}{\psi_p^j}, & w^j_p > \phi^j_p, \\
+                                  \min\{\ip{f^j(w^j)-\ell^j}{\psi_p^j},0\}, & w^j_p = \phi^j_p, \end{cases} \label{eq:cpresidual}
+\end{equation}
+where $w^j_p,\phi^j_p$ denote the nodal values of $w^j,\phi^j$ as $x_p^j$, respectively.  The \emph{upward nodal residual} $\hat r_{\text{up}}(w)$ is similarly defined using $\chi^j$.
+
+The nodal residuals $\hat r_{\text{down}}(w^j),\hat r_{\text{up}}(w^j)$ are nonzero if there are nodes where the NCP \eqref{eq:fe:ncp} is violated, either because $f^j(w^j)-\ell^j$ remains nonzero, in a weak sense, on the inactive set or because the same quantity is negative on the active set.
+If $w^j$ is admissible, for instance if $w^j = g^j+z^j$ for $z^j \in \mathcal{U}^j$, inside \pr{nmcd-vcycle}, and if $\|\hat r_{\text{up}}(w^j)\|$ is small in some appropriate norm then we regard the $j$th-level problem as being numerically solved.
+This provides stopping criteria for application of \pr{solver} on the coarsest mesh, and for terminating cycles of \pr{nmcd-vcycle} when addressing the original VI \eqref{eq:fe:vi}.
+
 FIXME in Appendix we show how NMCD reduces to FAS if the obstacle is removed and to MCD if $f$ is assumed to be linear
 
-FIXME observe that up-smoothing is more efficient
+FIXME again observe that up-smoothing is more efficient
 
-\section{Results for local variational inequalities} \label{sec:results}
+\section{Results} \label{sec:results}
 
@@
 
-% A BRIDGE TOO FAR:  \section{Results for a nonlocal variational inequality} \label{sec:resultsnonlocal}
+% A BRIDGE TOO FAR:  \section{Results for a nonlocal variational inequality}
 

### 2022-07-29 paper/mcd2.tex
make NMCD-VCYCLE in-place on w^J, the finest-level iterate

@@
 The starting point of a \emph{multilevel constraint decomposition} (MCD) \cite{GraeserKornhuber2009,Tai2003} method is to define constraint subsets, the sets generically denoted $\cK_i$ in sections \ref{sec:vi} and \ref{sec:convergence}, based on monotone restrictions of a defect constraint (see below).
-Our multilevel V-cycle algorithm will define two types of constraint subsets in this manner, distinctly for the downward and upward parts of the cycle.
+Our multilevel V-cycle algorithm will define two types of constraint subsets, distinct sets for the downward and upward parts of the cycle.
 We also diverge from the MCD algorithms by Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear \emph{full approximation scheme} (FAS) philosophy \cite{BrandtLivne2011}.
-When $f$ is linear, as in the implemented examples in \cite{GraeserKornhuber2009,Tai2003}, there is no need to coarsen the solution itself, as in classical linear multigrid only the residual and correction are coarsened, but in the FAS approach here we coarsen the solution iterate using the injection restriction.
+When $f$ is linear, as in the implemented examples in \cite{GraeserKornhuber2009,Tai2003}, there is no need to coarsen the solution itself, but in the FAS approach here we coarsen the solution iterate using the injection restriction.
+Note that in classical linear multigrid only the residual and correction are coarsened.
 
@@
 \end{center}
-\caption{Going down, the NMCD V-cycle computes corrections $y_j \in \mathcal{D}^j$.  The upward corrections $z_j$ can expand into larger constraint subsets $\mathcal{U}^j$.}
+\caption{An NMCD V-cycle computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j$ expand into larger constraint subsets $\mathcal{U}^j$.}
 \label{fig:nmcdvcycle}
@@
 
+These ideas come together in Algorithm \ref{alg:nmcd}, an NMCD V-cycle.  This procedure acts in-place on the final argument $w^J$, but it leaves the other inputs $\ell^J,\gamma^J$ unchanged.
+
 \begin{pseudofloat}[H]
 \begin{pseudo} \label{ps:nmcd-vcycle}
-\pr{nmcd-vcycle}(\ell^J,\gamma^J,w^J)\text{:} \\+
+\pr{nmcd-vcycle}(\ell^J,\gamma^J;w^J)\text{:} \\+
     $\chi^J = w^J - \gamma^J$ \\
@@
       $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\chi^j,g^j;z^j)$  \ct{smoothing of \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
-    return $w^J+z^J$
+    $w^J \gets w^J+z^J$
 \end{pseudo}
-\caption{Nonlinear multilevel constraint decomposition V-cycle for the finest-level VI problem \eqref{eq:fe:vi}, using mesh-level discretizations $f^j$ of the original nonlinear $f$ in problem \eqref{eq:vi}.}
+\caption{Nonlinear multilevel constraint decomposition V-cycle for the finest-level VI problem \eqref{eq:fe:vi}, as an in-place method on $w^J$, using mesh-level discretizations $f^j$ of the original nonlinear $f$ in problem \eqref{eq:vi}.}
 \label{alg:nmcd}
@@
 During one application of \pr{nmcd-vcycle}, storage must be allocated for $\chi^j$, $\phi^j$, $g^j$, $y^j$, and $\ell^j$ on each level.
-However, in algorithm lines 13 and 15 no allocation is strictly required because the $y^j$ space can be reused for $z^j$ as the coarser corrections are prolonged and accumulated.
-Strictly-speaking there another such optimization: in line 3 $g^J$ need not be a copy of $w^J$.
-The total memory usage of \pr{nmcd-vcycle}, excluding storage usage in the smoothers/solver but adding storage for the original obstacle $\gamma^J$, is about $8 m^J$ real numbers\footnote{That is: $8\approx 5(4/3) + 1$.} under the standard estimate that a 2D, uniform-mesh V-cycle uses $4/3$ the memory of the finest level alone \cite[Section 2.4]{Trottenbergetal2001}, whereas a single-level method needs only $3 m^J$ for vectors $\ell^J,w^J,\gamma^J$ themselves.
+However, line 3 is essentially notational in that the finest-level iterate $g^J$ is the same as $w^J$; no copy is needed.
+Similarly, in algorithm lines 13 and 15 no allocation is required because the $y^j$ space can be reused for $z^j$ as the coarser corrections are prolonged and accumulated.
+Therefore the total storage of \pr{nmcd-vcycle}, excluding any in the smoothers/solver, but including for the original obstacle $\gamma^J$, is about $8 m^J$ real numbers\footnote{That is: $8\approx 5(4/3) + 1$.} under the standard estimate that a 2D, uniform-mesh V-cycle uses $4/3$ the memory of the finest level alone \cite[Section 2.4]{Trottenbergetal2001}.
+For comparison, a single-level method needs at least $3 m^J$ storage for the vectors $\ell^J,\gamma^J,w^J$.
 

### 2022-07-29 paper/mcd2.tex
allow the obstacle to not have zero boundary values, so monotone restriction acts on general (non-zero boundary) FE functions

@@
 
-Let $\mathcal{V}^j$ be the continuous, piecewise-polynomial FE space $P_1$,\footnote{The FE method herein apparently extends to $Q_1$ spaces over quadrilateral meshes \cite{Elmanetal2014}.
+Let $\tilde{\mathcal{V}}^j$ be the continuous, piecewise-polynomial FE space $P_1$,\footnote{The FE method herein apparently extends to $Q_1$ spaces over quadrilateral meshes \cite{Elmanetal2014}.
 However, for $k>1$ the $P_k$ space seems not to be suitable because the monotone restriction operator, \eqref{eq:fe:monotonerestriction} below, depends on element-wise monotonicity.
 That is, for $P_1$ or $Q_1$ if the nodal values of an FE function (non-strictly) increase then the FE function also increases.
-As shown by Figure 1.7 of \cite{Elmanetal2014}, for example, such monotonicity is not true for $P_2$.} over $\mathcal{T}^j$, with zero boundary values.
-Note $\mathcal{V}^j$ is a subspace of $\mathcal{V}=W_0^{1,p}(\Omega)$ for any $p\ge 1$ \cite{Elmanetal2014}.
+As shown by Figure 1.7 of \cite{Elmanetal2014}, for example, such monotonicity is not true for $P_2$.} over $\mathcal{T}^j$, with arbitrary boundary values.
+Let $\mathcal{V}^j \subset \tilde{\mathcal{V}}^j$ be the subspace with zero boundary values.
+For any $p\ge 1$, $\tilde{\mathcal{V}}^j$ is a subspace of $W^{1,p}(\Omega)$ and $\mathcal{V}^j$ is a subspace of $\mathcal{V}=W_0^{1,p}(\Omega)$ \cite{Elmanetal2014}.
 Our choices imply the following nesting:
@@
 \end{equation}
-Suppose $\gamma^J \in \mathcal{V}^J$ denotes the \emph{obstacle} and define
+
+Our only need for the arbitrary-boundary-value spaces $\tilde{\mathcal{V}}^j$ is for the constraint.
+Suppose $\gamma^J \in \tilde{\mathcal{V}}^j$ denotes the \emph{obstacle}, a function satisfying $\gamma^J \le 0$ at boundary nodes of $\mathcal{T}^J$, and define
 \begin{equation}
@@
 \end{equation}
-the finest-level \emph{constraint set} of \emph{admissible} FE functions.
+the finest-level \emph{constraint set} of \emph{admissible} FE functions.  Note that admissible functions from $\mathcal{K}^J$ have zero boundary values.
 
-One might assume that $f$ in \eqref{eq:vi} is defined over $\mathcal{K}^J$, i.e.~$f:\mathcal{K}^J \to \mathcal{V}'$.
-However, especially for nontrivial nonlinear functionals, the finest-level, discretized FE functional $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ is likely to only approximate $f$ because of quadrature, or other implementation details.
-The functional $\ell$ in \eqref{eq:vi} is defined and continuous over $\mathcal{V}^J$, thus $\ell\in (\mathcal{V}^J)'$, and we assume for simplicity that it can be computed exactly, but we denote $\ell^J \in (\mathcal{V}^J)'$ to indicate that it must be represented in the FE sense.
+One might assume that the nonlinear functional $f$ in \eqref{eq:vi} is defined over $\mathcal{K}^J$, i.e.~$f:\mathcal{K}^J \to \mathcal{V}'$.
+However, especially for nontrivial nonlinear functionals, the finest-level, discretized FE functional $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ is likely to only approximate $f$ because of quadrature or other implementation details.
+The functional $\ell$ in \eqref{eq:vi} is defined and continuous over $\mathcal{V}^J$, thus $\ell\in (\mathcal{V}^J)'$, and we assume for simplicity that it can be computed exactly, but we denote $\ell^J \in (\mathcal{V}^J)'$ to indicate that it must also be given an FE representation.
 
@@
 
-While \eqref{eq:fe:vi} approximates \eqref{eq:vi}, even if quadrature and other implementation details are ignored it may not technically be conforming for \eqref{eq:vi}.
-That is, despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require $\mathcal{K}^J \subset \mathcal{K}$.
+While \eqref{eq:fe:vi} approximates \eqref{eq:vi}, it may not technically be conforming for \eqref{eq:vi}.
+That is, even if quadrature and other implementation details are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require $\mathcal{K}^J \subset \mathcal{K}$.
 Only the fine-level FE obstacle $\gamma^J$ needs to be defined, while $\mathcal{K}$ remains abstract.
-In fact, if a continuum obstacle exists at all then admissibility with respect to its FE interpolant $\gamma^J$ might not imply continuum admissiblity.
+In fact, if a continuum obstacle exists then admissibility with respect to its FE interpolant $\gamma^J$ might not imply continuum admissiblity.
 These caveats apply in geophysical situations where the obstacle is given by meshed elevation data, the continuum limit of which is undefined \cite{Bueler2016}.
@@
 The multilevel algorithm in the next section, designed to solve \eqref{eq:fe:vi}, requires transfer operators between meshes.
-We use four such operators, three of which are the standard and linear maps used for nonlinear multigrid methods \cite{Trottenbergetal2001}.
+We use four such operators, three of which are standard, linear maps used for multigrid methods \cite{Trottenbergetal2001}.
 Table \ref{tab:transfers} will help the reader keep track.
@@
 (nodal) injection restriction & $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ & \,\checkmark \\
-monotone restriction          & $\mR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ &  
+monotone restriction          & $\mR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ &
 \end{tabular}
@@
 \medskip
-\caption{Transfer operators used in this paper.}
+\caption{Transfer operators used in this paper.  FE vector space $\mathcal{V}^j$ has zero boundary values while $\tilde{\mathcal{V}}^j$ has arbitrary boundary values. $\mR$ is a nonlinear map.}
 \label{tab:transfers}
@@
 For $z\in\mathcal{V}^j$ we define the injection $\iR z$ as the element of $\mathcal{V}^{j-1}$ which has the same nodal values as $z$ at the nodes of $\mathcal{T}^{j-1}$.
-Note that injection clearly preserves ordering, that is, if $w \le z$ then $\iR w \le \iR z$.
+Note that injection preserves ordering, that is, if $w \le z$ then $\iR w \le \iR z$.
 
-Finally, monotone restriction $\mR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ is a nonlinear map which maximizes nodal values over the supports of coarse-mesh hat functions.
-To make the definition precise, first denote the \emph{interior nodes} of triangulation $\mathcal{T}^j$ by $x_q^j$ for $q=0,\dots,m_j-1$, and then the corresponding hat function \cite{Elmanetal2014} by $\psi_q^j$.
-Monotone restriction of $z\in \mathcal{V}^j$ is
+Finally, monotone restriction $\mR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ is defined as the nonlinear map which maximizes nodal values over the supports of coarse-mesh hat functions.
+To make the definition precise, first denote the \emph{nodes} of triangulation $\mathcal{T}^j$ by $x_q^j$ for $q=0,\dots,\tilde{m}_j-1$, and then the corresponding hat functions \cite{Elmanetal2014} by $\psi_q^j$.
+(Boundary nodes are included here.)  Monotone restriction of $z\in\tilde{\mathcal{V}}^j$ is
 \begin{equation}
-  \mR z = \sum_{p=1}^{m_{j-1}} \max \{z_q \,:\, \psi_p^{j-1}(x_q^j) > 0\}\, \psi_p^{j-1}. \label{eq:fe:monotonerestriction}
+  \mR z = \sum_{p=1}^{\tilde m_{j-1}} \max \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}\, \psi_p^{j-1}. \label{eq:fe:monotonerestriction}
 \end{equation}
-Recalling that the piecewise-linear functions in $\mathcal{V}^j$ have zero boundary values, note that if $z\le 0$ then $0\ge \mR z \ge z$ on $\Omega$.
+Note that if $z\in\tilde{\mathcal{V}}^j$ then $z \le \mR z$.
 
@@
 \end{equation}
-a nonpositive function.
+a nonpositive function in $\tilde{\mathcal{V}}^J$.
 The motivation here is that a corrected iterate is admissible, $w^J + z^J \ge \gamma^J$, if and only if the perturbation $z^J$ exceeds the defect constraint, $z^J \ge \chi^J$.
-Next, for $j=J,J-1,\dots,1$ we apply monotone restriction $\mR$ to define the \emph{level defect constraints}:
+Next, for $j=J,J-1,\dots,1$ we apply monotone restriction $\mR$ to define the \emph{level defect constraints},
 \begin{equation}
@@
 \end{equation}
-As observed after equation \eqref{eq:fe:monotonerestriction}, $0\ge \chi^{j-1} \ge \chi^j$.  A $J=4$ example is shown in Figure \ref{fig:chiphilevels}.
+so that $\chi^j \le \chi^{j-1}$.  A $J=4$ example is shown in Figure \ref{fig:chiphilevels}.
 
@@
 \begin{equation}
-\phi^j = \chi^j - \chi^{j-1}. \label{eq:fe:philevels}
+\phi^j = \chi^j - \chi^{j-1}, \label{eq:fe:philevels}
 \end{equation}
-(This uses the convenience definition $\chi^{-1}=0$, thus $\phi^0=\chi^0$ in particular.)  These are the gaps between the $\chi^j$ shown in Figure \ref{fig:chiphilevels}.
-Note the telescoping sum $\sum_{j=0}^J \phi^j = \chi^J$ decomposition.
+and these are nonpositive functions.
+(Definition \eqref{eq:fe:philevels} uses $\chi^{-1}=0$.)  Functions $\phi^j$ are the gaps between the $\chi^j$ shown in Figure \ref{fig:chiphilevels}, with $\phi^0=\chi^0$ in particular.
+Note the telescoping sum decomposition $\sum_{j=0}^J \phi^j = \chi^J$.
 

### 2022-07-29 paper/mcd2.tex
clean up nodal (NCP) residual presentation, and stopping criterion

@@
 Similarly, in algorithm lines 13 and 15 no allocation is required because the $y^j$ space can be reused for $z^j$ as the coarser corrections are prolonged and accumulated.
-Therefore the total storage of \pr{nmcd-vcycle}, excluding any in the smoothers/solver, but including for the original obstacle $\gamma^J$, is about $8 m^J$ real numbers\footnote{That is: $8\approx 5(4/3) + 1$.} under the standard estimate that a 2D, uniform-mesh V-cycle uses $4/3$ the memory of the finest level alone \cite[Section 2.4]{Trottenbergetal2001}.
+Let $m^j$ be the number of interior nodes in the $j$th-level mesh.
+The total storage of \pr{nmcd-vcycle}, excluding any in the smoothers/solver, but including for the original obstacle $\gamma^J$, is about $8 m^J$ real numbers\footnote{That is: $8\approx 5(4/3) + 1$.} under the standard estimate that a 2D, uniform-mesh V-cycle uses $4/3$ the memory of the finest level alone \cite[Section 2.4]{Trottenbergetal2001}.
 For comparison, a single-level method needs at least $3 m^J$ storage for the vectors $\ell^J,\gamma^J,w^J$.
 
-Regarding notation and semantics for the smoothers, note they may be iterated \id{down} and \id{up} times, respectively.
-Both the smoothers and the coarsest-level solver are supposed to do in-place modifications of their final arguments, and of course they are in-exact solvers.
-They must not modify their first three arguments.
+\pr{nmcd-vcycle} assumes certain semantics for the smoothers and the coarsest-level solver; the latter might also be a smoother algorithm.
+They do in-place modifications of their final arguments, and they do not modify their first three arguments.
+The smoothers are in-exact solvers of VIs \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}, respectively, and they are iterated \id{down} and \id{up} times.
+Further smoother implementation details and choices are addressed in the next section.
 
-Furthermore, repeated application of \pr{nmcd-vcycle}, or of the smoothers and solver at the levels, will not reduce the classical residual $f^J(u^J) - \ell^J$ to zero everywhere.
-Recall that a finite-dimensional VI problem is equivalent to a \emph{nonlinear complementarity problem} (NCP) \cite{FacchineiPang2003}.
-For example, corresponding to \eqref{eq:fe:vi} one could write
+Repeated application of \pr{nmcd-vcycle} to solve VI \eqref{eq:fe:vi} will not reduce the finest-level residual $f^J(u^J) - \ell^J$ to zero everywhere.
+However, recalling that a finite-dimensional VI is equivalent to a \emph{nonlinear complementarity problem} (NCP) \cite{FacchineiPang2003}, \eqref{eq:fe:vi} corresponds to the NCP
 \begin{equation}
@@
 \end{equation}
-Thus, at convergence we only expect $f^J(u^J) = \ell^J$ at those locations where the constraint is inactive, where $u^J > \gamma^J$.
-Therefore, recalling that a $j$th-level hat function based at node $x_p^j$ is denoted $\psi_p^j$, for an iterate $w^j$ define the \emph{downward nodal residual} by
+Thus, at convergence we only expect $f^J(u^J) = \ell^J$ at those locations where the constraint is inactive, namely where $u^J > \gamma^J$.
+Therefore, recalling that the hat function at node $x_q^J$ is denoted $\psi_q^J$, for $w^J \in \mathcal{K}^J$ we define the following vector in $\RR^{m^J}$ as the finest-level \emph{nodal residual}:
 \begin{equation}
-(\hat r_{\text{down}}^j(w))_p = \begin{cases} \ip{f^j(w^j)-\ell^j}{\psi_p^j}, & w^j_p > \phi^j_p, \\
-                                  \min\{\ip{f^j(w^j)-\ell^j}{\psi_p^j},0\}, & w^j_p = \phi^j_p, \end{cases} \label{eq:cpresidual}
+\hat r(w^J)_q = \begin{cases} \ip{f^J(w^J)-\ell^J}{\psi_q^J}, & w^J(x_q^J) > \gamma^J(x_q^J), \\
+                                  \min\{\ip{f^J(w^J)-\ell^J}{\psi_q^J},0\}, & w^J(x_q^J) = \gamma^J(x_q^J). \end{cases} \label{eq:cpresidual}
 \end{equation}
-where $w^j_p,\phi^j_p$ denote the nodal values of $w^j,\phi^j$ as $x_p^j$, respectively.  The \emph{upward nodal residual} $\hat r_{\text{up}}(w)$ is similarly defined using $\chi^j$.
+(Formally, $\hat r:\mathcal{V}^J \to \RR^{m^J}$ is a nonlinear map.)  Observe that $\hat r(w^J)$ is nonzero if there are nodes where the NCP \eqref{eq:fe:ncp} is violated, either because $f^J(w^J)-\ell^J$ remains nonzero on the inactive set or because the same quantity is negative on the active set, and that nodal values of $f^J(w^J)-\ell^J$ are in evaluated in a weak sense, by application to a hat function.
 
-The nodal residuals $\hat r_{\text{down}}(w^j),\hat r_{\text{up}}(w^j)$ are nonzero if there are nodes where the NCP \eqref{eq:fe:ncp} is violated, either because $f^j(w^j)-\ell^j$ remains nonzero, in a weak sense, on the inactive set or because the same quantity is negative on the active set.
-If $w^j$ is admissible, for instance if $w^j = g^j+z^j$ for $z^j \in \mathcal{U}^j$, inside \pr{nmcd-vcycle}, and if $\|\hat r_{\text{up}}(w^j)\|$ is small in some appropriate norm then we regard the $j$th-level problem as being numerically solved.
-This provides stopping criteria for application of \pr{solver} on the coarsest mesh, and for terminating cycles of \pr{nmcd-vcycle} when addressing the original VI \eqref{eq:fe:vi}.
+A solver for \eqref{eq:fe:vi} using repeated applications of \pr{nmcd-vcycle} would start with an initial iterate $w^{J,0}$, do V-cycles, and stop once the finest-level nodal residual $\hat r(w^J)$ was sufficiently small.
+For example, using tolerances \id{atol}$>0$ and \id{rtol}$>0$ the stopping criterion might be
+\begin{equation}
+\|\hat r(w^J)\| < \id{atol} \qquad \text{or} \qquad \frac{\|\hat r(w^J)\|}{\|\hat r(w^{J,0})\|} < \id{rtol},
+\end{equation}
+where $\|\cdot\|$ denotes a norm on $\RR^{m^J}$.  A step tolerance could also be used, measuring the difference between successive iterates.
 
@@
 
-FIXME again observe that up-smoothing is more efficient
+\section{Smoother implementations} \label{sec:smoothers}
+
+FIXME
+
+Recall that a $j$th-level hat function based at node $x_p^j$ is denoted $\psi_p^j$, for an iterate $g^j$ and an admissible correction $y^j \in \mathcal{D}^j$ we define the \emph{downward nodal residual} as
+\begin{equation}
+(\hat r_{\text{down}}(y^j))_p = \begin{cases} \ip{f^j(g^j+y^j)-\ell^j}{\psi_p^j}, & y^j(x_p^j) > \phi^j(x_p^j), \\
+                                  \min\{\ip{f^j(g^j+y^j)-\ell^j}{\psi_p^j},0\}, & y^j(x_p^j) = \phi^j(x_p^j). \end{cases} \label{eq:cpresidual}
+\end{equation}
+The \emph{upward nodal residual} $\hat r_{\text{up}}(z^j)$ for $z^j \in \mathcal{U}^j$ is similarly defined using $\chi^j$.
+
+FIXME stopping criteria for smoother
 
@@
 FIXME show results for 2D Firedrake implementation of NMCD for $p$-Laplacian obstacle problem (Example \ref{ex:plaplacian}), advection-diffusion obstacle problem (Example \ref{ex:advectiondiffusion}), and porous-media obstacle problem (Example \ref{ex:porous})
+
+FIXME observe that up-smoothing is more efficient
 

### 2022-07-29 paper/mcd2.tex
fix a bug: defect constraint wrong order in nmcd-vcycle

@@
 \pr{nmcd-vcycle}(\ell^J,\gamma^J;w^J)\text{:} \\+
-    $\chi^J = w^J - \gamma^J$ \\
+    $\chi^J = \gamma^J - w^J$ \\
     $g^J = w^J$ \\

### 2022-07-30 paper/mcd2.tex
I think I have the reduction to FAS

@@
 
-At each level in the downward part of the V-cycle (Figure \ref{fig:nmcdvcycle}) we apply a smoother, a partial solver of \eqref{eq:fe:downvi} which should rapidly reduce the high-frequencies in the error, to generate a correction $y^j \in \mathcal{D}^j$.
+At each level in the downward part of the V-cycle (Figure \ref{fig:nmcdvcycle}) we apply a smoother, an iterative partial solver of \eqref{eq:fe:downvi} which rapidly reduces the high-frequencies in the error.
+This smoother yields a correction $y^j \in \mathcal{D}^j$.
+It is important that we can initialize this iteration using an admissible first guess.
+However, by the monotone restriction construction the down-constraint is nonpositive, $\phi^j\le 0$ so $0\in \mathcal{D}^j$, and thus an initial value of $y^j=0$ is admissible.
+
 At the coarsest $j=0$ level we solve \eqref{eq:fe:downvi}, accurately if possible, yielding a correction $y^0$ in the coarsest defect constraint set $\mathcal{D}^0=\mathcal{U}^0 = \{v \ge \chi^0\}$.
@@
 
-In the upward part of the V-cycle the smoother starts from an initial iterate $z_0^j$ which has accumulated the coarser-level corrections.  Define $z_0^0 = y^0$ and
+In the upward part of the V-cycle the smoother starts from an initial iterate which has accumulated the coarser-level corrections.  Define $z^0 = y^0$ and
 \begin{equation}
-z_0^j = P z_0^{j-1} + y^j  \label{eq:fe:upwardaccumulation}
+z^j = P z^{j-1} + y^j  \label{eq:fe:upwardaccumulation}
 \end{equation}
-where $P$ is canonical prolongation; stated simply, $z_0^j = y^j + \dots + y^0$.
-By the decompositions $\mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i$, noted earlier, $z_0^j \in \mathcal{U}^j$.
-This initial $z_0^j$ is smoothed to generate a result $z^j \in \mathcal{U}^j$ which approximately solves the same VI as the downward version \eqref{eq:fe:downvi}, except now we can solver over a larger set:
+where $P$ is canonical prolongation; stated simply, $z^j = y^j + \dots + y^0$.
+By the decompositions $\mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i$, noted earlier, $z^j \in \mathcal{U}^j$.
+This initial $z^j$ is smoothed to generate a result, also denoted $z^j \in \mathcal{U}^j$.
+This result approximately solves the same VI as the downward version \eqref{eq:fe:downvi}, except now over a larger constraint set:
 \begin{equation}
@@
 \end{equation}
-The reason we can solve over $\mathcal{U}^j$, unlike during the downward phase, is that there is no forthcoming coarser correction which could violate admissibility.
+The reason we can solve over $\mathcal{U}^j$ is that, unlike during the downward phase, there is no forthcoming coarser correction which could violate admissibility.
 
@@
 \end{pseudo}
-\caption{Nonlinear multilevel constraint decomposition V-cycle for the finest-level VI problem \eqref{eq:fe:vi}, as an in-place method on $w^J$, using mesh-level discretizations $f^j$ of the original nonlinear $f$ in problem \eqref{eq:vi}.}
+\caption{Nonlinear multilevel constraint decomposition V-cycle for the finest-level VI problem \eqref{eq:fe:vi}, as an in-place method on $w^J$.
+$f^j$ denotes a discretization of $f$ in problem \eqref{eq:vi}.}
 \label{alg:nmcd}
@@
 
-FIXME by removing obstacle, reduce to ordinary FAS as in \cite{Bruneetal2015}
+\pr{nmcd-vcycle}, Algorithm \ref{alg:nmcd} in the main text, generalizes two multilevel algorithms, first the FAS multigrid V-cycle for nonlinear PDEs, which itself generalizes the standard linear multigrid V-cycle, and second the MCD V-cycle for linear variational inequalities.
+This Appendix helps readers understand \pr{nmcd-vcycle} by identifying the assumptions which reduces it to these existing methods.
 
-FIXME by assuming $N$ is linear, reduce to Algorithm 4.7 in \cite{GraeserKornhuber2009} but with $\text{V}(\nu_1,\nu_2)$ cycles which works for linear
+To derive the FAS multigrid V-cycle we need to remove the inequality constraints from \pr{nmcd-vcycle}.
+To do this, suppose that the finest-level obstacle $\gamma^J$ is very negative, with magnitude greater than $u^J$ or any prospective iterate $w^J$.
+Problem \eqref{eq:fe:vi} is then unconstrained, so by replacing ``$v-u^J$'' with an arbitrary $v \in \mathcal{V}^J$ we have the finest-level, weak form problem
+\begin{equation}
+\ip{f^J(u^J)}{v} = \ip{\ell^J}{v} \qquad \text{for all } v\in \mathcal{V}^J \label{eq:app:fas:pde}
+\end{equation}
+corresponding to the strong-form PDE $f(u)=\ell$.
+Now, \pr{nmcd-vcycle} would generate downward defect constraints by monotone restriction.
+However, all that is required for the downward system is that $\chi^{j-1} \ge \chi^j$, so that $\phi^j = \chi^j - P \chi^{j-1} \le 0$.
+In the unconstrained case we can assume, more simply, that the fine level defect constraint is a negative constant $\chi^J = -JZ$ for $Z>0$ such that $Z \gg \|u^J\|_\infty$.
+Defining $\chi^j = - jZ$, and calculating $\phi^j = \chi^j - P \chi^{j-1} = - Z$, we see that both the downward VI \eqref{eq:fe:downvi} and the upward VI \eqref{eq:fe:upvi} can also be made unconstrained.
+It then follows that lines 2, 5, and 6, can be removed from Algorithm \ref{alg:nmcd}, and further that the constraint arguments can be removed from the smoothers and the coarse-level solver.
+Because the downward corrections $y^j$ are no longer separately constrained, we may simply refer to the corrected solution $w^j=g^j+y^j$.
+The smoothers can be regarded as in-place modifications of $w^j$.
+Also recognizing that the initial iterate for the coarse-level smoother is $w^{j-1}=\iR w^j$, the definition of the coarsened source functional can be rewritten
+\begin{equation}
+\ell^{j-1} = f^{j-1}\left(w^{j-1}\right) + R\left(\ell^j-f^j(w^j)\right). \label{eq:app:fas:levelsource}
+\end{equation}
+However, now we must be careful that going upward only the correction gets prolonged;compare Remark 5.3.9 in \cite{Trottenbergetal2001}.
+We note that the coarse correction produces an updated $w^{j-1}$ (previously denoted $g^{j-1}+y^{j-1}$).
+If $w^j$ denotes the result of down-smoothing, and if $z^j = y^j + P z^{j-1} = y^j + y^{j-1} + \dots + y^0$ is the sum of all prolonged corrections, we want to set $w^j = g^j + z^j = (g^j+y^j) + P(y^{j-1} + \dots + y^0)$ as the iterate going into the up-smoother.
+For this we write $w^j \gets w^j + P(w^{j-1} - \iR w^j)$.
+With these modifications we get pseudocode in Algorithm \ref{alg:fas}.
+
+\begin{pseudofloat}[H]
+\begin{pseudo} \label{ps:nmcd-vcycle}
+\pr{fas-vcycle}(\ell^J; w^J)\text{:} \\+
+    for $j=J$ downto $j=1$ \\+
+      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j;w^j)$ \\
+      $w^{j-1} = \iR w^j$ \\
+      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j)\right)$ \\-
+    $\text{\pr{solve}}(\ell^0;w^0)$ \\
+    for $j=1$ to $j=J$ \\+
+      $w^j \gets w^j + P (w^{j-1} - \iR w^j)$ \\
+      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j;w^j)$ \\-
+\end{pseudo}
+\caption{An FAS V-cycle results from removing the inequality constraints from \pr{nmcd-vcycle}, and also grouping $w^j=g^j+y^j$ (downward) and $w^j=g^j+z^j$ (upward).}
+\label{alg:fas}
+\end{pseudofloat}
+
+a negative constant, $\gamma^J=-JZ$, where $Z>0$ is larger than the magnitude of $u^J$, or of any prospective iterate $w^J$, say $Z > \|w^J\|_\infty$.
+
+FIXME get Algorithm 14 \cite{Bruneetal2015}
+
+FIXME by assuming $f$ is linear, reduce to Algorithm 4.7 in \cite{GraeserKornhuber2009} but with $\text{V}(\nu_1,\nu_2)$ cycles
 

### 2022-07-31 py/bratu2patch.py
my attempt to follow the NGS implemented in the PCPatch paper; this attempt fails; put on slack for Lawrence

@@ -1,0 +1,98 @@
+"""Attempt to solve Bratu equation in 2D by single-level NGS using PatchSNES:
+   -triangle u - lambda e^u = 0
+The exact solution of this problem is given by Liouville (1853); see
+Exercise 7.12 in Bueler (2021), "PETSc for Partial Differential Equations"
+and c/ch7/solns/bratu2D.c in https://github.com/bueler/p4pdes.
+Dirichlet conditions on unit square are taken from exact solution.
+"""
+
+from firedrake import *
+
+newtoncg = {
+    "snes_type": "newtonls",
+    #"snes_view": None,
+    "snes_converged_reason": None,
+    "snes_monitor": None,
+    "ksp_type": "cg",
+    "ksp_converged_reason": None,
+    "pc_type": "jacobi",
+}
+
+npcnewtoncg = {
+    "mat_type": "matfree",
+    "snes_type": "nrichardson",
+    #"snes_view": None,
+    "snes_converged_reason": None,
+    "snes_monitor": None,
+    "snes_npc_side": "left",
+    "npc_snes_type": "newtonls",
+    "npc_snes_linesearch_type": "basic",
+    "npc_ksp_type": "cg",
+    "npc_ksp_converged_reason": None,
+    "npc_pc_type": "jacobi",
+    #"npc_pc_type": "python",
+    #"npc_pc_python_type": "firedrake.AssembledPC",
+    #"assembled_ksp_type": "preonly",
+    #"assembled_pc_type": "lu",
+}
+
+ngssweep = {
+    "mat_type": "matfree",
+    "snes_type": "nrichardson",
+    #"snes_view": None,
+    "snes_converged_reason": None,
+    "snes_monitor": None,
+    "snes_npc_side": "left",
+    "npc_snes_type": "python",
+    "npc_snes_python_type": "firedrake.PatchSNES",
+    "npc_snes_max_it": 1,
+    "npc_snes_convergence_test": "skip",
+    "npc_snes_linesearch_type": "basic",
+    "npc_snes_linesearch_damping": 1.0,
+    "npc_patch_snes_patch_construct_python_type": "star",
+    "npc_patch_snes_patch_partition_of_unity": True,
+    "npc_patch_snes_patch_sub_mat_type": "seqaij",
[…]

### 2022-08-07 paper/mcd2.tex
a couple of pages of edits

@@
 
-The goal of this paper is to extend the constraint decomposition (CD) method of X.-C.~Tai \cite{Tai2003} to nonlinear variational inequality (VI) problems, such as obstacle problems, to which it has not been applied.
+We extend the constraint decomposition (CD) method of X.-C.~Tai \cite{Tai2003} to nonlinear variational inequality (VI) problems to which it has not been applied.
 The convergence of this method, for several types of decompositions, was proven by Tai for coercive VI problems which arise from minimization of a convex functional over a convex set.
@@
 
-The theory presented in \cite{Tai2003} is extended in three particular directions:
+The theory presented in \cite{Tai2003} is extended as follows:
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
@@
 % THIS IS THE HOPE
-
-\item We make finite element (FE) implementations of the multilevel CD algorithm more practical by addressing storage of intermediate quantities.
-Specifically, in Section \ref{sec:results} we show results from a full approximation storage (FAS; see \cite{Brandt1977,Bruneetal2015}) implementation for nonlinear operators, one which avoids global (Newton) linearization; compare \cite{GraeserKornhuber2009}.
-
-\item We observe that multilevel ``up-smoothing'' is intrinsically more efficient than ``down-smoothing'' (Section \ref{sec:multilevel}).
+\end{enumerate}
+Furthermore the multilevel CD algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} are extended in two practical ways:
+\begin{enumerate}
+\setcounter{enumi}{1} % resume at (ii)
+\item We make finite element (FE) implementations of multilevel CD algorithm more effective for nonlinear problems by applying a full approximation storage (FAS; see \cite{Brandt1977,Bruneetal2015}) approach.
+Results from the resulting \emph{nonlinear multilevel constraint decomposition} (NMCD) algorithm, which avoids global (Newton) linearization (compare \cite{GraeserKornhuber2009}), are shown in Section \ref{sec:results} for several nonlinear problems.
+\item In Sections \ref{sec:fe} and \ref{sec:multilevel} we explain why, given the manner in which constraint sets are decomposed in the multilevel context \cite{GraeserKornhuber2009}, multilevel ``up-smoothing'' is more efficient than ``down-smoothing''.
 This observation seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}.
@@
 
-For problems of porous-media type the nonlinear operator $f$ is not known to be coercive, so the convergence theory in \emph{(i)} does not apply.
-However, a full-cycle implementation of the multilevel CD method is demonstrated numerically.
-By ``freezing' the solution-dependent coefficient, the operator is approximated by a coercive operator, nonlinear in general, for the duration of the V-cycle.
-The resulting scheme is highly-effective for doubly-nonlinear diffusion operators (Section \ref{sec:results}).
+The iterates from a CD method are always admissible, and thus the operator need only be defined for admissible states; Section \ref{sec:results} includes a nontrivial example.
+Admissible-iterate methods should permit direct solutions of certain VI problems, such as fluid-layer dynamics problems \cite{Bueler2021conservation,JouvetBueler2012}, for which non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
+Our NMCD scheme is also designed for these problems, some of which require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation, as discussed in \cite{Bueler2021conservation}.
 
-For the classical obstacle problem with a Laplacian operator, certain multilevel techniques are known to improve performance relative to the multilevel CD method \cite{GraeserKornhuber2009}.
-These improved methods either track the active set in the discretization or modify the nodal basis functions, and in this sense they are discrete algorithms.
-By contrast the CD method applies at the level of the continuum problem (Sections \ref{sec:cd} and \ref{sec:convergence}); compare the truncated monotone multigrid method \cite{Kornhuber1994}, for example.
-Acceleration of our nonlinear multilevel CD method via active-set and/or basis-level manipulations represents a potential extension of the method here, and is a topic for future research.
+One of our examples demonstrates the application of the NMCD method to a problem of porous-media type, to which the extended convergence theory, \emph{(i)} above, does not apply because the nonlinear operator is not known to be coercive.
+By ``freezing' the solution-dependent coefficient, the operator is approximated by a coercive operator for the duration of the V-cycle.
+The resulting scheme is highly-effective for doubly-nonlinear diffusion operators, another type of problem to which the extended theory does not apply (Section \ref{sec:results}).
 
-The iterates from a CD method are always admissible, and thus the operator need only be defined for admissible states; Section \ref{sec:results} includes a nontrivial example.
-Admissible-iterate methods should permit direct solutions of certain VI problems, such as fluid-layer dynamics problems \cite{Bueler2021conservation,JouvetBueler2012}, for which non-admissible methods, such as semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
-The full-cycle scheme in \emph{(ii)} above is also designed for these problems, which are characterized by the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation.
+For the classical obstacle problem with a Laplacian operator, certain multilevel techniques are known to improve performance relative to the multilevel CD method \cite{GraeserKornhuber2009}, such as the truncated monotone multigrid method \cite{Kornhuber1994}.
+These improved methods either track the active set in the discretization or modify the nodal basis functions, and in this sense they are discrete algorithms, while by contrast the CD method applies at the level of the continuum problem (Sections \ref{sec:cd} and \ref{sec:convergence}).
+Acceleration of our NMCD method via active-set and/or basis-level manipulations represents a potential extension of the method here, and is a topic for future research.
 
@@
 
-Suppose $\cV$ is a real, reflexive Banach space with norm $\|\cdot\|$ and topological dual space $\cV'$.
+Suppose $\cV$ is a real and reflexive Banach space with norm $\|\cdot\|$, and denote its topological dual space by $\cV'$.
 Denote the dual pairing of $\phi \in \cV'$ and $v\in\cV$ by $\ip{\phi}{v} = \phi(v)$, and note that $\|\phi\|_{\cV'} = \sup_{\|v\|=1} |\ip{\phi}{v}|$ defines a (Banach space) norm on $\cV'$.
+
 Let $\cK \subset \cV$ be a nonempty closed and convex subset, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.
-
 For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and a linear \emph{source functional} $\ell\in \cV'$ we consider the following \emph{variational inequality} (VI) for the (exact) solution $u^*\in \cK$, if it exists:
@@
 \end{equation}
-Because $f$ is (generally) nonlinear, the source term $\ell$ is not strictly needed for posing this VI,\footnote{By redefining $f$ we may take $\ell=0$.} but its presence is helpful to algorithms in Section \ref{sec:multilevel}.
+Because $f$ is (generally) nonlinear, the source term $\ell$ is not strictly needed, and indeed by redefining $f$ we may take $\ell=0$, but its presence clarifies algorithms in Section \ref{sec:multilevel}.
 
-VI \eqref{eq:vi} generalizes the nonlinear system of equations $f(u^*)=\ell$ to the constraint set $\cK$.
+VI \eqref{eq:vi} generalizes the nonlinear system of equations $f(u^*)=\ell$ to problems where $u^*$ is also required to be in a constraint set $\cK$.
 Informally, if we conceptualize the dual pairing as an inner product then \eqref{eq:vi} says that the angle between $f(u^*)-\ell$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
-That is, \eqref{eq:vi} says that $f(u^*)-\ell$ may not be zero but it points directly into $\cK$.
-However, if $u^*$ is in the interior $\cK^\circ$ then \eqref{eq:vi} implies $f(u^*)=\ell$.
+That is, while $f(u^*)-\ell$ may not be zero, it points directly into $\cK$.
+However, if $u^*$ is in the interior of the constraint set, $u^*\in\cK^\circ$, then \eqref{eq:vi} implies $f(u^*)=\ell$.
 
@@
 \end{equation}
-We say VI \eqref{eq:vi} is \emph{monotone} if $f$ is monotone, and likewise for strictly monotone and coercive. \end{definition}
+\end{definition}
 
-It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, and also that the solution $u^* \in \cK$ is unique when $f$ is strictly monotone.
-As in the calculus of variations \cite{Evans2010}, coercivity permits a compactness argument for unbounded sets $\cK$; recall that the bounded, closed subsets of a reflexive Banach space are weakly compact.
-The condition of continuity can be weakened to only apply on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but the stronger condition will apply in our examples.
+It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, uniquely so when $f$ is strictly monotone.
+As in the calculus of variations \cite{Evans2010}, coercivity permits a compactness argument for unbounded sets $\cK$.
+(Recall that the bounded, closed subsets of a reflexive Banach space are weakly compact.)  The condition of continuity can be weakened to only apply on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but the stronger condition will apply in our examples.
 
-The coercive VIs solved in this paper satisfy a stronger inequality than \eqref{eq:coercive}, and thus they are well-posed.
+The coercive VIs solved in this paper satisfy a stronger inequality than \eqref{eq:monotone} and \eqref{eq:coercive}.
 
@@
 It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive.
-Note that Tai \cite{Tai2003} uses ``coercive'' for what we call $2$-coercive.
-The following result holds.
 
@@
 The CD methods of Tai \cite{Tai2003} address problem \eqref{eq:minimization} under the hypothesis that $F'$ is 2-coercive.
+(Note that Tai \cite{Tai2003} uses ``coercive'' for what we call $2$-coercive.)
 
 From now on $\Omega \subset \RR^d$ denotes a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.
-Sobolev spaces \cite{Evans2010} are denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$, with $W^{k,2}$ denoted $H^k$.
+Sobolev spaces \cite{Evans2010} are denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$.
 The following example includes the classical obstacle problem for the linear Laplacian \cite{GraeserKornhuber2009} and the $p$-Laplacian for $p\ge 2$ \cite{ChoeLewis1991}.
@@
 \begin{equation}
-\ip{f(u)}{v} = \int_\Omega a(x) |\grad u|^{p-2} \grad u \cdot \grad v\,dx. \label{eq:plaplacian}
+\ip{f(u)}{v} = \int_\Omega a(x) |\grad u|^{p-2} \grad u \cdot \grad v\,dx, \label{eq:plaplacian}
 \end{equation}
+a continuous map \cite[Theorem A.0.6]{Peral1997}.
 Now, if $x,y\in\RR^d$ then $(|x|^{p-2} x - |y|^{p-2} y)\cdot (x-y) \ge 2^{2-p} |x-y|^p$ \cite[see Appendix A and references therein]{Bueler2021conservation}.
@@
     $$\ip{f(u) - f(v)}{u-v} \ge 2^{2-p} a_0 \|\grad u - \grad v\|_p^p \ge 2^{2-p} a_0 C \|u-v\|^p$$
-for some $C>0$, and thus $f$ is $p$-coercive.  On the other hand, for $g\in\cV'$ define
-    $$F(v) = \int_\Omega \frac{a(x)}{p} |\grad v|^p\,dx - \ip{g}{v}.$$
-Then $F'(v) = f(v) - g$ and $F$ is a convex functional (since $f$ is coercive).
+for some $C>0$, and thus $f$ is $p$-coercive.  On the other hand, for $\ell\in\cV'$ define
+    $$F(v) = \int_\Omega \frac{a(x)}{p} |\grad v|^p\,dx - \ip{\ell}{v}.$$
+Then $F'(v) = f(v) - \ell$ and $F$ is a convex functional (since $f$ is coercive).
 For any closed and convex $\cK\subset \cV$, VI problem \eqref{eq:vi} for is equivalent to optimization problem \eqref{eq:minimization}.\end{example}
@@
 The map in \eqref{eq:plaplacian} is also coercive if $1<p<2$, but the proof is somewhat different \cite[Theorem 4.4]{Bueler2021conservation}.
-In Section \ref{sec:results} we need only the $p\ge 2$ case.
+In Section \ref{sec:results} we consider only the $p\ge 2$ case.
 
@@
 \begin{lemma}  \label{lem:advectionskew}  \cite{Elmanetal2014}\,  Suppose $\bX :\Omega \to \RR^d$ is a bounded and boundedly-differentiable vector field on $\Omega$ with zero divergence ($\Div \bX=0$).
-For $u,v \in H^1(\Omega)$ let $b(u,v) = \int_\Omega (\bX \cdot \grad u) v\,dx$.
+For $u,v \in W^{1,2}(\Omega)$ let $b(u,v) = \int_\Omega (\bX \cdot \grad u) v\,dx$.
 Then $b(u,u) = \frac{1}{2} \int_{\partial \Omega} u^2 \bX\cdot \bn\,dx$ where $\bn$ is the outward normal on $\partial \Omega$.

### 2022-08-07 paper/genfigs/cartoon.py
add gray dots suggesting unboundedness

@@ -21,6 +21,7 @@ plt.figure(figsize=(8, 6))
 makeaxes(-0.25, 1.2, -0.1, 0.8)
 plt.text(-0.05, -0.07, r'$0$', fontsize=fsize)
 
+# filled gray box
 psi1 = 0.2
 psi2 = 0.35
 plt.fill([psi1, 1.2, 1.2, psi1, psi1],
@@ -28,10 +29,15 @@ plt.fill([psi1, 1.2, 1.2, psi1, psi1],
          'k', alpha = 0.2)
 plt.text(0.9, 0.6, r'$\mathcal{K}$', fontsize=fsize)
 
+# put gray dots suggesting unboundedness
+plt.plot(0.01 + np.array([1.25, 1.30, 1.35]), 0.6 * np.array([1,1,1]), 'k.', ms = 12.0, markeredgecolor='w', alpha = 0.3)
+plt.plot(0.7 * np.array([1,1,1]), 0.01 + np.array([0.85, 0.90, 0.95]), 'k.', ms = 12.0, markeredgecolor='w', alpha = 0.3)
+plt.plot([1.25, 1.30, 1.35], [0.85, 0.90, 0.95], 'k.', ms = 12.0, markeredgecolor='w', alpha = 0.3)
+
 plt.text(1.28, -0.02, r'$\mathcal{V}_1$', fontsize=fsize)
 plt.plot(psi1, 0.0, 'k.', ms = 12.0)
 plt.plot([psi1, 1.2], [0.0, 0.0], 'k', lw = 2.5)
-plt.text(psi1 - 0.07, - 0.085, r'$\psi(x_1)$', fontsize=fsize)
+plt.text(psi1 - 0.07, - 0.09, r'$\psi(x_1)$', fontsize=fsize)
 plt.text(0.7, - 0.09, r'$\mathcal{K}_1$', fontsize=fsize)
 plt.gca().annotate('', xy=(0.7, psi2 - 0.3), xytext=(0.7, psi2 - 0.05),
                    arrowprops=dict(arrowstyle='->, head_length=0.8, head_width=0.5'))
@@ -40,8 +46,8 @@ plt.text(0.72, 0.15, r'$R_1$', fontsize=fsize)
 plt.text(-0.02, 0.89, r'$\mathcal{V}_2$', fontsize=fsize)
 plt.plot(0.0, psi2, 'k.', ms = 12.0)
[…]

### 2022-08-07 paper/mcd2.tex
fix latex errors

@@
         \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$ \rm{we have} \\+
-            $\boxed{{\large \strut} \ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}}$ \\--
+            $\boxed{\ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}}$ \\--
     $\hat w = \sum_i \hat w_i\in\cK$ \\
@@
 (\hat r_{\text{down}}(y^j))_p = \begin{cases} \ip{f^j(g^j+y^j)-\ell^j}{\psi_p^j}, & y^j(x_p^j) > \phi^j(x_p^j), \\
-                                  \min\{\ip{f^j(g^j+y^j)-\ell^j}{\psi_p^j},0\}, & y^j(x_p^j) = \phi^j(x_p^j). \end{cases} \label{eq:cpresidual}
+                                  \min\{\ip{f^j(g^j+y^j)-\ell^j}{\psi_p^j},0\}, & y^j(x_p^j) = \phi^j(x_p^j). \end{cases} \label{eq:dncpresidual}
 \end{equation}
@@
 \begin{pseudofloat}[H]
-\begin{pseudo} \label{ps:nmcd-vcycle}
+\begin{pseudo} \label{ps:fas-vcycle}
 \pr{fas-vcycle}(\ell^J; w^J)\text{:} \\+

### 2022-08-07 paper/mcd2.tex
edits through page 7

@@
 \begin{example}  \label{ex:advectiondiffusion}  Suppose $\partial\Omega$ is partitioned into Dirichlet and Neumann portions, i.e.~$\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$, with $\partial_D\Omega$ of positive measure.
-Let $\cV = H_0^1(\Omega)$ be the space of functions which are zero along $\partial_D\Omega$.
+Let $\cV = W_0^{1,2}(\Omega)$ be the space of functions which are zero along $\partial_D\Omega$.
 Consider a divergence-free velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, but additionally assume that the flow is outward on the Neumann boundary, $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.
-For $u,v \in \cV = H_0^1(\Omega)$ and $\eps>0$ define
+For $u,v \in \cV = W_0^{1,2}(\Omega)$ and $\eps>0$ define
 \begin{equation}
@@
 Numerical solver performance for Examples \ref{ex:plaplacian}, \ref{ex:advectiondiffusion}, and \ref{ex:porous} will be considered in Section \ref{sec:results}.
-Further examples of nonlinear VI problems appear in ice sheet models and other geophysical fluids \cite{Bueler2021conservation,Calvoetal2002,JouvetBueler2012}.
+Further examples of nonlinear VI problems appear in ice sheet models \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids.
+In the above Examples $f(u)$ may be defined over all of $\mathcal{V}$, but in some geophysical fluid problems the value of $f(u)$ is only defined for admissible $u\in \mathcal{K}$ because $u$ defines the fluid layer thickness \cite{Bueler2021conservation}.
 
-\section{Constraint decomposition method} \label{sec:cd}
+\section{Constraint decomposition} \label{sec:cd}
 
@@
 The sum in \eqref{eq:constraintdecomp} is required to hold in two senses \cite{TaiTseng2002}: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.
-(Sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp} because the $\cV_i$ are subspaces.)  Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique.
+Note that sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp}, because the $\cV_i$ are subspaces, and that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique.
 Also, $\cK_i \not\subset \cK$ in many applications; see the cartoon in Figure \ref{fig:cartoon}.
@@
 \begin{figure}[ht]
-\includegraphics[width=0.55\textwidth]{genfigs/cartoon.pdf}
+\includegraphics[width=0.6\textwidth]{genfigs/cartoon.pdf}
 \caption{Suppose $\mathcal{V}$ is the space of real functions on a two-point set $\Omega=\{x_1,x_2\}$.
-A constraint decomposition (CD) for a one-sided obstacle problem with $\mathcal{K}=\{v\ge \psi\}$ looks like this.}
+A constraint decomposition (CD) for a one-sided obstacle problem with $\mathcal{K}=\{v\ge \psi\}$ might look like this.}
 \label{fig:cartoon}
@@
 
-For each $\cK_i$ we will also assume that there are bounded, generally-nonlinear restriction operators $R_i : \cK \to \cK_i$ such that if $v \in \cK$ then
+For each $\cK_i$ we will also assume that there are bounded, generally-nonlinear \emph{restriction} operators $R_i : \cK \to \cK_i$ such that if $v \in \cK$ then
 \begin{equation}
@@
 
-In Section \ref{sec:multilevel} we will introduce discretizations and practical algorithms, but we observe here that the CD concept applies at the level of the continuum problem.
-The following two examples illustrate this for obstacle problems \cite{GraeserKornhuber2009}.
+In Section \ref{sec:multilevel} we will introduce discretizations, and practical algorithms for finite-dimensional VI problems, but observe here that this CD concept applies at the level of the continuum problem.
+The following two obstacle problem examples illustrate this.
+The first is an overlapping domain decomposition.
 
-\begin{example}  \label{ex:domaindecomposition}  Consider an overlapping domain decomposition as follows.
-For a bounded domain $\Omega \subset \RR^d$, let $\cV = W_0^{k,p}(\Omega)$ for $k\ge 0$ and $p\ge 1$, and suppose the obstacle $\psi \in W^{k,p}(\Omega)$ satisfies $\psi|_{\partial \Omega} \le 0$.
-Let $\cK = \{v \ge \psi\} \subset \cV$.
-Suppose further that $\{\phi_i\}_{i=0}^{m-1}$ is a smooth partition of unity on $\Omega$, satisfying $0 \le \phi_i\le 1$ and $\sum_i \phi_i = 1$, and let $\Omega_i$ be the support of $\phi_i$.
+\begin{example}  \label{ex:domaindecomposition}  For a bounded domain $\Omega \subset \RR^d$, let $\cV = W_0^{k,p}(\Omega)$ for $k\ge 0$ and $p\ge 1$.
+Suppose an obstacle $\psi \in W^{k,p}(\Omega)$ satisfies $\psi|_{\partial \Omega} \le 0$, and let $\cK = \{v \ge \psi\} \subset \cV$.
+Suppose further that $\{\phi_i\}_{i=0}^{m-1}$ is a smooth partition of unity on $\Omega$, satisfying $0 \le \phi_i\le 1$ and $\sum_i \phi_i = 1$, and let $\Omega_i$ be the open support of $\phi_i$.
 Let $\cV_i = \{w \in \cV:w|_{\Omega \setminus \Omega_i} =0 \}$, $\cK_i = \{v \in \cV_i: v \ge \phi_i \psi\}$, and $R_i(v) = \phi_i v$.
@@
 
-Our second example is a disjoint frequency decomposition.  A multilevel FE CD, e.g.~the one proposed in Section \ref{sec:multilevel}, approximates such a frequency decomposition.
+Our second example is a disjoint frequency decomposition.  A multilevel CD, e.g.~the one proposed in Sections \ref{sec:fe} and \ref{sec:multilevel}, approximates such a frequency decomposition.
 
-\begin{example}  \label{ex:frequencydecomposition}  For simplicity suppose $\Omega = (0,a)^d \subset \RR^d$ is a cube, and let $\cV = H_{\text{per}}^k(\Omega)$, $k\ge 0$, be the periodic functions.
-Suppose $\psi \in H_{\text{per}}^k(\Omega)$ and let $\cK = \{v \ge \psi\} \subset \cV$.
-Without using any detailed notation for Fourier representation, but noting that the frequencies are discrete, suppose $\{\cV_i\}$ are $m<\infty$ subspaces of $\cV$ defined by an (nonoverlapping) partition by frequency, thus satisfying \eqref{eq:subspacedecomp} as an orthogonal decomposition.
+\begin{example}  \label{ex:frequencydecomposition}  For simplicity suppose $\Omega = (0,a)^d \subset \RR^d$ is a cube, and let $\cV = W_{\text{per}}^{p,k}(\Omega)$, for $p\ge 2$ and $k\ge 0$, be the periodic functions.
+Suppose $\psi \in W_{\text{per}}^{p,k}(\Omega)$ and let $\cK = \{v \ge \psi\} \subset \cV$.
+Without using any detailed notation for Fourier representation, but noting that the frequencies are discrete because $\Omega$ is compact, suppose $\{\cV_i\}$ are $m<\infty$ subspaces of $\cV$ defined by an (nonoverlapping) partition by frequency, thus satisfying \eqref{eq:subspacedecomp} as an orthogonal decomposition.
 Suppose $P_i:\cV \to \cV_i$ are the corresponding orthogonal projections, satisfying $I = \sum_i P_i$.
@@
 
-Note that $\cK_i \not\subset \cK$ in many cases.
+Again note that $\cK_i \not\subset \cK$ in these cases.
 Specifically in Example \ref{ex:domaindecomposition}, if $\psi$ is positive over portions of $\Omega$ where the decomposition into overlapping subdomains $\Omega_i$ is nontrivial, then $\cK_i \not\subset \cK$, and similarly for Example \ref{ex:frequencydecomposition}.
@@
 
-Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below state the basic CD method as functions which improve a current iterate $u \in \cK$, by solving ``smaller'' VI problems over each subset $\cK_i$, so that the output $w\in\cK$ should be closer to the solution $u^* \in \cK$ of \eqref{eq:vi}.
-While Algorithm \ref{alg:basiccd-add} generalizes a Jacobi iteration, Algorithm \ref{alg:basiccd-mult} generalizes Gauss-Seidel.
+\section{Iterations based on constraint decomposition} \label{sec:cditers}
+
+Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below record basic CD iterative methods as functions which improve a current iterate $u \in \cK$ by solving ``smaller'' VI problems over each subset $\cK_i$.
+Their outputs $w\in\cK$ should be closer to the solution $u^* \in \cK$ of \eqref{eq:vi}.
 
@@
 \begin{pseudo*}
-\pr{cd-add}(f,\ell,u,\cK,\cK_i,R_i)\text{:} \\+
-    for $i \in \{0,\dots,m-1\}$: \\+
-        \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$ \rm{we have} \\+
+\pr{cd-add}(u)\text{:} \\+
+    for all $i \in \{0,\dots,m-1\}$: \\+
+        \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$, \\+
             $\boxed{\ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}}$ \\--
     $\hat w = \sum_i \hat w_i\in\cK$ \\
-    return $w=(1-\alpha) u + \alpha \hat w\in\cK$
+    return $w=(1-\alpha) u + \alpha \hat w$
 \end{pseudo*}
@@
 
-The \textbf{for} loop in Algorithm \ref{alg:basiccd-add}, the additive or parallel CD method, can be computed in any order.
-Note that $u-R_iu+\hat w_i$ involves removing the part of $u$ which lies in $\mathcal{K}_i$ before adding-back the improved subset solution $\hat w_i$.
-By contrast, the following multiplicative (successive) Algorithm \ref{alg:basiccd-mult} orders the sets $\mathcal{K}_i$, and each subset solution updates the global iterate before proceeding to the next subset.
-The sum $\sum_{j>i} R_j u$ should be read as ``keep the parts of the current iterate which we have not yet improved.''
+In Algorithm \ref{alg:basiccd-add}, the additive or parallel CD method, the \textbf{for all} can be computed in any order.
+Note that $u-R_iu+\hat w_i$ removes the part of $u$ which lies in $\mathcal{K}_i$ before adding-back the improved subset solution $\hat w_i$, also from $\mathcal{K}_i$.
+By contrast, in the following multiplicative (successive) Algorithm \ref{alg:basiccd-mult} each subset solution updates the global iterate before proceeding to the next subset.
+The sets $\mathcal{K}_i$ are addressed in a fixed order, and the sum $\sum_{j>i} R_j u$ should be read as ``keep those parts of the input $u$ which we have not yet improved.''
 
@@
 \begin{pseudo*}
-\pr{cd-mult}(f,\ell,u,\cK,\cK_i,R_i)\text{:} \\+
+\pr{cd-mult}(u)\text{:} \\+
     for $i = 0,\dots,m-1$: \\+
-        \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$ \rm{we have} \\+
+        \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$, \\+
             $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}}$ \\-
@@
     $\hat w = \sum_i \hat w_i\in\cK$ \\
-    return $w=(1-\alpha) u + \alpha \hat w\in\cK$
+    return $w=(1-\alpha) u + \alpha \hat w$
 \end{pseudo*}
@@
 
-Note the use of a damping parameter $0<\alpha\le 1$.  In Algorithm \ref{alg:basiccd-mult} we must update the in-progress iterate $w_i$ using the same damping that is applied to the final output $w$.
+While Algorithm \ref{alg:basiccd-add} generalizes a Jacobi iteration, Algorithm \ref{alg:basiccd-mult} generalizes Gauss-Seidel.
+Note the use of a damping parameter $0<\alpha\le 1$ in both Algorithms.
+In Algorithm \ref{alg:basiccd-mult} we must update the in-progress iterate $w_i$ using the same damping that is applied to the final output $w$.
 
-In each boxed VI above the argument of $f$ is an element of $\cK$, as the reader should confirm.
-On the other hand, by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may also write the test vector $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors from $\cK$, namely, for the two Algorithms respectively:
+In each boxed VI the argument of $f$ is an element of $\cK$, as the reader should confirm.
+On the other hand, by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may also write the test vector $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors from $\cK$.
+For the two Algorithms respectively:
 \begin{align*}
@@
 
-Now define
+For either Algorithm we define
 \begin{equation}
@@
 \end{equation}
-the $i$th subset increment in either Algorithm.
-As the reader may check, $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$, and $\hat w = u^* + \sum_i \hat w_i - R_i u^*$ also holds.
+the $i$th \emph{subset update}.
+As the reader may check, $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.
+Note $\hat w = u^* + \sum_i \hat w_i - R_i u^*$ also holds.
 We will use these identities in Section \ref{sec:convergence}.
@@
 
-We refer to Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} as the \emph{basic CD method}.
+We refer to Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} as \emph{basic CD iterations}.
 To construct efficient FE implementations (Section \ref{sec:multilevel}) we will need additional notions.
-While the above Algorithms are indeed meaningful even when $f$ is nonlinear, non-local, or defined only on $\cK$, practical implementation for such general problems seems not to have been addressed in the literature.
-In particular, references \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem.
-The following example explains the implementation-relevant simplifications available in that case.
+While the basic iterations are meaningful even when $f$ is nonlinear, non-local, or defined only on $\cK$, practical implementation for such general problems seems not to have been addressed in the literature.
+In particular, references \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem for which $f$ is linear.
+The following example explains the implementation-relevant simplifications available in that simpler case.
 
@@
 
-\section{Convergence of the basic algorithm} \label{sec:convergence}
+\section{Convergence of constraint decomposition iterations} \label{sec:convergence}
 

### 2022-08-08 paper/mcd2.tex
complete a round of edits

@@
 
-We refer to Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} as \emph{basic CD iterations}.
-To construct efficient FE implementations (Section \ref{sec:multilevel}) we will need additional notions.
-While the basic iterations are meaningful even when $f$ is nonlinear, non-local, or defined only on $\cK$, practical implementation for such general problems seems not to have been addressed in the literature.
-In particular, references \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem for which $f$ is linear.
-The following example explains the implementation-relevant simplifications available in that simpler case.
+We refer to Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} as the \emph{basic CD iterations}.
+These basic iterations are meaningful even when $f$ is nonlinear, non-local, or defined only on $\cK$.
+However, practical and efficient FE implementation for such general problems seems not to have been addressed in the literature, and will need additional notions.
+In particular, references \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem for which $f$ is linear, local, and defined for all inputs.
+The following example exposes the implementation simplifications available in that case.
 
 \begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear and defined on all of $\cV$.
-Furthermore suppose $f$ is local in the sense that a basis $\{\phi_i\}$ of $\cV_i$ exists, with $\supp \phi_i$ compactly-contained in $\Omega$, such that for any $z\in\mathcal{V}$ the value $\ip{f(\phi_i)}{z}$ can be computed by an integral over the support of $\phi_i$.
-Considering only additive Algorithm \ref{alg:basiccd-add} for simplicity, by linearity the boxed VI over $\cK_i$ can be written as
+Furthermore suppose $f$ is a differential operator, thus local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral only over the support of $\phi \in \mathcal{V}$.
+Considering additive Algorithm \ref{alg:basiccd-add} for concreteness, by linearity the boxed VI can be written as
 \begin{equation}
-\ip{f(e_i)}{v_i-\hat w_i} \ge \ip{\tilde g}{v_i-\hat w_i} \label{eq:linearlocalvi}
+\ip{f(e_i)}{v_i-\hat w_i} \ge \ip{\tilde\ell}{v_i-\hat w_i} \label{eq:linearlocalvi}
 \end{equation}
-where $\tilde g = g - f(u)$.
-Note $e_i = \hat w_i - R_i u \in \cV_i \notin \cK$, in general, but VIs \eqref{eq:linearlocalvi} make sense because $f$ is defined over $\cV$.
-Each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $\ip{f(u)}{\cdot}$, namely the quantity included here into the source term.
-The VI \eqref{eq:vi} is then approximately solved in any incremental and efficient manner, by computations over the basis supports.
+where $\tilde\ell = \ell - f(u)$.
+Note $e_i = \hat w_i - R_i u \in \cV_i$ is not generally admissible ($e_i \notin \cK$), but VIs \eqref{eq:linearlocalvi} make sense because $f$ is defined over $\cV$.
+Each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $f(u)$, here included into a source term.
+If a hat function basis exists for $\cV_i$, namely of small-support functions, then by linearity of $f$ a solver for VI \eqref{eq:linearlocalvi} can be implemented using only integrals over the hat function supports.
 \end{example}
 
-The assumption that $\ip{f(\phi_i)}{z}$ can be evaluated over the support of $\phi_i$ is satisfied when $f$ is a differential operator, $\cV$ is an FE space, and $\phi_i$ are hat functions; see Section \ref{sec:multilevel}.
-Constrast non-local integral operators where $\ip{f(\phi_i)}{z}$ requires an integral over $\Omega$ even if $\phi_i$ is a hat function with small support.
+When $f$ has all the nice properties listed in Example \ref{ex:fnice} then a solver can take advantage of implementation efficiencies which are unavailable in general.
+The FE approximation of VI \eqref{eq:vi}---see Section \ref{sec:fe}---can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.
+The actual solution process for VI \eqref{eq:linearlocalvi} is not really the concern here; we are observing that the storage needed for problem \eqref{eq:linearlocalvi}, and the cost of residual evaluation, are smaller than in the general case.
+For more difficult problems certain additional ideas, illustrated in Section \ref{sec:multilevel}, are needed for practical implementation.
+Specifically, we will extend a multilevel CD iteration to a method for nonlinear functions $f:\cK\to\cV'$, i.e.~defined only on the constraint set $\cK$, by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
 
-When $f$ has all the nice properties listed in Example \ref{ex:fnice} then a solver can take advantage of implementation efficiencies which are unavailable in general.
-The actual solution process for VI \eqref{eq:linearlocalvi} is not the concern here; we are observing that the data of problem \eqref{eq:linearlocalvi}, and the cost of residual evaluation, are smaller than in the general case.
-For more difficult problems additional ideas, illustrated in Section \ref{sec:multilevel}, are needed for practical implementation.
-We will extend the basic method to nonlinear $f$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
-However, extending the algorithm to non-local functionals $f$ is a topic for future research.
+One may also contrast Example \ref{ex:fnice} with the case of a non-local integral operator $f$ \cite{Bueler2021conservation}, where $\ip{f(\phi)}{z}$ requires an integral over $\Omega$ even if $\phi$ is a function with small support such as a hat function.
+Extension of the multilevel CD method to such non-local functionals $f$ is a topic for future research.
 
-\section{Convergence of constraint decomposition iterations} \label{sec:convergence}
+\section{Convergence of a constraint decomposition iteration} \label{sec:convergence}
 
-We now prove the convergence of additive CD Algorithm \ref{alg:basiccd-add}. % ASPIRATION
-This will be possible if we restrict to $2$-coercive operators, require $\alpha \le 1/m$, and make certain other assumptions as in \cite{Tai2003}.
-However, our proof applies to VI problems which are not optimizations, and for which $f$ only applies to admissible elements, and thus extends the theory in \cite{Tai2003}.
-First we define a new quantity related to VI problem \eqref{eq:vi}.
+We now prove the convergence of the basic additive CD iteration, Algorithm \ref{alg:basiccd-add}. % ASPIRATION
+A convergence proof becomes possible if we restrict to $2$-coercive operators, require substantial damping, and make certain other assumptions as in \cite{Tai2003}.
+However, our proof applies to VI problems which are not optimizations, thus extending the theory in \cite{Tai2003}.
 
-\begin{definition} Suppose $f:\cK \to \cV'$ and $g \in \cV'$.  For $u,v \in \cK$ let
+First we define a new form related to VI problem \eqref{eq:vi}.
+
+\begin{definition} Suppose $f:\cK \to \cV'$ and $\ell \in \cV'$.  For $v,w \in \cK$ let
 \begin{equation}
-  E(v,u) = \ip{f(v)}{v-u} - \ip{g}{v-u}.  \label{eq:normlikedefn}
+  E(v,w) = \ip{f(v)-\ell}{v-w}.  \label{eq:normlikedefn}
 \end{equation}
@@
 
-If $u^*$ solves \eqref{eq:vi} then $E(v,u^*)$ is somewhat like an ``merit function'' for an iterate $v$, as used in solving nonlinear equations \cite{NocedalWright2006}, replacing the scalar objective which is assumed in \cite{Tai2003}.
+If $u^*$ solves \eqref{eq:vi} then $E(v,u^*)$ is a ``merit function'' for an iterate $v$, a concept used in solving nonlinear equations \cite{NocedalWright2006}, except that $u^*$ is not generally known so this meaning is only theoretical.
+In any case, our use of $E(v,u^*)$ replaces the scalar objective ``$F(v)$'' used in \cite{Tai2003}.
 The next lemma shows $E(v,u^*)$ is bounded below by a norm.
 
-\begin{lemma} \label{lem:normlike}  Suppose $f:\mathcal{K} \to \mathcal{V}'$ is $p$-coercive and $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.  For $v \in \mathcal{K}$,
+\begin{lemma} \label{lem:normlike}  Suppose $f:\mathcal{K} \to \mathcal{V}'$ is $p$-coercive with constant $\kappa>0$, and $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.  For  all $v \in \mathcal{K}$,
 \begin{equation}
@@
 \begin{align*}
-E(v,u^*) &= \ip{f(v)}{v-u^*} - \ip{f(u^*)}{v-u^*} + \ip{f(u^*)}{v-u^*} - \ip{g}{v-u^*} \\
+E(v,u^*) &= \ip{f(v)}{v-u^*} - \ip{f(u^*)}{v-u^*} + \ip{f(u^*)}{v-u^*} - \ip{\ell}{v-u^*} \\
    &\ge \ip{f(v)-f(u^*)}{v-u^*} + 0 \ge \kappa \|v-u^*\|^p.  \qedhere
@@
 Next we make two assumptions which are essentially the same as (7) and (8) in \cite{Tai2003}.
+These inequalities might be described as ``totally Lipschitz'' requirements, over the CD defined by \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, for the maps $R_i$ and $f$, respectively.
 
@@
 \end{equation}
-for all $u,v\in\cK$.  Furthermore, there exists a constant $C_2>0$ so that for all $v_i \in \cV_i$ and $y_j \in \cV_j$, and all $w_{ij} \in \cK$ such that $w_{ij} + v_i \in \cK$, it holds that
+for all $u,v\in\cK$.  There exists another constant $C_2>0$ so that for all $v_i \in \cV_i$ and $y_j \in \cV_j$, and all $w_{ij} \in \cK$ such that $w_{ij} + v_i \in \cK$, it holds that
 \begin{equation}
@@
 
-These inequalities might be described as ``totally Lipschitz'' requirements for the maps $R_i$ and $f$, respectively.  We make the following observations:
+We make the following observations:
 \begin{itemize}
-\item Assumption \eqref{as:lipschitzrestrictions} addresses only the CD, and not the VI problem $f$.
-Tai \cite{Tai2003} gives a value of $C_1$ for obstacle problem constraint decompositions using $P_1$ FE spaces over shape-regular and quasi-uniform triangulations.
-In particular, $C_1$ is known for overlapping domain decompositions, with or without an additional coarse mesh, and standard multilevel hierarchies.
-The cases we will demonstrate in Section \ref{sec:results} are covered.
-\item If $f$ is defined on all of $\mathcal{V}$ then one might be able to verify \eqref{as:lipschitzresidual} over a subspace decomposition, not the CD.
-For example, in the classical obstacle problem one may verify \eqref{as:lipschitzresidual} over arbitrary $w_{ij} \in \cV$ \cite{Tai2003}.
-\item If we assume $f$ itself is Lipschitz then the existence of $C_2$ is clear \cite{TaiXu2002}.
-However, our convergence rate bound is improved when $C_2$ is smaller, so the smallest $C_2$ for which \eqref{as:lipschitzresidual} holds is used.
-\item The constants $C_1,C_2$ generally depend on $m$.
-Furthermore, regarding the convergence proof below, Tai \cite{Tai2003} observes that if the iterates from Algorithm \ref{alg:basiccd-add} are bounded in $\mathcal{K}$ then constants $C_1,C_2$ are permitted to depend continuously on $u,v,w_{ij},v_i,y_i$.
-\item The unconstrained versions of \eqref{as:lipschitzrestrictions} and \eqref{as:lipschitzresidual} are (13), (14) in \cite{TaiXu2002}, respectively.
+\item The constants $C_1,C_2$ generally depend on $m$, the number of sets in the CD.
+In a multilevel CD method (Section \ref{sec:multilevel}) this number is logarithmic in the number of degrees of freedom (nodes) in the fine-level mesh.
+\item Regarding the convergence proof below, if the iterates from Algorithm \ref{alg:basiccd-add} are bounded in $\mathcal{K}$ then constants $C_1,C_2$ are permitted to depend continuously on $u,v,w_{ij},v_i,y_i$ \cite{Tai2003}.
+\item Bound \eqref{as:lipschitzrestrictions} addresses only the CD, and not the nonlinear operator $f$.
+Tai \cite{Tai2003} gives a value of $C_1$ for certain obstacle problem CDs using $P_1$ FE spaces over shape-regular and quasi-uniform triangulations.
+In particular, $C_1$ is known for the standard multilevel hierarchy, so the cases demonstrated in Section \ref{sec:results} are covered.
+\item If $f$ is defined on all of $\mathcal{V}$ then one may verify \eqref{as:lipschitzresidual} over a subspace, versus constraint, decomposition.
+For example, \eqref{as:lipschitzresidual} is verified for the classical obstacle problem over arbitrary $w_{ij} \in \cV$.
+(Unconstrained versions of \eqref{as:lipschitzrestrictions}, \eqref{as:lipschitzresidual} are (13), (14) in \cite{TaiXu2002}, respectively.)
+\item If $f$ is Lipschitz then the existence of $C_2$ is clear \cite{TaiXu2002}.
+However, our convergence rate bound is improved by taking the smallest $C_2$ for which \eqref{as:lipschitzresidual} holds.
 \end{itemize}
 
-From the above definitions and assumptions we have the following estimate for Algorithm \ref{alg:basiccd-add}.  Our proof follows \cite{Tai2003}.
+From the above definitions and assumptions we have the following estimate for Algorithm \ref{alg:basiccd-add}.  Our proof roughly follows that of Theorem 1 in \cite{Tai2003}.
 
-\begin{lemma} \label{lem:core}  Suppose $f$ is $2$-coercive and $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.
+\begin{lemma} \label{lem:core}  Suppose $f$ is $2$-coercive with constant $\kappa>0$, and that $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.
 Assume $\mathcal{V}_i$, $\mathcal{K}_i$, $R_i$ form a CD satisfying \eqref{as:lipschitzrestrictions}, and that \eqref{as:lipschitzresidual} is satisfied.
@@
 \begin{equation}
-   E(\hat w,u^*) \le C_2 \sum_{i=0}^{m-1} \|e_i\|^2 + \kappa^{-1} C_1 C_2 \left(\sum_{i=0}^{m-1} \|e_i\|^2\right)^{1/2} E(u,u^*)^{1/2}. \label{eq:core}
+   E(\hat w,u^*) \le C_2 \sum_{i=0}^{m-1} \|e_i\|^2 + \kappa^{-1} C_1 C_2 \left(\sum_{i=0}^{m-1} \|e_i\|^2\right)^{1/2} E(u,u^*)^{1/2} \label{eq:core}
 \end{equation}
+where again $e_i=\hat w_i - R_i u$ (equation \eqref{eq:ithupdate}).
 \end{lemma}
@@
 \begin{align}
-E(\hat w,u^*) &= \sum_{i=0}^{m-1} \ip{f(\hat w)}{\hat w_i - R_i u^*} - \ip{g}{\hat w_i - R_i u^*} \label{eq:startcore} \\
+E(\hat w,u^*) &= \sum_{i=0}^{m-1} \ip{f(\hat w)}{\hat w_i - R_i u^*} - \ip{\ell}{\hat w_i - R_i u^*} \notag \\
     &\le \sum_{i=0}^{m-1} \ip{f(\hat w)}{\hat w_i - R_i u^*} + \ip{f(u + e_i)}{R_i u^* - \hat w_i} \notag \\
-    &= \sum_{i=0}^{m-1} \ip{f(\hat w) - f(u + e_i)}{\hat w_i - R_i u^*}. \notag
+    &= \sum_{i=0}^{m-1} \ip{f(\hat w) - f(u + e_i)}{\hat w_i - R_i u^*}. \label{eq:startcore}
 \end{align}
@@
 
-Let $Z=\left(\sum_{i=0}^{m-1} \|e_i\|^2\right)^{1/2}$.  FIXME: MAKE THE RENUMBERING EXPLICIT
+Let $Z=\left(\sum_{i=0}^{m-1} \|e_i\|^2\right)^{1/2}$.  FIXME: MAKE THE RENUMBERING EXPLICIT AND NOTE $v_0=0$ IN APPLYING \eqref{as:lipschitzresidual}
 

### 2022-08-09 py/bratu1.py
remind myself where things stand with these Firedrake attempts

@@ -1,4 +1,5 @@
-"""Solve Bratu equation in 1D by FAS multigrid:
+"""Solve Bratu equation in 1D by FAS multigrid, using explicit
+calls to prolong, restrict, and inject:
    -u'' - lambda e^u = f(x)
 with Dirichlet conditions u(0)=u(1)=0 and
    f(x) = 9 pi^2 sin(3 pi x) − lambda exp(sin(3 pi x)).

### 2022-08-09 py/bratu2patch.py

@@ -1,9 +1,10 @@
-"""Attempt to solve Bratu equation in 2D by single-level NGS using PatchSNES:
+"""Solve Bratu equation in 2D by single-level NGS using PatchSNES:
    -triangle u - lambda e^u = 0
 The exact solution of this problem is given by Liouville (1853); see
 Exercise 7.12 in Bueler (2021), "PETSc for Partial Differential Equations"
 and c/ch7/solns/bratu2D.c in https://github.com/bueler/p4pdes.
 Dirichlet conditions on unit square are taken from exact solution.
+Compare to a Newton solve using Jacobi-preconditioned CG.
 """
 
 from firedrake import *
@@ -12,28 +13,10 @@ newtoncg = {
     "snes_type": "newtonls",
     #"snes_view": None,
     "snes_converged_reason": None,
-    "snes_monitor": None,
+    #"snes_monitor": None,
     "ksp_type": "cg",
     "ksp_converged_reason": None,
     "pc_type": "jacobi",
-}
-
-npcnewtoncg = {
-    "mat_type": "matfree",
-    "snes_type": "nrichardson",
-    #"snes_view": None,
-    "snes_converged_reason": None,
-    "snes_monitor": None,
-    "snes_npc_side": "left",
-    "npc_snes_type": "newtonls",
-    "npc_snes_linesearch_type": "basic",
-    "npc_ksp_type": "cg",
-    "npc_ksp_converged_reason": None,
-    "npc_pc_type": "jacobi",
-    #"npc_pc_type": "python",
-    #"npc_pc_python_type": "firedrake.AssembledPC",
-    #"assembled_ksp_type": "preonly",
-    #"assembled_pc_type": "lu",
 }
 
 ngssweep = {
@@ -41,7 +24,7 @@ ngssweep = {
     "snes_type": "nrichardson",
     #"snes_view": None,
     "snes_converged_reason": None,
-    "snes_monitor": None,
+    #"snes_monitor": None,
     "snes_npc_side": "left",
     "npc_snes_type": "python",
     "npc_snes_python_type": "firedrake.PatchSNES",
@@ -62,7 +45,7 @@ ngssweep = {
     #"npc_patch_snes_patch_sub_snes_rtol": 1.0e-11,
 }
 
-mesh = UnitSquareMesh(4,4)
+mesh = UnitSquareMesh(8,8)
 
 V = FunctionSpace(mesh, "CG", 1)
[…]

### 2022-08-09 py/poisson.py

@@ -1,4 +1,7 @@
-"""Solve poisson equation  - triangle u = f  by FAS multigrid.
+"""Solve poisson equation  - triangle u = f  by FAS multigrid, using explicit
+calls to prolong, restrict, and inject.
+
+WARNING: LOOKS LIKE THIS DOES NOT WORK.  COMPARE RESULT TO poisson0.py
 
 Poisson equation problem (weak form) is
     F(u)[v] = ell[v]

### 2022-08-09 py/obstacle.py
this was posted to slack for help and comments

@@ -1,0 +1,93 @@
+"""ATTEMPT to solve the following obstacle problem.  The plan is to solve by
+single-level projected NGS using PatchSNES.  (I know this is inefficient, but
+it is to be the smoother in an FAS scheme.)  I am currently blocked by
+not knowing how to implement *projected* NGS.
+
+ISSUES:
+  * I seem to only know how to get nonlinear *Jacobi* sweeps; setting 
+      "npc_patch_snes_patch_local_type": "multiplicative"
+    to get nonlinear GS sweeps generates a seg fault.
+  * I don't know how to intervene in the patch-wise Newton step to change it
+    to project back up to the obstacle if it over shoots.  So for now the
+    solver ignores the obstacle psi, and therefore just solves a Dirichlet
+    problem (PDE).
+
+Solve laplacian obstacle problem in 2D square Omega = (-2,2) x (-2,2):
+  u >= psi           on Omega
+  - triangle u = 0   where u > psi
+  u = g              on boundary of Omega
+Formulas for psi (a hemisphere) and g are given in the code.
+
+The exact solution of this problem is given in Chapter 12 of Bueler (2021),
+"PETSc for Partial Differential Equations".  See c/ch12/obstacle.c
+at https://github.com/bueler/p4pdes.  Dirichlet conditions are taken from
+the exact solution.
+"""
+
+import numpy as np
+from firedrake import *
+
+ngssweep = {
+    "mat_type": "matfree",
+    "snes_type": "nrichardson",
+    "snes_converged_reason": None,
+    "snes_npc_side": "left",
+    "npc_snes_type": "python",
+    "npc_snes_python_type": "firedrake.PatchSNES",
+    "npc_snes_max_it": 1,
+    "npc_snes_convergence_test": "skip",
+    "npc_snes_linesearch_type": "basic",
+    "npc_snes_linesearch_damping": 1.0,
+    "npc_patch_snes_patch_construct_type": "star",
+    "npc_patch_snes_patch_partition_of_unity": True,
+    "npc_patch_snes_patch_sub_mat_type": "seqaij",
+    "npc_patch_snes_patch_local_type": "additive",   # this is nonlinear Jacobi?
[…]

### 2022-08-09 py/obstacle.py
patch from patrick farrell (on petsc/src/ksp/pc/impls/patch/pcpatch.c) allows multiplicative (NGS)

@@ -1,16 +1,10 @@
 """ATTEMPT to solve the following obstacle problem.  The plan is to solve by
 single-level projected NGS using PatchSNES.  (I know this is inefficient, but
 it is to be the smoother in an FAS scheme.)  I am currently blocked by
-not knowing how to implement *projected* NGS.
-
-ISSUES:
-  * I seem to only know how to get nonlinear *Jacobi* sweeps; setting 
-      "npc_patch_snes_patch_local_type": "multiplicative"
-    to get nonlinear GS sweeps generates a seg fault.
-  * I don't know how to intervene in the patch-wise Newton step to change it
-    to project back up to the obstacle if it over shoots.  So for now the
-    solver ignores the obstacle psi, and therefore just solves a Dirichlet
-    problem (PDE).
+not knowing how to implement *projected* NGS.  That is, I don't know how to
+intervene in the patch-wise Newton step to change it to project back up to
+the obstacle if it over shoots.  So for now the solver ignores the obstacle
+psi, and therefore just solves a Dirichlet problem (PDE).
 
 Solve laplacian obstacle problem in 2D square Omega = (-2,2) x (-2,2):
   u >= psi           on Omega
@@ -41,8 +35,8 @@ ngssweep = {
     "npc_patch_snes_patch_construct_type": "star",
     "npc_patch_snes_patch_partition_of_unity": True,
     "npc_patch_snes_patch_sub_mat_type": "seqaij",
-    "npc_patch_snes_patch_local_type": "additive",   # this is nonlinear Jacobi?
-    #"npc_patch_snes_patch_local_type": "multiplicative",  # this would be NGS? seg faults!
+    #"npc_patch_snes_patch_local_type": "additive",   # nonlinear Jacobi
+    "npc_patch_snes_patch_local_type": "multiplicative",
     "npc_patch_snes_patch_symmetrise_sweep": False,
     "npc_patch_sub_snes_type": "newtonls",
     "npc_patch_sub_snes_linesearch_type": "basic",

### 2022-08-09 py/obstacle.py
minor thinking in comments

@@ -30,16 +30,18 @@ ngssweep = {
     "npc_snes_python_type": "firedrake.PatchSNES",
     "npc_snes_max_it": 1,
     "npc_snes_convergence_test": "skip",
+    #"npc_snes_converged_reason": None,
     "npc_snes_linesearch_type": "basic",
     "npc_snes_linesearch_damping": 1.0,
     "npc_patch_snes_patch_construct_type": "star",
-    "npc_patch_snes_patch_partition_of_unity": True,
+    "npc_patch_snes_patch_partition_of_unity": True,  # preferred value?
     "npc_patch_snes_patch_sub_mat_type": "seqaij",
     #"npc_patch_snes_patch_local_type": "additive",   # nonlinear Jacobi
     "npc_patch_snes_patch_local_type": "multiplicative",
     "npc_patch_snes_patch_symmetrise_sweep": False,
     "npc_patch_sub_snes_type": "newtonls",
     "npc_patch_sub_snes_linesearch_type": "basic",
+    #"npc_patch_sub_snes_converged_reason": None,  # no effect
     "npc_patch_sub_ksp_type": "preonly",
     "npc_patch_sub_pc_type": "lu",
 }

### 2022-08-09 paper/mcd2.tex
edits through page 10

@@
 \begin{document}
-\title[On multilevel constraint decomposition methods]{On multilevel constraint decomposition methods \\ for nonlinear variational inequalities}
+\title[Multilevel constraint decomposition methods]{Multilevel constraint decomposition methods \\ for nonlinear variational inequalities}
 
@@
 \begin{enumerate}
-\item We do not assume that the continuum VI problem arises from optimization of a scalar objective---examples are explored in Sections \ref{sec:vi} and \ref{sec:results}---but we prove convergence in $H^1$ norm at the same rate as the original method converges in energy (Section \ref{sec:convergence}).
+\item We do not assume that the continuum VI problem arises from optimization of a scalar objective---examples are explored in Sections \ref{sec:vi} and \ref{sec:results}---but we prove convergence in norm at the same rate as the original method converges in energy (Section \ref{sec:convergence}).
 % THIS IS THE HOPE
 \end{enumerate}
-Furthermore the multilevel CD algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} are extended in two practical ways:
+Furthermore the multilevel CD algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} are extended in three practical ways:
 \begin{enumerate}
@@
 Results from the resulting \emph{nonlinear multilevel constraint decomposition} (NMCD) algorithm, which avoids global (Newton) linearization (compare \cite{GraeserKornhuber2009}), are shown in Section \ref{sec:results} for several nonlinear problems.
+\item Our implementation permits ``box'' constraints, namely upper and lower obstacles.
 \item In Sections \ref{sec:fe} and \ref{sec:multilevel} we explain why, given the manner in which constraint sets are decomposed in the multilevel context \cite{GraeserKornhuber2009}, multilevel ``up-smoothing'' is more efficient than ``down-smoothing''.
@@
 
-The iterates from a CD method are always admissible, and thus the operator need only be defined for admissible states; Section \ref{sec:results} includes a nontrivial example.
+The iterates from a CD method are always admissible, and thus the operator need only be defined for admissible states.
 Admissible-iterate methods should permit direct solutions of certain VI problems, such as fluid-layer dynamics problems \cite{Bueler2021conservation,JouvetBueler2012}, for which non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
@@
 \end{equation}
-holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$.
-Equation \eqref{eq:subspacedecomp} is called a \emph{subspace decomposition} \cite{Xu1992}.
+holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$.  Equation \eqref{eq:subspacedecomp} is called a \emph{subspace decomposition} \cite{Xu1992}.
+
 Suppose further that $\cK_i \subset \cV_i$ are nonempty, closed, and convex subsets such that
@@
 \caption{Suppose $\mathcal{V}$ is the space of real functions on a two-point set $\Omega=\{x_1,x_2\}$.
-A constraint decomposition (CD) for a one-sided obstacle problem with $\mathcal{K}=\{v\ge \psi\}$ might look like this.}
+A constraint decomposition (CD) for a unilateral obstacle problem with $\mathcal{K}=\{v\ge \psi\}$ might look like this.}
 \label{fig:cartoon}
@@
 
-For each $\cK_i$ we will also assume that there are bounded, generally-nonlinear \emph{restriction} operators $R_i : \cK \to \cK_i$ such that if $v \in \cK$ then
+Finally, for each $\cK_i$ we assume there are bounded, generally-nonlinear \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.
+Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:multilevel}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
 \begin{equation}
-v = \sum_{i=0}^{m-1} R_i v;  \label{eq:constraintrestrictionsum}
+v = \sum_{i=0}^{m-1} \Pi_i v;  \label{eq:constraintrestrictionsum}
 \end{equation}
@@
 Clearly \eqref{eq:constraintrestrictionsum} implies sense \emph{(i)} for \eqref{eq:constraintdecomp}.
-A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,R_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} \cite{Tai2003}.
+A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} \cite{Tai2003}.
 
@@
 Suppose further that $\{\phi_i\}_{i=0}^{m-1}$ is a smooth partition of unity on $\Omega$, satisfying $0 \le \phi_i\le 1$ and $\sum_i \phi_i = 1$, and let $\Omega_i$ be the open support of $\phi_i$.
-Let $\cV_i = \{w \in \cV:w|_{\Omega \setminus \Omega_i} =0 \}$, $\cK_i = \{v \in \cV_i: v \ge \phi_i \psi\}$, and $R_i(v) = \phi_i v$.
+Let $\cV_i = \{w \in \cV:w|_{\Omega \setminus \Omega_i} =0 \}$, $\cK_i = \{v \in \cV_i: v \ge \phi_i \psi\}$, and $\Pi_i(v) = \phi_i v$.
 Then \eqref{eq:subspacedecomp}, \eqref{eq:constraintdecomp}, and \eqref{eq:constraintrestrictionsum} all hold.
@@
 Suppose $P_i:\cV \to \cV_i$ are the corresponding orthogonal projections, satisfying $I = \sum_i P_i$.
-Let $\cK_i = \{v \ge P_i \psi\} \subset \cV_i$ and $R_i = P_i$.
+Let $\cK_i = \{v \ge P_i \psi\} \subset \cV_i$ and $\Pi_i = P_i$.
 Then \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} also hold.
@@
 
-\section{Iterations based on constraint decomposition} \label{sec:cditers}
+\section{Basic constraint decomposition iterations} \label{sec:cditers}
 
@@
         \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$, \\+
-            $\boxed{\ip{f(u - R_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}}$ \\--
+            $\boxed{\ip{f(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}}$ \\--
     $\hat w = \sum_i \hat w_i\in\cK$ \\
@@
 In Algorithm \ref{alg:basiccd-add}, the additive or parallel CD method, the \textbf{for all} can be computed in any order.
-Note that $u-R_iu+\hat w_i$ removes the part of $u$ which lies in $\mathcal{K}_i$ before adding-back the improved subset solution $\hat w_i$, also from $\mathcal{K}_i$.
+Note that $u-\Pi_iu+\hat w_i$ removes the part of $u$ which lies in $\mathcal{K}_i$ before adding-back the improved subset solution $\hat w_i$, also from $\mathcal{K}_i$.
 By contrast, in the following multiplicative (successive) Algorithm \ref{alg:basiccd-mult} each subset solution updates the global iterate before proceeding to the next subset.
-The sets $\mathcal{K}_i$ are addressed in a fixed order, and the sum $\sum_{j>i} R_j u$ should be read as ``keep those parts of the input $u$ which we have not yet improved.''
+The sets $\mathcal{K}_i$ are addressed in a fixed order, and the sum $\sum_{j>i} \Pi_j u$ should be read as ``keep those parts of the input $u$ which we have not yet improved.''
 
@@
         \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$, \\+
-            $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}}$ \\-
-            $w_i = (1-\alpha) R_i u + \alpha \hat w_i\in\cK_i$ \\-
+            $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}}$ \\-
+            $w_i = (1-\alpha) \Pi_i u + \alpha \hat w_i\in\cK_i$ \\-
     $\hat w = \sum_i \hat w_i\in\cK$ \\
@@
 \begin{align*}
-[u - R_i u + v_i] - [u - R_i u + \hat w_i] &= v_i - \hat w_i, \label{eq:admissibledifference} \\
-\left[\sum_{j<i} w_j + v_i + \sum_{j>i} R_j u\right] - \left[\sum_{j<i} w_j + \hat w_i + \sum_{j>i} R_j u\right] &= v_i - \hat w_i.  \notag
+[u - \Pi_i u + v_i] - [u - \Pi_i u + \hat w_i] &= v_i - \hat w_i, \label{eq:admissibledifference} \\
+\left[\sum_{j<i} w_j + v_i + \sum_{j>i} \Pi_j u\right] - \left[\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\right] &= v_i - \hat w_i.  \notag
 \end{align*}
@@
 \begin{equation}
-e_i = \hat w_i - R_i u \in \cV_i, \label{eq:ithupdate}
+e_i = \hat w_i - \Pi_i u \in \cV_i, \label{eq:ithupdate}
 \end{equation}
@@
 As the reader may check, $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.
-Note $\hat w = u^* + \sum_i \hat w_i - R_i u^*$ also holds.
+Note $\hat w = u^* + \sum_i \hat w_i - \Pi_i u^*$ also holds.
 We will use these identities in Section \ref{sec:convergence}.
@@
 \begin{equation}
-\ip{f(e_i)}{v_i-\hat w_i} \ge \ip{\tilde\ell}{v_i-\hat w_i} \label{eq:linearlocalvi}
+\ip{f(e_i)}{v_i-\hat w_i} \ge \ip{\tilde\ell}{v_i-\hat w_i}, \label{eq:linearlocalvi}
 \end{equation}
-where $\tilde\ell = \ell - f(u)$.
-Note $e_i = \hat w_i - R_i u \in \cV_i$ is not generally admissible ($e_i \notin \cK$), but VIs \eqref{eq:linearlocalvi} make sense because $f$ is defined over $\cV$.
+for all $v_i \in \mathcal{K_i}$, where $\tilde\ell = \ell - f(u)$.
+Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible ($e_i \notin \cK$), but VIs \eqref{eq:linearlocalvi} make sense because $f$ is defined over $\cV$.
 Each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $f(u)$, here included into a source term.
@@
 Next we make two assumptions which are essentially the same as (7) and (8) in \cite{Tai2003}.
-These inequalities might be described as ``totally Lipschitz'' requirements, over the CD defined by \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, for the maps $R_i$ and $f$, respectively.
+These inequalities might be described as ``totally Lipschitz'' requirements, over the CD defined by \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, for the maps $\Pi_i$ and $f$, respectively.
 
@@
 \begin{equation}
-\left(\sum_{i=0}^{m-1} \|R_i u - R_i v\|^2\right)^{1/2} \le C_1 \|u-v\| \label{as:lipschitzrestrictions}
+\left(\sum_{i=0}^{m-1} \|\Pi_i u - \Pi_i v\|^2\right)^{1/2} \le C_1 \|u-v\| \label{as:lipschitzrestrictions}
 \end{equation}
@@
 \begin{lemma} \label{lem:core}  Suppose $f$ is $2$-coercive with constant $\kappa>0$, and that $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.
-Assume $\mathcal{V}_i$, $\mathcal{K}_i$, $R_i$ form a CD satisfying \eqref{as:lipschitzrestrictions}, and that \eqref{as:lipschitzresidual} is satisfied.
+Assume $\mathcal{V}_i$, $\mathcal{K}_i$, $\Pi_i$ form a CD satisfying \eqref{as:lipschitzrestrictions}, and that \eqref{as:lipschitzresidual} is satisfied.
 If $\hat w$ is computed from $u \in \mathcal{K}$ by Algorithm \ref{alg:basiccd-add} then
@@
 \end{equation}
-where again $e_i=\hat w_i - R_i u$ (equation \eqref{eq:ithupdate}).
+where again $e_i=\hat w_i - \Pi_i u$ (equation \eqref{eq:ithupdate}).
 \end{lemma}
 
-\begin{proof}  Recall that $\hat w = u^* + \sum_i \hat w_i - R_i u^*$.  Expand $E(\hat w,u^*)$ and substitute $v_i = R_i u^*$ into the boxed VI in Algorithm \ref{alg:basiccd-add} to find
+\begin{proof}  Recall that $\hat w = u^* + \sum_i \hat w_i - \Pi_i u^*$.  Expand $E(\hat w,u^*)$ and substitute $v_i = \Pi_i u^*$ into the boxed VI in Algorithm \ref{alg:basiccd-add} to find
 \begin{align}
-E(\hat w,u^*) &= \sum_{i=0}^{m-1} \ip{f(\hat w)}{\hat w_i - R_i u^*} - \ip{\ell}{\hat w_i - R_i u^*} \notag \\
-    &\le \sum_{i=0}^{m-1} \ip{f(\hat w)}{\hat w_i - R_i u^*} + \ip{f(u + e_i)}{R_i u^* - \hat w_i} \notag \\
-    &= \sum_{i=0}^{m-1} \ip{f(\hat w) - f(u + e_i)}{\hat w_i - R_i u^*}. \label{eq:startcore}
+E(\hat w,u^*) &= \sum_{i=0}^{m-1} \ip{f(\hat w)}{\hat w_i - \Pi_i u^*} - \ip{\ell}{\hat w_i - \Pi_i u^*} \notag \\
+    &\le \sum_{i=0}^{m-1} \ip{f(\hat w)}{\hat w_i - \Pi_i u^*} + \ip{f(u + e_i)}{\Pi_i u^* - \hat w_i} \notag \\
+    &= \sum_{i=0}^{m-1} \ip{f(\hat w) - f(u + e_i)}{\hat w_i - \Pi_i u^*}. \label{eq:startcore}
 \end{align}
@@
 \begin{equation}
-E(\hat w,u^*) = \sum_{i=0}^{m-1} \ip{f(\phi_{i,m-1}) - f(\phi_{i,0})}{\hat w_i - R_i u^*}. \label{eq:startcoreclean}
+E(\hat w,u^*) = \sum_{i=0}^{m-1} \ip{f(\phi_{i,m-1}) - f(\phi_{i,0})}{\hat w_i - \Pi_i u^*}. \label{eq:startcoreclean}
 \end{equation}
@@
 \begin{align}
-E(\hat w,u^*) &\le \sum_{i=0}^{m-1} \sum_{j=1}^{m-1} \ip{f(\phi_{i,j}) - f(\phi_{i,j-1})}{\hat w_i - R_i u^*} \label{eq:nextcore} \\
-  &\le \sum_{i=0}^{m-1} \sum_{j=1}^{m-1} \left|\ip{f(\phi_{i,j}) - f(\phi_{i,j-1})}{\hat w_i - R_i u^*}\right|. \notag
+E(\hat w,u^*) &\le \sum_{i=0}^{m-1} \sum_{j=1}^{m-1} \ip{f(\phi_{i,j}) - f(\phi_{i,j-1})}{\hat w_i - \Pi_i u^*} \label{eq:nextcore} \\
+  &\le \sum_{i=0}^{m-1} \sum_{j=1}^{m-1} \left|\ip{f(\phi_{i,j}) - f(\phi_{i,j-1})}{\hat w_i - \Pi_i u^*}\right|. \notag
 \end{align}
@@
 \begin{align}
-E(\hat w,u^*) &\le C_2 Z \left(\sum_{i=0}^{m-1} \|\hat w_i - R_i u^*\|^2\right)^{1/2} = C_2 Z \left(\sum_{i=0}^{m-1} \|\hat w_i - R_i u + R_i u - R_i u^*\|^2\right)^{1/2} \label{eq:nextnextcore} \\
-  &\le C_2 Z \left(Z + \left(\sum_{i=0}^{m-1} \|R_i u - R_i u^*\|^2\right)^{1/2}\right) \le C_2 Z \left(Z + C_1 \|u-u^*\|\right). \notag
+E(\hat w,u^*) &\le C_2 Z \left(\sum_{i=0}^{m-1} \|\hat w_i - \Pi_i u^*\|^2\right)^{1/2} = C_2 Z \left(\sum_{i=0}^{m-1} \|\hat w_i - \Pi_i u + \Pi_i u - \Pi_i u^*\|^2\right)^{1/2} \label{eq:nextnextcore} \\
+  &\le C_2 Z \left(Z + \left(\sum_{i=0}^{m-1} \|\Pi_i u - \Pi_i u^*\|^2\right)^{1/2}\right) \le C_2 Z \left(Z + C_1 \|u-u^*\|\right). \notag
 \end{align}

### 2022-08-09 paper/genfigs/cartoon.py

@@ -41,7 +41,7 @@ plt.text(psi1 - 0.07, - 0.09, r'$\psi(x_1)$', fontsize=fsize)
 plt.text(0.7, - 0.09, r'$\mathcal{K}_1$', fontsize=fsize)
 plt.gca().annotate('', xy=(0.7, psi2 - 0.3), xytext=(0.7, psi2 - 0.05),
                    arrowprops=dict(arrowstyle='->, head_length=0.8, head_width=0.5'))
-plt.text(0.72, 0.15, r'$R_1$', fontsize=fsize)
+plt.text(0.72, 0.15, r'$\Pi_1$', fontsize=fsize)
 
 plt.text(-0.02, 0.89, r'$\mathcal{V}_2$', fontsize=fsize)
 plt.plot(0.0, psi2, 'k.', ms = 12.0)
@@ -50,7 +50,7 @@ plt.text(- 0.22, psi2 - 0.02, r'$\psi(x_2)$', fontsize=fsize)
 plt.text(- 0.12, 0.6, r'$\mathcal{K}_2$', fontsize=fsize)
 plt.gca().annotate('', xy=(psi1 - 0.15, 0.6), xytext=(psi1 - 0.05, 0.6),
                    arrowprops=dict(arrowstyle='->, head_length=0.8, head_width=0.5'))
-plt.text(0.07, 0.65, r'$R_2$', fontsize=fsize)
+plt.text(0.07, 0.65, r'$\Pi_2$', fontsize=fsize)
 
 plt.axis('tight')
 plt.axis('off')

### 2022-08-09 paper/mcd2.tex
edits start of section 6

@@
 \item Our implementation permits ``box'' constraints, namely upper and lower obstacles.
-\item In Sections \ref{sec:fe} and \ref{sec:multilevel} we explain why, given the manner in which constraint sets are decomposed in the multilevel context \cite{GraeserKornhuber2009}, multilevel ``up-smoothing'' is more efficient than ``down-smoothing''.
+\item In Sections \ref{sec:multilevel} and \ref{sec:vcycle} we explain why, given the manner in which constraint sets are decomposed in the multilevel context \cite{GraeserKornhuber2009}, multilevel ``up-smoothing'' is more efficient than ``down-smoothing''.
 This observation seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}.
@@
 \end{equation}
-Because $f$ is (generally) nonlinear, the source term $\ell$ is not strictly needed, and indeed by redefining $f$ we may take $\ell=0$, but its presence clarifies algorithms in Section \ref{sec:multilevel}.
+Because $f$ is (generally) nonlinear, the source term $\ell$ is not strictly needed, and indeed by redefining $f$ we may take $\ell=0$, but its presence clarifies the algorithm in Section \ref{sec:vcycle}.
 
@@
 
-In Section \ref{sec:multilevel} we will introduce discretizations, and practical algorithms for finite-dimensional VI problems, but observe here that this CD concept applies at the level of the continuum problem.
+In Sections \ref{sec:multilevel} and \ref{sec:vcycle} we will introduce discretizations and a practical algorithms for finite-dimensional VI problems, but observe here that this CD concept applies at the level of the continuum problem.
 The following two obstacle problem examples illustrate this.
@@
 
-Our second example is a disjoint frequency decomposition.  A multilevel CD, e.g.~the one proposed in Sections \ref{sec:fe} and \ref{sec:multilevel}, approximates such a frequency decomposition.
+Our second example is a disjoint frequency decomposition.  A multilevel CD, e.g.~the one proposed in Section \ref{sec:multilevel}, approximates such a frequency decomposition.
 
@@
 When $f$ has all the nice properties listed in Example \ref{ex:fnice} then a solver can take advantage of implementation efficiencies which are unavailable in general.
-The FE approximation of VI \eqref{eq:vi}---see Section \ref{sec:fe}---can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.
+The FE approximation of VI \eqref{eq:vi}---see Sections \ref{sec:multilevel} and \ref{sec:vcycle}---can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.
 The actual solution process for VI \eqref{eq:linearlocalvi} is not really the concern here; we are observing that the storage needed for problem \eqref{eq:linearlocalvi}, and the cost of residual evaluation, are smaller than in the general case.
-For more difficult problems certain additional ideas, illustrated in Section \ref{sec:multilevel}, are needed for practical implementation.
+For more difficult problems certain additional ideas, illustrated in Section \ref{sec:vcycle}, are needed for practical implementation.
 Specifically, we will extend a multilevel CD iteration to a method for nonlinear functions $f:\cK\to\cV'$, i.e.~defined only on the constraint set $\cK$, by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
@@
 \item The constants $C_1,C_2$ generally depend on $m$, the number of sets in the CD.
-In a multilevel CD method (Section \ref{sec:multilevel}) this number is logarithmic in the number of degrees of freedom (nodes) in the fine-level mesh.
+In a multilevel CD method (Sections \ref{sec:multilevel} and \ref{sec:vcycle}) this number is logarithmic in the number of degrees of freedom (nodes) in the fine-level mesh.
 \item Regarding the convergence proof below, if the iterates from Algorithm \ref{alg:basiccd-add} are bounded in $\mathcal{K}$ then constants $C_1,C_2$ are permitted to depend continuously on $u,v,w_{ij},v_i,y_i$ \cite{Tai2003}.
@@
 
-\section{Multilevel finite elements} \label{sec:fe}
+\section{Multilevel finite element constraint decompositions} \label{sec:multilevel}
 
-We will approximately solve VI \eqref{eq:vi} for certain nonlinear obstacle problems, over two-dimensional computational domains $\Omega \subset \RR^2$ and Banach space $\mathcal{V}=W_0^{1,p}(\Omega)$, using piecewise-linear finite element (FE) spaces over triangular meshes.
+We will approximately solve VI \eqref{eq:vi} for certain nonlinear obstacle problems, over two-dimensional computational domains $\Omega \subset \RR^2$ and the Banach space $\mathcal{V}=W_0^{1,p}(\Omega)$, using piecewise-linear finite element (FE) spaces over triangular meshes.
 The targeted mesh will be the finest level in a hierarchy of meshes constructed by refinement of a coarsest mesh.
@@
 
-Let $\tilde{\mathcal{V}}^j$ be the continuous, piecewise-polynomial FE space $P_1$,\footnote{The FE method herein apparently extends to $Q_1$ spaces over quadrilateral meshes \cite{Elmanetal2014}.
-However, for $k>1$ the $P_k$ space seems not to be suitable because the monotone restriction operator, \eqref{eq:fe:monotonerestriction} below, depends on element-wise monotonicity.
-That is, for $P_1$ or $Q_1$ if the nodal values of an FE function (non-strictly) increase then the FE function also increases.
-As shown by Figure 1.7 of \cite{Elmanetal2014}, for example, such monotonicity is not true for $P_2$.} over $\mathcal{T}^j$, with arbitrary boundary values.
-Let $\mathcal{V}^j \subset \tilde{\mathcal{V}}^j$ be the subspace with zero boundary values.
-For any $p\ge 1$, $\tilde{\mathcal{V}}^j$ is a subspace of $W^{1,p}(\Omega)$ and $\mathcal{V}^j$ is a subspace of $\mathcal{V}=W_0^{1,p}(\Omega)$ \cite{Elmanetal2014}.
-Our choices imply the following nesting:
+Let $\mathcal{V}^j$ be the continuous, piecewise-linear FE space $P_1$ over $\mathcal{T}^j$ with zero boundary values.
+Note $\mathcal{V}^j$ is a subspace of $\mathcal{V}=W_0^{1,p}(\Omega)$ \cite{Elmanetal2014}, and observe the following nesting:
 \begin{equation}
@@
 
-Our only need for the arbitrary-boundary-value spaces $\tilde{\mathcal{V}}^j$ is for the constraint.
-Suppose $\gamma^J \in \tilde{\mathcal{V}}^j$ denotes the \emph{obstacle}, a function satisfying $\gamma^J \le 0$ at boundary nodes of $\mathcal{T}^J$, and define
+Let $\tilde{\mathcal{V}}^j$ denote the set of functions on the nodes of $\mathcal{T}^j$ which take their values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
+Our purpose for such extended-value functions is to define ``box constraints'' \cite{BensonMunson2006} as follows.
+At the finest-level, suppose $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ satisfy
 \begin{equation}
-\mathcal{K}^J = \left\{v \in \mathcal{V}^J\,:\,v \ge \gamma^J\right\}, \label{eq:fineconstraintset}
+\underline{\gamma}^J \le \overline{\gamma}^J, \qquad \underline{\gamma}^J < +\infty, \qquad \overline{\gamma}^J > -\infty, \qquad \text{and} \qquad \underline{\gamma}^J|_{\partial\Omega} \le 0 \le \overline{\gamma}^J|_{\partial\Omega}.
+\label{eq:fe:boxconstraintrequirements}
 \end{equation}
-the finest-level \emph{constraint set} of \emph{admissible} FE functions.  Note that admissible functions from $\mathcal{K}^J$ have zero boundary values.
+We call $\underline{\gamma}^J$ the \emph{lower obstacle} and $\overline{\gamma}^J$ the \emph{upper obstacle}.
+Where $\underline{\gamma}^J=-\infty$ there is no lower obstacle and where $\overline{\gamma}^J=+\infty$ there is no upper obstacle.
+Now define
+\begin{equation}
+\mathcal{K}^J = \left\{v \in \mathcal{V}^J\,:\,\underline{\gamma}^J \le v \le \overline{\gamma}^J\right\}, \label{eq:fineconstraintset}
+\end{equation}
+the finest-level \emph{constraint set} of \emph{admissible} FE functions.
+Note that admissible functions from $\mathcal{K}^J$ have zero boundary values and that elements of $\mathcal{K}^J$ never, themselves, take on values $\pm\infty$.
+
+The FE method herein apparently extends to $Q_1$ spaces over quadrilateral meshes \cite{Elmanetal2014}.
+However, for $k>1$ the $P_k$ space seems not to be suitable because the monotone restriction operator, \eqref{eq:fe:monotonerestriction} below, depends on element-wise monotonicity.
+That is, for $P_1$ or $Q_1$ if the nodal values of an FE function (non-strictly) increase then the FE function also increases.
+As shown by Figure 1.7 of \cite{Elmanetal2014}, for example, such monotonicity is not true for $P_2$.
 
@@
 
-\section{Nonlinear multilevel constraint decomposition} \label{sec:multilevel}
+\section{The nonlinear multilevel constraint decomposition V-cycle} \label{sec:vcycle}
 

### 2022-08-10 paper/mcd2.tex
fiddle with section headings

@@
 
-The iterates from a CD method are always admissible, and thus the operator need only be defined for admissible states.
+The iterates from a CD method are always admissible, and thus the operator can be defined only for admissible states.
 Admissible-iterate methods should permit direct solutions of certain VI problems, such as fluid-layer dynamics problems \cite{Bueler2021conservation,JouvetBueler2012}, for which non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
@@
 
-\section{Constraint decomposition} \label{sec:cd}
+\section{Constraint decomposition (CD)} \label{sec:cd}
 
@@
 
-\section{Basic constraint decomposition iterations} \label{sec:cditers}
+\section{Basic CD iterations} \label{sec:cditers}
 
@@
 \end{pseudo*}
-\caption{Additive constraint decomposition for VI problem \eqref{eq:vi}.}
+\caption{One additive CD iteration for VI problem \eqref{eq:vi}.}
 \label{alg:basiccd-add}
@@
 \end{pseudo*}
-\caption{Multiplicative constraint decomposition for VI problem \eqref{eq:vi}.}
+\caption{One multiplicative CD iteration for VI problem \eqref{eq:vi}.}
 \label{alg:basiccd-mult}
@@
 
-\section{Convergence of a constraint decomposition iteration} \label{sec:convergence}
+\section{Convergence of a basic CD iteration} \label{sec:convergence}
 
@@
 
-\section{Multilevel finite element constraint decompositions} \label{sec:multilevel}
+\section{Multilevel finite element CDs} \label{sec:multilevel}
 
@@
 
-\section{The nonlinear multilevel constraint decomposition V-cycle} \label{sec:vcycle}
+\section{The nonlinear multilevel CD V-cycle} \label{sec:vcycle}
 

### 2022-08-10 paper/mcd2.tex
edits on multilevel FE CD constructs

@@
 \newcommand{\bx}{\mathbf{x}}
+\newcommand{\by}{\mathbf{y}}
+\newcommand{\bz}{\mathbf{z}}
+
 \newcommand{\bF}{\mathbf{F}}
@@
 \newcommand{\bX}{\mathbf{X}}
+
 \newcommand{\bxi}{\bm{\xi}}
@@
 
-\newcommand{\mR}{R^{\bm{\oplus}}}
+\newcommand{\maxR}{R^{\bm{\oplus}}}
+\newcommand{\minR}{R^{\bm{\ominus}}}
 \newcommand{\iR}{R^{\bullet}}
@@
 
-We will approximately solve VI \eqref{eq:vi} for certain nonlinear obstacle problems, over two-dimensional computational domains $\Omega \subset \RR^2$ and the Banach space $\mathcal{V}=W_0^{1,p}(\Omega)$, using piecewise-linear finite element (FE) spaces over triangular meshes.
+We will approximately solve VI \eqref{eq:vi} for certain nonlinear obstacle problems, over two-dimensional computational domains $\Omega \subset \RR^2$, and the Sobolev space $\mathcal{V}=W_0^{1,p}(\Omega)$, using piecewise-linear finite element (FE) spaces over triangular meshes.
 The targeted mesh will be the finest level in a hierarchy of meshes constructed by refinement of a coarsest mesh.
+In this Section we construct the corresponding multilevel CD.
 
 The mesh hierarchy and FE spaces are constructed in a standard manner \cite{Elmanetal2014}.
-Suppose $\Omega$ is polygonal and triangulated by $\mathcal{T}^0$, the \emph{coarsest mesh}, a finite set of non-overlapping triangles with union $\overline{\Omega}$ and characteristic size (cell diameter) $h_0>0$.
-Starting with $j=1$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ by edge bisection, thus each triangle becomes four similar triangles in $\mathcal{T}^j$ with characteristic size $h_j = 2^{-j} h_0$.
-Refinement proceeds to level $J\ge 0$, called the \emph{depth} of the hierarchy, giving $\mathcal{T}^J$ as the \emph{finest mesh}.
-(In the single-level $J=0$ case all multilevel efficiency will be lost.)  Let $m_j$ be the number of interior nodes in $\mathcal{T}^j$, the \emph{degrees of freedom} at the $j$th level, and denote $h=h_J=2^{-J} h_0$ and $m=m_J$ for the finest level.
-
-Let $\mathcal{V}^j$ be the continuous, piecewise-linear FE space $P_1$ over $\mathcal{T}^j$ with zero boundary values.
-Note $\mathcal{V}^j$ is a subspace of $\mathcal{V}=W_0^{1,p}(\Omega)$ \cite{Elmanetal2014}, and observe the following nesting:
+Suppose $\Omega$ is polygonal and triangulated by $\mathcal{T}^0$, the \emph{coarsest mesh}, a finite set of non-overlapping triangles with union $\overline{\Omega}$ and characteristic size (e.g.~maximum edge length) $h_0>0$.
+Starting with $j=1$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ by edge bisection, thus each $\mathcal{T}^{j-1}$ triangle becomes four similar triangles in $\mathcal{T}^j$ with characteristic size $h_j = 2^{-j} h_0$.
+Refinement proceeds to level $J\ge 0$, the \emph{depth} of the hierarchy, giving $\mathcal{T}^J$ as the \emph{finest mesh}.
+Let $m_j$ be the number of interior nodes in $\mathcal{T}^j$, the \emph{degrees of freedom} at the $j$th level, and denote $h=h_J=2^{-J} h_0$ and $m=m_J$ for the finest level.
+Now let $\mathcal{V}^j$ be the continuous, piecewise-linear FE space $P_1$ \cite{Elmanetal2014} over $\mathcal{T}^j$ with zero boundary values, and observe the following nesting:
 \begin{equation}
@@
 
-Let $\tilde{\mathcal{V}}^j$ denote the set of functions on the nodes of $\mathcal{T}^j$ which take their values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
-Our purpose for such extended-value functions is to define ``box constraints'' \cite{BensonMunson2006} as follows.
-At the finest-level, suppose $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ satisfy
+In the above subspace construction we have chosen triangles and the $P_1$ FE space.
+Quadrilaterals and the $Q_1$ FE space will work just as well in what follows.
+However, use of higher-order Lagrange elements $P_k$ and $Q_k$ would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} specific to inequality constraints in FE schemes.
+Our FE construction of constraint sets requires a property we call \emph{nodal monotonicity}, namely that for all FE functions $w,z \in \mathcal{V}^j$,
 \begin{equation}
-\underline{\gamma}^J \le \overline{\gamma}^J, \qquad \underline{\gamma}^J < +\infty, \qquad \overline{\gamma}^J > -\infty, \qquad \text{and} \qquad \underline{\gamma}^J|_{\partial\Omega} \le 0 \le \overline{\gamma}^J|_{\partial\Omega}.
+\bw \ge \bz \quad \implies \quad w \ge z \label{eq:nodalmonotonicity}
+\end{equation}
+where $\bw,\bz$ denote the vectors of nodal values.
+For $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal value one at a single node, and otherwise zero nodal values, which take on negative values \cite[Figure 1.7]{Elmanetal2014}, whereas this does not occur for $k\le 1$.
+It is easy to show that \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of such sometimes-negative hat functions.
+
+Next we define ``box constraints'' \cite{BensonMunson2006}.
+Let $\tilde{\mathcal{V}}^j$ denote the set of functions on the $\mathcal{T}^j$ nodes with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
+At the finest-level, suppose that the \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ satisfy
+\begin{equation}
+\underline{\gamma}^J \le \overline{\gamma}^J, \qquad \underline{\gamma}^J < +\infty, \qquad -\infty < \overline{\gamma}^J, \qquad \text{and} \qquad \underline{\gamma}^J|_{\partial\Omega} \le 0 \le \overline{\gamma}^J|_{\partial\Omega}.
 \label{eq:fe:boxconstraintrequirements}
 \end{equation}
-We call $\underline{\gamma}^J$ the \emph{lower obstacle} and $\overline{\gamma}^J$ the \emph{upper obstacle}.
-Where $\underline{\gamma}^J=-\infty$ there is no lower obstacle and where $\overline{\gamma}^J=+\infty$ there is no upper obstacle.
-Now define
+Where $\underline{\gamma}^J=-\infty$ there will no lower constraint, and similarly where $\overline{\gamma}^J=+\infty$ no upper constraint.  Using nodal monotonicity, define
 \begin{equation}
@@
 \end{equation}
-the finest-level \emph{constraint set} of \emph{admissible} FE functions.
-Note that admissible functions from $\mathcal{K}^J$ have zero boundary values and that elements of $\mathcal{K}^J$ never, themselves, take on values $\pm\infty$.
+the finest-level \emph{constraint set} of \emph{admissible} FE functions, a closed and convex subset of $\mathcal{V}^J$.
+As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J = \pm \infty$.
 
-The FE method herein apparently extends to $Q_1$ spaces over quadrilateral meshes \cite{Elmanetal2014}.
-However, for $k>1$ the $P_k$ space seems not to be suitable because the monotone restriction operator, \eqref{eq:fe:monotonerestriction} below, depends on element-wise monotonicity.
-That is, for $P_1$ or $Q_1$ if the nodal values of an FE function (non-strictly) increase then the FE function also increases.
-As shown by Figure 1.7 of \cite{Elmanetal2014}, for example, such monotonicity is not true for $P_2$.
-
-One might assume that the nonlinear functional $f$ in \eqref{eq:vi} is defined over $\mathcal{K}^J$, i.e.~$f:\mathcal{K}^J \to \mathcal{V}'$.
-However, especially for nontrivial nonlinear functionals, the finest-level, discretized FE functional $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ is likely to only approximate $f$ because of quadrature or other implementation details.
-The functional $\ell$ in \eqref{eq:vi} is defined and continuous over $\mathcal{V}^J$, thus $\ell\in (\mathcal{V}^J)'$, and we assume for simplicity that it can be computed exactly, but we denote $\ell^J \in (\mathcal{V}^J)'$ to indicate that it must also be given an FE representation.
-
-With these understandings we seek the finest-level FE solution $u^J \in \mathcal{K}^J$ of this finite-dimensional, nonlinear VI
+Let $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ be a discretization of $f$ in \eqref{eq:vi}.
+Noting that the functional $\ell$ in \eqref{eq:vi} is already defined and continuous over $\mathcal{V}^J$, we denote $\ell^J \in (\mathcal{V}^J)'$ for its FE representation.
+We seek the finest-level FE solution $u^J \in \mathcal{K}^J$ of the finite-dimensional, nonlinear VI
 \begin{equation}
@@
 \end{equation}
-As for \eqref{eq:vi}, if $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, or $f^J$ is continuous, coercive and strictly monotone, then \eqref{eq:fe:vi} is well-posed \cite{KinderlehrerStampacchia1980}.
+As for \eqref{eq:vi}, if $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, then \eqref{eq:fe:vi} is well-posed.
 
 While \eqref{eq:fe:vi} approximates \eqref{eq:vi}, it may not technically be conforming for \eqref{eq:vi}.
-That is, even if quadrature and other implementation details are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require $\mathcal{K}^J \subset \mathcal{K}$.
-Only the fine-level FE obstacle $\gamma^J$ needs to be defined, while $\mathcal{K}$ remains abstract.
-In fact, if a continuum obstacle exists then admissibility with respect to its FE interpolant $\gamma^J$ might not imply continuum admissiblity.
-These caveats apply in geophysical situations where the obstacle is given by meshed elevation data, the continuum limit of which is undefined \cite{Bueler2016}.
+That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require $\mathcal{K}^J \subset \mathcal{K}$.
+The continuum constraint set $\mathcal{K}$ remains abstract in applications where only the fine-level FE obstacles $\underline{\gamma}^J, \overline{\gamma}^J$ are defined.
+In fact, even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants might not imply continuum admissiblity.
+These caveats specifically apply in geophysical situations where the obstacle is given by meshed elevation data, the continuum limit of which is undefined \cite{Bueler2016}.
 
-The multilevel algorithm in the next section, designed to solve \eqref{eq:fe:vi}, requires transfer operators between meshes.
-We use four such operators, three of which are standard, linear maps used for multigrid methods \cite{Trottenbergetal2001}.
+The multilevel algorithm in the next section, a solver for VI problem \eqref{eq:fe:vi}, requires transfer operators between meshes.
+We use five such operators, three of which are standard, linear maps used for multigrid methods \cite{Trottenbergetal2001}.
 Table \ref{tab:transfers} will help the reader keep track.
@@
 \begin{table}
-\begin{tabular}{lll}
+\begin{tabular}{llc}
 \emph{name}  & \emph{action}  & \emph{linear?} \\ \hline
@@
 canonical (dual) restriction  & $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ & \,\checkmark \\
-(nodal) injection restriction & $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ & \,\checkmark \\
-monotone restriction          & $\mR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ &
+(nodal) injection             & $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ & \,\checkmark \\
+maximum restriction           & $\maxR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ & \\
+minimum restriction           & $\minR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ & 
 \end{tabular}
@@
 \medskip
-\caption{Transfer operators used in this paper.  FE vector space $\mathcal{V}^j$ has zero boundary values while $\tilde{\mathcal{V}}^j$ has arbitrary boundary values. $\mR$ is a nonlinear map.}
+\caption{Transfer operators used in this paper.  The FE vector space $\mathcal{V}^j$ has zero boundary values while $\tilde{\mathcal{V}}^j$ has $\tilde \RR$ values, and arbitrary boundary values.}
 \label{tab:transfers}
@@
 
-Canonical prolongation $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ acts as the identity on coarser-mesh functions in $\mathcal{V}^{j-1}$; it is the linear vector space injection $\mathcal{V}^{j-1} \hookrightarrow \mathcal{V}^j$.
+Canonical prolongation $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ acts as the identity on coarser-mesh functions in $\mathcal{V}^{j-1}$; it is the linear vector space inclusion $\mathcal{V}^{j-1} \hookrightarrow \mathcal{V}^j$.
 Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ also acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $(R\sigma)[z] = \sigma[z]$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
 While these canonical transfers do not lose information, they typically involve a change of FE representation.
-For example, the practical formula for canonical restriction, derived by writing coarser-mesh hat functions in terms of finer-mesh ones, is a ``full-weighting'' average \cite{Trottenbergetal2001}.
-It is the transpose of canonical prolongation.
+For example, the practical formula for canonical restriction, derived by writing coarser-mesh hat functions in terms of finer-mesh ones, is a ``full-weighting'' average \cite{Trottenbergetal2001}, and its transpose is canonical prolongation.
 
 To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, observe that for any given values at the (interior) nodes of $\mathcal{T}^j$ there is a unique function in $\mathcal{V}^j$ with those values.
-Also observe that, by our refinement construction, the nodes of $\mathcal{T}^{j-1}$ are also nodes of $\mathcal{T}^j$.
+Also observe that, by our refinement construction, the nodes of $\mathcal{T}^{j-1}$ are nodes of $\mathcal{T}^j$.
 For $z\in\mathcal{V}^j$ we define the injection $\iR z$ as the element of $\mathcal{V}^{j-1}$ which has the same nodal values as $z$ at the nodes of $\mathcal{T}^{j-1}$.
-Note that injection preserves ordering, that is, if $w \le z$ then $\iR w \le \iR z$.
+Note that injection is monotonic, that is, if $w \le z$ then $\iR w \le \iR z$.
 
-Finally, monotone restriction $\mR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ is defined as the nonlinear map which maximizes nodal values over the supports of coarse-mesh hat functions.
-To make the definition precise, first denote the \emph{nodes} of triangulation $\mathcal{T}^j$ by $x_q^j$ for $q=0,\dots,\tilde{m}_j-1$, and then the corresponding hat functions \cite{Elmanetal2014} by $\psi_q^j$.
-(Boundary nodes are included here.)  Monotone restriction of $z\in\tilde{\mathcal{V}}^j$ is
-\begin{equation}
-  \mR z = \sum_{p=1}^{\tilde m_{j-1}} \max \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}\, \psi_p^{j-1}. \label{eq:fe:monotonerestriction}
-\end{equation}
-Note that if $z\in\tilde{\mathcal{V}}^j$ then $z \le \mR z$.
+Finally, the monotone restrictions $\maxR,\minR$ are nonlinear maps which maximize and minimize, respectively, nodal values over the supports of coarser-mesh hat functions.
+To make the definitions precise, first denote the \emph{nodes} of triangulation $\mathcal{T}^j$ by $x_q^j$ for $q=0,\dots,\tilde{m}_j-1$, and then the corresponding hat functions \cite{Elmanetal2014} by $\psi_q^j$.
+(Boundary nodes are included here.)  The monotone restrictions of $z\in\tilde{\mathcal{V}}^j$ are
+\begin{align}
+\maxR z &= \sum_{p=1}^{\tilde m_{j-1}} \max \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}\, \psi_p^{j-1}, \label{eq:fe:monotonerestrictions} \\
+\minR z &= \sum_{p=1}^{\tilde m_{j-1}} \min \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}\, \psi_p^{j-1}, \notag
+\end{align}
+with the result in $\tilde{\mathcal{V}}^{j-1}$.
+The usual ordering rules for the extended real line $\tilde\RR$ are applied.
+Note that if $z\in\tilde{\mathcal{V}}^j$ then $\minR z \le z \le \maxR z$.
 
@@
 The motivation here is that a corrected iterate is admissible, $w^J + z^J \ge \gamma^J$, if and only if the perturbation $z^J$ exceeds the defect constraint, $z^J \ge \chi^J$.
-Next, for $j=J,J-1,\dots,1$ we apply monotone restriction $\mR$ to define the \emph{level defect constraints},
+Next, for $j=J,J-1,\dots,1$ we apply monotone restriction $\maxR$ to define the \emph{level defect constraints},
 \begin{equation}
-\chi^{j-1} = \mR \chi^j. \label{eq:fe:chilevels}
+\chi^{j-1} = \maxR \chi^j. \label{eq:fe:chilevels}
 \end{equation}
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 \begin{center}
@@
     for $j=J$ downto $j=1$ \\+
-      $\chi^{j-1} = \mR \chi^j$ \\
+      $\chi^{j-1} = \maxR \chi^j$ \\
       $\phi^j = \chi^j - P\chi^{j-1}$ \\

### 2022-08-11 paper/mcd2.tex
done with edit sweep, so box constraints mostly documented

@@
 In this Section we construct the corresponding multilevel CD.
+Its constraint sets are designed to allow multigrid coarser-level corrections (next Section) in a manner which does not violate the inequality constraints, and which also does not introduce unnecessary high frequencies into the coarser-level problems.
 
@@
 \end{equation}
-the finest-level \emph{constraint set} of \emph{admissible} FE functions, a closed and convex subset of $\mathcal{V}^J$.
+the finest-level \emph{constraint set} of \emph{admissible} FE functions, a nonempty, closed, and convex subset of $\mathcal{V}^J$.
 As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J = \pm \infty$.
@@
 
-\section{The nonlinear multilevel CD V-cycle} \label{sec:vcycle}
+To define the constraint subsets, the sets generically denoted $\cK_i$ in Sections \ref{sec:cd}--\ref{sec:convergence}, in the current multilevel FE context, suppose $w^J \in \cK^J$ is an admissible fine-level iterate which approximates the solution $u^J$ of VI \eqref{eq:fe:vi}.
+From this iterate we define the finest-level \emph{lower and upper defect constraints} \cite{GraeserKornhuber2009}
+\begin{equation}
+\underline{\chi}^J = \underline{\gamma}^J - w^J, \qquad \overline{\chi}^J = \overline{\gamma}^J - w^J, \label{eq:fe:defectconstraints}
+\end{equation}
+functions in $\tilde{\mathcal{V}}^J$ satisfying $\underline{\chi}^J \le 0 \le \overline{\chi}^J$.
+The motivation here is that a corrected iterate is admissible, $\underline{\gamma}^J \le w^J + z^J \le \overline{\gamma}^J$, if and only if the perturbation $z^J$ is between the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$.
 
-The starting point of a \emph{multilevel constraint decomposition} (MCD) \cite{GraeserKornhuber2009,Tai2003} method is to define constraint subsets, the sets generically denoted $\cK_i$ in sections \ref{sec:vi} and \ref{sec:convergence}, based on monotone restrictions of a defect constraint (see below).
-Our multilevel V-cycle algorithm will define two types of constraint subsets, distinct sets for the downward and upward parts of the cycle.
-We also diverge from the MCD algorithms by Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear \emph{full approximation scheme} (FAS) philosophy \cite{BrandtLivne2011}.
-When $f$ is linear, as in the implemented examples in \cite{GraeserKornhuber2009,Tai2003}, there is no need to coarsen the solution itself, but in the FAS approach here we coarsen the solution iterate using the injection restriction.
-Note that in classical linear multigrid only the residual and correction are coarsened.
-
-To make this precise, suppose $w^J \in \cK^J$ is an admissible fine-level iterate which approximates the solution $u^J$ of VI \eqref{eq:fe:vi}.
-From this iterate we define the \emph{defect constraint} \cite{GraeserKornhuber2009}
+For $j=J,J-1,\dots,1$ we apply monotone restrictions to define the \emph{level defect constraints},
 \begin{equation}
-\chi^J = \gamma^J - w^J,  \label{eq:fe:defectconstraintfcn}
+\underline{\chi}^{j-1} = \maxR \underline{\chi}^j, \qquad \overline{\chi}^{j-1} = \minR \overline{\chi}^j. \label{eq:fe:chilevels}
 \end{equation}
-a nonpositive function in $\tilde{\mathcal{V}}^J$.
-The motivation here is that a corrected iterate is admissible, $w^J + z^J \ge \gamma^J$, if and only if the perturbation $z^J$ exceeds the defect constraint, $z^J \ge \chi^J$.
-Next, for $j=J,J-1,\dots,1$ we apply monotone restriction $\maxR$ to define the \emph{level defect constraints},
+Observe that
 \begin{equation}
-\chi^{j-1} = \maxR \chi^j. \label{eq:fe:chilevels}
+\underline{\chi}^{J} \le \dots \le \underline{\chi}^0 \le 0 \le \overline{\chi}^0 \le \dots \le \overline{\chi}^J. \label{eq:fe:chiordering}
 \end{equation}
-so that $\chi^j \le \chi^{j-1}$.  A $J=4$ example is shown in Figure \ref{fig:chiphilevels}.
-
-We also denote the differences between the level defect constraints by
-\begin{equation}
-\phi^j = \chi^j - \chi^{j-1}, \label{eq:fe:philevels}
-\end{equation}
-and these are nonpositive functions.
-(Definition \eqref{eq:fe:philevels} uses $\chi^{-1}=0$.)  Functions $\phi^j$ are the gaps between the $\chi^j$ shown in Figure \ref{fig:chiphilevels}, with $\phi^0=\chi^0$ in particular.
-Note the telescoping sum decomposition $\sum_{j=0}^J \phi^j = \chi^J$.
+A $J=4$ example, with no upper obstacles ($\overline{\chi}^j=+\infty$), is shown in Figure \ref{fig:chiphilevels}.
 
@@
 \begin{figure}
-\includegraphics[width=0.55\textwidth]{fixfigs/chiphilevels.pdf}
-\caption{Defect constraint $\chi^J = \gamma^J - w^J$ is decomposed using monotone restriction.}
+\includegraphics[width=0.65\textwidth]{fixfigs/chiphilevels.pdf}
+\caption{The lower defect constraint $\underline{\chi}^J = \underline{\gamma}^J - w^J$ is decomposed using maximum restriction: $\underline{\chi}^{j-1} = \maxR \underline{\chi}^j$.}
 \label{fig:chiphilevels}
@@
 
-For the downward part of the multigrid V-cycle we will use the $\phi^j$ as constraints, and for the upward part the $\chi^j$.  That is, let
+We also compute the differences between the level defect constraints:
 \begin{equation}
-\mathcal{D}^j = \{v \in \mathcal{V}^j \,:\, v\ge \phi^j\}, \qquad \mathcal{U}^j = \{v \in \mathcal{V}^j \,:\, v\ge \chi^j\} \label{eq:fe:constraintsets}
+\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1}, \qquad \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1},  \label{eq:fe:philevels}
 \end{equation}
-define \emph{downward} and \emph{upward constraint subsets} of the FE subspaces $\mathcal{V}^j$, respectively.
-The upward sets are larger, $\mathcal{D}^j \subset \mathcal{U}^j$, and set decompositions $\mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i$ apply in the same sense as \eqref{eq:constraintdecomp}.
+using $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$.  Again these constraints bracket zero,
+\begin{equation}
+\underline{\phi}^j \le 0 \le \overline{\phi}^j,  \label{eq:fe:phibrackets}
+\end{equation}
+but these functions are not ordered as in \eqref{eq:fe:chiordering}, but telescoping sums hold for $j=0,1,\dots,J$:
+\begin{equation}
+\sum_{i=0}^j \underline{\phi}^i = \underline{\chi}^j, \qquad \sum_{i=0}^j \overline{\phi}^i = \overline{\chi}^j.  \label{eq:fe:telescoping}
+\end{equation}
 
-Returning to the central MCD idea, we will correct the finest-level iterate $w^J\in \mathcal{K}^J$ by adding modifications from each coarser subspace $\mathcal{V}^j$.
+A multilevel V-cycle is defined in the next Section.
+For the downward direction we will use the differences $\underline{\phi}^j,\overline{\phi}^j$ as constraints, and for the upward part the defect constraints $\underline{\chi}^j,\overline{\chi}^j$.
+That is,
+\begin{align}
+\mathcal{D}^j = \{v \in \mathcal{V}^j \,:\, \underline{\phi}^j \le v \le \overline{\phi}^j\}, \label{eq:fe:constraintsets} \\
+\mathcal{U}^j = \{v \in \mathcal{V}^j \,:\, \underline{\chi}^j \le v \le \overline{\chi}^j\}, \notag
+\end{align}
+are the \emph{downward} and \emph{upward constraint subsets} of the FE subspaces $\mathcal{V}^j$, respectively.
+Observe that the upward sets are larger, $\mathcal{D}^j \subset \mathcal{U}^j$, and in fact by \eqref{eq:fe:telescoping} the set decompositions
+\begin{equation}
+\mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i \label{eq:fe:constraintdecomp}
+\end{equation}
+apply in the same sense as \eqref{eq:constraintdecomp}.
+Also note that the finest-level upward constraint set is equivalent to the original constraint set $\mathcal{K}^J$; for all $z^J \in \mathcal{V}^J$,
+\begin{equation}
+z^J \in \mathcal{U}^J \quad \iff \quad w^J+z^J \in \mathcal{K}^J. \label{eq:fe:finestlevelequivalent}
+\end{equation}
+
+FIXME what are the CDs?  what are the decomposition operators?
+
+\section{The nonlinear multilevel CD V-cycle} \label{sec:vcycle}
+
+Our \emph{nonlinear multilevel constraint decomposition} (NMCD) V-cycle algorithm in this section will use the two types of constraint subsets in \eqref{eq:fe:constraintsets}, with distinct sets for the downward and upward parts of the cycle.
+We also diverge from the multilevel CD algorithms by Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear \emph{full approximation scheme} (FAS) philosophy \cite{BrandtLivne2011}.
+In the FAS approach here we must coarsen the solution iterate itself, using the injection restriction, in constrast to classical linear multigrid, wherein only the residual and correction are coarsened, and in constrast to the implemented examples in \cite{GraeserKornhuber2009,Tai2003} in which $f$ is linear.
+
+FIXME from here
+
+We will correct the finest-level iterate $w^J\in \mathcal{K}^J$ by adding modifications from each coarser subspace $\mathcal{V}^j$.
 By nesting \eqref{eq:fe:nestedspaces} these corrections live in the fine-level subspace $\mathcal{V}^J$, but the corrected iterate needs to be admissible as well.
@@
 \end{center}
-\caption{An NMCD V-cycle computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j$ expand into larger constraint subsets $\mathcal{U}^j$.}
+\caption{An NMCD V-cycle computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j$ may expand into larger constraint subsets $\mathcal{U}^j$.}
 \label{fig:nmcdvcycle}
@@
 
-These ideas come together in Algorithm \ref{alg:nmcd}, an NMCD V-cycle.  This procedure acts in-place on the final argument $w^J$, but it leaves the other inputs $\ell^J,\gamma^J$ unchanged.
+These ideas come together in Algorithm \ref{alg:nmcd}, an NMCD V-cycle.  This procedure acts in-place on the final argument $w^J$, but it leaves the other vector inputs $\ell^J,\gamma^J$ unchanged.
 
@@
 \appendix
-\section{Reductions of the NMCD algorithm}
+\section{Reduction of NMCD to unconstrained FAS}
 
@@
 
+\section{Reduction of NMCD for linear operators}
+
 FIXME by assuming $f$ is linear, reduce to Algorithm 4.7 in \cite{GraeserKornhuber2009} but with $\text{V}(\nu_1,\nu_2)$ cycles

### 2022-08-11 c/bratufd.c
Copy over p4pdes/c/ch7/solutions/bratu2D.c for working FAS+NGS.
This copy has no dependence on c/ch6/poissonfunctions.h|c.
In particular it has no analytical jacobian.  It has full FAS+NGS
power, and it works fine with -snes_fd_color and standard
Newton-multigrid.

Next step:  Borrowing ideas from c/ch9/, convert to Q1 FE
and check full FAS+NGS power is retained.

@@ -1,0 +1,265 @@
+static char help[] =
+"Solve nonlinear Liouville-Bratu equation in 2D on a structured-grid.  Option prefix lb_.\n"
+"Solves\n"
+"  - nabla^2 u - lambda e^u = 0\n"
+"on the unit square [0,1]x[0,1] subject to zero Dirichlet boundary conditions.\n"
+"Critical value occurs about at lambda = 6.808.  Optional exact solution by\n"
+"Liouville (1853) for case lambda=1.0.\n\n";
+
+/*
+incredible performance: FAS+NGS, matrix-free, full-cycles, NGS for coarse solve
+400 million unknowns in 32 seconds, full 10 digit accuracy
+uses all memory of my Thelio massive machine (uses over 90% of 128 Gb; note 20 cores out of 40)
+note only 350 flops per degree of freedom!
+
+$ timer mpiexec -n 20 --map-by core --bind-to hwthread ./bratufd -da_grid_x 6 -da_grid_y 6 -lb_exact -snes_rtol 1.0e-10 -snes_converged_reason -lb_showcounts -snes_type fas -snes_fas_type full -fas_levels_snes_type ngs -fas_levels_snes_ngs_sweeps 2 -fas_levels_snes_max_it 1 -fas_coarse_snes_type ngs -fas_coarse_snes_ngs_sweeps 2 -fas_coarse_snes_max_it 4 -da_refine 12
+Nonlinear solve converged due to CONVERGED_FNORM_RELATIVE iterations 1
+flops = 1.467e+11,  residual calls = 416,  NGS calls = 208
+done on 20481 x 20481 grid:   error |u-uexact|_inf = 1.728e-10
+real 31.50
+*/
+
+/* excellent evidence of convergence and optimality in Liouville exact solution case (use opt PETSc build):
+NEWTON+MG with FD Jacobian matrix is fast
+$ for LEV in 6 7 8 9 10; do timer ./bratufd -da_refine $LEV -lb_exact -snes_rtol 1.0e-10 -snes_converged_reason -lb_showcounts -snes_type newtonls -snes_fd_color -pc_type mg; done
+FAS+NGS matrix-free full cycles are much faster
[…]

### 2022-08-11 c/testit.sh

@@ -1,0 +1,43 @@
+#!/bin/bash
+
+# A script to run regression tests.  "make test" runs this script as follows:
+#    ./testit.sh PROGRAM OPTS PROCESSES TESTNUM
+
+rm -f maketmp tmp difftmp
+
+make $1 > maketmp 2>&1;
+
+grep warning maketmp
+
+CURRDIR=${PWD##*/}
+
+if [ $3 -eq 1 ]; then
+    CMD="./$1 $2"
+elif [ $3 -eq -1 ]; then
+    CMD="$1 $2"
+else
+    CMD="mpiexec -n $3 ./$1 $2"
+fi
+
+if [[ ! -f output/$1.test$4 ]]; then
+    echo "FAIL: Test #$4 of $CURRDIR/$1"
+    echo "       command = '$CMD'"
+    echo "       OUTPUT MISSING"
+
+else
+
+    $CMD &> tmp
+
+    diff output/$1.test$4 tmp > difftmp
+
+    if [[ -s difftmp ]] ; then
+       echo "FAIL: Test #$4 of $CURRDIR/$1"
+       echo "       command = '$CMD'"
+       echo "       diffs follow:"
+       cat difftmp
+    else
+       echo "PASS: Test #$4 of $CURRDIR/$1"
+       rm -f maketmp tmp difftmp
+    fi
+
+fi

### 2022-08-11 c/bratufd.c
Bratu FD only needs star stencil.

@@ -1,6 +1,6 @@
 static char help[] =
-"Solve nonlinear Liouville-Bratu equation in 2D on a structured-grid.  Option prefix lb_.\n"
-"Solves\n"
+"Solve nonlinear Liouville-Bratu equation by finite differences\n"
+"in 2D on a structured-grid.  Option prefix lb_.  Solves\n"
 "  - nabla^2 u - lambda e^u = 0\n"
 "on the unit square [0,1]x[0,1] subject to zero Dirichlet boundary conditions.\n"
 "Critical value occurs about at lambda = 6.808.  Optional exact solution by\n"
@@ -73,7 +73,7 @@ int main(int argc,char **argv) {
                             "bratufd.c",bctx.lambda,&(bctx.lambda),NULL));
     PetscCall(PetscOptionsBool("-exact","use case of Liouville exact solution",
                             "bratufd.c",bctx.exact,&(bctx.exact),NULL));
-    PetscCall(PetscOptionsBool("-showcounts","at finish, print numbers of calls to call-back functions",
+    PetscCall(PetscOptionsBool("-showcounts","print counts for calls to call-back functions",
                             "bratufd.c",showcounts,&showcounts,NULL));
     PetscOptionsEnd();
     if (bctx.exact) {
@@ -84,7 +84,7 @@ int main(int argc,char **argv) {
     }
 
     PetscCall(DMDACreate2d(PETSC_COMM_WORLD, DM_BOUNDARY_NONE, DM_BOUNDARY_NONE,
-                        DMDA_STENCIL_BOX,  // contrast with fish
+                        DMDA_STENCIL_STAR,
                         3,3,PETSC_DECIDE,PETSC_DECIDE,1,1,NULL,NULL,&da));
     PetscCall(DMSetApplicationContext(da,&bctx));
     PetscCall(DMSetFromOptions(da));

### 2022-08-11 c/bratu.c
get started on Q1 FE version of FAS+NGS Bratu solver

@@ -1,0 +1,260 @@
+static char help[] =
+"Solve nonlinear Liouville-Bratu equation by Q1 finite elements\n"
+"in 2D on a structured-grid.  Option prefix lb_.  Solves\n"
+"  - nabla^2 u - lambda e^u = 0\n"
+"on the unit square [0,1]x[0,1] subject to zero Dirichlet boundary conditions.\n"
+"Critical value occurs about at lambda = 6.808.  Optional exact solution by\n"
+"Liouville (1853) for case lambda=1.0.\n\n";
+
+#include <petsc.h>
+#include "quadrature.h"
+#include "q1fem.h"
+
+typedef struct {
+  // Dirichlet boundary condition g(x,y,z)
+  PetscReal (*g_bdry)(PetscReal x, PetscReal y, PetscReal z, void *ctx);
+  PetscReal lambda;
+  PetscBool exact;
+  PetscInt  residualcount, ngscount, quadpts;
+} BratuCtx;
+
+static PetscReal g_zero(PetscReal x, PetscReal y, PetscReal z, void *ctx) {
+    return 0.0;
+}
+
+static PetscReal g_liouville(PetscReal x, PetscReal y, PetscReal z, void *ctx) {
+    PetscReal r2 = (x + 1.0) * (x + 1.0) + (y + 1.0) * (y + 1.0),
+              qq = r2 * r2 + 1.0,
+              omega = r2 / (qq * qq);
+    return PetscLogReal(32.0 * omega);
+}
+
+extern PetscErrorCode FormUExact(DMDALocalInfo*, Vec, BratuCtx*);
+extern PetscErrorCode FormFunctionLocal(DMDALocalInfo*, PetscReal **,
+                                        PetscReal**, BratuCtx*);
+extern PetscErrorCode NonlinearGS(SNES, Vec, Vec, void*);
+
+int main(int argc,char **argv) {
+    DM             da, da_after;
+    SNES           snes;
+    Vec            u, uexact;
+    BratuCtx       bctx;
+    DMDALocalInfo  info;
+    PetscBool      showcounts = PETSC_FALSE;
+    PetscLogDouble lflops, flops;
+    PetscReal      errinf;
+
+    PetscCall(PetscInitialize(&argc,&argv,NULL,help));
+
+    bctx.g_bdry = &g_zero;
+    bctx.lambda = 1.0;
+    bctx.exact = PETSC_FALSE;
+    bctx.residualcount = 0;
+    bctx.ngscount = 0;
[…]

### 2022-08-11 c/q1fem.h

@@ -1,0 +1,61 @@
+// Tools for Q1 finite elements in two dimensions.  Here xi,eta
+// denote coordinates on [-1,1]^2 reference element.  Node numbering is
+//   1 *---* 0
+//     |   |
+//   2 *---* 3
+// on reference element.  (E.g. L=0 is (1,1), ..., L=3 is (1,-1).)
+// Reference element hat functions are denoted chi_L(xi,eta),
+// with gradient (in reference coordinates) returned by dchi().
+// For documentation see Chapter 9 of Bueler, "PETSc for Partial
+// Differential Equations", SIAM Press 2021, and c/ch9/phelm.c at
+//   https://github.com/bueler/p4pdes
+
+#ifndef Q1FEM_H_
+#define Q1FEM_H_
+
+static PetscReal xiL[4]  = { 1.0, -1.0, -1.0,  1.0},
+                 etaL[4] = { 1.0,  1.0, -1.0, -1.0};
+
+static PetscReal chi(PetscInt L, PetscReal xi, PetscReal eta) {
+    return 0.25 * (1.0 + xiL[L] * xi) * (1.0 + etaL[L] * eta);
+}
+
+// evaluate v(xi,eta) on reference element using local node numbering
+static PetscReal eval(const PetscReal v[4], PetscReal xi, PetscReal eta) {
+    return   v[0] * chi(0,xi,eta) + v[1] * chi(1,xi,eta)
+           + v[2] * chi(2,xi,eta) + v[3] * chi(3,xi,eta);
+}
+
+typedef struct {
+    PetscReal  xi, eta;
+} gradRef;
+
+static gradRef dchi(PetscInt L, PetscReal xi, PetscReal eta) {
+    const gradRef result = {0.25 * xiL[L]  * (1.0 + etaL[L] * eta),
+                            0.25 * etaL[L] * (1.0 + xiL[L]  * xi)};
+    return result;
+}
+
+// evaluate partial derivs of v(xi,eta) on reference element
+static gradRef deval(const PetscReal v[4], PetscReal xi, PetscReal eta) {
+    gradRef   sum = {0.0,0.0}, tmp;
+    PetscInt  L;
+    for (L=0; L<4; L++) {
+        tmp = dchi(L,xi,eta);
+        sum.xi += v[L] * tmp.xi;  sum.eta += v[L] * tmp.eta;
+    }
+    return sum;
+}
+
[…]

### 2022-08-11 c/quadrature.h

@@ -1,0 +1,26 @@
+// One-dimensional Gauss-Legendre quadrature for interval [-1,1],
+// of degree 1,2,3.
+
+#ifndef QUADRATURE_H_
+#define QUADRATURE_H_
+
+#define MAXPTS 3
+
+typedef struct {
+    PetscInt   n;          // number of quadrature points for this rule
+    PetscReal  xi[MAXPTS], // locations in [-1,1]
+               w[MAXPTS];  // weights (sum to 2)
+} Quad1D;
+
+static const Quad1D gausslegendre[3]
+    = {  {1,
+          {0.0,                NAN,               NAN},
+          {2.0,                NAN,               NAN}},
+         {2,
+          {-0.577350269189626, 0.577350269189626, NAN},
+          {1.0,                1.0,               NAN}},
+         {3,
+          {-0.774596669241483, 0.0,               0.774596669241483},
+          {0.555555555555556,  0.888888888888889, 0.555555555555556}} };
+
+#endif

### 2022-08-11 c/bratu.c
WIP: converting to Q1 FEM

@@ -148,19 +148,24 @@ PetscErrorCode FormUExact(DMDALocalInfo *info, Vec u, BratuCtx* user) {
     return 0;
 }
 
-FIXME FROM HERE
+PetscReal IntegrandRef(PetscReal hx, PetscReal hy, PetscInt L,
+                       const PetscReal ff[4], const PetscReal uu[4],
+                       PetscReal xi, PetscReal eta, BratuCtx *user) {
+  const gradRef    du    = deval(uu,xi,eta),
+                   dchiL = dchi(L,xi,eta);
+  return GradInnerProd(hx,hy,du,dchiL)
+         - user->lambda * PetscExpScalar(eval(uu,xi,eta)) * chi(L,xi,eta);
+}
 
 // compute F(u), the residual of the discretized PDE on the given grid
 PetscErrorCode FormFunctionLocal(DMDALocalInfo *info, PetscReal **au,
                                  PetscReal **FF, BratuCtx *user) {
+    const Quad1D    q = gausslegendre[user->quadpts-1];
+    const PetscReal hx = 1.0 / (PetscReal)(info->mx - 1),
+                    hy = 1.0 / (PetscReal)(info->my - 1);
     PetscInt   i, j;
-    PetscReal  hx, hy, darea, hxhy, hyhx, x, y;
+    PetscReal  x, y;
 
-    hx = 1.0 / (PetscReal)(info->mx - 1);
-    hy = 1.0 / (PetscReal)(info->my - 1);
-    darea = hx * hy;
-    hxhy = hx / hy;
-    hyhx = hy / hx;
     for (j = info->ys; j < info->ys + info->ym; j++) {
         y = j * hy;
         for (i = info->xs; i < info->xs + info->xm; i++) {

### 2022-08-11 c/bratufd.c
does not need 3D args (holdover from using Ch 6 pieces)

@@ -29,18 +29,18 @@ $ for LEV in 6 7 8 9 10; do timer ./bratufd -da_refine $LEV -lb_exact -snes_rtol
 #include <petsc.h>
 
 typedef struct {
-  // Dirichlet boundary condition g(x,y,z)
-  PetscReal (*g_bdry)(PetscReal x, PetscReal y, PetscReal z, void *ctx);
+  // Dirichlet boundary condition g(x,y)
+  PetscReal (*g_bdry)(PetscReal x, PetscReal y, void *ctx);
   PetscReal lambda;
   PetscBool exact;
   int       residualcount, ngscount;
 } BratuCtx;
 
-static PetscReal g_zero(PetscReal x, PetscReal y, PetscReal z, void *ctx) {
+static PetscReal g_zero(PetscReal x, PetscReal y, void *ctx) {
     return 0.0;
 }
 
-static PetscReal g_liouville(PetscReal x, PetscReal y, PetscReal z, void *ctx) {
+static PetscReal g_liouville(PetscReal x, PetscReal y, void *ctx) {
     PetscReal r2 = (x + 1.0) * (x + 1.0) + (y + 1.0) * (y + 1.0),
               qq = r2 * r2 + 1.0,
               omega = r2 / (qq * qq);
@@ -148,7 +148,7 @@ PetscErrorCode FormUExact(DMDALocalInfo *info, Vec u, BratuCtx* user) {
         y = j * hy;
         for (i=info->xs; i<info->xs+info->xm; i++) {
             x = i * hx;
-            au[j][i] = user->g_bdry(x,y,0.0,user);
+            au[j][i] = user->g_bdry(x,y,user);
         }
     }
     PetscCall(DMDAVecRestoreArray(info->da, u, &au));
@@ -171,7 +171,7 @@ PetscErrorCode FormFunctionLocal(DMDALocalInfo *info, PetscReal **au,
         for (i = info->xs; i < info->xs + info->xm; i++) {
             if (j==0 || i==0 || i==info->mx-1 || j==info->my-1) {
                 x = i * hx;
-                FF[j][i] = au[j][i] - user->g_bdry(x,y,0.0,user);
+                FF[j][i] = au[j][i] - user->g_bdry(x,y,user);
             } else {
                 FF[j][i] =   hyhx * (2.0 * au[j][i] - au[j][i-1] - au[j][i+1])
                            + hxhy * (2.0 * au[j][i] - au[j-1][i] - au[j+1][i])
[…]

### 2022-08-11 c/bratufd.c
symmetrize Jacobian in bratufd.c; benefits Newton-Krylov but not FAS+NGS?

@@ -8,20 +8,20 @@ static char help[] =
 
 /*
 incredible performance: FAS+NGS, matrix-free, full-cycles, NGS for coarse solve
-400 million unknowns in 32 seconds, full 10 digit accuracy
-uses all memory of my Thelio massive machine (uses over 90% of 128 Gb; note 20 cores out of 40)
-note only 350 flops per degree of freedom!
+400 million unknowns in 22 seconds, 10-digit accuracy
+uses 66% of 128 Gb memory of my Thelio massive machine; note 20 cores out of 40
+note only 352 flops per degree of freedom!
 
-$ timer mpiexec -n 20 --map-by core --bind-to hwthread ./bratufd -da_grid_x 6 -da_grid_y 6 -lb_exact -snes_rtol 1.0e-10 -snes_converged_reason -lb_showcounts -snes_type fas -snes_fas_type full -fas_levels_snes_type ngs -fas_levels_snes_ngs_sweeps 2 -fas_levels_snes_max_it 1 -fas_coarse_snes_type ngs -fas_coarse_snes_ngs_sweeps 2 -fas_coarse_snes_max_it 4 -da_refine 12
+$ timer mpiexec -n 20 --map-by core --bind-to hwthread ./bratufd -da_grid_x 5 -da_grid_y 5 -lb_exact -snes_rtol 1.0e-10 -snes_converged_reason -lb_showcounts -snes_type fas -snes_fas_type full -fas_levels_snes_type ngs -fas_levels_snes_ngs_sweeps 2 -fas_levels_snes_max_it 1 -fas_coarse_snes_type ngs -fas_coarse_snes_ngs_sweeps 2 -fas_coarse_snes_max_it 4 -da_refine 12
 Nonlinear solve converged due to CONVERGED_FNORM_RELATIVE iterations 1
-flops = 1.467e+11,  residual calls = 416,  NGS calls = 208
-done on 20481 x 20481 grid:   error |u-uexact|_inf = 1.728e-10
-real 31.50
+flops = 9.448e+10,  residual calls = 416,  NGS calls = 208
+done on 16385 x 16385 grid:   error |u-uexact|_inf = 4.038e-10
+real 21.70
 */
 
 /* excellent evidence of convergence and optimality in Liouville exact solution case (use opt PETSc build):
 NEWTON+MG with FD Jacobian matrix is fast
[…]

### 2022-08-11 c/q1fem.h
avoid compiler warning about unused function

@@ -53,9 +53,12 @@ static PetscReal GradInnerProd(PetscReal hx, PetscReal hy,
     return cx * du.xi * dv.xi + cy * du.eta * dv.eta;
 }
 
+#if 0
+// unused in Bratu problem
 static PetscReal GradPow(PetscReal hx, PetscReal hy,
                          gradRef du, PetscReal P, PetscReal eps) {
     return PetscPowScalar(GradInnerProd(hx,hy,du,du) + eps*eps, P/2.0);
 }
+#endif
 
 #endif

### 2022-08-11 c/bratu.c
This seems to be a correctly implement Q1 FEM residual-only solver for Bratu.  (Next: convert NGS to Q1.)

@@ -11,18 +11,18 @@ static char help[] =
 #include "q1fem.h"
 
 typedef struct {
-  // Dirichlet boundary condition g(x,y,z)
-  PetscReal (*g_bdry)(PetscReal x, PetscReal y, PetscReal z, void *ctx);
+  // Dirichlet boundary condition g(x,y)
+  PetscReal (*g_bdry)(PetscReal x, PetscReal y, void *ctx);
   PetscReal lambda;
   PetscBool exact;
   PetscInt  residualcount, ngscount, quadpts;
 } BratuCtx;
 
-static PetscReal g_zero(PetscReal x, PetscReal y, PetscReal z, void *ctx) {
+static PetscReal g_zero(PetscReal x, PetscReal y, void *ctx) {
     return 0.0;
 }
 
-static PetscReal g_liouville(PetscReal x, PetscReal y, PetscReal z, void *ctx) {
+static PetscReal g_liouville(PetscReal x, PetscReal y, void *ctx) {
     PetscReal r2 = (x + 1.0) * (x + 1.0) + (y + 1.0) * (y + 1.0),
               qq = r2 * r2 + 1.0,
               omega = r2 / (qq * qq);
@@ -51,11 +51,13 @@ int main(int argc,char **argv) {
     bctx.exact = PETSC_FALSE;
     bctx.residualcount = 0;
     bctx.ngscount = 0;
+    bctx.quadpts = 2;
     PetscOptionsBegin(PETSC_COMM_WORLD,"lb_","Liouville-Bratu equation solver options","");
     PetscCall(PetscOptionsReal("-lambda","coefficient of e^u (reaction) term",
                             "bratu.c",bctx.lambda,&(bctx.lambda),NULL));
     PetscCall(PetscOptionsBool("-exact","use case of Liouville exact solution",
                             "bratu.c",bctx.exact,&(bctx.exact),NULL));
+    // WARNING: coarse problems are badly solved with -lb_quadpts 1, so avoid in MG
     PetscCall(PetscOptionsInt("-quadpts","number n of quadrature points (= 1,2,3 only)",
                             "bratu.c",bctx.quadpts,&(bctx.quadpts),NULL));
     PetscCall(PetscOptionsBool("-showcounts","print counts for calls to call-back functions",
@@ -141,49 +143,90 @@ PetscErrorCode FormUExact(DMDALocalInfo *info, Vec u, BratuCtx* user) {
         y = j * hy;
[…]

### 2022-08-11 c/q1fem.h

@@ -16,11 +16,13 @@
 static PetscReal xiL[4]  = { 1.0, -1.0, -1.0,  1.0},
                  etaL[4] = { 1.0,  1.0, -1.0, -1.0};
 
+// FLOPS: 6
 static PetscReal chi(PetscInt L, PetscReal xi, PetscReal eta) {
     return 0.25 * (1.0 + xiL[L] * xi) * (1.0 + etaL[L] * eta);
 }
 
 // evaluate v(xi,eta) on reference element using local node numbering
+// FLOPS: 7 + 4 * chi = 31
 static PetscReal eval(const PetscReal v[4], PetscReal xi, PetscReal eta) {
     return   v[0] * chi(0,xi,eta) + v[1] * chi(1,xi,eta)
            + v[2] * chi(2,xi,eta) + v[3] * chi(3,xi,eta);
@@ -47,6 +49,7 @@ static gradRef deval(const PetscReal v[4], PetscReal xi, PetscReal eta) {
     return sum;
 }
 
+// FLOPS: 9
 static PetscReal GradInnerProd(PetscReal hx, PetscReal hy,
                                gradRef du, gradRef dv) {
     const PetscReal cx = 4.0 / (hx * hx),  cy = 4.0 / (hy * hy);

### 2022-08-12 c/bratu.c
correct flops counting and document where I stand (next: NGS impl)

@@ -11,8 +11,7 @@ static char help[] =
 #include "q1fem.h"
 
 typedef struct {
-  // Dirichlet boundary condition g(x,y)
-  PetscReal (*g_bdry)(PetscReal x, PetscReal y, void *ctx);
+  PetscReal (*g_bdry)(PetscReal x, PetscReal y, void *ctx);  // Dirichlet b.c.
   PetscReal lambda;
   PetscBool exact;
   PetscInt  residualcount, ngscount, quadpts;
@@ -76,8 +75,8 @@ int main(int argc,char **argv) {
     }
 
     PetscCall(DMDACreate2d(PETSC_COMM_WORLD, DM_BOUNDARY_NONE, DM_BOUNDARY_NONE,
-                        DMDA_STENCIL_BOX,  // contrast with bratufd
-                        3,3,PETSC_DECIDE,PETSC_DECIDE,1,1,NULL,NULL,&da));
+                           DMDA_STENCIL_BOX,  // contrast with bratufd
+                           3,3,PETSC_DECIDE,PETSC_DECIDE,1,1,NULL,NULL,&da));
     PetscCall(DMSetApplicationContext(da,&bctx));
     PetscCall(DMSetFromOptions(da));
     PetscCall(DMSetUp(da));  // this must be called BEFORE SetUniformCoordinates
@@ -150,7 +149,7 @@ PetscErrorCode FormUExact(DMDALocalInfo *info, Vec u, BratuCtx* user) {
     return 0;
 }
 
-// FLOPS: 9 + 4 + 31 + 6 = 50
+// FLOPS: 4 + (48 + 8 + 9 + 31 + 6) = 106
 PetscReal IntegrandRef(PetscReal hx, PetscReal hy, PetscInt L,
                        const PetscReal uu[4],
                        PetscReal xi, PetscReal eta, BratuCtx *user) {
@@ -219,9 +218,9 @@ PetscErrorCode FormFunctionLocal(DMDALocalInfo *info, PetscReal **au,
         }
     }
     // only count quadrature-point residual computations:
-    //     4 + 50 = 54 flops per quadrature point
+    //     4 + 106 = 110 flops per quadrature point
     //     q.n^2 quadrature points per element
-    PetscCall(PetscLogFlops(54.0 * q.n * q.n * info->xm * info->ym));
+    PetscCall(PetscLogFlops(110.0 * q.n * q.n * info->xm * info->ym));
     (user->residualcount)++;
     return 0;
 }

### 2022-08-12 c/bratufd.c

@@ -6,20 +6,7 @@ static char help[] =
 "Critical value occurs about at lambda = 6.808.  Optional exact solution by\n"
 "Liouville (1853) for case lambda=1.0.\n\n";
 
-/*
-incredible performance: FAS+NGS, matrix-free, full-cycles, NGS for coarse solve
-400 million unknowns in 22 seconds, 10-digit accuracy
-uses 66% of 128 Gb memory of my Thelio massive machine; note 20 cores out of 40
-note only 352 flops per degree of freedom!
-
-$ timer mpiexec -n 20 --map-by core --bind-to hwthread ./bratufd -da_grid_x 5 -da_grid_y 5 -lb_exact -snes_rtol 1.0e-10 -snes_converged_reason -lb_showcounts -snes_type fas -snes_fas_type full -fas_levels_snes_type ngs -fas_levels_snes_ngs_sweeps 2 -fas_levels_snes_max_it 1 -fas_coarse_snes_type ngs -fas_coarse_snes_ngs_sweeps 2 -fas_coarse_snes_max_it 4 -da_refine 12
-Nonlinear solve converged due to CONVERGED_FNORM_RELATIVE iterations 1
-flops = 9.448e+10,  residual calls = 416,  NGS calls = 208
-done on 16385 x 16385 grid:   error |u-uexact|_inf = 4.038e-10
-real 21.70
-*/
-
-/* excellent evidence of convergence and optimality in Liouville exact solution case (use opt PETSc build):
+/* excellent evidence of convergence and optimality in Liouville exact solution case:
 NEWTON+MG with FD Jacobian matrix is fast
 $ for LEV in 6 7 8 9 10; do timer ./bratufd -da_refine $LEV -lb_exact -snes_rtol 1.0e-10 -snes_converged_reason -lb_showcounts -snes_type newtonls -snes_fd_color -ksp_type cg -pc_type mg; done
 FAS+NGS matrix-free full cycles are much faster

### 2022-08-12 c/q1fem.h

@@ -32,6 +32,7 @@ typedef struct {
     PetscReal  xi, eta;
 } gradRef;
 
+// FLOPS: 8
 static gradRef dchi(PetscInt L, PetscReal xi, PetscReal eta) {
     const gradRef result = {0.25 * xiL[L]  * (1.0 + etaL[L] * eta),
                             0.25 * etaL[L] * (1.0 + xiL[L]  * xi)};
@@ -39,6 +40,7 @@ static gradRef dchi(PetscInt L, PetscReal xi, PetscReal eta) {
 }
 
 // evaluate partial derivs of v(xi,eta) on reference element
+// FLOPS: 4 * (8 + 4) = 48
 static gradRef deval(const PetscReal v[4], PetscReal xi, PetscReal eta) {
     gradRef   sum = {0.0,0.0}, tmp;
     PetscInt  L;

### 2022-08-12 c/bratufd.c
minor: put get array in right spot

@@ -202,14 +202,14 @@ PetscErrorCode NonlinearGS(SNES snes, Vec u, Vec b, void *ctx) {
     hxhy = hx / hy;
     hyhx = hy / hx;
 
+    if (b) {
+        PetscCall(DMDAVecGetArrayRead(da,b,&ab));
+    }
     PetscCall(DMGetLocalVector(da,&uloc));
     for (l=0; l<sweeps; l++) {
         PetscCall(DMGlobalToLocalBegin(da,u,INSERT_VALUES,uloc));
         PetscCall(DMGlobalToLocalEnd(da,u,INSERT_VALUES,uloc));
         PetscCall(DMDAVecGetArray(da,uloc,&au));
-        if (b) {
-            PetscCall(DMDAVecGetArrayRead(da,b,&ab));
-        }
         for (j = info.ys; j < info.ys + info.ym; j++) {
             y = j * hy;
             for (i = info.xs; i < info.xs + info.xm; i++) {

### 2022-09-27 paper/mcd2.tex
remove convergence material and save as old fragment

@@
 
-The theory presented in \cite{Tai2003} is extended as follows:
+The multilevel CD algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} are extended in three practical ways:
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
 \begin{enumerate}
-\item We do not assume that the continuum VI problem arises from optimization of a scalar objective---examples are explored in Sections \ref{sec:vi} and \ref{sec:results}---but we prove convergence in norm at the same rate as the original method converges in energy (Section \ref{sec:convergence}).
-% THIS IS THE HOPE
-\end{enumerate}
-Furthermore the multilevel CD algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} are extended in three practical ways:
-\begin{enumerate}
-\setcounter{enumi}{1} % resume at (ii)
 \item We make finite element (FE) implementations of multilevel CD algorithm more effective for nonlinear problems by applying a full approximation storage (FAS; see \cite{Brandt1977,Bruneetal2015}) approach.
@@
 For the classical obstacle problem with a Laplacian operator, certain multilevel techniques are known to improve performance relative to the multilevel CD method \cite{GraeserKornhuber2009}, such as the truncated monotone multigrid method \cite{Kornhuber1994}.
-These improved methods either track the active set in the discretization or modify the nodal basis functions, and in this sense they are discrete algorithms, while by contrast the CD method applies at the level of the continuum problem (Sections \ref{sec:cd} and \ref{sec:convergence}).
+These improved methods either track the active set in the discretization or modify the nodal basis functions, and in this sense they are discrete algorithms, while by contrast the CD method applies essentially at the level of the continuum problem; see Section \ref{sec:cd}.
 Acceleration of our NMCD method via active-set and/or basis-level manipulations represents a potential extension of the method here, and is a topic for future research.
-
-% A BRIDGE TOO FAR:  In one example at the end of this paper (Section \ref{sec:resultsnonlocal}) we consider a nonlocal residual functional, that is, one which is not a partial differential operator.
-Each evaluation of this functional requires the solution of a Stokes problem for a layer of fluid \nocite{SayagWorster2013} on a substrate (which forms the obstacle), and thus the corresponding FE operator discretization is also not sparse.
-In this case we cannot prove coercivity but we nonetheless succeed in demonstrating near-optimal complexity of the Section \ref{sec:multilevel} multilevel CD algorithm in practice.
 
@@
 \end{equation}
-the $i$th \emph{subset update}.
-As the reader may check, $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.
-Note $\hat w = u^* + \sum_i \hat w_i - \Pi_i u^*$ also holds.
-We will use these identities in Section \ref{sec:convergence}.
-
-% FOLLOWING CONTAINS ASPIRATIONS
-The convergence results in Section \ref{sec:convergence} require substantial damping when applying additive Algorithm \ref{alg:basiccd-add}, namely we require $\alpha \le 1/m$ as in \cite{Tai2003}.
-By constrast, Algorithm \ref{alg:basiccd-mult} can be shown to converge without damping ($\alpha\le 1$).
-In Section \ref{sec:results} we demonstrate practical convergence for larger ranges of $\alpha$ than suggested by the theory.
+the $i$th \emph{subset update}.  As the reader may check, $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.  Note $\hat w = u^* + \sum_i \hat w_i - \Pi_i u^*$ also holds.
 
@@
 Extension of the multilevel CD method to such non-local functionals $f$ is a topic for future research.
-
-\section{Convergence of a basic CD iteration} \label{sec:convergence}
-
-We now prove the convergence of the basic additive CD iteration, Algorithm \ref{alg:basiccd-add}. % ASPIRATION
-A convergence proof becomes possible if we restrict to $2$-coercive operators, require substantial damping, and make certain other assumptions as in \cite{Tai2003}.
-However, our proof applies to VI problems which are not optimizations, thus extending the theory in \cite{Tai2003}.
-
-First we define a new form related to VI problem \eqref{eq:vi}.
-
-\begin{definition} Suppose $f:\cK \to \cV'$ and $\ell \in \cV'$.  For $v,w \in \cK$ let
-\begin{equation}
-  E(v,w) = \ip{f(v)-\ell}{v-w}.  \label{eq:normlikedefn}
-\end{equation}
-\end{definition}
-
-If $u^*$ solves \eqref{eq:vi} then $E(v,u^*)$ is a ``merit function'' for an iterate $v$, a concept used in solving nonlinear equations \cite{NocedalWright2006}, except that $u^*$ is not generally known so this meaning is only theoretical.
-In any case, our use of $E(v,u^*)$ replaces the scalar objective ``$F(v)$'' used in \cite{Tai2003}.
-The next lemma shows $E(v,u^*)$ is bounded below by a norm.
-
-\begin{lemma} \label{lem:normlike}  Suppose $f:\mathcal{K} \to \mathcal{V}'$ is $p$-coercive with constant $\kappa>0$, and $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.  For  all $v \in \mathcal{K}$,
-\begin{equation}
-  E(v,u^*) \ge \kappa \|v-u^*\|^p.  \label{eq:normlikebound}
-\end{equation}
-\end{lemma}
-
-\begin{proof}
-\begin{align*}
-E(v,u^*) &= \ip{f(v)}{v-u^*} - \ip{f(u^*)}{v-u^*} + \ip{f(u^*)}{v-u^*} - \ip{\ell}{v-u^*} \\
-   &\ge \ip{f(v)-f(u^*)}{v-u^*} + 0 \ge \kappa \|v-u^*\|^p.  \qedhere
-\end{align*}
-\end{proof}
-
-Next we make two assumptions which are essentially the same as (7) and (8) in \cite{Tai2003}.
-These inequalities might be described as ``totally Lipschitz'' requirements, over the CD defined by \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, for the maps $\Pi_i$ and $f$, respectively.
-
-\begin{assumptions*}  There exists a constant $C_1>0$ so that
-\begin{equation}
-\left(\sum_{i=0}^{m-1} \|\Pi_i u - \Pi_i v\|^2\right)^{1/2} \le C_1 \|u-v\| \label{as:lipschitzrestrictions}
-\end{equation}
-for all $u,v\in\cK$.  There exists another constant $C_2>0$ so that for all $v_i \in \cV_i$ and $y_j \in \cV_j$, and all $w_{ij} \in \cK$ such that $w_{ij} + v_i \in \cK$, it holds that
-\begin{equation}
-\sum_{i=0}^{m-1} \sum_{j=0}^{m-1} \left|\ip{f(w_{ij} + v_i) - f(w_{ij})}{y_j}\right| \le C_2 \left(\sum_{i=0}^{m-1} \|v_i\|^2\right)^{1/2} \left(\sum_{j=0}^{m-1} \|y_j\|^2\right)^{1/2}.
-\label{as:lipschitzresidual}
-\end{equation}
-\end{assumptions*}
-
-We make the following observations:
-\begin{itemize}
-\item The constants $C_1,C_2$ generally depend on $m$, the number of sets in the CD.
-In a multilevel CD method (Sections \ref{sec:multilevel} and \ref{sec:vcycle}) this number is logarithmic in the number of degrees of freedom (nodes) in the fine-level mesh.
-\item Regarding the convergence proof below, if the iterates from Algorithm \ref{alg:basiccd-add} are bounded in $\mathcal{K}$ then constants $C_1,C_2$ are permitted to depend continuously on $u,v,w_{ij},v_i,y_i$ \cite{Tai2003}.
-\item Bound \eqref{as:lipschitzrestrictions} addresses only the CD, and not the nonlinear operator $f$.
-Tai \cite{Tai2003} gives a value of $C_1$ for certain obstacle problem CDs using $P_1$ FE spaces over shape-regular and quasi-uniform triangulations.
-In particular, $C_1$ is known for the standard multilevel hierarchy, so the cases demonstrated in Section \ref{sec:results} are covered.
-\item If $f$ is defined on all of $\mathcal{V}$ then one may verify \eqref{as:lipschitzresidual} over a subspace, versus constraint, decomposition.
-For example, \eqref{as:lipschitzresidual} is verified for the classical obstacle problem over arbitrary $w_{ij} \in \cV$.
-(Unconstrained versions of \eqref{as:lipschitzrestrictions}, \eqref{as:lipschitzresidual} are (13), (14) in \cite{TaiXu2002}, respectively.)
-\item If $f$ is Lipschitz then the existence of $C_2$ is clear \cite{TaiXu2002}.
-However, our convergence rate bound is improved by taking the smallest $C_2$ for which \eqref{as:lipschitzresidual} holds.
-\end{itemize}
-
-From the above definitions and assumptions we have the following estimate for Algorithm \ref{alg:basiccd-add}.  Our proof roughly follows that of Theorem 1 in \cite{Tai2003}.
-
-\begin{lemma} \label{lem:core}  Suppose $f$ is $2$-coercive with constant $\kappa>0$, and that $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.
-Assume $\mathcal{V}_i$, $\mathcal{K}_i$, $\Pi_i$ form a CD satisfying \eqref{as:lipschitzrestrictions}, and that \eqref{as:lipschitzresidual} is satisfied.
-If $\hat w$ is computed from $u \in \mathcal{K}$ by Algorithm \ref{alg:basiccd-add} then
-\begin{equation}
-   E(\hat w,u^*) \le C_2 \sum_{i=0}^{m-1} \|e_i\|^2 + \kappa^{-1} C_1 C_2 \left(\sum_{i=0}^{m-1} \|e_i\|^2\right)^{1/2} E(u,u^*)^{1/2} \label{eq:core}
-\end{equation}
-where again $e_i=\hat w_i - \Pi_i u$ (equation \eqref{eq:ithupdate}).
-\end{lemma}
-
-\begin{proof}  Recall that $\hat w = u^* + \sum_i \hat w_i - \Pi_i u^*$.  Expand $E(\hat w,u^*)$ and substitute $v_i = \Pi_i u^*$ into the boxed VI in Algorithm \ref{alg:basiccd-add} to find
-\begin{align}
-E(\hat w,u^*) &= \sum_{i=0}^{m-1} \ip{f(\hat w)}{\hat w_i - \Pi_i u^*} - \ip{\ell}{\hat w_i - \Pi_i u^*} \notag \\
-    &\le \sum_{i=0}^{m-1} \ip{f(\hat w)}{\hat w_i - \Pi_i u^*} + \ip{f(u + e_i)}{\Pi_i u^* - \hat w_i} \notag \\
-    &= \sum_{i=0}^{m-1} \ip{f(\hat w) - f(u + e_i)}{\hat w_i - \Pi_i u^*}. \label{eq:startcore}
-\end{align}
-
-For $i\in \{0,1,\dots,m-1\}$ we now use wrapped indices to define progressive states $\phi_{i,j}$:
-\begin{align}
-\phi_{i,0} &= u + e_i, \label{eq:gridcore} \\
-\phi_{i,1} &= u + e_i + e_{i+1}, \notag \\
-  &\vdots \notag \\
-\phi_{i,m-1} &= u + e_i + e_{i+1} + \dots + e_{m-1} + e_0 + \dots + e_{i-1}. \notag
-\end{align}
-Observe that $\phi_{i,j} \in \cK$ by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum}, and that $\phi_{i,m-1} = u + \sum_i e_i = \hat w$, thus that \eqref{eq:startcore} can be written
-\begin{equation}
-E(\hat w,u^*) = \sum_{i=0}^{m-1} \ip{f(\phi_{i,m-1}) - f(\phi_{i,0})}{\hat w_i - \Pi_i u^*}. \label{eq:startcoreclean}
-\end{equation}
-From \eqref{eq:startcoreclean}, telescope into a sum over $j=1,\dots,m-1$, and then apply the triangle inequality:
-\begin{align}
-E(\hat w,u^*) &\le \sum_{i=0}^{m-1} \sum_{j=1}^{m-1} \ip{f(\phi_{i,j}) - f(\phi_{i,j-1})}{\hat w_i - \Pi_i u^*} \label{eq:nextcore} \\
-  &\le \sum_{i=0}^{m-1} \sum_{j=1}^{m-1} \left|\ip{f(\phi_{i,j}) - f(\phi_{i,j-1})}{\hat w_i - \Pi_i u^*}\right|. \notag
-\end{align}
-
-Let $Z=\left(\sum_{i=0}^{m-1} \|e_i\|^2\right)^{1/2}$.  FIXME: MAKE THE RENUMBERING EXPLICIT AND NOTE $v_0=0$ IN APPLYING \eqref{as:lipschitzresidual}
-
-renumbering the indices $i,j$ we may apply assumption \eqref{as:lipschitzresidual}, then the triangle inequality, and then assumption \eqref{as:lipschitzrestrictions}:
-\begin{align}
-E(\hat w,u^*) &\le C_2 Z \left(\sum_{i=0}^{m-1} \|\hat w_i - \Pi_i u^*\|^2\right)^{1/2} = C_2 Z \left(\sum_{i=0}^{m-1} \|\hat w_i - \Pi_i u + \Pi_i u - \Pi_i u^*\|^2\right)^{1/2} \label{eq:nextnextcore} \\
-  &\le C_2 Z \left(Z + \left(\sum_{i=0}^{m-1} \|\Pi_i u - \Pi_i u^*\|^2\right)^{1/2}\right) \le C_2 Z \left(Z + C_1 \|u-u^*\|\right). \notag
-\end{align}
-Finally apply Lemma \ref{lem:normlike} to give \eqref{eq:core}.
-\end{proof}
-
-FIXME NOW THE ENTIRE BATTLE IS TO GET A CONVEXITY RESULT $E(w,u) \ge C \sum_{j=0}^{m-1} \|e_j\|^2$ for some reasonable $C$
 
@@
 
-To define the constraint subsets, the sets generically denoted $\cK_i$ in Sections \ref{sec:cd}--\ref{sec:convergence}, in the current multilevel FE context, suppose $w^J \in \cK^J$ is an admissible fine-level iterate which approximates the solution $u^J$ of VI \eqref{eq:fe:vi}.
+To define the constraint subsets, the sets generically denoted $\cK_i$ in Section \ref{sec:cd}, in the current multilevel FE context, suppose $w^J \in \cK^J$ is an admissible fine-level iterate which approximates the solution $u^J$ of VI \eqref{eq:fe:vi}.
 From this iterate we define the finest-level \emph{lower and upper defect constraints} \cite{GraeserKornhuber2009}

### 2022-09-27 paper/old/convergence.tex
remove convergence material and save as old fragment

FIXME this fragment is an attempt to re-prove convergence a la Tai, but without using the objective function

We do not assume that the continuum VI problem arises from optimization of a scalar objective---examples are explored in Sections \ref{sec:vi} and \ref{sec:results}---but we prove convergence in norm at the same rate as the original method converges in energy (Section \ref{sec:convergence}). % THIS IS THE HOPE

The convergence results in Section \ref{sec:convergence} require substantial damping when applying additive Algorithm \ref{alg:basiccd-add}, namely we require $\alpha \le 1/m$ as in \cite{Tai2003}.  By constrast, Algorithm \ref{alg:basiccd-mult} can be shown to converge without damping ($\alpha\le 1$).  In Section \ref{sec:results} we demonstrate practical convergence for larger ranges of $\alpha$ than suggested by the theory.

\section{Convergence of a basic CD iteration} \label{sec:convergence}

We now prove the convergence of the basic additive CD iteration, Algorithm \ref{alg:basiccd-add}. % ASPIRATION
A convergence proof becomes possible if we restrict to $2$-coercive operators, require substantial damping, and make certain other assumptions as in \cite{Tai2003}.  However, our proof applies to VI problems which are not optimizations, thus extending the theory in \cite{Tai2003}.

First we define a new form related to VI problem \eqref{eq:vi}.

\begin{definition} Suppose $f:\cK \to \cV'$ and $\ell \in \cV'$.  For $v,w \in \cK$ let
\begin{equation}
  E(v,w) = \ip{f(v)-\ell}{v-w}.  \label{eq:normlikedefn}
\end{equation}
\end{definition}

If $u^*$ solves \eqref{eq:vi} then $E(v,u^*)$ is a ``merit function'' for an iterate $v$, a concept used in solving nonlinear equations \cite{NocedalWright2006}, except that $u^*$ is not generally known so this meaning is only theoretical.  In any case, our use of $E(v,u^*)$ replaces the scalar objective ``$F(v)$'' used in \cite{Tai2003}.  The next lemma shows $E(v,u^*)$ is bounded below by a norm.

\begin{lemma} \label{lem:normlike}  Suppose $f:\mathcal{K} \to \mathcal{V}'$ is $p$-coercive with constant $\kappa>0$, and $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.  For  all $v \in \mathcal{K}$,
\begin{equation}
  E(v,u^*) \ge \kappa \|v-u^*\|^p.  \label{eq:normlikebound}
\end{equation}
\end{lemma}

\begin{proof}
\begin{align*}
E(v,u^*) &= \ip{f(v)}{v-u^*} - \ip{f(u^*)}{v-u^*} + \ip{f(u^*)}{v-u^*} - \ip{\ell}{v-u^*} \\
   &\ge \ip{f(v)-f(u^*)}{v-u^*} + 0 \ge \kappa \|v-u^*\|^p.  \qedhere
\end{align*}
\end{proof}

Next we make two assumptions which are essentially the same as (7) and (8) in \cite{Tai2003}.  These inequalities might be described as ``totally Lipschitz'' requirements, over the CD defined by \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, for the maps $\Pi_i$ and $f$, respectively.

\begin{assumptions*}  There exists a constant $C_1>0$ so that
\begin{equation}
\left(\sum_{i=0}^{m-1} \|\Pi_i u - \Pi_i v\|^2\right)^{1/2} \le C_1 \|u-v\| \label{as:lipschitzrestrictions}
\end{equation}
for all $u,v\in\cK$.  There exists another constant $C_2>0$ so that for all $v_i \in \cV_i$ and $y_j \in \cV_j$, and all $w_{ij} \in \cK$ such that $w_{ij} + v_i \in \cK$, it holds that
\begin{equation}
\sum_{i=0}^{m-1} \sum_{j=0}^{m-1} \left|\ip{f(w_{ij} + v_i) - f(w_{ij})}{y_j}\right| \le C_2 \left(\sum_{i=0}^{m-1} \|v_i\|^2\right)^{1/2} \left(\sum_{j=0}^{m-1} \|y_j\|^2\right)^{1/2}. \label{as:lipschitzresidual}
\end{equation}
\end{assumptions*}

We make the following observations:
\begin{itemize}
\item The constants $C_1,C_2$ generally depend on $m$, the number of sets in the CD.  In a multilevel CD method (Sections \ref{sec:multilevel} and \ref{sec:vcycle}) this number is logarithmic in the number of degrees of freedom (nodes) in the fine-level mesh.
\item Regarding the convergence proof below, if the iterates from Algorithm \ref{alg:basiccd-add} are bounded in $\mathcal{K}$ then constants $C_1,C_2$ are permitted to depend continuously on $u,v,w_{ij},v_i,y_i$ \cite{Tai2003}.
\item Bound \eqref{as:lipschitzrestrictions} addresses only the CD, and not the nonlinear operator $f$.  Tai \cite{Tai2003} gives a value of $C_1$ for certain obstacle problem CDs using $P_1$ FE spaces over shape-regular and quasi-uniform triangulations.  In particular, $C_1$ is known for the standard multilevel hierarchy, so the cases demonstrated in Section \ref{sec:results} are covered.
\item If $f$ is defined on all of $\mathcal{V}$ then one may verify \eqref{as:lipschitzresidual} over a subspace, versus constraint, decomposition.  For example, \eqref{as:lipschitzresidual} is verified for the classical obstacle problem over arbitrary $w_{ij} \in \cV$.  (Unconstrained versions of \eqref{as:lipschitzrestrictions}, \eqref{as:lipschitzresidual} are (13), (14) in \cite{TaiXu2002}, respectively.)
\item If $f$ is Lipschitz then the existence of $C_2$ is clear \cite{TaiXu2002}.  However, our convergence rate bound is improved by taking the smallest $C_2$ for which \eqref{as:lipschitzresidual} holds.
\end{itemize}

From the above definitions and assumptions we have the following estimate for Algorithm \ref{alg:basiccd-add}.  Our proof roughly follows that of Theorem 1 in \cite{Tai2003}.

\begin{lemma} \label{lem:core}  Suppose $f$ is $2$-coercive with constant $\kappa>0$, and that $u^* \in \mathcal{K}$ solves \eqref{eq:vi}.  Assume $\mathcal{V}_i$, $\mathcal{K}_i$, $\Pi_i$ form a CD satisfying \eqref{as:lipschitzrestrictions}, and that \eqref{as:lipschitzresidual} is satisfied.  If $\hat w$ is computed from $u \in \mathcal{K}$ by Algorithm \ref{alg:basiccd-add} then
\begin{equation}
   E(\hat w,u^*) \le C_2 \sum_{i=0}^{m-1} \|e_i\|^2 + \kappa^{-1} C_1 C_2 \left(\sum_{i=0}^{m-1} \|e_i\|^2\right)^{1/2} E(u,u^*)^{1/2} \label{eq:core}
\end{equation}
where again $e_i=\hat w_i - \Pi_i u$ (equation \eqref{eq:ithupdate}).
\end{lemma}

\begin{proof}  Recall that $\hat w = u^* + \sum_i \hat w_i - \Pi_i u^*$.  Expand $E(\hat w,u^*)$ and substitute $v_i = \Pi_i u^*$ into the boxed VI in Algorithm \ref{alg:basiccd-add} to find
\begin{align}
E(\hat w,u^*) &= \sum_{i=0}^{m-1} \ip{f(\hat w)}{\hat w_i - \Pi_i u^*} - \ip{\ell}{\hat w_i - \Pi_i u^*} \notag \\
    &\le \sum_{i=0}^{m-1} \ip{f(\hat w)}{\hat w_i - \Pi_i u^*} + \ip{f(u + e_i)}{\Pi_i u^* - \hat w_i} \notag \\
    &= \sum_{i=0}^{m-1} \ip{f(\hat w) - f(u + e_i)}{\hat w_i - \Pi_i u^*}. \label{eq:startcore}
\end{align}

For $i\in \{0,1,\dots,m-1\}$ we now use wrapped indices to define progressive states $\phi_{i,j}$:
\begin{align}
\phi_{i,0} &= u + e_i, \label{eq:gridcore} \\
\phi_{i,1} &= u + e_i + e_{i+1}, \notag \\
  &\vdots \notag \\
\phi_{i,m-1} &= u + e_i + e_{i+1} + \dots + e_{m-1} + e_0 + \dots + e_{i-1}. \notag
\end{align}
Observe that $\phi_{i,j} \in \cK$ by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum}, and that $\phi_{i,m-1} = u + \sum_i e_i = \hat w$, thus that \eqref{eq:startcore} can be written
\begin{equation}
E(\hat w,u^*) = \sum_{i=0}^{m-1} \ip{f(\phi_{i,m-1}) - f(\phi_{i,0})}{\hat w_i - \Pi_i u^*}. \label{eq:startcoreclean}
\end{equation}
From \eqref{eq:startcoreclean}, telescope into a sum over $j=1,\dots,m-1$, and then apply the triangle inequality:
\begin{align}
E(\hat w,u^*) &\le \sum_{i=0}^{m-1} \sum_{j=1}^{m-1} \ip{f(\phi_{i,j}) - f(\phi_{i,j-1})}{\hat w_i - \Pi_i u^*} \label{eq:nextcore} \\
  &\le \sum_{i=0}^{m-1} \sum_{j=1}^{m-1} \left|\ip{f(\phi_{i,j}) - f(\phi_{i,j-1})}{\hat w_i - \Pi_i u^*}\right|. \notag
\end{align}

Let $Z=\left(\sum_{i=0}^{m-1} \|e_i\|^2\right)^{1/2}$.  FIXME: MAKE THE RENUMBERING EXPLICIT AND NOTE $v_0=0$ IN APPLYING \eqref{as:lipschitzresidual}

renumbering the indices $i,j$ we may apply assumption \eqref{as:lipschitzresidual}, then the triangle inequality, and then assumption \eqref{as:lipschitzrestrictions}:
\begin{align}
E(\hat w,u^*) &\le C_2 Z \left(\sum_{i=0}^{m-1} \|\hat w_i - \Pi_i u^*\|^2\right)^{1/2} = C_2 Z \left(\sum_{i=0}^{m-1} \|\hat w_i - \Pi_i u + \Pi_i u - \Pi_i u^*\|^2\right)^{1/2} \label{eq:nextnextcore} \\
  &\le C_2 Z \left(Z + \left(\sum_{i=0}^{m-1} \|\Pi_i u - \Pi_i u^*\|^2\right)^{1/2}\right) \le C_2 Z \left(Z + C_1 \|u-u^*\|\right). \notag
\end{align}
Finally apply Lemma \ref{lem:normlike} to give \eqref{eq:core}.
\end{proof}

FIXME NOW THE ENTIRE BATTLE IS TO GET A CONVEXITY RESULT $E(w,u) \ge C \sum_{j=0}^{m-1} \|e_j\|^2$ for some reasonable $C$

### 2022-09-27 paper/mcd2.tex
some edits in the first three sections

@@
 We extend the constraint decomposition (CD) method of X.-C.~Tai \cite{Tai2003} to nonlinear variational inequality (VI) problems to which it has not been applied.
-The convergence of this method, for several types of decompositions, was proven by Tai for coercive VI problems which arise from minimization of a convex functional over a convex set.
 In its multilevel form the method has been shown to have optimal complexity for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
+The convergence of this method, for several types of decompositions, was proven for coercive VI problems which arise from minimization of a convex functional over a convex set \cite{Tai2003}.
 
-The multilevel CD algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} are extended in three practical ways:
+The multilevel CD algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} are extended here in three practical ways:
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
@@
 \item We make finite element (FE) implementations of multilevel CD algorithm more effective for nonlinear problems by applying a full approximation storage (FAS; see \cite{Brandt1977,Bruneetal2015}) approach.
-Results from the resulting \emph{nonlinear multilevel constraint decomposition} (NMCD) algorithm, which avoids global (Newton) linearization (compare \cite{GraeserKornhuber2009}), are shown in Section \ref{sec:results} for several nonlinear problems.
-\item Our implementation permits ``box'' constraints, namely upper and lower obstacles.
-\item In Sections \ref{sec:multilevel} and \ref{sec:vcycle} we explain why, given the manner in which constraint sets are decomposed in the multilevel context \cite{GraeserKornhuber2009}, multilevel ``up-smoothing'' is more efficient than ``down-smoothing''.
+Section \ref{sec:results} shows results from this \emph{nonlinear multilevel constraint decomposition} (NMCD) algorithm  for several nonlinear problems.
+Note that this algorithm avoids global (Newton) linearization; compare \cite{GraeserKornhuber2009}.
+\item Our implementation permits ``box'' constraints, namely upper and lower obstacles; see Section \ref{sec:multilevel}.
+\item In Sections \ref{sec:multilevel} and \ref{sec:vcycle} we explain why, given the manner in which constraint sets are decomposed in the multilevel context \cite{GraeserKornhuber2009}, ``up-smoothing'' is more efficient than ``down-smoothing''.
 This observation seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}.
@@
 
-The iterates from a CD method are always admissible, and thus the operator can be defined only for admissible states.
-Admissible-iterate methods should permit direct solutions of certain VI problems, such as fluid-layer dynamics problems \cite{Bueler2021conservation,JouvetBueler2012}, for which non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
-Our NMCD scheme is also designed for these problems, some of which require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation, as discussed in \cite{Bueler2021conservation}.
+The iterates from a CD method are always admissible, and thus the operator need only be defined for admissible states.
+This property should permit direct solutions of certain VI problems, such as fluid-layer dynamics problems \cite{Bueler2021conservation,JouvetBueler2012}, for which non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
+Our NMCD scheme is designed for such problems, some of which require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation, as discussed in \cite{Bueler2021conservation}, but we defer this application to future work.
 
-One of our examples demonstrates the application of the NMCD method to a problem of porous-media type, to which the extended convergence theory, \emph{(i)} above, does not apply because the nonlinear operator is not known to be coercive.
-By ``freezing' the solution-dependent coefficient, the operator is approximated by a coercive operator for the duration of the V-cycle.
-The resulting scheme is highly-effective for doubly-nonlinear diffusion operators, another type of problem to which the extended theory does not apply (Section \ref{sec:results}).
+One of our examples demonstrates the application of the NMCD method to a problem of porous-media type, a nonlinear operator which is not known to be coercive.
+By ``freezing' the solution-dependent coefficient, the operator is approximated by a coercive operator for the duration of the V-cycle, and this scheme is also effective for doubly-nonlinear diffusion operators.
+See Section \ref{sec:results}.
 
@@
 Informally, if we conceptualize the dual pairing as an inner product then \eqref{eq:vi} says that the angle between $f(u^*)-\ell$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
-That is, while $f(u^*)-\ell$ may not be zero, it points directly into $\cK$.
-However, if $u^*$ is in the interior of the constraint set, $u^*\in\cK^\circ$, then \eqref{eq:vi} implies $f(u^*)=\ell$.
+That is, while $f(u^*)-\ell$ may not be zero, it must point directly into $\cK$.
+On the other hand, if $u^*$ is in the interior of the constraint set, $u^*\in\cK^\circ$, then \eqref{eq:vi} implies $f(u^*)=\ell$.
 
@@
 
-It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, uniquely so when $f$ is strictly monotone.
+It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, uniquely when $f$ is strictly monotone.
 As in the calculus of variations \cite{Evans2010}, coercivity permits a compactness argument for unbounded sets $\cK$.
@@
 
-When $f$ is monotone, \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
+When $f$ is monotone, VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
 In fact, suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cK \to \cV'$.
@@
 
-From now on $\Omega \subset \RR^d$ denotes a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.
+FIXME the problems we actually address are partial differential operators, and furthermore $\mathcal{K}$ incorporates both Dirichlet boundary conditions, as equality constraints, and inequality constraints
+
+From now on let $\Omega \subset \RR^d$ denote a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.
 Sobolev spaces \cite{Evans2010} are denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$.
@@
     $$F(v) = \int_\Omega \frac{a(x)}{p} |\grad v|^p\,dx - \ip{\ell}{v}.$$
-Then $F'(v) = f(v) - \ell$ and $F$ is a convex functional (since $f$ is coercive).
-For any closed and convex $\cK\subset \cV$, VI problem \eqref{eq:vi} for is equivalent to optimization problem \eqref{eq:minimization}.\end{example}
+Then $F$ is a convex functional and $F'(v) = f(v) - \ell$ for $f$ given in \eqref{eq:plaplacian}.
+For any closed and convex $\cK\subset \cV$, VI problem \eqref{eq:vi} for $f$ is equivalent to optimization problem \eqref{eq:minimization} for $F$.\end{example}
 
@@
 If $\bX \ne 0$ then VI \eqref{eq:vi} for $f$ in \eqref{eq:advectiondiffusion} does not correspond to a minimization problem.
-Indeed, $\ip{f(u)}{v}$ is not symmetric in that case,\footnote{Assume $\bX \ne 0$ is continuous for simplicity.
+Indeed, $\ip{f(u)}{v}$ is not symmetric in that case, so $f$ cannot be the gradient of a scalar objective.
+To see this, assume $\bX \ne 0$ is continuous for simplicity.
 For $u,v$ which are zero on $\partial \Omega$, note $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.
-By constructing $u,v$ locally near some point where $\bX$ is nonzero.
-one may show $b(u,v)\ne 0$.} so $f$ cannot be the gradient of a scalar objective.
+By constructing $u,v$ locally near some point where $\bX$ is nonzero one may show $b(u,v)\ne 0$.
 
@@
 Further examples of nonlinear VI problems appear in ice sheet models \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids.
-In the above Examples $f(u)$ may be defined over all of $\mathcal{V}$, but in some geophysical fluid problems the value of $f(u)$ is only defined for admissible $u\in \mathcal{K}$ because $u$ defines the fluid layer thickness \cite{Bueler2021conservation}.
 
@@
 \end{equation}
-holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$.  Equation \eqref{eq:subspacedecomp} is called a \emph{subspace decomposition} \cite{Xu1992}.
+holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$.  Equation \eqref{eq:subspacedecomp} is called a \emph{subspace decomposition} of $\cV$ \cite{Xu1992}.
 
@@
 \begin{equation}
-\cK = \sum_{i=0}^{m-1} \cK_i. \label{eq:constraintdecomp}
+\cK = \sum_{i=0}^{m-1} \cK_i \label{eq:constraintdecomp}
 \end{equation}
+for a closed, convex subset $\cK \subset \cV$, as considered in the previous section.
 The sum in \eqref{eq:constraintdecomp} is required to hold in two senses \cite{TaiTseng2002}: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.
-Note that sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp}, because the $\cV_i$ are subspaces, and that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique.
+Observe that sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp} because the $\cV_i$ are subspaces.
+Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique.
 Also, $\cK_i \not\subset \cK$ in many applications; see the cartoon in Figure \ref{fig:cartoon}.
+
+Finally, for each $\cK_i$ suppose there are bounded, generally-nonlinear \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.
+Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:multilevel}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
+\begin{equation}
+v = \sum_{i=0}^{m-1} \Pi_i v.  \label{eq:constraintrestrictionsum}
+\end{equation}
+Clearly \eqref{eq:constraintrestrictionsum} implies sense \emph{(i)} for \eqref{eq:constraintdecomp}.
+A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} \cite{Tai2003}.
+Figure \ref{fig:cartoon} suggests how a low-dimensional example might look.
 
@@
 
-Finally, for each $\cK_i$ we assume there are bounded, generally-nonlinear \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.
-Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:multilevel}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
-\begin{equation}
-v = \sum_{i=0}^{m-1} \Pi_i v;  \label{eq:constraintrestrictionsum}
-\end{equation}
-see Figure \ref{fig:cartoon}.
-Clearly \eqref{eq:constraintrestrictionsum} implies sense \emph{(i)} for \eqref{eq:constraintdecomp}.
-A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} \cite{Tai2003}.
-
-In Sections \ref{sec:multilevel} and \ref{sec:vcycle} we will introduce discretizations and a practical algorithms for finite-dimensional VI problems, but observe here that this CD concept applies at the level of the continuum problem.
+In Sections \ref{sec:multilevel} and \ref{sec:vcycle} we will introduce discretizations and a practical algorithms for finite-dimensional VI problems, but observe that this CD concept applies at the level of the continuum problem.
 The following two obstacle problem examples illustrate this.
@@
 
-Our second example is a disjoint frequency decomposition.  A multilevel CD, e.g.~the one proposed in Section \ref{sec:multilevel}, approximates such a frequency decomposition.
+Our second example is a disjoint frequency decomposition.  A multilevel CD, specifically the one proposed in Section \ref{sec:multilevel}, approximates such a frequency decomposition.
 
@@
 Suppose $\psi \in W_{\text{per}}^{p,k}(\Omega)$ and let $\cK = \{v \ge \psi\} \subset \cV$.
-Without using any detailed notation for Fourier representation, but noting that the frequencies are discrete because $\Omega$ is compact, suppose $\{\cV_i\}$ are $m<\infty$ subspaces of $\cV$ defined by an (nonoverlapping) partition by frequency, thus satisfying \eqref{eq:subspacedecomp} as an orthogonal decomposition.
+Without using any detailed notation for Fourier representation, but noting that the frequencies are discrete because $\Omega$ is compact, suppose $\{\cV_i\}$ are $m<\infty$ subspaces of $\cV$ defined by an (nonoverlapping) partition of $W_{\text{per}}^{p,k}(\Omega)$ by frequency, thus satisfying \eqref{eq:subspacedecomp} as an orthogonal decomposition.
 Suppose $P_i:\cV \to \cV_i$ are the corresponding orthogonal projections, satisfying $I = \sum_i P_i$.
@@
 
-Again note that $\cK_i \not\subset \cK$ in these cases.
-Specifically in Example \ref{ex:domaindecomposition}, if $\psi$ is positive over portions of $\Omega$ where the decomposition into overlapping subdomains $\Omega_i$ is nontrivial, then $\cK_i \not\subset \cK$, and similarly for Example \ref{ex:frequencydecomposition}.
-The important inclusion is $\cK_i \subset \cV_i$.
+Again note that $\cK_i \not\subset \cK$ in these cases; the important inclusion is $\cK_i \subset \cV_i$.
+Specifically in Example \ref{ex:domaindecomposition}, if $\psi$ is positive over portions of $\Omega$ where the decomposition into overlapping subdomains $\Omega_i$ is nontrivial, then $\cK_i \not\subset \cK$, and something similar holds for Example \ref{ex:frequencydecomposition}.
 

### 2022-10-14 c/src/ldc.c
slight change in LDC stdout

@@ -8,14 +8,12 @@ PetscErrorCode LDCCreateCoarsest(PetscBool verbose, DM cdmda, LDC *ldc) {
     PetscReal      xymin[2], xymax[2];
     ldc->_level = 0;
     ldc->_printinfo = verbose;
-    if (ldc->_printinfo)
-        PetscCall(PetscPrintf(PETSC_COMM_WORLD,
-        "  LDC info: creating LDC at level %d",ldc->_level));
     ldc->dal = cdmda;
     if (ldc->_printinfo) {
         PetscCall(DMDAGetLocalInfo(ldc->dal,&info));
         PetscCall(PetscPrintf(PETSC_COMM_WORLD,
-        " based on %d x %d DMDA\n",info.mx,info.my));
+        "  LDC info: creating LDC at level %d based on provided %d x %d coarse DMDA\n",
+        ldc->_level,info.mx,info.my));
     }
     PetscCall(DMGetBoundingBox(ldc->dal,xymin,xymax));
     ldc->_xmin = xymin[0];

### 2022-10-14 c/nmcd.c
fully draft NMCD V-cycle; need to revise PNGS

@@ -78,7 +78,8 @@ extern PetscErrorCode ProjectedNGS(LDC*, PetscBool, Vec, Vec, ObsCtx*);
 
 int main(int argc,char **argv) {
     Level          *levs;
-    Vec            gamlow, w, tmp, uexact, F;
+    Vec            gamlow, w, uexact, F,
+                   gplusy, tmpfine, tmpcoarse;
     ObsCtx         ctx;
     DMDALocalInfo  finfo;
     PetscInt       totlevs=2, cycles=1, jtop, j;
@@ -149,10 +150,11 @@ int main(int argc,char **argv) {
             levs[j].dmda = levs[j].ldc.dal;
         }
         PetscCall(DMSetApplicationContext(levs[j].dmda,&ctx));
-        PetscCall(DMCreateGlobalVector(levs[j].dmda,&(levs[j].g)));
-        PetscCall(VecDuplicate(levs[j].g,&(levs[j].ell)));
-        PetscCall(VecDuplicate(levs[j].g,&(levs[j].y)));
-        PetscCall(VecDuplicate(levs[j].g,&(levs[j].z)));
+        PetscCall(DMGetGlobalVector(levs[j].dmda,&(levs[j].g)));
+        PetscCall(DMGetGlobalVector(levs[j].dmda,&(levs[j].ell)));
+        PetscCall(DMGetGlobalVector(levs[j].dmda,&(levs[j].z)));
+        if (j > 0)
+            PetscCall(DMGetGlobalVector(levs[j].dmda,&(levs[j].y)));
     }
     jtop = totlevs - 1;
 
@@ -161,7 +163,7 @@ int main(int argc,char **argv) {
     PetscCall(AssertBoundaryAdmissible(levs[jtop].dmda,&ctx));
 
     // generate finest-level obstacle gamlow as Vec
-    PetscCall(DMCreateGlobalVector(levs[jtop].dmda,&gamlow));
+    PetscCall(DMGetGlobalVector(levs[jtop].dmda,&gamlow));
     PetscCall(LDCVecFromFormula(levs[jtop].ldc,gamma_lower,gamlow,&ctx));
 
     // create initial iterate w on finest level
@@ -177,6 +179,7 @@ int main(int argc,char **argv) {
 
     // complete set-up of LDC stack for initial iterate u
     PetscCall(LDCFinestUpDCsFromVecs(w,NULL,gamlow,&(levs[jtop].ldc)));
[…]

### 2022-10-14 c/nmcd.c
this version is acting like it solves the problem ... weirds me out

@@ -74,7 +74,7 @@ static PetscReal fg_zero(PetscReal x, PetscReal y, void *ctx) {
 extern PetscErrorCode AssertBoundaryAdmissible(DM da, ObsCtx*);
 extern PetscErrorCode FormExact(DM da, Vec, ObsCtx*);
 extern PetscErrorCode ApplyOperatorF(DM, Vec, Vec, ObsCtx*);
-extern PetscErrorCode ProjectedNGS(LDC*, PetscBool, Vec, Vec, ObsCtx*);
+extern PetscErrorCode ProjectedNGS(LDC*, PetscBool, Vec, Vec, Vec, ObsCtx*);
 
 int main(int argc,char **argv) {
     Level          *levs;
@@ -182,11 +182,18 @@ int main(int argc,char **argv) {
     PetscCall(DMRestoreGlobalVector(levs[jtop].dmda,&gamlow));
     PetscCall(LDCGenerateDCsVCycle(&(levs[jtop].ldc)));
 
-    // solve the problem
+    // report range on initial w and f^J(w) for initial iterate
     PetscCall(VecPrintRange(w,"initial iterate w",""));
-    PetscCall(PetscPrintf(PETSC_COMM_WORLD,"*NOT YET* SOLVING PROBLEM BY %d V-CYCLES\n",cycles));
+    PetscCall(DMGetGlobalVector(levs[jtop].dmda,&F));
+    PetscCall(ApplyOperatorF(levs[jtop].dmda,w,F,&ctx));
+    PetscCall(VecPrintRange(F,"initial residual f^J(w)",""));
+    PetscCall(DMRestoreGlobalVector(levs[jtop].dmda,&F));
 
-    // one NMCD V-cycle with one smoother iteration at each j>0 level
+    // one NMCD V-cycle with one smoother iteration at each level
+    if (cycles != 1) {
+        PetscCall(PetscPrintf(PETSC_COMM_WORLD,"IGNORING cycles=%d\n",cycles)); // FIXME
+    }
+    PetscCall(PetscPrintf(PETSC_COMM_WORLD,"solving with 1 V-cycle ...\n"));
     PetscCall(VecCopy(w,levs[jtop].g));
     PetscCall(VecSet(levs[jtop].ell,0.0));
     // downward direction
@@ -197,9 +204,8 @@ int main(int argc,char **argv) {
         PetscCall(DMGetGlobalVector(levs[j-1].dmda,&tmpcoarse));
[…]

### 2022-10-14 c/nmcd.c
fix some code which was clearly not correct in parallel

@@ -193,7 +193,8 @@ int main(int argc,char **argv) {
     if (cycles != 1) {
         PetscCall(PetscPrintf(PETSC_COMM_WORLD,"IGNORING cycles=%d\n",cycles)); // FIXME
     }
-    PetscCall(PetscPrintf(PETSC_COMM_WORLD,"solving with 1 V-cycle ...\n"));
+    PetscCall(PetscPrintf(PETSC_COMM_WORLD,"solving using 1 V-cycle with %d levels ...\n",
+                          totlevs));
     PetscCall(VecCopy(w,levs[jtop].g));
     PetscCall(VecSet(levs[jtop].ell,0.0));
     // downward direction
@@ -412,6 +413,7 @@ PetscReal _IntegrandRef(PetscReal hx, PetscReal hy, PetscInt L,
 //     F_ij = u_ij - g(x_i,y_j)
 PetscErrorCode ApplyOperatorF(DM da, Vec u, Vec F, ObsCtx *user) {
     DMDALocalInfo    info;
+    Vec              uloc;
     const Q1Quad1D   q = Q1gausslegendre[user->quadpts-1];
     const PetscInt   li[4] = {0,-1,-1,0},  lj[4] = {0,0,-1,-1};
     PetscInt         i, j, l, PP, QQ, r, s;
@@ -423,7 +425,9 @@ PetscErrorCode ApplyOperatorF(DM da, Vec u, Vec F, ObsCtx *user) {
     hx = 4.0 / (PetscReal)(info.mx - 1),
     hy = 4.0 / (PetscReal)(info.my - 1),
     detj = 0.25 * hx * hy;
-    PetscCall(DMDAVecGetArrayRead(da, u, &au));
+    PetscCall(DMGetLocalVector(da,&uloc));
+    PetscCall(DMGlobalToLocal(da,u,INSERT_VALUES,uloc));
+    PetscCall(DMDAVecGetArrayRead(da, uloc, &au));
     PetscCall(DMDAVecGetArray(da, F, &FF));
 
     // set up Q1 FEM tools for this grid
@@ -491,8 +495,9 @@ PetscErrorCode ApplyOperatorF(DM da, Vec u, Vec F, ObsCtx *user) {
         }
     }
 
-    PetscCall(DMDAVecRestoreArrayRead(da, u, &au));
+    PetscCall(DMDAVecRestoreArrayRead(da, uloc, &au));
     PetscCall(DMDAVecRestoreArray(da, F, &FF));
[…]

### 2022-10-14 c/src/utilities.c

@@ -14,29 +14,17 @@ PetscErrorCode VecPrintRange(Vec X, const char *name, const char *infcase) {
 }
 
 PetscErrorCode VecLessThanOrEqual(DM da, Vec u, Vec v, PetscBool *flg) {
-    PetscInt         i, j;
-    const PetscReal  **au, **av;
-    DMDALocalInfo info;
-    if ((!u) || (!v)) {
+    Vec       w;
+    PetscReal wmin;
+    if (u == NULL || v == NULL)
         *flg = PETSC_TRUE;
-        return 0;
+    else {
+        PetscCall(DMGetGlobalVector(da,&w));
+        PetscCall(VecWAXPY(w,-1.0,u,v));   // w = v - u  (should be nonnegative)
+        PetscCall(VecMin(w,NULL,&wmin));
+        PetscCall(DMRestoreGlobalVector(da,&w));
+        *flg = (wmin >= 0.0);
     }
-    PetscCall(DMDAGetLocalInfo(da,&info));
-    PetscCall(DMDAVecGetArrayRead(da, u, &au));
-    PetscCall(DMDAVecGetArrayRead(da, v, &av));
-    for (j=info.ys; j<info.ys+info.ym; j++) {
-        for (i=info.xs; i<info.xs+info.xm; i++) {
-            if (au[j][i] > av[j][i]) {
-                PetscCall(DMDAVecRestoreArrayRead(da, u, &au));
-                PetscCall(DMDAVecRestoreArrayRead(da, v, &av));
-                *flg = PETSC_FALSE;
-                return 0;
-            }
-        }
-    }
-    PetscCall(DMDAVecRestoreArrayRead(da, u, &au));
-    PetscCall(DMDAVecRestoreArrayRead(da, v, &av));
-    *flg = PETSC_TRUE;
     return 0;
 }
 

### 2022-10-14 c/nmcd.c
here is a bug: need to up-smooth on finest level to see any affect at all

@@ -229,7 +229,7 @@ int main(int argc,char **argv) {
     PetscCall(ProjectedNGS(&(levs[0].ldc),PETSC_TRUE,levs[0].ell,
                            levs[0].g,levs[0].z,&ctx));
     // upward direction
-    for (j=1; j<jtop; j++) {
+    for (j=1; j<=jtop; j++) {
         // compute z^j using prolongation:
         //   z^j = P z^j-1 + y^j
         PetscCall(Q1Interpolate(levs[j-1].dmda,levs[j].dmda,levs[j-1].z,&(levs[j].z)));

### 2022-10-14 c/src/ldc.h
edit comments to agree with current function signatures

@@ -27,12 +27,13 @@
 // ldc[0] is coarse while ldc[N] is finest:
 //
 //     LDC       ldc[N+1];
+//     DMDA      cdmda
 //     PetscInt  k;
-//     LDCCreateCoarsest(PETSC_TRUE,0,mx,my,xmin,xmax,ymin,ymax,&(ldc[0]));
+//     [  use DMDACreate2d() to create cdmda, and then fully configure it  ]
+//     LDCCreateCoarsest(PETSC_TRUE,cdmda,&(ldc[0]));
 //     for (k=0; k<N; k++)
 //         LDCRefine(ldc[k],&(ldc[k+1]));
-//     [  start a box-constrained solver on the finest level, and get a
-//        fine-level iterate w  ]
+//     [  a box-constrained solver on finest level returns fine-level iterate w  ]
 //     LDCFinestUpDCsFromVecs(w,vgamupp,vgamlow,&(ldc[N]));
 //     LDCGenerateDCsVCycle(&(ldc[N]));
 //     [  continue with the solver  ]

### 2022-10-14 c/nmcd.c
initialize from exact+bump; this version does something but V-cycles are not obviously converging

@@ -47,6 +47,12 @@ PetscReal gamma_lower(PetscReal x, PetscReal y, void *ctx) {
     }
 }
 
+// z = bump(x,y) is zero along boundary of Omega=(-2,2)^2 and reaches maximum
+// of bump(0,0) = 1.0 in center
+PetscReal bump(PetscReal x, PetscReal y, void *ctx) {
+    return (x + 2.0) * (2.0 - x) * (y + 2.0) * (2.0 - y) / 16.0;
+}
+
 /*  This exact solution solves a 1D radial free-boundary problem for the
 Laplace equation, on r >= 0, with hemispherical obstacle
 gamma_lower(r).  The Laplace equation applies where u(r) > gamma_lower(r),
@@ -82,11 +88,11 @@ int main(int argc,char **argv) {
                    gplusy, tmpfine, tmpcoarse;
     ObsCtx         ctx;
     DMDALocalInfo  finfo;
-    PetscInt       totlevs=2, cycles=1, jtop, j;
+    PetscInt       totlevs=2, cycles=1, csweeps=1, jtop, viter, j, k;
     PetscBool      ldcinfo = PETSC_FALSE, counts = PETSC_FALSE, view = PETSC_FALSE,
                    admis;
     PetscLogDouble lflops, flops;
-    PetscReal      errinf;
+    PetscReal      bumpsize=0.0, errinf;
     char           viewname[PETSC_MAX_PATH_LEN];
 
     PetscCall(PetscInitialize(&argc,&argv,NULL,help));
@@ -99,6 +105,10 @@ int main(int argc,char **argv) {
     ctx.residualcount = 0;
     ctx.ngscount = 0;
     PetscOptionsBegin(PETSC_COMM_WORLD,"nm_","obstacle problem NMCD solver options","");
+    PetscCall(PetscOptionsReal("-bumpsize","initialization from exact solution plus this much bump",
+                            "nmcd.c",bumpsize,&bumpsize,NULL));
+    PetscCall(PetscOptionsInt("-csweeps","number of PNGS sweeps used for coarse solver",
+                            "nmcd.c",csweeps,&csweeps,NULL));
     PetscCall(PetscOptionsBool("-counts","print counts for calls to call-back functions",
                             "nmcd.c",counts,&counts,NULL));
[…]

### 2022-10-15 c/nmcd.c
add to stdout, restore matlab view, and add a regression test

@@ -7,6 +7,13 @@ static char help[] =
 "solver is projected, nonlinear Gauss-Seidel (PNGS) sweeps.  Option prefix\n"
 "nm_.  Compare obstaclesl.c.\n\n";
 
+// FIXME possible ways to describe the current issues:
+//   * the iterate w is only slowly falling toward the obstacle
+//   * the corrections are always negative
+//   * the correction on the coarsest level is identically zero
+// see stdout and foo.m from
+//   $ ./nmcd -nm_monitor -nm_view foo.m -nm_levels 7 -nm_cycles 1 -nm_bumpsize 1.0
+
 #include <petsc.h>
 #include "src/utilities.h"
 #include "src/q1fem.h"
@@ -77,6 +84,20 @@ static PetscReal fg_zero(PetscReal x, PetscReal y, void *ctx) {
     return 0.0;
 }
 
+PetscErrorCode UpdateIndentPrintRange(Vec v, const char* name, PetscInt jtop, PetscInt j) {
+    char       strbuf[PETSC_MAX_PATH_LEN];
+    PetscInt   k;
+    PetscReal  vmin, vmax;
+    for (k = 0; k < jtop - j; k++)
+        PetscCall(PetscPrintf(PETSC_COMM_WORLD,"  "));
+    snprintf(strbuf,PETSC_MAX_PATH_LEN,"%s^%d",name,j);
+    PetscCall(VecMin(v,NULL,&vmin));
+    PetscCall(VecMax(v,NULL,&vmax));
+    PetscCall(PetscPrintf(PETSC_COMM_WORLD,"  %.4e <= %s <= %.4e\n",
+                          vmin,strbuf,vmax));
+    return 0;
+}
+
 extern PetscErrorCode AssertBoundaryAdmissible(DM da, ObsCtx*);
 extern PetscErrorCode FormExact(DM da, Vec, ObsCtx*);
 extern PetscErrorCode ApplyOperatorF(DM, Vec, Vec, ObsCtx*);
@@ -89,7 +110,8 @@ int main(int argc,char **argv) {
     ObsCtx         ctx;
     DMDALocalInfo  finfo;
     PetscInt       totlevs=2, cycles=1, csweeps=1, jtop, viter, j, k;
-    PetscBool      ldcinfo = PETSC_FALSE, counts = PETSC_FALSE, view = PETSC_FALSE,
+    PetscBool      ldcinfo = PETSC_FALSE, counts = PETSC_FALSE,
[…]

### 2022-10-15 paper/mcd2.tex
good edits through start section 5

@@
 
-Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below record basic CD iterative methods as functions which improve a current iterate $u \in \cK$ by solving ``smaller'' VI problems over each subset $\cK_i$.
-Their outputs $w\in\cK$ should be closer to the solution $u^* \in \cK$ of \eqref{eq:vi}.
+Associated to a constraint decomposition (CD) are certain basic iterative methods.
+Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below are functions which improve a current iterate $u \in \cK$ by solving smaller VI problems over each subset $\cK_i$ in the CD.
+The output $w\in\cK$ from either Algorithm should be closer to the solution $u^* \in \cK$ of \eqref{eq:vi}.
+Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} generalize the Jacobi and Gauss-Seidel iterations, respectively.
+
+In Algorithm \ref{alg:basiccd-add}, the additive or parallel CD method, ``\textbf{for all}'' can be computed in any order.
+Note that the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$ before adding-back the improved subset solution $\hat w_i \in \mathcal{K}_i$.
+In the multiplicative (successive) version, Algorithm \ref{alg:basiccd-mult}, each $\mathcal{K}_i$ solution corrects the global iterate immediately, and the $\mathcal{K}_i$ are addressed in a fixed order.
+The sum ``$\sum_{j>i} \Pi_j u$'' should be read as keeping the parts of the input $u$ which have not yet improved.
+Both Algorithms allow a damping parameter $0<\alpha\le 1$.
+In Algorithm \ref{alg:basiccd-mult} we must update the subset iterate $w_i$ using the same damping that is applied to the final output $w$.
 
@@
 
-In Algorithm \ref{alg:basiccd-add}, the additive or parallel CD method, the \textbf{for all} can be computed in any order.
-Note that $u-\Pi_iu+\hat w_i$ removes the part of $u$ which lies in $\mathcal{K}_i$ before adding-back the improved subset solution $\hat w_i$, also from $\mathcal{K}_i$.
-By contrast, in the following multiplicative (successive) Algorithm \ref{alg:basiccd-mult} each subset solution updates the global iterate before proceeding to the next subset.
-The sets $\mathcal{K}_i$ are addressed in a fixed order, and the sum $\sum_{j>i} \Pi_j u$ should be read as ``keep those parts of the input $u$ which we have not yet improved.''
-
-\begin{pseudofloat}[H]
+\begin{pseudofloat}
 \begin{pseudo*}
@@
 
-While Algorithm \ref{alg:basiccd-add} generalizes a Jacobi iteration, Algorithm \ref{alg:basiccd-mult} generalizes Gauss-Seidel.
-Note the use of a damping parameter $0<\alpha\le 1$ in both Algorithms.
-In Algorithm \ref{alg:basiccd-mult} we must update the in-progress iterate $w_i$ using the same damping that is applied to the final output $w$.
-
 In each boxed VI the argument of $f$ is an element of $\cK$, as the reader should confirm.
@@
 \end{align*}
+In this sense the boxed VIs are the same problem as the original VI \eqref{eq:vi}.
 
-For either Algorithm we define
+For either Algorithm we define the $i$th \emph{subset update} as
 \begin{equation}
-e_i = \hat w_i - \Pi_i u \in \cV_i, \label{eq:ithupdate}
+e_i = \hat w_i - \Pi_i u \in \cV_i. \label{eq:ithupdate}
 \end{equation}
-the $i$th \emph{subset update}.  As the reader may check, $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.  Note $\hat w = u^* + \sum_i \hat w_i - \Pi_i u^*$ also holds.
+The reader may check that $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.  Trivially, $\hat w = u^* + \sum_i \hat w_i - \Pi_i u^*$ also holds.
 
-We refer to Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} as the \emph{basic CD iterations}.
-These basic iterations are meaningful even when $f$ is nonlinear, non-local, or defined only on $\cK$.
-However, practical and efficient FE implementation for such general problems seems not to have been addressed in the literature, and will need additional notions.
-In particular, references \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem for which $f$ is linear, local, and defined for all inputs.
-The following example exposes the implementation simplifications available in that case.
+We will refer to Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} as the \emph{basic CD iterations}.
+These iterations are meaningful even when $f$ is nonlinear, non-local, or defined only on $\cK$.
+However, practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature, and doing so will require new notions.
+In particular, references \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem for which $f$ is linear, local, and defined for all inputs (over $\mathcal{V}$).
+The following example exposes the implementation simplifications available in that easier case.
 
 \begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear and defined on all of $\cV$.
-Furthermore suppose $f$ is a differential operator, thus local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral only over the support of $\phi \in \mathcal{V}$.
+Furthermore suppose $f$ is a differential operator, thus local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi \in \mathcal{V}$.
 Considering additive Algorithm \ref{alg:basiccd-add} for concreteness, by linearity the boxed VI can be written as
@@
 \end{equation}
-for all $v_i \in \mathcal{K_i}$, where $\tilde\ell = \ell - f(u)$.
+for all $v_i \in \mathcal{K}_i$, where $\tilde\ell = \ell - f(u)$.
 Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible ($e_i \notin \cK$), but VIs \eqref{eq:linearlocalvi} make sense because $f$ is defined over $\cV$.
-Each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $f(u)$, here included into a source term.
-If a hat function basis exists for $\cV_i$, namely of small-support functions, then by linearity of $f$ a solver for VI \eqref{eq:linearlocalvi} can be implemented using only integrals over the hat function supports.
+Each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $f(u)$, here included into the source term $\tilde\ell$.
+If a hat function basis exists for $\cV_i$, namely a basis of small-support functions, then by linearity of $f$ a solver for VI \eqref{eq:linearlocalvi} can be implemented using only integrals over the hat function supports; cf.~equations(2.31) and (2.32) in \cite{Farrelletal2021}.
 \end{example}
@@
 When $f$ has all the nice properties listed in Example \ref{ex:fnice} then a solver can take advantage of implementation efficiencies which are unavailable in general.
-The FE approximation of VI \eqref{eq:vi}---see Sections \ref{sec:multilevel} and \ref{sec:vcycle}---can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.
-The actual solution process for VI \eqref{eq:linearlocalvi} is not really the concern here; we are observing that the storage needed for problem \eqref{eq:linearlocalvi}, and the cost of residual evaluation, are smaller than in the general case.
-For more difficult problems certain additional ideas, illustrated in Section \ref{sec:vcycle}, are needed for practical implementation.
-Specifically, we will extend a multilevel CD iteration to a method for nonlinear functions $f:\cK\to\cV'$, i.e.~defined only on the constraint set $\cK$, by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
+An FE approximation of VI \eqref{eq:vi}---see Sections \ref{sec:multilevel} and \ref{sec:vcycle}---can then be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.
+The actual solution process for VI \eqref{eq:linearlocalvi} is not really the concern here; we are observing that the storage needed for problem \eqref{eq:linearlocalvi}, and the cost of residual evaluation, are substantially smaller than in the general case.
+For those general problems certain additional ideas, like those in Section \ref{sec:vcycle}, are needed for practical implementation.
+Specifically, we will construct a multilevel CD iteration for functions $f:\cK\to\cV'$, i.e.~defined only on the constraint set $\cK$, assuming neither linearity nor locality, by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
 
-One may also contrast Example \ref{ex:fnice} with the case of a non-local integral operator $f$ \cite{Bueler2021conservation}, where $\ip{f(\phi)}{z}$ requires an integral over $\Omega$ even if $\phi$ is a function with small support such as a hat function.
-Extension of the multilevel CD method to such non-local functionals $f$ is a topic for future research.
+In contrast with Example \ref{ex:fnice}, in the case of a non-local integral operator $f$ the expression $\ip{f(\phi)}{z}$ requires an integral over $\Omega$ even if $\phi$ is a function with small support (such as a hat function).
+While the algorithm proposed in Section \ref{sec:vcycle} applies as stated to such $f$, practical extension of multilevel CD methods to such non-local functionals $f$ is a topic for future research (cf.~subsection 4.5 in \cite{Bueler2021conservation}), and we will not demonstrate such a case here.
 
@@
 
-We will approximately solve VI \eqref{eq:vi} for certain nonlinear obstacle problems, over two-dimensional computational domains $\Omega \subset \RR^2$, and the Sobolev space $\mathcal{V}=W_0^{1,p}(\Omega)$, using piecewise-linear finite element (FE) spaces over triangular meshes.
+From now on we are concerned with the approximate solution of certain nonlinear obstacle problems over 2D computational domains.
+Suppose $\Omega \subset \RR^2$ is a (open) bounded polygon and that $\mathcal{V}=W^{1,p}(\Omega)$ is a Sobolev space of once-differentiable functions.
+
+We restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.
+Suppose \emph{lower} and \emph{upper obstacles} $\underline{\gamma} \le \overline{\gamma}$ are given measurable functions on $\overline{\Omega}$ which take their values in the extended real line $\hat\RR = [-\infty,+\infty]$.
+
+Furthermore we consider a Dirichlet (or mixed) boundary value problem.
+Suppose that the domain boundary is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive measure for simplicity, to permit well-posedness \cite{Evans2010}, and $\partial_N \Omega$ possibly empty.
+Suppose that the provided Dirichlet data $g_D:\partial_D \Omega \to \RR$ is sufficiently-regular to permit well-posedness, and that $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.
+
+Now define the constraint set
+\begin{equation}
+\cK = \left\{w \in \cV\,:\,\underline{\gamma} \le w \le \overline{\gamma} \,\text{ over }\, \Omega \, \text{ and }\, w\big|_{\partial_D \Omega} = g_D \, \text{ (trace sense \cite{Evans2010})}\right\} \subset \cV.
+\end{equation}
+It is easy to check that $\cK$ is closed and convex; note that the inequality and equality (boundary) constraints are treated in a unified manner by this definition of $\cK$.
+
+We will solve the same VI problem as the abstract problem \eqref{eq:vi}.  Given continuous $f:\cK \to \cV'$ and $\ell \in \cV'$, we assume that the following VI is well-posed for $u^*\in \cK$:
+\begin{equation}
+\ip{f(u^*)}{v-u^*} \ge \ip{\ell}{v-u^*} \qquad \text{for all } v\in \cK. \label{eq:boxdirichletvi}
+\end{equation}
+Note that $v-u^* \in W_0^{1,p}(\Omega)$, that is, the test functions in \eqref{eq:boxdirichletvi} have zero values on the Dirichlet boundary $\partial_D\Omega$.
+In general, problem \eqref{eq:boxdirichletvi} is a mixed boundary value problem with homogeneous Neumann, i.e.~natural, boundary conditions over $\partial_N \Omega = \Omega \setminus \partial_D \Omega$.
+
+In this Section we construct a multilevel finite element (FE) approximation to problem \eqref{eq:boxdirichletvi} using $P_1$ or $Q_1$ elements \cite{Elmanetal2014} and triangular or quadrilateral meshes, respectively, over $\Omega$.
 The targeted mesh will be the finest level in a hierarchy of meshes constructed by refinement of a coarsest mesh.
-In this Section we construct the corresponding multilevel CD.
-Its constraint sets are designed to allow multigrid coarser-level corrections (next Section) in a manner which does not violate the inequality constraints, and which also does not introduce unnecessary high frequencies into the coarser-level problems.
+This hierarchy generates a multilevel CD of $\cK$, as covered in previous Sections.
+The constraint subsets $\cK_i$  are designed to allow multigrid-type coarser-level corrections (next Section) in a manner which preserves admissibility while not introducing unnecessary high frequencies into the coarser-level problems, especially at the free boundary generated by the inequality constraints.
 
-The mesh hierarchy and FE spaces are constructed in a standard manner \cite{Elmanetal2014}.
+FIXME The mesh hierarchy and FE spaces are constructed in a standard manner \cite{Elmanetal2014}.
 Suppose $\Omega$ is polygonal and triangulated by $\mathcal{T}^0$, the \emph{coarsest mesh}, a finite set of non-overlapping triangles with union $\overline{\Omega}$ and characteristic size (e.g.~maximum edge length) $h_0>0$.

### 2022-10-15 paper/mcd2.tex
more good edits ... restart with FIXME in section 5

@@
 We restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.
-Suppose \emph{lower} and \emph{upper obstacles} $\underline{\gamma} \le \overline{\gamma}$ are given measurable functions on $\overline{\Omega}$ which take their values in the extended real line $\hat\RR = [-\infty,+\infty]$.
+Suppose \emph{lower} and \emph{upper obstacles} $\underline{\gamma} \le \overline{\gamma}$ are given measurable functions on $\overline{\Omega}$ which take their values in the extended real line $\tilde\RR = [-\infty,+\infty]$.
 
@@
 
-FIXME The mesh hierarchy and FE spaces are constructed in a standard manner \cite{Elmanetal2014}.
-Suppose $\Omega$ is polygonal and triangulated by $\mathcal{T}^0$, the \emph{coarsest mesh}, a finite set of non-overlapping triangles with union $\overline{\Omega}$ and characteristic size (e.g.~maximum edge length) $h_0>0$.
-Starting with $j=1$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ by edge bisection, thus each $\mathcal{T}^{j-1}$ triangle becomes four similar triangles in $\mathcal{T}^j$ with characteristic size $h_j = 2^{-j} h_0$.
-Refinement proceeds to level $J\ge 0$, the \emph{depth} of the hierarchy, giving $\mathcal{T}^J$ as the \emph{finest mesh}.
-Let $m_j$ be the number of interior nodes in $\mathcal{T}^j$, the \emph{degrees of freedom} at the $j$th level, and denote $h=h_J=2^{-J} h_0$ and $m=m_J$ for the finest level.
-Now let $\mathcal{V}^j$ be the continuous, piecewise-linear FE space $P_1$ \cite{Elmanetal2014} over $\mathcal{T}^j$ with zero boundary values, and observe the following nesting:
+The mesh hierarchy and FE spaces are constructed by nested refinement in the following standard manner \cite{Elmanetal2014}.
+Suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate elements (triangles or quadrilaterals) with union equal to $\overline{\Omega}$ and characteristic size (maximum edge length) $h_0>0$.
+For $j=1,\dots,J$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ by connecting the midpoints of edges.
+Each $\mathcal{T}^{j-1}$ element therefore becomes four similar elements in $\mathcal{T}^j$ with characteristic size $h_j = 2^{-j} h_0$.\footnote{For arbitrary triangles, and for rectangles, $h_j=2^{-j}h_0$ is the exact maximum edge length in $\mathcal{T}^j$.
+For general quadrilaterals this is proportional \cite{Zhang2004}.
+Only rectangular quadrilaterals are used in the Results section.}  We call $J\ge 0$ the \emph{depth} of the hierarchy and $\mathcal{T}^J$ the \emph{finest mesh}.
+Let $m_j$ be the number of nodes in $\mathcal{T}^j$, the \emph{degrees of freedom} at the $j$th level, and denote $h=h_J=2^{-J} h_0$ and $m=m_J$ for the finest level.
+Now let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space \cite{Elmanetal2014} over $\mathcal{T}^j$, with arbitrary boundary values, and observe the following vector space nesting:
 \begin{equation}
-\mathcal{V}^0 \subset \mathcal{V}^1 \subset \dots \subset \mathcal{V}^J \subset \mathcal{V}.  \label{eq:fe:nestedspaces}
+\mathcal{V}^0 \subset \mathcal{V}^1 \subset \dots \subset \mathcal{V}^J \subset \mathcal{V}=W^{1,p}(\Omega).  \label{eq:fe:nestedspaces}
 \end{equation}
 
-In the above subspace construction we have chosen triangles and the $P_1$ FE space.
-Quadrilaterals and the $Q_1$ FE space will work just as well in what follows.
-However, use of higher-order Lagrange elements $P_k$ and $Q_k$ would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} specific to inequality constraints in FE schemes.
-Our FE construction of constraint sets requires a property we call \emph{nodal monotonicity}, namely that for all FE functions $w,z \in \mathcal{V}^j$,
+In the following Sections, use of higher-order Lagrange elements $P_k$ or $Q_k$, $k\ge 2$, causes a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} specific to inequality constraints, and for this reason we restrict to $k\le 1$.
+Specifically, our FE construction of constraint sets will require \emph{nodal monotonicity}, namely that for all FE functions $w,z \in \mathcal{V}^j$,
 \begin{equation}
@@
 \end{equation}
-where $\bw,\bz$ denote the vectors of nodal values.
-For $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal value one at a single node, and otherwise zero nodal values, which take on negative values \cite[Figure 1.7]{Elmanetal2014}, whereas this does not occur for $k\le 1$.
-It is easy to show that \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of such sometimes-negative hat functions.
+where $\bw,\bz$ denote vectors of nodal values.
+For $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal value one at a single node, otherwise zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.
+It is easy to show that \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of sign-changing hat functions, and this implies $k\le 1$.
 
-Next we define ``box constraints'' \cite{BensonMunson2006}.
-Let $\tilde{\mathcal{V}}^j$ denote the set of functions on the $\mathcal{T}^j$ nodes with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
-At the finest-level, suppose that the \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ satisfy
+To allow box constraints in our multilevel FE scheme, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the nodes of $\mathcal{T}^j$, with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
+(Note that $\tilde{\mathcal{V}}^j$ functions are not in $\mathcal{V}^j$, and they are not defined everywhere on $\Omega$.)  On the finest-level mesh, let $g_D^J \in \mathcal{V}^J$ satisfy $g_D^J = g_D$ on $\partial_D \Omega$.\footnote{As is standard in conforming FE methods, this restricts the original Dirichlet boundary values $g_D$ to a piecewise-linear function \cite{Elmanetal2014}.}  Suppose that the \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ satisfy
 \begin{equation}
-\underline{\gamma}^J \le \overline{\gamma}^J, \qquad \underline{\gamma}^J < +\infty, \qquad -\infty < \overline{\gamma}^J, \qquad \text{and} \qquad \underline{\gamma}^J|_{\partial\Omega} \le 0 \le \overline{\gamma}^J|_{\partial\Omega}.
-\label{eq:fe:boxconstraintrequirements}
+\underline{\gamma}^J \le \overline{\gamma}^J, \qquad \underline{\gamma}^J < +\infty, \qquad -\infty < \overline{\gamma}^J \label{eq:fe:boxconstraintrequirements}
 \end{equation}
-Where $\underline{\gamma}^J=-\infty$ there will no lower constraint, and similarly where $\overline{\gamma}^J=+\infty$ no upper constraint.  Using nodal monotonicity, define
+on the nodes of $\mathcal{T}^J$ and $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.
+On nodes where $\underline{\gamma}^J=-\infty$ there is no lower constraint, and similarly where $\overline{\gamma}^J=+\infty$ no upper constraint.
+
+Define
 \begin{equation}
-\mathcal{K}^J = \left\{v \in \mathcal{V}^J\,:\,\underline{\gamma}^J \le v \le \overline{\gamma}^J\right\}, \label{eq:fineconstraintset}
+\mathcal{K}^J = \left\{w \in \mathcal{V}^J\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \, \text{ and } \, w|_{\partial_D\Omega} = g_D|_{\partial_D\Omega}\right\}, \label{eq:fe:fineconstraintset}
 \end{equation}
-the finest-level \emph{constraint set} of \emph{admissible} FE functions, a nonempty, closed, and convex subset of $\mathcal{V}^J$.
-As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J = \pm \infty$.
+as the finest-level \emph{constraint set} of \emph{admissible} FE functions.
+This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.
+As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fe:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J = \pm \infty$.
 
-Let $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ be a discretization of $f$ in \eqref{eq:vi}.
-Noting that the functional $\ell$ in \eqref{eq:vi} is already defined and continuous over $\mathcal{V}^J$, we denote $\ell^J \in (\mathcal{V}^J)'$ for its FE representation.
-We seek the finest-level FE solution $u^J \in \mathcal{K}^J$ of the finite-dimensional, nonlinear VI
+Let $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ be a discretization of $f$ in \eqref{eq:boxdirichletvi}.
+Noting that by inclusion \eqref{eq:fe:nestedspaces} the functional $\ell$ is already defined and continuous over $\mathcal{V}^J$, we denote $\ell^J \in (\mathcal{V}^J)'$ for its FE representation.
+We seek the FE solution $u^J \in \mathcal{K}^J$ of the finest-level, finite-dimensional, nonlinear VI
 \begin{equation}
@@
 
-While \eqref{eq:fe:vi} approximates \eqref{eq:vi}, it may not technically be conforming for \eqref{eq:vi}.
+While \eqref{eq:fe:vi} approximates \eqref{eq:boxdirichletvi}, it may not technically be conforming for that VI.
 That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require $\mathcal{K}^J \subset \mathcal{K}$.
-The continuum constraint set $\mathcal{K}$ remains abstract in applications where only the fine-level FE obstacles $\underline{\gamma}^J, \overline{\gamma}^J$ are defined.
-In fact, even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants might not imply continuum admissiblity.
-These caveats specifically apply in geophysical situations where the obstacle is given by meshed elevation data, the continuum limit of which is undefined \cite{Bueler2016}.
+The continuum constraint set $\mathcal{K}$ remains abstract in applications where only the nodal obstacles $\underline{\gamma}^J, \overline{\gamma}^J$ are defined.
+In fact, even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants might not imply continuum admissiblity.\footnote{Such caveats apply in geophysical situations where the obstacle is given by meshed elevation data, the continuum limit of which is undefined, for example \cite{Bueler2016}.}
 
 The multilevel algorithm in the next section, a solver for VI problem \eqref{eq:fe:vi}, requires transfer operators between meshes.
-We use five such operators, three of which are standard, linear maps used for multigrid methods \cite{Trottenbergetal2001}.
+We use five such operators, three of which are standard linear maps used for multigrid methods \cite{Trottenbergetal2001}.
 Table \ref{tab:transfers} will help the reader keep track.
@@
 Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ also acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $(R\sigma)[z] = \sigma[z]$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
-While these canonical transfers do not lose information, they typically involve a change of FE representation.
-For example, the practical formula for canonical restriction, derived by writing coarser-mesh hat functions in terms of finer-mesh ones, is a ``full-weighting'' average \cite{Trottenbergetal2001}, and its transpose is canonical prolongation.
+While these canonical transfers do not lose information, they typically do involve a change of FE representation.
+For example, canonical restriction, derived by writing coarser-mesh hat functions in terms of finer-mesh ones, is a ``full-weighting'' average \cite{Trottenbergetal2001}.
 
-To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, observe that for any given values at the (interior) nodes of $\mathcal{T}^j$ there is a unique function in $\mathcal{V}^j$ with those values.
+To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, observe that for any given values at the nodes of $\mathcal{T}^j$ there is a unique function in $\mathcal{V}^j$ with those values.
 Also observe that, by our refinement construction, the nodes of $\mathcal{T}^{j-1}$ are nodes of $\mathcal{T}^j$.
@@
 
-Finally, the monotone restrictions $\maxR,\minR$ are nonlinear maps which maximize and minimize, respectively, nodal values over the supports of coarser-mesh hat functions.
-To make the definitions precise, first denote the \emph{nodes} of triangulation $\mathcal{T}^j$ by $x_q^j$ for $q=0,\dots,\tilde{m}_j-1$, and then the corresponding hat functions \cite{Elmanetal2014} by $\psi_q^j$.
+FIXME READ FROM HERE Finally, the monotone restrictions $\maxR,\minR$ are nonlinear maps which maximize and minimize, respectively, nodal values over the supports of coarser-mesh hat functions.
+To make the definitions precise, first denote the \emph{nodes} of mesh $\mathcal{T}^j$ by $x_q^j$ for $q=0,\dots,m_j-1$, and then the corresponding hat functions \cite{Elmanetal2014} by $\psi_q^j$.
 (Boundary nodes are included here.)  The monotone restrictions of $z\in\tilde{\mathcal{V}}^j$ are

### 2022-10-16 paper/mcd2.tex
a few more edits

@@
 Note that $v-u^* \in W_0^{1,p}(\Omega)$, that is, the test functions in \eqref{eq:boxdirichletvi} have zero values on the Dirichlet boundary $\partial_D\Omega$.
-In general, problem \eqref{eq:boxdirichletvi} is a mixed boundary value problem with homogeneous Neumann, i.e.~natural, boundary conditions over $\partial_N \Omega = \Omega \setminus \partial_D \Omega$.
+
+Problem \eqref{eq:boxdirichletvi} is a mixed boundary value problem, in weak form, with Dirichlet conditions $u=g_D$ over $\partial_D\Omega$ and Neumann boundary conditions over $\partial_N \Omega = \Omega \setminus \partial_D \Omega$.
+The Neumann boundary conditions are determined by the formula for $f$.
+(That is, nonhomogeneous Neumann conditions imply that the nonlinear operator $f:\cK\to\cV'$ involves an integration over $\partial_N\Omega$.)
 
@@
 The continuum constraint set $\mathcal{K}$ remains abstract in applications where only the nodal obstacles $\underline{\gamma}^J, \overline{\gamma}^J$ are defined.
-In fact, even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants might not imply continuum admissiblity.\footnote{Such caveats apply in geophysical situations where the obstacle is given by meshed elevation data, the continuum limit of which is undefined, for example \cite{Bueler2016}.}
+In fact, even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants might not imply continuum admissiblity.\footnote{Such caveats apply in geophysical situations where the obstacle is given by meshed elevation data, the continuum limit of which is undefined \cite{Bueler2016}.}
 
@@
 
-FIXME READ FROM HERE Finally, the monotone restrictions $\maxR,\minR$ are nonlinear maps which maximize and minimize, respectively, nodal values over the supports of coarser-mesh hat functions.
-To make the definitions precise, first denote the \emph{nodes} of mesh $\mathcal{T}^j$ by $x_q^j$ for $q=0,\dots,m_j-1$, and then the corresponding hat functions \cite{Elmanetal2014} by $\psi_q^j$.
-(Boundary nodes are included here.)  The monotone restrictions of $z\in\tilde{\mathcal{V}}^j$ are
+Finally, the monotone restrictions $\maxR,\minR$ are nonlinear maps which maximize and minimize, respectively, nodal values over the supports of coarser-mesh hat functions.
+To make the definitions precise, first denote the \emph{nodes} of mesh $\mathcal{T}^j$ by $x_q^j$ for $q=0,1,\dots,m_j-1$, and then the corresponding hat functions \cite{Elmanetal2014} by $\psi_q^j(x)$.
+The monotone restrictions $\maxR z,\minR z$ of $z\in\tilde{\mathcal{V}}^j$ are the coarser-level functions in $\tilde{\mathcal{V}}^{j-1}$ defined by
 \begin{align}
-\maxR z &= \sum_{p=1}^{\tilde m_{j-1}} \max \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}\, \psi_p^{j-1}, \label{eq:fe:monotonerestrictions} \\
-\minR z &= \sum_{p=1}^{\tilde m_{j-1}} \min \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}\, \psi_p^{j-1}, \notag
+(\maxR z)(x_p^{j-1}) &= \max \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}, \label{eq:fe:monotonerestrictions} \\
+(\minR z)(x_p^{j-1}) &= \min \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}, \notag
 \end{align}
-with the result in $\tilde{\mathcal{V}}^{j-1}$.
-The usual ordering rules for the extended real line $\tilde\RR$ are applied.
-Note that if $z\in\tilde{\mathcal{V}}^j$ then $\minR z \le z \le \maxR z$.
+for $p=0,1,\dots,m_{j-1}-1$.
+In computing $\max$ and $\min$, the usual ordering rules for the extended real line $\tilde\RR$ are applied.
+Note that $\minR z \le z \le \maxR z$ because, by mesh refinement, node $x_p^{j-1}$ is in the set of $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$.
 
-To define the constraint subsets, the sets generically denoted $\cK_i$ in Section \ref{sec:cd}, in the current multilevel FE context, suppose $w^J \in \cK^J$ is an admissible fine-level iterate which approximates the solution $u^J$ of VI \eqref{eq:fe:vi}.
+To define the constraint subsets in the current multilevel FE context, namely the sets denoted $\cK_i$ in Section \ref{sec:cd}, suppose $w^J \in \cK^J$ is an admissible finest-level iterate which approximates the solution $u^J$ of VI \eqref{eq:fe:vi}.
 From this iterate we define the finest-level \emph{lower and upper defect constraints} \cite{GraeserKornhuber2009}
@@
 functions in $\tilde{\mathcal{V}}^J$ satisfying $\underline{\chi}^J \le 0 \le \overline{\chi}^J$.
-The motivation here is that a corrected iterate is admissible, $\underline{\gamma}^J \le w^J + z^J \le \overline{\gamma}^J$, if and only if the perturbation $z^J$ is between the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$.
+FIXME ADD DIRICHLET The motivation here is that a corrected iterate is admissible, $\underline{\gamma}^J \le w^J + z^J \le \overline{\gamma}^J$, if and only if the perturbation $z^J$ is between the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$.
 
@@
 \end{equation}
-A $J=4$ example, with no upper obstacles ($\overline{\chi}^j=+\infty$), is shown in Figure \ref{fig:chiphilevels}.
+A $J=4$ example, with no upper obstacles, namely $\overline{\chi}^j=+\infty$ for all $j$, is shown in Figure \ref{fig:chiphilevels}.
 
@@
 \end{equation}
-using $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$.  Again these constraints bracket zero,
+using $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ to define $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.  These constraints bracket zero,
 \begin{equation}
-\underline{\phi}^j \le 0 \le \overline{\phi}^j,  \label{eq:fe:phibrackets}
+\underline{\phi}^j \le 0 \le \overline{\phi}^j.  \label{eq:fe:phibrackets}
 \end{equation}
-but these functions are not ordered as in \eqref{eq:fe:chiordering}, but telescoping sums hold for $j=0,1,\dots,J$:
+Though these functions are not ordered as in \eqref{eq:fe:chiordering}, telescoping sums hold for $j=0,1,\dots,J$:
 \begin{equation}
@@
 
-A multilevel V-cycle is defined in the next Section.
+FIXME A multilevel V-cycle is defined in the next Section.
 For the downward direction we will use the differences $\underline{\phi}^j,\overline{\phi}^j$ as constraints, and for the upward part the defect constraints $\underline{\chi}^j,\overline{\chi}^j$.

### 2022-10-18 paper/mcd2.tex
clean existing text, through page 11

@@
 
-We extend the constraint decomposition (CD) method of X.-C.~Tai \cite{Tai2003} to nonlinear variational inequality (VI) problems to which it has not been applied.
+We extend the constraint decomposition (CD) method of X.-C.~Tai \cite{Tai2003} by implementing fully-nonlinear corrections, and we apply it to nonlinear variational inequality (VI) problems to which it has not been applied.
 In its multilevel form the method has been shown to have optimal complexity for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
@@
 
-The multilevel CD algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} are extended here in three practical ways:
+The multilevel CD algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} are extended here, for finite element (FE) solvers, in three practical ways:
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
 \begin{enumerate}
-\item We make finite element (FE) implementations of multilevel CD algorithm more effective for nonlinear problems by applying a full approximation storage (FAS; see \cite{Brandt1977,Bruneetal2015}) approach.
+\item We apply a full approximation storage (FAS; see \cite{Brandt1977,Bruneetal2015}) approach.
 Section \ref{sec:results} shows results from this \emph{nonlinear multilevel constraint decomposition} (NMCD) algorithm  for several nonlinear problems.
@@
 \item In Sections \ref{sec:multilevel} and \ref{sec:vcycle} we explain why, given the manner in which constraint sets are decomposed in the multilevel context \cite{GraeserKornhuber2009}, ``up-smoothing'' is more efficient than ``down-smoothing''.
-This observation seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}.
-A strong preference for up-smoothing is special to multilevel CD methods and does not arise in unconstrained problems.
+A strong preference for up-smoothing is special to multilevel CD methods and does not arise in unconstrained problems.\footnote{This observation seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}.}
 \end{enumerate}
 
-The iterates from a CD method are always admissible, and thus the operator need only be defined for admissible states.
-This property should permit direct solutions of certain VI problems, such as fluid-layer dynamics problems \cite{Bueler2021conservation,JouvetBueler2012}, for which non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
-Our NMCD scheme is designed for such problems, some of which require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation, as discussed in \cite{Bueler2021conservation}, but we defer this application to future work.
+The iterates from a CD method are always admissible, and thus the nonlinear operator in the VI need only be defined for admissible states.
+Our NMCD scheme is designed for such admissible-only problems, some of which require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation, as discussed in \cite{Bueler2021conservation}.
+NMCD methods should permit direct solutions of those VI problems for which non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
 
-One of our examples demonstrates the application of the NMCD method to a problem of porous-media type, a nonlinear operator which is not known to be coercive.
-By ``freezing' the solution-dependent coefficient, the operator is approximated by a coercive operator for the duration of the V-cycle, and this scheme is also effective for doubly-nonlinear diffusion operators.
-See Section \ref{sec:results}.
+One of our examples (Section \ref{sec:results}) demonstrates the application of the NMCD method to a problem of porous-media type, a nonlinear operator which is not known to be coercive.
+By ``freezing' the solution-dependent coefficient, the operator is approximated by a coercive operator for the duration of the V-cycle.
+This scheme is also effective for doubly-nonlinear diffusion operators.
 
-For the classical obstacle problem with a Laplacian operator, certain multilevel techniques are known to improve performance relative to the multilevel CD method \cite{GraeserKornhuber2009}, such as the truncated monotone multigrid method \cite{Kornhuber1994}.
+For the classical obstacle problem with a Laplacian operator, certain multilevel techniques, such as the truncated monotone multigrid method \cite{Kornhuber1994}, are known to improve performance relative to the original multilevel CD method from \cite{Tai2003}; see also \cite{GraeserKornhuber2009}.
 These improved methods either track the active set in the discretization or modify the nodal basis functions, and in this sense they are discrete algorithms, while by contrast the CD method applies essentially at the level of the continuum problem; see Section \ref{sec:cd}.
@@
 Let $\cK \subset \cV$ be a nonempty closed and convex subset, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.
-For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and a linear \emph{source functional} $\ell\in \cV'$ we consider the following \emph{variational inequality} (VI) for the (exact) solution $u^*\in \cK$, if it exists:
+For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and a linear \emph{source functional} $\ell\in \cV'$ we consider the following \emph{variational inequality} (VI) problem for the (exact) solution $u^*\in \cK$, if it exists:
 \begin{equation}
@@
 VI \eqref{eq:vi} generalizes the nonlinear system of equations $f(u^*)=\ell$ to problems where $u^*$ is also required to be in a constraint set $\cK$.
-Informally, if we conceptualize the dual pairing as an inner product then \eqref{eq:vi} says that the angle between $f(u^*)-\ell$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
+Informally, if we conceptualize the dual pairing as an inner product, then \eqref{eq:vi} says that the angle between $f(u^*)-\ell$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
 That is, while $f(u^*)-\ell$ may not be zero, it must point directly into $\cK$.
-On the other hand, if $u^*$ is in the interior of the constraint set, $u^*\in\cK^\circ$, then \eqref{eq:vi} implies $f(u^*)=\ell$.
+On the other hand, if $u^*$ is in the interior of the constraint set, i.e.~$u^*\in\cK^\circ$, then \eqref{eq:vi} implies $f(u^*)=\ell$.
 
@@
 
-It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, uniquely when $f$ is strictly monotone.
+It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, and that the solution is unique when $f$ is strictly monotone.
 As in the calculus of variations \cite{Evans2010}, coercivity permits a compactness argument for unbounded sets $\cK$.
-(Recall that the bounded, closed subsets of a reflexive Banach space are weakly compact.)  The condition of continuity can be weakened to only apply on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but the stronger condition will apply in our examples.
+(Recall that the closed and bounded subsets of a reflexive Banach space are weakly compact.)  The condition of continuity can be weakened, e.g.~to only apply on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but the stronger condition will apply in our examples.
 
-The coercive VIs solved in this paper satisfy a stronger inequality than \eqref{eq:monotone} and \eqref{eq:coercive}.
+The coercive VIs solved in this paper satisfy a stronger inequality than \eqref{eq:coercive}.
 
@@
 
-It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive.
+It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive, so the following well-posedness theorem holds.
 
@@
 Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
-Furthermore, Proposition II.2.1 in \cite{EkelandTemam1976} shows that if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ and $\ell=0$ if and only if
+Furthermore, Proposition II.2.1 in \cite{EkelandTemam1976} shows that if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ if and only if
 \begin{equation}
-u^* = \operatorname{arg-min}_{v\in\cK} F(v). \label{eq:minimization}
+u^* = \operatorname{arg-min}_{v\in\cK} F(v) - \ip{\ell}{v}. \label{eq:minimization}
 \end{equation}
-The CD methods of Tai \cite{Tai2003} address problem \eqref{eq:minimization} under the hypothesis that $F'$ is 2-coercive.
+The CD methods of Tai \cite{Tai2003} address problem \eqref{eq:minimization} under the hypotheses that such an objective $F$ exists, and that $F'$ is 2-coercive.
 (Note that Tai \cite{Tai2003} uses ``coercive'' for what we call $2$-coercive.)
 
-FIXME the problems we actually address are partial differential operators, and furthermore $\mathcal{K}$ incorporates both Dirichlet boundary conditions, as equality constraints, and inequality constraints
+The problems addressed later in Section \ref{sec:results} involve nonlinear and linear operators $f$ which are partial differential operators.
+As described in section \ref{sec:multilevel}, the constraint sets $\mathcal{K}$ incorporate both inequality box constraints and nonhomogeneous Dirichlet boundary conditions as equality constraints.
+However, the following examples use homogeneous (zero) Dirichlet conditions for ease of presentation.
 
-From now on let $\Omega \subset \RR^d$ denote a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.
+Let $\Omega \subset \RR^d$ denote a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.
 Sobolev spaces \cite{Evans2010} are denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$.
@@
 However, not all VI problems arise from optimization.
-We give two such examples next, first a coercive and linear advection-diffusion problem, and then a nonlinear porous-medium-type problem; each is important in applications.
+We give two such examples next, first a coercive and linear advection-diffusion problem, and then a nonlinear porous-medium-type problem.
 The first is preceded by a lemma.
@@
 
-In Sections \ref{sec:multilevel} and \ref{sec:vcycle} we will introduce discretizations and a practical algorithms for finite-dimensional VI problems, but observe that this CD concept applies at the level of the continuum problem.
-The following two obstacle problem examples illustrate this.
+In Sections \ref{sec:multilevel} and \ref{sec:vcycle} we will introduce discretizations and practical algorithms for finite-dimensional VI problems.
+However, one can observe that the CD concept applies at the level of the continuum problem, and the following two obstacle problem examples are illustrations.
 The first is an overlapping domain decomposition.
@@
 
-Our second example is a disjoint frequency decomposition.  A multilevel CD, specifically the one proposed in Section \ref{sec:multilevel}, approximates such a frequency decomposition.
+Our second example is a disjoint frequency decomposition.
+A multilevel CD over a finite-dimensional FE discretization, specifically the one proposed in Section \ref{sec:multilevel}, approximates such a frequency decomposition.
 
@@
 
-Again note that $\cK_i \not\subset \cK$ in these cases; the important inclusion is $\cK_i \subset \cV_i$.
-Specifically in Example \ref{ex:domaindecomposition}, if $\psi$ is positive over portions of $\Omega$ where the decomposition into overlapping subdomains $\Omega_i$ is nontrivial, then $\cK_i \not\subset \cK$, and something similar holds for Example \ref{ex:frequencydecomposition}.
+Again note that $\cK_i \not\subset \cK$ in these cases.\footnote{Informally speaking, the important inclusion is $\cK_i \subset \cV_i$.}  In Example \ref{ex:domaindecomposition}, if $\psi$ is positive over portions of $\Omega$ where the decomposition into overlapping subdomains $\Omega_i$ is nontrivial, then $\cK_i \not\subset \cK$, and something similar holds for Example \ref{ex:frequencydecomposition}.
 
@@
 Associated to a constraint decomposition (CD) are certain basic iterative methods.
-Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below are functions which improve a current iterate $u \in \cK$ by solving smaller VI problems over each subset $\cK_i$ in the CD.
-The output $w\in\cK$ from either Algorithm should be closer to the solution $u^* \in \cK$ of \eqref{eq:vi}.
-Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} generalize the Jacobi and Gauss-Seidel iterations, respectively.
+Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below are functions which improve a current iterate $u \in \cK$, generating output $w\in\cK$ which should be closer to the solution $u^* \in \cK$ of \eqref{eq:vi}, by solving smaller VI problems over each CD subset $\cK_i$.
+The two Algorithms each permit a damping parameter $0<\alpha\le 1$, and they generalize the Jacobi and Gauss-Seidel iterations, respectively.
 
@@
 Note that the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$ before adding-back the improved subset solution $\hat w_i \in \mathcal{K}_i$.
-In the multiplicative (successive) version, Algorithm \ref{alg:basiccd-mult}, each $\mathcal{K}_i$ solution corrects the global iterate immediately, and the $\mathcal{K}_i$ are addressed in a fixed order.
-The sum ``$\sum_{j>i} \Pi_j u$'' should be read as keeping the parts of the input $u$ which have not yet improved.
-Both Algorithms allow a damping parameter $0<\alpha\le 1$.
-In Algorithm \ref{alg:basiccd-mult} we must update the subset iterate $w_i$ using the same damping that is applied to the final output $w$.
 
@@
 \end{pseudofloat}
+
+In the multiplicative (successive) version, Algorithm \ref{alg:basiccd-mult}, each $\mathcal{K}_i$ solution corrects the global iterate immediately, and the $\mathcal{K}_i$ are addressed in a fixed order.
+The sum ``$\sum_{j>i} \Pi_j u$'' should be read as keeping the parts of the input $u$ which have not yet improved.
+Note we must update the subset iterate $w_i$ using the same damping that is applied to the final output $w$.
 
@@
 \end{align*}
-In this sense the boxed VIs are the same problem as the original VI \eqref{eq:vi}.
+In this sense the boxed VIs are for the same problem as the original VI \eqref{eq:vi}.
 
@@
 However, practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature, and doing so will require new notions.
-In particular, references \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem for which $f$ is linear, local, and defined for all inputs (over $\mathcal{V}$).
+In particular, references \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem for which $f$ is linear, local, and defined for all inputs (i.e.~$f$ is defined on all of $\mathcal{V}$).
 The following example exposes the implementation simplifications available in that easier case.
 
-\begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear and defined on all of $\cV$.
+\begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear.
 Furthermore suppose $f$ is a differential operator, thus local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi \in \mathcal{V}$.
@@
 for all $v_i \in \mathcal{K}_i$, where $\tilde\ell = \ell - f(u)$.
-Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible ($e_i \notin \cK$), but VIs \eqref{eq:linearlocalvi} make sense because $f$ is defined over $\cV$.
-Each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $f(u)$, here included into the source term $\tilde\ell$.
-If a hat function basis exists for $\cV_i$, namely a basis of small-support functions, then by linearity of $f$ a solver for VI \eqref{eq:linearlocalvi} can be implemented using only integrals over the hat function supports; cf.~equations(2.31) and (2.32) in \cite{Farrelletal2021}.
+Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible ($e_i \notin \cK$), but \eqref{eq:linearlocalvi} makes sense because $f$ is defined over $\cV$.
+Each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $f(u)$, now included into $\tilde\ell$.
+If a hat function basis exists for $\cV_i$, namely a basis of small-support functions, then by linearity of $f$ a solver for VI \eqref{eq:linearlocalvi} can be implemented using only integrals over the hat function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}
 \end{example}
@@
 The actual solution process for VI \eqref{eq:linearlocalvi} is not really the concern here; we are observing that the storage needed for problem \eqref{eq:linearlocalvi}, and the cost of residual evaluation, are substantially smaller than in the general case.
-For those general problems certain additional ideas, like those in Section \ref{sec:vcycle}, are needed for practical implementation.
-Specifically, we will construct a multilevel CD iteration for functions $f:\cK\to\cV'$, i.e.~defined only on the constraint set $\cK$, assuming neither linearity nor locality, by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
 
-In contrast with Example \ref{ex:fnice}, in the case of a non-local integral operator $f$ the expression $\ip{f(\phi)}{z}$ requires an integral over $\Omega$ even if $\phi$ is a function with small support (such as a hat function).
-While the algorithm proposed in Section \ref{sec:vcycle} applies as stated to such $f$, practical extension of multilevel CD methods to such non-local functionals $f$ is a topic for future research (cf.~subsection 4.5 in \cite{Bueler2021conservation}), and we will not demonstrate such a case here.
+For general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation.
+Section \ref{sec:vcycle} constructs a multilevel CD iteration for functions $f:\cK\to\cV'$, i.e.~defined only on the constraint set, and assuming neither linearity nor locality, by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
+
+Specifically for a non-local integral operator $f$, the expression $\ip{f(\phi)}{z}$ requires an integral over $\Omega$ even if $\phi$ is a function with small support (such as a hat function); cf.~subsection 4.5 in \cite{Bueler2021conservation}.
+While Algorithm \ref{alg:nmcd} applies, as stated, to such $f$, practical extension of multilevel CD methods to non-local operators is a topic for future research.
 
@@
 
-From now on we are concerned with the approximate solution of certain nonlinear obstacle problems over 2D computational domains.
+From now on we are concerned with the approximate solution of certain nonlinear VI problems over 2D computational domains.
 Suppose $\Omega \subset \RR^2$ is a (open) bounded polygon and that $\mathcal{V}=W^{1,p}(\Omega)$ is a Sobolev space of once-differentiable functions.
+Though the free boundaries are defined by inequality constraints (below), fixed boundary conditions may be Dirichlet or mixed.
+Suppose that the domain boundary is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive measure for simplicity, to permit well-posedness \cite{Evans2010}, and $\partial_N \Omega$ possibly empty.
+Suppose that the provided Dirichlet data $g_D:\partial_D \Omega \to \RR$ is sufficiently-regular to permit well-posedness of the VI problem stated below.
 
 We restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.
-Suppose \emph{lower} and \emph{upper obstacles} $\underline{\gamma} \le \overline{\gamma}$ are given measurable functions on $\overline{\Omega}$ which take their values in the extended real line $\tilde\RR = [-\infty,+\infty]$.
-
-Furthermore we consider a Dirichlet (or mixed) boundary value problem.
-Suppose that the domain boundary is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive measure for simplicity, to permit well-posedness \cite{Evans2010}, and $\partial_N \Omega$ possibly empty.
-Suppose that the provided Dirichlet data $g_D:\partial_D \Omega \to \RR$ is sufficiently-regular to permit well-posedness, and that $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.
+Suppose that the \emph{lower obstacle} $\underline{\gamma}$ and the \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions on $\overline{\Omega}$ which take their values in the extended real line $\tilde\RR = [-\infty,+\infty]$.
+Specifically suppose $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$ and $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$ are measurable, satisfying $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$ and $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.
 
@@
 \begin{equation}
-\cK = \left\{w \in \cV\,:\,\underline{\gamma} \le w \le \overline{\gamma} \,\text{ over }\, \Omega \, \text{ and }\, w\big|_{\partial_D \Omega} = g_D \, \text{ (trace sense \cite{Evans2010})}\right\} \subset \cV.
+\cK = \left\{w\,:\,\underline{\gamma} \le w \le \overline{\gamma} \,\text{ over }\, \Omega, \, \text{ and }\, w\big|_{\partial_D \Omega} = g_D\right\} \subset \cV =W^{1,p}(\Omega).
 \end{equation}
-It is easy to check that $\cK$ is closed and convex; note that the inequality and equality (boundary) constraints are treated in a unified manner by this definition of $\cK$.
+(The boundary values are, as usual, in the trace sense \cite{Evans2010}.)  It is easy to check that $\cK$ is closed and convex.
+Observe that this definition of $\cK$ treats the inequality and equality (boundary) constraints in a unified manner.
 
-We will solve the same VI problem as the abstract problem \eqref{eq:vi}.  Given continuous $f:\cK \to \cV'$ and $\ell \in \cV'$, we assume that the following VI is well-posed for $u^*\in \cK$:
+We will solve the same VI as in the abstract problem \eqref{eq:vi}.  Given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we assume that the following VI is well-posed for $u^*\in \cK$:
 \begin{equation}
@@
 
-Problem \eqref{eq:boxdirichletvi} is a mixed boundary value problem, in weak form, with Dirichlet conditions $u=g_D$ over $\partial_D\Omega$ and Neumann boundary conditions over $\partial_N \Omega = \Omega \setminus \partial_D \Omega$.
-The Neumann boundary conditions are determined by the formula for $f$.
-(That is, nonhomogeneous Neumann conditions imply that the nonlinear operator $f:\cK\to\cV'$ involves an integration over $\partial_N\Omega$.)
+Weak form problem \eqref{eq:boxdirichletvi} is both a free-boundary problem, because of the inequality constraints in defining $\cK$, and a mixed fixed-boundary value problem.
+Dirichlet conditions $u=g_D$ apply over $\partial_D\Omega$ and Neumann conditions over $\partial_N \Omega$.
+The latter are determined by the formula for $f$; nonhomogeneous Neumann conditions require that the nonlinear operator $f:\cK\to\cV'$ include an integration over $\partial_N\Omega$.
 
-In this Section we construct a multilevel finite element (FE) approximation to problem \eqref{eq:boxdirichletvi} using $P_1$ or $Q_1$ elements \cite{Elmanetal2014} and triangular or quadrilateral meshes, respectively, over $\Omega$.
-The targeted mesh will be the finest level in a hierarchy of meshes constructed by refinement of a coarsest mesh.
-This hierarchy generates a multilevel CD of $\cK$, as covered in previous Sections.
-The constraint subsets $\cK_i$  are designed to allow multigrid-type coarser-level corrections (next Section) in a manner which preserves admissibility while not introducing unnecessary high frequencies into the coarser-level problems, especially at the free boundary generated by the inequality constraints.
+Now, in this Section we construct a multilevel finite element (FE) approximation to problem \eqref{eq:boxdirichletvi} using either triangular or quadrilateral meshes over $\Omega$, and $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, respectively.
+The targeted solution is hosted on the finest-level mesh in a hierarchy of meshes constructed by refinement of a coarsest mesh.
+This hierarchy will also generate a multilevel CD of $\cK$.
+The subsets $\cK_i$ in this CD are designed to allow multigrid-type coarser-level corrections (next Section) in a manner which preserves admissibility, while not introducing unnecessary high frequencies into the coarser-level problems, especially at the free boundary.
 
 The mesh hierarchy and FE spaces are constructed by nested refinement in the following standard manner \cite{Elmanetal2014}.
-Suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate elements (triangles or quadrilaterals) with union equal to $\overline{\Omega}$ and characteristic size (maximum edge length) $h_0>0$.
+Suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate elements (triangles or quadrilaterals) with union equal to $\overline{\Omega}$ and maximum edge length (characteristic size) $h_0>0$.
 For $j=1,\dots,J$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ by connecting the midpoints of edges.
-Each $\mathcal{T}^{j-1}$ element therefore becomes four similar elements in $\mathcal{T}^j$ with characteristic size $h_j = 2^{-j} h_0$.\footnote{For arbitrary triangles, and for rectangles, $h_j=2^{-j}h_0$ is the exact maximum edge length in $\mathcal{T}^j$.
-For general quadrilaterals this is proportional \cite{Zhang2004}.
+For triangles and rectangles, each $\mathcal{T}^{j-1}$ element becomes four similar elements in $\mathcal{T}^j$ with size $h_j = 2^{-j} h_0$.\footnote{For arbitrary triangles, and for rectangles, the refinement generates similar elements and $h_j=2^{-j}h_0$ is the exact maximum edge length in $\mathcal{T}^j$.
+For general quadrilaterals, however, similarity is lost and maximum edge lengths are variable; cf.~\cite{Zhang2004}.
 Only rectangular quadrilaterals are used in the Results section.}  We call $J\ge 0$ the \emph{depth} of the hierarchy and $\mathcal{T}^J$ the \emph{finest mesh}.
 Let $m_j$ be the number of nodes in $\mathcal{T}^j$, the \emph{degrees of freedom} at the $j$th level, and denote $h=h_J=2^{-J} h_0$ and $m=m_J$ for the finest level.
-Now let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space \cite{Elmanetal2014} over $\mathcal{T}^j$, with arbitrary boundary values, and observe the following vector space nesting:
+
+Let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space \cite{Elmanetal2014} over $\mathcal{T}^j$, with arbitrary boundary values, and observe the following vector space nesting:
 \begin{equation}
@@
 For $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal value one at a single node, otherwise zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.
-It is easy to show that \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of sign-changing hat functions, and this implies $k\le 1$.
+It is easy to show that \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of sign-changing hat functions, so \eqref{eq:nodalmonotonicity} implies $k\le 1$.
 
-To allow box constraints in our multilevel FE scheme, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the nodes of $\mathcal{T}^j$, with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
-(Note that $\tilde{\mathcal{V}}^j$ functions are not in $\mathcal{V}^j$, and they are not defined everywhere on $\Omega$.)  On the finest-level mesh, let $g_D^J \in \mathcal{V}^J$ satisfy $g_D^J = g_D$ on $\partial_D \Omega$.\footnote{As is standard in conforming FE methods, this restricts the original Dirichlet boundary values $g_D$ to a piecewise-linear function \cite{Elmanetal2014}.}  Suppose that the \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ satisfy
+To allow box constraints, let $\tilde{\mathcal{V}}^j$ denote the set of all functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
+(Note that $\tilde{\mathcal{V}}^j$ functions are not in $\mathcal{V}^j$, and they are not even defined everywhere on $\Omega$.)  Suppose that the \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$, defined on the nodes of the finest-level mesh $\mathcal{T}^J$, satisfy
 \begin{equation}
-\underline{\gamma}^J \le \overline{\gamma}^J, \qquad \underline{\gamma}^J < +\infty, \qquad -\infty < \overline{\gamma}^J \label{eq:fe:boxconstraintrequirements}
+\underline{\gamma}^J \le \overline{\gamma}^J, \qquad \underline{\gamma}^J < +\infty, \qquad -\infty < \overline{\gamma}^J. \label{eq:fe:boxconstraintrequirements}
 \end{equation}
-on the nodes of $\mathcal{T}^J$ and $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.
-On nodes where $\underline{\gamma}^J=-\infty$ there is no lower constraint, and similarly where $\overline{\gamma}^J=+\infty$ no upper constraint.
+Where $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$ then there is no lower or upper constraint, respectively.
+
+To accomodate the fixed-boundary conditions, let $g_D^J \in \mathcal{V}^J$, on the finest-level mesh, satisfy $g_D^J = g_D$ on $\partial_D \Omega$.\footnote{As is standard in conforming FE methods, this restricts the original Dirichlet boundary values $g_D$ to a piecewise-linear function \cite{Elmanetal2014}.}  Suppose furthermore that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$, as otherwise the constraint set, defined next, is empty.
 
@@
 \begin{equation}
-\mathcal{K}^J = \left\{w \in \mathcal{V}^J\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \, \text{ and } \, w|_{\partial_D\Omega} = g_D|_{\partial_D\Omega}\right\}, \label{eq:fe:fineconstraintset}
+\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
 \end{equation}
@@
 The continuum constraint set $\mathcal{K}$ remains abstract in applications where only the nodal obstacles $\underline{\gamma}^J, \overline{\gamma}^J$ are defined.
-In fact, even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants might not imply continuum admissiblity.\footnote{Such caveats apply in geophysical situations where the obstacle is given by meshed elevation data, the continuum limit of which is undefined \cite{Bueler2016}.}
+In fact, even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants will generally not imply continuum admissiblity.\footnote{Such caveats apply in geophysical situations where the obstacle is given by meshed elevation data, the continuum limit of which is generally unknown \cite{Bueler2016}.}
 
@@
 Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ also acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $(R\sigma)[z] = \sigma[z]$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
-While these canonical transfers do not lose information, they typically do involve a change of FE representation.
-For example, canonical restriction, derived by writing coarser-mesh hat functions in terms of finer-mesh ones, is a ``full-weighting'' average \cite{Trottenbergetal2001}.
+Observe that these canonical transfers typically involve a change of FE representation.
+For example, the formula for canonical restriction, derived by writing coarser-mesh hat functions in terms of finer-mesh ones, is a ``full-weighting'' average \cite{Trottenbergetal2001}.
 
@@
 Also observe that, by our refinement construction, the nodes of $\mathcal{T}^{j-1}$ are nodes of $\mathcal{T}^j$.
-For $z\in\mathcal{V}^j$ we define the injection $\iR z$ as the element of $\mathcal{V}^{j-1}$ which has the same nodal values as $z$ at the nodes of $\mathcal{T}^{j-1}$.
+For $z\in\mathcal{V}^j$ we define the injection $\iR z$ as the element of $\mathcal{V}^{j-1}$ which has the same nodal values as $z$ does at the nodes of $\mathcal{T}^{j-1}$.
 Note that injection is monotonic, that is, if $w \le z$ then $\iR w \le \iR z$.
@@
 Finally, the monotone restrictions $\maxR,\minR$ are nonlinear maps which maximize and minimize, respectively, nodal values over the supports of coarser-mesh hat functions.
-To make the definitions precise, first denote the \emph{nodes} of mesh $\mathcal{T}^j$ by $x_q^j$ for $q=0,1,\dots,m_j-1$, and then the corresponding hat functions \cite{Elmanetal2014} by $\psi_q^j(x)$.
-The monotone restrictions $\maxR z,\minR z$ of $z\in\tilde{\mathcal{V}}^j$ are the coarser-level functions in $\tilde{\mathcal{V}}^{j-1}$ defined by
+To make the definitions precise, let $x_q^j$, $q=0,1,\dots,m_j-1$, denote the nodes of mesh $\mathcal{T}^j$, and denote the corresponding hat functions \cite{Elmanetal2014} by $\psi_q^j(x)$.
+The monotone restrictions $\maxR z,\minR z$ of an extended-real-valued nodal function $z\in\tilde{\mathcal{V}}^j$ are coarser-level functions in $\tilde{\mathcal{V}}^{j-1}$ defined by
 \begin{align}
@@
 for $p=0,1,\dots,m_{j-1}-1$.
-In computing $\max$ and $\min$, the usual ordering rules for the extended real line $\tilde\RR$ are applied.
-Note that $\minR z \le z \le \maxR z$ because, by mesh refinement, node $x_p^{j-1}$ is in the set of $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$.
+(In computing $\max$ and $\min$, the usual ordering of the extended real line is applied.)  Note that $\minR z \le z \le \maxR z$ because, by mesh refinement, node $x_p^{j-1}$ is in the set of $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$.
 

### 2022-10-18 paper/mcd2.tex
a few more edits; address FIXMEs pages 11-12

@@
 
-To define the constraint subsets in the current multilevel FE context, namely the sets denoted $\cK_i$ in Section \ref{sec:cd}, suppose $w^J \in \cK^J$ is an admissible finest-level iterate which approximates the solution $u^J$ of VI \eqref{eq:fe:vi}.
-From this iterate we define the finest-level \emph{lower and upper defect constraints} \cite{GraeserKornhuber2009}
+To define the CD constraint subsets in the current multilevel FE context, suppose $w^J \in \cK^J$ is an admissible finest-level iterate which approximates the solution $u^J$ of VI \eqref{eq:fe:vi}.
+From this iterate we define the \emph{finest-level lower/upper defect constraints}
 \begin{equation}
@@
 \end{equation}
-functions in $\tilde{\mathcal{V}}^J$ satisfying $\underline{\chi}^J \le 0 \le \overline{\chi}^J$.
-FIXME ADD DIRICHLET The motivation here is that a corrected iterate is admissible, $\underline{\gamma}^J \le w^J + z^J \le \overline{\gamma}^J$, if and only if the perturbation $z^J$ is between the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$.
+as functions in $\tilde{\mathcal{V}}^J$.
+Note that $\underline{\chi}^J \le 0 \le \overline{\chi}^J$ everywhere in $\overline{\Omega}$, including along the Dirichlet boundary $\partial_D\Omega$.
+The motivation for such defect constraints \cite{GraeserKornhuber2009} is that a corrected iterate is admissible with respect to the defining inequalities for $\cK^J$, namely $\underline{\gamma}^J \le w^J + z^J \le \overline{\gamma}^J$, if and only if the perturbation $z^J$ is between the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$.
 
-For $j=J,J-1,\dots,1$ we apply monotone restrictions to define the \emph{level defect constraints},
+For coarser levels $j=J-1,\dots,1,0$ we apply monotone restrictions to define the \emph{level defect constraints},
 \begin{equation}
-\underline{\chi}^{j-1} = \maxR \underline{\chi}^j, \qquad \overline{\chi}^{j-1} = \minR \overline{\chi}^j. \label{eq:fe:chilevels}
+\underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}, \qquad \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}. \label{eq:fe:chilevels}
 \end{equation}
-Observe that
+FIXME: VALUES ALONG $\partial_D\Omega$ ARE BY INJECTION  Observe that
 \begin{equation}

### 2022-10-19 c/nmcd.c
default problem is linear so default maxits (pointwise Newton its) should be 1

@@ -121,7 +121,7 @@ int main(int argc,char **argv) {
     ctx.gamma_lower = &gamma_lower;
     ctx.f_rhs = &fg_zero;
     ctx.g_bdry = &u_exact;
-    ctx.maxits = 2;
+    ctx.maxits = 1;
     ctx.quadpts = 2;
     ctx.sweeps = 1;
     ctx.residualcount = 0;

### 2022-10-19 c/nmcd.c
change interface and whitespace for VecPrintRange()

@@ -223,10 +223,10 @@ int main(int argc,char **argv) {
 
     // report ranges for initial w and corresponding residual f^J(w)
     PetscCall(PetscPrintf(PETSC_COMM_WORLD,"initial:\n"));
-    PetscCall(VecPrintRange(w,"iterate w",""));
+    PetscCall(VecPrintRange(w,"iterate w","",PETSC_TRUE));
     PetscCall(DMGetGlobalVector(levs[jtop].dmda,&F));
     PetscCall(ApplyOperatorF(levs[jtop].dmda,w,F,&ctx));
-    PetscCall(VecPrintRange(F,"residual f^J(w)",""));
+    PetscCall(VecPrintRange(F,"residual f^J(w)","",PETSC_TRUE));
     PetscCall(DMRestoreGlobalVector(levs[jtop].dmda,&F));
     // FIXME also CR residual?
 
@@ -288,10 +288,10 @@ int main(int argc,char **argv) {
         PetscCall(VecAXPY(w,1.0,levs[jtop].z));
         // report range on initial w and f^J(w) for initial iterate
         PetscCall(PetscPrintf(PETSC_COMM_WORLD,"iterate %2d result:\n",viter+1));
-        PetscCall(VecPrintRange(w,"iterate w",""));
+        PetscCall(VecPrintRange(w,"iterate w","",PETSC_TRUE));
         PetscCall(DMGetGlobalVector(levs[jtop].dmda,&F));
         PetscCall(ApplyOperatorF(levs[jtop].dmda,w,F,&ctx));
-        PetscCall(VecPrintRange(F,"residual f^J(w)",""));
+        PetscCall(VecPrintRange(F,"residual f^J(w)","",PETSC_TRUE));
         PetscCall(DMRestoreGlobalVector(levs[jtop].dmda,&F));
         // FIXME also CR residual?
     } // for viter ...

### 2022-10-19 c/src/utilities.c

@@ -1,15 +1,18 @@
 #include <petsc.h>
 
-PetscErrorCode VecPrintRange(Vec X, const char *name, const char *infcase) {
+PetscErrorCode VecPrintRange(Vec X, const char *name, const char *infcase,
+                             PetscBool newline) {
     PetscReal vmin, vmax;
     if (X) {
         PetscCall(VecMin(X,NULL,&vmin));
         PetscCall(VecMax(X,NULL,&vmax));
-        PetscCall(PetscPrintf(PETSC_COMM_WORLD,"  %9.6f <= %s <= %9.6f\n",
+        PetscCall(PetscPrintf(PETSC_COMM_WORLD,"  %10.6f <= %s <= %10.6f",
                               vmin,name,vmax));
     } else
-        PetscCall(PetscPrintf(PETSC_COMM_WORLD,"  [ %s=NULL is %s ]\n",
+        PetscCall(PetscPrintf(PETSC_COMM_WORLD,"  [[[   %s=NULL is %s   ]]]",
                               name,infcase));
+    if (newline)
+        PetscCall(PetscPrintf(PETSC_COMM_WORLD,"\n"));
     return 0;
 }
 

### 2022-10-19 c/src/utilities.h

@@ -6,7 +6,7 @@
 
 // print extended-reals min and max of a Vec; string infcase says what
 // to print if X==NULL
-PetscErrorCode VecPrintRange(Vec X, const char *name, const char *infcase);
+PetscErrorCode VecPrintRange(Vec X, const char *name, const char *infcase, PetscBool newline);
 
 // set flg=PETSC_TRUE if  u <= v  everywhere, otherwise flg=PETSC_FALSE
 // extended reals rule:  if u=NULL (-infty) or v=NULL (+infty) then flg=PETSC_TRUE

### 2022-10-19 c/ldctwolevel.c
change interface for LDC: LDCReportDCRanges() becomes LDCCheckDCRanges()

@@ -70,10 +70,6 @@ int main(int argc,char **argv) {
     PetscCall(VecViewMatlabStdout(ldc[1].philow));
 #endif
 
-    // ranges on Vecs on each level
-    PetscCall(LDCReportDCRanges(ldc[0]));
-    PetscCall(LDCReportDCRanges(ldc[1]));
-
     // check admissibility of constant z on level 0 and y and z on level 1
     PetscCall(PetscPrintf(PETSC_COMM_WORLD,"checking admissibility of constant vectors:\n"));
     PetscCall(DMCreateGlobalVector(ldc[0].dal,&v));

### 2022-10-19 c/src/ldc.c

@@ -79,14 +79,31 @@ PetscErrorCode LDCRefine(LDC *coarse, LDC *fine) {
     return 0;
 }
 
-PetscErrorCode LDCReportDCRanges(LDC ldc) {
-    PetscCall(PetscPrintf(PETSC_COMM_WORLD,"defect constraint ranges at level %d:\n",
-                          ldc._level));
-    PetscCall(VecPrintRange(ldc.chiupp,"chiupp","+infty"));
-    PetscCall(VecPrintRange(ldc.chilow,"chilow","-infty"));
-    PetscCall(VecPrintRange(ldc.phiupp,"phiupp","+infty"));
-    PetscCall(VecPrintRange(ldc.philow,"philow","-infty"));
-    return 0;
+PetscErrorCode LDCCheckDCRanges(LDC ldc) {
+    PetscReal clmax = PETSC_NINFINITY, cumin = PETSC_INFINITY,
+              plmax = PETSC_NINFINITY, pumin = PETSC_INFINITY;
+    PetscInt  retval = 0;
+    if (ldc.chilow)
+        PetscCall(VecMin(ldc.chilow,NULL,&clmax));
+    if (ldc.chiupp)
+        PetscCall(VecMin(ldc.chiupp,NULL,&cumin));
+    if (ldc.philow)
+        PetscCall(VecMin(ldc.philow,NULL,&plmax));
+    if (ldc.phiupp)
+        PetscCall(VecMin(ldc.phiupp,NULL,&pumin));
+    if ((clmax <= 0.0) && (0.0 <= cumin) && (plmax <= 0.0) && (0.0 <= pumin))
+        PetscCall(PetscPrintf(PETSC_COMM_WORLD,"defect constraint check PASSES (level %d):\n",
+                              ldc._level));
+    else {
+        PetscCall(PetscPrintf(PETSC_COMM_WORLD,"defect constraint check FAILS (level %d):\n",
+                              ldc._level));
+        retval = 1;
+    }
+    PetscCall(VecPrintRange(ldc.chilow,"chilow","-infty",PETSC_FALSE));
+    PetscCall(VecPrintRange(ldc.chiupp,"chiupp","+infty",PETSC_TRUE));
+    PetscCall(VecPrintRange(ldc.philow,"philow","-infty",PETSC_FALSE));
+    PetscCall(VecPrintRange(ldc.phiupp,"phiupp","+infty",PETSC_TRUE));
+    return retval;
 }
 
[…]

### 2022-10-19 c/src/ldc.h

@@ -100,8 +100,8 @@ PetscErrorCode LDCDownDCsCRFromResidual(LDC *ldc, Vec y, Vec F, Vec Fhat);
 PetscErrorCode LDCVecFromFormula(LDC ldc, PetscReal (*ufcn)(PetscReal,PetscReal,void*),
                                  Vec u, void *ctx);
 
-// for each of 4 DCs, report min and max
-PetscErrorCode LDCReportDCRanges(LDC ldc);
+// for each pair of DCs, check lower <= 0 <= upper, and report min and max
+PetscErrorCode LDCCheckDCRanges(LDC ldc);
 
 // return flg=PETSC_TRUE if  ldc.philow <= y <= ldc.phiupp,  otherwise flg=PETSC_FALSE
 PetscErrorCode LDCCheckAdmissibleDownDefect(LDC ldc, Vec y, PetscBool *flg);

### 2022-10-19 c/nmcd.c
this lame goal is absolutely necessary before I can proceed

@@ -7,6 +7,12 @@ static char help[] =
 "solver is projected, nonlinear Gauss-Seidel (PNGS) sweeps.  Option prefix\n"
 "nm_.  Compare obstaclesl.c.\n\n";
 
+// FIXME make it work as stupid single-level PNGS, i.e. make it reproduce
+//   ./obstaclesl -da_refine 2 -ob_counts -ob_pngs -snes_converged_reason
+// or similar by doing
+//   ./nmcd -da_grid_x 9 -da_grid_y 9 -nm_levels 1 -nm_cycles 26
+// currently there is NO progress with -nm_levels 1
+
 // FIXME possible ways to describe the current issues:
 //   * the iterate w is only slowly falling toward the obstacle
 //   * the corrections are always negative

### 2022-10-20 paper/mcd2.tex
edits and stuck

--- 2022-10-18 paper/mcd2.tex
@@
 
-Weak form problem \eqref{eq:boxdirichletvi} is both a free-boundary problem, because of the inequality constraints in defining $\cK$, and a mixed fixed-boundary value problem.
+Problem \eqref{eq:boxdirichletvi} is both a free-boundary problem, because of the inequality constraints in defining $\cK$, and a mixed fixed-boundary value problem.
 Dirichlet conditions $u=g_D$ apply over $\partial_D\Omega$ and Neumann conditions over $\partial_N \Omega$.
@@
 
-Now, in this Section we construct a multilevel finite element (FE) approximation to problem \eqref{eq:boxdirichletvi} using either triangular or quadrilateral meshes over $\Omega$, and $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, respectively.
+We now construct a multilevel finite element (FE) approximation to problem \eqref{eq:boxdirichletvi} using either triangular or quadrilateral meshes over $\Omega$, and $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, respectively.
 The targeted solution is hosted on the finest-level mesh in a hierarchy of meshes constructed by refinement of a coarsest mesh.
@@
 
-The multilevel algorithm in the next section, a solver for VI problem \eqref{eq:fe:vi}, requires transfer operators between meshes.
-We use five such operators, three of which are standard linear maps used for multigrid methods \cite{Trottenbergetal2001}.
+The multilevel algorithm in the next section, a solver for finite-dimensional problem \eqref{eq:fe:vi}, requires transfer operators between meshes.
+We use five such operators, three of which are standard within multigrid methods \cite{Trottenbergetal2001}.
 Table \ref{tab:transfers} will help the reader keep track.
@@
 Finally, the monotone restrictions $\maxR,\minR$ are nonlinear maps which maximize and minimize, respectively, nodal values over the supports of coarser-mesh hat functions.
+These restrictions will be applied only to nodal functions ($\tilde{\mathcal{V}}^j$), and only to functions which are either nonnegative or nonpositive, and not more generally.
 To make the definitions precise, let $x_q^j$, $q=0,1,\dots,m_j-1$, denote the nodes of mesh $\mathcal{T}^j$, and denote the corresponding hat functions \cite{Elmanetal2014} by $\psi_q^j(x)$.
@@
 for $p=0,1,\dots,m_{j-1}-1$.
-(In computing $\max$ and $\min$, the usual ordering of the extended real line is applied.)  Note that $\minR z \le z \le \maxR z$ because, by mesh refinement, node $x_p^{j-1}$ is in the set of $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$.
+(In computing $\max$ and $\min$, the usual ordering of the extended real line is applied.)  Because, by mesh refinement, a node $x_p^{j-1}$ is in the set of $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, the following formal property holds:
+\begin{equation}
+\minR z \le z \le \maxR z,  \label{eq:fe:monotonerestrictionprops}
+\end{equation}
+In particular, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These properties are used in the CD construction which follows.
 
-To define the constraint subsets in the current multilevel FE context, namely the sets denoted $\cK_i$ in Section \ref{sec:cd}, suppose $w^J \in \cK^J$ is an admissible finest-level iterate which approximates the solution $u^J$ of VI \eqref{eq:fe:vi}.
-From this iterate we define the finest-level \emph{lower and upper defect constraints} \cite{GraeserKornhuber2009}
+Suppose $w^J \in \cK^J$ is an admissible finest-level iterate which approximates the solution $u^J$ of VI \eqref{eq:fe:vi}.
+We will now define the CD constraint subsets in our multilevel FE context from $w^J$ and the defining inequalities for $\cK^J$ (equation \eqref{eq:fe:fineconstraintset}).
+Though the ideas are fundamentally the same as for unilateral constraints \cite{GraeserKornhuber2009}, the notation is more elaborate because we consider bi-obstacle (box-constrained) problems.
+
+From the iterate $w^J$ we define the \emph{finest-level defect constraints}
 \begin{equation}
@@
 \end{equation}
-functions in $\tilde{\mathcal{V}}^J$ satisfying $\underline{\chi}^J \le 0 \le \overline{\chi}^J$.
-FIXME ADD DIRICHLET The motivation here is that a corrected iterate is admissible, $\underline{\gamma}^J \le w^J + z^J \le \overline{\gamma}^J$, if and only if the perturbation $z^J$ is between the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$.
+as nodal functions in $\tilde{\mathcal{V}}^J$.
+Note that $\underline{\chi}^J \le 0 \le \overline{\chi}^J$ everywhere in $\overline{\Omega}$, including along the Dirichlet boundary $\partial_D\Omega$ where $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ and $w^J = g_D^J = g_D$.
+The motivation for creating such defect constraints \cite{GraeserKornhuber2009} is that a corrected iterate $w^J + z^J$, for $z^J\in\cV^J$, is admissible ($w^J + z^J \in \cK^J$) if and only if the perturbation $z^J$ is between the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$, and $z^J=0$ on $\partial_D\Omega$.
 
-For $j=J,J-1,\dots,1$ we apply monotone restrictions to define the \emph{level defect constraints},
+For coarser levels $j=J-1,\dots,1,0$ we apply the monotone restrictions to generate ``\emph{below}'' and ``\emph{above}'' \emph{level defect constraints} (LDCs):
 \begin{equation}
-\underline{\chi}^{j-1} = \maxR \underline{\chi}^j, \qquad \overline{\chi}^{j-1} = \minR \overline{\chi}^j. \label{eq:fe:chilevels}
+\underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}, \qquad \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}, \label{eq:fe:chilevels}
 \end{equation}
-Observe that
+thus $\underline{\chi}^{j},\overline{\chi}^{j} \in \tilde{\mathcal{V}}^J$.  Observe that
 \begin{equation}
-\underline{\chi}^{J} \le \dots \le \underline{\chi}^0 \le 0 \le \overline{\chi}^0 \le \dots \le \overline{\chi}^J. \label{eq:fe:chiordering}
+\underline{\chi}^{J} \le \dots \le \underline{\chi}^0 \le 0 \le \overline{\chi}^0 \le \dots \le \overline{\chi}^J, \label{eq:fe:chiordering}
 \end{equation}
-A $J=4$ example, with no upper obstacles, namely $\overline{\chi}^j=+\infty$ for all $j$, is shown in Figure \ref{fig:chiphilevels}.
+in the sense of comparing nodal values at nodes shared between levels.  A $J=4$ example of ``below'' LDCs $\underline{\chi}^j$ is shown in Figure \ref{fig:chiphilevels}.
 
@@
 \includegraphics[width=0.65\textwidth]{fixfigs/chiphilevels.pdf}
-\caption{The lower defect constraint $\underline{\chi}^J = \underline{\gamma}^J - w^J$ is decomposed using maximum restriction: $\underline{\chi}^{j-1} = \maxR \underline{\chi}^j$.}
+\caption{The ``below'' level defect constraint (LDC) $\underline{\chi}^J = \underline{\gamma}^J - w^J$ is decomposed using maximum restriction: $\underline{\chi}^{j-1} = \maxR \underline{\chi}^j$.}
 \label{fig:chiphilevels}
@@
 
-We also compute the differences between the level defect constraints:
+We then compute the differences between the previous LDCs:
 \begin{equation}
@@
 \end{equation}
-using $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ to define $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.  These constraints bracket zero,
+using $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ to define $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.
+These new nodal functions $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ are also LDCs, and again they bracket zero,
 \begin{equation}
-\underline{\phi}^j \le 0 \le \overline{\phi}^j.  \label{eq:fe:phibrackets}
+\underline{\phi}^j \le 0 \le \overline{\phi}^j,  \label{eq:fe:phibrackets}
 \end{equation}
-Though these functions are not ordered as in \eqref{eq:fe:chiordering}, telescoping sums hold for $j=0,1,\dots,J$:
+but they are not ordered as in \eqref{eq:fe:chiordering}.  However, telescoping sums hold for $j=0,1,\dots,J$:
 \begin{equation}
@@
 \end{equation}
+Table \ref{tab:ldcs} should help the reader keep track of these objects.
 
-FIXME A multilevel V-cycle is defined in the next Section.
-For the downward direction we will use the differences $\underline{\phi}^j,\overline{\phi}^j$ as constraints, and for the upward part the defect constraints $\underline{\chi}^j,\overline{\chi}^j$.
-That is,
+\begin{table}
+\begin{tabular}{llc}
+\emph{LDC name}        & \emph{formula} \\ \hline
+(finest) upward below & ${\Large \strut} \underline{\chi}^J = \underline{\gamma}^J - w^J$ \\
+(finest) upward above & ${\Large \strut} \overline{\chi}^J = \overline{\gamma}^J - w^J$ \\
+upward below          & ${\Large \strut} \underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}$ \\
+upward above          & ${\Large \strut} \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}$ \\
+downward below        & ${\Large \strut} \underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1}$ \\
+downward above        & ${\Large \strut} \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}$ \\
+\end{tabular}
+
+\medskip
+\caption{Level defect constraints (LDCs) are constructed from the defining box constraints ($\underline{\gamma}^J,\overline{\gamma}^J$) in problem \eqref{eq:fe:vi}, and a finest level iterate ($w^J \in \cK^J$), by starting at the finest level.
+Then monotone restrictions and differences generate LDCs for coarser levels $j=0,\dots,J-1$.
+LDCs $\underline{\chi}^j,\overline{\chi}^j$ are used for the upward direction in V-cycles, and $\underline{\phi}^j,\overline{\phi}^j$ for the downward direction.}
+\label{tab:ldcs}
+\end{table}
+
+A multilevel V-cycle is defined in the next Section.
+For the downward direction in a V-cycle we will use the LDCs $\underline{\phi}^j,\overline{\phi}^j$ as constraints, and for the upward part the LDCs $\underline{\chi}^j,\overline{\chi}^j$.
+In fact, let
 \begin{align}
-\mathcal{D}^j = \{v \in \mathcal{V}^j \,:\, \underline{\phi}^j \le v \le \overline{\phi}^j\}, \label{eq:fe:constraintsets} \\
-\mathcal{U}^j = \{v \in \mathcal{V}^j \,:\, \underline{\chi}^j \le v \le \overline{\chi}^j\}, \notag
+\mathcal{D}^j = \left\{v \,:\, \underline{\phi}^j \le v \le \overline{\phi}^j \text{ and } \, v|_{\partial_D\Omega} = 0\right\} \subset \mathcal{V}^j, \label{eq:fe:constraintsets} \\
+\mathcal{U}^j = \left\{v\,:\, \underline{\chi}^j \le v \le \overline{\chi}^j \text{ and } \, v|_{\partial_D\Omega} = 0\right\} \subset \mathcal{V}^j, \notag
 \end{align}
-are the \emph{downward} and \emph{upward constraint subsets} of the FE subspaces $\mathcal{V}^j$, respectively.
-Observe that the upward sets are larger, $\mathcal{D}^j \subset \mathcal{U}^j$, and in fact by \eqref{eq:fe:telescoping} the set decompositions
+be the \emph{downward} and \emph{upward defect constraint subsets} of the FE subspaces $\mathcal{V}^j$, respectively.  Again these subsets are closed and convex.
+
+Observe that the upward sets are larger, $\mathcal{D}^j \subset \mathcal{U}^j$, and by \eqref{eq:fe:telescoping} the set decompositions
 \begin{equation}
@@
 
-FIXME what are the CDs?  what are the decomposition operators?
+Recalling definitions \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, we now have two varieties of multilevel CDs of the finest-level upward defect constraint set $\mathcal{U}^J$.
+
+\begin{lemma}  Given upper and lower obstacles $\underline{\gamma}^J,\overline{\gamma}^J \in \cV^J$, the constraint set $\cK^J$ defined as in \eqref{eq:fe:fineconstraintset}, and a finest level iterate $w^J \in \cK^J$.
+Suppose that the downward and upward defect constraint subsets $\mathcal{D}^i,\mathcal{U}^j$ are defined as in \eqref{eq:fe:constraintsets}.
+Then for any $j=0,1,\dots,J$ the sums
+\begin{equation}
+\mathcal{U}^J \stackrel{\text{(i)}}{=} \sum_{i=0}^J \mathcal{D}^i \quad \text{ and } \quad  \mathcal{U}^J \stackrel{\text{(ii)}}{=} \mathcal{U}^j + \sum_{i={j+1}}^J \mathcal{D}^i
+\end{equation}
+are each CDs of $\mathcal{U}^J = \cK^J - w^J$, over subspace decompositions $\cV^J=\sum_{i=0}^J \cV^i$ and $\cV^J=\sum_{i=j}^J \cV^i$, respectively.
+\end{lemma}
+
+\begin{proof}  Definition \eqref{eq:fe:constraintsets} shows $\mathcal{D}^i \subset \cV^i$ and $\mathcal{U}^i \subset \cV^i$.  FIXME what are the decomposition operators?
+
+FIXME Define the coarsening decomposition (projection) operator
+    $$\Pi_j^{j-1} v = \min\{\max\{\iR v, \underline{\chi}^{j-1}\}, \overline{\chi}^{j-1}\}.$$
+It is easy to see that $\Pi_j^{j-1}:\mathcal{U}^j \to \mathcal{U}^{j-1}$ and $I-\Pi_j^{j-1}:\mathcal{U}^j \to \mathcal{D}^{j-1}$.
+\end{proof}
 

### 2022-10-22 paper/mcd2.tex
trying to be really clear on NMCD

@@
 
-\begin{table}
+\begin{table}[H]
 \begin{tabular}{llc}
@@
 \medskip
-\caption{Level defect constraints (LDCs) are constructed from the defining box constraints ($\underline{\gamma}^J,\overline{\gamma}^J$) in problem \eqref{eq:fe:vi}, and a finest level iterate ($w^J \in \cK^J$), by starting at the finest level.
-Then monotone restrictions and differences generate LDCs for coarser levels $j=0,\dots,J-1$.
-LDCs $\underline{\chi}^j,\overline{\chi}^j$ are used for the upward direction in V-cycles, and $\underline{\phi}^j,\overline{\phi}^j$ for the downward direction.}
+\caption{The box level defect constraints (LDCs) are constructed from the defining box constraints ($\underline{\gamma}^J,\overline{\gamma}^J$) in problem \eqref{eq:fe:vi}, and a finest level iterate ($w^J \in \cK^J$), by starting at the finest level.}
 \label{tab:ldcs}
@@
 
-\begin{lemma}  Given upper and lower obstacles $\underline{\gamma}^J,\overline{\gamma}^J \in \cV^J$, the constraint set $\cK^J$ defined as in \eqref{eq:fe:fineconstraintset}, and a finest level iterate $w^J \in \cK^J$.
-Suppose that the downward and upward defect constraint subsets $\mathcal{D}^i,\mathcal{U}^j$ are defined as in \eqref{eq:fe:constraintsets}.
-Then for any $j=0,1,\dots,J$ the sums
-\begin{equation}
-\mathcal{U}^J \stackrel{\text{(i)}}{=} \sum_{i=0}^J \mathcal{D}^i \quad \text{ and } \quad  \mathcal{U}^J \stackrel{\text{(ii)}}{=} \mathcal{U}^j + \sum_{i={j+1}}^J \mathcal{D}^i
-\end{equation}
-are each CDs of $\mathcal{U}^J = \cK^J - w^J$, over subspace decompositions $\cV^J=\sum_{i=0}^J \cV^i$ and $\cV^J=\sum_{i=j}^J \cV^i$, respectively.
+\begin{lemma}  Given upper and lower obstacles $\underline{\gamma}^J,\overline{\gamma}^J \in \cV^J$, the constraint set $\cK^J$ defined as in \eqref{eq:fe:fineconstraintset}, a finest level iterate $w^J \in \cK^J$, and downward and upward defect constraint subsets $\mathcal{D}^i,\mathcal{U}^j$ as defined in \eqref{eq:fe:constraintsets}.
+Then for any $j=0,1,\dots,J$ the descending sums
+\begin{align}
+\mathcal{U}^J &\stackrel{\text{(i)}}{=} \mathcal{D}^J + \mathcal{D}^{J-1} + \dots + \mathcal{D}^0, \label{eq:fe:downwardsumcd} \\
+\mathcal{U}^J &\stackrel{\text{(ii)}}{=} \mathcal{D}^J + \mathcal{D}^{J-1} + \dots + \mathcal{D}^{j+1} + \mathcal{U}^j \label{eq:fe:upwardsumcd}
+\end{align}
+are CDs of\, $\mathcal{U}^J = \cK^J - w^J$, over subspace decompositions $\cV^J=\sum_{i=0}^J \cV^i$ and $\cV^J=\sum_{i=j}^J \cV^i$, respectively.
 \end{lemma}
@@
 
-Our \emph{nonlinear multilevel constraint decomposition} (NMCD) V-cycle algorithm in this section will use the two types of constraint subsets in \eqref{eq:fe:constraintsets}, with distinct sets for the downward and upward parts of the cycle.
-We also diverge from the multilevel CD algorithms by Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear \emph{full approximation scheme} (FAS) philosophy \cite{BrandtLivne2011}.
-In the FAS approach here we must coarsen the solution iterate itself, using the injection restriction, in constrast to classical linear multigrid, wherein only the residual and correction are coarsened, and in constrast to the implemented examples in \cite{GraeserKornhuber2009,Tai2003} in which $f$ is linear.
+The \emph{nonlinear multilevel constraint decomposition} (NMCD) V-cycle algorithm in this section uses CDs \eqref{eq:fe:downwardsumcd} and \eqref{eq:fe:upwardsumcd}, thus applying coarse corrections from $\mathcal{D}^j$ for the downward and $\mathcal{U}^j$ for the upward directions, respectively.
+The algorithm generalizes the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear \emph{full approximation scheme} (FAS) approach \cite{BrandtLivne2011}.
+That is, we coarsen the solution iterate itself, using the injection restriction, in constrast to classical linear multigrid, wherein only the residual and correction are coarsened.
 
-FIXME from here
-
-We will correct the finest-level iterate $w^J\in \mathcal{K}^J$ by adding modifications from each coarser subspace $\mathcal{V}^j$.
-By nesting \eqref{eq:fe:nestedspaces} these corrections live in the fine-level subspace $\mathcal{V}^J$, but the corrected iterate needs to be admissible as well.
-The accumulated corrections, from the constraint subsets $\mathcal{D}^j$ and $\mathcal{U}^j$, over the whole multigrid V-cycle, compute a new iterate in $\mathcal{K}^J$.
-
-In detail, suppose $y^{j+1} \in \mathcal{D}^{j+1}, \dots, y^J \in \mathcal{D}^J$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:nmcdvcycle}).
-The so-far corrected iterate is admissible, namely $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$, because the sum of corrections is in the (finest) defect constraint set, i.e.~$y^J + \dots + y^{j+1} \in \mathcal{U}^J = \{v\ge \chi^J\}$.
+The finest-level iterate $w^J\in \mathcal{K}^J$ is corrected by perturbations from each coarser subspace $\mathcal{V}^j$---by nesting \eqref{eq:fe:nestedspaces} these corrections live in the fine-level subspace $\mathcal{V}^J$---but the corrected iterate needs to be admissible.
+In detail, suppose $y^J \in \mathcal{D}^J, y^{J-1} \in \mathcal{D}^{J-1}, \dots, y^{j+1} \in \mathcal{D}^{j+1}$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:nmcdvcycle}).
+By \eqref{eq:fe:downwardsumcd} the corrected iterate at this point is admissible, namely $y^J + \dots + y^{j+1} + 0 + \dots + 0 \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.
+Smoothing then happens in the next-coarser level, i.e.~we incompletely solve VI \eqref{eq:fe:downvi} below.
+The coarsest-level VI is then (accurately) solved to generate the downslash correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$, again by \eqref{eq:fe:downwardsumcd}.
+Starting upward, however, we can use $y^1+ z^0 \in \mathcal{U}^1$ as an initial iterate, then smooth by incompletely solving VI \eqref{eq:fe:upvi} in $\mathcal{U}^1$ (see below), yielding admissible total correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$ using \eqref{eq:fe:upwardsumcd}.
+Continuing in this way, solving in the larger upward constraint sets $\mathcal{U}^j$, eventually generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction, yielding the update $w^J \leftarrow w^J + z^J \in \mathcal{K}^J$.
+Each of the upward smoothing stages solves a less-restricted VI problem than the corresponding down-smoothing problem.
 
@@
 \end{center}
-\caption{An NMCD V-cycle computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j$ may expand into larger constraint subsets $\mathcal{U}^j$.}
+\caption{An NMCD V-cycle computes downward corrections $y_j \in \mathcal{D}^j$.  The upward corrections $z_j\in\mathcal{U}^j$ live in larger constraint subsets.}
 \label{fig:nmcdvcycle}
@@
 When computing the next-coarser correction one must use the coarser-mesh representation of the problem.
-As in nonlinear FAS multigrid schemes \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, this requires that we coarsen the iterate as well as the residual and the correction, and here injection is used \cite[section 5.3]{Trottenbergetal2001}.
+As in nonlinear FAS multigrid schemes \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, this requires that we coarsen the iterate as well as the residual and the correction.
+Injection is used \cite[section 5.3]{Trottenbergetal2001}.
 Define
@@
 
+FIXME constrast the implemented examples in \cite{GraeserKornhuber2009,Tai2003} in which $f$ is linear
+
 % A BRIDGE TOO FAR:  \section{Results for a nonlocal variational inequality}

### 2022-10-23 paper/mcd2.tex
commit a block of writing

@@
 Continuing in this way, solving in the larger upward constraint sets $\mathcal{U}^j$, eventually generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction, yielding the update $w^J \leftarrow w^J + z^J \in \mathcal{K}^J$.
-Each of the upward smoothing stages solves a less-restricted VI problem than the corresponding down-smoothing problem.
 
@@
 
-When computing the next-coarser correction one must use the coarser-mesh representation of the problem.
-As in nonlinear FAS multigrid schemes \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, this requires that we coarsen the iterate as well as the residual and the correction.
-Injection is used \cite[section 5.3]{Trottenbergetal2001}.
-Define
+A key idea, implicit in the above paragraph, is that the up-smoothing stages solve less-restricted VI problems than the down-smoothing stages.
+Intuitively-speaking, going downward one does not know what the coarse corrections will be, so they must be in sufficiently-small sets ($\mathcal{D}^j$) so that there sum is admissible once we get back to the current level.
+However, going upward the result of smoothing can be computed in the larger admissible-sum sets ($\mathcal{U}^j$).
+
+To state the NMCD algorithm---see Algorithm \ref{alg:nmcd} below---we must explain how to generate the coarsened problems solved at each mesh level.\footnote{Smoothers/solvers for these problems are proposed in Section \ref{sec:smoothers}.}  As in classical FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation (iterate) and the residual must be coarsened, and a but also a source functional must be created.
+Let $j$ denote the mesh level of the coarsened problem.
+We assume a (re-)discretized nonlinear operator $f^j$, approximating $f$ in \eqref{eq:boxdirichletvi}, with output in $(\cV^j)'$.
+Injection $\iR$ generates the coarsened iterate (denoted $g^j$ below; note injection is commonly used for classical FAS \cite[section 5.3]{Trottenbergetal2001}), and canonical restriction $R$ the coarsened source functional (denoted $\ell^j$).
+
+The formulas are as follows.
+At the finest level we have problem \eqref{eq:fe:vi} defined by the discretized operator $f^J$ and the original source functional $\ell^J$.
+A finest-level, admissible iterate $w^J \approx u^J$, $w^J \in \mathcal{K}^J$, has already\footnote{Conceptually-speaking, at least.
+In practice one generates LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ during the descent stage; see Algorithm \ref{alg:nmcd}.
+Observe that the LDCs depend neither on coarsened iterates $g^j$ nor corrections $y^j,z^j$.} been used to general the LDCs and the constraint sets $\mathcal{D}^j,\mathcal{U}^j$ at each level (Section \ref{sec:multilevel}).
+As we descend in the V-cycle, we recursively define $j$th-level iterates
 \begin{equation}
@@
 \end{equation}
-Function $g^j$ is the $j$th-level mesh approximation to the solution of \eqref{eq:fe:vi}, before smoothing and correction at that level.\footnote{If we were to inject the finest-mesh obstacle downward to coarser meshes, $\gamma^{j-1} = \iR \gamma^j$, then, because $\iR$ preserves inequalities and $y^j \in \mathcal{D}^j$, it would follow that $g^j \ge \gamma^j$.
-In this sense $g^j$ is admissible on the $j$th-level mesh.
-However, following \cite{GraeserKornhuber2009}, our use of defect constraints $\chi^j$ avoids the need to generate the $\gamma^j$.}
+These are the iterates \emph{before} smoothing and/or correction has been applied at level $j$.
+(Note that the problem solved by the finer-level, down-smoother correction $y^{j+1}$ is stated momentarily.)  From $g^j$ we define
+\begin{equation}
+\ell^j = \begin{cases} \ell^J, & j=J \\
+                       f^j(g^j) + R\left(\ell^{j+1}-f^{j+1}(g^{j+1}+y^{j+1})\right), & j<J. \end{cases} \label{eq:fe:levelsource}
+\end{equation}
+On descent we solve the following VI for $y^j \in \mathcal{D}^j$:
+\begin{equation}
+\ip{f^j(g^j + y^j)}{v-y^j} \ge \ip{\ell^j}{v-y^j} \qquad \text{for all } v\in \mathcal{D}^j, \label{eq:fe:downvi}
+\end{equation}
+for which the initial iterate is
+\begin{equation}
+(y^j)^{(0)}=0.  \label{eq:fe:downwardinitial}
+\end{equation}
+(Zero is admissible by \eqref{eq:fe:phibrackets}.)  On ascent the same VI is solved, but in a larger admissible set.  We find $z^j \in \mathcal{U}^j$ so that
+\begin{equation}
+\ip{f^j(g^j + z^j)}{v-z^j} \ge \ip{\ell^j}{v-z^j} \qquad \text{for all } v\in \mathcal{U}^j. \label{eq:fe:upvi}
+\end{equation}
+The nontrivial initial iterate for this problem is found by prolonging the coarser-level solution and using it to correct the down-smoothed iterate:
+\begin{equation}
+(z^j)^{(0)} = y^j + P z^{j-1}.  \label{eq:fe:upwardinitial}
+\end{equation}
+This is admissible since $\mathcal{U}^j = \mathcal{D}^j + \mathcal{U}^{j-1}$; see \eqref{eq:fe:constraintdecomp} or \eqref{eq:fe:upwardsumcd}.
 
-Down-smoothing makes progress toward solving the following VI for $y^j \in \mathcal{D}^j$:
+Regarding formula \eqref{eq:fe:defineg}, note that if we were to inject the bi-obstacles downward to coarser meshes, i.e.~$\underline{\gamma}^j = \iR \underline{\gamma}^{j+1}$ and $\overline{\gamma}^j = \iR \overline{\gamma}^{j+1}$, then, because injection preserves inequalities, and because $y^{j+1} \in \mathcal{D}^{j+1}$, it would follow that $\underline{\gamma}^j \le g^j \le \overline{\gamma}^j$, just like $\underline{\gamma}^J \le w^J \le \overline{\gamma}^J$ for the finest-level iterate.
+In this narrow sense $g^j$ is admissible with respect to the original (finest-level) box constraints.\footnote{In fact, while $g^j \in \cV^j \subset \cV^J$ certainly holds, generally $g^j \notin \cK^J$.}  However, following \cite{GraeserKornhuber2009}, our use of defect constraints (Table \ref{tab:ldcs}) avoids the need to generate functions $\underline{\gamma}^j$ and $\overline{\gamma}^j$, and it yields a final, finest-level iterate $g^J+z^J \in \cK^J$ which is genuinely admissible on that level.
+
+Regarding formula \eqref{eq:fe:levelsource}, the source function $\ell^j$ is determined by a classical FAS-type formula.
+(Compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.)  It has the following explanation, mimicking the PDE version from textbooks, which simultaneously justifies VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.
+
+First, the $j+1$ level problem \eqref{eq:fe:downvi} is not solved exactly by the finer-level correction $y^{j+1}$; the down-smoother was only approximate.
+Conceptually, we seek a further correction $y$ on the $j+1$ level so that
 \begin{equation}
-\ip{f^j(g^j + y^j)}{v-y^j} \ge \ip{\ell^j}{v-y^j} \qquad \text{for all } v\in \mathcal{D}^j. \label{eq:fe:downvi}
+\ip{f^{j+1}(g^{j+1}+y^{j+1}+y)}{v-y^{j+1}-y} \ge \ip{\ell^{j+1}}{v-y^{j+1}-y}, \, \forall v\in \mathcal{D}^j \quad \text{\emph{(notional)}} \label{eq:fe:downvinotional}
 \end{equation}
-Here $f^j$ is a discretization of $f$ in \eqref{eq:vi},\footnote{Again we diverge from the algorithm by Tai \cite{Tai2003}, who essentially assumes that ``$f(g^j + y^j)$'' is used here.
-If $f$ is linear then efficient implementation via that formula is possible, but otherwise one must apply $f^j$.} using quadrature etc.~as addressed for $f^J$ in \eqref{eq:fe:vi}.
-
-For $j<J$ the source function $\ell^j$ is determined by an FAS-type formula, as follows.
-First note that the $j+1$ level version of \eqref{eq:fe:downvi} is not solved exactly by the correction $y^{j+1}$.
-Fixing $y^{j+1} \in \mathcal{D}^{j+1}$, we would want to find a further correction $y$ so that $g^{j+1}+y^{j+1}+y$ is still admissible, and so that
-\begin{equation}
-\ip{f^{j+1}(g^{j+1}+y^{j+1}+y)}{v-y^{j+1}-y} \ge \ip{\ell^{j+1}}{v-y^{j+1}-y} \qquad \text{\emph{(notional)}} \label{eq:fe:downvinotional}
-\end{equation}
-held exactly.
-However, in a multilevel method we actually compute this next correction $y$ on the next-coarser level $j$.
-To do this we modify the notional finer-level VI \eqref{eq:fe:downvinotional} in five steps which mimic the standard argument for the FAS correction equations for PDEs \cite{BrandtLivne2011,Trottenbergetal2001}:
+held exactly.  However, $y$ is actually approximated from a problem on the next-coarser level.  This $j$th-level problem comes from modifying notional VI \eqref{eq:fe:downvinotional} in 4 steps:
 \begin{enumerate}
-\item group $v-y^{j+1}$ as a coarser-level test function $v\in \mathcal{D}^j$,
 \item subtract the known $f^{j+1}(g^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides,
 \item replace the residual $\ell^{j+1}-f^{j+1}(g^{j+1}+y^{j+1})$ by its restriction,
-\item replace $g^{j+1}+y^{j+1}$ where it appears by its restriction,
+\item replace $g^{j+1}+y^{j+1}$, where it appears on the left, by its restriction,
 \item and replace $f^{j+1}$ by its coarser rediscretization $f^j$ where it appears on the left.
 \end{enumerate}
-Steps \emph{(iii)}--\emph{(v)} are justified, as usual, by the assumption that the residual and the correction should be smooth as a result of the progress made by the smoother on the $j+1$ level.
-These steps now yield the precise $j$th-level VI problem, in cluttered form but the same as \eqref{eq:fe:downvi}:
+Steps \emph{(ii)}--\emph{(iv)} are justified, as for PDEs \cite[subsection 5.3.4, for example]{Trottenbergetal2001}, because the residual and the error should be smooth as a result of the progress made by the $j+1$ level down-smoother.
+These steps now yield a problem for the $j$th-level correction $y^j \in \mathcal{D}^j$:
 \begin{align}
@@
 \end{align}
+This is the same as VI problem \eqref{eq:fe:downvi}, but in cluttered form; application of definitions \eqref{eq:fe:defineg} and \eqref{eq:fe:levelsource} yields \eqref{eq:fe:downvi}.
+The explanation for VI \eqref{eq:fe:upvi} is essentially identical, but with different initial iterate and admissible set, as justified already.
 
-Thus we define $\ell^j \in (\mathcal{V}^j)'$ in \eqref{eq:fe:downvi} by the following recursive definition, starting from $\ell^J$ already defined for the finest-level problem \eqref{eq:fe:vi}, and incorporating definition \eqref{eq:fe:defineg}:
-\begin{equation}
-\ell^j = f^j(g^j) + R\left(\ell^{j+1}-f^{j+1}(g^{j+1}+y^{j+1})\right). \label{eq:fe:levelsource}
-\end{equation}
-This is identical, even if the notation differs, to the standard FAS formula for a PDE source term, for example equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.
+FIXME? Down-smoothing solves \eqref{eq:fe:downvi}, with $f^j$ is a discretization of $f$ in \eqref{eq:vi}, using quadrature etc.~as addressed for $f^J$ in \eqref{eq:fe:vi}.
+
+FIXME? smoothers generate $y^j$, $z^j$ in the given sets but these are not the exact solutions to VIs \eqref{eq:fe:downvi}, \eqref{eq:fe:upvi} resp
 
@@
 
-In the upward part of the V-cycle the smoother starts from an initial iterate which has accumulated the coarser-level corrections.  Define $z^0 = y^0$ and
-\begin{equation}
-z^j = P z^{j-1} + y^j  \label{eq:fe:upwardaccumulation}
-\end{equation}
-where $P$ is canonical prolongation; stated simply, $z^j = y^j + \dots + y^0$.
+FIXME upward initial iterate \emph{accumulates} the coarser-level corrections
+
+FIXME stated simply, $z^j = y^j + \dots + y^0$.
 By the decompositions $\mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i$, noted earlier, $z^j \in \mathcal{U}^j$.
 This initial $z^j$ is smoothed to generate a result, also denoted $z^j \in \mathcal{U}^j$.
-This result approximately solves the same VI as the downward version \eqref{eq:fe:downvi}, except now over a larger constraint set:
-\begin{equation}
-\ip{f^j(g^j + z^j)}{v-z^j} \ge \ip{\ell^j}{v-z^j} \qquad \text{for all } v\in \mathcal{U}^j. \label{eq:fe:upvi}
-\end{equation}
+This result approximately solves the same VI as the downward version \eqref{eq:fe:downvi}, except now over a larger constraint set; get \ref{eq:fe:upvi}.
 The reason we can solve over $\mathcal{U}^j$ is that, unlike during the downward phase, there is no forthcoming coarser correction which could violate admissibility.

### 2022-10-23 paper/mcd2.tex
edits through revised (box constraints) Alg 7.1

@@
 
-\section{Multilevel finite element CDs} \label{sec:multilevel}
+\section{Multilevel finite elements} \label{sec:multilevel}
 
@@
 
-The multilevel algorithm in the next section, a solver for finite-dimensional problem \eqref{eq:fe:vi}, requires transfer operators between meshes.
+The multilevel algorithm in the next two Sections, a solver for finite-dimensional problem \eqref{eq:fe:vi}, requires transfer operators between meshes.
 We use five such operators, three of which are standard within multigrid methods \cite{Trottenbergetal2001}.
@@
 
+\section{Level defect constraints and the multilevel CD} \label{sec:ldcs}
+
 Suppose $w^J \in \cK^J$ is an admissible finest-level iterate which approximates the solution $u^J$ of VI \eqref{eq:fe:vi}.
@@
 %fine level 5 (m=63): using 20 V(1,0) cycles -> 38.750 WU
-\begin{figure}
+\begin{figure}[ht]
 \includegraphics[width=0.65\textwidth]{fixfigs/chiphilevels.pdf}
@@
 using $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ to define $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.
-These new nodal functions $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ are also LDCs, and again they bracket zero,
-\begin{equation}
-\underline{\phi}^j \le 0 \le \overline{\phi}^j,  \label{eq:fe:phibrackets}
-\end{equation}
-but they are not ordered as in \eqref{eq:fe:chiordering}.  However, telescoping sums hold for $j=0,1,\dots,J$:
+These new nodal functions $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ are also LDCs, and again they bracket zero, $\underline{\phi}^j \le 0 \le \overline{\phi}^j$, but they are not ordered as in \eqref{eq:fe:chiordering}.
+However, telescoping sums hold for $j=0,1,\dots,J$:
 \begin{equation}
@@
 \medskip
-\caption{The box level defect constraints (LDCs) are constructed from the defining box constraints ($\underline{\gamma}^J,\overline{\gamma}^J$) in problem \eqref{eq:fe:vi}, and a finest level iterate ($w^J \in \cK^J$), by starting at the finest level.}
+\caption{Level defect constraints (LDCs) are constructed from the box constraints ($\underline{\gamma}^J,\overline{\gamma}^J$) in problem \eqref{eq:fe:vi}, and a finest level iterate ($w^J \in \cK^J$).}
 \label{tab:ldcs}
@@
 
-The \emph{nonlinear multilevel constraint decomposition} (NMCD) V-cycle algorithm in this section uses CDs \eqref{eq:fe:downwardsumcd} and \eqref{eq:fe:upwardsumcd}, thus applying coarse corrections from $\mathcal{D}^j$ for the downward and $\mathcal{U}^j$ for the upward directions, respectively.
+The \emph{nonlinear multilevel constraint decomposition} (NMCD) V-cycle algorithm in this section uses CDs \eqref{eq:fe:downwardsumcd} and \eqref{eq:fe:upwardsumcd}.
+It generates coarse corrections from $\mathcal{D}^j$ for the downward direction, and $\mathcal{U}^j$ for the upward direction, respectively, in the V-cycle.
 The algorithm generalizes the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear \emph{full approximation scheme} (FAS) approach \cite{BrandtLivne2011}.
-That is, we coarsen the solution iterate itself, using the injection restriction, in constrast to classical linear multigrid, wherein only the residual and correction are coarsened.
+We coarsen the solution iterate itself, using injection, in constrast to classical linear multigrid wherein only the residual and correction are coarsened.
 
-The finest-level iterate $w^J\in \mathcal{K}^J$ is corrected by perturbations from each coarser subspace $\mathcal{V}^j$---by nesting \eqref{eq:fe:nestedspaces} these corrections live in the fine-level subspace $\mathcal{V}^J$---but the corrected iterate needs to be admissible.
+The finest-level iterate $w^J = (w^J)^{(k)} \in \mathcal{K}^J$ will be corrected by perturbations from each coarser subspace $\mathcal{V}^j$---by nesting \eqref{eq:fe:nestedspaces} these corrections live in the fine-level subspace $\mathcal{V}^J$---but the corrected iterate needs to be admissible.
 In detail, suppose $y^J \in \mathcal{D}^J, y^{J-1} \in \mathcal{D}^{J-1}, \dots, y^{j+1} \in \mathcal{D}^{j+1}$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:nmcdvcycle}).
-By \eqref{eq:fe:downwardsumcd} the corrected iterate at this point is admissible, namely $y^J + \dots + y^{j+1} + 0 + \dots + 0 \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.
-Smoothing then happens in the next-coarser level, i.e.~we incompletely solve VI \eqref{eq:fe:downvi} below.
-The coarsest-level VI is then (accurately) solved to generate the downslash correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$, again by \eqref{eq:fe:downwardsumcd}.
-Starting upward, however, we can use $y^1+ z^0 \in \mathcal{U}^1$ as an initial iterate, then smooth by incompletely solving VI \eqref{eq:fe:upvi} in $\mathcal{U}^1$ (see below), yielding admissible total correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$ using \eqref{eq:fe:upwardsumcd}.
-Continuing in this way, solving in the larger upward constraint sets $\mathcal{U}^j$, eventually generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction, yielding the update $w^J \leftarrow w^J + z^J \in \mathcal{K}^J$.
+By \eqref{eq:fe:downwardsumcd} the corrected iterate at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.
+Smoothing then happens at the next-coarser level $j$, i.e.~we incompletely-solve VI problem \eqref{eq:fe:downvi} below.
+The coarsest-level VI is then (accurately) solved to generate the full downslash correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.
+Starting upward, however, we can use $y^1+ z^0 \in \mathcal{U}^1$ as an initial iterate on level $j=1$, and then smooth by incompletely-solving VI \eqref{eq:fe:upvi} in $\mathcal{U}^1$ (see below), yielding admissible total correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$ using \eqref{eq:fe:upwardsumcd}.
+Continuing in this way, up-smoothing on each level $j$, in the upward constraint sets $\mathcal{U}^j$, eventually generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction, yielding the new iterate $(w^J)^{(k+1)}  = w^J + z^J \in \mathcal{K}^J$.
 
@@
 \end{center}
-\caption{An NMCD V-cycle computes downward corrections $y_j \in \mathcal{D}^j$.  The upward corrections $z_j\in\mathcal{U}^j$ live in larger constraint subsets.}
+\caption{An NMCD V-cycle computes downward corrections $y_j \in \mathcal{D}^j$ and upward corrections $z_j\in\mathcal{U}^j$.  The latter live in larger constraint subsets.}
 \label{fig:nmcdvcycle}
@@
 A key idea, implicit in the above paragraph, is that the up-smoothing stages solve less-restricted VI problems than the down-smoothing stages.
-Intuitively-speaking, going downward one does not know what the coarse corrections will be, so they must be in sufficiently-small sets ($\mathcal{D}^j$) so that there sum is admissible once we get back to the current level.
-However, going upward the result of smoothing can be computed in the larger admissible-sum sets ($\mathcal{U}^j$).
+Intuitively-speaking, going downward one does not know what the coarse corrections will be, so they must be in such small sets ($\mathcal{D}^j$) that their sum is admissible once we return to the current level.
+However, going upward, once the coarser correction has been added, the smoothing can be computed in the larger admissible-sum sets ($\mathcal{U}^j$) because there is no forthcoming coarser correction which could violate admissibility.
 
-To state the NMCD algorithm---see Algorithm \ref{alg:nmcd} below---we must explain how to generate the coarsened problems solved at each mesh level.\footnote{Smoothers/solvers for these problems are proposed in Section \ref{sec:smoothers}.}  As in classical FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation (iterate) and the residual must be coarsened, and a but also a source functional must be created.
+To justify the NMCD algorithm---see Algorithm \ref{alg:nmcd} below---we must explain how to generate the coarsened problems solved at each mesh level.\footnote{Smoothers/solvers for these problems are proposed in Section \ref{sec:smoothers}.}  As in classical FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation (iterate) and the residual must be coarsened, and along the way a source functional is created.
+Recall that Table \ref{tab:transfers} summarizes the needed transfer operators.
+
 Let $j$ denote the mesh level of the coarsened problem.
-We assume a (re-)discretized nonlinear operator $f^j$, approximating $f$ in \eqref{eq:boxdirichletvi}, with output in $(\cV^j)'$.
-Injection $\iR$ generates the coarsened iterate (denoted $g^j$ below; note injection is commonly used for classical FAS \cite[section 5.3]{Trottenbergetal2001}), and canonical restriction $R$ the coarsened source functional (denoted $\ell^j$).
+We assume a (re-)discretized nonlinear operator $f^j$ with output in $(\cV^j)'$.
+It is an FE discretization of $f$ in \eqref{eq:boxdirichletvi}, using quadrature and etc., presumably by the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.
+Injection $\iR$ generates the coarsened iterate (denoted $g^j$ below); note injection is commonly used for classical FAS \cite[section 5.3]{Trottenbergetal2001}.
+Canonical restriction $R$ generates the coarsened source functional ($\ell^j$ below).
 
 The formulas are as follows.
-At the finest level we have problem \eqref{eq:fe:vi} defined by the discretized operator $f^J$ and the original source functional $\ell^J$.
+Finest-level problem \eqref{eq:fe:vi} is defined by the discretized operator $f^J$ and the original source functional $\ell^J$.
 A finest-level, admissible iterate $w^J \approx u^J$, $w^J \in \mathcal{K}^J$, has already\footnote{Conceptually-speaking, at least.
 In practice one generates LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ during the descent stage; see Algorithm \ref{alg:nmcd}.
-Observe that the LDCs depend neither on coarsened iterates $g^j$ nor corrections $y^j,z^j$.} been used to general the LDCs and the constraint sets $\mathcal{D}^j,\mathcal{U}^j$ at each level (Section \ref{sec:multilevel}).
+Observe that the LDCs depend neither on coarsened iterates $g^j$ nor corrections $y^j,z^j$.} been used to general the LDCs and constraint sets $\mathcal{D}^j,\mathcal{U}^j$ at each level (Section \ref{sec:multilevel}).
 As we descend in the V-cycle, we recursively define $j$th-level iterates
@@
 \end{equation}
-These are the iterates \emph{before} smoothing and/or correction has been applied at level $j$.
-(Note that the problem solved by the finer-level, down-smoother correction $y^{j+1}$ is stated momentarily.)  From $g^j$ we define
+These are the iterates \emph{before} smoothing and/or correction has been applied at level $j$.  (The problem solved by $y^{j+1}$ is stated momentarily.)  From $g^j$ we define
 \begin{equation}
@@
 \end{equation}
-On descent we solve the following VI for $y^j \in \mathcal{D}^j$:
+On descent, $j=J,J-1,\dots,1,0$, we solve the following VI for $y^j \in \mathcal{D}^j$:
 \begin{equation}
@@
 \end{equation}
-(Zero is admissible by \eqref{eq:fe:phibrackets}.)  On ascent the same VI is solved, but in a larger admissible set.  We find $z^j \in \mathcal{U}^j$ so that
+(Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)  On ascent, $j=1,2,\dots,J$, the same VI is solved, but in a larger admissible set.  We find $z^j \in \mathcal{U}^j$ so that
 \begin{equation}
@@
 \end{equation}
-The nontrivial initial iterate for this problem is found by prolonging the coarser-level solution and using it to correct the down-smoothed iterate:
+The nontrivial initial iterate for this problem is found by prolonging the coarser-level correction $z^{j-1}$, and using it to correct the down-smoothed iterate:
 \begin{equation}
@@
 \end{equation}
-This is admissible since $\mathcal{U}^j = \mathcal{D}^j + \mathcal{U}^{j-1}$; see \eqref{eq:fe:constraintdecomp} or \eqref{eq:fe:upwardsumcd}.
+(We define $z^0=y^0$ in the $j=1$ case.)  This initial iterate is admissible since $\mathcal{U}^j = \mathcal{D}^j + \mathcal{U}^{j-1}$; see \eqref{eq:fe:constraintdecomp} or \eqref{eq:fe:upwardsumcd}.
 
-Regarding formula \eqref{eq:fe:defineg}, note that if we were to inject the bi-obstacles downward to coarser meshes, i.e.~$\underline{\gamma}^j = \iR \underline{\gamma}^{j+1}$ and $\overline{\gamma}^j = \iR \overline{\gamma}^{j+1}$, then, because injection preserves inequalities, and because $y^{j+1} \in \mathcal{D}^{j+1}$, it would follow that $\underline{\gamma}^j \le g^j \le \overline{\gamma}^j$, just like $\underline{\gamma}^J \le w^J \le \overline{\gamma}^J$ for the finest-level iterate.
-In this narrow sense $g^j$ is admissible with respect to the original (finest-level) box constraints.\footnote{In fact, while $g^j \in \cV^j \subset \cV^J$ certainly holds, generally $g^j \notin \cK^J$.}  However, following \cite{GraeserKornhuber2009}, our use of defect constraints (Table \ref{tab:ldcs}) avoids the need to generate functions $\underline{\gamma}^j$ and $\overline{\gamma}^j$, and it yields a final, finest-level iterate $g^J+z^J \in \cK^J$ which is genuinely admissible on that level.
+Regarding formula \eqref{eq:fe:defineg}, if we were to inject the bi-obstacles downward to coarser meshes, i.e.~$\underline{\gamma}^j = \iR \underline{\gamma}^{j+1}$ and $\overline{\gamma}^j = \iR \overline{\gamma}^{j+1}$, then, because injection preserves inequalities, and because $y^{j+1} \in \mathcal{D}^{j+1}$, it would follow that $\underline{\gamma}^j \le g^j \le \overline{\gamma}^j$, just like $\underline{\gamma}^J \le w^J \le \overline{\gamma}^J$ for the finest-level iterate.
+In this narrow sense $g^j$ is admissible with respect to the original (finest-level) box constraints.\footnote{In fact, while $g^j \in \cV^j \subset \cV^J$ certainly holds, generally $g^j \notin \cK^J$.
+As noted earlier, even the finest-level FE VI \eqref{eq:fe:vi} is not generally conforming for the corresponding continuum VI \eqref{eq:boxdirichletvi}.}  However, following \cite{GraeserKornhuber2009}, our use of defect constraints (Table \ref{tab:ldcs}) avoids the need to generate functions $\underline{\gamma}^j$ and $\overline{\gamma}^j$, and it yields a final, finest-level iterate $g^J+z^J \in \cK^J$ which is genuinely admissible for problem \eqref{eq:fe:vi}.
 
 Regarding formula \eqref{eq:fe:levelsource}, the source function $\ell^j$ is determined by a classical FAS-type formula.
-(Compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.)  It has the following explanation, mimicking the PDE version from textbooks, which simultaneously justifies VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.
+(Compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.)  It has the following explanation, mimicking the PDE story from textbooks, which simultaneously justifies formula \eqref{eq:fe:levelsource} and VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.
 
@@
 \end{equation}
-held exactly.  However, $y$ is actually approximated from a problem on the next-coarser level.  This $j$th-level problem comes from modifying notional VI \eqref{eq:fe:downvinotional} in 4 steps:
+holds exactly.  However, $y$ will actually be approximated from a problem on the next-coarser level.  This $j$th-level problem comes from modifying notional VI \eqref{eq:fe:downvinotional} in 4 steps:
 \begin{enumerate}
-\item subtract the known $f^{j+1}(g^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides,
-\item replace the residual $\ell^{j+1}-f^{j+1}(g^{j+1}+y^{j+1})$ by its restriction,
+\item subtract the known quantity $f^{j+1}(g^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides,
+\item replace the residual $\ell^{j+1}-f^{j+1}(g^{j+1}+y^{j+1})$, now on the right, by its restriction,
 \item replace $g^{j+1}+y^{j+1}$, where it appears on the left, by its restriction,
@@
 \end{align}
-This is the same as VI problem \eqref{eq:fe:downvi}, but in cluttered form; application of definitions \eqref{eq:fe:defineg} and \eqref{eq:fe:levelsource} yields \eqref{eq:fe:downvi}.
+This is the same as VI problem \eqref{eq:fe:downvi}, but in cluttered form; application of definitions \eqref{eq:fe:defineg} and \eqref{eq:fe:levelsource} yields clean form \eqref{eq:fe:downvi}.
+
 The explanation for VI \eqref{eq:fe:upvi} is essentially identical, but with different initial iterate and admissible set, as justified already.
 
-FIXME? Down-smoothing solves \eqref{eq:fe:downvi}, with $f^j$ is a discretization of $f$ in \eqref{eq:vi}, using quadrature etc.~as addressed for $f^J$ in \eqref{eq:fe:vi}.
+As is standard in multigrid, at the coarsest $j=0$ level we solve problem \eqref{eq:fe:downvi} accurately if possible.
+This coarsest-level solver may be the same as the other smoothers, perhaps with a stronger tolerance.
 
-FIXME? smoothers generate $y^j$, $z^j$ in the given sets but these are not the exact solutions to VIs \eqref{eq:fe:downvi}, \eqref{eq:fe:upvi} resp
+These ideas come together in Algorithm \ref{alg:nmcd}.  This procedure acts in-place on $w^J$, but it leaves the other inputs unchanged.
 
-At each level in the downward part of the V-cycle (Figure \ref{fig:nmcdvcycle}) we apply a smoother, an iterative partial solver of \eqref{eq:fe:downvi} which rapidly reduces the high-frequencies in the error.
-This smoother yields a correction $y^j \in \mathcal{D}^j$.
-It is important that we can initialize this iteration using an admissible first guess.
-However, by the monotone restriction construction the down-constraint is nonpositive, $\phi^j\le 0$ so $0\in \mathcal{D}^j$, and thus an initial value of $y^j=0$ is admissible.
-
-At the coarsest $j=0$ level we solve \eqref{eq:fe:downvi}, accurately if possible, yielding a correction $y^0$ in the coarsest defect constraint set $\mathcal{D}^0=\mathcal{U}^0 = \{v \ge \chi^0\}$.
-Note that the coarsest-level solver may be the same as the other smoothers.
-
-FIXME upward initial iterate \emph{accumulates} the coarser-level corrections
-
-FIXME stated simply, $z^j = y^j + \dots + y^0$.
-By the decompositions $\mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i$, noted earlier, $z^j \in \mathcal{U}^j$.
-This initial $z^j$ is smoothed to generate a result, also denoted $z^j \in \mathcal{U}^j$.
-This result approximately solves the same VI as the downward version \eqref{eq:fe:downvi}, except now over a larger constraint set; get \ref{eq:fe:upvi}.
-The reason we can solve over $\mathcal{U}^j$ is that, unlike during the downward phase, there is no forthcoming coarser correction which could violate admissibility.
-
-These ideas come together in Algorithm \ref{alg:nmcd}, an NMCD V-cycle.  This procedure acts in-place on the final argument $w^J$, but it leaves the other vector inputs $\ell^J,\gamma^J$ unchanged.
-
-\begin{pseudofloat}[H]
+\begin{pseudofloat}[ht]
 \begin{pseudo} \label{ps:nmcd-vcycle}
-\pr{nmcd-vcycle}(\ell^J,\gamma^J;w^J)\text{:} \\+
-    $\chi^J = \gamma^J - w^J$ \\
+\pr{nmcd-vcycle}(\ell^J,\underline{\gamma}^J,\overline{\gamma}^J;w^J)\text{:} \\+
+    $\underline{\chi}^J = \underline{\gamma}^J - w^J {\large \strut}$ \\
+    $\overline{\chi}^J = \overline{\gamma}^J - w^J {\large \strut}$ \\
     $g^J = w^J$ \\
     for $j=J$ downto $j=1$ \\+
-      $\chi^{j-1} = \maxR \chi^j$ \\
-      $\phi^j = \chi^j - P\chi^{j-1}$ \\
+      $\underline{\chi}^{j-1} = \maxR \underline{\chi}^j {\large \strut}$ \\
+      $\overline{\chi}^{j-1} = \minR \overline{\chi}^j {\large \strut}$ \\
+      $\underline{\phi}^j = \underline{\chi}^j - P\underline{\chi}^{j-1} {\large \strut}$ \\
+      $\overline{\phi}^j = \overline{\chi}^j - P\overline{\chi}^{j-1} {\large \strut}$ \\
       $y^j = 0$ \\
-      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\phi^j,g^j;y^j)$  \ct{smoothing of \eqref{eq:fe:downvi} in $\mathcal{D}^j$}\\
+      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\underline{\phi}^j,\overline{\phi}^j,g^j;y^j)$  \ct{smoothing of \eqref{eq:fe:downvi} in $\mathcal{D}^j$}\\
       $g^{j-1} = \iR(g^j + y^j)$ \\
@@
     $y^0 = 0$ \\
-    $\text{\pr{solve}}(\ell^0,\chi^0,g^0;y^0)$ \ct{solving of \eqref{eq:fe:downvi} in $\mathcal{D}^0=\mathcal{U}^0$} \\
-    $z^0 = y^0$ \\
+    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,\overline{\chi}^0,g^0;y^0)$ \hspace{1.0cm} \ct{solving of \eqref{eq:fe:downvi} in $\mathcal{D}^0=\mathcal{U}^0$} \\
     for $j=1$ to $j=J$ \\+
-      $z^j = P z^{j-1} + y^{j}$ \\
-      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\chi^j,g^j;z^j)$  \ct{smoothing of \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
+      $z^j = y^{j} + P z^{j-1}$ \\
+      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,\overline{\chi}^j,g^j;z^j)$  \ct{smoothing of \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
     $w^J \gets w^J+z^J$
 \end{pseudo}
-\caption{Nonlinear multilevel constraint decomposition V-cycle for the finest-level VI problem \eqref{eq:fe:vi}, as an in-place method on $w^J$.
-$f^j$ denotes a discretization of $f$ in problem \eqref{eq:vi}.}
+\caption{Nonlinear multilevel constraint decomposition V-cycle as an iterative solver for VI problem \eqref{eq:fe:vi}.  $f^j$ denotes a discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
 \label{alg:nmcd}

### 2022-10-25 paper/mcd2.tex
edits re the FE problem

@@
 We restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.
-Suppose that the \emph{lower obstacle} $\underline{\gamma}$ and the \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions on $\overline{\Omega}$ which take their values in the extended real line $\tilde\RR = [-\infty,+\infty]$.
+The \emph{lower obstacle} $\underline{\gamma}$ and the \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions on $\overline{\Omega}$ which take their values in the extended real line $\tilde\RR = [-\infty,+\infty]$.
 Specifically suppose $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$ and $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$ are measurable, satisfying $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$ and $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.
@@
 (The boundary values are, as usual, in the trace sense \cite{Evans2010}.)  It is easy to check that $\cK$ is closed and convex.
-Observe that this definition of $\cK$ treats the inequality and equality (boundary) constraints in a unified manner.
+This definition of $\cK$ treats the inequality and equality (boundary) constraints in a unified manner.
 
-We will solve the same VI as in the abstract problem \eqref{eq:vi}.  Given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we assume that the following VI is well-posed for $u^*\in \cK$:
+We will solve the same VI as in abstract problem \eqref{eq:vi}.  Given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we assume that the following VI is well-posed for $u^*\in \cK$:
 \begin{equation}
@@
 \end{equation}
-Note that $v-u^* \in W_0^{1,p}(\Omega)$, that is, the test functions in \eqref{eq:boxdirichletvi} have zero values on the Dirichlet boundary $\partial_D\Omega$.
+Note that $v-u^* \in W_0^{1,p}(\Omega)$, that is, the test functions used in \eqref{eq:boxdirichletvi} have zero values on the Dirichlet boundary $\partial_D\Omega$.
 
-Problem \eqref{eq:boxdirichletvi} is both a free-boundary problem, because of the inequality constraints in defining $\cK$, and a mixed fixed-boundary value problem.
+Problem \eqref{eq:boxdirichletvi} is both a free-boundary problem, because of the inequality constraints in defining $\cK$, and a (mixed) fixed-boundary value problem.
 Dirichlet conditions $u=g_D$ apply over $\partial_D\Omega$ and Neumann conditions over $\partial_N \Omega$.
-The latter are determined by the formula for $f$; nonhomogeneous Neumann conditions require that the nonlinear operator $f:\cK\to\cV'$ include an integration over $\partial_N\Omega$.
+The latter are determined by the formula for $f$; for nonhomogeneous Neumann conditions, the nonlinear operator $f:\cK\to\cV'$ will include an integration over $\partial_N\Omega$.
 
 We now construct a multilevel finite element (FE) approximation to problem \eqref{eq:boxdirichletvi} using either triangular or quadrilateral meshes over $\Omega$, and $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, respectively.
-The targeted solution is hosted on the finest-level mesh in a hierarchy of meshes constructed by refinement of a coarsest mesh.
-This hierarchy will also generate a multilevel CD of $\cK$.
+The solution on the finest-level mesh, in a hierarchy of meshes constructed by refinement of a coarsest mesh, is the target.
+This hierarchy also generates a multilevel CD of $\cK$.
 The subsets $\cK_i$ in this CD are designed to allow multigrid-type coarser-level corrections (next Section) in a manner which preserves admissibility, while not introducing unnecessary high frequencies into the coarser-level problems, especially at the free boundary.
@@
 Let $m_j$ be the number of nodes in $\mathcal{T}^j$, the \emph{degrees of freedom} at the $j$th level, and denote $h=h_J=2^{-J} h_0$ and $m=m_J$ for the finest level.
-
 Let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space \cite{Elmanetal2014} over $\mathcal{T}^j$, with arbitrary boundary values, and observe the following vector space nesting:
@@
 
-In the following Sections, use of higher-order Lagrange elements $P_k$ or $Q_k$, $k\ge 2$, causes a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} specific to inequality constraints, and for this reason we restrict to $k\le 1$.
+Use of higher-order Lagrange elements $P_k$ or $Q_k$, $k\ge 2$, would be a ``variational crime'' \cite[Chapter 10]{BrennerScott2007}, specific to inequality constraints, in the upcoming material.
+For this reason we restrict to $k = 1$.
 Specifically, our FE construction of constraint sets will require \emph{nodal monotonicity}, namely that for all FE functions $w,z \in \mathcal{V}^j$,
@@
 
-To allow box constraints, let $\tilde{\mathcal{V}}^j$ denote the set of all functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
-(Note that $\tilde{\mathcal{V}}^j$ functions are not in $\mathcal{V}^j$, and they are not even defined everywhere on $\Omega$.)  Suppose that the \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$, defined on the nodes of the finest-level mesh $\mathcal{T}^J$, satisfy
+To accomodate the fixed-boundary conditions, let $g_D^J \in \mathcal{V}^J$, on the finest-level mesh, satisfy $g_D^J = g_D$ on $\partial_D \Omega$.\footnote{As is standard in conforming FE methods, this restricts the original Dirichlet boundary values $g_D$ to be a piecewise-linear function \cite{Elmanetal2014}.}  To allow box constraints, let $\tilde{\mathcal{V}}^j$ denote the set of all functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
+(Note that $\tilde{\mathcal{V}}^j$ functions are not in $\mathcal{V}^j$.
+They are not even defined everywhere on $\Omega$, but only on nodes.)  Suppose that the \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$, defined on the nodes of the finest-level mesh $\mathcal{T}^J$, satisfy
 \begin{equation}
@@
 Where $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$ then there is no lower or upper constraint, respectively.
-
-To accomodate the fixed-boundary conditions, let $g_D^J \in \mathcal{V}^J$, on the finest-level mesh, satisfy $g_D^J = g_D$ on $\partial_D \Omega$.\footnote{As is standard in conforming FE methods, this restricts the original Dirichlet boundary values $g_D$ to a piecewise-linear function \cite{Elmanetal2014}.}  Suppose furthermore that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$, as otherwise the constraint set, defined next, is empty.
+Finally assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.
 
@@
 Let $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ be a discretization of $f$ in \eqref{eq:boxdirichletvi}.
-Noting that by inclusion \eqref{eq:fe:nestedspaces} the functional $\ell$ is already defined and continuous over $\mathcal{V}^J$, we denote $\ell^J \in (\mathcal{V}^J)'$ for its FE representation.
+By inclusion \eqref{eq:fe:nestedspaces} the functional $\ell$ is already defined and continuous over $\mathcal{V}^J$, but we denote $\ell^J \in (\mathcal{V}^J)'$ for its FE representation.
 We seek the FE solution $u^J \in \mathcal{K}^J$ of the finest-level, finite-dimensional, nonlinear VI
@@
 \end{equation}
-As for \eqref{eq:vi}, if $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, then \eqref{eq:fe:vi} is well-posed.
+As is true for \eqref{eq:vi}, if $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, then \eqref{eq:fe:vi} is well-posed.
 
 While \eqref{eq:fe:vi} approximates \eqref{eq:boxdirichletvi}, it may not technically be conforming for that VI.
-That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require $\mathcal{K}^J \subset \mathcal{K}$.
-The continuum constraint set $\mathcal{K}$ remains abstract in applications where only the nodal obstacles $\underline{\gamma}^J, \overline{\gamma}^J$ are defined.
+That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we still do not require $\mathcal{K}^J \subset \mathcal{K}$.
+The continuum constraint set $\mathcal{K}$ remains abstract in those applications where only the nodal obstacles $\underline{\gamma}^J, \overline{\gamma}^J$ are defined.
 In fact, even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants will generally not imply continuum admissiblity.\footnote{Such caveats apply in geophysical situations where the obstacle is given by meshed elevation data, the continuum limit of which is generally unknown \cite{Bueler2016}.}
+
+Finite-dimensional, box-constrained VI problem \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003,FerrisPang1997}.
+Stating this strong-form problem may be clarifying for some readers.
+Denoting $x_p$ for nodes of $\mathcal{T}^J$, let $X=\left\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\right\}$ so $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap X$.
+Problem \eqref{eq:fe:vi} might be stated as VI$\left(f^J(\cdot)-\ell^J,\left[\underline{\gamma}^J,\overline{\gamma}^J\right] \cap X\right)$ in the operations research literature \cite{FerrisPang1997}, and it is equivalent to the equality-constrained MCP
+\begin{align}
+\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J + I_D^\top \lambda \label{eq:fe:mcp} \\
+I_D u^J &= g_D^J\big|_{\partial_D\Omega}. \notag
+\end{align}
+Here the $m_J\times m_D$ matrix $I_D$ indicates Dirichlet nodes, $m_D$ is the number of $\mathcal{T}^J$ nodes in the Dirichlet boundary, and the Lagrange multipliers $\lambda \in \RR^{m_D}$ are free.
+(The transpose $I_D^\top$ injects Dirichlet values into $(\mathcal{V}^J)'$.)  That is, at each node $x_p \in \mathcal{T}^J$ in the finest-level mesh, exactly one of the following conditions holds for the exact solution $u^J$ of \eqref{eq:fe:vi}, equivalently \eqref{eq:fe:mcp}:
+\begin{itemize}
+\item FIXME
+\end{itemize}
 

### 2022-10-25 paper/mcd2.tex
I think this is stating the MCP the right way

@@
 Stating this strong-form problem may be clarifying for some readers.
-Denoting $x_p$ for nodes of $\mathcal{T}^J$, let $X=\left\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\right\}$ so $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap X$.
-Problem \eqref{eq:fe:vi} might be stated as VI$\left(f^J(\cdot)-\ell^J,\left[\underline{\gamma}^J,\overline{\gamma}^J\right] \cap X\right)$ in the operations research literature \cite{FerrisPang1997}, and it is equivalent to the equality-constrained MCP
+Denoting $x_p$ for nodes of the finest-level mesh $\mathcal{T}^J$, let $X=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap X$.
+If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let us define $I_D$, an the $m_D\times m_J$ matrix, as the matrix which indicates (extracts) Dirichlet nodal values, so its transpose $I_D^\top$ injects Dirichlet values into $(\mathcal{V}^J)'$.
+Problem \eqref{eq:fe:vi} could be stated as VI$\left(f^J(\cdot)-\ell^J,\left[\underline{\gamma}^J,\overline{\gamma}^J\right] \cap X\right)$, for instance in the operations research literature \cite{FerrisPang1997}, and it is equivalent to the equality-constrained MCP
 \begin{align}
-\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J + I_D^\top \lambda \label{eq:fe:mcp} \\
+\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J - I_D^\top \lambda \label{eq:fe:mcp} \\
 I_D u^J &= g_D^J\big|_{\partial_D\Omega}. \notag
 \end{align}
-Here the $m_J\times m_D$ matrix $I_D$ indicates Dirichlet nodes, $m_D$ is the number of $\mathcal{T}^J$ nodes in the Dirichlet boundary, and the Lagrange multipliers $\lambda \in \RR^{m_D}$ are free.
-(The transpose $I_D^\top$ injects Dirichlet values into $(\mathcal{V}^J)'$.)  That is, at each node $x_p \in \mathcal{T}^J$ in the finest-level mesh, exactly one of the following conditions holds for the exact solution $u^J$ of \eqref{eq:fe:vi}, equivalently \eqref{eq:fe:mcp}:
+The Lagrange multipliers $\lambda \in \RR^{m_D}$ are free (unconstrained).
+Let $\psi_p(x)$ denote the hat function for a node $x_p \in \mathcal{T}^J$.
+MCP \eqref{eq:fe:mcp} says that at each node exactly\footnote{The stated conditions are actually exclusive only where $\underline{\gamma}^J(x_p) < \overline{\gamma}^J(x_p)$.
+Where the box constraints meet at $x_p \notin \partial_D\Omega$, both of the first conditions hold.} one of the following conditions holds for the solution $u^J$:
 \begin{itemize}
-\item FIXME
+\item $x_p \notin \partial_D\Omega$ and $u^J(x_p)=\underline{\gamma}^J(x_p)$ and $f^J(u^J)(\psi_p)-\ell^J(\psi_p) \ge 0$, or
+\item $x_p \notin \partial_D\Omega$ and $u^J(x_p)=\overline{\gamma}^J(x_p)$ and $f^J(u^J)(\psi_p)-\ell^J(\psi_p) \le 0$, or
+\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)<u^J(x_p)<\overline{\gamma}^J(x_p)$ and $f^J(u^J)(\psi_p)-\ell^J(\psi_p) = 0$, or
+\item $x_p \in \partial_D\Omega$ and $u^J(x_p)=g_D^J(x_p)$.
 \end{itemize}
+In the final case $\lambda_p=f^J(u^J)(\psi_p)-\ell^J(\psi_p)$ has any value.
 
-The multilevel algorithm in the next two Sections, a solver for finite-dimensional problem \eqref{eq:fe:vi}, requires transfer operators between meshes.
+The multilevel algorithm defined in the next two Sections, a solver for finite-dimensional problem \eqref{eq:fe:vi}, requires transfer operators between meshes.
 We use five such operators, three of which are standard within multigrid methods \cite{Trottenbergetal2001}.

### 2022-10-25 paper/mcd2.tex
edits on the NMCD V-cycle, including rewrite of storage estimate

@@
 Denoting $x_p$ for nodes of the finest-level mesh $\mathcal{T}^J$, let $X=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap X$.
-If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let us define $I_D$, an the $m_D\times m_J$ matrix, as the matrix which indicates (extracts) Dirichlet nodal values, so its transpose $I_D^\top$ injects Dirichlet values into $(\mathcal{V}^J)'$.
+If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let us define $I_D$ as the $m_D\times m_J$ matrix which indicates (extracts) Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet values as nodes in $(\mathcal{V}^J)'$.
 Problem \eqref{eq:fe:vi} could be stated as VI$\left(f^J(\cdot)-\ell^J,\left[\underline{\gamma}^J,\overline{\gamma}^J\right] \cap X\right)$, for instance in the operations research literature \cite{FerrisPang1997}, and it is equivalent to the equality-constrained MCP
@@
 \end{itemize}
-In the final case $\lambda_p=f^J(u^J)(\psi_p)-\ell^J(\psi_p)$ has any value.
+Regarding the final case, $\lambda_p=f^J(u^J)(\psi_p)-\ell^J(\psi_p)$ has any value if $x_p \in \partial_D\Omega$.
 
@@
 The \emph{nonlinear multilevel constraint decomposition} (NMCD) V-cycle algorithm in this section uses CDs \eqref{eq:fe:downwardsumcd} and \eqref{eq:fe:upwardsumcd}.
-It generates coarse corrections from $\mathcal{D}^j$ for the downward direction, and $\mathcal{U}^j$ for the upward direction, respectively, in the V-cycle.
+It does smoothing in $\mathcal{D}^j$ for the downward direction, and $\mathcal{U}^j$ for the upward direction, respectively, in the V-cycle.
 The algorithm generalizes the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear \emph{full approximation scheme} (FAS) approach \cite{BrandtLivne2011}.
-We coarsen the solution iterate itself, using injection, in constrast to classical linear multigrid wherein only the residual and correction are coarsened.
+We coarsen the solution iterate itself, using injection, in constrast to classical linear multigrid wherein only the residual is coarsened.
 
-The finest-level iterate $w^J = (w^J)^{(k)} \in \mathcal{K}^J$ will be corrected by perturbations from each coarser subspace $\mathcal{V}^j$---by nesting \eqref{eq:fe:nestedspaces} these corrections live in the fine-level subspace $\mathcal{V}^J$---but the corrected iterate needs to be admissible.
+The finest-level iterate $w^J \in \mathcal{K}^J$ will be corrected by perturbations from each coarser subspace $\mathcal{V}^j$,\footnote{By nesting \eqref{eq:fe:nestedspaces} these corrections also live in the fine-level subspace $\mathcal{V}^J$.
+Later we will use prolongation $P$ to indicate where a change in computer representation would be needed.} but the corrected iterate needs to be admissible.
 In detail, suppose $y^J \in \mathcal{D}^J, y^{J-1} \in \mathcal{D}^{J-1}, \dots, y^{j+1} \in \mathcal{D}^{j+1}$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:nmcdvcycle}).
@@
 Smoothing then happens at the next-coarser level $j$, i.e.~we incompletely-solve VI problem \eqref{eq:fe:downvi} below.
-The coarsest-level VI is then (accurately) solved to generate the full downslash correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.
-Starting upward, however, we can use $y^1+ z^0 \in \mathcal{U}^1$ as an initial iterate on level $j=1$, and then smooth by incompletely-solving VI \eqref{eq:fe:upvi} in $\mathcal{U}^1$ (see below), yielding admissible total correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$ using \eqref{eq:fe:upwardsumcd}.
-Continuing in this way, up-smoothing on each level $j$, in the upward constraint sets $\mathcal{U}^j$, eventually generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction, yielding the new iterate $(w^J)^{(k+1)}  = w^J + z^J \in \mathcal{K}^J$.
+At the bottom the coarsest-level VI is then (accurately) solved, to generate the full downslash correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.
+Starting upward, however, we can use $y^1+ z^0 \in \mathcal{U}^1$ as an initial iterate on level $j=1$, and then smooth by incompletely-solving VI problem \eqref{eq:fe:upvi} (below) in $\mathcal{U}^1$, yielding admissible total correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$ using \eqref{eq:fe:upwardsumcd}.
+Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$ from initial iterate $y^j+z^{j-1}$, eventually generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction, yielding the new iterate $w^J + z^J \in \mathcal{K}^J$.
 
@@
 
-A key idea, implicit in the above paragraph, is that the up-smoothing stages solve less-restricted VI problems than the down-smoothing stages.
-Intuitively-speaking, going downward one does not know what the coarse corrections will be, so they must be in such small sets ($\mathcal{D}^j$) that their sum is admissible once we return to the current level.
-However, going upward, once the coarser correction has been added, the smoothing can be computed in the larger admissible-sum sets ($\mathcal{U}^j$) because there is no forthcoming coarser correction which could violate admissibility.
+A key idea, implicit in the above paragraph, is that the up-smoothing stages solve less-restricted VI problems than down-smoothing stages.
+Intuitively-speaking, downward one does not know what the coarse corrections will be, so they must be in such small sets ($\mathcal{D}^j$) that their sum is admissible once we return to the current level.
+However, going upward smoothing can occur in the larger admissible-sum sets ($\mathcal{U}^j$; see \eqref{eq:fe:constraintdecomp}) because there are no forthcoming coarser corrections which could violate admissibility.
 
-To justify the NMCD algorithm---see Algorithm \ref{alg:nmcd} below---we must explain how to generate the coarsened problems solved at each mesh level.\footnote{Smoothers/solvers for these problems are proposed in Section \ref{sec:smoothers}.}  As in classical FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation (iterate) and the residual must be coarsened, and along the way a source functional is created.
+To justify the NMCD algorithm---see Algorithm \ref{alg:nmcd} below---we must state the problems solved at each mesh level.\footnote{Smoothers/solvers for these problems are proposed in Section \ref{sec:smoothers}.}  As in classical FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, along with creating a source functional.
 Recall that Table \ref{tab:transfers} summarizes the needed transfer operators.
@@
 Let $j$ denote the mesh level of the coarsened problem.
-We assume a (re-)discretized nonlinear operator $f^j$ with output in $(\cV^j)'$.
-It is an FE discretization of $f$ in \eqref{eq:boxdirichletvi}, using quadrature and etc., presumably by the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.
-Injection $\iR$ generates the coarsened iterate (denoted $g^j$ below); note injection is commonly used for classical FAS \cite[section 5.3]{Trottenbergetal2001}.
-Canonical restriction $R$ generates the coarsened source functional ($\ell^j$ below).
+We assume a (re-)discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, an FE discretization of $f$ in \eqref{eq:boxdirichletvi}, using quadrature and etc., presumably by the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.
 
@@
 \end{equation}
-These are the iterates \emph{before} smoothing and/or correction has been applied at level $j$.  (The problem solved by $y^{j+1}$ is stated momentarily.)  From $g^j$ we define
+These are the iterates \emph{before} smoothing and/or correction has been applied at level $j$.  (The problem solved by $y^{j+1}$ is stated momentarily.)  From $g^j$ we define $\ell^j \in (\cV^j)'$:
 \begin{equation}
@@
 \end{equation}
-On descent, $j=J,J-1,\dots,1,0$, we solve the following VI for $y^j \in \mathcal{D}^j$:
+Observe that injection $\iR$ coarsens the iterates $g^j$---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---while canonical restriction $R$ coarsens the source functionals.
+
+On descent, $j=J,J-1,\dots,1$, we solve the following VI for $y^j \in \mathcal{D}^j$:
 \begin{equation}
@@
 \end{equation}
-(Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)  On ascent, $j=1,2,\dots,J$, the same VI is solved, but in a larger admissible set.  We find $z^j \in \mathcal{U}^j$ so that
+(Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)  On ascent, $j=0,1,\dots,J$, the same VI is solved, but in a larger admissible set.  We find $z^j \in \mathcal{U}^j$ so that
 \begin{equation}
@@
 \end{equation}
-The nontrivial initial iterate for this problem is found by prolonging the coarser-level correction $z^{j-1}$, and using it to correct the down-smoothed iterate:
+The nontrivial initial iterate for this problem is found by prolonging the coarser-level correction $z^{j-1}$, and using it to correct the current (down-smoothed) iterate on the level:\footnote{Prolongation $P$ is used in \eqref{eq:fe:upwardinitial} to reflect the change in mesh/FE representation, from $\mathcal{T}^{j-1}$ to $\mathcal{T}^{j}$.}
 \begin{equation}
@@
 \end{equation}
-(We define $z^0=y^0$ in the $j=1$ case.)  This initial iterate is admissible since $\mathcal{U}^j = \mathcal{D}^j + \mathcal{U}^{j-1}$; see \eqref{eq:fe:constraintdecomp} or \eqref{eq:fe:upwardsumcd}.
+For the $j=0$ coarsest level, $(z^0)^{(0)}=0$.
+Initial iterate \eqref{eq:fe:upwardinitial} is admissible since $\mathcal{U}^j = \mathcal{D}^j + \mathcal{U}^{j-1}$; see \eqref{eq:fe:constraintdecomp} or \eqref{eq:fe:upwardsumcd}.
 
-Regarding formula \eqref{eq:fe:defineg}, if we were to inject the bi-obstacles downward to coarser meshes, i.e.~$\underline{\gamma}^j = \iR \underline{\gamma}^{j+1}$ and $\overline{\gamma}^j = \iR \overline{\gamma}^{j+1}$, then, because injection preserves inequalities, and because $y^{j+1} \in \mathcal{D}^{j+1}$, it would follow that $\underline{\gamma}^j \le g^j \le \overline{\gamma}^j$, just like $\underline{\gamma}^J \le w^J \le \overline{\gamma}^J$ for the finest-level iterate.
-In this narrow sense $g^j$ is admissible with respect to the original (finest-level) box constraints.\footnote{In fact, while $g^j \in \cV^j \subset \cV^J$ certainly holds, generally $g^j \notin \cK^J$.
-As noted earlier, even the finest-level FE VI \eqref{eq:fe:vi} is not generally conforming for the corresponding continuum VI \eqref{eq:boxdirichletvi}.}  However, following \cite{GraeserKornhuber2009}, our use of defect constraints (Table \ref{tab:ldcs}) avoids the need to generate functions $\underline{\gamma}^j$ and $\overline{\gamma}^j$, and it yields a final, finest-level iterate $g^J+z^J \in \cK^J$ which is genuinely admissible for problem \eqref{eq:fe:vi}.
+If we were to inject the original bi-obstacles from \eqref{eq:fe:fineconstraintset} downward to coarser meshes, i.e.~$\underline{\gamma}^j = \iR \underline{\gamma}^{j+1}$ and $\overline{\gamma}^j = \iR \overline{\gamma}^{j+1}$, then, because injection preserves inequalities, and because $y^{j+1} \in \mathcal{D}^{j+1}$, it would follow from \eqref{eq:fe:defineg} that $\underline{\gamma}^j \le g^j \le \overline{\gamma}^j$, just like $\underline{\gamma}^J \le w^J \le \overline{\gamma}^J$ for the finest-level iterate.
+In this narrow sense $g^j$ is admissible with respect to the original (finest-level) box constraints.
+(In fact, while $g^j \in \cV^j \subset \cV^J$ certainly holds, generally $g^j \notin \cK^J$.)  However, following \cite{GraeserKornhuber2009}, our use of defect constraints (Table \ref{tab:ldcs}) avoids the need to generate functions $\underline{\gamma}^j$ and $\overline{\gamma}^j$, and it yields a final, finest-level iterate $g^J+z^J \in \cK^J$ which is genuinely admissible for problem \eqref{eq:fe:vi}.
 
 Regarding formula \eqref{eq:fe:levelsource}, the source function $\ell^j$ is determined by a classical FAS-type formula.
-(Compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.)  It has the following explanation, mimicking the PDE story from textbooks, which simultaneously justifies formula \eqref{eq:fe:levelsource} and VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.
-
+(Compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.)  It has the following explanation, mimicking the PDE story from textbooks, which simultaneously justifies \eqref{eq:fe:levelsource} and VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.
 First, the $j+1$ level problem \eqref{eq:fe:downvi} is not solved exactly by the finer-level correction $y^{j+1}$; the down-smoother was only approximate.
@@
 \begin{enumerate}
-\item subtract the known quantity $f^{j+1}(g^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides,
+\item subtract the computable quantity $f^{j+1}(g^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides,
 \item replace the residual $\ell^{j+1}-f^{j+1}(g^{j+1}+y^{j+1})$, now on the right, by its restriction,
 \item replace $g^{j+1}+y^{j+1}$, where it appears on the left, by its restriction,
-\item and replace $f^{j+1}$ by its coarser rediscretization $f^j$ where it appears on the left.
+\item and, where it appears on the left, replace $f^{j+1}$ by its coarser rediscretization $f^j$.
 \end{enumerate}
@@
 \end{align}
-This is the same as VI problem \eqref{eq:fe:downvi}, but in cluttered form; application of definitions \eqref{eq:fe:defineg} and \eqref{eq:fe:levelsource} yields clean form \eqref{eq:fe:downvi}.
+This is the same as VI problem \eqref{eq:fe:downvi}, but in cluttered form; application of definitions \eqref{eq:fe:defineg} and \eqref{eq:fe:levelsource} yields \eqref{eq:fe:downvi}.
+The explanation for VI \eqref{eq:fe:upvi} is essentially identical.
 
-The explanation for VI \eqref{eq:fe:upvi} is essentially identical, but with different initial iterate and admissible set, as justified already.
+We will regard the smoothers for VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} as approximate, in-place procedures, acting on the corrections $y^j$ and $z^j$, respectively.
+Also, as is standard in multigrid, at the coarsest $j=0$ level we suppose that problem \eqref{eq:fe:upvi} is solved quite accurately.
+However, this coarsest-level solver may be the same as the other smoothers, perhaps with a stronger tolerance.
 
-As is standard in multigrid, at the coarsest $j=0$ level we solve problem \eqref{eq:fe:downvi} accurately if possible.
-This coarsest-level solver may be the same as the other smoothers, perhaps with a stronger tolerance.
-
-These ideas come together in Algorithm \ref{alg:nmcd}.  This procedure acts in-place on $w^J$, but it leaves the other inputs unchanged.
+These ideas come together in Algorithm \ref{alg:nmcd}, below.  The \pr{nmcd-vcycle} procedure acts in-place on $w^J$, but it leaves its other inputs unchanged.
 
@@
       $\ell^{j-1} = f^{j-1}(g^{j-1}) + R \left(\ell^j - f^j(g^j+y^j)\right)$ \\-
-    $y^0 = 0$ \\
-    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,\overline{\chi}^0,g^0;y^0)$ \hspace{1.0cm} \ct{solving of \eqref{eq:fe:downvi} in $\mathcal{D}^0=\mathcal{U}^0$} \\
+    $z^0 = 0$ \\
+    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,\overline{\chi}^0,g^0;z^0)$ \hspace{1.0cm} \ct{solving of \eqref{eq:fe:upvi} in $\mathcal{U}^0$} \\
     for $j=1$ to $j=J$ \\+
@@
 \end{pseudo}
-\caption{Nonlinear multilevel constraint decomposition V-cycle as an iterative solver for VI problem \eqref{eq:fe:vi}.  $f^j$ denotes a discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
+\caption{Nonlinear multilevel constraint decomposition V-cycle as an iterative solver for FE VI problem \eqref{eq:fe:vi}.
+$f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
 \label{alg:nmcd}
@@
 
-During one application of \pr{nmcd-vcycle}, storage must be allocated for $\chi^j$, $\phi^j$, $g^j$, $y^j$, and $\ell^j$ on each level.
-However, line 3 is essentially notational in that the finest-level iterate $g^J$ is the same as $w^J$; no copy is needed.
-Similarly, in algorithm lines 13 and 15 no allocation is required because the $y^j$ space can be reused for $z^j$ as the coarser corrections are prolonged and accumulated.
-Let $m^j$ be the number of interior nodes in the $j$th-level mesh.
-The total storage of \pr{nmcd-vcycle}, excluding any in the smoothers/solver, but including for the original obstacle $\gamma^J$, is about $8 m^J$ real numbers\footnote{That is: $8\approx 5(4/3) + 1$.} under the standard estimate that a 2D, uniform-mesh V-cycle uses $4/3$ the memory of the finest level alone \cite[Section 2.4]{Trottenbergetal2001}.
-For comparison, a single-level method needs at least $3 m^J$ storage for the vectors $\ell^J,\gamma^J,w^J$.
+Regarding its efficiency, to begin we consider storage.
+One application of \pr{nmcd-vcycle}, assuming all box constraints are nontrivial, storage must be allocated for the seven vectors $\underline{\chi}^j$, $\overline{\chi}^j$, $\underline{\phi}^j$, $\overline{\phi}^j$, $g^j$, $\ell^j$, and $y^j$ on each level.
+(Note $z^J$ can use the same storage as $y^j$.)  On the finest level one must also store $\underline{\gamma}^J$ and $\overline{\gamma}^J$, but note $g^J=w^J$ are the same.
+On the coarsest level there are only five vectors to store.
+Therefore, using estimates for 2D refinement plus a standard geometric series argument \cite{Trottenbergetal2001}, the total storage of a deep $V$-cycle, excluding the implementations of smoothers and the coarse-level solver, is about
+\begin{equation}
+9 m_J + 7 m_{J-1} + \dots + 7 m_1 + 5 m_0 \approx \left(9 + 7 \sum_{k=1}^\infty \frac{1}{4^k}\right) m_J \approx 11 m_J
+\end{equation}
+real numbers.
+For comparison, a single-level method needs at least $4 m_J$ storage for the vectors $\underline{\gamma}^J,\overline{\gamma}^J,\ell^J,w^J$.
+If some of the obstacles are trivial, for instance if $\overline{\gamma}^J=+\infty$ in a unilateral obstacle problem, storage is correspondingly reduced at all levels.
 
 \pr{nmcd-vcycle} assumes certain semantics for the smoothers and the coarsest-level solver; the latter might also be a smoother algorithm.
-They do in-place modifications of their final arguments, and they do not modify their first three arguments.
+These solvers do in-place modifications of their final arguments, and they do not modify their first three arguments.
 The smoothers are in-exact solvers of VIs \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}, respectively, and they are iterated \id{down} and \id{up} times.

### 2022-10-25 paper/mcd2.tex
reduce to FAS ... it is now very clear

@@
 \item In Sections \ref{sec:multilevel} and \ref{sec:vcycle} we explain why, given the manner in which constraint sets are decomposed in the multilevel context \cite{GraeserKornhuber2009}, ``up-smoothing'' is more efficient than ``down-smoothing''.
-A strong preference for up-smoothing is special to multilevel CD methods and does not arise in unconstrained problems.\footnote{This observation seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}.}
+A strong preference for up-smoothing is special to multilevel CD methods and does not arise in unconstrained problems.
 \end{enumerate}
@@
 Let $\psi_p(x)$ denote the hat function for a node $x_p \in \mathcal{T}^J$.
-MCP \eqref{eq:fe:mcp} says that at each node exactly\footnote{The stated conditions are actually exclusive only where $\underline{\gamma}^J(x_p) < \overline{\gamma}^J(x_p)$.
-Where the box constraints meet at $x_p \notin \partial_D\Omega$, both of the first conditions hold.} one of the following conditions holds for the solution $u^J$:
+MCP \eqref{eq:fe:mcp} says that at each node exactly\footnote{The stated conditions are correct anywhere $\underline{\gamma}^J(x_p) < \overline{\gamma}^J(x_p)$.
+Where the box constraints meet (pinch) at $x_p \notin \partial_D\Omega$ then $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$ holds and $\ip{f^J(u^J)-\ell^J}{\psi_p}$ can have any value.} one of the following conditions holds for the solution $u^J$:
 \begin{itemize}
-\item $x_p \notin \partial_D\Omega$ and $u^J(x_p)=\underline{\gamma}^J(x_p)$ and $f^J(u^J)(\psi_p)-\ell^J(\psi_p) \ge 0$, or
-\item $x_p \notin \partial_D\Omega$ and $u^J(x_p)=\overline{\gamma}^J(x_p)$ and $f^J(u^J)(\psi_p)-\ell^J(\psi_p) \le 0$, or
-\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)<u^J(x_p)<\overline{\gamma}^J(x_p)$ and $f^J(u^J)(\psi_p)-\ell^J(\psi_p) = 0$, or
+\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)<u^J(x_p)<\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} = 0$, or
+\item $x_p \notin \partial_D\Omega$ and $u^J(x_p)=\underline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \ge 0$, or
+\item $x_p \notin \partial_D\Omega$ and $u^J(x_p)=\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \le 0$, or
 \item $x_p \in \partial_D\Omega$ and $u^J(x_p)=g_D^J(x_p)$.
 \end{itemize}
-Regarding the final case, $\lambda_p=f^J(u^J)(\psi_p)-\ell^J(\psi_p)$ has any value if $x_p \in \partial_D\Omega$.
+These cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, and \emph{boundary}, respectively.
+Where case \emph{boundary} applies, $\lambda_p=\ip{f^J(u^J)-\ell^J}{\psi_p}\in\RR$ has any value.
 
@@
 \end{figure}
+
+Intuitively-speaking, below LDC $\underline{\chi}^j$ generated by maximum restriction $\maxR$, as in \eqref{eq:fe:chilevels}, is the most-negative function from $\mathcal{V}^j$ such that if another element of $\mathcal{V}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$ at the finer level.
+(Similar comments apply for above LDCs $\overline{\chi}^{j}$.)  In this way, monotone restriction construction of the coarser-level constraints is designed to permit the largest corrections from the coarsest levels \cite{GraeserKornhuber2009}, which promotes V-cycle efficiency.
+However, any system satisfying property \eqref{eq:fe:chiordering} will suffice; the particular formula \eqref{eq:fe:chilevels} is not required when generating the LDCs $\underline{\chi}^{j},\overline{\chi}^{j}$.
 
@@
 
-\pr{nmcd-vcycle} assumes certain semantics for the smoothers and the coarsest-level solver; the latter might also be a smoother algorithm.
-These solvers do in-place modifications of their final arguments, and they do not modify their first three arguments.
-The smoothers are in-exact solvers of VIs \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}, respectively, and they are iterated \id{down} and \id{up} times.
-Further smoother implementation details and choices are addressed in the next section.
+\pr{nmcd-vcycle} assumes certain semantics for the smoothers \pr{smooth} and the coarsest-level solver \pr{solve}.
+These solvers do in-place modifications of their final arguments, without modifying their first four arguments.
+$\pr{smooth}^{\id{down}}$ and $\pr{smooth}^{\id{up}}$ are in-exact solvers of VIs \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}, respectively, iterated \id{down} and \id{up} times.
+Further implementation details are addressed in Section \ref{sec:smoothers}.
 
-Repeated application of \pr{nmcd-vcycle} to solve VI \eqref{eq:fe:vi} will not reduce the finest-level residual $f^J(u^J) - \ell^J$ to zero everywhere.
-However, recalling that a finite-dimensional VI is equivalent to a \emph{nonlinear complementarity problem} (NCP) \cite{FacchineiPang2003}, \eqref{eq:fe:vi} corresponds to the NCP
+Repeated application of \pr{nmcd-vcycle} to solve VI \eqref{eq:fe:vi} will not reduce the finest-level residual $f^J(w^J) - \ell^J$ to zero everywhere, and we need to identify the vector quantity whose norm must be controlled by a tolerance.
+Recalling the comments following mixed complementarity problem \eqref{eq:fe:mcp}, and denoting the hat function at node $x_p \in \mathcal{T}^J$ by $\psi_p$, for $w^J \in \mathcal{K}^J$ we define the following vector in $\RR^{m_J}$ as the finest-level \emph{nodal complementarity residual} which is defined as follows in the four cases (inactive, lower active, upper active, boundary):\footnote{Formally-speaking, $\hat r:\mathcal{V}^J \to \RR^{m_J}$ is simply a computable nonlinear map derived from $f^J$ and the vectors $\ell^J$, $\underline{\gamma}^J$, and $\overline{\gamma}^J$.
+It is an important part of the implementation.}
 \begin{equation}
-u^J \ge \gamma^J, \qquad f^J(u^J) \ge \ell^J, \qquad (u^J - \gamma^J)\left(f^J(u^J) - \ell^J\right) = 0.  \label{eq:fe:ncp}
+\hat r(w^J)_p = \begin{cases}
+    \ip{f^J(w^J)-\ell^J}{\psi_p}, & \underline{\gamma}^J(x_p) < w^J(x_p) < \overline{\gamma}^J(x_p), \\
+    \min\{\ip{f^J(w^J)-\ell^J}{\psi_p},0\}, & w^J(x_p) = \underline{\gamma}^J(x_p), \\
+    \max\{\ip{f^J(w^J)-\ell^J}{\psi_p},0\}, & w^J(x_p) = \overline{\gamma}^J(x_p), \\
+    0, & x_p \in \partial_D\Omega. \end{cases} \label{eq:cpresidual}
 \end{equation}
-Thus, at convergence we only expect $f^J(u^J) = \ell^J$ at those locations where the constraint is inactive, namely where $u^J > \gamma^J$.
-Therefore, recalling that the hat function at node $x_q^J$ is denoted $\psi_q^J$, for $w^J \in \mathcal{K}^J$ we define the following vector in $\RR^{m^J}$ as the finest-level \emph{nodal residual}:
-\begin{equation}
-\hat r(w^J)_q = \begin{cases} \ip{f^J(w^J)-\ell^J}{\psi_q^J}, & w^J(x_q^J) > \gamma^J(x_q^J), \\
-                                  \min\{\ip{f^J(w^J)-\ell^J}{\psi_q^J},0\}, & w^J(x_q^J) = \gamma^J(x_q^J). \end{cases} \label{eq:cpresidual}
-\end{equation}
-(Formally, $\hat r:\mathcal{V}^J \to \RR^{m^J}$ is a nonlinear map.)  Observe that $\hat r(w^J)$ is nonzero if there are nodes where the NCP \eqref{eq:fe:ncp} is violated, either because $f^J(w^J)-\ell^J$ remains nonzero on the inactive set or because the same quantity is negative on the active set, and that nodal values of $f^J(w^J)-\ell^J$ are in evaluated in a weak sense, by application to a hat function.
+The point is that $\hat r(w^J)\in \RR^{m_J}$ is nonzero if there are nodes where MCP \eqref{eq:fe:mcp} is violated, either because the pointwise classical residual $\ip{f^J(w^J)-\ell^J}{\psi_p}$ remains nonzero where the constraints are inactive, or because it has the wrong sign on an actively-constrained degree of freedom.
 
-A solver for \eqref{eq:fe:vi} using repeated applications of \pr{nmcd-vcycle} would start with an initial iterate $w^{J,0}$, do V-cycles, and stop once the finest-level nodal residual $\hat r(w^J)$ was sufficiently small.
+A solver for \eqref{eq:fe:vi} using repeated applications of \pr{nmcd-vcycle} would start with an initial iterate $(w^J)^{(0)}$, do V-cycles, and stop once the finest-level nodal residual $\hat r(w^J)$ was sufficiently small.
 For example, using tolerances \id{atol}$>0$ and \id{rtol}$>0$ the stopping criterion might be
 \begin{equation}
-\|\hat r(w^J)\| < \id{atol} \qquad \text{or} \qquad \frac{\|\hat r(w^J)\|}{\|\hat r(w^{J,0})\|} < \id{rtol},
+\|\hat r(w^J)\| < \id{atol} \qquad \text{or} \qquad \frac{\|\hat r(w^J)\|}{\|\hat r((w^J)^{(0)})\|} < \id{rtol},
 \end{equation}
-where $\|\cdot\|$ denotes a norm on $\RR^{m^J}$.  A step tolerance could also be used, measuring the difference between successive iterates.
-
-FIXME in Appendix we show how NMCD reduces to FAS if the obstacle is removed and to MCD if $f$ is assumed to be linear
+for some norm $\|\cdot\|$ on $\RR^{m_J}$.  A step tolerance could also be used, measuring the difference between successive iterates.
 
@@
 
-FIXME
-
-Recall that a $j$th-level hat function based at node $x_p^j$ is denoted $\psi_p^j$, for an iterate $g^j$ and an admissible correction $y^j \in \mathcal{D}^j$ we define the \emph{downward nodal residual} as
+FIXME  Recall that a $j$th-level hat function based at node $x_p^j$ is denoted $\psi_p^j$, for an iterate $g^j$ and an admissible correction $y^j \in \mathcal{D}^j$ we define the \emph{downward nodal residual} as
 \begin{equation}
@@
 
-\section{Results} \label{sec:results}
+\section{Results and discussion} \label{sec:results}
 
@@
 
-FIXME observe that up-smoothing is more efficient
+FIXME observe that up-smoothing is more efficient; this observation seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}
 
 FIXME constrast the implemented examples in \cite{GraeserKornhuber2009,Tai2003} in which $f$ is linear
+
+FIXME in Appendix we show how NMCD reduces to FAS if the obstacle is removed and to MCD if $f$ is assumed to be linear
 
@@
 \appendix
-\section{Reduction of NMCD to unconstrained FAS}
+\section{Reductions of NMCD to familiar multilevel algorithms}
 
-\pr{nmcd-vcycle}, Algorithm \ref{alg:nmcd} in the main text, generalizes two multilevel algorithms, first the FAS multigrid V-cycle for nonlinear PDEs, which itself generalizes the standard linear multigrid V-cycle, and second the MCD V-cycle for linear variational inequalities.
-This Appendix helps readers understand \pr{nmcd-vcycle} by identifying the assumptions which reduces it to these existing methods.
+Algorithm \ref{alg:nmcd} in the main text, \pr{nmcd-vcycle}, generalizes two better-known multilevel algorithms.
+First we show that it reduces to the FAS multigrid V-cycle for nonlinear PDEs (which itself generalizes the standard linear multigrid V-cycle \cite{Trottenbergetal2001}).
+Second we reduce NMCD to an existing MCD V-cycle for unilateral linear variational inequalities \cite{GraeserKornhuber2009}.
+This Appendix helps readers understand Algorithm \ref{alg:nmcd} by separating those assumptions which relate to nonlinear and FAS aspects from the nontrivial modifications of standard multigrid which are caused by inequality constraints.
+The later arise even when $f$ in \eqref{eq:vi} is a linear operator.
 
-To derive the FAS multigrid V-cycle we need to remove the inequality constraints from \pr{nmcd-vcycle}.
-To do this, suppose that the finest-level obstacle $\gamma^J$ is very negative, with magnitude greater than $u^J$ or any prospective iterate $w^J$.
-Problem \eqref{eq:fe:vi} is then unconstrained, so by replacing ``$v-u^J$'' with an arbitrary $v \in \mathcal{V}^J$ we have the finest-level, weak form problem
+To derive the classical FAS multigrid V-cycle from Algorithm \ref{alg:nmcd}, we remove the inequality constraints from VI problem \eqref{eq:fe:vi} and then simplify the pseudocode accordingly.
+To do this we suppose that the finest-level obstacles $\underline{\gamma}^J$ and $\overline{\gamma}^J$ are very negative and positive, respectively, with magnitude exceeding $u^J$ or any prospective iterate $w^J$.
+Finest-level problem \eqref{eq:fe:vi} is then unconstrained except for its Dirichlet boundary conditions.
+Let $\mathcal{V}_0^J \subset \mathcal{V}^J$ denote the space with zero values on $\partial_D\Omega$.
+By replacing ``$v-u^J$'' with an arbitrary $v \in \mathcal{V}_0^J$ we have the finest-level, weak form problem
 \begin{equation}
-\ip{f^J(u^J)}{v} = \ip{\ell^J}{v} \qquad \text{for all } v\in \mathcal{V}^J \label{eq:app:fas:pde}
+\ip{f^J(u^J)}{v} = \ip{\ell^J}{v} \qquad \text{for all } v\in \mathcal{V}_0^J \label{eq:app:fas:pde}
 \end{equation}
 corresponding to the strong-form PDE $f(u)=\ell$.
-Now, \pr{nmcd-vcycle} would generate downward defect constraints by monotone restriction.
-However, all that is required for the downward system is that $\chi^{j-1} \ge \chi^j$, so that $\phi^j = \chi^j - P \chi^{j-1} \le 0$.
-In the unconstrained case we can assume, more simply, that the fine level defect constraint is a negative constant $\chi^J = -JZ$ for $Z>0$ such that $Z \gg \|u^J\|_\infty$.
-Defining $\chi^j = - jZ$, and calculating $\phi^j = \chi^j - P \chi^{j-1} = - Z$, we see that both the downward VI \eqref{eq:fe:downvi} and the upward VI \eqref{eq:fe:upvi} can also be made unconstrained.
-It then follows that lines 2, 5, and 6, can be removed from Algorithm \ref{alg:nmcd}, and further that the constraint arguments can be removed from the smoothers and the coarse-level solver.
-Because the downward corrections $y^j$ are no longer separately constrained, we may simply refer to the corrected solution $w^j=g^j+y^j$.
-The smoothers can be regarded as in-place modifications of $w^j$.
-Also recognizing that the initial iterate for the coarse-level smoother is $w^{j-1}=\iR w^j$, the definition of the coarsened source functional can be rewritten
+
+Furthermore, the downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets can be arbitrarily enlarged so that the coarser-level corrections $y^j$ and $z^j$ in Algorithm \ref{alg:nmcd} are also never constrained.
+Formally-speaking we might do this, still satisfying the LDC ordering property \eqref{eq:fe:chiordering}, by choosing LDCs with all gaps being large so that $\underline{\chi}^j \ll 0 \ll \overline{\chi}^j$ and also $\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1} \ll 0 \ll \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}$.\footnote{As noted after \eqref{eq:fe:chiordering}, the application of monotone restriction operators $\maxR,\minR$ is not required to generate LDCs $\underline{\chi}^j$ and $\overline{\chi}^j$ satisfying the ordering property.}
+
+Now the correction VIs \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} are also unconstrained, and lines 2--3 and 6--9 can be removed from Algorithm \ref{alg:nmcd}.
+Also, because the downward corrections $y^j$ are no longer separately constrained, we may simply refer to the corrected solution $w^j=g^j+y^j$ instead of $g^j$ and $y^j$ separately.
+Then the smoothers and the coarsest-level solver can be regarded as in-place modifications of $w^j$.
+Now noting $g^{j-1}=\iR w^j$ in line 12, after down-smoothing, the coarsened source functional can be rewritten using notation $w^{j-1}=\iR w^j$:
 \begin{equation}
@@
 \end{equation}
-However, now we must be careful that going upward only the correction gets prolonged;compare Remark 5.3.9 in \cite{Trottenbergetal2001}.
-We note that the coarse correction produces an updated $w^{j-1}$ (previously denoted $g^{j-1}+y^{j-1}$).
-If $w^j$ denotes the result of down-smoothing, and if $z^j = y^j + P z^{j-1} = y^j + y^{j-1} + \dots + y^0$ is the sum of all prolonged corrections, we want to set $w^j = g^j + z^j = (g^j+y^j) + P(y^{j-1} + \dots + y^0)$ as the iterate going into the up-smoother.
-For this we write $w^j \gets w^j + P(w^{j-1} - \iR w^j)$.
-With these modifications we get pseudocode in Algorithm \ref{alg:fas}.
+(Compare line 13 of Algorithm \ref{alg:nmcd}.)
 
-\begin{pseudofloat}[H]
+At the completion of down-smoothing the solution iterate on the $j$th level is $w^j=g^j+y^j$.
+Now we must be careful that going upward only the coarse correction, and not the coarse solution, gets prolonged.\footnote{Compare Remark 5.3.9 in \cite{Trottenbergetal2001}, a warning to novices.}  The coarser-level iterate is $w^{j-1} = g^{j-1}+z^{j-1}$ after up-smoothing on the $j-1$ level.
+From Algorithm \ref{alg:nmcd} the initial value of the correction $z^j$, before smoothing, is $z^j = y^j + P z^{j-1}$.
+These facts imply that the $j$th-level iterate before up-smoothing must be set to
+    $$w^j \gets w^j + Pz^j = (g^j+y^j)+P(w^{j-1} - g^{j-1}) = w^j + P(w^{j-1} - \iR w^j).$$
+This is a standard FAS statement.
+In fact, with the given modifications, we have the much-simplified pseudocode in Algorithm \ref{alg:fas}.
+This FAS V-cycle algorithm appears in various places, for example Algorithm 14 of \cite{Bruneetal2015} or in section 5.3.4 of \cite{Trottenbergetal2001}.
+
+\begin{pseudofloat}[ht]
 \begin{pseudo} \label{ps:fas-vcycle}
@@
 \end{pseudo}
-\caption{An FAS V-cycle results from removing the inequality constraints from \pr{nmcd-vcycle}, and also grouping $w^j=g^j+y^j$ (downward) and $w^j=g^j+z^j$ (upward).}
+\caption{An FAS V-cycle results from removing the inequality constraints from \pr{nmcd-vcycle}, and simplifying accordingly.}
 \label{alg:fas}
 \end{pseudofloat}
-
-a negative constant, $\gamma^J=-JZ$, where $Z>0$ is larger than the magnitude of $u^J$, or of any prospective iterate $w^J$, say $Z > \|w^J\|_\infty$.
-
-FIXME get Algorithm 14 \cite{Bruneetal2015}
-
-\section{Reduction of NMCD for linear operators}
 

### 2022-11-25 paper/mcd2.tex
still working on the lemma

@@
 
-%\tableofcontents
+\tableofcontents
 
@@
 Note that this algorithm avoids global (Newton) linearization; compare \cite{GraeserKornhuber2009}.
-\item Our implementation permits ``box'' constraints, namely upper and lower obstacles; see Section \ref{sec:multilevel}.
-\item In Sections \ref{sec:multilevel} and \ref{sec:vcycle} we explain why, given the manner in which constraint sets are decomposed in the multilevel context \cite{GraeserKornhuber2009}, ``up-smoothing'' is more efficient than ``down-smoothing''.
+\item Our implementation permits ``box'' constraints, namely upper and lower obstacles; see Section \ref{sec:femultilevel}.
+\item In Sections \ref{sec:femultilevel} and \ref{sec:vcycle} we explain why, given the manner in which constraint sets are decomposed in the multilevel context \cite{GraeserKornhuber2009}, ``up-smoothing'' is more efficient than ``down-smoothing''.
 A strong preference for up-smoothing is special to multilevel CD methods and does not arise in unconstrained problems.
@@
 The problems addressed later in Section \ref{sec:results} involve nonlinear and linear operators $f$ which are partial differential operators.
-As described in section \ref{sec:multilevel}, the constraint sets $\mathcal{K}$ incorporate both inequality box constraints and nonhomogeneous Dirichlet boundary conditions as equality constraints.
+As described in section \ref{sec:femultilevel}, the constraint sets $\mathcal{K}$ incorporate both inequality box constraints and nonhomogeneous Dirichlet boundary conditions as equality constraints.
 However, the following examples use homogeneous (zero) Dirichlet conditions for ease of presentation.
@@
 Finally, for each $\cK_i$ suppose there are bounded, generally-nonlinear \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.
-Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:multilevel}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
+Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:femultilevel}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
 \begin{equation}
@@
 
-In Sections \ref{sec:multilevel} and \ref{sec:vcycle} we will introduce discretizations and practical algorithms for finite-dimensional VI problems.
+In Sections \ref{sec:femultilevel} and \ref{sec:vcycle} we will introduce discretizations and practical algorithms for finite-dimensional VI problems.
 However, one can observe that the CD concept applies at the level of the continuum problem, and the following two obstacle problem examples are illustrations.
@@
 Our second example is a disjoint frequency decomposition.
-A multilevel CD over a finite-dimensional FE discretization, specifically the one proposed in Section \ref{sec:multilevel}, approximates such a frequency decomposition.
+A multilevel CD over a finite-dimensional FE discretization, specifically the one proposed in Section \ref{sec:femultilevel}, approximates such a frequency decomposition.
 
@@
 When $f$ has all the nice properties listed in Example \ref{ex:fnice} then a solver can take advantage of implementation efficiencies which are unavailable in general.
-An FE approximation of VI \eqref{eq:vi}---see Sections \ref{sec:multilevel} and \ref{sec:vcycle}---can then be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.
+An FE approximation of VI \eqref{eq:vi}---see Sections \ref{sec:femultilevel} and \ref{sec:vcycle}---can then be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.
 The actual solution process for VI \eqref{eq:linearlocalvi} is not really the concern here; we are observing that the storage needed for problem \eqref{eq:linearlocalvi}, and the cost of residual evaluation, are substantially smaller than in the general case.
@@
 
-\section{Multilevel finite elements} \label{sec:multilevel}
+\section{Multilevel finite elements} \label{sec:femultilevel}
 
@@
 
-\section{Level defect constraints and the multilevel CD} \label{sec:ldcs}
+\section{Level defect constraints} \label{sec:ldcs}
 
@@
 
+\section{Multilevel CD} \label{sec:cdmultilevel}
+
 A multilevel V-cycle is defined in the next Section.
-For the downward direction in a V-cycle we will use the LDCs $\underline{\phi}^j,\overline{\phi}^j$ as constraints, and for the upward part the LDCs $\underline{\chi}^j,\overline{\chi}^j$.
-In fact, let
+For the downward side of the V-cycle we will use the LDCs $\underline{\phi}^j,\overline{\phi}^j$ as constraints, and for the upward side the LDCs $\underline{\chi}^j,\overline{\chi}^j$; see \eqref{eq:fe:philevels} and \eqref{eq:fe:chilevels}, respectively.
+In the current Section we show how these level defect constraint functions generate constraint sets and constraint decompositions (CDs).
+
+Let
 \begin{align}
-\mathcal{D}^j = \left\{v \,:\, \underline{\phi}^j \le v \le \overline{\phi}^j \text{ and } \, v|_{\partial_D\Omega} = 0\right\} \subset \mathcal{V}^j, \label{eq:fe:constraintsets} \\
-\mathcal{U}^j = \left\{v\,:\, \underline{\chi}^j \le v \le \overline{\chi}^j \text{ and } \, v|_{\partial_D\Omega} = 0\right\} \subset \mathcal{V}^j, \notag
+\mathcal{D}^j = \left\{v \in \mathcal{V}^j \,:\, \underline{\phi}^j \le v \le \overline{\phi}^j \text{ and } \, v|_{\partial_D\Omega} = 0\right\}, \label{eq:fe:constraintsets} \\
+\mathcal{U}^j = \left\{v \in \mathcal{V}^j \,:\, \underline{\chi}^j \le v \le \overline{\chi}^j \text{ and } \, v|_{\partial_D\Omega} = 0\right\}, \notag
 \end{align}
-be the \emph{downward} and \emph{upward defect constraint subsets} of the FE subspaces $\mathcal{V}^j$, respectively.  Again these subsets are closed and convex.
-
-Observe that the upward sets are larger, $\mathcal{D}^j \subset \mathcal{U}^j$, and by \eqref{eq:fe:telescoping} the set decompositions
+be the \emph{downward} and \emph{upward defect constraint subsets}, respectively, of the FE subspaces $\mathcal{V}^j$, for $j=0,1,\dots,J$.
+Again these subsets are closed and convex.
+Observe that the upward sets are larger.
+By \eqref{eq:fe:telescoping}, we have $\mathcal{D}^j \subset \mathcal{U}^j$ and set decompositions
 \begin{equation}
@@
 \end{equation}
-apply in the same sense as \eqref{eq:constraintdecomp}.
-Also note that the finest-level upward constraint set is equivalent to the original constraint set $\mathcal{K}^J$; for all $z^J \in \mathcal{V}^J$,
+in the sense of \eqref{eq:constraintdecomp}.
+Also note that the finest-level upward constraint set $\mathcal{U}^J$ is equivalent to the original constraint set $\mathcal{K}^J$: for all $z^J \in \mathcal{V}^J$,
 \begin{equation}
@@
 
-Recalling definitions \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, we now have two varieties of multilevel CDs of the finest-level upward defect constraint set $\mathcal{U}^J$.
+Recalling definitions \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, we now have two varieties of multilevel CDs of each upward defect constraint set $\mathcal{U}^j$.
 
-\begin{lemma}  Given upper and lower obstacles $\underline{\gamma}^J,\overline{\gamma}^J \in \cV^J$, the constraint set $\cK^J$ defined as in \eqref{eq:fe:fineconstraintset}, a finest level iterate $w^J \in \cK^J$, and downward and upward defect constraint subsets $\mathcal{D}^i,\mathcal{U}^j$ as defined in \eqref{eq:fe:constraintsets}.
-Then for any $j=0,1,\dots,J$ the descending sums
+\begin{lemma}  For any $j=0,1,\dots,J$ and $0\le k\le j$ the decomposition sums
 \begin{align}
-\mathcal{U}^J &\stackrel{\text{(i)}}{=} \mathcal{D}^J + \mathcal{D}^{J-1} + \dots + \mathcal{D}^0, \label{eq:fe:downwardsumcd} \\
-\mathcal{U}^J &\stackrel{\text{(ii)}}{=} \mathcal{D}^J + \mathcal{D}^{J-1} + \dots + \mathcal{D}^{j+1} + \mathcal{U}^j \label{eq:fe:upwardsumcd}
+\mathcal{U}^j &= \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^0, \label{eq:fe:downwardsumcd} \\
+\mathcal{U}^j &= \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^{k+1} + \mathcal{U}^k \label{eq:fe:upwardsumcd}
 \end{align}
-are CDs of\, $\mathcal{U}^J = \cK^J - w^J$, over subspace decompositions $\cV^J=\sum_{i=0}^J \cV^i$ and $\cV^J=\sum_{i=j}^J \cV^i$, respectively.
+are CDs of\, $\mathcal{U}^j$ over subspace decompositions $\cV^j=\sum_{i=0}^j \cV^i$ and $\cV^j=\sum_{i=k}^j \cV^i$, respectively.
 \end{lemma}
 
-\begin{proof}  Definition \eqref{eq:fe:constraintsets} shows $\mathcal{D}^i \subset \cV^i$ and $\mathcal{U}^i \subset \cV^i$.  FIXME what are the decomposition operators?
+\begin{proof}  Subspace decomposition \eqref{eq:subspacedecomp} holds, definition \eqref{eq:fe:constraintsets} shows $\mathcal{D}^i \subset \cV^i$ and $\mathcal{U}^i \subset \cV^i$, and \eqref{eq:fe:constraintdecomp} implies \eqref{eq:constraintdecomp}.
+It remains to find decomposition operators $\Pi_i$ and show \eqref{eq:constraintrestrictionsum}.
 
-FIXME Define the coarsening decomposition (projection) operator
+For $v \in \mathcal{V}^j$ define the coarsening decomposition (projection) operator
     $$\Pi_j^{j-1} v = \min\{\max\{\iR v, \underline{\chi}^{j-1}\}, \overline{\chi}^{j-1}\}.$$
-It is easy to see that $\Pi_j^{j-1}:\mathcal{U}^j \to \mathcal{U}^{j-1}$ and $I-\Pi_j^{j-1}:\mathcal{U}^j \to \mathcal{D}^{j-1}$.
+It is easy to see that $\Pi_j^{j-1}:\mathcal{U}^j \to \mathcal{U}^{j-1}$ because $\iR v \in \mathcal{V}^{j-1}$ and $\underline{\chi}^{j-1} \le \Pi_j^{j-1} v \le \overline{\chi}^{j-1}$.
+It also holds that $I-\Pi_j^{j-1}:\mathcal{U}^j \to \mathcal{D}^{j}$ because
+    $$v - \Pi_j^{j-1} v FIXME$$
 \end{proof}
 
-\section{The nonlinear multilevel CD V-cycle} \label{sec:vcycle}
+\section{Nonlinear multilevel constraint decomposition (NMCD) V-cycle} \label{sec:vcycle}
 
@@
 In practice one generates LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ during the descent stage; see Algorithm \ref{alg:nmcd}.
-Observe that the LDCs depend neither on coarsened iterates $g^j$ nor corrections $y^j,z^j$.} been used to general the LDCs and constraint sets $\mathcal{D}^j,\mathcal{U}^j$ at each level (Section \ref{sec:multilevel}).
+Observe that the LDCs depend neither on coarsened iterates $g^j$ nor corrections $y^j,z^j$.} been used to general the LDCs and constraint sets $\mathcal{D}^j,\mathcal{U}^j$ at each level (Section \ref{sec:femultilevel}).
 As we descend in the V-cycle, we recursively define $j$th-level iterates

### 2022-11-26 paper/mcd2.tex
work in progress on lemma 8.1

@@
 \newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{corollary}[theorem]{Corollary}
 \newtheorem{assumptions}[theorem]{Assumptions}
@@
 
+\section{Multilevel transfer operators} \label{sec:transfers}
+
 The multilevel algorithm defined in the next two Sections, a solver for finite-dimensional problem \eqref{eq:fe:vi}, requires transfer operators between meshes.
@@
 A multilevel V-cycle is defined in the next Section.
-For the downward side of the V-cycle we will use the LDCs $\underline{\phi}^j,\overline{\phi}^j$ as constraints, and for the upward side the LDCs $\underline{\chi}^j,\overline{\chi}^j$; see \eqref{eq:fe:philevels} and \eqref{eq:fe:chilevels}, respectively.
-In the current Section we show how these level defect constraint functions generate constraint sets and constraint decompositions (CDs).
+For the downward side of the V-cycle we will use the LDCs $\underline{\phi}^j,\overline{\phi}^j$ as constraints, and for the upward side the LDCs $\underline{\chi}^j,\overline{\chi}^j$.
+Here we consider the corresponding constraint sets and decompositions (CDs; Section \ref{sec:cd}).
 
@@
 \end{align}
-be the \emph{downward} and \emph{upward defect constraint subsets}, respectively, of the FE subspaces $\mathcal{V}^j$, for $j=0,1,\dots,J$.
-Again these subsets are closed and convex.
-Observe that the upward sets are larger.
-By \eqref{eq:fe:telescoping}, we have $\mathcal{D}^j \subset \mathcal{U}^j$ and set decompositions
-\begin{equation}
-\mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i \label{eq:fe:constraintdecomp}
-\end{equation}
-in the sense of \eqref{eq:constraintdecomp}.
-Also note that the finest-level upward constraint set $\mathcal{U}^J$ is equivalent to the original constraint set $\mathcal{K}^J$: for all $z^J \in \mathcal{V}^J$,
+be the \emph{downward} and \emph{upward defect constraint sets}, respectively, which are closed and convex subsets of the FE subspaces $\mathcal{V}^j$, for $j=0,1,\dots,J$.
+Because $\underline{\chi}^j \le \underline{\phi}^j \le 0 \le \overline{\phi}^j \le \overline{\chi}^j$, it follows that $\mathcal{D}^j \subset \mathcal{U}^j$; the upward sets are larger.
+
+The finest-level upward constraint set $\mathcal{U}^J$ is equivalent to the original constraint set $\mathcal{K}^J$: for all $z^J \in \mathcal{V}^J$,
 \begin{equation}
@@
 
-Recalling definitions \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, we now have two varieties of multilevel CDs of each upward defect constraint set $\mathcal{U}^j$.
+Our multilevel strategy is based on CDs of $\mathcal{U}^J$, or incomplete CDs thereof.
+First, each upward set $\mathcal{U}^j$ can be decomposed in an inclusion sense, which becomes a true CD in unilateral cases.
+The following Lemma is based on the CD construction for Algorithm 4.7 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.
+It is the explanation of downward-side admissibility of the V-cycle.
 
-\begin{lemma}  For any $j=0,1,\dots,J$ and $0\le k\le j$ the decomposition sums
-\begin{align}
-\mathcal{U}^j &= \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^0, \label{eq:fe:downwardsumcd} \\
-\mathcal{U}^j &= \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^{k+1} + \mathcal{U}^k \label{eq:fe:upwardsumcd}
-\end{align}
-are CDs of\, $\mathcal{U}^j$ over subspace decompositions $\cV^j=\sum_{i=0}^j \cV^i$ and $\cV^j=\sum_{i=k}^j \cV^i$, respectively.
+\begin{lemma}  For arbitrary box constraints $\underline{\gamma}^J,\overline{\gamma}^J$ and any $j=0,1,\dots,J$,
+\begin{equation}
+\mathcal{U}^j \supseteq \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^0. \label{eq:fe:downwardsuminclusion}
+\end{equation}
+In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, inclusion \eqref{eq:fe:downwardsuminclusion} is a full CD \eqref{eq:constraintdecomp} of\, $\mathcal{U}^j$ over the subspace decomposition $\cV^j=\sum_{i=0}^j \cV^i$.
 \end{lemma}
 
-\begin{proof}  Subspace decomposition \eqref{eq:subspacedecomp} holds, definition \eqref{eq:fe:constraintsets} shows $\mathcal{D}^i \subset \cV^i$ and $\mathcal{U}^i \subset \cV^i$, and \eqref{eq:fe:constraintdecomp} implies \eqref{eq:constraintdecomp}.
-It remains to find decomposition operators $\Pi_i$ and show \eqref{eq:constraintrestrictionsum}.
+\begin{proof}  Subspace decomposition \eqref{eq:subspacedecomp} already holds by nesting \eqref{eq:fe:nestedspaces}, and definition \eqref{eq:fe:constraintsets} shows $\mathcal{D}^i \subset \cV^i$ and $\mathcal{U}^i \subset \cV^i$.
+If $y^i \in \mathcal{D}^i$ for $i=0,\dots,J$ then
+   $$\underline{\chi}^j = \sum_{i=0}^j \underline{\phi}^i \le \sum_{i=0}^j y^i \le \sum_{i=0}^j \overline{\phi}^i \le \overline{\chi}^j$$
+by \eqref{eq:fe:telescoping} and \eqref{eq:fe:constraintsets}, thus \eqref{eq:fe:downwardsuminclusion} holds for any box constraints.
 
-For $v \in \mathcal{V}^j$ define the coarsening decomposition (projection) operator
-    $$\Pi_j^{j-1} v = \min\{\max\{\iR v, \underline{\chi}^{j-1}\}, \overline{\chi}^{j-1}\}.$$
-It is easy to see that $\Pi_j^{j-1}:\mathcal{U}^j \to \mathcal{U}^{j-1}$ because $\iR v \in \mathcal{V}^{j-1}$ and $\underline{\chi}^{j-1} \le \Pi_j^{j-1} v \le \overline{\chi}^{j-1}$.
-It also holds that $I-\Pi_j^{j-1}:\mathcal{U}^j \to \mathcal{D}^{j}$ because
-    $$v - \Pi_j^{j-1} v FIXME$$
+It remains to consider unilateral cases.
+Here we can construct decomposition operators and show \eqref{eq:constraintrestrictionsum}.
+For example, suppose $\overline{\gamma}^J=+\infty$, thus $\overline{\chi}^j=+\infty$ and $\overline{\phi}^i = +\infty$ for all levels.
+(The case where $\underline{\gamma}^J=-\infty$ is similar.)  For $v\in \mathcal{V}^j$ and $0\le i \le j$, let $I_{j\to i}^\ominus$ denote the operation applying the minimum (monotone) restriction $j-i$ times to map (nonlinearly) into $\mathcal{V}^i$:
+\begin{equation}
+I_{j\to i}^\ominus v = \minR(\dots(\minR v)\dots)
+\end{equation}
+Then we define decomposition operators $\Pi_i:\mathcal{U}^j \to \mathcal{D}^j$ which will make \eqref{eq:fe:downwardsuminclusion} a true CD:
+\begin{equation}
+\Pi_i z^j = I_{j\to i}^\ominus(z^j - \underline{\chi}^j) - I_{j\to i}^\ominus(z^j - \underline{\chi}^j) + \underline{\phi}^j.  \label{eq:fe:unilateraldecompositionoperator}
+\end{equation}
+(Compare equation (4.9) in \cite{GraeserKornhuber2009}.)  Property FIXME implies $\Pi_i z^j \in \mathcal{D}^i$.  On the other hand, observe that \eqref{eq:fe:telescoping} generates a telescoping sum:
+\begin{align*}
+\sum_{i=0}^j \Pi_i z^j &= 
+\end{align*}
+
+FIXME finish up
 \end{proof}
+
+FIXME it would be tempting to try to show a full CD for arbitrary box constraints.
+however it should be observed that $\Pi_i z^j$ may exceed $z^j$ pointwise, so the strategy of splitting $z^j = \min\{z_j,0\} + \max\{z_j,0\} = (z_j)_- + (z_j)_+$ and applying the unilateral formula \eqref{eq:fe:unilateraldecompositionoperator} to each part will not apparently work
+
+Second, logically trivial but important to state, each upward defect constraint set $\mathcal{U}^j$ can also be be decomposed down to an arbitrary level where an upward set $\mathcal{U}^k$ is the base.
+This Corollary is the foundation of the upward side of the V-cycle.
+
+\begin{corollary}  For any $j=0,1,\dots,J$ and $0\le k\le j$,
+\begin{equation}
+\mathcal{U}^j \supseteq \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^{k+1} + \mathcal{U}^k \label{eq:fe:upwardsuminclusion}
+\end{equation}
+In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, inclusion \eqref{eq:fe:upwardsumcd} is a full CD.
+\end{corollary}
 

### 2022-11-27 paper/mcd2.tex
got a handle on the CDs, I think

--- 2022-11-25 paper/mcd2.tex
@@
 
-\section{Multilevel CD} \label{sec:cdmultilevel}
+\section{Multilevel CDs} \label{sec:cdmultilevel}
 
-A multilevel V-cycle is defined in the next Section.
-For the downward side of the V-cycle we will use the LDCs $\underline{\phi}^j,\overline{\phi}^j$ as constraints, and for the upward side the LDCs $\underline{\chi}^j,\overline{\chi}^j$; see \eqref{eq:fe:philevels} and \eqref{eq:fe:chilevels}, respectively.
-In the current Section we show how these level defect constraint functions generate constraint sets and constraint decompositions (CDs).
-
-Let
+A multilevel V-cycle solver is defined in the next Section.
+For the downward side of the V-cycle we will use the LDCs $\underline{\phi}^j,\overline{\phi}^j$ as constraints, and $\underline{\chi}^j,\overline{\chi}^j$ for the upward side.
+The corresponding constraint sets are
 \begin{align}
 \mathcal{D}^j = \left\{v \in \mathcal{V}^j \,:\, \underline{\phi}^j \le v \le \overline{\phi}^j \text{ and } \, v|_{\partial_D\Omega} = 0\right\}, \label{eq:fe:constraintsets} \\
-\mathcal{U}^j = \left\{v \in \mathcal{V}^j \,:\, \underline{\chi}^j \le v \le \overline{\chi}^j \text{ and } \, v|_{\partial_D\Omega} = 0\right\}, \notag
+\mathcal{U}^j = \left\{v \in \mathcal{V}^j \,:\, \underline{\chi}^j \le v \le \overline{\chi}^j \text{ and } \, v|_{\partial_D\Omega} = 0\right\}. \notag
 \end{align}
-be the \emph{downward} and \emph{upward defect constraint subsets}, respectively, of the FE subspaces $\mathcal{V}^j$, for $j=0,1,\dots,J$.
-Again these subsets are closed and convex.
-Observe that the upward sets are larger.
-By \eqref{eq:fe:telescoping}, we have $\mathcal{D}^j \subset \mathcal{U}^j$ and set decompositions
-\begin{equation}
-\mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i \label{eq:fe:constraintdecomp}
-\end{equation}
-in the sense of \eqref{eq:constraintdecomp}.
-Also note that the finest-level upward constraint set $\mathcal{U}^J$ is equivalent to the original constraint set $\mathcal{K}^J$: for all $z^J \in \mathcal{V}^J$,
+These are the \emph{downward} and \emph{upward defect constraint sets}, respectively, closed and convex subsets of the FE spaces $\mathcal{V}^j$, for $j=0,1,\dots,J$.
+Because $\underline{\chi}^j \le \underline{\phi}^j \le 0 \le \overline{\phi}^j \le \overline{\chi}^j$, it follows that $\mathcal{D}^j \subset \mathcal{U}^j$; the upward sets are larger.
+Note that the finest-level set $\mathcal{U}^J$ is equivalent to the original constraint set $\mathcal{K}^J$: for all $z^J \in \mathcal{V}^J$,
 \begin{equation}
@@
 
-Recalling definitions \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, we now have two varieties of multilevel CDs of each upward defect constraint set $\mathcal{U}^j$.
+Our multilevel strategy is based on constraint decompositions (CDs; Section \ref{sec:cd}) of $\mathcal{U}^J$, or rather incomplete CDs as we now explain.
+Each upward set $\mathcal{U}^j$ can be decomposed in an inclusion sense, which becomes a true CD in unilateral cases.
+The following Lemma is based on the unilateral CD construction underlying Algorithm 4.7 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.
+It explains downward admissibility in the V-cycle Algorithm \ref{alg:nmcd} in Section \ref{sec:vcycle}.
 
-\begin{lemma}  For any $j=0,1,\dots,J$ and $0\le k\le j$ the decomposition sums
-\begin{align}
-\mathcal{U}^j &= \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^0, \label{eq:fe:downwardsumcd} \\
-\mathcal{U}^j &= \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^{k+1} + \mathcal{U}^k \label{eq:fe:upwardsumcd}
-\end{align}
-are CDs of\, $\mathcal{U}^j$ over subspace decompositions $\cV^j=\sum_{i=0}^j \cV^i$ and $\cV^j=\sum_{i=k}^j \cV^i$, respectively.
+\begin{lemma}  \label{lem:downwardadmissibility}  For arbitrary upper and lower obstacles $\underline{\gamma}^J,\overline{\gamma}^J$ satisfying \eqref{eq:fe:boxconstraintrequirements}, and any $j=0,1,\dots,J$,
+\begin{equation}
+\mathcal{U}^j \supseteq \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^0. \label{eq:fe:downwardsuminclusion}
+\end{equation}
+In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, inclusion \eqref{eq:fe:downwardsuminclusion} becomes a full CD \eqref{eq:constraintdecomp}, namely $\mathcal{U}^j=\sum_{i=0}^j \mathcal{D}^i$, over the subspace decomposition $\cV^j=\sum_{i=0}^j \cV^i$.
 \end{lemma}
 
-\begin{proof}  Subspace decomposition \eqref{eq:subspacedecomp} holds, definition \eqref{eq:fe:constraintsets} shows $\mathcal{D}^i \subset \cV^i$ and $\mathcal{U}^i \subset \cV^i$, and \eqref{eq:fe:constraintdecomp} implies \eqref{eq:constraintdecomp}.
-It remains to find decomposition operators $\Pi_i$ and show \eqref{eq:constraintrestrictionsum}.
+\begin{proof}  Subspace decomposition \eqref{eq:subspacedecomp} holds---see nesting \eqref{eq:fe:nestedspaces}---and definition \eqref{eq:fe:constraintsets} shows $\mathcal{D}^i \subset \cV^i$ and $\mathcal{U}^i \subset \cV^i$.
+If $y^i \in \mathcal{D}^i$ for $i=0,\dots,J$ then
+\begin{equation}
+\underline{\chi}^j = \sum_{i=0}^j \underline{\phi}^i \le \sum_{i=0}^j y^i \le \sum_{i=0}^j \overline{\phi}^i \le \overline{\chi}^j \label{eq:fe:lemmaordering}
+\end{equation}
+by \eqref{eq:fe:telescoping} and \eqref{eq:fe:constraintsets}, thus \eqref{eq:fe:downwardsuminclusion} holds for any box constraints.
 
-For $v \in \mathcal{V}^j$ define the coarsening decomposition (projection) operator
-    $$\Pi_j^{j-1} v = \min\{\max\{\iR v, \underline{\chi}^{j-1}\}, \overline{\chi}^{j-1}\}.$$
-It is easy to see that $\Pi_j^{j-1}:\mathcal{U}^j \to \mathcal{U}^{j-1}$ because $\iR v \in \mathcal{V}^{j-1}$ and $\underline{\chi}^{j-1} \le \Pi_j^{j-1} v \le \overline{\chi}^{j-1}$.
-It also holds that $I-\Pi_j^{j-1}:\mathcal{U}^j \to \mathcal{D}^{j}$ because
-    $$v - \Pi_j^{j-1} v FIXME$$
+It remains to consider unilateral cases.
+Here we can construct decomposition operators and show \eqref{eq:constraintrestrictionsum}.
+Suppose $\overline{\gamma}^J=+\infty$, thus $\overline{\chi}^j=+\infty$ and $\overline{\phi}^i = +\infty$ for all levels.
+For $v\in \mathcal{V}^j$ and $0\le i \le j$, let $I_{j\to i}^\ominus$ be the operator applying the minimum (monotone) restriction $j-i$ times, mapping into $\mathcal{V}^i$:
+\begin{equation}
+I_{j\to i}^\ominus v = \minR(\dots(\minR v)\dots)  \label{eq:fe:minimummaps}
+\end{equation}
+Note that $I_{j\to j}^\ominus v = v$, and set $I_{j\to -1}^\ominus=0$ by definition.
+From operators \eqref{eq:fe:minimummaps} we define the nonlinear decomposition operators $\Pi_i:\mathcal{U}^j \to \mathcal{D}^i$:
+\begin{equation}
+\Pi_i z^j = I_{j\to i}^\ominus(z^j - \underline{\chi}^j) - I_{j\to i-1}^\ominus(z^j - \underline{\chi}^j) + \underline{\phi}^i.  \label{eq:fe:unilateraldecompositionoperator}
+\end{equation}
+(Compare equation (4.9) in \cite{GraeserKornhuber2009}.)  Property \eqref{eq:fe:monotonerestrictionprops} implies $I_{j\to i}^\ominus(z^j - \underline{\chi}^j) \ge I_{j\to i-1}^\ominus(z^j - \underline{\chi}^j)$, thus $\Pi_i z^j \ge \underline{\phi}^i$, so $\Pi_i z^j \in \mathcal{D}^i$ (in this unilateral case).
+On the other hand, the following sum telescopes, leaving only its first term plus the sum in \eqref{eq:fe:telescoping}:
+\begin{align*}
+\sum_{i=0}^j \Pi_i z^j &= z^j - \underline{\chi}^j + \sum_{i=0}^j \underline{\phi}^i = z^j.
+\end{align*}
+This shows \eqref{eq:constraintrestrictionsum}, thus that \eqref{eq:fe:downwardsuminclusion} is equality, a true CD \eqref{eq:constraintdecomp}.
+
+The case where $\underline{\gamma}^J=-\infty$ is similar.
 \end{proof}
+
+It is tempting to try to turn \eqref{eq:fe:downwardsuminclusion} into a full CD for arbitrary box constraints by using decomposition maps like $\Pi_i$.
+Specifically, one might split $z^j = \min\{z^j,0\} + \max\{z^j,0\}$ and apply unilateral formulas like \eqref{eq:fe:unilateraldecompositionoperator} to $\min\{z^j,0\}$ and $\max\{z^j,0\}$ separately, with $\Pi_i$ applied to $\min\{z^j,0\}$ and a straightforward variant operator to $\max\{z^j,0\}$.
+In fact, the lower bound $\Pi_i (\min\{z^j,0\}) \ge \underline{\phi}^i$ holds, as shown in the above proof.
+However, the following Example shows why $\Pi_i(\min\{z^j,0\})$ may have an arbitrarily large maximum, and thus exceed any finite upper LDC $\overline{\phi}^i$, so unfortunately $\Pi_i$ does not map even non-positive functions to $\mathcal{D}^i$ in the presence of a finite upper LDC.
+While the proposed splitting strategy does not work, without redefinition of a formula which is adequate in the unilateral case, the existence of an entirely-different strategy for showing a full CD is also not excluded.
+
+\begin{example}  \label{ex:notfullcd}
+Let $a\ge 0$ be an arbitrary positive number.
+Consider a 2-level hierarchy ($J=1$), and suppose the lower defect constraint is the constant function $\underline{\chi}^1=-a$.
+Recalling Table \ref{tab:ldcs}, note $\underline{\chi}^0=\maxR \underline{\chi}^1=-a$ also, so $\underline{\phi}^1=0$ and $\underline{\phi}^0=-a$.
+Suppose $z^1$ is a ``sawtooth'' function which takes on value $-a$ on the coarse nodes and value $0$ on the other fine-level nodes.
+Then $\overline{\chi}^1 \ge 0 \ge z^1\ge \underline{\chi}^1$, for any valid $\overline{\chi}^1$, so $z^j \in \mathcal{U}^1$.
+However, $I_{1\to 0}^\ominus(z^1 - \underline{\chi}^1) = \minR(z^1 - \underline{\chi}^1) = 0$ identically.
+From \eqref{eq:fe:unilateraldecompositionoperator}, $\Pi_1 z^1$ simplifies to $\Pi_1 z^1 = z^1 - \underline{\chi}^1$.
+But then $\Pi_1 z^1$ attains a maximum $a\ge 0$.
+This maximum can be chosen to exceed any upper defect obstacle $\overline{\phi}^1$.
+Thus, for general $w^1 \in \mathcal{U}^1$ and general LDCs, $\Pi_1$ does not map $\min\{w^1,0\}$ into $\mathcal{D}^1$.
+\end{example}
+
+Each upward defect constraint set $\mathcal{U}^j$ can also be be decomposed down to an arbitrary level $k$ where an upward set $\mathcal{U}^k$ is used as the coarsest set.
+The next Lemma, similar to Lemma \ref{lem:downwardadmissibility} but not following from it, explains upward admissibility in V-cycle Algorithm \ref{alg:nmcd}.
+
+\begin{lemma}  \label{lem:upwardadmissibility}  For any $j=0,1,\dots,J$ and $0\le k\le j$,
+\begin{equation}
+\mathcal{U}^j \supseteq \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^{k+1} + \mathcal{U}^k \label{eq:fe:upwardsuminclusion}
+\end{equation}
+In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, inclusion \eqref{eq:fe:upwardsuminclusion} is a full CD \eqref{eq:constraintdecomp}. \end{lemma}
+
+\begin{proof}  Definition \eqref{eq:fe:philevels} shows that $\underline{\chi}^j = \underline{\phi}^j + \dots + \underline{\phi}^{k+1} + \underline{\chi}^k$ and $\overline{\chi}^j = \overline{\phi}^j + \dots + \overline{\phi}^{k+1} + \overline{\chi}^k$.
+(These facts replace \eqref{eq:fe:telescoping}.)  If $y^i \in \mathcal{D}^i$ for $i=k+1,\dots,j$ and $z^k \in \mathcal{U}^k$ then it is easy to show, as in \eqref{eq:fe:lemmaordering}, that $\underline{\chi}^j \le y^j + \dots + y^{k+1} + z^k \le \overline{\chi}^j$, thus \eqref{eq:fe:upwardsuminclusion} holds.
+
+In unilateral cases we follow the proof of Lemma \ref{lem:downwardadmissibility}.
+Definition \eqref{eq:fe:minimummaps} is unchanged.
+We use \eqref{eq:fe:unilateraldecompositionoperator} unchanged for $i=k+1,\dots,j$, but modify it in the $i=k$ case: $\Pi_k z^j = I_{j\to k}^\ominus(z^j - \underline{\chi}^j) + \underline{\chi}^k$.
+The remainder of the proof goes through as for Lemma \ref{lem:downwardadmissibility}.
+\end{proof}
+
+In summary, inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in V-cycle Algorithm \ref{alg:nmcd}.
+However, a convergence analysis following \cite{Tai2003} or \cite{GraeserKornhuber2009}, somehow extending their arguments to VIs \eqref{eq:vi} which are not of minimization type, or to box constraints, would seem to require full CDs, and perhaps new ideas.
 

### 2022-12-07 paper/mcd2.tex
edits though end section 5

@@
 \begin{document}
-\title[Multilevel constraint decomposition methods]{Multilevel constraint decomposition methods \\ for nonlinear variational inequalities}
+\title[FAS for nonlinear variational inequalities]{A full approximation storage multilevel method \\ for nonlinear variational inequalities}
 
@@
 
-We extend the constraint decomposition (CD) method of X.-C.~Tai \cite{Tai2003} by implementing fully-nonlinear corrections, and we apply it to nonlinear variational inequality (VI) problems to which it has not been applied.
-In its multilevel form the method has been shown to have optimal complexity for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
+We extend the constraint decomposition (CD) method of Tai \cite{Tai2003} by implementing fully-nonlinear corrections, and we apply it to a new set of nonlinear variational inequality (VI) problems.
+The method has been shown to have almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
 The convergence of this method, for several types of decompositions, was proven for coercive VI problems which arise from minimization of a convex functional over a convex set \cite{Tai2003}.
@@
 For the classical obstacle problem with a Laplacian operator, certain multilevel techniques, such as the truncated monotone multigrid method \cite{Kornhuber1994}, are known to improve performance relative to the original multilevel CD method from \cite{Tai2003}; see also \cite{GraeserKornhuber2009}.
-These improved methods either track the active set in the discretization or modify the nodal basis functions, and in this sense they are discrete algorithms, while by contrast the CD method applies essentially at the level of the continuum problem; see Section \ref{sec:cd}.
+These improved methods either track the active set in the discretization or modify the nodal basis functions, and in this sense they are discrete algorithms.
+By contrast the CD method applies essentially at the level of the continuum problem; see Section \ref{sec:cd}.
 Acceleration of our NMCD method via active-set and/or basis-level manipulations represents a potential extension of the method here, and is a topic for future research.
@@
 
-Suppose there are $m<\infty$ closed subspaces $\cV_i \subset \cV$ so that the sum
+If there are $m<\infty$ closed subspaces $\cV_i \subset \cV$ so that the sum
 \begin{equation}
@@
 \end{equation}
-holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$.  Equation \eqref{eq:subspacedecomp} is called a \emph{subspace decomposition} of $\cV$ \cite{Xu1992}.
+holds in the sense that for $w \in \cV$ there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$ then we say \eqref{eq:subspacedecomp} is a \emph{subspace decomposition} of $\cV$ \cite{Xu1992}.
 
@@
 Finally, for each $\cK_i$ suppose there are bounded, generally-nonlinear \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.
-Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:femultilevel}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
+Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:transfers}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
 \begin{equation}
@@
 
-In Sections \ref{sec:femultilevel} and \ref{sec:vcycle} we will introduce discretizations and practical algorithms for finite-dimensional VI problems.
-However, one can observe that the CD concept applies at the level of the continuum problem, and the following two obstacle problem examples are illustrations.
-The first is an overlapping domain decomposition.
+In Sections \ref{sec:femultilevel}--\ref{sec:smoothers} we will introduce multilevel CD discretizations and algorithms for finite-dimensional VI problems.
+However, one can observe here that the CD concept applies at the level of the continuum problem.
+The following two obstacle problem examples are illustrations; the first is an overlapping domain decomposition.
 
@@
 Suppose $\Omega \subset \RR^2$ is a (open) bounded polygon and that $\mathcal{V}=W^{1,p}(\Omega)$ is a Sobolev space of once-differentiable functions.
-Though the free boundaries are defined by inequality constraints (below), fixed boundary conditions may be Dirichlet or mixed.
+In addition to the free boundaries defined by inequality constraints (below), fixed boundary conditions may be Dirichlet or mixed.
 Suppose that the domain boundary is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive measure for simplicity, to permit well-posedness \cite{Evans2010}, and $\partial_N \Omega$ possibly empty.
-Suppose that the provided Dirichlet data $g_D:\partial_D \Omega \to \RR$ is sufficiently-regular to permit well-posedness of the VI problem stated below.
+Suppose that the provided Dirichlet data $g_D:\partial_D \Omega \to \RR$ is sufficiently-regular so that the VI problem \eqref{eq:boxdirichletvi} below is well posed.
 
 We restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.
-The \emph{lower obstacle} $\underline{\gamma}$ and the \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions on $\overline{\Omega}$ which take their values in the extended real line $\tilde\RR = [-\infty,+\infty]$.
+The \emph{lower obstacle} $\underline{\gamma}$ and the \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions on the closure $\overline{\Omega}$ which take their values in the extended real line $\tilde\RR = [-\infty,+\infty]$.
 Specifically suppose $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$ and $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$ are measurable, satisfying $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$ and $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.
@@
 \end{equation}
-Note that $v-u^* \in W_0^{1,p}(\Omega)$, that is, the test functions used in \eqref{eq:boxdirichletvi} have zero values on the Dirichlet boundary $\partial_D\Omega$.
-
-Problem \eqref{eq:boxdirichletvi} is both a free-boundary problem, because of the inequality constraints in defining $\cK$, and a (mixed) fixed-boundary value problem.
+Note that $v-u^* \in W_0^{1,p}(\Omega)$, that is, test functions in \eqref{eq:boxdirichletvi} have zero values on the Dirichlet boundary $\partial_D\Omega$.
+
+Problem \eqref{eq:boxdirichletvi} is both a free-boundary problem, because of the inequality constraints in defining $\cK$, and a fixed-boundary value problem.
 Dirichlet conditions $u=g_D$ apply over $\partial_D\Omega$ and Neumann conditions over $\partial_N \Omega$.
-The latter are determined by the formula for $f$; for nonhomogeneous Neumann conditions, the nonlinear operator $f:\cK\to\cV'$ will include an integration over $\partial_N\Omega$.
+The latter are determined by the formula for $f$; for nonhomogeneous Neumann conditions, the nonlinear operator $f:\cK\to\cV'$ must include an integration over $\partial_N\Omega$.
 
 We now construct a multilevel finite element (FE) approximation to problem \eqref{eq:boxdirichletvi} using either triangular or quadrilateral meshes over $\Omega$, and $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, respectively.
-The solution on the finest-level mesh, in a hierarchy of meshes constructed by refinement of a coarsest mesh, is the target.
-This hierarchy also generates a multilevel CD of $\cK$.
+A hierarchy of meshes will be constructed by refinement of a coarsest mesh.
+A solution on the finest-level mesh is the goal.
+This hierarchy also allows the creation of a multilevel CD of $\cK$.
 The subsets $\cK_i$ in this CD are designed to allow multigrid-type coarser-level corrections (next Section) in a manner which preserves admissibility, while not introducing unnecessary high frequencies into the coarser-level problems, especially at the free boundary.
@@
 For $j=1,\dots,J$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ by connecting the midpoints of edges.
-For triangles and rectangles, each $\mathcal{T}^{j-1}$ element becomes four similar elements in $\mathcal{T}^j$ with size $h_j = 2^{-j} h_0$.\footnote{For arbitrary triangles, and for rectangles, the refinement generates similar elements and $h_j=2^{-j}h_0$ is the exact maximum edge length in $\mathcal{T}^j$.
+For triangles and rectangles, each $\mathcal{T}^{j-1}$ element becomes four similar elements in $\mathcal{T}^j$ with size $h_j = 2^{-j} h_0$.\footnote{For arbitrary triangles, and for rectangles, edge midpoint refinement generates similar elements and $h_j=2^{-j}h_0$ is the exact maximum edge length in $\mathcal{T}^j$.
 For general quadrilaterals, however, similarity is lost and maximum edge lengths are variable; cf.~\cite{Zhang2004}.
-Only rectangular quadrilaterals are used in the Results section.}  We call $J\ge 0$ the \emph{depth} of the hierarchy and $\mathcal{T}^J$ the \emph{finest mesh}.
+Only rectangular quadrilaterals are used in the Results section.}  We call $J\ge 0$ the \emph{depth} of the hierarchy and $\mathcal{T}^J$ the \emph{finest-level mesh}.
 Let $m_j$ be the number of nodes in $\mathcal{T}^j$, the \emph{degrees of freedom} at the $j$th level, and denote $h=h_J=2^{-J} h_0$ and $m=m_J$ for the finest level.
+
 Let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space \cite{Elmanetal2014} over $\mathcal{T}^j$, with arbitrary boundary values, and observe the following vector space nesting:
@@
 
-Use of higher-order Lagrange elements $P_k$ or $Q_k$, $k\ge 2$, would be a ``variational crime'' \cite[Chapter 10]{BrennerScott2007}, specific to inequality constraints, in the upcoming material.
-For this reason we restrict to $k = 1$.
-Specifically, our FE construction of constraint sets will require \emph{nodal monotonicity}, namely that for all FE functions $w,z \in \mathcal{V}^j$,
+Use of higher-order Lagrange elements $P_k$ or $Q_k$, $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} special to inequality constraints.
+Namely, our FE construction of constraint sets will require \emph{nodal monotonicity},
 \begin{equation}
@@
 \end{equation}
-where $\bw,\bz$ denote vectors of nodal values.
+for all FE functions $w,z \in \mathcal{V}^j$, where $\bw,\bz$ denote vectors of nodal values.
 For $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal value one at a single node, otherwise zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.
 It is easy to show that \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of sign-changing hat functions, so \eqref{eq:nodalmonotonicity} implies $k\le 1$.
-
-To accomodate the fixed-boundary conditions, let $g_D^J \in \mathcal{V}^J$, on the finest-level mesh, satisfy $g_D^J = g_D$ on $\partial_D \Omega$.\footnote{As is standard in conforming FE methods, this restricts the original Dirichlet boundary values $g_D$ to be a piecewise-linear function \cite{Elmanetal2014}.}  To allow box constraints, let $\tilde{\mathcal{V}}^j$ denote the set of all functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
+For this reason we restrict to $k = 1$.
+
+Let $g_D^J \in \mathcal{V}^J$, on the finest-level mesh, satisfy $g_D^J = g_D$ on $\partial_D \Omega$.
+Typically of conforming FE methods, this restricts the Dirichlet boundary function $g_D$ to be a piecewise-linear \cite{Elmanetal2014}.
+
+To allow box constraints, let $\tilde{\mathcal{V}}^j$ denote the set of all functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
 (Note that $\tilde{\mathcal{V}}^j$ functions are not in $\mathcal{V}^j$.
-They are not even defined everywhere on $\Omega$, but only on nodes.)  Suppose that the \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$, defined on the nodes of the finest-level mesh $\mathcal{T}^J$, satisfy
+They are not defined everywhere on $\Omega$, but only on nodes.)  Suppose that the \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$, for nodes of the finest-level mesh $\mathcal{T}^J$, satisfy
 \begin{equation}
@@
 \end{equation}
-as the finest-level \emph{constraint set} of \emph{admissible} FE functions.
+as the finest-level \emph{constraint set} of admissible FE functions.
 This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.
@@
 Let $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ be a discretization of $f$ in \eqref{eq:boxdirichletvi}.
-By inclusion \eqref{eq:fe:nestedspaces} the functional $\ell$ is already defined and continuous over $\mathcal{V}^J$, but we denote $\ell^J \in (\mathcal{V}^J)'$ for its FE representation.
+By inclusion \eqref{eq:fe:nestedspaces} the functional $\ell$ is already defined and continuous over $\mathcal{V}^J$, but we denote $\ell^J \in (\mathcal{V}^J)'$.
 We seek the FE solution $u^J \in \mathcal{K}^J$ of the finest-level, finite-dimensional, nonlinear VI
@@
 While \eqref{eq:fe:vi} approximates \eqref{eq:boxdirichletvi}, it may not technically be conforming for that VI.
-That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we still do not require $\mathcal{K}^J \subset \mathcal{K}$.
-The continuum constraint set $\mathcal{K}$ remains abstract in those applications where only the nodal obstacles $\underline{\gamma}^J, \overline{\gamma}^J$ are defined.
-In fact, even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants will generally not imply continuum admissiblity.\footnote{Such caveats apply in geophysical situations where the obstacle is given by meshed elevation data, the continuum limit of which is generally unknown \cite{Bueler2016}.}
+That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require $\mathcal{K}^J \subset \mathcal{K}$.
+Even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants will generally not imply continuum admissiblity.
+Such caveats apply in geophysical situations where only nodal values of the obstacle are known, and given by meshed elevation data.
+The continuum limit of nodal obstacles $\underline{\gamma}^J$, $\overline{\gamma}^J$ may be unknown \cite{Bueler2016}.
 
@@
 If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let us define $I_D$ as the $m_D\times m_J$ matrix which indicates (extracts) Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet values as nodes in $(\mathcal{V}^J)'$.
-Problem \eqref{eq:fe:vi} could be stated as VI$\left(f^J(\cdot)-\ell^J,\left[\underline{\gamma}^J,\overline{\gamma}^J\right] \cap X\right)$, for instance in the operations research literature \cite{FerrisPang1997}, and it is equivalent to the equality-constrained MCP
+Problem \eqref{eq:fe:vi} must be stated as VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap X\right)$, in the operations research literature \cite{FerrisPang1997}, and it is equivalent to the equality-constrained MCP
 \begin{align}
@@
 MCP \eqref{eq:fe:mcp} says that at each node exactly\footnote{The stated conditions are correct anywhere $\underline{\gamma}^J(x_p) < \overline{\gamma}^J(x_p)$.
-Where the box constraints meet (pinch) at $x_p \notin \partial_D\Omega$ then $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$ holds and $\ip{f^J(u^J)-\ell^J}{\psi_p}$ can have any value.} one of the following conditions holds for the solution $u^J$:
+Where the box constraints meet (pinch) at $x_p \notin \partial_D\Omega$ then $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$ holds.} one of the following conditions holds for the solution $u^J$:
 \begin{itemize}

### 2022-12-07 paper/mcd2.tex
a few more edits

@@
 \medskip
-\caption{Transfer operators used in this paper.  The FE vector space $\mathcal{V}^j$ has zero boundary values while $\tilde{\mathcal{V}}^j$ has $\tilde \RR$ values, and arbitrary boundary values.}
+\caption{Transfer operators used in this paper.}
 \label{tab:transfers}
@@
 To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, observe that for any given values at the nodes of $\mathcal{T}^j$ there is a unique function in $\mathcal{V}^j$ with those values.
-Also observe that, by our refinement construction, the nodes of $\mathcal{T}^{j-1}$ are nodes of $\mathcal{T}^j$.
+Also observe that, by construction, the nodes of $\mathcal{T}^{j-1}$ are nodes of $\mathcal{T}^j$.
 For $z\in\mathcal{V}^j$ we define the injection $\iR z$ as the element of $\mathcal{V}^{j-1}$ which has the same nodal values as $z$ does at the nodes of $\mathcal{T}^{j-1}$.
-Note that injection is monotonic, that is, if $w \le z$ then $\iR w \le \iR z$.
-
-Finally, the monotone restrictions $\maxR,\minR$ are nonlinear maps which maximize and minimize, respectively, nodal values over the supports of coarser-mesh hat functions.
-These restrictions will be applied only to nodal functions ($\tilde{\mathcal{V}}^j$), and only to functions which are either nonnegative or nonpositive, and not more generally.
-To make the definitions precise, let $x_q^j$, $q=0,1,\dots,m_j-1$, denote the nodes of mesh $\mathcal{T}^j$, and denote the corresponding hat functions \cite{Elmanetal2014} by $\psi_q^j(x)$.
-The monotone restrictions $\maxR z,\minR z$ of an extended-real-valued nodal function $z\in\tilde{\mathcal{V}}^j$ are coarser-level functions in $\tilde{\mathcal{V}}^{j-1}$ defined by
+Note that injection is monotonic: if $w \le z$ then $\iR w \le \iR z$.
+
+The monotone restrictions $\maxR,\minR$ are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh hat functions.
+These restrictions will be applied only to nodal functions ($\tilde{\mathcal{V}}^j$).
+To make the definitions precise, let $x_q^j$, $q=0,1,\dots,m_j-1$, denote the nodes of mesh $\mathcal{T}^j$, and denote the corresponding hat functions by $\psi_q^j(x)$.
+The monotone restrictions $\maxR z,\minR z$ of an $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ are coarser-level functions in $\tilde{\mathcal{V}}^{j-1}$ defined by
 \begin{align}
@@
 for $p=0,1,\dots,m_{j-1}-1$.
-(In computing $\max$ and $\min$, the usual ordering of the extended real line is applied.)  Because, by mesh refinement, a node $x_p^{j-1}$ is in the set of $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, the following formal property holds:
+(In computing $\max$ and $\min$, the usual $\tilde\RR$ ordering is applied.)  Because, by mesh refinement, a node $x_p^{j-1}$ is in the set of $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, the following formal property holds:
 \begin{equation}
@@
 \end{equation}
-In particular, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These properties are used in the CD construction which follows.
+In particular, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$, properties used in the CD construction which follows.
 
@@
 Suppose $w^J \in \cK^J$ is an admissible finest-level iterate which approximates the solution $u^J$ of VI \eqref{eq:fe:vi}.
-We will now define the CD constraint subsets in our multilevel FE context from $w^J$ and the defining inequalities for $\cK^J$ (equation \eqref{eq:fe:fineconstraintset}).
-Though the ideas are fundamentally the same as for unilateral constraints \cite{GraeserKornhuber2009}, the notation is more elaborate because we consider bi-obstacle (box-constrained) problems.
+We will now define the CD constraint subsets, in our multilevel FE context, using $w^J$ and the defining inequalities for $\cK^J$, namely \eqref{eq:fe:fineconstraintset}.
+Though the ideas are fundamentally the same as for unilateral constraints \cite{GraeserKornhuber2009}, the notation is necessarily more elaborate because we consider bi-obstacle (box-constrained) problems.
 
@@
 as nodal functions in $\tilde{\mathcal{V}}^J$.
-Note that $\underline{\chi}^J \le 0 \le \overline{\chi}^J$ everywhere in $\overline{\Omega}$, including along the Dirichlet boundary $\partial_D\Omega$ where $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ and $w^J = g_D^J = g_D$.
+Note that $\underline{\chi}^J \le 0 \le \overline{\chi}^J$ everywhere in $\overline{\Omega}$, including along the Dirichlet boundary $\partial_D\Omega$.
+(Recall that on $\partial_D\Omega$ we have $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ and $w^J = g_D^J = g_D$.)  If $\underline{\gamma}^J=-\infty$ then $\underline{\chi}^J=-\infty$ by definition, and similarly if $\overline{\gamma}^J=+\infty$ then $\overline{\chi}^J=+\infty$.
 The motivation for creating such defect constraints \cite{GraeserKornhuber2009} is that a corrected iterate $w^J + z^J$, for $z^J\in\cV^J$, is admissible ($w^J + z^J \in \cK^J$) if and only if the perturbation $z^J$ is between the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$, and $z^J=0$ on $\partial_D\Omega$.
 
-For coarser levels $j=J-1,\dots,1,0$ we apply the monotone restrictions to generate ``\emph{below}'' and ``\emph{above}'' \emph{level defect constraints} (LDCs):
+Counting down the coarser levels $j=J-1,\dots,1,0$, we apply the monotone restrictions to generate ``\emph{below}'' and ``\emph{above}'' \emph{level defect constraints} (LDCs):
 \begin{equation}
@@
 \end{equation}
-in the sense of comparing nodal values at nodes shared between levels.  A $J=4$ example of ``below'' LDCs $\underline{\chi}^j$ is shown in Figure \ref{fig:chiphilevels}.
+(in the sense of nodal values).  A $J=4$ example of ``below'' LDCs $\underline{\chi}^j$ is shown in Figure \ref{fig:chiphilevels}.
 
@@
 \includegraphics[width=0.65\textwidth]{fixfigs/chiphilevels.pdf}
-\caption{The ``below'' level defect constraint (LDC) $\underline{\chi}^J = \underline{\gamma}^J - w^J$ is decomposed using maximum restriction: $\underline{\chi}^{j-1} = \maxR \underline{\chi}^j$.}
+\caption{The ``below'' level defect constraints (LDC) $\underline{\chi}^j$ are generated from $\underline{\chi}^J = \underline{\gamma}^J - w^J$ using maximum restriction: $\underline{\chi}^j = \maxR \underline{\chi}^{j+1}$.}
 \label{fig:chiphilevels}
@@
 
-Intuitively-speaking, below LDC $\underline{\chi}^j$ generated by maximum restriction $\maxR$, as in \eqref{eq:fe:chilevels}, is the most-negative function from $\mathcal{V}^j$ such that if another element of $\mathcal{V}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$ at the finer level.
-(Similar comments apply for above LDCs $\overline{\chi}^{j}$.)  In this way, monotone restriction construction of the coarser-level constraints is designed to permit the largest corrections from the coarsest levels \cite{GraeserKornhuber2009}, which promotes V-cycle efficiency.
+Intuitively-speaking, below LDC is the most-negative function from $\mathcal{V}^j$ such that if another element of $\mathcal{V}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$ at the finer level.
+(Similar comments apply for above LDCs $\overline{\chi}^{j}$.)  In this way, monotone restriction construction of the coarser-level box constraints is designed to permit the largest corrections from the coarsest levels \cite{GraeserKornhuber2009}, which promotes V-cycle efficiency.
 However, any system satisfying property \eqref{eq:fe:chiordering} will suffice; the particular formula \eqref{eq:fe:chilevels} is not required when generating the LDCs $\underline{\chi}^{j},\overline{\chi}^{j}$.

### 2022-12-20 paper/mcd2.tex
edits sections 1-3 and in section 7

@@
 
-We extend the constraint decomposition (CD) method of Tai \cite{Tai2003} by implementing fully-nonlinear corrections, and we apply it to a new set of nonlinear variational inequality (VI) problems.
-The method has been shown to have almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
-The convergence of this method, for several types of decompositions, was proven for coercive VI problems which arise from minimization of a convex functional over a convex set \cite{Tai2003}.
-
-The multilevel CD algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} are extended here, for finite element (FE) solvers, in three practical ways:
+For coercive variational inequality (VI) problems which arise from minimization of a convex functional over a convex set, the constraint decomposition (CD) method of Tai \cite{Tai2003} converges for multilevel and domain decompositions.
+Furthermore the method has been shown to have almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
+
+We extend the multilevel CD method by implementing fully-nonlinear corrections, and we apply it to a new set of nonlinear VI problems which are coercive but which do not necessarily arise from minimization.
+Specifically, for finite element (FE) solvers the algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} are extended here in three practical ways:
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
 \begin{enumerate}
-\item We apply a full approximation storage (FAS; see \cite{Brandt1977,Bruneetal2015}) approach.
+\item We apply a full approximation storage (FAS; see \cite{Brandt1977,Bruneetal2015}) approach which avoids global (Newton) linearization.
 Section \ref{sec:results} shows results from this \emph{nonlinear multilevel constraint decomposition} (NMCD) algorithm  for several nonlinear problems.
-Note that this algorithm avoids global (Newton) linearization; compare \cite{GraeserKornhuber2009}.
 \item Our implementation permits ``box'' constraints, namely upper and lower obstacles; see Section \ref{sec:femultilevel}.
@@
 
-The iterates from a CD method are always admissible, and thus the nonlinear operator in the VI need only be defined for admissible states.
+The iterates from a CD method are always admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
 Our NMCD scheme is designed for such admissible-only problems, some of which require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation, as discussed in \cite{Bueler2021conservation}.
 NMCD methods should permit direct solutions of those VI problems for which non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
-
-One of our examples (Section \ref{sec:results}) demonstrates the application of the NMCD method to a problem of porous-media type, a nonlinear operator which is not known to be coercive.
-By ``freezing' the solution-dependent coefficient, the operator is approximated by a coercive operator for the duration of the V-cycle.
-This scheme is also effective for doubly-nonlinear diffusion operators.
+% FIXME and convergence is monotone
+
+One of our examples (Section \ref{sec:results}) demonstrates the application of the NMCD method to a nonlinear operator of porous-media type.
+By ``freezing'' the solution-dependent coefficient, the operator is approximated by a coercive operator for the duration of the V-cycle.
+This scheme is also effective for doubly-nonlinear diffusion operators \cite[for example]{Bueler2021conservation,Calvoetal2002}.
 
@@
 These improved methods either track the active set in the discretization or modify the nodal basis functions, and in this sense they are discrete algorithms.
-By contrast the CD method applies essentially at the level of the continuum problem; see Section \ref{sec:cd}.
-Acceleration of our NMCD method via active-set and/or basis-level manipulations represents a potential extension of the method here, and is a topic for future research.
+By contrast a CD method applies essentially at the level of the continuum problem; see Section \ref{sec:cd}.
+Acceleration of the NMCD method via active-set and/or basis-level manipulations is a topic for future research.
 
@@
 
-Suppose $\cV$ is a real and reflexive Banach space with norm $\|\cdot\|$, and denote its topological dual space by $\cV'$.
-Denote the dual pairing of $\phi \in \cV'$ and $v\in\cV$ by $\ip{\phi}{v} = \phi(v)$, and note that $\|\phi\|_{\cV'} = \sup_{\|v\|=1} |\ip{\phi}{v}|$ defines a (Banach space) norm on $\cV'$.
-
-Let $\cK \subset \cV$ be a nonempty closed and convex subset, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.
-For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and a linear \emph{source functional} $\ell\in \cV'$ we consider the following \emph{variational inequality} (VI) problem for the (exact) solution $u^*\in \cK$, if it exists:
+Suppose $\cV$ is a real and reflexive Banach space with norm $\|\cdot\|$ and topological dual space $\cV'$.
+Denote the dual pairing of $\phi \in \cV'$ and $v\in\cV$ by $\ip{\phi}{v} = \phi(v)$, and define the (Banach space) norm on $\cV'$ by $\|\phi\|_{\cV'} = \sup_{\|v\|=1} |\ip{\phi}{v}|$.
+Let $\cK \subset \cV$ be a nonempty, closed, and convex subset, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.
+For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and a linear \emph{source functional} $\ell\in \cV'$ we consider the following \emph{variational inequality} (VI) problem, with solution $u^*\in \cK$ if it exists:
 \begin{equation}
@@
 \end{equation}
-Because $f$ is (generally) nonlinear, the source term $\ell$ is not strictly needed, and indeed by redefining $f$ we may take $\ell=0$, but its presence clarifies the algorithm in Section \ref{sec:vcycle}.
-
-VI \eqref{eq:vi} generalizes the nonlinear system of equations $f(u^*)=\ell$ to problems where $u^*$ is also required to be in a constraint set $\cK$.
-Informally, if we conceptualize the dual pairing as an inner product, then \eqref{eq:vi} says that the angle between $f(u^*)-\ell$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
-That is, while $f(u^*)-\ell$ may not be zero, it must point directly into $\cK$.
-On the other hand, if $u^*$ is in the interior of the constraint set, i.e.~$u^*\in\cK^\circ$, then \eqref{eq:vi} implies $f(u^*)=\ell$.
-
-\begin{definition} The following definitions are standard \cite{KinderlehrerStampacchia1980}.  A map $f:\cK \to \cV'$ is \emph{monotone} if
+Because $f$ is (generally) nonlinear, the source term $\ell$ is not strictly needed, and by redefining $f$ we may take $\ell=0$, but its presence clarifies the algorithm in Section \ref{sec:vcycle}.
+
+VI \eqref{eq:vi} generalizes nonlinear systems of equations $f(u^*)=\ell$ to problems where $u^*$ is also constrained to be in $\cK$.
+Informally, regarding the dual pairing as an inner product, \eqref{eq:vi} says that the angle between $f(u^*)-\ell$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
+While $f(u^*)-\ell$ may not be zero, it must point directly into $\cK$.
+On the other hand, if $u^*\in\cK^\circ$, in the interior of the constraint set, then \eqref{eq:vi} implies $f(u^*)=\ell$.
+
+The following definitions are standard \cite{KinderlehrerStampacchia1980}.
+
+\begin{definition}  A map $f:\cK \to \cV'$ is \emph{monotone} if
 \begin{equation}
@@
 
-When $f$ is monotone, VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
-In fact, suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cK \to \cV'$.
-Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
-Furthermore, Proposition II.2.1 in \cite{EkelandTemam1976} shows that if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ if and only if
-\begin{equation}
-u^* = \operatorname{arg-min}_{v\in\cK} F(v) - \ip{\ell}{v}. \label{eq:minimization}
-\end{equation}
-The CD methods of Tai \cite{Tai2003} address problem \eqref{eq:minimization} under the hypotheses that such an objective $F$ exists, and that $F'$ is 2-coercive.
-(Note that Tai \cite{Tai2003} uses ``coercive'' for what we call $2$-coercive.)
-
-The problems addressed later in Section \ref{sec:results} involve nonlinear and linear operators $f$ which are partial differential operators.
-As described in section \ref{sec:femultilevel}, the constraint sets $\mathcal{K}$ incorporate both inequality box constraints and nonhomogeneous Dirichlet boundary conditions as equality constraints.
-However, the following examples use homogeneous (zero) Dirichlet conditions for ease of presentation.
-
 Let $\Omega \subset \RR^d$ denote a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.
-Sobolev spaces \cite{Evans2010} are denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$.
-The following example includes the classical obstacle problem for the linear Laplacian \cite{GraeserKornhuber2009} and the $p$-Laplacian for $p\ge 2$ \cite{ChoeLewis1991}.
+Sobolev spaces \cite{Evans2010} are denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$, with $W^{1,p}_0(\Omega)$ denoting the subspace having zero boundary values.
+
+The following example includes the classical and $p$-Laplacian obstacle problems \cite{ChoeLewis1991}.
 
@@
     $$\ip{f(u) - f(v)}{u-v} \ge 2^{2-p} a_0 \|\grad u - \grad v\|_p^p \ge 2^{2-p} a_0 C \|u-v\|^p$$
-for some $C>0$, and thus $f$ is $p$-coercive.  On the other hand, for $\ell\in\cV'$ define
-    $$F(v) = \int_\Omega \frac{a(x)}{p} |\grad v|^p\,dx - \ip{\ell}{v}.$$
+for some $C>0$, and thus $f$ is $p$-coercive.
+The map in \eqref{eq:plaplacian} is also coercive if $1<p<2$, but the proof is somewhat different \cite[Theorem 4.4]{Bueler2021conservation}, and our results (Section \ref{sec:results}) only use $p\ge 2$.
+\end{example}
+
+For monotone operators $f$, VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
+In fact, suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cK \to \cV'$.
+Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
+Furthermore, Proposition II.2.1 in \cite{EkelandTemam1976} shows that if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ if and only if
+\begin{equation}
+u^* = \operatorname{arg-min}_{v\in\cK} F(v) - \ip{\ell}{v}. \label{eq:minimization}
+\end{equation}
+The CD methods of Tai \cite{Tai2003} address problem \eqref{eq:minimization}; the analysis supposes that an objective $F$ exists, with $F'$ 2-coercive.
+
+Not all VI problems arise from optimization, but for Example \ref{ex:plaplacian} we may define
+\begin{equation}
+F(v) = \int_\Omega \frac{a(x)}{p} |\grad v|^p\,dx - \ip{\ell}{v} \label{eq:plaplacianobjective}
+\end{equation}
+for any $\ell\in\cV'$.
 Then $F$ is a convex functional and $F'(v) = f(v) - \ell$ for $f$ given in \eqref{eq:plaplacian}.
-For any closed and convex $\cK\subset \cV$, VI problem \eqref{eq:vi} for $f$ is equivalent to optimization problem \eqref{eq:minimization} for $F$.\end{example}
-
-The map in \eqref{eq:plaplacian} is also coercive if $1<p<2$, but the proof is somewhat different \cite[Theorem 4.4]{Bueler2021conservation}.
-In Section \ref{sec:results} we consider only the $p\ge 2$ case.
-
-However, not all VI problems arise from optimization.
-We give two such examples next, first a coercive and linear advection-diffusion problem, and then a nonlinear porous-medium-type problem.
-The first is preceded by a lemma.
+For any closed and convex $\cK\subset \cV$, VI problem \eqref{eq:vi} for \eqref{eq:plaplacian} is equivalent to optimization problem \eqref{eq:minimization} for \eqref{eq:plaplacianobjective}.
+
+Next we give two examples which are \emph{not} of optimization type, first a linear advection-diffusion problem and then a nonlinear porous-medium-type problem.
 
@@
 \begin{proof}
-Integration by parts gives $b(u,v) = - b(v,u) + \int_{\partial \Omega} uv \bX\cdot \bn\,dx$, so the result follows.
+Integration by parts gives $b(u,v) = - b(v,u) + \int_{\partial \Omega} uv \bX\cdot \bn\,dx$.
 \end{proof}
@@
 Let $\cV = W_0^{1,2}(\Omega)$ be the space of functions which are zero along $\partial_D\Omega$.
-Consider a divergence-free velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, but additionally assume that the flow is outward on the Neumann boundary, $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.
+Consider a divergence-free velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, and assume that the flow is outward on the Neumann boundary, $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.
 For $u,v \in \cV = W_0^{1,2}(\Omega)$ and $\eps>0$ define
@@
 Indeed, $\ip{f(u)}{v}$ is not symmetric in that case, so $f$ cannot be the gradient of a scalar objective.
-To see this, assume $\bX \ne 0$ is continuous for simplicity.
-For $u,v$ which are zero on $\partial \Omega$, note $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.
-By constructing $u,v$ locally near some point where $\bX$ is nonzero one may show $b(u,v)\ne 0$.
+(For $u,v$ which are zero on $\partial \Omega$, note that $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.)
 
 \begin{example}  \label{ex:porous}  Suppose $\phi:[0,\infty) \to [\phi_0,\infty)$ is continuous for some constant $\phi_0>0$.
-Then the following operator, of porous medium type, is not known to be monotone (or coercive) when $\phi$ is not constant:
+As with Example \ref{ex:advectiondiffusion}, the following operator, of porous medium type, does not have the symmetry necessary to be the gradient of a scalar objective:
 \begin{equation}
@@
 \end{equation}
-As with Example \ref{ex:advectiondiffusion}, this form does not have the symmetry necessary to be the gradient of a scalar objective.
+However, \eqref{eq:porous} is not known to be monotone (or coercive) when $\phi$ is not constant.
 \end{example}
@@
 
-If there are $m<\infty$ closed subspaces $\cV_i \subset \cV$ so that the sum
+Suppose there are $m<\infty$ closed subspaces $\cV_i \subset \cV$ so that the sum
 \begin{equation}
@@
 \end{equation}
-holds in the sense that for $w \in \cV$ there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$ then we say \eqref{eq:subspacedecomp} is a \emph{subspace decomposition} of $\cV$ \cite{Xu1992}.
-
+holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$.
+We say \eqref{eq:subspacedecomp} is a \emph{subspace decomposition} of $\cV$ \cite{Xu1992}.
 Suppose further that $\cK_i \subset \cV_i$ are nonempty, closed, and convex subsets such that
@@
 \end{equation}
-for a closed, convex subset $\cK \subset \cV$, as considered in the previous section.
+for a closed, convex subset $\cK \subset \cV$.
 The sum in \eqref{eq:constraintdecomp} is required to hold in two senses \cite{TaiTseng2002}: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.
-Observe that sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp} because the $\cV_i$ are subspaces.
-Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique.
-Also, $\cK_i \not\subset \cK$ in many applications; see the cartoon in Figure \ref{fig:cartoon}.
-
-Finally, for each $\cK_i$ suppose there are bounded, generally-nonlinear \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.
+(Sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp} because the $\cV_i$ are subspaces.)  Finally, for each $\cK_i$ suppose there are bounded, generally-nonlinear \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.
 Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:transfers}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
@@
 A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} \cite{Tai2003}.
+
 Figure \ref{fig:cartoon} suggests how a low-dimensional example might look.
+(Typically $\cK_i \not\subset \cK$.)  Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique.
 
@@
 
-In Sections \ref{sec:femultilevel}--\ref{sec:smoothers} we will introduce multilevel CD discretizations and algorithms for finite-dimensional VI problems.
+In Sections \ref{sec:femultilevel}--\ref{sec:vcycle} we will introduce multilevel CD discretizations and algorithms for finite-dimensional VI problems.
 However, one can observe here that the CD concept applies at the level of the continuum problem.
@@
 Our second example is a disjoint frequency decomposition.
-A multilevel CD over a finite-dimensional FE discretization, specifically the one proposed in Section \ref{sec:femultilevel}, approximates such a frequency decomposition.
+A multilevel CD over a finite-dimensional FE discretization, specifically the one proposed in Section \ref{sec:cdmultilevel}, approximates such a frequency decomposition.
 
@@
 
-Again note that $\cK_i \not\subset \cK$ in these cases.\footnote{Informally speaking, the important inclusion is $\cK_i \subset \cV_i$.}  In Example \ref{ex:domaindecomposition}, if $\psi$ is positive over portions of $\Omega$ where the decomposition into overlapping subdomains $\Omega_i$ is nontrivial, then $\cK_i \not\subset \cK$, and something similar holds for Example \ref{ex:frequencydecomposition}.
+Again note that $\cK_i \not\subset \cK$ in these cases; informally speaking, the important inclusion is $\cK_i \subset \cV_i$.
+In Example \ref{ex:domaindecomposition}, if $\psi$ is positive over portions of $\Omega$ where the decomposition into overlapping subdomains $\Omega_i$ is nontrivial, then $\cK_i \not\subset \cK$, and something similar holds for Example \ref{ex:frequencydecomposition}.
 
@@
 (Similar comments apply for above LDCs $\overline{\chi}^{j}$.)  In this way, monotone restriction construction of the coarser-level box constraints is designed to permit the largest corrections from the coarsest levels \cite{GraeserKornhuber2009}, which promotes V-cycle efficiency.
-However, any system satisfying property \eqref{eq:fe:chiordering} will suffice; the particular formula \eqref{eq:fe:chilevels} is not required when generating the LDCs $\underline{\chi}^{j},\overline{\chi}^{j}$.
-
-We then compute the differences between the previous LDCs:
+Any system satisfying property \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not required when generating the LDCs $\underline{\chi}^{j},\overline{\chi}^{j}$.
+
+We now compute the differences between the LDCs:
 \begin{equation}
@@
 using $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ to define $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.
+Regarding subtraction of extended real-valued quantities, if $\underline{\chi}^j(x_p)=-\infty$ then we set $\underline{\phi}^j(x_p)=-\infty$ regardless of the value of $\underline{\chi}^{j-1}(x_p)$, and similarly $\overline{\phi}^j(x_p)=+\infty$ if $\overline{\chi}^j(x_p)=+\infty$.
+
 These new nodal functions $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ are also LDCs, and again they bracket zero, $\underline{\phi}^j \le 0 \le \overline{\phi}^j$, but they are not ordered as in \eqref{eq:fe:chiordering}.
@@
 \end{equation}
-Table \ref{tab:ldcs} should help the reader keep track of these objects.
+Table \ref{tab:ldcs} should help the reader keep track of our box-constrained LDC system.
 

### 2022-12-21 paper/mcd2.tex
a few more edits ... see FIXME in section 4

@@
 Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below are functions which improve a current iterate $u \in \cK$, generating output $w\in\cK$ which should be closer to the solution $u^* \in \cK$ of \eqref{eq:vi}, by solving smaller VI problems over each CD subset $\cK_i$.
-The two Algorithms each permit a damping parameter $0<\alpha\le 1$, and they generalize the Jacobi and Gauss-Seidel iterations, respectively.
-
-In Algorithm \ref{alg:basiccd-add}, the additive or parallel CD method, ``\textbf{for all}'' can be computed in any order.
+Note that each Algorithm permits a damping parameter $0<\alpha\le 1$.
+
+Algorithm \ref{alg:basiccd-add}, the additive or parallel CD method, generalizes the classical Jacobi iteration.
+Here ``\textbf{for all}'' can be computed in any order.
 Note that the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$ before adding-back the improved subset solution $\hat w_i \in \mathcal{K}_i$.
@@
 
-In the multiplicative (successive) version, Algorithm \ref{alg:basiccd-mult}, each $\mathcal{K}_i$ solution corrects the global iterate immediately, and the $\mathcal{K}_i$ are addressed in a fixed order.
-The sum ``$\sum_{j>i} \Pi_j u$'' should be read as keeping the parts of the input $u$ which have not yet improved.
-Note we must update the subset iterate $w_i$ using the same damping that is applied to the final output $w$.
+Algorithm \ref{alg:basiccd-mult} is the multiplicative (successive) version which generalizes the Gauss-Seidel iteration.
+Each $\mathcal{K}_i$ solution corrects the global iterate immediately, and the order in which the $\mathcal{K}_i$ are addressed matters.
+The sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``parts of the input $u$ which have not yet improved''.
 
@@
 In each boxed VI the argument of $f$ is an element of $\cK$, as the reader should confirm.
-On the other hand, by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may also write the test vector $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors from $\cK$.
+By \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may also write the test vector $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors from $\cK$.
 For the two Algorithms respectively:
@@
 In this sense the boxed VIs are for the same problem as the original VI \eqref{eq:vi}.
-
-For either Algorithm we define the $i$th \emph{subset update} as
-\begin{equation}
-e_i = \hat w_i - \Pi_i u \in \cV_i. \label{eq:ithupdate}
-\end{equation}
-The reader may check that $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.  Trivially, $\hat w = u^* + \sum_i \hat w_i - \Pi_i u^*$ also holds.
+We define the $i$th \emph{subset update} as $e_i = \hat w_i - \Pi_i u \in \cV_i$, and the reader may check that $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.
 
 We will refer to Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} as the \emph{basic CD iterations}.
-These iterations are meaningful even when $f$ is nonlinear, non-local, or defined only on $\cK$.
-However, practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature, and doing so will require new notions.
-In particular, references \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem for which $f$ is linear, local, and defined for all inputs (i.e.~$f$ is defined on all of $\mathcal{V}$).
+
+FIXME show that if $f=F'$ then $F$ is reduced by each Alg \ref{alg:basiccd-mult} iteration; also reduced by Alg \ref{alg:basiccd-add} if $m\alpha=1$ by Jensen's inequality?; $\cK_i$ problems are a type of line search when the corresponding subspaces $\cV_i$ are one-dimensional
+
+The basic CD iterations are meaningful even when $f$ is nonlinear, non-local, or defined only on $\cK$.
+However, practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.
+References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem for which $f$ is linear, local, and defined for all inputs (i.e.~$f$ is defined on all of $\mathcal{V}$).
 The following example exposes the implementation simplifications available in that easier case.
@@
 \begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear.
-Furthermore suppose $f$ is a differential operator, thus local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi \in \mathcal{V}$.
+Furthermore suppose $f$ is a differential operator, thus local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi$.
 Considering additive Algorithm \ref{alg:basiccd-add} for concreteness, by linearity the boxed VI can be written as
@@
 for all $v_i \in \mathcal{K}_i$, where $\tilde\ell = \ell - f(u)$.
-Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible ($e_i \notin \cK$), but \eqref{eq:linearlocalvi} makes sense because $f$ is defined over $\cV$.
 Each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $f(u)$, now included into $\tilde\ell$.
+Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible ($e_i \notin \cK$), but \eqref{eq:linearlocalvi} is meaningful because $f$ is defined over $\cV$.
 If a hat function basis exists for $\cV_i$, namely a basis of small-support functions, then by linearity of $f$ a solver for VI \eqref{eq:linearlocalvi} can be implemented using only integrals over the hat function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}
@@
 For general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation.
+With a non-local integral operator $f$, for example, the expression $\ip{f(\phi)}{z}$ requires an integral over $\Omega$ even if $\phi$ is a function with small support, such as a hat function; cf.~subsection 4.5 in \cite{Bueler2021conservation}.
+
 Section \ref{sec:vcycle} constructs a multilevel CD iteration for functions $f:\cK\to\cV'$, i.e.~defined only on the constraint set, and assuming neither linearity nor locality, by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
 
-Specifically for a non-local integral operator $f$, the expression $\ip{f(\phi)}{z}$ requires an integral over $\Omega$ even if $\phi$ is a function with small support (such as a hat function); cf.~subsection 4.5 in \cite{Bueler2021conservation}.
-While Algorithm \ref{alg:nmcd} applies, as stated, to such $f$, practical extension of multilevel CD methods to non-local operators is a topic for future research.
-
 \section{Multilevel finite elements} \label{sec:femultilevel}
@@
 From now on we are concerned with the approximate solution of certain nonlinear VI problems over 2D computational domains.
-Suppose $\Omega \subset \RR^2$ is a (open) bounded polygon and that $\mathcal{V}=W^{1,p}(\Omega)$ is a Sobolev space of once-differentiable functions.
+Suppose $\Omega \subset \RR^2$ is a (open) bounded polygon and let $\mathcal{V}=W^{1,p}(\Omega)$.
 In addition to the free boundaries defined by inequality constraints (below), fixed boundary conditions may be Dirichlet or mixed.

### 2022-12-21 paper/mcd2.tex
most of the monotonicity lemma completed

@@
 
-Algorithm \ref{alg:basiccd-add}, the additive or parallel CD method, generalizes the classical Jacobi iteration.
+\pr{cd-add}, the additive or parallel CD method, generalizes the classical Jacobi iteration.
 Here ``\textbf{for all}'' can be computed in any order.
@@
 
-Algorithm \ref{alg:basiccd-mult} is the multiplicative (successive) version which generalizes the Gauss-Seidel iteration.
+\pr{cd-mult} is the multiplicative (successive) version which generalizes the Gauss-Seidel iteration.
 Each $\mathcal{K}_i$ solution corrects the global iterate immediately, and the order in which the $\mathcal{K}_i$ are addressed matters.
@@
 By \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may also write the test vector $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors from $\cK$.
-For the two Algorithms respectively:
-\begin{align*}
-[u - \Pi_i u + v_i] - [u - \Pi_i u + \hat w_i] &= v_i - \hat w_i, \label{eq:admissibledifference} \\
-\left[\sum_{j<i} w_j + v_i + \sum_{j>i} \Pi_j u\right] - \left[\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\right] &= v_i - \hat w_i.  \notag
-\end{align*}
+For \pr{cd-add} and \pr{cd-mult} respectively:
+\begin{align}
+[u - \Pi_i u + v_i] - [u - \Pi_i u + \hat w_i] &= v_i - \hat w_i, \label{eq:admissibledifferenceadd} \\
+\left[\sum_{j<i} w_j + v_i + \sum_{j>i} \Pi_j u\right] - \left[\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\right] &= v_i - \hat w_i.  \label{eq:admissibledifferencemult}
+\end{align}
 In this sense the boxed VIs are for the same problem as the original VI \eqref{eq:vi}.
@@
 
-We will refer to Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} as the \emph{basic CD iterations}.
-
-FIXME show that if $f=F'$ then $F$ is reduced by each Alg \ref{alg:basiccd-mult} iteration; also reduced by Alg \ref{alg:basiccd-add} if $m\alpha=1$ by Jensen's inequality?; $\cK_i$ problems are a type of line search when the corresponding subspaces $\cV_i$ are one-dimensional
+We will refer to \pr{cd-add} and \pr{cd-mult} as the \emph{basic CD iterations}.
+Under appropriate assumptions on the damping, these iterations generate monotonically-decreasing objective function values whenever VI problem \eqref{eq:vi} corresponds to convex optimization as in problem \eqref{eq:minimization}.
+In this sense the boxed VI problems in the basic CD iterations are analogous to constrained line searches; they reduce to line searches when the $\cV_i$ are one-dimensional.
+
+\begin{lemma}  Suppose $f=F'$ for $F:\cK\to\RR$ convex and differentiable and let $u\in\cK$.  For any damping $0<\alpha\le 1$, the output $w \in \cK$ from \pr{cd-mult} satisfies
+\begin{equation}
+F(w) - \ip{\ell}{w} \le F(u) - \ip{\ell}{u}.  \label{eq:monotone}
+\end{equation}
+If also $\alpha \le 1/m$ then the output of \pr{cd-add} likewise satisfies \eqref{eq:monotone}.
+\end{lemma}
+
+\begin{proof}  We assume $\ell=0$ without loss of generality; otherwise define $\tilde F(u)=F(u) - \ip{\ell}{u}$ and observe that the proof goes through for $\tilde F$ essentially unchanged.
+
+If $w$ is computed from \pr{cd-mult} then actually each iteration of the \textbf{for} loop lowers the objective value.
+To show this let $z_{i} = \sum_{j\le i} w_j + \sum_{j > i} \Pi_j u \in \cK$ be the state at the end of the $i$th iteration.
+Using convexity, observation \eqref{eq:admissibledifferencemult} with $v_i=\Pi_i u$, and the boxed inequality we find:
+\begin{equation*}
+F(z_{i-1}) - F\left(\sum_{j<i} w_j + \hat w_i + \sum_{j > i} \Pi_j u\right) \ge \ip{f\left(\sum_{j<i} w_j + \hat w_i + \sum_{j > i} \Pi_j u\right)}{\Pi_i u - \hat w_i} \ge 0.
+\end{equation*}
+Using convexity (again), and the above inequality, gives
+\begin{align*}
+F(z_i) &= F\left((1-\alpha) z_{i-1} + \alpha \left(\sum_{j<i} w_j + \hat w_i + \sum_{j > i} \Pi_j u\right)\right) \\
+&\le  (1-\alpha) F(z_{i-1}) + \alpha F\left(\sum_{j<i} w_j + \hat w_i + \sum_{j > i} \Pi_j u\right) \le F(z_{i-1}).
+\end{align*}
+Noting that $w = \sum_{i=0}^{m-1} w_i = z_{m-1}$, the result for \pr{cd-mult} follows.
+
+FIXME $\alpha \le 1/m$ argument
+\end{proof}
+
+When $\alpha=1/m$ exactly then the logic of the proof of monotonicity for \pr{cd-add} is clearer.
+In this case each VI solution over $\cK_i$ lowers the objective value from $F(u)$ to $F(u+e_i)$.
+(As noted, these VI solutions can occur in parallel.)  Then the convex combination of points $\{u+e_i\} \subset \cK$, with equal coefficients $1/m$, is $w=u+\frac{1}{m}\sum_i e_i$.
+By convexity of $F$ it follows that $F(w)$ is lower than the average $\frac{1}{m} \sum_i F(u+e_i)$, which is lower than $F(u)$.
 
@@
 Furthermore suppose $f$ is a differential operator, thus local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi$.
-Considering additive Algorithm \ref{alg:basiccd-add} for concreteness, by linearity the boxed VI can be written as
+Considering \pr{cd-add} for concreteness, by linearity the boxed VI can be written as
 \begin{equation}
@@
 
-When $f$ has all the nice properties listed in Example \ref{ex:fnice} then a solver can take advantage of implementation efficiencies which are unavailable in general.
+When $f$ has all the nice properties listed in Example \ref{ex:fnice} then a solver based on CD iterations can take advantage of implementation efficiencies which are unavailable in general.
 An FE approximation of VI \eqref{eq:vi}---see Sections \ref{sec:femultilevel} and \ref{sec:vcycle}---can then be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.
@@
 
-For general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation.
+For general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of \pr{cd-add} or \pr{cd-mult}.
 With a non-local integral operator $f$, for example, the expression $\ip{f(\phi)}{z}$ requires an integral over $\Omega$ even if $\phi$ is a function with small support, such as a hat function; cf.~subsection 4.5 in \cite{Bueler2021conservation}.
 
-Section \ref{sec:vcycle} constructs a multilevel CD iteration for functions $f:\cK\to\cV'$, i.e.~defined only on the constraint set, and assuming neither linearity nor locality, by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
+Section \ref{sec:vcycle} constructs a multiplicative multilevel CD iteration for functions $f:\cK\to\cV'$, and assuming neither linearity nor locality, by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
 

### 2022-12-22 paper/mcd2.tex
finalize proof of lemma on cd-add/mult

@@
 We will refer to \pr{cd-add} and \pr{cd-mult} as the \emph{basic CD iterations}.
-Under appropriate assumptions on the damping, these iterations generate monotonically-decreasing objective function values whenever VI problem \eqref{eq:vi} corresponds to convex optimization as in problem \eqref{eq:minimization}.
-In this sense the boxed VI problems in the basic CD iterations are analogous to constrained line searches; they reduce to line searches when the $\cV_i$ are one-dimensional.
+According to damping assumptions, these iterations generate monotonically-decreasing objective values if VI problem \eqref{eq:vi} corresponds to convex optimization (problem \eqref{eq:minimization}), as shown in the following Lemma.
+Thus the boxed VI problems in the \pr{cd-add} and \pr{cd-mult} algorithms are analogous to constrained line searches, and in fact they reduce to such when the $\cV_i$ are one-dimensional.
 
@@
 
-If $w$ is computed from \pr{cd-mult} then actually each iteration of the \textbf{for} loop lowers the objective value.
+Each iteration of the \textbf{for} loop in \pr{cd-mult} lowers the objective value.
 To show this let $z_{i} = \sum_{j\le i} w_j + \sum_{j > i} \Pi_j u \in \cK$ be the state at the end of the $i$th iteration.
-Using convexity, observation \eqref{eq:admissibledifferencemult} with $v_i=\Pi_i u$, and the boxed inequality we find:
+Using convexity, observation \eqref{eq:admissibledifferencemult} with $v_i=\Pi_i u$, and the boxed VI we find:
 \begin{equation*}
@@
 
-FIXME $\alpha \le 1/m$ argument
+Suppose $w=u+\alpha\sum_i e_i$ is computed by \pr{cd-add}.  Considering each $e_i=\hat w_i - \Pi_i u$ separately we have, by convexity and the boxed VI,
+\begin{equation*}
+F(u) - F(u+e_i) \ge \ip{f(u+e_i)}{u-(u+e_i)} = \ip{f(u-\Pi_i u + \hat w_i)}{\Pi_i u - \hat w_i} \ge 0.
+\end{equation*}
+If $z = u+ \frac{1}{m} \sum_i e_i$ then, again using convexity (Jensen's inequality),
+\begin{align*}
+F(u) - F(z) &= \frac{1}{m} \sum_{i=0}^{m-1} F(u) - F\left(u + \frac{1}{m} \sum_{i=0}^{m-1} e_i\right) \\
+&\ge \frac{1}{m} \sum_{i=0}^{m-1} F(u+e_i) - F\left(\frac{1}{m} \sum_{i=0}^{m-1} (u+e_i)\right) \ge 0.
+\end{align*}
+Now suppose $\alpha \le 1/m$ and define $\lambda = m\alpha$ so $0<\lambda\le 1$ and $w = (1-\lambda) u + \lambda z$.
+Thus by convexity and the above result, $F(w) \le (1-\lambda) F(u) + \lambda F(z) \le F(u)$.
 \end{proof}
 
-When $\alpha=1/m$ exactly then the logic of the proof of monotonicity for \pr{cd-add} is clearer.
-In this case each VI solution over $\cK_i$ lowers the objective value from $F(u)$ to $F(u+e_i)$.
-(As noted, these VI solutions can occur in parallel.)  Then the convex combination of points $\{u+e_i\} \subset \cK$, with equal coefficients $1/m$, is $w=u+\frac{1}{m}\sum_i e_i$.
-By convexity of $F$ it follows that $F(w)$ is lower than the average $\frac{1}{m} \sum_i F(u+e_i)$, which is lower than $F(u)$.
+When $\alpha=1/m$ then the logic of the above proof for \pr{cd-add} is simpler.
+Because each boxed VI over $\cK_i$ reduces the objective value from $F(u)$ to $F(u+e_i)$, we have $F(w)$ less than $\frac{1}{m} \sum_i F(u+e_i)$, which is lower than $F(u)$.
 

### 2022-12-22 paper/mcd2.tex
edits sections 4 and 5

@@
 
-Associated to a constraint decomposition (CD) are certain basic iterative methods.
-Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below are functions which improve a current iterate $u \in \cK$, generating output $w\in\cK$ which should be closer to the solution $u^* \in \cK$ of \eqref{eq:vi}, by solving smaller VI problems over each CD subset $\cK_i$.
-Note that each Algorithm permits a damping parameter $0<\alpha\le 1$.
-
-\pr{cd-add}, the additive or parallel CD method, generalizes the classical Jacobi iteration.
+Associated to a constraint decomposition (CD) are certain iterative methods.
+Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below solve ``smaller'' VI problems, over each subset $\cK_i$, to replace $u \in \cK$ with $w\in\cK$ which should be closer to $u^* \in \cK$ solving \eqref{eq:vi}.
+First is \pr{cd-add}, the additive or parallel method, which generalizes the classical Jacobi iteration.
 Here ``\textbf{for all}'' can be computed in any order.
-Note that the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$ before adding-back the improved subset solution $\hat w_i \in \mathcal{K}_i$.
+Note that the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$, and then adds-back the improved subset solution $\hat w_i \in \mathcal{K}_i$.
+Next is \pr{cd-mult}, the multiplicative or successive version, which generalizes the Gauss-Seidel iteration.
+Each subset solution corrects the global iterate immediately, so the $\mathcal{K}_i$ ordering is important.
+The sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``parts of the input $u$ which have not yet improved''.
 
@@
 \end{pseudofloat}
-
-\pr{cd-mult} is the multiplicative (successive) version which generalizes the Gauss-Seidel iteration.
-Each $\mathcal{K}_i$ solution corrects the global iterate immediately, and the order in which the $\mathcal{K}_i$ are addressed matters.
-The sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``parts of the input $u$ which have not yet improved''.
 
@@
 In each boxed VI the argument of $f$ is an element of $\cK$, as the reader should confirm.
-By \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may also write the test vector $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors from $\cK$.
+Also, by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write the test vector $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors from $\cK$.
 For \pr{cd-add} and \pr{cd-mult} respectively:
@@
 \end{align}
-In this sense the boxed VIs are for the same problem as the original VI \eqref{eq:vi}.
-We define the $i$th \emph{subset update} as $e_i = \hat w_i - \Pi_i u \in \cV_i$, and the reader may check that $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.
+In this sense the boxed VIs solve the same problem as \eqref{eq:vi}.
+
+We define the $i$th \emph{subset update} as $e_i = \hat w_i - \Pi_i u \in \cV_i$.  The reader may check that $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.
 
 We will refer to \pr{cd-add} and \pr{cd-mult} as the \emph{basic CD iterations}.
-According to damping assumptions, these iterations generate monotonically-decreasing objective values if VI problem \eqref{eq:vi} corresponds to convex optimization (problem \eqref{eq:minimization}), as shown in the following Lemma.
+Note that each permits a damping parameter $\alpha$.
+Subject to a bound on $\alpha$, which is restrictive for \pr{cd-add} as noted by Tai \cite{Tai2003}, these iterations generate monotonically-decreasing objective values whenever VI problem \eqref{eq:vi} corresponds to convex optimization (problem \eqref{eq:minimization}).
+This is shown in the following Lemma.
 Thus the boxed VI problems in the \pr{cd-add} and \pr{cd-mult} algorithms are analogous to constrained line searches, and in fact they reduce to such when the $\cV_i$ are one-dimensional.
 
-\begin{lemma}  Suppose $f=F'$ for $F:\cK\to\RR$ convex and differentiable and let $u\in\cK$.  For any damping $0<\alpha\le 1$, the output $w \in \cK$ from \pr{cd-mult} satisfies
+\begin{lemma}  Suppose $f=F'$ for $F:\cK\to\RR$ convex and differentiable, and let $u\in\cK$.  For $0<\alpha\le 1$ the output $w \in \cK$ from \pr{cd-mult} satisfies
 \begin{equation}
@@
 \end{equation}
-If also $\alpha \le 1/m$ then the output of \pr{cd-add} likewise satisfies \eqref{eq:monotone}.
+If $0 < \alpha \le 1/m$ then \pr{cd-add} likewise satisfies \eqref{eq:monotone}.
 \end{lemma}
@@
 Each iteration of the \textbf{for} loop in \pr{cd-mult} lowers the objective value.
-To show this let $z_{i} = \sum_{j\le i} w_j + \sum_{j > i} \Pi_j u \in \cK$ be the state at the end of the $i$th iteration.
-Using convexity, observation \eqref{eq:admissibledifferencemult} with $v_i=\Pi_i u$, and the boxed VI we find:
+To show this let $z_{i} = \sum_{j\le i} w_j + \sum_{j > i} \Pi_j u \in \cK$, $i=0,\dots,m-1$, be the state at the end of the $i$th iteration.
+Note that $z_0=u$, $z_{m-1}=w$, and $z_{i-1} = \sum_{j < i} w_j + \Pi_i u + \sum_{j > i} \Pi_j u$ for $i>0$.
+Using this last fact, convexity of $F$, observation \eqref{eq:admissibledifferencemult} with $v_i=\Pi_i u$, and the boxed VI we find:
 \begin{equation*}
@@
 \end{equation*}
-Using convexity (again), and the above inequality, gives
+Using convexity (again) with the above inequality gives
 \begin{align*}
@@
 \end{align*}
-Noting that $w = \sum_{i=0}^{m-1} w_i = z_{m-1}$, the result for \pr{cd-mult} follows.
-
-Suppose $w=u+\alpha\sum_i e_i$ is computed by \pr{cd-add}.  Considering each $e_i=\hat w_i - \Pi_i u$ separately we have, by convexity and the boxed VI,
+As $w = z_{m-1}$, the result for \pr{cd-mult} follows.
+
+Suppose $w=u+\alpha\sum_i e_i$ is computed by \pr{cd-add} using $0<\alpha \le 1/m$.  Considering each $e_i=\hat w_i - \Pi_i u$ separately, convexity of $F$ and the boxed VI with $v_i=\Pi_i u$ yields
+\begin{equation}
+F(u) - F(u+e_i) \ge \ip{f(u+e_i)}{u-(u+e_i)} = \ip{f(u-\Pi_i u + \hat w_i)}{\Pi_i u - \hat w_i} \ge 0. \label{eq:cdaddonereduction}
+\end{equation}
+Let $v = u+ \frac{1}{m} \sum_i e_i$.  Again using convexity (Jensen's inequality), plus the above inequality, we have
 \begin{equation*}
-F(u) - F(u+e_i) \ge \ip{f(u+e_i)}{u-(u+e_i)} = \ip{f(u-\Pi_i u + \hat w_i)}{\Pi_i u - \hat w_i} \ge 0.
+F(v) = F\left(\frac{1}{m} \sum_{i=0}^{m-1} (u+e_i)\right) \le \frac{1}{m} \sum_{i=0}^{m-1} F(u+e_i) \le \frac{1}{m} \sum_{i=0}^{m-1} F(u) = F(u).
 \end{equation*}
-If $z = u+ \frac{1}{m} \sum_i e_i$ then, again using convexity (Jensen's inequality),
-\begin{align*}
-F(u) - F(z) &= \frac{1}{m} \sum_{i=0}^{m-1} F(u) - F\left(u + \frac{1}{m} \sum_{i=0}^{m-1} e_i\right) \\
-&\ge \frac{1}{m} \sum_{i=0}^{m-1} F(u+e_i) - F\left(\frac{1}{m} \sum_{i=0}^{m-1} (u+e_i)\right) \ge 0.
-\end{align*}
-Now suppose $\alpha \le 1/m$ and define $\lambda = m\alpha$ so $0<\lambda\le 1$ and $w = (1-\lambda) u + \lambda z$.
-Thus by convexity and the above result, $F(w) \le (1-\lambda) F(u) + \lambda F(z) \le F(u)$.
+Now suppose $\alpha \le 1/m$ and define $\lambda = m\alpha$ so $0<\lambda\le 1$ and $w = (1-\lambda) u + \lambda v$.
+Thus by convexity and the above result, $F(w) \le (1-\lambda) F(u) + \lambda F(v) \le F(u)$.
 \end{proof}
 
-When $\alpha=1/m$ then the logic of the above proof for \pr{cd-add} is simpler.
-Because each boxed VI over $\cK_i$ reduces the objective value from $F(u)$ to $F(u+e_i)$, we have $F(w)$ less than $\frac{1}{m} \sum_i F(u+e_i)$, which is lower than $F(u)$.
-
-The basic CD iterations are meaningful even when $f$ is nonlinear, non-local, or defined only on $\cK$.
+When $\alpha=1/m$ then the logic of the above proof for \pr{cd-add} is simpler because $v=w$ and $\lambda=1$.
+The value $F(w)$ is less than $\frac{1}{m} \sum_i F(u+e_i)$ by Jensen's inequality, which is lower than $F(u)$ by the boxed VIs, i.e.~inequality \eqref{eq:cdaddonereduction}.
+
+The basic CD iterations are meaningful whether or not $f=F'$.
+Furthermore they are meaningful even when $f$ is nonlinear, non-local, or defined only on $\cK$.
 However, practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.
-References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem for which $f$ is linear, local, and defined for all inputs (i.e.~$f$ is defined on all of $\mathcal{V}$).
+References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem for which $f=F'$ is linear, local, and defined for all inputs (i.e.~$f$ is defined on all of $\mathcal{V}$).
 The following example exposes the implementation simplifications available in that easier case.
@@
 for all $v_i \in \mathcal{K}_i$, where $\tilde\ell = \ell - f(u)$.
-Each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $f(u)$, now included into $\tilde\ell$.
+Thus each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $f(u)$ included into $\tilde\ell$.
 Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible ($e_i \notin \cK$), but \eqref{eq:linearlocalvi} is meaningful because $f$ is defined over $\cV$.
-If a hat function basis exists for $\cV_i$, namely a basis of small-support functions, then by linearity of $f$ a solver for VI \eqref{eq:linearlocalvi} can be implemented using only integrals over the hat function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}
+If a hat function basis exists for $\cV_i$, namely a basis of small-support functions such as hat functions, then by locality of $f$ a solver for VI \eqref{eq:linearlocalvi} can be implemented using only integrals over these small support sets.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}
 \end{example}
@@
 When $f$ has all the nice properties listed in Example \ref{ex:fnice} then a solver based on CD iterations can take advantage of implementation efficiencies which are unavailable in general.
-An FE approximation of VI \eqref{eq:vi}---see Sections \ref{sec:femultilevel} and \ref{sec:vcycle}---can then be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.
-The actual solution process for VI \eqref{eq:linearlocalvi} is not really the concern here; we are observing that the storage needed for problem \eqref{eq:linearlocalvi}, and the cost of residual evaluation, are substantially smaller than in the general case.
+That is, an FE approximation of VI \eqref{eq:vi}---see Sections \ref{sec:femultilevel} and \ref{sec:vcycle}---can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.
+Whatever is the actual solution process for VI \eqref{eq:linearlocalvi}, both storage and the computational cost of residual evaluation are substantially smaller than in the general case.
 
@@
 With a non-local integral operator $f$, for example, the expression $\ip{f(\phi)}{z}$ requires an integral over $\Omega$ even if $\phi$ is a function with small support, such as a hat function; cf.~subsection 4.5 in \cite{Bueler2021conservation}.
-
-Section \ref{sec:vcycle} constructs a multiplicative multilevel CD iteration for functions $f:\cK\to\cV'$, and assuming neither linearity nor locality, by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
+In any case, Section \ref{sec:vcycle} constructs a multiplicative multilevel CD iteration for functions $f:\cK\to\cV'$, based on additive or multiplicative smoothers in Section \ref{sec:smoothers}, and assuming neither linearity nor locality, by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
 
@@
 
-From now on we are concerned with the approximate solution of certain nonlinear VI problems over 2D computational domains.
+From now on we seek the approximate solution of certain nonlinear VI problems over 2D computational domains.
 Suppose $\Omega \subset \RR^2$ is a (open) bounded polygon and let $\mathcal{V}=W^{1,p}(\Omega)$.
-In addition to the free boundaries defined by inequality constraints (below), fixed boundary conditions may be Dirichlet or mixed.
-Suppose that the domain boundary is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive measure for simplicity, to permit well-posedness \cite{Evans2010}, and $\partial_N \Omega$ possibly empty.
+In addition to the free boundaries defined by inequality constraints (below), the additional fixed boundary conditions are allowed to be either Dirichlet or mixed.
+Suppose that $\partial\Omega$ is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive measure for simplicity, to permit well-posedness \cite{Evans2010}, and $\partial_N \Omega$ possibly empty.
 Suppose that the provided Dirichlet data $g_D:\partial_D \Omega \to \RR$ is sufficiently-regular so that the VI problem \eqref{eq:boxdirichletvi} below is well posed.
 
-We restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.
+We restrict to the case of inequality \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.
 The \emph{lower obstacle} $\underline{\gamma}$ and the \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions on the closure $\overline{\Omega}$ which take their values in the extended real line $\tilde\RR = [-\infty,+\infty]$.
@@
 \begin{equation}
-\cK = \left\{w\,:\,\underline{\gamma} \le w \le \overline{\gamma} \,\text{ over }\, \Omega, \, \text{ and }\, w\big|_{\partial_D \Omega} = g_D\right\} \subset \cV =W^{1,p}(\Omega).
-\end{equation}
-(The boundary values are, as usual, in the trace sense \cite{Evans2010}.)  It is easy to check that $\cK$ is closed and convex.
-This definition of $\cK$ treats the inequality and equality (boundary) constraints in a unified manner.
+\cK = \left\{w\,:\,\underline{\gamma} \le w \le \overline{\gamma} \, \text{ and }\, w\big|_{\partial_D \Omega} = g_D\right\} \subset \cV =W^{1,p}(\Omega).
+\end{equation}
+(The boundary values are, as usual, in the trace sense \cite{Evans2010}.)  This definition treats the inequality and equality (boundary) constraints in a unified manner.
+Note that $\cK$ is closed and convex.
 
@@
 Note that $v-u^* \in W_0^{1,p}(\Omega)$, that is, test functions in \eqref{eq:boxdirichletvi} have zero values on the Dirichlet boundary $\partial_D\Omega$.
-
-Problem \eqref{eq:boxdirichletvi} is both a free-boundary problem, because of the inequality constraints in defining $\cK$, and a fixed-boundary value problem.
-Dirichlet conditions $u=g_D$ apply over $\partial_D\Omega$ and Neumann conditions over $\partial_N \Omega$.
-The latter are determined by the formula for $f$; for nonhomogeneous Neumann conditions, the nonlinear operator $f:\cK\to\cV'$ must include an integration over $\partial_N\Omega$.
+While Dirichlet conditions apply on $\partial_D\Omega$, Neumann conditions over $\partial_N \Omega$ are determined by the formula for $f$.
+(E.g.~operator $f:\cK\to\cV'$ may include an integration over $\partial_N\Omega$.)  In summary, VI \eqref{eq:boxdirichletvi} is both a free- and fixed-boundary value problem.
 
 We now construct a multilevel finite element (FE) approximation to problem \eqref{eq:boxdirichletvi} using either triangular or quadrilateral meshes over $\Omega$, and $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, respectively.
-A hierarchy of meshes will be constructed by refinement of a coarsest mesh.
-A solution on the finest-level mesh is the goal.
-This hierarchy also allows the creation of a multilevel CD of $\cK$.
-The subsets $\cK_i$ in this CD are designed to allow multigrid-type coarser-level corrections (next Section) in a manner which preserves admissibility, while not introducing unnecessary high frequencies into the coarser-level problems, especially at the free boundary.
+A hierarchy of meshes will be constructed by refinement of a coarsest mesh, with the goal being a solution on the finest-level mesh.
+This hierarchy also allows the creation of a multilevel CD of $\cK$ in Section \ref{sec:cdmultilevel}.
+The subsets $\cK_i$ in this CD permit coarser-level corrections in a manner which preserves admissibility while not introducing unnecessary high frequencies into the coarser-level problems, especially at the free boundary.
 
@@
 
-Use of higher-order Lagrange elements $P_k$ or $Q_k$, $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} special to inequality constraints.
+Use of higher-order Lagrange elements $P_k$ or $Q_k$, $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} special to methods for inequality constraints.
 Namely, our FE construction of constraint sets will require \emph{nodal monotonicity},
@@
 \end{equation}
-for all FE functions $w,z \in \mathcal{V}^j$, where $\bw,\bz$ denote vectors of nodal values.
+for all $w,z \in \mathcal{V}^j$, where $\bw,\bz$ denote vectors of nodal values.
 For $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal value one at a single node, otherwise zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.
-It is easy to show that \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of sign-changing hat functions, so \eqref{eq:nodalmonotonicity} implies $k\le 1$.
-For this reason we restrict to $k = 1$.
+It is easy to show that \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of sign-changing hat functions, so use of \eqref{eq:nodalmonotonicity} implies $k\le 1$.
 
 Let $g_D^J \in \mathcal{V}^J$, on the finest-level mesh, satisfy $g_D^J = g_D$ on $\partial_D \Omega$.
-Typically of conforming FE methods, this restricts the Dirichlet boundary function $g_D$ to be a piecewise-linear \cite{Elmanetal2014}.
-
-To allow box constraints, let $\tilde{\mathcal{V}}^j$ denote the set of all functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
-(Note that $\tilde{\mathcal{V}}^j$ functions are not in $\mathcal{V}^j$.
-They are not defined everywhere on $\Omega$, but only on nodes.)  Suppose that the \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$, for nodes of the finest-level mesh $\mathcal{T}^J$, satisfy
+Typically of conforming FE methods, this restricts the Dirichlet boundary function $g_D$ to be piecewise-linear \cite{Elmanetal2014}.
+
+To allow box constraints, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
+(Note that $\tilde{\mathcal{V}}^j$ functions are not in $\mathcal{V}^j$ and not defined everywhere on $\Omega$.)  Suppose that the \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$, for nodes of the finest-level mesh $\mathcal{T}^J$, satisfy
 \begin{equation}
@@
 Finally assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.
-
-Define
-\begin{equation}
-\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
+Then define
+\begin{equation}
+\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \, \text{ and } \, w|_{\partial_D\Omega} = g_D|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
 \end{equation}
@@
 Let $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ be a discretization of $f$ in \eqref{eq:boxdirichletvi}.
-By inclusion \eqref{eq:fe:nestedspaces} the functional $\ell$ is already defined and continuous over $\mathcal{V}^J$, but we denote $\ell^J \in (\mathcal{V}^J)'$.
+By inclusion \eqref{eq:fe:nestedspaces} the functional $\ell^J=\ell$ is defined and continuous over $\mathcal{V}^J$.
 We seek the FE solution $u^J \in \mathcal{K}^J$ of the finest-level, finite-dimensional, nonlinear VI
@@
 \end{equation}
-As is true for \eqref{eq:vi}, if $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, then \eqref{eq:fe:vi} is well-posed.
+As for \eqref{eq:vi}, if $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, then \eqref{eq:fe:vi} is well-posed.
 
@@
 That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require $\mathcal{K}^J \subset \mathcal{K}$.
-Even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants will generally not imply continuum admissiblity.
-Such caveats apply in geophysical situations where only nodal values of the obstacle are known, and given by meshed elevation data.
-The continuum limit of nodal obstacles $\underline{\gamma}^J$, $\overline{\gamma}^J$ may be unknown \cite{Bueler2016}.
+Even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants will generally not imply continuum admissiblity.\footnote{Such caveats are relevant in geophysical situations where only nodal values of the obstacle are known and nodal elevation data misses detail in between; see the Greenland ice sheet application in \cite{Bueler2016}, for example.}
 
@@
 Stating this strong-form problem may be clarifying for some readers.
-Denoting $x_p$ for nodes of the finest-level mesh $\mathcal{T}^J$, let $X=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap X$.
+Denote $x_p$ for a node of the finest-level mesh $\mathcal{T}^J$ and $\psi_p(x)$ for the corresponding hat function.
+Let $X=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap X$.
 If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let us define $I_D$ as the $m_D\times m_J$ matrix which indicates (extracts) Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet values as nodes in $(\mathcal{V}^J)'$.
-Problem \eqref{eq:fe:vi} must be stated as VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap X\right)$, in the operations research literature \cite{FerrisPang1997}, and it is equivalent to the equality-constrained MCP
+Problem \eqref{eq:fe:vi} can be stated as VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap X\right)$, e.g.~in the operations research literature \cite{FerrisPang1997}.
+It is equivalent to the following MCP with additional equality constraints:
 \begin{align}
 \underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J - I_D^\top \lambda \label{eq:fe:mcp} \\
-I_D u^J &= g_D^J\big|_{\partial_D\Omega}. \notag
+I_D u^J &= g_D^J\big|_{\partial_D\Omega} \notag
 \end{align}
-The Lagrange multipliers $\lambda \in \RR^{m_D}$ are free (unconstrained).
-Let $\psi_p(x)$ denote the hat function for a node $x_p \in \mathcal{T}^J$.
-MCP \eqref{eq:fe:mcp} says that at each node exactly\footnote{The stated conditions are correct anywhere $\underline{\gamma}^J(x_p) < \overline{\gamma}^J(x_p)$.
-Where the box constraints meet (pinch) at $x_p \notin \partial_D\Omega$ then $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$ holds.} one of the following conditions holds for the solution $u^J$:
+for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
+If $\underline{\gamma}^J(x_p) < \overline{\gamma}^J(x_p)$ then MCP \eqref{eq:fe:mcp} says that exactly one of the following conditions holds for the solution $u^J$:
 \begin{itemize}
@@
 \end{itemize}
-These cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, and \emph{boundary}, respectively.
+(Where the box constraints meet (pinch) at $x_p \notin \partial_D\Omega$ then $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$, and the value $\ip{f^J(u^J)-\ell^J}{\psi_p}$ is unconstrained.)  The four cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, and \emph{boundary}, respectively.
 Where case \emph{boundary} applies, $\lambda_p=\ip{f^J(u^J)-\ell^J}{\psi_p}\in\RR$ has any value.

### 2022-12-22 paper/mcd2.tex
edits through end section 6

@@
 NMCD methods should permit direct solutions of those VI problems for which non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
-% FIXME and convergence is monotone
 
@@
 (Sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp} because the $\cV_i$ are subspaces.)  Finally, for each $\cK_i$ suppose there are bounded, generally-nonlinear \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.
-Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:transfers}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
+Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:femultilevel}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
 \begin{equation}
@@
 
-\section{Multilevel transfer operators} \label{sec:transfers}
-
 The multilevel algorithm defined in the next two Sections, a solver for finite-dimensional problem \eqref{eq:fe:vi}, requires transfer operators between meshes.
-We use five such operators, three of which are standard within multigrid methods \cite{Trottenbergetal2001}.
+We require five such operators, but three are standard multigrid tools \cite{Trottenbergetal2001}.
 Table \ref{tab:transfers} will help the reader keep track.
@@
 Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ also acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $(R\sigma)[z] = \sigma[z]$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
-Observe that these canonical transfers typically involve a change of FE representation.
+However, these canonical transfers involve a change of FE representation.
 For example, the formula for canonical restriction, derived by writing coarser-mesh hat functions in terms of finer-mesh ones, is a ``full-weighting'' average \cite{Trottenbergetal2001}.
@@
 The monotone restrictions $\maxR,\minR$ are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh hat functions.
-These restrictions will be applied only to nodal functions ($\tilde{\mathcal{V}}^j$).
+These restrictions will be applied only to nodal functions $\tilde{\mathcal{V}}^j$.
 To make the definitions precise, let $x_q^j$, $q=0,1,\dots,m_j-1$, denote the nodes of mesh $\mathcal{T}^j$, and denote the corresponding hat functions by $\psi_q^j(x)$.
-The monotone restrictions $\maxR z,\minR z$ of an $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ are coarser-level functions in $\tilde{\mathcal{V}}^{j-1}$ defined by
+The monotone restrictions $\maxR z,\minR z$ of an $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ are the nodal functions in $\tilde{\mathcal{V}}^{j-1}$ defined by
 \begin{align}
@@
 for $p=0,1,\dots,m_{j-1}-1$.
-(In computing $\max$ and $\min$, the usual $\tilde\RR$ ordering is applied.)  Because, by mesh refinement, a node $x_p^{j-1}$ is in the set of $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, the following formal property holds:
+(In computing $\max$ and $\min$, the usual $\tilde\RR$ ordering is applied.)  By mesh refinement a node $x_p^{j-1}$ is in the set of $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so it follows that
 \begin{equation}
@@
 Suppose $w^J \in \cK^J$ is an admissible finest-level iterate which approximates the solution $u^J$ of VI \eqref{eq:fe:vi}.
-We will now define the CD constraint subsets, in our multilevel FE context, using $w^J$ and the defining inequalities for $\cK^J$, namely \eqref{eq:fe:fineconstraintset}.
-Though the ideas are fundamentally the same as for unilateral constraints \cite{GraeserKornhuber2009}, the notation is necessarily more elaborate because we consider bi-obstacle (box-constrained) problems.
-
-From the iterate $w^J$ we define the \emph{finest-level defect constraints}
-\begin{equation}
-\underline{\chi}^J = \underline{\gamma}^J - w^J, \qquad \overline{\chi}^J = \overline{\gamma}^J - w^J, \label{eq:fe:defectconstraints}
-\end{equation}
-as nodal functions in $\tilde{\mathcal{V}}^J$.
+From $w^J$ we now define the CD constraint subsets, in our multilevel FE context, using the defining inequalities for $\cK^J$.
+Though the ideas are fundamentally the same as for unilateral constraints, e.g.~as in \cite{GraeserKornhuber2009}, the notation is necessarily more elaborate because we consider box-constrained (bi-obstacle) problems.
+
+We define the \emph{finest-level defect constraints} \cite{GraeserKornhuber2009} as the following nodal functions in $\tilde{\mathcal{V}}^J$:
+\begin{equation}
+\underline{\chi}^J = \underline{\gamma}^J - w^J, \qquad \overline{\chi}^J = \overline{\gamma}^J - w^J. \label{eq:fe:defectconstraints}
+\end{equation}
 Note that $\underline{\chi}^J \le 0 \le \overline{\chi}^J$ everywhere in $\overline{\Omega}$, including along the Dirichlet boundary $\partial_D\Omega$.
-(Recall that on $\partial_D\Omega$ we have $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ and $w^J = g_D^J = g_D$.)  If $\underline{\gamma}^J=-\infty$ then $\underline{\chi}^J=-\infty$ by definition, and similarly if $\overline{\gamma}^J=+\infty$ then $\overline{\chi}^J=+\infty$.
-The motivation for creating such defect constraints \cite{GraeserKornhuber2009} is that a corrected iterate $w^J + z^J$, for $z^J\in\cV^J$, is admissible ($w^J + z^J \in \cK^J$) if and only if the perturbation $z^J$ is between the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$, and $z^J=0$ on $\partial_D\Omega$.
-
-Counting down the coarser levels $j=J-1,\dots,1,0$, we apply the monotone restrictions to generate ``\emph{below}'' and ``\emph{above}'' \emph{level defect constraints} (LDCs):
-\begin{equation}
-\underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}, \qquad \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}, \label{eq:fe:chilevels}
-\end{equation}
-thus $\underline{\chi}^{j},\overline{\chi}^{j} \in \tilde{\mathcal{V}}^J$.  Observe that
+If $\underline{\gamma}^J=-\infty$ then $\underline{\chi}^J=-\infty$ by definition, and similarly if $\overline{\gamma}^J=+\infty$ then $\overline{\chi}^J=+\infty$.
+
+Recall that on $\partial_D\Omega$ we have $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ and $w^J = g_D^J$.
+The motivation for creating defect constraints $\underline{\chi}^J,\overline{\chi}^J$ is that a corrected iterate $w^J + z^J$ is admissible, $w^J + z^J \in \cK^J$, if and only if the perturbation $z^J$ is between the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$, and $z^J=0$ on $\partial_D\Omega$.
+
+Next, counting down the coarser levels $j=J-1,\dots,1,0$, we apply the monotone restrictions to generate box \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:
+\begin{equation}
+\underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}, \qquad \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}. \label{eq:fe:chilevels}
+\end{equation}
+Observe that
 \begin{equation}
@@
 \end{equation}
-(in the sense of nodal values).  A $J=4$ example of ``below'' LDCs $\underline{\chi}^j$ is shown in Figure \ref{fig:chiphilevels}.
+i.e.~as nodal values.  A $J=4$ example of LDCs $\underline{\chi}^j$ is shown in Figure \ref{fig:chiphilevels}.
 
@@
 \includegraphics[width=0.65\textwidth]{fixfigs/chiphilevels.pdf}
-\caption{The ``below'' level defect constraints (LDC) $\underline{\chi}^j$ are generated from $\underline{\chi}^J = \underline{\gamma}^J - w^J$ using maximum restriction: $\underline{\chi}^j = \maxR \underline{\chi}^{j+1}$.}
+\caption{The level defect constraints (LDC) $\underline{\chi}^j$ are generated from $\underline{\chi}^J = \underline{\gamma}^J - w^J$ using maximum restriction: $\underline{\chi}^j = \maxR \underline{\chi}^{j+1}$.}
 \label{fig:chiphilevels}
@@
 
-Intuitively-speaking, below LDC is the most-negative function from $\mathcal{V}^j$ such that if another element of $\mathcal{V}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$ at the finer level.
-(Similar comments apply for above LDCs $\overline{\chi}^{j}$.)  In this way, monotone restriction construction of the coarser-level box constraints is designed to permit the largest corrections from the coarsest levels \cite{GraeserKornhuber2009}, which promotes V-cycle efficiency.
-Any system satisfying property \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not required when generating the LDCs $\underline{\chi}^{j},\overline{\chi}^{j}$.
-
-We now compute the differences between the LDCs:
-\begin{equation}
-\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1}, \qquad \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1},  \label{eq:fe:philevels}
-\end{equation}
-using $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ to define $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.
+Intuitively-speaking, the LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if another element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$, and similar comments apply to $\overline{\chi}^{j}$.
+This construction of the box constraints on each coarser level is designed to permit the largest corrections from the coarsest levels \cite{GraeserKornhuber2009}, which promotes V-cycle efficiency.
+Note that for the multilevel scheme in the following Sections, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not required when generating the LDCs $\underline{\chi}^{j},\overline{\chi}^{j}$.
+
+We now compute the LDC differences:
+\begin{equation}
+\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1}, \qquad \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}.  \label{eq:fe:philevels}
+\end{equation}
+We take $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ so that $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.
 Regarding subtraction of extended real-valued quantities, if $\underline{\chi}^j(x_p)=-\infty$ then we set $\underline{\phi}^j(x_p)=-\infty$ regardless of the value of $\underline{\chi}^{j-1}(x_p)$, and similarly $\overline{\phi}^j(x_p)=+\infty$ if $\overline{\chi}^j(x_p)=+\infty$.
-
-These new nodal functions $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ are also LDCs, and again they bracket zero, $\underline{\phi}^j \le 0 \le \overline{\phi}^j$, but they are not ordered as in \eqref{eq:fe:chiordering}.
-However, telescoping sums hold for $j=0,1,\dots,J$:
+These new LDCs $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ again bracket zero: $\underline{\phi}^j \le 0 \le \overline{\phi}^j$.
+Though  they are not ordered as in \eqref{eq:fe:chiordering}, telescoping sums hold for any $j=0,1,\dots,J$:
 \begin{equation}
@@
 \begin{tabular}{llc}
-\emph{LDC name}        & \emph{formula} \\ \hline
-(finest) upward below & ${\Large \strut} \underline{\chi}^J = \underline{\gamma}^J - w^J$ \\
-(finest) upward above & ${\Large \strut} \overline{\chi}^J = \overline{\gamma}^J - w^J$ \\
-upward below          & ${\Large \strut} \underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}$ \\
-upward above          & ${\Large \strut} \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}$ \\
-downward below        & ${\Large \strut} \underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1}$ \\
-downward above        & ${\Large \strut} \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}$ \\
+\emph{name}        & \emph{formulas} \\ \hline
+(finest) upward & ${\Large \strut} \underline{\chi}^J = \underline{\gamma}^J - w^J,\quad \overline{\chi}^J = \overline{\gamma}^J - w^J$ \\
+upward          & ${\Large \strut} \underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}, \quad \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}$ \\
+downward        & ${\Large \strut} \underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1}, \quad \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}$
 \end{tabular}
@@
 \medskip
-\caption{Level defect constraints (LDCs) are constructed from the box constraints ($\underline{\gamma}^J,\overline{\gamma}^J$) in problem \eqref{eq:fe:vi}, and a finest level iterate ($w^J \in \cK^J$).}
+\caption{Box level defect constraints (LDCs).}
 \label{tab:ldcs}

### 2022-12-23 paper/mcd2.tex
a couple of edits at start of section 7

@@
 
-A multilevel V-cycle solver is defined in the next Section.
-For the downward side of the V-cycle we will use the LDCs $\underline{\phi}^j,\overline{\phi}^j$ as constraints, and $\underline{\chi}^j,\overline{\chi}^j$ for the upward side.
-The corresponding constraint sets are
+A multilevel V-cycle solver is defined in the next Section, which uses LDCs $\underline{\phi}^j,\overline{\phi}^j$ for the downward side of the V-cycle and $\underline{\chi}^j,\overline{\chi}^j$ for the upward side.
+The corresponding \emph{downward} and \emph{upward defect constraint sets}, respectively, are
 \begin{align}
@@
 \end{align}
-These are the \emph{downward} and \emph{upward defect constraint sets}, respectively, closed and convex subsets of the FE spaces $\mathcal{V}^j$, for $j=0,1,\dots,J$.
+These are closed and convex subsets of the FE spaces $\mathcal{V}^j$, for $j=0,1,\dots,J$.
 Because $\underline{\chi}^j \le \underline{\phi}^j \le 0 \le \overline{\phi}^j \le \overline{\chi}^j$, it follows that $\mathcal{D}^j \subset \mathcal{U}^j$; the upward sets are larger.

### 2022-12-23 paper/mcd2.tex
edits section 7; starting nmcd->fascd rename

@@
 \usepackage[dvipsnames]{xcolor}
-\usepackage{longtable}
-\usepackage{booktabs}
+\usepackage{longtable,booktabs,xspace}
 
@@
 \item We apply a full approximation storage (FAS; see \cite{Brandt1977,Bruneetal2015}) approach which avoids global (Newton) linearization.
-Section \ref{sec:results} shows results from this \emph{nonlinear multilevel constraint decomposition} (NMCD) algorithm  for several nonlinear problems.
-\item Our implementation permits ``box'' constraints, namely upper and lower obstacles; see Section \ref{sec:femultilevel}.
-\item In Sections \ref{sec:femultilevel} and \ref{sec:vcycle} we explain why, given the manner in which constraint sets are decomposed in the multilevel context \cite{GraeserKornhuber2009}, ``up-smoothing'' is more efficient than ``down-smoothing''.
+Section \ref{sec:results} shows results from this \pr{fascd} algorithm for several nonlinear problems.
+\item The \pr{fascd} implementation permits ``box'' constraints, namely upper and lower obstacles; see Section \ref{sec:femultilevel}.
+\item In Sections \ref{sec:ldcs}--\ref{sec:vcycle} we explain why, given the manner in which constraint sets are decomposed in the multilevel context \cite{GraeserKornhuber2009}, ``up-smoothing'' is more efficient than ``down-smoothing''.
 A strong preference for up-smoothing is special to multilevel CD methods and does not arise in unconstrained problems.
@@
 The iterates from a CD method are always admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
-Our NMCD scheme is designed for such admissible-only problems, some of which require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation, as discussed in \cite{Bueler2021conservation}.
-NMCD methods should permit direct solutions of those VI problems for which non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
-
-One of our examples (Section \ref{sec:results}) demonstrates the application of the NMCD method to a nonlinear operator of porous-media type.
+\pr{fascd} is designed for such admissible-only problems, some of which require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation, as discussed in \cite{Bueler2021conservation}.
+\pr{fascd} should permit direct solutions of those VI problems for which non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
+
+One of our examples (Section \ref{sec:results}) demonstrates the application of the \pr{fascd} method to a nonlinear operator of porous-media type.
 By ``freezing'' the solution-dependent coefficient, the operator is approximated by a coercive operator for the duration of the V-cycle.
@@
 By contrast a CD method applies essentially at the level of the continuum problem; see Section \ref{sec:cd}.
-Acceleration of the NMCD method via active-set and/or basis-level manipulations is a topic for future research.
+Acceleration of the \pr{fascd} method via active-set and/or basis-level manipulations is a topic for future research.
 
@@
 
-A multilevel V-cycle solver is defined in the next Section, which uses LDCs $\underline{\phi}^j,\overline{\phi}^j$ for the downward side of the V-cycle and $\underline{\chi}^j,\overline{\chi}^j$ for the upward side.
+A multilevel V-cycle solver is defined in the next Section.
+This V-cycle uses LDCs $\underline{\phi}^j,\overline{\phi}^j$ for the downward side and $\underline{\chi}^j,\overline{\chi}^j$ for the upward side.
 The corresponding \emph{downward} and \emph{upward defect constraint sets}, respectively, are
@@
 \end{align}
-These are closed and convex subsets of the FE spaces $\mathcal{V}^j$, for $j=0,1,\dots,J$.
-Because $\underline{\chi}^j \le \underline{\phi}^j \le 0 \le \overline{\phi}^j \le \overline{\chi}^j$, it follows that $\mathcal{D}^j \subset \mathcal{U}^j$; the upward sets are larger.
+These are closed and convex subsets of the FE spaces $\mathcal{V}^j$.
+Because $\underline{\chi}^j \le \underline{\phi}^j \le 0 \le \overline{\phi}^j \le \overline{\chi}^j$, it follows that $\mathcal{D}^j \subset \mathcal{U}^j$.
 Note that the finest-level set $\mathcal{U}^J$ is equivalent to the original constraint set $\mathcal{K}^J$: for all $z^J \in \mathcal{V}^J$,
@@
 
-Our multilevel strategy is based on constraint decompositions (CDs; Section \ref{sec:cd}) of $\mathcal{U}^J$, or rather incomplete CDs as we now explain.
-Each upward set $\mathcal{U}^j$ can be decomposed in an inclusion sense, which becomes a true CD in unilateral cases.
+Our multilevel strategy is based on incomplete constraint decompositions (CDs; Section \ref{sec:cd}) of $\mathcal{U}^J$, as we now explain.
+Each upward set $\mathcal{U}^j$ can be decomposed in an inclusion sense, becoming a true CD in unilateral cases.
 The following Lemma is based on the unilateral CD construction underlying Algorithm 4.7 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.
-It explains downward admissibility in the V-cycle Algorithm \ref{alg:nmcd} in Section \ref{sec:vcycle}.
-
-\begin{lemma}  \label{lem:downwardadmissibility}  For arbitrary upper and lower obstacles $\underline{\gamma}^J,\overline{\gamma}^J$ satisfying \eqref{eq:fe:boxconstraintrequirements}, and any $j=0,1,\dots,J$,
-\begin{equation}
-\mathcal{U}^j \supseteq \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^0. \label{eq:fe:downwardsuminclusion}
-\end{equation}
-In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, inclusion \eqref{eq:fe:downwardsuminclusion} becomes a full CD \eqref{eq:constraintdecomp}, namely $\mathcal{U}^j=\sum_{i=0}^j \mathcal{D}^i$, over the subspace decomposition $\cV^j=\sum_{i=0}^j \cV^i$.
+
+\begin{lemma}  \label{lem:downwardadmissibility}  For arbitrary upper and lower obstacles $\underline{\gamma}^J,\overline{\gamma}^J$ satisfying \eqref{eq:fe:boxconstraintrequirements}, and $j=0,1,\dots,J$,
+\begin{equation}
+\mathcal{U}^j \supset \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^0. \label{eq:fe:downwardsuminclusion}
+\end{equation}
+In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, inclusion \eqref{eq:fe:downwardsuminclusion} becomes a full CD \eqref{eq:constraintdecomp}, namely $\mathcal{U}^j=\sum_{i=0}^j \mathcal{D}^i$.
 \end{lemma}
 
-\begin{proof}  Subspace decomposition \eqref{eq:subspacedecomp} holds---see nesting \eqref{eq:fe:nestedspaces}---and definition \eqref{eq:fe:constraintsets} shows $\mathcal{D}^i \subset \cV^i$ and $\mathcal{U}^i \subset \cV^i$.
-If $y^i \in \mathcal{D}^i$ for $i=0,\dots,J$ then
+\begin{proof}  Subspace decomposition \eqref{eq:subspacedecomp} holds---see nesting \eqref{eq:fe:nestedspaces}---and definition \eqref{eq:fe:constraintsets} shows $\mathcal{D}^i \subset \mathcal{U}^i \subset \cV^i$.
+If $y^i \in \mathcal{D}^i$ for $0 \le i \le j$ then by \eqref{eq:fe:telescoping},
 \begin{equation}
@@
 \end{equation}
-by \eqref{eq:fe:telescoping} and \eqref{eq:fe:constraintsets}, thus \eqref{eq:fe:downwardsuminclusion} holds for any box constraints.
-
-It remains to consider unilateral cases.
-Here we can construct decomposition operators and show \eqref{eq:constraintrestrictionsum}.
+thus \eqref{eq:fe:downwardsuminclusion} holds for any box constraints.
+
+In unilateral cases we can also construct decomposition operators $\Pi_i$ and show \eqref{eq:constraintrestrictionsum}.
 Suppose $\overline{\gamma}^J=+\infty$, thus $\overline{\chi}^j=+\infty$ and $\overline{\phi}^i = +\infty$ for all levels.
@@
 \end{align*}
-This shows \eqref{eq:constraintrestrictionsum}, thus that \eqref{eq:fe:downwardsuminclusion} is equality, a true CD \eqref{eq:constraintdecomp}.
-
-The case where $\underline{\gamma}^J=-\infty$ is similar.
+This shows \eqref{eq:constraintrestrictionsum}, thus that \eqref{eq:fe:downwardsuminclusion} is equality, a true CD \eqref{eq:constraintdecomp}.  The $\underline{\gamma}^J=-\infty$ case is similar.
 \end{proof}
@@
 However, the following Example shows why $\Pi_i(\min\{z^j,0\})$ may have an arbitrarily large maximum, and thus exceed any finite upper LDC $\overline{\phi}^i$, so unfortunately $\Pi_i$ does not map even non-positive functions to $\mathcal{D}^i$ in the presence of a finite upper LDC.
-While the proposed splitting strategy does not work, without redefinition of a formula which is adequate in the unilateral case, the existence of an entirely-different strategy for showing a full CD is also not excluded.
+On the other hand, the existence of an entirely-different strategy for showing a full CD is not excluded.
 
@@
 Each upward defect constraint set $\mathcal{U}^j$ can also be be decomposed down to an arbitrary level $k$ where an upward set $\mathcal{U}^k$ is used as the coarsest set.
-The next Lemma, similar to Lemma \ref{lem:downwardadmissibility} but not following from it, explains upward admissibility in V-cycle Algorithm \ref{alg:nmcd}.
+While similar to Lemma \ref{lem:downwardadmissibility}, the next Lemma does not follow logically from it.
 
@@
 \begin{proof}  Definition \eqref{eq:fe:philevels} shows that $\underline{\chi}^j = \underline{\phi}^j + \dots + \underline{\phi}^{k+1} + \underline{\chi}^k$ and $\overline{\chi}^j = \overline{\phi}^j + \dots + \overline{\phi}^{k+1} + \overline{\chi}^k$.
-(These facts replace \eqref{eq:fe:telescoping}.)  If $y^i \in \mathcal{D}^i$ for $i=k+1,\dots,j$ and $z^k \in \mathcal{U}^k$ then it is easy to show, as in \eqref{eq:fe:lemmaordering}, that $\underline{\chi}^j \le y^j + \dots + y^{k+1} + z^k \le \overline{\chi}^j$, thus \eqref{eq:fe:upwardsuminclusion} holds.
+(These facts replace \eqref{eq:fe:telescoping}.)  If $y^i \in \mathcal{D}^i$ for $k+1 \le i \le j$ and $z^k \in \mathcal{U}^k$ then it is easy to show, as in \eqref{eq:fe:lemmaordering}, that $\underline{\chi}^j \le y^j + \dots + y^{k+1} + z^k \le \overline{\chi}^j$, thus \eqref{eq:fe:upwardsuminclusion} holds.
 
@@
 Definition \eqref{eq:fe:minimummaps} is unchanged.
-We use \eqref{eq:fe:unilateraldecompositionoperator} unchanged for $i=k+1,\dots,j$, but modify it in the $i=k$ case: $\Pi_k z^j = I_{j\to k}^\ominus(z^j - \underline{\chi}^j) + \underline{\chi}^k$.
+We can use \eqref{eq:fe:unilateraldecompositionoperator} unchanged for $i=k+1,\dots,j$, but it must be modified in the $i=k$ case: $\Pi_k z^j = I_{j\to k}^\ominus(z^j - \underline{\chi}^j) + \underline{\chi}^k$.
 The remainder of the proof goes through as for Lemma \ref{lem:downwardadmissibility}.
@@
 In summary, inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in V-cycle Algorithm \ref{alg:nmcd}.
-However, a convergence analysis following \cite{Tai2003} or \cite{GraeserKornhuber2009}, somehow extending their arguments to VIs \eqref{eq:vi} which are not of minimization type, or to box constraints, would seem to require full CDs, and perhaps new ideas.
-
-\section{Nonlinear multilevel constraint decomposition (NMCD) V-cycle} \label{sec:vcycle}
-
-The \emph{nonlinear multilevel constraint decomposition} (NMCD) V-cycle algorithm in this section uses CDs \eqref{eq:fe:downwardsumcd} and \eqref{eq:fe:upwardsumcd}.
-It does smoothing in $\mathcal{D}^j$ for the downward direction, and $\mathcal{U}^j$ for the upward direction, respectively, in the V-cycle.
-The algorithm generalizes the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear \emph{full approximation scheme} (FAS) approach \cite{BrandtLivne2011}.
+Lemma \ref{lem:downwardadmissibility} explains downward admissibility, and Lemma \ref{lem:upwardadmissibility} upward admissibility, in Algorithm \ref{alg:nmcd}.
+However, a convergence analysis following \cite{Tai2003} or \cite{GraeserKornhuber2009}, somehow extending their arguments to VIs \eqref{eq:vi} which are not of minimization type, and/or to box constraints, would seem to require full CDs, and perhaps new ideas.
+
+\newcommand{\fascd}{\pr{fascd}\xspace}
+
+\section{Full approximation scheme constraint decomposition (\fascd) V-cycle} \label{sec:vcycle}
+
+The \fascd V-cycle algorithm in this section uses Lemmas \ref{lem:downwardadmissibility} and \ref{lem:upwardadmissibility}, that is, it smooths in $\mathcal{D}^j$ during the downward stages, and in $\mathcal{U}^j$ for the upward stages, respectively.
+The algorithm extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear FAS approach \cite{BrandtLivne2011}.
 We coarsen the solution iterate itself, using injection, in constrast to classical linear multigrid wherein only the residual is coarsened.
@@
 \begin{center}
-\input{tikz/nmcdvcycle.tex}
+\input{tikz/fascdvcycle.tex}
 \end{center}
-\caption{An NMCD V-cycle computes downward corrections $y_j \in \mathcal{D}^j$ and upward corrections $z_j\in\mathcal{U}^j$.  The latter live in larger constraint subsets.}
-\label{fig:nmcdvcycle}
+\caption{A \fascd V-cycle computes downward corrections $y_j \in \mathcal{D}^j$ and upward corrections $z_j\in\mathcal{U}^j$.  The latter live in larger constraint subsets.}
+\label{fig:fascdvcycle}
 \end{figure}
@@
 
-To justify the NMCD algorithm---see Algorithm \ref{alg:nmcd} below---we must state the problems solved at each mesh level.\footnote{Smoothers/solvers for these problems are proposed in Section \ref{sec:smoothers}.}  As in classical FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, along with creating a source functional.
+To justify \fascd, namely Algorithm \ref{alg:nmcd} below, we must state the problems solved at each mesh level.\footnote{Smoothers/solvers for these problems are proposed in Section \ref{sec:smoothers}.}  As in classical FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, along with creating a source functional.
 Recall that Table \ref{tab:transfers} summarizes the needed transfer operators.
@@
 \begin{pseudofloat}[ht]
-\begin{pseudo} \label{ps:nmcd-vcycle}
-\pr{nmcd-vcycle}(\ell^J,\underline{\gamma}^J,\overline{\gamma}^J;w^J)\text{:} \\+
+\begin{pseudo} \label{ps:fascd-vcycle}
+\pr{fascd-vcycle}(\ell^J,\underline{\gamma}^J,\overline{\gamma}^J;w^J)\text{:} \\+
     $\underline{\chi}^J = \underline{\gamma}^J - w^J {\large \strut}$ \\
@@
 \end{pseudo}
-\caption{Nonlinear multilevel constraint decomposition V-cycle as an iterative solver for FE VI problem \eqref{eq:fe:vi}.
-$f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
+\caption{The \fascd V-cycle as an iterative solver for FE VI problem \eqref{eq:fe:vi}.  $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
 \label{alg:nmcd}

### 2022-12-23 paper/fascd.tex
rename files: mcd2->fascd
edits through start section 8

--- 2022-12-23 paper/mcd2.tex
@@
 \newcommand{\Span}{\operatorname{span}}
+
+\newcommand{\fascd}{\pr{fascd}\xspace}
 
@@
 \begin{equation}
-F(w) - \ip{\ell}{w} \le F(u) - \ip{\ell}{u}.  \label{eq:monotone}
-\end{equation}
-If $0 < \alpha \le 1/m$ then \pr{cd-add} likewise satisfies \eqref{eq:monotone}.
+F(w) - \ip{\ell}{w} \le F(u) - \ip{\ell}{u}.  \label{eq:objectivemonotone}
+\end{equation}
+If $0 < \alpha \le 1/m$ then \pr{cd-add} likewise satisfies \eqref{eq:objectivemonotone}.
 \end{lemma}
@@
 
-\newcommand{\fascd}{\pr{fascd}\xspace}
-
 \section{Full approximation scheme constraint decomposition (\fascd) V-cycle} \label{sec:vcycle}
 
-The \fascd V-cycle algorithm in this section uses Lemmas \ref{lem:downwardadmissibility} and \ref{lem:upwardadmissibility}, that is, it smooths in $\mathcal{D}^j$ during the downward stages, and in $\mathcal{U}^j$ for the upward stages, respectively.
-The algorithm extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear FAS approach \cite{BrandtLivne2011}.
-We coarsen the solution iterate itself, using injection, in constrast to classical linear multigrid wherein only the residual is coarsened.
-
-The finest-level iterate $w^J \in \mathcal{K}^J$ will be corrected by perturbations from each coarser subspace $\mathcal{V}^j$,\footnote{By nesting \eqref{eq:fe:nestedspaces} these corrections also live in the fine-level subspace $\mathcal{V}^J$.
-Later we will use prolongation $P$ to indicate where a change in computer representation would be needed.} but the corrected iterate needs to be admissible.
-In detail, suppose $y^J \in \mathcal{D}^J, y^{J-1} \in \mathcal{D}^{J-1}, \dots, y^{j+1} \in \mathcal{D}^{j+1}$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:nmcdvcycle}).
-By \eqref{eq:fe:downwardsumcd} the corrected iterate at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.
-Smoothing then happens at the next-coarser level $j$, i.e.~we incompletely-solve VI problem \eqref{eq:fe:downvi} below.
-At the bottom the coarsest-level VI is then (accurately) solved, to generate the full downslash correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.
-Starting upward, however, we can use $y^1+ z^0 \in \mathcal{U}^1$ as an initial iterate on level $j=1$, and then smooth by incompletely-solving VI problem \eqref{eq:fe:upvi} (below) in $\mathcal{U}^1$, yielding admissible total correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$ using \eqref{eq:fe:upwardsumcd}.
-Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$ from initial iterate $y^j+z^{j-1}$, eventually generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction, yielding the new iterate $w^J + z^J \in \mathcal{K}^J$.
+The \fascd V-cycle algorithm in this section uses Lemmas \ref{lem:downwardadmissibility} and \ref{lem:upwardadmissibility}, that is, it smooths in $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.
+\fascd extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear FAS approach \cite{BrandtLivne2011}.
+Note that we coarsen both the residual and the solution iterate itself, in constrast to classical linear multigrid wherein only the residual is coarsened.
+In fact, \fascd reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).
+
+The V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ based on perturbations from each coarser FE subspace $\mathcal{V}^j$.
+By nesting \eqref{eq:fe:nestedspaces} these corrections also live in the fine-level subspace $\mathcal{V}^J$, but we will use prolongation $P$ to indicate where the computer representation changes.
+
+However, the corrected iterate needs to be admissible.
+In detail, suppose $y^i \in \mathcal{D}^i$ for $i=J,\dots,j-1$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).
+By \eqref{eq:fe:downwardsuminclusion} the corrected iterate at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.
+Smoothing, an incomplete solution process on a coarser-level VI problem (defined below), then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.
+At the bottom the coarsest-level VI is then solved, resulting in a downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.
+Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$.
+Smoothing in $\mathcal{U}^1$ yields admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$ using \eqref{eq:fe:upwardsuminclusion}.
+Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction.
+Thus the next admissible iterate is $w^J + z^J \in \mathcal{K}^J$.
 
@@
 \end{center}
-\caption{A \fascd V-cycle computes downward corrections $y_j \in \mathcal{D}^j$ and upward corrections $z_j\in\mathcal{U}^j$.  The latter live in larger constraint subsets.}
+\caption{The \fascd V-cycle computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.}
 \label{fig:fascdvcycle}
@@
 
-A key idea, implicit in the above paragraph, is that the up-smoothing stages solve less-restricted VI problems than down-smoothing stages.
-Intuitively-speaking, downward one does not know what the coarse corrections will be, so they must be in such small sets ($\mathcal{D}^j$) that their sum is admissible once we return to the current level.
-However, going upward smoothing can occur in the larger admissible-sum sets ($\mathcal{U}^j$; see \eqref{eq:fe:constraintdecomp}) because there are no forthcoming coarser corrections which could violate admissibility.
+A key idea is that, because $\mathcal{U}^j\supset \mathcal{D}^j$, up-smoothing solves less-restricted VI problems than down-smoothing.
+Intuitively-speaking, going downward one does not know what the upcoming coarse corrections will be, so they must be in the sufficiently-small sets $\mathcal{D}^j$ so that their sum is admissible once we return to a level.
+However, going upward no further \emph{coarser} corrections, relative to that level, can violate admissibility.
 
@@
 
-\bibliography{mcd2}
+\bibliography{fascd}
 \bibliographystyle{siam}
@@
 \appendix
-\section{Reductions of NMCD to familiar multilevel algorithms}
+\section{Reductions of NMCD to familiar multilevel algorithms} \label{app:reductions}
 

### 2022-12-28 paper/fascd.tex
edits section 8

@@
 
-\section{Full approximation scheme constraint decomposition (\fascd) V-cycle} \label{sec:vcycle}
+\section{Full approximation storage constraint decomposition (\fascd) V-cycle} \label{sec:vcycle}
 
 The \fascd V-cycle algorithm in this section uses Lemmas \ref{lem:downwardadmissibility} and \ref{lem:upwardadmissibility}, that is, it smooths in $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.
-\fascd extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear FAS approach \cite{BrandtLivne2011}.
-Note that we coarsen both the residual and the solution iterate itself, in constrast to classical linear multigrid wherein only the residual is coarsened.
-In fact, \fascd reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).
+\fascd extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear FAS approach \cite{BrandtLivne2011}, and in fact it reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).
 
@@
 By nesting \eqref{eq:fe:nestedspaces} these corrections also live in the fine-level subspace $\mathcal{V}^J$, but we will use prolongation $P$ to indicate where the computer representation changes.
+Note that we coarsen both the residual and the solution iterate, in constrast to classical linear multigrid wherein only the residual is coarsened.
 
@@
 In detail, suppose $y^i \in \mathcal{D}^i$ for $i=J,\dots,j-1$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).
-By \eqref{eq:fe:downwardsuminclusion} the corrected iterate at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.
-Smoothing, an incomplete solution process on a coarser-level VI problem (defined below), then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.
-At the bottom the coarsest-level VI is then solved, resulting in a downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.
+By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.
+Smoothing, an incomplete solution process on a VI problem (Section \ref{sec:smoothers}), then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.
+At the bottom the coarsest-level VI is then solved (also Section \ref{sec:smoothers}), resulting in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.
 Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$.
@@
 Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction.
-Thus the next admissible iterate is $w^J + z^J \in \mathcal{K}^J$.
+Then the next admissible iterate is $w^J + z^J \in \mathcal{K}^J$.
 
@@
 
-To justify \fascd, namely Algorithm \ref{alg:nmcd} below, we must state the problems solved at each mesh level.\footnote{Smoothers/solvers for these problems are proposed in Section \ref{sec:smoothers}.}  As in classical FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, along with creating a source functional.
-Recall that Table \ref{tab:transfers} summarizes the needed transfer operators.
-
-Let $j$ denote the mesh level of the coarsened problem.
-We assume a (re-)discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, an FE discretization of $f$ in \eqref{eq:boxdirichletvi}, using quadrature and etc., presumably by the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.
-
-The formulas are as follows.
-Finest-level problem \eqref{eq:fe:vi} is defined by the discretized operator $f^J$ and the original source functional $\ell^J$.
-A finest-level, admissible iterate $w^J \approx u^J$, $w^J \in \mathcal{K}^J$, has already\footnote{Conceptually-speaking, at least.
-In practice one generates LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ during the descent stage; see Algorithm \ref{alg:nmcd}.
-Observe that the LDCs depend neither on coarsened iterates $g^j$ nor corrections $y^j,z^j$.} been used to general the LDCs and constraint sets $\mathcal{D}^j,\mathcal{U}^j$ at each level (Section \ref{sec:femultilevel}).
-As we descend in the V-cycle, we recursively define $j$th-level iterates
-\begin{equation}
-g^j = \begin{cases} w^J, & j=J \\
-                    \iR(g^{j+1} + y^{j+1}), & j < J.
-      \end{cases}  \label{eq:fe:defineg}
-\end{equation}
-These are the iterates \emph{before} smoothing and/or correction has been applied at level $j$.  (The problem solved by $y^{j+1}$ is stated momentarily.)  From $g^j$ we define $\ell^j \in (\cV^j)'$:
+To state \fascd Algorithm \ref{alg:nmcd} below we must clarify the problems solved at each mesh level.
+As in FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, and a source functional created.
+On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:boxdirichletvi}, presumably using quadrature and etc.~in the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.
+Note that finest-level problem \eqref{eq:fe:vi} uses the original source functional $\ell^J$, and an admissible iterate $w^J \in \mathcal{K}^J$ has already\footnote{Conceptually-speaking this is true.
+In practice the LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ are generated during the descent stage; see Algorithm \ref{alg:nmcd}.
+Observe that the LDCs do not depend on $w^j$ for $j<J$, nor on the corrections $y^j,z^j$.} been used to general the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:ldcs}).
+
+As we descend $j<J$, we define new $j$th-level iterates
+\begin{equation}
+w^j = \iR(w^{j+1} + y^{j+1}),  \label{eq:fe:definew}
+\end{equation}
+the value \emph{before} smoothing and/or correction at level $j$.  From $w^j$ we define $\ell^j \in (\cV^j)'$:
 \begin{equation}
 \ell^j = \begin{cases} \ell^J, & j=J \\
-                       f^j(g^j) + R\left(\ell^{j+1}-f^{j+1}(g^{j+1}+y^{j+1})\right), & j<J. \end{cases} \label{eq:fe:levelsource}
-\end{equation}
-Observe that injection $\iR$ coarsens the iterates $g^j$---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---while canonical restriction $R$ coarsens the source functionals.
-
-On descent, $j=J,J-1,\dots,1$, we solve the following VI for $y^j \in \mathcal{D}^j$:
-\begin{equation}
-\ip{f^j(g^j + y^j)}{v-y^j} \ge \ip{\ell^j}{v-y^j} \qquad \text{for all } v\in \mathcal{D}^j, \label{eq:fe:downvi}
-\end{equation}
-for which the initial iterate is
-\begin{equation}
-(y^j)^{(0)}=0.  \label{eq:fe:downwardinitial}
-\end{equation}
-(Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)  On ascent, $j=0,1,\dots,J$, the same VI is solved, but in a larger admissible set.  We find $z^j \in \mathcal{U}^j$ so that
-\begin{equation}
-\ip{f^j(g^j + z^j)}{v-z^j} \ge \ip{\ell^j}{v-z^j} \qquad \text{for all } v\in \mathcal{U}^j. \label{eq:fe:upvi}
-\end{equation}
-The nontrivial initial iterate for this problem is found by prolonging the coarser-level correction $z^{j-1}$, and using it to correct the current (down-smoothed) iterate on the level:\footnote{Prolongation $P$ is used in \eqref{eq:fe:upwardinitial} to reflect the change in mesh/FE representation, from $\mathcal{T}^{j-1}$ to $\mathcal{T}^{j}$.}
-\begin{equation}
-(z^j)^{(0)} = y^j + P z^{j-1}.  \label{eq:fe:upwardinitial}
-\end{equation}
-For the $j=0$ coarsest level, $(z^0)^{(0)}=0$.
-Initial iterate \eqref{eq:fe:upwardinitial} is admissible since $\mathcal{U}^j = \mathcal{D}^j + \mathcal{U}^{j-1}$; see \eqref{eq:fe:constraintdecomp} or \eqref{eq:fe:upwardsumcd}.
-
-If we were to inject the original bi-obstacles from \eqref{eq:fe:fineconstraintset} downward to coarser meshes, i.e.~$\underline{\gamma}^j = \iR \underline{\gamma}^{j+1}$ and $\overline{\gamma}^j = \iR \overline{\gamma}^{j+1}$, then, because injection preserves inequalities, and because $y^{j+1} \in \mathcal{D}^{j+1}$, it would follow from \eqref{eq:fe:defineg} that $\underline{\gamma}^j \le g^j \le \overline{\gamma}^j$, just like $\underline{\gamma}^J \le w^J \le \overline{\gamma}^J$ for the finest-level iterate.
-In this narrow sense $g^j$ is admissible with respect to the original (finest-level) box constraints.
-(In fact, while $g^j \in \cV^j \subset \cV^J$ certainly holds, generally $g^j \notin \cK^J$.)  However, following \cite{GraeserKornhuber2009}, our use of defect constraints (Table \ref{tab:ldcs}) avoids the need to generate functions $\underline{\gamma}^j$ and $\overline{\gamma}^j$, and it yields a final, finest-level iterate $g^J+z^J \in \cK^J$ which is genuinely admissible for problem \eqref{eq:fe:vi}.
+                       f^j(w^j) + R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right), & j<J. \end{cases} \label{eq:fe:levelsource}
+\end{equation}
+Observe that injection $\iR$ coarsens the iterates---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---while canonical restriction $R$ coarsens linear functionals.
+(Recall Table \ref{tab:transfers} summarizes our transfer operators.)
+
+On descent, at levels $j=J,J-1,\dots,1$, we solve the following VI for $y^j \in \mathcal{D}^j$:
+\begin{equation}
+\ip{f^j(w^j + y^j)}{v-y^j} \ge \ip{\ell^j}{v-y^j} \qquad \text{for all } v\in \mathcal{D}^j, \label{eq:fe:downvi}
+\end{equation}
+for which the initial iterate is $y_{(0)}^j=0$.
+(Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)  On ascent, $j=0,1,\dots,J$, the same VI is solved, but in a larger admissible set, for $z^j \in \mathcal{U}^j$:
+\begin{equation}
+\ip{f^j(w^j + z^j)}{v-z^j} \ge \ip{\ell^j}{v-z^j} \qquad \text{for all } v\in \mathcal{U}^j. \label{eq:fe:upvi}
+\end{equation}
+The nontrivial initial iterate for this problem is found by prolonging the coarser-level correction $z^{j-1}$, and using it to correct the current (down-smoothed) correction on the level,\footnote{Prolongation $P$ is used in \eqref{eq:fe:upwardinitial} to reflect the change in FE representation, from $\mathcal{V}^{j-1}$ to $\mathcal{V}^{j}$.}
+\begin{equation}
+z_{(0)}^j = y^j + P z^{j-1},  \label{eq:fe:upwardinitial}
+\end{equation}
+but with $z_{(0)}^0=0$ on the coarsest level.  Note $z_{(0)}^j$ is admissible since $\mathcal{D}^j + \mathcal{U}^{j-1} \subset \mathcal{U}^j$; see \eqref{eq:fe:upwardsuminclusion}.
+
+If we were to inject the original box constraints \eqref{eq:fe:fineconstraintset} downward to coarser meshes, i.e.~$\underline{\gamma}^j = \iR \underline{\gamma}^{j+1}$ and $\overline{\gamma}^j = \iR \overline{\gamma}^{j+1}$, then, because injection preserves inequalities, and because $y^{j+1} \in \mathcal{D}^{j+1}$, it would follow from \eqref{eq:fe:definew} that $\underline{\gamma}^j \le w^j \le \overline{\gamma}^j$.
+In this sense $w^j$ is admissible with respect to the box constraints, but, following \cite{GraeserKornhuber2009}, our use of LDCs (Table \ref{tab:ldcs}) avoids the need to generate such box constraints $\underline{\gamma}^j$ and $\overline{\gamma}^j$ on each level.
+(Note that while $w^j \in \cV^j \subset \cV^J$ holds, generally $w^j \notin \cK^J$.)  In any case, Algorithm \ref{alg:nmcd} will yield a final, finest-level iterate $w^J+z^J \in \cK^J$ which is genuinely admissible for problem \eqref{eq:fe:vi}.
 

### 2022-12-28 paper/fascd.tex
done with edits section 8

@@
 
-In summary, inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in V-cycle Algorithm \ref{alg:nmcd}.
-Lemma \ref{lem:downwardadmissibility} explains downward admissibility, and Lemma \ref{lem:upwardadmissibility} upward admissibility, in Algorithm \ref{alg:nmcd}.
+In summary, inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in V-cycle Algorithm \ref{alg:fascd}.
+Lemma \ref{lem:downwardadmissibility} explains downward admissibility, and Lemma \ref{lem:upwardadmissibility} upward admissibility, in Algorithm \ref{alg:fascd}.
 However, a convergence analysis following \cite{Tai2003} or \cite{GraeserKornhuber2009}, somehow extending their arguments to VIs \eqref{eq:vi} which are not of minimization type, and/or to box constraints, would seem to require full CDs, and perhaps new ideas.
@@
 
-To state \fascd Algorithm \ref{alg:nmcd} below we must clarify the problems solved at each mesh level.
+To state \fascd Algorithm \ref{alg:fascd} below we must clarify the problems solved at each mesh level.
 As in FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, and a source functional created.
@@
 Note that finest-level problem \eqref{eq:fe:vi} uses the original source functional $\ell^J$, and an admissible iterate $w^J \in \mathcal{K}^J$ has already\footnote{Conceptually-speaking this is true.
-In practice the LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ are generated during the descent stage; see Algorithm \ref{alg:nmcd}.
+In practice the LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ are generated during the descent stage; see Algorithm \ref{alg:fascd}.
 Observe that the LDCs do not depend on $w^j$ for $j<J$, nor on the corrections $y^j,z^j$.} been used to general the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:ldcs}).
@@
 In this sense $w^j$ is admissible with respect to the box constraints, but, following \cite{GraeserKornhuber2009}, our use of LDCs (Table \ref{tab:ldcs}) avoids the need to generate such box constraints $\underline{\gamma}^j$ and $\overline{\gamma}^j$ on each level.
-(Note that while $w^j \in \cV^j \subset \cV^J$ holds, generally $w^j \notin \cK^J$.)  In any case, Algorithm \ref{alg:nmcd} will yield a final, finest-level iterate $w^J+z^J \in \cK^J$ which is genuinely admissible for problem \eqref{eq:fe:vi}.
-
-Regarding formula \eqref{eq:fe:levelsource}, the source function $\ell^j$ is determined by a classical FAS-type formula.
-(Compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.)  It has the following explanation, mimicking the PDE story from textbooks, which simultaneously justifies \eqref{eq:fe:levelsource} and VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.
-First, the $j+1$ level problem \eqref{eq:fe:downvi} is not solved exactly by the finer-level correction $y^{j+1}$; the down-smoother was only approximate.
-Conceptually, we seek a further correction $y$ on the $j+1$ level so that
-\begin{equation}
-\ip{f^{j+1}(g^{j+1}+y^{j+1}+y)}{v-y^{j+1}-y} \ge \ip{\ell^{j+1}}{v-y^{j+1}-y}, \, \forall v\in \mathcal{D}^j \quad \text{\emph{(notional)}} \label{eq:fe:downvinotional}
-\end{equation}
-holds exactly.  However, $y$ will actually be approximated from a problem on the next-coarser level.  This $j$th-level problem comes from modifying notional VI \eqref{eq:fe:downvinotional} in 4 steps:
+(Note that while $w^j \in \cV^j \subset \cV^J$ holds, generally $w^j \notin \cK^J$.)  In any case, Algorithm \ref{alg:fascd} will yield a final, finest-level iterate $w^J+z^J \in \cK^J$ which is genuinely admissible for problem \eqref{eq:fe:vi}.
+
+Regarding formula \eqref{eq:fe:levelsource}, the formula for source functional $\ell^j$ is a classical FAS construction.
+(Compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.)  It has the following explanation, mimicking the PDE version, which simultaneously justifies \eqref{eq:fe:levelsource} and VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.
+First, the $j+1$ level problem \eqref{eq:fe:downvi} is not solved exactly by the finer-level correction $y^{j+1}$---the down-smoother was only approximate---so we seek a further correction $y$ on the $j+1$ level so that
+\begin{equation}
+\ip{f^{j+1}(w^{j+1}+y^{j+1}+y)}{v-y^{j+1}-y} \ge \ip{\ell^{j+1}}{v-y^{j+1}-y} \quad \text{\emph{(notional)}} \label{eq:fe:downvinotional}
+\end{equation}
+for all $v\in \mathcal{D}^j$.  However, $y$ will actually be approximated from solving a coarser, $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps:
 \begin{enumerate}
-\item subtract the computable quantity $f^{j+1}(g^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides,
-\item replace the residual $\ell^{j+1}-f^{j+1}(g^{j+1}+y^{j+1})$, now on the right, by its restriction,
-\item replace $g^{j+1}+y^{j+1}$, where it appears on the left, by its restriction,
-\item and, where it appears on the left, replace $f^{j+1}$ by its coarser rediscretization $f^j$.
+\item subtract the computable quantity $f^{j+1}(w^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides,
+\item replace the residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction,
+\item replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its restriction,
+\item and, where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.
 \end{enumerate}
-Steps \emph{(ii)}--\emph{(iv)} are justified, as for PDEs \cite[subsection 5.3.4, for example]{Trottenbergetal2001}, because the residual and the error should be smooth as a result of the progress made by the $j+1$ level down-smoother.
+Steps \emph{(ii)}--\emph{(iv)} are justified when the residual and the error are smooth as a result of the progress made by the $j+1$ level down-smoother \cite[subsection 5.3.4, for example]{Trottenbergetal2001}.
 These steps now yield a problem for the $j$th-level correction $y^j \in \mathcal{D}^j$:
 \begin{align}
-&\ip{f^j\left(\iR(g^{j+1}+y^{j+1})+y^j\right)}{v-y^j} - \ip{f^j\left(\iR(g^{j+1}+y^{j+1})\right)}{v-y^j} \label{eq:fe:downvicluttered} \\
-&\qquad \ge \ip{R\left(\ell^{j+1}-f^{j+1}(g^{j+1}+y^{j+1})\right)}{v-y^j} \notag
+&\ip{f^j\left(\iR(w^{j+1}+y^{j+1})+y^j\right)}{v-y^j} - \ip{f^j\left(\iR(w^{j+1}+y^{j+1})\right)}{v-y^j} \label{eq:fe:downvicluttered} \\
+&\qquad \ge \ip{R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right)}{v-y^j} \notag
 \end{align}
-This is the same as VI problem \eqref{eq:fe:downvi}, but in cluttered form; application of definitions \eqref{eq:fe:defineg} and \eqref{eq:fe:levelsource} yields \eqref{eq:fe:downvi}.
+This is the same as VI problem \eqref{eq:fe:downvi}, but in cluttered form; definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource} yields \eqref{eq:fe:downvi}.
 The explanation for VI \eqref{eq:fe:upvi} is essentially identical.
 
-We will regard the smoothers for VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} as approximate, in-place procedures, acting on the corrections $y^j$ and $z^j$, respectively.
+We will regard the smoothers for VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} as approximate, in-place procedures, acting on the corrections $y^j$ and $z^j$, respectively (Section \ref{sec:smoothers}).
 Also, as is standard in multigrid, at the coarsest $j=0$ level we suppose that problem \eqref{eq:fe:upvi} is solved quite accurately.
@@
 
-These ideas come together in Algorithm \ref{alg:nmcd}, below.  The \pr{nmcd-vcycle} procedure acts in-place on $w^J$, but it leaves its other inputs unchanged.
+These ideas come together in Algorithm \ref{alg:fascd}.
+\pr{fascd-vcycle} acts in-place on $w^J$, and it assumes certain semantics for the \pr{smooth} and the coarsest-level \pr{solve} procedures (Section \ref{sec:smoothers}).
+These do in-place modifications of their final arguments, without modifying their first four arguments.
+$\pr{smooth}^{\id{down}}$ and $\pr{smooth}^{\id{up}}$ denote \id{down} and \id{up} iterations, respectively.
 
 \begin{pseudofloat}[ht]
-\begin{pseudo} \label{ps:fascd-vcycle}
+\begin{pseudo}
 \pr{fascd-vcycle}(\ell^J,\underline{\gamma}^J,\overline{\gamma}^J;w^J)\text{:} \\+
@@
     $\overline{\chi}^J = \overline{\gamma}^J - w^J {\large \strut}$ \\
-    $g^J = w^J$ \\
     for $j=J$ downto $j=1$ \\+
@@
       $y^j = 0$ \\
-      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\underline{\phi}^j,\overline{\phi}^j,g^j;y^j)$  \ct{smoothing of \eqref{eq:fe:downvi} in $\mathcal{D}^j$}\\
-      $g^{j-1} = \iR(g^j + y^j)$ \\
-      $\ell^{j-1} = f^{j-1}(g^{j-1}) + R \left(\ell^j - f^j(g^j+y^j)\right)$ \\-
+      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\underline{\phi}^j,\overline{\phi}^j,w^j;y^j)$  \ct{smoothing of \eqref{eq:fe:downvi} in $\mathcal{D}^j$}\\
+      $w^{j-1} = \iR(w^j + y^j)$ \\
+      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j+y^j)\right)$ \\-
     $z^0 = 0$ \\
-    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,\overline{\chi}^0,g^0;z^0)$ \hspace{1.0cm} \ct{solving of \eqref{eq:fe:upvi} in $\mathcal{U}^0$} \\
+    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,\overline{\chi}^0,w^0;z^0)$ \hspace{1.0cm} \ct{solving of \eqref{eq:fe:upvi} in $\mathcal{U}^0$} \\
     for $j=1$ to $j=J$ \\+
       $z^j = y^{j} + P z^{j-1}$ \\
-      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,\overline{\chi}^j,g^j;z^j)$  \ct{smoothing of \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
+      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,\overline{\chi}^j,w^j;z^j)$  \ct{smoothing of \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
     $w^J \gets w^J+z^J$
 \end{pseudo}
-\caption{The \fascd V-cycle as an iterative solver for FE VI problem \eqref{eq:fe:vi}.  $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
-\label{alg:nmcd}
+\caption{The \fascd V-cycle as an iteration for solving FE VI problem \eqref{eq:fe:vi}.  $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
+\label{alg:fascd}
 \end{pseudofloat}
 
-Regarding its efficiency, to begin we consider storage.
-One application of \pr{nmcd-vcycle}, assuming all box constraints are nontrivial, storage must be allocated for the seven vectors $\underline{\chi}^j$, $\overline{\chi}^j$, $\underline{\phi}^j$, $\overline{\phi}^j$, $g^j$, $\ell^j$, and $y^j$ on each level.
-(Note $z^J$ can use the same storage as $y^j$.)  On the finest level one must also store $\underline{\gamma}^J$ and $\overline{\gamma}^J$, but note $g^J=w^J$ are the same.
-On the coarsest level there are only five vectors to store.
-Therefore, using estimates for 2D refinement plus a standard geometric series argument \cite{Trottenbergetal2001}, the total storage of a deep $V$-cycle, excluding the implementations of smoothers and the coarse-level solver, is about
-\begin{equation}
-9 m_J + 7 m_{J-1} + \dots + 7 m_1 + 5 m_0 \approx \left(9 + 7 \sum_{k=1}^\infty \frac{1}{4^k}\right) m_J \approx 11 m_J
-\end{equation}
-real numbers.
-For comparison, a single-level method needs at least $4 m_J$ storage for the vectors $\underline{\gamma}^J,\overline{\gamma}^J,\ell^J,w^J$.
+Regarding storage required for \pr{fascd-vcycle}, and assuming all box constraints are nontrivial, 7 vectors must be allocated on each level: $\underline{\chi}^j$, $\overline{\chi}^j$, $\underline{\phi}^j$, $\overline{\phi}^j$, $w^j$, $\ell^j$, $y^j$.
+Note $z^j$ can use the same storage as $y^j$.
+On the finest level one must also store $\underline{\gamma}^J$ and $\overline{\gamma}^J$, while on the coarsest level there are only 5 vectors because $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.
+Therefore, using standard estimates for 2D refinement plus a geometric series argument \cite{Trottenbergetal2001}, the total floating-point storage, excluding the implementations of smoothers and the coarse-level solver, is
+\begin{equation}
+9 m_J + 7 m_{J-1} + \dots + 7 m_1 + 5 m_0 \approx \left(9 + 7 \sum_{k=1}^\infty \frac{1}{4^k}\right) m_J \approx 11 m_J.
+\end{equation}
+For comparison, a single-level method needs $4 m_J$ storage for the vectors $\underline{\gamma}^J,\overline{\gamma}^J,\ell^J,w^J$.
+
 If some of the obstacles are trivial, for instance if $\overline{\gamma}^J=+\infty$ in a unilateral obstacle problem, storage is correspondingly reduced at all levels.
-
-\pr{nmcd-vcycle} assumes certain semantics for the smoothers \pr{smooth} and the coarsest-level solver \pr{solve}.
-These solvers do in-place modifications of their final arguments, without modifying their first four arguments.
-$\pr{smooth}^{\id{down}}$ and $\pr{smooth}^{\id{up}}$ are in-exact solvers of VIs \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}, respectively, iterated \id{down} and \id{up} times.
-Further implementation details are addressed in Section \ref{sec:smoothers}.
-
-Repeated application of \pr{nmcd-vcycle} to solve VI \eqref{eq:fe:vi} will not reduce the finest-level residual $f^J(w^J) - \ell^J$ to zero everywhere, and we need to identify the vector quantity whose norm must be controlled by a tolerance.
-Recalling the comments following mixed complementarity problem \eqref{eq:fe:mcp}, and denoting the hat function at node $x_p \in \mathcal{T}^J$ by $\psi_p$, for $w^J \in \mathcal{K}^J$ we define the following vector in $\RR^{m_J}$ as the finest-level \emph{nodal complementarity residual} which is defined as follows in the four cases (inactive, lower active, upper active, boundary):\footnote{Formally-speaking, $\hat r:\mathcal{V}^J \to \RR^{m_J}$ is simply a computable nonlinear map derived from $f^J$ and the vectors $\ell^J$, $\underline{\gamma}^J$, and $\overline{\gamma}^J$.
-It is an important part of the implementation.}
+Furthermore, we will see in Section \ref{sec:results} that, though \pr{fascd-vcycle} defines a general $V(\pr{down},\pr{up})$ cycle for VI problems, in fact the $V(0,\pr{up})$ case generally produces the best performance.
+In that case the storage is reduced to about $9 m_J$ by neglecting the downward LDCs $\underline{\phi}^j$, $\overline{\phi}^j$ at all levels.
+
+Repeated application of \pr{fascd-vcycle} to solve VI \eqref{eq:fe:vi} will not reduce the finest-level residual $f^J(w^J) - \ell^J$ to zero everywhere because the problem is inequality-constrained.
+We need to identify the vector quantity whose norm must be controlled by a tolerance.
+Recalling mixed complementarity problem \eqref{eq:fe:mcp}, and denoting the hat function at node $x_p \in \mathcal{T}^J$ by $\psi_p$, for $w^J \in \mathcal{K}^J$ we define the finest-level \emph{nodal complementarity residual} in $\RR^{m_J}$.
+It is defined in the four cases listed after \ref{eq:fe:mcp}, namely inactive, lower active, upper active, and boundary:
 \begin{equation}
@@
     \max\{\ip{f^J(w^J)-\ell^J}{\psi_p},0\}, & w^J(x_p) = \overline{\gamma}^J(x_p), \\
-    0, & x_p \in \partial_D\Omega. \end{cases} \label{eq:cpresidual}
-\end{equation}
-The point is that $\hat r(w^J)\in \RR^{m_J}$ is nonzero if there are nodes where MCP \eqref{eq:fe:mcp} is violated, either because the pointwise classical residual $\ip{f^J(w^J)-\ell^J}{\psi_p}$ remains nonzero where the constraints are inactive, or because it has the wrong sign on an actively-constrained degree of freedom.
-
-A solver for \eqref{eq:fe:vi} using repeated applications of \pr{nmcd-vcycle} would start with an initial iterate $(w^J)^{(0)}$, do V-cycles, and stop once the finest-level nodal residual $\hat r(w^J)$ was sufficiently small.
+    w^J(x_p) - g_D^J(x_p), & x_p \in \partial_D\Omega. \end{cases} \label{eq:cpresidual}
+\end{equation}
+Where the box constraints pinch at $x_p \notin \partial_D\Omega$, i.e.~where $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$, we define $\hat r(w^J)_p=0$.
+Note that $\hat r:\mathcal{V}^J \to \RR^{m_J}$ is a computable nonlinear map derived from $f^J$ and the vectors $\ell^J$, $\underline{\gamma}^J$, and $\overline{\gamma}^J$.
+For unconstrained PDEs one would use only the inactive and boundary cases.
+
+The idea here is that $\hat r(w^J)$ is nonzero only if there are nodes where MCP \eqref{eq:fe:mcp} is violated, either because the pointwise residual $\ip{f^J(w^J)-\ell^J}{\psi_p}$ remains nonzero where the constraints are inactive, or because it has the wrong sign on an actively-constrained degree of freedom, or because the Dirichlet boundary condition is not satisfied.
+
+A solver for \eqref{eq:fe:vi} using repeated applications of \pr{fascd-vcycle} would start with an initial iterate $w_{(0)}^J$, do V-cycles, and stop once the finest-level nodal residual $\hat r(w^J)$ was sufficiently small.
 For example, using tolerances \id{atol}$>0$ and \id{rtol}$>0$ the stopping criterion might be
@@
 
-Algorithm \ref{alg:nmcd} in the main text, \pr{nmcd-vcycle}, generalizes two better-known multilevel algorithms.
+Algorithm \ref{alg:fascd} in the main text, \pr{nmcd-vcycle}, generalizes two better-known multilevel algorithms.
 First we show that it reduces to the FAS multigrid V-cycle for nonlinear PDEs (which itself generalizes the standard linear multigrid V-cycle \cite{Trottenbergetal2001}).
 Second we reduce NMCD to an existing MCD V-cycle for unilateral linear variational inequalities \cite{GraeserKornhuber2009}.
-This Appendix helps readers understand Algorithm \ref{alg:nmcd} by separating those assumptions which relate to nonlinear and FAS aspects from the nontrivial modifications of standard multigrid which are caused by inequality constraints.
+This Appendix helps readers understand Algorithm \ref{alg:fascd} by separating those assumptions which relate to nonlinear and FAS aspects from the nontrivial modifications of standard multigrid which are caused by inequality constraints.
 The later arise even when $f$ in \eqref{eq:vi} is a linear operator.
 
-To derive the classical FAS multigrid V-cycle from Algorithm \ref{alg:nmcd}, we remove the inequality constraints from VI problem \eqref{eq:fe:vi} and then simplify the pseudocode accordingly.
+To derive the classical FAS multigrid V-cycle from Algorithm \ref{alg:fascd}, we remove the inequality constraints from VI problem \eqref{eq:fe:vi} and then simplify the pseudocode accordingly.
 To do this we suppose that the finest-level obstacles $\underline{\gamma}^J$ and $\overline{\gamma}^J$ are very negative and positive, respectively, with magnitude exceeding $u^J$ or any prospective iterate $w^J$.
@@
 
-Furthermore, the downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets can be arbitrarily enlarged so that the coarser-level corrections $y^j$ and $z^j$ in Algorithm \ref{alg:nmcd} are also never constrained.
+Furthermore, the downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets can be arbitrarily enlarged so that the coarser-level corrections $y^j$ and $z^j$ in Algorithm \ref{alg:fascd} are also never constrained.
 Formally-speaking we might do this, still satisfying the LDC ordering property \eqref{eq:fe:chiordering}, by choosing LDCs with all gaps being large so that $\underline{\chi}^j \ll 0 \ll \overline{\chi}^j$ and also $\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1} \ll 0 \ll \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}$.\footnote{As noted after \eqref{eq:fe:chiordering}, the application of monotone restriction operators $\maxR,\minR$ is not required to generate LDCs $\underline{\chi}^j$ and $\overline{\chi}^j$ satisfying the ordering property.}
 
-Now the correction VIs \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} are also unconstrained, and lines 2--3 and 6--9 can be removed from Algorithm \ref{alg:nmcd}.
+Now the correction VIs \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} are also unconstrained, and lines 2--3 and 6--9 can be removed from Algorithm \ref{alg:fascd}.
 Also, because the downward corrections $y^j$ are no longer separately constrained, we may simply refer to the corrected solution $w^j=g^j+y^j$ instead of $g^j$ and $y^j$ separately.
@@
 \end{equation}
-(Compare line 13 of Algorithm \ref{alg:nmcd}.)
+(Compare line 13 of Algorithm \ref{alg:fascd}.)
 
@@
 Now we must be careful that going upward only the coarse correction, and not the coarse solution, gets prolonged.\footnote{Compare Remark 5.3.9 in \cite{Trottenbergetal2001}, a warning to novices.}  The coarser-level iterate is $w^{j-1} = g^{j-1}+z^{j-1}$ after up-smoothing on the $j-1$ level.
-From Algorithm \ref{alg:nmcd} the initial value of the correction $z^j$, before smoothing, is $z^j = y^j + P z^{j-1}$.
+From Algorithm \ref{alg:fascd} the initial value of the correction $z^j$, before smoothing, is $z^j = y^j + P z^{j-1}$.
 These facts imply that the $j$th-level iterate before up-smoothing must be set to

### 2022-12-28 paper/fascd.tex
a few replacements

@@
 
-FIXME show results for 2D Firedrake implementation of NMCD for $p$-Laplacian obstacle problem (Example \ref{ex:plaplacian}), advection-diffusion obstacle problem (Example \ref{ex:advectiondiffusion}), and porous-media obstacle problem (Example \ref{ex:porous})
+FIXME show results for 2D Firedrake implementation of FASCD for $p$-Laplacian obstacle problem (Example \ref{ex:plaplacian}), advection-diffusion obstacle problem (Example \ref{ex:advectiondiffusion}), and porous-media obstacle problem (Example \ref{ex:porous})
 
@@
 FIXME constrast the implemented examples in \cite{GraeserKornhuber2009,Tai2003} in which $f$ is linear
-
-FIXME in Appendix we show how NMCD reduces to FAS if the obstacle is removed and to MCD if $f$ is assumed to be linear
 
@@
 \appendix
-\section{Reductions of NMCD to familiar multilevel algorithms} \label{app:reductions}
-
-Algorithm \ref{alg:fascd} in the main text, \pr{nmcd-vcycle}, generalizes two better-known multilevel algorithms.
+\section{Reductions of FASCD to familiar multilevel algorithms} \label{app:reductions}
+
+Algorithm \ref{alg:fascd} in the main text, \pr{fascd-vcycle}, generalizes two better-known multilevel algorithms.
 First we show that it reduces to the FAS multigrid V-cycle for nonlinear PDEs (which itself generalizes the standard linear multigrid V-cycle \cite{Trottenbergetal2001}).
-Second we reduce NMCD to an existing MCD V-cycle for unilateral linear variational inequalities \cite{GraeserKornhuber2009}.
+Second we reduce FASCD to an existing MCD V-cycle for unilateral linear variational inequalities \cite{GraeserKornhuber2009}.
 This Appendix helps readers understand Algorithm \ref{alg:fascd} by separating those assumptions which relate to nonlinear and FAS aspects from the nontrivial modifications of standard multigrid which are caused by inequality constraints.
@@
 \end{pseudo}
-\caption{An FAS V-cycle results from removing the inequality constraints from \pr{nmcd-vcycle}, and simplifying accordingly.}
+\caption{An FAS V-cycle results from removing the inequality constraints from \pr{fascd-vcycle}, and simplifying accordingly.}
 \label{alg:fas}

### 2022-12-30 paper/fascd.tex
I think I have explained FASCD->FAS reduction

@@
 
-Algorithm \ref{alg:fascd} in the main text, \pr{fascd-vcycle}, generalizes two better-known multilevel algorithms.
-First we show that it reduces to the FAS multigrid V-cycle for nonlinear PDEs (which itself generalizes the standard linear multigrid V-cycle \cite{Trottenbergetal2001}).
-Second we reduce FASCD to an existing MCD V-cycle for unilateral linear variational inequalities \cite{GraeserKornhuber2009}.
-This Appendix helps readers understand Algorithm \ref{alg:fascd} by separating those assumptions which relate to nonlinear and FAS aspects from the nontrivial modifications of standard multigrid which are caused by inequality constraints.
-The later arise even when $f$ in \eqref{eq:vi} is a linear operator.
-
-To derive the classical FAS multigrid V-cycle from Algorithm \ref{alg:fascd}, we remove the inequality constraints from VI problem \eqref{eq:fe:vi} and then simplify the pseudocode accordingly.
-To do this we suppose that the finest-level obstacles $\underline{\gamma}^J$ and $\overline{\gamma}^J$ are very negative and positive, respectively, with magnitude exceeding $u^J$ or any prospective iterate $w^J$.
-Finest-level problem \eqref{eq:fe:vi} is then unconstrained except for its Dirichlet boundary conditions.
-Let $\mathcal{V}_0^J \subset \mathcal{V}^J$ denote the space with zero values on $\partial_D\Omega$.
-By replacing ``$v-u^J$'' with an arbitrary $v \in \mathcal{V}_0^J$ we have the finest-level, weak form problem
-\begin{equation}
-\ip{f^J(u^J)}{v} = \ip{\ell^J}{v} \qquad \text{for all } v\in \mathcal{V}_0^J \label{eq:app:fas:pde}
-\end{equation}
-corresponding to the strong-form PDE $f(u)=\ell$.
-
-Furthermore, the downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets can be arbitrarily enlarged so that the coarser-level corrections $y^j$ and $z^j$ in Algorithm \ref{alg:fascd} are also never constrained.
-Formally-speaking we might do this, still satisfying the LDC ordering property \eqref{eq:fe:chiordering}, by choosing LDCs with all gaps being large so that $\underline{\chi}^j \ll 0 \ll \overline{\chi}^j$ and also $\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1} \ll 0 \ll \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}$.\footnote{As noted after \eqref{eq:fe:chiordering}, the application of monotone restriction operators $\maxR,\minR$ is not required to generate LDCs $\underline{\chi}^j$ and $\overline{\chi}^j$ satisfying the ordering property.}
-
-Now the correction VIs \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} are also unconstrained, and lines 2--3 and 6--9 can be removed from Algorithm \ref{alg:fascd}.
-Also, because the downward corrections $y^j$ are no longer separately constrained, we may simply refer to the corrected solution $w^j=g^j+y^j$ instead of $g^j$ and $y^j$ separately.
-Then the smoothers and the coarsest-level solver can be regarded as in-place modifications of $w^j$.
-Now noting $g^{j-1}=\iR w^j$ in line 12, after down-smoothing, the coarsened source functional can be rewritten using notation $w^{j-1}=\iR w^j$:
-\begin{equation}
-\ell^{j-1} = f^{j-1}\left(w^{j-1}\right) + R\left(\ell^j-f^j(w^j)\right). \label{eq:app:fas:levelsource}
-\end{equation}
-(Compare line 13 of Algorithm \ref{alg:fascd}.)
-
-At the completion of down-smoothing the solution iterate on the $j$th level is $w^j=g^j+y^j$.
-Now we must be careful that going upward only the coarse correction, and not the coarse solution, gets prolonged.\footnote{Compare Remark 5.3.9 in \cite{Trottenbergetal2001}, a warning to novices.}  The coarser-level iterate is $w^{j-1} = g^{j-1}+z^{j-1}$ after up-smoothing on the $j-1$ level.
-From Algorithm \ref{alg:fascd} the initial value of the correction $z^j$, before smoothing, is $z^j = y^j + P z^{j-1}$.
-These facts imply that the $j$th-level iterate before up-smoothing must be set to
-    $$w^j \gets w^j + Pz^j = (g^j+y^j)+P(w^{j-1} - g^{j-1}) = w^j + P(w^{j-1} - \iR w^j).$$
-This is a standard FAS statement.
-In fact, with the given modifications, we have the much-simplified pseudocode in Algorithm \ref{alg:fas}.
+\pr{fascd-vcycle} (Algorithm \ref{alg:fascd}) in the main text generalizes two better-known multilevel algorithms.
+First we show that it reduces to the FAS multigrid V-cycle for nonlinear PDEs \cite{Trottenbergetal2001}, which itself generalizes linear multigrid, when inequality constraints are removed.
+Second we reduce it to an existing V-cycle for unilateral linear variational inequalities \cite{GraeserKornhuber2009} in the case that the operator is linear and local.
+Thus this Appendix separates FAS/nonlinear aspects of the algorithm from those modifications of linear multigrid which are caused by the inequality constraints.
+
+To derive the classical FAS multigrid V-cycle we remove the inequality constraints from the finest-level VI problem \eqref{eq:fe:vi} and then simplify the \pr{fascd-vcycle} pseudocode accordingly.
+In fact, suppose that the finest-level obstacles $\underline{\gamma}^J$ and $\overline{\gamma}^J$ are very negative and positive, respectively, thus not constraining the solution $u^J$ or any possible iterate.
+The admissible set $\mathcal{K}^J$ of \eqref{eq:fe:vi} is then determined only by the Dirichlet boundary conditions.
+Let $\mathcal{V}_0^J \subset \mathcal{V}^J$ denote the linear subspace with zero values on $\partial_D\Omega$.
+By replacing the (now) arbitrary element $v-u^J\in\mathcal{V}_0^J$ in \eqref{eq:fe:vi} with ``$v$'' we have the finest-level problem
+\begin{equation}
+\ip{f^J(u^J)}{v} = \ip{\ell^J}{v} \qquad \text{for all } v\in \mathcal{V}_0^J, \label{eq:app:fas:pde}
+\end{equation}
+the FE discretization of strong-form PDE $f(u)=\ell$.
+The downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets can also be enlarged so that the corrections $y^j$ and $z^j$ in \pr{fascd-vcycle} are never constrained.
+Formally we might do this by choosing constant LDCs with $\underline{\chi}^j \ll 0 \ll \overline{\chi}^j$, but so that ordering property \eqref{eq:fe:chiordering} still holds, and also with large gaps so that $\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1} \ll 0 \ll \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}$.
+Thus the corrections $y^j$ and $z^j$ are unconstrained, and lines 2--3 and 5--8 can be removed from Algorithm \ref{alg:fascd}.
+
+Now the downward/upward VIs \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} are also unconstrained, and can be rewritten as weak forms $\ip{f^j(w^j + y^j)}{v} = \ip{\ell^j}{v}$ and $\ip{f^j(w^j + z^j)}{v} = \ip{\ell^j}{v}$, respectively, for all $v\in \mathcal{V}_0^j$.
+However, we may simply refer to the unconstrained, approximate solution $u^j=w^j+y^j$, instead of $w^j$ and $y^j$ separately.
+The smoothers and the coarsest-level solver can be regarded as unconstrained, in-place modifications of $u^j$, so we denote them simply as ``$\text{\pr{smooth}}(\ell^j;u^j)$.''  Noting line 11 of Algorithm \ref{alg:fascd} says $u^{j-1}=\iR u^j$, after down-smoothing the coarsened source functional computed by line 12 can be rewritten as
+\begin{equation}
+\ell^{j-1} = f^{j-1}\left(u^{j-1}\right) + R\left(\ell^j-f^j(u^j)\right). \label{eq:app:fas:levelsource}
+\end{equation}
+
+Going upward we must be careful that only the coarse correction, and not the coarse solution, is prolonged.\footnote{Compare Remark 5.3.9 in \cite{Trottenbergetal2001}, a warning to novices.}  After up-smoothing on the $j-1$ level the coarse iterate is $u^{j-1}$, but at this point $u^j=w^j + y^j$ contains the result of down-smoothing on the $j$th level.
+On the other hand, in Algorithm \ref{alg:fascd} the initial value of the correction $z^j$, before up-smoothing on the $j$th level, is $z^j = y^j + P z^{j-1}$, where $z^{j-1}$ contains the total correction from coarser corrections and up-smoothing at the $j-1$ level.
+Thus the $j$th-level iterate \emph{before} up-smoothing must be set to
+    $$w^j + y^j + P z^{j-1} = u^j + P(u^{j-1} - \iR u^j),$$
+a standard FAS assignment.
+
+With the given modifications, we have the much-simplified pseudocode in Algorithm \ref{alg:fas}.
 This FAS V-cycle algorithm appears in various places, for example Algorithm 14 of \cite{Bruneetal2015} or in section 5.3.4 of \cite{Trottenbergetal2001}.
@@
 \begin{pseudo} \label{ps:fas-vcycle}
-\pr{fas-vcycle}(\ell^J; w^J)\text{:} \\+
+\pr{fas-vcycle}(\ell^J; u^J)\text{:} \\+
     for $j=J$ downto $j=1$ \\+
-      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j;w^j)$ \\
-      $w^{j-1} = \iR w^j$ \\
-      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j)\right)$ \\-
-    $\text{\pr{solve}}(\ell^0;w^0)$ \\
+      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j;u^j)$ \\
+      $u^{j-1} = \iR u^j$ \\
+      $\ell^{j-1} = f^{j-1}(u^{j-1}) + R \left(\ell^j - f^j(u^j)\right)$ \\-
+    $\text{\pr{solve}}(\ell^0;u^0)$ \\
     for $j=1$ to $j=J$ \\+
-      $w^j \gets w^j + P (w^{j-1} - \iR w^j)$ \\
-      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j;w^j)$ \\-
+      $u^j \gets u^j + P (u^{j-1} - \iR u^j)$ \\
+      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j;u^j)$ \\-
 \end{pseudo}
-\caption{An FAS V-cycle results from removing the inequality constraints from \pr{fascd-vcycle}, and simplifying accordingly.}
+\caption{Removing inequality constraints from \pr{fascd-vcycle} yields the standard FAS V-cycle.}
 \label{alg:fas}

### 2023-01-24 talk/slides.tex
a skeleton for my 2nd talk

\documentclass[svgnames,
               hyperref={colorlinks,citecolor=DeepPink4,linkcolor=FireBrick,urlcolor=Maroon},
               usepdftitle=false]  % see \hypersetup{} below
               {beamer}

\mode<presentation>{
  \usetheme{Madrid}
  %\usecolortheme{seagull}
  \usecolortheme{seagull}
  \setbeamercovered{transparent}
  \setbeamerfont{frametitle}{size=\large}
}

\setbeamercolor*{block title}{bg=red!10}
\setbeamercolor*{block body}{bg=red!5}

%\usepackage[svgnames]{xcolor}
\usepackage{hyperref}
\hypersetup{
    pdftitle = {Toward nonlinear multigrid for nonlinear variational inequalities},
    pdfauthor = {Ed Bueler},
    pdfsubject = {},
    pdfkeywords = {}
}

\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{empheq,bm,xspace,fancyvrb,soul}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows.meta,decorations.markings,decorations.pathreplacing,fadings,positioning}
\usepackage[kw]{pseudo}
\pseudoset{left-margin=15mm,topsep=5mm,idfont=\texttt,st-left=,st-right=}

\newcommand{\eps}{\epsilon}
\newcommand{\RR}{\mathbb{R}}

\newcommand{\grad}{\nabla}
\newcommand{\Div}{\nabla\cdot}
\newcommand{\trace}{\operatorname{tr}}

\newcommand{\hbn}{\hat{\mathbf{n}}}

\newcommand{\bb}{\mathbf{b}}
\newcommand{\be}{\mathbf{e}}
\newcommand{\bbf}{\mathbf{f}}
\newcommand{\bg}{\mathbf{g}}
\newcommand{\bn}{\mathbf{n}}
\newcommand{\br}{\mathbf{r}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\bv}{\mathbf{v}}
\newcommand{\bw}{\mathbf{w}}
\newcommand{\bx}{\mathbf{x}}

\newcommand{\bF}{\mathbf{F}}
\newcommand{\bQ}{\mathbf{Q}}
\newcommand{\bU}{\mathbf{U}}
\newcommand{\bV}{\mathbf{V}}
\newcommand{\bX}{\mathbf{X}}

\newcommand{\btau}{\bm{\tau}}
\newcommand{\bxi}{\bm{\xi}}

\newcommand{\bzero}{\bm{0}}

\newcommand{\rhoi}{\rho_{\text{i}}}

\newcommand{\ip}[2]{\left(#1,#2\right)}

\newcommand{\mR}{R^{\bm{\oplus}}}
\newcommand{\iR}{R^{\bullet}}

\newcommand{\nn}{{\text{n}}}
\newcommand{\pp}{{\text{p}}}
\newcommand{\qq}{{\text{q}}}
\newcommand{\rr}{{\text{r}}}

\newcommand{\bus}{\bu|_s}
\newcommand{\oo}[1]{\displaystyle O\left(#1\right)}
\newcommand{\sold}{s_{\text{o}}}

\title[Multigrid for nonlinear VI]{Toward nonlinear multigrid \\ for nonlinear variational inequalities}

%\subtitle{\emph{x}}

\author{Ed Bueler}

\institute[UAF]{University of Alaska Fairbanks}

\date[]{February 2023}

%\titlegraphic{\begin{picture}(0,0)
%    \put(0,180){\makebox(0,0)[rt]{\includegraphics[width=4cm]{figs/software.png}}}
%  \end{picture}
%}

\titlegraphic{\hfill \includegraphics[width=0.15\textwidth]{images/uafbw.png}}

%% to start section counter at 0 see
%% https://tex.stackexchange.com/questions/170222/change-the-numbering-in-beamers-table-of-content

\begin{document}
\beamertemplatenavigationsymbolsempty

%\begin{frame}
%  \maketitle
%\end{frame}

{
  %\usebackgroundtemplate{\includegraphics[width=\paperwidth]{images/gray-british-clark2022.png}}
  \begin{frame}
    \titlepage
  \end{frame}
}

\begin{frame}{Outline}
  \tableofcontents[hideallsubsections]
\end{frame}

\section{obstacle problems, variational inequalities (VIs)}

\section{nonlocal VIs for fluid layers}

\section{what is FAS multigrid?  (for PDEs)}

\section{scalable approaches to VIs}

\begin{frame}{the Stokes problem for ice}

\begin{itemize}
\item whether explicit or implicit, a non-shallow model solves a Stokes problem at each step:
\begin{align*}
- \nabla \cdot \left(2 \nu_\eps(D\bu)\, D\bu\right) + \nabla p - \rhoi \mathbf{g} &= \bzero && \text{in $\Lambda$} \\
\nabla \cdot \bu &= 0 && \text{''} \\
\btau_b - \bbf(\bu|_b) &= \bzero && \text{on $\Gamma_b$} \\
\bu|_b \cdot \bn_b &= 0 && \text{''} \\
\left(2 \nu_\eps(D\bu) D\bu - pI\right) \bn &= \bzero && \text{on $\Gamma_s$}
\end{align*}
\item this is the \alert{stress balance} (conservation of momentum) problem which determines velocity $\bu$ and pressure $p$
\item how fast is the numerical solution process?
    \begin{itemize}
    \item[$\circ$] how do solution algorithms \alert{scale with increasing resolution}?
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{multilevel NCP-coupled-to-Stokes solvers}

\begin{itemize}
\item direct attack on the problem seems to require a \alert{multilevel} solver for \alert{variational inequalities} (VIs), but in the \alert{non-local residual case}
    \begin{itemize}
    \item[$\circ$] this seems not to exist
    \item[$\circ$] the \alert{smoother} must compute an implicit-step SKE residual including surface-motion term $\Phi(s) = - \bu|_s\cdot \bn_s$ from a scalable Stokes solver
    \end{itemize}
\item near-optimal multilevel VI solvers exist for toy problems
    \begin{itemize}
    \item[$\circ$] Poisson equation obstacle problem (Brandt \& Cryer 1983; Gr\"aeser \& Kornhuber 2009)
    \end{itemize}
\item<2> \alert{help me}
\end{itemize}
\end{frame}

\section{conclusion}

\begin{frame}{\alert{summary}}

\begin{itemize}
\item x
\end{itemize}
\end{frame}

\begin{frame}{references}

{\scriptsize
\begin{itemize}
\item T.~Isaac, G.~Stadler, \& O.~Ghattas (2015). \emph{Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics}, SIAM J.~Sci.~Comput.~37 (6), B804--B833, \href{https://doi.org/10.1137/140974407}{10.1137/140974407}
\end{itemize}
}
\end{frame}

\end{document}

### 2023-02-04 talk/slides.tex
will need this ref
bunch of figures from book, and fluid layes problem

@@
 
-\section{obstacle problems, variational inequalities (VIs)}
+\section{variational inequalities (VIs) and obstacle problems}
 
-\section{nonlocal VIs for fluid layers}
+\begin{frame}{the classical obstacle problem}
 
-\section{what is FAS multigrid?  (for PDEs)}
+\begin{center}
+\includegraphics[width=0.6\textwidth]{images/obstacle65.pdf}
+\end{center}
 
-\section{scalable approaches to VIs}
+\begin{itemize}
+\item x
+\end{itemize}
+\end{frame}
+
+\begin{frame}{active sets, free boundaries}
+
+\begin{columns}
+\begin{column}{0.6\textwidth}
+\begin{itemize}
+\item x
+\end{itemize}
+\end{column}
+\begin{column}{0.4\textwidth}
+\includegraphics[width=0.6\textwidth]{images/obstacle-sets.png}
+\end{column}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Newton iterations for VIs}
+
+\begin{columns}
+\begin{column}{0.6\textwidth}
+\begin{itemize}
+\item x
+\end{itemize}
+\end{column}
+\begin{column}{0.4\textwidth}
+\includegraphics[width=\textwidth]{images/vi-newton-gmg-bad.png}
+
+\includegraphics[width=\textwidth]{images/vi-newton-gmg-good.png}
+\end{column}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Newton-multigrid for VIs}
+
+xxx
+\end{frame}
+
+\begin{frame}{optimality}
+
+\begin{columns}
+\begin{column}{0.6\textwidth}
+\begin{itemize}
+\item x
+\end{itemize}
+\end{column}
+\begin{column}{0.4\textwidth}
+\includegraphics[width=\textwidth]{images/obstacle-flops-per-n.png}
+\end{column}
+\end{columns}
+\end{frame}
+
+\section{the nonlocal VI for a fluid layer in a climate}
+
+\begin{frame}{problem: fluid layer in a climate}
+
+\begin{center}
+\includegraphics[width=0.6\textwidth]{images/fluid-in-climate.png}
+\end{center}
+
+\begin{itemize}
+\item x
+\end{itemize}
+\end{frame}
+
+\section{FAS multigrid for PDEs}
+
+\section{a possibly scalable approach to nonlocal VIs}
 
@@
 
-{\scriptsize
+{\tiny
 \begin{itemize}
-\item T.~Isaac, G.~Stadler, \& O.~Ghattas (2015).
+\item[] E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032, \href{https://doi.org/10.1137/20M135217X}{10.1137/20M135217X}
+\item[] E.~Bueler (2021). \emph{PETSc for Partial Differential Equations}, SIAM Press, Philadelphia, PA.
+\item[] E.~Bueler (2022). \emph{Performance analysis of high-resolution ice-sheet simulations}, J.~Glaciol., \href{https://doi.org/10.1017/jog.2022.113}{10.1017/jog.2022.113}
+\item[] T.~Isaac, G.~Stadler, \& O.~Ghattas (2015).
 \emph{Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics}, SIAM J.~Sci.~Comput.~37 (6), B804--B833, \href{https://doi.org/10.1137/140974407}{10.1137/140974407}

### 2023-02-05 c/bratuscaling/figure.py
experiment: FAS performance from bratu.c

@@ -1,0 +1,50 @@
+#!/usr/bin/env python
+
+# see run.sh (and output rawFD.txt), to generate input data data.txt
+
+import numpy as np
+import matplotlib.pyplot as plt
+import sys
+
+INFILE = 'data.txt'
+#PROBMARKER = ['*','o']
+#PARMARKERSIZE = [10.0,14.0,18.0,8.0,12.0,16.0]
+
+SHOW = False
+def writeout(outname):
+    if SHOW:
+        plt.show()
+    else:
+        print('writing file ' + outname)
+        plt.savefig(outname,bbox_inches='tight')
+
+print('reading %s ...' % INFILE)
+v = np.array(np.loadtxt(INFILE))
+
+assert len(v[:,0]) == 5
+
+N = v[:,0]**2
+flops = v[:,1]
+exps = v[:,2]
+time = v[:,3]
+
+# fit to show  (flops) = O(N^1):
+#qflops = np.polyfit(np.log(N),np.log(flops),1)
+
+# exps/N versus N
+plt.figure(figsize=(7,6))
+plt.semilogx(N,exps/N,'ko:',ms=14.0)
+plt.xlabel('N = degrees of freedom',fontsize=18.0)
+plt.ylabel('exps / N',fontsize=18.0)
+plt.grid(True)
+plt.axis([1e5, 1e8, 0, 50])
+writeout('bratuexps.png')
+
+# time/N versus N
+plt.figure(figsize=(7,6))
+plt.semilogx(N,1.0e6 * time/N,'ko:',ms=14.0)
+plt.xlabel('N = degrees of freedom',fontsize=18.0)
+plt.ylabel('$\mu s$ / N',fontsize=18.0)
+plt.grid(True)
+plt.axis([1e5, 1e8, 0, 2.5])
+writeout('bratutime.png')

### 2023-02-05 c/bratuscaling/run.sh

@@ -1,0 +1,30 @@
+#!/bin/bash
+
+# run a sequence of FAS Bratu solves to get scaling data, e.g.:
+#   ./run.sh >& rawFD.txt
+
+# remember to do in parent directory:
+#export PETSC_ARCH=linux-c-opt;  make bratu
+
+# compare Firedrake Poisson with P1 elements on 2049^2 grid using p4pdes/python/ch13/fish.py:
+# $ time mpiexec -n 4 --map-by core --bind-to hwthread ./fish.py -refine 10 -k 1 -s_ksp_converged_reason -s_ksp_rtol 1.0e-10 -s_pc_type mg
+#  Linear s_ solve converged due to CONVERGED_RTOL iterations 10
+#done on 2049 x 2049 grid with P_1 elements:
+#  error |u-uexact|_inf = 3.785e-09, |u-uexact|_h = 2.088e-09
+#real 84.97
+# = 20.2 microseconds per degree of freedom
+
+NP=4
+DEPTHLIST="7 8 9 10 11"
+DISC="-lb_fd"   # versus "-lb_fem"
+
+echo "NP=$NP"
+echo "DEPTHLIST=\"$DEPTHLIST\""
+echo "DISC=$DISC"
+echo
+
+for DEPTH in $DEPTHLIST; do
+    echo "DEPTH=$DEPTH"
+    time mpiexec -n $NP --map-by core --bind-to hwthread ../bratu $DISC -da_grid_x 5 -da_grid_y 5 -lb_exact -snes_rtol 1.0e-12 -snes_converged_reason -lb_counts -snes_type fas -snes_fas_type full -fas_levels_snes_type ngs -fas_levels_snes_ngs_sweeps 2 -fas_levels_snes_ngs_max_it 1 -fas_levels_snes_norm_schedule none -fas_coarse_snes_type ngs -fas_coarse_snes_max_it 1 -fas_coarse_snes_ngs_sweeps 4 -fas_coarse_snes_ngs_max_it 1 -da_refine $DEPTH
+    echo
+done

### 2023-02-05 talk/references.tex
clean up (and factor out) reference slide

% inputed at end of slides.tex

\newcommand{\shref}[2]{\,{\tiny \href{#1}{#2}}}
\begin{itemize}
\item[] E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032 \shref{https://doi.org/10.1137/20M135217X}{doi:10.1137/20M135217X}
\item[] E.~Bueler (2021). \emph{PETSc for Partial Differential Equations}, SIAM Press, Philadelphia
%\item[] E.~Bueler (2022). \emph{Performance analysis of high-resolution ice-sheet simulations}, J.~Glaciol., \href{https://doi.org/10.1017/jog.2022.113}{doi:10.1017/jog.2022.113}
\item[] P.~Farrell, M.~Croci, \& T.~Surowiec (2020). \emph{Deflation for semismooth equations}, Optimization Methods \& Software 35 (6), 1248--1271 \shref{https://doi.org/10.1080/10556788.2019.1613655}{doi:10.1080/10556788.2019.1613655}
\item[] C.~Gr{\"a}ser \& R.~Kornhuber (2009). \emph{Multigrid methods for obstacle problems}, J.~Comput.~Math., 1--44
\item[] G. Jouvet \& E. Bueler (2012). \emph{Steady, shallow ice sheets as obstacle problems: well-posedness and finite element approximation}, SIAM J.~Appl.~Math.~72 (4), 1292--1314 \shref{https://doi.org/10.1137/110856654}{doi:10.1137/110856654}
\item[] D. Kinderlehrer \& G. Stampacchia (1980). \emph{An Introduction to Variational Inequalities and their Applications}, Academic Press, New York
%\item[] T.~Isaac, G.~Stadler, \& O.~Ghattas (2015). \emph{Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics}, SIAM J.~Sci.~Comput.~37 (6), B804--B833 \shref{https://doi.org/10.1137/140974407}{doi:10.1137/140974407}
\item[] U.~Trottenberg, C.~Oosterlee, \& A. Schuller (2001).  \emph{Multigrid}, Elsevier, Oxford
\end{itemize}

### 2023-02-05 talk/slides.tex
clean up (and factor out) reference slide

@@
 \pseudoset{left-margin=15mm,topsep=5mm,idfont=\texttt,st-left=,st-right=}
+
+\makeatletter
+%\newcommand\notsotiny{\@setfontsize\notsotiny\@vipt\@viipt}
+\newcommand\notsotiny{\@setfontsize\notsotiny\@viipt\@viiipt}
+\makeatother
 
@@
 
-{\tiny
-\begin{itemize}
-\item[] E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032, \href{https://doi.org/10.1137/20M135217X}{10.1137/20M135217X}
-\item[] E.~Bueler (2021). \emph{PETSc for Partial Differential Equations}, SIAM Press, Philadelphia, PA.
-\item[] E.~Bueler (2022). \emph{Performance analysis of high-resolution ice-sheet simulations}, J.~Glaciol., \href{https://doi.org/10.1017/jog.2022.113}{10.1017/jog.2022.113}
-\item[] T.~Isaac, G.~Stadler, \& O.~Ghattas (2015).
-\emph{Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics}, SIAM J.~Sci.~Comput.~37 (6), B804--B833, \href{https://doi.org/10.1137/140974407}{10.1137/140974407}
-\end{itemize}
+{\scriptsize
+%{\notsotiny
+\input{references.tex}
 }

### 2023-02-05 talk/references.tex
add references I will need

@@
 \begin{itemize}
+\item[] A.~Brandt (1977). \emph{Multi-level adaptive solutions to boundary-value problems}, Mathematics of Computation 31 (138), 333--390
+\item[] A.~Brandt \& C.~Cryer (1983).
+\emph{Multigrid algorithms for the solution of linear complementarity problems \dots}, SIAM J.~Sci.~Stat.~Comput.~4 (4), 655--684 \shref{https://doi.org/10.1137/0904046}{doi:10.1137/0904046}
+%FTITLE Multigrid algorithms for the solution of linear complementarity problems arising from free boundary problems
 \item[] E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032 \shref{https://doi.org/10.1137/20M135217X}{doi:10.1137/20M135217X}
@@
 Bueler (2012).
-\emph{Steady, shallow ice sheets as obstacle problems: well-posedness and finite element approximation}, SIAM J.~Appl.~Math.~72 (4), 1292--1314 \shref{https://doi.org/10.1137/110856654}{doi:10.1137/110856654}
+\emph{Steady, shallow ice sheets as obstacle problems \dots}, SIAM J.~Appl.~Math.~72 (4), 1292--1314 \shref{https://doi.org/10.1137/110856654}{doi:10.1137/110856654}
+%FTITLE Steady, shallow ice sheets as obstacle problems: well-posedness and finite element approximation
 \item[] D. Kinderlehrer \& G. Stampacchia (1980). \emph{An Introduction to Variational Inequalities and their Applications}, Academic Press, New York
 %\item[] T.~Isaac, G.~Stadler, \& O.~Ghattas (2015).
-\emph{Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics}, SIAM J.~Sci.~Comput.~37 (6), B804--B833 \shref{https://doi.org/10.1137/140974407}{doi:10.1137/140974407}
+\emph{Solution of nonlinear Stokes equations \dots, with application to ice sheet dynamics}, SIAM J.~Sci.~Comput.~37 (6), B804--B833 \shref{https://doi.org/10.1137/140974407}{doi:10.1137/140974407}
+%FTITLE Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics
 \item[] U.~Trottenberg, C.~Oosterlee, \& A. Schuller (2001).  \emph{Multigrid}, Elsevier, Oxford

### 2023-02-05 talk/slides.tex
talk progress finally

@@
 
-\section{variational inequalities (VIs) and obstacle problems}
+\section{variational inequalities (VIs)?}
 
-\begin{frame}{the classical obstacle problem}
+\begin{frame}{example: a classical obstacle problem}
 
-\begin{center}
-\includegraphics[width=0.6\textwidth]{images/obstacle65.pdf}
-\end{center}
+\includegraphics[width=0.6\textwidth]{images/obstacle65.pdf} \qquad \includegraphics[width=0.3\textwidth]{images/obstacle-sets.png}
+
+\bigskip
+\only<1>{
+\begin{itemize}
+\item \emph{problem.} on a domain $\Omega \subset \RR^2$, find the displacement $u(x)$ of a membrane, with fixed value $u = g$ on $\partial \Omega$, above an \emph{obstacle} $\psi(x)$, which minimizes the elastic energy
+    $$J(u) = \int_\Omega \frac{1}{2} |\grad u|^2 - f\, u$$
+\item shown above: $\Omega=(-2,2)^2$, $\psi(x)$ a hemisphere
+\end{itemize}
+}
+\only<2>{
+\begin{itemize}
+\item i.e.~constrained optimization over an \emph{admissible set}
+	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \text{ and } v \ge \psi\right\}$$
+\item informally, $J'(u)$ points directly into $\mathcal{K}$ (\emph{variational inequality}; VI):
+    $$\ip{J'(u)}{v-u} = \int_\Omega \grad u\cdot \grad (v-u) - f (v-u) \ge 0 \quad \text{for all } v \in \mathcal{K}$$
+\end{itemize}
+}
+\only<3>{
+\begin{itemize}
+\item the solution defines \emph{active} $A_u = \{u = \psi\}$ and \emph{inactive} $R_u = \{u> \psi\}$ subsets of $\Omega$, and a \emph{free boundary} $\Gamma_u=\partial R_u \cap \Omega$
+\item a naive strong form poses the problem in terms of its solution:
+{\small
+\begin{align*}
+-\grad^2 u &= f \text{ on $R_u$} \\
+u &= \psi \text{ on $A_u$}
+\end{align*}
+}
+\item by contrast, the weak form VI is meaningful
+\end{itemize}
+}
+\only<4>{
+\begin{itemize}
+\item a \emph{complementarity problem}, essentially KKT conditions, is a meaningful strong form:
+{\small
+\begin{align*}
+u - \psi &\ge 0 \\
+-\grad^2 u - f &\ge 0 \\
+(u - \psi)(-\grad^2 u - f) &= 0
+\end{align*}
+}
+\end{itemize}
+}
+\end{frame}
+
+\begin{frame}{general VI problems}
 
 \begin{itemize}
-\item x
+\item let $\mathcal{K}$ be a closed and convex subset of a Banach space $\mathcal{V}$
+\item suppose $F:\mathcal{K} \to \mathcal{V}'$ is a continuous, generally nonlinear \emph{operator}
+    \begin{itemize}
+    \item[$\circ$] $F$ may be defined only on $\mathcal{K}$
     \end{itemize}
-\end{frame}
+\item the general problem VI($F$,$\mathcal{K}$) is
+	$$\ip{F(u)}{v-u} \ge 0 \text{ for all } v \in \mathcal{K}$$
 
-\begin{frame}{active sets, free boundaries}
+\bigskip
+\item analogy:
 
-\begin{columns}
-\begin{column}{0.6\textwidth}
-\begin{itemize}
-\item x
+\begin{tabular}{l|l}
+x & y \\ \hline
+z & w
+\end{tabular}
 \end{itemize}
-\end{column}
-\begin{column}{0.4\textwidth}
-\includegraphics[width=0.6\textwidth]{images/obstacle-sets.png}
-\end{column}
-\end{columns}
 \end{frame}
@@
 
+\section{full approximation scheme (FAS) multigrid for PDEs?}
+
+\begin{frame}{Bratu model problem: optimality}
+
+\begin{columns}
+\begin{column}{0.6\textwidth}
+\begin{itemize}
+\item x
+\end{itemize}
+\end{column}
+\begin{column}{0.4\textwidth}
+\includegraphics[width=\textwidth]{images/bratu-exps.png}
+
+\includegraphics[width=\textwidth]{images/bratu-time.png}
+\end{column}
+\end{columns}
+\end{frame}
+
 \section{the nonlocal VI for a fluid layer in a climate}
@@
 \end{frame}
-
-\section{FAS multigrid for PDEs}
 
@@
     \end{itemize}
-\item<2> \alert{help me}
 \end{itemize}

### 2023-02-05 talk/slides.tex
a bit more about VIs

@@
 
-\begin{frame}{general VI problems}
+\begin{frame}{general variational inequalities}
 
@@
     \item[$\circ$] $F$ may be defined only on $\mathcal{K}$
+    \item[$\circ$] $F$ may \emph{not} be the derivative of a $J(u)$
     \end{itemize}
-\item the general problem VI($F$,$\mathcal{K}$) is
-	$$\ip{F(u)}{v-u} \ge 0 \text{ for all } v \in \mathcal{K}$$
+\item the general problem {\color{FireBrick} VI($F$,$\mathcal{K}$)} is
+	$${\color{FireBrick} \ip{F(u)}{v-u} \ge 0 \quad \text{ for all } v \in \mathcal{K}}$$
+\end{itemize}
+\end{frame}
 
-\bigskip
-\item analogy:
+\begin{frame}{VI $=$ ``constrained equation''}
 
+\begin{center}
 \begin{tabular}{l|l}
-x & y \\ \hline
-z & w
+\begin{minipage}[t][16mm][t]{0.4\textwidth}
+unconstrained optimization:
+$$\min_{u\in\mathcal{V}} J(u)$$
+\end{minipage}
+&
+\begin{minipage}[t][16mm][t]{0.4\textwidth}
+constrained optimization:
+$$\min_{u\in\mathcal{K}} J(u)$$
+\end{minipage}
+\\ \hline
+\begin{minipage}[t][16mm][t]{0.4\textwidth}
+solve equation for $u \in \mathcal{V}$: {\LARGE \strut}
+
+$$F(u)=0$$
+\end{minipage}
+&
+\begin{minipage}[t][16mm][t]{0.4\textwidth}
+
+\vspace{-2mm}
+solve {\color{FireBrick} VI($F$,$\mathcal{K}$)} for $u \in \mathcal{K}$:
+$${\color{FireBrick} \ip{F(u)}{v-u} \ge 0 \quad \forall v \in \mathcal{K}}$$
+\end{minipage}
 \end{tabular}
+\end{center}
+\end{frame}
+
+\begin{frame}{most familiar case}
+
+\begin{itemize}
+\item we even teach VIs when we introduce calculus
+
+\includegraphics[width=\textwidth]{images/calcone.png}
 \end{itemize}

### 2023-02-06 talk/references.tex
add de Diego and KikuchiOden, and clean doi command

@@
 
-\newcommand{\shref}[2]{\,{\tiny \href{#1}{#2}}}
+\newcommand{\sdoi}[1]{\,{\tiny \href{https://doi.org/#1}{doi:#1}}}
 \begin{itemize}
 \item[] A.~Brandt (1977). \emph{Multi-level adaptive solutions to boundary-value problems}, Mathematics of Computation 31 (138), 333--390
-\item[] A.~Brandt \& C.~Cryer (1983).
-\emph{Multigrid algorithms for the solution of linear complementarity problems \dots}, SIAM J.~Sci.~Stat.~Comput.~4 (4), 655--684 \shref{https://doi.org/10.1137/0904046}{doi:10.1137/0904046}
+\item[] A.~Brandt \& C.~Cryer (1983). \emph{Multigrid algorithms for the solution of linear complementarity problems \dots}, SIAM J.~Sci.~Stat.~Comput.~4 (4), 655--684 \sdoi{10.1137/0904046}
 %FTITLE Multigrid algorithms for the solution of linear complementarity problems arising from free boundary problems
-\item[] E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032 \shref{https://doi.org/10.1137/20M135217X}{doi:10.1137/20M135217X}
+\item[] E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032 \sdoi{10.1137/20M135217X}
 \item[] E.~Bueler (2021). \emph{PETSc for Partial Differential Equations}, SIAM Press, Philadelphia
-%\item[] E.~Bueler (2022). \emph{Performance analysis of high-resolution ice-sheet simulations}, J.~Glaciol., \href{https://doi.org/10.1017/jog.2022.113}{doi:10.1017/jog.2022.113}
-\item[] P.~Farrell, M.~Croci, \& T.~Surowiec (2020).
-\emph{Deflation for semismooth equations}, Optimization Methods \& Software 35 (6), 1248--1271 \shref{https://doi.org/10.1080/10556788.2019.1613655}{doi:10.1080/10556788.2019.1613655}
+%\item[] E.~Bueler (2022). \emph{Performance analysis of high-resolution ice-sheet simulations}, J.~Glaciol., \sdoi{10.1017/jog.2022.113}
+\item[] G.~de Diego, P.~Farrell, \& I.~Hewitt (2022). \emph{Numerical approximation of viscous contact problems applied to glacial sliding}, J.~Fluid Mech.~938 (A21). \sdoi{10.1017/jfm.2022.178}
+\item[] P.~Farrell, M.~Croci, \& T.~Surowiec (2020). \emph{Deflation for semismooth equations}, Optimization Methods \& Software 35 (6), 1248--1271 \sdoi{10.1080/10556788.2019.1613655}
 \item[] C.~Gr{\"a}ser \& R.~Kornhuber (2009). \emph{Multigrid methods for obstacle problems}, J.~Comput.~Math., 1--44
-\item[] G.
-Jouvet \& E.
-Bueler (2012).
-\emph{Steady, shallow ice sheets as obstacle problems \dots}, SIAM J.~Appl.~Math.~72 (4), 1292--1314 \shref{https://doi.org/10.1137/110856654}{doi:10.1137/110856654}
+\item[] G. Jouvet \& E. Bueler (2012). \emph{Steady, shallow ice sheets as obstacle problems \dots}, SIAM J.~Appl.~Math.~72 (4), 1292--1314 \sdoi{10.1137/110856654}
 %FTITLE Steady, shallow ice sheets as obstacle problems: well-posedness and finite element approximation
-\item[] D. Kinderlehrer \& G. Stampacchia (1980). \emph{An Introduction to Variational Inequalities and their Applications}, Academic Press, New York
+\item[] N.~Kikuchi \& J.~Oden (1988).  \emph{Contact problems in elasticity: a study of variational inequalities and finite element methods}, SIAM Press, Philadelphia
+\item[] D.~Kinderlehrer \& G.~Stampacchia (1980). \emph{An Introduction to Variational Inequalities and their Applications}, Academic Press, New York
 %\item[] T.~Isaac, G.~Stadler, \& O.~Ghattas (2015).
-\emph{Solution of nonlinear Stokes equations \dots, with application to ice sheet dynamics}, SIAM J.~Sci.~Comput.~37 (6), B804--B833 \shref{https://doi.org/10.1137/140974407}{doi:10.1137/140974407}
+\emph{Solution of nonlinear Stokes equations \dots, with application to ice sheet dynamics}, SIAM J.~Sci.~Comput.~37 (6), B804--B833 \sdoi{10.1137/140974407}
 %FTITLE Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics

### 2023-02-06 talk/slides.tex
start multigrid material; other structure

@@
 \end{itemize}
+
+\phantom{x}
 }
@@
 	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \text{ and } v \ge \psi\right\}$$
-\item informally, $J'(u)$ points directly into $\mathcal{K}$ (\emph{variational inequality}; VI):
+\item informally, $J'(u)$ points directly into $\mathcal{K}$ ({\color{FireBrick} \emph{variational inequality}; VI}):
     $$\ip{J'(u)}{v-u} = \int_\Omega \grad u\cdot \grad (v-u) - f (v-u) \ge 0 \quad \text{for all } v \in \mathcal{K}$$
@@
 \item by contrast, the weak form VI is meaningful
+
+\phantom{x}
 \end{itemize}
@@
 \begin{itemize}
-\item a \emph{complementarity problem}, essentially KKT conditions, is a meaningful strong form:
+\item a \emph{complementarity problem} (CP) is a meaningful strong form:
 {\small
@@
 }
+\item for optimization problems: CP $=$ KKT conditions
+
+\phantom{x}
+
+\phantom{x}
 \end{itemize}
@@
 \item let $\mathcal{K}$ be a closed and convex subset of a Banach space $\mathcal{V}$
-\item suppose $F:\mathcal{K} \to \mathcal{V}'$ is a continuous, generally nonlinear \emph{operator}
+\item suppose $F:\mathcal{K} \to \mathcal{V}'$ is a continuous, generally nonlinear operator
     \begin{itemize}
@@
 \begin{minipage}[t][16mm][t]{0.4\textwidth}
-solve equation for $u \in \mathcal{V}$: {\LARGE \strut}
+equation for $u \in \mathcal{V}$: {\LARGE \strut}
 
@@
 \vspace{-2mm}
-solve {\color{FireBrick} VI($F$,$\mathcal{K}$)} for $u \in \mathcal{K}$:
+{\color{FireBrick} VI($F$,$\mathcal{K}$)} for $u \in \mathcal{K}$:
 $${\color{FireBrick} \ip{F(u)}{v-u} \ge 0 \quad \forall v \in \mathcal{K}}$$
@@
 
-\begin{frame}{most familiar case}
+\begin{frame}{applications of VIs}
 
 \begin{itemize}
-\item we even teach VIs when we introduce calculus
+\item elastic contact, Signorini problems (e.g.~Kikuchi \& Oden 1988)
+\item viscous contact problems (de Diego et al.~2022)
+\item pricing of American options in the Black-Scholes model
+\item the geometry of glaciers \dots \emph{more soon}
 
-\includegraphics[width=\textwidth]{images/calcone.png}
+\bigskip
+\item first-semester calculus
+\end{itemize}
+
+\begin{center}
+\includegraphics[width=0.9\textwidth]{images/calcone.png}
+\end{center}
+\end{frame}
+
+\AtBeginSection[]
+{
+  \begin{frame}{Outline}
+    \tableofcontents[currentsection]
+  \end{frame}
+}
+
+\section{full approximation scheme (FAS) multigrid for PDEs?}
+
+\begin{frame}{a nonlinear 2-grid scheme}
+
+\begin{center}
+\includegraphics[width=0.3\textwidth]{images/two-grids.png}
+\end{center}
+
+\bigskip
+\only<1>{
+\begin{itemize}
+\item suppose we have a nonlinear elliptic PDE problem:
+	$$F(u) = 0$$
+\end{itemize}
+
+\phantom{x}
+}
+\only<2>{
+\begin{itemize}
+\item x
+\end{itemize}
+}
+\end{frame}
+
+\begin{frame}{nonlinear multigrid}
+
+\mbox{
+\includegraphics[height=0.10\textheight]{images/mg-grids.png} \qquad \includegraphics[height=0.15\textheight]{images/mg-vcycle.png} \quad
+\includegraphics[height=0.15\textheight]{images/mg-fcycle.png}
+}
+
+\bigskip
+\only<1>{
+\begin{itemize}
+\item suppose we have a nonlinear elliptic PDE problem:
+	$$F(u) = 0$$
+\end{itemize}
+
+\phantom{x}
+}
+\only<2>{
+\begin{itemize}
+\item x
+\end{itemize}
+}
+\end{frame}
+
+\begin{frame}{nonlinear elliptic PDE problem}
+
+\begin{itemize}
+\item x
 \end{itemize}
 \end{frame}
+
+\begin{frame}{Bratu model problem: optimality}
+
+\begin{columns}
+\begin{column}{0.6\textwidth}
+\begin{itemize}
+\item x
+\end{itemize}
+\end{column}
+\begin{column}{0.4\textwidth}
+\includegraphics[width=\textwidth]{images/bratu-exps.png}
+
+\includegraphics[width=\textwidth]{images/bratu-time.png}
+\end{column}
+\end{columns}
+\end{frame}
+
+\section{the nonlocal VI for a fluid layer in a climate}
+
+\begin{frame}{problem: fluid layer in a climate}
+
+\begin{center}
+\includegraphics[width=0.6\textwidth]{images/fluid-in-climate.png}
+\end{center}
+
+\begin{itemize}
+\item x
+\end{itemize}
+\end{frame}
+
+\section{a possibly scalable approach to nonlocal VIs}
 
@@
 \end{frame}
-
-\section{full approximation scheme (FAS) multigrid for PDEs?}
-
-\begin{frame}{Bratu model problem: optimality}
-
-\begin{columns}
-\begin{column}{0.6\textwidth}
-\begin{itemize}
-\item x
-\end{itemize}
-\end{column}
-\begin{column}{0.4\textwidth}
-\includegraphics[width=\textwidth]{images/bratu-exps.png}
-
-\includegraphics[width=\textwidth]{images/bratu-time.png}
-\end{column}
-\end{columns}
-\end{frame}
-
-\section{the nonlocal VI for a fluid layer in a climate}
-
-\begin{frame}{problem: fluid layer in a climate}
-
-\begin{center}
-\includegraphics[width=0.6\textwidth]{images/fluid-in-climate.png}
-\end{center}
-
-\begin{itemize}
-\item x
-\end{itemize}
-\end{frame}
-
-\section{a possibly scalable approach to nonlocal VIs}
 

### 2023-02-06 talk/background.tex
progress talk, plus reference structure

--- 2023-02-05 talk/references.tex
@@
 
-\newcommand{\shref}[2]{\,{\tiny \href{#1}{#2}}}
+\newcommand{\sdoi}[1]{\,{\tiny \href{https://doi.org/#1}{doi:#1}}}
 \begin{itemize}
-\item[] E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032 \shref{https://doi.org/10.1137/20M135217X}{doi:10.1137/20M135217X}
+\item[] A.~Brandt (1977). \emph{Multi-level adaptive solutions to boundary-value problems}, Mathematics of Computation 31 (138), 333--390
 \item[] E.~Bueler (2021). \emph{PETSc for Partial Differential Equations}, SIAM Press, Philadelphia
-%\item[] E.~Bueler (2022). \emph{Performance analysis of high-resolution ice-sheet simulations}, J.~Glaciol., \href{https://doi.org/10.1017/jog.2022.113}{doi:10.1017/jog.2022.113}
-\item[] P.~Farrell, M.~Croci, \& T.~Surowiec (2020).
-\emph{Deflation for semismooth equations}, Optimization Methods \& Software 35 (6), 1248--1271 \shref{https://doi.org/10.1080/10556788.2019.1613655}{doi:10.1080/10556788.2019.1613655}
-\item[] C.~Gr{\"a}ser \& R.~Kornhuber (2009). \emph{Multigrid methods for obstacle problems}, J.~Comput.~Math., 1--44
-\item[] G.
-Jouvet \& E.
-Bueler (2012).
-\emph{Steady, shallow ice sheets as obstacle problems: well-posedness and finite element approximation}, SIAM J.~Appl.~Math.~72 (4), 1292--1314 \shref{https://doi.org/10.1137/110856654}{doi:10.1137/110856654}
-\item[] D. Kinderlehrer \& G. Stampacchia (1980). \emph{An Introduction to Variational Inequalities and their Applications}, Academic Press, New York
-%\item[] T.~Isaac, G.~Stadler, \& O.~Ghattas (2015).
-\emph{Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics}, SIAM J.~Sci.~Comput.~37 (6), B804--B833 \shref{https://doi.org/10.1137/140974407}{doi:10.1137/140974407}
+\item[] N.~Kikuchi \& J.~Oden (1988).  \emph{Contact Problems in Elasticity: A Study of Variational Inequalities and Finite Element Methods}, SIAM Press, Philadelphia
+\item[] D.~Kinderlehrer \& G.~Stampacchia (1980). \emph{An Introduction to Variational Inequalities and their Applications}, Academic Press, New York
 \item[] U.~Trottenberg, C.~Oosterlee, \& A. Schuller (2001).  \emph{Multigrid}, Elsevier, Oxford

### 2023-02-06 talk/references.tex
progress talk, plus reference structure

@@
 \begin{itemize}
-\item[] A.~Brandt (1977). \emph{Multi-level adaptive solutions to boundary-value problems}, Mathematics of Computation 31 (138), 333--390
 \item[] A.~Brandt \& C.~Cryer (1983). \emph{Multigrid algorithms for the solution of linear complementarity problems \dots}, SIAM J.~Sci.~Stat.~Comput.~4 (4), 655--684 \sdoi{10.1137/0904046}
@@
 \item[] E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032 \sdoi{10.1137/20M135217X}
-\item[] E.~Bueler (2021). \emph{PETSc for Partial Differential Equations}, SIAM Press, Philadelphia
 %\item[] E.~Bueler (2022). \emph{Performance analysis of high-resolution ice-sheet simulations}, J.~Glaciol., \sdoi{10.1017/jog.2022.113}
@@
 %FTITLE Steady, shallow ice sheets as obstacle problems: well-posedness and finite element approximation
-\item[] N.~Kikuchi \& J.~Oden (1988).  \emph{Contact problems in elasticity: a study of variational inequalities and finite element methods}, SIAM Press, Philadelphia
-\item[] D.~Kinderlehrer \& G.~Stampacchia (1980). \emph{An Introduction to Variational Inequalities and their Applications}, Academic Press, New York
-%\item[] T.~Isaac, G.~Stadler, \& O.~Ghattas (2015).
-\emph{Solution of nonlinear Stokes equations \dots, with application to ice sheet dynamics}, SIAM J.~Sci.~Comput.~37 (6), B804--B833 \sdoi{10.1137/140974407}
-%FTITLE Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics
-\item[] U.~Trottenberg, C.~Oosterlee, \& A. Schuller (2001).  \emph{Multigrid}, Elsevier, Oxford
+\item[] A.~Reusken (1987). \emph{Convergence of the multigrid full approximation scheme \dots}, Numer.~Math.~52, 251--277 \sdoi{10.1007/BF01398879}
+%FTITLE Convergence of the multigrid full approximation scheme for a class of elliptic mildly nonlinear boundary value problems
 \end{itemize}

### 2023-02-07 talk/slides.tex
progress talk, plus reference structure
progress ... now dinner

@@
 \usepackage[kw]{pseudo}
-\pseudoset{left-margin=15mm,topsep=5mm,idfont=\texttt,st-left=,st-right=}
+\pseudoset{left-margin=15mm,topsep=5mm,label=,idfont=\texttt,st-left=,st-right=}
 
@@
 
-\newcommand{\mR}{R^{\bm{\oplus}}}
-\newcommand{\iR}{R^{\bullet}}
-
 \newcommand{\nn}{{\text{n}}}
@@
 \newcommand{\sold}{s_{\text{o}}}
+
+\newcommand{\maxR}{R^{\bm{\oplus}}}
+\newcommand{\minR}{R^{\bm{\ominus}}}
+\newcommand{\iR}{R^{\bullet}}
 
@@
 
-\includegraphics[width=0.6\textwidth]{images/obstacle65.pdf} \qquad \includegraphics[width=0.3\textwidth]{images/obstacle-sets.png}
+\includegraphics[width=0.55\textwidth]{images/obstacle65.pdf} \qquad \includegraphics[width=0.35\textwidth]{images/obstacle-sets.png}
 
@@
 \item \emph{problem.} on a domain $\Omega \subset \RR^2$, find the displacement $u(x)$ of a membrane, with fixed value $u = g$ on $\partial \Omega$, above an \emph{obstacle} $\psi(x)$, which minimizes the elastic energy
-    $$J(u) = \int_\Omega \frac{1}{2} |\grad u|^2 - f\, u$$
-\item shown above: $\Omega=(-2,2)^2$, $\psi(x)$ a hemisphere
+    $$J(v) = \int_\Omega \frac{1}{2} |\grad v|^2 - f\, v$$
+\item shown above: \quad $\Omega=(-2,2)^2$, $\psi(x)$ a hemisphere, $f(x)=0$
 \end{itemize}
@@
 	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \text{ and } v \ge \psi\right\}$$
-\item informally, $J'(u)$ points directly into $\mathcal{K}$ ({\color{FireBrick} \emph{variational inequality}; VI}):
+\item $J'(u)$ points directly into $\mathcal{K}$, the {\color{FireBrick} \emph{variational inequality} (VI)}:
     $$\ip{J'(u)}{v-u} = \int_\Omega \grad u\cdot \grad (v-u) - f (v-u) \ge 0 \quad \text{for all } v \in \mathcal{K}$$
@@
 \item the solution defines \emph{active} $A_u = \{u = \psi\}$ and \emph{inactive} $R_u = \{u> \psi\}$ subsets of $\Omega$, and a \emph{free boundary} $\Gamma_u=\partial R_u \cap \Omega$
-\item a naive strong form poses the problem in terms of its solution:
-{\small
+\item naive strong form would pose the problem in terms of its solution:
 \begin{align*}
--\grad^2 u &= f \text{ on $R_u$} \\
-u &= \psi \text{ on $A_u$}
+-\grad^2 u &= f \quad \text{ on $R_u$} \\
+u &= \psi \quad \text{ on $A_u$}
 \end{align*}
-}
-\item by contrast, the weak form VI is meaningful
+
+\phantom{x}
 
@@
 \begin{itemize}
-\item a \emph{complementarity problem} (CP) is a meaningful strong form:
-{\small
+\item the \emph{complementarity problem} (CP) is meaningful as a strong form:
 \begin{align*}
@@
 \end{align*}
-}
-\item for optimization problems: CP $=$ KKT conditions
+   \begin{itemize}
+   \item[$\circ$] for optimization problems: \quad CP $=$ KKT conditions
+   \end{itemize}
 
@@
     \item[$\circ$] $F$ may be defined only on $\mathcal{K}$
-    \item[$\circ$] $F$ may \emph{not} be the derivative of a $J(u)$
+    \item[$\circ$] $F$ may \emph{not}\, be the derivative of an objective function $J$
     \end{itemize}
@@
 \vspace{-2mm}
-{\color{FireBrick} VI($F$,$\mathcal{K}$)} for $u \in \mathcal{K}$:
+{\color{FireBrick} VI} for $u \in \mathcal{K}$:
 $${\color{FireBrick} \ip{F(u)}{v-u} \ge 0 \quad \forall v \in \mathcal{K}}$$
@@
 \item pricing of American options in the Black-Scholes model
-\item the geometry of glaciers \dots \emph{more soon}
+\item the geometry of glaciers \hfill $\longleftarrow$ \emph{more soon}
 
@@
 
-\begin{frame}{a nonlinear 2-grid scheme}
+\begin{frame}{nonlinear 2-grid scheme}
 
 \begin{center}
-\includegraphics[width=0.3\textwidth]{images/two-grids.png}
+$\Omega^h$\, \includegraphics[height=0.16\textheight]{images/fine-grid.png} \hspace{25mm} \includegraphics[height=0.16\textheight]{images/coarse-grid.png} \,$\Omega^H$
 \end{center}
 
-\bigskip
 \only<1>{
 \begin{itemize}
-\item suppose we have a nonlinear elliptic PDE problem:
-	$$F(u) = 0$$
+\item consider a nonlinear elliptic PDE problem:
+	$$F(u) = \ell$$
+
+	\begin{itemize}
+	\item for example, $F : \mathcal{V} \to \mathcal{V}'$ for $\mathcal{V}=H^1(\Omega)$, with $\ell\in \mathcal{V}'$
+	\end{itemize}
+\item FEM/FD discretization gives algebraic system on fine grid $\Omega^h$:
+    $$F^h(u^h) = \ell^h$$
+\item suppose $w^h$ yields residual norm $\|\ell^h - F^h(w^h)\| > \text{TOL}$
 \end{itemize}
@@
 }
-\only<2>{
+\only<2-3>{
 \begin{itemize}
-\item x
+\item how can we improve $w^h$ \emph{without} globally linearizing $F^h$?
+
+are there alternatives to Newton's method?
+\item note the \emph{residual} $r^h(w^h) = \ell^h - F^h(w^h)$ is computable,
+
+while the \emph{error} $e^h = w^h-u^h$ is unknown
+\item the residual definition can be rewritten
+    $$F^h(w^h) - F^h(u^h) \stackrel{\ast}{=} -r^h(w^h)$$
+\only<2>{\item for $F^h$ linear, try to solve this \emph{error eqn} $F^h(e^h) = -r^h(w^h)$ for $\tilde e^h$, and correct $w^h \leftarrow w^h-\tilde e^h$ to improve $w^h$?}
+\only<3>{\item \st{for $F^h$ linear, try to solve this \emph{error eqn} $F^h(e^h) = -r^h(w^h)$ for $\tilde e^h$, and correct $w^h \leftarrow w^h-\tilde e^h$ to improve $w^h$?}}
+\end{itemize}
+}
+\only<4>{
+\begin{itemize}
+\item \emph{nodewise problem}: for $\psi_j^h$ a hat function or dof, solve for $c\in\RR$:
+	$${\color{FireBrick} \phi_j(c) = r^h(w^h + c \psi_j^h)[\psi_j^h] = 0}$$
+\item sweeping through and solving nodewise problems is a \emph{smoother}
+    \begin{itemize}
+    \item[$\circ$] Fourier analysis on linear PDEs shows smoothing property
+    \item[$\circ$] post-smoothing, $w^h-u^h$ and $r^h(w^h)$ have small high-frequency
+    \end{itemize}
+\item post-smoothing, both sides of $\ast$ should be accurately approximate-able on a coarser grid
+\end{itemize}
+}
+\only<5>{
+\begin{itemize}
+\item Brandt (1977) proposed to replace $F^h(w^h) - F^h(u^h) = -r^h(w^h)$ with a coarse-grid, re-discretized version
+\item \emph{full approximation storage} (FAS) \emph{equation}:
+	$${\color{FireBrick} F^H(w^H) - F^H(\iR w^h) = -R r^h(w^h)}$$
+
+    \begin{itemize}
+    \item[$\circ$] $\iR:\mathcal{V}^h \to \mathcal{V}^H$ is \emph{injection}
+    \item[$\circ$] $R:(\mathcal{V}^h)' \to (\mathcal{V}^H)'$ is \emph{canonical restriction}
+    \item[$\circ$] if $w^h=u^h$ exactly then $w^H = \iR w^h$
+    \end{itemize}
+
+FIXME CHECK CAREFULLY FROM HERE
+\item equivalently ${\color{FireBrick} F^H(w^H) = \ell^H}$ where ${\color{FireBrick} \ell^H = FIXME R\,F^h(u^h) - \iR\, r^h(w^h)}$
 \end{itemize}
@@
 
-\begin{frame}{nonlinear multigrid}
+\begin{frame}{FAS 2-grid solver}
 
-\mbox{
-\includegraphics[height=0.10\textheight]{images/mg-grids.png} \qquad \includegraphics[height=0.15\textheight]{images/mg-vcycle.png} \quad
+\begin{align*}
+&\text{smooth by sweeps over $j$:}  & &w^h \leftarrow \left[\phi_j(c) = 0 \,\forall j\right] \\
+&\text{evaluate:}                   & &\ell^H = R\,F^h(u^h) - \iR\, r^h(w^h) \\
+&\text{solve:}                       & &F^H(w^H) = \ell^H \\
+&\text{correct}:                    & &w^h \leftarrow w^h + P(w^H - \iR w^h) \\
+&\text{smooth by sweeps over $j$:}  & &w^h \leftarrow \left[\phi_j(c) = 0 \,\forall j\right]
+\end{align*}
+
+\bigskip
+\begin{itemize}
+\item evaluate$+$solve$+$correct \, $=$ \, \emph{coarse grid correction}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{nonlinear multigrid = FAS V-cycle or F-cycle}
+
+\includegraphics[height=0.15\textheight]{images/mg-grids.png}
+
+\bigskip
+
+\hfill \mbox{
+\includegraphics[height=0.15\textheight]{images/mg-vcycle.png} \quad
 \includegraphics[height=0.15\textheight]{images/mg-fcycle.png}
@@
 
-\bigskip
+\vspace{-3mm}
+
 \only<1>{
-\begin{itemize}
-\item suppose we have a nonlinear elliptic PDE problem:
-	$$F(u) = 0$$
-\end{itemize}
-
-\phantom{x}
+{\small
+\begin{pseudo}
+\pr{fas-vcycle}: \\+
+    for $j=J$ downto $j=1$ \\+
+      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j;u^j)$ \\
+      $u^{j-1} = \iR u^j$ \\
+      $\ell^{j-1} = f^{j-1}(u^{j-1}) + R \left(\ell^j - f^j(u^j)\right)$ \\-
+    $\text{\pr{solve}}(\ell^0;u^0)$ \\
+    for $j=1$ to $j=J$ \\+
+      $u^j \gets u^j + P (u^{j-1} - \iR u^j)$ \\
+      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j;u^j)$ \\-
+\end{pseudo}
+}
 }
@@
 
+\begin{frame}{\emph{background} references}
+
+{\scriptsize
+%{\notsotiny
+\input{background.tex}
+}
+\end{frame}
+
 \end{document}

### 2023-02-07 talk/slides.tex
progress talk

@@
 	\end{itemize}
-\item FEM/FD discretization gives algebraic system on fine grid $\Omega^h$:
+\item discretization gives algebraic system on fine grid $\Omega^h$:
     $$F^h(u^h) = \ell^h$$
@@
 
-are there alternatives to Newton's method?
+(are there alternatives to Newton's method?)
 \item note the \emph{residual} $r^h(w^h) = \ell^h - F^h(w^h)$ is computable,
@@
 \item the residual definition can be rewritten
-    $$F^h(w^h) - F^h(u^h) \stackrel{\ast}{=} -r^h(w^h)$$
-\only<2>{\item for $F^h$ linear, try to solve this \emph{error eqn} $F^h(e^h) = -r^h(w^h)$ for $\tilde e^h$, and correct $w^h \leftarrow w^h-\tilde e^h$ to improve $w^h$?}
-\only<3>{\item \st{for $F^h$ linear, try to solve this \emph{error eqn} $F^h(e^h) = -r^h(w^h)$ for $\tilde e^h$, and correct $w^h \leftarrow w^h-\tilde e^h$ to improve $w^h$?}}
+    $$F^h(u^h) - F^h(w^h) = r^h(w^h)$$
+\only<2>{\item for $F^h$ linear, try to solve this \emph{error equation} $F^h(e^h) = -r^h(w^h)$ for $\tilde e^h$, and correct $w^h \leftarrow w^h-\tilde e^h$ to improve $w^h$?}
+\only<3>{\item \st{for $F^h$ linear, try to solve this \emph{error equation} $F^h(e^h) = -r^h(w^h)$ for $\tilde e^h$, and correct $w^h \leftarrow w^h-\tilde e^h$ to improve $w^h$?}}
 \end{itemize}
@@
 \begin{itemize}
-\item \emph{nodewise problem}: for $\psi_j^h$ a hat function or dof, solve for $c\in\RR$:
-	$${\color{FireBrick} \phi_j(c) = r^h(w^h + c \psi_j^h)[\psi_j^h] = 0}$$
+\item goal: use a coarser mesh to estimate the error
+\item \emph{nodewise problem}: for $\psi_i^h$ a hat function or dof, solve for $c\in\RR$:
+	$${\color{FireBrick} \phi_i(c) = r^h(w^h + c \psi_i^h)[\psi_i^h] = 0}$$
 \item sweeping through and solving nodewise problems is a \emph{smoother}
@@
     \item[$\circ$] Fourier analysis on linear PDEs shows smoothing property
-    \item[$\circ$] post-smoothing, $w^h-u^h$ and $r^h(w^h)$ have small high-frequency
+    \item[$\circ$] post-smoothing, $e^h$ and $r^h(w^h)$ have smaller high-frequencies
     \end{itemize}
-\item post-smoothing, both sides of $\ast$ should be accurately approximate-able on a coarser grid
+\item Brandt (1977): post-smoothing, $F^h(u^h) - F^h(w^h) = r^h(w^h)$ should be accurately approximate-able on a coarser grid
 \end{itemize}
@@
 \begin{itemize}
-\item Brandt (1977) proposed to replace $F^h(w^h) - F^h(u^h) = -r^h(w^h)$ with a coarse-grid, re-discretized version
-\item \emph{full approximation storage} (FAS) \emph{equation}:
-	$${\color{FireBrick} F^H(w^H) - F^H(\iR w^h) = -R r^h(w^h)}$$
+\item \emph{full approximation storage} (FAS) equation:
+	$${\color{FireBrick} F^H(w^H) - F^H(\iR w^h) = R \, r^h(w^h)}$$
 
@@
     \item[$\circ$] $R:(\mathcal{V}^h)' \to (\mathcal{V}^H)'$ is \emph{canonical restriction}
-    \item[$\circ$] if $w^h=u^h$ exactly then $w^H = \iR w^h$
+    \item[$\circ$] if $w^h=u^h$ exactly then $w^H = \iR w^h$ by well-posedness
     \end{itemize}
 
-FIXME CHECK CAREFULLY FROM HERE
-\item equivalently ${\color{FireBrick} F^H(w^H) = \ell^H}$ where ${\color{FireBrick} \ell^H = FIXME R\,F^h(u^h) - \iR\, r^h(w^h)}$
+\item rewritten: \quad ${\color{FireBrick} F^H(w^H) = \ell^H}$ where ${\color{FireBrick} \ell^H = F^H(\iR w^h) + R\, r^h(w^h)}$
 \end{itemize}
@@
 \begin{align*}
-&\text{smooth by sweeps over $j$:}  & &w^h \leftarrow \left[\phi_j(c) = 0 \,\forall j\right] \\
-&\text{evaluate:}                   & &\ell^H = R\,F^h(u^h) - \iR\, r^h(w^h) \\
-&\text{solve:}                       & &F^H(w^H) = \ell^H \\
+&\text{smooth by sweeps over grid:} & &w^h \leftarrow \left[\phi_i(c) = 0 \,\forall i\right] \\
+&\text{restrict:}                   & &\ell^H = F^H(\iR w^h) + R\, r^h(w^h) \\
+&\text{solve coarse:}                      & &F^H(w^H) = \ell^H \\
 &\text{correct}:                    & &w^h \leftarrow w^h + P(w^H - \iR w^h) \\
-&\text{smooth by sweeps over $j$:}  & &w^h \leftarrow \left[\phi_j(c) = 0 \,\forall j\right]
+&\text{smooth by sweeps over grid:} & &w^h \leftarrow \left[\phi_i(c) = 0 \,\forall i\right]
 \end{align*}
@@
 \bigskip
+{\small
 \begin{itemize}
-\item evaluate$+$solve$+$correct \, $=$ \, \emph{coarse grid correction}
+\item $P: \mathcal{V}^H \to \mathcal{V}^h$ is \emph{prolongation}
+\item recall: \quad $\phi_i(c) = r^h(w^h + c \psi_i^h)[\psi_i^h]$
+\item restrict$+$(solve coarse)$+$correct \, $=$ \, \emph{coarse grid correction}
 \end{itemize}
+}
 \end{frame}
 
-\begin{frame}{nonlinear multigrid = FAS V-cycle or F-cycle}
+\begin{frame}{nonlinear multigrid by FAS V-cycle or F-cycle}
+
+\bigskip
 
@@
 
-\only<1>{
 {\small
 \begin{pseudo}
-\pr{fas-vcycle}: \\+
+\pr{fas-vcycle}$(\ell^J;w^J)$: \\+
     for $j=J$ downto $j=1$ \\+
-      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j;u^j)$ \\
-      $u^{j-1} = \iR u^j$ \\
-      $\ell^{j-1} = f^{j-1}(u^{j-1}) + R \left(\ell^j - f^j(u^j)\right)$ \\-
-    $\text{\pr{solve}}(\ell^0;u^0)$ \\
+      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j; w^j)$ \\
+      $w^{j-1} \gets \iR w^j$ \\
+      $\ell^{j-1} = F^{j-1}(w^{j-1}) + R \left(\ell^j - F^j(w^j)\right)$ \\-
+    $\text{\pr{solve}}(\ell^0;w^0)$ \\
     for $j=1$ to $j=J$ \\+
-      $u^j \gets u^j + P (u^{j-1} - \iR u^j)$ \\
-      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j;u^j)$ \\-
+      $w^j \gets w^j + P (w^{j-1} - \iR w^j)$ \\
+      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j;w^j)$ \\-
 \end{pseudo}
-}
-}
-\only<2>{
-\begin{itemize}
-\item x
-\end{itemize}
 }
@@
 
-\begin{frame}{nonlinear elliptic PDE problem}
+\begin{frame}{does it work?}
 
 \begin{itemize}
-\item x
+\item FAS multigrid works well on the right nonlinear PDE problem!
+\item example: Liouville-Bratu equation\footnote{exact solution by Liouville (1853) makes a nice test case}
+    $$-\nabla^2 u - e^u = 0$$
+with Dirichlet boundary conditions on $\Omega=(0,1)^2$
+\item minimal problem-specific code:
+    \begin{enumerate}
+    \item[1.] residual evaluation on grid level: $F^j(\cdot)$
+    \item[2.] pointwise smoother: $\phi_i(c) = 0 \,\forall i$
+        \begin{itemize}
+        \item[$\circ$] e.g.~nonlinear Jacobi or Gauss-Seidel iteration
         \end{itemize}
+    \item[3.] coarse solve can be same as smoother, or use Newton etc.
+    \end{enumerate}
+\end{itemize}
+\end{frame}
+
+\begin{frame}[fragile]
+\frametitle{multigrid solver composition in PETSc}
+
+\begin{itemize}
+\item implement using e.g.~the PETSc library\footnote{Portable Extensible Toolkit for Scientific computing \quad \href{https://petsc.org/release/}{\texttt{petsc.org}} \quad \includegraphics[height=3mm]{images/petsc.png}}
+    \begin{itemize}
+    \item[$\circ$] FAS multigrid is a nonlinear solver (SNES) type
+    \item[$\circ$] implement in C/Fortran/python/Julia(?)
+    \end{itemize}
+\item multigrid solvers in PETSc are \emph{composed} from smoothers on each level, and a solver on the coarse mesh
+\item example parallel run with $J$ grid levels:
+\end{itemize}
+\begin{Verbatim}[xleftmargin=15mm,fontsize=\scriptsize]
+mpiexec -n P ./bratu -da_grid_x 5 -da_grid_y 5 -da_refine J \
+    -snes_rtol 1.0e-12 \
+    -snes_type fas \
+    -snes_fas_type full \
+    -fas_levels_snes_type ngs \
+    -fas_levels_snes_ngs_sweeps 2 \
+    -fas_levels_snes_ngs_max_it 1 \
+    -fas_levels_snes_norm_schedule none \
+    -fas_coarse_snes_type ngs \
+    -fas_coarse_snes_max_it 1 \
+    -fas_coarse_snes_ngs_sweeps 4 \
+    -fas_coarse_snes_ngs_max_it 1 
+\end{Verbatim}
+
+%NOT IN MY BOOK \hfill \includegraphics[width=8mm]{images/frontcover.jpg}
+\end{frame}
+
+\begin{frame}{Bratu model problem: optimality}
+
+\begin{columns}
+\begin{column}{0.45\textwidth}
+\begin{itemize}
+\item runs on my laptop up to almost $N=10^8$ dofs
+\item observed optimality:
+\begin{align*}
+\text{flops} &= O(N^1) \\
+\text{exp evaluations} &= O(N^1) \\
+\text{processor time} &= O(N^1)
+\end{align*}
+\end{itemize}
+\end{column}
+\begin{column}{0.55\textwidth}
+\includegraphics<1>[width=\textwidth]{images/bratu-exps.png}
+
+\includegraphics<2>[width=\textwidth]{images/bratu-time.png}
+\end{column}
+\end{columns}
 \end{frame}

### 2023-02-07 talk/slides.tex
lots of progress ... and I have more than 24 hours left

@@
 \hypersetup{
-    pdftitle = {Toward nonlinear multigrid for nonlinear variational inequalities},
+    pdftitle = {Toward nonlinear multigrid for nonlinear and nonlocal variational inequalities},
     pdfauthor = {Ed Bueler},
@@
 \newcommand{\bn}{\mathbf{n}}
+\newcommand{\bq}{\mathbf{q}}
 \newcommand{\br}{\mathbf{r}}
@@
 
-\title[Multigrid for nonlinear VI]{Toward nonlinear multigrid \\ for nonlinear variational inequalities}
+\title[Multigrid for nonlinear VI]{Toward nonlinear multigrid \\ for nonlinear and nonlocal variational inequalities}
 
@@
 	$${\color{FireBrick} \ip{F(u)}{v-u} \ge 0 \quad \text{ for all } v \in \mathcal{K}}$$
+\item when $\mathcal{K}$ is nontrivial the problem {\color{FireBrick} VI($F$,$\mathcal{K}$)} is nonlinear \emph{even when $F$ is a linear operator}
 \end{itemize}
@@
 \item multigrid solvers in PETSc are \emph{composed} from smoothers on each level, and a solver on the coarse mesh
-\item example parallel run with $J$ grid levels:
+\item example run using FD discretization and FAS F-cycle:
 \end{itemize}
 \begin{Verbatim}[xleftmargin=15mm,fontsize=\scriptsize]
-mpiexec -n P ./bratu -da_grid_x 5 -da_grid_y 5 -da_refine J \
+./bratu -da_grid_x 5 -da_grid_y 5 -da_refine J \
     -snes_rtol 1.0e-12 \
@@
 \begin{itemize}
-\item runs on my laptop up to almost $N=10^8$ dofs
 \item observed optimality:
@@
 \end{align*}
+\item<1-> up to almost $N=10^8$ dofs at $J=11$ refinements
+\item<2> on my 2022 laptop
+\item<2> compare $\approx 20\,\mu\,\text{s}/\text{N}$ for Poisson equation, $P_1$ elements, and geometric multigrid (in Firedrake)
 \end{itemize}
@@
 
-\begin{frame}{Bratu model problem: optimality}
+\begin{frame}{benefits of FAS multigrid for nonlinear PDEs?}
 
-\begin{columns}
-\begin{column}{0.6\textwidth}
 \begin{itemize}
-\item x
+\item \alert{benefits} of FAS multigrid?
+    \begin{enumerate}
+    \item[1.] minimal code, esp.~in from-scratch implementations
+        \begin{itemize}
+        \item[$\circ$] just write residual plus pointwise smoother!
         \end{itemize}
-\end{column}
-\begin{column}{0.4\textwidth}
-\includegraphics[width=\textwidth]{images/bratu-exps.png}
+    \item[2.] composition with nonlinear preconditioners (Brune et al.~2015)
+    \end{enumerate}
+\item \alert{disadvantages?}
+    \begin{enumerate}
+    \item[1.] Firedrake/FENiCs \emph{do} automatically provide linearizations from UFL statements of weak forms
+    \item[2.] small literature of convergence or descriptive performance (see Trottenberg et al.~(2001), Reusken (1987))
+    \item[3.] not enough tutorial literature?
+    \end{enumerate}
 
-\includegraphics[width=\textwidth]{images/bratu-time.png}
-\end{column}
-\end{columns}
+\bigskip
+\item notable disadvantages relative to global linearization followed by linear multigrid (Newton-multigrid) for PDEs
+\end{itemize}
 \end{frame}
@@
 
-\begin{center}
-\includegraphics[width=0.6\textwidth]{images/fluid-in-climate.png}
-\end{center}
+\begin{itemize}
+\item let's not get stuck on textbook example problems!
+\item multigrid for a real-world VI problem?
+\item consider the steady state of a viscous layer of surface elevation $s(x,y)$ flowing with velocity $\bu(x,y,z)$ under its own weight, over fixed bed topography with elevation $b(x,y)$, in a \emph{climate} which adds or removes fluid at a signed rate $a(x,y)$ [$\text{m}\,\text{s}^{-1}$]
+    \begin{itemize}
+    \item[$\circ$] data $a,b$ defined on domain $\Omega \subset \RR^2$ (left figure)
+    \end{itemize}
+\item geophysical examples: \alert{glaciers and ice sheets}, sea ice, inland seas and lakes
+\end{itemize}
+
+\bigskip
+\hfill \mbox{\includegraphics[height=0.25\textheight]{images/domain-data.png} \hspace{7mm} \includegraphics[height=0.25\textheight]{images/domain-velocity.png}}
+\end{frame}
+
+\begin{frame}{glacier example}
+
+xxx
+
+\includegraphics[width=0.8\textwidth]{images/alps-seguinot2018.png}
+\end{frame}
+
+\begin{frame}{strong form}
 
 \begin{itemize}
+\item inequality constraint generates free boundary: $s \ge b$
+    \begin{itemize}
+    \item[$\circ$] equivalently: layer thickness must be nonnegative
+    \item[$\circ$] a free boundary arises where ablative climate ($a(x) < 0$) forces surface down to bed
+    \end{itemize}
+\item the fluid velocity $\bu$ is a generally non-local function of the fluid geometry
+\item naive strong form of the steady model: % ($\bn_s = \left<-s_x,-s_y,1\right>$ is surface normal):
+\begin{align*}
+s &\ge b                    & &\text{everywhere in } \Omega \\
+-\bu|_s \cdot \bn_s &= a    & &\text{where } s(x) > b(x)
+\end{align*}
+\end{itemize}
+
+\bigskip
+\hfill \mbox{\includegraphics[height=0.2\textheight]{images/domain-data.png} \hspace{7mm} \includegraphics[height=0.2\textheight]{images/domain-velocity.png}}
+\end{frame}
+
+\begin{frame}{VI for fluid layer in a climate}
+
+\begin{itemize}
+\item define admissible thicknesses: $\mathcal{K} = \left\{v \in \mathcal{V} \,:\, v \ge 0\right\}$
+
+    \begin{itemize}
+    \item[$\circ$] $\mathcal{V}$ to be determined by viscosity model\footnote{in lubrication approximation for glaciers, $u^{8/3} \in W^{1,4}(\Omega)$ (Jouvet \& Bueler, 2012)}
+    \end{itemize}
 \item x
 \end{itemize}
-\end{frame}
-
-\section{a possibly scalable approach to nonlocal VIs}
-
-\begin{frame}{Newton iterations for VIs}
-
-\begin{columns}
-\begin{column}{0.6\textwidth}
-\begin{itemize}
-\item x
-\end{itemize}
-\end{column}
-\begin{column}{0.4\textwidth}
-\includegraphics[width=\textwidth]{images/vi-newton-gmg-bad.png}
-
-\includegraphics[width=\textwidth]{images/vi-newton-gmg-good.png}
-\end{column}
-\end{columns}
-\end{frame}
-
-\begin{frame}{Newton-multigrid for VIs}
-
-xxx
-\end{frame}
-
-\begin{frame}{optimality}
-
-\begin{columns}
-\begin{column}{0.6\textwidth}
-\begin{itemize}
-\item x
-\end{itemize}
-\end{column}
-\begin{column}{0.4\textwidth}
-\includegraphics[width=\textwidth]{images/obstacle-flops-per-n.png}
-\end{column}
-\end{columns}
 \end{frame}
@@
 \begin{itemize}
-\item whether explicit or implicit, a non-shallow model solves a Stokes problem at each step:
+\item non-shallow model solves a Stokes problem at each step:
 \begin{align*}
@@
 \item this is the \alert{stress balance} (conservation of momentum) problem which determines velocity $\bu$ and pressure $p$
-\item how fast is the numerical solution process?
-    \begin{itemize}
-    \item[$\circ$] how do solution algorithms \alert{scale with increasing resolution}?
-    \end{itemize}
 \end{itemize}
@@
 
-\begin{frame}{multilevel NCP-coupled-to-Stokes solvers}
+\section{scalable approaches for VIs}
+
+\begin{frame}[fragile]
+\frametitle{Newton-multigrid for VIs}
 
 \begin{itemize}
-\item direct attack on the problem seems to require a \alert{multilevel} solver for \alert{variational inequalities} (VIs), but in the \alert{non-local residual case}
+\item multigrid for the classical obstacle problem?
     \begin{itemize}
-    \item[$\circ$] this seems not to exist
-    \item[$\circ$] the \alert{smoother} must compute an implicit-step SKE residual including surface-motion term $\Phi(s) = - \bu|_s\cdot \bn_s$ from a scalable Stokes solver
+    \item[$\circ$] a nonlinear problem, though $-\nabla^2$ is linear
     \end{itemize}
-\item near-optimal multilevel VI solvers exist for toy problems
+\item Newton-multigrid straightforward in PETSc:
+\begin{Verbatim}[xleftmargin=13mm,fontsize=\scriptsize]
+./obstacle -da_grid_x 3 -da_grid_y 3 \
+    -snes_type vinewtonrsls -ksp_type cg \
+    -pc_type mg -da_refine J
+\end{Verbatim}
     \begin{itemize}
-    \item[$\circ$] Poisson equation obstacle problem (Brandt \& Cryer 1983; Gr\"aeser \& Kornhuber 2009)
+    \item[$\circ$] linear solver applies to inactive variables
+    \end{itemize}
+\item the outer Newton iteration must determine the active set \alert{before} V-cycle multigrid can provide effective preconditioning
+    \begin{itemize}
+    \item[$\circ$] growing Newton (SNES) iterations
     \end{itemize}
 \end{itemize}
+
+\hfill \mbox{\includegraphics[height=0.2\textheight]{images/obstacle65.pdf} \hspace{15mm}
+\includegraphics[height=0.25\textheight]{images/vi-newton-gmg-bad.png}}
 \end{frame}
 
-\section{conclusion}
+\begin{frame}[fragile]
+\frametitle{Newton-multigrid for VIs: nested iteration}
+
+\begin{itemize}
+\item applying nested iteration (nonlinear F-cycle) resolves this:
+
+\vspace{2mm}
+\begin{Verbatim}[xleftmargin=13mm,fontsize=\scriptsize]
+./obstacle -da_grid_x 3 -da_grid_y 3 \
+    -snes_type vinewtonrsls -ksp_type cg \
+    -pc_type mg
+\end{Verbatim}
+
+\vspace{-4.9mm}
+\hspace{39mm} {\scriptsize \color{FireBrick} \texttt{-snes\_grid\_sequence J}}
+
+\vspace{2mm}
+    \begin{itemize}
+    \item[$\circ$] constant Newton (SNES) iterations
+    \item[$\circ$] grid-independent flops/time restored
+    \item[$\circ$] optimal $O(N^1)$ flops
+    \item[$\circ$] FIXME example in my book
+    \end{itemize}
+
+\vspace{-12mm}
+\hfill \includegraphics[width=0.25\textwidth]{images/mg-fcycle.png}
+\end{itemize}
+
+\bigskip
+\hspace{10mm} \includegraphics[height=0.25\textheight]{images/vi-newton-gmg-good.png}
+
+\vspace{-22mm}
+\hfill \includegraphics[height=0.45\textheight]{images/obstacle-flops-per-n.png}
+\end{frame}
+
+\begin{frame}{Newton-multigrid for VIs: semi-smooth}
+
+xxx
+
+semi-smooth Newton with penalty
+
+scaling also restores optimal
+
+performance (Farrell et al.~2020)
+
+FIXME Ulrich (20xx)
+
+\end{frame}
+
+\section{scalable approaches for \emph{nonlocal} VIs}
 

### 2023-02-08 talk/references.tex
lots of progress ... and I have more than 24 hours left
some progrees

@@
 %FTITLE Multigrid algorithms for the solution of linear complementarity problems arising from free boundary problems
+\item[] P.~Brune, M.~Knepley, B.~Smith, \& X.~Tu (2015). \emph{Composing scalable nonlinear algebraic solvers}, SIAM Review~57 (4), 535--565 \sdoi{10.1137/130936725}
 \item[] E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032 \sdoi{10.1137/20M135217X}
@@
 \item[] C.~Gr{\"a}ser \& R.~Kornhuber (2009). \emph{Multigrid methods for obstacle problems}, J.~Comput.~Math., 1--44
+\item[] T.~Isaac, G.~Stadler, \& O.~Ghattas (2015). \emph{Solution of nonlinear Stokes equations \dots ice sheet dynamics}, SIAM J.~Sci.~Comput., 37 (6), B804--B833 \sdoi{10.1137/140974407}
+%FTITLE Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics
+
 \item[] G. Jouvet \& E. Bueler (2012). \emph{Steady, shallow ice sheets as obstacle problems \dots}, SIAM J.~Appl.~Math.~72 (4), 1292--1314 \sdoi{10.1137/110856654}
@@
 %FTITLE Convergence of the multigrid full approximation scheme for a class of elliptic mildly nonlinear boundary value problems
+\item[] J.~Seguinot \& 5 others (2018).  \emph{Modelling last glacial cycle ice dynamics in the Alps}, The Cryosphere, 12 (10), 3265--3285 \sdoi{10.5194/tc-12-3265-2018}
+%FAUTHORS Seguinot, J., Ivy-Ochs, S., Jouvet, G., Huss, M., Funk, M., & Preusser, F.
 \end{itemize}

### 2023-02-08 talk/slides.tex
some progrees

@@
 
-\title[Multigrid for nonlinear VI]{Toward nonlinear multigrid \\ for nonlinear and nonlocal variational inequalities}
+\title[Multigrid for nonlinear VI]{Toward nonlinear multigrid \\ for nonlinear and non-local variational inequalities}
 
@@
 \begin{itemize}
-\item implement using e.g.~the PETSc library\footnote{Portable Extensible Toolkit for Scientific computing \quad \href{https://petsc.org/release/}{\texttt{petsc.org}} \quad \includegraphics[height=3mm]{images/petsc.png}}
+\item implemented using an FD discretization and PETSc\footnote{Portable Extensible Toolkit for Scientific computing \quad \href{https://petsc.org/release/}{\texttt{petsc.org}} \quad \includegraphics[height=3mm]{images/petsc.png}}
     \begin{itemize}
@@
 \item multigrid solvers in PETSc are \emph{composed} from smoothers on each level, and a solver on the coarse mesh
-\item example run using FD discretization and FAS F-cycle:
+\item FAS F-cycle run:
 \end{itemize}
@@
 \end{align*}
-\item<1-> up to almost $N=10^8$ dofs at $J=11$ refinements
-\item<2> on my 2022 laptop
-\item<2> compare $\approx 20\,\mu\,\text{s}/\text{N}$ for Poisson equation, $P_1$ elements, and geometric multigrid (in Firedrake)
+\item<1-> up to almost $N=10^8$ dofs ($J=11$)
+\item<2> compare $\approx 20\,\mu\,\text{s}/\text{N}$ for Poisson equation using Firedrake $P_1$ elements and geometric multigrid
+    \begin{itemize}
+    \item<2>[$\circ$] on my 2022 laptop
+    \end{itemize}
 \end{itemize}
@@
     \end{enumerate}
+
+\bigskip
 \item \alert{disadvantages?}
@@
     \item[1.] Firedrake/FENiCs \emph{do} automatically provide linearizations from UFL statements of weak forms
-    \item[2.] small literature of convergence or descriptive performance (see Trottenberg et al.~(2001), Reusken (1987))
+    \item[2.] small literature of convergence or descriptive performance for FAS (Trottenberg et al.~(2001), Reusken (1987))
     \item[3.] not enough tutorial literature?
     \end{enumerate}
-
-\bigskip
-\item notable disadvantages relative to global linearization followed by linear multigrid (Newton-multigrid) for PDEs
 \end{itemize}
@@
     \end{itemize}
-\item geophysical examples: \alert{glaciers and ice sheets}, sea ice, inland seas and lakes
+\item geophysical examples: \alert{glaciers and ice sheets}, sea ice, oceans and lakes
 \end{itemize}
@@
 
-\begin{frame}{glacier example}
+\begin{frame}{example: glacier ice coverage of the Alps in prior climates}
 
-xxx
+\includegraphics[width=1.02\textwidth]{images/alps-seguinot2018.png}
 
-\includegraphics[width=0.8\textwidth]{images/alps-seguinot2018.png}
+\vspace{-2mm}
+\hfill {\tiny Sequinot et al.~(2018)}
+
+%\vspace{-5mm}
+{\footnotesize
+\begin{itemize}
+\item<2> more ice sheet modeling at my Math.~Geoscience Seminar tomorrow 2pm L5
+\end{itemize}
+}
 \end{frame}
@@
     \begin{itemize}
-    \item[$\circ$] equivalently: layer thickness must be nonnegative
-    \item[$\circ$] a free boundary arises where ablative climate ($a(x) < 0$) forces surface down to bed
+    \item[$\circ$] ablative climate $a(x) < 0$ forces surface down to bed
+    \item[$\circ$] $s\ge b$ $\iff$ thickness $s-b$ must be nonnegative
     \end{itemize}
-\item the fluid velocity $\bu$ is a generally non-local function of the fluid geometry
 \item naive strong form of the steady model: % ($\bn_s = \left<-s_x,-s_y,1\right>$ is surface normal):
@@
 \end{align*}
+
+    \begin{itemize}
+    \item[$\circ$] surface velocity $\bu|_s$ is determined by fluid domain geometry
+    \item[$\circ$] true $\Phi(s) = -\bu|_s \cdot \bn_s$ is a \emph{non-local} function of $s$
+    \end{itemize}
 \end{itemize}
@@
 
+\begin{frame}{$\Phi(s) = - \bu|_s \cdot \bn_s$ for glacier ice?}
+
+\begin{itemize}
+\only<1>{
+\item \alert{Stokes model}
+
+solve the Stokes problem for $\bu,p$, then evaluate at surface:
+    $$\int_{\Lambda = \{b < z <s\}} 2 \nu(D\bu) D\bu : D\bv - p \Div\bv - (\Div\bu) q - \rhoi \bg \cdot \bv\,d\bx = 0 \quad \forall v,q$$
+    $$\Phi(s) = - \bu|_s \cdot \bn_s$$
+
+    \begin{itemize}
+    \item[$\circ$] also extend $\Phi(s)$ by 0 to all of $\Omega$
+    \item[$\circ$] non-Newtonian viscosity: $\nu(D\bu) = \frac{1}{2} \Gamma |D\bu|^{\pp-2}$ with e.g.~$\pp = \frac{4}{3}$
+    \item[$\circ$] well-posed problem for $\bu,p$
+    \item[$\circ$] near-optimal solvers available (Isaac et al 2015)
+    \end{itemize}
+}
+
+\only<2>{
+\item \alert{lubrication approximation model}
+
+apply a nonlinear differential operator to $s$:
+    $$\Phi(s) = - \frac{\gamma}{\qq} (s-b)^{\qq} |\grad_{\bx} s|^{\qq} - \grad_{\bx} \cdot\left(\frac{\gamma}{\qq+1} (s-b)^{\qq+1} |\grad_{\bx} s|^{\qq-2} \grad_{\bx} s\right)$$
+where $\qq = \nn+1$ and $\gamma>0$ is related to ice softness and density
+    \begin{itemize}
+    \item e.g.~$\qq = 4$
+    \item $\Phi(s)$ is a nonlinear differential operator only because membrane (longitudinal) stresses are not balanced in lubrication approximation
+    \end{itemize}
+}
+\end{itemize}
+\end{frame}
+
 \begin{frame}{VI for fluid layer in a climate}
@@
 \begin{itemize}
-\item define admissible thicknesses: $\mathcal{K} = \left\{v \in \mathcal{V} \,:\, v \ge 0\right\}$
+\item admissible surface elevations: \quad $\mathcal{K} = \left\{r \in \mathcal{V} \,:\, r \ge b\right\}$
 
@@
     \end{itemize}
-\item x
+\item VI problem:
+
 \end{itemize}

### 2023-02-08 talk/slides.tex
more talk work ... more work todo

@@
 
-\section{variational inequalities (VIs)?}
+\section{variational inequalities (VIs)}
 
@@
 
-\section{full approximation scheme (FAS) multigrid for PDEs?}
+\section{full approximation scheme (FAS) multigrid for PDEs}
 
@@
 \item multigrid for a real-world VI problem?
-\item consider the steady state of a viscous layer of surface elevation $s(x,y)$ flowing with velocity $\bu(x,y,z)$ under its own weight, over fixed bed topography with elevation $b(x,y)$, in a \emph{climate} which adds or removes fluid at a signed rate $a(x,y)$ [$\text{m}\,\text{s}^{-1}$]
+\item consider a viscous layer with surface elevation $s(x,y)$ flowing with velocity $\bu(x,y,z)$, driven by gravity, over fixed bed topography with elevation $b(x,y)$, in a \emph{climate} which adds or removes fluid at a signed rate $a(x,y)$ [$\text{m}\,\text{s}^{-1}$]
     \begin{itemize}
-    \item[$\circ$] data $a,b$ defined on domain $\Omega \subset \RR^2$ (left figure)
+    \item[$\circ$] data $a,b$ defined on domain $\Omega \subset \RR^2$
     \end{itemize}
-\item geophysical examples: \alert{glaciers and ice sheets}, sea ice, oceans and lakes
+\item geophysical examples: \alert{glaciers and ice sheets}, sea ice, lakes
 \end{itemize}
@@
 \begin{itemize}
-\item<2> more ice sheet modeling at my Math.~Geoscience Seminar tomorrow 2pm L5
+\item<2> more ice sheet modeling at my Math.~Geosci.~Seminar tomorrow 2pm L5
 \end{itemize}
@@
 
-\begin{frame}{strong form}
+\begin{frame}{naive strong form}
 
 \begin{itemize}
-\item inequality constraint generates free boundary: $s \ge b$
-    \begin{itemize}
-    \item[$\circ$] ablative climate $a(x) < 0$ forces surface down to bed
-    \item[$\circ$] $s\ge b$ $\iff$ thickness $s-b$ must be nonnegative
-    \end{itemize}
 \item naive strong form of the steady model: % ($\bn_s = \left<-s_x,-s_y,1\right>$ is surface normal):
@@
 s &\ge b                    & &\text{everywhere in } \Omega \\
--\bu|_s \cdot \bn_s &= a    & &\text{where } s(x) > b(x)
+-\bu|_s \cdot \bn_s &= a    & &\text{where } s(x,y) > b(x,y)
 \end{align*}
@@
     \begin{itemize}
-    \item[$\circ$] surface velocity $\bu|_s$ is determined by fluid domain geometry
-    \item[$\circ$] true $\Phi(s) = -\bu|_s \cdot \bn_s$ is a \emph{non-local} function of $s$
+    \item[$\circ$] surface velocity $\bu|_s$ is determined by fluid domain geometry, including $s$ itself
+    \item[$\circ$] $\bn_s=\left<-\grad s,1\right>$ is upward surface normal
+    \item[$\circ$] generally: $-\bu|_s \cdot \bn_s$ is a \emph{non-local} function of $s$
     \end{itemize}
+\item<2> the inequality constraint $s \ge b$ \alert{generates a free boundary}
+
+where the ablative climate $a < 0$ forces surface down to bed
 \end{itemize}
@@
 
-\begin{frame}{$\Phi(s) = - \bu|_s \cdot \bn_s$ for glacier ice?}
+\begin{frame}{how to evaluate $\Phi(s) = - \bu|_s \cdot \bn_s$ for glacier ice?}
 
@@
 
-solve the Stokes problem for $\bu,p$, then evaluate at surface:
-    $$\int_{\Lambda = \{b < z <s\}} 2 \nu(D\bu) D\bu : D\bv - p \Div\bv - (\Div\bu) q - \rhoi \bg \cdot \bv\,d\bx = 0 \quad \forall v,q$$
+solve the Stokes problem, then evaluate velocity at surface:
+    $$\int_{\Lambda(s) = \{b < z <s\}} 2 \nu(D\bu) D\bu : D\bv - p \Div\bv - (\Div\bu) q - \rhoi \bg \cdot \bv = 0 \quad \forall \bv,q$$
     $$\Phi(s) = - \bu|_s \cdot \bn_s$$
@@
     \begin{itemize}
-    \item[$\circ$] also extend $\Phi(s)$ by 0 to all of $\Omega$
-    \item[$\circ$] non-Newtonian viscosity: $\nu(D\bu) = \frac{1}{2} \Gamma |D\bu|^{\pp-2}$ with e.g.~$\pp = \frac{4}{3}$
-    \item[$\circ$] well-posed problem for $\bu,p$
+    \item[$\circ$] assuming incompressibility and non-Newtonian viscosity: $\nu(D\bu) = \frac{1}{2} \Gamma |D\bu|^{\pp-2}$ with $\pp = \frac{4}{3}$
+    \item[$\circ$] given $s$, this is a well-posed problem for velocity $\bu\in \mathbf{W}^{1,\pp}$ and pressure $p \in L^\qq$ on domain $\Lambda(s)$
     \item[$\circ$] near-optimal solvers available (Isaac et al 2015)
@@
 \only<2>{
-\item \alert{lubrication approximation model}
+\item \alert{lubrication approximation}\footnote{also known as the \emph{shallow ice approximation}} \alert{model}
 
-apply a nonlinear differential operator to $s$:
-    $$\Phi(s) = - \frac{\gamma}{\qq} (s-b)^{\qq} |\grad_{\bx} s|^{\qq} - \grad_{\bx} \cdot\left(\frac{\gamma}{\qq+1} (s-b)^{\qq+1} |\grad_{\bx} s|^{\qq-2} \grad_{\bx} s\right)$$
-where $\qq = \nn+1$ and $\gamma>0$ is related to ice softness and density
+apply a nonlinear elliptic differential operator to $s$:
+    $$\Phi(s) = - \frac{\gamma}{\qq} (s-b)^{\qq} |\grad s|^{\qq} - \grad \cdot\left(\frac{\gamma}{\qq+1} (s-b)^{\qq+1} |\grad s|^{\qq-2} \grad s\right)$$
     \begin{itemize}
-    \item e.g.~$\qq = 4$
-    \item $\Phi(s)$ is a nonlinear differential operator only because membrane (longitudinal) stresses are not balanced in lubrication approximation
+    \item $\qq = 4$
+    \item $\grad$ is in $x,y$ only
+    \item $\Phi(s)$ is a nonlinear differential operator because membrane stresses are \emph{not} balanced in the approximation
+    \item $\Phi(s)$ is doubly-degenerate
     \end{itemize}
@@
 \begin{itemize}
-\item admissible surface elevations: \quad $\mathcal{K} = \left\{r \in \mathcal{V} \,:\, r \ge b\right\}$
+\item admissible surface elevations:
+    $$\mathcal{K} = \left\{r \in \mathcal{V} \,:\, r \ge b\right\}$$
 
     \begin{itemize}
-    \item[$\circ$] $\mathcal{V}$ to be determined by viscosity model\footnote{in lubrication approximation for glaciers, $u^{8/3} \in W^{1,4}(\Omega)$ (Jouvet \& Bueler, 2012)}
+    \item[$\circ$] $\mathcal{V}$ to be determined by viscous fluid model\footnote{in shallow ice approximation, $(s-b)^{8/3} \in W^{1,4}(\Omega)$ (Jouvet \& Bueler, 2012)}
     \end{itemize}
-\item VI problem:
+\item VI problem for surface elevation $s\in\mathcal{K}$:
+	$$\ip{\Phi(s)}{r-s} \ge \ip{a}{r-s} \quad \text{ for all } r \in \mathcal{K}$$
+where
+    $$\Phi(s)=- \bu|_s \cdot \bn_s,$$
+with extension by 0 to all of $\Omega$, and $\bu$ is the velocity solution on
+    $$\Lambda(s) = \{(x,y,z) : b(x,y) < z < s(x,y)\}$$
 
@@
 
-\begin{frame}{the Stokes problem for ice}
+\begin{frame}{CP form (glacier ice in a climate)}
 
 \begin{itemize}
-\item non-shallow model solves a Stokes problem at each step:
+\item VI form on previous slide is too abstracted for clarity
+\item the strong form of the same problem is a \alert{complementarity problem (CP) coupled to a Stokes problem}:
 \begin{align*}
-- \nabla \cdot \left(2 \nu_\eps(D\bu)\, D\bu\right) + \nabla p - \rhoi \mathbf{g} &= \bzero && \text{in $\Lambda$} \\
+s - b &\ge 0 && \text{ in $\Omega$} \\
+- \bu|_s \cdot \bn_s - a &\ge 0 && \text{''} \\
+(s-b)(- \bu|_s \cdot \bn_s - a) &= 0 && \text{''} \\
+- \nabla \cdot \left(2 \nu(D\bu)\, D\bu\right) + \nabla p - \rhoi \mathbf{g} &= \bzero && \text{in $\Lambda(s)$} \\
 \nabla \cdot \bu &= 0 && \text{''} \\
-\btau_b - \bbf(\bu|_b) &= \bzero && \text{on $\Gamma_b$} \\
-\bu|_b \cdot \bn_b &= 0 && \text{''} \\
-\left(2 \nu_\eps(D\bu) D\bu - pI\right) \bn &= \bzero && \text{on $\Gamma_s$}
+\left(2 \nu(D\bu) D\bu - pI\right) \bn &= \bzero && \{z=s\} \subset \partial \Lambda(s) \\
+\bu &= \bzero && \{z=b\} \subset \partial \Lambda(s)
 \end{align*}
-\item this is the \alert{stress balance} (conservation of momentum) problem which determines velocity $\bu$ and pressure $p$
+\item solve this for $s$ on $\Omega$, and simultaneously for $\bu,p$ on $\Lambda(s) = \{(x,y,z) : b(x,y) < z < s(x,y)\}$
 \end{itemize}
@@
 
-\section{scalable approaches for VIs}
+\begin{frame}{a non-local VI problem}
+
+\begin{itemize}
+\item in the Stokes case, the residual $r(s) = a - \Phi(s)$ $= a + \bu|_s \cdot \bn_s$ \alert{depends non-locally on $s$}
+\item for example, consider $\bu_{(s+\psi)} - \bu_{(s)}$ from surface perturbation
+
+(hat function) $\psi$
+\end{itemize}
+
+\medskip
+\begin{center}
+\includegraphics[width=0.85\textwidth]{images/stokes-greens-arndt.png}
+\end{center}
+
+\vspace{-5mm}
+\hfill {\tiny P.~Arndt figure}
+\end{frame}
+
+\section{multigrid approaches for VIs?}
 
@@
     \end{itemize}
-\item Newton-multigrid straightforward in PETSc:
+\item Newton-multigrid is straightforward in PETSc:
 \begin{Verbatim}[xleftmargin=13mm,fontsize=\scriptsize]
@@
     \begin{itemize}
-    \item[$\circ$] linear solver applies to inactive variables
+    \item[$\circ$] CG-GMG V-cycles linear solver applies to inactive variables
     \end{itemize}
-\item the outer Newton iteration must determine the active set \alert{before} V-cycle multigrid can provide effective preconditioning
+\item the outer Newton iteration must converge on the active set \alert{before} multigrid can provide effective preconditioning
     \begin{itemize}
-    \item[$\circ$] growing Newton (SNES) iterations
+    \item[$\circ$] thus growing Newton (SNES) iterations!
     \end{itemize}
@@
 
-\section{scalable approaches for \emph{nonlocal} VIs}
+\section{multigrid approaches for \emph{nonlocal} VIs?}
 
@@
 
-\begin{frame}{\emph{background} references}
+\begin{frame}{additional background references}
 

### 2023-02-08 talk/references.tex
more talk work ... more work todo
fast and furious

@@
 \begin{itemize}
-\item[] A.~Brandt \& C.~Cryer (1983). \emph{Multigrid algorithms for the solution of linear complementarity problems \dots}, SIAM J.~Sci.~Stat.~Comput.~4 (4), 655--684 \sdoi{10.1137/0904046}
+\item A.~Brandt \& C.~Cryer (1983). \emph{Multigrid algorithms for the solution of linear complementarity problems \dots}, SIAM J.~Sci.~Stat.~Comput.~4 (4), 655--684 \sdoi{10.1137/0904046}
 %FTITLE Multigrid algorithms for the solution of linear complementarity problems arising from free boundary problems
-\item[] P.~Brune, M.~Knepley, B.~Smith, \& X.~Tu (2015). \emph{Composing scalable nonlinear algebraic solvers}, SIAM Review~57 (4), 535--565 \sdoi{10.1137/130936725}
-\item[] E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032 \sdoi{10.1137/20M135217X}
-%\item[] E.~Bueler (2022). \emph{Performance analysis of high-resolution ice-sheet simulations}, J.~Glaciol., \sdoi{10.1017/jog.2022.113}
-\item[] G.~de Diego, P.~Farrell, \& I.~Hewitt (2022). \emph{Numerical approximation of viscous contact problems applied to glacial sliding}, J.~Fluid Mech.~938 (A21). \sdoi{10.1017/jfm.2022.178}
-\item[] P.~Farrell, M.~Croci, \& T.~Surowiec (2020). \emph{Deflation for semismooth equations}, Optimization Methods \& Software 35 (6), 1248--1271 \sdoi{10.1080/10556788.2019.1613655}
-\item[] C.~Gr{\"a}ser \& R.~Kornhuber (2009). \emph{Multigrid methods for obstacle problems}, J.~Comput.~Math., 1--44
-\item[] T.~Isaac, G.~Stadler, \& O.~Ghattas (2015). \emph{Solution of nonlinear Stokes equations \dots ice sheet dynamics}, SIAM J.~Sci.~Comput., 37 (6), B804--B833 \sdoi{10.1137/140974407}
+\item P.~Brune, M.~Knepley, B.~Smith, \& X.~Tu (2015). \emph{Composing scalable nonlinear algebraic solvers}, SIAM Review~57 (4), 535--565 \sdoi{10.1137/130936725}
+\item E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032 \sdoi{10.1137/20M135217X}
+%\item E.~Bueler (2022). \emph{Performance analysis of high-resolution ice-sheet simulations}, J.~Glaciol., \sdoi{10.1017/jog.2022.113}
+\item G.~de Diego, P.~Farrell, \& I.~Hewitt (2022). \emph{Numerical approximation of viscous contact problems applied to glacial sliding}, J.~Fluid Mech.~938 (A21). \sdoi{10.1017/jfm.2022.178}
+\item P.~Farrell, M.~Croci, \& T.~Surowiec (2020). \emph{Deflation for semismooth equations}, Optimization Methods \& Software 35 (6), 1248--1271 \sdoi{10.1080/10556788.2019.1613655}
+\item C.~Gr{\"a}ser \& R.~Kornhuber (2009). \emph{Multigrid methods for obstacle problems}, J.~Comput.~Math., 1--44
+\item T.~Isaac, G.~Stadler, \& O.~Ghattas (2015). \emph{Solution of nonlinear Stokes equations \dots ice sheet dynamics}, SIAM J.~Sci.~Comput., 37 (6), B804--B833 \sdoi{10.1137/140974407}
 %FTITLE Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics
-
-\item[] G. Jouvet \& E. Bueler (2012). \emph{Steady, shallow ice sheets as obstacle problems \dots}, SIAM J.~Appl.~Math.~72 (4), 1292--1314 \sdoi{10.1137/110856654}
+\item G.~Jouvet \& E.~Bueler (2012). \emph{Steady, shallow ice sheets as obstacle problems \dots}, SIAM J.~Appl.~Math.~72 (4), 1292--1314 \sdoi{10.1137/110856654}
 %FTITLE Steady, shallow ice sheets as obstacle problems: well-posedness and finite element approximation
-\item[] A.~Reusken (1987). \emph{Convergence of the multigrid full approximation scheme \dots}, Numer.~Math.~52, 251--277 \sdoi{10.1007/BF01398879}
+\item G.~Jouvet \& J.~Rappaz (2011). \emph{Analysis and finite element approximation of a nonlinear stationary {S}tokes problem \dots}, Adv.~Numer.~Analysis 2011 (164581) \sdoi{10.1155/2011/164581}
+%FTITLE Analysis and finite element approximation of a nonlinear stationary {S}tokes problem arising in glaciology
+\item R.~Kornhuber (1994). \emph{Monotone multigrid methods for elliptic variational inequalities I}, Numer.~Math.~69, 167--184 \sdoi{10.1007/BF03325426}
+\item A.~Reusken (1987). \emph{Convergence of the multigrid full approximation scheme \dots}, Numer.~Math.~52, 251--277 \sdoi{10.1007/BF01398879}
 %FTITLE Convergence of the multigrid full approximation scheme for a class of elliptic mildly nonlinear boundary value problems
-\item[] J.~Seguinot \& 5 others (2018).  \emph{Modelling last glacial cycle ice dynamics in the Alps}, The Cryosphere, 12 (10), 3265--3285 \sdoi{10.5194/tc-12-3265-2018}
+\item J.~Seguinot \& 5 others (2018).  \emph{Modelling last glacial cycle ice dynamics in the Alps}, The Cryosphere, 12 (10), 3265--3285 \sdoi{10.5194/tc-12-3265-2018}
 %FAUTHORS Seguinot, J., Ivy-Ochs, S., Jouvet, G., Huss, M., Funk, M., & Preusser, F.
+\item X.~Tai (2003). \emph{Rate of convergence for some constraint decomposition methods for nonlinear variational inequalities}, Numer.~Math.~93 (4), 755--786 \sdoi{10.1007/s002110200404}
 \end{itemize}

### 2023-02-08 talk/slides.tex
fast and furious

@@
     pdftitle = {Toward nonlinear multigrid for nonlinear and nonlocal variational inequalities},
-    pdfauthor = {Ed Bueler},
+    pdfauthor = {Ed Bueler and Patrick Farrell},
     pdfsubject = {},
@@
 
-\title[Multigrid for nonlinear VI]{Toward nonlinear multigrid \\ for nonlinear and non-local variational inequalities}
+\title[Multigrid for nonlinear and nonlocal VIs]{Toward nonlinear multigrid \\ for nonlinear and nonlocal variational inequalities}
 
@@
 
-\author{Ed Bueler}
-
-\institute[UAF]{University of Alaska Fairbanks}
+\author[Bueler and Farrell]{Ed Bueler \inst{1} \and Patrick Farrell \inst{2}}
+\institute[]{\inst{1} University of Alaska Fairbanks \and %
+             \inst{2} Mathematical Institute, Oxford University}
 
@@
 
-\titlegraphic{\hfill \includegraphics[width=0.15\textwidth]{images/uafbw.png}}
+\titlegraphic{\hfill \includegraphics[width=0.15\textwidth]{images/uafbw.png} \quad \includegraphics[width=0.10\textwidth]{images/oxford.png}}
 
@@
 
-\section{the nonlocal VI for a fluid layer in a climate}
+\section{the nonlinear and nonlocal VI for a fluid layer in a climate}
 
@@
     $$\Lambda(s) = \{(x,y,z) : b(x,y) < z < s(x,y)\}$$
-
 \end{itemize}
@@
 
-\section{multigrid approaches for VIs?}
+\begin{frame}{what's needed for multigrid to work here?}
+
+\begin{block}{viscous fluid layer geometry problem}
+	$$\ip{\Phi(s)}{r-s} \ge \ip{a}{r-s} \quad \text{ for all } r \in \mathcal{K}$$
+where
+\begin{itemize}
+\item $\mathcal{K} = \left\{r \in \mathcal{V} \,:\, r \ge b\right\}$
+\item $\Phi(s)=- \bu|_s \cdot \bn_s$
+\item $s$ is solution surface elevation
+\item $\bu$ is Stokes solution on $\Lambda(s) = \{b < z < s\}$
+\item $a$ (climate) and $b$ (bed elevation) are the input data
+\end{itemize} 
+\end{block}
+
+\bigskip
+\begin{itemize}
+\item what is needed for scalable multilevel solutions?
+    \begin{enumerate}
+    \item<2>[1.] iterates must be admissible
+    \item<2>[2.] global linearization must be avoided
+    \item<2>[3.] smoother cost must be comparable to one residual    
+    \end{enumerate}
+\end{itemize}
+\end{frame}
+
+\section{multigrid approaches for VIs}
 
 \begin{frame}[fragile]
-\frametitle{Newton-multigrid for VIs}
+\frametitle{Newton-multigrid for the classical obstacle problem}
 
 \begin{itemize}
-\item multigrid for the classical obstacle problem?
-    \begin{itemize}
-    \item[$\circ$] a nonlinear problem, though $-\nabla^2$ is linear
-    \end{itemize}
+\item VIs are nonlinear problems, even for linear operators like $-\nabla^2$
 \item Newton-multigrid is straightforward in PETSc:
@@
     \begin{itemize}
-    \item[$\circ$] CG-GMG V-cycles linear solver applies to inactive variables
+    \item[$\circ$] linear solver applies to inactive variables
+        \begin{itemize}
+        \item[] \texttt{rsls} = reduced space line search
         \end{itemize}
-\item the outer Newton iteration must converge on the active set \alert{before} multigrid can provide effective preconditioning
+    \item[$\circ$] Newton step equations solved by CG with GMG V-cycles
+    \end{itemize}
+\item issue: the outer Newton iteration must converge on the active set \alert{before} multigrid can provide effective preconditioning
     \begin{itemize}
-    \item[$\circ$] thus growing Newton (SNES) iterations!
+    \item[$\circ$] grid-dependent (growing) Newton iterations
     \end{itemize}
@@
 
-\hfill \mbox{\includegraphics[height=0.2\textheight]{images/obstacle65.pdf} \hspace{15mm}
-\includegraphics[height=0.25\textheight]{images/vi-newton-gmg-bad.png}}
+\medskip
+\hspace{10mm} \includegraphics[height=0.2\textheight]{images/obstacle65.pdf}
+
+\vspace{-20mm}
+\hfill \includegraphics[height=0.25\textheight]{images/vi-newton-gmg-bad.png}
 \end{frame}
@@
 \begin{frame}[fragile]
-\frametitle{Newton-multigrid for VIs: nested iteration}
+\frametitle{nested iteration}
 
@@
     \begin{itemize}
-    \item[$\circ$] constant Newton (SNES) iterations
-    \item[$\circ$] grid-independent flops/time restored
-    \item[$\circ$] optimal $O(N^1)$ flops
-    \item[$\circ$] FIXME example in my book
+    \item[$\circ$] grid-independent Newton iterations
+    \item[$\circ$] optimal $O(N^1)$ flops and time
+    \item<2>[$\circ$] Chapter 12 example in my new book
     \end{itemize}
@@
 \bigskip
-\hspace{10mm} \includegraphics[height=0.25\textheight]{images/vi-newton-gmg-good.png}
+\includegraphics[height=0.25\textheight]{images/vi-newton-gmg-good.png}
 
 \vspace{-22mm}
-\hfill \includegraphics[height=0.45\textheight]{images/obstacle-flops-per-n.png}
+\hspace{40mm} \includegraphics[height=0.45\textheight]{images/obstacle-flops-per-n.png} \qquad \includegraphics<2>[width=20mm]{images/frontcover.jpg}
 \end{frame}
 
-\begin{frame}{Newton-multigrid for VIs: semi-smooth}
+\begin{frame}{multigrid strategies for VIs}
 
-xxx
-
-semi-smooth Newton with penalty
-
-scaling also restores optimal
-
-performance (Farrell et al.~2020)
-
-FIXME Ulrich (20xx)
-
+\begin{itemize}
+\item semi-smooth Newton can also yield mesh-independent iterations (Farrell et al.~2020)
+\item other VI multilevel strategies:
+    \begin{itemize}
+    \item[$\circ$] projected FAS multigrid for linear CPs (Brandt \& Cryer, 1983)
+    \item[$\circ$] monotone multigrid (Kornhuber, 1994)
+    \item[$\circ$] multilevel constraint decomposition (Tai, 2003)
+    \end{itemize}
+\end{itemize}
 \end{frame}
 
-\section{multigrid approaches for \emph{nonlocal} VIs?}
+\section{multigrid approaches for nonlinear and \emph{nonlocal} VIs?}
+
+\newcommand{\stacktwo}[2]{\begin{tabular}{c} #1 \\ #2 \end{tabular}}
+
+\begin{frame}{multigrid strategies for nonlinear VIs: feature table}
+
+{\small
+\begin{tabular}{l|ccccc}
+   & \stacktwo{admissible}{iterates} & \stacktwo{mesh-indep.}{rates}
+       & \stacktwo{no global}{linearization}
+         & \stacktwo{PETSc or}{Firedrake}  \\ \hline
+RS NM                      & $\checkmark$ &  & & $\checkmark$ \\
+\quad $+$ NI & $\checkmark$ & $\checkmark$ & & $\checkmark$ \\
+SS NM             &              & $\checkmark$ & & $\checkmark$ \\
+{\color{FireBrick} FASCD}  & $\checkmark$ & {\color{FireBrick} \textbf{?}} & $\checkmark$ & {\color{FireBrick} {\large $\checkmark$}}
+\end{tabular}
+}
+
+\bigskip
+{\scriptsize
+\begin{center}
+RS = reduced space, SS = semi-smooth, NM = Newton-multigrid, NI = nested iteration
+\end{center}
+}
+
+\vspace{5mm}
+\begin{itemize}
+\item for the non-local fluid layer VI problem we need all 4 checked
+\item we are trying-out a new algorithm,
+
+{\color{FireBrick} FASCD = full approximation storage constraint decomposition}
+
+    \begin{itemize}
+    \item[$\circ$] Firedrake implementation \only<2>{\dots as of yesterday}
+    \end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{constraint decomposition}
+
+CD gloss
+
+\begin{itemize}
+\item x
+\end{itemize}
+\end{frame}
+
+\begin{frame}{constraint decomposition}
+
+dishonest attempt figure
+
+\begin{itemize}
+\item x
+\end{itemize}
+
+\includegraphics[width=0.5\textwidth]{images/icedecomp.png}
+\end{frame}
+
+\begin{frame}{full approximation storage constraint decomposition}
+
+pseudocode
+\end{frame}
+
+\begin{frame}{evidence of mesh independence}
+
+\begin{itemize}
+\item 1D SIA
+\end{itemize}
+
+\includegraphics[width=0.5\textwidth]{images/sia-asymprates.pdf}
+\end{frame}
+
+\begin{frame}{2D SIA results (\emph{very fresh})}
+
+\begin{itemize}
+\item x
+\end{itemize}
+
+\includegraphics[width=0.5\textwidth]{images/fascd-firedrake-dome.png}
+\end{frame}
 
@@
 
-{\scriptsize
-%{\notsotiny
+%{\scriptsize
+{\notsotiny
 \input{references.tex}
@@
 
-{\scriptsize
-%{\notsotiny
+%{\scriptsize
+{\notsotiny
 \input{background.tex}

### 2023-02-09 talk/slides.tex
a bit of progress ... now for a practice run

@@
 
-\bigskip
 \begin{itemize}
-\item what is needed for scalable multilevel solutions?
+\item[]
+    \begin{itemize}
+    \item[$\circ$] for more on this problem see (Bueler, 2021)
+    \end{itemize}
+\end{itemize}
+
+\begin{itemize}
+\item {\color{FireBrick} what is needed for scalable multilevel solutions?}
     \begin{enumerate}
     \item<2>[1.] iterates must be admissible
-    \item<2>[2.] global linearization must be avoided
+    \item<2>[2.] global linearization of $\Phi(s)$ must be avoided
     \item<2>[3.] smoother cost must be comparable to one residual    
@@
 \begin{itemize}
-\item semi-smooth Newton can also yield mesh-independent iterations (Farrell et al.~2020)
+\item semi-smooth Newton also yields mesh-independent iterations
+    \begin{itemize}
+    \item[$\circ$] penalty scaling argument (Farrell et al.~2020)
+    \end{itemize}
 \item other VI multilevel strategies:
@@
     \item[$\circ$] monotone multigrid (Kornhuber, 1994)
-    \item[$\circ$] multilevel constraint decomposition (Tai, 2003)
+    \item[$\circ$] multilevel constraint decomposition (Tai, 2003)  \hfill $\gets$ \emph{more below}
     \end{itemize}
@@
 
-\begin{frame}{multigrid strategies for nonlinear VIs: feature table}
+\begin{frame}{multigrid strategies for VIs: feature table}
 
@@
 SS NM             &              & $\checkmark$ & & $\checkmark$ \\
-{\color{FireBrick} FASCD}  & $\checkmark$ & {\color{FireBrick} \textbf{?}} & $\checkmark$ & {\color{FireBrick} {\large $\checkmark$}}
+{\color{FireBrick} FASCD}  & $\checkmark$ & {\color{FireBrick} \textbf{?}} & $\checkmark$ & \only<2>{{\Large {\color{FireBrick} $\checkmark$}}} {\Large \strut}
 \end{tabular}
@@
 
-CD gloss
-
 \begin{itemize}
-\item x
+\item Tai's (2003) \emph{constraint decomposition} (CD) for VIs follows subspace decomposition for PDEs (Xu 1992)
+\item 
 \end{itemize}
@@
 
-\begin{frame}{constraint decomposition}
+\begin{frame}{as decomposition of the fluid layer}
 
@@
 \begin{itemize}
-\item x
+\item what does 
 \end{itemize}
 
-\includegraphics[width=0.5\textwidth]{images/icedecomp.png}
+\begin{center}
+\includegraphics[width=\textwidth]{images/icedecomp.png}
+\end{center}
 \end{frame}
@@
 
-pseudocode
+% case of lower obstacle only
+\begin{pseudo}
+\pr{fascd-vcycle}(\ell^J,\underline{\gamma}^J; w^J)\text{:} \\+
+    $\underline{\chi}^J = \underline{\gamma}^J - w^J {\large \strut}$ \\
+    %$\overline{\chi}^J = \overline{\gamma}^J - w^J {\large \strut}$ \\
+    for $j=J$ downto $j=1$ \\+
+      $\underline{\chi}^{j-1} = \maxR \underline{\chi}^j {\large \strut}$ \\
+      %$\overline{\chi}^{j-1} = \minR \overline{\chi}^j {\large \strut}$ \\
+      $\underline{\phi}^j = \underline{\chi}^j - P\underline{\chi}^{j-1} {\large \strut}$ \\
+      %$\overline{\phi}^j = \overline{\chi}^j - P\overline{\chi}^{j-1} {\large \strut}$ \\
+      $y^j = 0$ \\
+      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\underline{\phi}^j,w^j;y^j)$ \hspace{8mm} \ct{smoothing in $\mathcal{D}^j$}\\
+      $w^{j-1} = \iR(w^j + y^j)$ \\
+      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j+y^j)\right)$ \\-
+    $z^0 = 0$ \\
+    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,w^0;z^0)$ \hspace{10mm} \ct{coarse solve in $\mathcal{U}^0$} \\
+    for $j=1$ to $j=J$ \\+
+      $z^j = y^{j} + P z^{j-1}$ \\
+      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,w^j;z^j)$ \hspace{8mm} \ct{smoothing in $\mathcal{U}^j$} \\-
+    $w^J \gets w^J+z^J$
+\end{pseudo}
+\end{frame}
+
+\begin{frame}{2D SIA results (\emph{very fresh})}
+
+\begin{itemize}
+\item preliminary results
+\item dome test case in lubrication approximation
+    \begin{itemize}
+    \item[$\circ$] \emph{shallow ice approximation} = SIA
+    \item[$\circ$] $\Phi(s) = -\bu|_s \cdot \bn_s$ is a differential operator
+    \item[$\circ$] $s^{8/3} \in W^{1,p}(\Omega)$ but not in $C^2$
+    \end{itemize}
+\item FASCD algorithm using a strong smoother (\texttt{vinewtonrsls})
+    \begin{itemize}
+    \item[$\circ$] Firedrake $P_1$ elements
+    \end{itemize}
+\end{itemize}
+
+\begin{center}
+\includegraphics[width=0.7\textwidth]{images/fascd-firedrake-dome.png}
+\end{center}
 \end{frame}
@@
 \begin{itemize}
-\item 1D SIA
+\item again in the lubrication approximation,
 \end{itemize}
@@
 \includegraphics[width=0.5\textwidth]{images/sia-asymprates.pdf}
-\end{frame}
-
-\begin{frame}{2D SIA results (\emph{very fresh})}
-
-\begin{itemize}
-\item x
-\end{itemize}
-
-\includegraphics[width=0.5\textwidth]{images/fascd-firedrake-dome.png}
 \end{frame}

### 2023-02-09 talk/slides.tex
changes from trial run

@@
 
-\hfill \mbox{
-\includegraphics[height=0.15\textheight]{images/mg-vcycle.png} \quad
-\includegraphics[height=0.15\textheight]{images/mg-fcycle.png}
+\only<1>{
+\hfill \includegraphics[height=0.15\textheight]{images/mg-vcycle.png}
+}
+\only<2>{\hfill \includegraphics[height=0.15\textheight]{images/mg-fcycle.png} \quad \begin{tabular}{l} F-cycle = \\ nested iteration \end{tabular}
 }
@@
     \item[$\circ$] FAS multigrid is a nonlinear solver (SNES) type
-    \item[$\circ$] implement in C/Fortran/python/Julia(?)
+    \item[$\circ$] implement in C/Fortran/python
     \end{itemize}
@@
 \item multigrid for a real-world VI problem?
-\item consider a viscous layer with surface elevation $s(x,y)$ flowing with velocity $\bu(x,y,z)$, driven by gravity, over fixed bed topography with elevation $b(x,y)$, in a \emph{climate} which adds or removes fluid at a signed rate $a(x,y)$ [$\text{m}\,\text{s}^{-1}$]
+\item consider an incompressible, viscous layer with surface elevation $s(x,y)$, flowing with velocity $\bu(x,y,z)$, driven by gravity, over fixed bed topography with elevation $b(x,y)$, in a \emph{climate} which adds or removes fluid at a signed rate $a(x,y)$ [$\text{m}\,\text{s}^{-1}$]
     \begin{itemize}
@@
     \end{itemize}
-\item<2> the inequality constraint $s \ge b$ \alert{generates a free boundary}
+\item the inequality constraint $s \ge b$ \alert{generates a free boundary}
 
@@
 
-\begin{frame}{CP form (glacier ice in a climate)}
+\begin{frame}{CP form of viscous fluid layer in a climate}
 
@@
 \begin{align*}
-s - b &\ge 0 && \text{ in $\Omega$} \\
+s - b &\ge 0 && \text{ in $\Omega \subset \RR^2$} \\
 - \bu|_s \cdot \bn_s - a &\ge 0 && \text{''} \\
 (s-b)(- \bu|_s \cdot \bn_s - a) &= 0 && \text{''} \\
-- \nabla \cdot \left(2 \nu(D\bu)\, D\bu\right) + \nabla p - \rhoi \mathbf{g} &= \bzero && \text{in $\Lambda(s)$} \\
+- \nabla \cdot \left(2 \nu(D\bu)\, D\bu\right) + \nabla p - \rhoi \mathbf{g} &= \bzero && \text{in $\Lambda(s) \subset \RR^3$} \\
 \nabla \cdot \bu &= 0 && \text{''} \\
@@
 \end{align*}
-\item solve this for $s$ on $\Omega$, and simultaneously for $\bu,p$ on $\Lambda(s) = \{(x,y,z) : b(x,y) < z < s(x,y)\}$
+\item solve this for $s$ on $\Omega$, and simultaneously for $\bu,p$ on $\Lambda(s) = \{b < z < s\}$
 \end{itemize}
@@
     \begin{itemize}
-    \item[$\circ$] for more on this problem see (Bueler, 2021)
+    \item[$\circ$] for more on this problem class see (Bueler, 2021)
     \end{itemize}
@@
     \begin{enumerate}
-    \item<2>[1.] iterates must be admissible
-    \item<2>[2.] global linearization of $\Phi(s)$ must be avoided
-    \item<2>[3.] smoother cost must be comparable to one residual    
+    \item[1.] iterates must be admissible
+    \item[2.] global linearization of $\Phi(s)$ must be avoided
+    \item[3.] smoother cost must be comparable to one residual    
     \end{enumerate}
@@
 ./obstacle -da_grid_x 3 -da_grid_y 3 \
-    -snes_type vinewtonrsls -ksp_type cg \
-    -pc_type mg -da_refine J
+    -snes_type vinewtonrsls -ksp_type cg -pc_type mg \
+    -da_refine J
 \end{Verbatim}
@@
 ./obstacle -da_grid_x 3 -da_grid_y 3 \
-    -snes_type vinewtonrsls -ksp_type cg \
-    -pc_type mg
+    -snes_type vinewtonrsls -ksp_type cg -pc_type mg \
 \end{Verbatim}
 
-\vspace{-4.9mm}
-\hspace{39mm} {\scriptsize \color{FireBrick} \texttt{-snes\_grid\_sequence J}}
+\vspace{-1.5mm}
+\hspace{18.7mm} {\scriptsize \color{FireBrick} \texttt{-snes\_grid\_sequence J}}
 
@@
 \end{frame}
-
-\section{multigrid approaches for nonlinear and \emph{nonlocal} VIs?}
 
@@
 \end{frame}
+
+\section{multigrid approaches for nonlinear and \emph{nonlocal} VIs?}
 

### 2023-02-09 talk/references.tex
everything but the summary

@@
 \item C.~Gr{\"a}ser \& R.~Kornhuber (2009). \emph{Multigrid methods for obstacle problems}, J.~Comput.~Math., 1--44
-\item T.~Isaac, G.~Stadler, \& O.~Ghattas (2015). \emph{Solution of nonlinear Stokes equations \dots ice sheet dynamics}, SIAM J.~Sci.~Comput., 37 (6), B804--B833 \sdoi{10.1137/140974407}
+%\item T.~Isaac, G.~Stadler, \& O.~Ghattas (2015). \emph{Solution of nonlinear Stokes equations \dots ice sheet dynamics}, SIAM J.~Sci.~Comput., 37 (6), B804--B833 \sdoi{10.1137/140974407}
 %FTITLE Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics
@@
 %FTITLE Steady, shallow ice sheets as obstacle problems: well-posedness and finite element approximation
-\item G.~Jouvet \& J.~Rappaz (2011). \emph{Analysis and finite element approximation of a nonlinear stationary {S}tokes problem \dots}, Adv.~Numer.~Analysis 2011 (164581) \sdoi{10.1155/2011/164581}
+%\item G.~Jouvet \& J.~Rappaz (2011). \emph{Analysis and finite element approximation of a nonlinear stationary {S}tokes problem \dots}, Adv.~Numer.~Analysis 2011 (164581) \sdoi{10.1155/2011/164581}
 %FTITLE Analysis and finite element approximation of a nonlinear stationary {S}tokes problem arising in glaciology
@@
 %FTITLE Convergence of the multigrid full approximation scheme for a class of elliptic mildly nonlinear boundary value problems
-\item J.~Seguinot \& 5 others (2018).  \emph{Modelling last glacial cycle ice dynamics in the Alps}, The Cryosphere, 12 (10), 3265--3285 \sdoi{10.5194/tc-12-3265-2018}
+%\item J.~Seguinot \& 5 others (2018).  \emph{Modelling last glacial cycle ice dynamics in the Alps}, The Cryosphere, 12 (10), 3265--3285 \sdoi{10.5194/tc-12-3265-2018}
 %FAUTHORS Seguinot, J., Ivy-Ochs, S., Jouvet, G., Huss, M., Funk, M., & Preusser, F.

### 2023-02-09 talk/slides.tex
everything but the summary

@@
 
+\newcommand{\cK}{\mathcal{K}}
+
 \begin{frame}{constraint decomposition}
@@
 \begin{itemize}
-\item Tai's (2003) \emph{constraint decomposition} (CD) for VIs follows subspace decomposition for PDEs (Xu 1992)
-\item 
+\item Tai's (2003) \emph{constraint decomposition} (CD) for VIs follows the subspace decomposition idea (Xu 1992)
+\item suppose $\mathcal{K} \subset \mathcal{V}$ is a closed and convex admissible subset
+\item for a subspace decomposition $\mathcal{V} = \sum_i \mathcal{V}_i$, write the admissible subset as a sum
+    $$\mathcal{K} = \sum_i \mathcal{K}_i$$
+where $\mathcal{K}^i \subset \mathcal{V}^i$, with projections $\Pi_i : \mathcal{K} \to \mathcal{K}_i$
+\item CD additive and multiplicative iterations exist for $VI(F,\ell,\mathcal{K})$
+
+\begin{pseudo}
+\pr{cd-add}(u)\text{:} \\+
+    for $i \in \{0,\dots,m-1\}$: \\+
+        \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$, \\+
+            $\boxed{\ip{F(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}}$ \\--
+    $\hat w = \sum_i \hat w_i\in\cK$ \\
+    return $w=(1-\alpha) u + \alpha \hat w$
+\end{pseudo}
 \end{itemize}
+\end{frame}
+
+\begin{frame}{multilevel constraint decomposition}
+
+\begin{itemize}
+\item recall $\mathcal{K} = \{v \ge \psi\}$ in classical obstacle problem
+\item define \alert{defect obstacle} for a fine-level iterate $w^J$:
+    $$\chi^J = \psi^J - w^J$$
+\item \alert{monotone restriction} generates obstacles on each level:
+	$$\chi^j = R^{\oplus} \chi^{j+1}$$
+\item let $\mathcal{U}^j = \{z \ge \chi^j\}$, $\mathcal{D}^j = \{y \ge \chi^j - \chi^{j-1}\}$
+\item get CD of fine-level constraint set:
+	$$\mathcal{U}^J = \sum_{i=0}^J \mathcal{D}^i \phantom{KLDFJSKJDS SDF}$$
+\item multiplicative CD iteration $\to$ V-cycle
+\end{itemize}
+
+\vspace{-25mm}
+\hfill \includegraphics[width=0.4\textwidth]{images/fascd-vcycle.png}
 \end{frame}
@@
 
-dishonest attempt figure
-
 \begin{itemize}
-\item what does 
+\item again this is too abstract
+\item what does it look like for the fluid layer?
+    \begin{itemize}
+    \item[$\circ$] coarse grids have \emph{admissible} pieces of the fine-grid iterate
+    \end{itemize}
 \end{itemize}
 
+\bigskip
 \begin{center}
@@
 \begin{pseudo}
-\pr{fascd-vcycle}(\ell^J,\underline{\gamma}^J; w^J)\text{:} \\+
-    $\underline{\chi}^J = \underline{\gamma}^J - w^J {\large \strut}$ \\
-    %$\overline{\chi}^J = \overline{\gamma}^J - w^J {\large \strut}$ \\
+\pr{fascd-vcycle}(\ell^J,\psi^J; w^J)\text{:} \\+
+    $\chi^J = \psi^J - w^J {\large \strut}$ \\
     for $j=J$ downto $j=1$ \\+
-      $\underline{\chi}^{j-1} = \maxR \underline{\chi}^j {\large \strut}$ \\
-      %$\overline{\chi}^{j-1} = \minR \overline{\chi}^j {\large \strut}$ \\
-      $\underline{\phi}^j = \underline{\chi}^j - P\underline{\chi}^{j-1} {\large \strut}$ \\
-      %$\overline{\phi}^j = \overline{\chi}^j - P\overline{\chi}^{j-1} {\large \strut}$ \\
+      $\chi^{j-1} = \maxR \chi^j {\large \strut}$ \\
+      $\phi^j = \chi^j - P\chi^{j-1} {\large \strut}$ \\
       $y^j = 0$ \\
-      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\underline{\phi}^j,w^j;y^j)$ \hspace{8mm} \ct{smoothing in $\mathcal{D}^j$}\\
+      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\phi^j,w^j;y^j)$ \hspace{8mm} \ct{smoothing in $\mathcal{D}^j$}\\
       $w^{j-1} = \iR(w^j + y^j)$ \\
@@
     $z^0 = 0$ \\
-    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,w^0;z^0)$ \hspace{10mm} \ct{coarse solve in $\mathcal{U}^0$} \\
+    $\text{\pr{solve}}(\ell^0,\chi^0,w^0;z^0)$ \hspace{10mm} \ct{coarse solve in $\mathcal{U}^0$} \\
     for $j=1$ to $j=J$ \\+
       $z^j = y^{j} + P z^{j-1}$ \\
-      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,w^j;z^j)$ \hspace{8mm} \ct{smoothing in $\mathcal{U}^j$} \\-
+      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\chi^j,w^j;z^j)$ \hspace{8mm} \ct{smoothing in $\mathcal{U}^j$} \\-
     $w^J \gets w^J+z^J$
@@
 
-\begin{frame}{2D SIA results (\emph{very fresh})}
+\begin{frame}{2D shallow ice approximation results (\emph{very fresh})}
 
@@
     \begin{itemize}
-    \item[$\circ$] \emph{shallow ice approximation} = SIA
-    \item[$\circ$] $\Phi(s) = -\bu|_s \cdot \bn_s$ is a differential operator
-    \item[$\circ$] $s^{8/3} \in W^{1,p}(\Omega)$ but not in $C^2$
+    \item[$\circ$] here $\Phi(s)$ \emph{is} a differential operator
+    \item[$\circ$] note $s^{8/3} \in W^{1,4}(\Omega)$ but not in $C^2$
     \end{itemize}
-\item FASCD algorithm using a strong smoother (\texttt{vinewtonrsls})
+\item FASCD algorithm result
     \begin{itemize}
     \item[$\circ$] Firedrake $P_1$ elements
+    \item[$\circ$] strong smoother (\texttt{vinewtonrsls})
     \end{itemize}
@@
 
+\medskip
 \begin{center}
-\includegraphics[width=0.7\textwidth]{images/fascd-firedrake-dome.png}
+\includegraphics[width=0.8\textwidth]{images/fascd-firedrake-dome.png}
 \end{center}
@@
 \begin{itemize}
-\item again in the lubrication approximation,
+\item same lubrication approximation, but in 1D
+\item FASCD V-cycles with NGS and NJacobi smoothers
+\item up-smoothing preferred: V(0,2) beats V(1,1)
+\item evidence of mesh independence of factors $\|r^{(k+1)}\| / \|r^{(k)}\|$
 \end{itemize}
 
+\begin{center}
 \includegraphics[width=0.5\textwidth]{images/sia-asymprates.pdf}
+\end{center}
 \end{frame}

### 2023-02-09 talk/slides.tex
minimally acceptable talk

@@
 
-\begin{frame}{\alert{summary}}
+\begin{frame}{summary and outlook}
 
 \begin{itemize}
-\item x
+\item the \alert{variational inequality (VI)} problem class is good to know
+\item likewise \alert{full approximation storage (FAS) multigrid}
+    \begin{itemize}
+    \item[$\circ$] need for better support and documentation in PETSc/Firedrake
+    \end{itemize}
+\item multigrid treatment of nonlinear and \alert{nonlocal VIs}?
+    \begin{itemize}
+    \item[$\circ$] smoothers not obvious in nonlocal cases
+    \item[$\circ$] seeking practical evidence of mesh-independent convergence
+    \end{itemize}
+\item glacier evolution, as fluid-layer-in-climate problems, \alert{needs attention} from applied mathematicians and numerical analysts
+    \begin{itemize}
+    \item[$\circ$] VI form not widely recognized
+    \item[$\circ$] current state of the art $=$ explicit time stepping of surface
+        \begin{itemize}
+        \item[$\vartriangleright$] slow for science, intrinsically \emph{not} scalable
+        \end{itemize}
+    \item[$\circ$] to do: \alert{steady-state and implicit step VI problems}
+    \item[$\circ$] more on this view in my Math.~Geosci.~seminar tomorrow 2pm L5
+    \end{itemize}
 \end{itemize}

### 2023-02-09 talk/slides.tex
final fiddles

@@
 \begin{itemize}
-\item i.e.~constrained optimization over an \emph{admissible set}
+\item i.e.~constrained optimization over a convex \emph{admissible set}
 	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \text{ and } v \ge \psi\right\}$$
@@
 \item the residual definition can be rewritten
-    $$F^h(u^h) - F^h(w^h) = r^h(w^h)$$
+    $$F^h(u^h) - F^h(w^h) \stackrel{*}{=} r^h(w^h)$$
 \only<2>{\item for $F^h$ linear, try to solve this \emph{error equation} $F^h(e^h) = -r^h(w^h)$ for $\tilde e^h$, and correct $w^h \leftarrow w^h-\tilde e^h$ to improve $w^h$?}
@@
 \begin{itemize}
-\item goal: use a coarser mesh to estimate the error
+\item \emph{goal}: use a coarser mesh to estimate the error in $\ast$
 \item \emph{nodewise problem}: for $\psi_i^h$ a hat function or dof, solve for $c\in\RR$:
@@
     \end{itemize}
-\item Brandt (1977): post-smoothing, $F^h(u^h) - F^h(w^h) = r^h(w^h)$ should be accurately approximate-able on a coarser grid
+\item Brandt (1977): after smoothing, $F^h(u^h) - F^h(w^h) = r^h(w^h)$ should be accurately approximate-able on a coarser grid
 \end{itemize}
@@
 \begin{itemize}
+\item \emph{goal}: use a coarser mesh to estimate the error in $\ast$
 \item \emph{full approximation storage} (FAS) equation:
@@
 \item rewritten: \quad ${\color{FireBrick} F^H(w^H) = \ell^H}$ where ${\color{FireBrick} \ell^H = F^H(\iR w^h) + R\, r^h(w^h)}$
+
+\phantom{x}
+
+\phantom{x}
 \end{itemize}
@@
 
-\vspace{-3mm}
+\only<2>{\vspace{-3mm}}
 
@@
 \begin{itemize}
-\item FAS multigrid works well on the right nonlinear PDE problem!
+\item FAS multigrid works well on the right nonlinear PDE problem
 \item example: Liouville-Bratu equation\footnote{exact solution by Liouville (1853) makes a nice test case}
@@
 with Dirichlet boundary conditions on $\Omega=(0,1)^2$
-\item minimal problem-specific code:
+\item implement with minimal problem-specific code:
     \begin{enumerate}
@@
 \begin{itemize}
-\item implemented using an FD discretization and PETSc\footnote{Portable Extensible Toolkit for Scientific computing \quad \href{https://petsc.org/release/}{\texttt{petsc.org}} \quad \includegraphics[height=3mm]{images/petsc.png}}
+\item implemented here using an FD discretization and PETSc:\footnote{Portable Extensible Toolkit for Scientific computing \quad \href{https://petsc.org/release/}{\texttt{petsc.org}} \quad \includegraphics[height=3mm]{images/petsc.png}}
     \begin{itemize}
+    \item[$\circ$] multigrid solvers in PETSc are \emph{composed} from smoothers on each level, and a coarse-level solver \dots here these are nonlinear GS
     \item[$\circ$] FAS multigrid is a nonlinear solver (SNES) type
-    \item[$\circ$] implement in C/Fortran/python
+    \item[$\circ$] PETSc $=$ C/Fortran/python
     \end{itemize}
-\item multigrid solvers in PETSc are \emph{composed} from smoothers on each level, and a solver on the coarse mesh
-\item FAS F-cycle run:
+\item FAS multigrid F-cycle:
 \end{itemize}
@@
 \end{align*}
-\item<1-> up to almost $N=10^8$ dofs ($J=11$)
+\item<1-> up to $N\approx 10^8$ dofs
+    \begin{itemize}
+    \item[$\circ$] $J=11$ refinements
+    \item[$\circ$] laptop is memory-limited
+    \end{itemize}
 \item<2> compare $\approx 20\,\mu\,\text{s}/\text{N}$ for Poisson equation using Firedrake $P_1$ elements and geometric multigrid
-    \begin{itemize}
-    \item<2>[$\circ$] on my 2022 laptop
-    \end{itemize}
 \end{itemize}
@@
     \begin{itemize}
-    \item[$\circ$] surface velocity $\bu|_s$ is determined by fluid domain geometry, including $s$ itself
+    \item[$\circ$] surface velocity $\bu|_s$ is determined by fluid domain geometry $s$
     \item[$\circ$] $\bn_s=\left<-\grad s,1\right>$ is upward surface normal
@@
 
-where the ablative climate $a < 0$ forces surface down to bed
+if an ablative climate $a < 0$ forces surface down to bed
 \end{itemize}
@@
     \item $\grad$ is in $x,y$ only
-    \item $\Phi(s)$ is a nonlinear differential operator because membrane stresses are \emph{not} balanced in the approximation
+    \item $\Phi(s)$ is a nonlinear \alert{differential operator} in this model because membrane stresses are \emph{not} balanced
     \item $\Phi(s)$ is doubly-degenerate
@@
 
-\section{multigrid approaches for nonlinear and \emph{nonlocal} VIs?}
+\section{FAS multigrid for VIs?}
 

### 2023-02-19 paper/genfigs/cartoon.py
simplify cartoon: no dots needed for suggesting unboundedness

@@ -24,15 +24,10 @@ plt.text(-0.05, -0.07, r'$0$', fontsize=fsize)
 # filled gray box
 psi1 = 0.2
 psi2 = 0.35
-plt.fill([psi1, 1.2, 1.2, psi1, psi1],
-         [psi2, psi2, 0.8, 0.8, psi2],
+plt.fill([psi1, 1.4, 1.4, psi1, psi1],
+         [psi2, psi2, 1.0, 1.0, psi2],
          'k', alpha = 0.2)
 plt.text(0.9, 0.6, r'$\mathcal{K}$', fontsize=fsize)
-
-# put gray dots suggesting unboundedness
-plt.plot(0.01 + np.array([1.25, 1.30, 1.35]), 0.6 * np.array([1,1,1]), 'k.', ms = 12.0, markeredgecolor='w', alpha = 0.3)
-plt.plot(0.7 * np.array([1,1,1]), 0.01 + np.array([0.85, 0.90, 0.95]), 'k.', ms = 12.0, markeredgecolor='w', alpha = 0.3)
-plt.plot([1.25, 1.30, 1.35], [0.85, 0.90, 0.95], 'k.', ms = 12.0, markeredgecolor='w', alpha = 0.3)
 
 plt.text(1.28, -0.02, r'$\mathcal{V}_1$', fontsize=fsize)
 plt.plot(psi1, 0.0, 'k.', ms = 12.0)

### 2023-02-19 paper/fascd.tex
edits to language through end section 3

@@
 
-We extend the multilevel CD method by implementing fully-nonlinear corrections, and we apply it to a new set of nonlinear VI problems which are coercive but which do not necessarily arise from minimization.
+We extend Tai's multilevel CD method by implementing fully-nonlinear corrections, and we apply it to a new set of nonlinear VI problems which do not arise from minimization.
 Specifically, for finite element (FE) solvers the algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} are extended here in three practical ways:
@@
 Section \ref{sec:results} shows results from this \pr{fascd} algorithm for several nonlinear problems.
-\item The \pr{fascd} implementation permits ``box'' constraints, namely upper and lower obstacles; see Section \ref{sec:femultilevel}.
+\item Our \pr{fascd} implementation permits ``box'' constraints, namely upper and lower obstacles; see Section \ref{sec:femultilevel}.
 \item In Sections \ref{sec:ldcs}--\ref{sec:vcycle} we explain why, given the manner in which constraint sets are decomposed in the multilevel context \cite{GraeserKornhuber2009}, ``up-smoothing'' is more efficient than ``down-smoothing''.
@@
 
-The iterates from a CD method are always admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
-\pr{fascd} is designed for such admissible-only problems, some of which require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation, as discussed in \cite{Bueler2021conservation}.
-\pr{fascd} should permit direct solutions of those VI problems for which non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula.
+For the classical obstacle problem, based on the linear, symmetric Laplacian operator, certain multilevel techniques, such as the truncated monotone multigrid method \cite{GraeserKornhuber2009,Kornhuber1994}, are known to improve performance relative to the original multilevel CD method of Tai \cite{Tai2003}.
+These improved methods either track the active set in the discretization or modify the nodal basis functions, and in this sense they are discrete algorithms.
+By contrast a CD method applies essentially at the level of the continuum problem; see Section \ref{sec:cd}.
+Acceleration of the \pr{fascd} method via active-set and/or basis-level manipulations is an obvious topic for future research.
 
@@
 
-For the classical obstacle problem with a Laplacian operator, certain multilevel techniques, such as the truncated monotone multigrid method \cite{Kornhuber1994}, are known to improve performance relative to the original multilevel CD method from \cite{Tai2003}; see also \cite{GraeserKornhuber2009}.
-These improved methods either track the active set in the discretization or modify the nodal basis functions, and in this sense they are discrete algorithms.
-By contrast a CD method applies essentially at the level of the continuum problem; see Section \ref{sec:cd}.
-Acceleration of the \pr{fascd} method via active-set and/or basis-level manipulations is a topic for future research.
+The iterates from a CD method are admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
+\pr{fascd} is designed for such admissible-only problems.
+Our target fluid-layer problems require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation \cite{Bueler2021conservation}, but these problems are not addressed herein.
+Note that non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula in such a fluid-layer context.
 
@@
 \end{equation}
-Because $f$ is (generally) nonlinear, the source term $\ell$ is not strictly needed, and by redefining $f$ we may take $\ell=0$, but its presence clarifies the algorithm in Section \ref{sec:vcycle}.
+Because $f$ is (generally) nonlinear, the source term $\ell$ is not strictly needed in stating this class of problems.
+In fact, by redefining $f$ we may take $\ell=0$.
+However, allowing $\ell$ clarifies the algorithm in Section \ref{sec:vcycle}.
 
 VI \eqref{eq:vi} generalizes nonlinear systems of equations $f(u^*)=\ell$ to problems where $u^*$ is also constrained to be in $\cK$.
-Informally, regarding the dual pairing as an inner product, \eqref{eq:vi} says that the angle between $f(u^*)-\ell$ and any arbitrary vector $v-u$ pointing from $u$ into $\cK$ is at most $90^\circ$.
-While $f(u^*)-\ell$ may not be zero, it must point directly into $\cK$.
-On the other hand, if $u^*\in\cK^\circ$, in the interior of the constraint set, then \eqref{eq:vi} implies $f(u^*)=\ell$.
+Informally, pretending the dual pairing is an inner product, \eqref{eq:vi} says that the angle between $f(u^*)-\ell$ and any arbitrary vector $v-u^*$ pointing from $u^*$ into $\cK$ is at most $90^\circ$; while $f(u^*)-\ell$ may not be zero, it must point directly into $\cK$.
+On the other hand, if $u^*\in\cK^\circ$ is in the interior of the constraint set then \eqref{eq:vi} implies $f(u^*)=\ell$.
 
@@
 It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, and that the solution is unique when $f$ is strictly monotone.
-As in the calculus of variations \cite{Evans2010}, coercivity permits a compactness argument for unbounded sets $\cK$.
-(Recall that the closed and bounded subsets of a reflexive Banach space are weakly compact.)  The condition of continuity can be weakened, e.g.~to only apply on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but the stronger condition will apply in our examples.
+As is standard in the calculus of variations \cite{Evans2010}, coercivity permits a compactness argument for unbounded sets $\cK$.
+(Recall that the closed and bounded subsets of a reflexive Banach space are weakly compact.)  The condition of continuity can be weakened, e.g.~to only apply on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but it will apply in our examples.
 
@@
 Let $\Omega \subset \RR^d$ denote a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.
-Sobolev spaces \cite{Evans2010} are denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$, with $W^{1,p}_0(\Omega)$ denoting the subspace having zero boundary values.
+Sobolev spaces \cite{Evans2010} will be denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$, with $W^{1,p}_0(\Omega)$ denoting the subspace having zero boundary values.
 
@@
 a continuous map \cite[Theorem A.0.6]{Peral1997}.
-Now, if $x,y\in\RR^d$ then $(|x|^{p-2} x - |y|^{p-2} y)\cdot (x-y) \ge 2^{2-p} |x-y|^p$ \cite[see Appendix A and references therein]{Bueler2021conservation}.
-Thus it follows from the Poincar\'e inequality that
+Now, if $x,y\in\RR^d$ then $(|x|^{p-2} x - |y|^{p-2} y)\cdot (x-y) \ge 2^{2-p} |x-y|^p$ \cite[Appendix A]{Bueler2021conservation}.
+It follows from the Poincar\'e inequality that
     $$\ip{f(u) - f(v)}{u-v} \ge 2^{2-p} a_0 \|\grad u - \grad v\|_p^p \ge 2^{2-p} a_0 C \|u-v\|^p$$
@@
 For monotone operators $f$, VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
-In fact, suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cK \to \cV'$.
+Suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cK \to \cV'$.
 Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
@@
 \end{equation}
-The CD methods of Tai \cite{Tai2003} address problem \eqref{eq:minimization}; the analysis supposes that an objective $F$ exists, with $F'$ 2-coercive.
+The CD methods of Tai \cite{Tai2003} address problem \eqref{eq:minimization}; the analysis in \cite{Tai2003} supposes that an objective $F$ exists with $F'$ 2-coercive.
 
@@
 \begin{equation}
-F(v) = \int_\Omega \frac{a(x)}{p} |\grad v|^p\,dx - \ip{\ell}{v} \label{eq:plaplacianobjective}
-\end{equation}
-for any $\ell\in\cV'$.
-Then $F$ is a convex functional and $F'(v) = f(v) - \ell$ for $f$ given in \eqref{eq:plaplacian}.
-For any closed and convex $\cK\subset \cV$, VI problem \eqref{eq:vi} for \eqref{eq:plaplacian} is equivalent to optimization problem \eqref{eq:minimization} for \eqref{eq:plaplacianobjective}.
-
-Next we give two examples which are \emph{not} of optimization type, first a linear advection-diffusion problem and then a nonlinear porous-medium-type problem.
+F(v) = \frac{1}{p} \int_\Omega a(x) |\grad v|^p\,dx. \label{eq:plaplacianobjective}
+\end{equation}
+Then $F$ is a convex functional and $F'(v) = f(v)$ for $f$ given in \eqref{eq:plaplacian}.
+For any closed and convex $\cK\subset \cV$, VI problem \eqref{eq:vi} is then equivalent to optimization problem \eqref{eq:minimization}.
+
+Next we give two examples which are \emph{not} of optimization type, first a linear advection-diffusion problem, which needs the following Lemma, and then a nonlinear problem of porous-medium type.
 
@@
 \begin{example}  \label{ex:advectiondiffusion}  Suppose $\partial\Omega$ is partitioned into Dirichlet and Neumann portions, i.e.~$\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$, with $\partial_D\Omega$ of positive measure.
-Let $\cV = W_0^{1,2}(\Omega)$ be the space of functions which are zero along $\partial_D\Omega$.
-Consider a divergence-free velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, and assume that the flow is outward on the Neumann boundary, $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.
-For $u,v \in \cV = W_0^{1,2}(\Omega)$ and $\eps>0$ define
+Let $\cV \subset W^{1,2}(\Omega)$ be the space of functions which are zero along $\partial_D\Omega$.
+Consider a divergence-free velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, and assume that the flow is outward on the Neumann boundary: $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.
+For $u,v \in \cV$ and $\eps>0$ define
 \begin{equation}
@@
 
-References \cite{Bueler2021conservation,ChangNakshatrala2017} consider advection-diffusion VI problems like Example \ref{ex:advectiondiffusion}, specifically over the set $\cK = \{v\ge 0\}$.
-If $\bX \ne 0$ then VI \eqref{eq:vi} for $f$ in \eqref{eq:advectiondiffusion} does not correspond to a minimization problem.
+If $\bX \ne 0$ then $f$ in Example \ref{ex:advectiondiffusion} does not correspond to a minimization problem.
 Indeed, $\ip{f(u)}{v}$ is not symmetric in that case, so $f$ cannot be the gradient of a scalar objective.
-(For $u,v$ which are zero on $\partial \Omega$, note that $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.)
+(For $u,v$ which are zero on $\partial \Omega$, note that $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.)  References \cite{Bueler2021conservation,ChangNakshatrala2017} consider advection-diffusion VI problems like this, specifically over the set $\cK = \{v\ge 0\}$.
 
@@
 Numerical solver performance for Examples \ref{ex:plaplacian}, \ref{ex:advectiondiffusion}, and \ref{ex:porous} will be considered in Section \ref{sec:results}.
-Further examples of nonlinear VI problems appear in ice sheet models \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids.
+Further examples of nonlinear VI problems appear in ice sheet models \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids \cite{Bueler2021conservation}.
 
@@
 \end{equation}
-holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$.
-We say \eqref{eq:subspacedecomp} is a \emph{subspace decomposition} of $\cV$ \cite{Xu1992}.
+holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$, then we say \eqref{eq:subspacedecomp} is a \emph{space decomposition} of $\cV$ \cite{Xu1992}.
 Suppose further that $\cK_i \subset \cV_i$ are nonempty, closed, and convex subsets such that
@@
 for a closed, convex subset $\cK \subset \cV$.
-The sum in \eqref{eq:constraintdecomp} is required to hold in two senses \cite{TaiTseng2002}: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.
-(Sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp} because the $\cV_i$ are subspaces.)  Finally, for each $\cK_i$ suppose there are bounded, generally-nonlinear \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.
+The sum in \eqref{eq:constraintdecomp} must hold in two senses \cite{TaiTseng2002}: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.
+(Sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp} because the $\cV_i$ are subspaces.)  Finally, suppose there are bounded, generally-nonlinear \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.
 Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:femultilevel}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
@@
 A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} \cite{Tai2003}.
-
-Figure \ref{fig:cartoon} suggests how a low-dimensional example might look.
-(Typically $\cK_i \not\subset \cK$.)  Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique.
+Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique.
+Figure \ref{fig:cartoon} suggests how a low-dimensional example might look; as is typical, note $\cK_i \not\subset \cK$.
 
@@
 \includegraphics[width=0.6\textwidth]{genfigs/cartoon.pdf}
-\caption{Suppose $\mathcal{V}$ is the space of real functions on a two-point set $\Omega=\{x_1,x_2\}$.
-A constraint decomposition (CD) for a unilateral obstacle problem with $\mathcal{K}=\{v\ge \psi\}$ might look like this.}
+\caption{A constraint decomposition (CD) for a unilateral obstacle problem with $\Omega=\{x_1,x_2\}$, $\mathcal{V}=\left\{\text{real functions on } \Omega\right\}$, and $\mathcal{K}=\{v\ge \psi\}$.}
 \label{fig:cartoon}
@@
 In Sections \ref{sec:femultilevel}--\ref{sec:vcycle} we will introduce multilevel CD discretizations and algorithms for finite-dimensional VI problems.
-However, one can observe here that the CD concept applies at the level of the continuum problem.
-The following two obstacle problem examples are illustrations; the first is an overlapping domain decomposition.
+However, one can observe here that the CD concept applies at the level of the continuum problem, as illustrated in the following two examples.
 
@@
 Our second example is a disjoint frequency decomposition.
-A multilevel CD over a finite-dimensional FE discretization, specifically the one proposed in Section \ref{sec:cdmultilevel}, approximates such a frequency decomposition.
+A multilevel CD over a finite-dimensional FE discretization, proposed in Sections \ref{sec:femultilevel}--\ref{sec:vcycle}, approximates such a frequency decomposition.
 
@@
 
-Again note that $\cK_i \not\subset \cK$ in these cases; informally speaking, the important inclusion is $\cK_i \subset \cV_i$.
-In Example \ref{ex:domaindecomposition}, if $\psi$ is positive over portions of $\Omega$ where the decomposition into overlapping subdomains $\Omega_i$ is nontrivial, then $\cK_i \not\subset \cK$, and something similar holds for Example \ref{ex:frequencydecomposition}.
+Note that $\cK_i \not\subset \cK$ in these cases; informally speaking, the important inclusion is $\cK_i \subset \cV_i$.
 

### 2023-02-19 paper/fascd.tex
edits through end section 4

@@
 Associated to a constraint decomposition (CD) are certain iterative methods.
-Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below solve ``smaller'' VI problems, over each subset $\cK_i$, to replace $u \in \cK$ with $w\in\cK$ which should be closer to $u^* \in \cK$ solving \eqref{eq:vi}.
+Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below solve smaller VI problems over each subset $\cK_i$, to replace $u \in \cK$ with $w\in\cK$ which should be closer to $u^* \in \cK$ solving \eqref{eq:vi}.
 First is \pr{cd-add}, the additive or parallel method, which generalizes the classical Jacobi iteration.
 Here ``\textbf{for all}'' can be computed in any order.
-Note that the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$, and then adds-back the improved subset solution $\hat w_i \in \mathcal{K}_i$.
-Next is \pr{cd-mult}, the multiplicative or successive version, which generalizes the Gauss-Seidel iteration.
-Each subset solution corrects the global iterate immediately, so the $\mathcal{K}_i$ ordering is important.
+Note that the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$, and then adds-back the subset solution $\hat w_i \in \mathcal{K}_i$.
+Second is \pr{cd-mult}, the multiplicative or successive version, which generalizes the Gauss-Seidel iteration.
+Each subset solution corrects the global iterate immediately, so here ordering is important.
 The sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``parts of the input $u$ which have not yet improved''.
@@
 
-We define the $i$th \emph{subset update} as $e_i = \hat w_i - \Pi_i u \in \cV_i$.  The reader may check that $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.
-
-We will refer to \pr{cd-add} and \pr{cd-mult} as the \emph{basic CD iterations}.
-Note that each permits a damping parameter $\alpha$.
-Subject to a bound on $\alpha$, which is restrictive for \pr{cd-add} as noted by Tai \cite{Tai2003}, these iterations generate monotonically-decreasing objective values whenever VI problem \eqref{eq:vi} corresponds to convex optimization (problem \eqref{eq:minimization}).
+Let $e_i = \hat w_i - \Pi_i u \in \cV_i$.  The reader may check that $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.
+
+\pr{cd-add} and \pr{cd-mult} are the \emph{basic CD iterations}.
+Each one permits a damping parameter $\alpha$.
+Subject to a bound on $\alpha$, which is restrictive for \pr{cd-add} as noted by Tai \cite{Tai2003}, these iterations generate monotonically-decreasing objective values whenever VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization}.
 This is shown in the following Lemma.
@@
 
-\begin{lemma}  Suppose $f=F'$ for $F:\cK\to\RR$ convex and differentiable, and let $u\in\cK$.  For $0<\alpha\le 1$ the output $w \in \cK$ from \pr{cd-mult} satisfies
+\begin{lemma}  Suppose $f=F'$ for $F:\cK\to\RR$ convex and differentiable, and let $u\in\cK$.  For $0 \le \alpha \le 1$ the output $w \in \cK$ from \pr{cd-mult} satisfies
 \begin{equation}
@@
 \end{equation}
-If $0 < \alpha \le 1/m$ then \pr{cd-add} likewise satisfies \eqref{eq:objectivemonotone}.
+If $0 \le \alpha \le 1/m$ then \pr{cd-add} likewise satisfies \eqref{eq:objectivemonotone}.
 \end{lemma}
@@
 
-Each iteration of the \textbf{for} loop in \pr{cd-mult} lowers the objective value.
-To show this let $z_{i} = \sum_{j\le i} w_j + \sum_{j > i} \Pi_j u \in \cK$, $i=0,\dots,m-1$, be the state at the end of the $i$th iteration.
+To show that each iteration of the \textbf{for} loop in \pr{cd-mult} lowers the objective value, let $z_{i} = \sum_{j\le i} w_j + \sum_{j > i} \Pi_j u \in \cK$, $i=0,\dots,m-1$, be the state at the end of the $i$th iteration.
 Note that $z_0=u$, $z_{m-1}=w$, and $z_{i-1} = \sum_{j < i} w_j + \Pi_i u + \sum_{j > i} \Pi_j u$ for $i>0$.
-Using this last fact, convexity of $F$, observation \eqref{eq:admissibledifferencemult} with $v_i=\Pi_i u$, and the boxed VI we find:
+Therefore by convexity of $F$, observation \eqref{eq:admissibledifferencemult} with $v_i=\Pi_i u$, and the boxed VI we find:
 \begin{equation*}
@@
 
-Suppose $w=u+\alpha\sum_i e_i$ is computed by \pr{cd-add} using $0<\alpha \le 1/m$.  Considering each $e_i=\hat w_i - \Pi_i u$ separately, convexity of $F$ and the boxed VI with $v_i=\Pi_i u$ yields
+Suppose $w=u+\alpha\sum_i e_i$ is computed by \pr{cd-add} using $\alpha \le 1/m$.  Considering each $e_i=\hat w_i - \Pi_i u$ separately, convexity of $F$ and the boxed VI with $v_i=\Pi_i u$ yields
 \begin{equation}
@@
 \end{equation*}
-Now suppose $\alpha \le 1/m$ and define $\lambda = m\alpha$ so $0<\lambda\le 1$ and $w = (1-\lambda) u + \lambda v$.
-Thus by convexity and the above result, $F(w) \le (1-\lambda) F(u) + \lambda F(v) \le F(u)$.
+Now define $\lambda = m\alpha$ so $0 \le \lambda \le 1$ and $w = (1-\lambda) u + \lambda v$.  Thus by convexity and the above result, $F(w) \le (1-\lambda) F(u) + \lambda F(v) \le F(u)$.
 \end{proof}
@@
 When $\alpha=1/m$ then the logic of the above proof for \pr{cd-add} is simpler because $v=w$ and $\lambda=1$.
-The value $F(w)$ is less than $\frac{1}{m} \sum_i F(u+e_i)$ by Jensen's inequality, which is lower than $F(u)$ by the boxed VIs, i.e.~inequality \eqref{eq:cdaddonereduction}.
-
-The basic CD iterations are meaningful whether or not $f=F'$.
+Namely, the value $F(w)$ is less than $\frac{1}{m} \sum_i F(u+e_i)$ by Jensen's inequality, which is lower than $F(u)$ by the boxed VIs, which is inequality \eqref{eq:cdaddonereduction}.
+
+The basic CD iterations are meaningful whether we are in the optimization case $f=F'$ or not.
 Furthermore they are meaningful even when $f$ is nonlinear, non-local, or defined only on $\cK$.
-However, practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.
+Practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.
 References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem for which $f=F'$ is linear, local, and defined for all inputs (i.e.~$f$ is defined on all of $\mathcal{V}$).
@@
 Furthermore suppose $f$ is a differential operator, thus local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi$.
-Considering \pr{cd-add} for concreteness, by linearity the boxed VI can be written as
+Considering \pr{cd-add} for concreteness, by linearity the boxed VI can be rewritten
 \begin{equation}
@@
 Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible ($e_i \notin \cK$), but \eqref{eq:linearlocalvi} is meaningful because $f$ is defined over $\cV$.
-If a hat function basis exists for $\cV_i$, namely a basis of small-support functions such as hat functions, then by locality of $f$ a solver for VI \eqref{eq:linearlocalvi} can be implemented using only integrals over these small support sets.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}
+If a hat function basis exists for $\cV_i$, namely a basis of small-support functions, then by locality of $f$ a solver for VI \eqref{eq:linearlocalvi} can be implemented using only integrals over these small support sets.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}
 \end{example}
 
-When $f$ has all the nice properties listed in Example \ref{ex:fnice} then a solver based on CD iterations can take advantage of implementation efficiencies which are unavailable in general.
-That is, an FE approximation of VI \eqref{eq:vi}---see Sections \ref{sec:femultilevel} and \ref{sec:vcycle}---can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.
-Whatever is the actual solution process for VI \eqref{eq:linearlocalvi}, both storage and the computational cost of residual evaluation are substantially smaller than in the general case.
+When $f$ has all the nice properties listed in Example \ref{ex:fnice} then a solver based on CD iterations can take advantage of implementation efficiencies.
+That is, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.
+Whatever is the actual solution process for VI \eqref{eq:linearlocalvi}, both storage and the arithmetical cost of residual evaluation are substantially smaller than in the general case.
 
 For general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of \pr{cd-add} or \pr{cd-mult}.
-With a non-local integral operator $f$, for example, the expression $\ip{f(\phi)}{z}$ requires an integral over $\Omega$ even if $\phi$ is a function with small support, such as a hat function; cf.~subsection 4.5 in \cite{Bueler2021conservation}.
-In any case, Section \ref{sec:vcycle} constructs a multiplicative multilevel CD iteration for functions $f:\cK\to\cV'$, based on additive or multiplicative smoothers in Section \ref{sec:smoothers}, and assuming neither linearity nor locality, by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
+With a non-local integral operator $f$, for example, the expression $\ip{f(\phi)}{z}$ requires an integral over $\Omega$ even if $\phi$ is a function with small support; cf.~subsection 4.5 in \cite{Bueler2021conservation}.
+Section \ref{sec:vcycle} constructs a multiplicative multilevel CD iteration for general functions $f:\cK\to\cV'$, based on additive or multiplicative smoothers in Section \ref{sec:smoothers}, and assuming neither linearity nor locality, by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
 

### 2023-03-12 paper/fascd.tex
done with a round of edits

@@
 
-From now on we seek the approximate solution of certain nonlinear VI problems over 2D computational domains.
+From now on we seek the approximate solution of certain nonlinear VI problems \eqref{eq:vi} over 2D computational domains.
 Suppose $\Omega \subset \RR^2$ is a (open) bounded polygon and let $\mathcal{V}=W^{1,p}(\Omega)$.
-In addition to the free boundaries defined by inequality constraints (below), the additional fixed boundary conditions are allowed to be either Dirichlet or mixed.
-Suppose that $\partial\Omega$ is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive measure for simplicity, to permit well-posedness \cite{Evans2010}, and $\partial_N \Omega$ possibly empty.
-Suppose that the provided Dirichlet data $g_D:\partial_D \Omega \to \RR$ is sufficiently-regular so that the VI problem \eqref{eq:boxdirichletvi} below is well posed.
+In addition to the free boundaries generated by inequality constraints, fixed boundary conditions will be allowed, of either Dirichlet or mixed type.
 
 We restrict to the case of inequality \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.
-The \emph{lower obstacle} $\underline{\gamma}$ and the \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions on the closure $\overline{\Omega}$ which take their values in the extended real line $\tilde\RR = [-\infty,+\infty]$.
-Specifically suppose $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$ and $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$ are measurable, satisfying $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$ and $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.
-
-Now define the constraint set
-\begin{equation}
-\cK = \left\{w\,:\,\underline{\gamma} \le w \le \overline{\gamma} \, \text{ and }\, w\big|_{\partial_D \Omega} = g_D\right\} \subset \cV =W^{1,p}(\Omega).
-\end{equation}
-(The boundary values are, as usual, in the trace sense \cite{Evans2010}.)  This definition treats the inequality and equality (boundary) constraints in a unified manner.
-Note that $\cK$ is closed and convex.
-
-We will solve the same VI as in abstract problem \eqref{eq:vi}.  Given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we assume that the following VI is well-posed for $u^*\in \cK$:
+The \emph{lower obstacle} $\underline{\gamma}$ and the \emph{upper obstacle} $\overline{\gamma}$ are given functions, defined on the closure $\overline{\Omega}$, with values in the extended real line $\tilde\RR = [-\infty,+\infty]$.
+Specifically, we suppose $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$ and $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$ are measurable, $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$, and $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.
+
+Also, we assume that $\partial\Omega$ is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive measure for simplicity to permit well-posedness \cite{Evans2010}, and $\partial_N \Omega$ possibly empty.
+Suppose that the provided Dirichlet data $g_D:\partial_D \Omega \to \RR$ is sufficiently-regular so that the VI problem below is well posed.
+
+This data defines the constraint set
+\begin{equation}
+\cK = \left\{w\,:\,\underline{\gamma} \le w \le \overline{\gamma} \, \text{ and }\, w\big|_{\partial_D \Omega} = g_D\right\} \subset \cV =W^{1,p}(\Omega),
+\end{equation}
+a closed and convex set.  (Dirichlet boundary values are in the trace sense \cite{Evans2010}.)  Note that this definition treats the inequality and equality (boundary) constraints in a unified manner.
+
+We will solve the same VI as abstract problem \eqref{eq:vi}.  That is, given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we assume that the following VI is well-posed for $u^*\in \cK$:
 \begin{equation}
@@
 Note that $v-u^* \in W_0^{1,p}(\Omega)$, that is, test functions in \eqref{eq:boxdirichletvi} have zero values on the Dirichlet boundary $\partial_D\Omega$.
-While Dirichlet conditions apply on $\partial_D\Omega$, Neumann conditions over $\partial_N \Omega$ are determined by the formula for $f$.
-(E.g.~operator $f:\cK\to\cV'$ may include an integration over $\partial_N\Omega$.)  In summary, VI \eqref{eq:boxdirichletvi} is both a free- and fixed-boundary value problem.
+Of course, any Neumann conditions over $\partial_N \Omega$ are determined by the operator formula for $f$, which may include an integration over $\partial_N\Omega$.
+In summary, VI \eqref{eq:boxdirichletvi} is both a free- and fixed-boundary value problem.
 
@@
 A hierarchy of meshes will be constructed by refinement of a coarsest mesh, with the goal being a solution on the finest-level mesh.
-This hierarchy also allows the creation of a multilevel CD of $\cK$ in Section \ref{sec:cdmultilevel}.
+This hierarchy also allows the creation of a multilevel CD of $\cK$ (Section \ref{sec:cdmultilevel}).
 The subsets $\cK_i$ in this CD permit coarser-level corrections in a manner which preserves admissibility while not introducing unnecessary high frequencies into the coarser-level problems, especially at the free boundary.
@@
 The mesh hierarchy and FE spaces are constructed by nested refinement in the following standard manner \cite{Elmanetal2014}.
-Suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate elements (triangles or quadrilaterals) with union equal to $\overline{\Omega}$ and maximum edge length (characteristic size) $h_0>0$.
+Suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate elements (triangles or quadrilaterals) with union equal to $\overline{\Omega}$ and maximum edge length (characteristic mesh size) $h_0>0$.
+We assume $\mathcal{T}^0$ is conforming, with no hanging nodes.
 For $j=1,\dots,J$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ by connecting the midpoints of edges.
-For triangles and rectangles, each $\mathcal{T}^{j-1}$ element becomes four similar elements in $\mathcal{T}^j$ with size $h_j = 2^{-j} h_0$.\footnote{For arbitrary triangles, and for rectangles, edge midpoint refinement generates similar elements and $h_j=2^{-j}h_0$ is the exact maximum edge length in $\mathcal{T}^j$.
-For general quadrilaterals, however, similarity is lost and maximum edge lengths are variable; cf.~\cite{Zhang2004}.
-Only rectangular quadrilaterals are used in the Results section.}  We call $J\ge 0$ the \emph{depth} of the hierarchy and $\mathcal{T}^J$ the \emph{finest-level mesh}.
-Let $m_j$ be the number of nodes in $\mathcal{T}^j$, the \emph{degrees of freedom} at the $j$th level, and denote $h=h_J=2^{-J} h_0$ and $m=m_J$ for the finest level.
-
-Let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space \cite{Elmanetal2014} over $\mathcal{T}^j$, with arbitrary boundary values, and observe the following vector space nesting:
+For triangles and rectangles, each $\mathcal{T}^{j-1}$ element becomes four similar elements\footnote{For arbitrary triangles, and for rectangles, edge midpoint refinement generates similar elements and $h_j=2^{-j}h_0$ is the exact maximum edge length in $\mathcal{T}^j$.
+For general quadrilaterals, similarity is lost and maximum edge lengths are variable; cf.~\cite{Zhang2004}.
+However, only rectangular quadrilaterals are used in the Results section.} in $\mathcal{T}^j$, with characteristic mesh size $h_j = 2^{-j} h_0$.
+We call $J\ge 0$ the \emph{depth} of the hierarchy, $J+1$ the \emph{number of levels}, and $\mathcal{T}^J$ the \emph{finest-level mesh}.
+Let $m_j$ be the number of nodes in $\mathcal{T}^j$, so there are $m_j$ \emph{degrees of freedom} at the $j$th level.
+Denote $h=h_J=2^{-J} h_0$ and $m=m_J$ for the finest level.
+
+Let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space \cite{Elmanetal2014} over $\mathcal{T}^j$, with arbitrary boundary values.  Observe the following vector space nesting:
 \begin{equation}
@@
 
-Use of higher-order Lagrange elements $P_k$ or $Q_k$, $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} special to methods for inequality constraints.
+In the VI material which follows, use of higher-order Lagrange elements $P_k$ or $Q_k$, $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} of a type special to methods for inequality constraints.
 Namely, our FE construction of constraint sets will require \emph{nodal monotonicity},
@@
 \end{equation}
-for all $w,z \in \mathcal{V}^j$, where $\bw,\bz$ denote vectors of nodal values.
+for all $w,z \in \mathcal{V}^j$, where $\bw,\bz$ denote the vectors formed by the nodal values.
 For $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal value one at a single node, otherwise zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.
-It is easy to show that \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of sign-changing hat functions, so use of \eqref{eq:nodalmonotonicity} implies $k\le 1$.
+It is easy to show that \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of indefinite-sign hat functions, so use of \eqref{eq:nodalmonotonicity} implies $k\le 1$.
 
 Let $g_D^J \in \mathcal{V}^J$, on the finest-level mesh, satisfy $g_D^J = g_D$ on $\partial_D \Omega$.
-Typically of conforming FE methods, this restricts the Dirichlet boundary function $g_D$ to be piecewise-linear \cite{Elmanetal2014}.
-
-To allow box constraints, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
-(Note that $\tilde{\mathcal{V}}^j$ functions are not in $\mathcal{V}^j$ and not defined everywhere on $\Omega$.)  Suppose that the \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$, for nodes of the finest-level mesh $\mathcal{T}^J$, satisfy
+As is typical of conforming FE methods, this restricts the Dirichlet boundary function $g_D$ to be piecewise-linear \cite{Elmanetal2014}.
+
+To allow box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
+(Note that $\tilde{\mathcal{V}}^j$ functions are not defined everywhere on $\Omega$.)  Suppose that the finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$, for nodes of $\mathcal{T}^J$, satisfy
 \begin{equation}
@@
 \end{equation}
-Where $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$ then there is no lower or upper constraint, respectively.
-Finally assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.
-Then define
-\begin{equation}
-\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \, \text{ and } \, w|_{\partial_D\Omega} = g_D|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
+Assuming that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.  Then define
+\begin{equation}
+\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \text{ on nodes of } \mathcal{T}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
 \end{equation}
@@
 This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.
-As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fe:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J = \pm \infty$.
+As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fe:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J$ are infinite; there is no lower or upper constraint, respectively, at such nodes.
 
 Let $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ be a discretization of $f$ in \eqref{eq:boxdirichletvi}.
-By inclusion \eqref{eq:fe:nestedspaces} the functional $\ell^J=\ell$ is defined and continuous over $\mathcal{V}^J$.
+Note that the functional $\ell^J=\ell$ is defined and continuous over $\mathcal{V}^J$.
 We seek the FE solution $u^J \in \mathcal{K}^J$ of the finest-level, finite-dimensional, nonlinear VI
@@
 While \eqref{eq:fe:vi} approximates \eqref{eq:boxdirichletvi}, it may not technically be conforming for that VI.
-That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require $\mathcal{K}^J \subset \mathcal{K}$.
-Even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants will generally not imply continuum admissiblity.\footnote{Such caveats are relevant in geophysical situations where only nodal values of the obstacle are known and nodal elevation data misses detail in between; see the Greenland ice sheet application in \cite{Bueler2016}, for example.}
-
-Finite-dimensional, box-constrained VI problem \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003,FerrisPang1997}.
-Stating this strong-form problem may be clarifying for some readers.
+That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require or generally expect $\mathcal{K}^J \subset \mathcal{K}$.
+Even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants will generally not imply continuum admissiblity.
+Such caveats are especially relevant in geophysical situations where only nodal values of the obstacle are known and nodal elevation data misses detail in between; see the Greenland ice sheet application in \cite{Bueler2016}, for example.
+
+Finite-dimensional, box-constrained VI problem \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003,FerrisPang1997}, a strong form problem whose statement may be clarifying.
 Denote $x_p$ for a node of the finest-level mesh $\mathcal{T}^J$ and $\psi_p(x)$ for the corresponding hat function.
-Let $X=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap X$.
-If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let us define $I_D$ as the $m_D\times m_J$ matrix which indicates (extracts) Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet values as nodes in $(\mathcal{V}^J)'$.
-Problem \eqref{eq:fe:vi} can be stated as VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap X\right)$, e.g.~in the operations research literature \cite{FerrisPang1997}.
-It is equivalent to the following MCP with additional equality constraints:
+Let $\mathcal{X}=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}$.
+If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let us define $I_D$ as the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.
+Problem \eqref{eq:fe:vi}, which would be stated as VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap X\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP with additional equality constraints:
 \begin{align}
@@
 for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
-If $\underline{\gamma}^J(x_p) < \overline{\gamma}^J(x_p)$ then MCP \eqref{eq:fe:mcp} says that exactly one of the following conditions holds for the solution $u^J$:
+If $\underline{\gamma}^J(x_p) < \overline{\gamma}^J(x_p)$ then MCP \eqref{eq:fe:mcp} says that exactly one of the following five conditions holds for the solution $u^J$:
 \begin{itemize}
 \item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)<u^J(x_p)<\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} = 0$, or
-\item $x_p \notin \partial_D\Omega$ and $u^J(x_p)=\underline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \ge 0$, or
-\item $x_p \notin \partial_D\Omega$ and $u^J(x_p)=\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \le 0$, or
-\item $x_p \in \partial_D\Omega$ and $u^J(x_p)=g_D^J(x_p)$.
+\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)=u^J(x_p)<\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \ge 0$, or
+\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)<u^J(x_p)=\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \le 0$, or
+\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$, or
+\item $x_p \in \partial_D\Omega$ and $u^J(x_p)=g_D^J(x_p)$ and $\lambda_p=\ip{f^J(u^J)-\ell^J}{\psi_p}$.
 \end{itemize}
-(Where the box constraints meet (pinch) at $x_p \notin \partial_D\Omega$ then $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$, and the value $\ip{f^J(u^J)-\ell^J}{\psi_p}$ is unconstrained.)  The four cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, and \emph{boundary}, respectively.
-Where case \emph{boundary} applies, $\lambda_p=\ip{f^J(u^J)-\ell^J}{\psi_p}\in\RR$ has any value.
+The cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, \emph{pinched}, and \emph{boundary}, respectively.
+Where case \emph{pinched} applies $\ip{f^J(u^J)-\ell^J}{\psi_p}$ can have any real value, and similarly if \emph{boundary} applies then $\lambda_p$ is unconstrained.
 
 The multilevel algorithm defined in the next two Sections, a solver for finite-dimensional problem \eqref{eq:fe:vi}, requires transfer operators between meshes.
-We require five such operators, but three are standard multigrid tools \cite{Trottenbergetal2001}.
+We require five such operators, of which three are standard multigrid tools \cite{Trottenbergetal2001}.
 Table \ref{tab:transfers} will help the reader keep track.
@@
 Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ also acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $(R\sigma)[z] = \sigma[z]$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
-However, these canonical transfers involve a change of FE representation.
+However, these canonical transfers generally involve a change of FE representation.
 For example, the formula for canonical restriction, derived by writing coarser-mesh hat functions in terms of finer-mesh ones, is a ``full-weighting'' average \cite{Trottenbergetal2001}.

### 2023-03-24 paper/exact/plap1d.tex
start up exact solution reference document

\documentclass[11pt]{amsart}
%prepared in AMSLaTeX, under LaTeX2e
\addtolength{\oddsidemargin}{-.9in} 
\addtolength{\evensidemargin}{-.9in}
\addtolength{\topmargin}{-.9in}
\addtolength{\textwidth}{1.5in}
\addtolength{\textheight}{1.5in}

\renewcommand{\baselinestretch}{1.05}

\usepackage{verbatim} % for "comment" environment

\usepackage{palatino}

\usepackage[final]{graphicx}

\usepackage{tikz}
\usetikzlibrary{positioning}

\usepackage{enumitem,xspace,fancyvrb}

\newtheorem*{thm}{Theorem}
\newtheorem*{defn}{Definition}
\newtheorem*{example}{Example}
\newtheorem*{problem}{Problem}
\newtheorem*{remark}{Remark}

\DefineVerbatimEnvironment{mVerb}{Verbatim}{numbersep=2mm,frame=lines,framerule=0.1mm,framesep=2mm,xleftmargin=4mm,fontsize=\footnotesize}

% macros
\usepackage{amssymb}
\newcommand{\bA}{\mathbf{A}}
\newcommand{\bB}{\mathbf{B}}
\newcommand{\bE}{\mathbf{E}}
\newcommand{\bF}{\mathbf{F}}
\newcommand{\bJ}{\mathbf{J}}

\newcommand{\bb}{\mathbf{b}}
\newcommand{\bi}{\mathbf{i}}
\newcommand{\bj}{\mathbf{j}}
\newcommand{\bk}{\mathbf{k}}
\newcommand{\br}{\mathbf{r}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\bv}{\mathbf{v}}
\newcommand{\bw}{\mathbf{w}}
\newcommand{\bx}{\mathbf{x}}

\newcommand{\ppr}[1]{\frac{\partial #1}{\partial r}}
\newcommand{\ppt}[1]{\frac{\partial #1}{\partial t}}
\newcommand{\ppx}[1]{\frac{\partial #1}{\partial x}}
\newcommand{\ppy}[1]{\frac{\partial #1}{\partial y}}
\newcommand{\ppz}[1]{\frac{\partial #1}{\partial z}}

\newcommand{\Div}{\ensuremath{\nabla\cdot}}
\newcommand{\Curl}{\ensuremath{\nabla\times}}

\newcommand{\eps}{\epsilon}
\newcommand{\grad}{\nabla}
\newcommand{\ip}[2]{\ensuremath{\left<#1,#2\right>}}
\newcommand{\lam}{\lambda}
\newcommand{\lap}{\triangle}

\newcommand{\RR}{\mathbb{R}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\prob}[1]{\bigskip\noindent\textbf{#1.}\quad }

\newcommand{\Matlab}{\textsc{Matlab}\xspace}

\newcommand{\ds}{\displaystyle}

\begin{document}
\scriptsize \hfill \today

\Large
\bigskip
\centerline{\textbf{Exact solution to a 1D $p$-Laplacian obstacle problem}}
\bigskip

\normalsize

\thispagestyle{empty}

For the purposes of building a nonlinear variational inequality verification case, consider the $p$-Laplacian obstacle problem
\begin{align*}
-\Div(|\grad u|^{p-2} \grad u) &= f(x) \\
u(x) &\ge \psi(x).
\end{align*}
For $\Omega\subset \RR^d$ open, $p>1$, $f \in L^1(\Omega)$, and $\psi \in W^{1,p}(\Omega)$, the weak form of this problem is well-posed over $\mathcal{X} = W_0^{1,p}(\Omega)$ \cite{KinderlehrerStampacchia1980}.  The same problem is to minimize the coercive functional $J[u] = \int_\Omega \frac{1}{p} |\grad u|^p - f u$ over the closed and convex admissible set $\mathcal{K}=\{u \in \mathcal{X}\,:\, u \ge \psi\}$.

To be specific, for $p=1.5,2,4$ we will solve the one-dimensional case
\begin{align}
-\left(|u'(x)|^{p-2} u'(x)\right)' &= f(x) = \begin{cases} \alpha, & |x| < 1 \\ -\alpha, & |x| > 1\end{cases} \label{concrete} \\
u(x) &\ge 0 \notag
\end{align}
on the interval $\Omega = (-3,3) \subset \RR^1$.  Later, $\alpha>0$ will chosen for convenience.

By symmetry and well-posedness, $u(x)$ is an even function and $u'(0)=0$.  We will see that the solution satisfies $u'(x)\le 0$ on $0 \le x \le 3$.  Therefore, integrating \eqref{concrete} once gives
    $$\left(-u'(x)\right)^{p-1} = \int_0^x f(t)\,dt = \begin{cases} \alpha x, & |x| < 1 \\ \alpha (1 - FIXME), & |x| > 1\end{cases}$$
so

\begin{thebibliography}{1}

\bibitem{KinderlehrerStampacchia1980}
{\sc D.~Kinderlehrer and G.~Stampacchia}, {\em An {I}ntroduction to
  {V}ariational {I}nequalities and their {A}pplications}, Academic Press, New
  York, 1980.

\end{thebibliography}
\end{document}

### 2023-03-25 paper/exact/plap1d.tex
note completed

@@
 
-\renewcommand{\baselinestretch}{1.05}
+\renewcommand{\baselinestretch}{1.02}
 
@@
 For the purposes of building a nonlinear variational inequality verification case, consider the $p$-Laplacian obstacle problem
+    $$-\Div(|\grad u|^{p-2} \grad u) = f(x), \qquad u(x) \ge \psi(x)$$
+over $\Omega\subset \RR^d$ open, with $p>1$, $f \in L^1(\Omega)$, and $\psi \in W^{1,p}(\Omega)$.
+The weak form of this problem is well-posed over $\mathcal{X} = W_0^{1,p}(\Omega)$ \cite{KinderlehrerStampacchia1980}.
+Equivalently one may minimize the coercive functional $J[u] = \int_\Omega \frac{1}{p} |\grad u|^p - f u$ over the closed and convex admissible set $\mathcal{K}=\{u \in \mathcal{X}\,:\, u \ge \psi\}$.
+
+We will solve the one-dimensional case
+\begin{equation}
+-\left(|u'(x)|^{p-2} u'(x)\right)' = f(x) = \begin{cases} \alpha, & |x| < 1 \\ -\alpha, & |x| > 1\end{cases} \label{concrete}
+\end{equation}
+on the interval $\Omega = (-3,3) \subset \RR^1$, with obstacle $\psi(x)=0$, for any $\alpha>0$.
+
+By symmetry and well-posedness, $u(x)$ is an even function and $u'(0)=0$.  Restricting attention to $x\ge 0$, we will see that $u'(x)\le 0$.  Integrating \eqref{concrete} once gives
+    $$\left(-u'(x)\right)^{p-1} = \int_0^x f(t)\,dt = \begin{cases} \alpha x, & 0 \le x \le 1 \\ \alpha (2 - x), & x \ge 1\end{cases}$$
+The function on the right is continuous, and nonnegative for $0 \le x \le 2$.  Solving for $u'(x)$ by raising to the $(p-1)^{-1} = q-1$ power, where $\frac{1}{p} + \frac{1}{q} = 1$ defines $q$, gives
+    $$u'(x) = \begin{cases} - \alpha^{q-1} x^{q-1}, & 0 \le x \le 1 \\ -\alpha^{q-1} (2 - x)^{q-1}, & x \ge 1\end{cases}$$
+Integrating again, and recalling that $u(x)$ has even symmetry, we have
 \begin{align*}
--\Div(|\grad u|^{p-2} \grad u) &= f(x) \\
-u(x) &\ge \psi(x).
+u(x) &= \begin{cases} u(0) - \frac{\alpha^{q-1}}{q} |x|^q, & |x| \le 1 \\
+                      u(0) - \frac{\alpha^{q-1}}{q} -\alpha^{q-1} \int_1^{|x|} (2 - y)^{q-1}\,dy, & |x| \ge 1\end{cases} \\
+     &= \begin{cases} u(0) - \frac{\alpha^{q-1}}{q} |x|^q, & |x| \le 1 \\
+                      u(0) + \frac{\alpha^{q-1}}{q} \left( (2 - |x|)^q - 2 \right), & |x| \ge 1\end{cases}
 \end{align*}
-For $\Omega\subset \RR^d$ open, $p>1$, $f \in L^1(\Omega)$, and $\psi \in W^{1,p}(\Omega)$, the weak form of this problem is well-posed over $\mathcal{X} = W_0^{1,p}(\Omega)$ \cite{KinderlehrerStampacchia1980}.
-The same problem is to minimize the coercive functional $J[u] = \int_\Omega \frac{1}{p} |\grad u|^p - f u$ over the closed and convex admissible set $\mathcal{K}=\{u \in \mathcal{X}\,:\, u \ge \psi\}$.
 
-To be specific, for $p=1.5,2,4$ we will solve the one-dimensional case
-\begin{align}
--\left(|u'(x)|^{p-2} u'(x)\right)' &= f(x) = \begin{cases} \alpha, & |x| < 1 \\ -\alpha, & |x| > 1\end{cases} \label{concrete} \\
-u(x) &\ge 0 \notag
-\end{align}
-on the interval $\Omega = (-3,3) \subset \RR^1$.  Later, $\alpha>0$ will chosen for convenience.
+Now, where is the free boundary?
+It must occur at $x=\xi>0$ where simultaneously $u(\xi)=0$ and $u'(\xi)=0$.
+However, $u'(2)=0$, which reflects the symmetry that the integral of $f(x)$ on $[0,1]$ is cancelled by its integral on $[1,2]$, and $u'(x)$ is otherwise nonzero for $|x|>0$.
+Thus the free boundary must be at $\xi=2$.
+This determines $u(0)$:
+    $$0=u(2) = u(0) - 2 \frac{\alpha^{q-1}}{q} \qquad \iff \qquad u(0) = 2 \frac{\alpha^{q-1}}{q}.$$
+In conclusion, our exact solution is
+\begin{equation}
+\boxed{u(x) = \frac{\alpha^{q-1}}{q} \begin{cases} 2 - |x|^q, & |x| \le 1 \\
+                                                   (2 - |x|)^q, & |x| \ge 1.\end{cases}}
+\end{equation}
 
-By symmetry and well-posedness, $u(x)$ is an even function and $u'(0)=0$.  We will see that the solution satisfies $u'(x)\le 0$ on $0 \le x \le 3$.  Therefore, integrating \eqref{concrete} once gives
-    $$\left(-u'(x)\right)^{p-1} = \int_0^x f(t)\,dt = \begin{cases} \alpha x, & |x| < 1 \\ \alpha (1 - FIXME), & |x| > 1\end{cases}$$
-so
+We will numerically test powers $p=3/2,2,4$, corresponding to $q=3,2,4/3$, respectively, with parameter $\alpha=1$ for convenience.
 

### 2023-05-14 paper/fascd.tex
patrick is most definitely an author
move from slack, and edit, a summary of differences to G&K(2009)

@@
 \author{Ed Bueler}
+
+\author{Patrick E.~Farrell}
 
@@
 
+\bigskip
+From slack, to not lose.
+
+\noindent Patrick : \emph{Interesting that you think Graeser and Kornhuber were barking up the wrong tree.
+I still don't quite have my finger on the exact difference in the scheme between GK and the FAS approach.
+Is there a clean summary somewhere?
+(I looked in the mcd-extended manuscript, but couldn't see it in a quick skim.)  Is the difference to GK the monotone multigrid bit, or the particular smoothers used?}
+
+\medskip\noindent
+Ed: \cite{GraeserKornhuber2009} only considers the Laplacian lower obstacle case, and they write everything in terms of nonlinear Gauss-Seidel (i.e.
+optimize one hat function at a time).
+They first lay out the \cite{Tai2003} CD strategy as Alg.
+4.7, only for V(1,0) cycles.
+Then they observe that the convergence rates from \cite{Tai2003} are slower (worse) than if one does "monotone multigrid".
+This seems to mean to them that one modifies the coarse hat functions according to whether they overlap the current fine-grid estimate of the coincidence set, "truncated nodal basis functions".
+(See the start of \cite{GraeserKornhuber2009} section 5.2.)  I think that making these modifications in Firedrake, so we could try monotone multigrid V-cycles, would be a nightmare.
+FASCD does not attempt the monotone multigrid idea at all.
+
+By contrast, FASCD is essentially just the \cite{Tai2003} CD strategy, but (1) without assuming an objective, and (2) with a coarse correction which is FAS in the nonlinear PDE case.
+If you apply \cite{Tai2003} CD to a nonquadratic (e.g.
+p=1.5 p-laplacian) objective then the coarse-grid corrections are optimization problems over the coarse space.
+I don't even know how to interpret this in non-optimization VIs?
+Or maybe I should say that one can re-interpret this as FAS coarse corrections, and I have done so.
+
+But I have now concluded that everything discussed in \cite{GraeserKornhuber2009} is angels-on-the-head-of-a-pin stuff because the various differences of implementation don't produce any really good multigrid convergence if your smoother is NGS on each level.
+I think the convergence proofs in \cite{GraeserKornhuber2009} all depend on this?
+There are indeed differences in convergence rates reported in the results at the end of \cite{GraeserKornhuber2009}, but they are all not good compared to how multigrid works on PDEs.
+This NGS-based stuff is what I was trying before February, in 1d mostly, with hand-made NGS smoothers, and no Firedrake.
+
+Now, the good thing I was already doing before February was building a version of VI multigrid which makes minimal assumptions on the problem:  FAS (residual need not be linear) and CD (build-up the total V-cycle update in an admissible manner on every level).
+Thus it has a chance of working for the ice problems I care about.
+
+As I now see, and can easily demo on the spiral problem in \cite{GraeserKornhuber2009} and on any other classical obstacle problem I have tried, the thing which makes it fly is to realize, as you did, or perhaps as you just take for granted because you are so comfortable with Firedrake, that the smoother can be a grown-up but still O(N) VI solver.
+E.g.~a fixed number of iterations of vinewtonrsls+gmres+ilu or whatever.
+The role of the FASCD is simply to generate coarse VI problems which, once solved, make helpful admissible additions to the eventual V-cycle update.
+
+So the difference between FASCD and \cite{GraeserKornhuber2009} are: (1) FAS coarse corrections, (2) no attempt to do monotone/truncated multigrid which would be a mess (I think) across the FE infrastructure, (3) clarity on what sets you can use in $V(\nu_1,\nu_2)$ cycles, with bigger up-smoother sets in particular, and most importantly (4) implementation within the mature Firedrake/PETSc library framework where one can easily hook up an active-set Newton step as a smoother.
+
 \end{document}

### 2023-05-21 paper/exact/plap1d.tex
expanded family of exact solutions to p-lap obstacle problem

--- 2023-03-24 paper/exact/plap1d.tex
@@
 % macros
-\usepackage{amssymb}
+\usepackage{amssymb,amsmath,bbm}
 \newcommand{\bA}{\mathbf{A}}
@@
 \bigskip
-\centerline{\textbf{Exact solution to a 1D $p$-Laplacian obstacle problem}}
+\centerline{\textbf{A family of exact solutions to the 1D $p$-Laplacian obstacle problem}}
 \bigskip
@@
 
-For the purposes of building a nonlinear variational inequality verification case, consider the $p$-Laplacian obstacle problem
+For the purposes of building a nonlinear variational inequality verification case we will construct exact solutions to the one-dimensional case of the \emph{$p$-Laplacian obstacle problem}.
+Over an open set $\Omega\subset \RR^d$, for $p>1$, $f \in L^1(\Omega)$, and $\psi \in W^{1,p}(\Omega)$, recall that the problem is to find $u(x)$ satisfying
+\begin{equation}
+-\Div\left(|\grad u(x)|^{p-2} \grad u(x)\right) = f(x) \qquad \text{and} \qquad u(x) \ge \psi(x). \label{plap}
+\end{equation}
+
+The weak form of this problem is well-posed with Dirichlet boundary conditions, and in particular there is a unique solution $u \in \mathcal{X} = W_0^{1,p}(\Omega)$ \cite{KinderlehrerStampacchia1980}.
+Equivalently, the coercive functional $J[u] = \int_\Omega \frac{1}{p} |\grad u|^p - f u$ has a unique minimum over $\mathcal{K}=\{v \ge \psi\} \subset \mathcal{X}$, a closed and convex subset.
+Observe that if $p=2$ then the operator is the ordinary Laplacian, if $1<p<2$ then $D=|\grad u|^{p-2}>0$ is unbounded as $|\grad u| \to 0$ (\emph{fast diffusion}), while if $p>2$ then $D\to 0$ as $|\grad u| \to 0$ (\emph{degenerate diffusion}).
+The degenerate case violates uniform ellipticity, despite coercivity.
+
+We will solve a one-dimensional case of \eqref{plap}, with
+\begin{equation}
+f(x) = \begin{cases} \alpha, & |x| < 1 \\ -\alpha, & |x| > 1\end{cases} \qquad \text{and} \qquad \psi(x) = - \beta |x| \label{data}
+\end{equation}
+on the interval $\Omega = (-3,3) \subset \RR^1$, for $\alpha > 0$ and $\beta \ge 0$.
+% FIXME  I think the cases where \beta < 0$ are fine too, though eventually the free boundary will be outside $(-3,3)$
+That is, we seek $u(x) \ge -\beta |x|$ so that
+\begin{equation}
+-\left(|u'(x)|^{p-2} u'(x)\right)' = \alpha \mathbbm{1}_{(-1,1)}(x) - 2 \mathbbm{1}_{(-3,-1)\cup(1,3)}(x).  \label{concrete}
+\end{equation}
+
+By symmetry and well-posedness, $u(x)$ is an even function and $u'(0)=0$.  Restricting attention to $x\ge 0$, we will see that $u'(x)\le 0$ for $x\le 2$.  Integrating \eqref{concrete} once gives
+    $$\left(-u'(x)\right)^{p-1} = \int_0^x f(t)\,dt = \begin{cases} \alpha x, & 0 \le x \le 1 \\ \alpha (2 - x), & x \ge 1\end{cases}$$
+The function on the right is continuous.  Solving for $u'(x)$ by raising to the $(p-1)^{-1} = q-1$ power, where $\frac{1}{p} + \frac{1}{q} = 1$ defines the dual exponent $q>1$, gives
+    $$u'(x) = \begin{cases} - \alpha^{q-1} x^{q-1}, & 0 \le x \le 1 \\ -\alpha^{q-1} (2 - x)^{q-1}, & x \ge 1\end{cases}$$
+Note that $u''(x) < 0$ (concave) for $0 < x < 1$ and $u''(x)>0$ (convex) for $x>1$.
+
+Integrating again, and recalling that $u(x)$ has even symmetry, for some $u(0)$ to be determined, we have
 \begin{align*}
--\Div(|\grad u|^{p-2} \grad u) &= f(x) \\
-u(x) &\ge \psi(x).
+u(x) &= \begin{cases} u(0) - \frac{\alpha^{q-1}}{q} |x|^q, & |x| \le 1 \\
+                      u(0) - \frac{\alpha^{q-1}}{q} -\alpha^{q-1} \int_1^{|x|} (2 - y)^{q-1}\,dy, & |x| \ge 1\end{cases} \\
+     &= \begin{cases} u(0) - \frac{\alpha^{q-1}}{q} |x|^q, & |x| \le 1 \\
+                      u(0) + \frac{\alpha^{q-1}}{q} \left( (2 - |x|)^q - 2 \right), & |x| \ge 1\end{cases}
 \end{align*}
-For $\Omega\subset \RR^d$ open, $p>1$, $f \in L^1(\Omega)$, and $\psi \in W^{1,p}(\Omega)$, the weak form of this problem is well-posed over $\mathcal{X} = W_0^{1,p}(\Omega)$ \cite{KinderlehrerStampacchia1980}.
-The same problem is to minimize the coercive functional $J[u] = \int_\Omega \frac{1}{p} |\grad u|^p - f u$ over the closed and convex admissible set $\mathcal{K}=\{u \in \mathcal{X}\,:\, u \ge \psi\}$.
 
-To be specific, for $p=1.5,2,4$ we will solve the one-dimensional case
-\begin{align}
--\left(|u'(x)|^{p-2} u'(x)\right)' &= f(x) = \begin{cases} \alpha, & |x| < 1 \\ -\alpha, & |x| > 1\end{cases} \label{concrete} \\
-u(x) &\ge 0 \notag
-\end{align}
-on the interval $\Omega = (-3,3) \subset \RR^1$.  Later, $\alpha>0$ will chosen for convenience.
+Now, where is the free boundary?
+It must occur at $x=\xi>0$ where simultaneously $u(\xi)=-\beta \xi$ and $u'(\xi)=-\beta\le 0$.
+However, $u'(2)=0$, which reflects the symmetry that the integral of $f(x)$ on $[0,1]$ is cancelled by its integral on $[1,2]$.
+Furthermore $u'(x)<0$ for $0<x<2$ and $u'(x)$ is minimum at $x=1$, with $u'(1) = -\alpha^{q-1}$.
+By a convexity/concavity argument, existence of the free boundary requires $\alpha^{q-1}/q>\beta$, which implies $\beta^{p-1}/\alpha < 1$ (see below).
+Then, because $\beta \ge 0$, tangency implies that the free boundary is at some $1 < \xi \le 2$, and the above calculation is justified.
 
-By symmetry and well-posedness, $u(x)$ is an even function and $u'(0)=0$.  We will see that the solution satisfies $u'(x)\le 0$ on $0 \le x \le 3$.  Therefore, integrating \eqref{concrete} once gives
-    $$\left(-u'(x)\right)^{p-1} = \int_0^x f(t)\,dt = \begin{cases} \alpha x, & |x| < 1 \\ \alpha (1 - FIXME), & |x| > 1\end{cases}$$
-so
+Thus, given $\alpha>0$ and $0 \le \beta < \alpha^{q-1}/q$, we solve the system $u'(\xi)=-\beta$ and $u(\xi)=-\beta \xi$:
+\begin{align*}
+-\alpha^{q-1} (2 - \xi)^{q-1} &= -\beta \\
+u(0) + \frac{\alpha^{q-1}}{q} \left( (2 - \xi)^q - 2 \right) &= -\beta \xi
+\end{align*}
+The first equation determines $\xi$ and then the second will yield $u(0)$.  If $\beta=0$ then the free boundary is at $\xi=2$ and thus $u(0) = 2 \alpha^{q-1} / q$.  Generally
+\begin{equation}
+\boxed{\xi = 2 - \frac{\beta^{p-1}}{\alpha} \qquad \text{and} \qquad u(0) = \frac{\beta^p}{p \alpha} + 2 \left(\frac{\alpha^{q-1}}{q} - \beta\right).} \label{params}
+\end{equation}
+
+In conclusion, our exact solution family starts with a choice of $p>1$ and $\alpha>0$ arbitrary.
+Let $q>1$ be the dual exponent and choose $\beta \in [0,\alpha^{q-1}/q)$.
+Compute $\xi \in (1,2]$ and $u(0)>0$ by \eqref{params}.
+Then
+\begin{equation}
+\boxed{u(x) = \begin{cases} u(0) - \frac{\alpha^{q-1}}{q} |x|^q, & |x| \le 1 \\
+                      u(0) + \frac{\alpha^{q-1}}{q} \left( (2 - |x|)^q - 2 \right), & 1 \le |x| \le \xi \\
+                      -\beta |x|, & \xi \le |x| \le 3\end{cases}}  \label{exact}
+\end{equation}
+This function is in $W^{1,p}(\Omega) \cap W^{2,\infty}(\Omega)$, where $\Omega = (-3,3)$, and specifically it is continuously-differentiable.  It satisfies \eqref{plap} strongly a.e.
+
+We will set parameters $\alpha=1$ and $\beta = 0.2$ for convenience, and numerically test powers $p=3/2,2,4$, corresponding to $q=3,2,4/3$, respectively.
+Note that for our chosen parameters $\alpha,\beta$, the requirement $\beta < \alpha^{q-1}/q$ is satisfied in the range $p \in [3/2,4]$.
 

### 2023-05-21 paper/exact/plap1d.tex
include solution figures, and improve text

@@
 For the purposes of building a nonlinear variational inequality verification case we will construct exact solutions to the one-dimensional case of the \emph{$p$-Laplacian obstacle problem}.
-Over an open set $\Omega\subset \RR^d$, for $p>1$, $f \in L^1(\Omega)$, and $\psi \in W^{1,p}(\Omega)$, recall that the problem is to find $u(x)$ satisfying
+Over an open set $\Omega\subset \RR^d$, for $p>1$, $f \in L^1(\Omega)$, and $\psi \in W^{1,p}(\Omega)$, recall that the problem is to find $u(x) \ge \psi(x)$ satisfying
 \begin{equation}
--\Div\left(|\grad u(x)|^{p-2} \grad u(x)\right) = f(x) \qquad \text{and} \qquad u(x) \ge \psi(x). \label{plap}
+-\Div\left(|\grad u(x)|^{p-2} \grad u(x)\right) = f(x) \qquad \text{on the set where } u(x) > \psi(x). \label{plap}
 \end{equation}
 
-The weak form of this problem is well-posed with Dirichlet boundary conditions, and in particular there is a unique solution $u \in \mathcal{X} = W_0^{1,p}(\Omega)$ \cite{KinderlehrerStampacchia1980}.
-Equivalently, the coercive functional $J[u] = \int_\Omega \frac{1}{p} |\grad u|^p - f u$ has a unique minimum over $\mathcal{K}=\{v \ge \psi\} \subset \mathcal{X}$, a closed and convex subset.
-Observe that if $p=2$ then the operator is the ordinary Laplacian, if $1<p<2$ then $D=|\grad u|^{p-2}>0$ is unbounded as $|\grad u| \to 0$ (\emph{fast diffusion}), while if $p>2$ then $D\to 0$ as $|\grad u| \to 0$ (\emph{degenerate diffusion}).
+The weak form of this problem is well-posed with Dirichlet boundary conditions, and in particular there is a unique solution $u \in \mathcal{K}=\{v \ge \psi\} \subset W_0^{1,p}(\Omega)$ \cite{KinderlehrerStampacchia1980}.
+Equivalently, the coercive functional $J[u] = \int_\Omega \frac{1}{p} |\grad u|^p - f u$ has a unique minimum over $\mathcal{K}$, a closed and convex subset.
+Observe that if $p=2$ then the operator is the ordinary Laplacian.
+If $1<p<2$ then $D=|\grad u|^{p-2}>0$ is unbounded as $|\grad u| \to 0$ (\emph{fast diffusion}), while if $p>2$ then $D\to 0$ as $|\grad u| \to 0$ (\emph{degenerate diffusion}).
 The degenerate case violates uniform ellipticity, despite coercivity.
@@
 \begin{equation}
--\left(|u'(x)|^{p-2} u'(x)\right)' = \alpha \mathbbm{1}_{(-1,1)}(x) - 2 \mathbbm{1}_{(-3,-1)\cup(1,3)}(x).  \label{concrete}
+-\left(|u'(x)|^{p-2} u'(x)\right)' = \alpha \left(\mathbbm{1}_{(-1,1)}(x) - \mathbbm{1}_{(-3,-1)\cup(1,3)}(x)\right).  \label{concrete}
 \end{equation}
@@
 
-We will set parameters $\alpha=1$ and $\beta = 0.2$ for convenience, and numerically test powers $p=3/2,2,4$, corresponding to $q=3,2,4/3$, respectively.
-Note that for our chosen parameters $\alpha,\beta$, the requirement $\beta < \alpha^{q-1}/q$ is satisfied in the range $p \in [3/2,4]$.
+The $\beta=0$ case, in which the obstacle $\psi(x)=0$ is flat, implies that the solution has zero gradient at the free boundary ($u'(\xi)=0$).
+Note that this implies a loss of uniform ellipticity at the free boundary in the $p>2$ degenerate diffusion cases, which makes the obstacle problem very challenging numerically, so we avoid that difficulty by setting $\beta>0$.
+
+For numerical tests we set parameters $\alpha=1$ and $\beta = 0.2$, and consider powers $p=3/2,2,4$, corresponding to $q=3,2,4/3$ respectively.
+(For these $\alpha,\beta$ values, the requirement $\beta < \alpha^{q-1}/q$ is satisfied when $p > 5/4$.)  Numerical and exact solutions for the $p=3/2$ and $p=4$ cases are shown in Figure \ref{fig:solutions}.
+
+\begin{figure}
+\mbox{\includegraphics[width=0.45\textwidth]{solution1p5.pdf} \qquad \includegraphics[width=0.45\textwidth]{solution4.pdf}}
+\caption{Low-resolution $p$-Laplacian obstacle problem solutions for $\alpha=1$, $\beta=0.2$, and $p=1.5$ (left) or $p=4$ (right).}
+\label{fig:solutions}
+\end{figure}
 

### 2023-05-22 paper/exact/plap1d.tex
tidy up again

@@
 
-For the purposes of building a nonlinear variational inequality verification case we will construct exact solutions to the one-dimensional case of the \emph{$p$-Laplacian obstacle problem}.
+For the purposes of building a nonlinear variational inequality verification case we will construct exact solutions to the one-dimensional \emph{$p$-Laplacian obstacle problem}.
 Over an open set $\Omega\subset \RR^d$, for $p>1$, $f \in L^1(\Omega)$, and $\psi \in W^{1,p}(\Omega)$, recall that the problem is to find $u(x) \ge \psi(x)$ satisfying
@@
 
-Integrating again, and recalling that $u(x)$ has even symmetry, for some $u(0)$ to be determined, we have
-\begin{align*}
-u(x) &= \begin{cases} u(0) - \frac{\alpha^{q-1}}{q} |x|^q, & |x| \le 1 \\
-                      u(0) - \frac{\alpha^{q-1}}{q} -\alpha^{q-1} \int_1^{|x|} (2 - y)^{q-1}\,dy, & |x| \ge 1\end{cases} \\
-     &= \begin{cases} u(0) - \frac{\alpha^{q-1}}{q} |x|^q, & |x| \le 1 \\
+Integrating again, and recalling that $u(x)$ has even symmetry, then on $\Omega=(-3,3)$,
+\begin{equation*}
+u(x) = \begin{cases} u(0) - \frac{\alpha^{q-1}}{q} |x|^q, & |x| \le 1 \\
                       u(0) + \frac{\alpha^{q-1}}{q} \left( (2 - |x|)^q - 2 \right), & |x| \ge 1\end{cases}
-\end{align*}
+\end{equation*}
+for some $u(0)$ to be determined.
 
-Now, where is the free boundary?
-It must occur at $x=\xi>0$ where simultaneously $u(\xi)=-\beta \xi$ and $u'(\xi)=-\beta\le 0$.
-However, $u'(2)=0$, which reflects the symmetry that the integral of $f(x)$ on $[0,1]$ is cancelled by its integral on $[1,2]$.
+Now, where are the free boundaries?
+One must occur at $x=\xi>0$ where simultaneously $u(\xi)=-\beta \xi$ and $u'(\xi)=-\beta\le 0$.
+However, $u'(2)=0$ because $\int_0^2 f(x)\,dx = 0$.
 Furthermore $u'(x)<0$ for $0<x<2$ and $u'(x)$ is minimum at $x=1$, with $u'(1) = -\alpha^{q-1}$.
-By a convexity/concavity argument, existence of the free boundary requires $\alpha^{q-1}/q>\beta$, which implies $\beta^{p-1}/\alpha < 1$ (see below).
-Then, because $\beta \ge 0$, tangency implies that the free boundary is at some $1 < \xi \le 2$, and the above calculation is justified.
+By a convexity/concavity argument, existence of the free boundary requires $\alpha^{q-1}/q>\beta$.
+(Note this implies $\beta^{p-1} < \alpha$; in \eqref{params} below this shows $\xi>1$.)  Then, because $\beta \ge 0$, tangency implies that the free boundary is at some $1 < \xi \le 2$, and the above calculation is justified.
 
@@
 \end{align*}
-The first equation determines $\xi$ and then the second will yield $u(0)$.  If $\beta=0$ then the free boundary is at $\xi=2$ and thus $u(0) = 2 \alpha^{q-1} / q$.  Generally
+These equations determine $\xi$ and $u(0)$ in turn:
 \begin{equation}
@@
 
-In conclusion, our exact solution family starts with a choice of $p>1$ and $\alpha>0$ arbitrary.
+In conclusion, the solution family allows arbitrary choice of $p>1$ and $\alpha>0$.
 Let $q>1$ be the dual exponent and choose $\beta \in [0,\alpha^{q-1}/q)$.
-Compute $\xi \in (1,2]$ and $u(0)>0$ by \eqref{params}.
+From \eqref{params} compute $\xi \in (1,2]$ and $u(0)>0$.
 Then
@@
 \end{equation}
-This function is in $W^{1,p}(\Omega) \cap W^{2,\infty}(\Omega)$, where $\Omega = (-3,3)$, and specifically it is continuously-differentiable.  It satisfies \eqref{plap} strongly a.e.
+This function is in $W^{1,p}(\Omega) \cap W^{2,\infty}(\Omega)$, $\Omega = (-3,3)$, and specifically it is continuously-differentiable.  It satisfies \eqref{plap} strongly a.e.
 
-The $\beta=0$ case, in which the obstacle $\psi(x)=0$ is flat, implies that the solution has zero gradient at the free boundary ($u'(\xi)=0$).
-Note that this implies a loss of uniform ellipticity at the free boundary in the $p>2$ degenerate diffusion cases, which makes the obstacle problem very challenging numerically, so we avoid that difficulty by setting $\beta>0$.
+The $\beta=0$ case, in which the obstacle $\psi(x)=0$ is flat and $\xi=2$, implies that the solution has zero gradient at the free boundary ($u'(\xi)=0$).
+This implies a loss of uniform ellipticity \emph{at} the free boundary in the $p>2$ degenerate diffusion cases, which makes the obstacle problem very challenging numerically.
+We avoid the difficulty by only considering $\beta>0$.
 
 For numerical tests we set parameters $\alpha=1$ and $\beta = 0.2$, and consider powers $p=3/2,2,4$, corresponding to $q=3,2,4/3$ respectively.
-(For these $\alpha,\beta$ values, the requirement $\beta < \alpha^{q-1}/q$ is satisfied when $p > 5/4$.)  Numerical and exact solutions for the $p=3/2$ and $p=4$ cases are shown in Figure \ref{fig:solutions}.
+(For these $\alpha,\beta$ values, the requirement $\beta < \alpha^{q-1}/q$ is satisfied when $p > 5/4$.)  Solutions for $p=3/2,4$ are shown in Figure \ref{fig:solutions}.
 
 \begin{figure}
-\mbox{\includegraphics[width=0.45\textwidth]{solution1p5.pdf} \qquad \includegraphics[width=0.45\textwidth]{solution4.pdf}}
-\caption{Low-resolution $p$-Laplacian obstacle problem solutions for $\alpha=1$, $\beta=0.2$, and $p=1.5$ (left) or $p=4$ (right).}
+\mbox{\includegraphics[height=0.35\textwidth]{figs/solution1p5.pdf} \qquad \includegraphics[height=0.35\textwidth]{figs/solution4.pdf}}
+\caption{$p$-Laplacian obstacle problem solutions for $\alpha=1$ and $\beta=0.2$, with $p=1.5$ (left) and $p=4$ (right).}
 \label{fig:solutions}

### 2023-06-01 paper/fascd.tex
edits sections 1 and 2

@@
 
-\tableofcontents
+%\tableofcontents
 
@@
 We extend Tai's multilevel CD method by implementing fully-nonlinear corrections, and we apply it to a new set of nonlinear VI problems which do not arise from minimization.
-Specifically, for finite element (FE) solvers the algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} are extended here in three practical ways:
+Specifically, for finite element (FE) solvers the algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} are extended here in the following directions:
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
@@
 \item Our \pr{fascd} implementation permits ``box'' constraints, namely upper and lower obstacles; see Section \ref{sec:femultilevel}.
-\item In Sections \ref{sec:ldcs}--\ref{sec:vcycle} we explain why, given the manner in which constraint sets are decomposed in the multilevel context \cite{GraeserKornhuber2009}, ``up-smoothing'' is more efficient than ``down-smoothing''.
+\item We observe that ``up-smoothing'' is more effective than ``down-smoothing'' in for this class of algorithms.
 A strong preference for up-smoothing is special to multilevel CD methods and does not arise in unconstrained problems.
+\item By using strong, yet $O(N)$-per-application, smoothers, namely Newton-type active set methods with a fixed number of preconditioned Krylov iterations per step, in Section \ref{sec:results} we demonstrate mesh-independent convergence rates substantially better than in the current literature.
 \end{enumerate}
@@
 
-\section{Coercive variational inequalities} \label{sec:vi}
+\section{Variational inequalities} \label{sec:vi}
 
@@
 As is standard in the calculus of variations \cite{Evans2010}, coercivity permits a compactness argument for unbounded sets $\cK$.
-(Recall that the closed and bounded subsets of a reflexive Banach space are weakly compact.)  The condition of continuity can be weakened, e.g.~to only apply on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but it will apply in our examples.
-
-The coercive VIs solved in this paper satisfy a stronger inequality than \eqref{eq:coercive}.
+(Recall that the closed and bounded subsets of a reflexive Banach space are weakly compact.)  The condition of continuity can be weakened, e.g.~to only apply on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but this is unnecessary in our examples.
+
+Certain of the VIs solved in this paper satisfy a stronger inequality than \eqref{eq:coercive}.
 
@@
 
-\begin{example}  \label{ex:plaplacian}  Suppose $a\in L^\infty(\Omega)$ such that $a(x)\ge a_0$ a.e.~for some constant $a_0>0$, and $p\ge 2$.
-For $u,v \in \cV = W^{1,p}_0(\Omega)$ define $f:\cV \to \cV'$ by
-\begin{equation}
-\ip{f(u)}{v} = \int_\Omega a(x) |\grad u|^{p-2} \grad u \cdot \grad v\,dx, \label{eq:plaplacian}
+\begin{example}  \label{ex:plaplacian}  Suppose $p>1$.  For $u,v \in \cV = W^{1,p}_0(\Omega)$ define $f:\cV \to \cV'$ by
+\begin{equation}
+\ip{f(u)}{v} = \int_\Omega |\grad u|^{p-2} \grad u \cdot \grad v\,dx, \label{eq:plaplacian}
 \end{equation}
 a continuous map \cite[Theorem A.0.6]{Peral1997}.
-Now, if $x,y\in\RR^d$ then $(|x|^{p-2} x - |y|^{p-2} y)\cdot (x-y) \ge 2^{2-p} |x-y|^p$ \cite[Appendix A]{Bueler2021conservation}.
-It follows from the Poincar\'e inequality that
+For $p\ge 2$, if $x,y\in\RR^d$ then $(|x|^{p-2} x - |y|^{p-2} y)\cdot (x-y) \ge 2^{2-p} |x-y|^p$ \cite[Appendix A]{Bueler2021conservation}, hence it follows from the Poincar\'e inequality that
     $$\ip{f(u) - f(v)}{u-v} \ge 2^{2-p} a_0 \|\grad u - \grad v\|_p^p \ge 2^{2-p} a_0 C \|u-v\|^p$$
 for some $C>0$, and thus $f$ is $p$-coercive.
-The map in \eqref{eq:plaplacian} is also coercive if $1<p<2$, but the proof is somewhat different \cite[Theorem 4.4]{Bueler2021conservation}, and our results (Section \ref{sec:results}) only use $p\ge 2$.
+The map \eqref{eq:plaplacian} is also coercive if $1<p<2$, but the proof is somewhat different \cite[Theorem 4.4, for example]{Bueler2021conservation}.
 \end{example}
@@
 For monotone operators $f$, VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
-Suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cK \to \cV'$.
+In fact, suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cK \to \cV'$.
 Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
@@
 The CD methods of Tai \cite{Tai2003} address problem \eqref{eq:minimization}; the analysis in \cite{Tai2003} supposes that an objective $F$ exists with $F'$ 2-coercive.
-
-Not all VI problems arise from optimization, but for Example \ref{ex:plaplacian} we may define
-\begin{equation}
-F(v) = \frac{1}{p} \int_\Omega a(x) |\grad v|^p\,dx. \label{eq:plaplacianobjective}
+For Example \ref{ex:plaplacian} we may define
+\begin{equation}
+F(v) = \frac{1}{p} \int_\Omega |\grad v|^p\,dx. \label{eq:plaplacianobjective}
 \end{equation}
@@
 
-Next we give two examples which are \emph{not} of optimization type, first a linear advection-diffusion problem, which needs the following Lemma, and then a nonlinear problem of porous-medium type.
+Next we give two examples which are \emph{not} of optimization type \eqref{eq:minimization}, the first being an advection-diffusion problem, which needs the following Lemma, and the second a nonlinear problem for ice sheet geometry.
 
@@
 Consider a divergence-free velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, and assume that the flow is outward on the Neumann boundary: $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.
-For $u,v \in \cV$ and $\eps>0$ define
-\begin{equation}
-\ip{f(u)}{v} = \eps \left(\grad u, \grad v\right)_{L^2(\Omega)} - b(u,v). \label{eq:advectiondiffusion}
-\end{equation}
-Consider VI \eqref{eq:vi} for any closed and convex $\cK \subset \cV$ and $g\in\cV'$.
-It is easy to see that $|\ip{f(u)}{v}| \le (\eps + \|\bX\|_\infty) \|u\| \|v\|$, thus that $f:\cK \to \cV'$ is continuous.
-Lemma \ref{lem:advectionskew} says that the bilinear form $s(u,v)$ is skew-symmetric up to a nonnegative term.
+For $u,v \in \cV$, $\eps>0$, and $\phi \in L^\infty(\Omega)$, define
+\begin{equation}
+\ip{f(u)}{v} = \eps \left(\grad u, \grad v\right)_{L^2(\Omega)} + b(u,v) - \ip{\phi}{v}. \label{eq:advectiondiffusion}
+\end{equation}
+With this operator, consider VI \eqref{eq:vi} for any closed and convex $\cK \subset \cV$ and $g\in\cV'$.
+The interior condition, i.e.~PDE strong form, is the following linear advection-diffusion equation
+\begin{equation}
+-\eps \grad^2 u + \bX\cdot \grad u = \phi.
+\label{eq:pdeadvectiondiffusion}
+\end{equation}
+On the other hand, it is easy to see that $|\ip{f(u)}{v}| \le (\eps + \|\bX\|_\infty + \|\phi\|_\infty) \|u\| \|v\|$, thus that $f:\cK \to \cV'$ is continuous.
+Lemma \ref{lem:advectionskew} says that the bilinear form $b(u,v)$ is skew-symmetric up to a nonnegative term.
 By the outward flow assumption and the Poincar\'e inequality,
@@
 
-If $\bX \ne 0$ then $f$ in Example \ref{ex:advectiondiffusion} does not correspond to a minimization problem.
-Indeed, $\ip{f(u)}{v}$ is not symmetric in that case, so $f$ cannot be the gradient of a scalar objective.
+If $\bX \ne 0$ then $f$ in Example \ref{ex:advectiondiffusion} does not correspond to a minimization problem because $\ip{f(u)}{v}$ does not possess the symmetry of the gradient of a scalar objective.
 (For $u,v$ which are zero on $\partial \Omega$, note that $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.)  References \cite{Bueler2021conservation,ChangNakshatrala2017} consider advection-diffusion VI problems like this, specifically over the set $\cK = \{v\ge 0\}$.
+
+FIXME replace following with SIA
 

### 2023-06-02 paper/fascd.tex
edits and simplifications through end section 4

@@
 
-If there are $m<\infty$ closed subspaces $\cV_i \subset \cV$ so that the sum
+Suppose that there are $m<\infty$ closed subspaces $\cV_i \subset \cV$, so that the sum
 \begin{equation}
@@
 \end{equation}
-holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$, then we say \eqref{eq:subspacedecomp} is a \emph{space decomposition} of $\cV$ \cite{Xu1992}.
-Suppose further that $\cK_i \subset \cV_i$ are nonempty, closed, and convex subsets such that
-\begin{equation}
-\cK = \sum_{i=0}^{m-1} \cK_i \label{eq:constraintdecomp}
-\end{equation}
-for a closed, convex subset $\cK \subset \cV$.
+holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$.
+We say \eqref{eq:subspacedecomp} is a \emph{space decomposition} of $\cV$ \cite{Xu1992}.
+For a closed, convex subset $\cK \subset \cV$, suppose further that $\cK_i \subset \cV_i$ are nonempty, closed, and convex subsets such that
+\begin{equation}
+\cK = \sum_{i=0}^{m-1} \cK_i. \label{eq:constraintdecomp}
+\end{equation}
 The sum in \eqref{eq:constraintdecomp} must hold in two senses \cite{TaiTseng2002}: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.
-(Sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp} because the $\cV_i$ are subspaces.)  Finally, suppose there are bounded, generally-nonlinear \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.
+Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique, and also notice that sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp} because the $\cV_i$ are subspaces.
+Finally, suppose there are bounded \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.
 Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:femultilevel}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
@@
 \end{equation}
+The operators $\Pi_i$ are generally nonlinear.
 Clearly \eqref{eq:constraintrestrictionsum} implies sense \emph{(i)} for \eqref{eq:constraintdecomp}.
-A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} \cite{Tai2003}.
-Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique.
-Figure \ref{fig:cartoon} suggests how a low-dimensional example might look; as is typical, note $\cK_i \not\subset \cK$.
+Tai \cite{Tai2003} defined a \emph{constraint decomposition} (CD) of $\cK$ to be a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}.
+Figure \ref{fig:cartoon} suggests how a low-dimensional example might look; note $\cK_i \not\subset \cK$ in this, and in most nontrivial cases as well.
 
@@
 \includegraphics[width=0.6\textwidth]{genfigs/cartoon.pdf}
-\caption{A constraint decomposition (CD) for a unilateral obstacle problem with $\Omega=\{x_1,x_2\}$, $\mathcal{V}=\left\{\text{real functions on } \Omega\right\}$, and $\mathcal{K}=\{v\ge \psi\}$.}
+\caption{A constraint decomposition (CD) for a unilateral obstacle problem on a two-point space $\Omega=\{x_1,x_2\}$, with $\mathcal{V}=\left\{v \,:\, \Omega \to \RR\right\}$ and $\mathcal{K}=\{v\ge \psi\}$.}
 \label{fig:cartoon}
@@
 In Sections \ref{sec:femultilevel}--\ref{sec:vcycle} we will introduce multilevel CD discretizations and algorithms for finite-dimensional VI problems.
-However, one can observe here that the CD concept applies at the level of the continuum problem, as illustrated in the following two examples.
+However, one can observe here that the CD concept applies at the level of the continuum problem, as illustrated in the following two examples, the first of which is an overlapping domain decomposition of a Sobolev space.
 
 \begin{example}  \label{ex:domaindecomposition}  For a bounded domain $\Omega \subset \RR^d$, let $\cV = W_0^{k,p}(\Omega)$ for $k\ge 0$ and $p\ge 1$.
-Suppose an obstacle $\psi \in W^{k,p}(\Omega)$ satisfies $\psi|_{\partial \Omega} \le 0$, and let $\cK = \{v \ge \psi\} \subset \cV$.
+Let $\psi$ be a measurable, real-valued obstacle, defined on $\Omega$, which satisfies $\psi|_{\partial \Omega} \le 0$, and let $\cK = \{v \ge \psi\} \subset \cV$.
 Suppose further that $\{\phi_i\}_{i=0}^{m-1}$ is a smooth partition of unity on $\Omega$, satisfying $0 \le \phi_i\le 1$ and $\sum_i \phi_i = 1$, and let $\Omega_i$ be the open support of $\phi_i$.
@@
 
-Our second example is a disjoint frequency decomposition.
-A multilevel CD over a finite-dimensional FE discretization, proposed in Sections \ref{sec:femultilevel}--\ref{sec:vcycle}, approximates such a frequency decomposition.
-
-\begin{example}  \label{ex:frequencydecomposition}  For simplicity suppose $\Omega = (0,a)^d \subset \RR^d$ is a cube, and let $\cV = W_{\text{per}}^{p,k}(\Omega)$, for $p\ge 2$ and $k\ge 0$, be the periodic functions.
-Suppose $\psi \in W_{\text{per}}^{p,k}(\Omega)$ and let $\cK = \{v \ge \psi\} \subset \cV$.
-Without using any detailed notation for Fourier representation, but noting that the frequencies are discrete because $\Omega$ is compact, suppose $\{\cV_i\}$ are $m<\infty$ subspaces of $\cV$ defined by an (nonoverlapping) partition of $W_{\text{per}}^{p,k}(\Omega)$ by frequency, thus satisfying \eqref{eq:subspacedecomp} as an orthogonal decomposition.
+Our second example is a disjoint frequency decomposition of a Hilbert space.
+A multilevel CD over a finite-dimensional FE discretization, as proposed in Sections \ref{sec:femultilevel}--\ref{sec:cdmultilevel}, approximates such a decomposition.
+
+\begin{example}  \label{ex:frequencydecomposition}  For simplicity suppose $\Omega = (0,a)^d \subset \RR^d$ is a cube, and let $\cV = W_{\text{per}}^{2,k}(\Omega)$ be the periodic functions with square-integrable $k$th derivatives.
+Suppose $\psi \in W_{\text{per}}^{2,k}(\Omega)$ and let $\cK = \{v \ge \psi\} \subset \cV$.
+Without using any detailed notation for Fourier representation, but noting that the frequencies are discrete because $\Omega$ is compact, suppose $\{\cV_i\}$ are $m<\infty$ subspaces of $\cV$ defined by an (nonoverlapping) partition of $W_{\text{per}}^{2,k}(\Omega)$ by frequency, thus satisfying \eqref{eq:subspacedecomp} as an orthogonal decomposition.
 Suppose $P_i:\cV \to \cV_i$ are the corresponding orthogonal projections, satisfying $I = \sum_i P_i$.
@@
 
-Associated to a constraint decomposition (CD) are certain iterative methods.
-Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below solve smaller VI problems over each subset $\cK_i$, to replace $u \in \cK$ with $w\in\cK$ which should be closer to $u^* \in \cK$ solving \eqref{eq:vi}.
-First is \pr{cd-add}, the additive or parallel method, which generalizes the classical Jacobi iteration.
-Here ``\textbf{for all}'' can be computed in any order.
-Note that the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$, and then adds-back the subset solution $\hat w_i \in \mathcal{K}_i$.
-Second is \pr{cd-mult}, the multiplicative or successive version, which generalizes the Gauss-Seidel iteration.
-Each subset solution corrects the global iterate immediately, so here ordering is important.
-The sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``parts of the input $u$ which have not yet improved''.
+As is true for space decompositions \cite{Xu1992}, associated to a constraint decomposition (CD) are certain iterative methods \cite{Tai2003}.
+Let $0\le \alpha \le 1$ be the \emph{damping parameter}.
+Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below, which we now discuss, solve smaller VI problems over each subset $\cK_i$.
+These Algorithms replace $u \in \cK$ with a new iterate $w\in\cK$ which should be closer to $u^* \in \cK$ solving \eqref{eq:vi}.
 
@@
         \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$, \\+
-            $\boxed{\ip{f(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}}$ \\--
+            $\ip{f(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}$ \\--
     $\hat w = \sum_i \hat w_i\in\cK$ \\
@@
 
-\begin{pseudofloat}
+The additive/parallel \pr{cd-add} method (Algorithm \ref{alg:basiccd-add}) generalizes the classical Jacobi iteration.
+Here ``\textbf{for all}'' can be computed in any order.
+In the VI over $\mathcal{K}_i$ solved here,
+\begin{equation}
+\ip{f(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \qquad \forall v_i \in \mathcal{K}_i, \label{eq:cdaddvi}
+\end{equation}
+the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$, and then adds-back the subset solution $\hat w_i \in \mathcal{K}_i$.
+
+\begin{pseudofloat}[H]
 \begin{pseudo*}
@@
         \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$, \\+
-            $\displaystyle \boxed{\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}}$ \\-
+            $\displaystyle \ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}$ \\-
             $w_i = (1-\alpha) \Pi_i u + \alpha \hat w_i\in\cK_i$ \\-
@@
 
-In each boxed VI the argument of $f$ is an element of $\cK$, as the reader should confirm.
-Also, by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write the test vector $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors from $\cK$.
-For \pr{cd-add} and \pr{cd-mult} respectively:
+The multiplicative/successive \pr{cd-mult} version (Algorithm \ref{alg:basiccd-mult}) generalizes the Gauss-Seidel iteration.
+Each subset solution corrects the global iterate immediately, so here the ordering used in the ``\textbf{for}'' loop is important.
+In the VI over $\mathcal{K}_i$,
+\begin{equation}
+\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \qquad \forall v_i \in \mathcal{K}_i, \label{eq:cdmultvi}
+\end{equation}
+the sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``those portions of $u$ which have not yet improved''.
+
+In VIs \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} the argument of $f$ is an element of $\cK$, as the reader should confirm.
+
+By \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write the test vector $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors from $\cK$.
+For \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} respectively:
 \begin{align}
@@
 \end{align}
-In this sense the boxed VIs solve the same problem as \eqref{eq:vi}.
-
-Let $e_i = \hat w_i - \Pi_i u \in \cV_i$.  The reader may check that $\hat w = u + \sum_{i} e_i$ and $w = u + \alpha \sum_i e_i$.
-
-\pr{cd-add} and \pr{cd-mult} are the \emph{basic CD iterations}.
-Each one permits a damping parameter $\alpha$.
-Subject to a bound on $\alpha$, which is restrictive for \pr{cd-add} as noted by Tai \cite{Tai2003}, these iterations generate monotonically-decreasing objective values whenever VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization}.
-This is shown in the following Lemma.
-Thus the boxed VI problems in the \pr{cd-add} and \pr{cd-mult} algorithms are analogous to constrained line searches, and in fact they reduce to such when the $\cV_i$ are one-dimensional.
+In this sense the VIs over the subsets $\mathcal{K}_i$ solve the same problem as \eqref{eq:vi}.
+Furthermore, let $e_i = \hat w_i - \Pi_i u \in \cV_i$.
+The reader may confirm that $\hat w = u + \sum_{i} e_i$ and thus that $w = u + \alpha \sum_i e_i$.
+The new iterate $w$ is computed by adding a damped correction from each sub\emph{space} $\mathcal{V}_i$, but in such a manner that $w \in \mathcal{K}$; the corrections preserve admissibility.
+
+We call \pr{cd-add} and \pr{cd-mult} the \emph{basic CD iterations}.
+VI problems \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} are analogous to constrained line searches, and in fact they reduce to such line searches when the $\cV_i$ are one-dimensional.
+In general, as shown in the next Lemma \cite{Tai2003}, subject to a bound on the damping $\alpha$, which is quite restrictive for \pr{cd-add}, these iterations generate monotonically-decreasing objective values when VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization}.
 
@@
 
-To show that each iteration of the \textbf{for} loop in \pr{cd-mult} lowers the objective value, let $z_{i} = \sum_{j\le i} w_j + \sum_{j > i} \Pi_j u \in \cK$, $i=0,\dots,m-1$, be the state at the end of the $i$th iteration.
-Note that $z_0=u$, $z_{m-1}=w$, and $z_{i-1} = \sum_{j < i} w_j + \Pi_i u + \sum_{j > i} \Pi_j u$ for $i>0$.
-Therefore by convexity of $F$, observation \eqref{eq:admissibledifferencemult} with $v_i=\Pi_i u$, and the boxed VI we find:
+To show that each iteration of the \textbf{for} loop in \pr{cd-mult} lowers the objective value, let $z_{i} = \sum_{j\le i} w_j + \sum_{j > i} \Pi_j u \in \cK$, $i=0,\dots,m-1$; it is the state at the end of the $i$th iteration.
+Note that $z_{i-1} = \sum_{j < i} w_j + \Pi_i u + \sum_{j > i} \Pi_j u$ for $i>0$.
+Therefore by convexity of $F$, observation \eqref{eq:admissibledifferencemult} with $v_i=\Pi_i u$, and \eqref{eq:cdmultvi} we find:
 \begin{equation*}
@@
 \end{equation*}
-Using convexity (again) with the above inequality gives
+Using convexity (again) now gives
 \begin{align*}
@@
 \end{align*}
-As $w = z_{m-1}$, the result for \pr{cd-mult} follows.
-
-Suppose $w=u+\alpha\sum_i e_i$ is computed by \pr{cd-add} using $\alpha \le 1/m$.  Considering each $e_i=\hat w_i - \Pi_i u$ separately, convexity of $F$ and the boxed VI with $v_i=\Pi_i u$ yields
+As $u=z_0$ and $w = z_{m-1}$, the result for \pr{cd-mult} follows.
+
+Suppose $w=u+\alpha\sum_i e_i$ is computed by \pr{cd-add} using $\alpha \le 1/m$.
+Considering each $e_i=\hat w_i - \Pi_i u$ separately, convexity of $F$ and VI \eqref{eq:cdaddvi} with $v_i=\Pi_i u$ yields
 \begin{equation}
@@
 \end{equation}
-Let $v = u+ \frac{1}{m} \sum_i e_i$.  Again using convexity (Jensen's inequality), plus the above inequality, we have
+Let $v = u+ \frac{1}{m} \sum_i e_i$.  Again using convexity (Jensen's inequality), and \eqref{eq:cdaddonereduction}, we have
 \begin{equation*}
@@
 
-When $\alpha=1/m$ then the logic of the above proof for \pr{cd-add} is simpler because $v=w$ and $\lambda=1$.
-Namely, the value $F(w)$ is less than $\frac{1}{m} \sum_i F(u+e_i)$ by Jensen's inequality, which is lower than $F(u)$ by the boxed VIs, which is inequality \eqref{eq:cdaddonereduction}.
+When $\alpha=1/m$ then the above proof for \pr{cd-add} can be simplified because $v=w$ and $\lambda=1$.
+Then the value $F(w)$ is less than $\frac{1}{m} \sum_i F(u+e_i)$ by Jensen's inequality, which is lower than $F(u)$ by \eqref{eq:cdaddvi} or \eqref{eq:cdmultvi}, which gives inequality \eqref{eq:cdaddonereduction}.
 
@@
 Furthermore they are meaningful even when $f$ is nonlinear, non-local, or defined only on $\cK$.
-Practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.
-References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem for which $f=F'$ is linear, local, and defined for all inputs (i.e.~$f$ is defined on all of $\mathcal{V}$).
-The following example exposes the implementation simplifications available in that easier case.
+However, practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.
+References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem, and in that case $f=F'$ is linear, local, and defined for all inputs (i.e.~on all of $\mathcal{V}$).
+The following example exposes the implementation simplifications available in such easier cases.
 
@@
 Furthermore suppose $f$ is a differential operator, thus local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi$.
-Considering \pr{cd-add} for concreteness, by linearity the boxed VI can be rewritten
+Considering \pr{cd-add} for concreteness, by linearity \eqref{eq:cdaddvi} can be rewritten
 \begin{equation}
@@
 Thus each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $f(u)$ included into $\tilde\ell$.
-Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible ($e_i \notin \cK$), but \eqref{eq:linearlocalvi} is meaningful because $f$ is defined over $\cV$.
+Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible---generally $e_i \notin \cK$---but \eqref{eq:linearlocalvi} is meaningful because $f$ is defined over $\cV$.
 If a hat function basis exists for $\cV_i$, namely a basis of small-support functions, then by locality of $f$ a solver for VI \eqref{eq:linearlocalvi} can be implemented using only integrals over these small support sets.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}
@@
 
-When $f$ has all the nice properties listed in Example \ref{ex:fnice} then a solver based on CD iterations can take advantage of implementation efficiencies.
-That is, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.
-Whatever is the actual solution process for VI \eqref{eq:linearlocalvi}, both storage and the arithmetical cost of residual evaluation are substantially smaller than in the general case.
-
-For general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of \pr{cd-add} or \pr{cd-mult}.
-With a non-local integral operator $f$, for example, the expression $\ip{f(\phi)}{z}$ requires an integral over $\Omega$ even if $\phi$ is a function with small support; cf.~subsection 4.5 in \cite{Bueler2021conservation}.
-Section \ref{sec:vcycle} constructs a multiplicative multilevel CD iteration for general functions $f:\cK\to\cV'$, based on additive or multiplicative smoothers in Section \ref{sec:smoothers}, and assuming neither linearity nor locality, by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
+When $f$ has all the nice properties listed in Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.
+However, for general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of \pr{cd-add} or \pr{cd-mult}.
+Section \ref{sec:vcycle} constructs a multiplicative multilevel CD iteration for a general operator $f:\cK\to\cV'$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
 

### 2023-06-12 paper/fascd.tex
visualize a V-cycle; see comments before fig:imagesvcycle regarding generation

@@
 \usetikzlibrary{decorations.pathreplacing}
+\usetikzlibrary{graphs}
 
@@
 
+% figures in the following visualization were generated using branch elbueler/simplefigs
+% of git@bitbucket.org:pefarrell/fascd.git, and the following commands:
+%   cd examples/
+%   python3 plap1d.py -figures -fascd_max_it 2 -fascd_cycle_type multiplicative -mcoarse 2
+\begin{figure}[ht]
+\begin{center}
+\input{tikz/imagesvcycle.tex}
+\end{center}
+\caption{Visualization of one iteration of \textsc{FASCD-VCYCLE} (Algorithm \ref{alg:fascd}) in the case of Example FIXME with $p=1.5$ and $J=3$.
+Level iterates and solutions ($w^j$, $y^j$, $z^j$) are lines while obstacles and defect constraints ($\underline{\gamma}^j$, $\underline{\phi}^j\le 0$, $\underline{\chi}^j\le 0$) are strong dots.
+The coarsest mesh $\mathcal{T}^0$ has two elements.}
+\label{fig:imagesvcycle}
+\end{figure}
+
 FIXME observe that up-smoothing is more efficient; this observation seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}

### 2023-06-12 paper/tikz/imagesvcycle.tex
visualize a V-cycle; see comments before fig:imagesvcycle regarding generation

\begin{tikzpicture}[scale=1.0]
    \pgfmathsetmacro{\HEIGHT}{2}
    \pgfmathsetmacro{\XSTEP}{3.7}
    \pgfmathsetmacro{\YSTEP}{3.7}
    \pgfmathsetmacro{\YLABEL}{-0.4}
    \pgfmathsetmacro{\NUDGE}{0.5}
    \node (start3) [] at (0,3*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/start3.png}};
    \node (start3label) [below of = start3, yshift=\YLABEL cm] {$w^3 \ge \underline{\gamma}^3$};
    \node (down3) [] at (1.0*\XSTEP,3*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/down3.png}};
    \node (down3label) [below of = down3, yshift=\YLABEL cm] {$y^3 \ge \underline{\phi}^3$};
    \node (down2) [] at (1.2*\XSTEP,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/down2.png}};
    \node (down2label) [below of = down2, yshift=\YLABEL cm] {$y^2 \ge \underline{\phi}^2$};
    \node (down1) [] at (1.4*\XSTEP,1*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/down1.png}};
    \node (down1label) [below of = down1, yshift=\YLABEL cm] {$y^1 \ge \underline{\phi}^1$};
    \node (up0) [] at (1.9*\XSTEP,0) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/up0.png}};
    \node (up0label) [below of = up0, yshift=\YLABEL cm] {$z^0 \ge \underline{\chi}^0$};
    \node (up1) [] at (2.4*\XSTEP,1*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/up1.png}};
    \node (up1label) [below of = up1, yshift=\YLABEL cm] {$z^1 \ge \underline{\chi}^1$};
    \node (up2) [] at (2.6*\XSTEP,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/up2.png}};
    \node (up2label) [below of = up2, yshift=\YLABEL cm] {$z^2 \ge \underline{\chi}^2$};
    \node (up3) [] at (2.8*\XSTEP,3*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/up3.png}};
    \node (up3label) [below of = up3, yshift=\YLABEL cm] {$z^3 \ge \underline{\chi}^3$};
    \node (next3) [] at (3.8*\XSTEP,3*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/next3.png}};
    \node (next3label) [below of = next3, yshift=\YLABEL cm] {$\tilde{w}^3 \ge \underline{\gamma}^3$};
    \graph [edges = {very thick}] {
        (start3) -> (down3)
                 -!- [] (down3label) -> (down2)
                 -!- [] (down2label) -> (down1)
                 -!- [] (down1label) -> (up0)
                 -> (up1label) -!- (up1)
                 -> (up2label) -!- (up2)
                 -> (up3label) -!- (up3)
                 -> (next3)
    };
\end{tikzpicture}

### 2023-06-13 paper/fascd.tex
comprehensive edits and refactor; better!

--- 2023-06-02 paper/fascd.tex
@@
 
-As is true for space decompositions \cite{Xu1992}, associated to a constraint decomposition (CD) are certain iterative methods \cite{Tai2003}.
+Associated to a constraint decomposition (CD; equations \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}) are certain iterative methods \cite{Tai2003,Xu1992}.
 Let $0\le \alpha \le 1$ be the \emph{damping parameter}.
-Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below, which we now discuss, solve smaller VI problems over each subset $\cK_i$.
-These Algorithms replace $u \in \cK$ with a new iterate $w\in\cK$ which should be closer to $u^* \in \cK$ solving \eqref{eq:vi}.
+Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below, discussed next, solve smaller VI problems over each subset $\cK_i$.
+Each Algorithm replaces $u \in \cK$ with a new iterate $w\in\cK$ which should be closer to $u^* \in \cK$ solving \eqref{eq:vi}.
+
+The additive/parallel \pr{cd-add} Algorithm \ref{alg:basiccd-add} generalizes the classical Jacobi iteration.  Here ``\textbf{for all}'' can be computed in any order.  In its VI,
+\begin{equation}
+\ip{f(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \qquad \forall v_i \in \mathcal{K}_i, \label{eq:cdaddvi}
+\end{equation}
+the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$, and then adds-back the subset solution $\hat w_i \in \mathcal{K}_i$.
 
@@
 
-The additive/parallel \pr{cd-add} method (Algorithm \ref{alg:basiccd-add}) generalizes the classical Jacobi iteration.
-Here ``\textbf{for all}'' can be computed in any order.
-In the VI over $\mathcal{K}_i$ solved here,
-\begin{equation}
-\ip{f(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \qquad \forall v_i \in \mathcal{K}_i, \label{eq:cdaddvi}
-\end{equation}
-the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$, and then adds-back the subset solution $\hat w_i \in \mathcal{K}_i$.
+The multiplicative/successive \pr{cd-mult} Algorithm \ref{alg:basiccd-mult} generalizes the Gauss-Seidel iteration.
+Each subset solution corrects the global iterate immediately, so here the ordering used in the ``\textbf{for}'' loop is important.
+In the VI,
+\begin{equation}
+\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \qquad \forall v_i \in \mathcal{K}_i, \label{eq:cdmultvi}
+\end{equation}
+the sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``those portions of $u$ which have not yet improved''.
 
@@
 
-The multiplicative/successive \pr{cd-mult} version (Algorithm \ref{alg:basiccd-mult}) generalizes the Gauss-Seidel iteration.
-Each subset solution corrects the global iterate immediately, so here the ordering used in the ``\textbf{for}'' loop is important.
-In the VI over $\mathcal{K}_i$,
-\begin{equation}
-\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \qquad \forall v_i \in \mathcal{K}_i, \label{eq:cdmultvi}
-\end{equation}
-the sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``those portions of $u$ which have not yet improved''.
-
 In VIs \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} the argument of $f$ is an element of $\cK$, as the reader should confirm.
@@
 The reader may confirm that $\hat w = u + \sum_{i} e_i$ and thus that $w = u + \alpha \sum_i e_i$.
-The new iterate $w$ is computed by adding a damped correction from each sub\emph{space} $\mathcal{V}_i$, but in such a manner that $w \in \mathcal{K}$; the corrections preserve admissibility.
+The new iterate $w$ is therefore computed by adding a damped correction from each sub\emph{space} $\mathcal{V}_i$, but in such a manner that $w \in \mathcal{K}$; the corrections preserve admissibility.
 
@@
 VI problems \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} are analogous to constrained line searches, and in fact they reduce to such line searches when the $\cV_i$ are one-dimensional.
-In general, as shown in the next Lemma \cite{Tai2003}, subject to a bound on the damping $\alpha$, which is quite restrictive for \pr{cd-add}, these iterations generate monotonically-decreasing objective values when VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization}.
-
-\begin{lemma}  Suppose $f=F'$ for $F:\cK\to\RR$ convex and differentiable, and let $u\in\cK$.  For $0 \le \alpha \le 1$ the output $w \in \cK$ from \pr{cd-mult} satisfies
+As shown in the next Lemma, when VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization} then, subject to bounds on the damping $\alpha$, these iterations generate monotonically-decreasing objective values.
+
+\begin{lemma} \cite{Tai2003}  Suppose $f=F'$ for $F:\cK\to\RR$ convex and differentiable, and let $u\in\cK$.  For $0 \le \alpha \le 1$ the output $w \in \cK$ from \pr{cd-mult} satisfies
 \begin{equation}
@@
 
-To show that each iteration of the \textbf{for} loop in \pr{cd-mult} lowers the objective value, let $z_{i} = \sum_{j\le i} w_j + \sum_{j > i} \Pi_j u \in \cK$, $i=0,\dots,m-1$; it is the state at the end of the $i$th iteration.
+To show that each iteration of the \textbf{for} loop in \pr{cd-mult} lowers the objective value, let $z_{i} = \sum_{j\le i} w_j + \sum_{j > i} \Pi_j u \in \cK$, $i=0,\dots,m-1$; this is the state at the end of the $i$th iteration.
 Note that $z_{i-1} = \sum_{j < i} w_j + \Pi_i u + \sum_{j > i} \Pi_j u$ for $i>0$.
@@
 
-When $\alpha=1/m$ then the above proof for \pr{cd-add} can be simplified because $v=w$ and $\lambda=1$.
-Then the value $F(w)$ is less than $\frac{1}{m} \sum_i F(u+e_i)$ by Jensen's inequality, which is lower than $F(u)$ by \eqref{eq:cdaddvi} or \eqref{eq:cdmultvi}, which gives inequality \eqref{eq:cdaddonereduction}.
-
-The basic CD iterations are meaningful whether we are in the optimization case $f=F'$ or not.
-Furthermore they are meaningful even when $f$ is nonlinear, non-local, or defined only on $\cK$.
+When $\alpha=1/m$ then the \pr{cd-add} proof can be simplified because $v=w$ and $\lambda=1$.
+In this case $F(w)$ is less than $\frac{1}{m} \sum_i F(u+e_i)$ by Jensen's inequality, which is lower than $F(u)$ by \eqref{eq:cdaddvi} or \eqref{eq:cdmultvi}, which gives inequality \eqref{eq:cdaddonereduction}.
+
+The basic CD iterations are meaningful whether or not we are in the optimization case $f=F'$, and when $f$ is nonlinear, non-local, or defined only on $\cK$.
 However, practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.
-References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem, and in that case $f=F'$ is linear, local, and defined for all inputs (i.e.~on all of $\mathcal{V}$).
-The following example exposes the implementation simplifications available in such easier cases.
-
-\begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear.
-Furthermore suppose $f$ is a differential operator, thus local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi$.
+References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem, and which case $f=F'$ is linear, local, and defined for all inputs (on all of $\mathcal{V}$).
+Substantial implementation simplifications are available in such cases.
+
+\begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear and a differential operator, thus local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi$.
 Considering \pr{cd-add} for concreteness, by linearity \eqref{eq:cdaddvi} can be rewritten
@@
 \end{equation}
-for all $v_i \in \mathcal{K}_i$, where $\tilde\ell = \ell - f(u)$.
-Thus each problem \eqref{eq:linearlocalvi} can be solved using a stored residual $f(u)$ included into $\tilde\ell$.
-Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible---generally $e_i \notin \cK$---but \eqref{eq:linearlocalvi} is meaningful because $f$ is defined over $\cV$.
-If a hat function basis exists for $\cV_i$, namely a basis of small-support functions, then by locality of $f$ a solver for VI \eqref{eq:linearlocalvi} can be implemented using only integrals over these small support sets.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}
+for all $v_i \in \mathcal{K}_i$, where $\tilde\ell = \ell - f(u)$ incorporates the initial residual $f(u)$.
+Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible---generally $e_i \notin \cK$---but \eqref{eq:linearlocalvi} is meaningful for this $f$.
 \end{example}
 
-When $f$ has all the nice properties listed in Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.
-However, for general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of \pr{cd-add} or \pr{cd-mult}.
+When $f$ has the locality of Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}  However, for general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of \pr{cd-add} or \pr{cd-mult}.
 Section \ref{sec:vcycle} constructs a multiplicative multilevel CD iteration for a general operator $f:\cK\to\cV'$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
 
-\section{Multilevel finite elements} \label{sec:femultilevel}
-
-From now on we seek the approximate solution of certain nonlinear VI problems \eqref{eq:vi} over 2D computational domains.
-Suppose $\Omega \subset \RR^2$ is a (open) bounded polygon and let $\mathcal{V}=W^{1,p}(\Omega)$.
-In addition to the free boundaries generated by inequality constraints, fixed boundary conditions will be allowed, of either Dirichlet or mixed type.
-
-We restrict to the case of inequality \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.
+From now on we restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.
+Suppose $\Omega \subset \RR^d$ is open and let $\mathcal{V}=W^{1,p}(\Omega)$.
 The \emph{lower obstacle} $\underline{\gamma}$ and the \emph{upper obstacle} $\overline{\gamma}$ are given functions, defined on the closure $\overline{\Omega}$, with values in the extended real line $\tilde\RR = [-\infty,+\infty]$.
 Specifically, we suppose $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$ and $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$ are measurable, $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$, and $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.
-
-Also, we assume that $\partial\Omega$ is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive measure for simplicity to permit well-posedness \cite{Evans2010}, and $\partial_N \Omega$ possibly empty.
-Suppose that the provided Dirichlet data $g_D:\partial_D \Omega \to \RR$ is sufficiently-regular so that the VI problem below is well posed.
-
-This data defines the constraint set
-\begin{equation}
-\cK = \left\{w\,:\,\underline{\gamma} \le w \le \overline{\gamma} \, \text{ and }\, w\big|_{\partial_D \Omega} = g_D\right\} \subset \cV =W^{1,p}(\Omega),
-\end{equation}
-a closed and convex set.  (Dirichlet boundary values are in the trace sense \cite{Evans2010}.)  Note that this definition treats the inequality and equality (boundary) constraints in a unified manner.
-
-We will solve the same VI as abstract problem \eqref{eq:vi}.  That is, given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we assume that the following VI is well-posed for $u^*\in \cK$:
+Also, we assume that $\partial\Omega$ is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive measure, to facilitate well-posedness \cite{Evans2010}, and $\partial_N \Omega$ possibly empty.
+Dirichlet data\footnote{Dirichlet boundary values are in the trace sense \cite{Evans2010}, and $g_D$ is assumed to be as regular as needed.} $g_D:\partial_D \Omega \to \RR$ then defines
+\begin{equation}
+\cK = \left\{w\,:\,\underline{\gamma} \le w \le \overline{\gamma} \, \text{ and }\, w\big|_{\partial_D \Omega} = g_D\right\} \subset \cV =W^{1,p}(\Omega), \label{eq:originalconstraintset}
+\end{equation}
+the closed and convex constraint set.  Note that we are treating inequality and equality (boundary) constraints in a unified manner.
+
+Recalling abstract VI problem \eqref{eq:vi}, we now consider box-constrained cases only, using set \eqref{eq:originalconstraintset}.
+Given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we will assume that the following is well-posed for $u^*\in \cK$:
 \begin{equation}
@@
 \end{equation}
-Note that $v-u^* \in W_0^{1,p}(\Omega)$, that is, test functions in \eqref{eq:boxdirichletvi} have zero values on the Dirichlet boundary $\partial_D\Omega$.
-Of course, any Neumann conditions over $\partial_N \Omega$ are determined by the operator formula for $f$, which may include an integration over $\partial_N\Omega$.
-In summary, VI \eqref{eq:boxdirichletvi} is both a free- and fixed-boundary value problem.
-
-We now construct a multilevel finite element (FE) approximation to problem \eqref{eq:boxdirichletvi} using either triangular or quadrilateral meshes over $\Omega$, and $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, respectively.
-A hierarchy of meshes will be constructed by refinement of a coarsest mesh, with the goal being a solution on the finest-level mesh.
-This hierarchy also allows the creation of a multilevel CD of $\cK$ (Section \ref{sec:cdmultilevel}).
-The subsets $\cK_i$ in this CD permit coarser-level corrections in a manner which preserves admissibility while not introducing unnecessary high frequencies into the coarser-level problems, especially at the free boundary.
-
-The mesh hierarchy and FE spaces are constructed by nested refinement in the following standard manner \cite{Elmanetal2014}.
-Suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate elements (triangles or quadrilaterals) with union equal to $\overline{\Omega}$ and maximum edge length (characteristic mesh size) $h_0>0$.
+Note that $v-u^* \in W_0^{1,p}(\Omega)$, so test functions in \eqref{eq:boxdirichletvi} have zero Dirichlet values.
+Any Neumann conditions over $\partial_N \Omega$ are determined by the operator formula for $f$, i.e.~by an integration over $\partial_N\Omega$.
+In summary, from now on we consider well-posed VI problems \eqref{eq:boxdirichletvi}, namely box-constrained boundary value problems.
+
+\section{Multilevel finite elements} \label{sec:femultilevel}
+
+We seek to rapidly compute the approximate solution of box-constrained VI problems using finite elements (FE).
+Suppose $\Omega \subset \RR^d$ is a (open) bounded polygon.\footnote{Section \ref{sec:results} illustrates $d=1,2,3$ cases.}  We will construct a multilevel FE approximation to problem \eqref{eq:boxdirichletvi} over $\Omega$ using $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, by refining a coarsest mesh into a mesh hierarchy; the goal is an accurate solution on the finest-level mesh.
+The hierarchy allows the creation of a multilevel CD of $\cK$, using the ideas in the next two Sections, which permits admissibility-preserving coarse corrections.
+
+The mesh hierarchy and FE spaces can be constructed by standard nested refinement.\footnote{The dimension-dependent details of mesh refinement are suppressed here.
+Section \ref{sec:results} only uses $P_1$ elements for intervals, triangles, and tetrahedra, in dimensions $d=1,2,3$ respectively, within the Firedrake framework \cite{Rathgeberetal2016}.}  Suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate elements with union equal to $\overline{\Omega}$ and maximum edge length (characteristic mesh size) $h_0>0$.
 We assume $\mathcal{T}^0$ is conforming, with no hanging nodes.
-For $j=1,\dots,J$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ by connecting the midpoints of edges.
-For triangles and rectangles, each $\mathcal{T}^{j-1}$ element becomes four similar elements\footnote{For arbitrary triangles, and for rectangles, edge midpoint refinement generates similar elements and $h_j=2^{-j}h_0$ is the exact maximum edge length in $\mathcal{T}^j$.
-For general quadrilaterals, similarity is lost and maximum edge lengths are variable; cf.~\cite{Zhang2004}.
-However, only rectangular quadrilaterals are used in the Results section.} in $\mathcal{T}^j$, with characteristic mesh size $h_j = 2^{-j} h_0$.
+For $j=1,\dots,J$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ so that each $\mathcal{T}^{j-1}$ element becomes $2^d$ elements in $\mathcal{T}^j$ with maximum edge length $h_j = 2^{-j} h_0$.
 We call $J\ge 0$ the \emph{depth} of the hierarchy, $J+1$ the \emph{number of levels}, and $\mathcal{T}^J$ the \emph{finest-level mesh}.
 Let $m_j$ be the number of nodes in $\mathcal{T}^j$, so there are $m_j$ \emph{degrees of freedom} at the $j$th level.
-Denote $h=h_J=2^{-J} h_0$ and $m=m_J$ for the finest level.
-
-Let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space \cite{Elmanetal2014} over $\mathcal{T}^j$, with arbitrary boundary values.  Observe the following vector space nesting:
+Let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space over $\mathcal{T}^j$, with arbitrary boundary values, and observe the following vector space nesting:
 \begin{equation}
@@
 \end{equation}
-
-In the VI material which follows, use of higher-order Lagrange elements $P_k$ or $Q_k$, $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} of a type special to methods for inequality constraints.
-Namely, our FE construction of constraint sets will require \emph{nodal monotonicity},
+On the finest-level mesh let $g_D^J \in \mathcal{V}^J$ satisfy $g_D^J = g_D$ on $\partial_D \Omega$; this restricts $g_D$ in \eqref{eq:originalconstraintset} to be piecewise-linear.
+
+In the VI material which follows, use of higher-order Lagrange elements, i.e.~$P_k$ or $Q_k$ with $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007}, special to inequality constraints, which we avoid.
+Namely, our FE construction of constraint sets (Section \ref{sec:cdmultilevel}) will require \emph{nodal monotonicity},
 \begin{equation}
@@
 \end{equation}
-for all $w,z \in \mathcal{V}^j$, where $\bw,\bz$ denote the vectors formed by the nodal values.
-For $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal value one at a single node, otherwise zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.
-It is easy to show that \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of indefinite-sign hat functions, so use of \eqref{eq:nodalmonotonicity} implies $k\le 1$.
-
-Let $g_D^J \in \mathcal{V}^J$, on the finest-level mesh, satisfy $g_D^J = g_D$ on $\partial_D \Omega$.
-As is typical of conforming FE methods, this restricts the Dirichlet boundary function $g_D$ to be piecewise-linear \cite{Elmanetal2014}.
-
-To allow box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
-(Note that $\tilde{\mathcal{V}}^j$ functions are not defined everywhere on $\Omega$.)  Suppose that the finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$, for nodes of $\mathcal{T}^J$, satisfy
+for all $w,z \in \mathcal{V}^j$, where $\bw,\bz \in \RR^{m_j}$ denote vectors of nodal values.
+However, for $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal values one or zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.
+Evidently \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of such indefinite-sign hat functions.
+
+To allow box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions\footnote{Note that $\tilde{\mathcal{V}}^j$ functions need not be defined everywhere on $\Omega$.} on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
+Suppose that the finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ satisfy
 \begin{equation}
@@
 \end{equation}
-Assuming that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.  Then define
-\begin{equation}
-\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \text{ on nodes of } \mathcal{T}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
-\end{equation}
-as the finest-level \emph{constraint set} of admissible FE functions.
+Also assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.  Define
+\begin{equation}
+\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \text{ on nodes of } \mathcal{T}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D^J|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
+\end{equation}
+the finest-level \emph{constraint set} of admissible FE functions.
 This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.
@@
 \end{equation}
-As for \eqref{eq:vi}, if $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, then \eqref{eq:fe:vi} is well-posed.
-
-While \eqref{eq:fe:vi} approximates \eqref{eq:boxdirichletvi}, it may not technically be conforming for that VI.
+If $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, then, as for \eqref{eq:vi}, FE VI \eqref{eq:fe:vi} is well-posed.
+
+While \eqref{eq:fe:vi} approximates \eqref{eq:boxdirichletvi}, we warn the reader that it may not technically be conforming for that VI.
 That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require or generally expect $\mathcal{K}^J \subset \mathcal{K}$.
-Even if smooth continuum obstacles exist, admissibility with respect to their FE interpolants will generally not imply continuum admissiblity.
-Such caveats are especially relevant in geophysical situations where only nodal values of the obstacle are known and nodal elevation data misses detail in between; see the Greenland ice sheet application in \cite{Bueler2016}, for example.
-
-Finite-dimensional, box-constrained VI problem \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003,FerrisPang1997}, a strong form problem whose statement may be clarifying.
-Denote $x_p$ for a node of the finest-level mesh $\mathcal{T}^J$ and $\psi_p(x)$ for the corresponding hat function.
-Let $\mathcal{X}=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}$.
-If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let us define $I_D$ as the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.
-Problem \eqref{eq:fe:vi}, which would be stated as VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap X\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP with additional equality constraints:
-\begin{align}
-\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J - I_D^\top \lambda \label{eq:fe:mcp} \\
-I_D u^J &= g_D^J\big|_{\partial_D\Omega} \notag
-\end{align}
-for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
-If $\underline{\gamma}^J(x_p) < \overline{\gamma}^J(x_p)$ then MCP \eqref{eq:fe:mcp} says that exactly one of the following five conditions holds for the solution $u^J$:
-\begin{itemize}
-\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)<u^J(x_p)<\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} = 0$, or
-\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)=u^J(x_p)<\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \ge 0$, or
-\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)<u^J(x_p)=\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \le 0$, or
-\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$, or
-\item $x_p \in \partial_D\Omega$ and $u^J(x_p)=g_D^J(x_p)$ and $\lambda_p=\ip{f^J(u^J)-\ell^J}{\psi_p}$.
-\end{itemize}
-The cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, \emph{pinched}, and \emph{boundary}, respectively.
-Where case \emph{pinched} applies $\ip{f^J(u^J)-\ell^J}{\psi_p}$ can have any real value, and similarly if \emph{boundary} applies then $\lambda_p$ is unconstrained.
-
-The multilevel algorithm defined in the next two Sections, a solver for finite-dimensional problem \eqref{eq:fe:vi}, requires transfer operators between meshes.
+Regardless of the smoothness of the continuum obstacles $\underline{\gamma}, \overline{\gamma}$, admissibility with respect to their FE interpolants $\underline{\gamma}^J, \overline{\gamma}^J$ will generally not imply continuum admissiblity.\footnote{This caveat is especially relevant in geophysical situations where only nodal values of the obstacle are known.
+The nodal elevation data misses the actual topographic detail in-between the nodes.
+See the Greenland ice sheet application in \cite{Bueler2016}, for example.}
+
+The multilevel algorithm defined in the next two Sections, a solver for FE VI \eqref{eq:fe:vi}, requires transfer operators between meshes.
 We require five such operators, of which three are standard multigrid tools \cite{Trottenbergetal2001}.
@@
 
-Canonical prolongation $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ acts as the identity on coarser-mesh functions in $\mathcal{V}^{j-1}$; it is the linear vector space inclusion $\mathcal{V}^{j-1} \hookrightarrow \mathcal{V}^j$.
+Canonical prolongation $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ is the vector space inclusion $\mathcal{V}^{j-1} \hookrightarrow \mathcal{V}^j$ of coarser-mesh functions.
 Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ also acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $(R\sigma)[z] = \sigma[z]$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
-However, these canonical transfers generally involve a change of FE representation.
-For example, the formula for canonical restriction, derived by writing coarser-mesh hat functions in terms of finer-mesh ones, is a ``full-weighting'' average \cite{Trottenbergetal2001}.
-
-To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, observe that for any given values at the nodes of $\mathcal{T}^j$ there is a unique function in $\mathcal{V}^j$ with those values.
-Also observe that, by construction, the nodes of $\mathcal{T}^{j-1}$ are nodes of $\mathcal{T}^j$.
-For $z\in\mathcal{V}^j$ we define the injection $\iR z$ as the element of $\mathcal{V}^{j-1}$ which has the same nodal values as $z$ does at the nodes of $\mathcal{T}^{j-1}$.
+Note that these canonical transfers generally involve a change of FE representation.
+For example, the formula for canonical restriction, derived by writing coarser-mesh hat functions in terms of finer-mesh ones, is ``full-weighting'' \cite{Trottenbergetal2001}.
+
+To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, observe that the nodes of $\mathcal{T}^{j-1}$ are nodes of $\mathcal{T}^j$, and that for any values at $\mathcal{T}^j$ nodes there is a unique $\mathcal{V}^j$ function assuming those values.
+For $z\in\mathcal{V}^j$ we define the injection $\iR z$ as the element of $\mathcal{V}^{j-1}$ with the same (coarse) nodal values.
 Note that injection is monotonic: if $w \le z$ then $\iR w \le \iR z$.
 
-The monotone restrictions $\maxR,\minR$ are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh hat functions.
-These restrictions will be applied only to nodal functions $\tilde{\mathcal{V}}^j$.
+Finally, the monotone restrictions \cite{GraeserKornhuber2009} $\maxR,\minR$ are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh hat functions.
+These monotone operators will only be applied to nodal functions $\tilde{\mathcal{V}}^j$.
 To make the definitions precise, let $x_q^j$, $q=0,1,\dots,m_j-1$, denote the nodes of mesh $\mathcal{T}^j$, and denote the corresponding hat functions by $\psi_q^j(x)$.
-The monotone restrictions $\maxR z,\minR z$ of an $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ are the nodal functions in $\tilde{\mathcal{V}}^{j-1}$ defined by
+The monotone restrictions of an $\tilde\RR$-valued\footnote{In computing $\max$ and $\min$ to define \eqref{eq:fe:monotonerestrictions}, the usual $\tilde\RR$ ordering is applied.} nodal function $z\in\tilde{\mathcal{V}}^j$ are the following functions in $\tilde{\mathcal{V}}^{j-1}$:
 \begin{align}
@@
 \end{align}
-for $p=0,1,\dots,m_{j-1}-1$.
-(In computing $\max$ and $\min$, the usual $\tilde\RR$ ordering is applied.)  By mesh refinement a node $x_p^{j-1}$ is in the set of $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so it follows that
+for $p=0,1,\dots,m_{j-1}-1$.  By standard nested refinement, a coarse-mesh node $x_p^{j-1}$ is in the set of fine nodes $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so it follows that
 \begin{equation}
@@
 \end{equation}
-In particular, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$, properties used in the CD construction which follows.
-
-\section{Level defect constraints} \label{sec:ldcs}
-
-Suppose $w^J \in \cK^J$ is an admissible finest-level iterate which approximates the solution $u^J$ of VI \eqref{eq:fe:vi}.
-From $w^J$ we now define the CD constraint subsets, in our multilevel FE context, using the defining inequalities for $\cK^J$.
-Though the ideas are fundamentally the same as for unilateral constraints, e.g.~as in \cite{GraeserKornhuber2009}, the notation is necessarily more elaborate because we consider box-constrained (bi-obstacle) problems.
+On the other hand, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These ordering properties used in the CD construction of the next two sections.
+
+\section{Multilevel CDs from level defect constraints} \label{sec:cdmultilevel}
+
+Suppose $w^J \in \cK^J$ is an admissible finest-level iterate which approximates the solution $u^J$ of FE VI \eqref{eq:fe:vi}.
+From $w^J$ we will now define CD constraint subsets using the defining inequalities \eqref{eq:fe:fineconstraintset} for $\cK^J$ and the above monotone restrictions $\minR, \maxR$.
+Though the ideas are fundamentally the same as for unilateral constraints \cite{GraeserKornhuber2009}, the notation is necessarily more elaborate for box-constrained problems.
 
@@
 \end{equation}
-Note that $\underline{\chi}^J \le 0 \le \overline{\chi}^J$ everywhere in $\overline{\Omega}$, including along the Dirichlet boundary $\partial_D\Omega$.
-If $\underline{\gamma}^J=-\infty$ then $\underline{\chi}^J=-\infty$ by definition, and similarly if $\overline{\gamma}^J=+\infty$ then $\overline{\chi}^J=+\infty$.
-
-Recall that on $\partial_D\Omega$ we have $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ and $w^J = g_D^J$.
-The motivation for creating defect constraints $\underline{\chi}^J,\overline{\chi}^J$ is that a corrected iterate $w^J + z^J$ is admissible, $w^J + z^J \in \cK^J$, if and only if the perturbation $z^J$ is between the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$, and $z^J=0$ on $\partial_D\Omega$.
+From definition \eqref{eq:fe:fineconstraintset}, note that $\underline{\chi}^J \le 0 \le \overline{\chi}^J$ everywhere in $\overline{\Omega}$, including along the Dirichlet boundary $\partial_D\Omega$.
+Also, if $\underline{\gamma}^J=-\infty$ then $\underline{\chi}^J=-\infty$ by definition, and similarly if $\overline{\gamma}^J=+\infty$ then $\overline{\chi}^J=+\infty$.
+The motivation for creating $\underline{\chi}^J,\overline{\chi}^J$ is that a corrected iterate $w^J + z^J$ is admissible, $w^J + z^J \in \cK^J$, if and only if the perturbation $z^J$ is between the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$, and $z^J=0$ on $\partial_D\Omega$.
 
@@
 \begin{equation}
-\underline{\chi}^{J} \le \dots \le \underline{\chi}^0 \le 0 \le \overline{\chi}^0 \le \dots \le \overline{\chi}^J, \label{eq:fe:chiordering}
-\end{equation}
-i.e.~as nodal values.  A $J=4$ example of LDCs $\underline{\chi}^j$ is shown in Figure \ref{fig:chiphilevels}.
+\underline{\chi}^{J} \le \dots \le \underline{\chi}^0 \le 0 \le \overline{\chi}^0 \le \dots \le \overline{\chi}^J. \label{eq:fe:chiordering}
+\end{equation}
+A unilateral $J=4$ example of LDCs $\underline{\chi}^j$ is shown in Figure \ref{fig:chiphilevels}.
+Intuitively-speaking, the LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if another element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$.
+Similar comments apply to $\overline{\chi}^{j}$.
 
@@
 \includegraphics[width=0.65\textwidth]{fixfigs/chiphilevels.pdf}
-\caption{The level defect constraints (LDC) $\underline{\chi}^j$ are generated from $\underline{\chi}^J = \underline{\gamma}^J - w^J$ using maximum restriction: $\underline{\chi}^j = \maxR \underline{\chi}^{j+1}$.}
+\caption{Unilateral LDCs $\underline{\chi}^j$ are generated from $\underline{\chi}^J = \underline{\gamma}^J - w^J$ using maximum restriction: $\underline{\chi}^j = \maxR \underline{\chi}^{j+1}$.}
 \label{fig:chiphilevels}
@@
 
-Intuitively-speaking, the LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if another element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$, and similar comments apply to $\overline{\chi}^{j}$.
-This construction of the box constraints on each coarser level is designed to permit the largest corrections from the coarsest levels \cite{GraeserKornhuber2009}, which promotes V-cycle efficiency.
-Note that for the multilevel scheme in the following Sections, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not required when generating the LDCs $\underline{\chi}^{j},\overline{\chi}^{j}$.
+This construction of multilevel box constraints is designed to permit the largest admissible coarse corrections, which promotes multilevel solver efficiency.
+Note that for the multilevel schemes in Sections \ref{sec:vcycle} and \ref{sec:implementation}, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not required.
 
@@
 Regarding subtraction of extended real-valued quantities, if $\underline{\chi}^j(x_p)=-\infty$ then we set $\underline{\phi}^j(x_p)=-\infty$ regardless of the value of $\underline{\chi}^{j-1}(x_p)$, and similarly $\overline{\phi}^j(x_p)=+\infty$ if $\overline{\chi}^j(x_p)=+\infty$.
-These new LDCs $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ again bracket zero: $\underline{\phi}^j \le 0 \le \overline{\phi}^j$.
-Though  they are not ordered as in \eqref{eq:fe:chiordering}, telescoping sums hold for any $j=0,1,\dots,J$:
-\begin{equation}
-\sum_{i=0}^j \underline{\phi}^i = \underline{\chi}^j, \qquad \sum_{i=0}^j \overline{\phi}^i = \overline{\chi}^j.  \label{eq:fe:telescoping}
-\end{equation}
+These new LDCs $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ again bracket zero, $\underline{\phi}^j \le 0 \le \overline{\phi}^j$, but they are not ordered as in \eqref{eq:fe:chiordering}.
 Table \ref{tab:ldcs} should help the reader keep track of our box-constrained LDC system.
@@
 \begin{table}[H]
-\begin{tabular}{llc}
-\emph{name}        & \emph{formulas} \\ \hline
-(finest) upward & ${\Large \strut} \underline{\chi}^J = \underline{\gamma}^J - w^J,\quad \overline{\chi}^J = \overline{\gamma}^J - w^J$ \\
-upward          & ${\Large \strut} \underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}, \quad \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}$ \\
-downward        & ${\Large \strut} \underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1}, \quad \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}$
+\begin{tabular}{lll}
+\emph{name}         & \emph{formulas} & \emph{indices} \\ \hline
+finest-level defect constraints & ${\Large \strut} \underline{\chi}^J = \underline{\gamma}^J - w^J,\, \overline{\chi}^J = \overline{\gamma}^J - w^J$ \\
+upward LDCs  & ${\Large \strut} \underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}, \, \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}$ & $j=0,\dots,J-1$ \\
+downward LDCs       & ${\Large \strut} \underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1}, \, \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}$ & $j=0,\dots,J$
 \end{tabular}
@@
 \medskip
-\caption{Box level defect constraints (LDCs).}
+\caption{Level defect constraints (LDCs) for box-constrained VI problems.}
 \label{tab:ldcs}
@@
 
-\section{Multilevel CDs} \label{sec:cdmultilevel}
-
-A multilevel V-cycle solver is defined in the next Section.
-This V-cycle uses LDCs $\underline{\phi}^j,\overline{\phi}^j$ for the downward side and $\underline{\chi}^j,\overline{\chi}^j$ for the upward side.
-The corresponding \emph{downward} and \emph{upward defect constraint sets}, respectively, are
+The multilevel V-cycle solver defined in the next Section uses LDCs $\underline{\phi}^j,\overline{\phi}^j$ for the downward side and $\underline{\chi}^j,\overline{\chi}^j$ for the upward side.
+The corresponding \emph{downward/upward constraint sets}, respectively, are
 \begin{align}
@@
 These are closed and convex subsets of the FE spaces $\mathcal{V}^j$.
-Because $\underline{\chi}^j \le \underline{\phi}^j \le 0 \le \overline{\phi}^j \le \overline{\chi}^j$, it follows that $\mathcal{D}^j \subset \mathcal{U}^j$.
+Because $\underline{\chi}^j \le \underline{\phi}^j \le 0 \le \overline{\phi}^j \le \overline{\chi}^j$, it follows that $\mathcal{D}^j \subseteq \mathcal{U}^j$.
 Note that the finest-level set $\mathcal{U}^J$ is equivalent to the original constraint set $\mathcal{K}^J$: for all $z^J \in \mathcal{V}^J$,
@@
 
-Our multilevel strategy is based on incomplete constraint decompositions (CDs; Section \ref{sec:cd}) of $\mathcal{U}^J$, as we now explain.
+Our multilevel strategy is, however, based on \emph{incomplete} CDs (Section \ref{sec:cd}), in the general box-constrained cases, as we now explain.
 Each upward set $\mathcal{U}^j$ can be decomposed in an inclusion sense, becoming a true CD in unilateral cases.
@@
 
-\begin{lemma}  \label{lem:downwardadmissibility}  For arbitrary upper and lower obstacles $\underline{\gamma}^J,\overline{\gamma}^J$ satisfying \eqref{eq:fe:boxconstraintrequirements}, and $j=0,1,\dots,J$,
-\begin{equation}
-\mathcal{U}^j \supset \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^0. \label{eq:fe:downwardsuminclusion}
-\end{equation}
-In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, inclusion \eqref{eq:fe:downwardsuminclusion} becomes a full CD \eqref{eq:constraintdecomp}, namely $\mathcal{U}^j=\sum_{i=0}^j \mathcal{D}^i$.
+\begin{lemma}  \label{lem:downwardadmissibility}  For arbitrary finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$ satisfying \eqref{eq:fe:boxconstraintrequirements}, it follows that
+\begin{equation}
+\mathcal{U}^j \supseteq \mathcal{D}^j + \mathcal{D}^{j-1} + \dots + \mathcal{D}^0 \label{eq:fe:downwardsuminclusion}
+\end{equation}
+for $j=0,1,\dots,J$.
+In unilateral cases, where either $\underline{\gamma}^J=-\infty$ identically or $\overline{\gamma}^J=+\infty$ identically, inclusion \eqref{eq:fe:downwardsuminclusion} becomes a full CD \eqref{eq:constraintdecomp}, namely $\mathcal{U}^j=\sum_{i=0}^j \mathcal{D}^i$.
 \end{lemma}
@@
 \begin{proof}  Subspace decomposition \eqref{eq:subspacedecomp} holds---see nesting \eqref{eq:fe:nestedspaces}---and definition \eqref{eq:fe:constraintsets} shows $\mathcal{D}^i \subset \mathcal{U}^i \subset \cV^i$.
-If $y^i \in \mathcal{D}^i$ for $0 \le i \le j$ then by \eqref{eq:fe:telescoping},
-\begin{equation}
-\underline{\chi}^j = \sum_{i=0}^j \underline{\phi}^i \le \sum_{i=0}^j y^i \le \sum_{i=0}^j \overline{\phi}^i \le \overline{\chi}^j \label{eq:fe:lemmaordering}
-\end{equation}
-thus \eqref{eq:fe:downwardsuminclusion} holds for any box constraints.
+Furthermore, telescoping sums hold for any $j=0,1,\dots,J$:
+\begin{equation}
+\sum_{i=0}^j \underline{\phi}^i = \underline{\chi}^j, \qquad \sum_{i=0}^j \overline{\phi}^i = \overline{\chi}^j.  \label{eq:fe:telescoping}
+\end{equation}
+Thus if $y^i \in \mathcal{D}^i$ for $0 \le i \le j$ then
+\begin{equation}
+\underline{\chi}^j = \sum_{i=0}^j \underline{\phi}^i \le \sum_{i=0}^j y^i \le \sum_{i=0}^j \overline{\phi}^i \le \overline{\chi}^j, \label{eq:fe:lemmaordering}
+\end{equation}
+so \eqref{eq:fe:downwardsuminclusion} holds for any box constraints.
 
 In unilateral cases we can also construct decomposition operators $\Pi_i$ and show \eqref{eq:constraintrestrictionsum}.
-Suppose $\overline{\gamma}^J=+\infty$, thus $\overline{\chi}^j=+\infty$ and $\overline{\phi}^i = +\infty$ for all levels.
+For concreteness suppose $\overline{\gamma}^J=+\infty$, thus $\overline{\chi}^j=+\infty$ and $\overline{\phi}^i = +\infty$ for all levels.
 For $v\in \mathcal{V}^j$ and $0\le i \le j$, let $I_{j\to i}^\ominus$ be the operator applying the minimum (monotone) restriction $j-i$ times, mapping into $\mathcal{V}^i$:
@@
 Note that $I_{j\to j}^\ominus v = v$, and set $I_{j\to -1}^\ominus=0$ by definition.
-From operators \eqref{eq:fe:minimummaps} we define the nonlinear decomposition operators $\Pi_i:\mathcal{U}^j \to \mathcal{D}^i$:
+From operators \eqref{eq:fe:minimummaps} we define the following nonlinear decomposition operators $\Pi_i:\mathcal{U}^j \to \mathcal{D}^i$:
 \begin{equation}
@@
 \end{align*}
-This shows \eqref{eq:constraintrestrictionsum}, thus that \eqref{eq:fe:downwardsuminclusion} is equality, a true CD \eqref{eq:constraintdecomp}.  The $\underline{\gamma}^J=-\infty$ case is similar.
+This shows \eqref{eq:constraintrestrictionsum}, thus that \eqref{eq:fe:downwardsuminclusion} is equality, yielding a true CD \eqref{eq:constraintdecomp}.
+
+The $\underline{\gamma}^J=-\infty$ case is similar.
 \end{proof}
@@
 However, the following Example shows why $\Pi_i(\min\{z^j,0\})$ may have an arbitrarily large maximum, and thus exceed any finite upper LDC $\overline{\phi}^i$, so unfortunately $\Pi_i$ does not map even non-positive functions to $\mathcal{D}^i$ in the presence of a finite upper LDC.
-On the other hand, the existence of an entirely-different strategy for showing a full CD is not excluded.
 
@@
 
-Each upward defect constraint set $\mathcal{U}^j$ can also be be decomposed down to an arbitrary level $k$ where an upward set $\mathcal{U}^k$ is used as the coarsest set.
+The existence of an entirely-different strategy for showing a full CD in the general box-constrained case is, of course, not excluded.
+
+Each upward defect constraint set $\mathcal{U}^j$ can also be decomposed down to an arbitrary level $k$ where an upward set $\mathcal{U}^k$ is used as the coarsest set.
+FIXME this will be used.
 While similar to Lemma \ref{lem:downwardadmissibility}, the next Lemma does not follow logically from it.
@@
 
-\section{Full approximation storage constraint decomposition (\fascd) V-cycle} \label{sec:vcycle}
+\section{FASCD V-cycle} \label{sec:vcycle}
 
@@
 By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.
-Smoothing, an incomplete solution process on a VI problem (Section \ref{sec:smoothers}), then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.
-At the bottom the coarsest-level VI is then solved (also Section \ref{sec:smoothers}), resulting in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.
+Smoothing, an incomplete solution process on a VI problem (Section \ref{sec:implementation}), then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.
+At the bottom the coarsest-level VI is then solved, resulting in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.
 Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$.
@@
 In practice the LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ are generated during the descent stage; see Algorithm \ref{alg:fascd}.
-Observe that the LDCs do not depend on $w^j$ for $j<J$, nor on the corrections $y^j,z^j$.} been used to general the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:ldcs}).
+Observe that the LDCs do not depend on $w^j$ for $j<J$, nor on the corrections $y^j,z^j$.} been used to general the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}).
 
@@
 
-We will regard the smoothers for VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} as approximate, in-place procedures, acting on the corrections $y^j$ and $z^j$, respectively (Section \ref{sec:smoothers}).
+We will regard the smoothers for VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} as approximate, in-place procedures, acting on the corrections $y^j$ and $z^j$, respectively (Section \ref{sec:implementation}).
 Also, as is standard in multigrid, at the coarsest $j=0$ level we suppose that problem \eqref{eq:fe:upvi} is solved quite accurately.
@@
 These ideas come together in Algorithm \ref{alg:fascd}.
-\pr{fascd-vcycle} acts in-place on $w^J$, and it assumes certain semantics for the \pr{smooth} and the coarsest-level \pr{solve} procedures (Section \ref{sec:smoothers}).
+\pr{fascd-vcycle} acts in-place on $w^J$, and it assumes certain semantics for the \pr{smooth} and the coarsest-level \pr{solve} procedures.
 These do in-place modifications of their final arguments, without modifying their first four arguments.
@@
 
+\section{Implementation and smoothers} \label{sec:implementation}
+
+Finite-dimensional, box-constrained FE VI problem \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003,FerrisPang1997}, a strong form problem whose statement may be clarifying.
+Denote $x_p$ for a node of the finest-level mesh $\mathcal{T}^J$ and $\psi_p(x)$ for the corresponding hat function.
+Let $\mathcal{X}=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}$.
+If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let us define $I_D$ as the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.
+Problem \eqref{eq:fe:vi}, which would be stated as VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap X\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP with additional equality constraints:
+\begin{align}
+\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J - I_D^\top \lambda \label{eq:fe:mcp} \\
+I_D u^J &= g_D^J\big|_{\partial_D\Omega} \notag
+\end{align}
+for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
+If $\underline{\gamma}^J(x_p) < \overline{\gamma}^J(x_p)$ then MCP \eqref{eq:fe:mcp} says that exactly one of the following five conditions holds for the solution $u^J$:
+\begin{itemize}
+\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)<u^J(x_p)<\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} = 0$, or
+\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)=u^J(x_p)<\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \ge 0$, or
+\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)<u^J(x_p)=\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \le 0$, or
+\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$, or
+\item $x_p \in \partial_D\Omega$ and $u^J(x_p)=g_D^J(x_p)$ and $\lambda_p=\ip{f^J(u^J)-\ell^J}{\psi_p}$.
+\end{itemize}
+The cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, \emph{pinched}, and \emph{boundary}, respectively.
+Where case \emph{pinched} applies $\ip{f^J(u^J)-\ell^J}{\psi_p}$ can have any real value, and similarly if \emph{boundary} applies then $\lambda_p$ is unconstrained.
+
 Repeated application of \pr{fascd-vcycle} to solve VI \eqref{eq:fe:vi} will not reduce the finest-level residual $f^J(w^J) - \ell^J$ to zero everywhere because the problem is inequality-constrained.
@@
 
-\section{Smoother implementations} \label{sec:smoothers}
-
 FIXME  Recall that a $j$th-level hat function based at node $x_p^j$ is denoted $\psi_p^j$, for an iterate $g^j$ and an admissible correction $y^j \in \mathcal{D}^j$ we define the \emph{downward nodal residual} as
@@
 
+FIXME F-cycles
+
 \section{Results and discussion} \label{sec:results}
@@
 FIXME show results for 2D Firedrake implementation of FASCD for $p$-Laplacian obstacle problem (Example \ref{ex:plaplacian}), advection-diffusion obstacle problem (Example \ref{ex:advectiondiffusion}), and porous-media obstacle problem (Example \ref{ex:porous})
+
+% figures in the following visualization were generated using branch elbueler/simplefigs
+% of git@bitbucket.org:pefarrell/fascd.git, and the following commands:
+%   cd examples/
+%   python3 plap1d.py -figures -fascd_max_it 2 -fascd_cycle_type multiplicative -mcoarse 2
+\begin{figure}[ht]
+\begin{center}
+\input{tikz/imagesvcycle.tex}
+\end{center}
+\caption{Visualization of one iteration of \textsc{FASCD-VCYCLE} (Algorithm \ref{alg:fascd}) for Example FIXME, with $p=1.5$, $J=3$, and a coarsest mesh of two elements.
+Solutions ($w^j$, $y^j$, $z^j$) are lines while obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots.}
+\label{fig:imagesvcycle}
+\end{figure}
 
@@
 
-\bigskip
-From slack, to not lose.
+\section{From slack, to not lose} \label{app:slack}
+
+FIXME move this from this appendix, after using it in the discussion etc.
 

### 2023-06-14 paper/fascd.tex
edits through end of V-cycle section

@@
 for $j=0,1,\dots,J$.
-In unilateral cases, where either $\underline{\gamma}^J=-\infty$ identically or $\overline{\gamma}^J=+\infty$ identically, inclusion \eqref{eq:fe:downwardsuminclusion} becomes a full CD \eqref{eq:constraintdecomp}, namely $\mathcal{U}^j=\sum_{i=0}^j \mathcal{D}^i$.
+In unilateral cases, where either $\underline{\gamma}^J=-\infty$ identically or $\overline{\gamma}^J=+\infty$ identically, inclusion \eqref{eq:fe:downwardsuminclusion} becomes a full CD \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} with $\mathcal{U}^j=\sum_{i=0}^j \mathcal{D}^i$.
 \end{lemma}
@@
 
-It is tempting to try to turn \eqref{eq:fe:downwardsuminclusion} into a full CD for arbitrary box constraints by using decomposition maps like $\Pi_i$.
-Specifically, one might split $z^j = \min\{z^j,0\} + \max\{z^j,0\}$ and apply unilateral formulas like \eqref{eq:fe:unilateraldecompositionoperator} to $\min\{z^j,0\}$ and $\max\{z^j,0\}$ separately, with $\Pi_i$ applied to $\min\{z^j,0\}$ and a straightforward variant operator to $\max\{z^j,0\}$.
+One might try to turn \eqref{eq:fe:downwardsuminclusion} into a full CD for arbitrary box constraints by using decomposition maps like $\Pi_i$, for example by splitting $z^j = \min\{z^j,0\} + \max\{z^j,0\}$ and applying unilateral formulas like \eqref{eq:fe:unilateraldecompositionoperator} to $\min\{z^j,0\}$ and $\max\{z^j,0\}$ separately.
 In fact, the lower bound $\Pi_i (\min\{z^j,0\}) \ge \underline{\phi}^i$ holds, as shown in the above proof.
-However, the following Example shows why $\Pi_i(\min\{z^j,0\})$ may have an arbitrarily large maximum, and thus exceed any finite upper LDC $\overline{\phi}^i$, so unfortunately $\Pi_i$ does not map even non-positive functions to $\mathcal{D}^i$ in the presence of a finite upper LDC.
+However, the following Example shows why $\Pi_i(\min\{z^j,0\})$ may have an arbitrarily large maximum, and thus exceed any finite upper LDC $\overline{\phi}^i$.
+In fact, $\Pi_i$ does not map all non-positive functions into $\mathcal{D}^i$ in the presence of a finite upper LDC.
 
@@
 Recalling Table \ref{tab:ldcs}, note $\underline{\chi}^0=\maxR \underline{\chi}^1=-a$ also, so $\underline{\phi}^1=0$ and $\underline{\phi}^0=-a$.
-Suppose $z^1$ is a ``sawtooth'' function which takes on value $-a$ on the coarse nodes and value $0$ on the other fine-level nodes.
-Then $\overline{\chi}^1 \ge 0 \ge z^1\ge \underline{\chi}^1$, for any valid $\overline{\chi}^1$, so $z^j \in \mathcal{U}^1$.
+Suppose $z^1$ is a ``sawtooth'' function which takes on value $-a$ on the coarse nodes and value $0$ on the other nodes.
+Then, for any valid $\overline{\chi}^1$, we have $\overline{\chi}^1 \ge 0 \ge z^1\ge \underline{\chi}^1$, so $z^j \in \mathcal{U}^1$.
 However, $I_{1\to 0}^\ominus(z^1 - \underline{\chi}^1) = \minR(z^1 - \underline{\chi}^1) = 0$ identically.
@@
 But then $\Pi_1 z^1$ attains a maximum $a\ge 0$.
-This maximum can be chosen to exceed any upper defect obstacle $\overline{\phi}^1$.
-Thus, for general $w^1 \in \mathcal{U}^1$ and general LDCs, $\Pi_1$ does not map $\min\{w^1,0\}$ into $\mathcal{D}^1$.
+This maximum can be chosen to exceed any upper defect obstacle $\overline{\phi}^1$, so $\Pi_1 z^1 \notin \mathcal{D}^1$.
 \end{example}
@@
 
-Each upward defect constraint set $\mathcal{U}^j$ can also be decomposed down to an arbitrary level $k$ where an upward set $\mathcal{U}^k$ is used as the coarsest set.
-FIXME this will be used.
-While similar to Lemma \ref{lem:downwardadmissibility}, the next Lemma does not follow logically from it.
+Each upward defect constraint set $\mathcal{U}^j$ can also be decomposed down to an arbitrary coarser level.
+The incomplete CD in the next Lemma, which does not follow from Lemma \ref{lem:downwardadmissibility}, will justify upward admissibility in the V-cycle in the next Section.
 
@@
 \end{equation}
-In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, inclusion \eqref{eq:fe:upwardsuminclusion} is a full CD \eqref{eq:constraintdecomp}. \end{lemma}
+In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, inclusion \eqref{eq:fe:upwardsuminclusion} becomes a full CD. \end{lemma}
 
 \begin{proof}  Definition \eqref{eq:fe:philevels} shows that $\underline{\chi}^j = \underline{\phi}^j + \dots + \underline{\phi}^{k+1} + \underline{\chi}^k$ and $\overline{\chi}^j = \overline{\phi}^j + \dots + \overline{\phi}^{k+1} + \overline{\chi}^k$.
-(These facts replace \eqref{eq:fe:telescoping}.)  If $y^i \in \mathcal{D}^i$ for $k+1 \le i \le j$ and $z^k \in \mathcal{U}^k$ then it is easy to show, as in \eqref{eq:fe:lemmaordering}, that $\underline{\chi}^j \le y^j + \dots + y^{k+1} + z^k \le \overline{\chi}^j$, thus \eqref{eq:fe:upwardsuminclusion} holds.
+If $y^i \in \mathcal{D}^i$ for $k+1 \le i \le j$ and $z^k \in \mathcal{U}^k$ then it is easy to show, as in \eqref{eq:fe:lemmaordering}, that $\underline{\chi}^j \le y^j + \dots + y^{k+1} + z^k \le \overline{\chi}^j$, thus \eqref{eq:fe:upwardsuminclusion} holds.
 
 In unilateral cases we follow the proof of Lemma \ref{lem:downwardadmissibility}.
-Definition \eqref{eq:fe:minimummaps} is unchanged.
 We can use \eqref{eq:fe:unilateraldecompositionoperator} unchanged for $i=k+1,\dots,j$, but it must be modified in the $i=k$ case: $\Pi_k z^j = I_{j\to k}^\ominus(z^j - \underline{\chi}^j) + \underline{\chi}^k$.
-The remainder of the proof goes through as for Lemma \ref{lem:downwardadmissibility}.
+The remainder of the proof goes through as before.
 \end{proof}
 
-In summary, inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in V-cycle Algorithm \ref{alg:fascd}.
-Lemma \ref{lem:downwardadmissibility} explains downward admissibility, and Lemma \ref{lem:upwardadmissibility} upward admissibility, in Algorithm \ref{alg:fascd}.
-However, a convergence analysis following \cite{Tai2003} or \cite{GraeserKornhuber2009}, somehow extending their arguments to VIs \eqref{eq:vi} which are not of minimization type, and/or to box constraints, would seem to require full CDs, and perhaps new ideas.
+In summary, inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in the V-cycle algorithm which follows.
+However, a convergence analysis, perhaps extending the arguments of \cite{Tai2003} or \cite{GraeserKornhuber2009} to VIs \eqref{eq:vi} which are not of minimization type, and/or to box constraints, would seem to require full CDs, and perhaps new ideas, and we cannot offer that analysis.
 
@@
 
-The \fascd V-cycle algorithm in this section uses Lemmas \ref{lem:downwardadmissibility} and \ref{lem:upwardadmissibility}, that is, it smooths in $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.
-\fascd extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear FAS approach \cite{BrandtLivne2011}, and in fact it reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).
-
-The V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ based on perturbations from each coarser FE subspace $\mathcal{V}^j$.
-By nesting \eqref{eq:fe:nestedspaces} these corrections also live in the fine-level subspace $\mathcal{V}^J$, but we will use prolongation $P$ to indicate where the computer representation changes.
-Note that we coarsen both the residual and the solution iterate, in constrast to classical linear multigrid wherein only the residual is coarsened.
-
-However, the corrected iterate needs to be admissible.
-In detail, suppose $y^i \in \mathcal{D}^i$ for $i=J,\dots,j-1$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).
+The V-cycle algorithm in this section applies smoothers in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.
+It extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear full approximation storage (FAS) approach \cite{BrandtLivne2011}, and in fact it reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).
+
+Our V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ based on perturbations from each coarser FE subspace $\mathcal{V}^j$.
+By nesting \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$,\footnote{Note we use prolongation $P$ to indicate where the computer representation changes.} but the corrected iterate needs to be admissible.
+Suppose $y^i \in \mathcal{D}^i$ for $i=J,\dots,j-1$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).
 By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.
@@
 At the bottom the coarsest-level VI is then solved, resulting in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.
-Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$.
-Smoothing in $\mathcal{U}^1$ yields admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$ using \eqref{eq:fe:upwardsuminclusion}.
+Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$ using \eqref{eq:fe:upwardsuminclusion}.
 Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction.
-Then the next admissible iterate is $w^J + z^J \in \mathcal{K}^J$.
+Then the next admissible iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.
 
@@
 A key idea is that, because $\mathcal{U}^j\supset \mathcal{D}^j$, up-smoothing solves less-restricted VI problems than down-smoothing.
-Intuitively-speaking, going downward one does not know what the upcoming coarse corrections will be, so they must be in the sufficiently-small sets $\mathcal{D}^j$ so that their sum is admissible once we return to a level.
-However, going upward no further \emph{coarser} corrections, relative to that level, can violate admissibility.
-
-To state \fascd Algorithm \ref{alg:fascd} below we must clarify the problems solved at each mesh level.
+Intuitively-speaking, going downward one does not know what the next coarse corrections will be, thus they must be in small sets $\mathcal{D}^j$, so that their sum is admissible once we return to a given level.
+On the other hand, going upward no further coarse corrections can violate admissibility.
+
+To state the V-cycle algorithm below we must also clarify the problems solved at each mesh level.
 As in FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, and a source functional created.
-On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:boxdirichletvi}, presumably using quadrature and etc.~in the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.
-Note that finest-level problem \eqref{eq:fe:vi} uses the original source functional $\ell^J$, and an admissible iterate $w^J \in \mathcal{K}^J$ has already\footnote{Conceptually-speaking this is true.
-In practice the LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ are generated during the descent stage; see Algorithm \ref{alg:fascd}.
-Observe that the LDCs do not depend on $w^j$ for $j<J$, nor on the corrections $y^j,z^j$.} been used to general the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}).
-
-As we descend $j<J$, we define new $j$th-level iterates
-\begin{equation}
-w^j = \iR(w^{j+1} + y^{j+1}),  \label{eq:fe:definew}
-\end{equation}
-the value \emph{before} smoothing and/or correction at level $j$.  From $w^j$ we define $\ell^j \in (\cV^j)'$:
+On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:boxdirichletvi}, presumably using quadrature etc.~by the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.
+Note that finest-level problem \eqref{eq:fe:vi} uses the original source functional $\ell^J$.
+An admissible iterate $w^J \in \mathcal{K}^J$ has already\footnote{In practice the LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ are generated during the descent stage; see Algorithm \ref{alg:fascd}.} been used to generate the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}); these LDCs do not depend on the coarse-level corrections.
+
+As we descend we define new $j$th-level iterates
+\begin{equation}
+w^j = \iR(w^{j+1} + y^{j+1}).  \label{eq:fe:definew}
+\end{equation}
+(These are the states \emph{before} smoothing and/or correction.)  From $w^j$ we define $\ell^j \in (\cV^j)'$:
 \begin{equation}
@@
 \end{equation}
-Observe that injection $\iR$ coarsens the iterates---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---while canonical restriction $R$ coarsens linear functionals.
-(Recall Table \ref{tab:transfers} summarizes our transfer operators.)
+Observe that injection $\iR$ coarsens the iterates---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---but canonical restriction $R$ coarsens linear functionals.\footnote{Recall that Table \ref{tab:transfers} summarizes our transfer operators.}
 
@@
 for which the initial iterate is $y_{(0)}^j=0$.
-(Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)  On ascent, $j=0,1,\dots,J$, the same VI is solved, but in a larger admissible set, for $z^j \in \mathcal{U}^j$:
+(Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)  On ascent, $j=0,1,\dots,J$, the same VI is solved but for $z^j \in \mathcal{U}^j$, in a larger admissible set:
 \begin{equation}
@@
 \end{equation}
-The nontrivial initial iterate for this problem is found by prolonging the coarser-level correction $z^{j-1}$, and using it to correct the current (down-smoothed) correction on the level,\footnote{Prolongation $P$ is used in \eqref{eq:fe:upwardinitial} to reflect the change in FE representation, from $\mathcal{V}^{j-1}$ to $\mathcal{V}^{j}$.}
-\begin{equation}
-z_{(0)}^j = y^j + P z^{j-1},  \label{eq:fe:upwardinitial}
-\end{equation}
-but with $z_{(0)}^0=0$ on the coarsest level.  Note $z_{(0)}^j$ is admissible since $\mathcal{D}^j + \mathcal{U}^{j-1} \subset \mathcal{U}^j$; see \eqref{eq:fe:upwardsuminclusion}.
+The nontrivial initial iterate for this problem is found by prolonging $z^{j-1}$ and using it to correct the current (i.e.~down-smoothed) correction on the level:\footnote{Prolongation $P$ in \eqref{eq:fe:upwardinitial} reflects the change in FE representation.}
+\begin{equation}
+z_{(0)}^j = y^j + P z^{j-1}.  \label{eq:fe:upwardinitial}
+\end{equation}
+An exception is that $z_{(0)}^0=0$ on the coarsest level.  Note $z_{(0)}^j$ is admissible since $\mathcal{D}^j + \mathcal{U}^{j-1} \subset \mathcal{U}^j$; see \eqref{eq:fe:upwardsuminclusion}.
 
 If we were to inject the original box constraints \eqref{eq:fe:fineconstraintset} downward to coarser meshes, i.e.~$\underline{\gamma}^j = \iR \underline{\gamma}^{j+1}$ and $\overline{\gamma}^j = \iR \overline{\gamma}^{j+1}$, then, because injection preserves inequalities, and because $y^{j+1} \in \mathcal{D}^{j+1}$, it would follow from \eqref{eq:fe:definew} that $\underline{\gamma}^j \le w^j \le \overline{\gamma}^j$.
-In this sense $w^j$ is admissible with respect to the box constraints, but, following \cite{GraeserKornhuber2009}, our use of LDCs (Table \ref{tab:ldcs}) avoids the need to generate such box constraints $\underline{\gamma}^j$ and $\overline{\gamma}^j$ on each level.
-(Note that while $w^j \in \cV^j \subset \cV^J$ holds, generally $w^j \notin \cK^J$.)  In any case, Algorithm \ref{alg:fascd} will yield a final, finest-level iterate $w^J+z^J \in \cK^J$ which is genuinely admissible for problem \eqref{eq:fe:vi}.
-
-Regarding formula \eqref{eq:fe:levelsource}, the formula for source functional $\ell^j$ is a classical FAS construction.
-(Compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.)  It has the following explanation, mimicking the PDE version, which simultaneously justifies \eqref{eq:fe:levelsource} and VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.
-First, the $j+1$ level problem \eqref{eq:fe:downvi} is not solved exactly by the finer-level correction $y^{j+1}$---the down-smoother was only approximate---so we seek a further correction $y$ on the $j+1$ level so that
+In this sense $w^j$ is admissible with respect to the box constraints.
+However, while $w^j \in \cV^j \subset \cV^J$ holds, generally $w^j \notin \cK^J$, and furthermore, following \cite{GraeserKornhuber2009}, our use of LDCs (Table \ref{tab:ldcs}) avoids the need to generate these $\underline{\gamma}^j, \overline{\gamma}^j$.
+In any case, Algorithm \ref{alg:fascd} yields a new finest-level iterate $w^J+z^J \in \cK^J$ which is  admissible for problem \eqref{eq:fe:vi}.
+
+Observe that formula \eqref{eq:fe:levelsource} for functional $\ell^j$ is a classical FAS construction; compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.
+It has the following explanation, mimicking the PDE version, which simultaneously justifies VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.
+First, the $j+1$ level problem \eqref{eq:fe:downvi} is not solved exactly by the finer-level correction $y^{j+1}$---the down-smoother was only approximate---so we seek a further correction $y$ on the $j+1$ level such that
 \begin{equation}
@@
 \end{equation}
-for all $v\in \mathcal{D}^j$.  However, $y$ will actually be approximated from solving a coarser, $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps:
+for all $v\in \mathcal{D}^j$.  However, in fact $y$ will be approximated by solving a coarser, $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps:
 \begin{enumerate}
@@
 \item replace the residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction,
-\item replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its restriction,
-\item and, where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.
+\item replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its restriction, and
+\item where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.
 \end{enumerate}
-Steps \emph{(ii)}--\emph{(iv)} are justified when the residual and the error are smooth as a result of the progress made by the $j+1$ level down-smoother \cite[subsection 5.3.4, for example]{Trottenbergetal2001}.
-These steps now yield a problem for the $j$th-level correction $y^j \in \mathcal{D}^j$:
+Steps \emph{(ii)}--\emph{(iv)} are justified when the residual and the error are smooth; such smoothness should result from application of the down-smoother.
+These steps yield a problem for the $j$th-level correction $y^j \in \mathcal{D}^j$:
 \begin{align}
@@
 \end{align}
-This is the same as VI problem \eqref{eq:fe:downvi}, but in cluttered form; definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource} yields \eqref{eq:fe:downvi}.
-The explanation for VI \eqref{eq:fe:upvi} is essentially identical.
+This is the same as VI problem \eqref{eq:fe:downvi}, but in cluttered form; definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource} create \eqref{eq:fe:downvi}.
+The explanation for VI \eqref{eq:fe:upvi} is similar.
 
 We will regard the smoothers for VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} as approximate, in-place procedures, acting on the corrections $y^j$ and $z^j$, respectively (Section \ref{sec:implementation}).
-Also, as is standard in multigrid, at the coarsest $j=0$ level we suppose that problem \eqref{eq:fe:upvi} is solved quite accurately.
-However, this coarsest-level solver may be the same as the other smoothers, perhaps with a stronger tolerance.
-
-These ideas come together in Algorithm \ref{alg:fascd}.
-\pr{fascd-vcycle} acts in-place on $w^J$, and it assumes certain semantics for the \pr{smooth} and the coarsest-level \pr{solve} procedures.
-These do in-place modifications of their final arguments, without modifying their first four arguments.
-$\pr{smooth}^{\id{down}}$ and $\pr{smooth}^{\id{up}}$ denote \id{down} and \id{up} iterations, respectively.
+Also, as is standard in multigrid, at the coarsest $j=0$ level we suppose that problem \eqref{eq:fe:upvi} is solved accurately.
+
+These ideas come together in Algorithm \ref{alg:fascd}, \pr{fascd-vcycle}, which acts in-place on $w^J$.
+It assumes certain semantics for the \pr{smooth} and the coarsest-level \pr{solve} procedures; they do in-place modifications of their final arguments, without modifying their first four arguments.
+Parameters \id{down} and \id{up} determine the number of smoother iterations, respectively.
 
@@
 \end{pseudo}
-\caption{The \fascd V-cycle as an iteration for solving FE VI problem \eqref{eq:fe:vi}.  $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
+\caption{The \fascd V-cycle is an iteration for solving FE VI problem \eqref{eq:fe:vi}.  Here $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
 \label{alg:fascd}
@@
 On the finest level one must also store $\underline{\gamma}^J$ and $\overline{\gamma}^J$, while on the coarsest level there are only 5 vectors because $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.
-Therefore, using standard estimates for 2D refinement plus a geometric series argument \cite{Trottenbergetal2001}, the total floating-point storage, excluding the implementations of smoothers and the coarse-level solver, is
-\begin{equation}
-9 m_J + 7 m_{J-1} + \dots + 7 m_1 + 5 m_0 \approx \left(9 + 7 \sum_{k=1}^\infty \frac{1}{4^k}\right) m_J \approx 11 m_J.
-\end{equation}
-For comparison, a single-level method needs $4 m_J$ storage for the vectors $\underline{\gamma}^J,\overline{\gamma}^J,\ell^J,w^J$.
-
-If some of the obstacles are trivial, for instance if $\overline{\gamma}^J=+\infty$ in a unilateral obstacle problem, storage is correspondingly reduced at all levels.
-Furthermore, we will see in Section \ref{sec:results} that, though \pr{fascd-vcycle} defines a general $V(\pr{down},\pr{up})$ cycle for VI problems, in fact the $V(0,\pr{up})$ case generally produces the best performance.
-In that case the storage is reduced to about $9 m_J$ by neglecting the downward LDCs $\underline{\phi}^j$, $\overline{\phi}^j$ at all levels.
+Therefore, using $m_j=2^d m_{j-1}$, plus a geometric series argument, the total floating-point storage, excluding smoother and coarse-solver implementations, is
+\begin{equation}
+9 m_J + 7 m_{J-1} + \dots + 7 m_1 + 5 m_0 \le \left(9 + \frac{7}{2^d - 1}\right) m_J.
+\end{equation}
+This gives storage at most $16m_J$, $12m_J$, and $10m_J$ in dimensions $d=1,2,3$, respectively.
+For comparison, a single-level method needs $4 m_J$ storage.\footnote{That is, for vectors $\underline{\gamma}^J,\overline{\gamma}^J,\ell^J,w^J$.}  If some of the obstacles are trivial, for instance in the unilateral case $\overline{\gamma}^J=+\infty$, storage is reduced at all levels.
 
 \section{Implementation and smoothers} \label{sec:implementation}
+
+FIXME smoothers including semi-smooth and active set are motivated by equivalence of VIs and NCPs
 

### 2023-06-14 paper/fascd.tex
bits of progress and planning

@@
 \section{Multilevel CDs from level defect constraints} \label{sec:cdmultilevel}
+
+FIXME observe by example that using $\iR$ or $\minR/\maxR$ to directly generate coarse constaint sets will fail; this explains defect constraint idea
 
@@
 
-FIXME smoothers including semi-smooth and active set are motivated by equivalence of VIs and NCPs
-
-Finite-dimensional, box-constrained FE VI problem \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003,FerrisPang1997}, a strong form problem whose statement may be clarifying.
+In this section we consider several practical concerns regarding the implementation and application of Algorithm \ref{alg:fascd}.
+We address convergence criterion, smoothers and coarse solvers, and finally the F-cycle (sometimes called \emph{full multigrid cycle}) extension.
+
+Finite-dimensional, box-constrained FE VI problem \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}, a strong form whose statement may be clarifying.
 Denote $x_p$ for a node of the finest-level mesh $\mathcal{T}^J$ and $\psi_p(x)$ for the corresponding hat function.
-Let $\mathcal{X}=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}$.
+Let $\mathcal{X}^J=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}^J$.
 If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let us define $I_D$ as the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.
-Problem \eqref{eq:fe:vi}, which would be stated as VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap X\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP with additional equality constraints:
+Let us use the symbols $a\perp b$ to indicate that $a,b \in \RR^n$ are \emph{complementary}, that is, $a\ge 0$ and $b\ge 0$ and $ab = 0$ (entrywise).
+Problem \eqref{eq:fe:vi}, expressed as VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap \mathcal{X}^J\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP with additional equality constraints:
 \begin{align}
@@
 for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
-If $\underline{\gamma}^J(x_p) < \overline{\gamma}^J(x_p)$ then MCP \eqref{eq:fe:mcp} says that exactly one of the following five conditions holds for the solution $u^J$:
+
+MCP \eqref{eq:fe:mcp} says that exactly one of five conditions holds for the solution $u^J$ at $x_p \in \mathcal{T}^J$.  If $x_p \notin \partial_D\Omega$ then one of these four conditions hold:
 \begin{itemize}
-\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)<u^J(x_p)<\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} = 0$, or
-\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)=u^J(x_p)<\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \ge 0$, or
-\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)<u^J(x_p)=\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \le 0$, or
-\item $x_p \notin \partial_D\Omega$ and $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$, or
-\item $x_p \in \partial_D\Omega$ and $u^J(x_p)=g_D^J(x_p)$ and $\lambda_p=\ip{f^J(u^J)-\ell^J}{\psi_p}$.
+\item $\underline{\gamma}^J(x_p)<u^J(x_p)<\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} = 0$,
+\item $\underline{\gamma}^J(x_p)=u^J(x_p)<\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \ge 0$,
+\item $\underline{\gamma}^J(x_p)<u^J(x_p)=\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \le 0$,
+\item $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$.
+\end{itemize}
+Otherwise, if $x_p \in \partial_D\Omega$ then
+\begin{itemize}
+\item $u^J(x_p)=g_D^J(x_p)$ and $\lambda_p=\ip{f^J(u^J)-\ell^J}{\psi_p}$.
 \end{itemize}
 The cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, \emph{pinched}, and \emph{boundary}, respectively.
-Where case \emph{pinched} applies $\ip{f^J(u^J)-\ell^J}{\psi_p}$ can have any real value, and similarly if \emph{boundary} applies then $\lambda_p$ is unconstrained.
-
-Repeated application of \pr{fascd-vcycle} to solve VI \eqref{eq:fe:vi} will not reduce the finest-level residual $f^J(w^J) - \ell^J$ to zero everywhere because the problem is inequality-constrained.
-We need to identify the vector quantity whose norm must be controlled by a tolerance.
-Recalling mixed complementarity problem \eqref{eq:fe:mcp}, and denoting the hat function at node $x_p \in \mathcal{T}^J$ by $\psi_p$, for $w^J \in \mathcal{K}^J$ we define the finest-level \emph{nodal complementarity residual} in $\RR^{m_J}$.
+Where case \emph{pinched} applies, $\ip{f^J(u^J)-\ell^J}{\psi_p}$ can have any real value.
+
+Because the problem is inequality-constrained, repeated application of \pr{fascd-vcycle} to solve VI \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.
+However, MCP \eqref{eq:fe:mcp} identifies candidate quantities which should go to zero, and which we can compare to a tolerance.
+FIXME first, for $w^J \in \mathcal{K}^J$ we define the finest-level \emph{nodal complementarity residual} in $\RR^{m_J}$.
 It is defined in the four cases listed after \ref{eq:fe:mcp}, namely inactive, lower active, upper active, and boundary:
@@
 
+FIXME Fischer-Burmeister semismooth alternative for convergence
+
 A solver for \eqref{eq:fe:vi} using repeated applications of \pr{fascd-vcycle} would start with an initial iterate $w_{(0)}^J$, do V-cycles, and stop once the finest-level nodal residual $\hat r(w^J)$ was sufficiently small.
@@
 The \emph{upward nodal residual} $\hat r_{\text{up}}(z^j)$ for $z^j \in \mathcal{U}^j$ is similarly defined using $\chi^j$.
+
+FIXME smoothers including semi-smooth and active set \cite{BensonMunson2006}
 

### 2023-06-15 paper/fascd.tex
work on NCP residual text; we need a both-bounds test case

@@
 \newcommand{\fascd}{\pr{fascd}\xspace}
+
+\newcommand{\rNC}{r_{\text{NC}}}
+\newcommand{\rSS}{r_{\text{SS}}}
 
@@
 
-In this section we consider several practical concerns regarding the implementation and application of Algorithm \ref{alg:fascd}.
-We address convergence criterion, smoothers and coarse solvers, and finally the F-cycle (sometimes called \emph{full multigrid cycle}) extension.
-
-Finite-dimensional, box-constrained FE VI problem \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}, a strong form whose statement may be clarifying.
-Denote $x_p$ for a node of the finest-level mesh $\mathcal{T}^J$ and $\psi_p(x)$ for the corresponding hat function.
+In this section we consider several practical concerns when implementing and applying \pr{fascd-vcycle} Algorithm \ref{alg:fascd}.
+We address convergence criteria, smoothers, and coarse solvers, plus the F-cycle extension of the Algorithm, sometimes called a \emph{full multigrid cycle}.
+
+Because the problem is inequality-constrained, repeated application of \pr{fascd-vcycle} to solve the finite-dimensional, box-constrained VI \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.
+However, \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}, a strong form amenable to convergence criteria.
+Let $x_p \in \mathcal{T}^J$ be a finest-level node, and $\psi_p(x)$ the corresponding hat function.
 Let $\mathcal{X}^J=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}^J$.
-If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let us define $I_D$ as the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.
-Let us use the symbols $a\perp b$ to indicate that $a,b \in \RR^n$ are \emph{complementary}, that is, $a\ge 0$ and $b\ge 0$ and $ab = 0$ (entrywise).
-Problem \eqref{eq:fe:vi}, expressed as VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap \mathcal{X}^J\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP with additional equality constraints:
+If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let $I_D$ be the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.
+Finally, we will write $a\perp b$ when $a,b \in \RR^n$ are \emph{complementary}, that is, when $a_i \ge 0$ and $b_i \ge 0$ and $a_i b_i = 0$ for $i=0,\dots,n-1$.
+Then problem \eqref{eq:fe:vi}, which would be expressed VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap \mathcal{X}^J\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP for the solution $u^J \in \mathcal{K}^J$, but with additional equality constraints:
 \begin{align}
@@
 for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
-
-MCP \eqref{eq:fe:mcp} says that exactly one of five conditions holds for the solution $u^J$ at $x_p \in \mathcal{T}^J$.  If $x_p \notin \partial_D\Omega$ then one of these four conditions hold:
+By definition, MCP \eqref{eq:fe:mcp} says that certain node-wise conditions hold for $u^J$ at $x_p \in \mathcal{T}^J$.
+If $x_p \notin \partial_D\Omega$ then one of these four conditions hold:
 \begin{itemize}
@@
 \end{itemize}
-The cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, \emph{pinched}, and \emph{boundary}, respectively.
-Where case \emph{pinched} applies, $\ip{f^J(u^J)-\ell^J}{\psi_p}$ can have any real value.
-
-Because the problem is inequality-constrained, repeated application of \pr{fascd-vcycle} to solve VI \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.
-However, MCP \eqref{eq:fe:mcp} identifies candidate quantities which should go to zero, and which we can compare to a tolerance.
-FIXME first, for $w^J \in \mathcal{K}^J$ we define the finest-level \emph{nodal complementarity residual} in $\RR^{m_J}$.
-It is defined in the four cases listed after \ref{eq:fe:mcp}, namely inactive, lower active, upper active, and boundary:
-\begin{equation}
-\hat r(w^J)_p = \begin{cases}
-    \ip{f^J(w^J)-\ell^J}{\psi_p}, & \underline{\gamma}^J(x_p) < w^J(x_p) < \overline{\gamma}^J(x_p), \\
-    \min\{\ip{f^J(w^J)-\ell^J}{\psi_p},0\}, & w^J(x_p) = \underline{\gamma}^J(x_p), \\
-    \max\{\ip{f^J(w^J)-\ell^J}{\psi_p},0\}, & w^J(x_p) = \overline{\gamma}^J(x_p), \\
-    w^J(x_p) - g_D^J(x_p), & x_p \in \partial_D\Omega. \end{cases} \label{eq:cpresidual}
-\end{equation}
-Where the box constraints pinch at $x_p \notin \partial_D\Omega$, i.e.~where $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$, we define $\hat r(w^J)_p=0$.
-Note that $\hat r:\mathcal{V}^J \to \RR^{m_J}$ is a computable nonlinear map derived from $f^J$ and the vectors $\ell^J$, $\underline{\gamma}^J$, and $\overline{\gamma}^J$.
-For unconstrained PDEs one would use only the inactive and boundary cases.
-
-The idea here is that $\hat r(w^J)$ is nonzero only if there are nodes where MCP \eqref{eq:fe:mcp} is violated, either because the pointwise residual $\ip{f^J(w^J)-\ell^J}{\psi_p}$ remains nonzero where the constraints are inactive, or because it has the wrong sign on an actively-constrained degree of freedom, or because the Dirichlet boundary condition is not satisfied.
-
-FIXME Fischer-Burmeister semismooth alternative for convergence
+The cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, \emph{pinched} (\emph{both active}), and \emph{boundary}, respectively.
+Where \emph{pinched} applies, $\ip{f^J(u^J)-\ell^J}{\psi_p}$ can have any real value.
+
+Thus the MCP \eqref{eq:fe:mcp} identifies expressions whose norms should go below tolerance.
+For for $w^J \in \mathcal{K}^J$, let $r_p = \ip{f^J(w^J)-\ell^J}{\psi_p}$, the pointwise (ordinary) residual, as a notational convenience.
+First, interpreting the MCP directly, we define the finest-level \emph{nodal complementarity residual} $\rNC$ in $\RR^{m_J}$ from the five cases:
+\begin{equation}
+\rNC(w^J)_p = \begin{cases}
+    r_p, & \underline{\gamma}^J(x_p) < w^J(x_p) < \overline{\gamma}^J(x_p), \\
+    \min\{r_p,0\}, & w^J(x_p) = \underline{\gamma}^J(x_p), \\
+    \max\{r_p,0\}, & w^J(x_p) = \overline{\gamma}^J(x_p), \\
+    0, & \underline{\gamma}^J(x_p)=w^J(x_p)=\overline{\gamma}^J(x_p), \\
+    w^J(x_p) - g_D^J(x_p), & x_p \in \partial_D\Omega. \end{cases} \label{eq:rNC}
+\end{equation}
+(The first four cases assume $x_p \notin \partial_D\Omega$.)  Note that $\rNC : \mathcal{K}^J \to \RR^{m_J}$ is a computable nonlinear map defined from $f^J$ and the vectors $\ell^J$, $\underline{\gamma}^J$, and $\overline{\gamma}^J$, and that for PDEs without constraints one uses only the first and last cases.
+In summary, $\rNC(w^J)_p$ is nonzero at nodes where MCP \eqref{eq:fe:mcp} is violated, either because the pointwise residual $\ip{f^J(w^J)-\ell^J}{\psi_p}$ remains nonzero where the constraints are inactive, or because it has the wrong sign at an actively-constrained degree of freedom, or because the boundary condition is not satisfied.
+
+As a practical matter, however, direct implementation of \eqref{eq:rNC} is sensitive to rounding error because $\rNC$ is a discontinuous function of $w^J(x_p)$.
+For example, if $w^J(x_p)$ exceeds $\underline{\gamma}^J(x_p)$ by $10^{-16}$, then, instead of returning the minimum of $z_p = \left<f^J(w^J)-\ell^J,\psi_p\right>$ and zero (the second case), the first case applies and $\rNC(w^J)_p = z_p$, which might be large (positive).
+This would suggest that the iteration is far from converging.
+On the other hand, an advantage of formula \eqref{eq:rNC} is that it applies as stated if $\underline{\gamma}^J(x_p)=-\infty$ or $\overline{\gamma}^J(x_p)=+\infty$; compared to \eqref{eq:rSS} below, the difference $w^J(x_p) - \underline{\gamma}^J(x_p)$, for example, is never computed.
+
+%\newcommand{\phiFB}{\phi_{\text{FB}}}
+We will instead measure convergence using a \emph{semi-smooth residual} based on the Fischer-Burmeister function \cite{BensonMunson2006,Ulbrich2011} which is defined as follows for $a,b\in\RR$:
+\begin{equation}
+\phi(a,b) = a + b - \sqrt{a^2 + b^2}. \label{eq:phiFB}
+\end{equation}
+Observe that $\phi(a,b)=0$ if and only if $a,b$ are complementary, and also that $\phi$ semi-smooth (and continuous).  Now let
+\begin{equation}
+\rSS(w^J)_p = \begin{cases}
+r_p, & \underline{\gamma}^J(x_p) = -\infty \text{ and }\overline{\gamma}^J(x_p) = +\infty, \\
+\phi\left(w^J(x_p) - \underline{\gamma}^J(x_p), r_p\right), & \overline{\gamma}^J(x_p) = +\infty, \\
+\phi\left(\overline{\gamma}^J(x_p) - w^J(x_p), -r_p\right), & \underline{\gamma}^J(x_p) = -\infty, \\
+\text{FIXME: this case needs testing} \\
+w^J(x_p) - g_D^J(x_p), & x_p \in \partial_D\Omega.
+\end{cases} \label{eq:rSS}
+\end{equation}
 
@@
 
-FIXME  Recall that a $j$th-level hat function based at node $x_p^j$ is denoted $\psi_p^j$, for an iterate $g^j$ and an admissible correction $y^j \in \mathcal{D}^j$ we define the \emph{downward nodal residual} as
-\begin{equation}
-(\hat r_{\text{down}}(y^j))_p = \begin{cases} \ip{f^j(g^j+y^j)-\ell^j}{\psi_p^j}, & y^j(x_p^j) > \phi^j(x_p^j), \\
-                                  \min\{\ip{f^j(g^j+y^j)-\ell^j}{\psi_p^j},0\}, & y^j(x_p^j) = \phi^j(x_p^j). \end{cases} \label{eq:dncpresidual}
-\end{equation}
-The \emph{upward nodal residual} $\hat r_{\text{up}}(z^j)$ for $z^j \in \mathcal{U}^j$ is similarly defined using $\chi^j$.
-
-FIXME smoothers including semi-smooth and active set \cite{BensonMunson2006}
+FIXME smoother semantics; shift ``$w^j+y^j$'' etc.
+
+FIXME smoothers including Newton active set \cite{BensonMunson2006} or Newton semi-smooth \cite{BensonMunson2006}, with either direct linear solve or a fixed number of Krylov steps
 

### 2023-06-16 paper/fascd.tex
finish write-up of semismooth NCP residual, and convergence

@@
 \newcommand{\rSS}{r_{\text{SS}}}
+\newcommand{\rXX}{r_{\text{XX}}}
 
@@
 As a practical matter, however, direct implementation of \eqref{eq:rNC} is sensitive to rounding error because $\rNC$ is a discontinuous function of $w^J(x_p)$.
-For example, if $w^J(x_p)$ exceeds $\underline{\gamma}^J(x_p)$ by $10^{-16}$, then, instead of returning the minimum of $z_p = \left<f^J(w^J)-\ell^J,\psi_p\right>$ and zero (the second case), the first case applies and $\rNC(w^J)_p = z_p$, which might be large (positive).
-This would suggest that the iteration is far from converging.
-On the other hand, an advantage of formula \eqref{eq:rNC} is that it applies as stated if $\underline{\gamma}^J(x_p)=-\infty$ or $\overline{\gamma}^J(x_p)=+\infty$; compared to \eqref{eq:rSS} below, the difference $w^J(x_p) - \underline{\gamma}^J(x_p)$, for example, is never computed.
+For example, if $w^J(x_p)$ exceeds $\underline{\gamma}^J(x_p)$ by $10^{-16}$, then, instead of returning the minimum of $z_p = \left<f^J(w^J)-\ell^J,\psi_p\right>$ and zero (the second case), the first case applies and $\rNC(w^J)_p = z_p$, which might be large (positive), and this would suggest falsely that the iteration is far from converging.
+On the other hand, an advantage of formula \eqref{eq:rNC} is that it applies as stated if $\underline{\gamma}^J(x_p)=-\infty$ or $\overline{\gamma}^J(x_p)=+\infty$; compare \eqref{eq:rSS} below, wherein the difference $w^J(x_p) - \underline{\gamma}^J(x_p)$ is computed, for example.
 
 %\newcommand{\phiFB}{\phi_{\text{FB}}}
-We will instead measure convergence using a \emph{semi-smooth residual} based on the Fischer-Burmeister function \cite{BensonMunson2006,Ulbrich2011} which is defined as follows for $a,b\in\RR$:
+We will instead measure convergence using a \emph{semi-smooth residual} based on the Fischer-Burmeister function \cite{BensonMunson2006,Ulbrich2011}, a function defined as follows for $a,b\in\RR$:
 \begin{equation}
@@
 \end{equation}
-Observe that $\phi(a,b)=0$ if and only if $a,b$ are complementary, and also that $\phi$ semi-smooth (and continuous).  Now let
+Observe that $\phi(a,b)=0$ if and only if $a,b$ are complementary, and also that $\phi$ semi-smooth (and thus continuous).
+For convenience denote $\underline{g}_p = w^J(x_p) - \underline{\gamma}^J(x_p)$ and $\overline{g}_p = \overline{\gamma}^J(x_p) - w^J(x_p)$, the positive gaps at $x_p$.
+Finally define
 \begin{equation}
 \rSS(w^J)_p = \begin{cases}
-r_p, & \underline{\gamma}^J(x_p) = -\infty \text{ and }\overline{\gamma}^J(x_p) = +\infty, \\
-\phi\left(w^J(x_p) - \underline{\gamma}^J(x_p), r_p\right), & \overline{\gamma}^J(x_p) = +\infty, \\
-\phi\left(\overline{\gamma}^J(x_p) - w^J(x_p), -r_p\right), & \underline{\gamma}^J(x_p) = -\infty, \\
-\text{FIXME: this case needs testing} \\
-w^J(x_p) - g_D^J(x_p), & x_p \in \partial_D\Omega.
+r_p, & \underline{\gamma}^J(x_p) = -\infty \text{ and } \overline{\gamma}^J(x_p) = +\infty, \\
+\phi\big(\underline{g}_p, r_p\big), & \overline{\gamma}^J(x_p) = +\infty, \\
+\phi\left(\overline{g}_p, -r_p\right), & \underline{\gamma}^J(x_p) = -\infty, \\
+\max\big\{\phi\big(\underline{g}_p, r_p\big), \phi\left(\overline{g}_p, -r_p\right)\big\}, & \text{otherwise}, \\
+w^J(x_p) - g_D^J(x_p), & x_p \in \partial_D\Omega,
 \end{cases} \label{eq:rSS}
 \end{equation}
-
-A solver for \eqref{eq:fe:vi} using repeated applications of \pr{fascd-vcycle} would start with an initial iterate $w_{(0)}^J$, do V-cycles, and stop once the finest-level nodal residual $\hat r(w^J)$ was sufficiently small.
-For example, using tolerances \id{atol}$>0$ and \id{rtol}$>0$ the stopping criterion might be
-\begin{equation}
-\|\hat r(w^J)\| < \id{atol} \qquad \text{or} \qquad \frac{\|\hat r(w^J)\|}{\|\hat r((w^J)^{(0)})\|} < \id{rtol},
-\end{equation}
-for some norm $\|\cdot\|$ on $\RR^{m_J}$.  A step tolerance could also be used, measuring the difference between successive iterates.
+where again $x_p \notin \partial_D\Omega$ in the first four cases.
+
+Now suppose we repeatedly apply \pr{fascd-vcycle}, starting with $w^{J,0} \in \mathcal{K}^J$, generating (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.
+Let $\id{atol}\ge 0$, $\id{rtol}\ge 0$ be given tolerances.
+Then the criterion used in the results (Section \ref{sec:results}) is
+\begin{equation}
+\|\rSS(w^{J,k})\|_2 < \id{atol} \qquad \text{or} \qquad \frac{\|\rSS(w^{J,k})\|_2}{\|\rSS(w^{J,0})\|_2} < \id{rtol}
+\end{equation}
+A step tolerance, based on $\|w^{J,k} - w^{J,k-1}\|_2$, could also be used, but we did not find it necessary.
 

### 2023-06-16 paper/fascd.tex
clean appendix

@@
 \section{Introduction} \label{sec:intro}
+
+FIXME rewrite the entire introduction
 
@@
 
-FIXME smoothers including Newton active set \cite{BensonMunson2006} or Newton semi-smooth \cite{BensonMunson2006}, with either direct linear solve or a fixed number of Krylov steps
+FIXME smoothers including Newton active set \cite{BensonMunson2006} or Newton semi-smooth \cite{BensonMunson2006}, with either direct linear solve or a fixed number of Krylov steps; in latter case either fixed number of CG+ICC or fixed number of GMRES+ILU
 
@@
 
-FIXME show results for 2D Firedrake implementation of FASCD for $p$-Laplacian obstacle problem (Example \ref{ex:plaplacian}), advection-diffusion obstacle problem (Example \ref{ex:advectiondiffusion}), and porous-media obstacle problem (Example \ref{ex:porous})
+FIXME show results for Firedrake implementation of FASCD for the 2D classical obstacle problem (ball and spiral), 1D $p$-Laplacian obstacle problem (Example \ref{ex:plaplacian}), advection-diffusion obstacle problem (Example \ref{ex:advectiondiffusion}), and SIA obstacle problem
 
@@
 
-FIXME constrast the implemented examples in \cite{GraeserKornhuber2009,Tai2003} in which $f$ is linear
-
 % A BRIDGE TOO FAR:  \section{Results for a nonlocal variational inequality}
@@
 \appendix
-\section{Reductions of FASCD to familiar multilevel algorithms} \label{app:reductions}
-
-\pr{fascd-vcycle} (Algorithm \ref{alg:fascd}) in the main text generalizes two better-known multilevel algorithms.
-First we show that it reduces to the FAS multigrid V-cycle for nonlinear PDEs \cite{Trottenbergetal2001}, which itself generalizes linear multigrid, when inequality constraints are removed.
-Second we reduce it to an existing V-cycle for unilateral linear variational inequalities \cite{GraeserKornhuber2009} in the case that the operator is linear and local.
-Thus this Appendix separates FAS/nonlinear aspects of the algorithm from those modifications of linear multigrid which are caused by the inequality constraints.
-
-To derive the classical FAS multigrid V-cycle we remove the inequality constraints from the finest-level VI problem \eqref{eq:fe:vi} and then simplify the \pr{fascd-vcycle} pseudocode accordingly.
-In fact, suppose that the finest-level obstacles $\underline{\gamma}^J$ and $\overline{\gamma}^J$ are very negative and positive, respectively, thus not constraining the solution $u^J$ or any possible iterate.
+\section{Reduction of FASCD to FAS multigrid for PDEs} \label{app:reductions}
+
+\pr{fascd-vcycle} (Algorithm \ref{alg:fascd}) in the main text generalizes existing multilevel algorithms.
+Here we show that it reduces to the FAS multigrid V-cycle for nonlinear PDEs \cite{Trottenbergetal2001}, which itself generalizes linear multigrid, when all inequality constraints are removed.
+
+Suppose we remove the inequality constraints from the finest-level VI problem \eqref{eq:fe:vi} by replacing the finest-level obstacles $\underline{\gamma}^J$ and $\overline{\gamma}^J$ with very negative and positive constants, respectively, thus not constraining the solution $u^J$ or any possible iterate.
 The admissible set $\mathcal{K}^J$ of \eqref{eq:fe:vi} is then determined only by the Dirichlet boundary conditions.
@@
 the FE discretization of strong-form PDE $f(u)=\ell$.
-The downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets can also be enlarged so that the corrections $y^j$ and $z^j$ in \pr{fascd-vcycle} are never constrained.
-Formally we might do this by choosing constant LDCs with $\underline{\chi}^j \ll 0 \ll \overline{\chi}^j$, but so that ordering property \eqref{eq:fe:chiordering} still holds, and also with large gaps so that $\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1} \ll 0 \ll \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}$.
-Thus the corrections $y^j$ and $z^j$ are unconstrained, and lines 2--3 and 5--8 can be removed from Algorithm \ref{alg:fascd}.
+The downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets can also be enlarged so that the corrections $y^j$ and $z^j$ in \pr{fascd-vcycle} are never constrained.\footnote{Formally we might do this by choosing constant LDCs with $\underline{\chi}^j \ll 0 \ll \overline{\chi}^j$, but so that ordering property \eqref{eq:fe:chiordering} still holds, and also with large gaps so that $\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1} \ll 0 \ll \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}$.}  Thus the corrections $y^j$ and $z^j$ are unconstrained, and lines 2--3 and 5--8 can be removed from Algorithm \ref{alg:fascd}.
 
 Now the downward/upward VIs \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} are also unconstrained, and can be rewritten as weak forms $\ip{f^j(w^j + y^j)}{v} = \ip{\ell^j}{v}$ and $\ip{f^j(w^j + z^j)}{v} = \ip{\ell^j}{v}$, respectively, for all $v\in \mathcal{V}_0^j$.
-However, we may simply refer to the unconstrained, approximate solution $u^j=w^j+y^j$, instead of $w^j$ and $y^j$ separately.
-The smoothers and the coarsest-level solver can be regarded as unconstrained, in-place modifications of $u^j$, so we denote them simply as ``$\text{\pr{smooth}}(\ell^j;u^j)$.''  Noting line 11 of Algorithm \ref{alg:fascd} says $u^{j-1}=\iR u^j$, after down-smoothing the coarsened source functional computed by line 12 can be rewritten as
+However, we may simply store the unconstrained, approximate solution $u^j=w^j+y^j$, instead of $w^j$ and $y^j$ separately, and similarly for the upward solution $u^j=w^j+z^j$.
+The smoothers and the coarsest-level solver can be regarded as unconstrained, in-place modifications of $u^j$, so we denote them simply as ``$\text{\pr{smooth}}(\ell^j;u^j)$.''  Noting line 11 of Algorithm \ref{alg:fascd} now says $u^{j-1}=\iR u^j$, after down-smoothing the coarsened source functional computed by line 12 can be rewritten as
 \begin{equation}
@@
 
-Going upward we must be careful that only the coarse correction, and not the coarse solution, is prolonged.\footnote{Compare Remark 5.3.9 in \cite{Trottenbergetal2001}, a warning to novices.}  After up-smoothing on the $j-1$ level the coarse iterate is $u^{j-1}$, but at this point $u^j=w^j + y^j$ contains the result of down-smoothing on the $j$th level.
+Going upward we must be careful that only the coarse correction, and not the coarse solution, is prolonged.\footnote{Compare Remark 5.3.9 in \cite{Trottenbergetal2001}, a warning to novices and everyone else.}  After up-smoothing on the $j-1$ level the coarse iterate is $u^{j-1}$, but at this point $u^j=w^j + y^j$ contains the result of down-smoothing on the $j$th level.
 On the other hand, in Algorithm \ref{alg:fascd} the initial value of the correction $z^j$, before up-smoothing on the $j$th level, is $z^j = y^j + P z^{j-1}$, where $z^{j-1}$ contains the total correction from coarser corrections and up-smoothing at the $j-1$ level.
 Thus the $j$th-level iterate \emph{before} up-smoothing must be set to
-    $$w^j + y^j + P z^{j-1} = u^j + P(u^{j-1} - \iR u^j),$$
-a standard FAS assignment.
-
-With the given modifications, we have the much-simplified pseudocode in Algorithm \ref{alg:fas}.
-This FAS V-cycle algorithm appears in various places, for example Algorithm 14 of \cite{Bruneetal2015} or in section 5.3.4 of \cite{Trottenbergetal2001}.
+    $$w^j + y^j + P z^{j-1} = u^j + P(u^{j-1} - \iR u^j).$$
+
+With these modifications, we have a much-simplified pseudocode, Algorithm \ref{alg:fas}.
+This FAS V-cycle appears in various places, for example Algorithm 14 of \cite{Bruneetal2015} or in section 5.3.4 of \cite{Trottenbergetal2001}.
 
@@
 \end{pseudo}
-\caption{Removing inequality constraints from \pr{fascd-vcycle} yields the standard FAS V-cycle.}
+\caption{The standard FAS V-cycle for a discretized PDE $f^J(u^J)=\ell^J$.}
 \label{alg:fas}
@@
 
-FIXME by assuming $f$ is linear, reduce to Algorithm 4.7 in \cite{GraeserKornhuber2009} but with $\text{V}(\nu_1,\nu_2)$ cycles
-
 \section{From slack, to not lose} \label{app:slack}
 
-FIXME move this from this appendix, after using it in the discussion etc.
+FIXME remove this appendix, after using it in the introduction and discussion etc.
 

### 2023-06-16 paper/fascd.tex
text, example, figure which make Section 6 solid

@@
 
-FIXME observe by example that using $\iR$ or $\minR/\maxR$ to directly generate coarse constaint sets will fail; this explains defect constraint idea
-
-Suppose $w^J \in \cK^J$ is an admissible finest-level iterate which approximates the solution $u^J$ of FE VI \eqref{eq:fe:vi}.
-From $w^J$ we will now define CD constraint subsets using the defining inequalities \eqref{eq:fe:fineconstraintset} for $\cK^J$ and the above monotone restrictions $\minR, \maxR$.
-Though the ideas are fundamentally the same as for unilateral constraints \cite{GraeserKornhuber2009}, the notation is necessarily more elaborate for box-constrained problems.
-
-We define the \emph{finest-level defect constraints} \cite{GraeserKornhuber2009} as the following nodal functions in $\tilde{\mathcal{V}}^J$:
+Our practical multilevel CD scheme in this Section uses \emph{defect constraints} \cite{GraeserKornhuber2009} which are based on an admissible finest-level iterate.
+The next Example concretely illustrates why we do not apply restriction operators directly to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$ in definition \eqref{eq:fe:fineconstraintset}.
+Our straightforward iterate-dependent bounds, namely defect constraints \eqref{eq:fe:defectconstraints} below, bypass the difficulties exposed in Example \ref{ex:directRbad}.
+However, other successful schemes may exist.
+
+\begin{example}  \label{ex:directRbad}  
+For concreteness, suppose $\Omega = [0,L] \subset \RR^1$ is an interval and consider a two-level mesh hierarchy ($J=1$).
+Let $\underline{\gamma}^1$ be a sawtooth function with minimum value zero, and suppose $\overline{\gamma}^1=C+\underline{\gamma}^1$ with $C>0$ which is strictly less than the range of $\underline{\gamma}^1$.
+Supposing empty Dirichlet boundary for simplicity, the constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.
+Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ nodes, respectively.
+However, various restriction schemes applied directly to $\underline{\gamma}^1,\overline{\gamma}^1$ are problematical.
+Suppose we propose
+    $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \iR \underline{\gamma}^1, \iR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
+Then $\underline{\gamma}^0=0$ and $\overline{\gamma}^0=C$ identically.
+In that case the corresponding constraint set $\mathcal{K}^0$ is nonempty, but prolongation $Pw^0$ of an admissible solution of the coarse problem ($\underline{\gamma}^0 \le w^0 \le \overline{\gamma}^0$) is not admissible on the finer level: $Pw^0 \notin \mathcal{K}^1$.
+On the other hand, suppose we propose
+    $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
+The general motivation for this choice---compare \eqref{eq:fe:chilevels}---is that $w^0 \ge \underline{\gamma}^0 \, \implies \, Pw^0 \ge \underline{\gamma}^1$ and $w^0 \le \overline{\gamma}^0 \, \implies \, Pw^0 \le \overline{\gamma}^1$.
+However, while these facts are true, the difficulty should be clear: the coarse-level admissible set $\mathcal{K}^0$ is now empty because the proposed lower bound $\underline{\gamma}^0$ actually exceeds the proposed upper bound $\overline{\gamma}^0$.
+\end{example}
+
+\begin{figure}[ht]
+\begin{center}
+\input{tikz/directRbad.tex}
+\end{center}
+\caption{For finest-level box constraints $\underline{\gamma}^1,\overline{\gamma}^1$ like these, direct restriction using any of $\iR,\minR,\maxR$, to a coarser mesh (solid dots), is problematic for admissibility of coarse corrections.}
+\label{fig:directRbad}
+\end{figure}
+
+Subtraction of an admissible iterate provides a route out of these difficulties.
+Suppose $w^J \in \cK^J$ is an admissible finest-level iterate (i.e.~which approximates the solution $u^J$ of \eqref{eq:fe:vi}).
+We define the \emph{finest-level defect constraints} (DCs) \cite{GraeserKornhuber2009} as the following nodal functions in $\tilde{\mathcal{V}}^J$:
 \begin{equation}
@@
 \end{equation}
-From definition \eqref{eq:fe:fineconstraintset}, note that $\underline{\chi}^J \le 0 \le \overline{\chi}^J$ everywhere in $\overline{\Omega}$, including along the Dirichlet boundary $\partial_D\Omega$.
-Also, if $\underline{\gamma}^J=-\infty$ then $\underline{\chi}^J=-\infty$ by definition, and similarly if $\overline{\gamma}^J=+\infty$ then $\overline{\chi}^J=+\infty$.
-The motivation for creating $\underline{\chi}^J,\overline{\chi}^J$ is that a corrected iterate $w^J + z^J$ is admissible, $w^J + z^J \in \cK^J$, if and only if the perturbation $z^J$ is between the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$, and $z^J=0$ on $\partial_D\Omega$.
-
-Next, counting down the coarser levels $j=J-1,\dots,1,0$, we apply the monotone restrictions to generate box \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:
+Definition \eqref{eq:fe:fineconstraintset} implies that $\underline{\chi}^J \le 0 \le \overline{\chi}^J$ everywhere in $\overline{\Omega}$, including along the Dirichlet boundary $\partial_D\Omega$.
+If $\underline{\gamma}^J=-\infty$ then $\underline{\chi}^J=-\infty$ by definition, and similarly if $\overline{\gamma}^J=+\infty$ then $\overline{\chi}^J=+\infty$.
+Observe that a corrected iterate $w^J + z^J$ is admissible, $w^J + z^J \in \cK^J$, if and only if the perturbation $z^J$ is zero on $\partial_D\Omega$ and it is bounded by the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$.
+
+Descending coarser levels, $j=J-1,\dots,1,0$, we apply the monotone restrictions to generate box \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:
 \begin{equation}
@@
 A unilateral $J=4$ example of LDCs $\underline{\chi}^j$ is shown in Figure \ref{fig:chiphilevels}.
-Intuitively-speaking, the LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if another element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$.
-Similar comments apply to $\overline{\chi}^{j}$.
+The LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if another element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$, and similar comments apply to $\overline{\chi}^{j}$.
 
@@
 This construction of multilevel box constraints is designed to permit the largest admissible coarse corrections, which promotes multilevel solver efficiency.
-Note that for the multilevel schemes in Sections \ref{sec:vcycle} and \ref{sec:implementation}, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not required.
-
-We now compute the LDC differences:
+Note that for the multilevel schemes in Sections \ref{sec:vcycle} and \ref{sec:implementation}, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are actually not required.
+Also, though the system here is fundamentally the same as the unilateral defect constraints in \cite{GraeserKornhuber2009}, a system for box defect constraints has not been found in the literature.
+
+We now compute the LDC differences (compare \cite{GraeserKornhuber2009}):
 \begin{equation}
@@
 \emph{name}         & \emph{formulas} & \emph{indices} \\ \hline
-finest-level defect constraints & ${\Large \strut} \underline{\chi}^J = \underline{\gamma}^J - w^J,\, \overline{\chi}^J = \overline{\gamma}^J - w^J$ \\
+finest-level DCs & ${\Large \strut} \underline{\chi}^J = \underline{\gamma}^J - w^J,\, \overline{\chi}^J = \overline{\gamma}^J - w^J$ \\
 upward LDCs  & ${\Large \strut} \underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}, \, \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}$ & $j=0,\dots,J-1$ \\
@@
 
-The multilevel V-cycle solver defined in the next Section uses LDCs $\underline{\phi}^j,\overline{\phi}^j$ for the downward side and $\underline{\chi}^j,\overline{\chi}^j$ for the upward side.
 The corresponding \emph{downward/upward constraint sets}, respectively, are
@@
 
+The multilevel V-cycle solver defined in the next Section computes coarse corrections in $\mathcal{D}^j$ for the downward side and in $\mathcal{U}^j$ for the upward side.
 Our multilevel strategy is, however, based on \emph{incomplete} CDs (Section \ref{sec:cd}), in the general box-constrained cases, as we now explain.
 Each upward set $\mathcal{U}^j$ can be decomposed in an inclusion sense, becoming a true CD in unilateral cases.
-The following Lemma is based on the unilateral CD construction underlying Algorithm 4.7 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.
+The following Lemma is based in part on the unilateral CD construction underlying Algorithm 4.7 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.
 
@@
 
-In unilateral cases we can also construct decomposition operators $\Pi_i$ and show \eqref{eq:constraintrestrictionsum}.
+In unilateral cases we can also construct decomposition operators $\Pi_i$ and thereby show property \eqref{eq:constraintrestrictionsum}.
 For concreteness suppose $\overline{\gamma}^J=+\infty$, thus $\overline{\chi}^j=+\infty$ and $\overline{\phi}^i = +\infty$ for all levels.
@@
 However, the following Example shows why $\Pi_i(\min\{z^j,0\})$ may have an arbitrarily large maximum, and thus exceed any finite upper LDC $\overline{\phi}^i$.
-In fact, $\Pi_i$ does not map all non-positive functions into $\mathcal{D}^i$ in the presence of a finite upper LDC.
+In fact, $\Pi_i$ does not generally map all non-positive elements of $\mathcal{U}^J$ into $\mathcal{D}^i$ if a finite upper LDC is present.
+Note that this Example does not exclude the existence of an entirely-different strategy for showing a full CD in the general box-constrained case.
 
@@
 
-The existence of an entirely-different strategy for showing a full CD in the general box-constrained case is, of course, not excluded.
-
 Each upward defect constraint set $\mathcal{U}^j$ can also be decomposed down to an arbitrary coarser level.
@@
 In summary, inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in the V-cycle algorithm which follows.
-However, a convergence analysis, perhaps extending the arguments of \cite{Tai2003} or \cite{GraeserKornhuber2009} to VIs \eqref{eq:vi} which are not of minimization type, and/or to box constraints, would seem to require full CDs, and perhaps new ideas, and we cannot offer that analysis.
+However, a convergence analysis, perhaps extending the arguments of \cite{Tai2003} or \cite{GraeserKornhuber2009} to VIs \eqref{eq:vi} which are not of minimization type, and/or to box constraints, would seem to require full CDs, and probably further new ideas, and we cannot offer that analysis.
 

### 2023-06-17 paper/fascd.tex
SIA example replaces porous media (which was never implemented)

@@
 
-One of our examples (Section \ref{sec:results}) demonstrates the application of the \pr{fascd} method to a nonlinear operator of porous-media type.
-By ``freezing'' the solution-dependent coefficient, the operator is approximated by a coercive operator for the duration of the V-cycle.
-This scheme is also effective for doubly-nonlinear diffusion operators \cite[for example]{Bueler2021conservation,Calvoetal2002}.
+FIXME examples---see Sections \ref{sec:vi} and \ref{sec:results}
 
@@
 
-Next we give two examples which are \emph{not} of optimization type \eqref{eq:minimization}, the first being an advection-diffusion problem, which needs the following Lemma, and the second a nonlinear problem for ice sheet geometry.
+Next we give two examples which are \emph{not} of optimization type \eqref{eq:minimization}, the first being an advection-diffusion problem, which needs the following Lemma.
 
@@
 
-If $\bX \ne 0$ then $f$ in Example \ref{ex:advectiondiffusion} does not correspond to a minimization problem because $\ip{f(u)}{v}$ does not possess the symmetry of the gradient of a scalar objective.
-(For $u,v$ which are zero on $\partial \Omega$, note that $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.)  References \cite{Bueler2021conservation,ChangNakshatrala2017} consider advection-diffusion VI problems like this, specifically over the set $\cK = \{v\ge 0\}$.
-
-FIXME replace following with SIA
-
-\begin{example}  \label{ex:porous}  Suppose $\phi:[0,\infty) \to [\phi_0,\infty)$ is continuous for some constant $\phi_0>0$.
-As with Example \ref{ex:advectiondiffusion}, the following operator, of porous medium type, does not have the symmetry necessary to be the gradient of a scalar objective:
-\begin{equation}
-\ip{f(u)}{v} = \int_\Omega \phi(u) \grad u \cdot \grad v\,dx. \label{eq:porous}
-\end{equation}
-However, \eqref{eq:porous} is not known to be monotone (or coercive) when $\phi$ is not constant.
+If $\bX \ne 0$ then $f\ne F'$ in Example \ref{ex:advectiondiffusion}, for any objective $F$, because $\ip{f(u)}{v}$ does not possess the symmetry of the gradient of a scalar: for $u,v$ which are zero on $\partial \Omega$, note that $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.
+References \cite{Bueler2021conservation,ChangNakshatrala2017} consider inequality-constrained advection-diffusion VI problems like this, specifically over the set $\cK = \{v\ge 0\}$.
+
+The next VI example, nonlinear and generally not of optimization type, is a model for steady ice sheet geometry in a given climate.
+Related VI problems arise for any fluid layer which is subject to a climate which can remove the fluid \cite{Bueler2021conservation}.
+
+\begin{example}  \label{ex:sia}  Let $\Omega \subset \RR^2$ be a fixed region of land, and suppose $b \in C^1(\Omega)$ is the given bedrock elevation.
+Suppose $a \in L^\infty(\Omega)$ is the ``surface mass balance'' function, the given, annually-averaged rate of ice accumulation (as snow) minus melt and runoff.
+Then the surface elevation $s$ of a steady-state, isothermal, and non-sliding ice sheet (glacier) approximately satisfies the \emph{shallow ice approximation} (SIA) \cite{GreveBlatter2009} problem
+\begin{equation}
+\int_\Omega \Gamma (s-b)^{n+2} |\grad s|^{n-1} \grad s \cdot \grad (v-s) \,dx \ge \int_\Omega a (v-s)\,dx, \label{eq:siavi}
+\end{equation}
+over the set $\mathcal{K} = \{s\ge b\}$ of admissible surface elevations, where $n>1$ and $\Gamma>0$ are constants determined by the physical properties of ice \cite{GreveBlatter2009}; $n=3$ is typically taken.
+The doubly-nonlinear and doubly-degenerate operator in this case is
+\begin{equation}
+\ip{f(s)}{v} = \int_\Omega \Gamma (s-b)^{n+2} |\grad s|^{n-1} \grad s \cdot \grad v\,dx. \label{eq:sia}
+\end{equation}
+If the bed is flat ($b=0$) then a power transformation converts the problem into $p$-Laplacian form \eqref{eq:plaplacian} with $p=n+1$.
+For general beds $b$ less is understood about this operator, but existence of a solution to \eqref{eq:siavi} can be established \cite{JouvetBueler2012}, and the theory suggests that a solution power lives in a Sobolev space, namely $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.
+This theory, and also observations of ice sheets, shows that $|\grad s|$ will be large, or even unbounded, as one approaches the free boundary, also know as the \emph{ice margin}, from the icy side where $s>b$.
 \end{example}
 
-Numerical solver performance for Examples \ref{ex:plaplacian}, \ref{ex:advectiondiffusion}, and \ref{ex:porous} will be considered in Section \ref{sec:results}.
-Further examples of nonlinear VI problems appear in ice sheet models \cite{Calvoetal2002,JouvetBueler2012} and other geophysical fluids \cite{Bueler2021conservation}.
+Section \ref{sec:results} provides numerical results from applications of the \pr{fascd} method to Examples \ref{ex:plaplacian}, \ref{ex:advectiondiffusion}, and \ref{ex:sia}.
 
@@
 
-FIXME show results for Firedrake implementation of FASCD for the 2D classical obstacle problem (ball and spiral), 1D $p$-Laplacian obstacle problem (Example \ref{ex:plaplacian}), advection-diffusion obstacle problem (Example \ref{ex:advectiondiffusion}), and SIA obstacle problem
+FIXME show results for Firedrake implementation of FASCD for the 2D classical obstacle problem (ball and spiral; Example \ref{ex:plaplacian}), 1D $p$-Laplacian obstacle problem (also Example \ref{ex:plaplacian}), advection-diffusion obstacle problem (Example \ref{ex:advectiondiffusion}), and SIA obstacle problem (Example \ref{ex:sia})
 

### 2023-06-17 paper/fascd.tex
half-baked thoughts on icd-tele as another basic CD iteration

@@
 \end{proof}
+
+FIXME \pr{icd-tele} as third basic CD iteration
+
+\begin{pseudofloat}[H]
+\begin{pseudo*}
+\pr{icd-tele}(w^J)\text{:} \\+
+    for $j = 0,\dots,J$: \\+
+        \rm{find} $z_j\in \mathcal{U}_j$ \rm{so that for all} $v_j\in \mathcal{U}_j$, \\+
+            $\displaystyle \ip{f\Big(w^J + \alpha \sum_{i<j} z_i + \alpha z_j\Big)}{v_j-z_j} \ge \ip{\ell}{v_j-z_j}$ \\--
+    return $\tilde w^J=w^J + \alpha \sum_{i\le J} z_i$
+\end{pseudo*}
+\caption{One telescoping incomplete CD iteration for VI problem \eqref{eq:vi}, assuming $w^J\in\mathcal{K}^J$, property \eqref{eq:fe:finestlevelequivalent}, and incomplete CD \eqref{eq:fe:upwardsuminclusion}.}
+\label{alg:basiccd-tele}
+\end{pseudofloat}
 
@@
 \end{pseudofloat}
+
+Compared specifically to the $V(1,0)$ cycle presented as Algorithm 4.7 in \cite{GraeserKornhuber2009}, one might say informally that \pr{fascd-vcycle} is a $V(1,1)$ cycle.
+However, its structure relates directly to the basic CD iterations (Sections \ref{sec:cditers}, \ref{sec:cdmultilevel}) as well.
+In unilateral cases it is precise to say that one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} over CD \eqref{eq:fe:downwardsuminclusion}, using the defect constraint meaning of the downward constraint sets $\mathcal{D}^j$, followed by an application of \pr{icd-tele} over the incomplete CD \eqref{eq:fe:upwardsuminclusion}.
+FIXME: idea is that once you are using defect constraints you no longer depend on the literal availability of $\Pi_i$, so incomplete CD suffices, e.g.~even for \pr{cd-mult}
 

### 2023-06-17 paper/fascd.tex
start in on results section

@@
 
-FIXME F-cycles
+FIXME F-cycles in Algorithm \ref{alg:fascd-fcycle}
+
+\begin{pseudofloat}[ht]
+\begin{pseudo}
+\pr{fascd-fcycle} \\
+FIXME
+\end{pseudo}
+\caption{FIXME}
+\label{alg:fascd-fcycle}
+\end{pseudofloat}
 
@@
 
-FIXME show results for Firedrake implementation of FASCD for the 2D classical obstacle problem (ball and spiral; Example \ref{ex:plaplacian}), 1D $p$-Laplacian obstacle problem (also Example \ref{ex:plaplacian}), advection-diffusion obstacle problem (Example \ref{ex:advectiondiffusion}), and SIA obstacle problem (Example \ref{ex:sia})
+The FASCD V-cycle and F-cycle (Algorithms \ref{alg:fascd}, \ref{alg:fascd-fcycle}) have been implemented in Python using the Firedrake library \cite{Rathgeberetal2016}.
+In this Section we show numerical results corresponding to several VI examples from Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
+
+\begin{example}
+Consider the 2D classical, unilateral obstacle problem, namely Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$.
+Suppose $\Omega$ is a square.
+We consider two problems with very different coincidence (active) sets, the ``ball'' problem with $\psi$ equal to an upper hemisphere \cite[Chapter 12]{Bueler2021}, and the ``spiral'' problem in which $\psi$ is constructed to produce a nontrivial coincidence set \cite[Example 7.1.1]{GraeserKornhuber2009}.
+FIXME show table of results
+\end{example}
+
+%REGENERATE using pefarrell-fascd/examples/
+%$ timer python3 drape.py -prob ball|spiral -levs 9 -o ball|spiral.pvd -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
+%$ paraview ball|spiral.pvd
+%use preset "xray" and invert colors
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=0.35\textwidth]{fixfigs/ball-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/spiral-set.png}
+\end{center}
+\caption{Coincidence sets $u=\psi$ for the ball and spiral problems.}
+\label{fig:ballspiral}
+\end{figure}
+
+FIXME results for 1D $p$-Laplacian obstacle problem, Example \ref{ex:plaplacian} with $p=1.5,4$; this is a good problem to visualize the problems solved on each coarser level Figure \ref{fig:imagesvcycle}
 
@@
 \end{figure}
+
+FIXME results for advection-diffusion obstacle problem, Example \ref{ex:advectiondiffusion}
+
+FIXME results for SIA obstacle problem, Example \ref{ex:sia}
 

### 2023-06-21 paper/fascd.tex
drafting a better introduction, I hope

@@
 
-FIXME rewrite the entire introduction
-
-For coercive variational inequality (VI) problems which arise from minimization of a convex functional over a convex set, the constraint decomposition (CD) method of Tai \cite{Tai2003} converges for multilevel and domain decompositions.
-Furthermore the method has been shown to have almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
-
-We extend Tai's multilevel CD method by implementing fully-nonlinear corrections, and we apply it to a new set of nonlinear VI problems which do not arise from minimization.
-Specifically, for finite element (FE) solvers the algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} are extended here in the following directions:
+The constraint decomposition (CD) methods of Tai \cite{Tai2003} are designed to solve coercive variational inequality (VI) problems which arise from minimization of a convex functional over a convex set.
+These methods have been shown to converge for multilevel and domain decompositions in a finite element (FE) context.
+The multilevel CD method has been shown to have almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
+In this paper we extend the multilevel CD method by replacing the coarse-level corrections with a full approximation storage (FAS; \cite{Brandt1977,Bruneetal2015}) approach, and we apply this new solver to nonlinear VI problems which do not arise from minimization.
+
+Specifically, our \emph{full approximation storage constraint decomposition} (FASCD) method extends the algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} in the following directions:
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
 \begin{enumerate}
-\item We apply a full approximation storage (FAS; see \cite{Brandt1977,Bruneetal2015}) approach which avoids global (Newton) linearization.
-Section \ref{sec:results} shows results from this \pr{fascd} algorithm for several nonlinear problems.
-\item Our \pr{fascd} implementation permits ``box'' constraints, namely upper and lower obstacles; see Section \ref{sec:femultilevel}.
-\item We observe that ``up-smoothing'' is more effective than ``down-smoothing'' in for this class of algorithms.
-A strong preference for up-smoothing is special to multilevel CD methods and does not arise in unconstrained problems.
-\item By using strong, yet $O(N)$-per-application, smoothers, namely Newton-type active set methods with a fixed number of preconditioned Krylov iterations per step, in Section \ref{sec:results} we demonstrate mesh-independent convergence rates substantially better than in the current literature.
+\item The FAS coarse corrections are adapted for general VI problems.
+The needed computations do not require an objective function, nor do they require linearity of the residual operator.
+Section \ref{sec:results} shows results for several nonlinear problems.
+\item By using $O(N)$-per-application Newton-type active set methods as smoothers, with a fixed number of Krylov iterations per step, we demonstrate mesh-independent convergence rates substantially better than in the current literature.
+\item We implement both V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (also known as full multigrid; Algorithm \ref{alg:fascd-fcycle}).
+\item Box constraints, namely upper and lower obstacles, are permitted; see Section \ref{sec:femultilevel}.
+\item We observe an asymmetry special to multilevel CD methods, namely that up-smoothing is more effective than down-smoothing because a ``telescoping'' subset property can be exploited on the up-smoothing side.
 \end{enumerate}
 
-For the classical obstacle problem, based on the linear, symmetric Laplacian operator, certain multilevel techniques, such as the truncated monotone multigrid method \cite{GraeserKornhuber2009,Kornhuber1994}, are known to improve performance relative to the original multilevel CD method of Tai \cite{Tai2003}.
-These improved methods either track the active set in the discretization or modify the nodal basis functions, and in this sense they are discrete algorithms.
-By contrast a CD method applies essentially at the level of the continuum problem; see Section \ref{sec:cd}.
-Acceleration of the \pr{fascd} method via active-set and/or basis-level manipulations is an obvious topic for future research.
-
-FIXME examples---see Sections \ref{sec:vi} and \ref{sec:results}
-
-The iterates from a CD method are admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
-\pr{fascd} is designed for such admissible-only problems.
-Our target fluid-layer problems require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation \cite{Bueler2021conservation}, but these problems are not addressed herein.
-Note that non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, require unnatural modifications of the operator formula in such a fluid-layer context.
+FASCD exhibits essentially the same optimal, mesh-independent convergence rates for VI problems that FAS multigrid exhibits for the corresponding unconstrained PDE problems.
+The results in Section \ref{sec:results} include textbook multigrid efficiency, namely solutions within discretization error in work comparable to ten smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}, of F-cycles applied to classical Laplacian obstacle problems (Example \ref{ex:results:classical}), $p$-Laplacian obstacle problems (Example \ref{ex:results:plap}), and nonnegativity-constrained advection-diffusion equations (Example \ref{ex:results:advdiff}).
+Furthermore, these results are essentially insensitive to the length or structure of the free boundary, i.e.~the geometric complexity of the coincidence set (Examples \ref{ex:results:classical}, \ref{ex:results:advdiff}).
+
+The iterates from FASCD are admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
+For example, we solve an ice sheet geometry problem (Examples \ref{ex:sia}, \ref{ex:results:sia}) in which the ice velocity is only meaningful for surface elevation iterates which exceed the bed elevation.
+More generally, fluid-layer problems, with a non-negative thickness constraint, would require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation \cite{Bueler2021conservation}.
+(Though a target for FASCD, such problems, however, are not addressed here.)  Note that non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural modifications of the operator formula in such a fluid-layer context.
+In any case, FASCD solves the difficult ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a mesh-independent number of iterations, with near textbook efficiency.
+
+For the classical obstacle problem, based on the linear, symmetric Laplacian operator, certain multilevel techniques, such as the truncated monotone multigrid method \cite{GraeserKornhuber2009,Kornhuber1994}, are known to improve performance relative to the original multilevel CD method \cite{Tai2003}.
+These improved methods modify the nodal basis functions.
+By contrast FASCD makes essentially no assumptions about the FE discretization.
+Acceleration via basis-level manipulations is a topic for future research.
+
+FIXME organization of paper
 
@@
 
-\begin{example}
+\begin{example} \label{ex:results:classical}
 Consider the 2D classical, unilateral obstacle problem, namely Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$.
@@
 
+\begin{example}  \label{ex:results:plap}
 FIXME results for 1D $p$-Laplacian obstacle problem, Example \ref{ex:plaplacian} with $p=1.5,4$; this is a good problem to visualize the problems solved on each coarser level Figure \ref{fig:imagesvcycle}
+\end{example}
 
@@
 
+\begin{example}  \label{ex:results:advdiff}
 FIXME results for advection-diffusion obstacle problem, Example \ref{ex:advectiondiffusion}
-
+\end{example}
+
+\begin{example}   \label{ex:results:sia}
 FIXME results for SIA obstacle problem, Example \ref{ex:sia}
+\end{example}
 

### 2023-06-22 paper/fascd.tex
results for classical obstacle problem example

@@
 
-\usepackage{times,bm,bbm,empheq,fancyvrb,graphicx,amsthm,amssymb}
+\usepackage{times,bm,bbm,empheq,verbatim,fancyvrb,graphicx,amsthm,amssymb}
 \usepackage[dvipsnames]{xcolor}
-\usepackage{longtable,booktabs,xspace}
+\usepackage{booktabs,multirow,xspace}
 
@@
 \usetikzlibrary{decorations.pathreplacing}
-\usetikzlibrary{graphs}
+\usetikzlibrary{graphs,quotes}
 
@@
 \begin{equation}
-\|\rSS(w^{J,k})\|_2 < \id{atol} \qquad \text{or} \qquad \frac{\|\rSS(w^{J,k})\|_2}{\|\rSS(w^{J,0})\|_2} < \id{rtol}
+\|\rSS(w^{J,k})\|_2 < \id{atol} \qquad \text{or} \qquad \frac{\|\rSS(w^{J,k})\|_2}{\|\rSS(w^{J,0})\|_2} < \id{rtol} \label{eq:stoppingcriterion}
 \end{equation}
@@
 
-\section{Results and discussion} \label{sec:results}
-
-The FASCD V-cycle and F-cycle (Algorithms \ref{alg:fascd}, \ref{alg:fascd-fcycle}) have been implemented in Python using the Firedrake library \cite{Rathgeberetal2016}.
+\section{Results} \label{sec:results}
+
+Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}.
 In this Section we show numerical results corresponding to several VI examples from Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
@@
 \begin{example} \label{ex:results:classical}
-Consider the 2D classical, unilateral obstacle problem, namely Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$.
-Suppose $\Omega$ is a square.
-We consider two problems with very different coincidence (active) sets, the ``ball'' problem with $\psi$ equal to an upper hemisphere \cite[Chapter 12]{Bueler2021}, and the ``spiral'' problem in which $\psi$ is constructed to produce a nontrivial coincidence set \cite[Example 7.1.1]{GraeserKornhuber2009}.
-FIXME show table of results
+Consider the 2D classical, unilateral obstacle problem, namely Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$, with $\Omega$ square.
+We consider two problems with very different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$.
+The first is a standard ``ball'' problem, with an upper hemisphere obstacle $\psi$ \cite[Chapter 12]{Bueler2021}, and the second is the ``spiral'' problem in which $\psi$ is constructed to produce a more interesting coincidence set \cite[Example 7.1.1]{GraeserKornhuber2009}.
+For this experiment we generated uniform triangular meshs, from a coarsest mesh of 32 triangles and $N=4^2$ nodes, to a finest mesh of $N=2048^2 \approx 4.2 \times 10^{6}$ nodes after nine standard refinements.
+(The V- and F-cycles descended to the $4\times 4$ mesh.)  An RS$+$CG$+$ICC symmetric smoother (Section \ref{sec:implementation}) was applied, with fixed three iterations, and stopping tolerances \texttt{rtol} $=$ \texttt{atol} $= 10^{-6}$ in \eqref{eq:stoppingcriterion}.
+% FIXME set atol *small* and re-run
+
+Table \ref{tab:results:classical} shows the resulting iterations.
+The RS Newton method, applied by itself as the solver with direct (LU) solution of the Newton steps, shows rapidly increasing iterations, approximately proportionally to the resolution, as expected; the highest resolution cases were not attempted with this non-scalable solver.
+FASCD V-cycle iterations increase slowly, roughly logarithmically with the resolution.
+For F-cycles we counted the number of V-cycles after the ramp (Section \ref{sec:implementation}), which shows constant iterations.
+The spiral problem appears to be only slightly more challenging for these solvers.
+Note that three V-cycle iterations in the F-cycle application, which suffices to solve either problem at any resolution, does work proportional to approximately FIXME smoother applications on the finest level, so these problems are solved with nearly textbook multigrid efficiency despite the low solution regularity.
 \end{example}
@@
 \end{center}
-\caption{Coincidence sets $u=\psi$ for the ball and spiral problems.}
+\caption{Coincidence sets for the ball and spiral problems in Example \ref{ex:results:classical}.}
 \label{fig:ballspiral}
 \end{figure}
+
+\newcommand{\XX}{{\sffamily X}}
+% see pefarrell/examples/drape.py:
+%   $ python3 drape.py -prob PROB -levs LEVS
+% for PROB=ball,spiral and LEVS=1,...,10
+% note mcoarse=4, rtol=atol=1e-6
+% smoother is vinewtonrsls + 3 x (cg + icc)
+\begin{table}[ht]
+\begin{tabular}{lr rrr rrr}
+\multirow{2}{*}{\emph{levels}} & \multirow{2}{*}{$N$} & \multicolumn{3}{c}{\,\emph{ball}} & \multicolumn{3}{c}{\,\emph{spiral}} \\ \cmidrule(lr){3-5} \cmidrule(lr){6-8}
+             && \, V-cycle & F-cycle & RS only & \,V-cycle & F-cycle & RS only \\ \cmidrule{1-8}
+ 1 &    $4^2$ &  1 &  1 &  1  &  1 &  1 &  1 \\
+ 2 &    $8^2$ &  2 &  1 &  2  &  2 &  1 &  2 \\
+ 3 &   $16^2$ &  2 &  2 &  3  &  3 &  2 &  4 \\
+ 4 &   $32^2$ &  3 &  2 &  6  &  3 &  2 &  7 \\
+ 5 &   $64^2$ &  3 &  2 &  8  &  4 &  3 &  9 \\
+ 6 &  $128^2$ &  4 &  2 & 17  &  4 &  3 & 16 \\
+ 7 &  $256^2$ &  4 &  2 & 33  &  4 &  3 & 27 \\
+ 8 &  $512^2$ &  5 &  2 & 65  &  5 &  3 & 48 \\
+ 9 & $1024^2$ &  5 &  2 & \XX &  5 &  3 & \XX \\
+10 & $2048^2$ &  5 &  2 & \XX &  5 &  3 & \XX
+\end{tabular}
+
+\bigskip
+\caption{Iterations for FASCD V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}.}
+\label{tab:results:classical}
+\end{table}
 
@@
 \end{center}
-\caption{Visualization of one iteration of \textsc{FASCD-VCYCLE} (Algorithm \ref{alg:fascd}) for Example FIXME, with $p=1.5$, $J=3$, and a coarsest mesh of two elements.
-Solutions ($w^j$, $y^j$, $z^j$) are lines while obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots.}
+\caption{Visualization of one FASCD V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of two elements.
+Solutions (i.e.~$w^j$, $y^j$, $z^j$) are lines while obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots.
+The zero function (dotted) is the admissible initial iterate for $y^j$ and $z^j$.
+Vertical scale varies.}
 \label{fig:imagesvcycle}
@@
 % A BRIDGE TOO FAR:  \section{Results for a nonlocal variational inequality}
+
+\section{Discussion and Conclusion} \label{sec:discussion}
+
+FIXME
 
@@
 
-\section{From slack, to not lose} \label{app:slack}
-
-FIXME remove this appendix, after using it in the introduction and discussion etc.
+\begin{comment}
+%From slack, to not lose
 
@@
 So the difference between FASCD and \cite{GraeserKornhuber2009} are: (1) FAS coarse corrections, (2) no attempt to do monotone/truncated multigrid which would be a mess (I think) across the FE infrastructure, (3) clarity on what sets you can use in $V(\nu_1,\nu_2)$ cycles, with bigger up-smoother sets in particular, and most importantly (4) implementation within the mature Firedrake/PETSc library framework where one can easily hook up an active-set Newton step as a smoother.
+\end{comment}
 

### 2023-06-23 paper/fascd.tex
clean results example 9.1

@@
 The first is a standard ``ball'' problem, with an upper hemisphere obstacle $\psi$ \cite[Chapter 12]{Bueler2021}, and the second is the ``spiral'' problem in which $\psi$ is constructed to produce a more interesting coincidence set \cite[Example 7.1.1]{GraeserKornhuber2009}.
-For this experiment we generated uniform triangular meshs, from a coarsest mesh of 32 triangles and $N=4^2$ nodes, to a finest mesh of $N=2048^2 \approx 4.2 \times 10^{6}$ nodes after nine standard refinements.
-(The V- and F-cycles descended to the $4\times 4$ mesh.)  An RS$+$CG$+$ICC symmetric smoother (Section \ref{sec:implementation}) was applied, with fixed three iterations, and stopping tolerances \texttt{rtol} $=$ \texttt{atol} $= 10^{-6}$ in \eqref{eq:stoppingcriterion}.
-% FIXME set atol *small* and re-run
+For this experiment we generated uniform triangular meshs, from a coarsest mesh of 32 triangles and $N=4^2$ nodes, to a finest mesh of $N=2048^2 \approx 4.2 \times 10^{6}$ nodes after nine standard refinements, and note that the V- and F-cycles descended to the $4\times 4$ mesh.
+A symmetric RS smoother (Section \ref{sec:implementation}) with three (fixed) ICC-preconditioned CG iterations was applied, the coarse solver was RS with LU (direct) solution of the Newton step equations, and tolerances \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$ were used in stopping criterion \eqref{eq:stoppingcriterion}.
 
 Table \ref{tab:results:classical} shows the resulting iterations.
-The RS Newton method, applied by itself as the solver with direct (LU) solution of the Newton steps, shows rapidly increasing iterations, approximately proportionally to the resolution, as expected; the highest resolution cases were not attempted with this non-scalable solver.
-FASCD V-cycle iterations increase slowly, roughly logarithmically with the resolution.
-For F-cycles we counted the number of V-cycles after the ramp (Section \ref{sec:implementation}), which shows constant iterations.
-The spiral problem appears to be only slightly more challenging for these solvers.
-Note that three V-cycle iterations in the F-cycle application, which suffices to solve either problem at any resolution, does work proportional to approximately FIXME smoother applications on the finest level, so these problems are solved with nearly textbook multigrid efficiency despite the low solution regularity.
+The RS Newton method, i.e.~the smoother, applied by itself as the solver, shows rapidly increasing iterations, approximately proportionally to resolution, as expected.
+By contrast, FASCD V-cycle iterations increase slowly, roughly logarithmically with the resolution.
+For F-cycles we counted the number of V-cycles after the ramp (Section \ref{sec:implementation}), and this shows a constant number of V-cycle iterations.
+The spiral problem appears to be only slightly more challenging for these solvers; apparently the complexity of the coincidence set or free boundary does not strongly influence performance.
+
+Note that each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+ICC) iterations.
+For the F-cycle results, where three V-cycle iterations suffice in either problem, and at any finest-level resolution, we see textbook multigrid efficiency\footnote{This phrase is imperfectly defined, but one possible definition is given by \cite{BrownSmithAhmadia2013}: ``textbook multigrid efficiency [is] (an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation).''} despite their low solution regularity.
 \end{example}
@@
 \newcommand{\XX}{{\sffamily X}}
-% see pefarrell/examples/drape.py:
-%   $ python3 drape.py -prob PROB -levs LEVS
-% for PROB=ball,spiral and LEVS=1,...,10
-% note mcoarse=4, rtol=atol=1e-6
-% smoother is vinewtonrsls + 3 x (cg + icc)
+% see pefarrell/examples/drape.py and pefarrell/edplay/resultsdrape.sh|txt
+% note mcoarse=4, rtol=1e-6, atol=1e-12
+% smoother is vinewtonrsls + 3 x (cg + icc), coarse solver is vinewtonrsls + lu
 \begin{table}[ht]
@@
 \bigskip
-\caption{Iterations for FASCD V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}.}
+\caption{Iterations for FASCD V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}.  Runs marked \XX\xspace were not attempted.}
 \label{tab:results:classical}

### 2023-06-23 paper/fascd.tex
p=1.5 case of Example 9.2 done

@@
 \newcommand{\rXX}{r_{\text{XX}}}
+
+\newcommand{\XX}{{\sffamily X}}
 
@@
 We consider two problems with very different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$.
-The first is a standard ``ball'' problem, with an upper hemisphere obstacle $\psi$ \cite[Chapter 12]{Bueler2021}, and the second is the ``spiral'' problem in which $\psi$ is constructed to produce a more interesting coincidence set \cite[Example 7.1.1]{GraeserKornhuber2009}.
-For this experiment we generated uniform triangular meshs, from a coarsest mesh of 32 triangles and $N=4^2$ nodes, to a finest mesh of $N=2048^2 \approx 4.2 \times 10^{6}$ nodes after nine standard refinements, and note that the V- and F-cycles descended to the $4\times 4$ mesh.
-A symmetric RS smoother (Section \ref{sec:implementation}) with three (fixed) ICC-preconditioned CG iterations was applied, the coarse solver was RS with LU (direct) solution of the Newton step equations, and tolerances \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$ were used in stopping criterion \eqref{eq:stoppingcriterion}.
-
-Table \ref{tab:results:classical} shows the resulting iterations.
-The RS Newton method, i.e.~the smoother, applied by itself as the solver, shows rapidly increasing iterations, approximately proportionally to resolution, as expected.
+The first is a standard ``ball'' problem, with an upper hemisphere obstacle $\psi$ \cite[Chapter 12]{Bueler2021}, and the second is the ``spiral'' problem in which $\psi$ is constructed to produce a more interesting coincidence set \cite[Example 7.1.1]{GraeserKornhuber2009}; see Figure \ref{fig:results:classical}.
+For this experiment we generated uniform triangular meshs, from a coarsest mesh of 32 triangles and $m_0=5^2$ nodes, to a finest mesh of $m_9=2049^2 \approx 4.2 \times 10^{6}$ nodes after standard refinements; V- and F-cycles always descended to the $m_0=5^2$ mesh.
+As our $O(m_j)$ smoother we applied a single Newton iteration of the RS smoother (Section \ref{sec:implementation}), with three (fixed) ICC-preconditioned CG iterations to solve the Newton step equations.
+The coarse solver was RS$+$LU, iterated to convergence.
+Tolerances \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$ were used in stopping criterion \eqref{eq:stoppingcriterion}.
+Noting that $\psi\big|_{\partial \Omega} \le 0$ in both problems, the initial iterate was $w=\max\{0,\psi\}$.
+
+Table \ref{tab:results:classical} shows the resulting iteration counts.
+The RS Newton method, i.e.~the smoother, was also applied by itself as the solver, and it shows rapidly-increasing iterations as expected, approximately proportionally to resolution.
 By contrast, FASCD V-cycle iterations increase slowly, roughly logarithmically with the resolution.
@@
 Note that each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+ICC) iterations.
-For the F-cycle results, where three V-cycle iterations suffice in either problem, and at any finest-level resolution, we see textbook multigrid efficiency\footnote{This phrase is imperfectly defined, but one possible definition is given by \cite{BrownSmithAhmadia2013}: ``textbook multigrid efficiency [is] (an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation).''} despite their low solution regularity.
+For the F-cycle results, where three V-cycle iterations suffice in either problem, and at any finest-level resolution, we see textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''} despite their low solution regularity.
 \end{example}
 
-%REGENERATE using pefarrell-fascd/examples/
+%REGENERATE using git@bitbucket.org:pefarrell/fascd.git, in examples/
 %$ timer python3 drape.py -prob ball|spiral -levs 9 -o ball|spiral.pvd -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
@@
 \caption{Coincidence sets for the ball and spiral problems in Example \ref{ex:results:classical}.}
-\label{fig:ballspiral}
+\label{fig:results:classical}
 \end{figure}
 
-\newcommand{\XX}{{\sffamily X}}
-% see pefarrell/examples/drape.py and pefarrell/edplay/resultsdrape.sh|txt
+% see git@bitbucket.org:pefarrell/fascd.git: examples/drape.py and edplay/resultsdrape.sh|txt
 % note mcoarse=4, rtol=1e-6, atol=1e-12
@@
 \begin{table}[ht]
-\begin{tabular}{lr rrr rrr}
-\multirow{2}{*}{\emph{levels}} & \multirow{2}{*}{$N$} & \multicolumn{3}{c}{\,\emph{ball}} & \multicolumn{3}{c}{\,\emph{spiral}} \\ \cmidrule(lr){3-5} \cmidrule(lr){6-8}
+\begin{tabular}{lr ccc ccc}
+\multirow{2}{*}{\emph{levels}} & \multirow{2}{*}{$m_J$} & \multicolumn{3}{c}{\,\emph{ball}} & \multicolumn{3}{c}{\,\emph{spiral}} \\ \cmidrule(lr){3-5} \cmidrule(lr){6-8}
              && \, V-cycle & F-cycle & RS only & \,V-cycle & F-cycle & RS only \\ \cmidrule{1-8}
- 1 &    $4^2$ &  1 &  1 &  1  &  1 &  1 &  1 \\
- 2 &    $8^2$ &  2 &  1 &  2  &  2 &  1 &  2 \\
- 3 &   $16^2$ &  2 &  2 &  3  &  3 &  2 &  4 \\
- 4 &   $32^2$ &  3 &  2 &  6  &  3 &  2 &  7 \\
- 5 &   $64^2$ &  3 &  2 &  8  &  4 &  3 &  9 \\
- 6 &  $128^2$ &  4 &  2 & 17  &  4 &  3 & 16 \\
- 7 &  $256^2$ &  4 &  2 & 33  &  4 &  3 & 27 \\
- 8 &  $512^2$ &  5 &  2 & 65  &  5 &  3 & 48 \\
- 9 & $1024^2$ &  5 &  2 & \XX &  5 &  3 & \XX \\
-10 & $2048^2$ &  5 &  2 & \XX &  5 &  3 & \XX
+ 1 &    $5^2$ &  1 &  1 &  1  &  1 &  1 &  1 \\
+ 2 &    $9^2$ &  2 &  1 &  2  &  2 &  1 &  2 \\
+ 3 &   $17^2$ &  2 &  2 &  3  &  3 &  2 &  4 \\
+ 4 &   $33^2$ &  3 &  2 &  6  &  3 &  2 &  7 \\
+ 5 &   $65^2$ &  3 &  2 &  8  &  4 &  3 &  9 \\
+ 6 &  $129^2$ &  4 &  2 & 17  &  4 &  3 & 16 \\
+ 7 &  $257^2$ &  4 &  2 & 33  &  4 &  3 & 27 \\
+ 8 &  $513^2$ &  5 &  2 & 65  &  5 &  3 & 48 \\
+ 9 & $1025^2$ &  5 &  2 & \XX &  5 &  3 & \XX \\
+10 & $2049^2$ &  5 &  2 & \XX &  5 &  3 & \XX
 \end{tabular}
-
 \bigskip
@@
 \begin{example}  \label{ex:results:plap}
-FIXME results for 1D $p$-Laplacian obstacle problem, Example \ref{ex:plaplacian} with $p=1.5,4$; this is a good problem to visualize the problems solved on each coarser level Figure \ref{fig:imagesvcycle}
+Recall Example \ref{ex:plaplacian}, the nonlinear $p$-Laplacian operator.
+Here we solve a unilateral $p$-Laplacian obstacle problem, in one dimension for visualizability, considering both fast ($p=1.5$) and degenerate diffusivity ($p=4.0$) cases.\footnote{``Fast'' and ``degenerate'' refer to $D=|\grad u|^{p-2}\to \infty$ and $D\to 0$ as $|\grad u|\to 0$, respectively.
+Ellipticity is lost in the latter case.}  In this example we also contrast $V(1,0)$ (no up-smoothing) and $V(0,1)$ (no down-smoothing) results.
+
+Suppose $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, and $\ell(v) = \int_\Omega g v$ where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$.
+The exact solution of VI \eqref{eq:vi} is easily computed as a function of $p$; see Figure \ref{fig:results:plap}.
+For this experiment we generated a coarse mesh of $m_0=7$ nodes, refining nine times to a $m_{10}=3073$ finest mesh.
+The default smoother used fixed three iterations of RS, with LU solution of the Newton step equations;\footnote{This is an optimal $O(m_j)$ smoother in 1D.} the coarse solver was the same but iterated to convergence.
+Again \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$ were set in the stopping criterion.
+The initial iterate was $w=\psi$.
+
+Figure \ref{fig:imagesvcycle} illustrates an entire V-cycle, showing each of the nonlinear FAS coarse correction VI subproblem solutions.
+Each $y^j \in \mathcal{D}^j$ and $z^j \in \mathcal{U}^j$ is drawn, along with each level defect constraint $\underline{\phi}^j,\underline{\chi}^j$.
+
+Iteration results for the $p=1.5$ case are shown in Table \ref{tab:results:fastplap1d}.
+We see that $V(1,0)$ cycles, which correspond to an FAS implemention of Algorithm 4.7 in \cite{GraeserKornhuber2009}, but with a stronger smoother, often do not converge.
+By contrast, $V(0,1)$ cycles, using only up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are just as good as the default $V=V(1,1)$ cycles defined by Algorithm \ref{alg:fascd}.
+The default $V(1,1)$ cycles give convergence in the $\|\cdot\|_\infty$ norm (last column) at rate $O(h^{2.005})$.
+For this problem V-cycle iterations are seen to grow less consistently than in the Laplacian obstacle problem (e.g.~Example \ref{ex:results:classical}), not surprising for a singular diffusivity function.
+
+Turning to F-cycles, they are very efficient except when they fail entirely, but this too arises from singular diffusivity.
+If we regularize the diffusivity formula, from $D=|\grad u|^{p-2}$ to $D_\eps = \left(\eps + |\grad u|^2\right)^{(p-2)/2}$ with $\eps=10^{-8}$, then a single F-cycle suffices at every resolution (column ``F$_\eps$''), although numerical error is not as small on the finest meshes.
+
+Iteration results for the degenerate diffusion $p=4.0$ case are in Table \ref{tab:results:degenerateplap1d}.  FIXME V.   A single F-cycle suffices at every resolution.
+
 \end{example}
+
+%REGENERATE using elbueler/plapfigs branch of git@bitbucket.org:pefarrell/fascd.git, in examples/
+% $ python3 plap1d.py -p 1.5 -figures -levs 6 -fascd_cycle_type full && mv solution.pdf plap1d1p5.pdf
+% $ python3 plap1d.py -p 4.0 -figures -levs 6 -fascd_cycle_type full && mv solution.pdf plap1d4p0.pdf
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=0.45\textwidth]{fixfigs/plap1d1p5.pdf} \quad
+\includegraphics[width=0.45\textwidth]{fixfigs/plap1d4p0.pdf}
+\end{center}
+\caption{Solutions $u\ge \psi$ of the $p$-Laplacian VI in Example \ref{ex:results:plap}, for $p=1.5$ (left) and $p=4.0$ (right).}
+\label{fig:results:plap}
+\end{figure}
 
@@
 \end{center}
-\caption{Visualization of one FASCD V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of two elements.
+\caption{Visualization of one FASCD V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of $m_0=3$ nodes.
 Solutions (i.e.~$w^j$, $y^j$, $z^j$) are lines while obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots.
@@
 
+% for Tables \ref{tab:results:fastplap1d} and \ref{tab:results:degenerateplap1d},
+% see git@bitbucket.org:pefarrell/fascd.git: examples/plap1d.py and edplay/resultsplap1d.sh|txt
+
+\begin{table}[ht]
+\begin{tabular}{lr ccc cc | r}
+\emph{levels} & $m_J$ & V(1,0) & V(0,1) & V(1,1) & \phantom{m}F\phantom{m} & \phantom{i}F$_\eps$\phantom{i} & $\|$error$\|_\infty$ \\ \cmidrule{1-8}
+ 1 &    $7$ &  1 &  1 &  1 &  1 &  1 & \quad $2.3\times 10^{-1}$ \\
+ 2 &   $13$ &  3 &  3 &  2 &  2 &  1 & $3.3 \times 10^{-2}$ \\
+ 3 &   $25$ & 12 &  4 &  2 &  1 &  1 & $9.1 \times 10^{-3}$ \\
+ 4 &   $49$ & NC &  4 &  2 &  1 &  1 & $3.2 \times 10^{-3}$ \\
+ 5 &   $97$ & NC &  3 &  3 &  1 &  1 & $5.5 \times 10^{-4}$ \\
+ 6 &  $193$ & NC &  3 &  3 & NC &  1 & $1.7 \times 10^{-4}$ \\
+ 7 &  $385$ & NC &  3 &  3 &  1 &  1 & $4.7 \times 10^{-5}$ \\
+ 8 &  $769$ & NC &  4 &  7 &  1 &  1 & $9.5 \times 10^{-6}$ \\
+ 9 & $1537$ & NC &  5 &  6 & NC &  1 & $3.6 \times 10^{-6}$ \\
+10 & $3073$ & NC & 11 & 19 &  1 &  1 & $4.9 \times 10^{-7}$
+\end{tabular}
+\bigskip
+\caption{Iterations for V- and F-cycles on the $p=1.5$ problem in Example \ref{ex:results:plap}.  NC $=$ did not converge after 50 iterations.  See text regarding F$_\eps$ and error.}
+\label{tab:results:fastplap1d}
+\end{table}
+
+\begin{table}[ht]
+\begin{tabular}{lr ccc}
+\emph{levels} & $m_J$ & \phantom{m}V\phantom{m} & \phantom{m}V$_4$\phantom{m} & \phantom{m}F\phantom{m} \\ \cmidrule{1-5}
+ 1 &    $7$ &   \\
+ 2 &   $13$ &   \\
+ 3 &   $25$ &   \\
+ 4 &   $49$ &   \\
+ 5 &   $97$ &   \\
+ 6 &  $193$ &   \\
+ 7 &  $385$ &   \\
+ 8 &  $769$ &   \\
+ 9 & $1537$ &   \\
+10 & $3073$ &  
+\end{tabular}
+\bigskip
+\caption{Iterations for V- and F-cycles on the $p=4.0$ problem in Example \ref{ex:results:plap}.  NC $=$ did not converge after 50 iterations.  See text regarding V$_4$.}
+\label{tab:results:degenerateplap1d}
+\end{table}
+
 \begin{example}  \label{ex:results:advdiff}
-FIXME results for advection-diffusion obstacle problem, Example \ref{ex:advectiondiffusion}
+FIXME results for 3D advection-diffusion obstacle problem, Example \ref{ex:advectiondiffusion}
 \end{example}
@@
 
-FIXME observe that up-smoothing is more efficient; this observation seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}
-
 % A BRIDGE TOO FAR:  \section{Results for a nonlocal variational inequality}
@@
 FIXME
+
+FIXME the observation that up-smoothing is more efficient; this observation seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}
 

### 2023-06-24 paper/fascd.tex
completed proper write-up of 1D p-lap results, I think

--- 2023-06-23 paper/fascd.tex
@@
 
-FIXME smoothers including Newton active set \cite{BensonMunson2006} or Newton semi-smooth \cite{BensonMunson2006}, with either direct linear solve or a fixed number of Krylov steps; in latter case either fixed number of CG+ICC or fixed number of GMRES+ILU
+FIXME smoothers use Newton reduced space (RS) \cite{BensonMunson2006}with either direct linear solve or a fixed number of Krylov steps; in latter case either fixed number of CG+ICC for symmetric linear operator (Laplacian), otherwise fixed number of GMRES+ILU
+
+% FIXME try Newton semi-smooth \cite{BensonMunson2006}?
 
@@
 In this Section we show numerical results corresponding to several VI examples from Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
+Our essential observation is that, with strong $O(m_J)$ smoothers on each level, each V-cycle dramatically reduces the residual norm, and generally a couple of F-cycles solves the problem down to discretization error.
 
@@
 We consider two problems with very different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$.
-The first is a standard ``ball'' problem, with an upper hemisphere obstacle $\psi$ \cite[Chapter 12]{Bueler2021}, and the second is the ``spiral'' problem in which $\psi$ is constructed to produce a more interesting coincidence set \cite[Example 7.1.1]{GraeserKornhuber2009}.
-For this experiment we generated uniform triangular meshs, from a coarsest mesh of 32 triangles and $N=4^2$ nodes, to a finest mesh of $N=2048^2 \approx 4.2 \times 10^{6}$ nodes after nine standard refinements, and note that the V- and F-cycles descended to the $4\times 4$ mesh.
-A symmetric RS smoother (Section \ref{sec:implementation}) with three (fixed) ICC-preconditioned CG iterations was applied, the coarse solver was RS with LU (direct) solution of the Newton step equations, and tolerances \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$ were used in stopping criterion \eqref{eq:stoppingcriterion}.
-
-Table \ref{tab:results:classical} shows the resulting iterations.
-The RS Newton method, i.e.~the smoother, applied by itself as the solver, shows rapidly increasing iterations, approximately proportionally to resolution, as expected.
+The first is a standard ``ball'' problem, with an upper hemisphere obstacle $\psi$ \cite[Chapter 12]{Bueler2021}, and the second is the ``spiral'' problem from Example 7.1.1 in \cite{GraeserKornhuber2009}, wherein $\psi$ is constructed to produce a more interesting coincidence set; see Figure \ref{fig:results:classical}.
+For this experiment we used $P_1$ elements and uniform triangular meshs, from a coarsest mesh of $m_0=5^2$ nodes (32 triangles), to a finest mesh of $m_9=2049^2 \approx 4.2 \times 10^{6}$ nodes after standard refinements.
+As our $O(m_j)$ smoother we applied a single iteration of RS Newton (Section \ref{sec:implementation}), with three (fixed) ICC-preconditioned CG iterations to approximately solve the step equations.
+The coarse solver was RS$+$LU, iterated to convergence.
+Tolerances \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$ were used in stopping criterion \eqref{eq:stoppingcriterion}.
+Noting that $\psi\big|_{\partial \Omega} \le 0$ in both problems, the initial iterate was set to $w=\max\{0,\psi\}$.
+
+Table \ref{tab:results:classical} shows the resulting iteration counts.
+When the RS Newton smoother is applied by itself as a solver it shows rapidly-increasing iterations as expected, approximately proportionally to resolution.
 By contrast, FASCD V-cycle iterations increase slowly, roughly logarithmically with the resolution.
-For F-cycles we counted the number of V-cycles after the ramp (Section \ref{sec:implementation}), and this shows a constant number of V-cycle iterations.
+Noting that for F-cycles we are counting the number of V-cycles after the ramp (Section \ref{sec:implementation}), we see that a constant number of F-cycles solves the problem.
 The spiral problem appears to be only slightly more challenging for these solvers; apparently the complexity of the coincidence set or free boundary does not strongly influence performance.
@@
 Note that each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+ICC) iterations.
-For the F-cycle results, where three V-cycle iterations suffice in either problem, and at any finest-level resolution, we see textbook multigrid efficiency\footnote{This phrase is imperfectly defined, but one possible definition is given by \cite{BrownSmithAhmadia2013}: ``textbook multigrid efficiency [is] (an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation).''} despite their low solution regularity.
+For the F-cycle results, where three V-cycle iterations suffice in either problem, and at any finest-level resolution, this is nearly textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  However, work of at most ten residual evaluations is often mentioned.} despite the low solution regularity.
 \end{example}
 
-%REGENERATE using pefarrell-fascd/examples/
+%REGENERATE using git@bitbucket.org:pefarrell/fascd.git, in examples/
 %$ timer python3 drape.py -prob ball|spiral -levs 9 -o ball|spiral.pvd -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
@@
 \begin{table}[ht]
-\begin{tabular}{lr rrr rrr}
-\multirow{2}{*}{\emph{levels}} & \multirow{2}{*}{$N$} & \multicolumn{3}{c}{\,\emph{ball}} & \multicolumn{3}{c}{\,\emph{spiral}} \\ \cmidrule(lr){3-5} \cmidrule(lr){6-8}
-             && \, V-cycle & F-cycle & RS only & \,V-cycle & F-cycle & RS only \\ \cmidrule{1-8}
- 1 &    $4^2$ &  1 &  1 &  1  &  1 &  1 &  1 \\
- 2 &    $8^2$ &  2 &  1 &  2  &  2 &  1 &  2 \\
- 3 &   $16^2$ &  2 &  2 &  3  &  3 &  2 &  4 \\
- 4 &   $32^2$ &  3 &  2 &  6  &  3 &  2 &  7 \\
- 5 &   $64^2$ &  3 &  2 &  8  &  4 &  3 &  9 \\
- 6 &  $128^2$ &  4 &  2 & 17  &  4 &  3 & 16 \\
- 7 &  $256^2$ &  4 &  2 & 33  &  4 &  3 & 27 \\
- 8 &  $512^2$ &  5 &  2 & 65  &  5 &  3 & 48 \\
- 9 & $1024^2$ &  5 &  2 & \XX &  5 &  3 & \XX \\
-10 & $2048^2$ &  5 &  2 & \XX &  5 &  3 & \XX
+\begin{tabular}{lr@{\hskip 7mm}cccccc}
+\multirow{2}{*}{\emph{levels}} & \multirow{2}{*}{$m_J$} & \multicolumn{3}{c}{\,\emph{ball}} & \multicolumn{3}{c}{\,\emph{spiral}} \\ \cmidrule(lr){3-5} \cmidrule(lr){6-8}
+   &          & V-cycle & F-cycle & RS only & V-cycle & F-cycle & RS only \\ \cmidrule{1-8}
+ 1 &    $5^2$ &  1 &  1 &  1  &  1 &  1 &  1 \\
+ 2 &    $9^2$ &  2 &  1 &  2  &  2 &  1 &  2 \\
+ 3 &   $17^2$ &  2 &  2 &  3  &  3 &  2 &  4 \\
+ 4 &   $33^2$ &  3 &  2 &  6  &  3 &  2 &  7 \\
+ 5 &   $65^2$ &  3 &  2 &  8  &  4 &  3 &  9 \\
+ 6 &  $129^2$ &  4 &  2 & 17  &  4 &  3 & 16 \\
+ 7 &  $257^2$ &  4 &  2 & 33  &  4 &  3 & 27 \\
+ 8 &  $513^2$ &  5 &  2 & 65  &  5 &  3 & 48 \\
+ 9 & $1025^2$ &  5 &  2 & \XX &  5 &  3 & \XX \\
+10 & $2049^2$ &  5 &  2 & \XX &  5 &  3 & \XX
 \end{tabular}
-
 \bigskip
-\caption{Iterations for FASCD V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}.  Runs marked \XX\xspace were not attempted.}
+\caption{Iterations for FASCD V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}.  Smoother-only runs marked \XX\xspace were not attempted.}
 \label{tab:results:classical}
@@
 \begin{example}  \label{ex:results:plap}
-FIXME results for 1D $p$-Laplacian obstacle problem, Example \ref{ex:plaplacian} with $p=1.5,4$; this is a good problem to visualize the problems solved on each coarser level Figure \ref{fig:imagesvcycle}
-\end{example}
+Recall Example \ref{ex:plaplacian}, the nonlinear $p$-Laplacian operator.
+We will solve a unilateral obstacle problem, in one dimension for visualizability, considering both fast ($p=1.5$) and degenerate diffusivity ($p=4.0$) cases.\footnote{``Fast'' and ``degenerate'' refer to $D=|\grad u|^{p-2}\to \infty$ and $D\to 0$ as $|\grad u|\to 0$, respectively.
+Ellipticity is lost in the latter case.}  This example reveals a strong contrast between $V(1,0)$ and $V(0,1)$ results, namely \pr{up} $=0$ and \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively, and it emphasizes the need for robust smoothers.
+
+Suppose $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, $\mathcal{K} = \{v \ge \psi\} \subset W^{1,p}(\Omega)$, and $\ell(v) = \int_\Omega g v$ where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$.
+The exact solution of VI \eqref{eq:vi} is easily computed for any $p$; see Figure \ref{fig:results:plap}.
+For this experiment we used $P_1$ elements, a coarsest mesh of $m_0=7$ nodes, and nine standard refinements to a $m_9=3073$ finest mesh.
+The default smoother uses three (fixed) iterations of RS Newton, with LU solution of the step equations.\footnote{This is an optimal $O(m_j)$ smoother in 1D.}  The coarse solver is the same but iterated to convergence.
+The initial iterate was $w=\psi$, and again \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$.
+
+%REGENERATE using elbueler/plapfigs branch of git@bitbucket.org:pefarrell/fascd.git, in examples/
+% $ python3 plap1d.py -p 1.5 -figures -levs 6 -fascd_cycle_type full && mv solution.pdf plap1d1p5.pdf
+% $ python3 plap1d.py -p 4.0 -figures -levs 6 -fascd_cycle_type full && mv solution.pdf plap1d4p0.pdf
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=0.45\textwidth]{fixfigs/plap1d1p5.pdf} \quad
+\includegraphics[width=0.45\textwidth]{fixfigs/plap1d4p0.pdf}
+\end{center}
+\caption{Solutions $u\ge \psi$ of the $p$-Laplacian VI in Example \ref{ex:results:plap}, for $p=1.5$ (left) and $p=4.0$ (right).}
+\label{fig:results:plap}
+\end{figure}
+
+Figure \ref{fig:imagesvcycle} illustrates the major steps of an example V-cycle, showing each of the nonlinear FAS coarse correction VI subproblem solutions.
+That is, each $y^j \in \mathcal{D}^j$ and $z^j \in \mathcal{U}^j$ is drawn, along with each level defect constraint $\underline{\phi}^j,\underline{\chi}^j$.
 
@@
 
+Iteration results for the fast diffusion $p=1.5$ case are shown in Table \ref{tab:results:fastplap1d}.
+We see that $V(1,0)$ cycles, which correspond to an FAS implemention of Algorithm 4.7 in \cite{GraeserKornhuber2009}, but with a stronger smoother, often do not converge.
+By contrast, $V(0,1)$ cycles, using only up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are just as good as the default $V(1,1)$ cycles of Algorithm \ref{alg:fascd}.
+These default V-cycles give convergence in the $\|\cdot\|_\infty$ norm (last column) at rate $O(h^{2.005})$.
+
+Turning to the F-cycle results in Table \ref{tab:results:fastplap1d}, they are very efficient except when they fail entirely.
+This apparently arises from singular diffusivity.
+If we regularize the diffusivity formula, from $D=|\grad u|^{-0.5}$ to $D_\eps = \left(\eps + |\grad u|^2\right)^{-0.25}$ with $\eps=10^{-8}$, then a single F-cycle suffices at every resolution (column ``F$_\eps$'').
+
+Iteration results for the (unregularized) degenerate diffusion $p=4.0$ case are in Table \ref{tab:results:degenerateplap1d}.
+V-cycles using the default settings, specifically the fixed three RS Newton iterations, do not converge on the finer meshes; the semi-smooth NCP residual norms \eqref{eq:stoppingcriterion} are unbounded in the failed runs.
+Visualization shows an instability in which oscillations in the restricted residuals explode.
+Fortunately, either a slight increase in smoother power, to four RS Newton iterations, or improved finest-level initial iterates, through use of an F-cycle, stops the instability.
+In fact, a single F-cycle suffices at every resolution, giving convergence at rate $O(h^{1.420})$.
+\end{example}
+
+% for Tables \ref{tab:results:fastplap1d} and \ref{tab:results:degenerateplap1d},
+% see git@bitbucket.org:pefarrell/fascd.git: examples/plap1d.py and edplay/resultsplap1d.sh|txt
+
+\begin{table}[ht]
+\begin{tabular}{lr@{\hskip 7mm}c@{\hskip 3mm}c@{\hskip 3mm}c@{\hskip 4mm}c@{\hskip 5mm}c@{\hskip 6mm}c}
+\emph{levels} & $m_J$ & V(1,0) & V(0,1) & V(1,1) & F & F$_\eps$ & $\|$error$\|_\infty$ \\ \cmidrule{1-8}
+ 1 &    $7$ &  1 &  1 &  1 &  1 &  1 & $2.3\times 10^{-1}$ \\
+ 2 &   $13$ &  3 &  3 &  2 &  2 &  1 & $3.3 \times 10^{-2}$ \\
+ 3 &   $25$ & 12 &  4 &  2 &  1 &  1 & $9.1 \times 10^{-3}$ \\
+ 4 &   $49$ & NC &  4 &  2 &  1 &  1 & $3.2 \times 10^{-3}$ \\
+ 5 &   $97$ & NC &  3 &  3 &  1 &  1 & $5.5 \times 10^{-4}$ \\
+ 6 &  $193$ & NC &  3 &  3 & NC &  1 & $1.7 \times 10^{-4}$ \\
+ 7 &  $385$ & NC &  3 &  3 &  1 &  1 & $4.7 \times 10^{-5}$ \\
+ 8 &  $769$ & NC &  4 &  7 &  1 &  1 & $9.5 \times 10^{-6}$ \\
+ 9 & $1537$ & NC &  5 &  6 & NC &  1 & $3.6 \times 10^{-6}$ \\
+10 & $3073$ & NC & 11 & 19 &  1 &  1 & $4.9 \times 10^{-7}$
+\end{tabular}
+\bigskip
+\caption{Iterations for V- and F-cycles on the $p=1.5$ problem in Example \ref{ex:results:plap}.  NC $=$ did not converge after 50 iterations.  See text regarding F$_\eps$ and error.}
+\label{tab:results:fastplap1d}
+\end{table}
+
+\begin{table}[ht]
+\begin{tabular}{lr@{\hskip 7mm}c@{\hskip 4mm}c@{\hskip 4mm}c@{\hskip 6mm}c}
+\emph{levels} & $m_J$ & V & V$_4$ & F & $\|$error$\|_\infty$ \\ \cmidrule{1-6}
+ 1 &    $7$ &  1 &  1 &  1 & $8.7 \times 10^{-2}$ \\
+ 2 &   $13$ &  1 &  1 &  1 & $3.8 \times 10^{-2}$ \\
+ 3 &   $25$ &  2 &  1 &  1 & $1.6 \times 10^{-2}$ \\
+ 4 &   $49$ &  4 &  2 &  1 & $6.3 \times 10^{-3}$ \\
+ 5 &   $97$ & NC &  2 &  1 & $2.3 \times 10^{-3}$ \\
+ 6 &  $193$ & NC &  1 &  1 & $7.0 \times 10^{-4}$ \\
+ 7 &  $385$ & NC &  2 &  1 & $2.0 \times 10^{-4}$ \\
+ 8 &  $769$ & NC &  2 &  1 & $1.1 \times 10^{-4}$ \\
+ 9 & $1537$ & NC &  2 &  1 & $4.0 \times 10^{-5}$ \\
+10 & $3073$ & NC &  2 &  1 & $1.5 \times 10^{-5}$
+\end{tabular}
+\bigskip
+\caption{Iterations for V- and F-cycles on the $p=4.0$ problem in Example \ref{ex:results:plap}.  See text regarding V$_4$ and error.}
+\label{tab:results:degenerateplap1d}
+\end{table}
+
 \begin{example}  \label{ex:results:advdiff}
-FIXME results for advection-diffusion obstacle problem, Example \ref{ex:advectiondiffusion}
+FIXME results for 3D advection-diffusion obstacle problem, Example \ref{ex:advectiondiffusion}
 \end{example}

### 2023-06-24 paper/fascd.tex
plan to include weak scaling in SIA results

@@
 In this Section we show numerical results corresponding to several VI examples from Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
-Our essential observation is that, with strong $O(m_J)$ smoothers on each level, each V-cycle dramatically reduces the residual norm, and generally a couple of F-cycles solves the problem down to discretization error.
+Assuming strong $O(m_J)$ smoothers on each level, as illustrated in the Examples, our essential observations are that each V-cycle dramatically reduces the residual norm, and that a couple of F-cycles solves the problem down to discretization error.
 
@@
 
-% see git@bitbucket.org:pefarrell/fascd.git: examples/drape.py and edplay/resultsdrape.sh|txt
+% see git@bitbucket.org:pefarrell/fascd.git: examples/drape.py and edplay/results/drape.sh|txt
 % note mcoarse=4, rtol=1e-6, atol=1e-12
@@
 % for Tables \ref{tab:results:fastplap1d} and \ref{tab:results:degenerateplap1d},
-% see git@bitbucket.org:pefarrell/fascd.git: examples/plap1d.py and edplay/resultsplap1d.sh|txt
+% see git@bitbucket.org:pefarrell/fascd.git: examples/plap1d.py and edplay/results/plap1d.sh|txt
 
@@
 
+Solvers implemented in Firedrake apply parallel solvers from the PETSc library \cite{Balayetal2023}.
+All of the above Examples, which were run in serial so as to fix the smoother,\footnote{The precise action of $O(m_j)$ work parallel smoothers depends on the number of processes \cite[for example]{Bueler2021}.} can in fact be run efficiently in parallel.
+We start the next example with a demonstration of parallel weak scaling, and then explore the performance of the FASCD method up to memory-limited resolutions.
+
 \begin{example}   \label{ex:results:sia}
-FIXME results for SIA obstacle problem, Example \ref{ex:sia}
+FIXME results for SIA obstacle problem, Example \ref{ex:sia}; start with weak scaling; show Paraview-generated Figure
 \end{example}
@@
 
-FIXME
-
-FIXME the observation that up-smoothing is more efficient; this observation seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}
+FIXME the observation that up-smoothing is more efficient seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}
+
+FIXME smoother could be AMG
 

### 2023-06-24 paper/fascd.tex
avoid losing thoughts to unpaid slack

@@
 
+\begin{comment}
+from slack, regarding SIA example:
+Patrick:  The convergence criteria are relatively loose?
+Ed:  Yes, the convergence is loose, and furthermore there is a double-regularization in the nonlinear diffusivity.
+The problem is that only the 8/3 power of the solution u is actually in W^{1,p} for p=4; for the flat bed case u^{8/3} is actually C^1.
+But the derivative of u itself is not actually integrable, though u is C^0; you can see there is essentially a cliff all the way around the ice sheet.
+So as we converge to the exact solution, along the free boundary the P1 elements are trying to match this unbounded derivative, and it struggles at tighter tolerances.
+The best (but still not like textbook problems) results are with a Petrov-Galerkin method using Q1, for which the convergence rates are much closer to O(h^2).
+See Bueler (2016) for this P-G method, and Jouvet & Bueler (2012) for an existence proof giving u^{8/3} in W^{1,4}.
+If you really want to know ...
+\end{comment}
+
 % A BRIDGE TOO FAR:  \section{Results for a nonlocal variational inequality}

### 2023-06-24 paper/fascd.tex
add SIA eye candy

@@
 
-The next VI example, nonlinear and generally not of optimization type, is a model for steady ice sheet geometry in a given climate.
-Related VI problems arise for any fluid layer which is subject to a climate which can remove the fluid \cite{Bueler2021conservation}.
+The next VI example, nonlinear and not of optimization type, is a model for steady ice sheet geometry in a given climate.
+Related VI problems arise for any fluid layer which is subject to a source term which can remove the fluid \cite{Bueler2021conservation}.
 
 \begin{example}  \label{ex:sia}  Let $\Omega \subset \RR^2$ be a fixed region of land, and suppose $b \in C^1(\Omega)$ is the given bedrock elevation.
-Suppose $a \in L^\infty(\Omega)$ is the ``surface mass balance'' function, the given, annually-averaged rate of ice accumulation (as snow) minus melt and runoff.
-Then the surface elevation $s$ of a steady-state, isothermal, and non-sliding ice sheet (glacier) approximately satisfies the \emph{shallow ice approximation} (SIA) \cite{GreveBlatter2009} problem
+Let $a \in L^\infty(\Omega)$ denote a given ``surface mass balance'' function, the annually-averaged rate of ice accumulation (as snow) minus melt and runoff.
+Then the surface elevation $s\in C(\Omega)$ of a steady-state, isothermal, and non-sliding ice sheet (glacier) approximately satisfies the \emph{shallow ice approximation} (SIA) \cite{GreveBlatter2009} problem
 \begin{equation}
@@
 \end{equation}
-over the set $\mathcal{K} = \{s\ge b\}$ of admissible surface elevations, where $n>1$ and $\Gamma>0$ are constants determined by the physical properties of ice \cite{GreveBlatter2009}; $n=3$ is typically taken.
+over the set $\mathcal{K} = \{s\ge b\}$ of admissible surface elevations, where $n>1$ and $\Gamma>0$ are constants determined by the physical properties of ice; $n=3$ is a typical value.
 The doubly-nonlinear and doubly-degenerate operator in this case is
@@
 If the bed is flat ($b=0$) then a power transformation converts the problem into $p$-Laplacian form \eqref{eq:plaplacian} with $p=n+1$.
-For general beds $b$ less is understood about this operator, but existence of a solution to \eqref{eq:siavi} can be established \cite{JouvetBueler2012}, and the theory suggests that a solution power lives in a Sobolev space, namely $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.
-This theory, and also observations of ice sheets, shows that $|\grad s|$ will be large, or even unbounded, as one approaches the free boundary, also know as the \emph{ice margin}, from the icy side where $s>b$.
+For general beds $b$ less is understood about this operator, but existence holds for \eqref{eq:siavi} \cite{JouvetBueler2012}.
+The theory suggests that a power of the solution lives in a Sobolev space, namely $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.
+This theory, and also observations of ice sheets, shows that $|\grad s|$ is generally unbounded as one approaches the free boundary, the \emph{ice margin}, from the icy side (where $s>b$).
 \end{example}
 
-Section \ref{sec:results} provides numerical results from applications of the \pr{fascd} method to Examples \ref{ex:plaplacian}, \ref{ex:advectiondiffusion}, and \ref{ex:sia}.
+Numerical results are shown in Section \ref{sec:results} for applications of FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} to these Examples.
 
@@
 \end{center}
-\caption{The \fascd V-cycle computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.}
+\caption{The FASCD V-cycle computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.}
 \label{fig:fascdvcycle}
@@
 \end{pseudo}
-\caption{The \fascd V-cycle is an iteration for solving FE VI problem \eqref{eq:fe:vi}.  Here $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
+\caption{The FASCD V-cycle is an iteration for solving FE VI problem \eqref{eq:fe:vi}.  Here $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
 \label{alg:fascd}
@@
 Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}.
-In this Section we show numerical results corresponding to several VI examples from Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
-Assuming strong $O(m_J)$ smoothers on each level, as illustrated in the Examples, our essential observations are that each V-cycle dramatically reduces the residual norm, and that a couple of F-cycles solves the problem down to discretization error.
+In this Section we show numerical results corresponding to the several VI Examples in Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
+Assuming strong $O(m_J)$ smoothers on each level, as illustrated below, our essential observations are that each V-cycle dramatically reduces the residual norm, and that a couple of F-cycles solves most Examples down to discretization error.
 
@@
 Solvers implemented in Firedrake apply parallel solvers from the PETSc library \cite{Balayetal2023}.
-All of the above Examples, which were run in serial so as to fix the smoother,\footnote{The precise action of $O(m_j)$ work parallel smoothers depends on the number of processes \cite[for example]{Bueler2021}.} can in fact be run efficiently in parallel.
-We start the next example with a demonstration of parallel weak scaling, and then explore the performance of the FASCD method up to memory-limited resolutions.
+All of the above Examples, which were run in serial so as to fix the smoother,\footnote{The precise action of $O(m_j)$ work parallel smoothers depends on the number of processes \cite[for example]{Bueler2021}.} also run efficiently in parallel.
+We start the next Example with a demonstration of parallel weak scaling \cite{Bueler2021}, and then explore the performance of the FASCD up to memory-limited resolutions.
 
 \begin{example}   \label{ex:results:sia}
-FIXME results for SIA obstacle problem, Example \ref{ex:sia}; start with weak scaling; show Paraview-generated Figure
+FIXME results for SIA obstacle problem, Example \ref{ex:sia}; start with weak scaling
 \end{example}
+
+% to regenerate:
+%   $ mpiexec -n 16 --map-by core --bind-to hwthread python3 sia.py -levs 8 -bumps \
+%       -mcoarse 8 -fascd_cycle_type full -fascd_monitor -fascd_rtol 2.0e-4 \
+%       -fascd_atol 1.0e-8 -o sialev8.pvd
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=1.0\textwidth]{fixfigs/sialev8scene.png}
+\end{center}
+\caption{FIXME}
+\label{fig:results:siascene}
+\end{figure}
 

### 2023-06-25 paper/fascd.tex
text in intro, and cleanup near end

@@
 
-FIXME organization of paper
+This paper is organized as follows.
+In Section \ref{sec:vi} we recall the theory of coercive variational inequalities, with several example problems, and then Section \ref{sec:cd} reviews the basics of CDs, including the associated iterations, as proposed by Tai \cite{Tai2003}.
+Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, along with the special monotone restriction operators used here.
+Section \ref{sec:cdmultilevel} presents defect constraints, leading to an apparently new understanding of telescoping sets for up-smoothing.
+Section \ref{sec:vcycle} then presents the FASCD V-cycle, with Section \ref{sec:implementation} addressing implementation and smoother semantics, plus F-cycles.
+Section \ref{sec:results} gives FASCD numerical results for the Section \ref{sec:vi} problems, followed by discussion and outlook in Section \ref{sec:discussion}.
+An Appendix sketchs how FASCD reduces to classical FAS multigrid for PDEs.
 
@@
 
-\section{Basic CD iterations} \label{sec:cditers}
-
-Associated to a constraint decomposition (CD; equations \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}) are certain iterative methods \cite{Tai2003,Xu1992}.
+Associated to a constraint decomposition \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} are certain iterative methods \cite{Tai2003,Xu1992}.
 Let $0\le \alpha \le 1$ be the \emph{damping parameter}.
@@
 Compared specifically to the $V(1,0)$ cycle presented as Algorithm 4.7 in \cite{GraeserKornhuber2009}, one might say informally that \pr{fascd-vcycle} is a $V(1,1)$ cycle.
-However, its structure relates directly to the basic CD iterations (Sections \ref{sec:cditers}, \ref{sec:cdmultilevel}) as well.
-In unilateral cases it is precise to say that one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} over CD \eqref{eq:fe:downwardsuminclusion}, using the defect constraint meaning of the downward constraint sets $\mathcal{D}^j$, followed by an application of \pr{icd-tele} over the incomplete CD \eqref{eq:fe:upwardsuminclusion}.
+However, its structure relates directly to the basic CD iterations (Sections \ref{sec:cd}) as well.
+In unilateral cases it is precise to say that one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} over CD \eqref{eq:fe:downwardsuminclusion}, using the defect constraint meaning of the downward constraint sets $\mathcal{D}^j$, followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over the incomplete CD \eqref{eq:fe:upwardsuminclusion}.
 FIXME: idea is that once you are using defect constraints you no longer depend on the literal availability of $\Pi_i$, so incomplete CD suffices, e.g.~even for \pr{cd-mult}
@@
 
-FIXME stopping criteria for smoother
-
-FIXME F-cycles in Algorithm \ref{alg:fascd-fcycle}
+FIXME F-cycles
 
@@
 \end{pseudo}
-\caption{FIXME}
+\caption{The FASCD F-cycle for solving FE VI problem \eqref{eq:fe:vi}.}
 \label{alg:fascd-fcycle}
@@
 
-% A BRIDGE TOO FAR:  \section{Results for a nonlocal variational inequality}
-
 \section{Discussion and Conclusion} \label{sec:discussion}
@@
 FIXME smoother could be AMG
+
+FIXME desired extension to nonlocal variational inequalities
 
@@
 \appendix
-\section{Reduction of FASCD to FAS multigrid for PDEs} \label{app:reductions}
-
-\pr{fascd-vcycle} (Algorithm \ref{alg:fascd}) in the main text generalizes existing multilevel algorithms.
-Here we show that it reduces to the FAS multigrid V-cycle for nonlinear PDEs \cite{Trottenbergetal2001}, which itself generalizes linear multigrid, when all inequality constraints are removed.
-
-Suppose we remove the inequality constraints from the finest-level VI problem \eqref{eq:fe:vi} by replacing the finest-level obstacles $\underline{\gamma}^J$ and $\overline{\gamma}^J$ with very negative and positive constants, respectively, thus not constraining the solution $u^J$ or any possible iterate.
+\section{Reduction to FAS multigrid for PDEs} \label{app:reductions}
+
+\pr{fascd-vcycle} (Algorithm \ref{alg:fascd}) reduces to the classical FAS multigrid V-cycle for nonlinear PDEs \cite{Trottenbergetal2001}, which itself generalizes linear multigrid, when all inequality constraints are removed.
+In fact, suppose we remove the inequality constraints from the finest-level VI problem \eqref{eq:fe:vi} by replacing the finest-level obstacles $\underline{\gamma}^J$ and $\overline{\gamma}^J$ with very negative and positive constants, respectively, thus not constraining the solution $u^J$ or any possible iterate.
 The admissible set $\mathcal{K}^J$ of \eqref{eq:fe:vi} is then determined only by the Dirichlet boundary conditions.
@@
 Now the downward/upward VIs \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} are also unconstrained, and can be rewritten as weak forms $\ip{f^j(w^j + y^j)}{v} = \ip{\ell^j}{v}$ and $\ip{f^j(w^j + z^j)}{v} = \ip{\ell^j}{v}$, respectively, for all $v\in \mathcal{V}_0^j$.
-However, we may simply store the unconstrained, approximate solution $u^j=w^j+y^j$, instead of $w^j$ and $y^j$ separately, and similarly for the upward solution $u^j=w^j+z^j$.
+Then we combine $u^j=w^j+y^j$, instead of $w^j$ and $y^j$ separately, and similarly $u^j=w^j+z^j$ when up-smoothing.
 The smoothers and the coarsest-level solver can be regarded as unconstrained, in-place modifications of $u^j$, so we denote them simply as ``$\text{\pr{smooth}}(\ell^j;u^j)$.''  Noting line 11 of Algorithm \ref{alg:fascd} now says $u^{j-1}=\iR u^j$, after down-smoothing the coarsened source functional computed by line 12 can be rewritten as
@@
 
-Going upward we must be careful that only the coarse correction, and not the coarse solution, is prolonged.\footnote{Compare Remark 5.3.9 in \cite{Trottenbergetal2001}, a warning to novices and everyone else.}  After up-smoothing on the $j-1$ level the coarse iterate is $u^{j-1}$, but at this point $u^j=w^j + y^j$ contains the result of down-smoothing on the $j$th level.
-On the other hand, in Algorithm \ref{alg:fascd} the initial value of the correction $z^j$, before up-smoothing on the $j$th level, is $z^j = y^j + P z^{j-1}$, where $z^{j-1}$ contains the total correction from coarser corrections and up-smoothing at the $j-1$ level.
+Going upward we must be careful that only the coarse correction, and not the coarse solution, is prolonged \cite[remark 5.3.9]{Trottenbergetal2001}.
+After up-smoothing on the $j-1$ level the coarse iterate is $u^{j-1}$, but at this point $u^j=w^j + y^j$ contains the result of $j$th-level down-smoothing.
+The initial value of the correction $z^j$, before up-smoothing on the $j$th level, is $z^j = y^j + P z^{j-1}$, where $z^{j-1}$ contains the total correction from coarser corrections and $j-1$ level up-smoothing.
 Thus the $j$th-level iterate \emph{before} up-smoothing must be set to
@@
 
-With these modifications, we have a much-simplified pseudocode, Algorithm \ref{alg:fas}.
+With these modifications, we have a much-simplified Algorithm \ref{alg:fas}.
 This FAS V-cycle appears in various places, for example Algorithm 14 of \cite{Bruneetal2015} or in section 5.3.4 of \cite{Trottenbergetal2001}.

### 2023-06-25 paper/fascd.tex
reasonable SIA example

@@
 
-Solvers implemented in Firedrake apply parallel solvers from the PETSc library \cite{Balayetal2023}.
-All of the above Examples, which were run in serial so as to fix the smoother,\footnote{The precise action of $O(m_j)$ work parallel smoothers depends on the number of processes \cite[for example]{Bueler2021}.} also run efficiently in parallel.
-We start the next Example with a demonstration of parallel weak scaling \cite{Bueler2021}, and then explore the performance of the FASCD up to memory-limited resolutions.
+Solvers implemented in Firedrake exploit parallel solvers from the PETSc library \cite{Balayetal2023}.
+In fact, all of the above Examples, which were run in serial so as to fix the smoother,\footnote{The precise action of $O(m_j)$ work parallel smoothers depends on the number of processes \cite[for example]{Bueler2021}.} also run efficiently in parallel.
+The next Example demonstrates the parallel weak scaling \cite{Bueler2021} of FASCD F-cycle solvers.
 
 \begin{example}   \label{ex:results:sia}
-FIXME results for SIA obstacle problem, Example \ref{ex:sia}; start with weak scaling
-\end{example}
-
+Recall the steady shallow ice approximation (SIA) obstacle problem, Example \ref{ex:sia}, in which the surface elevation $s$ is constrained to exceed the bedrock elevation $b$.
+For a test case we set $\Omega=(0,L)^2$, with $L=1800$ km, a radial surface mass balance $a$ function \cite[equation (5.122)]{GreveBlatter2009}, and bumpy, synthetic values for $b\in C^1(\Omega)$, with an elevation range of $1800$ m.
+The numerical solution, colored by the ice flow, is shown in Figure \ref{fig:results:siascene}; this used the surface value of the SIA solution, namely $\mathbf{u} = (n+1)(n+2)^{-1} \Gamma (s-b)^{n+1} |\grad s|^{n-1} \grad s$.
+As noted in Example \ref{ex:sia} and seen in this Figure, $|\grad s|$ is unbounded at the free boundary, so the solution has low regularity.
+
+\vspace{-2mm}
 % to regenerate:
@@
 \begin{center}
-\includegraphics[width=1.0\textwidth]{fixfigs/sialev8scene.png}
+\includegraphics[width=0.95\textwidth]{fixfigs/sialev8scene.png}
 \end{center}
-\caption{FIXME}
+\caption{A high-resolution ($m^J\approx 10^6$) solution of Example \ref{ex:sia} over a bumpy bedrock topography, with the ice speed as the surface coloring.}
 \label{fig:results:siascene}
 \end{figure}
+
+We applied FASCD F-cycles using a smoother of four (fixed) iterations of RS Newton (Section \ref{sec:implementation}), with quadratic backtracking line search.
+The step equations were approximately solved, in parallel, by three (fixed) iterations of additive-Schwarz preconditioned GMRES, with ILU preconditioning on the blocks.
+The coarse solver was again RS Newton, with LU preconditioning redundantly over all processors.
+The coarse mesh was relatively fine, with $m_0=289$ nodes and 512 triangles.
+Because of the low regularity of the solution we used \id{rtol} $= 2 \times 10^{-4}$ and \id{atol} $= 10^{-8}$ in \eqref{eq:stoppingcriterion}.
+The initial iterate was zero ice: $s=b$.
+
+These solver parameters generate a weak-scaling solver.
+For $P$ process solutions we fixed the number of degrees of freedom per process at $m_J/P=1025^2$ and solved over $P=1,4,16$ processes.\footnote{Runs were on a single-node workstation with two Intel Xeon 4210 cpus (2.2GHz), a total of 20 physical cores, and 128 GB shared memory.}  The results in Table \ref{tab:results:siaweak} showing mesh independence in the number of FASCD cycles and clear weak scaling.
+The finest mesh result, with $m_J=4097^2=17 \times 10^6$ degrees of freedom, needed under six minutes to solve for the surface elevation of an ice sheet slightly larger than the current Greenland ice sheet ($1.7$ million $\text{km}^2$) at uniform $439$ m resolution.
+\end{example}
+
+% see pefarrell-fascd/edplay/siaweak.sh|txt
+% time in seconds:
+%>> 5*60+[11,36,55]
+%ans =      311      336      355
+\begin{table}[ht]
+\begin{tabular}{c@{\hskip 4mm}c@{\hskip 4mm}c@{\hskip 7mm}c@{\hskip 4mm}c}
+P & \emph{levels} & $m_J$ & \emph{F-cycles} & \emph{time} (s) \\ \cmidrule{1-5}
+ 1 & 7 & $1025^2$ & 4 & 311 \\
+ 4 & 8 & $2049^2$ & 4 & 336 \\
+16 & 9 & $4097^2$ & 4 & 355
+\end{tabular}
+\bigskip
+\caption{Iterations and run time (wall clock, in seconds) for parallel FASCD F-cycle solutions of Example \ref{ex:sia}.}
+\label{tab:results:siaweak}
+\end{table}
 

### 2023-06-26 paper/fascd.tex
draft abstract, and clean intro

@@
 
-\begin{abstract} FIXME
+\begin{abstract}
+The \emph{full approximation storage constraint decomposition} (FASCD) multilevel finite element method introduced here is a common extension of the full approximation storage multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and of the constraint decomposition method introduced by X.-C.~Tai for variational inequalities.
+We extend the existing constraint decomposition ideas by recognizing and exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
+Then, by using strong smoothers with work proportional to the number of unknowns on the finite element mesh level, we demonstrate that FASCD V-cycles show superior mesh-independent convergence rates and that FASCD F-cycles (full multigrid) are optimal solvers.
+The example problems include unilateral and bi-obstacle variational inequality problems with linear and nonlinear differential operators.
 \end{abstract}
@@
 These methods have been shown to converge for multilevel and domain decompositions in a finite element (FE) context.
-The multilevel CD method has been shown to have almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
+In the multilevel case, the CD method has been shown to have almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
 In this paper we extend the multilevel CD method by replacing the coarse-level corrections with a full approximation storage (FAS; \cite{Brandt1977,Bruneetal2015}) approach, and we apply this new solver to nonlinear VI problems which do not arise from minimization.
@@
 \begin{enumerate}
-\item The FAS coarse corrections are adapted for general VI problems.
-The needed computations do not require an objective function, nor do they require linearity of the residual operator.
-Section \ref{sec:results} shows results for several nonlinear problems.
-\item By using $O(N)$-per-application Newton-type active set methods as smoothers, with a fixed number of Krylov iterations per step, we demonstrate mesh-independent convergence rates substantially better than in the current literature.
-\item We implement both V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (also known as full multigrid; Algorithm \ref{alg:fascd-fcycle}).
+\item The FAS coarse corrections are adapted for general VI problems.  The needed computations require neither an objective function nor linearity of the residual operator.
 \item Box constraints, namely upper and lower obstacles, are permitted; see Section \ref{sec:femultilevel}.
 \item We observe an asymmetry special to multilevel CD methods, namely that up-smoothing is more effective than down-smoothing because a ``telescoping'' subset property can be exploited on the up-smoothing side.
+\item We implement V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle})
+\item By using a Newton-type active set method as a smoother, with a fixed number of Krylov iterations per step so that the work of the smoother is proportional to the number of unknowns on the given level, we demonstrate that F-cycle FASCD is an optimal solver.
 \end{enumerate}
 
-FASCD exhibits essentially the same optimal, mesh-independent convergence rates for VI problems that FAS multigrid exhibits for the corresponding unconstrained PDE problems.
-The results in Section \ref{sec:results} include textbook multigrid efficiency, namely solutions within discretization error in work comparable to ten smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}, of F-cycles applied to classical Laplacian obstacle problems (Example \ref{ex:results:classical}), $p$-Laplacian obstacle problems (Example \ref{ex:results:plap}), and nonnegativity-constrained advection-diffusion equations (Example \ref{ex:results:advdiff}).
-Furthermore, these results are essentially insensitive to the length or structure of the free boundary, i.e.~the geometric complexity of the coincidence set (Examples \ref{ex:results:classical}, \ref{ex:results:advdiff}).
+FASCD V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for the corresponding (unconstrained) PDE problems.
+The F-cycle results in Section \ref{sec:results} show nearly-textbook multigrid efficiency, namely solutions within discretization error in work comparable to few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}, for classical Laplacian obstacle problems, $p$-Laplacian obstacle problems, and nonnegativity-constrained advection-diffusion equations.
+Furthermore, these results are essentially insensitive to the length or structure of the free boundary, i.e.~the geometric complexity of the coincidence set.
 
 The iterates from FASCD are admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
-For example, we solve an ice sheet geometry problem (Examples \ref{ex:sia}, \ref{ex:results:sia}) in which the ice velocity is only meaningful for surface elevation iterates which exceed the bed elevation.
-More generally, fluid-layer problems, with a non-negative thickness constraint, would require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation \cite{Bueler2021conservation}.
-(Though a target for FASCD, such problems, however, are not addressed here.)  Note that non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural modifications of the operator formula in such a fluid-layer context.
-In any case, FASCD solves the difficult ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a mesh-independent number of iterations, with near textbook efficiency.
-
-For the classical obstacle problem, based on the linear, symmetric Laplacian operator, certain multilevel techniques, such as the truncated monotone multigrid method \cite{GraeserKornhuber2009,Kornhuber1994}, are known to improve performance relative to the original multilevel CD method \cite{Tai2003}.
-These improved methods modify the nodal basis functions.
-By contrast FASCD makes essentially no assumptions about the FE discretization.
-Acceleration via basis-level manipulations is a topic for future research.
+For example, we solve an ice sheet geometry problem (Examples \ref{ex:sia}, \ref{ex:results:sia}) in which the ice flow formulas are only meaningful for surface elevation iterates which exceed the bed elevation.
+(More generally, fluid-layer problems, with a non-negative thickness constraint, often require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation \cite{Bueler2021conservation}.
+However, though a target for future applications of FASCD, such problems are not addressed here.)  Note that non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural modifications of the operator formula in such a fluid-layer context.
+FASCD solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a mesh-independent number of F-cycle iterations.
 
 This paper is organized as follows.
-In Section \ref{sec:vi} we recall the theory of coercive variational inequalities, with several example problems, and then Section \ref{sec:cd} reviews the basics of CDs, including the associated iterations, as proposed by Tai \cite{Tai2003}.
-Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, along with the special monotone restriction operators used here.
-Section \ref{sec:cdmultilevel} presents defect constraints, leading to an apparently new understanding of telescoping sets for up-smoothing.
-Section \ref{sec:vcycle} then presents the FASCD V-cycle, with Section \ref{sec:implementation} addressing implementation and smoother semantics, plus F-cycles.
-Section \ref{sec:results} gives FASCD numerical results for the Section \ref{sec:vi} problems, followed by discussion and outlook in Section \ref{sec:discussion}.
-An Appendix sketchs how FASCD reduces to classical FAS multigrid for PDEs.
+In Section \ref{sec:vi} we recall the theory of coercive and monotone variational inequalities, with several example problems.
+Then Section \ref{sec:cd} reviews the basics of CDs, including the associated iterations, as proposed by Tai \cite{Tai2003}.
+Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, along with special monotone restriction operators \cite{GraeserKornhuber2009}.
+Section \ref{sec:cdmultilevel} presents defect constraints \cite{GraeserKornhuber2009}, leading to an apparently new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD.
+Section \ref{sec:vcycle} then states the FASCD V-cycle, with Section \ref{sec:implementation} addressing implementation concerns, smoother semantics, and F-cycles.
+Section \ref{sec:results} gives FASCD numerical results, for problems in Section \ref{sec:vi}, from a parallel Firedrake \cite{Rathgeberetal2016} implementation.
+Discussion and outlook in Section \ref{sec:discussion} and an Appendix, which sketchs how FASCD reduces to classical FAS multigrid for PDEs, finish the work.
 

### 2023-06-26 paper/fascd.tex
clean Examples 8.1, 8.2

@@
 Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}.
-In this Section we show numerical results corresponding to the several VI Examples in Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
-Assuming strong $O(m_J)$ smoothers on each level, as illustrated below, our essential observations are that each V-cycle dramatically reduces the residual norm, and that a couple of F-cycles solves most Examples down to discretization error.
+In this Section we show numerical results for the several VI Examples in Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
+Assuming strong, yet $O(m_J)$ arithmetic, smoothers on each level, as illustrated below, our essential observations are that each V-cycle dramatically reduces the residual norm, and that a couple of F-cycles solves most Examples down to discretization error.
 
 \begin{example} \label{ex:results:classical}
-Consider the 2D classical, unilateral obstacle problem, namely Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$, with $\Omega$ square.
-We consider two problems with very different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$.
-The first is a standard ``ball'' problem, with an upper hemisphere obstacle $\psi$ \cite[Chapter 12]{Bueler2021}, and the second is the ``spiral'' problem from Example 7.1.1 in \cite{GraeserKornhuber2009}, wherein $\psi$ is constructed to produce a more interesting coincidence set; see Figure \ref{fig:results:classical}.
-For this experiment we used $P_1$ elements and uniform triangular meshs, from a coarsest mesh of $m_0=5^2$ nodes (32 triangles), to a finest mesh of $m_9=2049^2 \approx 4.2 \times 10^{6}$ nodes after standard refinements.
-As our $O(m_j)$ smoother we applied a single iteration of RS Newton (Section \ref{sec:implementation}), with three (fixed) ICC-preconditioned CG iterations to approximately solve the step equations.
+Consider the 2D classical, i.e.~unilateral and Laplacian, obstacle problem, namely Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$.
+We consider two problems with very different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$, both over square $\Omega$.
+The standard ``ball'' problem has an upper hemisphere obstacle $\psi$ \cite[Chapter 12]{Bueler2021}.
+In the ``spiral'' problem from Example 7.1.1 in \cite{GraeserKornhuber2009}, $\psi$ is constructed to produce a more interesting coincidence set; see Figure \ref{fig:results:classical}.
+Note that $\psi\big|_{\partial \Omega} \le 0$ in both problems.
+
+For this experiment we used $P_1$ elements, standard uniform mesh refinement, and a coarsest mesh of $m_0=5^2$ nodes (32 triangles).
+The finest mesh had $m_9=2049^2 \approx 4.2 \times 10^{6}$ nodes.
+As our smoother we applied a single RS Newton iteration (Section \ref{sec:implementation}), with three (fixed) ICC-preconditioned CG iterations to approximately solve the step equations.
 The coarse solver was RS$+$LU, iterated to convergence.
 Tolerances \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$ were used in stopping criterion \eqref{eq:stoppingcriterion}.
-Noting that $\psi\big|_{\partial \Omega} \le 0$ in both problems, the initial iterate was set to $w=\max\{0,\psi\}$.
+The initial iterate was set to $w=\max\{0,\psi\}$.
 
 Table \ref{tab:results:classical} shows the resulting iteration counts.
-When the RS Newton smoother is applied by itself as a solver it shows rapidly-increasing iterations as expected, approximately proportionally to resolution.
-By contrast, FASCD V-cycle iterations increase slowly, roughly logarithmically with the resolution.
-Noting that for F-cycles we are counting the number of V-cycles after the ramp (Section \ref{sec:implementation}), we see that a constant number of F-cycles solves the problem.
-The spiral problem appears to be only slightly more challenging for these solvers; apparently the complexity of the coincidence set or free boundary does not strongly influence performance.
-
-Note that each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+ICC) iterations.
-For the F-cycle results, where three V-cycle iterations suffice in either problem, and at any finest-level resolution, this is nearly textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  However, work of at most ten residual evaluations is often mentioned.} despite the low solution regularity.
+When the RS Newton smoother is applied by itself as a solver it shows rapidly-increasing iterations as expected, roughly proportional to resolution.
+By contrast, V-cycle iterations increase slowly, perhaps logarithmically with the resolution.
+Noting that for F-cycles we are counting the number of V-cycles after the ramp (Section \ref{sec:implementation}), we see that a constant number of cycles solves the problem.
+The spiral problem appears to be only slightly more challenging for these solvers.
+The complexity of the coincidence set or free boundary does not strongly-influence FASCD performance.
+
+Each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+ICC) iterations.
+For the F-cycle results, where three iterations suffice at all resolutions in either problem, this is close to textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  However, work of at most ten residual evaluations is often mentioned.} despite the low solution regularity.
 \end{example}
@@
 \end{center}
-\caption{Coincidence sets for the ball and spiral problems in Example \ref{ex:results:classical}.}
+\caption{Coincidence sets (black) for the ball and spiral problems in Example \ref{ex:results:classical}.}
 \label{fig:results:classical}
@@
 % see git@bitbucket.org:pefarrell/fascd.git: examples/drape.py and edplay/results/drape.sh|txt
-% note mcoarse=4, rtol=1e-6, atol=1e-12
-% smoother is vinewtonrsls + 3 x (cg + icc), coarse solver is vinewtonrsls + lu
 \begin{table}[ht]
@@
 Recall Example \ref{ex:plaplacian}, the nonlinear $p$-Laplacian operator.
-We will solve a unilateral obstacle problem, in one dimension for visualizability, considering both fast ($p=1.5$) and degenerate diffusivity ($p=4.0$) cases.\footnote{``Fast'' and ``degenerate'' refer to $D=|\grad u|^{p-2}\to \infty$ and $D\to 0$ as $|\grad u|\to 0$, respectively.
-Ellipticity is lost in the latter case.}  This example reveals a strong contrast between $V(1,0)$ and $V(0,1)$ results, namely \pr{up} $=0$ and \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively, and it emphasizes the need for robust smoothers.
+Here we consider a unilateral obstacle problem, in one dimension for visualizability, for both fast ($p=1.5$) and degenerate ($p=4.0$) diffusivity cases.\footnote{``Fast'' and ``degenerate'' refer to $D=|\grad u|^{p-2}\to \infty$ and $D\to 0$ as $|\grad u|\to 0$, respectively.
+Ellipticity is lost in the latter case.
+For such $p$-Laplacian obstacle problems, well-posedness and ($p$-dependent) solution regularity are considered in \cite{ChoeLewis1991}.}  This example reveals the sometimes strong contrast between $V(1,0)$ and $V(0,1)$ results, namely \pr{up} $=0$ and \pr{down} $=0$ in Algorithm \ref{alg:fascd}, and it emphasizes the need for robust smoothers.
 
@@
 
-Iteration results for the fast diffusion $p=1.5$ case are shown in Table \ref{tab:results:fastplap1d}.
-We see that $V(1,0)$ cycles, which correspond to an FAS implemention of Algorithm 4.7 in \cite{GraeserKornhuber2009}, but with a stronger smoother, often do not converge.
-By contrast, $V(0,1)$ cycles, using only up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are just as good as the default $V(1,1)$ cycles of Algorithm \ref{alg:fascd}.
-These default V-cycles give convergence in the $\|\cdot\|_\infty$ norm (last column) at rate $O(h^{2.005})$.
-
-Turning to the F-cycle results in Table \ref{tab:results:fastplap1d}, they are very efficient except when they fail entirely.
-This apparently arises from singular diffusivity.
-If we regularize the diffusivity formula, from $D=|\grad u|^{-0.5}$ to $D_\eps = \left(\eps + |\grad u|^2\right)^{-0.25}$ with $\eps=10^{-8}$, then a single F-cycle suffices at every resolution (column ``F$_\eps$'').
+Iteration results are shown in Table \ref{tab:results:fastplap1d} for the fast diffusion $p=1.5$ case.
+We see that $V(1,0)$ cycles, essentially Algorithm 4.7 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.
+By contrast, $V(0,1)$ cycles, using only up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are just as good as the default $V(1,1)$ cycles (i.e.~Algorithm \ref{alg:fascd}).
+F-cycles are even more efficient except when they fail entirely, but the failures arise from singular diffusivity.
+If we regularize  $D=|\grad u|^{-0.5}$ to $D_\eps = \left(\eps + |\grad u|^2\right)^{-0.25}$, with $\eps=10^{-8}$, then a single F-cycle suffices at every resolution (column ``F$_\eps$'').
 
 Iteration results for the (unregularized) degenerate diffusion $p=4.0$ case are in Table \ref{tab:results:degenerateplap1d}.
-V-cycles using the default settings, specifically the fixed three RS Newton iterations, do not converge on the finer meshes; the semi-smooth NCP residual norms \eqref{eq:stoppingcriterion} are unbounded in the failed runs.
+V-cycles using the default settings, specifically three fixed RS Newton iterations, do not converge on the finer meshes; the semi-smooth NCP residual norms \eqref{eq:stoppingcriterion} are unbounded in the failed runs.
 Visualization shows an instability in which oscillations in the restricted residuals explode.
-Fortunately, either a slight increase in smoother power, to four RS Newton iterations, or improved finest-level initial iterates, through use of an F-cycle, stops the instability.
-In fact, a single F-cycle suffices at every resolution, giving convergence at rate $O(h^{1.420})$.
+However, either a slight increase in smoother power, to four RS Newton iterations, or improved finest-level initial iterates through use of an F-cycle, stops the instability.
+Even for the three-iteration solver, a single F-cycle suffices at every resolution.
+
+The $p=1.5$ and $p=4$ cases give different convergence rates in the $\|\cdot\|_\infty$ norm, namely at rates $O(h^{2.005})$ and $O(h^{1.420})$, respectively.
 \end{example}
 
-% for Tables \ref{tab:results:fastplap1d} and \ref{tab:results:degenerateplap1d},
 % see git@bitbucket.org:pefarrell/fascd.git: examples/plap1d.py and edplay/results/plap1d.sh|txt
-
 \begin{table}[ht]
@@
 
+% see git@bitbucket.org:pefarrell/fascd.git: examples/plap1d.py and edplay/results/plap1d.sh|txt
 \begin{table}[ht]
@@
 
+Solvers implemented in Firedrake can exploit parallel solvers from the PETSc library \cite{Balayetal2023}.
+In fact, the above Examples, which were run in serial so as to fix the smoother,\footnote{The action of parallel preconditioners generally depends on the number of processes \cite[for example]{Bueler2021}.} also run efficiently in parallel.
+The next Example applies a parallel solver, and Example \ref{ex:results:sia} also demonstrates the parallel weak scaling \cite{Bueler2021} of FASCD.
+
 \begin{example}  \label{ex:results:advdiff}

### 2023-06-26 paper/fascd.tex
sweet bi-obstacle advection-diffusion example

@@
 \begin{example}  \label{ex:results:advdiff}
-FIXME results for 3D advection-diffusion obstacle problem, Example \ref{ex:advectiondiffusion}
+The advection-diffusion operator in Example \ref{ex:advectiondiffusion} is coercive when the velocity $\bX$ is divergence-free.
+We now demonstrate FASCD F-cycles on such a VI problem over $\Omega=(-1,1)^d$ ($d=2,3$).
+The 2D problem satisfies the coercivity hypotheses ($\bX = (7+5y,-5x)$ is divergence-free), but the 3D problem has a more general velocity ($\grad\cdot\bX \ne 0$); note $u=0$ on all of $\partial \Omega$ in both problems.
+The discontinuous source term $\phi \in L^\infty(\Omega)$ has $\phi>0$ patches in half of the domain and $\phi<0$ values in the other half.
+While the other Examples in this Section are unilateral, here we constrained the solution with upper and lower obstacles: $0 \le u \le 1$.
+The $u=0$ and $u=1$ coincidence sets for the 2D problem are shown in Figure \ref{fig:results:advdiff}; the 3D sets are comparable in regularity.
+
+% REGENERATE using git@bitbucket.org:pefarrell/fascd.git, in examples/
+%   mpiexec -n 16 --map-by core --bind-to hwthread python3 pollutant.py -dim 2 -levs 7 -o poll2d.pvd
+%   paraview poll2d.pvd
+% to plot coincidence sets use ranges 0 <= u <= 0.00001 and 0.99999 <= u <= 1 and
+% preset "xray" and invert colors (for u=0) or keep colors (u=1)
+\begin{figure}[ht]
+\begin{center}
+\includegraphics[width=0.35\textwidth]{fixfigs/poll2d-zero-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/poll2d-one-set.png}
+\end{center}
+\caption{Coincidence sets (black) for the 2D bi-obstacle problem in Example \ref{ex:results:advdiff}: $u=0$ (left) and $u=1$ (right).}
+\label{fig:results:advdiff}
+\end{figure}
+
+For the experiment we used $P_1$ elements, uniform refinement, and a coarse mesh with $m_0=16^d$ nodes so as to minimally-resolve the coincidence set.
+The smoother was two (fixed) RS Newton iterations,  with three preconditioned GMRES iterations to approximately solve the step equations.
+The additive Schwarz method (ASM) preconditioner applied incomplete factorization (ILU) preconditioning on the blocks.
+The coarse solver was RS Newton and LU, iterated to convergence.
+Tolerances \id{rtol} $=10^{-5}$, \id{atol} $=10^{-9}$ were used in \eqref{eq:stoppingcriterion}.
+
+F-cycle iteration results are shown in Table \ref{tab:results:advdiff}.
+Despite the hint of growing iterations under mesh refinement, a few iterations suffice up to memory-limited resolutions.
+For any of the finer mesh runs here, the FASCD F-cycle solver takes less time than generating the $P_1$ mesh hierarchy.
 \end{example}
 
-Solvers implemented in Firedrake exploit parallel solvers from the PETSc library \cite{Balayetal2023}.
-In fact, all of the above Examples, which were run in serial so as to fix the smoother,\footnote{The precise action of $O(m_j)$ work parallel smoothers depends on the number of processes \cite[for example]{Bueler2021}.} also run efficiently in parallel.
-The next Example demonstrates the parallel weak scaling \cite{Bueler2021} of FASCD F-cycle solvers.
+% see git@bitbucket.org:pefarrell/fascd.git: examples/pollutant.py and edplay/results/pollutant.sh|txt
+\begin{table}[ht]
+\begin{tabular}{lr@{\hskip 7mm}c@{\hskip 4mm}c}
+\emph{levels} & $m_J$ & F ($d=2$) & F ($d=3$) \\ \cmidrule{1-4}
+ 1 &    $16^d$ & 1 & 1 \\
+ 2 &    $31^d$ & 2 & 2 \\
+ 3 &    $61^d$ & 2 & 3 \\
+ 4 &   $121^d$ & 2 & 3 \\
+ 5 &   $241^d$ & 2 & 4 \\
+ 6 &   $481^d$ & 2 \\
+ 7 &   $961^d$ & 3 \\
+ 8 &  $1921^d$ & 3 \\
+ 9 &  $3841^d$ & 3
+\end{tabular}
+\bigskip
+\caption{F-cycle iterations for 2D (serial) and 3D (16 processes) solutions of the advection-diffusion bi-obstacle VI problem in Example \ref{ex:results:advdiff}.}
+\label{tab:results:advdiff}
+\end{table}
 

### 2023-06-26 paper/fascd.tex
clean SIA example; text moved to discussion from intro

@@
 Recall the steady shallow ice approximation (SIA) obstacle problem, Example \ref{ex:sia}, in which the surface elevation $s$ is constrained to exceed the bedrock elevation $b$.
-For a test case we set $\Omega=(0,L)^2$, with $L=1800$ km, a radial surface mass balance $a$ function \cite[equation (5.122)]{GreveBlatter2009}, and bumpy, synthetic values for $b\in C^1(\Omega)$, with an elevation range of $1800$ m.
-The numerical solution, colored by the ice flow, is shown in Figure \ref{fig:results:siascene}; this used the surface value of the SIA solution, namely $\mathbf{u} = (n+1)(n+2)^{-1} \Gamma (s-b)^{n+1} |\grad s|^{n-1} \grad s$.
-As noted in Example \ref{ex:sia} and seen in this Figure, $|\grad s|$ is unbounded at the free boundary, so the solution has low regularity.
-
-\vspace{-2mm}
+For a test case we set $\Omega=(0,L)^2$ with $L=1800$ km.
+The data consists of a radial surface mass balance $a$ function \cite[equation (5.122)]{GreveBlatter2009} and a bumpy, synthetic map for $b\in C^1(\Omega)$, with an elevation range of $1800$ m.
+The numerical solution, colored by the solution surface flow speed, i.e.~the magnitude of $\mathbf{u} = (n+1)(n+2)^{-1} \Gamma (s-b)^{n+1} |\grad s|^{n-1} \grad s$, is shown in Figure \ref{fig:results:siascene}.
+As noted in Example \ref{ex:sia}, and clearly seen in this Figure, $|\grad s|$ is unbounded at the free boundary; the solution, while continuous, has very low regularity.
+
 % to regenerate:
@@
 \end{center}
-\caption{A high-resolution ($m^J\approx 10^6$) solution of Example \ref{ex:sia} over a bumpy bedrock topography, with the ice speed as the surface coloring.}
+\caption{Solution of Example \ref{ex:sia} over a bumpy bedrock topography, with the ice speed as the surface coloring, on a $P_1$ mesh of $m^J=1025^2 \approx 10^6$ nodes.}
 \label{fig:results:siascene}
@@
 We applied FASCD F-cycles using a smoother of four (fixed) iterations of RS Newton (Section \ref{sec:implementation}), with quadratic backtracking line search.
-The step equations were approximately solved, in parallel, by three (fixed) iterations of additive-Schwarz preconditioned GMRES, with ILU preconditioning on the blocks.
-The coarse solver was again RS Newton, with LU preconditioning redundantly over all processors.
-The coarse mesh was relatively fine, with $m_0=289$ nodes and 512 triangles.
+The step equations were approximately solved, in parallel, by three (fixed) iterations of preconditioned GMRES, again with a ASM+ILU preconditioner.
+The coarse solver was RS Newton with LU applied redundantly over all processors.
+The coarse mesh was relatively fine, with $m_0=17^2$ nodes and 512 triangles.
 Because of the low regularity of the solution we used \id{rtol} $= 2 \times 10^{-4}$ and \id{atol} $= 10^{-8}$ in \eqref{eq:stoppingcriterion}.
@@
 
-These solver parameters generate a weak-scaling solver.
-For $P$ process solutions we fixed the number of degrees of freedom per process at $m_J/P=1025^2$ and solved over $P=1,4,16$ processes.\footnote{Runs were on a single-node workstation with two Intel Xeon 4210 cpus (2.2GHz), a total of 20 physical cores, and 128 GB shared memory.}  The results in Table \ref{tab:results:siaweak} showing mesh independence in the number of FASCD cycles and clear weak scaling.
-The finest mesh result, with $m_J=4097^2=17 \times 10^6$ degrees of freedom, needed under six minutes to solve for the surface elevation of an ice sheet slightly larger than the current Greenland ice sheet ($1.7$ million $\text{km}^2$) at uniform $439$ m resolution.
+These solver parameters generate a solver with mesh-independent iterations.
+We fixed the number of degrees of freedom per process at $m_J/P=1025^2$ and solved over $P=1,4,16$ processes.\footnote{Runs were on a single-node workstation with two Intel Xeon 4210 cpus (2.2GHz), a total of 20 physical cores, and 128 GB shared memory.}  The results in Table \ref{tab:results:siaweak} show clear weak scaling.
+The finest mesh result, with $m_J=4097^2=17 \times 10^6$ degrees of freedom, needed less than six minutes to solve for the surface elevation of an ice sheet slightly larger than the current Greenland ice sheet ($1.7$ million $\text{km}^2$) at uniform $439$ m resolution.
 \end{example}
@@
 
+FIXME For the classical obstacle problem, based on the linear, symmetric Laplacian operator, certain multilevel techniques, such as the truncated monotone multigrid method \cite{GraeserKornhuber2009,Kornhuber1994}, are known to improve performance relative to the original multilevel CD method \cite{Tai2003}.
+These improved methods modify the nodal basis functions.
+By contrast FASCD makes essentially no assumptions about the FE discretization.
+Acceleration via basis-level manipulations is a topic for future research.
+
 \bibliography{fascd}

### 2023-06-26 paper/fascd.tex
draft F-cycle pseudocode and figure

@@
 \begin{abstract}
-The \emph{full approximation storage constraint decomposition} (FASCD) multilevel finite element method introduced here is a common extension of the full approximation storage multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and of the constraint decomposition method introduced by X.-C.~Tai for variational inequalities.
-We extend the existing constraint decomposition ideas by recognizing and exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
+The \emph{full approximation storage constraint decomposition} (FASCD) multilevel finite element method introduced here is a common extension of the full approximation storage multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and of the constraint decomposition method introduced by X.-C.~Tai for the variational inequalities associated to inequality-constrained optimization.
+We extend the constraint decomposition idea by recognizing and exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
 Then, by using strong smoothers with work proportional to the number of unknowns on the finite element mesh level, we demonstrate that FASCD V-cycles show superior mesh-independent convergence rates and that FASCD F-cycles (full multigrid) are optimal solvers.
-The example problems include unilateral and bi-obstacle variational inequality problems with linear and nonlinear differential operators.
+The several example problems include unilateral and bi-obstacle variational inequality problems with symmetric linear, nonsymmetric linear, and nonlinear differential operators.
 \end{abstract}
@@
 \begin{center}
-\input{tikz/fascdvcycle.tex}
+\input{tikz/vcycle.tex}
 \end{center}
@@
 
-We will regard the smoothers for VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} as approximate, in-place procedures, acting on the corrections $y^j$ and $z^j$, respectively (Section \ref{sec:implementation}).
+We will regard the smoothers for VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} as approximate solvers, acting on the corrections $y^j$ and $z^j$, respectively (Section \ref{sec:implementation}).
 Also, as is standard in multigrid, at the coarsest $j=0$ level we suppose that problem \eqref{eq:fe:upvi} is solved accurately.
 
-These ideas come together in Algorithm \ref{alg:fascd}, \pr{fascd-vcycle}, which acts in-place on $w^J$.
+These ideas come together in Algorithm \ref{alg:fascd}, \pr{fascd-vcycle}, which updates $w^J$.
 It assumes certain semantics for the \pr{smooth} and the coarsest-level \pr{solve} procedures; they do in-place modifications of their final arguments, without modifying their first four arguments.
@@
 \begin{pseudo}
-\pr{fascd-vcycle}(\ell^J,\underline{\gamma}^J,\overline{\gamma}^J;w^J)\text{:} \\+
+\pr{fascd-vcycle}(J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J;w^J)\text{:} \\+
     $\underline{\chi}^J = \underline{\gamma}^J - w^J {\large \strut}$ \\
@@
       $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,\overline{\chi}^j,w^j;z^j)$  \ct{smoothing of \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
-    $w^J \gets w^J+z^J$
+    return $w^J+z^J$
 \end{pseudo}
@@
 
-FIXME F-cycles
+FIXME a perfectly reasonable algorithm is to iterate V-cycles until \eqref{eq:stoppingcriterion} (see Section \ref{sec:results}), but F-cycles are even better; idea is to ``ramp up'' to an initial iterate on the finest level by solving coarse mesh problems, and only then doing full-depth V-cycles
+
+FIXME while the direct application of injection to the original fine-level constraints $\underline{\gamma}^J, \overline{\gamma}^J$ is not used in generating the LDCs for V-cycles (Section \ref{sec:cdmultilevel}), it is the right idea for the ramp in an F-cycle (as long as it is followed by truncation into the next-finest admissible set); here $\mathcal{K}^j=\{v \in \mathcal{V}^j \,:\, \underline{\gamma}^j \le v \le \overline{\gamma}^j\}$; see Algorithm \ref{alg:fascd-fcycle}; Figure \ref{fig:fcycle} illustrates the F-cycle.
 
@@
 \begin{pseudo}
-\pr{fascd-fcycle} \\
-FIXME
+\pr{fascd-fcycle}(J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J)\text{:} \\+
+    for $j=J$ downto $j=1$ \\+
+        $\ell^{j-1}, \underline{\gamma}^{j-1}, \overline{\gamma}^{j-1} = R(\ell^j), \iR \underline{\gamma}^{j}, \iR \overline{\gamma}^{j}$ {\large\strut} \\-
+    choose $w^0 \in \mathcal{K}^0$ \\
+    $\text{\pr{solve}}(\ell^0,\underline{\gamma}^0,\overline{\gamma}^0;w^0)$ \\
+    for $j=1$ to $j=J-1$ \\+
+        $w^j = \max\left\{\underline{\gamma}^{j},\min\{\overline{\gamma}^{j}, Pw^{j-1}\}\right\} \in \mathcal{K}^j$ \\
+        $\pr{fascd-vcycle}^{\text{\id{rampv}}}(j,\ell^j,\underline{\gamma}^j,\overline{\gamma}^j;w^j)$ \\-
+    $w^J = \max\left\{\underline{\gamma}^{J},\min\{\overline{\gamma}^{J}, Pw^{J-1}\}\right\} \in \mathcal{K}^J$ \\
+    while not \eqref{eq:stoppingcriterion} \\+
+        \pr{fascd-vcycle}(J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J;w^J) \\-
+    return $w^J$
 \end{pseudo}
@@
 \end{pseudofloat}
+
+\begin{figure}[ht]
+\begin{center}
+\input{tikz/fcycle.tex}
+\end{center}
+\caption{Algorithm \ref{alg:fascd-fcycle} initially restricts the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray), then prolongs and truncates (doubled; see lines 7 and 9 in Alg.~\ref{alg:fascd-fcycle}) to create the next-finer initial iterate.
+See Figure \ref{fig:fascdvcycle} for other symbols.}
+\label{fig:fcycle}
+\end{figure}
+
+FIXME though not shown in Algorithm \ref{alg:fascd-fcycle}, one can generate an admissible initial iterate for the first coarse solve in the F-cycle by injection and truncation down to coarsest level
 

### 2023-06-27 paper/fascd.tex
decide on box constraint assignment style, and use line refs

@@
 \pr{fascd-vcycle}(J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J;w^J)\text{:} \\+
-    $\underline{\chi}^J = \underline{\gamma}^J - w^J {\large \strut}$ \\
-    $\overline{\chi}^J = \overline{\gamma}^J - w^J {\large \strut}$ \\
+    $\underline{\chi}^J, \, \overline{\chi}^J = \underline{\gamma}^J - w^J, \, \overline{\gamma}^J - w^J {\large \strut}$ \label{line:vcyclegenchifinest} \\
     for $j=J$ downto $j=1$ \\+
-      $\underline{\chi}^{j-1} = \maxR \underline{\chi}^j {\large \strut}$ \\
-      $\overline{\chi}^{j-1} = \minR \overline{\chi}^j {\large \strut}$ \\
-      $\underline{\phi}^j = \underline{\chi}^j - P\underline{\chi}^{j-1} {\large \strut}$ \\
-      $\overline{\phi}^j = \overline{\chi}^j - P\overline{\chi}^{j-1} {\large \strut}$ \\
+      $\underline{\chi}^{j-1}, \, \overline{\chi}^{j-1} = \maxR \underline{\chi}^j, \, \minR \overline{\chi}^j {\large \strut}$ \label{line:vcyclegenchi} \\
+      $\underline{\phi}^j, \, \overline{\phi}^j = \underline{\chi}^j - P\underline{\chi}^{j-1}, \, \overline{\chi}^j - P\overline{\chi}^{j-1} {\large \strut}$ \label{line:vcyclegenphi} \\
       $y^j = 0$ \\
       $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\underline{\phi}^j,\overline{\phi}^j,w^j;y^j)$  \ct{smoothing of \eqref{eq:fe:downvi} in $\mathcal{D}^j$}\\
-      $w^{j-1} = \iR(w^j + y^j)$ \\
-      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j+y^j)\right)$ \\-
+      $w^{j-1} = \iR(w^j + y^j)$ \label{line:vcyclerestrictsolution} \\
+      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j+y^j)\right)$ \label{line:vcyclerestrictell} \\-
     $z^0 = 0$ \\
@@
     for $j=J$ downto $j=1$ \\+
-        $\ell^{j-1}, \underline{\gamma}^{j-1}, \overline{\gamma}^{j-1} = R(\ell^j), \iR \underline{\gamma}^{j}, \iR \overline{\gamma}^{j}$ {\large\strut} \\-
+        $\ell^{j-1} = R(\ell^j)$ \\
+        $\underline{\gamma}^{j-1}, \, \overline{\gamma}^{j-1} = \iR \underline{\gamma}^{j}, \, \iR \overline{\gamma}^{j}$ {\large\strut} \\-
     choose $w^0 \in \mathcal{K}^0$ \\
@@
     for $j=1$ to $j=J-1$ \\+
-        $w^j = \max\left\{\underline{\gamma}^{j},\min\{\overline{\gamma}^{j}, Pw^{j-1}\}\right\} \in \mathcal{K}^j$ \\
+        $w^j = \max\left\{\underline{\gamma}^{j},\min\{\overline{\gamma}^{j}, Pw^{j-1}\}\right\} \in \mathcal{K}^j$ \label{line:fcycleprolongtruncone} \\
         $\pr{fascd-vcycle}^{\text{\id{rampv}}}(j,\ell^j,\underline{\gamma}^j,\overline{\gamma}^j;w^j)$ \\-
-    $w^J = \max\left\{\underline{\gamma}^{J},\min\{\overline{\gamma}^{J}, Pw^{J-1}\}\right\} \in \mathcal{K}^J$ \\
+    $w^J = \max\left\{\underline{\gamma}^{J},\min\{\overline{\gamma}^{J}, Pw^{J-1}\}\right\} \in \mathcal{K}^J$ \label{line:fcycleprolongtrunctwo} \\
     while not \eqref{eq:stoppingcriterion} \\+
@@
 \end{center}
-\caption{Algorithm \ref{alg:fascd-fcycle} initially restricts the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray), then prolongs and truncates (doubled; see lines 7 and 9 in Alg.~\ref{alg:fascd-fcycle}) to create the next-finer initial iterate.
+\caption{Algorithm \ref{alg:fascd-fcycle} initially restricts the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray), then prolongs and truncates (doubled; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Alg.~\ref{alg:fascd-fcycle}) to create the next-finer initial iterate.
 See Figure \ref{fig:fascdvcycle} for other symbols.}
@@
 the FE discretization of strong-form PDE $f(u)=\ell$.
-The downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets can also be enlarged so that the corrections $y^j$ and $z^j$ in \pr{fascd-vcycle} are never constrained.\footnote{Formally we might do this by choosing constant LDCs with $\underline{\chi}^j \ll 0 \ll \overline{\chi}^j$, but so that ordering property \eqref{eq:fe:chiordering} still holds, and also with large gaps so that $\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1} \ll 0 \ll \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}$.}  Thus the corrections $y^j$ and $z^j$ are unconstrained, and lines 2--3 and 5--8 can be removed from Algorithm \ref{alg:fascd}.
+The downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets can also be enlarged so that the corrections $y^j$ and $z^j$ in \pr{fascd-vcycle} are never constrained.\footnote{Formally we might do this by choosing constant LDCs with $\underline{\chi}^j \ll 0 \ll \overline{\chi}^j$, but so that ordering property \eqref{eq:fe:chiordering} still holds, and also with large gaps so that $\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1} \ll 0 \ll \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}$.}  Thus the corrections $y^j$ and $z^j$ are unconstrained, and lines \ref{line:vcyclegenchifinest}, \ref{line:vcyclegenchi}, and \ref{line:vcyclegenphi} can be removed from Algorithm \ref{alg:fascd}.
 
@@
 Then we combine $u^j=w^j+y^j$, instead of $w^j$ and $y^j$ separately, and similarly $u^j=w^j+z^j$ when up-smoothing.
-The smoothers and the coarsest-level solver can be regarded as unconstrained, in-place modifications of $u^j$, so we denote them simply as ``$\text{\pr{smooth}}(\ell^j;u^j)$.''  Noting line 11 of Algorithm \ref{alg:fascd} now says $u^{j-1}=\iR u^j$, after down-smoothing the coarsened source functional computed by line 12 can be rewritten as
+The smoothers and the coarsest-level solver can be regarded as unconstrained, in-place modifications of $u^j$, so we denote them simply as ``$\text{\pr{smooth}}(\ell^j;u^j)$.''  Noting line \ref{line:vcyclerestrictsolution} of Algorithm \ref{alg:fascd} now says $u^{j-1}=\iR u^j$, after down-smoothing the coarsened source functional computed by line \ref{line:vcyclerestrictell} can be rewritten as
 \begin{equation}

### 2023-06-28 paper/fascd.tex
cleanup in section 7

@@
 
-\section{Implementation and smoothers} \label{sec:implementation}
+\section{Implementation} \label{sec:implementation}
 
 In this section we consider several practical concerns when implementing and applying \pr{fascd-vcycle} Algorithm \ref{alg:fascd}.
-We address convergence criteria, smoothers, and coarse solvers, plus the F-cycle extension of the Algorithm, sometimes called a \emph{full multigrid cycle}.
+We address convergence criteria and the F-cycle extension of the Algorithm.
+We also describe an active set Newton method used in Section \ref{sec:results} as a smoother and coarse solver.
 
 Because the problem is inequality-constrained, repeated application of \pr{fascd-vcycle} to solve the finite-dimensional, box-constrained VI \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.
-However, \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}, a strong form amenable to convergence criteria.
+Indeed \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}, a strong form amenable to constructing a convergence criteria for the discretization.
 Let $x_p \in \mathcal{T}^J$ be a finest-level node, and $\psi_p(x)$ the corresponding hat function.
@@
 If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let $I_D$ be the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.
-Finally, we will write $a\perp b$ when $a,b \in \RR^n$ are \emph{complementary}, that is, when $a_i \ge 0$ and $b_i \ge 0$ and $a_i b_i = 0$ for $i=0,\dots,n-1$.
+Finally, write $a\perp b$ when $a,b \in \RR^n$ are \emph{complementary}, that is, when $a_i \ge 0$ and $b_i \ge 0$ and $a_i b_i = 0$ for $i=0,\dots,n-1$.
 Then problem \eqref{eq:fe:vi}, which would be expressed VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap \mathcal{X}^J\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP for the solution $u^J \in \mathcal{K}^J$, but with additional equality constraints:
@@
 for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
+
 By definition, MCP \eqref{eq:fe:mcp} says that certain node-wise conditions hold for $u^J$ at $x_p \in \mathcal{T}^J$.
+Let $r(u^J)_p = \ip{f^J(w^J)-\ell^J}{\psi_p}$ be the pointwise (ordinary) residual.
 If $x_p \notin \partial_D\Omega$ then one of these four conditions hold:
 \begin{itemize}
-\item $\underline{\gamma}^J(x_p)<u^J(x_p)<\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} = 0$,
-\item $\underline{\gamma}^J(x_p)=u^J(x_p)<\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \ge 0$,
-\item $\underline{\gamma}^J(x_p)<u^J(x_p)=\overline{\gamma}^J(x_p)$ and $\ip{f^J(u^J)-\ell^J}{\psi_p} \le 0$,
+\item $\underline{\gamma}^J(x_p)<u^J(x_p)<\overline{\gamma}^J(x_p)$ and $r(u^J)_p = 0$,
+\item $\underline{\gamma}^J(x_p)=u^J(x_p)<\overline{\gamma}^J(x_p)$ and $r(u^J)_p \ge 0$,
+\item $\underline{\gamma}^J(x_p)<u^J(x_p)=\overline{\gamma}^J(x_p)$ and $r(u^J)_p \le 0$,
 \item $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$.
@@
 \begin{itemize}
-\item $u^J(x_p)=g_D^J(x_p)$ and $\lambda_p=\ip{f^J(u^J)-\ell^J}{\psi_p}$.
+\item $u^J(x_p)=g_D^J(x_p)$ and $\lambda_p=r(u^J)_p$.
 \end{itemize}
 The cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, \emph{pinched} (\emph{both active}), and \emph{boundary}, respectively.
-Where \emph{pinched} applies, $\ip{f^J(u^J)-\ell^J}{\psi_p}$ can have any real value.
-
-Thus the MCP \eqref{eq:fe:mcp} identifies expressions whose norms should go below tolerance.
-For for $w^J \in \mathcal{K}^J$, let $r_p = \ip{f^J(w^J)-\ell^J}{\psi_p}$, the pointwise (ordinary) residual, as a notational convenience.
-First, interpreting the MCP directly, we define the finest-level \emph{nodal complementarity residual} $\rNC$ in $\RR^{m_J}$ from the five cases:
+Where \emph{pinched} applies, $r(u^J)_p$ can have any real value.
+
+Thus \eqref{eq:fe:mcp} allows us to construct expressions whose norms should go below tolerance.
+First, interpreting the MCP directly, for $w^J \in \mathcal{K}^J$ we define the finest-level \emph{nodal complementarity residual} $\rNC(w^J)$ from the five cases:
 \begin{equation}
 \rNC(w^J)_p = \begin{cases}
-    r_p, & \underline{\gamma}^J(x_p) < w^J(x_p) < \overline{\gamma}^J(x_p), \\
-    \min\{r_p,0\}, & w^J(x_p) = \underline{\gamma}^J(x_p), \\
-    \max\{r_p,0\}, & w^J(x_p) = \overline{\gamma}^J(x_p), \\
+    r(w^J)_p, & \underline{\gamma}^J(x_p) < w^J(x_p) < \overline{\gamma}^J(x_p), \\
+    \min\{r(w^J)_p,0\}, & w^J(x_p) = \underline{\gamma}^J(x_p), \\
+    \max\{r(w^J)_p,0\}, & w^J(x_p) = \overline{\gamma}^J(x_p), \\
     0, & \underline{\gamma}^J(x_p)=w^J(x_p)=\overline{\gamma}^J(x_p), \\
@@
 \end{equation}
-(The first four cases assume $x_p \notin \partial_D\Omega$.)  Note that $\rNC : \mathcal{K}^J \to \RR^{m_J}$ is a computable nonlinear map defined from $f^J$ and the vectors $\ell^J$, $\underline{\gamma}^J$, and $\overline{\gamma}^J$, and that for PDEs without constraints one uses only the first and last cases.
-In summary, $\rNC(w^J)_p$ is nonzero at nodes where MCP \eqref{eq:fe:mcp} is violated, either because the pointwise residual $\ip{f^J(w^J)-\ell^J}{\psi_p}$ remains nonzero where the constraints are inactive, or because it has the wrong sign at an actively-constrained degree of freedom, or because the boundary condition is not satisfied.
-
-As a practical matter, however, direct implementation of \eqref{eq:rNC} is sensitive to rounding error because $\rNC$ is a discontinuous function of $w^J(x_p)$.
-For example, if $w^J(x_p)$ exceeds $\underline{\gamma}^J(x_p)$ by $10^{-16}$, then, instead of returning the minimum of $z_p = \left<f^J(w^J)-\ell^J,\psi_p\right>$ and zero (the second case), the first case applies and $\rNC(w^J)_p = z_p$, which might be large (positive), and this would suggest falsely that the iteration is far from converging.
-On the other hand, an advantage of formula \eqref{eq:rNC} is that it applies as stated if $\underline{\gamma}^J(x_p)=-\infty$ or $\overline{\gamma}^J(x_p)=+\infty$; compare \eqref{eq:rSS} below, wherein the difference $w^J(x_p) - \underline{\gamma}^J(x_p)$ is computed, for example.
-
-%\newcommand{\phiFB}{\phi_{\text{FB}}}
-We will instead measure convergence using a \emph{semi-smooth residual} based on the Fischer-Burmeister function \cite{BensonMunson2006,Ulbrich2011}, a function defined as follows for $a,b\in\RR$:
+The first four cases assume $x_p \notin \partial_D\Omega$.  For PDEs one uses only the first and last cases.
+
+In summary, $\rNC : \mathcal{K}^J \to \RR^{m_J}$ is a computable nonlinear map defined using the discretization $f^J$ and the vectors $\ell^J$, $\underline{\gamma}^J$, and $\overline{\gamma}^J$.
+A value $\rNC(w^J)_p$ is nonzero only at nodes where \eqref{eq:fe:mcp} is violated, either because $r(w^J)_p$ remains nonzero where the constraints are inactive, or because it has the wrong sign at an active node, or because the boundary condition is not satisfied.
+
+Direct implementation of \eqref{eq:rNC} is, however, sensitive to rounding error because $\rNC$ is a discontinuous function of $w^J(x_p)$.
+For example, if $w^J(x_p)$ exceeds $\underline{\gamma}^J(x_p)$ by $10^{-16}$, then, instead of returning the minimum of $r(w^J)_p$ and zero (the second case), the first case applies, so $\rNC(w^J)_p$ might be large and positive.
+This situation would falsely suggest that the iteration is far from converging.
+On the other hand, formula \eqref{eq:rNC} applies as stated if $\underline{\gamma}^J(x_p)=-\infty$ or $\overline{\gamma}^J(x_p)=+\infty$.
+Compare \eqref{eq:rSS} in this case, wherein the difference $w^J(x_p) - \underline{\gamma}^J(x_p)$ is computed.
+
+We will instead measure convergence using a \emph{semi-smooth residual} based on the Fischer-Burmeister function \cite{BensonMunson2006,Ulbrich2011} on $a,b\in\RR$:
 \begin{equation}
@@
 \end{equation}
-Observe that $\phi(a,b)=0$ if and only if $a,b$ are complementary, and also that $\phi$ semi-smooth (and thus continuous).
-For convenience denote $\underline{g}_p = w^J(x_p) - \underline{\gamma}^J(x_p)$ and $\overline{g}_p = \overline{\gamma}^J(x_p) - w^J(x_p)$, the positive gaps at $x_p$.
-Finally define
+Observe that $\phi$ is semi-smooth, thus continuous, and that $\phi(a,b)=0$ if and only if $a,b$ are complementary.
+For convenience denote $\underline{g}_p = w^J(x_p) - \underline{\gamma}^J(x_p)$ and $\overline{g}_p = \overline{\gamma}^J(x_p) - w^J(x_p)$, the nonnegative gaps at $x_p$.
+Now define\footnote{This semismooth residual is straightforwardly modified for unconstrained degrees of freedom, that is, where $\overline{\gamma}^J(x_p) = +\infty$ or $\underline{\gamma}^J(x_p) = -\infty$ or both.}
 \begin{equation}
 \rSS(w^J)_p = \begin{cases}
-r_p, & \underline{\gamma}^J(x_p) = -\infty \text{ and } \overline{\gamma}^J(x_p) = +\infty, \\
-\phi\big(\underline{g}_p, r_p\big), & \overline{\gamma}^J(x_p) = +\infty, \\
-\phi\left(\overline{g}_p, -r_p\right), & \underline{\gamma}^J(x_p) = -\infty, \\
-\max\big\{\phi\big(\underline{g}_p, r_p\big), \phi\left(\overline{g}_p, -r_p\right)\big\}, & \text{otherwise}, \\
-w^J(x_p) - g_D^J(x_p), & x_p \in \partial_D\Omega,
+\max\big\{\phi\big(\underline{g}_p, r(w^J)_p\big), \phi\left(\overline{g}_p, -r(w^J)_p\right)\big\}, & x_p \notin \partial_D\Omega, \\
+w^J(x_p) - g_D^J(x_p), & x_p \in \partial_D\Omega.
 \end{cases} \label{eq:rSS}
 \end{equation}
-where again $x_p \notin \partial_D\Omega$ in the first four cases.
-
-Now suppose we repeatedly apply \pr{fascd-vcycle}, starting with $w^{J,0} \in \mathcal{K}^J$, generating (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.
+Suppose an iteration starts with $w^{J,0} \in \mathcal{K}^J$, generating (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.
 Let $\id{atol}\ge 0$, $\id{rtol}\ge 0$ be given tolerances.
-Then the criterion used in the results (Section \ref{sec:results}) is
+Then the criterion used in Section \ref{sec:results} for computations is\footnote{A step tolerance $\|w^{J,k} - w^{J,k-1}\|_2 < \id{stol}$ or similar could also be used, but we did not find it necessary.}
 \begin{equation}
@@
 \end{equation}
-A step tolerance, based on $\|w^{J,k} - w^{J,k-1}\|_2$, could also be used, but we did not find it necessary.
-
-FIXME smoother semantics; shift ``$w^j+y^j$'' etc.
-
-FIXME smoothers use Newton reduced space (RS) \cite{BensonMunson2006}with either direct linear solve or a fixed number of Krylov steps; in latter case either fixed number of CG+ICC for symmetric linear operator (Laplacian), otherwise fixed number of GMRES+ILU
-
-% FIXME try Newton semi-smooth \cite{BensonMunson2006}?
-
-FIXME a perfectly reasonable algorithm is to iterate V-cycles until \eqref{eq:stoppingcriterion} (see Section \ref{sec:results}), but F-cycles are even better; idea is to ``ramp up'' to an initial iterate on the finest level by solving coarse mesh problems, and only then doing full-depth V-cycles
-
-FIXME while the direct application of injection to the original fine-level constraints $\underline{\gamma}^J, \overline{\gamma}^J$ is not used in generating the LDCs for V-cycles (Section \ref{sec:cdmultilevel}), it is the right idea for the ramp in an F-cycle (as long as it is followed by truncation into the next-finest admissible set); here $\mathcal{K}^j=\{v \in \mathcal{V}^j \,:\, \underline{\gamma}^j \le v \le \overline{\gamma}^j\}$; see Algorithm \ref{alg:fascd-fcycle}; Figure \ref{fig:fcycle} illustrates the F-cycle.
-
-\begin{pseudofloat}[ht]
+
+Straightforward iteration of V-cycle Algorithm \ref{alg:fascd}, starting with some finest-mesh initial iterate $w^{J,0}$ and repeated until \eqref{eq:stoppingcriterion} holds, is demonstrated as a solver in Section \ref{sec:results}.
+However, application of an FASCD F-cycle, also called a \emph{full multigrid cycle}, is seen to be even better.
+The well-known principle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate.
+Then an admissible initial iterate on the next-finer mesh is generated via prolongation and truncation, and so on.
+During this ``ramp'' up one approximately solves on each level by using a fixed number (\id{rampv}) of V-cycles, each down to the coarsest mesh.
+From the finest level one does full-depth V-cycles to convergence.
+
+This F-cycle approach is in Figure \ref{fig:fcycle} and Algorithm \ref{alg:fascd-fcycle}.
+Observe that injection of the original fine-level constraints $\underline{\gamma}^J, \overline{\gamma}^J$ is \emph{not} used in generating the LDCs for V-cycles; compare Section \ref{sec:cdmultilevel}.
+However, injection is appropriate for the ramp in an F-cycle, if followed by truncation into $\mathcal{K}^j=\{v \in \mathcal{V}^j \,:\, \underline{\gamma}^j \le v \le \overline{\gamma}^j, \, v|_{\partial\Omega} = g_D\}$; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fcycle}.
+Though not included in Algorithm \ref{alg:fascd-fcycle}, one may generate the coarsest-level initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation down to coarsest level; compare line \ref{line:fcyclecoarsestinitial}.
+
+\begin{figure}[h]
+\begin{center}
+\input{tikz/fcycle.tex}
+\end{center}
+\caption{Algorithm \ref{alg:fascd-fcycle} initially restricts the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray edges).
+Then it prolongs and truncates (doubled edges) to create an admissible initial iterate on the next-finer level.
+Solution on each level is by V-cycles; \id{rampv} $=1$ here.
+See Figure \ref{fig:fascdvcycle} for the other symbols.}
+\label{fig:fcycle}
+\end{figure}
+
+\begin{pseudofloat}[h]
 \begin{pseudo}
@@
         $\underline{\gamma}^{j-1}, \, \overline{\gamma}^{j-1} = \iR \underline{\gamma}^{j}, \, \iR \overline{\gamma}^{j}$ {\large\strut} \\-
-    choose $w^0 \in \mathcal{K}^0$ \\
+    choose $w^0 \in \mathcal{K}^0$ \label{line:fcyclecoarsestinitial} \\
     $\text{\pr{solve}}(\ell^0,\underline{\gamma}^0,\overline{\gamma}^0;w^0)$ \\
@@
 
-\begin{figure}[ht]
-\begin{center}
-\input{tikz/fcycle.tex}
-\end{center}
-\caption{Algorithm \ref{alg:fascd-fcycle} initially restricts the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray), then prolongs and truncates (doubled; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Alg.~\ref{alg:fascd-fcycle}) to create the next-finer initial iterate.
-See Figure \ref{fig:fascdvcycle} for other symbols.}
-\label{fig:fcycle}
-\end{figure}
-
-FIXME though not shown in Algorithm \ref{alg:fascd-fcycle}, one can generate an admissible initial iterate for the first coarse solve in the F-cycle by injection and truncation down to coarsest level
+FIXME smoother semantics; shift ``$w^j+y^j$'' etc.
+
+FIXME smoothers use Newton reduced space (RS) \cite{BensonMunson2006}with either direct linear solve or a fixed number of Krylov steps; in latter case either fixed number of CG+ICC for symmetric linear operator (Laplacian), otherwise fixed number of GMRES+ILU
+
+% FIXME try Newton semi-smooth \cite{BensonMunson2006}?
 

### 2023-06-28 paper/fascd.tex
section 7 fully drafted

@@
 In Section \ref{sec:vi} we recall the theory of coercive and monotone variational inequalities, with several example problems.
-Then Section \ref{sec:cd} reviews the basics of CDs, including the associated iterations, as proposed by Tai \cite{Tai2003}.
-Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, along with special monotone restriction operators \cite{GraeserKornhuber2009}.
-Section \ref{sec:cdmultilevel} presents defect constraints \cite{GraeserKornhuber2009}, leading to an apparently new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD.
-Section \ref{sec:vcycle} then states the FASCD V-cycle, with Section \ref{sec:implementation} addressing implementation concerns, smoother semantics, and F-cycles.
-Section \ref{sec:results} gives FASCD numerical results, for problems in Section \ref{sec:vi}, from a parallel Firedrake \cite{Rathgeberetal2016} implementation.
-Discussion and outlook in Section \ref{sec:discussion} and an Appendix, which sketchs how FASCD reduces to classical FAS multigrid for PDEs, finish the work.
+Section \ref{sec:cd} reviews the basics of CDs, as proposed by Tai \cite{Tai2003}, including the associated iterations.
+Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, and then Section \ref{sec:cdmultilevel} presents defect constraints generated via monotone restriction operators \cite{GraeserKornhuber2009}.
+Here we propose an apparently new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD.
+Section \ref{sec:vcycle} then states the FASCD V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, F-cycles, and our smoother choice.
+Section \ref{sec:results} gives FASCD numerical results from a parallel Firedrake \cite{Rathgeberetal2016} implementation for the problems introduced in Section \ref{sec:vi}.
+Discussion and outlook in Section \ref{sec:discussion}, and an Appendix on how FASCD reduces to classical FAS multigrid for PDEs, finish the work.
 
@@
 In this section we consider several practical concerns when implementing and applying \pr{fascd-vcycle} Algorithm \ref{alg:fascd}.
-We address convergence criteria and the F-cycle extension of the Algorithm.
-We also describe an active set Newton method used in Section \ref{sec:results} as a smoother and coarse solver.
+We address convergence criteria, the F-cycle extension, and an active set Newton method used in Section \ref{sec:results} as a smoother and coarse solver.
 
@@
 
-FIXME smoother semantics; shift ``$w^j+y^j$'' etc.
-
-FIXME smoothers use Newton reduced space (RS) \cite{BensonMunson2006}with either direct linear solve or a fixed number of Krylov steps; in latter case either fixed number of CG+ICC for symmetric linear operator (Laplacian), otherwise fixed number of GMRES+ILU
-
-% FIXME try Newton semi-smooth \cite{BensonMunson2006}?
+\begin{comment}
+smoother semantics: in our implementation of the V-cycle we actually shift to solve a VI for $\tilde w^j = w^j+y^j$, then compute $y^j = \tilde w^j - w^j$ after the solve, etc.; this is evidently equivalent
+\end{comment}
+
+While a large space of smoothers could be tested,\footnote{See comments in Section \ref{sec:discussion}.} Section \ref{sec:results} only demonstrates one powerful, $O(m_j)$ work per iteration, smoother choice, namely a fixed number of iterations of the active-set Newton method from \cite{BensonMunson2006}, as implemented in PETSc \cite{Balayetal2023}, with a fixed number of preconditioned Krylov iterations to solve the reduced-space step equations.
+We will denote this truncated active-set Newton method as \emph{RS Newton}.
+The coarse solver is the same, but with LU decomposition solution of the step equations.
+For the symmetric Laplacian in Example \ref{ex:results:classical} we used conjugate gradients (CG) with incomplete Cholesky (IC) preconditioning, but for other operators the generalized minimum residual (GMRES) method with incomplete LU (ILU) preconditioning is used.
 
@@
 
-Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}.
-In this Section we show numerical results for the several VI Examples in Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
-Assuming strong, yet $O(m_J)$ arithmetic, smoothers on each level, as illustrated below, our essential observations are that each V-cycle dramatically reduces the residual norm, and that a couple of F-cycles solves most Examples down to discretization error.
+Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}, with $P_1$ elements and standard, global mesh refinement.
+Here we show numerical results for the several VI Examples in Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
+Our essential observations are that each V-cycle dramatically reduces the residual norm, and that a few F-cycles solves most Examples down to discretization error.
 
@@
 
-For this experiment we used $P_1$ elements, standard uniform mesh refinement, and a coarsest mesh of $m_0=5^2$ nodes (32 triangles).
+For this experiment we used a coarsest mesh of $m_0=5^2$ nodes (32 triangles).
 The finest mesh had $m_9=2049^2 \approx 4.2 \times 10^{6}$ nodes.
-As our smoother we applied a single RS Newton iteration (Section \ref{sec:implementation}), with three (fixed) ICC-preconditioned CG iterations to approximately solve the step equations.
+As our smoother we applied a single RS Newton iteration (Section \ref{sec:implementation}), with three (fixed) IC-preconditioned CG iterations to approximately solve the step equations.
 The coarse solver was RS$+$LU, iterated to convergence.
@@
 
-Table \ref{tab:results:classical} shows the resulting iteration counts.
-When the RS Newton smoother is applied by itself as a solver it shows rapidly-increasing iterations as expected, roughly proportional to resolution.
+Table \ref{tab:results:classical} shows the resulting iteration counts for the various finest meshes of $m_J$ nodes.
+When the RS Newton smoother is applied by itself as a solver (``RS only'') it shows rapidly-increasing iterations as expected, roughly proportional to resolution.
 By contrast, V-cycle iterations increase slowly, perhaps logarithmically with the resolution.
@@
 
-Each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+ICC) iterations.
+Each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+IC) iterations.
 For the F-cycle results, where three iterations suffice at all resolutions in either problem, this is close to textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  However, work of at most ten residual evaluations is often mentioned.} despite the low solution regularity.
@@
 The exact solution of VI \eqref{eq:vi} is easily computed for any $p$; see Figure \ref{fig:results:plap}.
-For this experiment we used $P_1$ elements, a coarsest mesh of $m_0=7$ nodes, and nine standard refinements to a $m_9=3073$ finest mesh.
+For this experiment we used a coarsest mesh of $m_0=7$ nodes and an $m_9=3073$ finest mesh.
 The default smoother uses three (fixed) iterations of RS Newton, with LU solution of the step equations.\footnote{This is an optimal $O(m_j)$ smoother in 1D.}  The coarse solver is the same but iterated to convergence.
@@
 
-The $p=1.5$ and $p=4$ cases give different convergence rates in the $\|\cdot\|_\infty$ norm, namely at rates $O(h^{2.005})$ and $O(h^{1.420})$, respectively.
+The $p=1.5$ and $p=4$ cases give different convergence rates in the $\|\cdot\|_\infty$ norm, namely at rates $O(h^{2.005})$ and $O(h^{1.420})$, respectively; compare \cite{ChoeLewis1991}.
 \end{example}
@@
 
-For the experiment we used $P_1$ elements, uniform refinement, and a coarse mesh with $m_0=16^d$ nodes so as to minimally-resolve the coincidence set.
+For the experiment we used a coarse mesh with $m_0=16^d$ nodes so as to minimally-resolve the coincidence set.
 The smoother was two (fixed) RS Newton iterations,  with three preconditioned GMRES iterations to approximately solve the step equations.
@@
 \begin{tabular}{c@{\hskip 4mm}c@{\hskip 4mm}c@{\hskip 7mm}c@{\hskip 4mm}c}
-P & \emph{levels} & $m_J$ & \emph{F-cycles} & \emph{time} (s) \\ \cmidrule{1-5}
+$P$ & \emph{levels} & $m_J$ & \emph{F-cycles} & \emph{time} (s) \\ \cmidrule{1-5}
  1 & 7 & $1025^2$ & 4 & 311 \\
@@
 \bigskip
-\caption{Iterations and run time (wall clock, in seconds) for parallel FASCD F-cycle solutions of Example \ref{ex:sia}.}
+\caption{Iterations and run time (wall clock, in seconds) for parallel FASCD F-cycle solutions of Example \ref{ex:sia} on $P$ processes.}
 \label{tab:results:siaweak}
@@
 
-\section{Discussion and Conclusion} \label{sec:discussion}
+\section{Discussion and outlook} \label{sec:discussion}
 
@@
 
-FIXME smoother could be AMG
-
-FIXME desired extension to nonlocal variational inequalities
+FIXME smoother could be Newton semi-smooth \cite{BensonMunson2006}; smoother could be interior point method, etc.; smoother could use GMG or AMG on step equations but this is not seen to be necessary in applications so far
+
+FIXME desired extension to nonlocal variational inequalities not tried here
 

### 2023-06-28 paper/fascd.tex
section 5 refinished; no FIXMEs outside of Discussion section at end

@@
 However, one can observe here that the CD concept applies at the level of the continuum problem, as illustrated in the following two examples, the first of which is an overlapping domain decomposition of a Sobolev space.
+Note that $\cK_i \not\subset \cK$ in these cases; informally speaking, the important inclusion is $\cK_i \subset \cV_i$.
 
@@
 Our second example is a disjoint frequency decomposition of a Hilbert space.
-A multilevel CD over a finite-dimensional FE discretization, as proposed in Sections \ref{sec:femultilevel}--\ref{sec:cdmultilevel}, approximates such a decomposition.
+A multilevel CD over a finite-dimensional FE discretization, as proposed in Sections \ref{sec:femultilevel} and \ref{sec:cdmultilevel}, approximates such a decomposition.
 
@@
 
-Note that $\cK_i \not\subset \cK$ in these cases; informally speaking, the important inclusion is $\cK_i \subset \cV_i$.
-
 Associated to a constraint decomposition \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} are certain iterative methods \cite{Tai2003,Xu1992}.
-Let $0\le \alpha \le 1$ be the \emph{damping parameter}.
-Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below, discussed next, solve smaller VI problems over each subset $\cK_i$.
+Let $0\le \alpha \le 1$ be the \emph{damping parameter}.\footnote{FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} are presented without damping, and indeed it seems to be unnecessary.
+However, the general convergence of the basic CD iterations appears to depend upon it.}  Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below, discussed next, solve smaller VI problems over each subset $\cK_i$.
 Each Algorithm replaces $u \in \cK$ with a new iterate $w\in\cK$ which should be closer to $u^* \in \cK$ solving \eqref{eq:vi}.
@@
 In VIs \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} the argument of $f$ is an element of $\cK$, as the reader should confirm.
-
-By \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write the test vector $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors from $\cK$.
+Furthermore, by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write the test vector $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors from $\cK$.
 For \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} respectively:
@@
 In this sense the VIs over the subsets $\mathcal{K}_i$ solve the same problem as \eqref{eq:vi}.
-Furthermore, let $e_i = \hat w_i - \Pi_i u \in \cV_i$.
+
+Let $e_i = \hat w_i - \Pi_i u \in \cV_i$.
 The reader may confirm that $\hat w = u + \sum_{i} e_i$ and thus that $w = u + \alpha \sum_i e_i$.
@@
 
-The basic CD iterations are meaningful whether or not we are in the optimization case $f=F'$, and when $f$ is nonlinear, non-local, or defined only on $\cK$.
+The basic CD iterations are meaningful whether or not we are in the optimization case $f=F'$, and even when $f$ is nonlinear, non-local, or defined only on $\cK$.
 However, practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.
-References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem, and which case $f=F'$ is linear, local, and defined for all inputs (on all of $\mathcal{V}$).
+References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem, wherein $f=F'$ is linear, local, and defined for all inputs (on all of $\mathcal{V}$).
 Substantial implementation simplifications are available in such cases.
@@
 
-When $f$ has the locality of Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}  However, for general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of \pr{cd-add} or \pr{cd-mult}.
+When $f$ has the locality of Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}  However, for general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of CD-type iterations.
 Section \ref{sec:vcycle} constructs a multiplicative multilevel CD iteration for a general operator $f:\cK\to\cV'$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
 
-From now on we restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.
+From now on we will restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.
 Suppose $\Omega \subset \RR^d$ is open and let $\mathcal{V}=W^{1,p}(\Omega)$.
 The \emph{lower obstacle} $\underline{\gamma}$ and the \emph{upper obstacle} $\overline{\gamma}$ are given functions, defined on the closure $\overline{\Omega}$, with values in the extended real line $\tilde\RR = [-\infty,+\infty]$.
-Specifically, we suppose $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$ and $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$ are measurable, $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$, and $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.
-Also, we assume that $\partial\Omega$ is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive measure, to facilitate well-posedness \cite{Evans2010}, and $\partial_N \Omega$ possibly empty.
+We suppose $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$ and $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$ are measurable, $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$, and $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$, and that $\partial\Omega$ is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive measure, to facilitate well-posedness \cite{Evans2010}, and $\partial_N \Omega$ possibly empty.
 Dirichlet data\footnote{Dirichlet boundary values are in the trace sense \cite{Evans2010}, and $g_D$ is assumed to be as regular as needed.} $g_D:\partial_D \Omega \to \RR$ then defines
@@
 \end{equation}
-the closed and convex constraint set.  Note that we are treating inequality and equality (boundary) constraints in a unified manner.
-
-Recalling abstract VI problem \eqref{eq:vi}, we now consider box-constrained cases only, using set \eqref{eq:originalconstraintset}.
+the closed and convex \emph{constraint set}.
+Note that we are treating inequality constraints and boundary conditions in a unified manner.\footnote{Any Neumann conditions over $\partial_N \Omega$ are determined by the operator formula for $f$, i.e.~by an integration over $\partial_N\Omega$.}
+
+Recalling abstract VI problem \eqref{eq:vi}, from now on we only constraint sets \eqref{eq:originalconstraintset}.
 Given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we will assume that the following is well-posed for $u^*\in \cK$:
@@
 \end{equation}
-Note that $v-u^* \in W_0^{1,p}(\Omega)$, so test functions in \eqref{eq:boxdirichletvi} have zero Dirichlet values.
-Any Neumann conditions over $\partial_N \Omega$ are determined by the operator formula for $f$, i.e.~by an integration over $\partial_N\Omega$.
-In summary, from now on we consider well-posed VI problems \eqref{eq:boxdirichletvi}, namely box-constrained boundary value problems.
+Note that $v-u^* \in W_0^{1,p}(\Omega)$.
 
@@
 
-We seek to rapidly compute the approximate solution of box-constrained VI problems using finite elements (FE).
-Suppose $\Omega \subset \RR^d$ is a (open) bounded polygon.\footnote{Section \ref{sec:results} illustrates $d=1,2,3$ cases.}  We will construct a multilevel FE approximation to problem \eqref{eq:boxdirichletvi} over $\Omega$ using $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, by refining a coarsest mesh into a mesh hierarchy; the goal is an accurate solution on the finest-level mesh.
-The hierarchy allows the creation of a multilevel CD of $\cK$, using the ideas in the next two Sections, which permits admissibility-preserving coarse corrections.
+We seek to rapidly compute the approximate solution of box-constrained VI problems \eqref{eq:boxdirichletvi} using finite elements (FE).
+Suppose $\Omega \subset \RR^d$ is a (open) bounded polygon.\footnote{Section \ref{sec:results} illustrates $d=1,2,3$ cases.}  We will construct a multilevel FE approximation to \eqref{eq:boxdirichletvi} over $\Omega$ using $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, by refining a coarsest mesh into a mesh hierarchy; the goal is an accurate solution on the finest-level mesh.
+The hierarchy allows the creation of a multilevel CD of $\cK$, using the ideas in the next two Sections, which permits solvers based on admissibility-preserving coarse corrections.
 
@@
 
-In the VI material which follows, use of higher-order Lagrange elements, i.e.~$P_k$ or $Q_k$ with $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007}, special to inequality constraints, which we avoid.
+In the VI material which follows, use of higher-order Lagrange elements, i.e.~$P_k$ or $Q_k$ with $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} special to inequality constraints, which we avoid.
 Namely, our FE construction of constraint sets (Section \ref{sec:cdmultilevel}) will require \emph{nodal monotonicity},
@@
 
-To allow box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions\footnote{Note that $\tilde{\mathcal{V}}^j$ functions need not be defined everywhere on $\Omega$.} on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
+To allow box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
 Suppose that the finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ satisfy
@@
 That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require or generally expect $\mathcal{K}^J \subset \mathcal{K}$.
-Regardless of the smoothness of the continuum obstacles $\underline{\gamma}, \overline{\gamma}$, admissibility with respect to their FE interpolants $\underline{\gamma}^J, \overline{\gamma}^J$ will generally not imply continuum admissiblity.\footnote{This caveat is especially relevant in geophysical situations where only nodal values of the obstacle are known.
-The nodal elevation data misses the actual topographic detail in-between the nodes.
-See the Greenland ice sheet application in \cite{Bueler2016}, for example.}
-
-The multilevel algorithm defined in the next two Sections, a solver for FE VI \eqref{eq:fe:vi}, requires transfer operators between meshes.
-We require five such operators, of which three are standard multigrid tools \cite{Trottenbergetal2001}.
+Regardless of the smoothness of the continuum obstacles $\underline{\gamma}, \overline{\gamma}$, admissibility with respect to their FE interpolants $\underline{\gamma}^J, \overline{\gamma}^J$ will generally not imply continuum admissibility.\footnote{This caveat is especially relevant in geophysical situations where only nodal values of the obstacle are known.
+Nodal elevation data generally misses the actual topographic detail in-between the nodes; compare \cite{Bueler2016}.}
+
+The multilevel algorithm defined in Section \ref{sec:vcycle}, a solver for FE VI \eqref{eq:fe:vi}, requires transfer operators between meshes.
+We require five such operators, of which three are standard \cite{Trottenbergetal2001}.
 Table \ref{tab:transfers} will help the reader keep track.
@@
 Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ also acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $(R\sigma)[z] = \sigma[z]$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
-Note that these canonical transfers generally involve a change of FE representation.
-For example, the formula for canonical restriction, derived by writing coarser-mesh hat functions in terms of finer-mesh ones, is ``full-weighting'' \cite{Trottenbergetal2001}.
+These canonical transfers generally involve a change of FE representation.
+For example, $R$ is ``full-weighting'' \cite{Trottenbergetal2001}.
 
@@
 For $z\in\mathcal{V}^j$ we define the injection $\iR z$ as the element of $\mathcal{V}^{j-1}$ with the same (coarse) nodal values.
-Note that injection is monotonic: if $w \le z$ then $\iR w \le \iR z$.
-
-Finally, the monotone restrictions \cite{GraeserKornhuber2009} $\maxR,\minR$ are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh hat functions.
-These monotone operators will only be applied to nodal functions $\tilde{\mathcal{V}}^j$.
-To make the definitions precise, let $x_q^j$, $q=0,1,\dots,m_j-1$, denote the nodes of mesh $\mathcal{T}^j$, and denote the corresponding hat functions by $\psi_q^j(x)$.
-The monotone restrictions of an $\tilde\RR$-valued\footnote{In computing $\max$ and $\min$ to define \eqref{eq:fe:monotonerestrictions}, the usual $\tilde\RR$ ordering is applied.} nodal function $z\in\tilde{\mathcal{V}}^j$ are the following functions in $\tilde{\mathcal{V}}^{j-1}$:
+Injection is monotonic: if $w \le z$ then $\iR w \le \iR z$.
+
+The monotone restrictions $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to nodal functions $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh hat functions.
+To make this precise, let $x_q^j$ denote the nodes of mesh $\mathcal{T}^j$, with corresponding hat functions $\psi_q^j(x)$.
+For a $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$:
 \begin{align}
 (\maxR z)(x_p^{j-1}) &= \max \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}, \label{eq:fe:monotonerestrictions} \\
-(\minR z)(x_p^{j-1}) &= \min \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}, \notag
+(\minR z)(x_p^{j-1}) &= \min \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}. \notag
 \end{align}
-for $p=0,1,\dots,m_{j-1}-1$.  By standard nested refinement, a coarse-mesh node $x_p^{j-1}$ is in the set of fine nodes $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so it follows that
+For standard refinement a coarse-mesh node $x_p^{j-1}$ is in the set of fine nodes $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so it follows that
 \begin{equation}
@@
 \end{equation}
-On the other hand, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These ordering properties used in the CD construction of the next two sections.
+On the other hand, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These ordering properties are used in the CD construction of the next two Sections.
 
@@
 
-Our practical multilevel CD scheme in this Section uses \emph{defect constraints} \cite{GraeserKornhuber2009} which are based on an admissible finest-level iterate.
-The next Example concretely illustrates why we do not apply restriction operators directly to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$ in definition \eqref{eq:fe:fineconstraintset}.
-Our straightforward iterate-dependent bounds, namely defect constraints \eqref{eq:fe:defectconstraints} below, bypass the difficulties exposed in Example \ref{ex:directRbad}.
-However, other successful schemes may exist.
+Our practical multilevel CD scheme uses \emph{defect constraints} \cite{GraeserKornhuber2009} which are based on an admissible finest-level iterate.
+The next Example concretely illustrates why we do not apply restriction operators directly\footnote{An exception is in the F-cycle; see Section \ref{sec:implementation}.} to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.
+Our straightforward (iterate-dependent) bounds on the defect constraints, namely \eqref{eq:fe:chiordering} below, bypass the difficulties exposed in Example \ref{ex:directRbad}.
 
@@
 Let $\underline{\gamma}^1$ be a sawtooth function with minimum value zero, and suppose $\overline{\gamma}^1=C+\underline{\gamma}^1$ with $C>0$ which is strictly less than the range of $\underline{\gamma}^1$.
+Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ nodes, respectively.
 Supposing empty Dirichlet boundary for simplicity, the constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.
-Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ nodes, respectively.
 However, various restriction schemes applied directly to $\underline{\gamma}^1,\overline{\gamma}^1$ are problematical.
@@
 
-Subtraction of an admissible iterate provides a route out of these difficulties.
-Suppose $w^J \in \cK^J$ is an admissible finest-level iterate (i.e.~which approximates the solution $u^J$ of \eqref{eq:fe:vi}).
+Subtracting an admissible iterate from the original bounds provides a route out of the difficulties in Example \ref{ex:directRbad}.
+Suppose $w^J \in \cK^J$ is an admissible finest-level iterate for \eqref{eq:fe:vi}.
 We define the \emph{finest-level defect constraints} (DCs) \cite{GraeserKornhuber2009} as the following nodal functions in $\tilde{\mathcal{V}}^J$:
@@
 \end{equation}
-Definition \eqref{eq:fe:fineconstraintset} implies that $\underline{\chi}^J \le 0 \le \overline{\chi}^J$ everywhere in $\overline{\Omega}$, including along the Dirichlet boundary $\partial_D\Omega$.
+Definition \eqref{eq:fe:fineconstraintset} implies that $\underline{\chi}^J \le 0 \le \overline{\chi}^J$ everywhere in $\overline{\Omega}$, including on the Dirichlet boundary $\partial_D\Omega$.
 If $\underline{\gamma}^J=-\infty$ then $\underline{\chi}^J=-\infty$ by definition, and similarly if $\overline{\gamma}^J=+\infty$ then $\overline{\chi}^J=+\infty$.
-Observe that a corrected iterate $w^J + z^J$ is admissible, $w^J + z^J \in \cK^J$, if and only if the perturbation $z^J$ is zero on $\partial_D\Omega$ and it is bounded by the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$.
-
-Descending coarser levels, $j=J-1,\dots,1,0$, we apply the monotone restrictions to generate box \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:
+Observe that a corrected iterate $w^J + z^J$ is admissible, $w^J + z^J \in \cK^J$, if and only if the perturbation $z^J \in \mathcal{V}^J$ is zero on $\partial_D\Omega$ and bounded by the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$.
+
+Descending coarser levels, $j=J-1,\dots,1,0$, we apply the monotone restrictions to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:
 \begin{equation}
@@
 A unilateral $J=4$ example of LDCs $\underline{\chi}^j$ is shown in Figure \ref{fig:chiphilevels}.
-The LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if another element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$, and similar comments apply to $\overline{\chi}^{j}$.
+The LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if another element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$; similar comments apply to $\overline{\chi}^{j}$.
 
@@
 
-This construction of multilevel box constraints is designed to permit the largest admissible coarse corrections, which promotes multilevel solver efficiency.
-Note that for the multilevel schemes in Sections \ref{sec:vcycle} and \ref{sec:implementation}, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are actually not required.
-Also, though the system here is fundamentally the same as the unilateral defect constraints in \cite{GraeserKornhuber2009}, a system for box defect constraints has not been found in the literature.
-
-We now compute the LDC differences (compare \cite{GraeserKornhuber2009}):
+This construction of multilevel box constraints, fundamentally the same as the unilateral construction in \cite{GraeserKornhuber2009}, is designed to permit large-yet-admissible coarse corrections, which promotes multilevel solver efficiency.
+Note that for Algorithm \ref{alg:fascd} below, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are actually not required.
+
+We now compute the LDC differences \cite{GraeserKornhuber2009}:
 \begin{equation}
@@
 Regarding subtraction of extended real-valued quantities, if $\underline{\chi}^j(x_p)=-\infty$ then we set $\underline{\phi}^j(x_p)=-\infty$ regardless of the value of $\underline{\chi}^{j-1}(x_p)$, and similarly $\overline{\phi}^j(x_p)=+\infty$ if $\overline{\chi}^j(x_p)=+\infty$.
-These new LDCs $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ again bracket zero, $\underline{\phi}^j \le 0 \le \overline{\phi}^j$, but they are not ordered as in \eqref{eq:fe:chiordering}.
-Table \ref{tab:ldcs} should help the reader keep track of our box-constrained LDC system.
-
-\begin{table}[H]
-\begin{tabular}{lll}
-\emph{name}         & \emph{formulas} & \emph{indices} \\ \hline
-finest-level DCs & ${\Large \strut} \underline{\chi}^J = \underline{\gamma}^J - w^J,\, \overline{\chi}^J = \overline{\gamma}^J - w^J$ \\
-upward LDCs  & ${\Large \strut} \underline{\chi}^{j} = \maxR \underline{\chi}^{j+1}, \, \overline{\chi}^{j} = \minR \overline{\chi}^{j+1}$ & $j=0,\dots,J-1$ \\
-downward LDCs       & ${\Large \strut} \underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1}, \, \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}$ & $j=0,\dots,J$
-\end{tabular}
-
-\medskip
-\caption{Level defect constraints (LDCs) for box-constrained VI problems.}
-\label{tab:ldcs}
-\end{table}
-
-The corresponding \emph{downward/upward constraint sets}, respectively, are
+While $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ again bracket zero, $\underline{\phi}^j \le 0 \le \overline{\phi}^j$, they are not ordered as in \eqref{eq:fe:chiordering}.
+
+The corresponding \emph{downward/upward constraint sets}, respectively, are closed and convex subsets of the FE spaces $\mathcal{V}^j$:
 \begin{align}
@@
 \end{align}
-These are closed and convex subsets of the FE spaces $\mathcal{V}^j$.
 Because $\underline{\chi}^j \le \underline{\phi}^j \le 0 \le \overline{\phi}^j \le \overline{\chi}^j$, it follows that $\mathcal{D}^j \subseteq \mathcal{U}^j$.
@@
 
-The multilevel V-cycle solver defined in the next Section computes coarse corrections in $\mathcal{D}^j$ for the downward side and in $\mathcal{U}^j$ for the upward side.
-Our multilevel strategy is, however, based on \emph{incomplete} CDs (Section \ref{sec:cd}), in the general box-constrained cases, as we now explain.
-Each upward set $\mathcal{U}^j$ can be decomposed in an inclusion sense, becoming a true CD in unilateral cases.
-The following Lemma is based in part on the unilateral CD construction underlying Algorithm 4.7 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.
+Our V-cycle solver (Algorithm \ref{alg:fascd}) will compute coarse corrections in $\mathcal{D}^j$ for the downward side and in $\mathcal{U}^j$ for the upward side.
+This strategy is, however, based on \emph{incomplete} CDs, in the general box-constrained cases, as we now explain.
+Each upward set $\mathcal{U}^j$ is generally only a superset of a sum of downward sets $\mathcal{D}^j$, but this becomes a true CD, i.e.~\eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, in unilateral cases.
+The following Lemma, apparently new, is based in part on the unilateral CD construction underlying Algorithm 4.7 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.
 
@@
 for $j=0,1,\dots,J$.
-In unilateral cases, where either $\underline{\gamma}^J=-\infty$ identically or $\overline{\gamma}^J=+\infty$ identically, inclusion \eqref{eq:fe:downwardsuminclusion} becomes a full CD \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} with $\mathcal{U}^j=\sum_{i=0}^j \mathcal{D}^i$.
+In unilateral cases, where either $\underline{\gamma}^J=-\infty$ identically or $\overline{\gamma}^J=+\infty$ identically, inclusion \eqref{eq:fe:downwardsuminclusion} becomes a full CD with $\mathcal{U}^j=\sum_{i=0}^j \mathcal{D}^i$.
 \end{lemma}
@@
 \end{equation}
-Note that $I_{j\to j}^\ominus v = v$, and set $I_{j\to -1}^\ominus=0$ by definition.
-From operators \eqref{eq:fe:minimummaps} we define the following nonlinear decomposition operators $\Pi_i:\mathcal{U}^j \to \mathcal{D}^i$:
+(Set $I_{j\to j}^\ominus v = v$ and $I_{j\to -1}^\ominus=0$ by definition.)  Now define the following nonlinear decomposition operators $\Pi_i:\mathcal{U}^j \to \mathcal{D}^i$:
 \begin{equation}
@@
 \end{equation}
-(Compare equation (4.9) in \cite{GraeserKornhuber2009}.)  Property \eqref{eq:fe:monotonerestrictionprops} implies $I_{j\to i}^\ominus(z^j - \underline{\chi}^j) \ge I_{j\to i-1}^\ominus(z^j - \underline{\chi}^j)$, thus $\Pi_i z^j \ge \underline{\phi}^i$, so $\Pi_i z^j \in \mathcal{D}^i$ (in this unilateral case).
+(Compare equation (4.9) in \cite{GraeserKornhuber2009}.)  Property \eqref{eq:fe:monotonerestrictionprops} implies $I_{j\to i}^\ominus(z^j - \underline{\chi}^j) \ge I_{j\to i-1}^\ominus(z^j - \underline{\chi}^j)$, thus $\Pi_i z^j \ge \underline{\phi}^i$, so $\Pi_i z^j \in \mathcal{D}^i$.
 On the other hand, the following sum telescopes, leaving only its first term plus the sum in \eqref{eq:fe:telescoping}:
@@
 \end{align*}
-This shows \eqref{eq:constraintrestrictionsum}, thus that \eqref{eq:fe:downwardsuminclusion} is equality, yielding a true CD \eqref{eq:constraintdecomp}.
-
-The $\underline{\gamma}^J=-\infty$ case is similar.
+This shows \eqref{eq:constraintrestrictionsum}, thus that \eqref{eq:fe:downwardsuminclusion} is equality, yielding \eqref{eq:constraintdecomp}.  The $\underline{\gamma}^J=-\infty$ case is similar.
 \end{proof}
@@
 One might try to turn \eqref{eq:fe:downwardsuminclusion} into a full CD for arbitrary box constraints by using decomposition maps like $\Pi_i$, for example by splitting $z^j = \min\{z^j,0\} + \max\{z^j,0\}$ and applying unilateral formulas like \eqref{eq:fe:unilateraldecompositionoperator} to $\min\{z^j,0\}$ and $\max\{z^j,0\}$ separately.
-In fact, the lower bound $\Pi_i (\min\{z^j,0\}) \ge \underline{\phi}^i$ holds, as shown in the above proof.
-However, the following Example shows why $\Pi_i(\min\{z^j,0\})$ may have an arbitrarily large maximum, and thus exceed any finite upper LDC $\overline{\phi}^i$.
+The above proof shows that the lower bound $\Pi_i (\min\{z^j,0\}) \ge \underline{\phi}^i$ does hold.
+However, the following Example\footnote{Of course, this Example does not exclude the existence of an entirely-different strategy for showing a full CD in the general box-constrained case.} shows why $\Pi_i(\min\{z^j,0\})$ may have an arbitrarily large maximum, and thus exceed any finite upper LDC $\overline{\phi}^i$.
 In fact, $\Pi_i$ does not generally map all non-positive elements of $\mathcal{U}^J$ into $\mathcal{D}^i$ if a finite upper LDC is present.
-Note that this Example does not exclude the existence of an entirely-different strategy for showing a full CD in the general box-constrained case.
 
@@
 Consider a 2-level hierarchy ($J=1$), and suppose the lower defect constraint is the constant function $\underline{\chi}^1=-a$.
-Recalling Table \ref{tab:ldcs}, note $\underline{\chi}^0=\maxR \underline{\chi}^1=-a$ also, so $\underline{\phi}^1=0$ and $\underline{\phi}^0=-a$.
+Note $\underline{\chi}^0=\maxR \underline{\chi}^1=-a$ also, so $\underline{\phi}^1=0$ and $\underline{\phi}^0=-a$.
 Suppose $z^1$ is a ``sawtooth'' function which takes on value $-a$ on the coarse nodes and value $0$ on the other nodes.
-Then, for any valid $\overline{\chi}^1$, we have $\overline{\chi}^1 \ge 0 \ge z^1\ge \underline{\chi}^1$, so $z^j \in \mathcal{U}^1$.
+Then, for any $\overline{\chi}^1\ge 0$, we have $\overline{\chi}^1 \ge 0 \ge z^1\ge \underline{\chi}^1$, so $z^j \in \mathcal{U}^1$.
 However, $I_{1\to 0}^\ominus(z^1 - \underline{\chi}^1) = \minR(z^1 - \underline{\chi}^1) = 0$ identically.
@@
 
-Each upward defect constraint set $\mathcal{U}^j$ can also be decomposed down to an arbitrary coarser level.
-The incomplete CD in the next Lemma, which does not follow from Lemma \ref{lem:downwardadmissibility}, will justify upward admissibility in the V-cycle in the next Section.
+Each upward set $\mathcal{U}^j$ can also be decomposed (incompletely) down to an arbitrary coarser level.
+The next Lemma, which does not follow from Lemma \ref{lem:downwardadmissibility}, justifies upward admissibility in Algorithm \ref{alg:fascd}.
 
@@
 \end{equation}
-In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, inclusion \eqref{eq:fe:upwardsuminclusion} becomes a full CD. \end{lemma}
+In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, actually $\mathcal{U}^j = \mathcal{U}^k + \sum_{i=k+1}^j \mathcal{D}^i$. \end{lemma}
 
 \begin{proof}  Definition \eqref{eq:fe:philevels} shows that $\underline{\chi}^j = \underline{\phi}^j + \dots + \underline{\phi}^{k+1} + \underline{\chi}^k$ and $\overline{\chi}^j = \overline{\phi}^j + \dots + \overline{\phi}^{k+1} + \overline{\chi}^k$.
-If $y^i \in \mathcal{D}^i$ for $k+1 \le i \le j$ and $z^k \in \mathcal{U}^k$ then it is easy to show, as in \eqref{eq:fe:lemmaordering}, that $\underline{\chi}^j \le y^j + \dots + y^{k+1} + z^k \le \overline{\chi}^j$, thus \eqref{eq:fe:upwardsuminclusion} holds.
+If $y^i \in \mathcal{D}^i$ for $k+1 \le i \le j$ and $z^k \in \mathcal{U}^k$ then $\underline{\chi}^j \le y^j + \dots + y^{k+1} + z^k \le \overline{\chi}^j$ and \eqref{eq:fe:upwardsuminclusion}.
 
@@
 
-FIXME \pr{icd-tele} as third basic CD iteration
-
-\begin{pseudofloat}[H]
+Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion}, which are special to multilevel FE schemes based on defect constraints, hold regardless of whether a CD operator $\Pi_i$ can be constructed.
+Here there is, in fact, another ``basic'' iteration, namely \pr{icd-tele} in Algorithm \ref{alg:basiccd-tele}, which also does not need the $\Pi_i$ operators.
+It is a variant of \pr{cd-mult} (Algorithm \ref{alg:basiccd-mult}).
+
+\begin{pseudofloat}[ht]
 \begin{pseudo*}
@@
     for $j = 0,\dots,J$: \\+
-        \rm{find} $z_j\in \mathcal{U}_j$ \rm{so that for all} $v_j\in \mathcal{U}_j$, \\+
-            $\displaystyle \ip{f\Big(w^J + \alpha \sum_{i<j} z_i + \alpha z_j\Big)}{v_j-z_j} \ge \ip{\ell}{v_j-z_j}$ \\--
-    return $\tilde w^J=w^J + \alpha \sum_{i\le J} z_i$
+        \rm{find} $z^j\in \mathcal{U}^j$ \rm{so that for all} $v \in \mathcal{U}^j$, \\+
+            $\displaystyle \ip{f\Big(w^J + \alpha \sum_{i<j} z^i + \alpha z^j\Big)}{v-z^j} \ge \ip{\ell}{v-z^j}$ \\--
+    return $\tilde w^J=w^J + \alpha \sum_{i\le J} z^i$
 \end{pseudo*}
-\caption{One telescoping incomplete CD iteration for VI problem \eqref{eq:vi}, assuming $w^J\in\mathcal{K}^J$, property \eqref{eq:fe:finestlevelequivalent}, and incomplete CD \eqref{eq:fe:upwardsuminclusion}.}
+\caption{One telescoping incomplete CD iteration for VI problem \eqref{eq:fe:vi}, assuming $w^J\in\mathcal{K}^J$, property \eqref{eq:fe:finestlevelequivalent}, and incomplete CD \eqref{eq:fe:upwardsuminclusion}.}
 \label{alg:basiccd-tele}
@@
 
-In summary, inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in the V-cycle algorithm which follows.
-However, a convergence analysis, perhaps extending the arguments of \cite{Tai2003} or \cite{GraeserKornhuber2009} to VIs \eqref{eq:vi} which are not of minimization type, and/or to box constraints, would seem to require full CDs, and probably further new ideas, and we cannot offer that analysis.
+While inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in Algorithm \ref{alg:fascd} below, a convergence analysis, perhaps extending the arguments of \cite{Tai2003} or \cite{GraeserKornhuber2009} to VIs which are not of minimization type, and/or to box constraints, would seem to require full CDs.
+We cannot offer that analysis.
 
@@
 The V-cycle algorithm in this section applies smoothers in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.
-It extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear full approximation storage (FAS) approach \cite{BrandtLivne2011}, and in fact it reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).
-
-Our V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ based on perturbations from each coarser FE subspace $\mathcal{V}^j$.
-By nesting \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$,\footnote{Note we use prolongation $P$ to indicate where the computer representation changes.} but the corrected iterate needs to be admissible.
-Suppose $y^i \in \mathcal{D}^i$ for $i=J,\dots,j-1$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).
+It extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear full approximation storage (FAS) approach \cite{BrandtLivne2011}.
+In fact, it reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).
+
+Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ based on perturbations from each coarser FE subspace $\mathcal{V}^j$.
+By nesting \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$,\footnote{Nonetheless we use prolongation $P$ to indicate where the computer representation changes.} but the corrected iterate needs to be admissible.
+Proceeding inductively, suppose $y^i \in \mathcal{D}^i$ for $i=J,\dots,j-1$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).
 By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.
-Smoothing, an incomplete solution process on a VI problem (Section \ref{sec:implementation}), then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.
+Smoothing then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.
 At the bottom the coarsest-level VI is then solved, resulting in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.
-Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$ using \eqref{eq:fe:upwardsuminclusion}.
-Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction.
+Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using \eqref{eq:fe:upwardsuminclusion}.
+Continuing in this ``telescoping'' way (Algorithm \ref{alg:basiccd-tele}), up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction.
 Then the next admissible iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.
@@
 
-A key idea is that, because $\mathcal{U}^j\supset \mathcal{D}^j$, up-smoothing solves less-restricted VI problems than down-smoothing.
-Intuitively-speaking, going downward one does not know what the next coarse corrections will be, thus they must be in small sets $\mathcal{D}^j$, so that their sum is admissible once we return to a given level.
-On the other hand, going upward no further coarse corrections can violate admissibility.
+A key idea is that, because of inclusions \eqref{eq:fe:upwardsuminclusion}, up-smoothing solves less-restricted VI problems than down-smoothing.
+Intuitively-speaking, going downward one does not know what the next coarse corrections will be, and thus they must be in sufficiently small sets ($\mathcal{D}^j$) so that their sum is admissible once we return to a given level.
+Going upward, however, no further coarse corrections can violate admissibility.
 
@@
 As in FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, and a source functional created.
-On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:boxdirichletvi}, presumably using quadrature etc.~by the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.
-Note that finest-level problem \eqref{eq:fe:vi} uses the original source functional $\ell^J$.
-An admissible iterate $w^J \in \mathcal{K}^J$ has already\footnote{In practice the LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ are generated during the descent stage; see Algorithm \ref{alg:fascd}.} been used to generate the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}); these LDCs do not depend on the coarse-level corrections.
+On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:boxdirichletvi}.\footnote{Presumably it uses quadrature etc.~in the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.}  Finally, we assume that an admissible iterate $w^J \in \mathcal{K}^J$ has already\footnote{In practice the LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ are generated during the descent stage; see Algorithm \ref{alg:fascd}.} been used to generate the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}); these LDCs do not depend on the coarse-level corrections.
 
@@
 \end{equation}
-(These are the states \emph{before} smoothing and/or correction.)  From $w^j$ we define $\ell^j \in (\cV^j)'$:
+(These are the states before smoothing or correction.)  From $w^j$ we define $\ell^j \in (\cV^j)'$:
 \begin{equation}
@@
 \end{equation}
-Observe that injection $\iR$ coarsens the iterates---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---but canonical restriction $R$ coarsens linear functionals.\footnote{Recall that Table \ref{tab:transfers} summarizes our transfer operators.}
+Observe that injection $\iR$ coarsens the iterates---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---but canonical restriction $R$ coarsens linear functionals; recall Table \ref{tab:transfers}.
 
@@
 for which the initial iterate is $y_{(0)}^j=0$.
-(Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)  On ascent, $j=0,1,\dots,J$, the same VI is solved but for $z^j \in \mathcal{U}^j$, in a larger admissible set:
+(Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)  On ascent, $j=0,1,\dots,J$, the same VI is solved but for $z^j \in \mathcal{U}^j$:
 \begin{equation}
@@
 \end{equation}
-The nontrivial initial iterate for this problem is found by prolonging $z^{j-1}$ and using it to correct the current (i.e.~down-smoothed) correction on the level:\footnote{Prolongation $P$ in \eqref{eq:fe:upwardinitial} reflects the change in FE representation.}
+The nontrivial initial iterate for this problem is found by prolonging $z^{j-1}$ and using it to correct the current (i.e.~down-smoothed) correction on the level:
 \begin{equation}
@@
 \end{equation}
-An exception is that $z_{(0)}^0=0$ on the coarsest level.  Note $z_{(0)}^j$ is admissible since $\mathcal{D}^j + \mathcal{U}^{j-1} \subset \mathcal{U}^j$; see \eqref{eq:fe:upwardsuminclusion}.
+An exception is that $z_{(0)}^0=0$ on the coarsest level.  Note $z_{(0)}^j$ is admissible by \eqref{eq:fe:upwardsuminclusion}: $\mathcal{D}^j + \mathcal{U}^{j-1} \subset \mathcal{U}^j$.
 
@@
 In this sense $w^j$ is admissible with respect to the box constraints.
-However, while $w^j \in \cV^j \subset \cV^J$ holds, generally $w^j \notin \cK^J$, and furthermore, following \cite{GraeserKornhuber2009}, our use of LDCs (Table \ref{tab:ldcs}) avoids the need to generate these $\underline{\gamma}^j, \overline{\gamma}^j$.
-In any case, Algorithm \ref{alg:fascd} yields a new finest-level iterate $w^J+z^J \in \cK^J$ which is  admissible for problem \eqref{eq:fe:vi}.
-
-Observe that formula \eqref{eq:fe:levelsource} for functional $\ell^j$ is a classical FAS construction; compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.
-It has the following explanation, mimicking the PDE version, which simultaneously justifies VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.
+However, while $w^j \in \cV^j \subset \cV^J$ holds, generally $w^j \notin \cK^J$, and furthermore our use of LDCs (Section \ref{sec:cdmultilevel}) avoids the need to generate these $\underline{\gamma}^j, \overline{\gamma}^j$.
+
+Observe that formula \eqref{eq:fe:levelsource} for functional $\ell^j$ is a classical FAS construction for PDEs; compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.
+It has the following explanation, which simultaneously justifies VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.
 First, the $j+1$ level problem \eqref{eq:fe:downvi} is not solved exactly by the finer-level correction $y^{j+1}$---the down-smoother was only approximate---so we seek a further correction $y$ on the $j+1$ level such that
@@
 \end{align}
-This is the same as VI problem \eqref{eq:fe:downvi}, but in cluttered form; definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource} create \eqref{eq:fe:downvi}.
-The explanation for VI \eqref{eq:fe:upvi} is similar.
-
-We will regard the smoothers for VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} as approximate solvers, acting on the corrections $y^j$ and $z^j$, respectively (Section \ref{sec:implementation}).
-Also, as is standard in multigrid, at the coarsest $j=0$ level we suppose that problem \eqref{eq:fe:upvi} is solved accurately.
+This is the same as VI problem \eqref{eq:fe:downvi}, but in cluttered form, thus definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource}.  The explanation for VI \eqref{eq:fe:upvi} is similar.
 
 These ideas come together in Algorithm \ref{alg:fascd}, \pr{fascd-vcycle}, which updates $w^J$.
-It assumes certain semantics for the \pr{smooth} and the coarsest-level \pr{solve} procedures; they do in-place modifications of their final arguments, without modifying their first four arguments.
+It assumes that the \pr{smooth} and \pr{solve} procedures do in-place modifications of their final arguments, without modifying their first four arguments.
 Parameters \id{down} and \id{up} determine the number of smoother iterations, respectively.
@@
 
-Compared specifically to the $V(1,0)$ cycle presented as Algorithm 4.7 in \cite{GraeserKornhuber2009}, one might say informally that \pr{fascd-vcycle} is a $V(1,1)$ cycle.
-However, its structure relates directly to the basic CD iterations (Sections \ref{sec:cd}) as well.
-In unilateral cases it is precise to say that one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} over CD \eqref{eq:fe:downwardsuminclusion}, using the defect constraint meaning of the downward constraint sets $\mathcal{D}^j$, followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over the incomplete CD \eqref{eq:fe:upwardsuminclusion}.
-FIXME: idea is that once you are using defect constraints you no longer depend on the literal availability of $\Pi_i$, so incomplete CD suffices, e.g.~even for \pr{cd-mult}
-
-Regarding storage required for \pr{fascd-vcycle}, and assuming all box constraints are nontrivial, 7 vectors must be allocated on each level: $\underline{\chi}^j$, $\overline{\chi}^j$, $\underline{\phi}^j$, $\overline{\phi}^j$, $w^j$, $\ell^j$, $y^j$.
+Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle.\footnote{We could write ``$V(\id{down},\id{up})$'', but we prefer to emphasize the number of subsets used on each level rather than the number of smoother sweeps in a given set.}  If \id{up} $=0$ then it becomes a $V(1,0)$ cycle and if \id{down} $=0$ then it is a $V(0,1)$ cycle; in the former case it generalizes Algorithm 4.7 in \cite{GraeserKornhuber2009} by allowing nonlinear operators $f$, any smoother, and applying FAS-type corrections.
+In unilateral cases it is precise to say that one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} over CD \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over incomplete CD \eqref{eq:fe:upwardsuminclusion}; in this sense $V(1,1)=V(1,0)+V(0,1)$.
+
+Regarding the floating-point storage required, let us assuming all box constraints are nontrivial.
+Then seven vectors must be allocated on each level: $\underline{\chi}^j$, $\overline{\chi}^j$, $\underline{\phi}^j$, $\overline{\phi}^j$, $w^j$, $\ell^j$, $y^j$.
 Note $z^j$ can use the same storage as $y^j$.
-On the finest level one must also store $\underline{\gamma}^J$ and $\overline{\gamma}^J$, while on the coarsest level there are only 5 vectors because $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.
-Therefore, using $m_j=2^d m_{j-1}$, plus a geometric series argument, the total floating-point storage, excluding smoother and coarse-solver implementations, is
+On the finest level one must also store $\underline{\gamma}^J$ and $\overline{\gamma}^J$, while on the coarsest level there are only five vectors because $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.
+Therefore, using $m_j=2^d m_{j-1}$, plus a geometric series argument, the total storage, excluding smoother and coarse-solver implementations, is
 \begin{equation}
@@
 This gives storage at most $16m_J$, $12m_J$, and $10m_J$ in dimensions $d=1,2,3$, respectively.
-For comparison, a single-level method needs $4 m_J$ storage.\footnote{That is, for vectors $\underline{\gamma}^J,\overline{\gamma}^J,\ell^J,w^J$.}  If some of the obstacles are trivial, for instance in the unilateral case $\overline{\gamma}^J=+\infty$, storage is reduced at all levels.
+For comparison, a single-level method needs $4 m_J$ storage.\footnote{That is, for vectors $\underline{\gamma}^J,\overline{\gamma}^J,\ell^J,w^J$.}  If some of the obstacles are trivial, for instance in the unilateral case $\overline{\gamma}^J=+\infty$, storage is correspondingly reduced at all levels.
 
@@
 
+FIXME the fact that FASCD reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}) suggests it is a completely different strategy from PFAS in \cite{BrandtCryer1983}
+
 FIXME desired extension to nonlocal variational inequalities not tried here

### 2023-06-28 paper/fascd.tex
fiddle with text ... this version for Patrick

@@
 This gives storage at most $16m_J$, $12m_J$, and $10m_J$ in dimensions $d=1,2,3$, respectively.
-For comparison, a single-level method needs $4 m_J$ storage.\footnote{That is, for vectors $\underline{\gamma}^J,\overline{\gamma}^J,\ell^J,w^J$.}  If some of the obstacles are trivial, for instance in the unilateral case $\overline{\gamma}^J=+\infty$, storage is correspondingly reduced at all levels.
+For comparison, a single-level method needs $4 m_J$ storage (i.e.~$\underline{\gamma}^J,\overline{\gamma}^J,\ell^J,w^J$).
+If some of the obstacles are trivial, for instance in the unilateral case $\overline{\gamma}^J=+\infty$, storage can be correspondingly reduced at all levels.
 
@@
 
-In this section we consider several practical concerns when implementing and applying \pr{fascd-vcycle} Algorithm \ref{alg:fascd}.
-We address convergence criteria, the F-cycle extension, and an active set Newton method used in Section \ref{sec:results} as a smoother and coarse solver.
-
-Because the problem is inequality-constrained, repeated application of \pr{fascd-vcycle} to solve the finite-dimensional, box-constrained VI \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.
-Indeed \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}, a strong form amenable to constructing a convergence criteria for the discretization.
+In this section we consider the implementation of FASCD methods.  We address convergence criteria, the F-cycle extension, and an active set Newton method used as a smoother and coarse solver.
+
+Because the problem is inequality-constrained, repeated application of Algorithm \ref{alg:fascd} to solve the finite-dimensional, box-constrained VI \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.
+To construct an appropriate convergence criteria for the discretization, note \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}, a strong form.
 Let $x_p \in \mathcal{T}^J$ be a finest-level node, and $\psi_p(x)$ the corresponding hat function.
@@
 For convenience denote $\underline{g}_p = w^J(x_p) - \underline{\gamma}^J(x_p)$ and $\overline{g}_p = \overline{\gamma}^J(x_p) - w^J(x_p)$, the nonnegative gaps at $x_p$.
-Now define\footnote{This semismooth residual is straightforwardly modified for unconstrained degrees of freedom, that is, where $\overline{\gamma}^J(x_p) = +\infty$ or $\underline{\gamma}^J(x_p) = -\infty$ or both.}
+Now define
 \begin{equation}
@@
 \end{equation}
-Suppose an iteration starts with $w^{J,0} \in \mathcal{K}^J$, generating (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.
-Let $\id{atol}\ge 0$, $\id{rtol}\ge 0$ be given tolerances.
-Then the criterion used in Section \ref{sec:results} for computations is\footnote{A step tolerance $\|w^{J,k} - w^{J,k-1}\|_2 < \id{stol}$ or similar could also be used, but we did not find it necessary.}
+Straightforward modifications apply for unconstrained degrees of freedom, where $\overline{\gamma}^J(x_p) = +\infty$ or $\underline{\gamma}^J(x_p) = -\infty$ or both.
+
+Suppose an iteration starts with $w^{J,0} \in \mathcal{K}^J$, generating (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.  Let $\id{atol}\ge 0$, $\id{rtol}\ge 0$ be given tolerances.  We propose
 \begin{equation}
@@
 \end{equation}
-
-Straightforward iteration of V-cycle Algorithm \ref{alg:fascd}, starting with some finest-mesh initial iterate $w^{J,0}$ and repeated until \eqref{eq:stoppingcriterion} holds, is demonstrated as a solver in Section \ref{sec:results}.
-However, application of an FASCD F-cycle, also called a \emph{full multigrid cycle}, is seen to be even better.
+as the stopping criterion.  A step tolerance, $\|w^{J,k} - w^{J,k-1}\|_2 < \id{stol}$ or similar, could also be used, but for the results in Section \ref{sec:results} we did not find it necessary.
+
+Iteration of V-cycle Algorithm \ref{alg:fascd}, starting with some finest-mesh initial iterate $w^{J,0}$ and repeated until \eqref{eq:stoppingcriterion} holds, will be shown (Section \ref{sec:results}) to be an effective solver.
+However, an F-cycle (Figure \ref{fig:fcycle}), also called a \emph{full multigrid cycle}, is even better.
 The well-known principle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate.
 Then an admissible initial iterate on the next-finer mesh is generated via prolongation and truncation, and so on.
-During this ``ramp'' up one approximately solves on each level by using a fixed number (\id{rampv}) of V-cycles, each down to the coarsest mesh.
+During this ``ramp'' up one approximately solves on each level by using a fixed number of V-cycles, each down to the coarsest mesh.
 From the finest level one does full-depth V-cycles to convergence.
-
-This F-cycle approach is in Figure \ref{fig:fcycle} and Algorithm \ref{alg:fascd-fcycle}.
-Observe that injection of the original fine-level constraints $\underline{\gamma}^J, \overline{\gamma}^J$ is \emph{not} used in generating the LDCs for V-cycles; compare Section \ref{sec:cdmultilevel}.
-However, injection is appropriate for the ramp in an F-cycle, if followed by truncation into $\mathcal{K}^j=\{v \in \mathcal{V}^j \,:\, \underline{\gamma}^j \le v \le \overline{\gamma}^j, \, v|_{\partial\Omega} = g_D\}$; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fcycle}.
-Though not included in Algorithm \ref{alg:fascd-fcycle}, one may generate the coarsest-level initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation down to coarsest level; compare line \ref{line:fcyclecoarsestinitial}.
 
@@
 \end{center}
-\caption{Algorithm \ref{alg:fascd-fcycle} initially restricts the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray edges).
+\caption{F-cycle Algorithm \ref{alg:fascd-fcycle} initially restricts the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray edges).
 Then it prolongs and truncates (doubled edges) to create an admissible initial iterate on the next-finer level.
@@
 \end{figure}
+
+This F-cycle approach is in Algorithm \ref{alg:fascd-fcycle}.
+Observe that while injection of the original fine-level constraints $\underline{\gamma}^J, \overline{\gamma}^J$ is \emph{not} used for V-cycles, wherein the LDCs use an existing admissible iterate (Section \ref{sec:cdmultilevel}), when combined with truncation into $\mathcal{K}^j=\{v \in \mathcal{V}^j \,:\, \underline{\gamma}^j \le v \le \overline{\gamma}^j, \, v|_{\partial\Omega} = g_D\}$ injection is appropriate for the ramp in an F-cycle; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo}.
+One may generate the coarsest-level initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation down to coarsest level; compare line \ref{line:fcyclecoarsestinitial}.
 
@@
 
-While a large space of smoothers could be tested,\footnote{See comments in Section \ref{sec:discussion}.} Section \ref{sec:results} only demonstrates one powerful, $O(m_j)$ work per iteration, smoother choice, namely a fixed number of iterations of the active-set Newton method from \cite{BensonMunson2006}, as implemented in PETSc \cite{Balayetal2023}, with a fixed number of preconditioned Krylov iterations to solve the reduced-space step equations.
-We will denote this truncated active-set Newton method as \emph{RS Newton}.
-The coarse solver is the same, but with LU decomposition solution of the step equations.
-For the symmetric Laplacian in Example \ref{ex:results:classical} we used conjugate gradients (CG) with incomplete Cholesky (IC) preconditioning, but for other operators the generalized minimum residual (GMRES) method with incomplete LU (ILU) preconditioning is used.
+A large space of smoothers could be explored,\footnote{See comments in Section \ref{sec:discussion}.} but Section \ref{sec:results} only demonstrates one powerful, $O(m_j)$ work per iteration, smoother choice.
+We apply a fixed number of iterations of the active-set (reduced-space) Newton method from \cite{BensonMunson2006}, the \texttt{-snes\_type vinewtonrsls} solver in PETSc \cite{Balayetal2023}, denoted as \emph{RS Newton}.
+We use it with a fixed number of preconditioned Krylov iterations to solve the step equations.
+For the symmetric Laplacian in Example \ref{ex:results:classical} we apply conjugate gradients (CG) with incomplete Cholesky (IC) preconditioning, but for other operators we apply the generalized minimum residual (GMRES) method with incomplete LU (ILU) preconditioning.
+The coarse solver is also RS Newton, but iterated to convergence and with LU decomposition solution of the step equations.
 

### 2023-06-30 paper/fascd.tex
Start of some edits

@@
 
-\newcommand{\ip}[2]{\left<#1,#2\right>}
+\newcommand{\ip}[2]{\langle#1,#2\rangle}
 
@@
 
+\usepackage{todonotes}
+\newcommand{\pef}[1]{\todo[inline, color=red!20]{PF: #1}}
+
 \begin{document}
@@
 \begin{abstract}
-The \emph{full approximation storage constraint decomposition} (FASCD) multilevel finite element method introduced here is a common extension of the full approximation storage multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and of the constraint decomposition method introduced by X.-C.~Tai for the variational inequalities associated to inequality-constrained optimization.
+We present the \emph{full approximation storage constraint decomposition} (FASCD) multilevel method for solving variational inequalities.
+FASCD is a common extension of both the full approximation storage (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and of the constraint decomposition (CD) method introduced by X.-C.~Tai for variational inequalities arising in inequality-constrained optimization.
 We extend the constraint decomposition idea by recognizing and exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
-Then, by using strong smoothers with work proportional to the number of unknowns on the finite element mesh level, we demonstrate that FASCD V-cycles show superior mesh-independent convergence rates and that FASCD F-cycles (full multigrid) are optimal solvers.
-The several example problems include unilateral and bi-obstacle variational inequality problems with symmetric linear, nonsymmetric linear, and nonlinear differential operators.
+Then, by using strong smoothers with work proportional to the number of unknowns on the finite element mesh level, we demonstrate that FASCD V-cycles show excellent mesh-independent convergence rates and that FASCD F-cycles (full multigrid) are optimal solvers.
+The example problems include unilateral and bi-obstacle variational inequality problems with symmetric linear, nonsymmetric linear, and nonlinear differential operators; the examples demonstrate very good convergence even on problems where other multilevel VI solvers struggle.
+\pef{I want to say it's better than monotone multigrid. Is this too subtle/too much?}
+
 \end{abstract}
@@
 
-The constraint decomposition (CD) methods of Tai \cite{Tai2003} are designed to solve coercive variational inequality (VI) problems which arise from minimization of a convex functional over a convex set.
-These methods have been shown to converge for multilevel and domain decompositions in a finite element (FE) context.
-In the multilevel case, the CD method has been shown to have almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
-In this paper we extend the multilevel CD method by replacing the coarse-level corrections with a full approximation storage (FAS; \cite{Brandt1977,Bruneetal2015}) approach, and we apply this new solver to nonlinear VI problems which do not arise from minimization.
+The constraint decomposition (CD) methods of Tai \cite{Tai2003} are designed to solve coercive variational inequality (VI) problems arising as the Karush--Kuhn--Tucker optimality conditions for the minimization of a convex functional over a convex set.
+These methods have been shown to converge for multilevel and domain decompositions in a finite element (FE) context; in the multilevel case, the CD method has been shown to have almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
+In this paper we extend the multilevel CD method by replacing the coarser-level corrections with a full approximation storage (FAS; \cite{Brandt1977,Bruneetal2015}) approach, and we apply this new solver to nonlinear VI problems which do not arise from minimization.
+\pef{Note I've changed ``coarse-level corrections'' to ``coarser-level corrections'', to imply we're not just operating only on the coarsest grids, but also on intermediate grids.}
 
@@
 \item We observe an asymmetry special to multilevel CD methods, namely that up-smoothing is more effective than down-smoothing because a ``telescoping'' subset property can be exploited on the up-smoothing side.
-\item We implement V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle})
+\item We implement V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}).
 \item By using a Newton-type active set method as a smoother, with a fixed number of Krylov iterations per step so that the work of the smoother is proportional to the number of unknowns on the given level, we demonstrate that F-cycle FASCD is an optimal solver.
+\pef{Clarify/limit which classes of problems we're claiming this for.}
 \end{enumerate}
 
-FASCD V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for the corresponding (unconstrained) PDE problems.
+\pef{I think we need some text here hitting the nail on the head about what's different to monotone multigrid methods (even if it's just the absence of the invasive modifications of the basis functions).}
+
+In our experiments, FASCD V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for the corresponding (unconstrained) PDE problems.
 The F-cycle results in Section \ref{sec:results} show nearly-textbook multigrid efficiency, namely solutions within discretization error in work comparable to few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}, for classical Laplacian obstacle problems, $p$-Laplacian obstacle problems, and nonnegativity-constrained advection-diffusion equations.
@@
 The iterates from FASCD are admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
-For example, we solve an ice sheet geometry problem (Examples \ref{ex:sia}, \ref{ex:results:sia}) in which the ice flow formulas are only meaningful for surface elevation iterates which exceed the bed elevation.
-(More generally, fluid-layer problems, with a non-negative thickness constraint, often require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation \cite{Bueler2021conservation}.
-However, though a target for future applications of FASCD, such problems are not addressed here.)  Note that non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural modifications of the operator formula in such a fluid-layer context.
-FASCD solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a mesh-independent number of F-cycle iterations.
+For example, we solve an ice sheet geometry problem (Examples \ref{ex:sia}, \ref{ex:results:sia}) in which the ice flow formulas are only meaningful for surface elevation iterates that do not penetrate the bedrock\footnote{More generally, fluid-layer problems, with a non-negative thickness constraint, often require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation \cite{Bueler2021conservation}.}.
+Note that non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural and invasive modifications of the operator formula in such a fluid-layer context.
+FASCD solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, with a mesh-independent number of F-cycle iterations.
 
@@
 In Section \ref{sec:vi} we recall the theory of coercive and monotone variational inequalities, with several example problems.
-Section \ref{sec:cd} reviews the basics of CDs, as proposed by Tai \cite{Tai2003}, including the associated iterations.
+Section \ref{sec:cd} reviews the basics of CDs, as proposed by Tai \cite{Tai2003}, and their induced iterations.
 Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, and then Section \ref{sec:cdmultilevel} presents defect constraints generated via monotone restriction operators \cite{GraeserKornhuber2009}.
@@
 Section \ref{sec:results} gives FASCD numerical results from a parallel Firedrake \cite{Rathgeberetal2016} implementation for the problems introduced in Section \ref{sec:vi}.
-Discussion and outlook in Section \ref{sec:discussion}, and an Appendix on how FASCD reduces to classical FAS multigrid for PDEs, finish the work.
+The work concludes with discussion and outlook in Section \ref{sec:discussion}, and an appendix on how FASCD reduces to classical FAS multigrid for PDEs.
 
@@
 
-Suppose $\cV$ is a real and reflexive Banach space with norm $\|\cdot\|$ and topological dual space $\cV'$.
+Let $\cV$ be a real and reflexive Banach space with norm $\|\cdot\|$ and topological dual space $\cV'$.
 Denote the dual pairing of $\phi \in \cV'$ and $v\in\cV$ by $\ip{\phi}{v} = \phi(v)$, and define the (Banach space) norm on $\cV'$ by $\|\phi\|_{\cV'} = \sup_{\|v\|=1} |\ip{\phi}{v}|$.
 Let $\cK \subset \cV$ be a nonempty, closed, and convex subset, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.
-For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and a linear \emph{source functional} $\ell\in \cV'$ we consider the following \emph{variational inequality} (VI) problem, with solution $u^*\in \cK$ if it exists:
-\begin{equation}
-\ip{f(u^*)}{v-u^*} \ge \ip{\ell}{v-u^*} \qquad \text{for all } v\in \cK. \label{eq:vi}
+For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and a linear \emph{source functional} $\ell\in \cV'$ we consider the following \emph{variational inequality} (VI) problem: find $u^*\in \cK$ such that
+\begin{equation}
+\ip{f(u^*)}{v-u^*} \ge \ip{\ell}{v-u^*} \quad \text{for all } v\in \cK. \label{eq:vi}
 \end{equation}
@@
 In fact, by redefining $f$ we may take $\ell=0$.
-However, allowing $\ell$ clarifies the algorithm in Section \ref{sec:vcycle}.
+However, introducing $\ell$ clarifies the algorithm in Section \ref{sec:vcycle}.
 

### 2023-06-30 paper/fascd.tex
More edits

@@
 
+\newcommand{\dx}{\, \mathrm{d}x}
+\DeclareMathOperator*{\argmin}{arg\,min}
+
 \usepackage{todonotes}
@@
 Let $\cK \subset \cV$ be a nonempty, closed, and convex subset, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.
-For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and a linear \emph{source functional} $\ell\in \cV'$ we consider the following \emph{variational inequality} (VI) problem: find $u^*\in \cK$ such that
-\begin{equation}
-\ip{f(u^*)}{v-u^*} \ge \ip{\ell}{v-u^*} \quad \text{for all } v\in \cK. \label{eq:vi}
+For a continuous, but generally nonlinear, operator $f:\cK \to \cV'$ and a linear \emph{source functional} $\ell\in \cV'$ we consider the following \emph{variational inequality} (VI) problem: find $u^\star\in \cK$ such that
+\begin{equation}
+\ip{f(u^\star)}{v-u^\star} \ge \ip{\ell}{v-u^\star} \quad \text{for all } v\in \cK. \label{eq:vi}
 \end{equation}
@@
 
-VI \eqref{eq:vi} generalizes nonlinear systems of equations $f(u^*)=\ell$ to problems where $u^*$ is also constrained to be in $\cK$.
-Informally, pretending the dual pairing is an inner product, \eqref{eq:vi} says that the angle between $f(u^*)-\ell$ and any arbitrary vector $v-u^*$ pointing from $u^*$ into $\cK$ is at most $90^\circ$; while $f(u^*)-\ell$ may not be zero, it must point directly into $\cK$.
-On the other hand, if $u^*\in\cK^\circ$ is in the interior of the constraint set then \eqref{eq:vi} implies $f(u^*)=\ell$.
+VI \eqref{eq:vi} generalizes nonlinear systems of equations $f(u^\star)=\ell$ to problems where $u^\star$ is also constrained to be in $\cK$.
+Informally, pretending the dual pairing is an inner product, \eqref{eq:vi} says that the angle between $f(u^\star)-\ell$ and any arbitrary vector $v-u^\star$ pointing from $u^\star$ into $\cK$ is at most $90^\circ$; while $f(u^\star)-\ell$ may not be zero, it must point directly into $\cK$.
+On the other hand, if $u^\star$ is in the interior of the constraint set then \eqref{eq:vi} implies $f(u^\star)=\ell$.
+Variational inequalities may generally permit multiple solutions~\cite{Farrell2019}.
 
@@
 
-Certain of the VIs solved in this paper satisfy a stronger inequality than \eqref{eq:coercive}.
+Some of the VIs solved in this paper satisfy a stronger inequality than \eqref{eq:coercive}.
 
@@
 
-\begin{theorem}  \label{thm:viwellposed}  If $f:\cK \to \cV'$ is continuous and $p$-coercive then there exists a unique $u^*\in \cK$ solving VI \eqref{eq:vi}.
+\begin{theorem}  \label{thm:viwellposed}  If $f:\cK \to \cV'$ is continuous and $p$-coercive then there exists a unique $u^\star\in \cK$ solving VI \eqref{eq:vi}.
 \end{theorem}
@@
 
-\begin{example}  \label{ex:plaplacian}  Suppose $p>1$.  For $u,v \in \cV = W^{1,p}_0(\Omega)$ define $f:\cV \to \cV'$ by
-\begin{equation}
-\ip{f(u)}{v} = \int_\Omega |\grad u|^{p-2} \grad u \cdot \grad v\,dx, \label{eq:plaplacian}
+\begin{example}  \label{ex:plaplacian}  Let $p>1$.  For $u,v \in \cV = W^{1,p}_0(\Omega)$, define $f:\cV \to \cV'$ by
+\begin{equation}
+\ip{f(u)}{v} = \int_\Omega |\grad u|^{p-2} \grad u \cdot \grad v\dx, \label{eq:plaplacian}
 \end{equation}
@@
 \begin{equation}
-u^* = \operatorname{arg-min}_{v\in\cK} F(v) - \ip{\ell}{v}. \label{eq:minimization}
+u^\star = \argmin_{v\in\cK} F(v) - \ip{\ell}{v}. \label{eq:minimization}
 \end{equation}
@@
 \begin{equation}
-F(v) = \frac{1}{p} \int_\Omega |\grad v|^p\,dx. \label{eq:plaplacianobjective}
+F(v) = \frac{1}{p} \int_\Omega |\grad v|^p\dx. \label{eq:plaplacianobjective}
 \end{equation}
@@
 \begin{lemma}  \label{lem:advectionskew}  \cite{Elmanetal2014}\,  Suppose $\bX :\Omega \to \RR^d$ is a bounded and boundedly-differentiable vector field on $\Omega$ with zero divergence ($\Div \bX=0$).
-For $u,v \in W^{1,2}(\Omega)$ let $b(u,v) = \int_\Omega (\bX \cdot \grad u) v\,dx$.
-Then $b(u,u) = \frac{1}{2} \int_{\partial \Omega} u^2 \bX\cdot \bn\,dx$ where $\bn$ is the outward normal on $\partial \Omega$.
+For $u,v \in W^{1,2}(\Omega)$ let $b(u,v) = \int_\Omega (\bX \cdot \grad u) v\dx$.
+Then $b(u,u) = \frac{1}{2} \int_{\partial \Omega} u^2 \bX\cdot \bn\dx$ where $\bn$ is the outward normal on $\partial \Omega$.
 \end{lemma}
@@
 \begin{proof}
-Integration by parts gives $b(u,v) = - b(v,u) + \int_{\partial \Omega} uv \bX\cdot \bn\,dx$.
+Integration by parts gives $b(u,v) = - b(v,u) + \int_{\partial \Omega} uv \bX\cdot \bn\dx$.
 \end{proof}
@@
 \end{equation}
-With this operator, consider VI \eqref{eq:vi} for any closed and convex $\cK \subset \cV$ and $g\in\cV'$.
+\pef{Why is $\phi \in L^\infty(\Omega)$? If that is the case, then should we have an $L^2$ inner product in the last term, rather than a dual pairing?}
+With this operator, consider VI \eqref{eq:vi} with $\ell = 0$ for any closed and convex $\cK \subset \cV$.
 The interior condition, i.e.~PDE strong form, is the following linear advection-diffusion equation
@@
 \begin{align*}
-\ip{f(u)-f(v)}{u-v} &= \eps \int_\Omega |\grad u - \grad v|^2\,dx + b(u-v,u-v) \\
-                    &= \eps \int_\Omega |\grad u - \grad v|^2\,dx + \frac{1}{2} \int_{\partial_N\Omega} (u-v)^2 \bX\cdot\bn \ge \eps C \|u-v\|^2.
+\ip{f(u)-f(v)}{u-v} &= \eps \int_\Omega |\grad u - \grad v|^2\dx + b(u-v,u-v) \\
+                    &= \eps \int_\Omega |\grad u - \grad v|^2\dx + \frac{1}{2} \int_{\partial_N\Omega} (u-v)^2 \bX\cdot\bn \ge \eps C \|u-v\|^2.
 \end{align*}
-Thus $f$ is 2-coercive, and so VI problem \eqref{eq:vi} is well-posed.
+Thus $f$ is 2-coercive, and so the associated VI problem is well-posed.
 \end{example}
 
-If $\bX \ne 0$ then $f\ne F'$ in Example \ref{ex:advectiondiffusion}, for any objective $F$, because $\ip{f(u)}{v}$ does not possess the symmetry of the gradient of a scalar: for $u,v$ which are zero on $\partial \Omega$, note that $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.
+If $\bX \ne 0$ then $f\ne F'$ for any objective $F$, because $\ip{f(u)}{v}$ does not possess the symmetry of the gradient of a scalar: for $u,v$ which are zero on $\partial \Omega$, note that $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.
 References \cite{Bueler2021conservation,ChangNakshatrala2017} consider inequality-constrained advection-diffusion VI problems like this, specifically over the set $\cK = \{v\ge 0\}$.
 
-The next VI example, nonlinear and not of optimization type, is a model for steady ice sheet geometry in a given climate.
-Related VI problems arise for any fluid layer which is subject to a source term which can remove the fluid \cite{Bueler2021conservation}.
+The next VI example is a model for steady ice sheet geometry in a given climate.
+It is nonlinear and not of optimization type.
+Related VI problems arise for any fluid layer which is subject to processes that can remove the fluid \cite{Bueler2021conservation}.
 
@@
 Let $a \in L^\infty(\Omega)$ denote a given ``surface mass balance'' function, the annually-averaged rate of ice accumulation (as snow) minus melt and runoff.
-Then the surface elevation $s\in C(\Omega)$ of a steady-state, isothermal, and non-sliding ice sheet (glacier) approximately satisfies the \emph{shallow ice approximation} (SIA) \cite{GreveBlatter2009} problem
-\begin{equation}
-\int_\Omega \Gamma (s-b)^{n+2} |\grad s|^{n-1} \grad s \cdot \grad (v-s) \,dx \ge \int_\Omega a (v-s)\,dx, \label{eq:siavi}
-\end{equation}
-over the set $\mathcal{K} = \{s\ge b\}$ of admissible surface elevations, where $n>1$ and $\Gamma>0$ are constants determined by the physical properties of ice; $n=3$ is a typical value.
-The doubly-nonlinear and doubly-degenerate operator in this case is
-\begin{equation}
-\ip{f(s)}{v} = \int_\Omega \Gamma (s-b)^{n+2} |\grad s|^{n-1} \grad s \cdot \grad v\,dx. \label{eq:sia}
+Then the surface elevation $s\in C(\Omega)$ of a steady-state, isothermal, and non-sliding ice sheet (glacier) approximately satisfies the \emph{shallow ice approximation} (SIA) \cite{GreveBlatter2009} problem: find $s \in \mathcal{K} = \{s\ge b\}$ such that
+\begin{equation}
+\int_\Omega \Gamma (s-b)^{n+2} |\grad s|^{n-1} \grad s \cdot \grad (v-s) \dx \ge \int_\Omega a (v-s)\dx \quad \text{ for all } v \in \mathcal{K}, \label{eq:siavi}
+\end{equation}
+where $n>1$ and $\Gamma>0$ are constants determined by the physical properties of ice; $n=3$ is a typical value.  The doubly-nonlinear and doubly-degenerate operator in this case is
+\begin{equation}
+\ip{f(s)}{v} = \int_\Omega \Gamma (s-b)^{n+2} |\grad s|^{n-1} \grad s \cdot \grad v\dx. \label{eq:sia}
 \end{equation}
@@
 The theory suggests that a power of the solution lives in a Sobolev space, namely $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.
-This theory, and also observations of ice sheets, shows that $|\grad s|$ is generally unbounded as one approaches the free boundary, the \emph{ice margin}, from the icy side (where $s>b$).
+\pef{`suggests' is quite weak; can we use a stronger word?} This theory, and also observations of ice sheets, shows that $|\grad s|$ is generally unbounded as one approaches the free boundary, the \emph{ice margin}, from the icy side (where $s>b$).
 \end{example}
@@
 Numerical results are shown in Section \ref{sec:results} for applications of FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} to these Examples.
+
+\pef{I'm not sure starting with the example problems we want to solve is the right move. Let me read more and think about it.}
 
@@
 
-\begin{example}  \label{ex:frequencydecomposition}  For simplicity suppose $\Omega = (0,a)^d \subset \RR^d$ is a cube, and let $\cV = W_{\text{per}}^{2,k}(\Omega)$ be the periodic functions with square-integrable $k$th derivatives.
+\begin{example}  \label{ex:frequencydecomposition}  For simplicity suppose $\Omega = (0,a)^d \subset \RR^d$ is a cube, and let $\cV = W_{\text{per}}^{2,k}(\Omega)$ be the periodic functions with square-integrable $k$th derivatives for $k \ge 0$.
 Suppose $\psi \in W_{\text{per}}^{2,k}(\Omega)$ and let $\cK = \{v \ge \psi\} \subset \cV$.
@@
 Let $0\le \alpha \le 1$ be the \emph{damping parameter}.\footnote{FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} are presented without damping, and indeed it seems to be unnecessary.
-However, the general convergence of the basic CD iterations appears to depend upon it.}  Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below, discussed next, solve smaller VI problems over each subset $\cK_i$.
-Each Algorithm replaces $u \in \cK$ with a new iterate $w\in\cK$ which should be closer to $u^* \in \cK$ solving \eqref{eq:vi}.
-
-The additive/parallel \pr{cd-add} Algorithm \ref{alg:basiccd-add} generalizes the classical Jacobi iteration.  Here ``\textbf{for all}'' can be computed in any order.  In its VI,
-\begin{equation}
-\ip{f(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \qquad \forall v_i \in \mathcal{K}_i, \label{eq:cdaddvi}
-\end{equation}
-the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$, and then adds-back the subset solution $\hat w_i \in \mathcal{K}_i$.
-
+However, the convergence of the more general CD iterations appears to depend upon it.}  Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below, discussed next, solve smaller VI problems over each subset $\cK_i$.
+Each Algorithm replaces $u \in \cK$ with a new iterate $w\in\cK$ which should be closer to $u^\star \in \cK$ solving \eqref{eq:vi}.
+
+The additive/parallel \pr{cd-add} Algorithm \ref{alg:basiccd-add} generalizes the classical Jacobi iteration.
 \begin{pseudofloat}[H]
@@
     for all $i \in \{0,\dots,m-1\}$: \\+
-        \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$, \\+
-            $\ip{f(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}$ \\--
+        \rm{find} $\hat w_i\in \cK_i$ \rm{such that} \\+
+            $\ip{f(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \textnormal{ for all } v_i \in \cK_i$ \\--
     $\hat w = \sum_i \hat w_i\in\cK$ \\
@@
 \end{pseudofloat}
-
-The multiplicative/successive \pr{cd-mult} Algorithm \ref{alg:basiccd-mult} generalizes the Gauss-Seidel iteration.
-Each subset solution corrects the global iterate immediately, so here the ordering used in the ``\textbf{for}'' loop is important.
-In the VI,
-\begin{equation}
-\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \qquad \forall v_i \in \mathcal{K}_i, \label{eq:cdmultvi}
-\end{equation}
-the sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``those portions of $u$ which have not yet improved''.
-
-\begin{pseudofloat}[H]
+Here ``\textbf{for all}'' can be computed in any order.  In its VI,
+\begin{equation}
+\ip{f(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \textnormal{ for all } v_i \in \mathcal{K}_i, \label{eq:cdaddvi}
+\end{equation}
+the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$, and then adds-back the subset solution $\hat w_i \in \mathcal{K}_i$.
+\pef{Could remove this explanation, but it's nice to keep it in if we can fit it}
+
+The multiplicative/successive \pr{cd-mult} Algorithm \ref{alg:basiccd-mult} generalizes the Gauss--Seidel iteration.  \begin{pseudofloat}[H]
 \begin{pseudo*}
@@
     for $i = 0,\dots,m-1$: \\+
-        \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$, \\+
-            $\displaystyle \ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}$ \\-
+        \rm{find} $\hat w_i\in \cK_i$ \rm{such that} \\+
+            $\displaystyle \ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \textnormal{ for all } v_i \in \cK_i$ \\-
             $w_i = (1-\alpha) \Pi_i u + \alpha \hat w_i\in\cK_i$ \\-
@@
 \end{pseudofloat}
-
-In VIs \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} the argument of $f$ is an element of $\cK$, as the reader should confirm.
+Each subset solution corrects the global iterate immediately, so here the ordering used in the ``\textbf{for}'' loop is important.  In the VI,
+\begin{equation}
+\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \qquad \forall v_i \in \mathcal{K}_i, \label{eq:cdmultvi}
+\end{equation}
+the sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``those portions of $u$ which have not yet improved''.
+
+In VIs \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} the argument of $f$ is an element of $\cK$.
 Furthermore, by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write the test vector $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors from $\cK$.
@@
 \end{align}
-In this sense the VIs over the subsets $\mathcal{K}_i$ solve the same problem as \eqref{eq:vi}.
+In this sense the VIs over the subsets $\mathcal{K}_i$ solve the same problem as the original one.
 
@@
 We call \pr{cd-add} and \pr{cd-mult} the \emph{basic CD iterations}.
-VI problems \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} are analogous to constrained line searches, and in fact they reduce to such line searches when the $\cV_i$ are one-dimensional.
+VI problems \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} reduce to constrained line searches when the $\cV_i$ are one-dimensional.
 As shown in the next Lemma, when VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization} then, subject to bounds on the damping $\alpha$, these iterations generate monotonically-decreasing objective values.
@@
 
-To show that each iteration of the \textbf{for} loop in \pr{cd-mult} lowers the objective value, let $z_{i} = \sum_{j\le i} w_j + \sum_{j > i} \Pi_j u \in \cK$, $i=0,\dots,m-1$; this is the state at the end of the $i$th iteration.
+To show that each iteration of the {for} loop in \pr{cd-mult} lowers the objective value, let $z_{i} = \sum_{j\le i} w_j + \sum_{j > i} \Pi_j u \in \cK$, $i=0,\dots,m-1$; this is the state at the end of the $i$th iteration.
 Note that $z_{i-1} = \sum_{j < i} w_j + \Pi_i u + \sum_{j > i} \Pi_j u$ for $i>0$.
@@
 Recalling abstract VI problem \eqref{eq:vi}, from now on we only constraint sets \eqref{eq:originalconstraintset}.
-Given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we will assume that the following is well-posed for $u^*\in \cK$:
-\begin{equation}
-\ip{f(u^*)}{v-u^*} \ge \ip{\ell}{v-u^*} \qquad \text{for all } v\in \cK. \label{eq:boxdirichletvi}
-\end{equation}
-Note that $v-u^* \in W_0^{1,p}(\Omega)$.
+Given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we will assume that the following is well-posed for $u^\star\in \cK$:
+\begin{equation}
+\ip{f(u^\star)}{v-u^\star} \ge \ip{\ell}{v-u^\star} \qquad \text{for all } v\in \cK. \label{eq:boxdirichletvi}
+\end{equation}
+Note that $v-u^\star \in W_0^{1,p}(\Omega)$.
 

### 2023-06-30 paper/fascd.tex
Some edits and removals; we don't appear to need the proof of Lemma 3.3, but I've just commented it out for now

@@
 
-\begin{example}  \label{ex:advectiondiffusion}  Suppose $\partial\Omega$ is partitioned into Dirichlet and Neumann portions, i.e.~$\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$, with $\partial_D\Omega$ of positive measure.
+\begin{example}  \label{ex:advectiondiffusion}  Suppose $\partial\Omega$ is partitioned into Dirichlet and Neumann portions, i.e.~$\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$, with $\partial_D\Omega$ of positive Hausdorff measure.
 Let $\cV \subset W^{1,2}(\Omega)$ be the space of functions which are zero along $\partial_D\Omega$.
@@
 
-In VIs \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} the argument of $f$ is an element of $\cK$.
+%In VIs \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} the argument of $f$ is an element of $\cK$.
 Furthermore, by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write the test vector $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors from $\cK$.
 For \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} respectively:
-\begin{align}
-[u - \Pi_i u + v_i] - [u - \Pi_i u + \hat w_i] &= v_i - \hat w_i, \label{eq:admissibledifferenceadd} \\
-\left[\sum_{j<i} w_j + v_i + \sum_{j>i} \Pi_j u\right] - \left[\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\right] &= v_i - \hat w_i.  \label{eq:admissibledifferencemult}
-\end{align}
-In this sense the VIs over the subsets $\mathcal{K}_i$ solve the same problem as the original one.
+%\begin{align}
+%[u - \Pi_i u + v_i] - [u - \Pi_i u + \hat w_i] &= v_i - \hat w_i, \label{eq:admissibledifferenceadd} \\
+%\left[\sum_{j<i} w_j + v_i + \sum_{j>i} \Pi_j u\right] - \left[\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\right] &= v_i - \hat w_i.  \label{eq:admissibledifferencemult}
+%\end{align}
+%In this sense the VIs over the subsets $\mathcal{K}_i$ solve the same problem as the original one.
 
@@
 As shown in the next Lemma, when VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization} then, subject to bounds on the damping $\alpha$, these iterations generate monotonically-decreasing objective values.
-
 \begin{lemma} \cite{Tai2003}  Suppose $f=F'$ for $F:\cK\to\RR$ convex and differentiable, and let $u\in\cK$.  For $0 \le \alpha \le 1$ the output $w \in \cK$ from \pr{cd-mult} satisfies
@@
 
-\begin{proof}  We assume $\ell=0$ without loss of generality; otherwise define $\tilde F(u)=F(u) - \ip{\ell}{u}$ and observe that the proof goes through for $\tilde F$ essentially unchanged.
-
-To show that each iteration of the {for} loop in \pr{cd-mult} lowers the objective value, let $z_{i} = \sum_{j\le i} w_j + \sum_{j > i} \Pi_j u \in \cK$, $i=0,\dots,m-1$; this is the state at the end of the $i$th iteration.
+%\begin{proof}  We assume $\ell=0$ without loss of generality; otherwise define $\tilde F(u)=F(u) - \ip{\ell}{u}$ and observe that the proof goes through for $\tilde F$ essentially unchanged.
+%
+%To show that each iteration of the {for} loop in \pr{cd-mult} lowers the objective value, let $z_{i} = \sum_{j\le i} w_j + \sum_{j > i} \Pi_j u \in \cK$, $i=0,\dots,m-1$; this is the state at the end of the $i$th iteration.
 Note that $z_{i-1} = \sum_{j < i} w_j + \Pi_i u + \sum_{j > i} \Pi_j u$ for $i>0$.
 Therefore by convexity of $F$, observation \eqref{eq:admissibledifferencemult} with $v_i=\Pi_i u$, and \eqref{eq:cdmultvi} we find:
-\begin{equation*}
-F(z_{i-1}) - F\left(\sum_{j<i} w_j + \hat w_i + \sum_{j > i} \Pi_j u\right) \ge \ip{f\left(\sum_{j<i} w_j + \hat w_i + \sum_{j > i} \Pi_j u\right)}{\Pi_i u - \hat w_i} \ge 0.
-\end{equation*}
-Using convexity (again) now gives
-\begin{align*}
-F(z_i) &= F\left((1-\alpha) z_{i-1} + \alpha \left(\sum_{j<i} w_j + \hat w_i + \sum_{j > i} \Pi_j u\right)\right) \\
-&\le  (1-\alpha) F(z_{i-1}) + \alpha F\left(\sum_{j<i} w_j + \hat w_i + \sum_{j > i} \Pi_j u\right) \le F(z_{i-1}).
-\end{align*}
-As $u=z_0$ and $w = z_{m-1}$, the result for \pr{cd-mult} follows.
-
-Suppose $w=u+\alpha\sum_i e_i$ is computed by \pr{cd-add} using $\alpha \le 1/m$.
+%\begin{equation*}
+%F(z_{i-1}) - F\left(\sum_{j<i} w_j + \hat w_i + \sum_{j > i} \Pi_j u\right) \ge \ip{f\left(\sum_{j<i} w_j + \hat w_i + \sum_{j > i} \Pi_j u\right)}{\Pi_i u - \hat w_i} \ge 0.
+%\end{equation*}
+%Using convexity (again) now gives
+%\begin{align*}
+%F(z_i) &= F\left((1-\alpha) z_{i-1} + \alpha \left(\sum_{j<i} w_j + \hat w_i + \sum_{j > i} \Pi_j u\right)\right) \\
+%&\le  (1-\alpha) F(z_{i-1}) + \alpha F\left(\sum_{j<i} w_j + \hat w_i + \sum_{j > i} \Pi_j u\right) \le F(z_{i-1}).
+%\end{align*}
+%As $u=z_0$ and $w = z_{m-1}$, the result for \pr{cd-mult} follows.
+%
+%Suppose $w=u+\alpha\sum_i e_i$ is computed by \pr{cd-add} using $\alpha \le 1/m$.
 Considering each $e_i=\hat w_i - \Pi_i u$ separately, convexity of $F$ and VI \eqref{eq:cdaddvi} with $v_i=\Pi_i u$ yields
-\begin{equation}
-F(u) - F(u+e_i) \ge \ip{f(u+e_i)}{u-(u+e_i)} = \ip{f(u-\Pi_i u + \hat w_i)}{\Pi_i u - \hat w_i} \ge 0. \label{eq:cdaddonereduction}
-\end{equation}
-Let $v = u+ \frac{1}{m} \sum_i e_i$.  Again using convexity (Jensen's inequality), and \eqref{eq:cdaddonereduction}, we have
-\begin{equation*}
-F(v) = F\left(\frac{1}{m} \sum_{i=0}^{m-1} (u+e_i)\right) \le \frac{1}{m} \sum_{i=0}^{m-1} F(u+e_i) \le \frac{1}{m} \sum_{i=0}^{m-1} F(u) = F(u).
-\end{equation*}
-Now define $\lambda = m\alpha$ so $0 \le \lambda \le 1$ and $w = (1-\lambda) u + \lambda v$.  Thus by convexity and the above result, $F(w) \le (1-\lambda) F(u) + \lambda F(v) \le F(u)$.
-\end{proof}
-
-When $\alpha=1/m$ then the \pr{cd-add} proof can be simplified because $v=w$ and $\lambda=1$.
+%\begin{equation}
+%F(u) - F(u+e_i) \ge \ip{f(u+e_i)}{u-(u+e_i)} = \ip{f(u-\Pi_i u + \hat w_i)}{\Pi_i u - \hat w_i} \ge 0. \label{eq:cdaddonereduction}
+%\end{equation}
+%Let $v = u+ \frac{1}{m} \sum_i e_i$.  Again using convexity (Jensen's inequality), and \eqref{eq:cdaddonereduction}, we have
+%\begin{equation*}
+%F(v) = F\left(\frac{1}{m} \sum_{i=0}^{m-1} (u+e_i)\right) \le \frac{1}{m} \sum_{i=0}^{m-1} F(u+e_i) \le \frac{1}{m} \sum_{i=0}^{m-1} F(u) = F(u).
+%\end{equation*}
+%Now define $\lambda = m\alpha$ so $0 \le \lambda \le 1$ and $w = (1-\lambda) u + \lambda v$.  Thus by convexity and the above result, $F(w) \le (1-\lambda) F(u) + \lambda F(v) \le F(u)$.
+%\end{proof}
+%When $\alpha=1/m$ then the \pr{cd-add} proof can be simplified because $v=w$ and $\lambda=1$.
 In this case $F(w)$ is less than $\frac{1}{m} \sum_i F(u+e_i)$ by Jensen's inequality, which is lower than $F(u)$ by \eqref{eq:cdaddvi} or \eqref{eq:cdmultvi}, which gives inequality \eqref{eq:cdaddonereduction}.
@@
 However, practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.
-References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem, wherein $f=F'$ is linear, local, and defined for all inputs (on all of $\mathcal{V}$).
+References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem, wherein $f=F'$ is linear, local, and defined on all of $\mathcal{V}$.
 Substantial implementation simplifications are available in such cases.
 
-\begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear and a differential operator, thus local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi$.
+\begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear and is local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi$.
 Considering \pr{cd-add} for concreteness, by linearity \eqref{eq:cdaddvi} can be rewritten
@@
 Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible---generally $e_i \notin \cK$---but \eqref{eq:linearlocalvi} is meaningful for this $f$.
+\pef{Can you expand a little bit here? I confess I don't follow.}
 \end{example}
@@
 
-From now on we will restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.
+From now on we restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.
 Suppose $\Omega \subset \RR^d$ is open and let $\mathcal{V}=W^{1,p}(\Omega)$.
 The \emph{lower obstacle} $\underline{\gamma}$ and the \emph{upper obstacle} $\overline{\gamma}$ are given functions, defined on the closure $\overline{\Omega}$, with values in the extended real line $\tilde\RR = [-\infty,+\infty]$.
-We suppose $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$ and $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$ are measurable, $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$, and $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$, and that $\partial\Omega$ is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive measure, to facilitate well-posedness \cite{Evans2010}, and $\partial_N \Omega$ possibly empty.
+We suppose $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$ and $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$ are measurable, $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$, and $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$, and that $\partial\Omega$ is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive Hausdorff measure, and $\partial_N \Omega$ possibly empty.
 Dirichlet data\footnote{Dirichlet boundary values are in the trace sense \cite{Evans2010}, and $g_D$ is assumed to be as regular as needed.} $g_D:\partial_D \Omega \to \RR$ then defines
@@
 the closed and convex \emph{constraint set}.
-Note that we are treating inequality constraints and boundary conditions in a unified manner.\footnote{Any Neumann conditions over $\partial_N \Omega$ are determined by the operator formula for $f$, i.e.~by an integration over $\partial_N\Omega$.}
-
-Recalling abstract VI problem \eqref{eq:vi}, from now on we only constraint sets \eqref{eq:originalconstraintset}.
+Note that we are treating inequality constraints and boundary conditions in a unified manner.\footnote{Any Neumann conditions over $\partial_N \Omega$ are imposed weakly in the definition of $f$, i.e.~by an integration over $\partial_N\Omega$.}
+
+We recall the abstract VI problem \eqref{eq:vi} with constraint sets \eqref{eq:originalconstraintset}.
 Given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we will assume that the following is well-posed for $u^\star\in \cK$:

### 2023-07-01 paper/results/drape.sh
scripts and text output moved from https://bitbucket.org/pefarrell/fascd/src/master/

@@ -1,0 +1,19 @@
+# generate results for Example 9.1 in paper
+
+for PROB in ball spiral; do
+    echo ""
+    echo "******** -nofas PROB=${PROB} *******"
+    for LEVS in 1 2 3 4 5 6 7 8; do
+        python3 drape.py -prob $PROB -levs $LEVS -nofas
+    done
+done
+
+for CYCLE in multiplicative full; do
+    for PROB in ball spiral; do
+        echo ""
+        echo "******** PROB=${PROB} CYCLE=${CYCLE} *******"
+        for LEVS in 1 2 3 4 5 6 7 8 9 10; do
+            python3 drape.py -prob $PROB -levs $LEVS -fascd_cycle_type $CYCLE
+        done
+    done
+done

### 2023-07-01 paper/results/plap1d.m

@@ -1,0 +1,15 @@
+% generate convergence rates from results in plap1d.txt
+
+h = 2.^(-(0:9));
+fasterr = [2.263e-01 3.255e-02 9.108e-03 3.248e-03 5.505e-04 ...
+           1.690e-04 4.717e-05 9.522e-06 3.591e-06 4.948e-07];
+degerr = [8.700e-02 3.812e-02 1.589e-02 6.318e-03 2.330e-03 ...
+          6.992e-04 2.047e-04 1.078e-04 3.997e-05 1.506e-05];
+loglog(h,fasterr,'*',h,degerr,'o')
+grid on,  xlabel h,  ylabel('|error|_{inf}')
+legend('p=1.5','p=4.0')
+fastp = polyfit(log(h),log(fasterr),1)
+degp = polyfit(log(h),log(degerr),1)
+
+%   fastp = 2.0047  -1.7579
+%   degp  = 1.4203  -2.2802

### 2023-07-01 paper/results/plap1d.sh

@@ -1,0 +1,37 @@
+# generate results for Example 9.2 in paper
+# run as
+#   $ bash PATHTOHERE/plap1d.sh &> plap1d.txt
+# from git@bitbucket.org:pefarrell/fascd.gitexamples/
+# for convergence rates, run plap1d.m in octave
+
+LLIST="1 2 3 4 5 6 7 8 9 10"
+
+for VOPT in "-fascd_skip_upsmooth" "-fascd_skip_downsmooth" ""; do
+    echo "******** p = 1.5 V-cycle VOPT=${VOPT} *******"
+    for LEVS in $LLIST; do
+        python3 plap1d.py -p 1.5 $VOPT -levs $LEVS
+    done
+    echo ""
+done
+
+for EPSREG in 0.0 1.0e-8; do
+    echo "******** p = 1.5 F-cycle EPSREG=${EPSREG} *******"
+    for LEVS in $LLIST; do
+        python3 plap1d.py -fascd_cycle_type full -p 1.5 -epsreg $EPSREG -levs $LEVS
+    done
+    echo ""
+done
+
+for ITS in 3 4; do
+    echo "******** p = 4.0 V-cycle ITS=${ITS} *******"
+    for LEVS in $LLIST; do
+        python3 plap1d.py -p 4.0 -fascd_levels_snes_max_it $ITS -levs $LEVS
+    done
+    echo ""
+done
+
+echo "******** p = 4.0 F-cycle *******"
+for LEVS in $LLIST; do
+    python3 plap1d.py -fascd_cycle_type full -p 4.0 -levs $LEVS
+done
+echo ""

### 2023-07-01 paper/results/pollutant.sh

@@ -1,0 +1,21 @@
+# generate results for Example 9.3 in paper
+# run as
+#   $ bash PATHTOHERE/pollutant.sh &> pollutant.txt
+# from git@bitbucket.org:pefarrell/fascd.git in examples/
+
+LIST2D="1 2 3 4 5 6 7 8 9"
+LIST3D="1 2 3 4 5"  # level 5 almost fills 128GB memory machine
+
+echo "******** -dim 2 *******"
+for LEVS in $LIST2D; do
+    python3 pollutant.py -dim 2 -levs $LEVS
+done
+echo ""
+
+# which mpiexec
+echo "******** -dim 3; P = 16 parallel *******"
+for LEVS in $LIST3D; do
+    mpiexec -n 16 --map-by core --bind-to hwthread python3 pollutant.py -dim 3 -levs $LEVS
+done
+echo ""
+

### 2023-07-01 paper/results/siaweak.sh

@@ -1,0 +1,6 @@
+# generate weak-scaling results for SIA Example 9.4 in paper
+
+OPTS="-bumps -mcoarse 20 -fascd_cycle_type full -fascd_monitor -fascd_rtol 2.0e-4 -fascd_atol 1.0e-8 -quad"
+time mpiexec -n  1 --map-by core --bind-to hwthread python3 sia.py $OPTS -levs 7
+time mpiexec -n  4 --map-by core --bind-to hwthread python3 sia.py $OPTS -levs 8
+time mpiexec -n 16 --map-by core --bind-to hwthread python3 sia.py $OPTS -levs 9

### 2023-07-01 paper/results/drape.sh
update results from Q1 on 3D pollutant.py

@@ -1,4 +1,4 @@
-# generate results for Example 9.1 in paper
+# generate results for Example 8.1 in paper
 
 for PROB in ball spiral; do
     echo ""

### 2023-07-01 paper/results/plap1d.sh

@@ -1,4 +1,4 @@
-# generate results for Example 9.2 in paper
+# generate results for Example 8.2 in paper
 # run as
 #   $ bash PATHTOHERE/plap1d.sh &> plap1d.txt
 # from git@bitbucket.org:pefarrell/fascd.gitexamples/

### 2023-07-01 paper/results/pollutant.sh

@@ -1,4 +1,4 @@
-# generate results for Example 9.3 in paper
+# generate results for Example 8.3 in paper
 # run as
 #   $ bash PATHTOHERE/pollutant.sh &> pollutant.txt
 # from git@bitbucket.org:pefarrell/fascd.git in examples/
@@ -6,16 +6,16 @@
 LIST2D="1 2 3 4 5 6 7 8 9"
 LIST3D="1 2 3 4 5"  # level 5 almost fills 128GB memory machine
 
-echo "******** -dim 2 *******"
+echo "******** -dim 2: serial, P1 *******"
 for LEVS in $LIST2D; do
     python3 pollutant.py -dim 2 -levs $LEVS
 done
 echo ""
 
 # which mpiexec
-echo "******** -dim 3; P = 16 parallel *******"
+echo "******** -dim 3: P = 16 parallel, Q1 *******"
 for LEVS in $LIST3D; do
-    mpiexec -n 16 --map-by core --bind-to hwthread python3 pollutant.py -dim 3 -levs $LEVS
+    mpiexec -n 16 --map-by core --bind-to hwthread python3 pollutant.py -dim 3 -q1 -levs $LEVS
 done
 echo ""
 

### 2023-07-01 paper/results/siaweak.sh

@@ -1,4 +1,4 @@
-# generate weak-scaling results for SIA Example 9.4 in paper
+# generate weak-scaling results for SIA Example 8.4 in paper
 
 OPTS="-bumps -mcoarse 20 -fascd_cycle_type full -fascd_monitor -fascd_rtol 2.0e-4 -fascd_atol 1.0e-8 -quad"
 time mpiexec -n  1 --map-by core --bind-to hwthread python3 sia.py $OPTS -levs 7

### 2023-07-02 paper/fascd.tex
update Results section: Q1 results, better text

@@
 -\eps \grad^2 u + \bX\cdot \grad u = \phi.
-\label{eq:pdeadvectiondiffusion}
+\label{eq:advectiondiffusionstrong}
 \end{equation}
@@
 
-Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}, with $P_1$ elements and standard, global mesh refinement.
-Here we show numerical results for the several VI Examples in Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
-Our essential observations are that each V-cycle dramatically reduces the residual norm, and that a few F-cycles solves most Examples down to discretization error.
+Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}, allowing $P_1$ or $Q_1$ elements and based on standard, global mesh refinement.
+Here we show numerical results for the several example problems in Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
+Our essential observations are that each V-cycle dramatically reduces the residual norm, and that very few F-cycles solve most examples to within discretization error.
 
@@
 In the ``spiral'' problem from Example 7.1.1 in \cite{GraeserKornhuber2009}, $\psi$ is constructed to produce a more interesting coincidence set; see Figure \ref{fig:results:classical}.
-Note that $\psi\big|_{\partial \Omega} \le 0$ in both problems.
 
@@
 
-Table \ref{tab:results:classical} shows the resulting iteration counts for the various finest meshes of $m_J$ nodes.
+Table \ref{tab:results:classical} shows the resulting iteration counts for finest meshes of $m_J$ nodes.
 When the RS Newton smoother is applied by itself as a solver (``RS only'') it shows rapidly-increasing iterations as expected, roughly proportional to resolution.
@@
 Noting that for F-cycles we are counting the number of V-cycles after the ramp (Section \ref{sec:implementation}), we see that a constant number of cycles solves the problem.
-The spiral problem appears to be only slightly more challenging for these solvers.
-The complexity of the coincidence set or free boundary does not strongly-influence FASCD performance.
-
-Each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+IC) iterations.
-For the F-cycle results, where three iterations suffice at all resolutions in either problem, this is close to textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  However, work of at most ten residual evaluations is often mentioned.} despite the low solution regularity.
+Comparing the ball and spiral results, the complexity of the coincidence set or free boundary does not strongly-influence FASCD performance.
+Note that each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+IC) iterations.
+Thus for the F-cycle results, where three iterations suffice at all resolutions, we see close to textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  Work of less than ten residual evaluations is often mentioned as the standard.} despite the low solution regularity.
 \end{example}
@@
 \end{center}
-\caption{Coincidence sets (black) for the ball and spiral problems in Example \ref{ex:results:classical}.}
+\caption{Coincidence sets for the ball and spiral problems in Example \ref{ex:results:classical}.}
 \label{fig:results:classical}
@@
 
-% see git@bitbucket.org:pefarrell/fascd.git: examples/drape.py and edplay/results/drape.sh|txt
+% see git@bitbucket.org:pefarrell/fascd.git: examples/drape.py
+% and https://github.com/bueler/mcd-extended: paper/results/drape.sh|txt
 \begin{table}[ht]
@@
 Recall Example \ref{ex:plaplacian}, the nonlinear $p$-Laplacian operator.
-Here we consider a unilateral obstacle problem, in one dimension for visualizability, for both fast ($p=1.5$) and degenerate ($p=4.0$) diffusivity cases.\footnote{``Fast'' and ``degenerate'' refer to $D=|\grad u|^{p-2}\to \infty$ and $D\to 0$ as $|\grad u|\to 0$, respectively.
-Ellipticity is lost in the latter case.
-For such $p$-Laplacian obstacle problems, well-posedness and ($p$-dependent) solution regularity are considered in \cite{ChoeLewis1991}.}  This example reveals the sometimes strong contrast between $V(1,0)$ and $V(0,1)$ results, namely \pr{up} $=0$ and \pr{down} $=0$ in Algorithm \ref{alg:fascd}, and it emphasizes the need for robust smoothers.
+Here we consider a unilateral obstacle problem, in one dimension for visualizability, for both fast ($p=1.5$) and degenerate ($p=4.0$) diffusivity cases.\footnote{``Fast'' and ``degenerate'' refer to $D=|\grad u|^{p-2}\to \infty$ and $D\to 0$ as $|\grad u|\to 0$, respectively; ellipticity is lost in the latter case.
+Well-posedness and ($p$-dependent) solution regularity for $p$-Laplacian obstacle problems is considered in \cite{ChoeLewis1991}.}  This example reveals the sometimes strong contrast between $V(1,0)$ and $V(0,1)$ results (\pr{up}, \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively), and it emphasizes the need for robust smoothers.
 
@@
 The default smoother uses three (fixed) iterations of RS Newton, with LU solution of the step equations.\footnote{This is an optimal $O(m_j)$ smoother in 1D.}  The coarse solver is the same but iterated to convergence.
-The initial iterate was $w=\psi$, and again \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$.
+The initial iterate was $w=\psi$, and \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$.
+Figure \ref{fig:imagesvcycle} illustrates the major steps of an example V-cycle, showing each of the coarse-correction VI subproblems and solutions.
 
@@
 
-Figure \ref{fig:imagesvcycle} illustrates the major steps of an example V-cycle, showing each of the nonlinear FAS coarse correction VI subproblem solutions.
-That is, each $y^j \in \mathcal{D}^j$ and $z^j \in \mathcal{U}^j$ is drawn, along with each level defect constraint $\underline{\phi}^j,\underline{\chi}^j$.
-
 % figures in the following visualization were generated using branch elbueler/simplefigs
@@
 Solutions (i.e.~$w^j$, $y^j$, $z^j$) are lines while obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots.
-The zero function (dotted) is the admissible initial iterate for $y^j$ and $z^j$.
-Vertical scale varies.}
+The zero function is shown (dotted), but vertical scale varies.}
 \label{fig:imagesvcycle}
@@
 We see that $V(1,0)$ cycles, essentially Algorithm 4.7 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.
-By contrast, $V(0,1)$ cycles, using only up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are just as good as the default $V(1,1)$ cycles (i.e.~Algorithm \ref{alg:fascd}).
-F-cycles are even more efficient except when they fail entirely, but the failures arise from singular diffusivity.
+By contrast, $V(0,1)$ cycles, using only up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are just as good as the default $V(1,1)$ cycles ($=$ Algorithm \ref{alg:fascd}).
+F-cycles are even more efficient except when they fail entirely, but these failures arise from singular diffusivity.
 If we regularize  $D=|\grad u|^{-0.5}$ to $D_\eps = \left(\eps + |\grad u|^2\right)^{-0.25}$, with $\eps=10^{-8}$, then a single F-cycle suffices at every resolution (column ``F$_\eps$'').
@@
 However, either a slight increase in smoother power, to four RS Newton iterations, or improved finest-level initial iterates through use of an F-cycle, stops the instability.
-Even for the three-iteration solver, a single F-cycle suffices at every resolution.
+Even for the three-iteration smoother, a single F-cycle suffices at every resolution.
 
@@
 
-% see git@bitbucket.org:pefarrell/fascd.git: examples/plap1d.py and edplay/results/plap1d.sh|txt
+% see git@bitbucket.org:pefarrell/fascd.git: examples/plap1d.py
+% and https://github.com/bueler/mcd-extended: paper/results/plap1d.sh|txt
 \begin{table}[ht]
@@
 Solvers implemented in Firedrake can exploit parallel solvers from the PETSc library \cite{Balayetal2023}.
-In fact, the above Examples, which were run in serial so as to fix the smoother,\footnote{The action of parallel preconditioners generally depends on the number of processes \cite[for example]{Bueler2021}.} also run efficiently in parallel.
-The next Example applies a parallel solver, and Example \ref{ex:results:sia} also demonstrates the parallel weak scaling \cite{Bueler2021} of FASCD.
+The above Examples, run in serial so as to fix the smoother,\footnote{The action of parallel preconditioners generally depends on the number of processes \cite[for example]{Bueler2021}.} also run efficiently in parallel.
+The next Example applies a parallel solver, and Example \ref{ex:results:sia} demonstrates the parallel weak scaling \cite{Bueler2021} of FASCD.
 
 \begin{example}  \label{ex:results:advdiff}
-The advection-diffusion operator in Example \ref{ex:advectiondiffusion} is coercive when the velocity $\bX$ is divergence-free.
-We now demonstrate FASCD F-cycles on such a VI problem over $\Omega=(-1,1)^d$ ($d=2,3$).
-The 2D problem satisfies the coercivity hypotheses ($\bX = (7+5y,-5x)$ is divergence-free), but the 3D problem has a more general velocity ($\grad\cdot\bX \ne 0$); note $u=0$ on all of $\partial \Omega$ in both problems.
-The discontinuous source term $\phi \in L^\infty(\Omega)$ has $\phi>0$ patches in half of the domain and $\phi<0$ values in the other half.
-While the other Examples in this Section are unilateral, here we constrained the solution with upper and lower obstacles: $0 \le u \le 1$.
-The $u=0$ and $u=1$ coincidence sets for the 2D problem are shown in Figure \ref{fig:results:advdiff}; the 3D sets are comparable in regularity.
+We applied FASCD F-cycles to the advection-diffusion VI problem in Example \ref{ex:advectiondiffusion}.
+The domain was $\Omega=(-1,1)^d$, for $d=2,3$, diffusivity was $\eps=0.1$ (equation \eqref{eq:advectiondiffusionstrong}), and there were zero Dirichlet boundary conditions.
+The 2D problem satisfied the coercivity hypothese with a divergence-free velocity field $\bX = (7+5y,-5x)$, but in 3D we set $\bX = (7+5y,-5x,2z)$ so $\grad\cdot\bX \ne 0$.
+The discontinuous source term $\phi \in L^\infty(\Omega)$ satisfied $\phi\ge 0$ in half of the domain, with positive circular patches, while values were negative in the other half.
+While the other Examples in this Section are unilateral, here the solution was box-constrained: $0 \le u \le 1$.
+Figure \ref{fig:results:advdiff} shows coincidence sets for the 2D problem; the 3D sets are comparable in regularity and general pattern.
 
@@
 \end{center}
-\caption{Coincidence sets (black) for the 2D bi-obstacle problem in Example \ref{ex:results:advdiff}: $u=0$ (left) and $u=1$ (right).}
+\caption{Coincidence sets for the 2D bi-obstacle problem in Example \ref{ex:results:advdiff}: $u=0$ (left) and $u=1$ (right).}
 \label{fig:results:advdiff}
@@
 
-For the experiment we used a coarse mesh with $m_0=16^d$ nodes so as to minimally-resolve the coincidence set.
+To minimally-resolve the coincidence set we used a coarse mesh with $m_0=16^d$ nodes.
+The 2D solver applied $P_1$ elements and a single processor, but in 3D we used $Q_1$ elements (hexahedra) and $P=16$ processes.
 The smoother was two (fixed) RS Newton iterations,  with three preconditioned GMRES iterations to approximately solve the step equations.
@@
 F-cycle iteration results are shown in Table \ref{tab:results:advdiff}.
-Despite the hint of growing iterations under mesh refinement, a few iterations suffice up to memory-limited resolutions.
-For any of the finer mesh runs here, the FASCD F-cycle solver takes less time than generating the $P_1$ mesh hierarchy.
+Despite the hint of growing iterations under mesh refinement, a few iterations sufficed up to memory-limited resolutions.
+For any of the finer mesh runs here, the FASCD F-cycle solver took less time than generating the $P_1$ mesh hierarchy.
+The local Peclet number $\text{Pe}=\|\bX(x,y,z)\|_2 L \eps^{-1}$, with $L=2$ the width of $\Omega$, varied from 40 to 260.
+Thus the $m_J\le 121^d$ problems were significantly advection-dominated while those with $m_J\ge 481^d$ were diffusion-dominated.
+Nonetheless the smoother was the same and F-cycle iterations were nearly constant.
 \end{example}
 
-% see git@bitbucket.org:pefarrell/fascd.git: examples/pollutant.py and edplay/results/pollutant.sh|txt
+% see git@bitbucket.org:pefarrell/fascd.git: examples/pollutant.py
+% and https://github.com/bueler/mcd-extended: paper/results/pollutant.sh|txt
 \begin{table}[ht]
@@
  2 &    $31^d$ & 2 & 2 \\
- 3 &    $61^d$ & 2 & 3 \\
- 4 &   $121^d$ & 2 & 3 \\
- 5 &   $241^d$ & 2 & 4 \\
+ 3 &    $61^d$ & 2 & 2 \\
+ 4 &   $121^d$ & 2 & 2 \\
+ 5 &   $241^d$ & 2 & 3 \\
  6 &   $481^d$ & 2 \\
@@
 \bigskip
-\caption{F-cycle iterations for 2D (serial) and 3D (16 processes) solutions of the advection-diffusion bi-obstacle VI problem in Example \ref{ex:results:advdiff}.}
+\caption{F-cycle iterations for 2D ($P_1$, serial) and 3D ($Q_1$, 16 processes) solutions of the advection-diffusion bi-obstacle VI problem in Example \ref{ex:results:advdiff}.}
 \label{tab:results:advdiff}
@@
 The data consists of a radial surface mass balance $a$ function \cite[equation (5.122)]{GreveBlatter2009} and a bumpy, synthetic map for $b\in C^1(\Omega)$, with an elevation range of $1800$ m.
-The numerical solution, colored by the solution surface flow speed, i.e.~the magnitude of $\mathbf{u} = (n+1)(n+2)^{-1} \Gamma (s-b)^{n+1} |\grad s|^{n-1} \grad s$, is shown in Figure \ref{fig:results:siascene}.
-As noted in Example \ref{ex:sia}, and clearly seen in this Figure, $|\grad s|$ is unbounded at the free boundary; the solution, while continuous, has very low regularity.
-
-% to regenerate:
+A numerical solution is shown in Figure \ref{fig:results:siascene}, colored by the solution surface flow speed, i.e.~the magnitude of $\mathbf{u} = (n+1)(n+2)^{-1} \Gamma (s-b)^{n+1} |\grad s|^{n-1} \grad s$.
+As noted in Example \ref{ex:sia}, and seen in this Figure, $|\grad s|$ is unbounded at the free boundary; the solution $s$, while continuous, has very low regularity.
+
+% to regenerate from git@bitbucket.org:pefarrell/fascd.git: examples/:
 %   $ mpiexec -n 16 --map-by core --bind-to hwthread python3 sia.py -levs 8 -bumps \
@@
 \end{center}
-\caption{Solution of Example \ref{ex:sia} over a bumpy bedrock topography, with the ice speed as the surface coloring, on a $P_1$ mesh of $m^J=1025^2 \approx 10^6$ nodes.}
+\caption{Solution of Example \ref{ex:sia} over a bumpy bedrock topography, with the ice speed as surface coloring, and 100-times vertical exageration.}
 \label{fig:results:siascene}
@@
 
-We applied FASCD F-cycles using a smoother of four (fixed) iterations of RS Newton (Section \ref{sec:implementation}), with quadratic backtracking line search.
+We applied FASCD F-cycles using $Q_1$ elements and a smoother of four (fixed) iterations of RS Newton (Section \ref{sec:implementation}), with quadratic backtracking line search.
 The step equations were approximately solved, in parallel, by three (fixed) iterations of preconditioned GMRES, again with a ASM+ILU preconditioner.
 The coarse solver was RS Newton with LU applied redundantly over all processors.
-The coarse mesh was relatively fine, with $m_0=17^2$ nodes and 512 triangles.
+The coarse mesh was relatively fine, with $m_0=21^2$ nodes and 400 elements.
 Because of the low regularity of the solution we used \id{rtol} $= 2 \times 10^{-4}$ and \id{atol} $= 10^{-8}$ in \eqref{eq:stoppingcriterion}.
@@
 These solver parameters generate a solver with mesh-independent iterations.
-We fixed the number of degrees of freedom per process at $m_J/P=1025^2$ and solved over $P=1,4,16$ processes.\footnote{Runs were on a single-node workstation with two Intel Xeon 4210 cpus (2.2GHz), a total of 20 physical cores, and 128 GB shared memory.}  The results in Table \ref{tab:results:siaweak} show clear weak scaling.
-The finest mesh result, with $m_J=4097^2=17 \times 10^6$ degrees of freedom, needed less than six minutes to solve for the surface elevation of an ice sheet slightly larger than the current Greenland ice sheet ($1.7$ million $\text{km}^2$) at uniform $439$ m resolution.
+Then we fixed the number of degrees of freedom per process at $m_J/P=1281^2 \approx 1.6 \times 10^6$ and solved over $P=1,4,16$ processes.\footnote{Runs were on a single-node workstation with two Intel Xeon 4210 cpus (2.2GHz), a total of 20 physical cores, and 128 GB shared memory.}  The results in Table \ref{tab:results:siaweak} show high-quality weak scaling, that is, nearly-constant run time for mesh refinement under constant $m_J/P$.
+The finest mesh result, with $m_J=5121^2=2.6 \times 10^7$ degrees of freedom, needed six minutes to solve for the surface elevation of an ice sheet slightly larger than the current Greenland ice sheet ($1.7$ million $\text{km}^2$) at uniform $350$ m resolution.
 \end{example}
 
-% see pefarrell-fascd/edplay/siaweak.sh|txt
-% time in seconds:
-%>> 5*60+[11,36,55]
-%ans =      311      336      355
+% see git@bitbucket.org:pefarrell/fascd.git: examples/sia.py
+% and https://github.com/bueler/mcd-extended: paper/results/siaweak.sh|txt
 \begin{table}[ht]
@@
 $P$ & \emph{levels} & $m_J$ & \emph{F-cycles} & \emph{time} (s) \\ \cmidrule{1-5}
- 1 & 7 & $1025^2$ & 4 & 311 \\
- 4 & 8 & $2049^2$ & 4 & 336 \\
-16 & 9 & $4097^2$ & 4 & 355
+ 1 & 7 & $1281^2$ & 3 & 309 \\
+ 4 & 8 & $2561^2$ & 3 & 368 \\
+16 & 9 & $5121^2$ & 3 & 374
 \end{tabular}
@@
 \end{table}
-
-\begin{comment}
-from slack, regarding SIA example:
-Patrick:  The convergence criteria are relatively loose?
-Ed:  Yes, the convergence is loose, and furthermore there is a double-regularization in the nonlinear diffusivity.
-The problem is that only the 8/3 power of the solution u is actually in W^{1,p} for p=4; for the flat bed case u^{8/3} is actually C^1.
-But the derivative of u itself is not actually integrable, though u is C^0; you can see there is essentially a cliff all the way around the ice sheet.
-So as we converge to the exact solution, along the free boundary the P1 elements are trying to match this unbounded derivative, and it struggles at tighter tolerances.
-The best (but still not like textbook problems) results are with a Petrov-Galerkin method using Q1, for which the convergence rates are much closer to O(h^2).
-See Bueler (2016) for this P-G method, and Jouvet & Bueler (2012) for an existence proof giving u^{8/3} in W^{1,4}.
-If you really want to know ...
-\end{comment}
 

### 2023-07-02 paper/fascd.tex
remove an unneeded footnote

@@
 
-The mesh hierarchy and FE spaces can be constructed by standard nested refinement.\footnote{The dimension-dependent details of mesh refinement are suppressed here.
-Section \ref{sec:results} only uses $P_1$ elements for intervals, triangles, and tetrahedra, in dimensions $d=1,2,3$ respectively, within the Firedrake framework \cite{Rathgeberetal2016}.}  Suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate elements with union equal to $\overline{\Omega}$ and maximum edge length (characteristic mesh size) $h_0>0$.
+The mesh hierarchy and FE spaces can be constructed by standard nested refinement.
+Suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate elements with union equal to $\overline{\Omega}$ and maximum edge length (characteristic mesh size) $h_0>0$.
 We assume $\mathcal{T}^0$ is conforming, with no hanging nodes.
@@
 
-In the VI material which follows, use of higher-order Lagrange elements, i.e.~$P_k$ or $Q_k$ with $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} special to inequality constraints, which we avoid.
+In the material which follows, use of higher-order Lagrange elements, i.e.~$P_k$ or $Q_k$ with $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} special to inequality constraints, which we avoid.
 Namely, our FE construction of constraint sets (Section \ref{sec:cdmultilevel}) will require \emph{nodal monotonicity},

### 2023-07-04 paper/fascd.tex
Some more edits

@@
 \usepackage{booktabs,multirow,xspace}
+\usepackage{pifont}
 
@@
 We seek to rapidly compute the approximate solution of box-constrained VI problems \eqref{eq:boxdirichletvi} using finite elements (FE).
-Suppose $\Omega \subset \RR^d$ is a (open) bounded polygon.\footnote{Section \ref{sec:results} illustrates $d=1,2,3$ cases.}  We will construct a multilevel FE approximation to \eqref{eq:boxdirichletvi} over $\Omega$ using $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, by refining a coarsest mesh into a mesh hierarchy; the goal is an accurate solution on the finest-level mesh.
+Let $\Omega \subset \RR^d$ be a (open) bounded polygon.
+We will construct a multilevel FE approximation to \eqref{eq:boxdirichletvi} over $\Omega$ using $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, by refining a coarsest mesh into a nested mesh hierarchy; the goal is an accurate solution on the finest-level mesh.
 The hierarchy allows the creation of a multilevel CD of $\cK$, using the ideas in the next two Sections, which permits solvers based on admissibility-preserving coarse corrections.
@@
 The mesh hierarchy and FE spaces can be constructed by standard nested refinement.
-Suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate elements with union equal to $\overline{\Omega}$ and maximum edge length (characteristic mesh size) $h_0>0$.
+Let $\mathcal{T}^0$ be the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate cells with union equal to $\overline{\Omega}$ and maximum cell diameter (characteristic mesh size) $h_0>0$.
 We assume $\mathcal{T}^0$ is conforming, with no hanging nodes.
-For $j=1,\dots,J$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ so that each $\mathcal{T}^{j-1}$ element becomes $2^d$ elements in $\mathcal{T}^j$ with maximum edge length $h_j = 2^{-j} h_0$.
+For $j=1,\dots,J$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ so that each $\mathcal{T}^{j-1}$ cell becomes $2^d$ cells in $\mathcal{T}^j$ with mesh size $h_j = 2^{-j} h_0$.
 We call $J\ge 0$ the \emph{depth} of the hierarchy, $J+1$ the \emph{number of levels}, and $\mathcal{T}^J$ the \emph{finest-level mesh}.
-Let $m_j$ be the number of nodes in $\mathcal{T}^j$, so there are $m_j$ \emph{degrees of freedom} at the $j$th level.
-Let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space over $\mathcal{T}^j$, with arbitrary boundary values, and observe the following vector space nesting:
+Let $m_j$ be the number of vertices in $\mathcal{T}^j$, so there are $m_j$ \emph{degrees of freedom} at the $j$th level.
+Let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space over $\mathcal{T}^j$, with arbitrary boundary values, with the standard vector space nesting:
 \begin{equation}
@@
 Evidently \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of such indefinite-sign hat functions.
+\pef{I'm not sure this constitutes a variational crime?
+We enforce the RHS because it's tractable, and this is only equivalent to what we actually want (the LHS) for P1 and Q1.
+I don't see where we're departing from the variational framework, however.}
 
@@
 \end{equation}
-Also assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.  Define
+\pef{Shouldn't there be some relation between the continuous bounds and the discrete ones?}
+In analogy to the continuous data, also assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.  Define
 \begin{equation}
@@
 This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.
-As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fe:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J$ are infinite; there is no lower or upper constraint, respectively, at such nodes.
+%As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fe:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J$ are infinite; there is no lower or upper constraint, respectively, at such nodes.
 
@@
 \end{equation}
-If $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, then, as for \eqref{eq:vi}, FE VI \eqref{eq:fe:vi} is well-posed.
+If $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, then the discrete VI \eqref{eq:fe:vi} is well-posed.
 
@@
 
-The multilevel algorithm defined in Section \ref{sec:vcycle}, a solver for FE VI \eqref{eq:fe:vi}, requires transfer operators between meshes.
-We require five such operators, of which three are standard \cite{Trottenbergetal2001}.
-Table \ref{tab:transfers} will help the reader keep track.
+The multilevel algorithm defined in Section \ref{sec:vcycle}, a solver for \eqref{eq:fe:vi}, requires five transfer operators between meshes.
+Of these five operators, three are employed in standard FAS \cite{Trottenbergetal2001}.
+We include Table \ref{tab:transfers} to help the reader keep track.
 
@@
 \begin{tabular}{llc}
-\emph{name}  & \emph{action}  & \emph{linear?} \\ \hline
+\toprule
+\emph{name}  & \emph{action}  & \emph{linear?} \\ \midrule
 canonical prolongation        & $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ & \,\checkmark \\
@@
 (nodal) injection             & $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ & \,\checkmark \\
-maximum restriction           & $\maxR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ & \\
-minimum restriction           & $\minR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ &
+maximum restriction           & $\maxR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ & \ding{55} \\
+minimum restriction           & $\minR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ & \ding{55} \\
+\bottomrule
 \end{tabular}
-
-\medskip
 \caption{Transfer operators used in this paper.}
@@
 
-Canonical prolongation $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ is the vector space inclusion $\mathcal{V}^{j-1} \hookrightarrow \mathcal{V}^j$ of coarser-mesh functions.
+Canonical prolongation $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ implements the vector space inclusion $\mathcal{V}^{j-1} \hookrightarrow \mathcal{V}^j$ of coarser-mesh functions.
 Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ also acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $(R\sigma)[z] = \sigma[z]$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
@@
 
-The monotone restrictions $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to nodal functions $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh hat functions.
-To make this precise, let $x_q^j$ denote the nodes of mesh $\mathcal{T}^j$, with corresponding hat functions $\psi_q^j(x)$.
+\pef{I don't like the name ``monotone restrictions''.
+The word restriction is strongly associated with operating on the dual space.
+I would prefer to call them ``monotone injections'', since that is the matching operator.}
+The monotone restrictions $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to nodal functions $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the \emph{stars} of the coarse vertices).
+These operators are constructed to enforce the property
+\begin{equation}
+\minR z \le z \le \maxR z,  \label{eq:fe:monotonerestrictionprops}
+\end{equation}
+which does not hold for the standard injection operator $\iR$.
+To make the construction precise, let $x_q^j$ denote the nodes of mesh $\mathcal{T}^j$, with corresponding basis functions $\psi_q^j(x)$.
 For a $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$:
@@
 \end{align}
-For standard refinement a coarse-mesh node $x_p^{j-1}$ is in the set of fine nodes $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so it follows that
-\begin{equation}
-\minR z \le z \le \maxR z,  \label{eq:fe:monotonerestrictionprops}
-\end{equation}
+For standard refinement a coarse-mesh node $x_p^{j-1}$ is in the set of fine nodes $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so \eqref{eq:fe:monotonerestrictionprops} follows.
 On the other hand, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These ordering properties are used in the CD construction of the next two Sections.
+\pef{It would be good to have a diagram here. I could write one in tikz, if you don't have one handy.}
 
@@
 
-Our practical multilevel CD scheme uses \emph{defect constraints} \cite{GraeserKornhuber2009} which are based on an admissible finest-level iterate.
+A key choice in devising a multilevel scheme for VIs is how to construct coarse-grid bounds.
+Our multilevel CD scheme uses \emph{defect constraints} \cite{GraeserKornhuber2009} which are based on an admissible finest-level iterate.
 The next Example concretely illustrates why we do not apply restriction operators directly\footnote{An exception is in the F-cycle; see Section \ref{sec:implementation}.} to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.
@@
 \begin{example}  \label{ex:directRbad}
-For concreteness, suppose $\Omega = [0,L] \subset \RR^1$ is an interval and consider a two-level mesh hierarchy ($J=1$).
-Let $\underline{\gamma}^1$ be a sawtooth function with minimum value zero, and suppose $\overline{\gamma}^1=C+\underline{\gamma}^1$ with $C>0$ which is strictly less than the range of $\underline{\gamma}^1$.
+
+Let $\Omega = (0,L) \subset \RR^1$ and consider a two-level mesh hierarchy ($J=1$).
+Let $\underline{\gamma}^1$ be a sawtooth function with minimum value zero, and suppose $\overline{\gamma}^1=C+\underline{\gamma}^1$ with $C>0$ which is strictly less than the maximum of $\underline{\gamma}^1$.
 Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ nodes, respectively.
-Supposing empty Dirichlet boundary for simplicity, the constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.
-However, various restriction schemes applied directly to $\underline{\gamma}^1,\overline{\gamma}^1$ are problematical.
+Ignoring Dirichlet boundary for simplicity, the constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.
+However, various restriction schemes applied directly to $\underline{\gamma}^1,\overline{\gamma}^1$ are problematic.
 Suppose we propose
@@
     $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
-The general motivation for this choice---compare \eqref{eq:fe:chilevels}---is that $w^0 \ge \underline{\gamma}^0 \, \implies \, Pw^0 \ge \underline{\gamma}^1$ and $w^0 \le \overline{\gamma}^0 \, \implies \, Pw^0 \le \overline{\gamma}^1$.
-However, while these facts are true, the difficulty should be clear: the coarse-level admissible set $\mathcal{K}^0$ is now empty because the proposed lower bound $\underline{\gamma}^0$ actually exceeds the proposed upper bound $\overline{\gamma}^0$.
+The general motivation for this choice---compare \eqref{eq:fe:chilevels} below---is that $w^0 \ge \underline{\gamma}^0 \, \implies \, Pw^0 \ge \underline{\gamma}^1$ and $w^0 \le \overline{\gamma}^0 \, \implies \, Pw^0 \le \overline{\gamma}^1$.
+However, while these facts are true, the difficulty with this choice is that the coarse-level admissible set $\mathcal{K}^0$ is now empty because the proposed lower bound $\underline{\gamma}^0$ actually exceeds the proposed upper bound $\overline{\gamma}^0$.
 \end{example}
@@
 \end{figure}
+\pef{Can we put another figure beside it illustrating the choice $\underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1$?}
+\pef{Maybe use different colours for upper and lower bounds?}
 
@@
 This construction of multilevel box constraints, fundamentally the same as the unilateral construction in \cite{GraeserKornhuber2009}, is designed to permit large-yet-admissible coarse corrections, which promotes multilevel solver efficiency.
-Note that for Algorithm \ref{alg:fascd} below, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are actually not required.
+Note that for Algorithm \ref{alg:fascd} below, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not required.
 
 We now compute the LDC differences \cite{GraeserKornhuber2009}:
+\pef{Say a few words of motivation first for introducing this concept.}
 \begin{equation}
@@
 
-Our V-cycle solver (Algorithm \ref{alg:fascd}) will compute coarse corrections in $\mathcal{D}^j$ for the downward side and in $\mathcal{U}^j$ for the upward side.
+Our V-cycle solver (Algorithm \ref{alg:fascd}) will compute corrections in $\mathcal{D}^j$ for the downward stage and in $\mathcal{U}^j$ for the upward stage.
 This strategy is, however, based on \emph{incomplete} CDs, in the general box-constrained cases, as we now explain.
-Each upward set $\mathcal{U}^j$ is generally only a superset of a sum of downward sets $\mathcal{D}^j$, but this becomes a true CD, i.e.~\eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, in unilateral cases.
+Each upward set $\mathcal{U}^j$ is generally only a superset of a sum of downward sets $\mathcal{D}^j$, but this becomes a true CD, i.e.~satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, in unilateral cases.
 The following Lemma, apparently new, is based in part on the unilateral CD construction underlying Algorithm 4.7 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.
@@
 \begin{example}  \label{ex:notfullcd}
-Let $a\ge 0$ be an arbitrary positive number.
+Let $a > 0$ be an arbitrary positive number.
 Consider a 2-level hierarchy ($J=1$), and suppose the lower defect constraint is the constant function $\underline{\chi}^1=-a$.
@@
 
-Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion}, which are special to multilevel FE schemes based on defect constraints, hold regardless of whether a CD operator $\Pi_i$ can be constructed.
-Here there is, in fact, another ``basic'' iteration, namely \pr{icd-tele} in Algorithm \ref{alg:basiccd-tele}, which also does not need the $\Pi_i$ operators.
+Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion}, which are special to multilevel schemes based on defect constraints, hold regardless of whether a CD operator $\Pi_i$ can be constructed.
+Here there is, in fact, another ``basic'' iteration, an incomplete variant of \pr{cd-mult} (Algorithm \ref{alg:basiccd-mult}) that does not need the $\Pi_i$ operators.
 It is a variant of \pr{cd-mult} (Algorithm \ref{alg:basiccd-mult}).
@@
 While inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in Algorithm \ref{alg:fascd} below, a convergence analysis, perhaps extending the arguments of \cite{Tai2003} or \cite{GraeserKornhuber2009} to VIs which are not of minimization type, and/or to box constraints, would seem to require full CDs.
-We cannot offer that analysis.
+We cannot currently offer that analysis.
 
@@
 
-Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ based on perturbations from each coarser FE subspace $\mathcal{V}^j$.
-By nesting \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$,\footnote{Nonetheless we use prolongation $P$ to indicate where the computer representation changes.} but the corrected iterate needs to be admissible.
-Proceeding inductively, suppose $y^i \in \mathcal{D}^i$ for $i=J,\dots,j-1$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).
+Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ by incorporating information from each coarser FE subspace $\mathcal{V}^j$.
+By nesting \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$,\footnote{Nonetheless we use prolongation $P$ to indicate where the computer representation changes.} and are constructed so that the corrected iterate remains admissible.
+Proceeding inductively, suppose $y^i \in \mathcal{D}^i$ for $i=J,\dots,j+1$ \pef{I changed the sign here from $j-1$ to $j+1$, double-check} are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).
 By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.
@@
 Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using \eqref{eq:fe:upwardsuminclusion}.
-Continuing in this ``telescoping'' way (Algorithm \ref{alg:basiccd-tele}), up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction.
+Continuing in this ``telescoping'' way (Algorithm \ref{alg:basiccd-tele}), up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.
 Then the next admissible iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.
@@
 A key idea is that, because of inclusions \eqref{eq:fe:upwardsuminclusion}, up-smoothing solves less-restricted VI problems than down-smoothing.
-Intuitively-speaking, going downward one does not know what the next coarse corrections will be, and thus they must be in sufficiently small sets ($\mathcal{D}^j$) so that their sum is admissible once we return to a given level.
+Intuitively-speaking, going downward one does not know what the next coarse corrections will be, and thus on any level one must choose conservative constraint sets $\mathcal{D}^j$ so that the sum of the corrections is admissible once we return to a given level.
 Going upward, however, no further coarse corrections can violate admissibility.
@@
 \end{equation}
-Observe that injection $\iR$ coarsens the iterates---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---but canonical restriction $R$ coarsens linear functionals; recall Table \ref{tab:transfers}.
+Observe that injection $\iR$ coarsens the (primal) iterates---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---but canonical restriction $R$ coarsens (dual) linear functionals; recall Table \ref{tab:transfers}.
 
@@
 \end{equation}
-An exception is that $z_{(0)}^0=0$ on the coarsest level.  Note $z_{(0)}^j$ is admissible by \eqref{eq:fe:upwardsuminclusion}: $\mathcal{D}^j + \mathcal{U}^{j-1} \subset \mathcal{U}^j$.
-
-If we were to inject the original box constraints \eqref{eq:fe:fineconstraintset} downward to coarser meshes, i.e.~$\underline{\gamma}^j = \iR \underline{\gamma}^{j+1}$ and $\overline{\gamma}^j = \iR \overline{\gamma}^{j+1}$, then, because injection preserves inequalities, and because $y^{j+1} \in \mathcal{D}^{j+1}$, it would follow from \eqref{eq:fe:definew} that $\underline{\gamma}^j \le w^j \le \overline{\gamma}^j$.
+An exception is that $z_{(0)}^0=0$ on the coarsest level.  Note $z_{(0)}^j$ is admissible by \eqref{eq:fe:upwardsuminclusion}.
+
+%If we were to inject the original box constraints \eqref{eq:fe:fineconstraintset} downward to coarser meshes, i.e.~$\underline{\gamma}^j = \iR \underline{\gamma}^{j+1}$ and $\overline{\gamma}^j = \iR \overline{\gamma}^{j+1}$, then, because injection preserves inequalities, and because $y^{j+1} \in \mathcal{D}^{j+1}$, it would follow from \eqref{eq:fe:definew} that $\underline{\gamma}^j \le w^j \le \overline{\gamma}^j$.
 In this sense $w^j$ is admissible with respect to the box constraints.
@@
 \end{align}
-This is the same as VI problem \eqref{eq:fe:downvi}, but in cluttered form, thus definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource}.  The explanation for VI \eqref{eq:fe:upvi} is similar.
+This is the same as VI problem \eqref{eq:fe:downvi}, but in cluttered form, motivating definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource}.
+The explanation for VI \eqref{eq:fe:upvi} is similar.
 
@@
 
-Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle.\footnote{We could write ``$V(\id{down},\id{up})$'', but we prefer to emphasize the number of subsets used on each level rather than the number of smoother sweeps in a given set.}  If \id{up} $=0$ then it becomes a $V(1,0)$ cycle and if \id{down} $=0$ then it is a $V(0,1)$ cycle; in the former case it generalizes Algorithm 4.7 in \cite{GraeserKornhuber2009} by allowing nonlinear operators $f$, any smoother, and applying FAS-type corrections.
+Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle.\footnote{We could write ``$V(\id{down},\id{up})$'', but we prefer to emphasize the number of subsets used on each level rather than the number of smoother sweeps in a given set.
+\pef{I don't understand this.}}  If \id{up} $=0$ then it becomes a $V(1,0)$ cycle and if \id{down} $=0$ then it is a $V(0,1)$ cycle; in the former case it generalizes Algorithm 4.7 in \cite{GraeserKornhuber2009} by allowing nonlinear operators $f$, any smoother, and applying FAS-type corrections.
 In unilateral cases it is precise to say that one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} over CD \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over incomplete CD \eqref{eq:fe:upwardsuminclusion}; in this sense $V(1,1)=V(1,0)+V(0,1)$.

### 2023-07-04 paper/fascd.tex
Edits through Section 7

@@
 
-When $f$ has the locality of Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over hat function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}  However, for general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of CD-type iterations.
+When $f$ has the locality of Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over basis function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}  However, for general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of CD-type iterations.
 Section \ref{sec:vcycle} constructs a multiplicative multilevel CD iteration for a general operator $f:\cK\to\cV'$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
@@
 for all $w,z \in \mathcal{V}^j$, where $\bw,\bz \in \RR^{m_j}$ denote vectors of nodal values.
-However, for $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal values one or zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.
-Evidently \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of such indefinite-sign hat functions.
+However, for $k\ge 2$ there are basis functions in $P_k$ and $Q_k$, i.e.~functions with nodal values one or zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.
+Evidently \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of such indefinite-sign basis functions.
 \pef{I'm not sure this constitutes a variational crime?
@@
 \end{pseudo}
-\caption{The FASCD V-cycle is an iteration for solving FE VI problem \eqref{eq:fe:vi}.  Here $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
+\caption{The FASCD V-cycle is an iteration for solving VI problem \eqref{eq:fe:vi}.  Here $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
 \label{alg:fascd}
@@
 Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle.\footnote{We could write ``$V(\id{down},\id{up})$'', but we prefer to emphasize the number of subsets used on each level rather than the number of smoother sweeps in a given set.
-\pef{I don't understand this.}}  If \id{up} $=0$ then it becomes a $V(1,0)$ cycle and if \id{down} $=0$ then it is a $V(0,1)$ cycle; in the former case it generalizes Algorithm 4.7 in \cite{GraeserKornhuber2009} by allowing nonlinear operators $f$, any smoother, and applying FAS-type corrections.
-In unilateral cases it is precise to say that one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} over CD \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over incomplete CD \eqref{eq:fe:upwardsuminclusion}; in this sense $V(1,1)=V(1,0)+V(0,1)$.
-
-Regarding the floating-point storage required, let us assuming all box constraints are nontrivial.
-Then seven vectors must be allocated on each level: $\underline{\chi}^j$, $\overline{\chi}^j$, $\underline{\phi}^j$, $\overline{\phi}^j$, $w^j$, $\ell^j$, $y^j$.
+\pef{I don't understand what the number of subsets used on each level means.}}  If \id{up} $=0$ then it becomes a $V(1,0)$ cycle and if \id{down} $=0$ then it is a $V(0,1)$ cycle; in the former case it generalizes Algorithm 4.7 in \cite{GraeserKornhuber2009} by allowing nonlinear operators $f$, any smoother, and applying FAS-type corrections.
+In unilateral cases, one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} over CD \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over incomplete CD \eqref{eq:fe:upwardsuminclusion}; in this sense $V(1,1)=V(1,0)+V(0,1)$.
+
+Regarding the storage required, let us assuming all box constraints are nontrivial.
+In this case, seven vectors must be allocated on each level: $\underline{\chi}^j$, $\overline{\chi}^j$, $\underline{\phi}^j$, $\overline{\phi}^j$, $w^j$, $\ell^j$, $y^j$.
 Note $z^j$ can use the same storage as $y^j$.
@@
 Because the problem is inequality-constrained, repeated application of Algorithm \ref{alg:fascd} to solve the finite-dimensional, box-constrained VI \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.
-To construct an appropriate convergence criteria for the discretization, note \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}, a strong form.
-Let $x_p \in \mathcal{T}^J$ be a finest-level node, and $\psi_p(x)$ the corresponding hat function.
+To construct appropriate convergence criteria for the discretization, note \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}.
+Let $x_p \in \mathcal{T}^J$ be a finest-level node, and $\psi_p(x)$ the corresponding basis function.
 Let $\mathcal{X}^J=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}^J$.
@@
 for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
+\pef{I wonder if this definition will confuse people, because they won't know how to plug in $\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ into your definition of $\perp$ ($\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ is not in $\RR^n$).
+Maybe it would be easier to just use the spelled-out version below, and not use this notation?}
 
@@
 
+\pef{I think this is all fairly standard and we could abbreviate this section substantially.}
+
 Direct implementation of \eqref{eq:rNC} is, however, sensitive to rounding error because $\rNC$ is a discontinuous function of $w^J(x_p)$.
@@
 
-Suppose an iteration starts with $w^{J,0} \in \mathcal{K}^J$, generating (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.  Let $\id{atol}\ge 0$, $\id{rtol}\ge 0$ be given tolerances.  We propose
+Suppose an iteration starts with $w^{J,0} \in \mathcal{K}^J$, generating (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.  Let $\id{atol}\ge 0$, $\id{rtol}\ge 0$ be given tolerances.  We employ
 \begin{equation}
@@
 \end{equation}
-as the stopping criterion.  A step tolerance, $\|w^{J,k} - w^{J,k-1}\|_2 < \id{stol}$ or similar, could also be used, but for the results in Section \ref{sec:results} we did not find it necessary.
-
-Iteration of V-cycle Algorithm \ref{alg:fascd}, starting with some finest-mesh initial iterate $w^{J,0}$ and repeated until \eqref{eq:stoppingcriterion} holds, will be shown (Section \ref{sec:results}) to be an effective solver.
+as the stopping criterion.  %A step tolerance, $\|w^{J,k} - w^{J,k-1}\|_2 < \id{stol}$ or similar, could also be used, but for the results in Section \ref{sec:results} we did not find it necessary.
+
+In Section \ref{sec:results} we will show that iteration of the V-cycle Algorithm \ref{alg:fascd}, starting with some finest-mesh initial iterate $w^{J,0}$ and repeated until \eqref{eq:stoppingcriterion} holds, is an effective solver.
 However, an F-cycle (Figure \ref{fig:fcycle}), also called a \emph{full multigrid cycle}, is even better.
 The well-known principle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate.
-Then an admissible initial iterate on the next-finer mesh is generated via prolongation and truncation, and so on.
+Then an admissible initial iterate on the next-finer mesh is generated via prolongation and truncation\pef{Clarify truncation}, and so on.
 During this ``ramp'' up one approximately solves on each level by using a fixed number of V-cycles, each down to the coarsest mesh.
@@
 Then it prolongs and truncates (doubled edges) to create an admissible initial iterate on the next-finer level.
-Solution on each level is by V-cycles; \id{rampv} $=1$ here.
+Solution on each level is by V-cycles.
 See Figure \ref{fig:fascdvcycle} for the other symbols.}
@@
 
-This F-cycle approach is in Algorithm \ref{alg:fascd-fcycle}.
-Observe that while injection of the original fine-level constraints $\underline{\gamma}^J, \overline{\gamma}^J$ is \emph{not} used for V-cycles, wherein the LDCs use an existing admissible iterate (Section \ref{sec:cdmultilevel}), when combined with truncation into $\mathcal{K}^j=\{v \in \mathcal{V}^j \,:\, \underline{\gamma}^j \le v \le \overline{\gamma}^j, \, v|_{\partial\Omega} = g_D\}$ injection is appropriate for the ramp in an F-cycle; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo}.
+This F-cycle approach is described in Algorithm \ref{alg:fascd-fcycle}.
+Observe that while injection of the original fine-level constraints $\underline{\gamma}^J, \overline{\gamma}^J$ is \emph{not} used for V-cycles, wherein the LDCs use an existing admissible iterate (Section \ref{sec:cdmultilevel}), when combined with truncation into $\mathcal{K}^j=\{v \in \mathcal{V}^j \,:\, \underline{\gamma}^j \le v \le \overline{\gamma}^j, \, v|_{\partial\Omega} = g_D\}$, injection \emph{is} appropriate for the ramp in an F-cycle\pef{Explain why?}; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo}.
 One may generate the coarsest-level initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation down to coarsest level; compare line \ref{line:fcyclecoarsestinitial}.
@@
 \end{pseudo}
-\caption{The FASCD F-cycle for solving FE VI problem \eqref{eq:fe:vi}.}
+\caption{The FASCD F-cycle for solving VI problem \eqref{eq:fe:vi}.}
 \label{alg:fascd-fcycle}
@@
 
-A large space of smoothers could be explored,\footnote{See comments in Section \ref{sec:discussion}.} but Section \ref{sec:results} only demonstrates one powerful, $O(m_j)$ work per iteration, smoother choice.
-We apply a fixed number of iterations of the active-set (reduced-space) Newton method from \cite{BensonMunson2006}, the \texttt{-snes\_type vinewtonrsls} solver in PETSc \cite{Balayetal2023}, denoted as \emph{RS Newton}.
-We use it with a fixed number of preconditioned Krylov iterations to solve the step equations.
+A large space of smoothers could be explored; in Section \ref{sec:results} we apply a fixed number of iterations of the active-set (reduced-space) Newton method from \cite{BensonMunson2006}, the \texttt{VINEWTONRSLS} solver in PETSc \cite{Balayetal2023}, denoted as \emph{RS Newton}.
+We use it with a fixed number of preconditioned Krylov iterations to solve the arising linear systems.
 For the symmetric Laplacian in Example \ref{ex:results:classical} we apply conjugate gradients (CG) with incomplete Cholesky (IC) preconditioning, but for other operators we apply the generalized minimum residual (GMRES) method with incomplete LU (ILU) preconditioning.
-The coarse solver is also RS Newton, but iterated to convergence and with LU decomposition solution of the step equations.
+This requires $O(m_j)$ work per iteration.
+The coarse solver is also RS Newton, but iterated to convergence and with direct solvers applied to solve the arising linear systems.
 
@@
 \medskip\noindent
-Ed: \cite{GraeserKornhuber2009} only considers the Laplacian lower obstacle case, and they write everything in terms of nonlinear Gauss-Seidel (i.e.
-optimize one hat function at a time).
+Ed: \cite{GraeserKornhuber2009} only considers the Laplacian lower obstacle case, and they write everything in terms of nonlinear Gauss-Seidel (i.e.~optimize one basis function at a time).
 They first lay out the \cite{Tai2003} CD strategy as Alg.
@@
 Then they observe that the convergence rates from \cite{Tai2003} are slower (worse) than if one does "monotone multigrid".
-This seems to mean to them that one modifies the coarse hat functions according to whether they overlap the current fine-grid estimate of the coincidence set, "truncated nodal basis functions".
+This seems to mean to them that one modifies the coarse basis functions according to whether they overlap the current fine-grid estimate of the coincidence set, "truncated nodal basis functions".
 (See the start of \cite{GraeserKornhuber2009} section 5.2.)  I think that making these modifications in Firedrake, so we could try monotone multigrid V-cycles, would be a nightmare.

### 2023-07-07 paper/fascd.tex
First editing pass complete

@@
 
-\newcommand{\XX}{{\sffamily X}}
+%\newcommand{\XX}{{\sffamily X}}
+\newcommand{\XX}{\ding{55}}
 
@@
 
-To allow box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
+To allow box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the vertices of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
 Suppose that the finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ satisfy
@@
 \pef{Shouldn't there be some relation between the continuous bounds and the discrete ones?}
-In analogy to the continuous data, also assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.  Define
-\begin{equation}
-\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \text{ on nodes of } \mathcal{T}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D^J|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
+In analogy to the continuous data, also assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those vertices of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.  Define
+\begin{equation}
+\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \text{ on vertices of } \mathcal{T}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D^J|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
 \end{equation}
@@
 This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.
-%As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fe:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J$ are infinite; there is no lower or upper constraint, respectively, at such nodes.
+%As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fe:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J$ are infinite; there is no lower or upper constraint, respectively, at such vertices.
 
@@
 Regardless of the smoothness of the continuum obstacles $\underline{\gamma}, \overline{\gamma}$, admissibility with respect to their FE interpolants $\underline{\gamma}^J, \overline{\gamma}^J$ will generally not imply continuum admissibility.\footnote{This caveat is especially relevant in geophysical situations where only nodal values of the obstacle are known.
-Nodal elevation data generally misses the actual topographic detail in-between the nodes; compare \cite{Bueler2016}.}
+Nodal elevation data generally misses the actual topographic detail in-between the vertices; compare \cite{Bueler2016}.}
 
@@
 
-To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, observe that the nodes of $\mathcal{T}^{j-1}$ are nodes of $\mathcal{T}^j$, and that for any values at $\mathcal{T}^j$ nodes there is a unique $\mathcal{V}^j$ function assuming those values.
+To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, observe that the vertices of $\mathcal{T}^{j-1}$ are vertices of $\mathcal{T}^j$, and that for any values at $\mathcal{T}^j$ vertices there is a unique $\mathcal{V}^j$ function assuming those values.
 For $z\in\mathcal{V}^j$ we define the injection $\iR z$ as the element of $\mathcal{V}^{j-1}$ with the same (coarse) nodal values.
@@
 which does not hold for the standard injection operator $\iR$.
-To make the construction precise, let $x_q^j$ denote the nodes of mesh $\mathcal{T}^j$, with corresponding basis functions $\psi_q^j(x)$.
+To make the construction precise, let $x_q^j$ denote the vertices of mesh $\mathcal{T}^j$, with corresponding basis functions $\psi_q^j(x)$.
 For a $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$:
@@
 \end{align}
-For standard refinement a coarse-mesh node $x_p^{j-1}$ is in the set of fine nodes $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so \eqref{eq:fe:monotonerestrictionprops} follows.
+For standard refinement a coarse-mesh node $x_p^{j-1}$ is in the set of fine vertices $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so \eqref{eq:fe:monotonerestrictionprops} follows.
 On the other hand, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These ordering properties are used in the CD construction of the next two Sections.
@@
 Let $\underline{\gamma}^1$ be a sawtooth function with minimum value zero, and suppose $\overline{\gamma}^1=C+\underline{\gamma}^1$ with $C>0$ which is strictly less than the maximum of $\underline{\gamma}^1$.
-Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ nodes, respectively.
+Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ vertices, respectively.
 Ignoring Dirichlet boundary for simplicity, the constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.
@@
 Note $\underline{\chi}^0=\maxR \underline{\chi}^1=-a$ also, so $\underline{\phi}^1=0$ and $\underline{\phi}^0=-a$.
-Suppose $z^1$ is a ``sawtooth'' function which takes on value $-a$ on the coarse nodes and value $0$ on the other nodes.
+Suppose $z^1$ is a ``sawtooth'' function which takes on value $-a$ on the coarse vertices and value $0$ on the other vertices.
 Then, for any $\overline{\chi}^1\ge 0$, we have $\overline{\chi}^1 \ge 0 \ge z^1\ge \underline{\chi}^1$, so $z^j \in \mathcal{U}^1$.
@@
 Let $\mathcal{X}^J=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}^J$.
-If $m_D$ is the number of nodes $x_p$ in the Dirichlet boundary, let $I_D$ be the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.
+If $m_D$ is the number of vertices $x_p$ in the Dirichlet boundary, let $I_D$ be the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.
 Finally, write $a\perp b$ when $a,b \in \RR^n$ are \emph{complementary}, that is, when $a_i \ge 0$ and $b_i \ge 0$ and $a_i b_i = 0$ for $i=0,\dots,n-1$.
@@
 In summary, $\rNC : \mathcal{K}^J \to \RR^{m_J}$ is a computable nonlinear map defined using the discretization $f^J$ and the vectors $\ell^J$, $\underline{\gamma}^J$, and $\overline{\gamma}^J$.
-A value $\rNC(w^J)_p$ is nonzero only at nodes where \eqref{eq:fe:mcp} is violated, either because $r(w^J)_p$ remains nonzero where the constraints are inactive, or because it has the wrong sign at an active node, or because the boundary condition is not satisfied.
+A value $\rNC(w^J)_p$ is nonzero only at vertices where \eqref{eq:fe:mcp} is violated, either because $r(w^J)_p$ remains nonzero where the constraints are inactive, or because it has the wrong sign at an active node, or because the boundary condition is not satisfied.
 
@@
 
-Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}, allowing $P_1$ or $Q_1$ elements and based on standard, global mesh refinement.
-Here we show numerical results for the several example problems in Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
-Our essential observations are that each V-cycle dramatically reduces the residual norm, and that very few F-cycles solve most examples to within discretization error.
+Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}, allowing $P_1$ or $Q_1$ elements and based on standard nested mesh refinement.
+Here we show numerical results for the example problems presented in Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
+Our main observations are that each V-cycle dramatically reduces the residual norm, and that very few F-cycles solve most examples to within discretization error.
 
 \begin{example} \label{ex:results:classical}
-Consider the 2D classical, i.e.~unilateral and Laplacian, obstacle problem, namely Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$.
+We consider the unilateral 2D Laplacian obstacle problem presented in Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$.
 We consider two problems with very different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$, both over square $\Omega$.
 The standard ``ball'' problem has an upper hemisphere obstacle $\psi$ \cite[Chapter 12]{Bueler2021}.
-In the ``spiral'' problem from Example 7.1.1 in \cite{GraeserKornhuber2009}, $\psi$ is constructed to produce a more interesting coincidence set; see Figure \ref{fig:results:classical}.
-
-For this experiment we used a coarsest mesh of $m_0=5^2$ nodes (32 triangles).
-The finest mesh had $m_9=2049^2 \approx 4.2 \times 10^{6}$ nodes.
-As our smoother we applied a single RS Newton iteration (Section \ref{sec:implementation}), with three (fixed) IC-preconditioned CG iterations to approximately solve the step equations.
+In the ``spiral'' problem \cite[Example 7.1.1]{GraeserKornhuber2009}, $\psi$ is constructed to produce a more interesting coincidence set; see Figure \ref{fig:results:classical}.
+
+For this experiment we used a coarsest mesh of $m_0=5^2$ vertices (32 triangles).
+The finest mesh had $m_9=2049^2 \approx 4.2 \times 10^{6}$ vertices.
+As our smoother we applied a single RS Newton iteration (Section \ref{sec:implementation}), with three (fixed) IC-preconditioned CG iterations to approximately solve the arising linear systems.
 The coarse solver was RS$+$LU, iterated to convergence.
@@
 
-Table \ref{tab:results:classical} shows the resulting iteration counts for finest meshes of $m_J$ nodes.
+Table \ref{tab:results:classical} shows the resulting iteration counts for finest meshes of $m_J$ vertices.
 When the RS Newton smoother is applied by itself as a solver (``RS only'') it shows rapidly-increasing iterations as expected, roughly proportional to resolution.
@@
 Noting that for F-cycles we are counting the number of V-cycles after the ramp (Section \ref{sec:implementation}), we see that a constant number of cycles solves the problem.
-Comparing the ball and spiral results, the complexity of the coincidence set or free boundary does not strongly-influence FASCD performance.
-Note that each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+IC) iterations.
-Thus for the F-cycle results, where three iterations suffice at all resolutions, we see close to textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  Work of less than ten residual evaluations is often mentioned as the standard.} despite the low solution regularity.
+Comparing the ball and spiral results, the complexity of the coincidence set or free boundary does not appear to strongly influence FASCD performance \pef{unlike \cite{GraeserKornhuber2009}?}.
+Each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+IC) iterations; for the F-cycle results, where three iterations suffice at all resolutions, we see close to textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  Work of less than ten residual evaluations is often mentioned as the standard.} despite the low solution regularity.
 \end{example}
@@
 \end{figure}
+\pef{The spiral problem looks under-resolved; is that right?}
 
@@
 \begin{example}  \label{ex:results:plap}
-Recall Example \ref{ex:plaplacian}, the nonlinear $p$-Laplacian operator.
-Here we consider a unilateral obstacle problem, in one dimension for visualizability, for both fast ($p=1.5$) and degenerate ($p=4.0$) diffusivity cases.\footnote{``Fast'' and ``degenerate'' refer to $D=|\grad u|^{p-2}\to \infty$ and $D\to 0$ as $|\grad u|\to 0$, respectively; ellipticity is lost in the latter case.
-Well-posedness and ($p$-dependent) solution regularity for $p$-Laplacian obstacle problems is considered in \cite{ChoeLewis1991}.}  This example reveals the sometimes strong contrast between $V(1,0)$ and $V(0,1)$ results (\pr{up}, \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively), and it emphasizes the need for robust smoothers.
-
-Suppose $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, $\mathcal{K} = \{v \ge \psi\} \subset W^{1,p}(\Omega)$, and $\ell(v) = \int_\Omega g v$ where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$.
+We again consider Example \ref{ex:plaplacian} with a unilateral constraint, but now for the nonlinear $p$-Laplacian.
+We investigate both fast ($p=1.5$) and degenerate ($p=4.0$) diffusivities.\footnote{``Fast'' and ``degenerate'' refer to $|\grad u|^{p-2}\to \infty$ and $\to 0$, respectively; ellipticity is lost in the latter case.
+Well-posedness and ($p$-dependent) solution regularity for $p$-Laplacian obstacle problems is considered in \cite{ChoeLewis1991}.} We consider the problem in one dimension for ease of visualization.
+This example reveals the sometimes strong contrast between $V(1,0)$ and $V(0,1)$ results (\pr{up}, \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively), and it emphasizes the need for robust smoothers.
+
+Let $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, $\mathcal{K} = \{v \ge \psi\} \subset W^{1,p}(\Omega)$, and $\ell(v) = \int_\Omega g v$ where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$.
 The exact solution of VI \eqref{eq:vi} is easily computed for any $p$; see Figure \ref{fig:results:plap}.
-For this experiment we used a coarsest mesh of $m_0=7$ nodes and an $m_9=3073$ finest mesh.
-The default smoother uses three (fixed) iterations of RS Newton, with LU solution of the step equations.\footnote{This is an optimal $O(m_j)$ smoother in 1D.}  The coarse solver is the same but iterated to convergence.
+\pef{Write down the exact solution?}  For this experiment we used a coarsest mesh of $m_0=7$ vertices and a finest mesh with $m_9=3073$.
+The default smoother uses three (fixed) iterations of RS Newton, with LU solution of the arising linear systems.\footnote{This is an optimal $O(m_j)$ smoother in 1D.}  The coarse solver is the same but iterated to convergence.
 The initial iterate was $w=\psi$, and \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$.
@@
 \end{center}
-\caption{Visualization of one FASCD V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of $m_0=3$ nodes.
+\caption{Visualization of one FASCD V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of $m_0=3$ vertices.
 Solutions (i.e.~$w^j$, $y^j$, $z^j$) are lines while obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots.
@@
 Solvers implemented in Firedrake can exploit parallel solvers from the PETSc library \cite{Balayetal2023}.
-The above Examples, run in serial so as to fix the smoother,\footnote{The action of parallel preconditioners generally depends on the number of processes \cite[for example]{Bueler2021}.} also run efficiently in parallel.
+The above Examples, run in serial so as to fix the smoother,\footnote{The action of parallel preconditioners can depend on the number of processes \cite[for example]{Bueler2021}.} also run efficiently in parallel.
 The next Example applies a parallel solver, and Example \ref{ex:results:sia} demonstrates the parallel weak scaling \cite{Bueler2021} of FASCD.
@@
 \begin{example}  \label{ex:results:advdiff}
-We applied FASCD F-cycles to the advection-diffusion VI problem in Example \ref{ex:advectiondiffusion}.
-The domain was $\Omega=(-1,1)^d$, for $d=2,3$, diffusivity was $\eps=0.1$ (equation \eqref{eq:advectiondiffusionstrong}), and there were zero Dirichlet boundary conditions.
-The 2D problem satisfied the coercivity hypothese with a divergence-free velocity field $\bX = (7+5y,-5x)$, but in 3D we set $\bX = (7+5y,-5x,2z)$ so $\grad\cdot\bX \ne 0$.
-The discontinuous source term $\phi \in L^\infty(\Omega)$ satisfied $\phi\ge 0$ in half of the domain, with positive circular patches, while values were negative in the other half.
-While the other Examples in this Section are unilateral, here the solution was box-constrained: $0 \le u \le 1$.
+We consider the advection-diffusion VI problem in Example \ref{ex:advectiondiffusion}.
+Let $\Omega=(-1,1)^d$ for $d=2,3$, the diffusivity $\eps=0.1$ (equation \eqref{eq:advectiondiffusionstrong}), and we impose zero Dirichlet boundary conditions.
+The 2D problem satisfied the coercivity hypothesis with a divergence-free velocity field $\bX = (7+5y,-5x)$, but in 3D we set $\bX = (7+5y,-5x,2z)$ so $\grad\cdot\bX \ne 0$.
+\pef{Why?} The discontinuous source term $\phi \in L^\infty(\Omega)$ satisfied $\phi\ge 0$ in half of the domain, with positive circular patches, while values were negative in the other half.
+\pef{State where to find the full specification of the source term.} While the other Examples in this Section are unilateral, here the solution was box-constrained: $0 \le u \le 1$.
 Figure \ref{fig:results:advdiff} shows coincidence sets for the 2D problem; the 3D sets are comparable in regularity and general pattern.
@@
 
-To minimally-resolve the coincidence set we used a coarse mesh with $m_0=16^d$ nodes.
+To minimally-resolve the coincidence set we used a coarse mesh with $m_0=16^d$ vertices.
 The 2D solver applied $P_1$ elements and a single processor, but in 3D we used $Q_1$ elements (hexahedra) and $P=16$ processes.
-The smoother was two (fixed) RS Newton iterations,  with three preconditioned GMRES iterations to approximately solve the step equations.
-The additive Schwarz method (ASM) preconditioner applied incomplete factorization (ILU) preconditioning on the blocks.
+The smoother was two (fixed) RS Newton iterations,  with three preconditioned GMRES iterations to approximately solve the arising linear systems.
+The additive Schwarz method (ASM) preconditioner applied incomplete factorization (ILU) preconditioning on each core.
 The coarse solver was RS Newton and LU, iterated to convergence.
-Tolerances \id{rtol} $=10^{-5}$, \id{atol} $=10^{-9}$ were used in \eqref{eq:stoppingcriterion}.
+Tolerances \id{rtol} $=10^{-5}$ and \id{atol} $=10^{-9}$ were used in \eqref{eq:stoppingcriterion}.
 
@@
 For any of the finer mesh runs here, the FASCD F-cycle solver took less time than generating the $P_1$ mesh hierarchy.
-The local Peclet number $\text{Pe}=\|\bX(x,y,z)\|_2 L \eps^{-1}$, with $L=2$ the width of $\Omega$, varied from 40 to 260.
+The local Péclet number $\text{Pe}=\|\bX(x,y,z)\|_2 L \eps^{-1}$, with $L=2$ the width of $\Omega$, varied from 40 to 260.
 Thus the $m_J\le 121^d$ problems were significantly advection-dominated while those with $m_J\ge 481^d$ were diffusion-dominated.
-Nonetheless the smoother was the same and F-cycle iterations were nearly constant.
+\pef{I would have said you need $Pe < 1$ to be diffusion-dominated, not have $Pe = 40$?} Nonetheless the smoother was the same and F-cycle iterations were nearly constant.
 \end{example}
@@
 \begin{example}   \label{ex:results:sia}
-Recall the steady shallow ice approximation (SIA) obstacle problem, Example \ref{ex:sia}, in which the surface elevation $s$ is constrained to exceed the bedrock elevation $b$.
-For a test case we set $\Omega=(0,L)^2$ with $L=1800$ km.
-The data consists of a radial surface mass balance $a$ function \cite[equation (5.122)]{GreveBlatter2009} and a bumpy, synthetic map for $b\in C^1(\Omega)$, with an elevation range of $1800$ m.
+We consider the steady shallow ice approximation (SIA) obstacle problem presented in Example \ref{ex:sia}, in which the surface elevation $s$ is constrained to exceed the bedrock elevation $b$.
+We set $\Omega=(0,L)^2$ with $L=1800$ km.
+The data consist of a radial surface mass balance function $a$ \cite[equation (5.122)]{GreveBlatter2009} and a bumpy, synthetic map $b\in C^1(\Omega)$, with an elevation range of $1800$ m.
 A numerical solution is shown in Figure \ref{fig:results:siascene}, colored by the solution surface flow speed, i.e.~the magnitude of $\mathbf{u} = (n+1)(n+2)^{-1} \Gamma (s-b)^{n+1} |\grad s|^{n-1} \grad s$.
@@
 The coarse solver was RS Newton with LU applied redundantly over all processors.
-The coarse mesh was relatively fine, with $m_0=21^2$ nodes and 400 elements.
+The coarse mesh was relatively fine, with $m_0=21^2$ vertices and 400 cells.
 Because of the low regularity of the solution we used \id{rtol} $= 2 \times 10^{-4}$ and \id{atol} $= 10^{-8}$ in \eqref{eq:stoppingcriterion}.
@@
 \pr{fascd-vcycle} (Algorithm \ref{alg:fascd}) reduces to the classical FAS multigrid V-cycle for nonlinear PDEs \cite{Trottenbergetal2001}, which itself generalizes linear multigrid, when all inequality constraints are removed.
-In fact, suppose we remove the inequality constraints from the finest-level VI problem \eqref{eq:fe:vi} by replacing the finest-level obstacles $\underline{\gamma}^J$ and $\overline{\gamma}^J$ with very negative and positive constants, respectively, thus not constraining the solution $u^J$ or any possible iterate.
+In fact, suppose we remove the inequality constraints from the finest-level VI problem \eqref{eq:fe:vi} by replacing the finest-level obstacles $\underline{\gamma}^J$ and $\overline{\gamma}^J$ with $-\infty$ and $+\infty$ respectively, thus not constraining the solution $u^J$ or any possible iterate.
 The admissible set $\mathcal{K}^J$ of \eqref{eq:fe:vi} is then determined only by the Dirichlet boundary conditions.
 Let $\mathcal{V}_0^J \subset \mathcal{V}^J$ denote the linear subspace with zero values on $\partial_D\Omega$.
-By replacing the (now) arbitrary element $v-u^J\in\mathcal{V}_0^J$ in \eqref{eq:fe:vi} with ``$v$'' we have the finest-level problem
+Since $v-u^J\in\mathcal{V}_0^J$ in \eqref{eq:fe:vi} is now an arbitrary element, we can rewrite the finest-level problem as: find $u^J \in \mathcal{V}^J$ such that
 \begin{equation}
@@
 the FE discretization of strong-form PDE $f(u)=\ell$.
-The downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets can also be enlarged so that the corrections $y^j$ and $z^j$ in \pr{fascd-vcycle} are never constrained.\footnote{Formally we might do this by choosing constant LDCs with $\underline{\chi}^j \ll 0 \ll \overline{\chi}^j$, but so that ordering property \eqref{eq:fe:chiordering} still holds, and also with large gaps so that $\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1} \ll 0 \ll \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}$.}  Thus the corrections $y^j$ and $z^j$ are unconstrained, and lines \ref{line:vcyclegenchifinest}, \ref{line:vcyclegenchi}, and \ref{line:vcyclegenphi} can be removed from Algorithm \ref{alg:fascd}.
-
-Now the downward/upward VIs \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} are also unconstrained, and can be rewritten as weak forms $\ip{f^j(w^j + y^j)}{v} = \ip{\ell^j}{v}$ and $\ip{f^j(w^j + z^j)}{v} = \ip{\ell^j}{v}$, respectively, for all $v\in \mathcal{V}_0^j$.
+The downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets are also enlarged so that the corrections $y^j$ and $z^j$ in \pr{fascd-vcycle} are never constrained; lines \ref{line:vcyclegenchifinest}, \ref{line:vcyclegenchi}, and \ref{line:vcyclegenphi} always set $-\infty, +\infty$ by the arithmetic on $\tilde\RR$.
+The corrections $y^j$ and $z^j$ are thus unconstrained.
+
+Since the downward/upward VIs \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} are unconstrained, they can be rewritten as weak forms $\ip{f^j(w^j + y^j)}{v} = \ip{\ell^j}{v}$ and $\ip{f^j(w^j + z^j)}{v} = \ip{\ell^j}{v}$, respectively, for all $v\in \mathcal{V}_0^j$.
 Then we combine $u^j=w^j+y^j$, instead of $w^j$ and $y^j$ separately, and similarly $u^j=w^j+z^j$ when up-smoothing.
@@
 
-With these modifications, we have a much-simplified Algorithm \ref{alg:fas}.
-This FAS V-cycle appears in various places, for example Algorithm 14 of \cite{Bruneetal2015} or in section 5.3.4 of \cite{Trottenbergetal2001}.
+Modifying Algorithm \ref{alg:fascd} appropriately, we have a much-simplified Algorithm \ref{alg:fas}.
+This FAS V-cycle appears in various sources in the literature, for example \cite[Algorithm 14]{Bruneetal2015} and \cite[section 5.3.4]{Trottenbergetal2001}.
 

### 2023-07-13 paper/fascd.tex
fix underresolved spiral set ... easy
Branding: FAScd.  And address some easier PF comments.

@@
 \begin{document}
-\title[FAS for nonlinear variational inequalities]{A full approximation storage multilevel method \\ for nonlinear variational inequalities}
+\title[FAS for nonlinear variational inequalities]{A full approximation scheme multilevel method \\ for nonlinear variational inequalities}
 
@@
 \begin{abstract}
-We present the \emph{full approximation storage constraint decomposition} (FASCD) multilevel method for solving variational inequalities.
-FASCD is a common extension of both the full approximation storage (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and of the constraint decomposition (CD) method introduced by X.-C.~Tai for variational inequalities arising in inequality-constrained optimization.
-We extend the constraint decomposition idea by recognizing and exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
-Then, by using strong smoothers with work proportional to the number of unknowns on the finite element mesh level, we demonstrate that FASCD V-cycles show excellent mesh-independent convergence rates and that FASCD F-cycles (full multigrid) are optimal solvers.
-The example problems include unilateral and bi-obstacle variational inequality problems with symmetric linear, nonsymmetric linear, and nonlinear differential operators; the examples demonstrate very good convergence even on problems where other multilevel VI solvers struggle.
-\pef{I want to say it's better than monotone multigrid. Is this too subtle/too much?}
-
+We present the \emph{full approximation scheme constraint decomposition} (FAScd) multilevel method for solving variational inequalities (VIs).
+FAScd is a common extension of both the full approximation scheme (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and the constraint decomposition (CD) method introduced by X.-C.~Tai for VIs arising in optimization.
+We extend the CD idea by recognizing and exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
+Then, by applying strong smoothers, with work proportional to the number of unknowns on a finite element mesh level, we demonstrate that FAScd V-cycles exhibit excellent mesh-independent convergence rates and that FAScd F-cycles (full multigrid) are optimal solvers.
+The example problems include unilateral and bi-obstacle VI problems with symmetric linear, nonsymmetric linear, and nonlinear differential operators.
+Our examples demonstrate very good convergence even on problems where other multilevel VI solvers struggle.
+%\pef{I want to say it's better than monotone multigrid. Is this too subtle/too much?}
+% I vote for not saying this in the abstract.
+It is hard to say which aspect of FAScd is more important to beating monotone multigrid: stronger smoother or improved up-smooth CD.
+Probably the former, but for now we don't have any illuminating analysis.
 \end{abstract}
@@
 These methods have been shown to converge for multilevel and domain decompositions in a finite element (FE) context; in the multilevel case, the CD method has been shown to have almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
-In this paper we extend the multilevel CD method by replacing the coarser-level corrections with a full approximation storage (FAS; \cite{Brandt1977,Bruneetal2015}) approach, and we apply this new solver to nonlinear VI problems which do not arise from minimization.
-\pef{Note I've changed ``coarse-level corrections'' to ``coarser-level corrections'', to imply we're not just operating only on the coarsest grids, but also on intermediate grids.}
-
-Specifically, our \emph{full approximation storage constraint decomposition} (FASCD) method extends the algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} in the following directions:
+In this paper we extend the multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach, and we apply this new solver to nonlinear VI problems which do not arise from minimization.
+%\pef{Note I've changed ``coarse-level corrections'' to ``coarser-level corrections'', to imply we're not just operating only on the coarsest grids, but also on intermediate grids.}
+% Cool.
+
+Specifically, our \emph{full approximation scheme constraint decomposition} (FAScd) method extends the algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} in the following directions:
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
@@
 \item We implement V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}).
-\item By using a Newton-type active set method as a smoother, with a fixed number of Krylov iterations per step so that the work of the smoother is proportional to the number of unknowns on the given level, we demonstrate that F-cycle FASCD is an optimal solver.
+\item By using a Newton-type active set method as a smoother, with a fixed number of Krylov iterations per step so that the work of the smoother is proportional to the number of unknowns on the given level, we demonstrate that F-cycle FAScd is an optimal solver.
 \pef{Clarify/limit which classes of problems we're claiming this for.}
@@
 
-In our experiments, FASCD V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for the corresponding (unconstrained) PDE problems.
+In our experiments, FAScd V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for the corresponding (unconstrained) PDE problems.
 The F-cycle results in Section \ref{sec:results} show nearly-textbook multigrid efficiency, namely solutions within discretization error in work comparable to few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}, for classical Laplacian obstacle problems, $p$-Laplacian obstacle problems, and nonnegativity-constrained advection-diffusion equations.
@@
 
-The iterates from FASCD are admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
+The iterates from FAScd are admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
 For example, we solve an ice sheet geometry problem (Examples \ref{ex:sia}, \ref{ex:results:sia}) in which the ice flow formulas are only meaningful for surface elevation iterates that do not penetrate the bedrock\footnote{More generally, fluid-layer problems, with a non-negative thickness constraint, often require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation \cite{Bueler2021conservation}.}.
 Note that non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural and invasive modifications of the operator formula in such a fluid-layer context.
-FASCD solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, with a mesh-independent number of F-cycle iterations.
+FAScd solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, with a mesh-independent number of F-cycle iterations.
 
@@
 Here we propose an apparently new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD.
-Section \ref{sec:vcycle} then states the FASCD V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, F-cycles, and our smoother choice.
-Section \ref{sec:results} gives FASCD numerical results from a parallel Firedrake \cite{Rathgeberetal2016} implementation for the problems introduced in Section \ref{sec:vi}.
-The work concludes with discussion and outlook in Section \ref{sec:discussion}, and an appendix on how FASCD reduces to classical FAS multigrid for PDEs.
+Section \ref{sec:vcycle} then states the FAScd V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, F-cycles, and our smoother choice.
+Section \ref{sec:results} gives FAScd numerical results from a parallel Firedrake \cite{Rathgeberetal2016} implementation for the problems introduced in Section \ref{sec:vi}.
+The work concludes with discussion and outlook in Section \ref{sec:discussion}, and an appendix on how FAScd reduces to classical FAS multigrid for PDEs.
 
@@
 Let $\cV$ be a real and reflexive Banach space with norm $\|\cdot\|$ and topological dual space $\cV'$.
-Denote the dual pairing of $\phi \in \cV'$ and $v\in\cV$ by $\ip{\phi}{v} = \phi(v)$, and define the (Banach space) norm on $\cV'$ by $\|\phi\|_{\cV'} = \sup_{\|v\|=1} |\ip{\phi}{v}|$.
+Denote the dual pairing of $\phi \in \cV'$ and $v\in\cV$ by $\ip{\phi}{v} = \phi(v)$, and define the (Banach space) norm on $\cV'$ by $\|\phi\|' = \sup_{\|v\|=1} |\ip{\phi}{v}|$.
 Let $\cK \subset \cV$ be a nonempty, closed, and convex subset, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.
@@
 Consider a divergence-free velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, and assume that the flow is outward on the Neumann boundary: $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.
-For $u,v \in \cV$, $\eps>0$, and $\phi \in L^\infty(\Omega)$, define
+For $u,v \in \cV$, $\eps>0$, and $\phi \in \cV'$, define
 \begin{equation}
@@
 \end{equation}
-\pef{Why is $\phi \in L^\infty(\Omega)$? If that is the case, then should we have an $L^2$ inner product in the last term, rather than a dual pairing?}
+%\pef{Why is $\phi \in L^\infty(\Omega)$? If that is the case, then should we have an $L^2$ inner product in the last term, rather than a dual pairing?}
+% Fixed.  \phi can be in the dual space
 With this operator, consider VI \eqref{eq:vi} with $\ell = 0$ for any closed and convex $\cK \subset \cV$.
@@
 \end{equation}
-On the other hand, it is easy to see that $|\ip{f(u)}{v}| \le (\eps + \|\bX\|_\infty + \|\phi\|_\infty) \|u\| \|v\|$, thus that $f:\cK \to \cV'$ is continuous.
+On the other hand, it is easy to see that $|\ip{f(u)}{v}| \le \left((\eps + \|\bX\|_\infty) \|u\| + \|\phi\|'\right) \|v\|$, thus that $f:\cK \to \cV'$ is continuous.
 Lemma \ref{lem:advectionskew} says that the bilinear form $b(u,v)$ is skew-symmetric up to a nonnegative term.
@@
 It is nonlinear and not of optimization type.
-Related VI problems arise for any fluid layer which is subject to processes that can remove the fluid \cite{Bueler2021conservation}.
-
-\begin{example}  \label{ex:sia}  Let $\Omega \subset \RR^2$ be a fixed region of land, and suppose $b \in C^1(\Omega)$ is the given bedrock elevation.
-Let $a \in L^\infty(\Omega)$ denote a given ``surface mass balance'' function, the annually-averaged rate of ice accumulation (as snow) minus melt and runoff.
-Then the surface elevation $s\in C(\Omega)$ of a steady-state, isothermal, and non-sliding ice sheet (glacier) approximately satisfies the \emph{shallow ice approximation} (SIA) \cite{GreveBlatter2009} problem: find $s \in \mathcal{K} = \{s\ge b\}$ such that
+Related VI problems arise for any fluid layer which is subject to boundary processes that can remove fluid mass \cite{Bueler2021conservation}.
+
+\begin{example}  \label{ex:sia}  Let $\Omega \subset \RR^2$ be a fixed region of land, and suppose $b \in C^1(\Omega)$ is a given bedrock elevation.
+Let $a \in L^2(\Omega)$ denote a given ``surface mass balance'' function, the annually-averaged rate of ice accumulation (as snow) minus melt and runoff.
+Then the surface elevation $s\in C(\Omega)$ of a steady-state, isothermal, and non-sliding ice sheet (glacier) approximately satisfies the following \emph{shallow ice approximation} (SIA) \cite{GreveBlatter2009} VI problem: find $s \in \mathcal{K} = \{s\ge b\}$ such that
 \begin{equation}
@@
 If the bed is flat ($b=0$) then a power transformation converts the problem into $p$-Laplacian form \eqref{eq:plaplacian} with $p=n+1$.
-For general beds $b$ less is understood about this operator, but existence holds for \eqref{eq:siavi} \cite{JouvetBueler2012}.
-The theory suggests that a power of the solution lives in a Sobolev space, namely $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.
-\pef{`suggests' is quite weak; can we use a stronger word?} This theory, and also observations of ice sheets, shows that $|\grad s|$ is generally unbounded as one approaches the free boundary, the \emph{ice margin}, from the icy side (where $s>b$).
+For general beds $b$ less is understood about this operator, but existence holds for \eqref{eq:siavi} \cite{JouvetBueler2012} and a power of the solution lives in a Sobolev space: $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.
+This theory, and also observations of ice sheets, shows that $|\grad s|$ is generally unbounded as one approaches the free boundary from the $s>b$ side.
 \end{example}
-
-Numerical results are shown in Section \ref{sec:results} for applications of FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} to these Examples.
-
-\pef{I'm not sure starting with the example problems we want to solve is the right move. Let me read more and think about it.}
+%\pef{`suggests' is quite weak; can we use a stronger word?}
+% Agreed.  Last sentences made brief and direct.
+
+Numerical results are shown in Section \ref{sec:results} for applications of FAScd Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} to these Examples.
+%\pef{I'm not sure starting with the example problems we want to solve is the right move. Let me read more and think about it.}
+% I think it is o.k.
 
@@
 Associated to a constraint decomposition \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} are certain iterative methods \cite{Tai2003,Xu1992}.
-Let $0\le \alpha \le 1$ be the \emph{damping parameter}.\footnote{FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} are presented without damping, and indeed it seems to be unnecessary.
+Let $0\le \alpha \le 1$ be the \emph{damping parameter}.\footnote{FAScd Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} are presented without damping, and indeed it seems to be unnecessary.
 However, the convergence of the more general CD iterations appears to depend upon it.}  Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below, discussed next, solve smaller VI problems over each subset $\cK_i$.
@@
 the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$, and then adds-back the subset solution $\hat w_i \in \mathcal{K}_i$.
-\pef{Could remove this explanation, but it's nice to keep it in if we can fit it}
+%\pef{Could remove this explanation, but it's nice to keep it in if we can fit it}
+% Proposing to keep for now.
 
@@
 When $f$ has the locality of Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over basis function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}  However, for general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of CD-type iterations.
-Section \ref{sec:vcycle} constructs a multiplicative multilevel CD iteration for a general operator $f:\cK\to\cV'$ by applying the full approximation storage (FAS) idea of Brandt \cite{Brandt1977}.
+Section \ref{sec:vcycle} constructs a multiplicative multilevel CD iteration for a general operator $f:\cK\to\cV'$ by applying the full approximation scheme (FAS) idea of Brandt \cite{Brandt1977}.
 
@@
 
-\section{FASCD V-cycle} \label{sec:vcycle}
+\section{FAScd V-cycle} \label{sec:vcycle}
 
 The V-cycle algorithm in this section applies smoothers in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.
-It extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear full approximation storage (FAS) approach \cite{BrandtLivne2011}.
+It extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}.
 In fact, it reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).
@@
 \end{center}
-\caption{The FASCD V-cycle computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.}
+\caption{The FAScd V-cycle computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.}
 \label{fig:fascdvcycle}
@@
 \end{pseudo}
-\caption{The FASCD V-cycle is an iteration for solving VI problem \eqref{eq:fe:vi}.  Here $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
+\caption{The FAScd V-cycle is an iteration for solving VI problem \eqref{eq:fe:vi}.  Here $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
 \label{alg:fascd}
@@
 
-In this section we consider the implementation of FASCD methods.  We address convergence criteria, the F-cycle extension, and an active set Newton method used as a smoother and coarse solver.
+In this section we consider the implementation of FAScd methods.  We address convergence criteria, the F-cycle extension, and an active set Newton method used as a smoother and coarse solver.
 
@@
 \end{pseudo}
-\caption{The FASCD F-cycle for solving VI problem \eqref{eq:fe:vi}.}
+\caption{The FAScd F-cycle for solving VI problem \eqref{eq:fe:vi}.}
 \label{alg:fascd-fcycle}
@@
 Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}, allowing $P_1$ or $Q_1$ elements and based on standard nested mesh refinement.
-Here we show numerical results for the example problems presented in Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
+Here we show numerical results for the example problems presented in Section \ref{sec:vi}, with a focus on the scaling of FAScd iterations as the mesh resolution increases.
 Our main observations are that each V-cycle dramatically reduces the residual norm, and that very few F-cycles solve most examples to within discretization error.
@@
 Noting that for F-cycles we are counting the number of V-cycles after the ramp (Section \ref{sec:implementation}), we see that a constant number of cycles solves the problem.
-Comparing the ball and spiral results, the complexity of the coincidence set or free boundary does not appear to strongly influence FASCD performance \pef{unlike \cite{GraeserKornhuber2009}?}.
+Comparing the ball and spiral results, the complexity of the coincidence set or free boundary does not appear to strongly influence FAScd performance \pef{unlike \cite{GraeserKornhuber2009}?}.
 Each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+IC) iterations; for the F-cycle results, where three iterations suffice at all resolutions, we see close to textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  Work of less than ten residual evaluations is often mentioned as the standard.} despite the low solution regularity.
@@
 %REGENERATE using git@bitbucket.org:pefarrell/fascd.git, in examples/
-%$ timer python3 drape.py -prob ball|spiral -levs 9 -o ball|spiral.pvd -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
+%$ python3 drape.py -prob ball -levs 9 -o ball -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
+%$ mpiexec -n 16 python3 drape.py -prob spiral -levs 10 -o spiral.pvd -fascd_levels_snes_converged_reason -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
 %$ paraview ball|spiral.pvd
-%use preset "xray" and invert colors
+%use preset "xray" and invert colors; for spiral.pvd, show lgap = u - lb as black if above 0.02; for ball i used 0.001
 \begin{figure}[ht]
@@
 \end{figure}
-\pef{The spiral problem looks under-resolved; is that right?}
+%\pef{The spiral problem looks under-resolved; is that right?}
+% Fixed.
 
@@
 \bigskip
-\caption{Iterations for FASCD V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}.  Smoother-only runs marked \XX\xspace were not attempted.}
+\caption{Iterations for FAScd V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}.  Smoother-only runs marked \XX\xspace were not attempted.}
 \label{tab:results:classical}
@@
 \end{center}
-\caption{Visualization of one FASCD V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of $m_0=3$ vertices.
+\caption{Visualization of one FAScd V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of $m_0=3$ vertices.
 Solutions (i.e.~$w^j$, $y^j$, $z^j$) are lines while obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots.
@@
 The above Examples, run in serial so as to fix the smoother,\footnote{The action of parallel preconditioners can depend on the number of processes \cite[for example]{Bueler2021}.} also run efficiently in parallel.
-The next Example applies a parallel solver, and Example \ref{ex:results:sia} demonstrates the parallel weak scaling \cite{Bueler2021} of FASCD.
+The next Example applies a parallel solver, and Example \ref{ex:results:sia} demonstrates the parallel weak scaling \cite{Bueler2021} of FAScd.
 
@@
 Despite the hint of growing iterations under mesh refinement, a few iterations sufficed up to memory-limited resolutions.
-For any of the finer mesh runs here, the FASCD F-cycle solver took less time than generating the $P_1$ mesh hierarchy.
+For any of the finer mesh runs here, the FAScd F-cycle solver took less time than generating the $P_1$ mesh hierarchy.
 The local Péclet number $\text{Pe}=\|\bX(x,y,z)\|_2 L \eps^{-1}$, with $L=2$ the width of $\Omega$, varied from 40 to 260.
@@
 
-We applied FASCD F-cycles using $Q_1$ elements and a smoother of four (fixed) iterations of RS Newton (Section \ref{sec:implementation}), with quadratic backtracking line search.
+We applied FAScd F-cycles using $Q_1$ elements and a smoother of four (fixed) iterations of RS Newton (Section \ref{sec:implementation}), with quadratic backtracking line search.
 The step equations were approximately solved, in parallel, by three (fixed) iterations of preconditioned GMRES, again with a ASM+ILU preconditioner.
@@
 \bigskip
-\caption{Iterations and run time (wall clock, in seconds) for parallel FASCD F-cycle solutions of Example \ref{ex:sia} on $P$ processes.}
+\caption{Iterations and run time (wall clock, in seconds) for parallel FAScd F-cycle solutions of Example \ref{ex:sia} on $P$ processes.}
 \label{tab:results:siaweak}
@@
 
-FIXME the fact that FASCD reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}) suggests it is a completely different strategy from PFAS in \cite{BrandtCryer1983}
+FIXME the fact that FAScd reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}) suggests it is a completely different strategy from PFAS in \cite{BrandtCryer1983}
 
@@
 These improved methods modify the nodal basis functions.
-By contrast FASCD makes essentially no assumptions about the FE discretization.
+By contrast FAScd makes essentially no assumptions about the FE discretization.
 Acceleration via basis-level manipulations is a topic for future research.
@@
 (See the start of \cite{GraeserKornhuber2009} section 5.2.)  I think that making these modifications in Firedrake, so we could try monotone multigrid V-cycles, would be a nightmare.
-FASCD does not attempt the monotone multigrid idea at all.
-
-By contrast, FASCD is essentially just the \cite{Tai2003} CD strategy, but (1) without assuming an objective, and (2) with a coarse correction which is FAS in the nonlinear PDE case.
+FAScd does not attempt the monotone multigrid idea at all.
+
+By contrast, FAScd is essentially just the \cite{Tai2003} CD strategy, but (1) without assuming an objective, and (2) with a coarse correction which is FAS in the nonlinear PDE case.
 If you apply \cite{Tai2003} CD to a nonquadratic (e.g.
@@
 E.g.~a fixed number of iterations of vinewtonrsls+gmres+ilu or whatever.
-The role of the FASCD is simply to generate coarse VI problems which, once solved, make helpful admissible additions to the eventual V-cycle update.
-
-So the difference between FASCD and \cite{GraeserKornhuber2009} are: (1) FAS coarse corrections, (2) no attempt to do monotone/truncated multigrid which would be a mess (I think) across the FE infrastructure, (3) clarity on what sets you can use in $V(\nu_1,\nu_2)$ cycles, with bigger up-smoother sets in particular, and most importantly (4) implementation within the mature Firedrake/PETSc library framework where one can easily hook up an active-set Newton step as a smoother.
+The role of the FAScd is simply to generate coarse VI problems which, once solved, make helpful admissible additions to the eventual V-cycle update.
+
+So the difference between FAScd and \cite{GraeserKornhuber2009} are: (1) FAS coarse corrections, (2) no attempt to do monotone/truncated multigrid which would be a mess (I think) across the FE infrastructure, (3) clarity on what sets you can use in $V(\nu_1,\nu_2)$ cycles, with bigger up-smoother sets in particular, and most importantly (4) implementation within the mature Firedrake/PETSc library framework where one can easily hook up an active-set Newton step as a smoother.
 \end{comment}

### 2023-07-13 paper/fascd.tex
address more PF comments

@@
 Section \ref{sec:cd} reviews the basics of CDs, as proposed by Tai \cite{Tai2003}, and their induced iterations.
-Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, and then Section \ref{sec:cdmultilevel} presents defect constraints generated via monotone restriction operators \cite{GraeserKornhuber2009}.
+Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, and then Section \ref{sec:cdmultilevel} presents defect constraints generated via monotone injection operators \cite{GraeserKornhuber2009}.
 Here we propose an apparently new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD.
@@
 \end{equation}
-the finest-level \emph{constraint set} of admissible FE functions.
-This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.
+the finest-level \emph{constraint set} of admissible FE functions. This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.
 %As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fe:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J$ are infinite; there is no lower or upper constraint, respectively, at such vertices.
@@
 (nodal) injection             & $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ & \,\checkmark \\
-maximum restriction           & $\maxR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ & \ding{55} \\
-minimum restriction           & $\minR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ & \ding{55} \\
+maximum injection           & $\maxR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ & \ding{55} \\
+minimum injection           & $\minR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ & \ding{55} \\
 \bottomrule
@@
 
-\pef{I don't like the name ``monotone restrictions''.
+%\pef{I don't like the name ``monotone restrictions''.
 The word restriction is strongly associated with operating on the dual space.
 I would prefer to call them ``monotone injections'', since that is the matching operator.}
-The monotone restrictions $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to nodal functions $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the \emph{stars} of the coarse vertices).
+% Agreed.  "monotone restriction" is used in GraeserKornhuber2009, but happy not to follow.
+
+The monotone injections $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to nodal functions $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the \emph{stars} of the coarse vertices).
 These operators are constructed to enforce the property
@@
 
-Descending coarser levels, $j=J-1,\dots,1,0$, we apply the monotone restrictions to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:
+Descending coarser levels, $j=J-1,\dots,1,0$, we apply the monotone injections to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:
 \begin{equation}
@@
 \includegraphics[width=0.65\textwidth]{fixfigs/chiphilevels.pdf}
-\caption{Unilateral LDCs $\underline{\chi}^j$ are generated from $\underline{\chi}^J = \underline{\gamma}^J - w^J$ using maximum restriction: $\underline{\chi}^j = \maxR \underline{\chi}^{j+1}$.}
+\caption{Unilateral LDCs $\underline{\chi}^j$ are generated from $\underline{\chi}^J = \underline{\gamma}^J - w^J$ using maximum injection: $\underline{\chi}^j = \maxR \underline{\chi}^{j+1}$.}
 \label{fig:chiphilevels}
@@
 
-The V-cycle algorithm in this section applies smoothers in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.
-It extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}.
-In fact, it reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).
+The V-cycle applies smoothers in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.
+It extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}, and in fact it reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).
 
 Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ by incorporating information from each coarser FE subspace $\mathcal{V}^j$.
-By nesting \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$,\footnote{Nonetheless we use prolongation $P$ to indicate where the computer representation changes.} and are constructed so that the corrected iterate remains admissible.
-Proceeding inductively, suppose $y^i \in \mathcal{D}^i$ for $i=J,\dots,j+1$ \pef{I changed the sign here from $j-1$ to $j+1$, double-check} are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).
+By nesting \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$.\footnote{Nonetheless we use prolongation $P$ to indicate where the computer representation changes.}  The corrections are constructed, inductively, so that iterates remain admissible at every level.
+Suppose $y^i \in \mathcal{D}^i$, for $i=J,\dots,j+1$, are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).
 By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.
 Smoothing then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.
-At the bottom the coarsest-level VI is then solved, resulting in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.
+Solving the coarsest-level VI results in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.
 Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using \eqref{eq:fe:upwardsuminclusion}.
 Continuing in this ``telescoping'' way (Algorithm \ref{alg:basiccd-tele}), up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.
-Then the next admissible iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.
+Then the next finest-level, admissible iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.
 
@@
 Observe that injection $\iR$ coarsens the (primal) iterates---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---but canonical restriction $R$ coarsens (dual) linear functionals; recall Table \ref{tab:transfers}.
-
-On descent, at levels $j=J,J-1,\dots,1$, we solve the following VI for $y^j \in \mathcal{D}^j$:
+At levels $j=J,J-1,\dots,1$, we solve the following VI for $y^j \in \mathcal{D}^j$:
 \begin{equation}
@@
 \end{equation}
-for which the initial iterate is $y_{(0)}^j=0$.
-(Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)  On ascent, $j=0,1,\dots,J$, the same VI is solved but for $z^j \in \mathcal{U}^j$:
+for which the initial iterate is $y_{(0)}^j=0$.  (Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)
+
+On ascent, $j=0,1,\dots,J$, the same VI is solved but for $z^j \in \mathcal{U}^j$:
 \begin{equation}
@@
 
-Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle.\footnote{We could write ``$V(\id{down},\id{up})$'', but we prefer to emphasize the number of subsets used on each level rather than the number of smoother sweeps in a given set.
-\pef{I don't understand what the number of subsets used on each level means.}}  If \id{up} $=0$ then it becomes a $V(1,0)$ cycle and if \id{down} $=0$ then it is a $V(0,1)$ cycle; in the former case it generalizes Algorithm 4.7 in \cite{GraeserKornhuber2009} by allowing nonlinear operators $f$, any smoother, and applying FAS-type corrections.
-In unilateral cases, one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} over CD \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over incomplete CD \eqref{eq:fe:upwardsuminclusion}; in this sense $V(1,1)=V(1,0)+V(0,1)$.
+Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle with \id{down}, \id{up} smoother iterations on each level.
+For \id{up} $=0$ it becomes a $V(1,0)$ cycle, and for \id{down} $=0$ a $V(0,1)$ cycle.
+The $V(1,0)$ cycle is similar to Algorithm 4.7 in \cite{GraeserKornhuber2009}, but generalizes it by applying FAS-type corrections and allowing any smoother or nonlinear operator $f$.
+One iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} (Algorithm \ref{alg:basiccd-mult}) over CD\footnote{Strictly speaking, this statement is restricted to unilateral cases.} \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Algorithm \ref{alg:basiccd-tele}) over incomplete CD \eqref{eq:fe:upwardsuminclusion}, and in this sense $V(1,1)=V(1,0)+V(0,1)$.
 
@@
 
-In this section we consider the implementation of FAScd methods.  We address convergence criteria, the F-cycle extension, and an active set Newton method used as a smoother and coarse solver.
-
-Because the problem is inequality-constrained, repeated application of Algorithm \ref{alg:fascd} to solve the finite-dimensional, box-constrained VI \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.
-To construct appropriate convergence criteria for the discretization, note \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}.
-Let $x_p \in \mathcal{T}^J$ be a finest-level node, and $\psi_p(x)$ the corresponding basis function.
+Implementations of FAScd methods must choose a concrete convergence criterion.
+In this section we also construct the F-cycle extension, and we describe the active set Newton method we will use as a smoother and coarse solver.
+
+%Let $x_p \in \mathcal{T}^J$ be a finest-level node, and $\psi_p(x)$ the corresponding basis function.
 Let $\mathcal{X}^J=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}^J$.
@@
 Then problem \eqref{eq:fe:vi}, which would be expressed VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap \mathcal{X}^J\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP for the solution $u^J \in \mathcal{K}^J$, but with additional equality constraints:
-\begin{align}
-\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J - I_D^\top \lambda \label{eq:fe:mcp} \\
-I_D u^J &= g_D^J\big|_{\partial_D\Omega} \notag
-\end{align}
-for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
-\pef{I wonder if this definition will confuse people, because they won't know how to plug in $\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ into your definition of $\perp$ ($\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ is not in $\RR^n$).
+%\begin{align}
+%\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J - I_D^\top \lambda \label{eq:fe:mcp} \\
+%I_D u^J &= g_D^J\big|_{\partial_D\Omega} \notag
+%\end{align}
+%for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
+%\pef{I wonder if this definition will confuse people, because they won't know how to plug in $\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ into your definition of $\perp$ ($\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ is not in $\RR^n$).
 Maybe it would be easier to just use the spelled-out version below, and not use this notation?}
-
-By definition, MCP \eqref{eq:fe:mcp} says that certain node-wise conditions hold for $u^J$ at $x_p \in \mathcal{T}^J$.
+% Agreed.  Greatly simplified
+
+Because the problem is inequality-constrained, repeated application of Algorithm \ref{alg:fascd} to solve the finite-dimensional, box-constrained VI \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.
+To construct appropriate convergence criteria for the discretization, note \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}.
+By definition, this says that certain node-wise conditions hold for $u^J$ at $x_p \in \mathcal{T}^J$.
 Let $r(u^J)_p = \ip{f^J(w^J)-\ell^J}{\psi_p}$ be the pointwise (ordinary) residual.
@@
 \begin{itemize}
-\item $u^J(x_p)=g_D^J(x_p)$ and $\lambda_p=r(u^J)_p$.
+\item $u^J(x_p)=g_D^J(x_p)$.
 \end{itemize}
-The cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, \emph{pinched} (\emph{both active}), and \emph{boundary}, respectively.
-Where \emph{pinched} applies, $r(u^J)_p$ can have any real value.
-
-Thus \eqref{eq:fe:mcp} allows us to construct expressions whose norms should go below tolerance.
-First, interpreting the MCP directly, for $w^J \in \mathcal{K}^J$ we define the finest-level \emph{nodal complementarity residual} $\rNC(w^J)$ from the five cases:
+These five cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, \emph{pinched} (\emph{both active}), and \emph{boundary}, respectively.
+Where \emph{pinched} or \emph{boundary} apply, $r(u^J)_p$ can have any value.
+
+This MCP interpretation allows us to construct expressions whose norms should go below tolerance.
+For $w^J \in \mathcal{K}^J$ we define the finest-level \emph{nodal complementarity residual} $\rNC(w^J)$ from the five cases:
 \begin{equation}
@@
 
-In summary, $\rNC : \mathcal{K}^J \to \RR^{m_J}$ is a computable nonlinear map defined using the discretization $f^J$ and the vectors $\ell^J$, $\underline{\gamma}^J$, and $\overline{\gamma}^J$.
-A value $\rNC(w^J)_p$ is nonzero only at vertices where \eqref{eq:fe:mcp} is violated, either because $r(w^J)_p$ remains nonzero where the constraints are inactive, or because it has the wrong sign at an active node, or because the boundary condition is not satisfied.
-
-\pef{I think this is all fairly standard and we could abbreviate this section substantially.}
-
-Direct implementation of \eqref{eq:rNC} is, however, sensitive to rounding error because $\rNC$ is a discontinuous function of $w^J(x_p)$.
-For example, if $w^J(x_p)$ exceeds $\underline{\gamma}^J(x_p)$ by $10^{-16}$, then, instead of returning the minimum of $r(w^J)_p$ and zero (the second case), the first case applies, so $\rNC(w^J)_p$ might be large and positive.
-This situation would falsely suggest that the iteration is far from converging.
-On the other hand, formula \eqref{eq:rNC} applies as stated if $\underline{\gamma}^J(x_p)=-\infty$ or $\overline{\gamma}^J(x_p)=+\infty$.
-Compare \eqref{eq:rSS} in this case, wherein the difference $w^J(x_p) - \underline{\gamma}^J(x_p)$ is computed.
-
-We will instead measure convergence using a \emph{semi-smooth residual} based on the Fischer-Burmeister function \cite{BensonMunson2006,Ulbrich2011} on $a,b\in\RR$:
+%\pef{I think this is all fairly standard and we could abbreviate this section substantially.}
+% Yes.
+
+The computable nonlinear map $\rNC : \mathcal{K}^J \to \RR^{m_J}$ in \eqref{eq:rNC} is defined using only the finest-level objects $f^J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$.
+A value $\rNC(w^J)_p$ is nonzero only at vertices where the MCP is violated, and the formula applies as stated if $\underline{\gamma}^J(x_p)=-\infty$ or $\overline{\gamma}^J(x_p)=+\infty$.
+However, direct implementation of \eqref{eq:rNC} is sensitive to rounding error because $\rNC$ is a discontinuous function of $w^J(x_p)$.
+For example, if $w^J(x_p)$ exceeds $\underline{\gamma}^J(x_p)$ by $10^{-16}$, then evaluation of $\rNC$ would falsely suggest that the iteration is far from converging.
+Thus we instead measure convergence using the Fischer-Burmeister function \cite{BensonMunson2006,Ulbrich2011} on $a,b\in\RR$:
 \begin{equation}
@@
 For convenience denote $\underline{g}_p = w^J(x_p) - \underline{\gamma}^J(x_p)$ and $\overline{g}_p = \overline{\gamma}^J(x_p) - w^J(x_p)$, the nonnegative gaps at $x_p$.
-Now define
+Now define the \emph{semi-smooth residual}
 \begin{equation}
@@
 \end{equation}
-as the stopping criterion.  %A step tolerance, $\|w^{J,k} - w^{J,k-1}\|_2 < \id{stol}$ or similar, could also be used, but for the results in Section \ref{sec:results} we did not find it necessary.
-
+as the stopping criterion.
 In Section \ref{sec:results} we will show that iteration of the V-cycle Algorithm \ref{alg:fascd}, starting with some finest-mesh initial iterate $w^{J,0}$ and repeated until \eqref{eq:stoppingcriterion} holds, is an effective solver.
+
 However, an F-cycle (Figure \ref{fig:fcycle}), also called a \emph{full multigrid cycle}, is even better.
-The well-known principle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate.
-Then an admissible initial iterate on the next-finer mesh is generated via prolongation and truncation\pef{Clarify truncation}, and so on.
-During this ``ramp'' up one approximately solves on each level by using a fixed number of V-cycles, each down to the coarsest mesh.
-From the finest level one does full-depth V-cycles to convergence.
+The well-known principle of such a cycle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate.
+Then an admissible initial iterate on the next-finer mesh is generated via prolongation, and so on to finer levels.
+(In the VI context one must also truncate after the prolongation; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fcycle}.)  During this ``ramp'' one approximately solves on each level by using a fixed number of V-cycles, each down to the coarsest mesh, and then one does full-depth V-cycles from the finest level, to convergence.
+%\pef{Clarify truncation}
+% Attempted.
 
@@
 \end{center}
-\caption{F-cycle Algorithm \ref{alg:fascd-fcycle} initially restricts the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray edges).
+\caption{F-cycle Algorithm \ref{alg:fascd-fcycle} initially restricts and injects the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray).
 Then it prolongs and truncates (doubled edges) to create an admissible initial iterate on the next-finer level.
@@
 This F-cycle approach is described in Algorithm \ref{alg:fascd-fcycle}.
-Observe that while injection of the original fine-level constraints $\underline{\gamma}^J, \overline{\gamma}^J$ is \emph{not} used for V-cycles, wherein the LDCs use an existing admissible iterate (Section \ref{sec:cdmultilevel}), when combined with truncation into $\mathcal{K}^j=\{v \in \mathcal{V}^j \,:\, \underline{\gamma}^j \le v \le \overline{\gamma}^j, \, v|_{\partial\Omega} = g_D\}$, injection \emph{is} appropriate for the ramp in an F-cycle\pef{Explain why?}; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo}.
-One may generate the coarsest-level initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation down to coarsest level; compare line \ref{line:fcyclecoarsestinitial}.
+Injection is used to generate the $\underline{\gamma}^j,\overline{\gamma}^j$ on each level.\footnote{Contrast the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, based on an existing admissible iterate.}  Regarding line \ref{line:fcyclecoarsestinitial}, one may generate an initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation of a finest-level iterate, if one is available.
+%\pef{Explain why?}
+% Mostly not done.  But restated for clarity?
 
@@
 
-A large space of smoothers could be explored; in Section \ref{sec:results} we apply a fixed number of iterations of the active-set (reduced-space) Newton method from \cite{BensonMunson2006}, the \texttt{VINEWTONRSLS} solver in PETSc \cite{Balayetal2023}, denoted as \emph{RS Newton}.
-We use it with a fixed number of preconditioned Krylov iterations to solve the arising linear systems.
+A large space of smoothers could be explored, but in Section \ref{sec:results} we apply a fixed number of iterations of the active-set (reduced-space) Newton method from \cite{BensonMunson2006}, the \texttt{VINEWTONRSLS} solver in PETSc \cite{Balayetal2023}, denoted as \emph{RS Newton}.
+A fixed number of preconditioned Krylov iterations are applied to approximately solve the arising linear systems.
 For the symmetric Laplacian in Example \ref{ex:results:classical} we apply conjugate gradients (CG) with incomplete Cholesky (IC) preconditioning, but for other operators we apply the generalized minimum residual (GMRES) method with incomplete LU (ILU) preconditioning.
-This requires $O(m_j)$ work per iteration.
-The coarse solver is also RS Newton, but iterated to convergence and with direct solvers applied to solve the arising linear systems.
+This smoother requires $O(m_j)$ work per iteration.
+The coarse solver is also RS Newton, but iterated to convergence, with a direct solver for the arising linear systems.
 

### 2023-07-13 paper/fascd.tex
work on draft of discussion section

@@
 
-FIXME the observation that up-smoothing is more efficient seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}
-
-FIXME smoother could be Newton semi-smooth \cite{BensonMunson2006}; smoother could be interior point method, etc.; smoother could use GMG or AMG on step equations but this is not seen to be necessary in applications so far
-
-FIXME the fact that FAScd reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}) suggests it is a completely different strategy from PFAS in \cite{BrandtCryer1983}
-
-FIXME desired extension to nonlocal variational inequalities not tried here
-
 FIXME For the classical obstacle problem, based on the linear, symmetric Laplacian operator, certain multilevel techniques, such as the truncated monotone multigrid method \cite{GraeserKornhuber2009,Kornhuber1994}, are known to improve performance relative to the original multilevel CD method \cite{Tai2003}.
@@
 Acceleration via basis-level manipulations is a topic for future research.
+
+FIXME So the difference between FAScd and \cite{GraeserKornhuber2009} are: (1) FAS coarse corrections, (2) no attempt to do monotone/truncated multigrid which would be a mess (I think) across the FE infrastructure, (3) clarity on what sets you can use in $V(\nu_1,\nu_2)$ cycles, with bigger up-smoother sets in particular, and most importantly (4) implementation within the mature Firedrake/PETSc library framework where one can easily hook up an active-set Newton step as a smoother.
+
+FIXME the fact that FAScd reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}) suggests it is a completely different strategy from PFAS in \cite{BrandtCryer1983}
+
+FIXME main difference relative to monotone multigrid \cite{Kornhuber1994} and PFAS \cite{BrandtCryer1983} is that method is not tied to, or presented in terms of, an PNGS type smoother
+
+FIXME at a high level one might say FAScd merely generates coarse VI problems which, once solved, make helpful admissible additions to the eventual V-cycle update; the method is smoother-agnostic; smoother could be Newton semi-smooth \cite{BensonMunson2006}, interior point method, etc.
+
+FIXME VINewton smoother could use GMG or AMG on linear step equations, but this was not seen to be necessary in applications so far and was not even tested
+
+FIXME the observation that up-smoothing is more efficient seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}
+
+FIXME desired extension to nonlocal variational inequalities not tried here
 

### 2023-07-14 paper/fascd.tex
address all PF comments re Results section

@@
 
-Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}, allowing $P_1$ or $Q_1$ elements and based on standard nested mesh refinement.
+Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python, for $P_1$ and $Q_1$ elements, using the Firedrake library \cite{Rathgeberetal2016}.
 Here we show numerical results for the example problems presented in Section \ref{sec:vi}, with a focus on the scaling of FAScd iterations as the mesh resolution increases.
-Our main observations are that each V-cycle dramatically reduces the residual norm, and that very few F-cycles solve most examples to within discretization error.
+Our primary observations are that each V-cycle dramatically reduces the residual norm, and that very few F-cycles solve most examples to within discretization error.
 
 \begin{example} \label{ex:results:classical}
-We consider the unilateral 2D Laplacian obstacle problem presented in Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$.
-We consider two problems with very different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$, both over square $\Omega$.
+Consider the unilateral 2D Laplacian obstacle problem, Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$.
+We consider two problems, both over square domains $\Omega$, with the very different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$ shown in Figure \ref{fig:results:classical}.
 The standard ``ball'' problem has an upper hemisphere obstacle $\psi$ \cite[Chapter 12]{Bueler2021}.
-In the ``spiral'' problem \cite[Example 7.1.1]{GraeserKornhuber2009}, $\psi$ is constructed to produce a more interesting coincidence set; see Figure \ref{fig:results:classical}.
-
-For this experiment we used a coarsest mesh of $m_0=5^2$ vertices (32 triangles).
-The finest mesh had $m_9=2049^2 \approx 4.2 \times 10^{6}$ vertices.
+For the ``spiral'' problem \cite[problem 7.1.1]{GraeserKornhuber2009}, $\psi$ is constructed to produce the coincidence set shown.
+For this experiment we used a coarsest regular-triangulation mesh with $m_0=5^2$ vertices (32 triangles).
+The finest mesh, built by standard refinement, had $m_9=2049^2 \approx 4.2 \times 10^{6}$ vertices.
 As our smoother we applied a single RS Newton iteration (Section \ref{sec:implementation}), with three (fixed) IC-preconditioned CG iterations to approximately solve the arising linear systems.
@@
 
-Table \ref{tab:results:classical} shows the resulting iteration counts for finest meshes of $m_J$ vertices.
-When the RS Newton smoother is applied by itself as a solver (``RS only'') it shows rapidly-increasing iterations as expected, roughly proportional to resolution.
+Table \ref{tab:results:classical} shows the resulting iteration counts.
+When the RS Newton smoother is applied by itself as a solver (``RS only''), as expected it shows rapidly-increasing iterations, roughly proportional to resolution.
 By contrast, V-cycle iterations increase slowly, perhaps logarithmically with the resolution.
-Noting that for F-cycles we are counting the number of V-cycles after the ramp (Section \ref{sec:implementation}), we see that a constant number of cycles solves the problem.
-Comparing the ball and spiral results, the complexity of the coincidence set or free boundary does not appear to strongly influence FAScd performance \pef{unlike \cite{GraeserKornhuber2009}?}.
-Each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+IC) iterations; for the F-cycle results, where three iterations suffice at all resolutions, we see close to textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  Work of less than ten residual evaluations is often mentioned as the standard.} despite the low solution regularity.
+For F-cycles we report the number of V-cycles after the ramp (Section \ref{sec:implementation}); a constant number solves the problem.
+Comparing the ball and spiral results, the complexity of the coincidence set or free boundary does not appear to strongly influence FAScd performance.
+For the spiral results, note that \cite{GraeserKornhuber2009} reports more than 30 iterations of all their tested methods to achieve similar tolerance for $m_J \approx 5\times 10^5$.
+
+Each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+IC) iterations.
+Thus for the F-cycle results, where three iterations suffice at all resolutions, we see close to textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  Work of less than ten residual evaluations is often mentioned as the standard.} despite the low solution regularity.
 \end{example}
+
+%\pef{unlike \cite{GraeserKornhuber2009}?}
+% Hard to say.  They only look at one resolution, and they don't show something like "ball".  But their iterations are in the 40--80 range; I've noted that.
 
@@
 \begin{example}  \label{ex:results:plap}
-We again consider Example \ref{ex:plaplacian} with a unilateral constraint, but now for the nonlinear $p$-Laplacian.
-We investigate both fast ($p=1.5$) and degenerate ($p=4.0$) diffusivities.\footnote{``Fast'' and ``degenerate'' refer to $|\grad u|^{p-2}\to \infty$ and $\to 0$, respectively; ellipticity is lost in the latter case.
+We again consider Example \ref{ex:plaplacian} with a unilateral constraint, but now for the nonlinear $p$-Laplacian with both fast ($p=1.5$) and degenerate ($p=4.0$) diffusivities.\footnote{``Fast'' and ``degenerate'' refer to $|\grad u|^{p-2}\to \infty,0$, respectively; ellipticity is lost in the latter case.
 Well-posedness and ($p$-dependent) solution regularity for $p$-Laplacian obstacle problems is considered in \cite{ChoeLewis1991}.} We consider the problem in one dimension for ease of visualization.
@@
 
-Let $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, $\mathcal{K} = \{v \ge \psi\} \subset W^{1,p}(\Omega)$, and $\ell(v) = \int_\Omega g v$ where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$.
-The exact solution of VI \eqref{eq:vi} is easily computed for any $p$; see Figure \ref{fig:results:plap}.
-\pef{Write down the exact solution?}  For this experiment we used a coarsest mesh of $m_0=7$ vertices and a finest mesh with $m_9=3073$.
-The default smoother uses three (fixed) iterations of RS Newton, with LU solution of the arising linear systems.\footnote{This is an optimal $O(m_j)$ smoother in 1D.}  The coarse solver is the same but iterated to convergence.
+Set $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, $\mathcal{K} = \{v \ge \psi\} \subset W^{1,p}(\Omega)$, and $\ip\ell v = \int_\Omega g v$, where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$, in VI \eqref{eq:vi}.
+The exact solution is easily computed for any $p$; see Figure \ref{fig:results:plap} and the documentation of the code associated to this paper.
+For this experiment we used a coarsest mesh of $m_0=7$ vertices and a finest mesh with $m_9=3073$.
+The default smoother uses three (fixed) iterations of RS Newton, with LU solution of the arising linear systems, an $O(m_j)$ smoother in 1D.
+The coarse solver is the same but iterated to convergence.
 The initial iterate was $w=\psi$, and \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$.
 Figure \ref{fig:imagesvcycle} illustrates the major steps of an example V-cycle, showing each of the coarse-correction VI subproblems and solutions.
+
+%\pef{Write down the exact solution?}
+% see code reproducibility
 
@@
 
-Solvers implemented in Firedrake can exploit parallel solvers from the PETSc library \cite{Balayetal2023}.
-The above Examples, run in serial so as to fix the smoother,\footnote{The action of parallel preconditioners can depend on the number of processes \cite[for example]{Bueler2021}.} also run efficiently in parallel.
-The next Example applies a parallel solver, and Example \ref{ex:results:sia} demonstrates the parallel weak scaling \cite{Bueler2021} of FAScd.
+Solvers implemented in Firedrake can exploit parallel solvers from the PETSc library \cite{Balayetal2023}, and in fact the above Examples, which were run in serial so as to fix the smoother, also run efficiently in parallel.
+The next Examples apply parallel solvers, and Example \ref{ex:results:sia} demonstrates the parallel weak scaling \cite{Bueler2021} of FAScd.
 
@@
 Let $\Omega=(-1,1)^d$ for $d=2,3$, the diffusivity $\eps=0.1$ (equation \eqref{eq:advectiondiffusionstrong}), and we impose zero Dirichlet boundary conditions.
-The 2D problem satisfied the coercivity hypothesis with a divergence-free velocity field $\bX = (7+5y,-5x)$, but in 3D we set $\bX = (7+5y,-5x,2z)$ so $\grad\cdot\bX \ne 0$.
-\pef{Why?} The discontinuous source term $\phi \in L^\infty(\Omega)$ satisfied $\phi\ge 0$ in half of the domain, with positive circular patches, while values were negative in the other half.
-\pef{State where to find the full specification of the source term.} While the other Examples in this Section are unilateral, here the solution was box-constrained: $0 \le u \le 1$.
+The 2D problem satisfied the coercivity hypothesis with a divergence-free velocity field $\bX = (7+5y,-5x)$, but in 3D, for varieties sake, we set $\bX = (7+5y,-5x,2z)$ so $\grad\cdot\bX \ne 0$.
+The discontinuous source term $\phi \in L^\infty(\Omega)$ satisfied $\phi\ge 0$ in half of the domain, with positive circular patches, while values were negative in the other half; details are in the associated code.
+While the other Examples in this Section are unilateral, here the solution was box-constrained: $0 \le u \le 1$.
 Figure \ref{fig:results:advdiff} shows coincidence sets for the 2D problem; the 3D sets are comparable in regularity and general pattern.
@@
 Despite the hint of growing iterations under mesh refinement, a few iterations sufficed up to memory-limited resolutions.
-For any of the finer mesh runs here, the FAScd F-cycle solver took less time than generating the $P_1$ mesh hierarchy.
-The local Péclet number $\text{Pe}=\|\bX(x,y,z)\|_2 L \eps^{-1}$, with $L=2$ the width of $\Omega$, varied from 40 to 260.
-Thus the $m_J\le 121^d$ problems were significantly advection-dominated while those with $m_J\ge 481^d$ were diffusion-dominated.
-\pef{I would have said you need $Pe < 1$ to be diffusion-dominated, not have $Pe = 40$?} Nonetheless the smoother was the same and F-cycle iterations were nearly constant.
+For any of the higher resolution runs, the FAScd F-cycle solver took less time than generating the $P_1$ mesh hierarchy.
+The local Péclet number $\text{Pe}=\|\bX(x,y,z)\|_2 L \eps^{-1}$, with $L=2$ the width of $\Omega$, varied from 40 to 260; the mesh Péclet number $m_J^{-1/d}\,\text{Pe}$ varied accordingly.
+Thus the $m_J\le 121^d$ problems were advection-dominated while those with $m_J\ge 481^d$ were diffusion-dominated, but the smoother was the same in all cases, and the iterations were nearly constant.
 \end{example}
+
+%\pef{I would have said you need $Pe < 1$ to be diffusion-dominated, not have $Pe = 40$?}
+% Clarified that this is a mesh Peclet thing.
 
@@
 \begin{tabular}{lr@{\hskip 7mm}c@{\hskip 4mm}c}
-\emph{levels} & $m_J$ & F ($d=2$) & F ($d=3$) \\ \cmidrule{1-4}
+\emph{levels} & $m_J$ & $d=2$ & $d=3$ \\ \cmidrule{1-4}
  1 &    $16^d$ & 1 & 1 \\
@@
 \begin{example}   \label{ex:results:sia}
-We consider the steady shallow ice approximation (SIA) obstacle problem presented in Example \ref{ex:sia}, in which the surface elevation $s$ is constrained to exceed the bedrock elevation $b$.
+We consider the steady shallow ice approximation (SIA) obstacle problem presented in Example \ref{ex:sia}, in which the surface elevation is constrained to exceed the bedrock elevation, $s\ge b$.
 We set $\Omega=(0,L)^2$ with $L=1800$ km.
-The data consist of a radial surface mass balance function $a$ \cite[equation (5.122)]{GreveBlatter2009} and a bumpy, synthetic map $b\in C^1(\Omega)$, with an elevation range of $1800$ m.
+The data consist of a radial surface mass balance function $a$ \cite[equation (5.122)]{GreveBlatter2009} and a bumpy, synthetic map $b\in C^1(\Omega)$, with an elevation range of $1800$ m; details for $b$ are in the associated code.
 A numerical solution is shown in Figure \ref{fig:results:siascene}, colored by the solution surface flow speed, i.e.~the magnitude of $\mathbf{u} = (n+1)(n+2)^{-1} \Gamma (s-b)^{n+1} |\grad s|^{n-1} \grad s$.
@@
 We applied FAScd F-cycles using $Q_1$ elements and a smoother of four (fixed) iterations of RS Newton (Section \ref{sec:implementation}), with quadratic backtracking line search.
-The step equations were approximately solved, in parallel, by three (fixed) iterations of preconditioned GMRES, again with a ASM+ILU preconditioner.
-The coarse solver was RS Newton with LU applied redundantly over all processors.
+The linear step equations were approximately solved, in parallel, by three (fixed) iterations of preconditioned GMRES and a ASM+ILU preconditioner.
+The coarse solver was RS+LU, applied redundantly over all processors.
 The coarse mesh was relatively fine, with $m_0=21^2$ vertices and 400 cells.
@@
 The initial iterate was zero ice: $s=b$.
-
-These solver parameters generate a solver with mesh-independent iterations.
-Then we fixed the number of degrees of freedom per process at $m_J/P=1281^2 \approx 1.6 \times 10^6$ and solved over $P=1,4,16$ processes.\footnote{Runs were on a single-node workstation with two Intel Xeon 4210 cpus (2.2GHz), a total of 20 physical cores, and 128 GB shared memory.}  The results in Table \ref{tab:results:siaweak} show high-quality weak scaling, that is, nearly-constant run time for mesh refinement under constant $m_J/P$.
-The finest mesh result, with $m_J=5121^2=2.6 \times 10^7$ degrees of freedom, needed six minutes to solve for the surface elevation of an ice sheet slightly larger than the current Greenland ice sheet ($1.7$ million $\text{km}^2$) at uniform $350$ m resolution.
+These parameters generated a solver with mesh-independent iterations.
+
+For a weak-scaling study we fixed the number of degrees of freedom per process at $m_J/P=1281^2 \approx 1.6 \times 10^6$ and solved over $P=1,4,16$ processes.\footnote{Runs were on a single-node workstation with two Intel Xeon 4210 cpus (2.2GHz), a total of 20 physical cores, and 128 GB shared memory.}  The results in Table \ref{tab:results:siaweak} show nearly-constant run time for this parallel refinement path.
+The finest mesh result, with $m_J=5121^2=2.6 \times 10^7$ degrees of freedom, needed six minutes to solve for the steady surface elevation of an ice sheet larger than the current Greenland ice sheet ($1.7$ million $\text{km}^2$) at uniform $350$ m resolution.
 \end{example}
@@
 FIXME desired extension to nonlocal variational inequalities not tried here
+
+\subsection*{Code availablility} \label{sec:code}  For reproducibility we have archived the exact software versions used to produce the results in this paper, including the correct Firedrake version, at FIXME Zenodo?
+
+\subsection*{Acknowledgements} \label{sec:acknowledgements}  EB was supported by a Faculty Development Travel Award from United Academics, University of Alaska Fairbanks.  FIXME PF grant support?
 

### 2023-07-14 paper/fascd.tex
I think this addresses PF comments in Introduction

@@
 We present the \emph{full approximation scheme constraint decomposition} (FAScd) multilevel method for solving variational inequalities (VIs).
-FAScd is a common extension of both the full approximation scheme (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and the constraint decomposition (CD) method introduced by X.-C.~Tai for VIs arising in optimization.
-We extend the CD idea by recognizing and exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
+FAScd is a common extension of the full approximation scheme (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and the constraint decomposition (CD) method introduced by X.-C.~Tai for VIs arising in optimization.
+We extend the CD idea by exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
 Then, by applying strong smoothers, with work proportional to the number of unknowns on a finite element mesh level, we demonstrate that FAScd V-cycles exhibit excellent mesh-independent convergence rates and that FAScd F-cycles (full multigrid) are optimal solvers.
-The example problems include unilateral and bi-obstacle VI problems with symmetric linear, nonsymmetric linear, and nonlinear differential operators.
+The example problems include unilateral and bi-obstacle VI problems, with differential operators which are symmetric linear, nonsymmetric linear, and nonlinear.
 Our examples demonstrate very good convergence even on problems where other multilevel VI solvers struggle.
+\end{abstract}
+
 %\pef{I want to say it's better than monotone multigrid. Is this too subtle/too much?}
@@
 Probably the former, but for now we don't have any illuminating analysis.
-\end{abstract}
 
@@
 The constraint decomposition (CD) methods of Tai \cite{Tai2003} are designed to solve coercive variational inequality (VI) problems arising as the Karush--Kuhn--Tucker optimality conditions for the minimization of a convex functional over a convex set.
-These methods have been shown to converge for multilevel and domain decompositions in a finite element (FE) context; in the multilevel case, the CD method has been shown to have almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}; see also \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
-In this paper we extend the multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach, and we apply this new solver to nonlinear VI problems which do not arise from minimization.
-%\pef{Note I've changed ``coarse-level corrections'' to ``coarser-level corrections'', to imply we're not just operating only on the coarsest grids, but also on intermediate grids.}
-% Cool.
-
-Specifically, our \emph{full approximation scheme constraint decomposition} (FAScd) method extends the algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} in the following directions:
+These methods were shown to converge for multilevel and domain decompositions in a finite element (FE) context.
+In the multilevel case, the CD method has almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}, \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
+In this paper we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach, and we apply this new solver to a variety of general, i.e.~nonlinear and non-optimization, VI problems.
+
+Specifically, our \emph{full approximation scheme constraint decomposition} (FAScd) method extends the algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} in the following five directions:
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
@@
 \item The FAS coarse corrections are adapted for general VI problems. The needed computations require neither an objective function nor linearity of the residual operator.
-\item Box constraints, namely upper and lower obstacles, are permitted; see Section \ref{sec:femultilevel}.
-\item We observe an asymmetry special to multilevel CD methods, namely that up-smoothing is more effective than down-smoothing because a ``telescoping'' subset property can be exploited on the up-smoothing side.
+\item Box constraints, namely upper and lower obstacles, are permitted.
 \item We implement V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}).
-\item By using a Newton-type active set method as a smoother, with a fixed number of Krylov iterations per step so that the work of the smoother is proportional to the number of unknowns on the given level, we demonstrate that F-cycle FAScd is an optimal solver.
-\pef{Clarify/limit which classes of problems we're claiming this for.}
+\item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
+\item By using a Newton-type active set method as a smoother, with a fixed number of (Newton and Krylov) iterations per step, so that the work of the smoother is proportional to the number of unknowns on the given level, we demonstrate that F-cycle FAScd is an optimal solver for the examples tested (Section \ref{sec:results}).
 \end{enumerate}
 
-\pef{I think we need some text here hitting the nail on the head about what's different to monotone multigrid methods (even if it's just the absence of the invasive modifications of the basis functions).}
+Prior multilevel algorithms for VIs have various substantial limitations relative to the FAScd framework.
+The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} is apparently limited to linear complementarity problems and projected (truncated) Gauss-Seidel smoothers.
+While FAScd reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed (Appendix \ref{app:reductions}), PFAS apparently reduces to linear multigrid.
+The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so it only applies (as stated) to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
+The problematic latter property is corrected by reformulation using defect constraints and monotone injection operators (see \cite[Algorithm 4.7]{GraeserKornhuber2009} and Section \ref{sec:cdmultilevel} herein).
+The monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be implemented only with a Gauss-Seidel type smoother, and applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.
+Furthermore, monotone multigrid methods are more difficult to implement in FE libraries because modifications of the FE basis functions are needed to approximate the coarser-level coincidence sets.
+No such invasive basis modifications are used either in Tai's original multilevel CD or in FAScd.
+Our Firedrake library \cite{Rathgeberetal2016} implementation of FAScd requires no modifications of basis functions and yet it exhibits excellent performance, including on a problem used to demonstrate the benefits of monotone multigrid (Example \ref{ex:results:classical}; compare \cite[problem 7.1.1]{GraeserKornhuber2009}).
+
+%\pef{I think we need some text here hitting the nail on the head about what's different to monotone multigrid methods (even if it's just the absence of the invasive modifications of the basis functions).}
+% Gave it a shot.
 
 In our experiments, FAScd V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for the corresponding (unconstrained) PDE problems.
-The F-cycle results in Section \ref{sec:results} show nearly-textbook multigrid efficiency, namely solutions within discretization error in work comparable to few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}, for classical Laplacian obstacle problems, $p$-Laplacian obstacle problems, and nonnegativity-constrained advection-diffusion equations.
-Furthermore, these results are essentially insensitive to the length or structure of the free boundary, i.e.~the geometric complexity of the coincidence set.
-
-The iterates from FAScd are admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
-For example, we solve an ice sheet geometry problem (Examples \ref{ex:sia}, \ref{ex:results:sia}) in which the ice flow formulas are only meaningful for surface elevation iterates that do not penetrate the bedrock\footnote{More generally, fluid-layer problems, with a non-negative thickness constraint, often require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation \cite{Bueler2021conservation}.}.
-Note that non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural and invasive modifications of the operator formula in such a fluid-layer context.
-FAScd solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, with a mesh-independent number of F-cycle iterations.
-
-This paper is organized as follows.
+The F-cycle results in Section \ref{sec:results} show nearly-textbook multigrid efficiency, namely solutions within discretization error in work comparable to few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}.
+Specifically we see this performance for classical Laplacian obstacle problems, $p$-Laplacian obstacle problems, and nonnegativity-constrained advection-diffusion equations.
+Furthermore, these results appear to be insensitive to the length or structure of the free boundary, i.e.~the geometric complexity of the coincidence set.
+
+All iterates in a FAScd cycle are admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
+For example, we solve an ice sheet geometry problem (Examples \ref{ex:sia} and \ref{ex:results:sia}) in which the ice flow formulas are only meaningful for surface elevation iterates that do not penetrate the bedrock.
+(More generally, fluid layers subject to a non-negative thickness constraint typically require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation \cite{Bueler2021conservation}, but this is not needed in the ice sheet problem here.)  Non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural modifications of the operator formula.
+FAScd solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a mesh-independent number of F-cycle iterations.
+
+The paper is organized as follows.
 In Section \ref{sec:vi} we recall the theory of coercive and monotone variational inequalities, with several example problems.
-Section \ref{sec:cd} reviews the basics of CDs, as proposed by Tai \cite{Tai2003}, and their induced iterations.
+Section \ref{sec:cd} reviews the basics of CDs and their associated iterations.
 Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, and then Section \ref{sec:cdmultilevel} presents defect constraints generated via monotone injection operators \cite{GraeserKornhuber2009}.
-Here we propose an apparently new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD.
-Section \ref{sec:vcycle} then states the FAScd V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, F-cycles, and our smoother choice.
-Section \ref{sec:results} gives FAScd numerical results from a parallel Firedrake \cite{Rathgeberetal2016} implementation for the problems introduced in Section \ref{sec:vi}.
-The work concludes with discussion and outlook in Section \ref{sec:discussion}, and an appendix on how FAScd reduces to classical FAS multigrid for PDEs.
+Here we propose an apparently new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD and an associated iteration.
+Section \ref{sec:vcycle} then states the FAScd V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, F-cycles, and smoother construction.
+Section \ref{sec:results} gives numerical results, from a parallel Firedrake \cite{Rathgeberetal2016} implementation (see Code Availability), for each example problem introduced in Section \ref{sec:vi}.
+The work concludes with a discussion and outlook in Section \ref{sec:discussion}.
 
@@
 
-FIXME the observation that up-smoothing is more efficient seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}
+FIXME the observation that up-smoothing is more efficient, with consequent asymmetry between $V(1,0)$ and $V(0,1)$ performance, seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}
 

### 2023-07-14 paper/fascd.tex
clean Section 3; having floats for the basic CD iterations is not necessary

@@
 Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique, and also notice that sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp} because the $\cV_i$ are subspaces.
-Finally, suppose there are bounded \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.
+Finally, suppose there are bounded, generally nonlinear, \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.
 Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:femultilevel}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
@@
 \end{equation}
-The operators $\Pi_i$ are generally nonlinear.
 Clearly \eqref{eq:constraintrestrictionsum} implies sense \emph{(i)} for \eqref{eq:constraintdecomp}.
-Tai \cite{Tai2003} defined a \emph{constraint decomposition} (CD) of $\cK$ to be a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}.
-Figure \ref{fig:cartoon} suggests how a low-dimensional example might look; note $\cK_i \not\subset \cK$ in this, and in most nontrivial cases as well.
+A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} \cite{Tai2003}.
+Figure \ref{fig:cartoon} suggests how a low-dimensional example might look; note $\cK_i \not\subset \cK$ in this and most other cases.
 
@@
 
-In Sections \ref{sec:femultilevel}--\ref{sec:vcycle} we will introduce multilevel CD discretizations and algorithms for finite-dimensional VI problems.
-However, one can observe here that the CD concept applies at the level of the continuum problem, as illustrated in the following two examples, the first of which is an overlapping domain decomposition of a Sobolev space.
-Note that $\cK_i \not\subset \cK$ in these cases; informally speaking, the important inclusion is $\cK_i \subset \cV_i$.
+In Sections \ref{sec:femultilevel}--\ref{sec:vcycle} we will introduce multilevel FE CD discretizations and algorithms for finite-dimensional VI problems.
+However, the CD concept applies at the level of the continuum problem, as illustrated in the following two examples.
+The first example is an overlapping domain decomposition of a Sobolev space.
+Again note that $\cK_i \not\subset \cK$; informally speaking, the important inclusion is $\cK_i \subset \cV_i$.
 
@@
 Let $\psi$ be a measurable, real-valued obstacle, defined on $\Omega$, which satisfies $\psi|_{\partial \Omega} \le 0$, and let $\cK = \{v \ge \psi\} \subset \cV$.
-Suppose further that $\{\phi_i\}_{i=0}^{m-1}$ is a smooth partition of unity on $\Omega$, satisfying $0 \le \phi_i\le 1$ and $\sum_i \phi_i = 1$, and let $\Omega_i$ be the open support of $\phi_i$.
+Suppose that $\{\phi_i\}_{i=0}^{m-1}$ is a smooth partition of unity on $\Omega$, satisfying $0 \le \phi_i\le 1$ and $\sum_i \phi_i = 1$, and let $\Omega_i$ be the open support of $\phi_i$.
 Let $\cV_i = \{w \in \cV:w|_{\Omega \setminus \Omega_i} =0 \}$, $\cK_i = \{v \in \cV_i: v \ge \phi_i \psi\}$, and $\Pi_i(v) = \phi_i v$.
-Then \eqref{eq:subspacedecomp}, \eqref{eq:constraintdecomp}, and \eqref{eq:constraintrestrictionsum} all hold.
+Then \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} hold, giving a CD.
 \end{example}
@@
 Our second example is a disjoint frequency decomposition of a Hilbert space.
-A multilevel CD over a finite-dimensional FE discretization, as proposed in Sections \ref{sec:femultilevel} and \ref{sec:cdmultilevel}, approximates such a decomposition.
-
-\begin{example}  \label{ex:frequencydecomposition}  For simplicity suppose $\Omega = (0,a)^d \subset \RR^d$ is a cube, and let $\cV = W_{\text{per}}^{2,k}(\Omega)$ be the periodic functions with square-integrable $k$th derivatives for $k \ge 0$.
+A multilevel CD over a finite-dimensional FE discretization (Sections \ref{sec:femultilevel}--\ref{sec:cdmultilevel}) should approximate this CD.
+
+\begin{example}  \label{ex:frequencydecomposition}  Suppose $\Omega = (0,a)^d \subset \RR^d$ is a cube, and let $\cV = W_{\text{per}}^{2,k}(\Omega)$ be the periodic functions with square-integrable $k$th derivatives, for $k \ge 0$.
 Suppose $\psi \in W_{\text{per}}^{2,k}(\Omega)$ and let $\cK = \{v \ge \psi\} \subset \cV$.
@@
 
-Associated to a constraint decomposition \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} are certain iterative methods \cite{Tai2003,Xu1992}.
+Associated to any CD are certain iterative methods \cite{Tai2003,Xu1992}.
 Let $0\le \alpha \le 1$ be the \emph{damping parameter}.\footnote{FAScd Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} are presented without damping, and indeed it seems to be unnecessary.
-However, the convergence of the more general CD iterations appears to depend upon it.}  Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below, discussed next, solve smaller VI problems over each subset $\cK_i$.
-Each Algorithm replaces $u \in \cK$ with a new iterate $w\in\cK$ which should be closer to $u^\star \in \cK$ solving \eqref{eq:vi}.
-
-The additive/parallel \pr{cd-add} Algorithm \ref{alg:basiccd-add} generalizes the classical Jacobi iteration.
-\begin{pseudofloat}[H]
+However, the convergence of the general CD iterations appears to depend upon it.}  The following algorithms, which replace $u \in \cK$ with a new iterate $w\in\cK$ which should be closer to $u^\star \in \cK$ solving \eqref{eq:vi}, solve smaller VI problems over each subset $\cK_i$.
+
+The additive/parallel \pr{cd-add} algorithm generalizes the classical Jacobi iteration.
+In its VI problem the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$, and then adds-back the subset solution $\hat w_i \in \mathcal{K}_i$.
+
 \begin{pseudo*}
@@
 \end{pseudo*}
-\caption{One additive CD iteration for VI problem \eqref{eq:vi}.}
-\label{alg:basiccd-add}
-\end{pseudofloat}
-Here ``\textbf{for all}'' can be computed in any order.  In its VI,
-\begin{equation}
-\ip{f(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \textnormal{ for all } v_i \in \mathcal{K}_i, \label{eq:cdaddvi}
-\end{equation}
-the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$, and then adds-back the subset solution $\hat w_i \in \mathcal{K}_i$.
-%\pef{Could remove this explanation, but it's nice to keep it in if we can fit it}
-% Proposing to keep for now.
-
-The multiplicative/successive \pr{cd-mult} Algorithm \ref{alg:basiccd-mult} generalizes the Gauss--Seidel iteration.  \begin{pseudofloat}[H]
+
+The multiplicative/successive \pr{cd-mult} algorithm generalizes the Gauss--Seidel iteration.
+Each subset solution corrects the global iterate immediately, so here the ordering of sets $\{\cK_i\}$ is important.
+In its VI problem the sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``those portions of $u$ which have not yet improved''.
+
 \begin{pseudo*}
@@
 \end{pseudo*}
-\caption{One multiplicative CD iteration for VI problem \eqref{eq:vi}.}
-\label{alg:basiccd-mult}
-\end{pseudofloat}
-Each subset solution corrects the global iterate immediately, so here the ordering used in the ``\textbf{for}'' loop is important.  In the VI,
-\begin{equation}
-\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \qquad \forall v_i \in \mathcal{K}_i, \label{eq:cdmultvi}
-\end{equation}
-the sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``those portions of $u$ which have not yet improved''.
 
@@
 We call \pr{cd-add} and \pr{cd-mult} the \emph{basic CD iterations}.
-VI problems \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} reduce to constrained line searches when the $\cV_i$ are one-dimensional.
-As shown in the next Lemma, when VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization} then, subject to bounds on the damping $\alpha$, these iterations generate monotonically-decreasing objective values.
+The VI problems solved in these iterations reduce to constrained line searches when the $\cV_i$ are one-dimensional.
+When VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization} then, subject to bounds on the damping $\alpha$, these iterations generate monotonically-decreasing objective values.
 \begin{lemma} \cite{Tai2003}  Suppose $f=F'$ for $F:\cK\to\RR$ convex and differentiable, and let $u\in\cK$.  For $0 \le \alpha \le 1$ the output $w \in \cK$ from \pr{cd-mult} satisfies
@@
 
-The basic CD iterations are meaningful whether or not we are in the optimization case $f=F'$, and even when $f$ is nonlinear, non-local, or defined only on $\cK$.
+The basic CD iterations are meaningful whether or not we are in the optimization case $f=F'$, and when $f$ is nonlinear, non-local, or defined only on $\cK$.
 However, practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.
@@
 
-\begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear and is local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi$.
-Considering \pr{cd-add} for concreteness, by linearity \eqref{eq:cdaddvi} can be rewritten
+\begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear, and suppose it is local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi$.
+Considering \pr{cd-add} for concreteness, by linearity the VI can be rewritten
 \begin{equation}
@@
 for all $v_i \in \mathcal{K}_i$, where $\tilde\ell = \ell - f(u)$ incorporates the initial residual $f(u)$.
-Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible---generally $e_i \notin \cK$---but \eqref{eq:linearlocalvi} is meaningful for this $f$.
-\pef{Can you expand a little bit here? I confess I don't follow.}
+Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible---generally $e_i \notin \cK$---but ``$f(e_i)$'' in \eqref{eq:linearlocalvi} is meaningful because $f$ is defined on all of $\cV$.
 \end{example}
 
+%\pef{Can you expand a little bit here? I confess I don't follow.}
+% Tried
+
 When $f$ has the locality of Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over basis function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}  However, for general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of CD-type iterations.
-Section \ref{sec:vcycle} constructs a multiplicative multilevel CD iteration for a general operator $f:\cK\to\cV'$ by applying the full approximation scheme (FAS) idea of Brandt \cite{Brandt1977}.
+Section \ref{sec:vcycle} constructs such an iteration for a general operator $f:\cK\to\cV'$ by applying the full approximation scheme (FAS) idea of Brandt \cite{Brandt1977}.
 
@@
 Suppose $\Omega \subset \RR^d$ is open and let $\mathcal{V}=W^{1,p}(\Omega)$.
-The \emph{lower obstacle} $\underline{\gamma}$ and the \emph{upper obstacle} $\overline{\gamma}$ are given functions, defined on the closure $\overline{\Omega}$, with values in the extended real line $\tilde\RR = [-\infty,+\infty]$.
-We suppose $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$ and $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$ are measurable, $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$, and $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$, and that $\partial\Omega$ is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive Hausdorff measure, and $\partial_N \Omega$ possibly empty.
-Dirichlet data\footnote{Dirichlet boundary values are in the trace sense \cite{Evans2010}, and $g_D$ is assumed to be as regular as needed.} $g_D:\partial_D \Omega \to \RR$ then defines
+The \emph{lower obstacle} $\underline{\gamma}$ and \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions, defined on the closure $\overline{\Omega}$, with values in the extended real line $\tilde\RR = [-\infty,+\infty]$, satisfying the following properties: $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$, $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$, and $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$.
+Also suppose $\partial\Omega$ is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive Hausdorff measure, and $\partial_N \Omega$ possibly empty.
+Dirichlet data\footnote{Dirichlet boundary values are in the trace sense \cite{Evans2010}, and $g_D$ is assumed to be as regular as needed.} is given by $g_D:\partial_D \Omega \to \RR$, and we assume $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.
+Now define
 \begin{equation}
@@
 the closed and convex \emph{constraint set}.
-Note that we are treating inequality constraints and boundary conditions in a unified manner.\footnote{Any Neumann conditions over $\partial_N \Omega$ are imposed weakly in the definition of $f$, i.e.~by an integration over $\partial_N\Omega$.}
-
-We recall the abstract VI problem \eqref{eq:vi} with constraint sets \eqref{eq:originalconstraintset}.
-Given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we will assume that the following is well-posed for $u^\star\in \cK$:
+Note that we are treating inequality constraints and Dirichlet boundary conditions in a unified manner.
+Any Neumann conditions over $\partial_N \Omega$ are imposed weakly in the definition of the operator $f$, i.e.~by an integration over $\partial_N\Omega$.
+
+In the next Sections we numerically solve continuum VI problem \eqref{eq:vi} over a constraint set of the form \eqref{eq:originalconstraintset}.
+In fact, given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, from now on we assume that the following is well-posed for $u^\star\in \cK$:
 \begin{equation}
@@
 Let $\Omega \subset \RR^d$ be a (open) bounded polygon.
-We will construct a multilevel FE approximation to \eqref{eq:boxdirichletvi} over $\Omega$ using $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, by refining a coarsest mesh into a nested mesh hierarchy; the goal is an accurate solution on the finest-level mesh.
-The hierarchy allows the creation of a multilevel CD of $\cK$, using the ideas in the next two Sections, which permits solvers based on admissibility-preserving coarse corrections.
+A multilevel FE approximation to \eqref{eq:boxdirichletvi} is constructed, using $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, by refining a coarsest mesh into a mesh hierarchy.
+The hierarchy allows the creation of a multilevel CD of $\cK$ using the ideas in the next two Sections.
 
@@
 Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion}, which are special to multilevel schemes based on defect constraints, hold regardless of whether a CD operator $\Pi_i$ can be constructed.
-Here there is, in fact, another ``basic'' iteration, an incomplete variant of \pr{cd-mult} (Algorithm \ref{alg:basiccd-mult}) that does not need the $\Pi_i$ operators.
-It is a variant of \pr{cd-mult} (Algorithm \ref{alg:basiccd-mult}).
-
-\begin{pseudofloat}[ht]
+Here there is, in fact, another ``basic'' iteration, an incomplete variant of \pr{cd-mult} (Section \ref{sec:cd}) that does not need the $\Pi_i$ operators.
+One application of \pr{icd-tele} below assumes $w^J\in\mathcal{K}^J$, property \eqref{eq:fe:finestlevelequivalent}, and incomplete CD \eqref{eq:fe:upwardsuminclusion}.
+
 \begin{pseudo*}
@@
 \end{pseudo*}
-\caption{One telescoping incomplete CD iteration for VI problem \eqref{eq:fe:vi}, assuming $w^J\in\mathcal{K}^J$, property \eqref{eq:fe:finestlevelequivalent}, and incomplete CD \eqref{eq:fe:upwardsuminclusion}.}
-\label{alg:basiccd-tele}
-\end{pseudofloat}
 
@@
 Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using \eqref{eq:fe:upwardsuminclusion}.
-Continuing in this ``telescoping'' way (Algorithm \ref{alg:basiccd-tele}), up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.
+Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.
 Then the next finest-level, admissible iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.
@@
 The $V(1,0)$ cycle is similar to Algorithm 4.7 in \cite{GraeserKornhuber2009}, but generalizes it by applying FAS-type corrections and allowing any smoother or nonlinear operator $f$.
-One iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} (Algorithm \ref{alg:basiccd-mult}) over CD\footnote{Strictly speaking, this statement is restricted to unilateral cases.} \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Algorithm \ref{alg:basiccd-tele}) over incomplete CD \eqref{eq:fe:upwardsuminclusion}, and in this sense $V(1,1)=V(1,0)+V(0,1)$.
+One iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} (Section \ref{sec:cd}) over CD\footnote{Strictly speaking, this statement is restricted to unilateral cases.} \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over incomplete CD \eqref{eq:fe:upwardsuminclusion}, and in this sense $V(1,1)=V(1,0)+V(0,1)$.
 

### 2023-07-14 paper/fascd.tex
Small edits and a comment

@@
 We extend the CD idea by exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
-Then, by applying strong smoothers, with work proportional to the number of unknowns on a finite element mesh level, we demonstrate that FAScd V-cycles exhibit excellent mesh-independent convergence rates and that FAScd F-cycles (full multigrid) are optimal solvers.
-The example problems include unilateral and bi-obstacle VI problems, with differential operators which are symmetric linear, nonsymmetric linear, and nonlinear.
+Then, by applying strong smoothers, with work proportional to the number of unknowns on a given mesh level, we demonstrate that FAScd V-cycles exhibit excellent mesh-independent convergence rates and that FAScd F-cycles (full multigrid) are optimal solvers.
+The example problems include unilateral and bilateral VI problems, with differential operators which are symmetric linear, nonsymmetric linear, and nonlinear.
 Our examples demonstrate very good convergence even on problems where other multilevel VI solvers struggle.
@@
 \end{figure}
+\pef{Could we make the bedrock grey?}
 

### 2023-07-14 paper/fascd.tex
tighten appendix, address a PEF, remove commented-out clutter

@@
 
-%\newcommand{\XX}{{\sffamily X}}
 \newcommand{\XX}{\ding{55}}
@@
 
+\newfloat{pseudofloat}{t}{xyz}[section]
+\floatname{pseudofloat}{Algorithm}
+
 \usepackage{todonotes}
@@
 We present the \emph{full approximation scheme constraint decomposition} (FAScd) multilevel method for solving variational inequalities (VIs).
-FAScd is a common extension of the full approximation scheme (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and the constraint decomposition (CD) method introduced by X.-C.~Tai for VIs arising in optimization.
+FAScd is a common extension both of the full approximation scheme (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and the constraint decomposition (CD) method introduced by X.-C.~Tai for VIs arising in optimization.
 We extend the CD idea by exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
@@
 
-%\pef{I want to say it's better than monotone multigrid. Is this too subtle/too much?}
-% I vote for not saying this in the abstract.
-It is hard to say which aspect of FAScd is more important to beating monotone multigrid: stronger smoother or improved up-smooth CD.
-Probably the former, but for now we don't have any illuminating analysis.
-
 \maketitle
 
-%\tableofcontents
-
 \thispagestyle{empty}
-%\bigskip
-
-\newfloat{pseudofloat}{t}{xyz}[section]
-\floatname{pseudofloat}{Algorithm}
 
@@
 
-%\pef{I think we need some text here hitting the nail on the head about what's different to monotone multigrid methods (even if it's just the absence of the invasive modifications of the basis functions).}
-% Gave it a shot.
-
 In our experiments, FAScd V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for the corresponding (unconstrained) PDE problems.
@@
 \end{equation}
-%\pef{Why is $\phi \in L^\infty(\Omega)$? If that is the case, then should we have an $L^2$ inner product in the last term, rather than a dual pairing?}
-% Fixed.  \phi can be in the dual space
 With this operator, consider VI \eqref{eq:vi} with $\ell = 0$ for any closed and convex $\cK \subset \cV$.
@@
 \end{example}
-%\pef{`suggests' is quite weak; can we use a stronger word?}
-% Agreed.  Last sentences made brief and direct.
 
 Numerical results are shown in Section \ref{sec:results} for applications of FAScd Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} to these Examples.
-%\pef{I'm not sure starting with the example problems we want to solve is the right move. Let me read more and think about it.}
-% I think it is o.k.
 
@@
 
-%\pef{Can you expand a little bit here? I confess I don't follow.}
-% Tried
-
 When $f$ has the locality of Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over basis function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}  However, for general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of CD-type iterations.
@@
 the finest-level \emph{constraint set} of admissible FE functions. This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.
-%As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fe:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J$ are infinite; there is no lower or upper constraint, respectively, at such vertices.
 
@@
 
-%\pef{I don't like the name ``monotone restrictions''.
-The word restriction is strongly associated with operating on the dual space.
-I would prefer to call them ``monotone injections'', since that is the matching operator.}
-% Agreed.  "monotone restriction" is used in GraeserKornhuber2009, but happy not to follow.
-
 The monotone injections $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to nodal functions $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the \emph{stars} of the coarse vertices).
@@
 \begin{example}  \label{ex:directRbad}  
-
 Let $\Omega = (0,L) \subset \RR^1$ and consider a two-level mesh hierarchy ($J=1$).
@@
 The LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if another element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$; similar comments apply to $\overline{\chi}^{j}$.
+Note that for Algorithm \ref{alg:fascd} below, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not required.
 
@@
 
-This construction of multilevel box constraints, fundamentally the same as the unilateral construction in \cite{GraeserKornhuber2009}, is designed to permit large-yet-admissible coarse corrections, which promotes multilevel solver efficiency.
-Note that for Algorithm \ref{alg:fascd} below, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not required.
-
-We now compute the LDC differences \cite{GraeserKornhuber2009}:
-\pef{Say a few words of motivation first for introducing this concept.}
+This construction of multilevel box constraints, as for the unilateral construction in \cite{GraeserKornhuber2009}, is designed to permit large-yet-admissible coarse corrections, which promotes multilevel solver efficiency.
+However, when coarsening the problem one does not know what the next coarse corrections will be, and we require constraint sets from which any sum of corrections will be admissible.
+With this in mind, and looking forward to telescoping sums \eqref{eq:fe:telescoping}, we compute the LDC differences \cite{GraeserKornhuber2009}:
 \begin{equation}
@@
     for $j=1$ to $j=J$ \\+
-      $z^j = y^{j} + P z^{j-1}$ \\
+      $z^j = y^{j} + P z^{j-1}$ \label{line:vcycleupsmoothinitial} \\
       $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,\overline{\chi}^j,w^j;z^j)$  \ct{smoothing of \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
@@
 In this section we also construct the F-cycle extension, and we describe the active set Newton method we will use as a smoother and coarse solver.
-
-%Let $x_p \in \mathcal{T}^J$ be a finest-level node, and $\psi_p(x)$ the corresponding basis function.
-Let $\mathcal{X}^J=\big\{w\in\cV^J\,:\,w(x_p)=g_D^J(x_p) \text{ if } x_p\in \partial_D\Omega\big\}$, thus $\cK^J = \left\{\underline{\gamma}^J \le w \le \overline{\gamma}^J\right\} \cap \mathcal{X}^J$.
-If $m_D$ is the number of vertices $x_p$ in the Dirichlet boundary, let $I_D$ be the $m_D\times m_J$ matrix which extracts Dirichlet nodal values; its transpose $I_D^\top$ injects Dirichlet boundary degrees of freedom into $(\mathcal{V}^J)'$.
-Finally, write $a\perp b$ when $a,b \in \RR^n$ are \emph{complementary}, that is, when $a_i \ge 0$ and $b_i \ge 0$ and $a_i b_i = 0$ for $i=0,\dots,n-1$.
-Then problem \eqref{eq:fe:vi}, which would be expressed VI$\left(f^J(\cdot)-\ell^J,\left[\,\underline{\gamma}^J,\overline{\gamma}^J\right] \cap \mathcal{X}^J\right)$ in the operations research literature \cite{FerrisPang1997}, is equivalent to the following MCP for the solution $u^J \in \mathcal{K}^J$, but with additional equality constraints:
-%\begin{align}
-%\underline{\gamma}^J \le u^J \le \overline{\gamma}^J \quad &\perp \quad f^J(u^J) - \ell^J - I_D^\top \lambda \label{eq:fe:mcp} \\
-%I_D u^J &= g_D^J\big|_{\partial_D\Omega} \notag
-%\end{align}
-%for some Lagrange multipliers $\lambda \in \RR^{m_D}$.
-%\pef{I wonder if this definition will confuse people, because they won't know how to plug in $\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ into your definition of $\perp$ ($\underline{\gamma}^J \le u^J \le \overline{\gamma}^J$ is not in $\RR^n$).
-Maybe it would be easier to just use the spelled-out version below, and not use this notation?}
-% Agreed.  Greatly simplified
 
@@
 
-%\pef{I think this is all fairly standard and we could abbreviate this section substantially.}
-% Yes.
-
 The computable nonlinear map $\rNC : \mathcal{K}^J \to \RR^{m_J}$ in \eqref{eq:rNC} is defined using only the finest-level objects $f^J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$.
@@
 (In the VI context one must also truncate after the prolongation; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fcycle}.)  During this ``ramp'' one approximately solves on each level by using a fixed number of V-cycles, each down to the coarsest mesh, and then one does full-depth V-cycles from the finest level, to convergence.
-%\pef{Clarify truncation}
-% Attempted.
 
@@
 Injection is used to generate the $\underline{\gamma}^j,\overline{\gamma}^j$ on each level.\footnote{Contrast the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, based on an existing admissible iterate.}  Regarding line \ref{line:fcyclecoarsestinitial}, one may generate an initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation of a finest-level iterate, if one is available.
-%\pef{Explain why?}
-% Mostly not done.  But restated for clarity?
 
@@
 
-%\pef{unlike \cite{GraeserKornhuber2009}?}
-% Hard to say.  They only look at one resolution, and they don't show something like "ball".  But their iterations are in the 40--80 range; I've noted that.
-
 %REGENERATE using git@bitbucket.org:pefarrell/fascd.git, in examples/
@@
 \end{figure}
-%\pef{The spiral problem looks under-resolved; is that right?}
-% Fixed.
 
@@
 
-%\pef{Write down the exact solution?}
-% see code reproducibility
-
 %REGENERATE using elbueler/plapfigs branch of git@bitbucket.org:pefarrell/fascd.git, in examples/
@@
 
-%\pef{I would have said you need $Pe < 1$ to be diffusion-dominated, not have $Pe = 40$?}
-% Clarified that this is a mesh Peclet thing.
-
 % see git@bitbucket.org:pefarrell/fascd.git: examples/pollutant.py
@@
 
+FIXME  It is hard to say which aspect of FAScd is more important to beating monotone multigrid: stronger smoother or improved up-smooth CD.
+Probably the former, but for now we don't have any illuminating analysis, and a head-to-head race is comparing apples and oranges because we are using such a strong smoother.
+Various differences of implementation don't produce any really good multigrid convergence if your smoother is NGS on each level.
+I think the convergence proofs in \cite{GraeserKornhuber2009} all depend on this?
+
 FIXME the fact that FAScd reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}) suggests it is a completely different strategy from PFAS in \cite{BrandtCryer1983}
@@
 
-\pr{fascd-vcycle} (Algorithm \ref{alg:fascd}) reduces to the classical FAS multigrid V-cycle for nonlinear PDEs \cite{Trottenbergetal2001}, which itself generalizes linear multigrid, when all inequality constraints are removed.
-In fact, suppose we remove the inequality constraints from the finest-level VI problem \eqref{eq:fe:vi} by replacing the finest-level obstacles $\underline{\gamma}^J$ and $\overline{\gamma}^J$ with $-\infty$ and $+\infty$ respectively, thus not constraining the solution $u^J$ or any possible iterate.
-The admissible set $\mathcal{K}^J$ of \eqref{eq:fe:vi} is then determined only by the Dirichlet boundary conditions.
+When all inequality constraints are removed the FAScd V-cycle Algorithm \ref{alg:fascd} reduces to the classical FAS multigrid V-cycle for nonlinear PDEs \cite{Brandt1977,Trottenbergetal2001}, which itself generalizes linear multigrid.
+In fact, suppose we replace $\underline{\gamma}^J,\overline{\gamma}^J$ in definition \eqref{eq:fe:fineconstraintset} with $-\infty,+\infty$, respectively.
+The finest-level admissible set $\mathcal{K}^J$ is then determined only by the Dirichlet boundary conditions.
 Let $\mathcal{V}_0^J \subset \mathcal{V}^J$ denote the linear subspace with zero values on $\partial_D\Omega$.
-Since $v-u^J\in\mathcal{V}_0^J$ in \eqref{eq:fe:vi} is now an arbitrary element, we can rewrite the finest-level problem as: find $u^J \in \mathcal{V}^J$ such that
-\begin{equation}
-\ip{f^J(u^J)}{v} = \ip{\ell^J}{v} \qquad \text{for all } v\in \mathcal{V}_0^J, \label{eq:app:fas:pde}
-\end{equation}
-the FE discretization of strong-form PDE $f(u)=\ell$.
-The downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets are also enlarged so that the corrections $y^j$ and $z^j$ in \pr{fascd-vcycle} are never constrained; lines \ref{line:vcyclegenchifinest}, \ref{line:vcyclegenchi}, and \ref{line:vcyclegenphi} always set $-\infty, +\infty$ by the arithmetic on $\tilde\RR$.
-The corrections $y^j$ and $z^j$ are thus unconstrained.
-
-Since the downward/upward VIs \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} are unconstrained, they can be rewritten as weak forms $\ip{f^j(w^j + y^j)}{v} = \ip{\ell^j}{v}$ and $\ip{f^j(w^j + z^j)}{v} = \ip{\ell^j}{v}$, respectively, for all $v\in \mathcal{V}_0^j$.
-Then we combine $u^j=w^j+y^j$, instead of $w^j$ and $y^j$ separately, and similarly $u^j=w^j+z^j$ when up-smoothing.
-The smoothers and the coarsest-level solver can be regarded as unconstrained, in-place modifications of $u^j$, so we denote them simply as ``$\text{\pr{smooth}}(\ell^j;u^j)$.''  Noting line \ref{line:vcyclerestrictsolution} of Algorithm \ref{alg:fascd} now says $u^{j-1}=\iR u^j$, after down-smoothing the coarsened source functional computed by line \ref{line:vcyclerestrictell} can be rewritten as
-\begin{equation}
-\ell^{j-1} = f^{j-1}\left(u^{j-1}\right) + R\left(\ell^j-f^j(u^j)\right). \label{eq:app:fas:levelsource}
-\end{equation}
+Since $v-u^J\in\mathcal{V}_0^J$ is now arbitrary we can rewrite \eqref{eq:fe:vi} as the FE discretization of strong-form PDE $f(u)=\ell$: find $u^J \in \mathcal{V}^J$ such that
+\begin{equation}
+\ip{f^J(u^J)}{v} = \ip{\ell^J}{v} \qquad \text{for all } v\in \mathcal{V}_0^J. \label{eq:app:fas:pde}
+\end{equation}
+
+The downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets are also enlarged to vector spaces, and the corrections $y^j$ and $z^j$ are unconstrained; lines \ref{line:vcyclegenchifinest}, \ref{line:vcyclegenchi}, and \ref{line:vcyclegenphi} in Algorithm \ref{alg:fascd} can be removed.
+In fact, VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} can be rewritten as the weak form problem $\ip{f^j(u^j)}{v} = \ip{\ell^j}{v}$ over $v\in \mathcal{V}_0^j$---this groups $u^j=w^j+y^j$ when down-smoothing and $u^j=w^j+z^j$ when up-smoothing---and the smoothers can be regarded as unconstrained, in-place modifications of $u^j$.
+Noting line \ref{line:vcyclerestrictsolution} now says $u^{j-1}=\iR u^j$, after down-smoothing the coarsened source functional computed by line \ref{line:vcyclerestrictell} can be rewritten as $\ell^{j-1} = f^{j-1}\left(u^{j-1}\right) + R\left(\ell^j-f^j(u^j)\right)$.
 
@@
 The initial value of the correction $z^j$, before up-smoothing on the $j$th level, is $z^j = y^j + P z^{j-1}$, where $z^{j-1}$ contains the total correction from coarser corrections and $j-1$ level up-smoothing.
-Thus the $j$th-level iterate \emph{before} up-smoothing must be set to
-    $$w^j + y^j + P z^{j-1} = u^j + P(u^{j-1} - \iR u^j).$$
-
-Modifying Algorithm \ref{alg:fascd} appropriately, we have a much-simplified Algorithm \ref{alg:fas}.
-This FAS V-cycle appears in various sources in the literature, for example \cite[Algorithm 14]{Bruneetal2015} and \cite[section 5.3.4]{Trottenbergetal2001}.
-
-\begin{pseudofloat}[ht]
-\begin{pseudo} \label{ps:fas-vcycle}
+Thus the $j$th-level iterate \emph{before} up-smoothing must be set to $w^j + y^j + P z^{j-1} = u^j + P(u^{j-1} - \iR u^j)$, which replaces line \ref{line:vcycleupsmoothinitial}.
+
+With these modifications, Algorithm \ref{alg:fascd} becomes the FAS V-cycle below, as a solver for \eqref{eq:app:fas:pde}.
+This algorithm appears in various sources in the literature, for example \cite[Algorithm 14]{Bruneetal2015} and \cite[section 5.3.4]{Trottenbergetal2001}.
+
+\begin{pseudo*}
 \pr{fas-vcycle}(\ell^J; u^J)\text{:} \\+
@@
       $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j;u^j)$ \\-
-\end{pseudo}
-\caption{The standard FAS V-cycle for a discretized PDE $f^J(u^J)=\ell^J$.}
-\label{alg:fas}
-\end{pseudofloat}
-
-\begin{comment}
-%From slack, to not lose
-
-\noindent Patrick : \emph{Interesting that you think Graeser and Kornhuber were barking up the wrong tree.
-I still don't quite have my finger on the exact difference in the scheme between GK and the FAS approach.
-Is there a clean summary somewhere?
-(I looked in the mcd-extended manuscript, but couldn't see it in a quick skim.)  Is the difference to GK the monotone multigrid bit, or the particular smoothers used?}
-
-\medskip\noindent
-Ed: \cite{GraeserKornhuber2009} only considers the Laplacian lower obstacle case, and they write everything in terms of nonlinear Gauss-Seidel (i.e.~optimize one basis function at a time).
-They first lay out the \cite{Tai2003} CD strategy as Alg.
-4.7, only for V(1,0) cycles.
-Then they observe that the convergence rates from \cite{Tai2003} are slower (worse) than if one does "monotone multigrid".
-This seems to mean to them that one modifies the coarse basis functions according to whether they overlap the current fine-grid estimate of the coincidence set, "truncated nodal basis functions".
-(See the start of \cite{GraeserKornhuber2009} section 5.2.)  I think that making these modifications in Firedrake, so we could try monotone multigrid V-cycles, would be a nightmare.
-FAScd does not attempt the monotone multigrid idea at all.
-
-By contrast, FAScd is essentially just the \cite{Tai2003} CD strategy, but (1) without assuming an objective, and (2) with a coarse correction which is FAS in the nonlinear PDE case.
-If you apply \cite{Tai2003} CD to a nonquadratic (e.g.
-p=1.5 p-laplacian) objective then the coarse-grid corrections are optimization problems over the coarse space.
-I don't even know how to interpret this in non-optimization VIs?
-Or maybe I should say that one can re-interpret this as FAS coarse corrections, and I have done so.
-
-But I have now concluded that everything discussed in \cite{GraeserKornhuber2009} is angels-on-the-head-of-a-pin stuff because the various differences of implementation don't produce any really good multigrid convergence if your smoother is NGS on each level.
-I think the convergence proofs in \cite{GraeserKornhuber2009} all depend on this?
-There are indeed differences in convergence rates reported in the results at the end of \cite{GraeserKornhuber2009}, but they are all not good compared to how multigrid works on PDEs.
-This NGS-based stuff is what I was trying before February, in 1d mostly, with hand-made NGS smoothers, and no Firedrake.
-
-Now, the good thing I was already doing before February was building a version of VI multigrid which makes minimal assumptions on the problem:  FAS (residual need not be linear) and CD (build-up the total V-cycle update in an admissible manner on every level).
-Thus it has a chance of working for the ice problems I care about.
-
-As I now see, and can easily demo on the spiral problem in \cite{GraeserKornhuber2009} and on any other classical obstacle problem I have tried, the thing which makes it fly is to realize, as you did, or perhaps as you just take for granted because you are so comfortable with Firedrake, that the smoother can be a grown-up but still O(N) VI solver.
-E.g.~a fixed number of iterations of vinewtonrsls+gmres+ilu or whatever.
-The role of the FAScd is simply to generate coarse VI problems which, once solved, make helpful admissible additions to the eventual V-cycle update.
-
-So the difference between FAScd and \cite{GraeserKornhuber2009} are: (1) FAS coarse corrections, (2) no attempt to do monotone/truncated multigrid which would be a mess (I think) across the FE infrastructure, (3) clarity on what sets you can use in $V(\nu_1,\nu_2)$ cycles, with bigger up-smoother sets in particular, and most importantly (4) implementation within the mature Firedrake/PETSc library framework where one can easily hook up an active-set Newton step as a smoother.
-\end{comment}
+\end{pseudo*}
 

### 2023-07-14 paper/fascd.tex
grey bedrock cause I'm an advanced paraviewer
further image edit
do a PEF (color!) and clean some text

@@
 A key choice in devising a multilevel scheme for VIs is how to construct coarse-grid bounds.
-Our multilevel CD scheme uses \emph{defect constraints} \cite{GraeserKornhuber2009} which are based on an admissible finest-level iterate.
-The next Example concretely illustrates why we do not apply restriction operators directly\footnote{An exception is in the F-cycle; see Section \ref{sec:implementation}.} to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.
-Our straightforward (iterate-dependent) bounds on the defect constraints, namely \eqref{eq:fe:chiordering} below, bypass the difficulties exposed in Example \ref{ex:directRbad}.
+Our multilevel CD scheme uses \emph{defect constraints} \cite{GraeserKornhuber2009} derived from an admissible finest-level iterate.
+The next Example illustrates why we do \emph{not} apply restriction operators directly\footnote{An exception is in the F-cycle; see Section \ref{sec:implementation}.} to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.
+Iterate-dependent defect constraints, definition \eqref{eq:fe:defectconstraints} below, satisfy ordering \eqref{eq:fe:chiordering} and bypass these difficulties.
 
@@
 Let $\Omega = (0,L) \subset \RR^1$ and consider a two-level mesh hierarchy ($J=1$).
-Let $\underline{\gamma}^1$ be a sawtooth function with minimum value zero, and suppose $\overline{\gamma}^1=C+\underline{\gamma}^1$ with $C>0$ which is strictly less than the maximum of $\underline{\gamma}^1$.
+Let $\underline{\gamma}^1$ be a sawtooth function with minimum value zero, and suppose $\overline{\gamma}^1=C+\underline{\gamma}^1$ with $C>0$ strictly less than the maximum of $\underline{\gamma}^1$.
 Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ vertices, respectively.
-Ignoring Dirichlet boundary for simplicity, the constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.
-However, various restriction schemes applied directly to $\underline{\gamma}^1,\overline{\gamma}^1$ are problematic.
+Ignoring the boundary for simplicity, the fine-level constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.
+However, various restriction schemes to coarsen the constraints, applied directly to $\underline{\gamma}^1,\overline{\gamma}^1$, are problematic.
 Suppose we propose
@@
 The general motivation for this choice---compare \eqref{eq:fe:chilevels} below---is that $w^0 \ge \underline{\gamma}^0 \, \implies \, Pw^0 \ge \underline{\gamma}^1$ and $w^0 \le \overline{\gamma}^0 \, \implies \, Pw^0 \le \overline{\gamma}^1$.
-However, while these facts are true, the difficulty with this choice is that the coarse-level admissible set $\mathcal{K}^0$ is now empty because the proposed lower bound $\underline{\gamma}^0$ actually exceeds the proposed upper bound $\overline{\gamma}^0$.
+However, in this case the coarse-level set $\mathcal{K}^0$ is empty because the proposed lower bound $\underline{\gamma}^0$ exceeds the proposed upper bound $\overline{\gamma}^0$ (Figure \ref{fig:directRbad}).
 \end{example}
@@
 \begin{figure}[ht]
-\begin{center}
-\input{tikz/directRbad.tex}
-\end{center}
-\caption{For finest-level box constraints $\underline{\gamma}^1,\overline{\gamma}^1$ like these, direct restriction using any of $\iR,\minR,\maxR$, to a coarser mesh (solid dots), is problematic for admissibility of coarse corrections.}
+\input{tikz/directRbadfine.tex}
+\quad
+\input{tikz/directRbadcoarse.tex}
+\caption{For the finer-level constraints $\underline{\gamma}^1$ (blue) and $\overline{\gamma}^1$ (red) on the left, direct application of $\iR,\minR,\maxR$ is problematic.
+For example, if we apply the monotone injections then the resulting coarser-level constraint set $\mathcal{K}^0$ is empty because $\maxR \underline{\gamma}^1 > \minR \overline{\gamma}^1$ (right).
+Defect constraints avoid this difficulty.}
 \label{fig:directRbad}
 \end{figure}
-\pef{Can we put another figure beside it illustrating the choice $\underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1$?}
-\pef{Maybe use different colours for upper and lower bounds?}
-
-Subtracting an admissible iterate from the original bounds provides a route out of the difficulties in Example \ref{ex:directRbad}.
-Suppose $w^J \in \cK^J$ is an admissible finest-level iterate for \eqref{eq:fe:vi}.
+%\pef{Can we put another figure beside it illustrating the choice $\underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1$?}
+%\pef{Maybe use different colours for upper and lower bounds?}
+
+So suppose $w^J \in \cK^J$ is an admissible finest-level iterate for \eqref{eq:fe:vi}.
 We define the \emph{finest-level defect constraints} (DCs) \cite{GraeserKornhuber2009} as the following nodal functions in $\tilde{\mathcal{V}}^J$:
@@
 
-Descending coarser levels, $j=J-1,\dots,1,0$, we apply the monotone injections to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:
+Descending coarser levels, $j=J-1,\dots,1,0$, we next apply the monotone injections to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:
 \begin{equation}
@@
 This construction of multilevel box constraints, as for the unilateral construction in \cite{GraeserKornhuber2009}, is designed to permit large-yet-admissible coarse corrections, which promotes multilevel solver efficiency.
-However, when coarsening the problem one does not know what the next coarse corrections will be, and we require constraint sets from which any sum of corrections will be admissible.
+However, when coarsening the problem downward in a V-cycle one does not know what the next coarse corrections will be, and so we require constraint sets from which any sum will be admissible.
 With this in mind, and looking forward to telescoping sums \eqref{eq:fe:telescoping}, we compute the LDC differences \cite{GraeserKornhuber2009}:
@@
 \end{equation}
-We take $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ so that $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.
-Regarding subtraction of extended real-valued quantities, if $\underline{\chi}^j(x_p)=-\infty$ then we set $\underline{\phi}^j(x_p)=-\infty$ regardless of the value of $\underline{\chi}^{j-1}(x_p)$, and similarly $\overline{\phi}^j(x_p)=+\infty$ if $\overline{\chi}^j(x_p)=+\infty$.
-While $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ again bracket zero, $\underline{\phi}^j \le 0 \le \overline{\phi}^j$, they are not ordered as in \eqref{eq:fe:chiordering}.
+(Take $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ so that $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.
+Regarding subtraction of extended real-valued quantities, if $\underline{\chi}^j(x_p)=-\infty$ then we set $\underline{\phi}^j(x_p)=-\infty$ regardless of the value of $\underline{\chi}^{j-1}(x_p)$, and etc.)  While $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ again bracket zero, $\underline{\phi}^j \le 0 \le \overline{\phi}^j$, they are not ordered as in \eqref{eq:fe:chiordering}.
 
@@
 Our V-cycle solver (Algorithm \ref{alg:fascd}) will compute corrections in $\mathcal{D}^j$ for the downward stage and in $\mathcal{U}^j$ for the upward stage.
-This strategy is, however, based on \emph{incomplete} CDs, in the general box-constrained cases, as we now explain.
-Each upward set $\mathcal{U}^j$ is generally only a superset of a sum of downward sets $\mathcal{D}^j$, but this becomes a true CD, i.e.~satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, in unilateral cases.
+This strategy is, however, based on \emph{incomplete} CDs in general box-constrained cases, as we now explain.
+Each upward set $\mathcal{U}^j$ is generally only a superset of a sum of downward sets $\mathcal{D}^j$, though this becomes a true CD, i.e.~satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, in unilateral cases.
 The following Lemma, apparently new, is based in part on the unilateral CD construction underlying Algorithm 4.7 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.
@@
 
-One might try to turn \eqref{eq:fe:downwardsuminclusion} into a full CD for arbitrary box constraints by using decomposition maps like $\Pi_i$, for example by splitting $z^j = \min\{z^j,0\} + \max\{z^j,0\}$ and applying unilateral formulas like \eqref{eq:fe:unilateraldecompositionoperator} to $\min\{z^j,0\}$ and $\max\{z^j,0\}$ separately.
-The above proof shows that the lower bound $\Pi_i (\min\{z^j,0\}) \ge \underline{\phi}^i$ does hold.
-However, the following Example\footnote{Of course, this Example does not exclude the existence of an entirely-different strategy for showing a full CD in the general box-constrained case.} shows why $\Pi_i(\min\{z^j,0\})$ may have an arbitrarily large maximum, and thus exceed any finite upper LDC $\overline{\phi}^i$.
+One might attempt to convert \eqref{eq:fe:downwardsuminclusion} into a full CD for arbitrary box constraints by using decomposition maps like $\Pi_i$, for example by splitting $z^j = \min\{z^j,0\} + \max\{z^j,0\}$ and applying unilateral formulas like \eqref{eq:fe:unilateraldecompositionoperator} to $\min\{z^j,0\}$ and $\max\{z^j,0\}$ separately.
+The above proof shows that the lower bound $\Pi_i (\min\{z^j,0\}) \ge \underline{\phi}^i$ indeed holds.
+However, the following Example shows why $\Pi_i(\min\{z^j,0\})$ may have an arbitrarily large maximum, and thus exceed any finite upper LDC $\overline{\phi}^i$.
 In fact, $\Pi_i$ does not generally map all non-positive elements of $\mathcal{U}^J$ into $\mathcal{D}^i$ if a finite upper LDC is present.
+On the other hand, we can not exclude entirely-different strategies for showing full CDs in the general box-constrained case.
 
@@
 But then $\Pi_1 z^1$ attains a maximum $a\ge 0$.
-This maximum can be chosen to exceed any upper defect obstacle $\overline{\phi}^1$, so $\Pi_1 z^1 \notin \mathcal{D}^1$.
+This maximum can be chosen to exceed any upper defect obstacle $\overline{\phi}^1$, and $\Pi_1 z^1 \notin \mathcal{D}^1$ would occur.
 \end{example}
@@
 Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion}, which are special to multilevel schemes based on defect constraints, hold regardless of whether a CD operator $\Pi_i$ can be constructed.
-Here there is, in fact, another ``basic'' iteration, an incomplete variant of \pr{cd-mult} (Section \ref{sec:cd}) that does not need the $\Pi_i$ operators.
-One application of \pr{icd-tele} below assumes $w^J\in\mathcal{K}^J$, property \eqref{eq:fe:finestlevelequivalent}, and incomplete CD \eqref{eq:fe:upwardsuminclusion}.
+Correspondingly, there is another basic iteration that does not need the $\Pi_i$ operators, namely an incomplete variant of \pr{cd-mult} (Section \ref{sec:cd}).
+Application \pr{icd-tele} below requires $w^J\in\mathcal{K}^J$, property \eqref{eq:fe:finestlevelequivalent}, and incomplete CD \eqref{eq:fe:upwardsuminclusion}.
 
@@
 
-While inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in Algorithm \ref{alg:fascd} below, a convergence analysis, perhaps extending the arguments of \cite{Tai2003} or \cite{GraeserKornhuber2009} to VIs which are not of minimization type, and/or to box constraints, would seem to require full CDs.
-We cannot currently offer that analysis.
+While inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in Algorithm \ref{alg:fascd} below, a convergence analysis extending the arguments of \cite{GraeserKornhuber2009,Tai2003} to VIs which are not of minimization type, or to box constraints, would seem to require full CDs.
 
@@
 \begin{center}
-\includegraphics[width=0.95\textwidth]{fixfigs/sialev8scene.png}
+\includegraphics[width=1.0\textwidth]{fixfigs/sialev8scene.png}
 \end{center}
-\caption{Solution of Example \ref{ex:sia} over a bumpy bedrock topography, with the ice speed as surface coloring, and 100-times vertical exageration.}
+\caption{Solution of Example \ref{ex:sia} over a bumpy bedrock topography (grey), with the ice speed as surface coloring and 100-times vertical exageration.}
 \label{fig:results:siascene}
 \end{figure}
-\pef{Could we make the bedrock grey?}
 

### 2023-07-14 paper/tikz/directRbadcoarse.tex
do a PEF (color!) and clean some text

\begin{tikzpicture}[scale=1.5]
  \draw[blue,thick] (0.0,1.5) -- (4.0,1.5)
                     node [xshift=-6mm,yshift=4mm] {{\color{black} $\underline{\gamma}^0 = \maxR\underline{\gamma}^1$?}};
  \draw[red,thick] (0.0,1.0) -- (4.0,1.0)
                     node [xshift=-6mm,yshift=-4mm] {{\color{black} $\overline{\gamma}^1 = \minR\overline{\gamma}^1$?}};
  \draw[black,thin,->] (-0.3,0.0) -- (4.3,0.0) node [xshift=5mm] {$x$};
  \foreach \x in {0.0, 2.0, 4.0} {
      \filldraw (\x,0.0) circle (1.5pt);
  }
\end{tikzpicture}

### 2023-07-15 paper/fascd.tex
draft a discussion and outlook section

@@
 \item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
-\item By using a Newton-type active set method as a smoother, with a fixed number of (Newton and Krylov) iterations per step, so that the work of the smoother is proportional to the number of unknowns on the given level, we demonstrate that F-cycle FAScd is an optimal solver for the examples tested (Section \ref{sec:results}).
+\item By using a Newton-type active set method as a smoother, with a fixed number of (Newton and Krylov) iterations per step, so that the work of the smoother is proportional to the number of unknowns on the given level, we demonstrate that F-cycle FAScd is an optimal solver for the differential VI examples in Section \ref{sec:results}.
 \end{enumerate}
@@
 
-FIXME For the classical obstacle problem, based on the linear, symmetric Laplacian operator, certain multilevel techniques, such as the truncated monotone multigrid method \cite{GraeserKornhuber2009,Kornhuber1994}, are known to improve performance relative to the original multilevel CD method \cite{Tai2003}.
-These improved methods modify the nodal basis functions.
-By contrast FAScd makes essentially no assumptions about the FE discretization.
-Acceleration via basis-level manipulations is a topic for future research.
-
-FIXME So the difference between FAScd and \cite{GraeserKornhuber2009} are: (1) FAS coarse corrections, (2) no attempt to do monotone/truncated multigrid which would be a mess (I think) across the FE infrastructure, (3) clarity on what sets you can use in $V(\nu_1,\nu_2)$ cycles, with bigger up-smoother sets in particular, and most importantly (4) implementation within the mature Firedrake/PETSc library framework where one can easily hook up an active-set Newton step as a smoother.
-
-FIXME  It is hard to say which aspect of FAScd is more important to beating monotone multigrid: stronger smoother or improved up-smooth CD.
-Probably the former, but for now we don't have any illuminating analysis, and a head-to-head race is comparing apples and oranges because we are using such a strong smoother.
-Various differences of implementation don't produce any really good multigrid convergence if your smoother is NGS on each level.
-I think the convergence proofs in \cite{GraeserKornhuber2009} all depend on this?
-
-FIXME the fact that FAScd reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}) suggests it is a completely different strategy from PFAS in \cite{BrandtCryer1983}
-
-FIXME main difference relative to monotone multigrid \cite{Kornhuber1994} and PFAS \cite{BrandtCryer1983} is that method is not tied to, or presented in terms of, an PNGS type smoother
-
-FIXME at a high level one might say FAScd merely generates coarse VI problems which, once solved, make helpful admissible additions to the eventual V-cycle update; the method is smoother-agnostic; smoother could be Newton semi-smooth \cite{BensonMunson2006}, interior point method, etc.
-
-FIXME VINewton smoother could use GMG or AMG on linear step equations, but this was not seen to be necessary in applications so far and was not even tested
-
-FIXME the observation that up-smoothing is more efficient, with consequent asymmetry between $V(1,0)$ and $V(0,1)$ performance, seems to be new; compare the comments on V(1,0) and V(1,1) cycles in \cite{GraeserKornhuber2009,Tai2003}
-
-FIXME desired extension to nonlocal variational inequalities not tried here
+As a multilevel solver for VI problems, one might describe the new FAScd method as merely a strategy for generating coarser-level problems which, once solved, make helpful admissible additions to a final V-cycle update.
+The method is smoother-agnostic, and our implementation within the extensible Firedrake \cite{Rathgeberetal2016} and PETSc \cite{Balayetal2023} library framework should allow easy experimentation with other smoothers (projected Gauss-Seidel, semi-smooth Newton, interior-point, etc.), beyond the active-set Newton method \cite{BensonMunson2006} chosen for demonstration here.
+For box-constrained smoothers of Newton-Krylov type one could also add linear geometric or algebraic multigrid preconditioning \cite{Trottenbergetal2001}, instead of our simple choice of incomplete factorization and (in parallel) domain decomposition preconditioners.
+Thus there is a large smoother and coarse solver space to explore.
+However, performance of the particular method shown in Section \ref{sec:results} is already quite good.
+
+The observation that up-smoothing is more efficient in a multilevel CD context, which arises because of constraint set distinctions, seems to be new, but compare comments on $V(1,1)$ cycles in \cite{GraeserKornhuber2009,Tai2003}.
+The increased efficiency is shown in Example \ref{ex:results:plap} particularly.
+In summary, the distinction is that construction of constraint sets via monotone injection generates up-smoothing sets which can telescope the down-smoothing sets (Lemma \ref{lem:upwardadmissibility}); we call this an \emph{incomplete CD}.
+This structure can be exploited by a new CD iteration, algorithm \pr{icd-tele} in Section \ref{sec:cdmultilevel}.
+
+In contrast with monotone multigrid \cite{Kornhuber1994}, FAScd does not modify the FE discretization, but the possibility of acceleration via basis-level manipulations is a topic for future research.
+On the other hand,  the convergence proofs in \cite{GraeserKornhuber2009,Kornhuber1994}, which appear to depend on particular smoother and basis-level manipulation choices, have no analog here.
+
+%While an an extension to nonlocal, non-differential VI problems is of interest, it was not tried here
 

### 2023-07-15 paper/fascd.tex
the V-cycle visualization Fig 8.3 is better with one fewer level

@@
 % figures in the following visualization were generated using branch elbueler/simplefigs
-% of git@bitbucket.org:pefarrell/fascd.git, and the following commands:
+% of git@bitbucket.org:pefarrell/fascd.git, and the following commands (including imagemagick):
 %   cd examples/
-%   python3 plap1d.py -figures -fascd_max_it 2 -fascd_cycle_type multiplicative -mcoarse 2
+%   python3 plap1d.py -fascd_cycle_type multiplicative -mcoarse 4 -levs 3 -figures
+%   [save the images into mcd-extended/paper/fixfigs/vcycle/]
+%   for X in start2 down2 down1 up0 up1 up2 next2; do mogrify -trim $X.png $X.png; done
 \begin{figure}[ht]
@@
 \end{center}
-\caption{Visualization of one FAScd V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of $m_0=3$ vertices.
-Solutions (i.e.~$w^j$, $y^j$, $z^j$) are lines while obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots.
-The zero function is shown (dotted), but vertical scale varies.}
+\caption{Visualization of one FAScd V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of $m_0=5$ vertices.
+Solutions (i.e.~$w^j$, $y^j$, $z^j$) are lines, obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots, and the zero function is dotted.
+Vertical scale varies.}
 \label{fig:imagesvcycle}

### 2023-07-15 paper/tikz/imagesvcycle.tex
the V-cycle visualization Fig 8.3 is better with one fewer level

@@
     \pgfmathsetmacro{\NUDGE}{0.5}
-    \node (start3) [] at (0,3*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/start3.png}};
-    \node (start3label) [below of = start3, yshift=\YLABEL cm] {$w^3 \ge \underline{\gamma}^3$};
-    \node (down3) [] at (1.0*\XSTEP,3*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/down3.png}};
-    \node (down3label) [below of = down3, yshift=\YLABEL cm] {$y^3 \ge \underline{\phi}^3$};
-    \node (down2) [] at (1.2*\XSTEP,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/down2.png}};
+    \node (start2) [] at (0,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/start2.png}};
+    \node (start2label) [below of = start2, yshift=\YLABEL cm] {$w^2 \ge \underline{\gamma}^2$};
+    \node (down2) [] at (1.0*\XSTEP,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/down2.png}};
     \node (down2label) [below of = down2, yshift=\YLABEL cm] {$y^2 \ge \underline{\phi}^2$};
@@
     \node (up1label) [below of = up1, yshift=\YLABEL cm] {$z^1 \ge \underline{\chi}^1$};
-    \node (up2) [] at (2.6*\XSTEP,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/up2.png}};
+    \node (up2) [] at (2.8*\XSTEP,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/up2.png}};
     \node (up2label) [below of = up2, yshift=\YLABEL cm] {$z^2 \ge \underline{\chi}^2$};
-    \node (up3) [] at (2.8*\XSTEP,3*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/up3.png}};
-    \node (up3label) [below of = up3, yshift=\YLABEL cm] {$z^3 \ge \underline{\chi}^3$};
-    \node (next3) [] at (3.8*\XSTEP,3*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/next3.png}};
-    \node (next3label) [below of = next3, yshift=\YLABEL cm] {$\tilde{w}^3 \ge \underline{\gamma}^3$};
+    \node (next2) [] at (3.8*\XSTEP,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/next2.png}};
+    \node (next2label) [below of = next2, yshift=\YLABEL cm] {$\tilde{w}^2 \ge \underline{\gamma}^2$};
     \graph [edges = {very thick}] {
-        (start3) -> (down3)
-                 -!- [] (down3label) -> (down2)
+        (start2) -> (down2)
                  -!- [] (down2label) -> (down1)
@@
                  -> (up2label) -!- (up2)
-                 -> (up3label) -!- (up3)
-                 -> (next3)
+                 -> (next2)
     };

### 2023-07-17 paper/fascd.tex
Another editing pass

@@
 We extend the CD idea by exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
-Then, by applying strong smoothers, with work proportional to the number of unknowns on a given mesh level, we demonstrate that FAScd V-cycles exhibit excellent mesh-independent convergence rates and that FAScd F-cycles (full multigrid) are optimal solvers.
+We demonstrate that FAScd V-cycles exhibit excellent mesh-independent convergence rates and that FAScd F-cycles (full multigrid) are optimal solvers, when used in combination with reduced-space active set smoothers with work proportional to the number of unknowns on a given mesh level.
 The example problems include unilateral and bilateral VI problems, with differential operators which are symmetric linear, nonsymmetric linear, and nonlinear.
@@
 The constraint decomposition (CD) methods of Tai \cite{Tai2003} are designed to solve coercive variational inequality (VI) problems arising as the Karush--Kuhn--Tucker optimality conditions for the minimization of a convex functional over a convex set.
-These methods were shown to converge for multilevel and domain decompositions in a finite element (FE) context.
-In the multilevel case, the CD method has almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}, \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
-In this paper we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach, and we apply this new solver to a variety of general, i.e.~nonlinear and non-optimization, VI problems.
+These methods were shown to converge for multilevel and domain decompositions in a finite element (FE) context; in the multilevel case, the CD method has almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}, \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
+\pef{I prefer the semicolon here because the second sentence is filling in the details for the first; without it, at the end of the first sentence, I'm thinking ``there's a citation missing here''.} In this paper we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach, and we apply this new solver to a variety of general, i.e.~nonlinear and non-optimization, VI problems.
+\pef{Could Tai have applied his algorithm directly to non-optimization problems? I don't think so. If that is true, we're underselling our work by using the word `apply'.}
 
@@
 \item The FAS coarse corrections are adapted for general VI problems. The needed computations require neither an objective function nor linearity of the residual operator.
+\item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
 \item Box constraints, namely upper and lower obstacles, are permitted.
+\pef{Should we list this? \cite{GraeserKornhuber2009} say ``We emphasize that all algorithms and convergence results to be presented can be generalized to where an upper obstacle is present''}
 \item We implement V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}).
-\item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
-\item By using a Newton-type active set method as a smoother, with a fixed number of (Newton and Krylov) iterations per step, so that the work of the smoother is proportional to the number of unknowns on the given level, we demonstrate that F-cycle FAScd is an optimal solver for the differential VI examples in Section \ref{sec:results}.
+\pef{\dots for the first time for VIs beyond linear complementarity problems?}
+\item We demonstrate that F-cycle FAScd is an optimal solver for the different VI examples in Section \ref{sec:results}, when used in conjunction with reduced-space active set methods as smoothers.
+The number of Newton and Krylov iterations per step are fixed, so that the work of the smoother is proportional to the number of unknowns on a given level.
 \end{enumerate}
 
-Prior multilevel algorithms for VIs have various substantial limitations relative to the FAScd framework.
-The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} is apparently limited to linear complementarity problems and projected (truncated) Gauss-Seidel smoothers.
+Prior multilevel algorithms for VIs have various substantial limitations compared to the FAScd framework.
+The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} is apparently limited to linear complementarity problems and projected (truncated) Gauss--Seidel smoothers.
 While FAScd reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed (Appendix \ref{app:reductions}), PFAS apparently reduces to linear multigrid.
-The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so it only applies (as stated) to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
+\pef{I don't like the 2$\times$ apparently.} The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so as stated it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
 The problematic latter property is corrected by reformulation using defect constraints and monotone injection operators (see \cite[Algorithm 4.7]{GraeserKornhuber2009} and Section \ref{sec:cdmultilevel} herein).
-The monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be implemented only with a Gauss-Seidel type smoother, and applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.
+The monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be formulated only with Gauss--Seidel type smoothers, and applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.
 Furthermore, monotone multigrid methods are more difficult to implement in FE libraries because modifications of the FE basis functions are needed to approximate the coarser-level coincidence sets.
 No such invasive basis modifications are used either in Tai's original multilevel CD or in FAScd.
-Our Firedrake library \cite{Rathgeberetal2016} implementation of FAScd requires no modifications of basis functions and yet it exhibits excellent performance, including on a problem used to demonstrate the benefits of monotone multigrid (Example \ref{ex:results:classical}; compare \cite[problem 7.1.1]{GraeserKornhuber2009}).
+Our implementation of FAScd in the Firedrake library \cite{Rathgeberetal2016} requires no modifications of basis functions and yet exhibits excellent performance, including on a problem used to demonstrate the benefits of monotone multigrid (Example \ref{ex:results:classical}; compare \cite[problem 7.1.1]{GraeserKornhuber2009}).
 
@@
 The F-cycle results in Section \ref{sec:results} show nearly-textbook multigrid efficiency, namely solutions within discretization error in work comparable to few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}.
-Specifically we see this performance for classical Laplacian obstacle problems, $p$-Laplacian obstacle problems, and nonnegativity-constrained advection-diffusion equations.
+Specifically, we see this performance for classical Laplacian obstacle problems, $p$-Laplacian obstacle problems, and nonnegativity-constrained advection-diffusion equations.
 Furthermore, these results appear to be insensitive to the length or structure of the free boundary, i.e.~the geometric complexity of the coincidence set.
 
-All iterates in a FAScd cycle are admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
+All iterates in FAScd are admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
 For example, we solve an ice sheet geometry problem (Examples \ref{ex:sia} and \ref{ex:results:sia}) in which the ice flow formulas are only meaningful for surface elevation iterates that do not penetrate the bedrock.
 (More generally, fluid layers subject to a non-negative thickness constraint typically require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation \cite{Bueler2021conservation}, but this is not needed in the ice sheet problem here.)  Non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural modifications of the operator formula.
-FAScd solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a mesh-independent number of F-cycle iterations.
+FAScd solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a mesh-independent number of F-cycle iterations (Example \ref{ex:results:sia}).
 
 The paper is organized as follows.
-In Section \ref{sec:vi} we recall the theory of coercive and monotone variational inequalities, with several example problems.
+In Section \ref{sec:vi} we recall the theory of coercive and monotone variational inequalities, and describe several motivating example problems.
 Section \ref{sec:cd} reviews the basics of CDs and their associated iterations.
@@
 \end{equation}
-On the other hand, it is easy to see that $|\ip{f(u)}{v}| \le \left((\eps + \|\bX\|_\infty) \|u\| + \|\phi\|'\right) \|v\|$, thus that $f:\cK \to \cV'$ is continuous.
+On the other hand, it is easy to see that $|\ip{f(u)}{v}| \le \left((\eps + \|\bX\|_\infty) \|u\| + \|\phi\|'\right) \|v\|$, so that $f:\cK \to \cV'$ is continuous.
 Lemma \ref{lem:advectionskew} says that the bilinear form $b(u,v)$ is skew-symmetric up to a nonnegative term.
@@
 For general beds $b$ less is understood about this operator, but existence holds for \eqref{eq:siavi} \cite{JouvetBueler2012} and a power of the solution lives in a Sobolev space: $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.
-This theory, and also observations of ice sheets, shows that $|\grad s|$ is generally unbounded as one approaches the free boundary from the $s>b$ side.
+This theory, and also observations of ice sheets, shows that $|\grad s|$ is generally unbounded as one approaches the free boundary from the icy ($s>b$) side.
 \end{example}
@@
 \end{example}
-
-When $f$ has the locality of Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over basis function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}  However, for general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of CD-type iterations.
+\pef{I confess I just don't get the point of this example.
+Why is this a substantial implementation simplification?
+I am sorry I am being dumb.
+Maybe you can explain it to me on a video call and I will write it in a way that I would have understood it?}
+
+When $f$ has the locality of Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over basis function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}  However, for general problems lacking the convenience of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of CD-type iterations.
 Section \ref{sec:vcycle} constructs such an iteration for a general operator $f:\cK\to\cV'$ by applying the full approximation scheme (FAS) idea of Brandt \cite{Brandt1977}.
@@
 the closed and convex \emph{constraint set}.
-Note that we are treating inequality constraints and Dirichlet boundary conditions in a unified manner.
-Any Neumann conditions over $\partial_N \Omega$ are imposed weakly in the definition of the operator $f$, i.e.~by an integration over $\partial_N\Omega$.
+Note that we are treating inequality constraints and strongly-imposed Dirichlet boundary conditions in a unified manner.
+Any Neumann conditions over $\partial_N \Omega$ are imposed weakly in the definition of the operator $f$, i.e.~by an integral over $\partial_N\Omega$.
 
@@
 Our multilevel CD scheme uses \emph{defect constraints} \cite{GraeserKornhuber2009} derived from an admissible finest-level iterate.
-The next Example illustrates why we do \emph{not} apply restriction operators directly\footnote{An exception is in the F-cycle; see Section \ref{sec:implementation}.} to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.
+The next Example illustrates why we do \emph{not} apply injection operators directly\footnote{An exception is in the F-cycle; see Section \ref{sec:implementation}.} to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.
 Iterate-dependent defect constraints, definition \eqref{eq:fe:defectconstraints} below, satisfy ordering \eqref{eq:fe:chiordering} and bypass these difficulties.
@@
 Ignoring the boundary for simplicity, the fine-level constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.
-However, various restriction schemes to coarsen the constraints, applied directly to $\underline{\gamma}^1,\overline{\gamma}^1$, are problematic.
+However, various natural schemes to coarsen the constraints, applied directly to $\underline{\gamma}^1,\overline{\gamma}^1$, are problematic.
 Suppose we propose
@@
 \begin{figure}[ht]
+\begin{tabular}{cc}
 \input{tikz/directRbadfine.tex}
-\quad
+&
 \input{tikz/directRbadcoarse.tex}
+\end{tabular}
 \caption{For the finer-level constraints $\underline{\gamma}^1$ (blue) and $\overline{\gamma}^1$ (red) on the left, direct application of $\iR,\minR,\maxR$ is problematic.
@@
 So suppose $w^J \in \cK^J$ is an admissible finest-level iterate for \eqref{eq:fe:vi}.
-We define the \emph{finest-level defect constraints} (DCs) \cite{GraeserKornhuber2009} as the following nodal functions in $\tilde{\mathcal{V}}^J$:
+We define the \emph{finest-level defect constraints} (DCs) \cite{GraeserKornhuber2009} as the following functions in $\tilde{\mathcal{V}}^J$:
 \begin{equation}
@@
 This construction of multilevel box constraints, as for the unilateral construction in \cite{GraeserKornhuber2009}, is designed to permit large-yet-admissible coarse corrections, which promotes multilevel solver efficiency.
-However, when coarsening the problem downward in a V-cycle one does not know what the next coarse corrections will be, and so we require constraint sets from which any sum will be admissible.
+However, when coarsening the problem downward in a V-cycle, one does not know what the next coarse corrections will be, and so we require constraint sets from which any sum will be admissible.
 With this in mind, and looking forward to telescoping sums \eqref{eq:fe:telescoping}, we compute the LDC differences \cite{GraeserKornhuber2009}:
@@
 For concreteness suppose $\overline{\gamma}^J=+\infty$, thus $\overline{\chi}^j=+\infty$ and $\overline{\phi}^i = +\infty$ for all levels.
-For $v\in \mathcal{V}^j$ and $0\le i \le j$, let $I_{j\to i}^\ominus$ be the operator applying the minimum (monotone) restriction $j-i$ times, mapping into $\mathcal{V}^i$:
+For $v\in \mathcal{V}^j$ and $0\le i \le j$, let $I_{j\to i}^\ominus$ be the operator applying the minimum (monotone) injection $j-i$ times, mapping into $\mathcal{V}^i$:
 \begin{equation}
@@
 Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.
-Then the next finest-level, admissible iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.
+Then the corrected finest-level iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.
 
@@
 Observe that injection $\iR$ coarsens the (primal) iterates---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---but canonical restriction $R$ coarsens (dual) linear functionals; recall Table \ref{tab:transfers}.
-At levels $j=J,J-1,\dots,1$, we solve the following VI for $y^j \in \mathcal{D}^j$:
+On descent, at levels $j=J,J-1,\dots,1$, we solve the following VI for $y^j \in \mathcal{D}^j$:
 \begin{equation}
@@
 \item replace the residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction,
-\item replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its restriction, and
+\item replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its injection, and
 \item where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.
@@
 
-Regarding the storage required, let us assuming all box constraints are nontrivial.
+Regarding the storage required, let us assume all box constraints are nontrivial.
 In this case, seven vectors must be allocated on each level: $\underline{\chi}^j$, $\overline{\chi}^j$, $\underline{\phi}^j$, $\overline{\phi}^j$, $w^j$, $\ell^j$, $y^j$.
@@
 Let $r(u^J)_p = \ip{f^J(w^J)-\ell^J}{\psi_p}$ be the pointwise (ordinary) residual.
-If $x_p \notin \partial_D\Omega$ then one of these four conditions hold:
+If $x_p \notin \partial_D\Omega$ then one of these four conditions must hold:
 \begin{itemize}
@@
 
-This MCP interpretation allows us to construct expressions whose norms should go below tolerance.
+This MCP interpretation allows us to construct expressions whose norms should approach zero.
 For $w^J \in \mathcal{K}^J$ we define the finest-level \emph{nodal complementarity residual} $\rNC(w^J)$ from the five cases:
@@
 Then an admissible initial iterate on the next-finer mesh is generated via prolongation, and so on to finer levels.
-(In the VI context one must also truncate after the prolongation; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fcycle}.)  During this ``ramp'' one approximately solves on each level by using a fixed number of V-cycles, each down to the coarsest mesh, and then one does full-depth V-cycles from the finest level, to convergence.
+(In the VI context one must also truncate to the bounds after the prolongation; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fcycle}.)  During this ``ramp'' one approximately solves on each level by using a fixed number of V-cycles, each down to the coarsest mesh, and then one does full-depth V-cycles from the finest level, to convergence.
 
@@
 This F-cycle approach is described in Algorithm \ref{alg:fascd-fcycle}.
-Injection is used to generate the $\underline{\gamma}^j,\overline{\gamma}^j$ on each level.\footnote{Contrast the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, based on an existing admissible iterate.}  Regarding line \ref{line:fcyclecoarsestinitial}, one may generate an initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation of a finest-level iterate, if one is available.
+Injection is used to generate the $\underline{\gamma}^j,\overline{\gamma}^j$ on each level.\pef{Why is this fine for F-cycles, but not for V-cycles?
+I think we should spell it out.}\footnote{Contrast the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, based on an existing admissible iterate.}  Regarding line \ref{line:fcyclecoarsestinitial}, one may generate an initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation of a finest-level iterate, if one is available.
 
@@
 Here we show numerical results for the example problems presented in Section \ref{sec:vi}, with a focus on the scaling of FAScd iterations as the mesh resolution increases.
-Our primary observations are that each V-cycle dramatically reduces the residual norm, and that very few F-cycles solve most examples to within discretization error.
+Our primary observations are that each V-cycle greatly reduces the residual norm, and that very few F-cycles solve most examples to within discretization error.
 
@@
 Comparing the ball and spiral results, the complexity of the coincidence set or free boundary does not appear to strongly influence FAScd performance.
-For the spiral results, note that \cite{GraeserKornhuber2009} reports more than 30 iterations of all their tested methods to achieve similar tolerance for $m_J \approx 5\times 10^5$.
+For the spiral results, note that \cite{GraeserKornhuber2009} reports more than 30 iterations of all their tested methods were required to achieve similar tolerance for $m_J \approx 5\times 10^5$.
 
@@
 Let $\Omega=(-1,1)^d$ for $d=2,3$, the diffusivity $\eps=0.1$ (equation \eqref{eq:advectiondiffusionstrong}), and we impose zero Dirichlet boundary conditions.
-The 2D problem satisfied the coercivity hypothesis with a divergence-free velocity field $\bX = (7+5y,-5x)$, but in 3D, for varieties sake, we set $\bX = (7+5y,-5x,2z)$ so $\grad\cdot\bX \ne 0$.
+The 2D problem satisfied the coercivity hypothesis with a divergence-free velocity field $\bX = (7+5y,-5x)$, but in 3D, for variety's sake, we set $\bX = (7+5y,-5x,2z)$ so $\grad\cdot\bX \ne 0$.
 The discontinuous source term $\phi \in L^\infty(\Omega)$ satisfied $\phi\ge 0$ in half of the domain, with positive circular patches, while values were negative in the other half; details are in the associated code.
@@
 We applied FAScd F-cycles using $Q_1$ elements and a smoother of four (fixed) iterations of RS Newton (Section \ref{sec:implementation}), with quadratic backtracking line search.
-The linear step equations were approximately solved, in parallel, by three (fixed) iterations of preconditioned GMRES and a ASM+ILU preconditioner.
+The arising linear systems were approximately solved, in parallel, by three (fixed) iterations of preconditioned GMRES and a ASM+ILU preconditioner.
 The coarse solver was RS+LU, applied redundantly over all processors.
@@
 As a multilevel solver for VI problems, one might describe the new FAScd method as merely a strategy for generating coarser-level problems which, once solved, make helpful admissible additions to a final V-cycle update.
-The method is smoother-agnostic, and our implementation within the extensible Firedrake \cite{Rathgeberetal2016} and PETSc \cite{Balayetal2023} library framework should allow easy experimentation with other smoothers (projected Gauss-Seidel, semi-smooth Newton, interior-point, etc.), beyond the active-set Newton method \cite{BensonMunson2006} chosen for demonstration here.
-For box-constrained smoothers of Newton-Krylov type one could also add linear geometric or algebraic multigrid preconditioning \cite{Trottenbergetal2001}, instead of our simple choice of incomplete factorization and (in parallel) domain decomposition preconditioners.
+The method is smoother-agnostic, and our implementation within the extensible Firedrake \cite{Rathgeberetal2016} and PETSc \cite{Balayetal2023} library framework should allow easy experimentation with other smoothers (projected Gauss--Seidel, semi-smooth Newton, interior-point, etc.), beyond the active-set Newton method \cite{BensonMunson2006} chosen for demonstration here.
+For box-constrained smoothers of Newton-Krylov type one could also add linear geometric or algebraic multigrid preconditioning \cite{Trottenbergetal2001}, instead of our simple choice of incomplete factorization and (in parallel) additive Schwarz preconditioners.
 Thus there is a large smoother and coarse solver space to explore.
-However, performance of the particular method shown in Section \ref{sec:results} is already quite good.
+However, the performance of the particular method shown in Section \ref{sec:results} is already excellent.
 
 The observation that up-smoothing is more efficient in a multilevel CD context, which arises because of constraint set distinctions, seems to be new, but compare comments on $V(1,1)$ cycles in \cite{GraeserKornhuber2009,Tai2003}.
-The increased efficiency is shown in Example \ref{ex:results:plap} particularly.
+The increased efficiency is particularly shown in Example \ref{ex:results:plap}.
 In summary, the distinction is that construction of constraint sets via monotone injection generates up-smoothing sets which can telescope the down-smoothing sets (Lemma \ref{lem:upwardadmissibility}); we call this an \emph{incomplete CD}.
@@
 
-\subsection*{Acknowledgements} \label{sec:acknowledgements}  EB was supported by a Faculty Development Travel Award from United Academics, University of Alaska Fairbanks.  FIXME PF grant support?
+\subsection*{Acknowledgements} \label{sec:acknowledgements}  EB was supported by a Faculty Development Travel Award from United Academics, University of Alaska Fairbanks.
+PEF was supported by EPSRC grants EP/R029423/1 and EP/W026163/1.
 

### 2023-07-17 paper/tikz/directRbadcoarse.tex
Another editing pass

@@
+\definecolor{seabornblue}{rgb}{0.2980392156862745, 0.4470588235294118, 0.6901960784313725}
+\definecolor{seaborngreen}{rgb}{0.3333333333333333, 0.6588235294117647, 0.40784313725490196}
+\definecolor{seabornred}{rgb}{0.7686274509803922, 0.3058823529411765, 0.3215686274509804}
+
 \begin{tikzpicture}[scale=1.5]
-  \draw[blue,thick] (0.0,1.5) -- (4.0,1.5)
+  \draw[seabornblue,very thick] (0.0,1.5) -- (4.0,1.5)
                      node [xshift=-6mm,yshift=4mm] {{\color{black} $\underline{\gamma}^0 = \maxR\underline{\gamma}^1$?}};
-  \draw[red,thick] (0.0,1.0) -- (4.0,1.0)
+  \draw[seabornred,very thick] (0.0,1.0) -- (4.0,1.0)
                      node [xshift=-6mm,yshift=-4mm] {{\color{black} $\overline{\gamma}^1 = \minR\overline{\gamma}^1$?}};

### 2023-07-17 paper/fascd.tex
Add a diagram

@@
 On the other hand, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These ordering properties are used in the CD construction of the next two Sections.
+\begin{figure}[ht]
+\input{tikz/Rplusminus.tex}
+\caption{Evaluation of $\maxR$ and $\minR$.
+The coarse mesh is shown in black and the fine mesh in grey.
+The monotone injection operators $\maxR$ and $\minR$ assign to a coarse degree of freedom (in blue) the maximum and minimum respectively of the fine function over the patch of coarse cells sharing its vertex (in green; the fine degree of freedom at the same vertex is not shown).}
+\end{figure}
 \pef{It would be good to have a diagram here. I could write one in tikz, if you don't have one handy.}
@@
 \end{center}
-\caption{Solution of Example \ref{ex:sia} over a bumpy bedrock topography (grey), with the ice speed as surface coloring and 100-times vertical exageration.}
+\caption{Solution of Example \ref{ex:sia} over a bumpy bedrock topography (grey), with the ice speed as surface coloring and 100-times vertical exaggeration.}
+\pef{Why does the colour bar have grey in it?}
 \label{fig:results:siascene}
@@
 \subsection*{Acknowledgements} \label{sec:acknowledgements}  EB was supported by a Faculty Development Travel Award from United Academics, University of Alaska Fairbanks.
-PEF was supported by EPSRC grants EP/R029423/1 and EP/W026163/1.
+PEF was supported by the Engineering and Physical Sciences Research Council [EPSRC grants EP/R029423/1 and EP/W026163/1].
 

### 2023-07-17 paper/tikz/Rplusminus.tex
Add a diagram

\definecolor{seabornblue}{rgb}{0.2980392156862745, 0.4470588235294118, 0.6901960784313725}
\definecolor{seaborngreen}{rgb}{0.3333333333333333, 0.6588235294117647, 0.40784313725490196}
\definecolor{seabornred}{rgb}{0.7686274509803922, 0.3058823529411765, 0.3215686274509804}

\begin{tikzpicture}[scale=2.0]
  \foreach \k in {0, 1, 2, 3, 4} {
    %\filldraw[seabornblue, fill=seabornblue] ({cos(deg(2*\k*pi/5))}, {sin(deg(2*\k*pi/5))}) circle (1.5pt);
    \node (A) at ({0.5*cos(deg(2*\k*pi/5)) + 0.5*cos(deg(2*(\k+1)*pi/5))}, {0.5*sin(deg(2*\k*pi/5)) + 0.5*sin(deg(2*(\k+1)*pi/5))}) {};
    \node (B) at ({0.5*cos(deg(2*\k*pi/5))}, {0.5*sin(deg(2*\k*pi/5))}) {};
    \node (C) at ({0.5*cos(deg(2*(\k+1)*pi/5))}, {0.5*sin(deg(2*(\k+1)*pi/5))}) {};
    \node (D) at ({cos(deg(2*\k*pi/5))}, {sin(deg(2*\k*pi/5))}) {};
    \node (E) at ({cos(deg(2*(\k+1)*pi/5))}, {sin(deg(2*(\k+1)*pi/5))}) {};

    \draw [-,gray, very thick] (A.center) -- (B.center);
    \draw [-,gray, very thick] (B.center) -- (C.center);
    \draw [-,gray, very thick] (C.center) -- (A.center);

    \draw [black, ultra thick] (0, 0) -- (D.center);
    \draw [black, ultra thick] (D.center) -- (E.center);

    \filldraw[seaborngreen] (A) circle (1.5pt);
    \filldraw[seaborngreen] (B) circle (1.5pt);
    \filldraw[seaborngreen] (C) circle (1.5pt);
    \filldraw[seaborngreen] (D) circle (1.5pt);
    \filldraw[seaborngreen] (E) circle (1.5pt);

  }
  \filldraw[seabornblue] (0, 0) circle (1.5pt);
  \node (expl) at (1.5, 0.5) {};
  \filldraw[seabornblue] (expl) circle (1.5pt);
  \node at (2.25, 0.525) {$= \text{max/min of }$};  % I'm sure there's a better way to place this, but tikz mystifies me
  \filldraw[seaborngreen] (2.95, 0.5) circle (1.5pt);
\end{tikzpicture}

### 2023-07-18 paper/fascd.tex
A comment
accept many of PEF, and other fiddles; PEF remain

@@
 These methods were shown to converge for multilevel and domain decompositions in a finite element (FE) context; in the multilevel case, the CD method has almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}, \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
-\pef{I prefer the semicolon here because the second sentence is filling in the details for the first; without it, at the end of the first sentence, I'm thinking ``there's a citation missing here''.} In this paper we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach, and we apply this new solver to a variety of general, i.e.~nonlinear and non-optimization, VI problems.
-\pef{Could Tai have applied his algorithm directly to non-optimization problems? I don't think so. If that is true, we're underselling our work by using the word `apply'.}
-
+%\pef{I prefer the semicolon here because the second sentence is filling in the details for the first; without it, at the end of the first sentence, I'm thinking ``there's a citation missing here''.}
+% Agreed
+%\pef{Could Tai have applied his algorithm directly to non-optimization problems? I don't think so. If that is true, we're underselling our work by using the word `apply'.}
+% Agreed.  Text compressed; the first item on the list already says this.
+In this paper we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach.
 Specifically, our \emph{full approximation scheme constraint decomposition} (FAScd) method extends the algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} in the following five directions:
@@
 \item Box constraints, namely upper and lower obstacles, are permitted.
-\pef{Should we list this? \cite{GraeserKornhuber2009} say ``We emphasize that all algorithms and convergence results to be presented can be generalized to where an upper obstacle is present''}
+%\pef{Should we list this? \cite{GraeserKornhuber2009} say ``We emphasize that all algorithms and convergence results to be presented can be generalized to where an upper obstacle is present''}
+% Yes.
+"can be generalized" is true, but in doing so you discover that (see Section 4) the \Pi_i operators in the general box-constrained CD must have a very different formula.
+We actually address this by writing code that works at this level of generality.
+At a certain point, if G&K can't make something clear and demo that it works then they did not do it!
+We will see what we get from referees ...
 \item We implement V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}).
-\pef{\dots for the first time for VIs beyond linear complementarity problems?}
+%\pef{\dots for the first time for VIs beyond linear complementarity problems?}
+% For F-cycles, maybe.
+If something like this has been done before I want a referee to say "here it is ...".
+I'd rather not guess what unpublished, undocumented, and closed extensions people have to their methods.
+So I'm thinking just say what we did.
+See previous point.
 \item We demonstrate that F-cycle FAScd is an optimal solver for the different VI examples in Section \ref{sec:results}, when used in conjunction with reduced-space active set methods as smoothers.
@@
 The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} is apparently limited to linear complementarity problems and projected (truncated) Gauss--Seidel smoothers.
-While FAScd reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed (Appendix \ref{app:reductions}), PFAS apparently reduces to linear multigrid.
-\pef{I don't like the 2$\times$ apparently.} The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so as stated it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
+While FAScd reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed (Appendix \ref{app:reductions}), PFAS reduces to linear multigrid.
+%\pef{I don't like the 2$\times$ apparently.}
+% Me neither.  Fixed.
+The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so as stated it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
 The problematic latter property is corrected by reformulation using defect constraints and monotone injection operators (see \cite[Algorithm 4.7]{GraeserKornhuber2009} and Section \ref{sec:cdmultilevel} herein).
@@
 \end{example}
+
 \pef{I confess I just don't get the point of this example.
@@
 Maybe you can explain it to me on a video call and I will write it in a way that I would have understood it?}
+{\color{blue} ELB: Not being dumb.
+I am describing an application I am headed toward.
+FAScd applies to both differential \emph{and integral} operators.
+(We don't try it out in the latter case, but it is a feature that it does not care about the locality of your operator, unlike monotone multigrid.)  It applies to operators where $f(e_i)$ makes no sense because $e_i\notin \mathcal{K}$ (SIA is an example IMO).
+And it does FAS coarse corrections because we don't assume linearity of $f$.}
 
@@
 To make the construction precise, let $x_q^j$ denote the vertices of mesh $\mathcal{T}^j$, with corresponding basis functions $\psi_q^j(x)$.
-For a $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$:
+For a $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$ (Figure \ref{fig:Rplusminus}):
 \begin{align}
@@
 On the other hand, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These ordering properties are used in the CD construction of the next two Sections.
+
 \begin{figure}[ht]
 \input{tikz/Rplusminus.tex}
-\caption{Evaluation of $\maxR$ and $\minR$.
-The coarse mesh is shown in black and the fine mesh in grey.
-The monotone injection operators $\maxR$ and $\minR$ assign to a coarse degree of freedom (in blue) the maximum and minimum respectively of the fine function over the patch of coarse cells sharing its vertex (in green; the fine degree of freedom at the same vertex is not shown).}
+\caption{The monotone injection operators $\maxR$ and $\minR$ assign to a coarse degree of freedom (blue) the maximum and minimum respectively of the fine function over the patch of coarse cells sharing its vertex.
+The coarse mesh is in black and the fine in grey.}
+\label{fig:Rplusminus}
 \end{figure}
-\pef{It would be good to have a diagram here. I could write one in tikz, if you don't have one handy.}
+%\pef{It would be good to have a diagram here. I could write one in tikz, if you don't have one handy.}
+% Like it!  I shortened the caption by removing the parenthetical ("(in green; the fine degree of freedom at the same vertex is not shown)"); that's clear because of the coarse edges.
 
@@
 This F-cycle approach is described in Algorithm \ref{alg:fascd-fcycle}.
-Injection is used to generate the $\underline{\gamma}^j,\overline{\gamma}^j$ on each level.\pef{Why is this fine for F-cycles, but not for V-cycles?
-I think we should spell it out.}\footnote{Contrast the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, based on an existing admissible iterate.}  Regarding line \ref{line:fcyclecoarsestinitial}, one may generate an initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation of a finest-level iterate, if one is available.
+Injection is used to generate the $\underline{\gamma}^j,\overline{\gamma}^j$ on each level.\footnote{Contrast the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, based on an existing admissible iterate.}
+\pef{Why is this fine for F-cycles, but not for V-cycles? I think we should spell it out.}
+{\color{blue} ELB: This is the point of Example 5.1.}
+
+\noindent
+Regarding line \ref{line:fcyclecoarsestinitial}, one may generate an initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation of a finest-level iterate, if one is available.
 
@@
 \caption{Solution of Example \ref{ex:sia} over a bumpy bedrock topography (grey), with the ice speed as surface coloring and 100-times vertical exaggeration.}
-\pef{Why does the colour bar have grey in it?}
 \label{fig:results:siascene}
 \end{figure}
+%\pef{Why does the colour bar have grey in it?}
+% Cause Paraview does it.  But I have edited it out of the .png.
 
@@
 
+\pef{I wonder if there's a Firedrake installation we could use on a supercomputer somewhere. It would be nice to have better weak scaling (not exhausting memory bandwidth) and to go to more cores.}
+{\color{blue} ELB: Yes.  I don't have such access currently.}
+
 \section{Discussion and outlook} \label{sec:discussion}
@@
 
-\subsection*{Code availablility} \label{sec:code}  For reproducibility we have archived the exact software versions used to produce the results in this paper, including the correct Firedrake version, at FIXME Zenodo?
+\subsection*{Code availability} \label{sec:code}  For reproducibility we have archived the exact software versions used to produce the results in this paper, including the correct Firedrake version, at {\color{blue} ELB: Zenodo?}
 

### 2023-07-18 paper/fascd.tex
Remove some comments
Comments

@@
 \newcommand{\pef}[1]{\todo[inline, color=red!20]{PF: #1}}
+\newcommand{\elb}[1]{\todo[inline, color=blue!20]{EB: #1}}
 
@@
 These methods were shown to converge for multilevel and domain decompositions in a finite element (FE) context; in the multilevel case, the CD method has almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}, \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
-%\pef{I prefer the semicolon here because the second sentence is filling in the details for the first; without it, at the end of the first sentence, I'm thinking ``there's a citation missing here''.}
-% Agreed
-%\pef{Could Tai have applied his algorithm directly to non-optimization problems? I don't think so. If that is true, we're underselling our work by using the word `apply'.}
-% Agreed.  Text compressed; the first item on the list already says this.
 In this paper we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach.
@@
 \item Box constraints, namely upper and lower obstacles, are permitted.
-%\pef{Should we list this? \cite{GraeserKornhuber2009} say ``We emphasize that all algorithms and convergence results to be presented can be generalized to where an upper obstacle is present''}
-% Yes.
-"can be generalized" is true, but in doing so you discover that (see Section 4) the \Pi_i operators in the general box-constrained CD must have a very different formula.
-We actually address this by writing code that works at this level of generality.
-At a certain point, if G&K can't make something clear and demo that it works then they did not do it!
-We will see what we get from referees ...
 \item We implement V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}).
-%\pef{\dots for the first time for VIs beyond linear complementarity problems?}
-% For F-cycles, maybe.
-If something like this has been done before I want a referee to say "here it is ...".
-I'd rather not guess what unpublished, undocumented, and closed extensions people have to their methods.
-So I'm thinking just say what we did.
-See previous point.
 \item We demonstrate that F-cycle FAScd is an optimal solver for the different VI examples in Section \ref{sec:results}, when used in conjunction with reduced-space active set methods as smoothers.
@@
 While FAScd reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed (Appendix \ref{app:reductions}), PFAS reduces to linear multigrid.
-%\pef{I don't like the 2$\times$ apparently.}
-% Me neither.  Fixed.
 The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so as stated it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
@@
 Maybe you can explain it to me on a video call and I will write it in a way that I would have understood it?}
-{\color{blue} ELB: Not being dumb.
+\elb{Not being dumb.
 I am describing an application I am headed toward.
@@
 \end{figure}
-%\pef{It would be good to have a diagram here. I could write one in tikz, if you don't have one handy.}
-% Like it!  I shortened the caption by removing the parenthetical ("(in green; the fine degree of freedom at the same vertex is not shown)"); that's clear because of the coarse edges.
 
@@
 \pef{Why is this fine for F-cycles, but not for V-cycles? I think we should spell it out.}
-{\color{blue} ELB: This is the point of Example 5.1.}
+\elb{This is the point of Example 5.1.}
+\pef{The point of Example 5.1 is to show why it is not fine for V-cycles. We don't explain anywhere it is fine for F-cycles, as far as I can see.}
 
@@
 \end{figure}
-%\pef{Why does the colour bar have grey in it?}
-% Cause Paraview does it.  But I have edited it out of the .png.
 
@@
 \pef{I wonder if there's a Firedrake installation we could use on a supercomputer somewhere. It would be nice to have better weak scaling (not exhausting memory bandwidth) and to go to more cores.}
-{\color{blue} ELB: Yes.  I don't have such access currently.}
+\elb{Yes.  I don't have such access currently.}
 

### 2023-07-19 paper/fascd.tex
eliminate Example 3.4, and tighten adjacent text

@@
 
-The basic CD iterations are meaningful whether or not we are in the optimization case $f=F'$, and when $f$ is nonlinear, non-local, or defined only on $\cK$.
-However, practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.
+The basic CD iterations are meaningful whether or not we are in the optimization case $f=F'$, and even if $f$ is nonlinear, non-local, or defined only on $\cK$.
+However, practical and efficient FE implementation for general operators $f$ seems not to have been addressed in the literature.
 References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem, wherein $f=F'$ is linear, local, and defined on all of $\mathcal{V}$.
-Substantial implementation simplifications are available in such cases.
-
-\begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear, and suppose it is local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi$.
-Considering \pr{cd-add} for concreteness, by linearity the VI can be rewritten
-\begin{equation}
-\ip{f(e_i)}{v_i-\hat w_i} \ge \ip{\tilde\ell}{v_i-\hat w_i}, \label{eq:linearlocalvi}
-\end{equation}
-for all $v_i \in \mathcal{K}_i$, where $\tilde\ell = \ell - f(u)$ incorporates the initial residual $f(u)$.
-Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible---generally $e_i \notin \cK$---but ``$f(e_i)$'' in \eqref{eq:linearlocalvi} is meaningful because $f$ is defined on all of $\cV$.
-\end{example}
-
-\pef{I confess I just don't get the point of this example.
-Why is this a substantial implementation simplification?
-I am sorry I am being dumb.
-Maybe you can explain it to me on a video call and I will write it in a way that I would have understood it?}
-\elb{Not being dumb.
-I am describing an application I am headed toward.
-FAScd applies to both differential \emph{and integral} operators.
-(We don't try it out in the latter case, but it is a feature that it does not care about the locality of your operator, unlike monotone multigrid.)  It applies to operators where $f(e_i)$ makes no sense because $e_i\notin \mathcal{K}$ (SIA is an example IMO).
-And it does FAS coarse corrections because we don't assume linearity of $f$.}
-
-When $f$ has the locality of Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over basis function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}  However, for general problems lacking the convenience of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of CD-type iterations.
-Section \ref{sec:vcycle} constructs such an iteration for a general operator $f:\cK\to\cV'$ by applying the full approximation scheme (FAS) idea of Brandt \cite{Brandt1977}.
-
-From now on we restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.
+In Section \ref{sec:vcycle} we will extend CD-type iterations to nonlinear operators by applying the full approximation scheme (FAS) idea of Brandt \cite{Brandt1977}, and then illustrate its application to non-optimization operators in Examples \ref{ex:results:advdiff} and \ref{ex:results:sia}.
+In the latter Example $f$ is defined only over a subset $\cK$.
+
+From now on we restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bilateral obstacle problems}.
 Suppose $\Omega \subset \RR^d$ is open and let $\mathcal{V}=W^{1,p}(\Omega)$.
@@
 Note that we are treating inequality constraints and strongly-imposed Dirichlet boundary conditions in a unified manner.
-Any Neumann conditions over $\partial_N \Omega$ are imposed weakly in the definition of the operator $f$, i.e.~by an integral over $\partial_N\Omega$.
-
-In the next Sections we numerically solve continuum VI problem \eqref{eq:vi} over a constraint set of the form \eqref{eq:originalconstraintset}.
-In fact, given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, from now on we assume that the following is well-posed for $u^\star\in \cK$:
+(Any Neumann conditions over $\partial_N \Omega$ are imposed weakly by the operator $f$, via an integral over $\partial_N\Omega$.)  Next we will numerically solve continuum VI problem \eqref{eq:vi} over a constraint set \eqref{eq:originalconstraintset}.
+In fact, given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we will assume that the following is well-posed for $u^\star\in \cK$:
 \begin{equation}

### 2023-07-19 paper/fascd.tex
Write up SIA results

@@
 
-For a weak-scaling study we fixed the number of degrees of freedom per process at $m_J/P=1281^2 \approx 1.6 \times 10^6$ and solved over $P=1,4,16$ processes.\footnote{Runs were on a single-node workstation with two Intel Xeon 4210 cpus (2.2GHz), a total of 20 physical cores, and 128 GB shared memory.}  The results in Table \ref{tab:results:siaweak} show nearly-constant run time for this parallel refinement path.
-The finest mesh result, with $m_J=5121^2=2.6 \times 10^7$ degrees of freedom, needed six minutes to solve for the steady surface elevation of an ice sheet larger than the current Greenland ice sheet ($1.7$ million $\text{km}^2$) at uniform $350$ m resolution.
+For a weak-scaling study we fixed the number of degrees of freedom per process at $m_J/P=641^2 \approx 4.1 \times 10^5$ and solved with $6, 7, 8, 9, $ and 10 refinements.\footnote{The experiments were conducted on ARCHER2, the UK national supercomputer.
+A maximum of 32 cores per node were employed to avoid saturating the memory bandwidth.}  The results in Table \ref{tab:results:siaweak} show some degradation with increasing core counts, perhaps due to the use of one-level additive Schwarz in the solution of the arising linear systems.
+Nevertheless, the finest mesh result, with $m_J=10241^2=1.0 \times 10^8$ degrees of freedom, required less than three minutes to solve for the steady surface elevation of an ice sheet larger than the current Greenland ice sheet ($1.7$ million $\text{km}^2$) at uniform $175$ m resolution.
 \end{example}
@@
 \begin{tabular}{c@{\hskip 4mm}c@{\hskip 4mm}c@{\hskip 7mm}c@{\hskip 4mm}c}
+\toprule
 $P$ & \emph{levels} & $m_J$ & \emph{F-cycles} & \emph{time} (s) \\ \cmidrule{1-5}
- 1 & 7 & $1281^2$ & 3 & 309 \\
- 4 & 8 & $2561^2$ & 3 & 368 \\
-16 & 9 & $5121^2$ & 3 & 374
+\midrule
+ 1 & 6 & $641^2$ & 3 & 105 \\
+ 4 & 7 & $1281^2$ & 3 & 106 \\
+ 16 & 8 & $2561^2$ & 3 & 100 \\
+ 64 & 9 & $5121^2$ & 3 & 159 \\
+ 256 & 10 & $10241^2$ & 2 & 148 \\
+ \bottomrule
 \end{tabular}
@@
 
-\pef{I wonder if there's a Firedrake installation we could use on a supercomputer somewhere. It would be nice to have better weak scaling (not exhausting memory bandwidth) and to go to more cores.}
-\elb{Yes.  I don't have such access currently.}
-
 \section{Discussion and outlook} \label{sec:discussion}
@@
 PEF was supported by the Engineering and Physical Sciences Research Council [EPSRC grants EP/R029423/1 and EP/W026163/1].
+This work used the ARCHER2 UK National Supercomputing Service (https://www.archer2.ac.uk).
 

### 2023-07-20 paper/fascd.tex
Some edits to section 4

@@
 
-In the material which follows, use of higher-order Lagrange elements, i.e.~$P_k$ or $Q_k$ with $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} special to inequality constraints, which we avoid.
-Namely, our FE construction of constraint sets (Section \ref{sec:cdmultilevel}) will require \emph{nodal monotonicity},
+As is common in the discretization of variational inequalities, we employ lowest-order continuous Lagrange elements $P_1$ or $Q_1$ on simplices and tensor-product cells respectively.
+This choice is convenient because this choice enjoys \emph{nodal monotonicity}:
 \begin{equation}
@@
 for all $w,z \in \mathcal{V}^j$, where $\bw,\bz \in \RR^{m_j}$ denote vectors of nodal values.
-However, for $k\ge 2$ there are basis functions in $P_k$ and $Q_k$, i.e.~functions with nodal values one or zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.
-Evidently \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of such indefinite-sign basis functions.
-\pef{I'm not sure this constitutes a variational crime?
-We enforce the RHS because it's tractable, and this is only equivalent to what we actually want (the LHS) for P1 and Q1.
-I don't see where we're departing from the variational framework, however.}
+In other words, to enforce an inequality constraint, with $P_1$ or $Q_1$ we only need enforce it pointwise on the degrees of freedom.
+For $k\ge 2$ there are functions in the standard choices of bases for $P_k$ and $Q_k$ which take on negative values \cite[Figure 1.7]{Elmanetal2014}, violating \eqref{eq:nodalmonotonicity}.
 
 To allow box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the vertices of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
-Suppose that the finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ satisfy
+The finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ are constructed by suitable interpolation of $\underline{\gamma}$, $\overline{\gamma}$ (e.g.~by applying the Cl\'ement interpolation operator) and satisfy
 \begin{equation}
@@
 \end{equation}
-\pef{Shouldn't there be some relation between the continuous bounds and the discrete ones?}
-In analogy to the continuous data, also assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those vertices of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.  Define
+Analogously, we construct the discrete boundary data $g_D^J$ so that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those vertices of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.
+Define
 \begin{equation}
@@
 
-While \eqref{eq:fe:vi} approximates \eqref{eq:boxdirichletvi}, we warn the reader that it may not technically be conforming for that VI.
-That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require or generally expect $\mathcal{K}^J \subset \mathcal{K}$.
+While \eqref{eq:fe:vi} approximates \eqref{eq:boxdirichletvi}, we warn the reader that it is usually not conforming for that VI.
+That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require or expect $\mathcal{K}^J \subset \mathcal{K}$.
 Regardless of the smoothness of the continuum obstacles $\underline{\gamma}, \overline{\gamma}$, admissibility with respect to their FE interpolants $\underline{\gamma}^J, \overline{\gamma}^J$ will generally not imply continuum admissibility.\footnote{This caveat is especially relevant in geophysical situations where only nodal values of the obstacle are known.
@@
 
-The monotone injections $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to nodal functions $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the \emph{stars} of the coarse vertices).
+The monotone injections $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to nodal functions $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the \emph{stars} of the coarse vertices), Figure \ref{fig:Rplusminus}.
 These operators are constructed to enforce the property
@@
 To make the construction precise, let $x_q^j$ denote the vertices of mesh $\mathcal{T}^j$, with corresponding basis functions $\psi_q^j(x)$.
-For a $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$ (Figure \ref{fig:Rplusminus}):
+For a $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$:
 \begin{align}

### 2023-07-20 paper/tikz/Rplusminus.tex
Some edits to section 4

@@
   \filldraw[seabornblue] (expl) circle (1.5pt);
-  \node at (2.25, 0.525) {$= \text{max/min of }$};  % I'm sure there's a better way to place this, but tikz mystifies me
-  \filldraw[seaborngreen] (2.95, 0.5) circle (1.5pt);
+  \node at (2.9, 0.5125) {$= \text{max/min over patch of cells}$};  % I'm sure there's a better way to place this, but tikz mystifies me
+  \node at (2.25, 0.2925) {$= \text{max/min of }$};  % I'm sure there's a better way to place this, but tikz mystifies me
+  \filldraw[seaborngreen] (2.95, 0.29) circle (1.5pt);
 \end{tikzpicture}

### 2023-07-20 paper/fascd.tex
Rephrase F-cycle bound construction

@@
 Suppose we propose
-    $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \iR \underline{\gamma}^1, \iR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
+\begin{equation}
+    \label{eq:fe:standardinjection}
+    \underline{\gamma}^0, \overline{\gamma}^0 \equiv \iR \underline{\gamma}^1, \iR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}
+\end{equation}
 Then $\underline{\gamma}^0=0$ and $\overline{\gamma}^0=C$ identically.
@@
 On the other hand, suppose we propose
-    $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
+\begin{equation}
+    \label{eq:fe:monotoneinjection}
+    \underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}
+\end{equation}
 The general motivation for this choice---compare \eqref{eq:fe:chilevels} below---is that $w^0 \ge \underline{\gamma}^0 \, \implies \, Pw^0 \ge \underline{\gamma}^1$ and $w^0 \le \overline{\gamma}^0 \, \implies \, Pw^0 \le \overline{\gamma}^1$.
@@
 \end{figure}
-%\pef{Can we put another figure beside it illustrating the choice $\underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1$?}
-%\pef{Maybe use different colours for upper and lower bounds?}
 
@@
 This F-cycle approach is described in Algorithm \ref{alg:fascd-fcycle}.
-Injection is used to generate the $\underline{\gamma}^j,\overline{\gamma}^j$ on each level.\footnote{Contrast the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, based on an existing admissible iterate.}
-\pef{Why is this fine for F-cycles, but not for V-cycles? I think we should spell it out.}
-\elb{This is the point of Example 5.1.}
-\pef{The point of Example 5.1 is to show why it is not fine for V-cycles. We don't explain anywhere it is fine for F-cycles, as far as I can see.}
-
-\noindent
-Regarding line \ref{line:fcyclecoarsestinitial}, one may generate an initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation of a finest-level iterate, if one is available.
+In contrast to the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, standard injection is used to generate the $\underline{\gamma}^j,\overline{\gamma}^j$ on each level during the initial nested iteration.
+This is because nested iteration approximates the full solution on the coarse grids, not just a correction to an initial iterate; essentially, our choice is between \eqref{eq:fe:standardinjection} (where the coarse constraint set is always nonempty, but the solution may not be admissible on the next level) or \eqref{eq:fe:monotoneinjection} (where the coarse solution is always admissible if it exists, but the constraint set may be empty).
+We choose \eqref{eq:fe:standardinjection} so that the coarse problems always have nonempty constraint sets, and project pointwise to the bounds as we refine in the nested iteration.
+On line \ref{line:fcyclecoarsestinitial}, one may generate an initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation of a finest-level iterate, if one is available.
 
@@
 This work used the ARCHER2 UK National Supercomputing Service (https://www.archer2.ac.uk).
+PEF thanks Jack D.~Betteridge for assistance with the numerical experiments of Section \ref{ex:results:sia}.
 

### 2023-07-20 paper/fascd.tex
FAScd -> FASCD

@@
 \begin{abstract}
-We present the \emph{full approximation scheme constraint decomposition} (FAScd) multilevel method for solving variational inequalities (VIs).
-FAScd is a common extension both of the full approximation scheme (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and the constraint decomposition (CD) method introduced by X.-C.~Tai for VIs arising in optimization.
+We present the \emph{full approximation scheme constraint decomposition} (FASCD) multilevel method for solving variational inequalities (VIs).
+FASCD is a common extension both of the full approximation scheme (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and the constraint decomposition (CD) method introduced by X.-C.~Tai for VIs arising in optimization.
 We extend the CD idea by exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
-We demonstrate that FAScd V-cycles exhibit excellent mesh-independent convergence rates and that FAScd F-cycles (full multigrid) are optimal solvers, when used in combination with reduced-space active set smoothers with work proportional to the number of unknowns on a given mesh level.
+We demonstrate that FASCD V-cycles exhibit excellent mesh-independent convergence rates and that FASCD F-cycles (full multigrid) are optimal solvers, when used in combination with reduced-space active set smoothers with work proportional to the number of unknowns on a given mesh level.
 The example problems include unilateral and bilateral VI problems, with differential operators which are symmetric linear, nonsymmetric linear, and nonlinear.
@@
 In this paper we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach.
-Specifically, our \emph{full approximation scheme constraint decomposition} (FAScd) method extends the algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} in the following five directions:
+Specifically, our \emph{full approximation scheme constraint decomposition} (FASCD) method extends the algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} in the following five directions:
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
@@
 \item We implement V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}).
-\item We demonstrate that F-cycle FAScd is an optimal solver for the different VI examples in Section \ref{sec:results}, when used in conjunction with reduced-space active set methods as smoothers.
+\item We demonstrate that F-cycle FASCD is an optimal solver for the different VI examples in Section \ref{sec:results}, when used in conjunction with reduced-space active set methods as smoothers.
 The number of Newton and Krylov iterations per step are fixed, so that the work of the smoother is proportional to the number of unknowns on a given level.
@@
 
-Prior multilevel algorithms for VIs have various substantial limitations compared to the FAScd framework.
+Prior multilevel algorithms for VIs have various substantial limitations compared to the FASCD framework.
 The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} is apparently limited to linear complementarity problems and projected (truncated) Gauss--Seidel smoothers.
-While FAScd reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed (Appendix \ref{app:reductions}), PFAS reduces to linear multigrid.
+While FASCD reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed (Appendix \ref{app:reductions}), PFAS reduces to linear multigrid.
 The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so as stated it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
@@
 Furthermore, monotone multigrid methods are more difficult to implement in FE libraries because modifications of the FE basis functions are needed to approximate the coarser-level coincidence sets.
-No such invasive basis modifications are used either in Tai's original multilevel CD or in FAScd.
-Our implementation of FAScd in the Firedrake library \cite{Rathgeberetal2016} requires no modifications of basis functions and yet exhibits excellent performance, including on a problem used to demonstrate the benefits of monotone multigrid (Example \ref{ex:results:classical}; compare \cite[problem 7.1.1]{GraeserKornhuber2009}).
-
-In our experiments, FAScd V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for the corresponding (unconstrained) PDE problems.
+No such invasive basis modifications are used either in Tai's original multilevel CD or in FASCD.
+Our implementation of FASCD in the Firedrake library \cite{Rathgeberetal2016} requires no modifications of basis functions and yet exhibits excellent performance, including on a problem used to demonstrate the benefits of monotone multigrid (Example \ref{ex:results:classical}; compare \cite[problem 7.1.1]{GraeserKornhuber2009}).
+
+In our experiments, FASCD V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for the corresponding (unconstrained) PDE problems.
 The F-cycle results in Section \ref{sec:results} show nearly-textbook multigrid efficiency, namely solutions within discretization error in work comparable to few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}.
@@
 
-All iterates in FAScd are admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
+All iterates in FASCD are admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
 For example, we solve an ice sheet geometry problem (Examples \ref{ex:sia} and \ref{ex:results:sia}) in which the ice flow formulas are only meaningful for surface elevation iterates that do not penetrate the bedrock.
 (More generally, fluid layers subject to a non-negative thickness constraint typically require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation \cite{Bueler2021conservation}, but this is not needed in the ice sheet problem here.)  Non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural modifications of the operator formula.
-FAScd solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a mesh-independent number of F-cycle iterations (Example \ref{ex:results:sia}).
+FASCD solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a mesh-independent number of F-cycle iterations (Example \ref{ex:results:sia}).
 
@@
 Here we propose an apparently new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD and an associated iteration.
-Section \ref{sec:vcycle} then states the FAScd V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, F-cycles, and smoother construction.
+Section \ref{sec:vcycle} then states the FASCD V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, F-cycles, and smoother construction.
 Section \ref{sec:results} gives numerical results, from a parallel Firedrake \cite{Rathgeberetal2016} implementation (see Code Availability), for each example problem introduced in Section \ref{sec:vi}.
@@
 
-Numerical results are shown in Section \ref{sec:results} for applications of FAScd Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} to these Examples.
+Numerical results are shown in Section \ref{sec:results} for applications of FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} to these Examples.
 
@@
 Associated to any CD are certain iterative methods \cite{Tai2003,Xu1992}.
-Let $0\le \alpha \le 1$ be the \emph{damping parameter}.\footnote{FAScd Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} are presented without damping, and indeed it seems to be unnecessary.
+Let $0\le \alpha \le 1$ be the \emph{damping parameter}.\footnote{FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} are presented without damping, and indeed it seems to be unnecessary.
 However, the convergence of the general CD iterations appears to depend upon it.}  The following algorithms, which replace $u \in \cK$ with a new iterate $w\in\cK$ which should be closer to $u^\star \in \cK$ solving \eqref{eq:vi}, solve smaller VI problems over each subset $\cK_i$.
@@
 
-\section{FAScd V-cycle} \label{sec:vcycle}
+\section{FASCD V-cycle} \label{sec:vcycle}
 
@@
 \end{center}
-\caption{The FAScd V-cycle computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.}
+\caption{The FASCD V-cycle computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.}
 \label{fig:fascdvcycle}
@@
 \end{pseudo}
-\caption{The FAScd V-cycle is an iteration for solving VI problem \eqref{eq:fe:vi}.  Here $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
+\caption{The FASCD V-cycle is an iteration for solving VI problem \eqref{eq:fe:vi}.  Here $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
 \label{alg:fascd}
@@
 
-Implementations of FAScd methods must choose a concrete convergence criterion.
+Implementations of FASCD methods must choose a concrete convergence criterion.
 In this section we also construct the F-cycle extension, and we describe the active set Newton method we will use as a smoother and coarse solver.
@@
 \end{pseudo}
-\caption{The FAScd F-cycle for solving VI problem \eqref{eq:fe:vi}.}
+\caption{The FASCD F-cycle for solving VI problem \eqref{eq:fe:vi}.}
 \label{alg:fascd-fcycle}
@@
 Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python, for $P_1$ and $Q_1$ elements, using the Firedrake library \cite{Rathgeberetal2016}.
-Here we show numerical results for the example problems presented in Section \ref{sec:vi}, with a focus on the scaling of FAScd iterations as the mesh resolution increases.
+Here we show numerical results for the example problems presented in Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
 Our primary observations are that each V-cycle greatly reduces the residual norm, and that very few F-cycles solve most examples to within discretization error.
@@
 For F-cycles we report the number of V-cycles after the ramp (Section \ref{sec:implementation}); a constant number solves the problem.
-Comparing the ball and spiral results, the complexity of the coincidence set or free boundary does not appear to strongly influence FAScd performance.
+Comparing the ball and spiral results, the complexity of the coincidence set or free boundary does not appear to strongly influence FASCD performance.
 For the spiral results, note that \cite{GraeserKornhuber2009} reports more than 30 iterations of all their tested methods were required to achieve similar tolerance for $m_J \approx 5\times 10^5$.
@@
 \bigskip
-\caption{Iterations for FAScd V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}.  Smoother-only runs marked \XX\xspace were not attempted.}
+\caption{Iterations for FASCD V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}.  Smoother-only runs marked \XX\xspace were not attempted.}
 \label{tab:results:classical}
@@
 \end{center}
-\caption{Visualization of one FAScd V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of $m_0=5$ vertices.
+\caption{Visualization of one FASCD V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of $m_0=5$ vertices.
 Solutions (i.e.~$w^j$, $y^j$, $z^j$) are lines, obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots, and the zero function is dotted.
@@
 Solvers implemented in Firedrake can exploit parallel solvers from the PETSc library \cite{Balayetal2023}, and in fact the above Examples, which were run in serial so as to fix the smoother, also run efficiently in parallel.
-The next Examples apply parallel solvers, and Example \ref{ex:results:sia} demonstrates the parallel weak scaling \cite{Bueler2021} of FAScd.
+The next Examples apply parallel solvers, and Example \ref{ex:results:sia} demonstrates the parallel weak scaling \cite{Bueler2021} of FASCD.
 
@@
 Despite the hint of growing iterations under mesh refinement, a few iterations sufficed up to memory-limited resolutions.
-For any of the higher resolution runs, the FAScd F-cycle solver took less time than generating the $P_1$ mesh hierarchy.
+For any of the higher resolution runs, the FASCD F-cycle solver took less time than generating the $P_1$ mesh hierarchy.
 The local Péclet number $\text{Pe}=\|\bX(x,y,z)\|_2 L \eps^{-1}$, with $L=2$ the width of $\Omega$, varied from 40 to 260; the mesh Péclet number $m_J^{-1/d}\,\text{Pe}$ varied accordingly.
@@
 
-We applied FAScd F-cycles using $Q_1$ elements and a smoother of four (fixed) iterations of RS Newton (Section \ref{sec:implementation}), with quadratic backtracking line search.
+We applied FASCD F-cycles using $Q_1$ elements and a smoother of four (fixed) iterations of RS Newton (Section \ref{sec:implementation}), with quadratic backtracking line search.
 The arising linear systems were approximately solved, in parallel, by three (fixed) iterations of preconditioned GMRES and a ASM+ILU preconditioner.
@@
 \bigskip
-\caption{Iterations and run time (wall clock, in seconds) for parallel FAScd F-cycle solutions of Example \ref{ex:sia} on $P$ processes.}
+\caption{Iterations and run time (wall clock, in seconds) for parallel FASCD F-cycle solutions of Example \ref{ex:sia} on $P$ processes.}
 \label{tab:results:siaweak}
@@
 
-As a multilevel solver for VI problems, one might describe the new FAScd method as merely a strategy for generating coarser-level problems which, once solved, make helpful admissible additions to a final V-cycle update.
+As a multilevel solver for VI problems, one might describe the new FASCD method as merely a strategy for generating coarser-level problems which, once solved, make helpful admissible additions to a final V-cycle update.
 The method is smoother-agnostic, and our implementation within the extensible Firedrake \cite{Rathgeberetal2016} and PETSc \cite{Balayetal2023} library framework should allow easy experimentation with other smoothers (projected Gauss--Seidel, semi-smooth Newton, interior-point, etc.), beyond the active-set Newton method \cite{BensonMunson2006} chosen for demonstration here.
@@
 
-In contrast with monotone multigrid \cite{Kornhuber1994}, FAScd does not modify the FE discretization, but the possibility of acceleration via basis-level manipulations is a topic for future research.
+In contrast with monotone multigrid \cite{Kornhuber1994}, FASCD does not modify the FE discretization, but the possibility of acceleration via basis-level manipulations is a topic for future research.
 On the other hand,  the convergence proofs in \cite{GraeserKornhuber2009,Kornhuber1994}, which appear to depend on particular smoother and basis-level manipulation choices, have no analog here.
@@
 
-When all inequality constraints are removed the FAScd V-cycle Algorithm \ref{alg:fascd} reduces to the classical FAS multigrid V-cycle for nonlinear PDEs \cite{Brandt1977,Trottenbergetal2001}, which itself generalizes linear multigrid.
+When all inequality constraints are removed the FASCD V-cycle Algorithm \ref{alg:fascd} reduces to the classical FAS multigrid V-cycle for nonlinear PDEs \cite{Brandt1977,Trottenbergetal2001}, which itself generalizes linear multigrid.
 In fact, suppose we replace $\underline{\gamma}^J,\overline{\gamma}^J$ in definition \eqref{eq:fe:fineconstraintset} with $-\infty,+\infty$, respectively.

### 2023-07-20 paper/fascd.tex
Add some todo notes from discussion yesterday
Use rules in tables consistently

@@
 The number of Newton and Krylov iterations per step are fixed, so that the work of the smoother is proportional to the number of unknowns on a given level.
+\pef{The solver works for VIs where the residual is only defined on the admissible set \dots}
 \end{enumerate}
@@
 On the other hand, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These ordering properties are used in the CD construction of the next two Sections.
+\pef{Promote this to its own numbered equation and highlight its importance later on in the construction of LDCs}
 
@@
 \begin{tabular}{lr@{\hskip 7mm}cccccc}
+\toprule
 \multirow{2}{*}{\emph{levels}} & \multirow{2}{*}{$m_J$} & \multicolumn{3}{c}{\,\emph{ball}} & \multicolumn{3}{c}{\,\emph{spiral}} \\ \cmidrule(lr){3-5} \cmidrule(lr){6-8}
-   &          & V-cycle & F-cycle & RS only & V-cycle & F-cycle & RS only \\ \cmidrule{1-8}
+   &          & V-cycle & F-cycle & RS only & V-cycle & F-cycle & RS only \\
+\midrule
  1 &    $5^2$ &  1 &  1 &  1  &  1 &  1 &  1 \\
@@
  9 & $1025^2$ &  5 &  2 & \XX &  5 &  3 & \XX \\
-10 & $2049^2$ &  5 &  2 & \XX &  5 &  3 & \XX
+10 & $2049^2$ &  5 &  2 & \XX &  5 &  3 & \XX \\
+\bottomrule
 \end{tabular}
@@
 \begin{tabular}{lr@{\hskip 7mm}c@{\hskip 3mm}c@{\hskip 3mm}c@{\hskip 4mm}c@{\hskip 5mm}c@{\hskip 6mm}c}
-\emph{levels} & $m_J$ & V(1,0) & V(0,1) & V(1,1) & F & F$_\eps$ & $\|$error$\|_\infty$ \\ \cmidrule{1-8}
+\toprule
+\emph{levels} & $m_J$ & V(1,0) & V(0,1) & V(1,1) & F & F$_\eps$ & $\|$error$\|_\infty$ \\
+\midrule
  1 &    $7$ &  1 &  1 &  1 &  1 &  1 & $2.3\times 10^{-1}$ \\
@@
  9 & $1537$ & NC &  5 &  6 & NC &  1 & $3.6 \times 10^{-6}$ \\
-10 & $3073$ & NC & 11 & 19 &  1 &  1 & $4.9 \times 10^{-7}$
+10 & $3073$ & NC & 11 & 19 &  1 &  1 & $4.9 \times 10^{-7}$ \\
+\bottomrule
 \end{tabular}
@@
 \begin{tabular}{lr@{\hskip 7mm}c@{\hskip 4mm}c@{\hskip 4mm}c@{\hskip 6mm}c}
-\emph{levels} & $m_J$ & V & V$_4$ & F & $\|$error$\|_\infty$ \\ \cmidrule{1-6}
+\toprule
+\emph{levels} & $m_J$ & V & V$_4$ & F & $\|$error$\|_\infty$ \\
+\midrule
  1 &    $7$ &  1 &  1 &  1 & $8.7 \times 10^{-2}$ \\
@@
  9 & $1537$ & NC &  2 &  1 & $4.0 \times 10^{-5}$ \\
-10 & $3073$ & NC &  2 &  1 & $1.5 \times 10^{-5}$
+10 & $3073$ & NC &  2 &  1 & $1.5 \times 10^{-5}$ \\
+\bottomrule
 \end{tabular}
@@
 \begin{tabular}{lr@{\hskip 7mm}c@{\hskip 4mm}c}
-\emph{levels} & $m_J$ & $d=2$ & $d=3$ \\ \cmidrule{1-4}
+\toprule
+\emph{levels} & $m_J$ & $d=2$ & $d=3$ \\
+\midrule
  1 &    $16^d$ & 1 & 1 \\
@@
  8 &  $1921^d$ & 3 \\
- 9 &  $3841^d$ & 3
+ 9 &  $3841^d$ & 3 \\
+\bottomrule
 \end{tabular}
@@
 \toprule
-$P$ & \emph{levels} & $m_J$ & \emph{F-cycles} & \emph{time} (s) \\ \cmidrule{1-5}
+$P$ & \emph{levels} & $m_J$ & \emph{F-cycles} & \emph{time} (s) \\
 \midrule

### 2023-07-20 paper/fascd.tex
Add level 11 results

@@
 
-For a weak-scaling study we fixed the number of degrees of freedom per process at $m_J/P=641^2 \approx 4.1 \times 10^5$ and solved with $6, 7, 8, 9, $ and 10 refinements.\footnote{The experiments were conducted on ARCHER2, the UK national supercomputer.
-A maximum of 32 cores per node were employed to avoid saturating the memory bandwidth.}  The results in Table \ref{tab:results:siaweak} show some degradation with increasing core counts, perhaps due to the use of one-level additive Schwarz in the solution of the arising linear systems.
-Nevertheless, the finest mesh result, with $m_J=10241^2=1.0 \times 10^8$ degrees of freedom, required less than three minutes to solve for the steady surface elevation of an ice sheet larger than the current Greenland ice sheet ($1.7$ million $\text{km}^2$) at uniform $175$ m resolution.
+For a weak-scaling study we fixed the number of degrees of freedom per process at $m_J/P=641^2 \approx 4.1 \times 10^5$ and solved with $6, \dots, 11$ refinements.\footnote{The experiments were conducted on ARCHER2, the UK national supercomputer.
+Each node has 16 memory channels, so a maximum of 16 cores per node were employed.}  The results in Table \ref{tab:results:siaweak} show some degradation in time per cycle with increasing core counts, perhaps due to the use of one-level additive Schwarz in the solution of the arising linear systems.
+Nevertheless, the finest mesh result, with $m_J=20481^2=4.1 \times 10^8$ degrees of freedom, required less than three minutes to solve for the steady surface elevation of an ice sheet larger than the current Greenland ice sheet ($1.7$ million $\text{km}^2$) at uniform $87.5$ m resolution.
 \end{example}
@@
 \midrule
- 1 & 6 & $641^2$ & 3 & 105 \\
- 4 & 7 & $1281^2$ & 3 & 106 \\
- 16 & 8 & $2561^2$ & 3 & 100 \\
- 64 & 9 & $5121^2$ & 3 & 159 \\
- 256 & 10 & $10241^2$ & 2 & 148 \\
+ 1 & 6 & $641^2$ & 3 & 98 \\
+ 4 & 7 & $1281^2$ & 3 & 98 \\
+ 16 & 8 & $2561^2$ & 3 & 124 \\
+ 64 & 9 & $5121^2$ & 3 & 136 \\
+ 256 & 10 & $10241^2$ & 2 & 95 \\
+ 1024 & 11 & $20481^2$ & 2 & 177 \\
  \bottomrule

### 2023-07-21 paper/results/siaweak.sh
updating SIA script to represent what actually happened for example 8.4

@@ -1,6 +1,12 @@
 # generate weak-scaling results for SIA Example 8.4 in paper
+# uncomment later lines if you are on a big machine
+# siaweak.txt is result from this script; see archer2/ for large machine logs
 
-OPTS="-bumps -mcoarse 20 -fascd_cycle_type full -fascd_monitor -fascd_rtol 2.0e-4 -fascd_atol 1.0e-8 -quad"
-time mpiexec -n  1 --map-by core --bind-to hwthread python3 sia.py $OPTS -levs 7
-time mpiexec -n  4 --map-by core --bind-to hwthread python3 sia.py $OPTS -levs 8
-time mpiexec -n 16 --map-by core --bind-to hwthread python3 sia.py $OPTS -levs 9
+MPI="mpiexec -map-by numa -bind-to core"
+OPTS="-bumps -mcoarse 20 -fascd_cycle_type full -fascd_monitor -fascd_rtol 1.0e-4 -fascd_atol 1.0e-8 -quad"
+time $MPI -n  1 python3 sia.py $OPTS -levs 6
+time $MPI -n  4 python3 sia.py $OPTS -levs 7
+time $MPI -n 16 python3 sia.py $OPTS -levs 8
+#time $MPI -n 64 python3 sia.py $OPTS -levs 9
+#time $MPI -n 256 python3 sia.py $OPTS -levs 10
+#time $MPI -n 1024 python3 sia.py $OPTS -levs 11

### 2023-07-21 paper/fascd.tex
update, and minor clean, of adv-diff example 8.3
no change in parallel iterations after halo exchange
FASCD bug fix

@@
 \begin{example}  \label{ex:results:advdiff}
-We consider the advection-diffusion VI problem in Example \ref{ex:advectiondiffusion}.
-Let $\Omega=(-1,1)^d$ for $d=2,3$, the diffusivity $\eps=0.1$ (equation \eqref{eq:advectiondiffusionstrong}), and we impose zero Dirichlet boundary conditions.
-The 2D problem satisfied the coercivity hypothesis with a divergence-free velocity field $\bX = (7+5y,-5x)$, but in 3D, for variety's sake, we set $\bX = (7+5y,-5x,2z)$ so $\grad\cdot\bX \ne 0$.
-The discontinuous source term $\phi \in L^\infty(\Omega)$ satisfied $\phi\ge 0$ in half of the domain, with positive circular patches, while values were negative in the other half; details are in the associated code.
-While the other Examples in this Section are unilateral, here the solution was box-constrained: $0 \le u \le 1$.
+Consider the advection-diffusion VI problem in Example \ref{ex:advectiondiffusion} with $\Omega=(-1,1)^d$, $d=2,3$, diffusivity $\eps=0.1$ in equation \eqref{eq:advectiondiffusionstrong}, and zero Dirichlet boundary conditions.
+We made the 2D problem satisfy the coercivity hypothesis with a divergence-free velocity field $\bX = (7+5y,-5x)$, but in 3D, for variety's sake, we set $\bX = (7+5y,-5x,2z)$ so $\grad\cdot\bX \ne 0$.
+The chosen discontinuous source term $\phi \in L^\infty(\Omega)$ satisfies $\phi\ge 0$ in half of the domain, with positive circular patches, with negative values in the other half; details are in the associated code.
+While the other Examples in this Section are unilateral, here the solution is box-constrained: $0 \le u \le 1$.
 Figure \ref{fig:results:advdiff} shows coincidence sets for the 2D problem; the 3D sets are comparable in regularity and general pattern.
@@
 
-To minimally-resolve the coincidence set we used a coarse mesh with $m_0=16^d$ vertices.
-The 2D solver applied $P_1$ elements and a single processor, but in 3D we used $Q_1$ elements (hexahedra) and $P=16$ processes.
-The smoother was two (fixed) RS Newton iterations,  with three preconditioned GMRES iterations to approximately solve the arising linear systems.
-The additive Schwarz method (ASM) preconditioner applied incomplete factorization (ILU) preconditioning on each core.
+The 2D solver applied $P_1$ elements and a single processor, but in 3D the elements were chosen to be $Q_1$ (hexahedra) and the run parallel ($P=16$ processes).
+A coarse mesh with $m_0=16^d$ vertices minimally-resolved the coincidence set.
+The smoother was two (fixed) RS Newton iterations, with three preconditioned GMRES iterations to approximately solve the arising linear systems; the preconditioning was additive Schwarz method (ASM) with incomplete factorization (ILU) on each process.
 The coarse solver was RS Newton and LU, iterated to convergence.
@@
 Despite the hint of growing iterations under mesh refinement, a few iterations sufficed up to memory-limited resolutions.
-For any of the higher resolution runs, the FASCD F-cycle solver took less time than generating the $P_1$ mesh hierarchy.
-The local Péclet number $\text{Pe}=\|\bX(x,y,z)\|_2 L \eps^{-1}$, with $L=2$ the width of $\Omega$, varied from 40 to 260; the mesh Péclet number $m_J^{-1/d}\,\text{Pe}$ varied accordingly.
-Thus the $m_J\le 121^d$ problems were advection-dominated while those with $m_J\ge 481^d$ were diffusion-dominated, but the smoother was the same in all cases, and the iterations were nearly constant.
+For any of the higher resolution runs, the FASCD F-cycle solver took less time than generating the mesh hierarchy.
+The local Péclet number $\text{Pe}=\|\bX(x,y,z)\|_2 L \eps^{-1}$, with $L=2$ the width of $\Omega$, varied from 40 to 260.
+The mesh Péclet number $m_J^{-1/d}\,\text{Pe}$ varied accordingly, and though the $m_J\le 121^d$ problems were advection-dominated, those with $m_J\ge 481^d$ were diffusion-dominated, and the smoother was the same in all cases, with nearly-constant iterations.
 \end{example}

### 2023-07-21 paper/results/pollutant.sh

@@ -15,7 +15,7 @@ echo ""
 # which mpiexec
 echo "******** -dim 3: P = 16 parallel, Q1 *******"
 for LEVS in $LIST3D; do
-    mpiexec -n 16 --map-by core --bind-to hwthread python3 pollutant.py -dim 3 -q1 -levs $LEVS
+    mpiexec -n 16 -map-by numa -bind-to core python3 pollutant.py -dim 3 -q1 -levs $LEVS
 done
 echo ""
 

### 2023-07-22 paper/fascd.tex
clean abstract and introduction; address a PEF

@@
 We extend the CD idea by exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
-We demonstrate that FASCD V-cycles exhibit excellent mesh-independent convergence rates and that FASCD F-cycles (full multigrid) are optimal solvers, when used in combination with reduced-space active set smoothers with work proportional to the number of unknowns on a given mesh level.
-The example problems include unilateral and bilateral VI problems, with differential operators which are symmetric linear, nonsymmetric linear, and nonlinear.
+We demonstrate that FASCD V-cycles exhibit excellent mesh-independent convergence rates.
+Furthermore, when a reduced-space active set method is applied as a smoother, with work proportional to the number of unknowns on a given mesh level, FASCD F-cycles (full multigrid) are optimal solvers.
+The example problems include differential operators which are symmetric linear, nonsymmetric linear, and nonlinear, in unilateral and bilateral VI problems.
 Our examples demonstrate very good convergence even on problems where other multilevel VI solvers struggle.
@@
 The constraint decomposition (CD) methods of Tai \cite{Tai2003} are designed to solve coercive variational inequality (VI) problems arising as the Karush--Kuhn--Tucker optimality conditions for the minimization of a convex functional over a convex set.
-These methods were shown to converge for multilevel and domain decompositions in a finite element (FE) context; in the multilevel case, the CD method has almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}, \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
-In this paper we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach.
-Specifically, our \emph{full approximation scheme constraint decomposition} (FASCD) method extends the algorithms in \cite{GraeserKornhuber2009} and \cite{Tai2003} in the following five directions:
+These methods were shown to converge for multilevel and domain decompositions in a finite element (FE) context.
+In the multilevel case, this CD method has almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}, \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
+
+Here we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach.
+Specifically, our \emph{full approximation scheme constraint decomposition} (FASCD) method extends the \cite{GraeserKornhuber2009,Tai2003} algorithms in the following directions:
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
@@
 \item The FAS coarse corrections are adapted for general VI problems. The needed computations require neither an objective function nor linearity of the residual operator.
+\item Box constraints, namely arbitrary upper and lower obstacles, are allowed.
+\item As with the original CD method \cite{Tai2003}, but unlike some implementations \cite[for example]{GraeserKornhuber2009}, and semismooth methods generally \cite{BensonMunson2006,Ulbrich2011}, the method can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
+\item V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}) are implemented.
 \item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
-\item Box constraints, namely upper and lower obstacles, are permitted.
-\item We implement V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}).
-\item We demonstrate that F-cycle FASCD is an optimal solver for the different VI examples in Section \ref{sec:results}, when used in conjunction with reduced-space active set methods as smoothers.
-The number of Newton and Krylov iterations per step are fixed, so that the work of the smoother is proportional to the number of unknowns on a given level.
-\pef{The solver works for VIs where the residual is only defined on the admissible set \dots}
+\item When used in conjunction with a reduced-space active set method as a smoother, we demonstrate that F-cycle FASCD is an optimal solver for the different VI examples in Section \ref{sec:results}.
+The work of the smoother is proportional to the number of unknowns on a given level because the number of Newton and Krylov iterations is fixed.
 \end{enumerate}
@@
 While FASCD reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed (Appendix \ref{app:reductions}), PFAS reduces to linear multigrid.
-The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so as stated it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
-The problematic latter property is corrected by reformulation using defect constraints and monotone injection operators (see \cite[Algorithm 4.7]{GraeserKornhuber2009} and Section \ref{sec:cdmultilevel} herein).
+The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
+The latter property is improved by reformulation using defect constraints and monotone injection operators (see \cite[Algorithm 4.7]{GraeserKornhuber2009}, or Section \ref{sec:cdmultilevel} herein).
 The monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be formulated only with Gauss--Seidel type smoothers, and applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.
 Furthermore, monotone multigrid methods are more difficult to implement in FE libraries because modifications of the FE basis functions are needed to approximate the coarser-level coincidence sets.
-No such invasive basis modifications are used either in Tai's original multilevel CD or in FASCD.
-Our implementation of FASCD in the Firedrake library \cite{Rathgeberetal2016} requires no modifications of basis functions and yet exhibits excellent performance, including on a problem used to demonstrate the benefits of monotone multigrid (Example \ref{ex:results:classical}; compare \cite[problem 7.1.1]{GraeserKornhuber2009}).
-
-In our experiments, FASCD V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for the corresponding (unconstrained) PDE problems.
+No such invasive basis modifications are used either in Tai's CD, nor in our Firedrake library \cite{Rathgeberetal2016} implementation of FASCD.
+However, our implementation exceeds monotone multigrid performance, including on a problem used to demonstrate the benefits of that method (Example \ref{ex:results:classical}; compare \cite[problem 7.1.1]{GraeserKornhuber2009}).
+
+Regarding its limitations, FASCD solves only bound-constrained VI problems.  The method differs from the original FAS method only if the constraint set is proper (Appendix \ref{app:reductions}).
+
+In our experiments, FASCD V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for corresponding unconstrained PDE problems.
 The F-cycle results in Section \ref{sec:results} show nearly-textbook multigrid efficiency, namely solutions within discretization error in work comparable to few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}.
-Specifically, we see this performance for classical Laplacian obstacle problems, $p$-Laplacian obstacle problems, and nonnegativity-constrained advection-diffusion equations.
+Specifically, we see this performance for classical (unilateral, Laplacian) obstacle problems, $p$-Laplacian obstacle problems, and box-constrained advection-diffusion equations.
 Furthermore, these results appear to be insensitive to the length or structure of the free boundary, i.e.~the geometric complexity of the coincidence set.
@@
 For example, we solve an ice sheet geometry problem (Examples \ref{ex:sia} and \ref{ex:results:sia}) in which the ice flow formulas are only meaningful for surface elevation iterates that do not penetrate the bedrock.
-(More generally, fluid layers subject to a non-negative thickness constraint typically require the solution of an auxiliary PDE on a domain determined inside the VI residual evaluation \cite{Bueler2021conservation}, but this is not needed in the ice sheet problem here.)  Non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural modifications of the operator formula.
-FASCD solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a mesh-independent number of F-cycle iterations (Example \ref{ex:results:sia}).
+Non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural modifications of the operator formula.
+FASCD solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a small, mesh-independent number of F-cycle iterations (Example \ref{ex:results:sia}).
+
+{\color{blue} ELB: beam 2d example?}
 
 The paper is organized as follows.
-In Section \ref{sec:vi} we recall the theory of coercive and monotone variational inequalities, and describe several motivating example problems.
+In Section \ref{sec:vi} we recall the theory of coercive and monotone variational inequalities, and describe several motivating examples.
 Section \ref{sec:cd} reviews the basics of CDs and their associated iterations.
-Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, and then Section \ref{sec:cdmultilevel} presents defect constraints generated via monotone injection operators \cite{GraeserKornhuber2009}.
+Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, and then Section \ref{sec:cdmultilevel} extends the CD method via defect constraints generated by monotone injection operators \cite{GraeserKornhuber2009}.
 Here we propose an apparently new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD and an associated iteration.
-Section \ref{sec:vcycle} then states the FASCD V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, F-cycles, and smoother construction.
-Section \ref{sec:results} gives numerical results, from a parallel Firedrake \cite{Rathgeberetal2016} implementation (see Code Availability), for each example problem introduced in Section \ref{sec:vi}.
+Section \ref{sec:vcycle} then states the FASCD V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, F-cycles, and our smoother choice.
+Section \ref{sec:results} gives numerical results from a parallel Firedrake \cite{Rathgeberetal2016} implementation (see Code Availability), for each example problem introduced in Section \ref{sec:vi}.
 The work concludes with a discussion and outlook in Section \ref{sec:discussion}.

### 2023-07-22 paper/fascd.tex
address PEF comments; see a few ELB comments remaining

@@
 
-{\color{blue} ELB: beam 2d example?}
+\elb{beam 2d example?}
 
@@
 
-To allow box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the vertices of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
-The finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ are constructed by suitable interpolation of $\underline{\gamma}$, $\overline{\gamma}$ (e.g.~by applying the Cl\'ement interpolation operator) and satisfy
+To construct box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the vertices of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
+The finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ are constructed by suitable interpolation of $\underline{\gamma}$, $\overline{\gamma}$ (e.g.~by applying the Cl\'ement interpolation operator \elb{cite?}).
+They must satisfy
 \begin{equation}
@@
 Analogously, we construct the discrete boundary data $g_D^J$ so that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those vertices of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.
-Define
+Then we define
 \begin{equation}
@@
 Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ also acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $(R\sigma)[z] = \sigma[z]$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
-These canonical transfers generally involve a change of FE representation.
-For example, $R$ is ``full-weighting'' \cite{Trottenbergetal2001}.
-
-To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, observe that the vertices of $\mathcal{T}^{j-1}$ are vertices of $\mathcal{T}^j$, and that for any values at $\mathcal{T}^j$ vertices there is a unique $\mathcal{V}^j$ function assuming those values.
-For $z\in\mathcal{V}^j$ we define the injection $\iR z$ as the element of $\mathcal{V}^{j-1}$ with the same (coarse) nodal values.
+These canonical transfers generally involve a change of FE representation; $R$ is ``full-weighting'' \cite{Trottenbergetal2001}.
+
+To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, note that the vertices of $\mathcal{T}^{j-1}$ are vertices of $\mathcal{T}^j$.
+For $z\in\mathcal{V}^j$ we define $\iR z$ as the element of $\mathcal{V}^{j-1}$ with the same (coarse) nodal values.
 Injection is monotonic: if $w \le z$ then $\iR w \le \iR z$.
 
-The monotone injections $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to nodal functions $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the \emph{stars} of the coarse vertices), Figure \ref{fig:Rplusminus}.
+The monotone injections $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to constraints in $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the stars of the coarse vertices), Figure \ref{fig:Rplusminus}.
 These operators are constructed to enforce the property
 \begin{equation}
-\minR z \le z \le \maxR z,  \label{eq:fe:monotonerestrictionprops}
+\minR z \le z \le \maxR z,  \label{eq:fe:monotoneproperty}
 \end{equation}
@@
 \begin{align}
-(\maxR z)(x_p^{j-1}) &= \max \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}, \label{eq:fe:monotonerestrictions} \\
+(\maxR z)(x_p^{j-1}) &= \max \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}, \label{eq:fe:monotoneinjections} \\
 (\minR z)(x_p^{j-1}) &= \min \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}. \notag
 \end{align}
-For standard refinement a coarse-mesh node $x_p^{j-1}$ is in the set of fine vertices $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so \eqref{eq:fe:monotonerestrictionprops} follows.
-On the other hand, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These ordering properties are used in the CD construction of the next two Sections.
-\pef{Promote this to its own numbered equation and highlight its importance later on in the construction of LDCs}
+For standard refinement a coarse-mesh node $x_p^{j-1}$ is in the set of fine vertices $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so \eqref{eq:fe:monotoneproperty} follows.
+In the next Section we use $\maxR,\minR$ to construct ``level defect constraints'' such that the zero function is always admissible.
+This construction depends on an additional ordering property of the monotone injections:
+\begin{equation}
+z\ge 0 \implies \minR z \ge 0, \qquad z \le 0 \implies \maxR z \le 0. \label{eq:fe:monotoneadditional}
+\end{equation}
 
@@
 \begin{equation}
-    \label{eq:fe:standardinjection}
+    \label{eq:fe:badstandard}
     \underline{\gamma}^0, \overline{\gamma}^0 \equiv \iR \underline{\gamma}^1, \iR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}
@@
 \begin{equation}
-    \label{eq:fe:monotoneinjection}
+    \label{eq:fe:badmonotone}
     \underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}
@@
 \end{equation}
-Observe that
+Observe that by \eqref{eq:fe:monotoneproperty} and \eqref{eq:fe:monotoneadditional},
 \begin{equation}
@@
 \end{equation}
-(Compare equation (4.9) in \cite{GraeserKornhuber2009}.)  Property \eqref{eq:fe:monotonerestrictionprops} implies $I_{j\to i}^\ominus(z^j - \underline{\chi}^j) \ge I_{j\to i-1}^\ominus(z^j - \underline{\chi}^j)$, thus $\Pi_i z^j \ge \underline{\phi}^i$, so $\Pi_i z^j \in \mathcal{D}^i$.
+(Compare equation (4.9) in \cite{GraeserKornhuber2009}.)  Property \eqref{eq:fe:monotoneproperty} implies $I_{j\to i}^\ominus(z^j - \underline{\chi}^j) \ge I_{j\to i-1}^\ominus(z^j - \underline{\chi}^j)$, thus $\Pi_i z^j \ge \underline{\phi}^i$, so $\Pi_i z^j \in \mathcal{D}^i$.
 On the other hand, the following sum telescopes, leaving only its first term plus the sum in \eqref{eq:fe:telescoping}:
@@
 In contrast to the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, standard injection is used to generate the $\underline{\gamma}^j,\overline{\gamma}^j$ on each level during the initial nested iteration.
-This is because nested iteration approximates the full solution on the coarse grids, not just a correction to an initial iterate; essentially, our choice is between \eqref{eq:fe:standardinjection} (where the coarse constraint set is always nonempty, but the solution may not be admissible on the next level) or \eqref{eq:fe:monotoneinjection} (where the coarse solution is always admissible if it exists, but the constraint set may be empty).
-We choose \eqref{eq:fe:standardinjection} so that the coarse problems always have nonempty constraint sets, and project pointwise to the bounds as we refine in the nested iteration.
+This is because nested iteration approximates the full solution on the coarse grids, not just a correction to an initial iterate; essentially, our choice is between \eqref{eq:fe:badstandard} (where the coarse constraint set is always nonempty, but the solution may not be admissible on the next level) or \eqref{eq:fe:badmonotone} (where the coarse solution is always admissible if it exists, but the constraint set may be empty).
+We choose \eqref{eq:fe:badstandard} so that the coarse problems always have nonempty constraint sets, and project pointwise to the bounds as we refine in the nested iteration.
 On line \ref{line:fcyclecoarsestinitial}, one may generate an initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation of a finest-level iterate, if one is available.
@@
 
+\elb{beam example?}
+
 \section{Discussion and outlook} \label{sec:discussion}
@@
 
-%While an an extension to nonlocal, non-differential VI problems is of interest, it was not tried here
-
-\subsection*{Code availability} \label{sec:code}  For reproducibility we have archived the exact software versions used to produce the results in this paper, including the correct Firedrake version, at {\color{blue} ELB: Zenodo?}
+\subsection*{Code availability} \label{sec:code}  The software used to produce the results is archived at tag \elb{FIX} in the repository \url{https://bitbucket.org/pefarrell/fascd/}, using the Firedrake version \elb{what version?}
 
@@
 The downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets are also enlarged to vector spaces, and the corrections $y^j$ and $z^j$ are unconstrained; lines \ref{line:vcyclegenchifinest}, \ref{line:vcyclegenchi}, and \ref{line:vcyclegenphi} in Algorithm \ref{alg:fascd} can be removed.
-In fact, VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} can be rewritten as the weak form problem $\ip{f^j(u^j)}{v} = \ip{\ell^j}{v}$ over $v\in \mathcal{V}_0^j$---this groups $u^j=w^j+y^j$ when down-smoothing and $u^j=w^j+z^j$ when up-smoothing---and the smoothers can be regarded as unconstrained, in-place modifications of $u^j$.
+Then VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} can be rewritten as the weak form problem $\ip{f^j(u^j)}{v} = \ip{\ell^j}{v}$ over $v\in \mathcal{V}_0^j$, where $u^j=w^j+y^j$ when down-smoothing and $u^j=w^j+z^j$ when up-smoothing, and the smoothers can be regarded as unconstrained, in-place modifications of $u^j$.
 Noting line \ref{line:vcyclerestrictsolution} now says $u^{j-1}=\iR u^j$, after down-smoothing the coarsened source functional computed by line \ref{line:vcyclerestrictell} can be rewritten as $\ell^{j-1} = f^{j-1}\left(u^{j-1}\right) + R\left(\ell^j-f^j(u^j)\right)$.
@@
 With these modifications, Algorithm \ref{alg:fascd} becomes the FAS V-cycle below, as a solver for \eqref{eq:app:fas:pde}.
-This algorithm appears in various sources in the literature, for example \cite[Algorithm 14]{Bruneetal2015} and \cite[section 5.3.4]{Trottenbergetal2001}.
+It appears in various literature, for example \cite[Algorithm 14]{Bruneetal2015} and \cite[section 5.3.4]{Trottenbergetal2001}.
 

### 2023-07-23 paper/fascd.tex
Some edits; not sure if I want to offer these yet

@@
 \item The FAS coarse corrections are adapted for general VI problems. The needed computations require neither an objective function nor linearity of the residual operator.
-\item Box constraints, namely arbitrary upper and lower obstacles, are allowed.
+\item Box constraints (i.e.~both upper and lower obstacles) are allowed.
 \item As with the original CD method \cite{Tai2003}, but unlike some implementations \cite[for example]{GraeserKornhuber2009}, and semismooth methods generally \cite{BensonMunson2006,Ulbrich2011}, the method can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
+\pef{I confess I find this explicit calling out of other algorithms a bit uncomfortable/too much. I wonder if the following is better?}
+\item The method is strictly admissible, and so can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
 \item V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}) are implemented.
@@
 Furthermore, monotone multigrid methods are more difficult to implement in FE libraries because modifications of the FE basis functions are needed to approximate the coarser-level coincidence sets.
-No such invasive basis modifications are used either in Tai's CD, nor in our Firedrake library \cite{Rathgeberetal2016} implementation of FASCD.
+\pef{My current understanding is that this refers to \emph{truncated} monotone multigrid; monotone multigrid just means using monotone injection operators appropriately.
+Let me double-check.
+And my understanding is that the `modification' of the basis functions is just zeroing the basis functions for vertices in the active set, so you don't have to modify the basis functions, just zero some rows/columns of the matrix.} No such invasive basis modifications are used either in Tai's CD, nor in our Firedrake library \cite{Rathgeberetal2016} implementation of FASCD.
 However, our implementation exceeds monotone multigrid performance, including on a problem used to demonstrate the benefits of that method (Example \ref{ex:results:classical}; compare \cite[problem 7.1.1]{GraeserKornhuber2009}).

### 2023-07-23 paper/fascd.tex
More edits

@@
 Specifically, our \emph{full approximation scheme constraint decomposition} (FASCD) method extends the \cite{GraeserKornhuber2009,Tai2003} algorithms in the following directions:
+\pef{These aren't all really extensions. Maybe we could say `enjoys the following advantages'? We distinguish it from prior work below, so this section does not need to do that.}
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
@@
 While FASCD reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed (Appendix \ref{app:reductions}), PFAS reduces to linear multigrid.
-The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
+The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so as stated it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
 The latter property is improved by reformulation using defect constraints and monotone injection operators (see \cite[Algorithm 4.7]{GraeserKornhuber2009}, or Section \ref{sec:cdmultilevel} herein).
 The monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be formulated only with Gauss--Seidel type smoothers, and applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.
-Furthermore, monotone multigrid methods are more difficult to implement in FE libraries because modifications of the FE basis functions are needed to approximate the coarser-level coincidence sets.
-\pef{My current understanding is that this refers to \emph{truncated} monotone multigrid; monotone multigrid just means using monotone injection operators appropriately.
-Let me double-check.
-And my understanding is that the `modification' of the basis functions is just zeroing the basis functions for vertices in the active set, so you don't have to modify the basis functions, just zero some rows/columns of the matrix.} No such invasive basis modifications are used either in Tai's CD, nor in our Firedrake library \cite{Rathgeberetal2016} implementation of FASCD.
-However, our implementation exceeds monotone multigrid performance, including on a problem used to demonstrate the benefits of that method (Example \ref{ex:results:classical}; compare \cite[problem 7.1.1]{GraeserKornhuber2009}).
-
-Regarding its limitations, FASCD solves only bound-constrained VI problems.  The method differs from the original FAS method only if the constraint set is proper (Appendix \ref{app:reductions}).
+Unlike the fastest variants of monotone multigrid, no zeroing of the basis functions associated with the active set (truncation) is employed in our method, both simplifying the implementation and allowing the coarse grid corrections to update the estimate of the active set on the fine grid.
+Numerical evidence indicates that the FAS construction of coarse grid corrections is useful for this purpose, even when the coarse grids are too coarse to usefully resolve the active set.
+\pef{Insert comparison against the spiral problem once we have set rtol = 1e-12}
+A limitation of FASCD is that, as formulated, it only solves bound-constrained VI problems.
 
@@
 To construct box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the vertices of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
-The finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ are constructed by suitable interpolation of $\underline{\gamma}$, $\overline{\gamma}$ (e.g.~by applying the Cl\'ement interpolation operator \elb{cite?}).
+The finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ are constructed by suitable interpolation of $\underline{\gamma}$, $\overline{\gamma}$ (e.g.~by applying the Cl\'ement interpolation operator~\cite{Carstensen2006}).
 They must satisfy
@@
 
-The monotone injections $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to constraints in $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the stars of the coarse vertices), Figure \ref{fig:Rplusminus}.
+The monotone injections $\maxR,\minR$ \cite{Tai2003,GraeserKornhuber2009}, which will only be applied to constraints in $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the stars of the coarse vertices), Figure \ref{fig:Rplusminus}.
 These operators are constructed to enforce the property
@@
 The initial iterate was set to $w=\max\{0,\psi\}$.
+\pef{For a fair comparison with the spiral problem in \cite{GraeserKornhuber2009}, I think we should set rtol = 1e-12}
 
@@
 
-The 2D solver applied $P_1$ elements and a single processor, but in 3D the elements were chosen to be $Q_1$ (hexahedra) and the run parallel ($P=16$ processes).
+The 2D solver applied $P_1$ elements on a single core, but in 3D the elements were chosen to be $Q_1$ (hexahedra) on $P=16$ cores.
 A coarse mesh with $m_0=16^d$ vertices minimally-resolved the coincidence set.
@@
 
-In contrast with monotone multigrid \cite{Kornhuber1994}, FASCD does not modify the FE discretization, but the possibility of acceleration via basis-level manipulations is a topic for future research.
-On the other hand,  the convergence proofs in \cite{GraeserKornhuber2009,Kornhuber1994}, which appear to depend on particular smoother and basis-level manipulation choices, have no analog here.
-
 \subsection*{Code availability} \label{sec:code}  The software used to produce the results is archived at tag \elb{FIX} in the repository \url{https://bitbucket.org/pefarrell/fascd/}, using the Firedrake version \elb{what version?}
@@
 With these modifications, Algorithm \ref{alg:fascd} becomes the FAS V-cycle below, as a solver for \eqref{eq:app:fas:pde}.
-It appears in various literature, for example \cite[Algorithm 14]{Bruneetal2015} and \cite[section 5.3.4]{Trottenbergetal2001}.
+This algorithm appears in various sources in the literature, for example \cite[Algorithm 14]{Bruneetal2015} and \cite[section 5.3.4]{Trottenbergetal2001}.
 

### 2023-07-24 paper/fascd.tex
Minor edits

@@
 \item The method is strictly admissible, and so can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
-\item V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}) are implemented.
+\item V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}) are proposed.
 \item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
@@
 This structure can be exploited by a new CD iteration, algorithm \pr{icd-tele} in Section \ref{sec:cdmultilevel}.
+\pef{This last paragraph feels like a flat way to end. Perhaps we could highlight the importance of ICD-TELE in the text, and remove this last paragraph?}
 
@@
 This work used the ARCHER2 UK National Supercomputing Service (https://www.archer2.ac.uk).
-PEF thanks Jack D.~Betteridge for assistance with the numerical experiments of Section \ref{ex:results:sia}.
+PEF thanks Jack D.~Betteridge for assistance with the numerical experiments of Example \ref{ex:results:sia}.
 

### 2023-07-24 paper/fascd.tex
switch over to SIAM style ... will sweep through and clean figure/code appearance

@@
-\documentclass[letterpaper,final,12pt,reqno]{amsart}
-
-\usepackage[total={6.3in,9.2in},top=1.1in,left=1.1in]{geometry}
-
-\usepackage{times,bm,bbm,empheq,verbatim,fancyvrb,graphicx,amsthm,amssymb}
-\usepackage[dvipsnames]{xcolor}
+\documentclass[review,hidelinks,onefignum,onetabnum]{siamart220329}
+
+\usepackage{amsfonts}
+\usepackage{graphicx}
+\usepackage{epstopdf}
+\ifpdf
+  \DeclareGraphicsExtensions{.eps,.pdf,.png,.jpg}
+\else
+  \DeclareGraphicsExtensions{.eps}
+\fi
+
+% Add a serial/Oxford comma by default.
+\newcommand{\creflastconjunction}{, and~}
+
+% Used for creating new theorem and remark environments
+\newsiamremark{remark}{Remark}
+\newsiamremark{hypothesis}{Hypothesis}
+\crefname{hypothesis}{Hypothesis}{Hypotheses}
+\newsiamthm{claim}{Claim}
+\newsiamremark{example}{Example}
+
+\usepackage{amsopn}
+\DeclareMathOperator{\diag}{diag}
+
+\usepackage{bm,bbm,empheq,verbatim,fancyvrb,amssymb}
 \usepackage{booktabs,multirow,xspace}
 \usepackage{pifont}
-
-\usepackage{tabto}
-\TabPositions{1.5cm}
 
@@
 
-\usepackage{float}
-
-% hyperref should be the last package we load
-\usepackage[pdftex,
-colorlinks=true,
-plainpages=false, % only if colorlinks=true
-linkcolor=blue,   % ...
-citecolor=Red,    % ...
-urlcolor=black    % ...
-]{hyperref}
-
-\renewcommand{\baselinestretch}{1.05}
-
-\allowdisplaybreaks[1]  % allow display breaks in align environments, if they avoid major underfulls
-
-\newtheoremstyle{cstyle}% name
-  {5pt}% space above
-  {5pt}% space below
-  {\itshape}% body font
-  {}% indent amount
-  {\itshape}% theorem head font
-  {.}% punctuation after theorem head
-  {.5em}% space after theorem head
-  {\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}% theorem head spec
-\theoremstyle{cstyle}
-
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{assumptions}[theorem]{Assumptions}
-
-\newtheoremstyle{cstyle*}% name
-  {5pt}% space above
-  {5pt}% space below
-  {\itshape}% body font
-  {}% indent amount
-  {\itshape}% theorem head font
-  {.}% punctuation after theorem head
-  {.5em}% space after theorem head
-  {\thmname{#1}}% theorem head spec
-\theoremstyle{cstyle*}
-\newtheorem{assumptions*}{Assumptions}
-
-\newtheoremstyle{dstyle}% name
-  {5pt}% space above
-  {5pt}% space below
-  {}%{\itshape}% body font
-  {}% indent amount
-  {\itshape}% theorem head font
-  {.}% punctuation after theorem head
-  {.5em}% space after theorem head
-  {\thmname{#1}\thmnumber{ #2}\thmnote{ (#3)}}% theorem head spec
-\theoremstyle{dstyle}
-
-\newtheorem{definition}[theorem]{Definition}
-\newtheorem{example}[theorem]{Example}
-
-% numbering
-\numberwithin{equation}{section}
-\numberwithin{figure}{section}
-\numberwithin{table}{section}
-\numberwithin{theorem}{section}
-
 \newcommand{\eps}{\epsilon}
@@
 \newcommand{\trace}{\operatorname{tr}}
-
-\newcommand{\hbn}{\hat{\mathbf{n}}}
 
@@
 
+% Sets running headers as well as PDF title and authors
+\headers{FAS for nonlinear variational inequalities}{E. Bueler and P. E. Farrell}
+
+\title{A full approximation scheme multilevel method \\ for nonlinear variational inequalities\thanks{Submitted to the editors DATE.
+\funding{EB was supported by a Faculty Development Travel Award from United Academics, University of Alaska Fairbanks.
+PEF was supported by the Engineering and Physical Sciences Research Council [EPSRC grants EP/R029423/1 and EP/W026163/1].
+This work used the ARCHER2 UK National Supercomputing Service (https://www.archer2.ac.uk).
+PEF thanks Jack D.~Betteridge for assistance with the numerical experiments of Example \ref{ex:results:sia}.}}}
+
+\author{Ed Bueler\thanks{Department of Mathematics and Statistics, University of Alaska Fairbanks, USA
+  (\email{elbueler@alaska.edu}).}
+\and Patrick E. Farrell\thanks{Mathematical Institute, University of Oxford, Oxford, UK
+  (\email{patrick.farrell@maths.ox.ac.uk}).}}
+
 \begin{document}
-\title[FAS for nonlinear variational inequalities]{A full approximation scheme multilevel method \\ for nonlinear variational inequalities}
-
-\author{Ed Bueler}
-
-\author{Patrick E.~Farrell}
-
-\date{\today}
+
+\maketitle
 
@@
 
-\maketitle
-
-\thispagestyle{empty}
+% REQUIRED
+\begin{keywords}
+FIXME
+\end{keywords}
+
+% REQUIRED
+\begin{MSCcodes}
+FIXME
+\end{MSCcodes}
 
@@
 
-\subsection*{Acknowledgements} \label{sec:acknowledgements}  EB was supported by a Faculty Development Travel Award from United Academics, University of Alaska Fairbanks.
-PEF was supported by the Engineering and Physical Sciences Research Council [EPSRC grants EP/R029423/1 and EP/W026163/1].
-This work used the ARCHER2 UK National Supercomputing Service (https://www.archer2.ac.uk).
-PEF thanks Jack D.~Betteridge for assistance with the numerical experiments of Example \ref{ex:results:sia}.
-
+\elb{acknowledge Lawrence for help with monotone injections?}
+
+\bibliographystyle{siamplain}
 \bibliography{fascd}
-\bibliographystyle{siam}
 

### 2023-07-24 paper/results/drape.sh
will run this

@@ -4,7 +4,7 @@ for PROB in ball spiral; do
     echo ""
     echo "******** -nofas PROB=${PROB} *******"
     for LEVS in 1 2 3 4 5 6 7 8; do
-        python3 drape.py -prob $PROB -levs $LEVS -nofas
+        python3 drape.py -prob $PROB -levs $LEVS -nofas -snes_rtol 1.0e-12 -snes_atol 1.0e-12
     done
 done
 
@@ -12,8 +12,8 @@ for CYCLE in multiplicative full; do
     for PROB in ball spiral; do
         echo ""
         echo "******** PROB=${PROB} CYCLE=${CYCLE} *******"
-        for LEVS in 1 2 3 4 5 6 7 8 9 10; do
-            python3 drape.py -prob $PROB -levs $LEVS -fascd_cycle_type $CYCLE
+        for LEVS in 1 2 3 4 5 6 7 8 9 10 11; do
+            python3 drape.py -prob $PROB -levs $LEVS -fascd_cycle_type $CYCLE -fascd_rtol 1.0e-12 -fascd_atol 1.0e-12
         done
     done
 done

### 2023-07-25 paper/fascd.tex
keywords and MSC codes
Some small edits to reduce the length of the text, more are necessary

@@
 \begin{keywords}
-FIXME
+multigrid, variational inequalities, full approximation scheme, constraint decomposition
 \end{keywords}
@@
 \begin{MSCcodes}
-FIXME
+65K15, 35M86, 90C33
 \end{MSCcodes}
@@
 These methods were shown to converge for multilevel and domain decompositions in a finite element (FE) context.
-In the multilevel case, this CD method has almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}, \cite[Theorem 4.6 and Algorithm 4.7]{GraeserKornhuber2009}.
+In the multilevel case, this CD method has almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}, \cite[Theorem 4.6]{GraeserKornhuber2009}.
 
 Here we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach.
-Specifically, our \emph{full approximation scheme constraint decomposition} (FASCD) method extends the \cite{GraeserKornhuber2009,Tai2003} algorithms in the following directions:
+Specifically, our \emph{full approximation scheme constraint decomposition} (FASCD) method extends the algorithms of \cite{GraeserKornhuber2009,Tai2003} in the following directions:
 \pef{These aren't all really extensions. Maybe we could say `enjoys the following advantages'? We distinguish it from prior work below, so this section does not need to do that.}
@@
 \begin{enumerate}
-\item The FAS coarse corrections are adapted for general VI problems. The needed computations require neither an objective function nor linearity of the residual operator.
+\item The FAS coarse corrections are adapted for general VI problems. The method requires neither an objective function nor linearity of the residual.
 \item Box constraints (i.e.~both upper and lower obstacles) are allowed.
-\item As with the original CD method \cite{Tai2003}, but unlike some implementations \cite[for example]{GraeserKornhuber2009}, and semismooth methods generally \cite{BensonMunson2006,Ulbrich2011}, the method can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
-\pef{I confess I find this explicit calling out of other algorithms a bit uncomfortable/too much. I wonder if the following is better?}
+%\item As with the original CD method \cite{Tai2003}, but unlike some implementations \cite[for example]{GraeserKornhuber2009}, and semismooth methods generally \cite{BensonMunson2006,Ulbrich2011}, the method can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
+%\pef{I confess I find this explicit calling out of other algorithms a bit uncomfortable/too much. I wonder if the following is better?}
 \item The method is strictly admissible, and so can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
-\item V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (full multigrid; Algorithm \ref{alg:fascd-fcycle}) are proposed.
+\item V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (Algorithm \ref{alg:fascd-fcycle}) are proposed.
 \item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
-\item When used in conjunction with a reduced-space active set method as a smoother, we demonstrate that F-cycle FASCD is an optimal solver for the different VI examples in Section \ref{sec:results}.
+\item When a (mesh-dependent) reduced-space active set method is used as relaxation, we demonstrate that F-cycle FASCD is an optimal solver for the different VI examples in Section \ref{sec:results}.
 The work of the smoother is proportional to the number of unknowns on a given level because the number of Newton and Krylov iterations is fixed.
@@
 
-\elb{beam 2d example?}
+%\elb{beam 2d example?}
 
@@
 
-\begin{lemma}  \label{lem:advectionskew}  \cite{Elmanetal2014}\,  Suppose $\bX :\Omega \to \RR^d$ is a bounded and boundedly-differentiable vector field on $\Omega$ with zero divergence ($\Div \bX=0$).
+\begin{lemma}  \label{lem:advectionskew}  \cite{Elmanetal2014}\,  Suppose $\bX :\Omega \to \RR^d$ is a bounded and boundedly-differentiable vector field on $\Omega$ with $\Div \bX=0$.
 For $u,v \in W^{1,2}(\Omega)$ let $b(u,v) = \int_\Omega (\bX \cdot \grad u) v\dx$.
@@
 Let $\cV \subset W^{1,2}(\Omega)$ be the space of functions which are zero along $\partial_D\Omega$.
-Consider a divergence-free velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, and assume that the flow is outward on the Neumann boundary: $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.
+Consider a velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, and assume that the flow is outward on the Neumann boundary: $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.
 For $u,v \in \cV$, $\eps>0$, and $\phi \in \cV'$, define
@@
 We seek to rapidly compute the approximate solution of box-constrained VI problems \eqref{eq:boxdirichletvi} using finite elements (FE).
-Let $\Omega \subset \RR^d$ be a (open) bounded polygon.
-A multilevel FE approximation to \eqref{eq:boxdirichletvi} is constructed, using $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, by refining a coarsest mesh into a mesh hierarchy.
-The hierarchy allows the creation of a multilevel CD of $\cK$ using the ideas in the next two Sections.
+Let $\Omega \subset \RR^d$ be an open bounded polygon.
+A multilevel FE approximation to \eqref{eq:boxdirichletvi} is constructed, using $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, by refinement of a coarse mesh.
+The hierarchy of meshes allows the creation of a multilevel CD of $\cK$ using the ideas in the next two Sections.
 
@@
 Nevertheless, the finest mesh result, with $m_J=20481^2=4.1 \times 10^8$ degrees of freedom, required less than three minutes to solve for the steady surface elevation of an ice sheet larger than the current Greenland ice sheet ($1.7$ million $\text{km}^2$) at uniform $87.5$ m resolution.
-\end{example}
 
@@
 \end{table}
+\end{example}
 

### 2023-07-25 paper/fascd.tex
Some centering, some acknowledging Lawrence

@@
 This work used the ARCHER2 UK National Supercomputing Service (https://www.archer2.ac.uk).
-PEF thanks Jack D.~Betteridge for assistance with the numerical experiments of Example \ref{ex:results:sia}.}}}
+PEF thanks Lawrence Mitchell for advice on the implementation of FASCD, and thanks Jack D.~Betteridge for assistance with the numerical experiments of Example \ref{ex:results:sia}.}}}
 
@@
 \begin{figure}[ht]
+\centering
 \includegraphics[width=0.6\textwidth]{genfigs/cartoon.pdf}
@@
 \begin{table}[H]
+\centering
 \begin{tabular}{llc}
@@
 \begin{figure}[ht]
+\centering
 \input{tikz/Rplusminus.tex}
@@
 \begin{figure}[ht]
+\centering
 \begin{tabular}{cc}
@@
 \begin{figure}[ht]
+\centering
 \includegraphics[width=0.65\textwidth]{fixfigs/chiphilevels.pdf}
@@
 \begin{figure}[ht]
-\begin{center}
+\centering
 \input{tikz/vcycle.tex}
-\end{center}
 \caption{The FASCD V-cycle computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.}
@@
 \begin{figure}[h]
-\begin{center}
+\centering
 \input{tikz/fcycle.tex}
-\end{center}
 \caption{F-cycle Algorithm \ref{alg:fascd-fcycle} initially restricts and injects the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray).
@@
 \begin{figure}[ht]
-\begin{center}
+\centering
 \includegraphics[width=0.35\textwidth]{fixfigs/ball-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/spiral-set.png}
-\end{center}
 \caption{Coincidence sets for the ball and spiral problems in Example \ref{ex:results:classical}.}
@@
 \begin{table}[ht]
+\centering
 \begin{tabular}{lr@{\hskip 7mm}cccccc}
@@
 \begin{figure}[ht]
-\begin{center}
+\centering
 \includegraphics[width=0.45\textwidth]{fixfigs/plap1d1p5.pdf} \quad
 \includegraphics[width=0.45\textwidth]{fixfigs/plap1d4p0.pdf}
-\end{center}
 \caption{Solutions $u\ge \psi$ of the $p$-Laplacian VI in Example \ref{ex:results:plap}, for $p=1.5$ (left) and $p=4.0$ (right).}
@@
 \begin{figure}[ht]
-\begin{center}
+\centering
 \input{tikz/imagesvcycle.tex}
-\end{center}
 \caption{Visualization of one FASCD V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of $m_0=5$ vertices.
@@
 \begin{table}[ht]
+\centering
 \begin{tabular}{lr@{\hskip 7mm}c@{\hskip 3mm}c@{\hskip 3mm}c@{\hskip 4mm}c@{\hskip 5mm}c@{\hskip 6mm}c}
@@
 \begin{table}[ht]
+\centering
 \begin{tabular}{lr@{\hskip 7mm}c@{\hskip 4mm}c@{\hskip 4mm}c@{\hskip 6mm}c}
@@
 \begin{figure}[ht]
-\begin{center}
+\centering
 \includegraphics[width=0.35\textwidth]{fixfigs/poll2d-zero-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/poll2d-one-set.png}
-\end{center}
 \caption{Coincidence sets for the 2D bi-obstacle problem in Example \ref{ex:results:advdiff}: $u=0$ (left) and $u=1$ (right).}
@@
 \begin{table}[ht]
+\centering
 \begin{tabular}{lr@{\hskip 7mm}c@{\hskip 4mm}c}
@@
 \begin{figure}[ht]
-\begin{center}
+\centering
 \includegraphics[width=1.0\textwidth]{fixfigs/sialev8scene.png}
-\end{center}
 \caption{Solution of Example \ref{ex:sia} over a bumpy bedrock topography (grey), with the ice speed as surface coloring and 100-times vertical exaggeration.}
@@
 \begin{table}[ht]
+\centering
 \begin{tabular}{c@{\hskip 4mm}c@{\hskip 4mm}c@{\hskip 7mm}c@{\hskip 4mm}c}
@@
 
-\elb{acknowledge Lawrence for help with monotone injections?}
-
 \bibliographystyle{siamplain}

### 2023-07-25 paper/results/drape.sh
there is an irritatingly-robust barrier on certain fine grids to residual reduction more that 10^-8 ... I do not understand it

@@ -1,19 +1,0 @@
-# generate results for Example 8.1 in paper
-
-for PROB in ball spiral; do
-    echo ""
-    echo "******** -nofas PROB=${PROB} *******"
-    for LEVS in 1 2 3 4 5 6 7 8; do
-        python3 drape.py -prob $PROB -levs $LEVS -nofas -snes_rtol 1.0e-12 -snes_atol 1.0e-12
-    done
-done
-
-for CYCLE in multiplicative full; do
-    for PROB in ball spiral; do
-        echo ""
-        echo "******** PROB=${PROB} CYCLE=${CYCLE} *******"
-        for LEVS in 1 2 3 4 5 6 7 8 9 10 11; do
-            python3 drape.py -prob $PROB -levs $LEVS -fascd_cycle_type $CYCLE -fascd_rtol 1.0e-12 -fascd_atol 1.0e-12
-        done
-    done
-done

### 2023-07-26 paper/results/drape_fascd.sh
using fascd branch elbueler/stol, drape results look good

@@ -4,7 +4,7 @@ for PROB in ball spiral; do
     echo ""
     echo "******** PROB=${PROB}, V-cycles, tight tolerance (over-solving) *******"
     for LEVS in 1 2 3 4 5 6 7 8 9 10; do
-        python3 drape.py -prob $PROB -levs $LEVS -fascd_cycle_type multiplicative -fascd_rtol 1.0e-12 -fascd_atol 1.0e-12 -fascd_max_it 50
+        python3 drape.py -prob $PROB -levs $LEVS -fascd_cycle_type multiplicative -fascd_rtol 1.0e-12
     done
 done
 
@@ -12,6 +12,6 @@ for PROB in ball spiral; do
     echo ""
     echo "******** PROB=${PROB}, F-cycles, to discretization error *******"
     for LEVS in 1 2 3 4 5 6 7 8 9 10; do
-        python3 drape.py -prob $PROB -levs $LEVS -fascd_cycle_type full -fascd_rtol 1.0e-6 -fascd_atol 1.0e-12 -fascd_max_it 50
+        python3 drape.py -prob $PROB -levs $LEVS -fascd_cycle_type full
     done
 done

### 2023-07-26 paper/results/drape_nofas.sh

@@ -4,6 +4,6 @@ for PROB in ball spiral; do
     echo ""
     echo "******** -nofas PROB=${PROB} *******"
     for LEVS in 1 2 3 4 5 6 7 8; do
-        python3 drape.py -prob $PROB -levs $LEVS -nofas -snes_rtol 1.0e-12 -snes_atol 1.0e-12
+        python3 drape.py -prob $PROB -levs $LEVS -nofas -snes_rtol 1.0e-12
     done
 done

### 2023-07-26 paper/fascd.tex
WIP: updating Example 8.1 results

@@
 
-A large space of smoothers could be explored, but in Section \ref{sec:results} we apply a fixed number of iterations of the active-set (reduced-space) Newton method from \cite{BensonMunson2006}, the \texttt{VINEWTONRSLS} solver in PETSc \cite{Balayetal2023}, denoted as \emph{RS Newton}.
-A fixed number of preconditioned Krylov iterations are applied to approximately solve the arising linear systems.
-For the symmetric Laplacian in Example \ref{ex:results:classical} we apply conjugate gradients (CG) with incomplete Cholesky (IC) preconditioning, but for other operators we apply the generalized minimum residual (GMRES) method with incomplete LU (ILU) preconditioning.
-This smoother requires $O(m_j)$ work per iteration.
-The coarse solver is also RS Newton, but iterated to convergence, with a direct solver for the arising linear systems.
+A large space of smoothers could be explored, but in the next Section we will only apply an active-set (reduced-space) Newton method \cite{Balayetal2023,BensonMunson2006}, denoted as \emph{RS Newton}.
+A fixed number of preconditioned Krylov iterations approximately solve the arising linear systems, thus $O(m_j)$ work per smoother application.
+For the symmetric Laplacian (Example \ref{ex:results:classical}) we apply conjugate gradients (CG) with incomplete Cholesky (IC) preconditioning.
+The generalized minimum residual (GMRES) with incomplete LU (ILU) preconditioning is used in other cases.
+The coarse solver is also RS Newton, but iterated to convergence using a direct linear solver.
 
@@
 
-Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python, for $P_1$ and $Q_1$ elements, using the Firedrake library \cite{Rathgeberetal2016}.
-Here we show numerical results for the example problems presented in Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
-Our primary observations are that each V-cycle greatly reduces the residual norm, and that very few F-cycles solve most examples to within discretization error.
+For $P_1$ and $Q_1$ elements, Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}.
+The numerical results in this Section follow the Section \ref{sec:vi} examples, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
+Our primary observations are that each V-cycle greatly reduces the residual norm, and that a few F-cycles solve most examples to within discretization error.
 
 \begin{example} \label{ex:results:classical}
-Consider the unilateral 2D Laplacian obstacle problem, Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$.
-We consider two problems, both over square domains $\Omega$, with the very different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$ shown in Figure \ref{fig:results:classical}.
-The standard ``ball'' problem has an upper hemisphere obstacle $\psi$ \cite[Chapter 12]{Bueler2021}.
-For the ``spiral'' problem \cite[problem 7.1.1]{GraeserKornhuber2009}, $\psi$ is constructed to produce the coincidence set shown.
-For this experiment we used a coarsest regular-triangulation mesh with $m_0=5^2$ vertices (32 triangles).
-The finest mesh, built by standard refinement, had $m_9=2049^2 \approx 4.2 \times 10^{6}$ vertices.
+Consider the unilateral 2D Laplacian obstacle problem, Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\}$.
+We consider two problems over square domains $\Omega$, with the different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$ shown in Figure \ref{fig:results:classical}.
+The standard ``ball'' problem has an upper hemisphere obstacle \cite[Chapter 12]{Bueler2021}, but in the ``spiral'' problem \cite[problem 7.1.1]{GraeserKornhuber2009} the obstacle is constructed to produce the coincidence set shown.
+For this experiment we used a coarsest regular-triangulation mesh with $m_0=41$ nodes (64 triangles).
+The finest mesh had $m_{10} \approx 3.4 \times 10^7$ nodes ($6.7\times 10^7$ triangles).
 As our smoother we applied a single RS Newton iteration (Section \ref{sec:implementation}), with three (fixed) IC-preconditioned CG iterations to approximately solve the arising linear systems.
@@
 The initial iterate was set to $w=\max\{0,\psi\}$.
-\pef{For a fair comparison with the spiral problem in \cite{GraeserKornhuber2009}, I think we should set rtol = 1e-12}
 
 Table \ref{tab:results:classical} shows the resulting iteration counts.
-When the RS Newton smoother is applied by itself as a solver (``RS only''), as expected it shows rapidly-increasing iterations, roughly proportional to resolution.
-By contrast, V-cycle iterations increase slowly, perhaps logarithmically with the resolution.
-For F-cycles we report the number of V-cycles after the ramp (Section \ref{sec:implementation}); a constant number solves the problem.
+When RS Newton is applied by itself as a solver, as expected it shows rapidly-increasing iterations, roughly proportional to resolution.
+By contrast, V-cycle iterations increase slowly, logarithmically with the resolution; note this experiment over-solves, with \texttt{rtol} $= 10^{-12}$ in stopping criterion \eqref{eq:stoppingcriterion}.
+For F-cycles, using \texttt{rtol} $= 10^{-12}$ to reach discretization error, we report the number of V-cycles after the ramp (Section \ref{sec:implementation}); a constant number solves the problem.
 Comparing the ball and spiral results, the complexity of the coincidence set or free boundary does not appear to strongly influence FASCD performance.
-For the spiral results, note that \cite{GraeserKornhuber2009} reports more than 30 iterations of all their tested methods were required to achieve similar tolerance for $m_J \approx 5\times 10^5$.
+For the spiral results, note that for $m_J \approx 5\times 10^5$ resolution, \cite{GraeserKornhuber2009} report FIXME.
 
 Each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+IC) iterations.
-Thus for the F-cycle results, where three iterations suffice at all resolutions, we see close to textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  Work of less than ten residual evaluations is often mentioned as the standard.} despite the low solution regularity.
+Thus the F-cycle results, where FIXME iterations suffice at all resolutions, yield close to textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  Work of less than ten residual evaluations is often mentioned as the standard.} despite the low solution regularity.
 \end{example}
@@
 % see git@bitbucket.org:pefarrell/fascd.git: examples/drape.py
-% and https://github.com/bueler/mcd-extended: paper/results/drape.sh|txt
+% and https://github.com/bueler/mcd-extended: paper/results/drape_{nofas,fascd}.sh|txt
 \begin{table}[ht]
@@
 \midrule
- 1 &    $5^2$ &  1 &  1 &  1  &  1 &  1 &  1 \\
- 2 &    $9^2$ &  2 &  1 &  2  &  2 &  1 &  2 \\
- 3 &   $17^2$ &  2 &  2 &  3  &  3 &  2 &  4 \\
- 4 &   $33^2$ &  3 &  2 &  6  &  3 &  2 &  7 \\
- 5 &   $65^2$ &  3 &  2 &  8  &  4 &  3 &  9 \\
- 6 &  $129^2$ &  4 &  2 & 17  &  4 &  3 & 16 \\
- 7 &  $257^2$ &  4 &  2 & 33  &  4 &  3 & 27 \\
- 8 &  $513^2$ &  5 &  2 & 65  &  5 &  3 & 48 \\
- 9 & $1025^2$ &  5 &  2 & \XX &  5 &  3 & \XX \\
-10 & $2049^2$ &  5 &  2 & \XX &  5 &  3 & \XX \\
+ 1 &             $41$ &  1 & FIXME  &  1  &  1 &   &  2 \\
+ 2 &            $145$ &  3 &   &  1  &  4 &   &  4 \\
+ 3 &            $545$ &  6 &   &  3  &  5 &   &  6 \\
+ 4 & $2.1\times 10^3$ &  7 &   &  6  &  7 &   &  7 \\
+ 5 & $8.3\times 10^3$ &  9 &   & 13  &  9 &   & 11 \\
+ 6 & $3.3\times 10^4$ & 11 &   & 23  & 10 &   & 20 \\
+ 7 & $1.3\times 10^5$ & 11 &   & 46  & 11 &   & 35 \\
+ 8 & $5.3\times 10^5$ & 12 &   & 90  & 12 &   & 67 \\
+ 9 & $2.1\times 10^6$ & 13 &   & \XX & 13 &   & \XX \\
+10 & $8.4\times 10^6$ & 13 &   & \XX & 14 &   & \XX \\
+11 & $3.4\times 10^7$ & 14 &    & \XX & FIXME   &    & \XX \\
 \bottomrule
@@
 \bigskip
-\caption{Iterations for FASCD V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}.  Smoother-only runs marked \XX\xspace were not attempted.}
+\caption{Iterations for FASCD V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}.  Runs marked \XX\xspace were not attempted.}
 \label{tab:results:classical}

### 2023-07-26 paper/results/drape_fascd.sh

@@ -1,17 +1,19 @@
 # generate FASCD results for Example 8.1 in paper
+# warning: the level 10,11 runs assume decent memory on you machine
+# note: these runs are serial so as to fix the smoother
 
 for PROB in ball spiral; do
     echo ""
     echo "******** PROB=${PROB}, V-cycles, tight tolerance (over-solving) *******"
-    for LEVS in 1 2 3 4 5 6 7 8 9 10; do
+    for LEVS in 1 2 3 4 5 6 7 8 9 10 11; do
         python3 drape.py -prob $PROB -levs $LEVS -fascd_cycle_type multiplicative -fascd_rtol 1.0e-12
     done
 done
 
 for PROB in ball spiral; do
     echo ""
-    echo "******** PROB=${PROB}, F-cycles, to discretization error *******"
-    for LEVS in 1 2 3 4 5 6 7 8 9 10; do
+    echo "******** PROB=${PROB}, F-cycles, to discretization error, with default rtol 1.0e-8 *******"
+    for LEVS in 1 2 3 4 5 6 7 8 9 10 11; do
         python3 drape.py -prob $PROB -levs $LEVS -fascd_cycle_type full
     done
 done

### 2023-07-26 paper/fascd.tex
WIP: update Example 8.1 results

@@
 For $P_1$ and $Q_1$ elements, Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}.
-The numerical results in this Section follow the Section \ref{sec:vi} examples, with a focus on the scaling of FASCD iterations as the mesh resolution increases.
+The numerical results in this Section, which follow the Section \ref{sec:vi} examples, focus on the scaling of FASCD iterations as the mesh resolution increases.
 Our primary observations are that each V-cycle greatly reduces the residual norm, and that a few F-cycles solve most examples to within discretization error.
@@
 \begin{example} \label{ex:results:classical}
-Consider the unilateral 2D Laplacian obstacle problem, Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\}$.
-We consider two problems over square domains $\Omega$, with the different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$ shown in Figure \ref{fig:results:classical}.
-The standard ``ball'' problem has an upper hemisphere obstacle \cite[Chapter 12]{Bueler2021}, but in the ``spiral'' problem \cite[problem 7.1.1]{GraeserKornhuber2009} the obstacle is constructed to produce the coincidence set shown.
-For this experiment we used a coarsest regular-triangulation mesh with $m_0=41$ nodes (64 triangles).
-The finest mesh had $m_{10} \approx 3.4 \times 10^7$ nodes ($6.7\times 10^7$ triangles).
-As our smoother we applied a single RS Newton iteration (Section \ref{sec:implementation}), with three (fixed) IC-preconditioned CG iterations to approximately solve the arising linear systems.
-The coarse solver was RS$+$LU, iterated to convergence.
-Tolerances \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$ were used in stopping criterion \eqref{eq:stoppingcriterion}.
+Consider the classical unilateral 2D Laplacian obstacle problem, Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\}$.
+We consider two problems over square domains $\Omega$, with different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$ (Figure \ref{fig:results:classical}).
+The standard ``ball'' problem has an upper hemisphere obstacle \cite[Chapter 12]{Bueler2021}, but in the ``spiral'' problem the obstacle is constructed to produce the coincidence set shown \cite[problem 7.1.1]{GraeserKornhuber2009}.
+For this experiment we used a coarsest regular-triangulation mesh with $m_0=41$ nodes.
+The finest mesh had $m_{10} \approx 3.4 \times 10^7$ nodes.
+The smoother applied a single RS Newton iteration (Section \ref{sec:implementation}), with three IC-preconditioned CG iterations.
 The initial iterate was set to $w=\max\{0,\psi\}$.
@@
 Table \ref{tab:results:classical} shows the resulting iteration counts.
-When RS Newton is applied by itself as a solver, as expected it shows rapidly-increasing iterations, roughly proportional to resolution.
-By contrast, V-cycle iterations increase slowly, logarithmically with the resolution; note this experiment over-solves, with \texttt{rtol} $= 10^{-12}$ in stopping criterion \eqref{eq:stoppingcriterion}.
-For F-cycles, using \texttt{rtol} $= 10^{-12}$ to reach discretization error, we report the number of V-cycles after the ramp (Section \ref{sec:implementation}); a constant number solves the problem.
-Comparing the ball and spiral results, the complexity of the coincidence set or free boundary does not appear to strongly influence FASCD performance.
-For the spiral results, note that for $m_J \approx 5\times 10^5$ resolution, \cite{GraeserKornhuber2009} report FIXME.
-
-Each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+IC) iterations.
-Thus the F-cycle results, where FIXME iterations suffice at all resolutions, yield close to textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  Work of less than ten residual evaluations is often mentioned as the standard.} despite the low solution regularity.
+As expected, RS Newton applied by itself as a solver shows rapidly-increasing iterations, roughly proportional to resolution, but V-cycle iterations increase slowly, logarithmically with the resolution.
+These experiments over-solve, with \texttt{rtol} $=$ \texttt{atol} $= 10^{-12}$ in stopping criterion \eqref{eq:stoppingcriterion}.
+The complexity of the coincidence set or free boundary does not appear to strongly influence FASCD V-cycle performance.
+For the spiral problem \cite{GraeserKornhuber2009} reports that for $m_J \approx 5\times 10^5$ resolution FIXME.
+FIXME: asymp rate figure?
+
+For F-cycles we instead used \texttt{rtol} $= 10^{-8}$, sufficient to reach discretization error, and we report the number of V-cycles after the ramp (Section \ref{sec:implementation}).
+A small, constant number solves the problem, though spiral needs more V-cycles to clean-up error remaining after the ramp.
+Since each V-cycle does the work of about eight finest-level Krylov iterations, the ball results show nearly textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  Work of less than ten residual evaluations is often mentioned as the standard.} at the highest resolutions despite the low solution regularity.
 \end{example}
@@
 \multirow{2}{*}{\emph{levels}} & \multirow{2}{*}{$m_J$} & \multicolumn{3}{c}{\,\emph{ball}} & \multicolumn{3}{c}{\,\emph{spiral}} \\ \cmidrule(lr){3-5} \cmidrule(lr){6-8}
-   &          & V-cycle & F-cycle & RS only & V-cycle & F-cycle & RS only \\
+   &                  & RS only & V-cycle & F-cycle & RS only & V-cycle & F-cycle \\
 \midrule
- 1 &             $41$ &  1 & FIXME  &  1  &  1 &   &  2 \\
- 2 &            $145$ &  3 &   &  1  &  4 &   &  4 \\
- 3 &            $545$ &  6 &   &  3  &  5 &   &  6 \\
- 4 & $2.1\times 10^3$ &  7 &   &  6  &  7 &   &  7 \\
- 5 & $8.3\times 10^3$ &  9 &   & 13  &  9 &   & 11 \\
- 6 & $3.3\times 10^4$ & 11 &   & 23  & 10 &   & 20 \\
- 7 & $1.3\times 10^5$ & 11 &   & 46  & 11 &   & 35 \\
- 8 & $5.3\times 10^5$ & 12 &   & 90  & 12 &   & 67 \\
- 9 & $2.1\times 10^6$ & 13 &   & \XX & 13 &   & \XX \\
-10 & $8.4\times 10^6$ & 13 &   & \XX & 14 &   & \XX \\
-11 & $3.4\times 10^7$ & 14 &    & \XX & FIXME   &    & \XX \\
+ 1 &             $41$ &   1 &  1 &  1 &   2 &  1 &  1 \\
+ 2 &            $145$ &   1 &  3 &  2 &   4 &  4 &  2 \\
+ 3 &            $545$ &   3 &  6 &  3 &   6 &  5 &  3 \\
+ 4 & $2.1\times 10^3$ &   6 &  7 &  3 &   7 &  7 &  4 \\
+ 5 & $8.3\times 10^3$ &  13 &  9 &  4 &  11 &  9 &  4 \\
+ 6 & $3.3\times 10^4$ &  23 & 11 &  5 &  20 & 10 &  5 \\
+ 7 & $1.3\times 10^5$ &  46 & 11 &  4 &  35 & 11 &  5 \\
+ 8 & $5.3\times 10^5$ &  90 & 12 &  4 &  67 & 12 &  5 \\
+ 9 & $2.1\times 10^6$ & \XX & 13 &  3 & \XX & 13 &  6 \\
+10 & $8.4\times 10^6$ & \XX & 13 &  2 & \XX & 14 &  6 \\
+11 & $3.4\times 10^7$ & \XX & 14 &  2 & \XX & 18 &  6 \\
 \bottomrule
@@
 \bigskip
-\caption{Iterations for FASCD V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}.  Runs marked \XX\xspace were not attempted.}
+\caption{Iterations for FASCD V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}; see text regarding tolerances.  Runs marked \XX\xspace were not attempted.}
 \label{tab:results:classical}
@@
 \begin{example}  \label{ex:results:plap}
-We again consider Example \ref{ex:plaplacian} with a unilateral constraint, but now for the nonlinear $p$-Laplacian with both fast ($p=1.5$) and degenerate ($p=4.0$) diffusivities.\footnote{``Fast'' and ``degenerate'' refer to $|\grad u|^{p-2}\to \infty,0$, respectively; ellipticity is lost in the latter case.
-Well-posedness and ($p$-dependent) solution regularity for $p$-Laplacian obstacle problems is considered in \cite{ChoeLewis1991}.} We consider the problem in one dimension for ease of visualization.
-This example reveals the sometimes strong contrast between $V(1,0)$ and $V(0,1)$ results (\pr{up}, \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively), and it emphasizes the need for robust smoothers.
-
+We again consider Example \ref{ex:plaplacian} with a unilateral constraint, but now for nonlinear $p$-Laplacians with both fast ($p=1.5$) and degenerate ($p=4.0$) diffusivities.
+``Fast'' and ``degenerate'' refer to $|\grad u|^{p-2}\to \infty,0$, respectively; ellipticity is lost in the latter case.
+Well-posedness and solution regularity for $p$-Laplacian obstacle problems is considered in \cite{ChoeLewis1991}.
+
+This example reveals the strong contrast between $V(1,0)$ and $V(0,1)$ results (\pr{up}, \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively), and the need for robust smoothers when the operator is nonlinear.
 Set $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, $\mathcal{K} = \{v \ge \psi\} \subset W^{1,p}(\Omega)$, and $\ip\ell v = \int_\Omega g v$, where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$, in VI \eqref{eq:vi}.
-The exact solution is easily computed for any $p$; see Figure \ref{fig:results:plap} and the documentation of the code associated to this paper.
+The exact solution is easily computed for any $p$; see Figure \ref{fig:results:plap} and the accompanying this paper.
 For this experiment we used a coarsest mesh of $m_0=7$ vertices and a finest mesh with $m_9=3073$.
@@
 
-\elb{beam example?}
-
 \section{Discussion and outlook} \label{sec:discussion}

### 2023-07-26 paper/fascd.tex
implement figure reductions/removals, and some text edits; do more after finalizing results

@@
 \pseudoset{%
-left-margin=15mm,%
+left-margin=6mm,%
 topsep=5mm,%
@@
 % Sets running headers as well as PDF title and authors
-\headers{FAS for nonlinear variational inequalities}{E. Bueler and P. E. Farrell}
+\headers{FASCD for nonlinear variational inequalities}{E. Bueler and P. E. Farrell}
 
@@
 We extend the CD idea by exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
-We demonstrate that FASCD V-cycles exhibit excellent mesh-independent convergence rates.
-Furthermore, when a reduced-space active set method is applied as a smoother, with work proportional to the number of unknowns on a given mesh level, FASCD F-cycles (full multigrid) are optimal solvers.
+When a reduced-space active set method is applied as a smoother, with work proportional to the number of unknowns on a given mesh level, FASCD V-cycles exhibit mesh-independent convergence rates and F-cycles (full multigrid) are optimal solvers.
 The example problems include differential operators which are symmetric linear, nonsymmetric linear, and nonlinear, in unilateral and bilateral VI problems.
-Our examples demonstrate very good convergence even on problems where other multilevel VI solvers struggle.
 \end{abstract}
@@
 
-The constraint decomposition (CD) methods of Tai \cite{Tai2003} are designed to solve coercive variational inequality (VI) problems arising as the Karush--Kuhn--Tucker optimality conditions for the minimization of a convex functional over a convex set.
-These methods were shown to converge for multilevel and domain decompositions in a finite element (FE) context.
-In the multilevel case, this CD method has almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}, \cite[Theorem 4.6]{GraeserKornhuber2009}.
+The constraint decomposition (CD) methods of Tai \cite{Tai2003} are designed to solve variational inequality (VI) problems arising as the Karush--Kuhn--Tucker optimality conditions for the minimization of a convex functional over a convex set.
+In a finite element (FE) context, these methods converge for multilevel and domain decompositions.
+The multilevel CD method has almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}, \cite[Theorem 4.6]{GraeserKornhuber2009}.
 
 Here we extend Tai's multilevel CD method by replacing the coarser-level corrections with a full approximation scheme (FAS; \cite{Brandt1977,Bruneetal2015}) approach.
-Specifically, our \emph{full approximation scheme constraint decomposition} (FASCD) method extends the algorithms of \cite{GraeserKornhuber2009,Tai2003} in the following directions:
-\pef{These aren't all really extensions. Maybe we could say `enjoys the following advantages'? We distinguish it from prior work below, so this section does not need to do that.}
+Specifically, our \emph{full approximation scheme constraint decomposition} (FASCD) method enjoys the following advantages:
+
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
@@
 \item The FAS coarse corrections are adapted for general VI problems. The method requires neither an objective function nor linearity of the residual.
-\item Box constraints (i.e.~both upper and lower obstacles) are allowed.
-%\item As with the original CD method \cite{Tai2003}, but unlike some implementations \cite[for example]{GraeserKornhuber2009}, and semismooth methods generally \cite{BensonMunson2006,Ulbrich2011}, the method can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
-%\pef{I confess I find this explicit calling out of other algorithms a bit uncomfortable/too much. I wonder if the following is better?}
+\item Box constraints, i.e.~both upper and lower obstacles, are allowed.
 \item The method is strictly admissible, and so can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
@@
 \item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
-\item When a (mesh-dependent) reduced-space active set method is used as relaxation, we demonstrate that F-cycle FASCD is an optimal solver for the different VI examples in Section \ref{sec:results}.
-The work of the smoother is proportional to the number of unknowns on a given level because the number of Newton and Krylov iterations is fixed.
+\item When a (mesh-dependent) reduced-space active set method is used as relaxation, we demonstrate that F-cycles are an optimal solver for the examples in Section \ref{sec:results}; smoother effort is proportional to the number of unknowns on a given level because the number of Newton and Krylov iterations is fixed.
 \end{enumerate}
 
-Prior multilevel algorithms for VIs have various substantial limitations compared to the FASCD framework.
-The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} is apparently limited to linear complementarity problems and projected (truncated) Gauss--Seidel smoothers.
+Prior multilevel algorithms for VIs have limitations compared to the FASCD framework.
+The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} applies to linear complementarity problems and projected (truncated) Gauss--Seidel smoothers.
 While FASCD reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed (Appendix \ref{app:reductions}), PFAS reduces to linear multigrid.
@@
 The latter property is improved by reformulation using defect constraints and monotone injection operators (see \cite[Algorithm 4.7]{GraeserKornhuber2009}, or Section \ref{sec:cdmultilevel} herein).
-The monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be formulated only with Gauss--Seidel type smoothers, and applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.
+The monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be formulated only with Gauss--Seidel type smoothers, and may have been applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.
 Unlike the fastest variants of monotone multigrid, no zeroing of the basis functions associated with the active set (truncation) is employed in our method, both simplifying the implementation and allowing the coarse grid corrections to update the estimate of the active set on the fine grid.
-Numerical evidence indicates that the FAS construction of coarse grid corrections is useful for this purpose, even when the coarse grids are too coarse to usefully resolve the active set.
-\pef{Insert comparison against the spiral problem once we have set rtol = 1e-12}
-A limitation of FASCD is that, as formulated, it only solves bound-constrained VI problems.
+Numerical evidence indicates that our coarse grid corrections are useful for this purpose, even when the coarse grids are too coarse to usefully resolve the active set.
+On a problem for which the truncated Newton/monotone multigrid methods were designed \cite[problem 7.1.1]{GraeserKornhuber2009}, the performance of FASCD is at least as good (Example \ref{ex:results:classical}).
+
+A known limitation of FASCD is that, as formulated, it only solves bound-constrained VI problems.
 
@@
 Specifically, we see this performance for classical (unilateral, Laplacian) obstacle problems, $p$-Laplacian obstacle problems, and box-constrained advection-diffusion equations.
-Furthermore, these results appear to be insensitive to the length or structure of the free boundary, i.e.~the geometric complexity of the coincidence set.
+Furthermore, these results appear to be relatively insensitive to the geometric complexity of the coincidence set and free boundary.
 
@@
 
-Descending coarser levels, $j=J-1,\dots,1,0$, we next apply the monotone injections to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:
+Descending coarser levels, $j=J-1,\dots,1,0$, we next apply the monotone injections to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:\footnote{The definition of $\underline{\chi}^j$ matches that in Section 4 of \cite{GraeserKornhuber2009}; see (4.22) and Figure 4.1 therein.}
 \begin{equation}
@@
 \end{equation}
-A unilateral $J=4$ example of LDCs $\underline{\chi}^j$ is shown in Figure \ref{fig:chiphilevels}.
 The LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if another element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$; similar comments apply to $\overline{\chi}^{j}$.
 Note that for Algorithm \ref{alg:fascd} below, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not required.
-
-%REGENERATE using https://github.com/bueler/mg-glaciers:
-%$ cd mg-glaciers/py/
-%$ [modify decomposition() inside visualize.py to turn off axes and print \gamma for obstacle]
-%$ ./obstacle.py -J 5 -jcoarse 1 -random -randommodes 8 -diagnostics -o defect.pdf
-%fine level 5 (m=63): using 20 V(1,0) cycles -> 38.750 WU
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.65\textwidth]{fixfigs/chiphilevels.pdf}
-\caption{Unilateral LDCs $\underline{\chi}^j$ are generated from $\underline{\chi}^J = \underline{\gamma}^J - w^J$ using maximum injection: $\underline{\chi}^j = \maxR \underline{\chi}^{j+1}$.}
-\label{fig:chiphilevels}
-\end{figure}
 
@@
 
-Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion}, which are special to multilevel schemes based on defect constraints, hold regardless of whether a CD operator $\Pi_i$ can be constructed.
-Correspondingly, there is another basic iteration that does not need the $\Pi_i$ operators, namely an incomplete variant of \pr{cd-mult} (Section \ref{sec:cd}).
-Application \pr{icd-tele} below requires $w^J\in\mathcal{K}^J$, property \eqref{eq:fe:finestlevelequivalent}, and incomplete CD \eqref{eq:fe:upwardsuminclusion}.
+Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in Algorithm \ref{alg:fascd} below.
+These inclusions are special to multilevel schemes based on defect constraints, and they hold regardless of whether CD operators $\Pi_i$ can be constructed.
+A different ``basic'' iteration, an incomplete variant of \pr{cd-mult}, does not need $\Pi_i$ operators.
+Application of \pr{icd-tele} requires $w^J\in\mathcal{K}^J$, property \eqref{eq:fe:finestlevelequivalent}, and \eqref{eq:fe:upwardsuminclusion}.
 
@@
 
-While inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in Algorithm \ref{alg:fascd} below, a convergence analysis extending the arguments of \cite{GraeserKornhuber2009,Tai2003} to VIs which are not of minimization type, or to box constraints, would seem to require full CDs.
+A convergence analysis extending the arguments of \cite{GraeserKornhuber2009,Tai2003} to VIs which are not of minimization type, or to box constraints, would seem to require full CDs.
+On the other hand, the observation that up-smoothing is more efficient in a multilevel CD context, illustrated by Example \ref{ex:results:plap} in particular, seems to be new.\footnote{Compare comments on $V(1,1)$ cycles in \cite{GraeserKornhuber2009} and \cite{Tai2003}.}
 
@@
 By nesting \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$.\footnote{Nonetheless we use prolongation $P$ to indicate where the computer representation changes.}  The corrections are constructed, inductively, so that iterates remain admissible at every level.
-Suppose $y^i \in \mathcal{D}^i$, for $i=J,\dots,j+1$, are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).
+Suppose $y^i \in \mathcal{D}^i$, for $i=J,\dots,j+1$, are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdcycles}).
 By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.
@@
 \centering
-\input{tikz/vcycle.tex}
-\caption{The FASCD V-cycle computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.}
-\label{fig:fascdvcycle}
+\mbox{\input{tikz/vcycle.tex} \quad
+\input{tikz/fcycle.tex}}
+\caption{Left: The FASCD V-cycle Algorithm \ref{alg:fascd} computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.
+Right: F-cycle Algorithm \ref{alg:fascd-fcycle} initially restricts and injects the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray), then it prolongs and truncates (doubled edges) to create an admissible initial iterate on the next-finer level, and then it applies V-cycles.}
+\label{fig:fascdcycles}
 \end{figure}
@@
 
-However, an F-cycle (Figure \ref{fig:fcycle}), also called a \emph{full multigrid cycle}, is even better.
+However, an F-cycle (Figure \ref{fig:fascdcycles}), also called a \emph{full multigrid cycle}, is even better.
 The well-known principle of such a cycle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate.
@@
 (In the VI context one must also truncate to the bounds after the prolongation; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fcycle}.)  During this ``ramp'' one approximately solves on each level by using a fixed number of V-cycles, each down to the coarsest mesh, and then one does full-depth V-cycles from the finest level, to convergence.
-
-\begin{figure}[h]
-\centering
-\input{tikz/fcycle.tex}
-\caption{F-cycle Algorithm \ref{alg:fascd-fcycle} initially restricts and injects the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray).
-Then it prolongs and truncates (doubled edges) to create an admissible initial iterate on the next-finer level.
-Solution on each level is by V-cycles.
-See Figure \ref{fig:fascdvcycle} for the other symbols.}
-\label{fig:fcycle}
-\end{figure}
 
@@
 The complexity of the coincidence set or free boundary does not appear to strongly influence FASCD V-cycle performance.
-For the spiral problem \cite{GraeserKornhuber2009} reports that for $m_J \approx 5\times 10^5$ resolution FIXME.
-FIXME: asymp rate figure?
+For the spiral problem \cite{GraeserKornhuber2009} reports that for $m_J \approx 5\times 10^5$ resolution \elb{describe GK results}
+
+\elb{asymp rate figure?}
 
@@
 \begin{example}  \label{ex:results:plap}
-We again consider Example \ref{ex:plaplacian} with a unilateral constraint, but now for nonlinear $p$-Laplacians with both fast ($p=1.5$) and degenerate ($p=4.0$) diffusivities.
-``Fast'' and ``degenerate'' refer to $|\grad u|^{p-2}\to \infty,0$, respectively; ellipticity is lost in the latter case.
+We again consider Example \ref{ex:plaplacian} with a unilateral constraint, but now for the nonlinear $p=1.5$-Laplacian, a ``fast'' diffusion with diffusivity limit $D=|\grad u|^{p-2}\to \infty$ as $|\grad u|\to 0$.
+Note that Example \ref{ex:results:sia} applies a doubly-nonlinear version of a $p=4$ ``degenerate'' diffusion operator.
 Well-posedness and solution regularity for $p$-Laplacian obstacle problems is considered in \cite{ChoeLewis1991}.
@@
 Set $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, $\mathcal{K} = \{v \ge \psi\} \subset W^{1,p}(\Omega)$, and $\ip\ell v = \int_\Omega g v$, where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$, in VI \eqref{eq:vi}.
-The exact solution is easily computed for any $p$; see Figure \ref{fig:results:plap} and the accompanying this paper.
+The exact solution is easily computed for any $p$; see the codes accompanying this paper.
 For this experiment we used a coarsest mesh of $m_0=7$ vertices and a finest mesh with $m_9=3073$.
-The default smoother uses three (fixed) iterations of RS Newton, with LU solution of the arising linear systems, an $O(m_j)$ smoother in 1D.
+The default smoother uses three (fixed) iterations of RS Newton, with LU solution of the arising linear systems.
 The coarse solver is the same but iterated to convergence.
 The initial iterate was $w=\psi$, and \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$.
-Figure \ref{fig:imagesvcycle} illustrates the major steps of an example V-cycle, showing each of the coarse-correction VI subproblems and solutions.
-
-%REGENERATE using elbueler/plapfigs branch of git@bitbucket.org:pefarrell/fascd.git, in examples/
-% $ python3 plap1d.py -p 1.5 -figures -levs 6 -fascd_cycle_type full && mv solution.pdf plap1d1p5.pdf
-% $ python3 plap1d.py -p 4.0 -figures -levs 6 -fascd_cycle_type full && mv solution.pdf plap1d4p0.pdf
-\begin{figure}[ht]
-\centering
-\includegraphics[width=0.45\textwidth]{fixfigs/plap1d1p5.pdf} \quad
-\includegraphics[width=0.45\textwidth]{fixfigs/plap1d4p0.pdf}
-\caption{Solutions $u\ge \psi$ of the $p$-Laplacian VI in Example \ref{ex:results:plap}, for $p=1.5$ (left) and $p=4.0$ (right).}
-\label{fig:results:plap}
-\end{figure}
+Figure \ref{fig:imagesvcycle} illustrates the major steps of an example V-cycle, showing each of the coarse-correction VI subproblems and solutions; the result $\tilde w^2$ is visually close to the exact solution.
 
@@
 \input{tikz/imagesvcycle.tex}
-\caption{Visualization of one FASCD V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of $m_0=5$ vertices.
+\caption{Visualization of one FASCD V-cycle iteration for $p$-Laplacian Example \ref{ex:results:plap}, with $J=3$ and a coarsest mesh of $m_0=5$ vertices.
 Solutions (i.e.~$w^j$, $y^j$, $z^j$) are lines, obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots, and the zero function is dotted.
@@
 
-Iteration results are shown in Table \ref{tab:results:fastplap1d} for the fast diffusion $p=1.5$ case.
+Iteration results are shown in Table \ref{tab:results:fastplap1d}.
 We see that $V(1,0)$ cycles, essentially Algorithm 4.7 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.
-By contrast, $V(0,1)$ cycles, using only up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are just as good as the default $V(1,1)$ cycles ($=$ Algorithm \ref{alg:fascd}).
+By contrast, $V(0,1)$ cycles, up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are as good as the default $V(1,1)$ cycles ($=$ Algorithm \ref{alg:fascd}).
 F-cycles are even more efficient except when they fail entirely, but these failures arise from singular diffusivity.
@@
 
-Iteration results for the (unregularized) degenerate diffusion $p=4.0$ case are in Table \ref{tab:results:degenerateplap1d}.
-V-cycles using the default settings, specifically three fixed RS Newton iterations, do not converge on the finer meshes; the semi-smooth NCP residual norms \eqref{eq:stoppingcriterion} are unbounded in the failed runs.
-Visualization shows an instability in which oscillations in the restricted residuals explode.
-However, either a slight increase in smoother power, to four RS Newton iterations, or improved finest-level initial iterates through use of an F-cycle, stops the instability.
-Even for the three-iteration smoother, a single F-cycle suffices at every resolution.
-
-The $p=1.5$ and $p=4$ cases give different convergence rates in the $\|\cdot\|_\infty$ norm, namely at rates $O(h^{2.005})$ and $O(h^{1.420})$, respectively; compare \cite{ChoeLewis1991}.
+\elb{check if regularization is even necessary when you re-run with \texttt{-fascd\_levels\_snes\_vi\_zero\_tolerance 1.0e-12}}
+
+Note that convergence is at rate $O(h^{2.005})$; compare \cite{ChoeLewis1991}.
 \end{example}
@@
 \bigskip
-\caption{Iterations for V- and F-cycles on the $p=1.5$ problem in Example \ref{ex:results:plap}.  NC $=$ did not converge after 50 iterations.  See text regarding F$_\eps$ and error.}
+\caption{Iterations for V- and F-cycles in Example \ref{ex:results:plap}.  NC $=$ did not converge after 50 iterations.  See text regarding F$_\eps$.}
 \label{tab:results:fastplap1d}
-\end{table}
-
-% see git@bitbucket.org:pefarrell/fascd.git: examples/plap1d.py
-% and https://github.com/bueler/mcd-extended: paper/results/plap1d.sh|txt
-\begin{table}[ht]
-\centering
-\begin{tabular}{lr@{\hskip 7mm}c@{\hskip 4mm}c@{\hskip 4mm}c@{\hskip 6mm}c}
-\toprule
-\emph{levels} & $m_J$ & V & V$_4$ & F & $\|$error$\|_\infty$ \\
-\midrule
- 1 &    $7$ &  1 &  1 &  1 & $8.7 \times 10^{-2}$ \\
- 2 &   $13$ &  1 &  1 &  1 & $3.8 \times 10^{-2}$ \\
- 3 &   $25$ &  2 &  1 &  1 & $1.6 \times 10^{-2}$ \\
- 4 &   $49$ &  4 &  2 &  1 & $6.3 \times 10^{-3}$ \\
- 5 &   $97$ & NC &  2 &  1 & $2.3 \times 10^{-3}$ \\
- 6 &  $193$ & NC &  1 &  1 & $7.0 \times 10^{-4}$ \\
- 7 &  $385$ & NC &  2 &  1 & $2.0 \times 10^{-4}$ \\
- 8 &  $769$ & NC &  2 &  1 & $1.1 \times 10^{-4}$ \\
- 9 & $1537$ & NC &  2 &  1 & $4.0 \times 10^{-5}$ \\
-10 & $3073$ & NC &  2 &  1 & $1.5 \times 10^{-5}$ \\
-\bottomrule
-\end{tabular}
-\bigskip
-\caption{Iterations for V- and F-cycles on the $p=4.0$ problem in Example \ref{ex:results:plap}.  See text regarding V$_4$ and error.}
-\label{tab:results:degenerateplap1d}
 \end{table}
@@
 
-As a multilevel solver for VI problems, one might describe the new FASCD method as merely a strategy for generating coarser-level problems which, once solved, make helpful admissible additions to a final V-cycle update.
+As a multilevel solver for VI problems, the FASCD method is a strategy for generating coarser-level problems which, once solved, make helpful admissible additions to a V-cycle update.
 The method is smoother-agnostic, and our implementation within the extensible Firedrake \cite{Rathgeberetal2016} and PETSc \cite{Balayetal2023} library framework should allow easy experimentation with other smoothers (projected Gauss--Seidel, semi-smooth Newton, interior-point, etc.), beyond the active-set Newton method \cite{BensonMunson2006} chosen for demonstration here.
 For box-constrained smoothers of Newton-Krylov type one could also add linear geometric or algebraic multigrid preconditioning \cite{Trottenbergetal2001}, instead of our simple choice of incomplete factorization and (in parallel) additive Schwarz preconditioners.
-Thus there is a large smoother and coarse solver space to explore.
-However, the performance of the particular method shown in Section \ref{sec:results} is already excellent.
-
-The observation that up-smoothing is more efficient in a multilevel CD context, which arises because of constraint set distinctions, seems to be new, but compare comments on $V(1,1)$ cycles in \cite{GraeserKornhuber2009,Tai2003}.
-The increased efficiency is particularly shown in Example \ref{ex:results:plap}.
-In summary, the distinction is that construction of constraint sets via monotone injection generates up-smoothing sets which can telescope the down-smoothing sets (Lemma \ref{lem:upwardadmissibility}); we call this an \emph{incomplete CD}.
-This structure can be exploited by a new CD iteration, algorithm \pr{icd-tele} in Section \ref{sec:cdmultilevel}.
-\pef{This last paragraph feels like a flat way to end. Perhaps we could highlight the importance of ICD-TELE in the text, and remove this last paragraph?}
+However, the performance illustrated in Section \ref{sec:results} is already excellent.
 

### 2023-07-27 paper/fascd.tex
Get down to 26 pages

@@
 Specifically, our \emph{full approximation scheme constraint decomposition} (FASCD) method enjoys the following advantages:
-
 \renewcommand{\labelenumi}{\emph{(\roman{enumi})}}
@@
 The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} applies to linear complementarity problems and projected (truncated) Gauss--Seidel smoothers.
-While FASCD reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed (Appendix \ref{app:reductions}), PFAS reduces to linear multigrid.
+While FASCD reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed, PFAS reduces to linear multigrid.
 The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so as stated it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
@@
 On a problem for which the truncated Newton/monotone multigrid methods were designed \cite[problem 7.1.1]{GraeserKornhuber2009}, the performance of FASCD is at least as good (Example \ref{ex:results:classical}).
-
-A known limitation of FASCD is that, as formulated, it only solves bound-constrained VI problems.
+A limitation of FASCD is that, as formulated, it only solves bound-constrained VI problems.
 
@@
 \end{equation}
-With this operator, consider VI \eqref{eq:vi} with $\ell = 0$ for any closed and convex $\cK \subset \cV$.
-The interior condition, i.e.~PDE strong form, is the following linear advection-diffusion equation
+With this operator, consider VI \eqref{eq:vi} with $\ell = 0$ for any closed and convex $\cK \subset \cV$. The interior condition in strong form is the following linear advection-diffusion equation
 \begin{equation}
@@
 The V-cycle applies smoothers in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.
-It extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}, and in fact it reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).
+It extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}, and in fact it reduces to FAS for PDEs when inequality constraints are removed.
 
@@
 \centering
-\mbox{\input{tikz/vcycle.tex} \quad
-\input{tikz/fcycle.tex}}
+\begin{tabular}{cc}
+\input{tikz/vcycle.tex}
+&
+\input{tikz/fcycle.tex}
+\end{tabular}
 \caption{Left: The FASCD V-cycle Algorithm \ref{alg:fascd} computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.
@@
 \end{equation}
-The first four cases assume $x_p \notin \partial_D\Omega$.  For PDEs one uses only the first and last cases.
+The first four cases assume $x_p \notin \partial_D\Omega$.  For PDEs only the first and last apply.
 
@@
 The generalized minimum residual (GMRES) with incomplete LU (ILU) preconditioning is used in other cases.
-The coarse solver is also RS Newton, but iterated to convergence using a direct linear solver.
+The coarse solver is also RS Newton, but iterated to convergence using the MUMPS sparse direct linear solver~\cite{Amestoy2001}.
 
@@
 \begin{example}  \label{ex:results:advdiff}
-Consider the advection-diffusion VI problem in Example \ref{ex:advectiondiffusion} with $\Omega=(-1,1)^d$, $d=2,3$, diffusivity $\eps=0.1$ in equation \eqref{eq:advectiondiffusionstrong}, and zero Dirichlet boundary conditions.
-We made the 2D problem satisfy the coercivity hypothesis with a divergence-free velocity field $\bX = (7+5y,-5x)$, but in 3D, for variety's sake, we set $\bX = (7+5y,-5x,2z)$ so $\grad\cdot\bX \ne 0$.
+Consider the advection-diffusion VI problem in Example \ref{ex:advectiondiffusion} with $\Omega=(-1,1)^d$, $d=2,3$, diffusivity $\eps=0.1$ in \eqref{eq:advectiondiffusionstrong}, and zero Dirichlet boundary conditions.
+We made the 2D problem satisfy the coercivity hypothesis with a divergence-free velocity field $\bX = (7+5y,-5x)$, but in 3D set $\bX = (7+5y,-5x,2z)$ so $\grad\cdot\bX \ne 0$.
 The chosen discontinuous source term $\phi \in L^\infty(\Omega)$ satisfies $\phi\ge 0$ in half of the domain, with positive circular patches, with negative values in the other half; details are in the associated code.
@@
 A coarse mesh with $m_0=16^d$ vertices minimally-resolved the coincidence set.
-The smoother was two (fixed) RS Newton iterations, with three preconditioned GMRES iterations to approximately solve the arising linear systems; the preconditioning was additive Schwarz method (ASM) with incomplete factorization (ILU) on each process.
+The smoother was two (fixed) RS Newton iterations, with three preconditioned GMRES iterations for the arising linear systems; the preconditioning was additive Schwarz method (ASM) with incomplete factorization (ILU) on each process.
 The coarse solver was RS Newton and LU, iterated to convergence.
@@
 
-\appendix
-\section{Reduction to FAS multigrid for PDEs} \label{app:reductions}
-
-When all inequality constraints are removed the FASCD V-cycle Algorithm \ref{alg:fascd} reduces to the classical FAS multigrid V-cycle for nonlinear PDEs \cite{Brandt1977,Trottenbergetal2001}, which itself generalizes linear multigrid.
+%\appendix
+%\section{Reduction to FAS multigrid for PDEs} \label{app:reductions}
+%
+%When all inequality constraints are removed the FASCD V-cycle Algorithm \ref{alg:fascd} reduces to the classical FAS multigrid V-cycle for nonlinear PDEs \cite{Brandt1977,Trottenbergetal2001}, which itself generalizes linear multigrid.
 In fact, suppose we replace $\underline{\gamma}^J,\overline{\gamma}^J$ in definition \eqref{eq:fe:fineconstraintset} with $-\infty,+\infty$, respectively.
@@
 Since $v-u^J\in\mathcal{V}_0^J$ is now arbitrary we can rewrite \eqref{eq:fe:vi} as the FE discretization of strong-form PDE $f(u)=\ell$: find $u^J \in \mathcal{V}^J$ such that
-\begin{equation}
-\ip{f^J(u^J)}{v} = \ip{\ell^J}{v} \qquad \text{for all } v\in \mathcal{V}_0^J. \label{eq:app:fas:pde}
-\end{equation}
-
-The downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets are also enlarged to vector spaces, and the corrections $y^j$ and $z^j$ are unconstrained; lines \ref{line:vcyclegenchifinest}, \ref{line:vcyclegenchi}, and \ref{line:vcyclegenphi} in Algorithm \ref{alg:fascd} can be removed.
+%\begin{equation}
+%\ip{f^J(u^J)}{v} = \ip{\ell^J}{v} \qquad \text{for all } v\in \mathcal{V}_0^J. \label{eq:app:fas:pde}
+%\end{equation}
+%
+%The downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets are also enlarged to vector spaces, and the corrections $y^j$ and $z^j$ are unconstrained; lines \ref{line:vcyclegenchifinest}, \ref{line:vcyclegenchi}, and \ref{line:vcyclegenphi} in Algorithm \ref{alg:fascd} can be removed.
 Then VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} can be rewritten as the weak form problem $\ip{f^j(u^j)}{v} = \ip{\ell^j}{v}$ over $v\in \mathcal{V}_0^j$, where $u^j=w^j+y^j$ when down-smoothing and $u^j=w^j+z^j$ when up-smoothing, and the smoothers can be regarded as unconstrained, in-place modifications of $u^j$.
 Noting line \ref{line:vcyclerestrictsolution} now says $u^{j-1}=\iR u^j$, after down-smoothing the coarsened source functional computed by line \ref{line:vcyclerestrictell} can be rewritten as $\ell^{j-1} = f^{j-1}\left(u^{j-1}\right) + R\left(\ell^j-f^j(u^j)\right)$.
-
-Going upward we must be careful that only the coarse correction, and not the coarse solution, is prolonged \cite[remark 5.3.9]{Trottenbergetal2001}.
+%
+%Going upward we must be careful that only the coarse correction, and not the coarse solution, is prolonged \cite[remark 5.3.9]{Trottenbergetal2001}.
 After up-smoothing on the $j-1$ level the coarse iterate is $u^{j-1}$, but at this point $u^j=w^j + y^j$ contains the result of $j$th-level down-smoothing.
@@
 Thus the $j$th-level iterate \emph{before} up-smoothing must be set to $w^j + y^j + P z^{j-1} = u^j + P(u^{j-1} - \iR u^j)$, which replaces line \ref{line:vcycleupsmoothinitial}.
-
-With these modifications, Algorithm \ref{alg:fascd} becomes the FAS V-cycle below, as a solver for \eqref{eq:app:fas:pde}.
+%
+%With these modifications, Algorithm \ref{alg:fascd} becomes the FAS V-cycle below, as a solver for \eqref{eq:app:fas:pde}.
 This algorithm appears in various sources in the literature, for example \cite[Algorithm 14]{Bruneetal2015} and \cite[section 5.3.4]{Trottenbergetal2001}.
-
-\begin{pseudo*}
-\pr{fas-vcycle}(\ell^J; u^J)\text{:} \\+
-    for $j=J$ downto $j=1$ \\+
-      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j;u^j)$ \\
-      $u^{j-1} = \iR u^j$ \\
-      $\ell^{j-1} = f^{j-1}(u^{j-1}) + R \left(\ell^j - f^j(u^j)\right)$ \\-
-    $\text{\pr{solve}}(\ell^0;u^0)$ \\
-    for $j=1$ to $j=J$ \\+
-      $u^j \gets u^j + P (u^{j-1} - \iR u^j)$ \\
-      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j;u^j)$ \\-
-\end{pseudo*}
+%
+%\begin{pseudo*}
+%\pr{fas-vcycle}(\ell^J; u^J)\text{:} \\+
+%    for $j=J$ downto $j=1$ \\+
+%      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j;u^j)$ \\
+%      $u^{j-1} = \iR u^j$ \\
+%      $\ell^{j-1} = f^{j-1}(u^{j-1}) + R \left(\ell^j - f^j(u^j)\right)$ \\-
+%    $\text{\pr{solve}}(\ell^0;u^0)$ \\
+%    for $j=1$ to $j=J$ \\+
+%      $u^j \gets u^j + P (u^{j-1} - \iR u^j)$ \\
+%      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j;u^j)$ \\-
+%\end{pseudo*}
 

### 2023-07-27 paper/results/plap1d.m
p-Laplacian results: no change except -fascd_stol allows skipping regularization

@@ -3,13 +3,15 @@
 h = 2.^(-(0:9));
 fasterr = [2.263e-01 3.255e-02 9.108e-03 3.248e-03 5.505e-04 ...
            1.690e-04 4.717e-05 9.522e-06 3.591e-06 4.948e-07];
-degerr = [8.700e-02 3.812e-02 1.589e-02 6.318e-03 2.330e-03 ...
-          6.992e-04 2.047e-04 1.078e-04 3.997e-05 1.506e-05];
-loglog(h,fasterr,'*',h,degerr,'o')
+%degerr = [8.700e-02 3.812e-02 1.589e-02 6.318e-03 2.330e-03 ...
+%          6.992e-04 2.047e-04 1.078e-04 3.997e-05 1.506e-05];
+%loglog(h,fasterr,'*',h,degerr,'o')
+loglog(h,fasterr,'*')
 grid on,  xlabel h,  ylabel('|error|_{inf}')
-legend('p=1.5','p=4.0')
+%legend('p=1.5','p=4.0')
+legend('p=1.5')
 fastp = polyfit(log(h),log(fasterr),1)
-degp = polyfit(log(h),log(degerr),1)
+%degp = polyfit(log(h),log(degerr),1)
 
 %   fastp = 2.0047  -1.7579
 %   degp  = 1.4203  -2.2802

### 2023-07-27 paper/results/plap1d.sh

@@ -22,16 +22,16 @@ for EPSREG in 0.0 1.0e-8; do
     echo ""
 done
 
-for ITS in 3 4; do
-    echo "******** p = 4.0 V-cycle ITS=${ITS} *******"
-    for LEVS in $LLIST; do
-        python3 plap1d.py -p 4.0 -fascd_levels_snes_max_it $ITS -levs $LEVS
-    done
-    echo ""
-done
+# for ITS in 3 4; do
+#     echo "******** p = 4.0 V-cycle ITS=${ITS} *******"
+#     for LEVS in $LLIST; do
+#         python3 plap1d.py -p 4.0 -fascd_levels_snes_max_it $ITS -levs $LEVS
+#     done
+#     echo ""
+# done
 
-echo "******** p = 4.0 F-cycle *******"
-for LEVS in $LLIST; do
-    python3 plap1d.py -fascd_cycle_type full -p 4.0 -levs $LEVS
-done
-echo ""
+# echo "******** p = 4.0 F-cycle *******"
+# for LEVS in $LLIST; do
+#     python3 plap1d.py -fascd_cycle_type full -p 4.0 -levs $LEVS
+# done
+# echo ""

### 2023-07-28 paper/fascd.tex
precision
Edits in Results section cut 1/2 page.
But more edits to come, in all earlier sections.
24 pages is a reasonable target.  Note I need
to decide on final write-up of stol and of
Example 8.1.

@@
 Specifically, we see this performance for classical (unilateral, Laplacian) obstacle problems, $p$-Laplacian obstacle problems, and box-constrained advection-diffusion equations.
-Furthermore, these results appear to be relatively insensitive to the geometric complexity of the coincidence set and free boundary.
+Furthermore, V-cycle results appear to be insensitive to the geometric complexity of the coincidence set and free boundary.
 
@@
 
+\elb{add \id{stol}}
+
 However, an F-cycle (Figure \ref{fig:fascdcycles}), also called a \emph{full multigrid cycle}, is even better.
@@
 
-\begin{comment}
-smoother semantics: in our implementation of the V-cycle we actually shift to solve a VI for $\tilde w^j = w^j+y^j$, then compute $y^j = \tilde w^j - w^j$ after the solve, etc.; this is evidently equivalent
-\end{comment}
-
 A large space of smoothers could be explored, but in the next Section we will only apply an active-set (reduced-space) Newton method \cite{Balayetal2023,BensonMunson2006}, denoted as \emph{RS Newton}.
@@
 For $P_1$ and $Q_1$ elements, Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}.
-The numerical results in this Section, which follow the Section \ref{sec:vi} examples, focus on the scaling of FASCD iterations as the mesh resolution increases.
-Our primary observations are that each V-cycle greatly reduces the residual norm, and that a few F-cycles solve most examples to within discretization error.
+The numerical results in this Section, which follow the Section \ref{sec:vi} examples, focus on the scaling of FASCD iterations with mesh resolution.
+We see that each V-cycle greatly reduces the residual norm, and that a few F-cycles solve most of these problems to within discretization error.
 
@@
 Consider the classical unilateral 2D Laplacian obstacle problem, Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\}$.
-We consider two problems over square domains $\Omega$, with different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$ (Figure \ref{fig:results:classical}).
-The standard ``ball'' problem has an upper hemisphere obstacle \cite[Chapter 12]{Bueler2021}, but in the ``spiral'' problem the obstacle is constructed to produce the coincidence set shown \cite[problem 7.1.1]{GraeserKornhuber2009}.
-For this experiment we used a coarsest regular-triangulation mesh with $m_0=41$ nodes.
-The finest mesh had $m_{10} \approx 3.4 \times 10^7$ nodes.
+We consider two cases, over square domains $\Omega$, with different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$ (Figure \ref{fig:results:classical}).
+The standard ``ball'' problem has an upper hemisphere obstacle \cite[Chapter 12]{Bueler2021}, but the ``spiral'' problem's obstacle is constructed to produce the coincidence set shown \cite[problem 7.1.1]{GraeserKornhuber2009}.
+For this experiment we used a coarsest (regular triangulation) mesh with $m_0=41$ nodes, and the finest mesh had $m_{10} \approx 3.4 \times 10^7$ nodes.
 The smoother applied a single RS Newton iteration (Section \ref{sec:implementation}), with three IC-preconditioned CG iterations.
-The initial iterate was set to $w=\max\{0,\psi\}$.
+The initial iterate was $w=\max\{0,\psi\}$.
 
 Table \ref{tab:results:classical} shows the resulting iteration counts.
-As expected, RS Newton applied by itself as a solver shows rapidly-increasing iterations, roughly proportional to resolution, but V-cycle iterations increase slowly, logarithmically with the resolution.
+As expected, RS Newton, by itself as a solver, shows rapidly-increasing iterations, roughly proportional to resolution, but V-cycle iterations increase slowly, logarithmically with the resolution.
 These experiments over-solve, with \texttt{rtol} $=$ \texttt{atol} $= 10^{-12}$ in stopping criterion \eqref{eq:stoppingcriterion}.
@@
 
-For F-cycles we instead used \texttt{rtol} $= 10^{-8}$, sufficient to reach discretization error, and we report the number of V-cycles after the ramp (Section \ref{sec:implementation}).
-A small, constant number solves the problem, though spiral needs more V-cycles to clean-up error remaining after the ramp.
-Since each V-cycle does the work of about eight finest-level Krylov iterations, the ball results show nearly textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  Work of less than ten residual evaluations is often mentioned as the standard.} at the highest resolutions despite the low solution regularity.
+For F-cycles we instead used \texttt{rtol} $= 10^{-8}$, sufficient to reach discretization error, and we report the number of V-cycles after the ramp.
+A small, constant number of cycles solves the problem, though spiral needs more V-cycles to clean-up error remaining after the ramp.
+Since each V-cycle does the work of about eight finest-level Krylov iterations, the ball results show nearly textbook multigrid efficiency at the highest resolutions despite the low solution regularity.
 \end{example}
@@
 \begin{example}  \label{ex:results:plap}
-We again consider Example \ref{ex:plaplacian} with a unilateral constraint, but now for the nonlinear $p=1.5$-Laplacian, a ``fast'' diffusion with diffusivity limit $D=|\grad u|^{p-2}\to \infty$ as $|\grad u|\to 0$.
-Note that Example \ref{ex:results:sia} applies a doubly-nonlinear version of a $p=4$ ``degenerate'' diffusion operator.
-Well-posedness and solution regularity for $p$-Laplacian obstacle problems is considered in \cite{ChoeLewis1991}.
-
-This example reveals the strong contrast between $V(1,0)$ and $V(0,1)$ results (\pr{up}, \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively), and the need for robust smoothers when the operator is nonlinear.
+We again consider Example \ref{ex:plaplacian}, with a unilateral constraint, but now for the nonlinear $p$-Laplacian \cite{ChoeLewis1991}.
+Setting $p=1.5$ yields a fast diffusion ($D=|\grad u|^{p-2}\to \infty$ as $|\grad u|\to 0$).
+Note that Example \ref{ex:results:sia} below considers a doubly-nonlinear operator with $p=4$ degenerate diffusion.
+
+This example reveals the strong contrast between $V(1,0)$ and $V(0,1)$ results (\pr{up}, \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively).
 Set $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, $\mathcal{K} = \{v \ge \psi\} \subset W^{1,p}(\Omega)$, and $\ip\ell v = \int_\Omega g v$, where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$, in VI \eqref{eq:vi}.
 The exact solution is easily computed for any $p$; see the codes accompanying this paper.
-For this experiment we used a coarsest mesh of $m_0=7$ vertices and a finest mesh with $m_9=3073$.
-The default smoother uses three (fixed) iterations of RS Newton, with LU solution of the arising linear systems.
-The coarse solver is the same but iterated to convergence.
-The initial iterate was $w=\psi$, and \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$.
-Figure \ref{fig:imagesvcycle} illustrates the major steps of an example V-cycle, showing each of the coarse-correction VI subproblems and solutions; the result $\tilde w^2$ is visually close to the exact solution.
+For this experiment the coarsest mesh has $m_0=7$ nodes and the finest $m_9=3073$.
+The default smoother uses three (fixed) iterations of RS Newton, with LU.
+We set \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $=$ \texttt{stol} $= 10^{-12}$.
+Figure \ref{fig:imagesvcycle} illustrates the major steps of a V-cycle, showing each of the coarse-correction VI subproblems and solutions; the result $\tilde w^2$ is visually close to the exact solution.
 
@@
 Iteration results are shown in Table \ref{tab:results:fastplap1d}.
-We see that $V(1,0)$ cycles, essentially Algorithm 4.7 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.
-By contrast, $V(0,1)$ cycles, up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are as good as the default $V(1,1)$ cycles ($=$ Algorithm \ref{alg:fascd}).
-F-cycles are even more efficient except when they fail entirely, but these failures arise from singular diffusivity.
-If we regularize  $D=|\grad u|^{-0.5}$ to $D_\eps = \left(\eps + |\grad u|^2\right)^{-0.25}$, with $\eps=10^{-8}$, then a single F-cycle suffices at every resolution (column ``F$_\eps$'').
-
-\elb{check if regularization is even necessary when you re-run with \texttt{-fascd\_levels\_snes\_vi\_zero\_tolerance 1.0e-12}}
-
+We see that $V(1,0)$ cycles, Algorithm 4.7 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.
+By contrast, $V(0,1)$ cycles, up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are as good as the default $V(1,1)$ cycles (Algorithm \ref{alg:fascd}).
+F-cycles are even more efficient; a single cycle suffices at all higher resolutions.
 Note that convergence is at rate $O(h^{2.005})$; compare \cite{ChoeLewis1991}.
@@
 \centering
-\begin{tabular}{lr@{\hskip 7mm}c@{\hskip 3mm}c@{\hskip 3mm}c@{\hskip 4mm}c@{\hskip 5mm}c@{\hskip 6mm}c}
+\begin{tabular}{lr@{\hskip 7mm}c@{\hskip 3mm}c@{\hskip 3mm}c@{\hskip 4mm}c@{\hskip 6mm}c}
 \toprule
-\emph{levels} & $m_J$ & V(1,0) & V(0,1) & V(1,1) & F & F$_\eps$ & $\|$error$\|_\infty$ \\
+\emph{levels} & $m_J$ & V(1,0) & V(0,1) & V(1,1) & F & $\|$error$\|_\infty$ \\
 \midrule
- 1 &    $7$ &  1 &  1 &  1 &  1 &  1 & $2.3\times 10^{-1}$ \\
- 2 &   $13$ &  3 &  3 &  2 &  2 &  1 & $3.3 \times 10^{-2}$ \\
- 3 &   $25$ & 12 &  4 &  2 &  1 &  1 & $9.1 \times 10^{-3}$ \\
- 4 &   $49$ & NC &  4 &  2 &  1 &  1 & $3.2 \times 10^{-3}$ \\
- 5 &   $97$ & NC &  3 &  3 &  1 &  1 & $5.5 \times 10^{-4}$ \\
- 6 &  $193$ & NC &  3 &  3 & NC &  1 & $1.7 \times 10^{-4}$ \\
- 7 &  $385$ & NC &  3 &  3 &  1 &  1 & $4.7 \times 10^{-5}$ \\
- 8 &  $769$ & NC &  4 &  7 &  1 &  1 & $9.5 \times 10^{-6}$ \\
- 9 & $1537$ & NC &  5 &  6 & NC &  1 & $3.6 \times 10^{-6}$ \\
-10 & $3073$ & NC & 11 & 19 &  1 &  1 & $4.9 \times 10^{-7}$ \\
+ 1 &    $7$ &  1 &  1 &  1 &  1 & $2.3\times 10^{-1}$ \\
+ 2 &   $13$ &  3 &  3 &  2 &  2 & $3.3 \times 10^{-2}$ \\
+ 3 &   $25$ & 12 &  4 &  2 &  1 & $9.1 \times 10^{-3}$ \\
+ 4 &   $49$ & NC &  4 &  2 &  1 & $3.2 \times 10^{-3}$ \\
+ 5 &   $97$ & NC &  3 &  3 &  1 & $5.5 \times 10^{-4}$ \\
+ 6 &  $193$ & NC &  3 &  3 &  1 & $1.7 \times 10^{-4}$ \\
+ 7 &  $385$ & NC &  3 &  3 &  1 & $4.7 \times 10^{-5}$ \\
+ 8 &  $769$ & NC &  4 &  7 &  1 & $9.5 \times 10^{-6}$ \\
+ 9 & $1537$ & NC &  5 &  6 &  1 & $3.6 \times 10^{-6}$ \\
+10 & $3073$ & NC & 11 & 19 &  1 & $4.9 \times 10^{-7}$ \\
 \bottomrule
@@
 \bigskip
-\caption{Iterations for V- and F-cycles in Example \ref{ex:results:plap}.  NC $=$ did not converge after 50 iterations.  See text regarding F$_\eps$.}
+\caption{Iterations for V- and F-cycles in Example \ref{ex:results:plap}.  NC $=$ did not converge after 50 iterations.}
 \label{tab:results:fastplap1d}
@@
 
-Solvers implemented in Firedrake can exploit parallel solvers from the PETSc library \cite{Balayetal2023}, and in fact the above Examples, which were run in serial so as to fix the smoother, also run efficiently in parallel.
-The next Examples apply parallel solvers, and Example \ref{ex:results:sia} demonstrates the parallel weak scaling \cite{Bueler2021} of FASCD.
+Firedrake applications can exploit the PETSc \cite{Balayetal2023} library's parallel solvers.
+The above Examples, which were run in serial to fix the smoother, also run efficiently in parallel.
+The next Examples apply parallel solvers and then demonstrate the weak scaling of FASCD F-cycles.
 
@@
 Consider the advection-diffusion VI problem in Example \ref{ex:advectiondiffusion} with $\Omega=(-1,1)^d$, $d=2,3$, diffusivity $\eps=0.1$ in \eqref{eq:advectiondiffusionstrong}, and zero Dirichlet boundary conditions.
-We made the 2D problem satisfy the coercivity hypothesis with a divergence-free velocity field $\bX = (7+5y,-5x)$, but in 3D set $\bX = (7+5y,-5x,2z)$ so $\grad\cdot\bX \ne 0$.
-The chosen discontinuous source term $\phi \in L^\infty(\Omega)$ satisfies $\phi\ge 0$ in half of the domain, with positive circular patches, with negative values in the other half; details are in the associated code.
+The 2D problem satisfies the coercivity hypothesis, with a divergence-free velocity field $\bX = (7+5y,-5x)$, but in 3D we set $\bX = (7+5y,-5x,2z)$ so $\grad\cdot\bX \ne 0$.
+The discontinuous source term $\phi \in L^\infty(\Omega)$ satisfies $\phi\ge 0$ in half of the domain, with positive circular patches, but values are negative in the other half; details are in the associated code.
 While the other Examples in this Section are unilateral, here the solution is box-constrained: $0 \le u \le 1$.
@@
 \includegraphics[width=0.35\textwidth]{fixfigs/poll2d-zero-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/poll2d-one-set.png}
-\caption{Coincidence sets for the 2D bi-obstacle problem in Example \ref{ex:results:advdiff}: $u=0$ (left) and $u=1$ (right).}
+\caption{Coincidence sets for the 2D problem in Example \ref{ex:results:advdiff}: $u=0$ (left) and $u=1$ (right).}
 \label{fig:results:advdiff}
@@
 
-The 2D solver applied $P_1$ elements on a single core, but in 3D the elements were chosen to be $Q_1$ (hexahedra) on $P=16$ cores.
+The 2D solver applied $P_1$ elements on a single core, but in 3D the elements were chosen to be $Q_1$ (hexahedra) on $16$ cores.
 A coarse mesh with $m_0=16^d$ vertices minimally-resolved the coincidence set.
-The smoother was two (fixed) RS Newton iterations, with three preconditioned GMRES iterations for the arising linear systems; the preconditioning was additive Schwarz method (ASM) with incomplete factorization (ILU) on each process.
+The local Péclet number $\text{Pe}=\|\bX(x,y,z)\|_2 L \eps^{-1}$, with $L=2$ the width of $\Omega$, varied from 40 to 260, and the mesh Péclet number $m_J^{-1/d}\,\text{Pe}$ varied further with resolution, so the $m_J\le 121^d$ problems were advection-dominated, and those with $m_J\ge 481^d$ diffusion-dominated.
+However, the smoother was the same in all cases: two (fixed) RS Newton iterations, with three preconditioned (additive Schwarz method (ASM) with ILU on each process) GMRES iterations.
 The coarse solver was RS Newton and LU, iterated to convergence.
-Tolerances \id{rtol} $=10^{-5}$ and \id{atol} $=10^{-9}$ were used in \eqref{eq:stoppingcriterion}.
-
+Tolerances were set to \id{rtol} $=10^{-5}$ and \id{atol} $=$ \id{stol} $= 10^{-9}$.
 F-cycle iteration results are shown in Table \ref{tab:results:advdiff}.
-Despite the hint of growing iterations under mesh refinement, a few iterations sufficed up to memory-limited resolutions.
-For any of the higher resolution runs, the FASCD F-cycle solver took less time than generating the mesh hierarchy.
-The local Péclet number $\text{Pe}=\|\bX(x,y,z)\|_2 L \eps^{-1}$, with $L=2$ the width of $\Omega$, varied from 40 to 260.
-The mesh Péclet number $m_J^{-1/d}\,\text{Pe}$ varied accordingly, and though the $m_J\le 121^d$ problems were advection-dominated, those with $m_J\ge 481^d$ were diffusion-dominated, and the smoother was the same in all cases, with nearly-constant iterations.
+A few F-cycles sufficed up to these memory-limited resolutions.
 \end{example}
@@
 \begin{example}   \label{ex:results:sia}
-We consider the steady shallow ice approximation (SIA) obstacle problem presented in Example \ref{ex:sia}, in which the surface elevation is constrained to exceed the bedrock elevation, $s\ge b$.
-We set $\Omega=(0,L)^2$ with $L=1800$ km.
-The data consist of a radial surface mass balance function $a$ \cite[equation (5.122)]{GreveBlatter2009} and a bumpy, synthetic map $b\in C^1(\Omega)$, with an elevation range of $1800$ m; details for $b$ are in the associated code.
-A numerical solution is shown in Figure \ref{fig:results:siascene}, colored by the solution surface flow speed, i.e.~the magnitude of $\mathbf{u} = (n+1)(n+2)^{-1} \Gamma (s-b)^{n+1} |\grad s|^{n-1} \grad s$.
-As noted in Example \ref{ex:sia}, and seen in this Figure, $|\grad s|$ is unbounded at the free boundary; the solution $s$, while continuous, has very low regularity.
+Consider the steady shallow ice approximation (SIA) obstacle problem (Example \ref{ex:sia}), in which the surface elevation is constrained to exceed the bedrock elevation, $s\ge b$, on $\Omega=(0,L)^2$ with $L=1800$ km.
+The data consist of a radial surface mass balance function $a$ \cite[equation (5.122)]{GreveBlatter2009} and a bumpy, synthetic map $b\in C^1(\Omega)$, with an elevation range of $1800$ m; details are in the associated code.
+A numerical solution is shown in Figure \ref{fig:results:siascene}, colored by the solution surface flow speed, the magnitude of $\mathbf{u} = (n+1)(n+2)^{-1} \Gamma (s-b)^{n+1} |\grad s|^{n-1} \grad s$.
+Note $|\grad s|$ is unbounded at the free boundary; the solution $s$, while continuous, has very low regularity.
 
@@
 \includegraphics[width=1.0\textwidth]{fixfigs/sialev8scene.png}
-\caption{Solution of Example \ref{ex:sia} over a bumpy bedrock topography (grey), with the ice speed as surface coloring and 100-times vertical exaggeration.}
+\caption{Example \ref{ex:results:sia} solution over a bumpy bedrock topography (grey), with 100-times vertical exaggeration and surface coloring by flow speed.}
 \label{fig:results:siascene}
@@
 We applied FASCD F-cycles using $Q_1$ elements and a smoother of four (fixed) iterations of RS Newton (Section \ref{sec:implementation}), with quadratic backtracking line search.
-The arising linear systems were approximately solved, in parallel, by three (fixed) iterations of preconditioned GMRES and a ASM+ILU preconditioner.
-The coarse solver was RS+LU, applied redundantly over all processors.
+The arising linear systems were approximately solved by three (fixed) iterations of preconditioned GMRES and a ASM+ILU preconditioner.
 The coarse mesh was relatively fine, with $m_0=21^2$ vertices and 400 cells.
-Because of the low regularity of the solution we used \id{rtol} $= 2 \times 10^{-4}$ and \id{atol} $= 10^{-8}$ in \eqref{eq:stoppingcriterion}.
+Because of the low regularity of the solution we used \id{rtol} $= 2 \times 10^{-4}$ and \id{atol} $= 10^{-8}$.
 The initial iterate was zero ice: $s=b$.
-These parameters generated a solver with mesh-independent iterations.
+These choices generated a solver with mesh-independent iterations.
 
 For a weak-scaling study we fixed the number of degrees of freedom per process at $m_J/P=641^2 \approx 4.1 \times 10^5$ and solved with $6, \dots, 11$ refinements.\footnote{The experiments were conducted on ARCHER2, the UK national supercomputer.
-Each node has 16 memory channels, so a maximum of 16 cores per node were employed.}  The results in Table \ref{tab:results:siaweak} show some degradation in time per cycle with increasing core counts, perhaps due to the use of one-level additive Schwarz in the solution of the arising linear systems.
+Each node has 16 memory channels, so a maximum of 16 cores per node were employed.}  The results in Table \ref{tab:results:siaweak} show some degradation in time per cycle with increasing core counts, perhaps due to the use of one-level additive Schwarz.
 Nevertheless, the finest mesh result, with $m_J=20481^2=4.1 \times 10^8$ degrees of freedom, required less than three minutes to solve for the steady surface elevation of an ice sheet larger than the current Greenland ice sheet ($1.7$ million $\text{km}^2$) at uniform $87.5$ m resolution.
@@
 \bigskip
-\caption{Iterations and run time (wall clock, in seconds) for parallel FASCD F-cycle solutions of Example \ref{ex:sia} on $P$ processes.}
+\caption{Iterations and wall clock time for parallel FASCD solutions of Example \ref{ex:results:sia} on $P$ processes.}
 \label{tab:results:siaweak}
@@
 As a multilevel solver for VI problems, the FASCD method is a strategy for generating coarser-level problems which, once solved, make helpful admissible additions to a V-cycle update.
-The method is smoother-agnostic, and our implementation within the extensible Firedrake \cite{Rathgeberetal2016} and PETSc \cite{Balayetal2023} library framework should allow easy experimentation with other smoothers (projected Gauss--Seidel, semi-smooth Newton, interior-point, etc.), beyond the active-set Newton method \cite{BensonMunson2006} chosen for demonstration here.
-For box-constrained smoothers of Newton-Krylov type one could also add linear geometric or algebraic multigrid preconditioning \cite{Trottenbergetal2001}, instead of our simple choice of incomplete factorization and (in parallel) additive Schwarz preconditioners.
-However, the performance illustrated in Section \ref{sec:results} is already excellent.
-
-\subsection*{Code availability} \label{sec:code}  The software used to produce the results is archived at tag \elb{FIX} in the repository \url{https://bitbucket.org/pefarrell/fascd/}, using the Firedrake version \elb{what version?}
+The method is smoother-agnostic, and our implementation within the extensible Firedrake \cite{Rathgeberetal2016} and PETSc \cite{Balayetal2023} library framework will allow easy experimentation with other smoothers (projected Gauss--Seidel, semi-smooth Newton, interior-point, etc.), beyond the active-set Newton method chosen for demonstration here.
+For smoothers of Newton-Krylov type one could also add linear geometric or algebraic multigrid preconditioning \cite{Trottenbergetal2001}, instead of our simple choice of incomplete factorization preconditioners.
+However, the performance seen in Section \ref{sec:results} is already excellent.
+
+\subsection*{Code availability} \label{sec:code}  The software used to produce the results is archived at tag \texttt{v1.0} in the repository \texttt{\url{https://bitbucket.org/pefarrell/fascd/}}, using the Firedrake version \texttt{3fb16ad47}.
 
@@
 
-%\appendix
-%\section{Reduction to FAS multigrid for PDEs} \label{app:reductions}
-%
-%When all inequality constraints are removed the FASCD V-cycle Algorithm \ref{alg:fascd} reduces to the classical FAS multigrid V-cycle for nonlinear PDEs \cite{Brandt1977,Trottenbergetal2001}, which itself generalizes linear multigrid.
-In fact, suppose we replace $\underline{\gamma}^J,\overline{\gamma}^J$ in definition \eqref{eq:fe:fineconstraintset} with $-\infty,+\infty$, respectively.
-The finest-level admissible set $\mathcal{K}^J$ is then determined only by the Dirichlet boundary conditions.
-Let $\mathcal{V}_0^J \subset \mathcal{V}^J$ denote the linear subspace with zero values on $\partial_D\Omega$.
-Since $v-u^J\in\mathcal{V}_0^J$ is now arbitrary we can rewrite \eqref{eq:fe:vi} as the FE discretization of strong-form PDE $f(u)=\ell$: find $u^J \in \mathcal{V}^J$ such that
-%\begin{equation}
-%\ip{f^J(u^J)}{v} = \ip{\ell^J}{v} \qquad \text{for all } v\in \mathcal{V}_0^J. \label{eq:app:fas:pde}
-%\end{equation}
-%
-%The downward $\mathcal{D}^j$ and upward $\mathcal{U}^j$ constraint sets are also enlarged to vector spaces, and the corrections $y^j$ and $z^j$ are unconstrained; lines \ref{line:vcyclegenchifinest}, \ref{line:vcyclegenchi}, and \ref{line:vcyclegenphi} in Algorithm \ref{alg:fascd} can be removed.
-Then VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi} can be rewritten as the weak form problem $\ip{f^j(u^j)}{v} = \ip{\ell^j}{v}$ over $v\in \mathcal{V}_0^j$, where $u^j=w^j+y^j$ when down-smoothing and $u^j=w^j+z^j$ when up-smoothing, and the smoothers can be regarded as unconstrained, in-place modifications of $u^j$.
-Noting line \ref{line:vcyclerestrictsolution} now says $u^{j-1}=\iR u^j$, after down-smoothing the coarsened source functional computed by line \ref{line:vcyclerestrictell} can be rewritten as $\ell^{j-1} = f^{j-1}\left(u^{j-1}\right) + R\left(\ell^j-f^j(u^j)\right)$.
-%
-%Going upward we must be careful that only the coarse correction, and not the coarse solution, is prolonged \cite[remark 5.3.9]{Trottenbergetal2001}.
-After up-smoothing on the $j-1$ level the coarse iterate is $u^{j-1}$, but at this point $u^j=w^j + y^j$ contains the result of $j$th-level down-smoothing.
-The initial value of the correction $z^j$, before up-smoothing on the $j$th level, is $z^j = y^j + P z^{j-1}$, where $z^{j-1}$ contains the total correction from coarser corrections and $j-1$ level up-smoothing.
-Thus the $j$th-level iterate \emph{before} up-smoothing must be set to $w^j + y^j + P z^{j-1} = u^j + P(u^{j-1} - \iR u^j)$, which replaces line \ref{line:vcycleupsmoothinitial}.
-%
-%With these modifications, Algorithm \ref{alg:fascd} becomes the FAS V-cycle below, as a solver for \eqref{eq:app:fas:pde}.
-This algorithm appears in various sources in the literature, for example \cite[Algorithm 14]{Bruneetal2015} and \cite[section 5.3.4]{Trottenbergetal2001}.
-%
-%\begin{pseudo*}
-%\pr{fas-vcycle}(\ell^J; u^J)\text{:} \\+
-%    for $j=J$ downto $j=1$ \\+
-%      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j;u^j)$ \\
-%      $u^{j-1} = \iR u^j$ \\
-%      $\ell^{j-1} = f^{j-1}(u^{j-1}) + R \left(\ell^j - f^j(u^j)\right)$ \\-
-%    $\text{\pr{solve}}(\ell^0;u^0)$ \\
-%    for $j=1$ to $j=J$ \\+
-%      $u^j \gets u^j + P (u^{j-1} - \iR u^j)$ \\
-%      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j;u^j)$ \\-
-%\end{pseudo*}
-
 \end{document}

### 2023-07-28 paper/fascd.tex
edits sections 1 and 2 ... more to come

@@
 
-\title{A full approximation scheme multilevel method \\ for nonlinear variational inequalities\thanks{Submitted to the editors DATE.
+\title{A full approximation scheme multilevel method \\ for nonlinear variational inequalities\thanks{Draft \today.%Submitted to the editors DATE.
 \funding{EB was supported by a Faculty Development Travel Award from United Academics, University of Alaska Fairbanks.
@@
 This work used the ARCHER2 UK National Supercomputing Service (https://www.archer2.ac.uk).
-PEF thanks Lawrence Mitchell for advice on the implementation of FASCD, and thanks Jack D.~Betteridge for assistance with the numerical experiments of Example \ref{ex:results:sia}.}}}
+PEF thanks Lawrence Mitchell for advice on implementation, and Jack D.~Betteridge for assistance with numerical Example \ref{ex:results:sia}.}}}
 
@@
 We extend the CD idea by exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
-When a reduced-space active set method is applied as a smoother, with work proportional to the number of unknowns on a given mesh level, FASCD V-cycles exhibit mesh-independent convergence rates and F-cycles (full multigrid) are optimal solvers.
+When a reduced-space active set Newton method is applied as a smoother, with work proportional to the number of unknowns on a given mesh level, FASCD V-cycles exhibit mesh-independent convergence rates and F-cycles (full multigrid) are optimal solvers.
 The example problems include differential operators which are symmetric linear, nonsymmetric linear, and nonlinear, in unilateral and bilateral VI problems.
@@
 \item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
-\item When a (mesh-dependent) reduced-space active set method is used as relaxation, we demonstrate that F-cycles are an optimal solver for the examples in Section \ref{sec:results}; smoother effort is proportional to the number of unknowns on a given level because the number of Newton and Krylov iterations is fixed.
+\item When a (mesh-dependent) reduced-space active set Newton method is used as relaxation, we demonstrate that F-cycles are an optimal solver for the various examples in Section \ref{sec:results}.
 \end{enumerate}
@@
 The monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be formulated only with Gauss--Seidel type smoothers, and may have been applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.
-Unlike the fastest variants of monotone multigrid, no zeroing of the basis functions associated with the active set (truncation) is employed in our method, both simplifying the implementation and allowing the coarse grid corrections to update the estimate of the active set on the fine grid.
-Numerical evidence indicates that our coarse grid corrections are useful for this purpose, even when the coarse grids are too coarse to usefully resolve the active set.
+Unlike the fastest variants of monotone multigrid, no zeroing of the basis functions associated with the active set (truncation) is employed in our method.
+This simplifies the implementation and allows the coarse grid corrections to update the estimate of the active set on the fine grid.
+Numerical evidence indicates that our coarse corrections are effective even when the coarse meshes are too coarse to resolve the active set.
 On a problem for which the truncated Newton/monotone multigrid methods were designed \cite[problem 7.1.1]{GraeserKornhuber2009}, the performance of FASCD is at least as good (Example \ref{ex:results:classical}).
@@
 
-In our experiments, FASCD V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for corresponding unconstrained PDE problems.
-The F-cycle results in Section \ref{sec:results} show nearly-textbook multigrid efficiency, namely solutions within discretization error in work comparable to few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}.
+In our experiments, FASCD V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for the corresponding unconstrained PDE problems.
+The F-cycle results in Section \ref{sec:results} show near-textbook multigrid efficiency, namely solutions within discretization error in work comparable to few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}.
 Specifically, we see this performance for classical (unilateral, Laplacian) obstacle problems, $p$-Laplacian obstacle problems, and box-constrained advection-diffusion equations.
@@
 
-All iterates in FASCD are admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
+All iterates in FASCD, on all mesh levels, are admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
 For example, we solve an ice sheet geometry problem (Examples \ref{ex:sia} and \ref{ex:results:sia}) in which the ice flow formulas are only meaningful for surface elevation iterates that do not penetrate the bedrock.
-Non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural modifications of the operator formula.
-FASCD solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a small, mesh-independent number of F-cycle iterations (Example \ref{ex:results:sia}).
-
-%\elb{beam 2d example?}
+(Non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural modifications of the operator formula.)  FASCD solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a small, mesh-independent number of F-cycle iterations (Example \ref{ex:results:sia}).
+The same Example demonstrates the parallel weak scaling \cite{Bueler2021} of our implementation.
 
@@
 In Section \ref{sec:vi} we recall the theory of coercive and monotone variational inequalities, and describe several motivating examples.
-Section \ref{sec:cd} reviews the basics of CDs and their associated iterations.
+Section \ref{sec:cd} reviews CDs and their associated iterations.
 Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, and then Section \ref{sec:cdmultilevel} extends the CD method via defect constraints generated by monotone injection operators \cite{GraeserKornhuber2009}.
 Here we propose an apparently new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD and an associated iteration.
-Section \ref{sec:vcycle} then states the FASCD V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, F-cycles, and our smoother choice.
-Section \ref{sec:results} gives numerical results from a parallel Firedrake \cite{Rathgeberetal2016} implementation (see Code Availability), for each example problem introduced in Section \ref{sec:vi}.
-The work concludes with a discussion and outlook in Section \ref{sec:discussion}.
+Section \ref{sec:vcycle} then states the V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, F-cycles, and our smoother choice.
+Section \ref{sec:results} gives numerical results from our Firedrake \cite{Rathgeberetal2016} implementation (see Code Availability).
 
@@
 \end{equation}
-Because $f$ is (generally) nonlinear, the source term $\ell$ is not strictly needed in stating this class of problems.
-In fact, by redefining $f$ we may take $\ell=0$.
-However, introducing $\ell$ clarifies the algorithm in Section \ref{sec:vcycle}.
+Because $f$ is (generally) nonlinear, the source term $\ell$ is not strictly needed in stating this class of problems---by redefining $f$ we may take $\ell=0$---but introducing $\ell$ clarifies the algorithms.
 
 VI \eqref{eq:vi} generalizes nonlinear systems of equations $f(u^\star)=\ell$ to problems where $u^\star$ is also constrained to be in $\cK$.
-Informally, pretending the dual pairing is an inner product, \eqref{eq:vi} says that the angle between $f(u^\star)-\ell$ and any arbitrary vector $v-u^\star$ pointing from $u^\star$ into $\cK$ is at most $90^\circ$; while $f(u^\star)-\ell$ may not be zero, it must point directly into $\cK$.
-On the other hand, if $u^\star$ is in the interior of the constraint set then \eqref{eq:vi} implies $f(u^\star)=\ell$.
+Pretending the dual pairing is an inner product, by \eqref{eq:vi} the angle between $f(u^\star)-\ell$ and any arbitrary vector $v-u^\star$ pointing from $u^\star$ into $\cK$ is at most $90^\circ$; the dual vector $f(u^\star)-\ell$ must point into $\cK$.
+If $u^\star$ is in the interior of the constraint set then \eqref{eq:vi} implies $f(u^\star)=\ell$.
 Variational inequalities may generally permit multiple solutions~\cite{Farrell2019}.
@@
 It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, and that the solution is unique when $f$ is strictly monotone.
-As is standard in the calculus of variations \cite{Evans2010}, coercivity permits a compactness argument for unbounded sets $\cK$.
-(Recall that the closed and bounded subsets of a reflexive Banach space are weakly compact.)  The condition of continuity can be weakened, e.g.~to only apply on finite-dimensional subspaces \cite{KinderlehrerStampacchia1980}, but this is unnecessary in our examples.
-
-Some of the VIs solved in this paper satisfy a stronger inequality than \eqref{eq:coercive}.
+As is standard in the calculus of variations \cite{Evans2010}, coercivity permits a compactness argument for unbounded sets $\cK$; recall that the closed and bounded subsets of a reflexive Banach space are weakly compact.
+However, some of the VIs solved in this paper satisfy a stronger inequality than \eqref{eq:coercive}.
 
@@
 
-It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive, so the following well-posedness theorem holds.
-
-\begin{theorem}  \label{thm:viwellposed}  If $f:\cK \to \cV'$ is continuous and $p$-coercive then there exists a unique $u^\star\in \cK$ solving VI \eqref{eq:vi}.
-\end{theorem}
-
-Let $\Omega \subset \RR^d$ denote a bounded, open set with smooth or piecewise-smooth (e.g.~polygonal) boundary.
+It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive.
+Thus if $f:\cK \to \cV'$ is continuous and $p$-coercive then there exists a unique $u^\star\in \cK$ solving VI \eqref{eq:vi}.
+
+Let $\Omega \subset \RR^d$ denote a bounded, open set with piecewise-smooth boundary.
 Sobolev spaces \cite{Evans2010} will be denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$, with $W^{1,p}_0(\Omega)$ denoting the subspace having zero boundary values.
-
 The following example includes the classical and $p$-Laplacian obstacle problems \cite{ChoeLewis1991}.
 
-\begin{example}  \label{ex:plaplacian}  Let $p>1$.  For $u,v \in \cV = W^{1,p}_0(\Omega)$, define $f:\cV \to \cV'$ by
+\begin{example}  \label{ex:plaplacian}  Let $1<p<\infty$.  For $u,v \in \cV = W^{1,p}_0(\Omega)$, define $f:\cV \to \cV'$ by
 \begin{equation}
@@
 a continuous map \cite[Theorem A.0.6]{Peral1997}.
-For $p\ge 2$, if $x,y\in\RR^d$ then $(|x|^{p-2} x - |y|^{p-2} y)\cdot (x-y) \ge 2^{2-p} |x-y|^p$ \cite[Appendix A]{Bueler2021conservation}, hence it follows from the Poincar\'e inequality that
-    $$\ip{f(u) - f(v)}{u-v} \ge 2^{2-p} a_0 \|\grad u - \grad v\|_p^p \ge 2^{2-p} a_0 C \|u-v\|^p$$
-for some $C>0$, and thus $f$ is $p$-coercive.
-The map \eqref{eq:plaplacian} is also coercive if $1<p<2$, but the proof is somewhat different \cite[Theorem 4.4, for example]{Bueler2021conservation}.
+It follows from inequalities for $|\cdot|^p$ in $\RR^d$, and the Poincar\'e inequality, that $F$ is $p$-coercive \cite{ChoeLewis1991}; \cite[Appendix A]{Bueler2021conservation}.
 \end{example}
@@
 
-Next we give two examples which are \emph{not} of optimization type \eqref{eq:minimization}, the first being an advection-diffusion problem, which needs the following Lemma.
+Next we give two examples which are \emph{not} of optimization type \eqref{eq:minimization}, but which are nonetheless $p$-coercive.
+The first is an advection-diffusion problem; it needs the following Lemma.
 
@@
 
-\begin{example}  \label{ex:advectiondiffusion}  Suppose $\partial\Omega$ is partitioned into Dirichlet and Neumann portions, i.e.~$\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$, with $\partial_D\Omega$ of positive Hausdorff measure.
+\begin{example}  \label{ex:advectiondiffusion}  Suppose $\partial\Omega$ is partitioned into Dirichlet and Neumann portions, $\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$, with $\partial_D\Omega$ of positive Hausdorff measure.
 Let $\cV \subset W^{1,2}(\Omega)$ be the space of functions which are zero along $\partial_D\Omega$.
@@
 Thus $f$ is 2-coercive, and so the associated VI problem is well-posed.
+If $\bX \ne 0$ then $f\ne F'$ for any objective $F$, because $\ip{f(u)}{v}$ does not possess the symmetry of the gradient of a scalar: for $u,v$ which are zero on $\partial \Omega$, note that $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.
+References \cite{Bueler2021conservation,ChangNakshatrala2017} consider similar advection-diffusion VI problems over $\cK = \{v\ge 0\}$.
 \end{example}
-
-If $\bX \ne 0$ then $f\ne F'$ for any objective $F$, because $\ip{f(u)}{v}$ does not possess the symmetry of the gradient of a scalar: for $u,v$ which are zero on $\partial \Omega$, note that $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.
-References \cite{Bueler2021conservation,ChangNakshatrala2017} consider inequality-constrained advection-diffusion VI problems like this, specifically over the set $\cK = \{v\ge 0\}$.
 
@@
 \begin{example}  \label{ex:sia}  Let $\Omega \subset \RR^2$ be a fixed region of land, and suppose $b \in C^1(\Omega)$ is a given bedrock elevation.
-Let $a \in L^2(\Omega)$ denote a given ``surface mass balance'' function, the annually-averaged rate of ice accumulation (as snow) minus melt and runoff.
+Let $a \in L^2(\Omega)$ denote a given ``surface mass balance'' function, the annually-averaged rate of ice accumulation (snow) minus melt and runoff.
 Then the surface elevation $s\in C(\Omega)$ of a steady-state, isothermal, and non-sliding ice sheet (glacier) approximately satisfies the following \emph{shallow ice approximation} (SIA) \cite{GreveBlatter2009} VI problem: find $s \in \mathcal{K} = \{s\ge b\}$ such that
@@
 \end{equation}
-where $n>1$ and $\Gamma>0$ are constants determined by the physical properties of ice; $n=3$ is a typical value.  The doubly-nonlinear and doubly-degenerate operator in this case is
-\begin{equation}
-\ip{f(s)}{v} = \int_\Omega \Gamma (s-b)^{n+2} |\grad s|^{n-1} \grad s \cdot \grad v\dx. \label{eq:sia}
-\end{equation}
-If the bed is flat ($b=0$) then a power transformation converts the problem into $p$-Laplacian form \eqref{eq:plaplacian} with $p=n+1$.
-For general beds $b$ less is understood about this operator, but existence holds for \eqref{eq:siavi} \cite{JouvetBueler2012} and a power of the solution lives in a Sobolev space: $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.
-This theory, and also observations of ice sheets, shows that $|\grad s|$ is generally unbounded as one approaches the free boundary from the icy ($s>b$) side.
+where $n>1$ and $\Gamma>0$ are physical constants for ice; $n=3$ is a typical value \cite{GreveBlatter2009}.  The doubly-nonlinear and doubly-degenerate operator here is
+\begin{equation}
+\ip{f(s)}{v} = \int_\Omega \Gamma (s-b)^{n+2} |\grad s|^{n-1} \grad s \cdot \grad v\dx. \label{eq:siaoperator}
+\end{equation}
+If the bed is flat ($b=0$) then a power transformation converts \eqref{eq:siaoperator} into form \eqref{eq:plaplacian} with $p=n+1$.
+For general beds $b$ less is understood, but existence holds for \eqref{eq:siavi} \cite{JouvetBueler2012} and a power of the solution lives in a Sobolev space: $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.
+This theory, and also observations of ice sheets, shows that $|\grad s|$ may be unbounded as one approaches the free boundary from the icy ($s>b$) side.
 \end{example}
 
-Numerical results are shown in Section \ref{sec:results} for applications of FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} to these Examples.
+The numerical results in Section \ref{sec:results} apply FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} to the above Examples.
 
@@
 \centering
-\includegraphics[width=0.6\textwidth]{genfigs/cartoon.pdf}
+\includegraphics[width=0.5\textwidth]{genfigs/cartoon.pdf}
 \caption{A constraint decomposition (CD) for a unilateral obstacle problem on a two-point space $\Omega=\{x_1,x_2\}$, with $\mathcal{V}=\left\{v \,:\, \Omega \to \RR\right\}$ and $\mathcal{K}=\{v\ge \psi\}$.}

### 2023-07-28 paper/fascd.tex
Just change a few words

@@
 This work used the ARCHER2 UK National Supercomputing Service (https://www.archer2.ac.uk).
-PEF thanks Lawrence Mitchell for advice on implementation, and Jack D.~Betteridge for assistance with numerical Example \ref{ex:results:sia}.}}}
+PEF thanks Lawrence Mitchell for advice on implementation, and Jack D.~Betteridge for assistance with Example \ref{ex:results:sia}.}}}
 
@@
 (Non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural modifications of the operator formula.)  FASCD solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a small, mesh-independent number of F-cycle iterations (Example \ref{ex:results:sia}).
-The same Example demonstrates the parallel weak scaling \cite{Bueler2021} of our implementation.
+The same Example investigates the parallel weak scaling \cite{Bueler2021} of our implementation.
 
@@
 The above Examples, which were run in serial to fix the smoother, also run efficiently in parallel.
-The next Examples apply parallel solvers and then demonstrate the weak scaling of FASCD F-cycles.
+The next Examples apply parallel solvers and investigate the weak scaling of FASCD F-cycles.
 

### 2023-07-28 paper/fascd.tex
edits section 3: another 1/2 page shorter; more to come

@@
 \newcommand{\Div}{\nabla\cdot}
-\newcommand{\trace}{\operatorname{tr}}
 
@@
 
-\newcommand{\bxi}{\bm{\xi}}
-\newcommand{\bzero}{\bm{0}}
-
 \newcommand{\cK}{\mathcal{K}}
@@
 
-\newcommand{\rhoi}{\rho_{\text{i}}}
-
-\newcommand{\ip}[2]{\langle#1,#2\rangle}
+\newcommand{\ip}[2]{\left<#1,#2\right>}
 
@@
 
-\newcommand{\nn}{{\text{n}}}
-\newcommand{\pp}{{\text{p}}}
-\newcommand{\qq}{{\text{q}}}
-\newcommand{\rr}{{\text{r}}}
-
-\newcommand{\supp}{\operatorname{supp}}
-\newcommand{\Span}{\operatorname{span}}
-
 \newcommand{\fascd}{\pr{fascd}\xspace}
@@
 \newcommand{\rSS}{r_{\text{SS}}}
-\newcommand{\rXX}{r_{\text{XX}}}
 
@@
 Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique, and also notice that sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp} because the $\cV_i$ are subspaces.
-Finally, suppose there are bounded, generally nonlinear, \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.
-Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:femultilevel}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
+Finally, suppose there are bounded, generally nonlinear, \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}; our notation avoids conflict with the transfer operators in Section \ref{sec:femultilevel}.} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
 \begin{equation}
@@
 A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} \cite{Tai2003}.
-Figure \ref{fig:cartoon} suggests how a low-dimensional example might look; note $\cK_i \not\subset \cK$ in this and most other cases.
+Figure \ref{fig:cartoon} suggests how a low-dimensional example might look; note $\cK_i \not\subset \cK$ in this and many other cases.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{genfigs/cartoon.pdf}
+\includegraphics[width=0.45\textwidth]{genfigs/cartoon.pdf}
 \caption{A constraint decomposition (CD) for a unilateral obstacle problem on a two-point space $\Omega=\{x_1,x_2\}$, with $\mathcal{V}=\left\{v \,:\, \Omega \to \RR\right\}$ and $\mathcal{K}=\{v\ge \psi\}$.}
@@
 
-In Sections \ref{sec:femultilevel}--\ref{sec:vcycle} we will introduce multilevel FE CD discretizations and algorithms for finite-dimensional VI problems.
-However, the CD concept applies at the level of the continuum problem, as illustrated in the following two examples.
-The first example is an overlapping domain decomposition of a Sobolev space.
-Again note that $\cK_i \not\subset \cK$; informally speaking, the important inclusion is $\cK_i \subset \cV_i$.
-
-\begin{example}  \label{ex:domaindecomposition}  For a bounded domain $\Omega \subset \RR^d$, let $\cV = W_0^{k,p}(\Omega)$ for $k\ge 0$ and $p\ge 1$.
-Let $\psi$ be a measurable, real-valued obstacle, defined on $\Omega$, which satisfies $\psi|_{\partial \Omega} \le 0$, and let $\cK = \{v \ge \psi\} \subset \cV$.
-Suppose that $\{\phi_i\}_{i=0}^{m-1}$ is a smooth partition of unity on $\Omega$, satisfying $0 \le \phi_i\le 1$ and $\sum_i \phi_i = 1$, and let $\Omega_i$ be the open support of $\phi_i$.
-Let $\cV_i = \{w \in \cV:w|_{\Omega \setminus \Omega_i} =0 \}$, $\cK_i = \{v \in \cV_i: v \ge \phi_i \psi\}$, and $\Pi_i(v) = \phi_i v$.
-Then \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} hold, giving a CD.
-\end{example}
-
-Our second example is a disjoint frequency decomposition of a Hilbert space.
-A multilevel CD over a finite-dimensional FE discretization (Sections \ref{sec:femultilevel}--\ref{sec:cdmultilevel}) should approximate this CD.
-
-\begin{example}  \label{ex:frequencydecomposition}  Suppose $\Omega = (0,a)^d \subset \RR^d$ is a cube, and let $\cV = W_{\text{per}}^{2,k}(\Omega)$ be the periodic functions with square-integrable $k$th derivatives, for $k \ge 0$.
-Suppose $\psi \in W_{\text{per}}^{2,k}(\Omega)$ and let $\cK = \{v \ge \psi\} \subset \cV$.
-Without using any detailed notation for Fourier representation, but noting that the frequencies are discrete because $\Omega$ is compact, suppose $\{\cV_i\}$ are $m<\infty$ subspaces of $\cV$ defined by an (nonoverlapping) partition of $W_{\text{per}}^{2,k}(\Omega)$ by frequency, thus satisfying \eqref{eq:subspacedecomp} as an orthogonal decomposition.
-Suppose $P_i:\cV \to \cV_i$ are the corresponding orthogonal projections, satisfying $I = \sum_i P_i$.
-Let $\cK_i = \{v \ge P_i \psi\} \subset \cV_i$ and $\Pi_i = P_i$.
-Then \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} also hold.
-\end{example}
+Although a multilevel FE CD discretization is our purpose (Sections \ref{sec:femultilevel}--\ref{sec:implementation}), the CD concept can be applied at the level of the continuum problem.
+For example, in an obstacle problem ($\cK = \{v \ge \psi\}$) one can construct a overlapping domain decomposition of a Sobolev space $\cV$ by using a smooth, finite partition of unity $\{\varphi_i\}$, subspaces $\cV_i = \{w \in \cV : w = 0 \text{ if } \varphi_i = 0\}$, constraint sets $\cK_i = \{v \in \cV_i : v \ge \varphi_i \psi\}$, and decomposition operators $\Pi_i(v) = \varphi_i v$.
+One may also construct a CD which is a disjoint frequency decomposition of a Hilbert space of periodic functions on a cube $\Omega=(0,1)^d$.
+A multilevel CD over a finite-dimensional FE discretization (Sections \ref{sec:femultilevel}--\ref{sec:cdmultilevel}) will, in vague sense, approximate such a frequency decomposition.
 
 Associated to any CD are certain iterative methods \cite{Tai2003,Xu1992}.
-Let $0\le \alpha \le 1$ be the \emph{damping parameter}.\footnote{FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} are presented without damping, and indeed it seems to be unnecessary.
-However, the convergence of the general CD iterations appears to depend upon it.}  The following algorithms, which replace $u \in \cK$ with a new iterate $w\in\cK$ which should be closer to $u^\star \in \cK$ solving \eqref{eq:vi}, solve smaller VI problems over each subset $\cK_i$.
-
+The following algorithms, which replace $u \in \cK$ with a new iterate $w\in\cK$, which should be closer to $u^\star \in \cK$ solving \eqref{eq:vi}, solve smaller VI problems over each subset $\cK_i$.
 The additive/parallel \pr{cd-add} algorithm generalizes the classical Jacobi iteration.
-In its VI problem the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$, and then adds-back the subset solution $\hat w_i \in \mathcal{K}_i$.
+Note that ``$u-\Pi_iu+\hat w_i$'' removes the $\mathcal{K}_i$ part of $u$ and then adds-back the $\mathcal{K}_i$ solution $\hat w_i$.
+The multiplicative/successive \pr{cd-mult} algorithm generalizes the Gauss--Seidel iteration.
+Each subset solution corrects the global iterate immediately, so here the ordering of the $\cK_i$ is important.
+The sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``those parts of $u$ which need improvement''.
 
@@
 
-The multiplicative/successive \pr{cd-mult} algorithm generalizes the Gauss--Seidel iteration.
-Each subset solution corrects the global iterate immediately, so here the ordering of sets $\{\cK_i\}$ is important.
-In its VI problem the sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``those portions of $u$ which have not yet improved''.
-
 \begin{pseudo*}
@@
         \rm{find} $\hat w_i\in \cK_i$ \rm{such that} \\+
-            $\displaystyle \ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \textnormal{ for all } v_i \in \cK_i$ \\-
+            $\displaystyle \Big<f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big),\, v_i-\hat w_i\Big> \ge \ip{\ell}{v_i-\hat w_i} \textnormal{ for all } v_i \in \cK_i$ \\-
             $w_i = (1-\alpha) \Pi_i u + \alpha \hat w_i\in\cK_i$ \\-
@@
 
-%In VIs \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} the argument of $f$ is an element of $\cK$.
-Furthermore, by \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} one may write the test vector $v_i - \hat w_i \in \cV_i$ as a difference of admissible vectors from $\cK$.
-For \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} respectively:
-%\begin{align}
-%[u - \Pi_i u + v_i] - [u - \Pi_i u + \hat w_i] &= v_i - \hat w_i, \label{eq:admissibledifferenceadd} \\
-%\left[\sum_{j<i} w_j + v_i + \sum_{j>i} \Pi_j u\right] - \left[\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\right] &= v_i - \hat w_i.  \label{eq:admissibledifferencemult}
-%\end{align}
-%In this sense the VIs over the subsets $\mathcal{K}_i$ solve the same problem as the original one.
-
-Let $e_i = \hat w_i - \Pi_i u \in \cV_i$.
-The reader may confirm that $\hat w = u + \sum_{i} e_i$ and thus that $w = u + \alpha \sum_i e_i$.
-The new iterate $w$ is therefore computed by adding a damped correction from each sub\emph{space} $\mathcal{V}_i$, but in such a manner that $w \in \mathcal{K}$; the corrections preserve admissibility.
-
 We call \pr{cd-add} and \pr{cd-mult} the \emph{basic CD iterations}.
-The VI problems solved in these iterations reduce to constrained line searches when the $\cV_i$ are one-dimensional.
-When VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization} then, subject to bounds on the damping $\alpha$, these iterations generate monotonically-decreasing objective values.
+The VI problems in these iterations reduce to constrained line searches when the $\cV_i$ are one-dimensional.
+In fact, let $e_i = \hat w_i - \Pi_i u \in \cV_i$ be the correction from $\cK_i$.
+Noting the damping parameter\footnote{FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} are presented without damping, as it seems to be unnecessary, but the convergence of the general CD iterations depends upon damping.} $0\le \alpha \le 1$, the reader may confirm that $w = u + \alpha \sum_i e_i$.
+Thus the new iterate is computed by adding a correction from each sub\emph{space} $\cV_i$, but so that $w \in \cK$; the corrections preserve admissibility.
+If VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization} then, subject to bounds on $\alpha$, the basic iterations generate monotonically-decreasing objective values.
+
 \begin{lemma} \cite{Tai2003}  Suppose $f=F'$ for $F:\cK\to\RR$ convex and differentiable, and let $u\in\cK$.  For $0 \le \alpha \le 1$ the output $w \in \cK$ from \pr{cd-mult} satisfies
@@
 
-%\begin{proof}  We assume $\ell=0$ without loss of generality; otherwise define $\tilde F(u)=F(u) - \ip{\ell}{u}$ and observe that the proof goes through for $\tilde F$ essentially unchanged.
-%
-%To show that each iteration of the {for} loop in \pr{cd-mult} lowers the objective value, let $z_{i} = \sum_{j\le i} w_j + \sum_{j > i} \Pi_j u \in \cK$, $i=0,\dots,m-1$; this is the state at the end of the $i$th iteration.
-Note that $z_{i-1} = \sum_{j < i} w_j + \Pi_i u + \sum_{j > i} \Pi_j u$ for $i>0$.
-Therefore by convexity of $F$, observation \eqref{eq:admissibledifferencemult} with $v_i=\Pi_i u$, and \eqref{eq:cdmultvi} we find:
-%\begin{equation*}
-%F(z_{i-1}) - F\left(\sum_{j<i} w_j + \hat w_i + \sum_{j > i} \Pi_j u\right) \ge \ip{f\left(\sum_{j<i} w_j + \hat w_i + \sum_{j > i} \Pi_j u\right)}{\Pi_i u - \hat w_i} \ge 0.
-%\end{equation*}
-%Using convexity (again) now gives
-%\begin{align*}
-%F(z_i) &= F\left((1-\alpha) z_{i-1} + \alpha \left(\sum_{j<i} w_j + \hat w_i + \sum_{j > i} \Pi_j u\right)\right) \\
-%&\le  (1-\alpha) F(z_{i-1}) + \alpha F\left(\sum_{j<i} w_j + \hat w_i + \sum_{j > i} \Pi_j u\right) \le F(z_{i-1}).
-%\end{align*}
-%As $u=z_0$ and $w = z_{m-1}$, the result for \pr{cd-mult} follows.
-%
-%Suppose $w=u+\alpha\sum_i e_i$ is computed by \pr{cd-add} using $\alpha \le 1/m$.
-Considering each $e_i=\hat w_i - \Pi_i u$ separately, convexity of $F$ and VI \eqref{eq:cdaddvi} with $v_i=\Pi_i u$ yields
-%\begin{equation}
-%F(u) - F(u+e_i) \ge \ip{f(u+e_i)}{u-(u+e_i)} = \ip{f(u-\Pi_i u + \hat w_i)}{\Pi_i u - \hat w_i} \ge 0. \label{eq:cdaddonereduction}
-%\end{equation}
-%Let $v = u+ \frac{1}{m} \sum_i e_i$.  Again using convexity (Jensen's inequality), and \eqref{eq:cdaddonereduction}, we have
-%\begin{equation*}
-%F(v) = F\left(\frac{1}{m} \sum_{i=0}^{m-1} (u+e_i)\right) \le \frac{1}{m} \sum_{i=0}^{m-1} F(u+e_i) \le \frac{1}{m} \sum_{i=0}^{m-1} F(u) = F(u).
-%\end{equation*}
-%Now define $\lambda = m\alpha$ so $0 \le \lambda \le 1$ and $w = (1-\lambda) u + \lambda v$.  Thus by convexity and the above result, $F(w) \le (1-\lambda) F(u) + \lambda F(v) \le F(u)$.
-%\end{proof}
-%When $\alpha=1/m$ then the \pr{cd-add} proof can be simplified because $v=w$ and $\lambda=1$.
-In this case $F(w)$ is less than $\frac{1}{m} \sum_i F(u+e_i)$ by Jensen's inequality, which is lower than $F(u)$ by \eqref{eq:cdaddvi} or \eqref{eq:cdmultvi}, which gives inequality \eqref{eq:cdaddonereduction}.
-
 The basic CD iterations are meaningful whether or not we are in the optimization case $f=F'$, and even if $f$ is nonlinear, non-local, or defined only on $\cK$.
-However, practical and efficient FE implementation for general operators $f$ seems not to have been addressed in the literature.
-References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem, wherein $f=F'$ is linear, local, and defined on all of $\mathcal{V}$.
-In Section \ref{sec:vcycle} we will extend CD-type iterations to nonlinear operators by applying the full approximation scheme (FAS) idea of Brandt \cite{Brandt1977}, and then illustrate its application to non-optimization operators in Examples \ref{ex:results:advdiff} and \ref{ex:results:sia}.
-In the latter Example $f$ is defined only over a subset $\cK$.
+The practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.
+(References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem, wherein $f=F'$ is linear, local, and defined on all of $\mathcal{V}$.)  In Section \ref{sec:vcycle} we will extend multilevel CD iterations to general nonlinear operators by applying the full approximation scheme (FAS) idea of Brandt \cite{Brandt1977}, and then illustrate its application to general operators in Examples \ref{ex:results:plap}--\ref{ex:results:sia}.
 
@@
 Suppose $\Omega \subset \RR^d$ is open and let $\mathcal{V}=W^{1,p}(\Omega)$.
-The \emph{lower obstacle} $\underline{\gamma}$ and \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions, defined on the closure $\overline{\Omega}$, with values in the extended real line $\tilde\RR = [-\infty,+\infty]$, satisfying the following properties: $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$, $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$, and $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$.
-Also suppose $\partial\Omega$ is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive Hausdorff measure, and $\partial_N \Omega$ possibly empty.
-Dirichlet data\footnote{Dirichlet boundary values are in the trace sense \cite{Evans2010}, and $g_D$ is assumed to be as regular as needed.} is given by $g_D:\partial_D \Omega \to \RR$, and we assume $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.
+The \emph{lower obstacle} $\underline{\gamma}$ and \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions, defined on the closure $\overline{\Omega}$, with values in the extended real line $\tilde\RR = [-\infty,+\infty]$, satisfying the following properties: $\underline{\gamma} < +\infty$, $\overline{\gamma}>-\infty$, and $\underline{\gamma} \le \overline{\gamma}$.
+Also suppose $\partial\Omega$ is split into disjoint sets $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive Hausdorff measure, and $\partial_N \Omega$ possibly empty.
+Dirichlet data\footnote{Boundary values are in the trace sense \cite{Evans2010}, and $g_D$ is assumed to be as regular as needed.} is given by $g_D:\partial_D \Omega \to \RR$, and we assume $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.
 Now define
@@
 \end{equation}
-the closed and convex \emph{constraint set}.
-Note that we are treating inequality constraints and strongly-imposed Dirichlet boundary conditions in a unified manner.
-(Any Neumann conditions over $\partial_N \Omega$ are imposed weakly by the operator $f$, via an integral over $\partial_N\Omega$.)  Next we will numerically solve continuum VI problem \eqref{eq:vi} over a constraint set \eqref{eq:originalconstraintset}.
-In fact, given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we will assume that the following is well-posed for $u^\star\in \cK$:
-\begin{equation}
-\ip{f(u^\star)}{v-u^\star} \ge \ip{\ell}{v-u^\star} \qquad \text{for all } v\in \cK. \label{eq:boxdirichletvi}
-\end{equation}
-Note that $v-u^\star \in W_0^{1,p}(\Omega)$.
+the closed and convex \emph{constraint set}; this treats inequality constraints and Dirichlet conditions in a unified manner.
+(Any Neumann conditions over $\partial_N \Omega$ are imposed weakly by the operator $f$, via an integral over $\partial_N\Omega$.)
+
+Next we will numerically solve VI problem \eqref{eq:vi} over a constraint set \eqref{eq:originalconstraintset}.  Given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we will assume that
+\begin{equation}
+\ip{f(u^\star)}{v-u^\star} \ge \ip{\ell}{v-u^\star} \qquad \text{for all } v\in \cK \label{eq:boxdirichletvi}
+\end{equation}
+is well-posed for $u^\star\in \cK$.  Note that $v-u^\star \in W_0^{1,p}(\Omega)$.
 

### 2023-07-28 paper/fascd.tex
edits into section 5; more to come

@@
 We seek to rapidly compute the approximate solution of box-constrained VI problems \eqref{eq:boxdirichletvi} using finite elements (FE).
-Let $\Omega \subset \RR^d$ be an open bounded polygon.
-A multilevel FE approximation to \eqref{eq:boxdirichletvi} is constructed, using $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, by refinement of a coarse mesh.
-The hierarchy of meshes allows the creation of a multilevel CD of $\cK$ using the ideas in the next two Sections.
+In this Section a multilevel $P_1$ or $Q_1$ element \cite{Elmanetal2014} approximation to \eqref{eq:boxdirichletvi} is constructed by refinement of a coarse mesh.
+The mesh-refinement hierarchy permits the construction of a multilevel CD of the finest-level constraint set.
 
 The mesh hierarchy and FE spaces can be constructed by standard nested refinement.
-Let $\mathcal{T}^0$ be the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate cells with union equal to $\overline{\Omega}$ and maximum cell diameter (characteristic mesh size) $h_0>0$.
+Suppose $\Omega \subset \RR^d$ be an open bounded polygon.
+Let $\mathcal{T}^0$ be the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate cells with union equal to $\overline{\Omega}$ and maximum cell diameter (mesh size) $h_0>0$.
 We assume $\mathcal{T}^0$ is conforming, with no hanging nodes.
-For $j=1,\dots,J$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ so that each $\mathcal{T}^{j-1}$ cell becomes $2^d$ cells in $\mathcal{T}^j$ with mesh size $h_j = 2^{-j} h_0$.
-We call $J\ge 0$ the \emph{depth} of the hierarchy, $J+1$ the \emph{number of levels}, and $\mathcal{T}^J$ the \emph{finest-level mesh}.
-Let $m_j$ be the number of vertices in $\mathcal{T}^j$, so there are $m_j$ \emph{degrees of freedom} at the $j$th level.
-Let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space over $\mathcal{T}^j$, with arbitrary boundary values, with the standard vector space nesting:
+For $j=1,\dots,J$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$; each $\mathcal{T}^{j-1}$ cell becomes $2^d$ cells in $\mathcal{T}^j$ with mesh size $h_j = 2^{-j} h_0$.
+We call $\mathcal{T}^J$ the \emph{finest-level mesh} and $J+1$ the \emph{number of levels}.
+Let $m_j$ be the number of vertices in $\mathcal{T}^j$, the \emph{degrees of freedom} at the $j$th level.
+
+Let $\mathcal{V}^j$ be nested FE spaces over each mesh $\mathcal{T}^j$, with arbitrary boundary values:
 \begin{equation}
@@
 \end{equation}
-On the finest-level mesh let $g_D^J \in \mathcal{V}^J$ satisfy $g_D^J = g_D$ on $\partial_D \Omega$; this restricts $g_D$ in \eqref{eq:originalconstraintset} to be piecewise-linear.
-
 As is common in the discretization of variational inequalities, we employ lowest-order continuous Lagrange elements $P_1$ or $Q_1$ on simplices and tensor-product cells respectively.
-This choice is convenient because this choice enjoys \emph{nodal monotonicity}:
+This choice is convenient because it enjoys \emph{nodal monotonicity}:
 \begin{equation}
@@
 for all $w,z \in \mathcal{V}^j$, where $\bw,\bz \in \RR^{m_j}$ denote vectors of nodal values.
-In other words, to enforce an inequality constraint, with $P_1$ or $Q_1$ we only need enforce it pointwise on the degrees of freedom.
-For $k\ge 2$ there are functions in the standard choices of bases for $P_k$ and $Q_k$ which take on negative values \cite[Figure 1.7]{Elmanetal2014}, violating \eqref{eq:nodalmonotonicity}.
+In other words, to enforce an inequality constraint, with $P_1$ or $Q_1$ we only enforce it pointwise on the degrees of freedom.
+For $P_k$ and $Q_k$, $k\ge 2$, there are functions in the standard bases which take on negative values \cite[Figure 1.7]{Elmanetal2014}, violating \eqref{eq:nodalmonotonicity}.
 
@@
 The finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ are constructed by suitable interpolation of $\underline{\gamma}$, $\overline{\gamma}$ (e.g.~by applying the Cl\'ement interpolation operator~\cite{Carstensen2006}).
-They must satisfy
-\begin{equation}
-\underline{\gamma}^J \le \overline{\gamma}^J, \qquad \underline{\gamma}^J < +\infty, \qquad -\infty < \overline{\gamma}^J. \label{eq:fe:boxconstraintrequirements}
-\end{equation}
-Analogously, we construct the discrete boundary data $g_D^J$ so that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those vertices of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.
+Again they must satisfy $\underline{\gamma}^J < +\infty$, $-\infty < \overline{\gamma}^J$, and $\underline{\gamma}^J \le \overline{\gamma}^J$.
+On the finest-level mesh suppose the discrete boundary data $g_D^J \in \mathcal{V}^J$ satisfies $g_D^J = g_D$ on $\partial_D \Omega$; this restricts $g_D$ in \eqref{eq:originalconstraintset} to be piecewise-linear.
+We assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those vertices of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.
 Then we define
 \begin{equation}
-\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \text{ on vertices of } \mathcal{T}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D^J|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
-\end{equation}
-the finest-level \emph{constraint set} of admissible FE functions. This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.
+\mathcal{K}^J = \big\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \text{ on vertices of } \mathcal{T}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D^J|_{\partial_D\Omega}\big\} \subset \mathcal{V}^J.
+\label{eq:fe:fineconstraintset}
+\end{equation}
+This nonempty, closed, and convex subset is the finest-level \emph{constraint set}.
 
@@
 \end{equation}
-If $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, then the discrete VI \eqref{eq:fe:vi} is well-posed.
-
-While \eqref{eq:fe:vi} approximates \eqref{eq:boxdirichletvi}, we warn the reader that it is usually not conforming for that VI.
+If $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, then \eqref{eq:fe:vi} is well-posed.
+
+While \eqref{eq:fe:vi} approximates \eqref{eq:boxdirichletvi}, we warn the reader that it is usually not conforming.
 That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require or expect $\mathcal{K}^J \subset \mathcal{K}$.
-Regardless of the smoothness of the continuum obstacles $\underline{\gamma}, \overline{\gamma}$, admissibility with respect to their FE interpolants $\underline{\gamma}^J, \overline{\gamma}^J$ will generally not imply continuum admissibility.\footnote{This caveat is especially relevant in geophysical situations where only nodal values of the obstacle are known.
-Nodal elevation data generally misses the actual topographic detail in-between the vertices; compare \cite{Bueler2016}.}
-
-The multilevel algorithm defined in Section \ref{sec:vcycle}, a solver for \eqref{eq:fe:vi}, requires five transfer operators between meshes.
-Of these five operators, three are employed in standard FAS \cite{Trottenbergetal2001}.
-We include Table \ref{tab:transfers} to help the reader keep track.
+Even for continuous obstacles $\underline{\gamma}, \overline{\gamma}$,  admissibility with respect to the FE obstacles $\underline{\gamma}^J, \overline{\gamma}^J$ will generally not imply $\mathcal{K}$-admissibility.
+For a practical example, nodal elevation data in glacier application \cite{Bueler2016} moits the actual topographic detail in-between the vertices.
+
+The multilevel algorithm defined in Section \ref{sec:vcycle} requires five transfer operators between meshes.
+Of these, three are employed in standard FAS \cite{Trottenbergetal2001}; Table \ref{tab:transfers} will help the reader keep track.
+Canonical prolongation $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ implements the vector space inclusion.
+Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $(R\sigma)[z] = \sigma[z]$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
+These canonical transfers generally involve a change of FE representation, and in particular $R$ is ``full-weighting'' \cite{Trottenbergetal2001}.
+To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, note that the vertices of $\mathcal{T}^{j-1}$ are vertices of $\mathcal{T}^j$ and for $z\in\mathcal{V}^j$ define $\iR z$ as the element of $\mathcal{V}^{j-1}$ with the same (coarse) nodal values.
+Injection is monotonic: if $w \le z$ then $\iR w \le \iR z$.
 
@@
 
-Canonical prolongation $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ implements the vector space inclusion $\mathcal{V}^{j-1} \hookrightarrow \mathcal{V}^j$ of coarser-mesh functions.
-Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ also acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $(R\sigma)[z] = \sigma[z]$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
-These canonical transfers generally involve a change of FE representation; $R$ is ``full-weighting'' \cite{Trottenbergetal2001}.
-
-To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, note that the vertices of $\mathcal{T}^{j-1}$ are vertices of $\mathcal{T}^j$.
-For $z\in\mathcal{V}^j$ we define $\iR z$ as the element of $\mathcal{V}^{j-1}$ with the same (coarse) nodal values.
-Injection is monotonic: if $w \le z$ then $\iR w \le \iR z$.
-
 The monotone injections $\maxR,\minR$ \cite{Tai2003,GraeserKornhuber2009}, which will only be applied to constraints in $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the stars of the coarse vertices), Figure \ref{fig:Rplusminus}.
-These operators are constructed to enforce the property
-\begin{equation}
-\minR z \le z \le \maxR z,  \label{eq:fe:monotoneproperty}
-\end{equation}
-which does not hold for the standard injection operator $\iR$.
 To make the construction precise, let $x_q^j$ denote the vertices of mesh $\mathcal{T}^j$, with corresponding basis functions $\psi_q^j(x)$.
@@
 \end{align}
-For standard refinement a coarse-mesh node $x_p^{j-1}$ is in the set of fine vertices $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so \eqref{eq:fe:monotoneproperty} follows.
-In the next Section we use $\maxR,\minR$ to construct ``level defect constraints'' such that the zero function is always admissible.
-This construction depends on an additional ordering property of the monotone injections:
-\begin{equation}
-z\ge 0 \implies \minR z \ge 0, \qquad z \le 0 \implies \maxR z \le 0. \label{eq:fe:monotoneadditional}
+These operators are constructed to enforce the property
+\begin{equation}
+\minR z \le z \le \maxR z,  \label{eq:fe:monotoneproperty}
+\end{equation}
+which does not hold for injection operator $\iR$; that is, for standard refinement a coarse-mesh node $x_p^{j-1}$ is among the fine nodes $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$.
+In the next Section we will use $\maxR,\minR$ to construct ``level defect constraints'' such that the zero function is always admissible.
+This construction depends on additional ordering properties:
+\begin{equation}
+z\ge 0 \implies \minR z \ge 0 \quad \text{ and } \quad z \le 0 \implies \maxR z \le 0. \label{eq:fe:monotoneadditional}
 \end{equation}
@@
 
-A key choice in devising a multilevel scheme for VIs is how to construct coarse-grid bounds.
+A key choice in devising a multilevel scheme for VIs is how to construct coarser-mesh bounds.
 Our multilevel CD scheme uses \emph{defect constraints} \cite{GraeserKornhuber2009} derived from an admissible finest-level iterate.
-The next Example illustrates why we do \emph{not} apply injection operators directly\footnote{An exception is in the F-cycle; see Section \ref{sec:implementation}.} to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.
-Iterate-dependent defect constraints, definition \eqref{eq:fe:defectconstraints} below, satisfy ordering \eqref{eq:fe:chiordering} and bypass these difficulties.
+The next Example illustrates why we do \emph{not} apply injection operators directly\footnote{An exception occurs in the F-cycle; see Section \ref{sec:implementation}.} to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.
 
@@
 Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ vertices, respectively.
-Ignoring the boundary for simplicity, the fine-level constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.
-However, various natural schemes to coarsen the constraints, applied directly to $\underline{\gamma}^1,\overline{\gamma}^1$, are problematic.
+For admissible boundary conditions the fine-level constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.
+However, various schemes to coarsen the constraints by direct transfers of $\underline{\gamma}^1,\overline{\gamma}^1$ are doomed to failure.
 Suppose we propose
@@
 Then $\underline{\gamma}^0=0$ and $\overline{\gamma}^0=C$ identically.
-In that case the corresponding constraint set $\mathcal{K}^0$ is nonempty, but prolongation $Pw^0$ of an admissible solution of the coarse problem ($\underline{\gamma}^0 \le w^0 \le \overline{\gamma}^0$) is not admissible on the finer level: $Pw^0 \notin \mathcal{K}^1$.
-On the other hand, suppose we propose
+In that case the corresponding constraint set $\mathcal{K}^0$ is nonempty, but prolongation $Pw^0$ of an admissible solution of the coarse problem ($w^0\in\mathcal{K}^0$; $\underline{\gamma}^0 \le w^0 \le \overline{\gamma}^0$) is not admissible on the finer level ($Pw^0 \notin \mathcal{K}^1$), and attempting to fix this by projecting into $\mathcal{K}^1$ would introduce high frequencies not present in $w^0$.
+On the other hand, if
 \begin{equation}
@@
 \end{equation}
-The general motivation for this choice---compare \eqref{eq:fe:chilevels} below---is that $w^0 \ge \underline{\gamma}^0 \, \implies \, Pw^0 \ge \underline{\gamma}^1$ and $w^0 \le \overline{\gamma}^0 \, \implies \, Pw^0 \le \overline{\gamma}^1$.
-However, in this case the coarse-level set $\mathcal{K}^0$ is empty because the proposed lower bound $\underline{\gamma}^0$ exceeds the proposed upper bound $\overline{\gamma}^0$ (Figure \ref{fig:directRbad}).
+then $w^0 \ge \underline{\gamma}^0 \, \implies \, Pw^0 \ge \underline{\gamma}^1$ and $w^0 \le \overline{\gamma}^0 \, \implies \, Pw^0 \le \overline{\gamma}^1$.
+However, now the coarse-level constraint set $\mathcal{K}^0$ is empty because $\underline{\gamma}^0 > \overline{\gamma}^0$ (Figure \ref{fig:directRbad}).
 \end{example}
@@
 \caption{For the finer-level constraints $\underline{\gamma}^1$ (blue) and $\overline{\gamma}^1$ (red) on the left, direct application of $\iR,\minR,\maxR$ is problematic.
-For example, if we apply the monotone injections then the resulting coarser-level constraint set $\mathcal{K}^0$ is empty because $\maxR \underline{\gamma}^1 > \minR \overline{\gamma}^1$ (right).
+For example, if we apply the monotone injections then the resulting coarser-level constraint set $\mathcal{K}^0$ is empty (right).
 Defect constraints avoid this difficulty.}
@@
 
-So suppose $w^J \in \cK^J$ is an admissible finest-level iterate for \eqref{eq:fe:vi}.
-We define the \emph{finest-level defect constraints} (DCs) \cite{GraeserKornhuber2009} as the following functions in $\tilde{\mathcal{V}}^J$:
+Iterate-dependent defect constraints will satisfy ordering \eqref{eq:fe:chiordering} below and bypass the difficulties in the above Example.
+Suppose $w^J \in \cK^J$ is an admissible iterate for problem \eqref{eq:fe:vi}.
+We define the \emph{finest-level defect constraints} \cite{GraeserKornhuber2009} as functions in $\tilde{\mathcal{V}}^J$:
 \begin{equation}
@@
 \end{equation}
-Definition \eqref{eq:fe:fineconstraintset} implies that $\underline{\chi}^J \le 0 \le \overline{\chi}^J$ everywhere in $\overline{\Omega}$, including on the Dirichlet boundary $\partial_D\Omega$.
-If $\underline{\gamma}^J=-\infty$ then $\underline{\chi}^J=-\infty$ by definition, and similarly if $\overline{\gamma}^J=+\infty$ then $\overline{\chi}^J=+\infty$.
+(If $\underline{\gamma}^J=-\infty$ then $\underline{\chi}^J=-\infty$ by definition, and similarly if $\overline{\gamma}^J=+\infty$ then $\overline{\chi}^J=+\infty$.)  Definition \eqref{eq:fe:fineconstraintset} implies that $\underline{\chi}^J \le 0 \le \overline{\chi}^J$ everywhere in $\overline{\Omega}$, including on the Dirichlet boundary $\partial_D\Omega$.
 Observe that a corrected iterate $w^J + z^J$ is admissible, $w^J + z^J \in \cK^J$, if and only if the perturbation $z^J \in \mathcal{V}^J$ is zero on $\partial_D\Omega$ and bounded by the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$.
 
-Descending coarser levels, $j=J-1,\dots,1,0$, we next apply the monotone injections to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:\footnote{The definition of $\underline{\chi}^j$ matches that in Section 4 of \cite{GraeserKornhuber2009}; see (4.22) and Figure 4.1 therein.}
+Now descending to coarser levels, $j=J-1,\dots,1,0$, we apply the monotone injections to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:\footnote{The definition of $\underline{\chi}^j$ matches that in Section 4 of \cite{GraeserKornhuber2009}; see (4.22) and Figure 4.1 therein.}
 \begin{equation}
@@
 
-This construction of multilevel box constraints, as for the unilateral construction in \cite{GraeserKornhuber2009}, is designed to permit large-yet-admissible coarse corrections, which promotes multilevel solver efficiency.
-However, when coarsening the problem downward in a V-cycle, one does not know what the next coarse corrections will be, and so we require constraint sets from which any sum will be admissible.
+This construction of multilevel box constraints, likewise for the unilateral construction in \cite{GraeserKornhuber2009}, permits large, yet admissible, coarse corrections, which promotes multilevel solver efficiency.
+However, when coarsening the problem downward in a V-cycle, one cannot predict the upcoming coarse corrections and so one must construct constraint sets from which any sum will be admissible.
 With this in mind, and looking forward to telescoping sums \eqref{eq:fe:telescoping}, we compute the LDC differences \cite{GraeserKornhuber2009}:
@@
 (Take $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ so that $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.
-Regarding subtraction of extended real-valued quantities, if $\underline{\chi}^j(x_p)=-\infty$ then we set $\underline{\phi}^j(x_p)=-\infty$ regardless of the value of $\underline{\chi}^{j-1}(x_p)$, and etc.)  While $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ again bracket zero, $\underline{\phi}^j \le 0 \le \overline{\phi}^j$, they are not ordered as in \eqref{eq:fe:chiordering}.
-
-The corresponding \emph{downward/upward constraint sets}, respectively, are closed and convex subsets of the FE spaces $\mathcal{V}^j$:
+By definition, if $\underline{\chi}^j(x_p)=-\infty$ then we set $\underline{\phi}^j(x_p)=-\infty$ regardless of the value of $\underline{\chi}^{j-1}(x_p)$, and etc.)  Note that while $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ again bracket zero, they are not ordered as in \eqref{eq:fe:chiordering}.
+
+The corresponding \emph{downward} and \emph{upward constraint sets}, respectively, are closed and convex subsets of the FE spaces $\mathcal{V}^j$:
 \begin{align}
@@
 Because $\underline{\chi}^j \le \underline{\phi}^j \le 0 \le \overline{\phi}^j \le \overline{\chi}^j$, it follows that $\mathcal{D}^j \subseteq \mathcal{U}^j$.
-Note that the finest-level set $\mathcal{U}^J$ is equivalent to the original constraint set $\mathcal{K}^J$: for all $z^J \in \mathcal{V}^J$,
-\begin{equation}
-z^J \in \mathcal{U}^J \quad \iff \quad w^J+z^J \in \mathcal{K}^J. \label{eq:fe:finestlevelequivalent}
-\end{equation}
-
-Our V-cycle solver (Algorithm \ref{alg:fascd}) will compute corrections in $\mathcal{D}^j$ for the downward stage and in $\mathcal{U}^j$ for the upward stage.
-This strategy is, however, based on \emph{incomplete} CDs in general box-constrained cases, as we now explain.
-Each upward set $\mathcal{U}^j$ is generally only a superset of a sum of downward sets $\mathcal{D}^j$, though this becomes a true CD, i.e.~satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, in unilateral cases.
-The following Lemma, apparently new, is based in part on the unilateral CD construction underlying Algorithm 4.7 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.
-
-\begin{lemma}  \label{lem:downwardadmissibility}  For arbitrary finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$ satisfying \eqref{eq:fe:boxconstraintrequirements}, it follows that
+Note that the finest-level set $\mathcal{U}^J$ is equivalent to the original constraint set $\mathcal{K}^J$: for all $z^J \in \mathcal{V}^J$ , $z^J \in \mathcal{U}^J$ if and only if $w^J+z^J \in \mathcal{K}^J$.
+
+Our V-cycle solver (Algorithm \ref{alg:fascd}) will compute corrections in $\mathcal{D}^j$ and $\mathcal{U}^j$.
+For general box constraints the upward direction is, however, based on \emph{incomplete} CDs as we now explain.
+Each upward set $\mathcal{U}^j$ is generally only a superset of a sum of downward sets $\mathcal{D}^j$, though this becomes a true CD satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} in unilateral cases.
+The following Lemma, apparently new, extends the unilateral CD construction underlying Algorithm 4.7 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.
+
+\begin{lemma}  \label{lem:downwardadmissibility}
 \begin{equation}
@@
 
-\begin{proof}  Subspace decomposition \eqref{eq:subspacedecomp} holds---see nesting \eqref{eq:fe:nestedspaces}---and definition \eqref{eq:fe:constraintsets} shows $\mathcal{D}^i \subset \mathcal{U}^i \subset \cV^i$.
+\begin{proof}  From nesting \eqref{eq:fe:nestedspaces}, subspace decomposition \eqref{eq:subspacedecomp} holds, and definition \eqref{eq:fe:constraintsets} shows $\mathcal{D}^i \subset \mathcal{U}^i \subset \cV^i$.
 Furthermore, telescoping sums hold for any $j=0,1,\dots,J$:
@@
 
-In unilateral cases we can also construct decomposition operators $\Pi_i$ and thereby show property \eqref{eq:constraintrestrictionsum}.
-For concreteness suppose $\overline{\gamma}^J=+\infty$, thus $\overline{\chi}^j=+\infty$ and $\overline{\phi}^i = +\infty$ for all levels.
-For $v\in \mathcal{V}^j$ and $0\le i \le j$, let $I_{j\to i}^\ominus$ be the operator applying the minimum (monotone) injection $j-i$ times, mapping into $\mathcal{V}^i$:
-\begin{equation}
-I_{j\to i}^\ominus v = \minR(\dots(\minR v)\dots)  \label{eq:fe:minimummaps}
-\end{equation}
-(Set $I_{j\to j}^\ominus v = v$ and $I_{j\to -1}^\ominus=0$ by definition.)  Now define the following nonlinear decomposition operators $\Pi_i:\mathcal{U}^j \to \mathcal{D}^i$:
+In unilateral cases we can also construct decomposition operators $\Pi_i$ and thereby show \eqref{eq:constraintrestrictionsum}.
+For concreteness suppose $\overline{\gamma}^J=+\infty$, thus $\overline{\chi}^j=+\infty$ and $\overline{\phi}^i = +\infty$ for all levels; the $\underline{\gamma}^J=-\infty$ case is similar.
+For $v\in \mathcal{V}^j$ and $0\le i \le j$, let $I_{j\to i}^\ominus$ be the operator applying the minimum (monotone) injection $j-i$ times, mapping into $\mathcal{V}^i$: $I_{j\to i}^\ominus v = \minR(\dots(\minR v)\dots) $.
+(Set $I_{j\to j}^\ominus v = v$ and $I_{j\to -1}^\ominus=0$ by definition.)  Now define nonlinear decomposition operators $\Pi_i:\mathcal{U}^j \to \mathcal{D}^i$:
 \begin{equation}
@@
 \end{align*}
-This shows \eqref{eq:constraintrestrictionsum}, thus that \eqref{eq:fe:downwardsuminclusion} is equality, yielding \eqref{eq:constraintdecomp}.  The $\underline{\gamma}^J=-\infty$ case is similar.
+This shows \eqref{eq:constraintrestrictionsum}, thus \eqref{eq:constraintdecomp}, and that \eqref{eq:fe:downwardsuminclusion} is equality.
 \end{proof}
 
-One might attempt to convert \eqref{eq:fe:downwardsuminclusion} into a full CD for arbitrary box constraints by using decomposition maps like $\Pi_i$, for example by splitting $z^j = \min\{z^j,0\} + \max\{z^j,0\}$ and applying unilateral formulas like \eqref{eq:fe:unilateraldecompositionoperator} to $\min\{z^j,0\}$ and $\max\{z^j,0\}$ separately.
+One might attempt to convert \eqref{eq:fe:downwardsuminclusion} into a full CD for arbitrary box constraints by constructing different decomposition maps $\Pi_i$, for example by splitting $z^j = \min\{z^j,0\} + \max\{z^j,0\}$ and applying unilateral formulas \eqref{eq:fe:unilateraldecompositionoperator} to $\min\{z^j,0\}$ and $\max\{z^j,0\}$ separately.
 The above proof shows that the lower bound $\Pi_i (\min\{z^j,0\}) \ge \underline{\phi}^i$ indeed holds.

### 2023-07-29 paper/fascd.tex
edits in sections 5,6; now 24 pages; more to come

@@
 
-Each upward set $\mathcal{U}^j$ can also be decomposed (incompletely) down to an arbitrary coarser level.
+Each upward set $\mathcal{U}^j$ can also be decomposed, incompletely in the sense of \eqref{eq:fe:downwardsuminclusion}, down to an arbitrary coarser level.
 The next Lemma, which does not follow from Lemma \ref{lem:downwardadmissibility}, justifies upward admissibility in Algorithm \ref{alg:fascd}.
@@
 \end{equation}
-In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, actually $\mathcal{U}^j = \mathcal{U}^k + \sum_{i=k+1}^j \mathcal{D}^i$. \end{lemma}
+In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, actually $\mathcal{U}^j = \mathcal{U}^k + \sum_{i=k+1}^j \mathcal{D}^i$ as a full CD. \end{lemma}
 
 \begin{proof}  Definition \eqref{eq:fe:philevels} shows that $\underline{\chi}^j = \underline{\phi}^j + \dots + \underline{\phi}^{k+1} + \underline{\chi}^k$ and $\overline{\chi}^j = \overline{\phi}^j + \dots + \overline{\phi}^{k+1} + \overline{\chi}^k$.
-If $y^i \in \mathcal{D}^i$ for $k+1 \le i \le j$ and $z^k \in \mathcal{U}^k$ then $\underline{\chi}^j \le y^j + \dots + y^{k+1} + z^k \le \overline{\chi}^j$ and \eqref{eq:fe:upwardsuminclusion}.
-
-In unilateral cases we follow the proof of Lemma \ref{lem:downwardadmissibility}.
-We can use \eqref{eq:fe:unilateraldecompositionoperator} unchanged for $i=k+1,\dots,j$, but it must be modified in the $i=k$ case: $\Pi_k z^j = I_{j\to k}^\ominus(z^j - \underline{\chi}^j) + \underline{\chi}^k$.
-The remainder of the proof goes through as before.
+If $y^i \in \mathcal{D}^i$ for $k+1 \le i \le j$ and $z^k \in \mathcal{U}^k$ then $\underline{\chi}^j \le y^j + \dots + y^{k+1} + z^k \le \overline{\chi}^j$, thus \eqref{eq:fe:upwardsuminclusion} holds.
+In unilateral cases we may follow the proof of Lemma \ref{lem:downwardadmissibility} with \eqref{eq:fe:unilateraldecompositionoperator} unchanged for $i=k+1,\dots,j$, but modified in the $i=k$ case: $\Pi_k z^j = I_{j\to k}^\ominus(z^j - \underline{\chi}^j) + \underline{\chi}^k$.
+Now the proof goes through as before.
 \end{proof}
 
-Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in Algorithm \ref{alg:fascd} below.
-These inclusions are special to multilevel schemes based on defect constraints, and they hold regardless of whether CD operators $\Pi_i$ can be constructed.
-A different ``basic'' iteration, an incomplete variant of \pr{cd-mult}, does not need $\Pi_i$ operators.
-Application of \pr{icd-tele} requires $w^J\in\mathcal{K}^J$, property \eqref{eq:fe:finestlevelequivalent}, and \eqref{eq:fe:upwardsuminclusion}.
+Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} show the admissibility of all iterates and corrections in Algorithm \ref{alg:fascd} below.
+These inclusions are special to multilevel schemes based on defect constraints, and we have seen that they hold regardless of whether decomposition operators $\Pi_i$ can be constructed.
+Thus a new ``basic'' iteration, an incomplete variant of \pr{cd-mult}, can be applied to $w^J\in\mathcal{K}^J$:
 
@@
         \rm{find} $z^j\in \mathcal{U}^j$ \rm{so that for all} $v \in \mathcal{U}^j$, \\+
-            $\displaystyle \ip{f\Big(w^J + \alpha \sum_{i<j} z^i + \alpha z^j\Big)}{v-z^j} \ge \ip{\ell}{v-z^j}$ \\--
+            $\displaystyle \Big<f\Big(w^J + \alpha \sum_{i<j} z^i + \alpha z^j\Big),v-z^j\Big> \ge \ip{\ell}{v-z^j}$ \\--
     return $\tilde w^J=w^J + \alpha \sum_{i\le J} z^i$
@@
 
-A convergence analysis extending the arguments of \cite{GraeserKornhuber2009,Tai2003} to VIs which are not of minimization type, or to box constraints, would seem to require full CDs.
-On the other hand, the observation that up-smoothing is more efficient in a multilevel CD context, illustrated by Example \ref{ex:results:plap} in particular, seems to be new.\footnote{Compare comments on $V(1,1)$ cycles in \cite{GraeserKornhuber2009} and \cite{Tai2003}.}
+Algorithm \ref{alg:fascd} incorporates \pr{icd-tele} for up-smoothing in larger constraint sets.
+The observation that this is more efficient, illustrated by Example \ref{ex:results:plap} in particular, seems to be new; compare comments on $V(1,1)$ cycles in \cite{GraeserKornhuber2009} and \cite{Tai2003}.
+However, a convergence analysis extending the arguments of \cite{GraeserKornhuber2009,Tai2003}, either to non-optimization VIs or box constraints, or both, would seem to require full CDs.
 
@@
 
-The V-cycle applies smoothers in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.
-It extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}, and in fact it reduces to FAS for PDEs when inequality constraints are removed.
-
-Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ by incorporating information from each coarser FE subspace $\mathcal{V}^j$.
-By nesting \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$.\footnote{Nonetheless we use prolongation $P$ to indicate where the computer representation changes.}  The corrections are constructed, inductively, so that iterates remain admissible at every level.
-Suppose $y^i \in \mathcal{D}^i$, for $i=J,\dots,j+1$, are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdcycles}).
+The V-cycle in this Section extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}; it reduces to FAS for PDEs when inequality constraints are removed.
+Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$, via information from each coarser FE subspace $\mathcal{V}^j$, by approximately solving VI problems in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively; by \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$.
+The corrections are inductively constructed so that iterates remain admissible at every level.
+A key idea is that, by inclusion \eqref{eq:fe:upwardsuminclusion}, up-smoothing solves less-restricted VI problems than down-smoothing.
+Going downward one does not know what coarse corrections are coming, thus one must choose smaller constraint sets ($\mathcal{D}^j$) such that the correction sum is admissible once we return to a given level.
+Going upward, however, no further coarse corrections can violate admissibility.
+
+Suppose $y^i \in \mathcal{D}^i$, for $i=J,\dots,j+1$, are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdcycles}, left).
 By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.
-Smoothing then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.
+Relaxation then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.
 Solving the coarsest-level VI results in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.
-Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using \eqref{eq:fe:upwardsuminclusion}.
+Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for relaxation on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using \eqref{eq:fe:upwardsuminclusion}.
 Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.
-Then the corrected finest-level iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.
+The new finest-level iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.
 
@@
 \caption{Left: The FASCD V-cycle Algorithm \ref{alg:fascd} computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.
-Right: F-cycle Algorithm \ref{alg:fascd-fcycle} initially restricts and injects the data $\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$ down to level $j=0$ (gray), then it prolongs and truncates (doubled edges) to create an admissible initial iterate on the next-finer level, and then it applies V-cycles.}
+Right: F-cycle Algorithm \ref{alg:fascd-fcycle} initially restricts and injects the finest-level problem down to the coarsest level (gray).
+It generates each next-finer initial iterate by injection and truncation (doubled edges), then it iterates V-cycles.}
 \label{fig:fascdcycles}
@@
 
-A key idea is that, because of inclusions \eqref{eq:fe:upwardsuminclusion}, up-smoothing solves less-restricted VI problems than down-smoothing.
-Intuitively-speaking, going downward one does not know what the next coarse corrections will be, and thus on any level one must choose conservative constraint sets $\mathcal{D}^j$ so that the sum of the corrections is admissible once we return to a given level.
-Going upward, however, no further coarse corrections can violate admissibility.
-
-To state the V-cycle algorithm below we must also clarify the problems solved at each mesh level.
+To state the V-cycle algorithm we must specify the problems solved at each mesh level.
 As in FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, and a source functional created.
-On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:boxdirichletvi}.\footnote{Presumably it uses quadrature etc.~in the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.}  Finally, we assume that an admissible iterate $w^J \in \mathcal{K}^J$ has already\footnote{In practice the LDCs $\underline{\chi}^j,\overline{\chi}^j,\underline{\phi}^j,\overline{\phi}^j$ are generated during the descent stage; see Algorithm \ref{alg:fascd}.} been used to generate the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}); these LDCs do not depend on the coarse-level corrections.
-
-As we descend we define new $j$th-level iterates
+On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:boxdirichletvi}.\footnote{Presumably it uses quadrature etc.~in the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.}  At the start of each V-cycle we assume that an admissible iterate $w^J \in \mathcal{K}^J$ has been used to generate the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}); they do not depend on the coarse-level corrections.
+
+As we descend we define the initial $j$th-level iterates
 \begin{equation}
@@
 \end{equation}
-(These are the states before smoothing or correction.)  From $w^j$ we define $\ell^j \in (\cV^j)'$:
+From $w^j$ we define $\ell^j \in (\cV^j)'$:
 \begin{equation}
@@
 \end{equation}
-Observe that injection $\iR$ coarsens the (primal) iterates---injection is commonly used in classical FAS \cite[section 5.3]{Trottenbergetal2001}---but canonical restriction $R$ coarsens (dual) linear functionals; recall Table \ref{tab:transfers}.
-On descent, at levels $j=J,J-1,\dots,1$, we solve the following VI for $y^j \in \mathcal{D}^j$:
-\begin{equation}
-\ip{f^j(w^j + y^j)}{v-y^j} \ge \ip{\ell^j}{v-y^j} \qquad \text{for all } v\in \mathcal{D}^j, \label{eq:fe:downvi}
-\end{equation}
-for which the initial iterate is $y_{(0)}^j=0$.  (Zero is admissible; see the comment after \eqref{eq:fe:philevels}.)
+As is typical of FAS \cite[section 5.3]{Trottenbergetal2001}, injection $\iR$ coarsens the (primal) iterates, but canonical restriction $R$ coarsens (dual) linear functionals.
+Descending through the levels $j=J,J-1,\dots,1$ we solve the following VI problems for $y^j \in \mathcal{D}^j$:
+\begin{equation}
+\ip{f^j(w^j + y^j)}{v-y^j} \ge \ip{\ell^j}{v-y^j} \qquad \text{for all } v\in \mathcal{D}^j. \label{eq:fe:downvi}
+\end{equation}
+The initial iterate is $y_{(0)}^j=0$; this is admissible because $\underline{\phi}^j \le 0 \le \overline{\phi}^j$.
 
@@
 \end{equation}
-The nontrivial initial iterate for this problem is found by prolonging $z^{j-1}$ and using it to correct the current (i.e.~down-smoothed) correction on the level:
+However, the nontrivial initial iterate for this problem is found by prolonging $z^{j-1}$ and then updating the current (i.e.~down-smoothed) correction:
 \begin{equation}
@@
 
-%If we were to inject the original box constraints \eqref{eq:fe:fineconstraintset} downward to coarser meshes, i.e.~$\underline{\gamma}^j = \iR \underline{\gamma}^{j+1}$ and $\overline{\gamma}^j = \iR \overline{\gamma}^{j+1}$, then, because injection preserves inequalities, and because $y^{j+1} \in \mathcal{D}^{j+1}$, it would follow from \eqref{eq:fe:definew} that $\underline{\gamma}^j \le w^j \le \overline{\gamma}^j$.
-In this sense $w^j$ is admissible with respect to the box constraints.
-However, while $w^j \in \cV^j \subset \cV^J$ holds, generally $w^j \notin \cK^J$, and furthermore our use of LDCs (Section \ref{sec:cdmultilevel}) avoids the need to generate these $\underline{\gamma}^j, \overline{\gamma}^j$.
-
-Observe that formula \eqref{eq:fe:levelsource} for functional $\ell^j$ is a classical FAS construction for PDEs; compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.
-It has the following explanation, which simultaneously justifies VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.
-First, the $j+1$ level problem \eqref{eq:fe:downvi} is not solved exactly by the finer-level correction $y^{j+1}$---the down-smoother was only approximate---so we seek a further correction $y$ on the $j+1$ level such that
-\begin{equation}
-\ip{f^{j+1}(w^{j+1}+y^{j+1}+y)}{v-y^{j+1}-y} \ge \ip{\ell^{j+1}}{v-y^{j+1}-y} \quad \text{\emph{(notional)}} \label{eq:fe:downvinotional}
-\end{equation}
-for all $v\in \mathcal{D}^j$.  However, in fact $y$ will be approximated by solving a coarser, $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps:
-\begin{enumerate}
-\item subtract the computable quantity $f^{j+1}(w^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides,
-\item replace the residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction,
-\item replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its injection, and
-\item where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.
-\end{enumerate}
-Steps \emph{(ii)}--\emph{(iv)} are justified when the residual and the error are smooth; such smoothness should result from application of the down-smoother.
+Formula \eqref{eq:fe:levelsource} for $\ell^j$ is a classical FAS construction; compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.
+It has the following explanation, which also justifies VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.
+First, problem \eqref{eq:fe:downvi} was not solved exactly by the finer-level correction $y^{j+1}$---the down-smoother was only approximate---so we seek a further correction $y$ on the $j+1$ level such that
+\begin{equation}
+\ip{f^{j+1}(w^{j+1}+y^{j+1}+y)}{v-y^{j+1}-y} \ge \ip{\ell^{j+1}}{v-y^{j+1}-y} \quad \text{\emph{[notional]}} \label{eq:fe:downvinotional}
+\end{equation}
+for all $v\in \mathcal{D}^{j+1}$.
+However, because the down-smoother has made progress, $y$ can be approximated by solving a coarser, $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps: \emph{(i)} subtract the computable quantity $f^{j+1}(w^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides, \emph{(ii)} replace the residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction, \emph{(iii)} replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its injection, and \emph{(iv)} where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.
 These steps yield a problem for the $j$th-level correction $y^j \in \mathcal{D}^j$:
@@
 &\ip{f^j\left(\iR(w^{j+1}+y^{j+1})+y^j\right)}{v-y^j} - \ip{f^j\left(\iR(w^{j+1}+y^{j+1})\right)}{v-y^j} \label{eq:fe:downvicluttered} \\
-&\qquad \ge \ip{R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right)}{v-y^j} \notag
+&\qquad \ge \ip{R\left(\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})\right)}{v-y^j}, \notag
 \end{align}
-This is the same as VI problem \eqref{eq:fe:downvi}, but in cluttered form, motivating definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource}.
-The explanation for VI \eqref{eq:fe:upvi} is similar.
+for all $v\in \mathcal{D}^{j}$.
+This is VI problem \eqref{eq:fe:downvi}, but in cluttered form, motivating definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource}.
+The explanation for problem \eqref{eq:fe:upvi} is similar.
 

### 2023-07-29 paper/fascd.tex
edits to end of section 7; 23 pages, essentially; I have just a bit more work to do in Section 8

@@
 These ideas come together in Algorithm \ref{alg:fascd}, \pr{fascd-vcycle}, which updates $w^J$.
-It assumes that the \pr{smooth} and \pr{solve} procedures do in-place modifications of their final arguments, without modifying their first four arguments.
-Parameters \id{down} and \id{up} determine the number of smoother iterations, respectively.
+The \pr{smooth} and \pr{solve} procedures are assumed to do in-place modifications of their final arguments, without modifying the others.
 
@@
       $y^j = 0$ \\
-      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\underline{\phi}^j,\overline{\phi}^j,w^j;y^j)$  \ct{smoothing of \eqref{eq:fe:downvi} in $\mathcal{D}^j$}\\
+      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\underline{\phi}^j,\overline{\phi}^j,w^j;y^j)$  \ct{problem \eqref{eq:fe:downvi} in $\mathcal{D}^j$}\\
       $w^{j-1} = \iR(w^j + y^j)$ \label{line:vcyclerestrictsolution} \\
@@
     $z^0 = 0$ \\
-    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,\overline{\chi}^0,w^0;z^0)$ \hspace{1.0cm} \ct{solving of \eqref{eq:fe:upvi} in $\mathcal{U}^0$} \\
+    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,\overline{\chi}^0,w^0;z^0)$ \hspace{1.0cm} \ct{problem \eqref{eq:fe:upvi} in $\mathcal{U}^0$} \\
     for $j=1$ to $j=J$ \\+
       $z^j = y^{j} + P z^{j-1}$ \label{line:vcycleupsmoothinitial} \\
-      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,\overline{\chi}^j,w^j;z^j)$  \ct{smoothing of \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
+      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,\overline{\chi}^j,w^j;z^j)$  \ct{problem \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
     return $w^J+z^J$
 \end{pseudo}
-\caption{The FASCD V-cycle is an iteration for solving VI problem \eqref{eq:fe:vi}.  Here $f^j$ denotes an FE discretization of $f$ in problem \eqref{eq:boxdirichletvi}.}
+\caption{The FASCD V-cycle, an iteration for solving VI problem \eqref{eq:fe:vi}.}
 \label{alg:fascd}
@@
 Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle with \id{down}, \id{up} smoother iterations on each level.
-For \id{up} $=0$ it becomes a $V(1,0)$ cycle, and for \id{down} $=0$ a $V(0,1)$ cycle.
-The $V(1,0)$ cycle is similar to Algorithm 4.7 in \cite{GraeserKornhuber2009}, but generalizes it by applying FAS-type corrections and allowing any smoother or nonlinear operator $f$.
-One iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} (Section \ref{sec:cd}) over CD\footnote{Strictly speaking, this statement is restricted to unilateral cases.} \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over incomplete CD \eqref{eq:fe:upwardsuminclusion}, and in this sense $V(1,1)=V(1,0)+V(0,1)$.
-
-Regarding the storage required, let us assume all box constraints are nontrivial.
-In this case, seven vectors must be allocated on each level: $\underline{\chi}^j$, $\overline{\chi}^j$, $\underline{\phi}^j$, $\overline{\phi}^j$, $w^j$, $\ell^j$, $y^j$.
-Note $z^j$ can use the same storage as $y^j$.
-On the finest level one must also store $\underline{\gamma}^J$ and $\overline{\gamma}^J$, while on the coarsest level there are only five vectors because $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.
-Therefore, using $m_j=2^d m_{j-1}$, plus a geometric series argument, the total storage, excluding smoother and coarse-solver implementations, is
-\begin{equation}
-9 m_J + 7 m_{J-1} + \dots + 7 m_1 + 5 m_0 \le \left(9 + \frac{7}{2^d - 1}\right) m_J.
-\end{equation}
-This gives storage at most $16m_J$, $12m_J$, and $10m_J$ in dimensions $d=1,2,3$, respectively.
+The $V(1,0)$ cycle, \id{up} $=0$, is similar to Algorithm 4.7 in \cite{GraeserKornhuber2009}, but generalized by applying FAS-type corrections and allowing any smoother or nonlinear operator $f$.
+In unilateral cases one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} (Section \ref{sec:cd}) over CD \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over CDs \eqref{eq:fe:upwardsuminclusion}, and in this sense $V(1,1)=V(1,0)+V(0,1)$.
+
+Regarding the storage required, if all box constraints are nontrivial then 7 vectors must be allocated on each level: $\underline{\chi}^j$, $\overline{\chi}^j$, $\underline{\phi}^j$, $\overline{\phi}^j$, $w^j$, $\ell^j$, $y^j$; vectors $z^j$ and $y^j$ can use the same storage.
+On the finest level one must also store $\underline{\gamma}^J$ and $\overline{\gamma}^J$.
+Therefore, using $m_j=2^d m_{j-1}$ and a geometric series argument, the total storage is at most
+\begin{equation}
+9 m_J + 7 m_{J-1} + \dots + 7 m_1 + 7 m_0 \le \left(9 + \frac{7}{2^d - 1}\right) m_J,
+\end{equation}
+namely $16m_J$, $12m_J$, and $10m_J$ in dimensions $d=1,2,3$, respectively.
 For comparison, a single-level method needs $4 m_J$ storage (i.e.~$\underline{\gamma}^J,\overline{\gamma}^J,\ell^J,w^J$).
-If some of the obstacles are trivial, for instance in the unilateral case $\overline{\gamma}^J=+\infty$, storage can be correspondingly reduced at all levels.
+In unilateral cases the storage can be reduced at all levels.
 
@@
 
-Implementations of FASCD methods must choose a concrete convergence criterion.
-In this section we also construct the F-cycle extension, and we describe the active set Newton method we will use as a smoother and coarse solver.
-
-Because the problem is inequality-constrained, repeated application of Algorithm \ref{alg:fascd} to solve the finite-dimensional, box-constrained VI \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.
-To construct appropriate convergence criteria for the discretization, note \eqref{eq:fe:vi} is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}.
-By definition, this says that certain node-wise conditions hold for $u^J$ at $x_p \in \mathcal{T}^J$.
+Implementations of FASCD methods must choose a convergence criterion.
+In this section we also construct the F-cycle extension, and we describe the active set Newton method used as the smoother and coarse solver in Section \ref{sec:results}.
+
+Repeated application of Algorithm \ref{alg:fascd} to solve VI problem \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere, but the VI is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}, which says, by definition, that certain nodal conditions hold for $u^J$ at $x_p \in \mathcal{T}^J$.
 Let $r(u^J)_p = \ip{f^J(w^J)-\ell^J}{\psi_p}$ be the pointwise (ordinary) residual.
@@
 \end{itemize}
-These five cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, \emph{pinched} (\emph{both active}), and \emph{boundary}, respectively.
+These 5 cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, \emph{pinched} (\emph{both active}), and \emph{boundary}, respectively.
 Where \emph{pinched} or \emph{boundary} apply, $r(u^J)_p$ can have any value.
 
-This MCP interpretation allows us to construct expressions whose norms should approach zero.
-For $w^J \in \mathcal{K}^J$ we define the finest-level \emph{nodal complementarity residual} $\rNC(w^J)$ from the five cases:
-\begin{equation}
-\rNC(w^J)_p = \begin{cases}
-    r(w^J)_p, & \underline{\gamma}^J(x_p) < w^J(x_p) < \overline{\gamma}^J(x_p), \\
-    \min\{r(w^J)_p,0\}, & w^J(x_p) = \underline{\gamma}^J(x_p), \\
-    \max\{r(w^J)_p,0\}, & w^J(x_p) = \overline{\gamma}^J(x_p), \\
-    0, & \underline{\gamma}^J(x_p)=w^J(x_p)=\overline{\gamma}^J(x_p), \\
-    w^J(x_p) - g_D^J(x_p), & x_p \in \partial_D\Omega. \end{cases} \label{eq:rNC}
-\end{equation}
-The first four cases assume $x_p \notin \partial_D\Omega$.  For PDEs only the first and last apply.
-
-The computable nonlinear map $\rNC : \mathcal{K}^J \to \RR^{m_J}$ in \eqref{eq:rNC} is defined using only the finest-level objects $f^J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J$.
-A value $\rNC(w^J)_p$ is nonzero only at vertices where the MCP is violated, and the formula applies as stated if $\underline{\gamma}^J(x_p)=-\infty$ or $\overline{\gamma}^J(x_p)=+\infty$.
-However, direct implementation of \eqref{eq:rNC} is sensitive to rounding error because $\rNC$ is a discontinuous function of $w^J(x_p)$.
-For example, if $w^J(x_p)$ exceeds $\underline{\gamma}^J(x_p)$ by $10^{-16}$, then evaluation of $\rNC$ would falsely suggest that the iteration is far from converging.
-Thus we instead measure convergence using the Fischer-Burmeister function \cite{BensonMunson2006,Ulbrich2011} on $a,b\in\RR$:
+The MCP interpretation allows us to construct expressions whose norms should approach zero.  Define the Fischer-Burmeister function \cite{BensonMunson2006,Ulbrich2011} on $a,b\in\RR$:
 \begin{equation}
@@
 \end{equation}
-Observe that $\phi$ is semi-smooth, thus continuous, and that $\phi(a,b)=0$ if and only if $a,b$ are complementary.
-For convenience denote $\underline{g}_p = w^J(x_p) - \underline{\gamma}^J(x_p)$ and $\overline{g}_p = \overline{\gamma}^J(x_p) - w^J(x_p)$, the nonnegative gaps at $x_p$.
+Observe that $\phi$ is semi-smooth, thus continuous, and that $\phi(a,b)=0$ if and only if $a,b$ are complementary, i.e.~$a\ge 0$, $b\ge 0$, and $ab=0$.
+For $w^J\in \mathcal{K}^J$ denote $\underline{g}_p = w^J(x_p) - \underline{\gamma}^J(x_p)$ and $\overline{g}_p = \overline{\gamma}^J(x_p) - w^J(x_p)$, the nonnegative \emph{gaps} at $x_p$.
 Now define the \emph{semi-smooth residual}
@@
 Straightforward modifications apply for unconstrained degrees of freedom, where $\overline{\gamma}^J(x_p) = +\infty$ or $\underline{\gamma}^J(x_p) = -\infty$ or both.
-
-Suppose an iteration starts with $w^{J,0} \in \mathcal{K}^J$, generating (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.  Let $\id{atol}\ge 0$, $\id{rtol}\ge 0$ be given tolerances.  We employ
-\begin{equation}
-\|\rSS(w^{J,k})\|_2 < \id{atol} \qquad \text{or} \qquad \frac{\|\rSS(w^{J,k})\|_2}{\|\rSS(w^{J,0})\|_2} < \id{rtol} \label{eq:stoppingcriterion}
-\end{equation}
-as the stopping criterion.
-In Section \ref{sec:results} we will show that iteration of the V-cycle Algorithm \ref{alg:fascd}, starting with some finest-mesh initial iterate $w^{J,0}$ and repeated until \eqref{eq:stoppingcriterion} holds, is an effective solver.
-
-\elb{add \id{stol}}
-
-However, an F-cycle (Figure \ref{fig:fascdcycles}), also called a \emph{full multigrid cycle}, is even better.
-The well-known principle of such a cycle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate.
-Then an admissible initial iterate on the next-finer mesh is generated via prolongation, and so on to finer levels.
-(In the VI context one must also truncate to the bounds after the prolongation; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fcycle}.)  During this ``ramp'' one approximately solves on each level by using a fixed number of V-cycles, each down to the coarsest mesh, and then one does full-depth V-cycles from the finest level, to convergence.
-
-This F-cycle approach is described in Algorithm \ref{alg:fascd-fcycle}.
-In contrast to the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, standard injection is used to generate the $\underline{\gamma}^j,\overline{\gamma}^j$ on each level during the initial nested iteration.
-This is because nested iteration approximates the full solution on the coarse grids, not just a correction to an initial iterate; essentially, our choice is between \eqref{eq:fe:badstandard} (where the coarse constraint set is always nonempty, but the solution may not be admissible on the next level) or \eqref{eq:fe:badmonotone} (where the coarse solution is always admissible if it exists, but the constraint set may be empty).
-We choose \eqref{eq:fe:badstandard} so that the coarse problems always have nonempty constraint sets, and project pointwise to the bounds as we refine in the nested iteration.
-On line \ref{line:fcyclecoarsestinitial}, one may generate an initial iterate $w^0\in\mathcal{K}^0$ by injection and truncation of a finest-level iterate, if one is available.
+VI problem \eqref{eq:fe:vi} is exactly solved by $u^J\in\cK^J$ if and only if $\rSS(u^J)=0$ identically.
+
+Suppose that iteration of Algorithm \ref{alg:fascd} starts with $w^{J,0} \in \mathcal{K}^J$ and generates (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.
+Let $\id{atol}$, $\id{rtol}$, $\id{stol} \ge 0$ be given absolute, relative, and step tolerances, respectively.
+We employ
+\begin{equation}
+\|\rSS(w^{J,k})\|_2 < \id{atol} \quad \text{or} \quad \frac{\|\rSS(w^{J,k})\|_2}{\|\rSS(w^{J,0})\|_2} < \id{rtol} \quad \text{or} \quad \frac{\|\Delta w^{J,k}\|_2}{\|w^{J,k}\|_2} < \id{stol}, \label{eq:stoppingcriterion}
+\end{equation}
+where $\Delta w^{J,k} = w^{J,k} -  w^{J,k-1}$, as the stopping criterion.\footnote{The defaults in our Firedrake implementation match those of PETSc SNES \cite{Balayetal2023}: $\id{atol}=10^{-50}$, $\id{rtol}=10^{-8}$, $\id{stol}=10^{-8}$.
+See Section \ref{sec:results} for specific settings in Examples.}
+
+In Section \ref{sec:results} we will show that V-cycle iterations alone, starting with some finest-mesh initial iterate $w^{J,0}$ and repeated until \eqref{eq:stoppingcriterion} holds, forms an effective solver.
+However, an F-cycle, Algorithm \ref{alg:fascd-fcycle} and Figure \ref{fig:fascdcycles} (right), also called a \emph{full multigrid cycle}, is even better.
+The well-known principle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate $w^0\in\mathcal{K}^0$; one may generate it by injection and projection/truncation of $w^{J,0}$.
+Then an admissible initial iterate on the next-finer mesh is generated via prolongation.
+In the VI context one must also project (truncate) to the bounds after the prolongation; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fcycle}.
+During this nested-iteration ``ramp'' one approximately solves on each level using a fixed number of V-cycles, each down to the coarsest mesh.
+When the ramp is complete, one does full-depth V-cycles to convergence.
 
@@
 
+In contrast to the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, standard injection is used to generate the constraints $\underline{\gamma}^j,\overline{\gamma}^j$ on each level because, during the ramp, a full solution iterate is immediately needed, not just a correction to an initial iterate.
+Essentially, our choice is between \eqref{eq:fe:badstandard} (where the coarse constraint set is always nonempty, but the solution may not be admissible on the next level) or \eqref{eq:fe:badmonotone} (where the coarse solution is always admissible if it exists, but the constraint set may be empty).
+We choose \eqref{eq:fe:badstandard} so that the coarse problems always have nonempty constraint sets, and project (truncate) pointwise to the bounds as we refine.
+
 A large space of smoothers could be explored, but in the next Section we will only apply an active-set (reduced-space) Newton method \cite{Balayetal2023,BensonMunson2006}, denoted as \emph{RS Newton}.
-A fixed number of preconditioned Krylov iterations approximately solve the arising linear systems, thus $O(m_j)$ work per smoother application.
-For the symmetric Laplacian (Example \ref{ex:results:classical}) we apply conjugate gradients (CG) with incomplete Cholesky (IC) preconditioning.
-The generalized minimum residual (GMRES) with incomplete LU (ILU) preconditioning is used in other cases.
-The coarse solver is also RS Newton, but iterated to convergence using the MUMPS sparse direct linear solver~\cite{Amestoy2001}.
+A fixed number of preconditioned Krylov iterations approximately solve the arising linear systems, thus $O(m_j)$ work is done per smoother application.
+For the symmetric Laplacian (Example \ref{ex:results:classical}) we apply conjugate gradients (CG) with incomplete Cholesky (IC) preconditioning, but otherwise generalized minimum residual (GMRES) with incomplete LU (ILU) preconditioning is used.
+The coarse solver is also RS Newton, iterated to convergence using the MUMPS sparse direct linear solver~\cite{Amestoy2001}.
 

### 2023-07-30 paper/fascd.tex
WIP: compare to G&K, and asymp rates figure; waiting on last re-run

@@
 
-\newcommand{\bb}{\mathbf{b}}
-\newcommand{\be}{\mathbf{e}}
-\newcommand{\bbf}{\mathbf{f}}
-\newcommand{\bg}{\mathbf{g}}
 \newcommand{\bn}{\mathbf{n}}
-\newcommand{\br}{\mathbf{r}}
-\newcommand{\bu}{\mathbf{u}}
-\newcommand{\bv}{\mathbf{v}}
 \newcommand{\bw}{\mathbf{w}}
-\newcommand{\bx}{\mathbf{x}}
-\newcommand{\by}{\mathbf{y}}
 \newcommand{\bz}{\mathbf{z}}
-
-\newcommand{\bF}{\mathbf{F}}
-\newcommand{\bV}{\mathbf{V}}
 \newcommand{\bX}{\mathbf{X}}
@@
 
-\newcommand{\rNC}{r_{\text{NC}}}
 \newcommand{\rSS}{r_{\text{SS}}}
@@
 
-\usepackage{todonotes}
-\newcommand{\pef}[1]{\todo[inline, color=red!20]{PF: #1}}
-\newcommand{\elb}[1]{\todo[inline, color=blue!20]{EB: #1}}
+%\usepackage{todonotes}
+%\newcommand{\pef}[1]{\todo[inline, color=red!20]{PF: #1}}
+%\newcommand{\elb}[1]{\todo[inline, color=blue!20]{EB: #1}}
 
@@
 The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so as stated it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
-The latter property is improved by reformulation using defect constraints and monotone injection operators (see \cite[Algorithm 4.7]{GraeserKornhuber2009}, or Section \ref{sec:cdmultilevel} herein).
+The latter property is improved by reformulation using defect constraints and monotone injection operators (see \cite[Algorithm 4.2]{GraeserKornhuber2009}, or Section \ref{sec:cdmultilevel} herein).
 The monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be formulated only with Gauss--Seidel type smoothers, and may have been applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.
@@
 Each upward set $\mathcal{U}^j$ is generally only a superset of a sum of downward sets $\mathcal{D}^j$, though this becomes a true CD satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} in unilateral cases.
-The following Lemma, apparently new, extends the unilateral CD construction underlying Algorithm 4.7 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.
+The following Lemma, apparently new, extends the unilateral CD construction underlying Algorithm 4.2 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.
 
@@
 Algorithm \ref{alg:fascd} incorporates \pr{icd-tele} for up-smoothing in larger constraint sets.
-The observation that this is more efficient, illustrated by Example \ref{ex:results:plap} in particular, seems to be new; compare comments on $V(1,1)$ cycles in \cite{GraeserKornhuber2009} and \cite{Tai2003}.
+The observation that this is more efficient, illustrated by Example \ref{ex:results:plap} in particular, seems to be new.
+Specifically, $V(1,1)$ cycles in \cite{GraeserKornhuber2009} use half-size versions of the smaller sets $\mathcal{D}^j$ both on descent and ascent; see also \cite[section 5.4]{Tai2003}.
 However, a convergence analysis extending the arguments of \cite{GraeserKornhuber2009,Tai2003}, either to non-optimization VIs or box constraints, or both, would seem to require full CDs.
@@
 
-The V-cycle in this Section extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}; it reduces to FAS for PDEs when inequality constraints are removed.
+The V-cycle in this Section extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.2]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}; it reduces to FAS for PDEs when inequality constraints are removed.
 Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$, via information from each coarser FE subspace $\mathcal{V}^j$, by approximately solving VI problems in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively; by \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$.
@@
 Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle with \id{down}, \id{up} smoother iterations on each level.
-The $V(1,0)$ cycle, \id{up} $=0$, is similar to Algorithm 4.7 in \cite{GraeserKornhuber2009}, but generalized by applying FAS-type corrections and allowing any smoother or nonlinear operator $f$.
+The $V(1,0)$ cycle, \id{up} $=0$, is similar to Algorithm 4.2 in \cite{GraeserKornhuber2009}, but generalized by applying FAS-type corrections and allowing any smoother or nonlinear operator $f$.
 In unilateral cases one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} (Section \ref{sec:cd}) over CD \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over CDs \eqref{eq:fe:upwardsuminclusion}, and in this sense $V(1,1)=V(1,0)+V(0,1)$.
@@
 The standard ``ball'' problem has an upper hemisphere obstacle \cite[Chapter 12]{Bueler2021}, but the ``spiral'' problem's obstacle is constructed to produce the coincidence set shown \cite[problem 7.1.1]{GraeserKornhuber2009}.
-For this experiment we used a coarsest (regular triangulation) mesh with $m_0=41$ nodes, and the finest mesh had $m_{10} \approx 3.4 \times 10^7$ nodes.
-The smoother applied a single RS Newton iteration (Section \ref{sec:implementation}), with three IC-preconditioned CG iterations.
+For this experiment we used a regular coarsest mesh with $m_0=41$ nodes (64 triangles), and the finest mesh had $m_{10} \approx 3.4 \times 10^7$ nodes.
+The smoother applied a single RS Newton iteration with three IC-preconditioned CG iterations.
 The initial iterate was $w=\max\{0,\psi\}$.
@@
 Table \ref{tab:results:classical} shows the resulting iteration counts.
-As expected, RS Newton, by itself as a solver, shows rapidly-increasing iterations, roughly proportional to resolution, but V-cycle iterations increase slowly, logarithmically with the resolution.
-These experiments over-solve, with \texttt{rtol} $=$ \texttt{atol} $= 10^{-12}$ in stopping criterion \eqref{eq:stoppingcriterion}.
-The complexity of the coincidence set or free boundary does not appear to strongly influence FASCD V-cycle performance.
-For the spiral problem \cite{GraeserKornhuber2009} reports that for $m_J \approx 5\times 10^5$ resolution \elb{describe GK results}
-
-\elb{asymp rate figure?}
+For a first experiment we over-solve, with \texttt{atol} $=$ \texttt{rtol} $=$ \texttt{stol} $= 10^{-12}$ in stopping criterion \eqref{eq:stoppingcriterion}, to facilitate comparison to results in \cite[section 7]{GraeserKornhuber2009}.
+As expected, RS Newton, by itself as a solver, shows rapidly-increasing iterations, roughly proportional to resolution.
+By contrast, V-cycle iterations increase slowly, logarithmically with the resolution.
+For the spiral problem, at fixed $m_J \approx 5.2 \times 10^5$ resolution, \cite[Figure 7.11]{GraeserKornhuber2009} reports approximately 25 V-cycle iterations for the best (truncated, TNMG/TMMG) monotone multilevel algorithms using a comparable stopping criterion.\footnote{The stopping criterion and asymptotic convergence rates in \cite{GraeserKornhuber2009} are based on norm differences from a last numerical iterate with unknown accuracy.
+Our method measures accuracy directly by a residual norm, namely $\|\rSS(w)\|_2$.}  The 12 V-cycle iterations seen here from comparable resolution (8 levels) represent somewhat more arithmetic because our smoother is RS Newton using CG+IC, rather than projected Gauss-Seidel.
+
+Resolution-dependent asymptotic convergence rates for V-cycles are shown in Figure \ref{fig:results:asymp}.
+We define this rate as as $\rho_J=(\|\rSS(w^{J,k})\|_2/\|\rSS(w^{J,0})\|_2)^{1/k}$ where $w^{J,k}$ is the first iterate to satisfy \eqref{eq:stoppingcriterion}.
+The complexity of the coincidence set or free boundary does not appear to strongly influence FASCD V-cycle performance, as both iterations and rates are similar for the ball and spiral problems.
+Contrast the rate of about $0.1$ for 8 levels with about $0.4$ reported by \cite[Figure 7.11]{GraeserKornhuber2009} for TNMG/TMMG at comparable resolution.
+These rates are all excellent, and our better rate mostly reflects a stronger smoother.
 
 For F-cycles we instead used \texttt{rtol} $= 10^{-8}$, sufficient to reach discretization error, and we report the number of V-cycles after the ramp.
-A small, constant number of cycles solves the problem, though spiral needs more V-cycles to clean-up error remaining after the ramp.
+A small, constant number of cycles apparently solves the problem.
+However, spiral needs more V-cycles to clean-up error remaining after the ramp, presumably from the initial-iterate injection scheme during the ramp (Section \ref{sec:implementation}).
 Since each V-cycle does the work of about eight finest-level Krylov iterations, the ball results show nearly textbook multigrid efficiency at the highest resolutions despite the low solution regularity.
@@
 \includegraphics[width=0.35\textwidth]{fixfigs/ball-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/spiral-set.png}
-\caption{Coincidence sets for the ball and spiral problems in Example \ref{ex:results:classical}.}
+\caption{Coincidence sets for the ball and spiral solutions in Example \ref{ex:results:classical}.}
 \label{fig:results:classical}
@@
 
+%REGENERATE using results/asymprates.py
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.65\textwidth]{genfigs/asymprates.pdf}
+\caption{V-cycle asymptotic convergence rates in Example \ref{ex:results:classical}.}
+\label{fig:results:asymp}
+\end{figure}
+
 \begin{example}  \label{ex:results:plap}
@@
 Iteration results are shown in Table \ref{tab:results:fastplap1d}.
-We see that $V(1,0)$ cycles, Algorithm 4.7 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.
+We see that $V(1,0)$ cycles, Algorithm 4.2 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.
 By contrast, $V(0,1)$ cycles, up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are as good as the default $V(1,1)$ cycles (Algorithm \ref{alg:fascd}).
@@
 \includegraphics[width=0.35\textwidth]{fixfigs/poll2d-zero-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/poll2d-one-set.png}
-\caption{Coincidence sets for the 2D problem in Example \ref{ex:results:advdiff}: $u=0$ (left) and $u=1$ (right).}
+\caption{Coincidence sets for the 2D results in Example \ref{ex:results:advdiff}: $u=0$ (left) and $u=1$ (right).}
 \label{fig:results:advdiff}

### 2023-07-30 paper/genfigs/asymprates.py

@@ -1,0 +1,56 @@
+#!/usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+from writeout import writeout
+
+# from V-cycle results, mcd-extended/paper/results/drape_fascd.txt
+# line J:  [k, |r_ss(w^0)|, |r_ss(w^k)|]
+resnormball = [[ 1, 2.977687712552e+00, 5.592593521511e-16],  # J=0, single level
+               [ 3, 4.597248896467e+00, 9.299029163684e-15],  # J=1, two-level
+               [ 6, 4.019378820472e+00, 3.548401366476e-14],  # ...
+               [ 7, 3.518652347563e+00, 4.898369539597e-13],
+               [ 9, 3.989265817522e+00, 5.346032356444e-13],
+               [11, 5.149110069199e+00, 1.683050131619e-12],
+               [11, 7.072193804686e+00, 6.219113662073e-12],
+               [12, 9.929467647243e+00, 6.630087376663e-12],
+               [13, 1.401524016881e+01, 5.806872030903e-12],
+               [13, 1.981030784949e+01, 1.876069712594e-11],
+               [14, 2.801197654214e+01, 1.004260356580e-11]]
+resnormspir = [[ 1, 1.250364911422e+01, 1.441420557646e-15],
+               [ 4, 3.581030295829e+01, 1.356576646565e-14],
+               [ 5, 5.195274377477e+01, 2.683842615342e-11],
+               [ 7, 7.324445644144e+01, 8.422103502963e-12],
+               [ 9, 1.069563682455e+02, 1.703081752902e-11],
+               [10, 1.517759488345e+02, 6.241169671924e-11],
+               [11, 2.119960943342e+02, 6.428954633272e-11],
+               [12, 3.016400256848e+02, 1.293809842620e-10],
+               [13, 4.275614401535e+02, 1.177739303183e-10],
+               [14, 6.031903763444e+02, 1.310547032348e-10]] # FIXME add final row
+
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["mathtext.fontset"] = "cm"
+
+fsize = 20.0
+
+plt.figure(figsize=(8, 6))
+J = np.arange(1,11) # lev = J + 1, but skip single level J=0
+mcoarse = 4
[…]

### 2023-07-30 paper/results/drape_fascd.sh

@@ -6,14 +6,14 @@ for PROB in ball spiral; do
     echo ""
     echo "******** PROB=${PROB}, V-cycles, tight tolerance (over-solving) *******"
     for LEVS in 1 2 3 4 5 6 7 8 9 10 11; do
-        python3 drape.py -prob $PROB -levs $LEVS -fascd_cycle_type multiplicative -fascd_rtol 1.0e-12
+        python3 drape.py -prob $PROB -levs $LEVS -fascd_cycle_type multiplicative -fascd_monitor -fascd_atol 1.0e-12 -fascd_rtol 1.0e-12 -fascd_stol 1.0e-12
     done
 done
 
 for PROB in ball spiral; do
     echo ""
-    echo "******** PROB=${PROB}, F-cycles, to discretization error, with default rtol 1.0e-8 *******"
+    echo "******** PROB=${PROB}, F-cycles, to discretization error, with defaults rtol=stol=1.0e-8 and atol=1.0e-50 *******"
     for LEVS in 1 2 3 4 5 6 7 8 9 10 11; do
-        python3 drape.py -prob $PROB -levs $LEVS -fascd_cycle_type full
+        python3 drape.py -prob $PROB -levs $LEVS -fascd_cycle_type full -fascd_monitor
     done
 done

### 2023-07-30 paper/fascd.tex
done

@@
 For a first experiment we over-solve, with \texttt{atol} $=$ \texttt{rtol} $=$ \texttt{stol} $= 10^{-12}$ in stopping criterion \eqref{eq:stoppingcriterion}, to facilitate comparison to results in \cite[section 7]{GraeserKornhuber2009}.
-As expected, RS Newton, by itself as a solver, shows rapidly-increasing iterations, roughly proportional to resolution.
-By contrast, V-cycle iterations increase slowly, logarithmically with the resolution.
-For the spiral problem, at fixed $m_J \approx 5.2 \times 10^5$ resolution, \cite[Figure 7.11]{GraeserKornhuber2009} reports approximately 25 V-cycle iterations for the best (truncated, TNMG/TMMG) monotone multilevel algorithms using a comparable stopping criterion.\footnote{The stopping criterion and asymptotic convergence rates in \cite{GraeserKornhuber2009} are based on norm differences from a last numerical iterate with unknown accuracy.
-Our method measures accuracy directly by a residual norm, namely $\|\rSS(w)\|_2$.}  The 12 V-cycle iterations seen here from comparable resolution (8 levels) represent somewhat more arithmetic because our smoother is RS Newton using CG+IC, rather than projected Gauss-Seidel.
-
-Resolution-dependent asymptotic convergence rates for V-cycles are shown in Figure \ref{fig:results:asymp}.
-We define this rate as as $\rho_J=(\|\rSS(w^{J,k})\|_2/\|\rSS(w^{J,0})\|_2)^{1/k}$ where $w^{J,k}$ is the first iterate to satisfy \eqref{eq:stoppingcriterion}.
-The complexity of the coincidence set or free boundary does not appear to strongly influence FASCD V-cycle performance, as both iterations and rates are similar for the ball and spiral problems.
-Contrast the rate of about $0.1$ for 8 levels with about $0.4$ reported by \cite[Figure 7.11]{GraeserKornhuber2009} for TNMG/TMMG at comparable resolution.
-These rates are all excellent, and our better rate mostly reflects a stronger smoother.
-
-For F-cycles we instead used \texttt{rtol} $= 10^{-8}$, sufficient to reach discretization error, and we report the number of V-cycles after the ramp.
-A small, constant number of cycles apparently solves the problem.
-However, spiral needs more V-cycles to clean-up error remaining after the ramp, presumably from the initial-iterate injection scheme during the ramp (Section \ref{sec:implementation}).
+As expected, RS Newton as a solver shows rapidly-increasing iterations, roughly proportional to resolution.
+By contrast, V-cycle iterations increase slowly, perhaps logarithmically with the resolution.
+For the spiral problem, \cite[Figure 7.11]{GraeserKornhuber2009} reports that a $m_J \approx 5.2 \times 10^5$ problem needs approximately 25 V-cycle iterations for the best (truncated, TNMG/TMMG) monotone multilevel algorithms.\footnote{The stopping criterion and asymptotic convergence rates in \cite{GraeserKornhuber2009} are based on norm differences from a last numerical iterate with unknown accuracy.
+Our method measures accuracy directly by a residual norm, namely $\|\rSS(w)\|_2$.}  The 12 V-cycle iterations needed here, at comparable resolution (8 levels), represent somewhat more arithmetic because our smoother is RS Newton using CG+IC, rather than projected Gauss-Seidel.
+Resolution-dependent asymptotic convergence rates for V-cycles, defined as $\rho_J=(\|\rSS(w^{J,k})\|_2/\|\rSS(w^{J,0})\|_2)^{1/k}$ where $w^{J,k}$ is the first iterate to satisfy \eqref{eq:stoppingcriterion}, are shown in Figure \ref{fig:results:asymp}.
+Both iterations and rates are similar for the ball and spiral problems suggesting that the geometric complexity of the coincidence set and free boundary minimally influences FASCD V-cycle performance.
+(The highest-resolution rate difference has no explanation for now.)  Contrast the rate of about $0.1$ for 8 levels here with about $0.4$ reported by \cite[Figure 7.11]{GraeserKornhuber2009} for TNMG/TMMG at comparable resolution.
+These rates are all excellent; our better rate mostly reflects a stronger smoother.
+
+For F-cycles we instead used the defaults \texttt{rtol} $=$ \texttt{stol} $= 10^{-8}$, sufficient to reach discretization error, and in Table \ref{tab:results:classical} we report the number of V-cycles after the ramp; a small, constant number solves the problem.
 Since each V-cycle does the work of about eight finest-level Krylov iterations, the ball results show nearly textbook multigrid efficiency at the highest resolutions despite the low solution regularity.
+More V-cycles are needed to clean-up error remaining after the ramp in the spiral problem, presumably from the F-cycle initial-iterate injection scheme (Section \ref{sec:implementation}).
 \end{example}

### 2023-07-30 paper/genfigs/asymprates.py

@@ -26,7 +26,8 @@ resnormspir = [[ 1, 1.250364911422e+01, 1.441420557646e-15],
                [11, 2.119960943342e+02, 6.428954633272e-11],
                [12, 3.016400256848e+02, 1.293809842620e-10],
                [13, 4.275614401535e+02, 1.177739303183e-10],
-               [14, 6.031903763444e+02, 1.310547032348e-10]] # FIXME add final row
+               [14, 6.031903763444e+02, 1.310547032348e-10],
+               [18, 8.529374614123e+02, 5.472658274103e-10]]
 
 plt.rcParams["font.family"] = "serif"
 plt.rcParams["mathtext.fontset"] = "cm"
@@ -41,11 +42,11 @@ mJ = (mcoarse * 2**J + 1)**2 + (mcoarse * 2**J)**2
 rnball = np.array(resnormball)
 kball = rnball[1:,0]
 rhoball = (rnball[1:,2] / rnball[1:,1])**(1.0/kball)
-plt.semilogx(mJ, rhoball, 'ko:', ms=10.0, label='ball')
+plt.semilogx(mJ, rhoball, 'ko:', ms=10.0, markerfacecolor='w', label='ball')
 rnspir = np.array(resnormspir)
 kspir = rnspir[1:,0]
 rhospir = (rnspir[1:,2] / rnspir[1:,1])**(1.0/kspir)
-plt.semilogx(mJ[:-1], rhospir, 'k*:', ms=12.0, label='spiral')  # FIXME
+plt.semilogx(mJ, rhospir, 'k*:', ms=12.0, label='spiral')
 plt.xlabel(r'degrees of freedom $m_J$', fontsize=fsize)
 plt.ylabel(r'asymptotic rate $\rho_J$', fontsize=fsize)
 #plt.gca().set_xticklabels(mJ)

### 2023-07-31 paper/fascd.tex
moits?
My final editing pass

@@
 
-%\usepackage{todonotes}
-%\newcommand{\pef}[1]{\todo[inline, color=red!20]{PF: #1}}
+\usepackage{todonotes}
+\newcommand{\pef}[1]{\todo[inline, color=red!20]{PF: #1}}
 %\newcommand{\elb}[1]{\todo[inline, color=blue!20]{EB: #1}}
@@
 We present the \emph{full approximation scheme constraint decomposition} (FASCD) multilevel method for solving variational inequalities (VIs).
-FASCD is a common extension both of the full approximation scheme (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and the constraint decomposition (CD) method introduced by X.-C.~Tai for VIs arising in optimization.
+FASCD is a common extension of both the full approximation scheme (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and the constraint decomposition (CD) method introduced by X.-C.~Tai for VIs arising in optimization.
 We extend the CD idea by exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
@@
 In our experiments, FASCD V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for the corresponding unconstrained PDE problems.
-The F-cycle results in Section \ref{sec:results} show near-textbook multigrid efficiency, namely solutions within discretization error in work comparable to few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}.
+The F-cycle results in Section \ref{sec:results} show near-textbook multigrid efficiency, yielding solutions within discretization error in work comparable to a few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}.
 Specifically, we see this performance for classical (unilateral, Laplacian) obstacle problems, $p$-Laplacian obstacle problems, and box-constrained advection-diffusion equations.
@@
 In Section \ref{sec:vi} we recall the theory of coercive and monotone variational inequalities, and describe several motivating examples.
-Section \ref{sec:cd} reviews CDs and their associated iterations.
+Section \ref{sec:cd} reviews CDs and the iterations they induce.
 Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, and then Section \ref{sec:cdmultilevel} extends the CD method via defect constraints generated by monotone injection operators \cite{GraeserKornhuber2009}.
@@
 Section \ref{sec:vcycle} then states the V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, F-cycles, and our smoother choice.
-Section \ref{sec:results} gives numerical results from our Firedrake \cite{Rathgeberetal2016} implementation (see Code Availability).
+Section \ref{sec:results} gives numerical results from our Firedrake \cite{Rathgeberetal2016} implementation.
 
@@
 
-Let $\cV$ be a real and reflexive Banach space with norm $\|\cdot\|$ and topological dual space $\cV'$.
-Denote the dual pairing of $\phi \in \cV'$ and $v\in\cV$ by $\ip{\phi}{v} = \phi(v)$, and define the (Banach space) norm on $\cV'$ by $\|\phi\|' = \sup_{\|v\|=1} |\ip{\phi}{v}|$.
+Let $\cV$ be a real reflexive Banach space with norm $\|\cdot\|$ and topological dual space $\cV'$.
+Denote the dual pairing of $\phi \in \cV'$ and $v\in\cV$ by $\ip{\phi}{v} = \phi(v)$, and define the (Banach space) norm on $\cV'$ by $\|\phi\|' = \sup_{\|v\|=1} |\!\ip{\phi}{v}\!|$.
 Let $\cK \subset \cV$ be a nonempty, closed, and convex subset, the \emph{constraint set}; elements of $\cK$ are said to be \emph{admissible}.
@@
 Let $\Omega \subset \RR^d$ denote a bounded, open set with piecewise-smooth boundary.
-Sobolev spaces \cite{Evans2010} will be denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$, with $W^{1,p}_0(\Omega)$ denoting the subspace having zero boundary values.
+Sobolev spaces \cite{Evans2010} will be denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$, with $W^{1,p}_0(\Omega)$ denoting the subspace with zero trace.
 The following example includes the classical and $p$-Laplacian obstacle problems \cite{ChoeLewis1991}.
@@
 a continuous map \cite[Theorem A.0.6]{Peral1997}.
-It follows from inequalities for $|\cdot|^p$ in $\RR^d$, and the Poincar\'e inequality, that $F$ is $p$-coercive \cite{ChoeLewis1991}; \cite[Appendix A]{Bueler2021conservation}.
+It follows from inequalities for $|\cdot|^p$ in $\RR^d$, and the Poincar\'e inequality, that $f$ is $p$-coercive \cite{ChoeLewis1991}; \cite[Appendix A]{Bueler2021conservation}.
 \end{example}
@@
 For monotone operators $f$, VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
-Suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateau) differentiable with continuous derivative $F':\cK \to \cV'$.
+Suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateaux) differentiable with continuous derivative $F':\cK \to \cV'$.
 Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
@@
 Next we give two examples which are \emph{not} of optimization type \eqref{eq:minimization}, but which are nonetheless $p$-coercive.
-The first is an advection-diffusion problem; it needs the following Lemma.
+The first is an advection-diffusion problem; it requires the following Lemma.
 
@@
 \begin{example}  \label{ex:advectiondiffusion}  Suppose $\partial\Omega$ is partitioned into Dirichlet and Neumann portions, $\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$, with $\partial_D\Omega$ of positive Hausdorff measure.
-Let $\cV \subset W^{1,2}(\Omega)$ be the space of functions which are zero along $\partial_D\Omega$.
+Let $\cV \subset W^{1,2}(\Omega)$ be the space of functions which have zero trace on $\partial_D\Omega$.
 Consider a velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, and assume that the flow is outward on the Neumann boundary: $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.
@@
 \end{equation}
-On the other hand, it is easy to see that $|\ip{f(u)}{v}| \le \left((\eps + \|\bX\|_\infty) \|u\| + \|\phi\|'\right) \|v\|$, so that $f:\cK \to \cV'$ is continuous.
+On the other hand, it is easy to see that $|\!\ip{f(u)}{v}\!| \le \left((\eps + \|\bX\|_\infty) \|u\| + \|\phi\|'\right) \|v\|$, so that $f:\cK \to \cV'$ is continuous.
 Lemma \ref{lem:advectionskew} says that the bilinear form $b(u,v)$ is skew-symmetric up to a nonnegative term.
@@
 Let $a \in L^2(\Omega)$ denote a given ``surface mass balance'' function, the annually-averaged rate of ice accumulation (snow) minus melt and runoff.
-Then the surface elevation $s\in C(\Omega)$ of a steady-state, isothermal, and non-sliding ice sheet (glacier) approximately satisfies the following \emph{shallow ice approximation} (SIA) \cite{GreveBlatter2009} VI problem: find $s \in \mathcal{K} = \{s\ge b\}$ such that
+Then the surface elevation $s\in C(\Omega)$ of a steady-state, isothermal, and non-sliding ice sheet (glacier) under the \emph{shallow ice approximation} (SIA) \cite{GreveBlatter2009} satisfies the following VI problem: find $s \in \mathcal{K} = \{s\ge b\}$ such that
 \begin{equation}
@@
 One may also construct a CD which is a disjoint frequency decomposition of a Hilbert space of periodic functions on a cube $\Omega=(0,1)^d$.
-A multilevel CD over a finite-dimensional FE discretization (Sections \ref{sec:femultilevel}--\ref{sec:cdmultilevel}) will, in vague sense, approximate such a frequency decomposition.
+A multilevel CD over a finite-dimensional FE discretization (Sections \ref{sec:femultilevel}--\ref{sec:cdmultilevel}) will, in a vague sense, approximate such a frequency decomposition.
 
@@
 is well-posed for $u^\star\in \cK$.  Note that $v-u^\star \in W_0^{1,p}(\Omega)$.
+\pef{Is this really true? Shouldn't the test function only match the solution on $\partial_D \Omega$?}
 
@@
 The mesh hierarchy and FE spaces can be constructed by standard nested refinement.
-Suppose $\Omega \subset \RR^d$ be an open bounded polygon.
+Let $\Omega \subset \RR^d$ be an open bounded polygon.
 Let $\mathcal{T}^0$ be the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate cells with union equal to $\overline{\Omega}$ and maximum cell diameter (mesh size) $h_0>0$.
@@
 \end{equation}
-As is common in the discretization of variational inequalities, we employ lowest-order continuous Lagrange elements $P_1$ or $Q_1$ on simplices and tensor-product cells respectively.
+As is common in the discretization of variational inequalities, we employ lowest-order continuous Lagrange elements $P_1$ or $Q_1$ on simplices and tensor-product cells respectively\footnote{Keith \& Surowiec have recently proposed a scheme for solving VIs using high-order finite elements~\cite{Keith2023}.}.
 This choice is convenient because it enjoys \emph{nodal monotonicity}:
@@
 Even for continuous obstacles $\underline{\gamma}, \overline{\gamma}$, admissibility with respect to the FE obstacles $\underline{\gamma}^J, \overline{\gamma}^J$ will generally not imply $\mathcal{K}$-admissibility.
-For a practical example, nodal elevation data in glacier application \cite{Bueler2016} moits the actual topographic detail in-between the vertices.
+For a practical example, nodal elevation data in a glacier application misses the topographic detail in-between the vertices \cite{Bueler2016}.
 
@@
 \toprule
-\emph{name}  & \emph{action}  & \emph{linear?} \\ \midrule
+\emph{name}  & \emph{mapping}  & \emph{linear?} \\ \midrule
 canonical prolongation        & $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ & \,\checkmark \\
@@
 \end{equation}
-which does not hold for injection operator $\iR$; that is, for standard refinement a coarse-mesh node $x_p^{j-1}$ is among the fine nodes $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$.
+which does not hold for the injection operator $\iR$.
 In the next Section we will use $\maxR,\minR$ to construct ``level defect constraints'' such that the zero function is always admissible.
@@
 The V-cycle in this Section extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.2]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}; it reduces to FAS for PDEs when inequality constraints are removed.
-Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$, via information from each coarser FE subspace $\mathcal{V}^j$, by approximately solving VI problems in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively; by \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$.
+Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$, via information from each coarser FE subspace $\mathcal{V}^j$, by approximately solving VI problems in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively; by \eqref{eq:fe:nestedspaces} these corrections are all elements of the fine-level subspace $\mathcal{V}^J$.
 The corrections are inductively constructed so that iterates remain admissible at every level.
@@
 Implementations of FASCD methods must choose a convergence criterion.
-In this section we also construct the F-cycle extension, and we describe the active set Newton method used as the smoother and coarse solver in Section \ref{sec:results}.
+In this section we also construct the F-cycle extension, and we describe the active set Newton method later used as the smoother and coarse solver.
 
@@
 For the spiral problem, \cite[Figure 7.11]{GraeserKornhuber2009} reports that a $m_J \approx 5.2 \times 10^5$ problem needs approximately 25 V-cycle iterations for the best (truncated, TNMG/TMMG) monotone multilevel algorithms.\footnote{The stopping criterion and asymptotic convergence rates in \cite{GraeserKornhuber2009} are based on norm differences from a last numerical iterate with unknown accuracy.
-Our method measures accuracy directly by a residual norm, namely $\|\rSS(w)\|_2$.}  The 12 V-cycle iterations needed here, at comparable resolution (8 levels), represent somewhat more arithmetic because our smoother is RS Newton using CG+IC, rather than projected Gauss-Seidel.
+Our method directly measures satisfaction of the VI by a residual norm, namely $\|\rSS(w)\|_2$.}  The 12 V-cycle iterations needed here, at comparable resolution (8 levels), represent somewhat more arithmetic because our smoother is RS Newton using CG+IC, rather than projected Gauss-Seidel.
 Resolution-dependent asymptotic convergence rates for V-cycles, defined as $\rho_J=(\|\rSS(w^{J,k})\|_2/\|\rSS(w^{J,0})\|_2)^{1/k}$ where $w^{J,k}$ is the first iterate to satisfy \eqref{eq:stoppingcriterion}, are shown in Figure \ref{fig:results:asymp}.
@@
 We applied FASCD F-cycles using $Q_1$ elements and a smoother of four (fixed) iterations of RS Newton (Section \ref{sec:implementation}), with quadratic backtracking line search.
-The arising linear systems were approximately solved by three (fixed) iterations of preconditioned GMRES and a ASM+ILU preconditioner.
+The arising linear systems were approximately solved by three (fixed) iterations of preconditioned GMRES and an ASM+ILU preconditioner.
 The coarse mesh was relatively fine, with $m_0=21^2$ vertices and 400 cells.
@@
 For smoothers of Newton-Krylov type one could also add linear geometric or algebraic multigrid preconditioning \cite{Trottenbergetal2001}, instead of our simple choice of incomplete factorization preconditioners.
-However, the performance seen in Section \ref{sec:results} is already excellent.
+However, the performance observed in Section \ref{sec:results} is already excellent on the tested examples.
 

### 2023-08-02 paper/fascd.tex
remove unneeded detail, and last PF
Remove the cd-add pseudocode, and clean accordingly.
* Remove cd-add.
* Remove concept of a basic iteration.
* Remove damping constant alpha entirely.
* Mention additive/parallel CD iteration from Tai,
  and mention BPX possibility in Discussion.

@@
 
-\usepackage{todonotes}
-\newcommand{\pef}[1]{\todo[inline, color=red!20]{PF: #1}}
+%\usepackage{todonotes}
+%\newcommand{\pef}[1]{\todo[inline, color=red!20]{PF: #1}}
 %\newcommand{\elb}[1]{\todo[inline, color=blue!20]{EB: #1}}
@@
 
-Suppose that there are $m<\infty$ closed subspaces $\cV_i \subset \cV$, so that the sum
+Suppose that there are $m<\infty$ closed linear subspaces $\cV_i \subset \cV$, so that the sum
 \begin{equation}
@@
 Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique, and also notice that sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp} because the $\cV_i$ are subspaces.
-Finally, suppose there are bounded, generally nonlinear, \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}; our notation avoids conflict with the transfer operators in Section \ref{sec:femultilevel}.} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
+Finally, suppose there are bounded, generally nonlinear, \emph{decomposition operators} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
 \begin{equation}
@@
 
-Although a multilevel FE CD discretization is our purpose (Sections \ref{sec:femultilevel}--\ref{sec:implementation}), the CD concept can be applied at the level of the continuum problem.
-For example, in an obstacle problem ($\cK = \{v \ge \psi\}$) one can construct a overlapping domain decomposition of a Sobolev space $\cV$ by using a smooth, finite partition of unity $\{\varphi_i\}$, subspaces $\cV_i = \{w \in \cV : w = 0 \text{ if } \varphi_i = 0\}$, constraint sets $\cK_i = \{v \in \cV_i : v \ge \varphi_i \psi\}$, and decomposition operators $\Pi_i(v) = \varphi_i v$.
+We will construct a multilevel FE CD discretization in Sections \ref{sec:femultilevel}--\ref{sec:implementation}.
+However, the CD concept can be applied at the level of the continuum problem.
+For example, in a unilateral obstacle problem ($\cK = \{v \ge \psi\}$) one can construct a overlapping domain decomposition of $\cV$ by using a smooth, finite partition of unity $\{\varphi_i\}$, subspaces $\cV_i = \{w \in \cV : w = 0 \text{ if } \varphi_i = 0\}$, constraint sets $\cK_i = \{v \in \cV_i : v \ge \varphi_i \psi\}$, and decomposition operators $\Pi_i(v) = \varphi_i v$.
 One may also construct a CD which is a disjoint frequency decomposition of a Hilbert space of periodic functions on a cube $\Omega=(0,1)^d$.
-A multilevel CD over a finite-dimensional FE discretization (Sections \ref{sec:femultilevel}--\ref{sec:cdmultilevel}) will, in a vague sense, approximate such a frequency decomposition.
+The multilevel CD, over a finite-dimensional FE discretization, will, in a vague sense, approximate such a frequency decomposition.
 
 Associated to any CD are certain iterative methods \cite{Tai2003,Xu1992}.
-The following algorithms, which replace $u \in \cK$ with a new iterate $w\in\cK$, which should be closer to $u^\star \in \cK$ solving \eqref{eq:vi}, solve smaller VI problems over each subset $\cK_i$.
-The additive/parallel \pr{cd-add} algorithm generalizes the classical Jacobi iteration.
-Note that ``$u-\Pi_iu+\hat w_i$'' removes the $\mathcal{K}_i$ part of $u$ and then adds-back the $\mathcal{K}_i$ solution $\hat w_i$.
-The multiplicative/successive \pr{cd-mult} algorithm generalizes the Gauss--Seidel iteration.
-Each subset solution corrects the global iterate immediately, so here the ordering of the $\cK_i$ is important.
+The following multiplicative/successive \pr{cd-mult} algorithm, which generalizes the Gauss--Seidel iteration, replaces $u \in \cK$ with a new iterate $w\in\cK$ which should be closer to $u^\star \in \cK$ solving \eqref{eq:vi}.
+The idea is to solve smaller VI problems over each subset $\cK_i$.
+Each subset solution corrects the global iterate immediately, so the ordering of the $\cK_i$ is important.
 The sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``those parts of $u$ which need improvement''.
-
-\begin{pseudo*}
-\pr{cd-add}(u)\text{:} \\+
-    for all $i \in \{0,\dots,m-1\}$: \\+
-        \rm{find} $\hat w_i\in \cK_i$ \rm{such that} \\+
-            $\ip{f(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \textnormal{ for all } v_i \in \cK_i$ \\--
-    $\hat w = \sum_i \hat w_i\in\cK$ \\
-    return $w=(1-\alpha) u + \alpha \hat w$
-\end{pseudo*}
+Tai \cite{Tai2003} also proposes an additive/parallel version of the algorithm, one which we will not use here and whose convergence proof requires substantial damping.
 
@@
     for $i = 0,\dots,m-1$: \\+
-        \rm{find} $\hat w_i\in \cK_i$ \rm{such that} \\+
-            $\displaystyle \Big<f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big),\, v_i-\hat w_i\Big> \ge \ip{\ell}{v_i-\hat w_i} \textnormal{ for all } v_i \in \cK_i$ \\-
-            $w_i = (1-\alpha) \Pi_i u + \alpha \hat w_i\in\cK_i$ \\-
-    $\hat w = \sum_i \hat w_i\in\cK$ \\
-    return $w=(1-\alpha) u + \alpha \hat w$
+        \rm{find} $w_i\in \cK_i$ \rm{such that} \\+
+            $\displaystyle \Big<f\Big(\sum_{j<i} w_j + w_i + \sum_{j>i} \Pi_j u\Big),\, v_i - w_i\Big> \ge \ip{\ell}{v_i - w_i} \textnormal{ for all } v_i \in \cK_i$ \\--
+    return $w=\sum_i w_i\in\cK$
 \end{pseudo*}
 
-We call \pr{cd-add} and \pr{cd-mult} the \emph{basic CD iterations}.
-The VI problems in these iterations reduce to constrained line searches when the $\cV_i$ are one-dimensional.
-In fact, let $e_i = \hat w_i - \Pi_i u \in \cV_i$ be the correction from $\cK_i$.
-Noting the damping parameter\footnote{FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} are presented without damping, as it seems to be unnecessary, but the convergence of the general CD iterations depends upon damping.} $0\le \alpha \le 1$, the reader may confirm that $w = u + \alpha \sum_i e_i$.
-Thus the new iterate is computed by adding a correction from each sub\emph{space} $\cV_i$, but so that $w \in \cK$; the corrections preserve admissibility.
-If VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization} then, subject to bounds on $\alpha$, the basic iterations generate monotonically-decreasing objective values.
-
-\begin{lemma} \cite{Tai2003}  Suppose $f=F'$ for $F:\cK\to\RR$ convex and differentiable, and let $u\in\cK$.  For $0 \le \alpha \le 1$ the output $w \in \cK$ from \pr{cd-mult} satisfies
+The VI problems in \pr{cd-mult} reduce to constrained line searches when the $\cV_i$ are one-dimensional.
+If $e_i = w_i - \Pi_i u \in \cV_i$ denotes the correction from $\cK_i$, then we see that the new iterate $w = u + \sum_i e_i$ is computed by adding a correction from each sub\emph{space} $\cV_i$, but so that $w \in \cK$; the corrections preserve admissibility.
+Furthermore, if VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization} then \pr{cd-mult} generates monotonically-decreasing objective values.
+
+\begin{lemma} \cite{Tai2003}  Suppose $f=F'$ for $F:\cK\to\RR$ convex and differentiable, and let $u\in\cK$.  The output $w \in \cK$ from \pr{cd-mult} satisfies
 \begin{equation}
@@
 \end{equation}
-If $0 \le \alpha \le 1/m$ then \pr{cd-add} likewise satisfies \eqref{eq:objectivemonotone}.
 \end{lemma}
 
-The basic CD iterations are meaningful whether or not we are in the optimization case $f=F'$, and even if $f$ is nonlinear, non-local, or defined only on $\cK$.
-The practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.
-(References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem, wherein $f=F'$ is linear, local, and defined on all of $\mathcal{V}$.)  In Section \ref{sec:vcycle} we will extend multilevel CD iterations to general nonlinear operators by applying the full approximation scheme (FAS) idea of Brandt \cite{Brandt1977}, and then illustrate its application to general operators in Examples \ref{ex:results:plap}--\ref{ex:results:sia}.
+\pr{cd-mult} is meaningful whether or not we are in the optimization case $f=F'$, and even if $f$ is nonlinear, non-local, or defined only on $\cK$, but its practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.
+In particular, references \cite{GraeserKornhuber2009,Tai2003} only apply the multilevel CD method to the classical obstacle problem, wherein $f=F'$ is linear, local, and defined on all of $\mathcal{V}$.
+In Section \ref{sec:vcycle} we will extend multilevel CD iterations to general nonlinear operators by applying the full approximation scheme (FAS) idea of Brandt \cite{Brandt1977}, and then illustrate its application to general operators in Examples \ref{ex:results:plap}--\ref{ex:results:sia}.
 
@@
 \end{equation}
-is well-posed for $u^\star\in \cK$.  Note that $v-u^\star \in W_0^{1,p}(\Omega)$.
-\pef{Is this really true? Shouldn't the test function only match the solution on $\partial_D \Omega$?}
+is well-posed for $u^\star\in \cK$.
 
@@
 These inclusions are special to multilevel schemes based on defect constraints, and we have seen that they hold regardless of whether decomposition operators $\Pi_i$ can be constructed.
-Thus a new ``basic'' iteration, an incomplete variant of \pr{cd-mult}, can be applied to $w^J\in\mathcal{K}^J$:
+Thus we have a new, incomplete variant of \pr{cd-mult}.
+For $w^J\in\mathcal{K}^J$, it applies in the multilevel CD case where \eqref{eq:fe:upwardsuminclusion} holds:
 
@@
         \rm{find} $z^j\in \mathcal{U}^j$ \rm{so that for all} $v \in \mathcal{U}^j$, \\+
-            $\displaystyle \Big<f\Big(w^J + \alpha \sum_{i<j} z^i + \alpha z^j\Big),v-z^j\Big> \ge \ip{\ell}{v-z^j}$ \\--
-    return $\tilde w^J=w^J + \alpha \sum_{i\le J} z^i$
+            $\displaystyle \Big<f\Big(w^J + \sum_{i<j} z^i + z^j\Big),v-z^j\Big> \ge \ip{\ell}{v-z^j}$ \\--
+    return $\tilde w^J=w^J + \sum_{i\le J} z^i$
 \end{pseudo*}
 
-Algorithm \ref{alg:fascd} incorporates \pr{icd-tele} for up-smoothing in larger constraint sets.
+Algorithm \ref{alg:fascd} below incorporates \pr{icd-tele} for up-smoothing in the larger constraint sets $\mathcal{U}^j$.
 The observation that this is more efficient, illustrated by Example \ref{ex:results:plap} in particular, seems to be new.
@@
 The method is smoother-agnostic, and our implementation within the extensible Firedrake \cite{Rathgeberetal2016} and PETSc \cite{Balayetal2023} library framework will allow easy experimentation with other smoothers (projected Gauss--Seidel, semi-smooth Newton, interior-point, etc.), beyond the active-set Newton method chosen for demonstration here.
-For smoothers of Newton-Krylov type one could also add linear geometric or algebraic multigrid preconditioning \cite{Trottenbergetal2001}, instead of our simple choice of incomplete factorization preconditioners.
+For smoothers of Newton-Krylov type one could also add linear geometric or algebraic multigrid preconditioning \cite{Trottenbergetal2001} for the linear step problems, instead of our simpler choice of incomplete factorization preconditioners.
+A restructured multilevel CD approach which applies the additive/parallel CD iteration \cite{Tai2003} might be considered, in pursuit of a VI algorithm analogous to the BPX multigrid method for PDEs \cite{BramblePasciakXu1990}.
 However, the performance observed in Section \ref{sec:results} is already excellent on the tested examples.

### 2023-08-03 paper/fascd.tex
Typos pointed out by Christiana Mavroyiakoumou

@@
 However, the CD concept can be applied at the level of the continuum problem.
-For example, in a unilateral obstacle problem ($\cK = \{v \ge \psi\}$) one can construct a overlapping domain decomposition of $\cV$ by using a smooth, finite partition of unity $\{\varphi_i\}$, subspaces $\cV_i = \{w \in \cV : w = 0 \text{ if } \varphi_i = 0\}$, constraint sets $\cK_i = \{v \in \cV_i : v \ge \varphi_i \psi\}$, and decomposition operators $\Pi_i(v) = \varphi_i v$.
+For example, in a unilateral obstacle problem ($\cK = \{v \ge \psi\}$) one can construct an overlapping domain decomposition of $\cV$ by using a smooth, finite partition of unity $\{\varphi_i\}$, subspaces $\cV_i = \{w \in \cV : w = 0 \text{ if } \varphi_i = 0\}$, constraint sets $\cK_i = \{v \in \cV_i : v \ge \varphi_i \psi\}$, and decomposition operators $\Pi_i(v) = \varphi_i v$.
 One may also construct a CD which is a disjoint frequency decomposition of a Hilbert space of periodic functions on a cube $\Omega=(0,1)^d$.

### 2023-08-03 paper/tikz/directRbadcoarse.tex
Some centering, some acknowledging Lawrence
Typos pointed out by Christiana Mavroyiakoumou

@@
 
-\begin{tikzpicture}[scale=1.5]
+\begin{tikzpicture}[scale=1.0]
   \draw[seabornblue,very thick] (0.0,1.5) -- (4.0,1.5)
@@
   \draw[seabornred,very thick] (0.0,1.0) -- (4.0,1.0)
-                     node [xshift=-6mm,yshift=-4mm] {{\color{black} $\overline{\gamma}^1 = \minR\overline{\gamma}^1$?}};
+                     node [xshift=-6mm,yshift=-4mm] {{\color{black} $\overline{\gamma}^0 = \minR\overline{\gamma}^1$?}};
   \draw[black,thin,->] (-0.3,0.0) -- (4.3,0.0) node [xshift=5mm] {$x$};

### 2023-08-03 paper/fascd.tex
Add time-per cycle in Table 5 to make weak scaling evident
Reword a sentence that confused a reader

@@
 If the bed is flat ($b=0$) then a power transformation converts \eqref{eq:siaoperator} into form \eqref{eq:plaplacian} with $p=n+1$.
-For general beds $b$ less is understood, but existence holds for \eqref{eq:siavi} \cite{JouvetBueler2012} and a power of the solution lives in a Sobolev space: $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.
+Less is understood for general beds $b$, but existence holds for \eqref{eq:siavi} \cite{JouvetBueler2012} and a power of the solution lives in a Sobolev space: $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.
 This theory, and also observations of ice sheets, shows that $|\grad s|$ may be unbounded as one approaches the free boundary from the icy ($s>b$) side.
@@
 \centering
-\begin{tabular}{c@{\hskip 4mm}c@{\hskip 4mm}c@{\hskip 7mm}c@{\hskip 4mm}c}
+\begin{tabular}{c@{\hskip 4mm}c@{\hskip 4mm}c@{\hskip 7mm}c@{\hskip 4mm}cc}
 \toprule
-$P$ & \emph{levels} & $m_J$ & \emph{F-cycles} & \emph{time} (s) \\
+$P$ & \emph{levels} & $m_J$ & \emph{F-cycles} & \emph{time} (s) & \emph{time per cycle} (s) \\
 \midrule
- 1 & 6 & $641^2$ & 3 & 98 \\
- 4 & 7 & $1281^2$ & 3 & 98 \\
- 16 & 8 & $2561^2$ & 3 & 124 \\
- 64 & 9 & $5121^2$ & 3 & 136 \\
- 256 & 10 & $10241^2$ & 2 & 95 \\
- 1024 & 11 & $20481^2$ & 2 & 177 \\
+ 1 & 6 & $641^2$ & 3 & 98 & 32.7 \\
+ 4 & 7 & $1281^2$ & 3 & 98 & 32.7 \\
+ 16 & 8 & $2561^2$ & 3 & 124 & 41.3 \\
+ 64 & 9 & $5121^2$ & 3 & 136 & 45.3 \\
+ 256 & 10 & $10241^2$ & 2 & 95 & 47.5 \\
+ 1024 & 11 & $20481^2$ & 2 & 177 & 88.5 \\
  \bottomrule

### 2023-08-07 paper/fascd.tex
Edits in response to Maryam Parvizi's comments

@@
 \begin{equation}
-v = \sum_{i=0}^{m-1} \Pi_i v.  \label{eq:constraintrestrictionsum}
+v = \sum_{i=0}^{m-1} \Pi_i v,  \label{eq:constraintrestrictionsum}
+\end{equation}
+with the decomposition operators satisfying the stability property
+\begin{equation} \label{eq:decompositionstability}
+\left(\sum_{i=0}^{m-1} \| \Pi_i u - \Pi_i v \|^2\right)^{1/2} \lesssim \|u - v\|.
 \end{equation}
 Clearly \eqref{eq:constraintrestrictionsum} implies sense \emph{(i)} for \eqref{eq:constraintdecomp}.
-A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} \cite{Tai2003}.
+A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:decompositionstability} \cite{Tai2003}.
 Figure \ref{fig:cartoon} suggests how a low-dimensional example might look; note $\cK_i \not\subset \cK$ in this and many other cases.
@@
 We seek to rapidly compute the approximate solution of box-constrained VI problems \eqref{eq:boxdirichletvi} using finite elements (FE).
-In this Section a multilevel $P_1$ or $Q_1$ element \cite{Elmanetal2014} approximation to \eqref{eq:boxdirichletvi} is constructed by refinement of a coarse mesh.
+In this Section a nested multilevel $P_1$ or $Q_1$ element \cite{Elmanetal2014} approximation to \eqref{eq:boxdirichletvi} is constructed by refinement of a coarse mesh.
 The mesh-refinement hierarchy permits the construction of a multilevel CD of the finest-level constraint set.
@@
 
-The monotone injections $\maxR,\minR$ \cite{Tai2003,GraeserKornhuber2009}, which will only be applied to constraints in $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the stars of the coarse vertices), Figure \ref{fig:Rplusminus}.
-To make the construction precise, let $x_q^j$ denote the vertices of mesh $\mathcal{T}^j$, with corresponding basis functions $\psi_q^j(x)$.
+The monotone injections $\maxR,\minR$ \cite{Tai2003,GraeserKornhuber2009}, which will only be applied to constraints in $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the stars of the coarse vertices).
+To make the construction precise, let $x_q^j$ denote the vertices of mesh $\mathcal{T}^j$, with corresponding basis functions $\psi_q^j$.
 For a $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$:
@@
 which does not hold for the injection operator $\iR$.
+An illustration of the construction on nested meshes is given in Figure \ref{fig:Rplusminus}; implementing these operators on non-nested meshes could be efficiently done using a supermesh~\cite{Farrell2011}.
 In the next Section we will use $\maxR,\minR$ to construct ``level defect constraints'' such that the zero function is always admissible.
@@
 For concreteness suppose $\overline{\gamma}^J=+\infty$, thus $\overline{\chi}^j=+\infty$ and $\overline{\phi}^i = +\infty$ for all levels; the $\underline{\gamma}^J=-\infty$ case is similar.
-For $v\in \mathcal{V}^j$ and $0\le i \le j$, let $I_{j\to i}^\ominus$ be the operator applying the minimum (monotone) injection $j-i$ times, mapping into $\mathcal{V}^i$: $I_{j\to i}^\ominus v = \minR(\dots(\minR v)\dots) $.
+For $v\in \mathcal{V}^j$ and $0\le i \le j$, let $I_{j\to i}^\ominus$ be the operator applying the minimum (monotone) injection $j-i$ times, mapping into $\mathcal{V}^i$: $I_{j\to i}^\ominus v = \minR(\cdots(\minR v)\cdots) $.
 (Set $I_{j\to j}^\ominus v = v$ and $I_{j\to -1}^\ominus=0$ by definition.)  Now define nonlinear decomposition operators $\Pi_i:\mathcal{U}^j \to \mathcal{D}^i$:
 \begin{equation}
-\Pi_i z^j = I_{j\to i}^\ominus(z^j - \underline{\chi}^j) - I_{j\to i-1}^\ominus(z^j - \underline{\chi}^j) + \underline{\phi}^i.  \label{eq:fe:unilateraldecompositionoperator}
+\Pi_i z^j \coloneqq I_{j\to i}^\ominus(z^j - \underline{\chi}^j) - I_{j\to i-1}^\ominus(z^j - \underline{\chi}^j) + \underline{\phi}^i.  \label{eq:fe:unilateraldecompositionoperator}
 \end{equation}
@@
 Suppose $z^1$ is a ``sawtooth'' function which takes on value $-a$ on the coarse vertices and value $0$ on the other vertices.
-Then, for any $\overline{\chi}^1\ge 0$, we have $\overline{\chi}^1 \ge 0 \ge z^1\ge \underline{\chi}^1$, so $z^j \in \mathcal{U}^1$.
+Then, for any $\overline{\chi}^1\ge 0$, we have $\overline{\chi}^1 \ge 0 \ge z^1\ge \underline{\chi}^1$, so $z^1 \in \mathcal{U}^1$.
 However, $I_{1\to 0}^\ominus(z^1 - \underline{\chi}^1) = \minR(z^1 - \underline{\chi}^1) = 0$ identically.
@@
 \end{equation}
-However, the nontrivial initial iterate for this problem is found by prolonging $z^{j-1}$ and then updating the current (i.e.~down-smoothed) correction:
+However, the nontrivial initial iterate for this problem is found by prolonging $z^{j-1} \in \mathcal{U}^{j-1}$ and then updating the current (i.e.~down-smoothed) correction:
 \begin{equation}
@@
 \end{align}
+\pef{I don't see where in steps \emph{(i)-(iv)} the expression in the right of the angle brackets changes from $v - y^{j+1} - y$ (as in \eqref{eq:fe:downvinotional}) to $v - y^j$ (as in \eqref{eq:fe:downvicluttered}).
+From comparing the left hand sides, clearly $y^j$ in \eqref{eq:fe:downvicluttered} approximates $y$ in \eqref{eq:fe:downvinotional}, but from comparing the right hand sides, you want $y^j = y^{j+1} + y$.
+Maybe add a sentence, or another step, explaining what happened to the test expression?}
 for all $v\in \mathcal{D}^{j}$.
@@
 Repeated application of Algorithm \ref{alg:fascd} to solve VI problem \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere, but the VI is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}, which says, by definition, that certain nodal conditions hold for $u^J$ at $x_p \in \mathcal{T}^J$.
-Let $r(u^J)_p = \ip{f^J(w^J)-\ell^J}{\psi_p}$ be the pointwise (ordinary) residual.
+Let $r(u^J)_p = \ip{f^J(u^J)-\ell^J}{\psi_p}$ be the pointwise (ordinary) residual.
 If $x_p \notin \partial_D\Omega$ then one of these four conditions must hold:

### 2023-08-07 paper/fascd.tex
connect to stability (3.4) in proof of Lemma 5.2

@@
 \end{equation}
-(Compare equation (4.9) in \cite{GraeserKornhuber2009}.)  Property \eqref{eq:fe:monotoneproperty} implies $I_{j\to i}^\ominus(z^j - \underline{\chi}^j) \ge I_{j\to i-1}^\ominus(z^j - \underline{\chi}^j)$, thus $\Pi_i z^j \ge \underline{\phi}^i$, so $\Pi_i z^j \in \mathcal{D}^i$.
+(Compare equation (4.9) in \cite{GraeserKornhuber2009}.)  Property \eqref{eq:fe:monotoneproperty} implies $I_{j\to i}^\ominus(z^j - \underline{\chi}^j) \ge I_{j\to i-1}^\ominus(z^j - \underline{\chi}^j)$, thus $\Pi_i z^j \ge \underline{\phi}^i$, so $\Pi_i z^j \in \mathcal{D}^i$, and furthermore stability \eqref{eq:decompositionstability} holds \cite[Theorem 4.2]{GraeserKornhuber2009}.
 On the other hand, the following sum telescopes, leaving only its first term plus the sum in \eqref{eq:fe:telescoping}:

### 2023-08-07 paper/fascd.tex
decide that "FMG" is more cultural than "F-cycle" (thanks Max)

@@
 We extend the CD idea by exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
-When a reduced-space active set Newton method is applied as a smoother, with work proportional to the number of unknowns on a given mesh level, FASCD V-cycles exhibit mesh-independent convergence rates and F-cycles (full multigrid) are optimal solvers.
+When a reduced-space active set Newton method is applied as a smoother, with work proportional to the number of unknowns on a given mesh level, FASCD V-cycles exhibit nearly mesh-independent convergence rates, and full multigrid cycles are optimal solvers.
 The example problems include differential operators which are symmetric linear, nonsymmetric linear, and nonlinear, in unilateral and bilateral VI problems.
@@
 \item The method is strictly admissible, and so can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
-\item V-cycles (Algorithm \ref{alg:fascd}) and F-cycles (Algorithm \ref{alg:fascd-fcycle}) are proposed.
+\item V-cycles (Algorithm \ref{alg:fascd}) and full multigrid (FMG; \cite{Trottenbergetal2001}) cycles (Algorithm \ref{alg:fascd-fmg}) are proposed.
 \item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
-\item When a (mesh-dependent) reduced-space active set Newton method is used as relaxation, we demonstrate that F-cycles are an optimal solver for the various examples in Section \ref{sec:results}.
+\item When a (mesh-dependent) reduced-space active set Newton method is used as relaxation, we demonstrate that application of FMG is an optimal solver for the various examples in Section \ref{sec:results}.
 \end{enumerate}
@@
 In our experiments, FASCD V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for the corresponding unconstrained PDE problems.
-The F-cycle results in Section \ref{sec:results} show near-textbook multigrid efficiency, yielding solutions within discretization error in work comparable to a few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}.
+The FMG results in Section \ref{sec:results} show near-textbook multigrid efficiency, yielding solutions within discretization error in work comparable to a few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}.
 Specifically, we see this performance for classical (unilateral, Laplacian) obstacle problems, $p$-Laplacian obstacle problems, and box-constrained advection-diffusion equations.
@@
 For example, we solve an ice sheet geometry problem (Examples \ref{ex:sia} and \ref{ex:results:sia}) in which the ice flow formulas are only meaningful for surface elevation iterates that do not penetrate the bedrock.
-(Non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural modifications of the operator formula.)  FASCD solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a small, mesh-independent number of F-cycle iterations (Example \ref{ex:results:sia}).
+(Non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural modifications of the operator formula.)  FASCD solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a small, mesh-independent number of FMG iterations (Example \ref{ex:results:sia}).
 The same Example investigates the parallel weak scaling \cite{Bueler2021} of our implementation.
@@
 Here we propose an apparently new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD and an associated iteration.
-Section \ref{sec:vcycle} then states the V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, F-cycles, and our smoother choice.
+Section \ref{sec:vcycle} then states the V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, FMG, and our smoother choice.
 Section \ref{sec:results} gives numerical results from our Firedrake \cite{Rathgeberetal2016} implementation.
@@
 
-The numerical results in Section \ref{sec:results} apply FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} to the above Examples.
+The numerical results in Section \ref{sec:results} apply FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fmg} to the above Examples.
 
@@
 Our multilevel CD scheme uses \emph{defect constraints} \cite{GraeserKornhuber2009} derived from an admissible finest-level iterate.
-The next Example illustrates why we do \emph{not} apply injection operators directly\footnote{An exception occurs in the F-cycle; see Section \ref{sec:implementation}.} to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.
+The next Example illustrates why we do \emph{not} apply injection operators directly\footnote{An exception occurs in FMG; see Section \ref{sec:implementation}.} to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.
 
@@
 \caption{Left: The FASCD V-cycle Algorithm \ref{alg:fascd} computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.
-Right: F-cycle Algorithm \ref{alg:fascd-fcycle} initially restricts and injects the finest-level problem down to the coarsest level (gray).
-It generates each next-finer initial iterate by injection and truncation (doubled edges), then it iterates V-cycles.}
+Right: FMG Algorithm \ref{alg:fascd-fmg} generates each next-finer initial iterate by injection and truncation (doubled edges), then it iterates V-cycles.}
 \label{fig:fascdcycles}
@@
 Implementations of FASCD methods must choose a convergence criterion.
-In this section we also construct the F-cycle extension, and we describe the active set Newton method later used as the smoother and coarse solver.
+In this section we also construct the FMG extension, and we describe the active set Newton method later used as the smoother and coarse solver.
 
@@
 In Section \ref{sec:results} we will show that V-cycle iterations alone, starting with some finest-mesh initial iterate $w^{J,0}$ and repeated until \eqref{eq:stoppingcriterion} holds, forms an effective solver.
-However, an F-cycle, Algorithm \ref{alg:fascd-fcycle} and Figure \ref{fig:fascdcycles} (right), also called a \emph{full multigrid cycle}, is even better.
+However, an FMG cycle, Algorithm \ref{alg:fascd-fmg} and Figure \ref{fig:fascdcycles} (right), is even better.
 The well-known principle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate $w^0\in\mathcal{K}^0$; one may generate it by injection and projection/truncation of $w^{J,0}$.
 Then an admissible initial iterate on the next-finer mesh is generated via prolongation.
-In the VI context one must also project (truncate) to the bounds after the prolongation; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fcycle}.
+In the VI context one must also project (truncate) to the bounds after the prolongation; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fmg}.
 During this nested-iteration ``ramp'' one approximately solves on each level using a fixed number of V-cycles, each down to the coarsest mesh.
@@
 \begin{pseudo}
-\pr{fascd-fcycle}(J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J)\text{:} \\+
+\pr{fascd-fmg}(J,\ell^J,\underline{\gamma}^J,\overline{\gamma}^J)\text{:} \\+
     for $j=J$ downto $j=1$ \\+
@@
 \end{pseudo}
-\caption{The FASCD F-cycle for solving VI problem \eqref{eq:fe:vi}.}
-\label{alg:fascd-fcycle}
+\caption{The FASCD full multigrid (FMG) cycle for solving VI problem \eqref{eq:fe:vi}.}
+\label{alg:fascd-fmg}
 \end{pseudofloat}
@@
 
-For $P_1$ and $Q_1$ elements, Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}.
+For $P_1$ and $Q_1$ elements, Algorithms \ref{alg:fascd} and \ref{alg:fascd-fmg} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}.
 The numerical results in this Section, which follow the Section \ref{sec:vi} examples, focus on the scaling of FASCD iterations with mesh resolution.
-We see that each V-cycle greatly reduces the residual norm, and that a few F-cycles solve most of these problems to within discretization error.
+We see that each V-cycle greatly reduces the residual norm, and that a few FMG iterations solve most of these problems to within discretization error.
 
@@
 
-For F-cycles we instead used the defaults \texttt{rtol} $=$ \texttt{stol} $= 10^{-8}$, sufficient to reach discretization error, and in Table \ref{tab:results:classical} we report the number of V-cycles after the ramp; a small, constant number solves the problem.
+For FMG we instead used the defaults \texttt{rtol} $=$ \texttt{stol} $= 10^{-8}$, sufficient to reach discretization error, and in Table \ref{tab:results:classical} we report the number of V-cycles after the ramp; a small, constant number solves the problem.
 Since each V-cycle does the work of about eight finest-level Krylov iterations, the ball results show nearly textbook multigrid efficiency at the highest resolutions despite the low solution regularity.
-More V-cycles are needed to clean-up error remaining after the ramp in the spiral problem, presumably from the F-cycle initial-iterate injection scheme (Section \ref{sec:implementation}).
+More V-cycles are needed to clean-up error remaining after the ramp in the spiral problem, presumably from the FMG initial-iterate injection scheme (Section \ref{sec:implementation}).
 \end{example}
@@
 \multirow{2}{*}{\emph{levels}} & \multirow{2}{*}{$m_J$} & \multicolumn{3}{c}{\,\emph{ball}} & \multicolumn{3}{c}{\,\emph{spiral}} \\ \cmidrule(lr){3-5} \cmidrule(lr){6-8}
-   &                  & RS only & V-cycle & F-cycle & RS only & V-cycle & F-cycle \\
+   &                  & RS only & V-cycle & FMG & RS only & V-cycle & FMG \\
 \midrule
@@
 \bigskip
-\caption{Iterations for FASCD V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}; see text regarding tolerances.  Runs marked \XX\xspace were not attempted.}
+\caption{Iterations for FASCD V-cycle and FMG algorithms in Example \ref{ex:results:classical}; see text regarding tolerances.  Runs marked \XX\xspace were not attempted.}
 \label{tab:results:classical}
@@
 By contrast, $V(0,1)$ cycles, up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are as good as the default $V(1,1)$ cycles (Algorithm \ref{alg:fascd}).
-F-cycles are even more efficient; a single cycle suffices at all higher resolutions.
+FMG is even more efficient; a single iteration suffices at all higher resolutions.
 Note that convergence is at rate $O(h^{2.005})$; compare \cite{ChoeLewis1991}.
@@
 \toprule
-\emph{levels} & $m_J$ & V(1,0) & V(0,1) & V(1,1) & F & $\|$error$\|_\infty$ \\
+\emph{levels} & $m_J$ & V(1,0) & V(0,1) & V(1,1) & FMG & $\|$error$\|_\infty$ \\
 \midrule
@@
 \bigskip
-\caption{Iterations for V- and F-cycles in Example \ref{ex:results:plap}.  NC $=$ did not converge after 50 iterations.}
+\caption{Iterations for V-cycles and FMG in Example \ref{ex:results:plap}.  NC $=$ did not converge after 50 iterations.}
 \label{tab:results:fastplap1d}
@@
 The above Examples, which were run in serial to fix the smoother, also run efficiently in parallel.
-The next Examples apply parallel solvers and investigate the weak scaling of FASCD F-cycles.
+The next Examples apply parallel solvers and investigate the weak scaling of FASCD FMG.
 
@@
 Tolerances were set to \id{rtol} $=10^{-5}$ and \id{atol} $=$ \id{stol} $= 10^{-9}$.
-F-cycle iteration results are shown in Table \ref{tab:results:advdiff}.
-A few F-cycles sufficed up to these memory-limited resolutions.
+FMG iteration results are shown in Table \ref{tab:results:advdiff}; a few iterations sufficed up to these memory-limited resolutions.
 \end{example}
@@
 \bigskip
-\caption{F-cycle iterations for 2D ($P_1$, serial) and 3D ($Q_1$, 16 processes) solutions of the advection-diffusion bi-obstacle VI problem in Example \ref{ex:results:advdiff}.}
+\caption{FMG iterations for 2D ($P_1$, serial) and 3D ($Q_1$, 16 processes) solutions of the advection-diffusion bi-obstacle VI problem in Example \ref{ex:results:advdiff}.}
 \label{tab:results:advdiff}
@@
 
-We applied FASCD F-cycles using $Q_1$ elements and a smoother of four (fixed) iterations of RS Newton (Section \ref{sec:implementation}), with quadratic backtracking line search.
+We applied FASCD FMG using $Q_1$ elements and a smoother of four (fixed) iterations of RS Newton (Section \ref{sec:implementation}), with quadratic backtracking line search.
 The arising linear systems were approximately solved by three (fixed) iterations of preconditioned GMRES and an ASM+ILU preconditioner.
@@
 \toprule
-$P$ & \emph{levels} & $m_J$ & \emph{F-cycles} & \emph{time} (s) & \emph{time per cycle} (s) \\
+$P$ & \emph{levels} & $m_J$ & \emph{iterations} & \emph{time} (s) & \emph{time$/$iterations} (s) \\
 \midrule
@@
 \bigskip
-\caption{Iterations and wall clock time for parallel FASCD solutions of Example \ref{ex:results:sia} on $P$ processes.}
+\caption{Iterations and wall clock time for parallel FASCD FMG solutions of Example \ref{ex:results:sia} on $P$ processes.}
 \label{tab:results:siaweak}

### 2023-08-08 paper/fascd.tex
thank Max
I think this addresses the PF.

@@
 \title{A full approximation scheme multilevel method \\ for nonlinear variational inequalities\thanks{Draft \today.%Submitted to the editors DATE.
-\funding{EB was supported by a Faculty Development Travel Award from United Academics, University of Alaska Fairbanks.
+\funding{EB was supported by a Faculty Development Travel Award from United Academics, University of Alaska Fairbanks, and thanks Max Heldman for helpful comments on the draft.
 PEF was supported by the Engineering and Physical Sciences Research Council [EPSRC grants EP/R029423/1 and EP/W026163/1].
@@
 Formula \eqref{eq:fe:levelsource} for $\ell^j$ is a classical FAS construction; compare equation (8.5b) in \cite{BrandtLivne2011} or equation (5.3.14) in \cite{Trottenbergetal2001}.
-It has the following explanation, which also justifies VI problems \eqref{eq:fe:downvi} and \eqref{eq:fe:upvi}.
-First, problem \eqref{eq:fe:downvi} was not solved exactly by the finer-level correction $y^{j+1}$---the down-smoother was only approximate---so we seek a further correction $y$ on the $j+1$ level such that
-\begin{equation}
-\ip{f^{j+1}(w^{j+1}+y^{j+1}+y)}{v-y^{j+1}-y} \ge \ip{\ell^{j+1}}{v-y^{j+1}-y} \quad \text{\emph{[notional]}} \label{eq:fe:downvinotional}
-\end{equation}
-for all $v\in \mathcal{D}^{j+1}$.
-However, because the down-smoother has made progress, $y$ can be approximated by solving a coarser, $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps: \emph{(i)} subtract the computable quantity $f^{j+1}(w^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides, \emph{(ii)} replace the residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction, \emph{(iii)} replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its injection, and \emph{(iv)} where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.
-These steps yield a problem for the $j$th-level correction $y^j \in \mathcal{D}^j$:
+It has the following explanation, inductively on VI problems \eqref{eq:fe:downvi}.
+Because the down-smoother was only approximate, problem \eqref{eq:fe:downvi} was not solved exactly by the finer-level correction $y^{j+1} \in \mathcal{D}^{j+1}$.
+Suppose we seek an improved correction by adding $y \in \mathcal{D}^{j}$, to give $y^{j+1}+y \in \mathcal{D}^{j+1}+\mathcal{D}^j \subset \cV^{j+1}$, so that
+\begin{equation}
+\ip{f^{j+1}(w^{j+1}+y^{j+1}+y)}{v-y} \ge \ip{\ell^{j+1}}{v-y} \quad \text{\emph{[notional]}} \label{eq:fe:downvinotional}
+\end{equation}
+for all $v \in \mathcal{D}^j$.  (Note $v-y = (y^{j+1}+v)-(y^{j+1}+y)$ and that nesting $\cV^j \subset \cV^{j+1}$ is assumed here.)
+
+However, because the down-smoother has made progress, $y$ can be approximated by solving a coarser, $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps: \emph{(i)} subtract the computable quantity $f^{j+1}(w^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides, \emph{(ii)} replace the relaxed residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction, \emph{(iii)} replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its injection, and \emph{(iv)} where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.
+These steps yield a problem for the $j$th-level correction $y=y^j \in \mathcal{D}^j$:
 \begin{align}
@@
 \end{align}
-\pef{I don't see where in steps \emph{(i)-(iv)} the expression in the right of the angle brackets changes from $v - y^{j+1} - y$ (as in \eqref{eq:fe:downvinotional}) to $v - y^j$ (as in \eqref{eq:fe:downvicluttered}).
-From comparing the left hand sides, clearly $y^j$ in \eqref{eq:fe:downvicluttered} approximates $y$ in \eqref{eq:fe:downvinotional}, but from comparing the right hand sides, you want $y^j = y^{j+1} + y$.
-Maybe add a sentence, or another step, explaining what happened to the test expression?}
-for all $v\in \mathcal{D}^{j}$.
-This is VI problem \eqref{eq:fe:downvi}, but in cluttered form, motivating definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource}.
-The explanation for problem \eqref{eq:fe:upvi} is similar.
+for all $v\in \mathcal{D}^{j}$.  Though in cluttered form, this is VI problem \eqref{eq:fe:downvi}, once definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource} are applied.
+
+The explanation for \eqref{eq:fe:upvi} and \eqref{eq:fe:upwardinitial} is simpler.
+Whether or not problem \eqref{eq:fe:downvi} has been exactly solved by the computed $y^j\in\mathcal{D}^j$, when going upward we can expand problem \eqref{eq:fe:downvi} into the larger set $\mathcal{U}^j \supset \mathcal{D}^j$.
+(One may continue admissibly upward from any approximate solution to \eqref{eq:fe:upvi}.)  Generally $y^j$ is not the solution to \eqref{eq:fe:upvi}; in the optimization case $f=F'$ it would not be optimal in the enlarged constraint set.
+From \eqref{eq:fe:upwardsuminclusion} our current best estimate of the solution to \eqref{eq:fe:upvi} is $z^j \approx y^j + Pz^{j-1}\in \mathcal{U}^j$; this gives the initial iterate for the up-smoother relaxation.
 
@@
 
-Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle with \id{down}, \id{up} smoother iterations on each level.
-The $V(1,0)$ cycle, \id{up} $=0$, is similar to Algorithm 4.2 in \cite{GraeserKornhuber2009}, but generalized by applying FAS-type corrections and allowing any smoother or nonlinear operator $f$.
+Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle with \id{down} and \id{up} smoother iterations on a level, respectively.
+The $V(1,0)$ cycle with \id{up} $=0$ is similar to Algorithm 4.2 in \cite{GraeserKornhuber2009}, but generalized by applying FAS-type corrections and allowing any smoother or nonlinear operator $f$.
 In unilateral cases one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} (Section \ref{sec:cd}) over CD \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over CDs \eqref{eq:fe:upwardsuminclusion}, and in this sense $V(1,1)=V(1,0)+V(0,1)$.

### 2023-08-12 paper/fascd.tex
Fix small typo
final polish in sections 1-4

@@
 \title{A full approximation scheme multilevel method \\ for nonlinear variational inequalities\thanks{Draft \today.%Submitted to the editors DATE.
-\funding{EB was supported by a Faculty Development Travel Award from United Academics, University of Alaska Fairbanks, and thanks Max Heldman for helpful comments on the draft.
+\funding{EB was supported by a Faculty Development Travel Award from United Academics, University of Alaska Fairbanks, and thanks Max Heldman for helpful comments.
 PEF was supported by the Engineering and Physical Sciences Research Council [EPSRC grants EP/R029423/1 and EP/W026163/1].
@@
 We extend the CD idea by exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
-When a reduced-space active set Newton method is applied as a smoother, with work proportional to the number of unknowns on a given mesh level, FASCD V-cycles exhibit nearly mesh-independent convergence rates, and full multigrid cycles are optimal solvers.
+When a reduced-space (active set) Newton method is applied as a smoother, with work proportional to the number of unknowns on a given mesh level, FASCD V-cycles exhibit nearly mesh-independent convergence rates, and full multigrid cycles are optimal solvers.
 The example problems include differential operators which are symmetric linear, nonsymmetric linear, and nonlinear, in unilateral and bilateral VI problems.
@@
 \item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
-\item When a (mesh-dependent) reduced-space active set Newton method is used as relaxation, we demonstrate that application of FMG is an optimal solver for the various examples in Section \ref{sec:results}.
+\item The method is (nonlinear) smoother-agnostic.
+When a (mesh-dependent) reduced-space Newton method \cite{BensonMunson2006} is used as relaxation, we demonstrate that application of FMG is an optimal solver for the examples in Section \ref{sec:results}.
 \end{enumerate}
@@
 The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} applies to linear complementarity problems and projected (truncated) Gauss--Seidel smoothers.
-While FASCD reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed, PFAS reduces to linear multigrid.
-The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so as stated it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
+While FASCD reduces to general, nonlinear FAS for PDEs, i.e.~when the inequality constraints are removed, PFAS reduces to linear multigrid.
+On the other hand, the multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so as stated it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
 The latter property is improved by reformulation using defect constraints and monotone injection operators (see \cite[Algorithm 4.2]{GraeserKornhuber2009}, or Section \ref{sec:cdmultilevel} herein).
 The monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be formulated only with Gauss--Seidel type smoothers, and may have been applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.
-Unlike the fastest variants of monotone multigrid, no zeroing of the basis functions associated with the active set (truncation) is employed in our method.
+Unlike the fastest variants of monotone multigrid, no zeroing of the basis functions associated with the active set (truncation) is employed in FASCD.
 This simplifies the implementation and allows the coarse grid corrections to update the estimate of the active set on the fine grid.
-Numerical evidence indicates that our coarse corrections are effective even when the coarse meshes are too coarse to resolve the active set.
+Numerical evidence indicates that FASCD coarse corrections are effective even when the coarse meshes are too coarse to resolve the active set.
 On a problem for which the truncated Newton/monotone multigrid methods were designed \cite[problem 7.1.1]{GraeserKornhuber2009}, the performance of FASCD is at least as good (Example \ref{ex:results:classical}).
@@
 Specifically, we see this performance for classical (unilateral, Laplacian) obstacle problems, $p$-Laplacian obstacle problems, and box-constrained advection-diffusion equations.
-Furthermore, V-cycle results appear to be insensitive to the geometric complexity of the coincidence set and free boundary.
+Furthermore, V-cycle results are largely insensitive to the geometric complexity of the coincidence set and free boundary.
 
@@
 The paper is organized as follows.
-In Section \ref{sec:vi} we recall the theory of coercive and monotone variational inequalities, and describe several motivating examples.
+In Section \ref{sec:vi} we recall the theory of coercive and monotone VIs, and describe several motivating examples.
 Section \ref{sec:cd} reviews CDs and the iterations they induce.
@@
 Here we propose an apparently new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD and an associated iteration.
-Section \ref{sec:vcycle} then states the V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, FMG, and our smoother choice.
-Section \ref{sec:results} gives numerical results from our Firedrake \cite{Rathgeberetal2016} implementation.
+Section \ref{sec:vcycle} then states the V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, FMG cycles, and our smoother choice.
+Section \ref{sec:results} gives numerical results from our open source Python/Firedrake \cite{Rathgeberetal2016} implementation.
 
@@
 \end{equation}
-Because $f$ is (generally) nonlinear, the source term $\ell$ is not strictly needed in stating this class of problems---by redefining $f$ we may take $\ell=0$---but introducing $\ell$ clarifies the algorithms.
+Because $f$ is (generally) nonlinear, the source term $\ell$ is not strictly needed in stating this class of problems---by redefining $f$ we may take $\ell=0$---but introducing $\ell$ clarifies the algorithms which follow.
 
@@
 
-\begin{definition}  A map $f:\cK \to \cV'$ is \emph{monotone} if
+\begin{definition}  An operator $f:\cK \to \cV'$ is \emph{monotone} if
 \begin{equation}
@@
 As is standard in the calculus of variations \cite{Evans2010}, coercivity permits a compactness argument for unbounded sets $\cK$; recall that the closed and bounded subsets of a reflexive Banach space are weakly compact.
-However, some of the VIs solved in this paper satisfy a stronger inequality than \eqref{eq:coercive}.
+Some VIs solved in this paper satisfy a stronger inequality than \eqref{eq:coercive}.
 
@@
 
-Next we give two examples which are \emph{not} of optimization type \eqref{eq:minimization}, but which are nonetheless $p$-coercive.
-The first is an advection-diffusion problem; it requires the following Lemma.
+Next we give two examples which are \emph{not} of optimization type \eqref{eq:minimization}.  The first is an advection-diffusion problem; it requires the following Lemma.
 
@@
 On the other hand, it is easy to see that $|\!\ip{f(u)}{v}\!| \le \left((\eps + \|\bX\|_\infty) \|u\| + \|\phi\|'\right) \|v\|$, so that $f:\cK \to \cV'$ is continuous.
-Lemma \ref{lem:advectionskew} says that the bilinear form $b(u,v)$ is skew-symmetric up to a nonnegative term.
-By the outward flow assumption and the Poincar\'e inequality,
+Lemma \ref{lem:advectionskew} says that the bilinear form $b(u,v)$ is skew-symmetric up to a term which is nonnegative by the outward flow assumption.
+From the Poincar\'e inequality,
 \begin{align*}
@@
 Thus $f$ is 2-coercive, and so the associated VI problem is well-posed.
-If $\bX \ne 0$ then $f\ne F'$ for any objective $F$, because $\ip{f(u)}{v}$ does not possess the symmetry of the gradient of a scalar: for $u,v$ which are zero on $\partial \Omega$, note that $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.
+If $\bX \ne 0$ then $f\ne F'$ for any objective $F$, because $\ip{f(u)}{v}$ does not possess the symmetry of a gradient: for $u,v$ which are zero on $\partial \Omega$, note that $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.
 References \cite{Bueler2021conservation,ChangNakshatrala2017} consider similar advection-diffusion VI problems over $\cK = \{v\ge 0\}$.
@@
 The next VI example is a model for steady ice sheet geometry in a given climate.
-It is nonlinear and not of optimization type.
-Related VI problems arise for any fluid layer which is subject to boundary processes that can remove fluid mass \cite{Bueler2021conservation}.
+It is strongly nonlinear and not of optimization type.
+Related VI problems arise for any fluid layer subject to processes which can remove fluid mass \cite{Bueler2021conservation}.
 
@@
 If the bed is flat ($b=0$) then a power transformation converts \eqref{eq:siaoperator} into form \eqref{eq:plaplacian} with $p=n+1$.
-Less is understood for general beds $b$, but existence holds for \eqref{eq:siavi} \cite{JouvetBueler2012} and a power of the solution lives in a Sobolev space: $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.
-This theory, and also observations of ice sheets, shows that $|\grad s|$ may be unbounded as one approaches the free boundary from the icy ($s>b$) side.
+Less is understood for general beds $b$, but existence holds for the VI \cite{JouvetBueler2012}, and a power of the solution lives in a Sobolev space: $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.
+This theory, and also observations of ice sheets, shows that $|\grad s|$ is usually unbounded as one approaches the free boundary from the icy ($s>b$) side.
 \end{example}
 
-The numerical results in Section \ref{sec:results} apply FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fmg} to the above Examples.
+The numerical results in Section \ref{sec:results} will apply FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fmg} to the above Examples.
 
@@
 Each subset solution corrects the global iterate immediately, so the ordering of the $\cK_i$ is important.
-The sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``those parts of $u$ which need improvement''.
-Tai \cite{Tai2003} also proposes an additive/parallel version of the algorithm, one which we will not use here and whose convergence proof requires substantial damping.
 
@@
 
+\noindent The sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``those parts of $u$ which need improvement''.
+Tai \cite{Tai2003} also proposes an additive/parallel version of the algorithm, but it is not used here.
+
 The VI problems in \pr{cd-mult} reduce to constrained line searches when the $\cV_i$ are one-dimensional.
-If $e_i = w_i - \Pi_i u \in \cV_i$ denotes the correction from $\cK_i$, then we see that the new iterate $w = u + \sum_i e_i$ is computed by adding a correction from each sub\emph{space} $\cV_i$, but so that $w \in \cK$; the corrections preserve admissibility.
+If $e_i = w_i - \Pi_i u \in \cV_i$ denotes the correction from $\cK_i$, then the new iterate $w = u + \sum_i e_i$ is computed by adding a correction from each sub\emph{space} $\cV_i$, but so that $w \in \cK$; the corrections preserve admissibility.
 Furthermore, if VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization} then \pr{cd-mult} generates monotonically-decreasing objective values.
@@
 
-From now on we restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bilateral obstacle problems}.
-Suppose $\Omega \subset \RR^d$ is open and let $\mathcal{V}=W^{1,p}(\Omega)$.
+From now on we restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bilateral obstacle problems}, over $\mathcal{V}=W^{1,p}(\Omega)$.
 The \emph{lower obstacle} $\underline{\gamma}$ and \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions, defined on the closure $\overline{\Omega}$, with values in the extended real line $\tilde\RR = [-\infty,+\infty]$, satisfying the following properties: $\underline{\gamma} < +\infty$, $\overline{\gamma}>-\infty$, and $\underline{\gamma} \le \overline{\gamma}$.
 Also suppose $\partial\Omega$ is split into disjoint sets $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive Hausdorff measure, and $\partial_N \Omega$ possibly empty.
-Dirichlet data\footnote{Boundary values are in the trace sense \cite{Evans2010}, and $g_D$ is assumed to be as regular as needed.} is given by $g_D:\partial_D \Omega \to \RR$, and we assume $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.
+Dirichlet data is given by $g_D:\partial_D \Omega \to \RR$, and we assume $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.
 Now define
@@
 the closed and convex \emph{constraint set}; this treats inequality constraints and Dirichlet conditions in a unified manner.
-(Any Neumann conditions over $\partial_N \Omega$ are imposed weakly by the operator $f$, via an integral over $\partial_N\Omega$.)
-
-Next we will numerically solve VI problem \eqref{eq:vi} over a constraint set \eqref{eq:originalconstraintset}.  Given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we will assume that
-\begin{equation}
-\ip{f(u^\star)}{v-u^\star} \ge \ip{\ell}{v-u^\star} \qquad \text{for all } v\in \cK \label{eq:boxdirichletvi}
-\end{equation}
-is well-posed for $u^\star\in \cK$.
+Dirichlet values are in the trace sense \cite{Evans2010}, and $g_D$ is assumed to be as regular as needed.
+Any Neumann conditions over $\partial_N \Omega$ are imposed weakly by the operator $f$, via an integral over $\partial_N\Omega$.
+We will numerically solve VI problems \eqref{eq:vi} over constraint sets \eqref{eq:originalconstraintset}.
 
@@
 
-We seek to rapidly compute the approximate solution of box-constrained VI problems \eqref{eq:boxdirichletvi} using finite elements (FE).
-In this Section a nested multilevel $P_1$ or $Q_1$ element \cite{Elmanetal2014} approximation to \eqref{eq:boxdirichletvi} is constructed by refinement of a coarse mesh.
+We seek to rapidly compute the approximate solution of box-constrained VI problems \eqref{eq:vi} and \eqref{eq:originalconstraintset} using finite elements (FE).
+In this Section a nested multilevel $P_1$ or $Q_1$ element \cite{Elmanetal2014} approximation to \eqref{eq:vi} will be constructed by refinement of a coarse mesh.
 The mesh-refinement hierarchy permits the construction of a multilevel CD of the finest-level constraint set.
@@
 The mesh hierarchy and FE spaces can be constructed by standard nested refinement.
-Let $\Omega \subset \RR^d$ be an open bounded polygon.
-Let $\mathcal{T}^0$ be the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate cells with union equal to $\overline{\Omega}$ and maximum cell diameter (mesh size) $h_0>0$.
+Let $\Omega \subset \RR^d$ be an open bounded polygon and suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate cells with union equal to $\overline{\Omega}$ and maximum cell diameter (mesh size) $h_0>0$.
 We assume $\mathcal{T}^0$ is conforming, with no hanging nodes.
@@
 In other words, to enforce an inequality constraint, with $P_1$ or $Q_1$ we only enforce it pointwise on the degrees of freedom.
-For $P_k$ and $Q_k$, $k\ge 2$, there are functions in the standard bases which take on negative values \cite[Figure 1.7]{Elmanetal2014}, violating \eqref{eq:nodalmonotonicity}.
-
-To construct box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the vertices of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.
-The finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ are constructed by suitable interpolation of $\underline{\gamma}$, $\overline{\gamma}$ (e.g.~by applying the Cl\'ement interpolation operator~\cite{Carstensen2006}).
-Again they must satisfy $\underline{\gamma}^J < +\infty$, $-\infty < \overline{\gamma}^J$, and $\underline{\gamma}^J \le \overline{\gamma}^J$.
-On the finest-level mesh suppose the discrete boundary data $g_D^J \in \mathcal{V}^J$ satisfies $g_D^J = g_D$ on $\partial_D \Omega$; this restricts $g_D$ in \eqref{eq:originalconstraintset} to be piecewise-linear.
-We assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those vertices of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.
+For $k\ge 2$ there are functions in the standard bases of $P_k$ and $Q_k$ which take on negative values \cite[Figure 1.7]{Elmanetal2014}, violating \eqref{eq:nodalmonotonicity}.
+
+To construct box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the vertices of $\mathcal{T}^j$ with values in $\tilde{\RR} = [-\infty,+\infty]$.
+The finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ are constructed by suitable interpolation of $\underline{\gamma}$, $\overline{\gamma}$, e.g.~by applying the Cl\'ement interpolation operator~\cite{Carstensen2006}.
+Again the obstacles must satisfy $\underline{\gamma}^J < +\infty$, $-\infty < \overline{\gamma}^J$, and $\underline{\gamma}^J \le \overline{\gamma}^J$.
+On the finest-level mesh the discrete boundary data $g_D^J \in \mathcal{V}^J$ is assumed to satisfy $g_D^J = g_D$ on $\partial_D \Omega$---this restricts $g_D$ in \eqref{eq:originalconstraintset} to be piecewise-linear---and $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those vertices of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.
 Then we define
@@
 
-Let $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ be a discretization of $f$ in \eqref{eq:boxdirichletvi}.
+Let $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ be a discretization of $f$ in \eqref{eq:vi}.
 Note that the functional $\ell^J=\ell$ is defined and continuous over $\mathcal{V}^J$.
@@
 
-While \eqref{eq:fe:vi} approximates \eqref{eq:boxdirichletvi}, we warn the reader that it is usually not conforming.
-That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, we do not require or expect $\mathcal{K}^J \subset \mathcal{K}$.
+While \eqref{eq:fe:vi} approximates \eqref{eq:vi}, we warn the reader that it is usually not conforming.
+That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, from definitions \eqref{eq:originalconstraintset} and \eqref{eq:fe:fineconstraintset} we do not expect $\mathcal{K}^J \subset \mathcal{K}$ in general.
 Even for continuous obstacles $\underline{\gamma}, \overline{\gamma}$, admissibility with respect to the FE obstacles $\underline{\gamma}^J, \overline{\gamma}^J$ will generally not imply $\mathcal{K}$-admissibility.
@@
 Canonical prolongation $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ implements the vector space inclusion.
-Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $(R\sigma)[z] = \sigma[z]$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
+Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $\ip{R\sigma}{z} = \ip{\sigma}{z}$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
 These canonical transfers generally involve a change of FE representation, and in particular $R$ is ``full-weighting'' \cite{Trottenbergetal2001}.
-To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, note that the vertices of $\mathcal{T}^{j-1}$ are vertices of $\mathcal{T}^j$ and for $z\in\mathcal{V}^j$ define $\iR z$ as the element of $\mathcal{V}^{j-1}$ with the same (coarse) nodal values.
-Injection is monotonic: if $w \le z$ then $\iR w \le \iR z$.
+To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, note that the vertices of $\mathcal{T}^{j-1}$ are vertices of $\mathcal{T}^j$ and so for $z\in\mathcal{V}^j$ define $\iR z$ as the element of $\mathcal{V}^{j-1}$ with the same (coarse) nodal values.
+Injection is monotonic in the sense that if $w \le z$ then $\iR w \le \iR z$.
 
@@
 To make the construction precise, let $x_q^j$ denote the vertices of mesh $\mathcal{T}^j$, with corresponding basis functions $\psi_q^j$.
-For a $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$:
+For an $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$:
 \begin{align}
@@
 \end{equation}
-which does not hold for the injection operator $\iR$.
+which does not hold for the ordinary injection operator $\iR$.
 An illustration of the construction on nested meshes is given in Figure \ref{fig:Rplusminus}; implementing these operators on non-nested meshes could be efficiently done using a supermesh~\cite{Farrell2011}.
 In the next Section we will use $\maxR,\minR$ to construct ``level defect constraints'' such that the zero function is always admissible.
-This construction depends on additional ordering properties:
+This construction depends on additional ordering properties which follow from definition \eqref{eq:fe:monotoneinjections}:
 \begin{equation}
@@
 As in FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, and a source functional created.
-On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:boxdirichletvi}.\footnote{Presumably it uses quadrature etc.~in the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.}  At the start of each V-cycle we assume that an admissible iterate $w^J \in \mathcal{K}^J$ has been used to generate the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}); they do not depend on the coarse-level corrections.
+On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:vi}.\footnote{Presumably it uses quadrature etc.~in the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.}  At the start of each V-cycle we assume that an admissible iterate $w^J \in \mathcal{K}^J$ has been used to generate the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}); they do not depend on the coarse-level corrections.
 
@@
 \toprule
-$P$ & \emph{levels} & $m_J$ & \emph{iterations} & \emph{time} (s) & \emph{time$/$iterations} (s) \\
+$P$ & \emph{levels} & $m_J$ & \emph{iterations} & \emph{time} (s) & \emph{time$/$iteration} (s) \\
 \midrule

### 2023-08-13 paper/fascd.tex
Important bug fix: ICD-TELE should make sense now.
I don't know what I was thinking when I wrote the
previous version.  This should make sense.  The only way
that the FASCD V-cycle is different, on the upward side
that is, is because it does FAS-type coarse corrections,
whereas the "basic" ICD-TELE iteration refers to the original
operator f(.).

@@
 These inclusions are special to multilevel schemes based on defect constraints, and we have seen that they hold regardless of whether decomposition operators $\Pi_i$ can be constructed.
-Thus we have a new, incomplete variant of \pr{cd-mult}.
-For $w^J\in\mathcal{K}^J$, it applies in the multilevel CD case where \eqref{eq:fe:upwardsuminclusion} holds:
+In fact we can propose a new, incomplete variant of the \pr{cd-mult} iteration.
+For $w^J\in\mathcal{K}^J$, and existing downward perturbations $y^i\in\mathcal{D}^i$, the iteration applies in the special case where inclusion \eqref{eq:fe:upwardsuminclusion} holds:
 
 \begin{pseudo*}
-\pr{icd-tele}(w^J)\text{:} \\+
+\pr{icd-tele}(w^J, \{y^i\}_{i=1}^J)\text{:} \\+
     for $j = 0,\dots,J$: \\+
         \rm{find} $z^j\in \mathcal{U}^j$ \rm{so that for all} $v \in \mathcal{U}^j$, \\+
-            $\displaystyle \Big<f\Big(w^J + \sum_{i<j} z^i + z^j\Big),v-z^j\Big> \ge \ip{\ell}{v-z^j}$ \\--
-    return $\tilde w^J=w^J + \sum_{i\le J} z^i$
+            $\displaystyle \Big<f\Big(w^J + \sum_{i>j} y^i + z^j\Big),v-z^j\Big> \ge \ip{\ell}{v-z^j}$ \\--
+    return $\tilde w^J=w^J + z^J$
 \end{pseudo*}
 
+\noindent
+Note that the downward perturbations $y^i$, which apparently disappear as \pr{icd-tele} proceeds, can actually play a practical role in solving each VI problem, namely when constructing an initial iterate for $z^j$; see the next Section.
+
 Algorithm \ref{alg:fascd} below incorporates \pr{icd-tele} for up-smoothing in the larger constraint sets $\mathcal{U}^j$.
-The observation that this is more efficient, illustrated by Example \ref{ex:results:plap} in particular, seems to be new.
-Specifically, $V(1,1)$ cycles in \cite{GraeserKornhuber2009} use half-size versions of the smaller sets $\mathcal{D}^j$ both on descent and ascent; see also \cite[section 5.4]{Tai2003}.
-However, a convergence analysis extending the arguments of \cite{GraeserKornhuber2009,Tai2003}, either to non-optimization VIs or box constraints, or both, would seem to require full CDs.
+The observation that this can be more efficient, illustrated by Example \ref{ex:results:plap} in particular, seems to be new.
+For comparison, $V(1,1)$ cycles in \cite{GraeserKornhuber2009} use half-size versions of the smaller sets $\mathcal{D}^j$ both on descent and ascent; see also \cite[section 5.4]{Tai2003}.
+However, any convergence analysis extending the arguments of \cite{GraeserKornhuber2009,Tai2003} might require full CDs.
 

### 2023-08-13 paper/fascd.tex
address PF comments on wording

@@
 The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} applies to linear complementarity problems and projected (truncated) Gauss--Seidel smoothers.
-While FASCD reduces to general, nonlinear FAS for PDEs, i.e.~when the inequality constraints are removed, PFAS reduces to linear multigrid.
+While FASCD reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed, PFAS reduces to linear multigrid.
 On the other hand, the multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so as stated it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
@@
 The next VI example is a model for steady ice sheet geometry in a given climate.
-It is strongly nonlinear and not of optimization type.
+It is fully nonlinear and not of optimization type.
 Related VI problems arise for any fluid layer subject to processes which can remove fluid mass \cite{Bueler2021conservation}.

### 2023-08-13 paper/fascd.tex
finish the sweep of final polish edits; an EB remains ...

@@
 \newcommand{\pef}[1]{\todo[inline, color=red!20]{PF: #1}}
-%\newcommand{\elb}[1]{\todo[inline, color=blue!20]{EB: #1}}
+\newcommand{\elb}[1]{\todo[inline, color=blue!20]{EB: #1}}
 
@@
 \end{enumerate}
+
+\elb{address H\&K 2011 in following paragraph}
 
@@
 
-A key choice in devising a multilevel scheme for VIs is how to construct coarser-mesh bounds.
+A key choice in devising a multilevel scheme for VIs is how to construct bounds for the coarser levels.
 Our multilevel CD scheme uses \emph{defect constraints} \cite{GraeserKornhuber2009} derived from an admissible finest-level iterate.
-The next Example illustrates why we do \emph{not} apply injection operators directly\footnote{An exception occurs in FMG; see Section \ref{sec:implementation}.} to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.
+The next Example illustrates why we do \emph{not}\footnote{An exception occurs in FMG; see Section \ref{sec:implementation}.} apply injection operators directly to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.
 
@@
 Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ vertices, respectively.
-For admissible boundary conditions the fine-level constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.
+For appropriate boundary conditions the fine-level constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.
 However, various schemes to coarsen the constraints by direct transfers of $\underline{\gamma}^1,\overline{\gamma}^1$ are doomed to failure.
@@
 Then $\underline{\gamma}^0=0$ and $\overline{\gamma}^0=C$ identically.
-In that case the corresponding constraint set $\mathcal{K}^0$ is nonempty, but prolongation $Pw^0$ of an admissible solution of the coarse problem ($w^0\in\mathcal{K}^0$; $\underline{\gamma}^0 \le w^0 \le \overline{\gamma}^0$) is not admissible on the finer level ($Pw^0 \notin \mathcal{K}^1$), and attempting to fix this by projecting into $\mathcal{K}^1$ would introduce high frequencies not present in $w^0$.
+In that case the corresponding constraint set $\mathcal{K}^0$ is nonempty, but prolongation $Pw^0$ of an admissible solution of the coarse problem ($w^0\in\mathcal{K}^0$; $\underline{\gamma}^0 \le w^0 \le \overline{\gamma}^0$) is not admissible on the finer level ($Pw^0 \notin \mathcal{K}^1$).
+Attempting to fix this by projecting (truncating) into $\mathcal{K}^1$ would introduce high frequencies not present in $w^0$.
 On the other hand, if
@@
     \label{eq:fe:badmonotone}
-    \underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}
-\end{equation}
-then $w^0 \ge \underline{\gamma}^0 \, \implies \, Pw^0 \ge \underline{\gamma}^1$ and $w^0 \le \overline{\gamma}^0 \, \implies \, Pw^0 \le \overline{\gamma}^1$.
-However, now the coarse-level constraint set $\mathcal{K}^0$ is empty because $\underline{\gamma}^0 > \overline{\gamma}^0$ (Figure \ref{fig:directRbad}).
+    \underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1 \qquad \text{\emph{[problematic]}}
+\end{equation}
+then the coarse-level constraint set $\mathcal{K}^0$ is empty because $\underline{\gamma}^0 > \overline{\gamma}^0$ (Figure \ref{fig:directRbad}, right).
 \end{example}
@@
 \end{tabular}
-\caption{For the finer-level constraints $\underline{\gamma}^1$ (blue) and $\overline{\gamma}^1$ (red) on the left, direct application of $\iR,\minR,\maxR$ is problematic.
-For example, if we apply the monotone injections then the resulting coarser-level constraint set $\mathcal{K}^0$ is empty (right).
-Defect constraints avoid this difficulty.}
+\caption{For the finer-level constraints $\underline{\gamma}^1$ (blue) and $\overline{\gamma}^1$ (red) on the left, direct application of $\iR,\minR,\maxR$ to generate coarser-level constraints is problematic.
+Possibility \eqref{eq:fe:badmonotone}, which generates an empty coarser-level set $\mathcal{K}^0$, is at right.
+Defect constraints \eqref{eq:fe:defectconstraints} avoid this difficulty.}
 \label{fig:directRbad}
@@
 
-This construction of multilevel box constraints, likewise for the unilateral construction in \cite{GraeserKornhuber2009}, permits large, yet admissible, coarse corrections, which promotes multilevel solver efficiency.
-However, when coarsening the problem downward in a V-cycle, one cannot predict the upcoming coarse corrections and so one must construct constraint sets from which any sum will be admissible.
+This construction of multilevel box constraints, likewise the unilateral construction in \cite{GraeserKornhuber2009}, permits large, yet admissible, coarse corrections, which promotes multilevel solver efficiency.
+When coarsening the problem downward in a V-cycle, one cannot predict the upcoming coarser corrections and so one must construct constraint sets from which any sum will be admissible.
 With this in mind, and looking forward to telescoping sums \eqref{eq:fe:telescoping}, we compute the LDC differences \cite{GraeserKornhuber2009}:
@@
 
-Our V-cycle solver (Algorithm \ref{alg:fascd}) will compute corrections in $\mathcal{D}^j$ and $\mathcal{U}^j$.
+The V-cycle solver in the next Section (Algorithm \ref{alg:fascd}) will compute corrections from $\mathcal{D}^j$ and $\mathcal{U}^j$.
 For general box constraints the upward direction is, however, based on \emph{incomplete} CDs as we now explain.
-Each upward set $\mathcal{U}^j$ is generally only a superset of a sum of downward sets $\mathcal{D}^j$, though this becomes a true CD satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} in unilateral cases.
+Each upward set $\mathcal{U}^j$ is generally only a superset of a sum of downward sets $\mathcal{D}^j$, though this becomes a true CD satisfying \eqref{eq:subspacedecomp}--\eqref{eq:decompositionstability} in unilateral cases.
 The following Lemma, apparently new, extends the unilateral CD construction underlying Algorithm 4.2 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.
@@
 For concreteness suppose $\overline{\gamma}^J=+\infty$, thus $\overline{\chi}^j=+\infty$ and $\overline{\phi}^i = +\infty$ for all levels; the $\underline{\gamma}^J=-\infty$ case is similar.
-For $v\in \mathcal{V}^j$ and $0\le i \le j$, let $I_{j\to i}^\ominus$ be the operator applying the minimum (monotone) injection $j-i$ times, mapping into $\mathcal{V}^i$: $I_{j\to i}^\ominus v = \minR(\cdots(\minR v)\cdots) $.
+For $v\in \mathcal{V}^j$ and $0\le i \le j$, let $I_{j\to i}^\ominus\,:\,\tilde{\cV}^j \to \tilde{\cV}^i$ be the operator applying the minimum (monotone) injection $j-i$ times: $I_{j\to i}^\ominus v = \minR(\cdots(\minR v)\cdots) $.
 (Set $I_{j\to j}^\ominus v = v$ and $I_{j\to -1}^\ominus=0$ by definition.)  Now define nonlinear decomposition operators $\Pi_i:\mathcal{U}^j \to \mathcal{D}^i$:
@@
 However, the following Example shows why $\Pi_i(\min\{z^j,0\})$ may have an arbitrarily large maximum, and thus exceed any finite upper LDC $\overline{\phi}^i$.
-In fact, $\Pi_i$ does not generally map all non-positive elements of $\mathcal{U}^J$ into $\mathcal{D}^i$ if a finite upper LDC is present.
-On the other hand, we can not exclude entirely-different strategies for showing full CDs in the general box-constrained case.
+In fact, $\Pi_i$ defined above does not generally map all non-positive elements of $\mathcal{U}^J$ into $\mathcal{D}^i$ if a finite upper LDC is present.
+The Example does not exclude other strategies for showing full CDs in the general box-constrained case.
 
@@
 But then $\Pi_1 z^1$ attains a maximum $a\ge 0$.
-This maximum can be chosen to exceed any upper defect obstacle $\overline{\phi}^1$, and $\Pi_1 z^1 \notin \mathcal{D}^1$ would occur.
+This maximum can be chosen to exceed any upper defect obstacle $\overline{\phi}^1$, and $\Pi_1 z^1 \notin \mathcal{D}^1$ would then occur.
 \end{example}
@@
 \end{equation}
-In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, actually $\mathcal{U}^j = \mathcal{U}^k + \sum_{i=k+1}^j \mathcal{D}^i$ as a full CD. \end{lemma}
-
-\begin{proof}  Definition \eqref{eq:fe:philevels} shows that $\underline{\chi}^j = \underline{\phi}^j + \dots + \underline{\phi}^{k+1} + \underline{\chi}^k$ and $\overline{\chi}^j = \overline{\phi}^j + \dots + \overline{\phi}^{k+1} + \overline{\chi}^k$.
-If $y^i \in \mathcal{D}^i$ for $k+1 \le i \le j$ and $z^k \in \mathcal{U}^k$ then $\underline{\chi}^j \le y^j + \dots + y^{k+1} + z^k \le \overline{\chi}^j$, thus \eqref{eq:fe:upwardsuminclusion} holds.
+In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, actually $\mathcal{U}^j = \mathcal{U}^k + \sum_{i=k+1}^j \mathcal{D}^i$ is a full CD. \end{lemma}
+
+\begin{proof}  Definition \eqref{eq:fe:philevels} shows inclusion \eqref{eq:fe:upwardsuminclusion}.
 In unilateral cases we may follow the proof of Lemma \ref{lem:downwardadmissibility} with \eqref{eq:fe:unilateraldecompositionoperator} unchanged for $i=k+1,\dots,j$, but modified in the $i=k$ case: $\Pi_k z^j = I_{j\to k}^\ominus(z^j - \underline{\chi}^j) + \underline{\chi}^k$.
-Now the proof goes through as before.
+Otherwise the proof goes through as before.
 \end{proof}
@@
 The V-cycle in this Section extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.2]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}; it reduces to FAS for PDEs when inequality constraints are removed.
-Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$, via information from each coarser FE subspace $\mathcal{V}^j$, by approximately solving VI problems in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively; by \eqref{eq:fe:nestedspaces} these corrections are all elements of the fine-level subspace $\mathcal{V}^J$.
+Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$, via information from each coarser FE subspace $\mathcal{V}^j$, by approximately solving VI problems in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.
 The corrections are inductively constructed so that iterates remain admissible at every level.
-A key idea is that, by inclusion \eqref{eq:fe:upwardsuminclusion}, up-smoothing solves less-restricted VI problems than down-smoothing.
+A key idea is that up-smoothing solves less-restricted VI problems than down-smoothing.
 Going downward one does not know what coarse corrections are coming, thus one must choose smaller constraint sets ($\mathcal{D}^j$) such that the correction sum is admissible once we return to a given level.
-Going upward, however, no further coarse corrections can violate admissibility.
+Going upward, however, no further coarse corrections can violate admissibility, so we relax in the larger sets $\mathcal{U}^j$.
 
@@
 Solving the coarsest-level VI results in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.
-Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for relaxation on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using \eqref{eq:fe:upwardsuminclusion}.
+Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for relaxation on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using inclusion \eqref{eq:fe:upwardsuminclusion}.
 Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.
@@
 \caption{Left: The FASCD V-cycle Algorithm \ref{alg:fascd} computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.
-Right: FMG Algorithm \ref{alg:fascd-fmg} generates each next-finer initial iterate by injection and truncation (doubled edges), then it iterates V-cycles.}
+Right: FMG Algorithm \ref{alg:fascd-fmg} generates the initial iterate on finer levels by injection and truncation (doubled edges).}
 \label{fig:fascdcycles}
@@
 
-To state the V-cycle algorithm we must specify the problems solved at each mesh level.
+We must specify the problems solved at each mesh level.
 As in FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, and a source functional created.
-On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:vi}.\footnote{Presumably it uses quadrature etc.~in the same scheme which generated $f^J$ in \eqref{eq:fe:vi}.}  At the start of each V-cycle we assume that an admissible iterate $w^J \in \mathcal{K}^J$ has been used to generate the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}); they do not depend on the coarse-level corrections.
+On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:vi}.
+At the start of each V-cycle we assume that an admissible iterate $w^J \in \mathcal{K}^J$ has been used to generate the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}); they do not depend on the coarse-level corrections.
 
@@
 \begin{equation}
-w^j = \iR(w^{j+1} + y^{j+1}).  \label{eq:fe:definew}
-\end{equation}
-From $w^j$ we define $\ell^j \in (\cV^j)'$:
+w^j = \iR(w^{j+1} + y^{j+1}),  \label{eq:fe:definew}
+\end{equation}
+from which we define $\ell^j \in (\cV^j)'$:
 \begin{equation}
@@
 
-However, because the down-smoother has made progress, $y$ can be approximated by solving a coarser, $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps: \emph{(i)} subtract the computable quantity $f^{j+1}(w^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides, \emph{(ii)} replace the relaxed residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction, \emph{(iii)} replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its injection, and \emph{(iv)} where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.
+However, because the down-smoother has made progress, $y$ should be approximate-able by solving a coarser, $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps: \emph{(i)} subtract the computable quantity $f^{j+1}(w^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides, \emph{(ii)} replace the relaxed residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction, \emph{(iii)} replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its injection, and \emph{(iv)} where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.
 These steps yield a problem for the $j$th-level correction $y=y^j \in \mathcal{D}^j$:
@@
 The explanation for \eqref{eq:fe:upvi} and \eqref{eq:fe:upwardinitial} is simpler.
-Whether or not problem \eqref{eq:fe:downvi} has been exactly solved by the computed $y^j\in\mathcal{D}^j$, when going upward we can expand problem \eqref{eq:fe:downvi} into the larger set $\mathcal{U}^j \supset \mathcal{D}^j$.
-(One may continue admissibly upward from any approximate solution to \eqref{eq:fe:upvi}.)  Generally $y^j$ is not the solution to \eqref{eq:fe:upvi}; in the optimization case $f=F'$ it would not be optimal in the enlarged constraint set.
-From \eqref{eq:fe:upwardsuminclusion} our current best estimate of the solution to \eqref{eq:fe:upvi} is $z^j \approx y^j + Pz^{j-1}\in \mathcal{U}^j$; this gives the initial iterate for the up-smoother relaxation.
+Whether or not problem \eqref{eq:fe:downvi} has been exactly solved by the computed $y^j\in\mathcal{D}^j$, when going upward we can expand problem \eqref{eq:fe:downvi} into the larger set $\mathcal{U}^j \supset \mathcal{D}^j$, and \eqref{eq:fe:upwardsuminclusion} assures admissibly upward from any approximate solution to \eqref{eq:fe:upvi}.
+Generally $y^j$ is not the solution to \eqref{eq:fe:upvi}; e.g.~in the optimization case $f=F'$ it would not be optimal in the enlarged constraint set.
+Our current best estimate of the solution to \eqref{eq:fe:upvi} is $z^j \approx y^j + Pz^{j-1}\in \mathcal{U}^j$, so this is the initial iterate for relaxation.
 
@@
 
-Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle with \id{down} and \id{up} smoother iterations on a level, respectively.
-The $V(1,0)$ cycle with \id{up} $=0$ is similar to Algorithm 4.2 in \cite{GraeserKornhuber2009}, but generalized by applying FAS-type corrections and allowing any smoother or nonlinear operator $f$.
-In unilateral cases one iteration of \pr{fascd-vcycle} is an application of \pr{cd-mult} (Section \ref{sec:cd}) over CD \eqref{eq:fe:downwardsuminclusion} followed by an application of \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over CDs \eqref{eq:fe:upwardsuminclusion}, and in this sense $V(1,1)=V(1,0)+V(0,1)$.
-
-Regarding the storage required, if all box constraints are nontrivial then 7 vectors must be allocated on each level: $\underline{\chi}^j$, $\overline{\chi}^j$, $\underline{\phi}^j$, $\overline{\phi}^j$, $w^j$, $\ell^j$, $y^j$; vectors $z^j$ and $y^j$ can use the same storage.
+With \id{down} $=1$ and \id{up} $=1$ smoother iterations on each level, Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle.
+The $V(1,0)$ cycle with \id{up} $=0$ is similar to Algorithm 4.2 in \cite{GraeserKornhuber2009}, but here it is generalized by applying FAS-type corrections and allowing any smoother.
+In unilateral cases one $V(1,1)$ iteration of \pr{fascd-vcycle} approximates applications of \pr{cd-mult} (Section \ref{sec:cd}) over CD \eqref{eq:fe:downwardsuminclusion}, followed by \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over \eqref{eq:fe:upwardsuminclusion}, and in this sense $V(1,1)=V(1,0)+V(0,1)$.
+
+Regarding the storage required, if all box constraints are nontrivial then 7 vectors must be allocated on each level: $\underline{\chi}^j$, $\overline{\chi}^j$, $\underline{\phi}^j$, $\overline{\phi}^j$, $w^j$, $\ell^j$, $y^j$; note vectors $z^j$ and $y^j$ may use the same storage.
 On the finest level one must also store $\underline{\gamma}^J$ and $\overline{\gamma}^J$.
@@
 
-Implementations of FASCD methods must choose a convergence criterion.
-In this section we also construct the FMG extension, and we describe the active set Newton method later used as the smoother and coarse solver.
-
-Repeated application of Algorithm \ref{alg:fascd} to solve VI problem \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere, but the VI is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}, which says, by definition, that certain nodal conditions hold for $u^J$ at $x_p \in \mathcal{T}^J$.
-Let $r(u^J)_p = \ip{f^J(u^J)-\ell^J}{\psi_p}$ be the pointwise (ordinary) residual.
-If $x_p \notin \partial_D\Omega$ then one of these four conditions must hold:
+In this section we propose a convergence criterion, construct the full multigrid (FMG) extension of Algorithm \ref{alg:fascd}, and summarize the reduce-space Newton method used as the smoother and coarse solver.
+
+Repeated application of Algorithm \ref{alg:fascd} to solve finite-dimensional VI problem \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.
+However, the VI is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}, which says, by definition, that certain conditions hold for $u^J$ at nodes $x_p \in \mathcal{T}^J$.
+Let $r(u^J)_p = \ip{f^J(u^J)-\ell^J}{\psi_p}$ be the ordinary (pointwise) residual.
+If $x_p \notin \partial_D\Omega$ then exactly one of these four conditions must hold:
 \begin{itemize}
@@
 \item $\underline{\gamma}^J(x_p)=u^J(x_p)<\overline{\gamma}^J(x_p)$ and $r(u^J)_p \ge 0$,
-\item $\underline{\gamma}^J(x_p)<u^J(x_p)=\overline{\gamma}^J(x_p)$ and $r(u^J)_p \le 0$,
+\item $\underline{\gamma}^J(x_p)<u^J(x_p)=\overline{\gamma}^J(x_p)$ and $r(u^J)_p \le 0$, or
 \item $\underline{\gamma}^J(x_p)=u^J(x_p)=\overline{\gamma}^J(x_p)$.
@@
 
-The MCP interpretation allows us to construct expressions whose norms should approach zero.  Define the Fischer-Burmeister function \cite{BensonMunson2006,Ulbrich2011} on $a,b\in\RR$:
+This MCP interpretation allows us to construct expressions whose norms should approach zero.  Define the Fischer-Burmeister function \cite{BensonMunson2006,Ulbrich2011} on $a,b\in\RR$:
 \begin{equation}
@@
 Observe that $\phi$ is semi-smooth, thus continuous, and that $\phi(a,b)=0$ if and only if $a,b$ are complementary, i.e.~$a\ge 0$, $b\ge 0$, and $ab=0$.
-For $w^J\in \mathcal{K}^J$ denote $\underline{g}_p = w^J(x_p) - \underline{\gamma}^J(x_p)$ and $\overline{g}_p = \overline{\gamma}^J(x_p) - w^J(x_p)$, the nonnegative \emph{gaps} at $x_p$.
+For any $w^J\in \mathcal{K}^J$ denote $\underline{g}_p = w^J(x_p) - \underline{\gamma}^J(x_p)$ and $\overline{g}_p = \overline{\gamma}^J(x_p) - w^J(x_p)$, the nonnegative \emph{gaps} at $x_p$.
 Now define the \emph{semi-smooth residual}
@@
 \end{equation}
-Straightforward modifications apply for unconstrained degrees of freedom, where $\overline{\gamma}^J(x_p) = +\infty$ or $\underline{\gamma}^J(x_p) = -\infty$ or both.
+Straightforward modifications apply for unilaterally-constrained or unconstrained degrees of freedom, where $\underline{\gamma}^J(x_p) = -\infty$, $\overline{\gamma}^J(x_p) = +\infty$, or both.
 VI problem \eqref{eq:fe:vi} is exactly solved by $u^J\in\cK^J$ if and only if $\rSS(u^J)=0$ identically.
 
-Suppose that iteration of Algorithm \ref{alg:fascd} starts with $w^{J,0} \in \mathcal{K}^J$ and generates (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.
+Suppose that Algorithm \ref{alg:fascd} is started with $w^{J,0} \in \mathcal{K}^J$ and generates (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.
 Let $\id{atol}$, $\id{rtol}$, $\id{stol} \ge 0$ be given absolute, relative, and step tolerances, respectively.
@@
 \begin{equation}
-\|\rSS(w^{J,k})\|_2 < \id{atol} \quad \text{or} \quad \frac{\|\rSS(w^{J,k})\|_2}{\|\rSS(w^{J,0})\|_2} < \id{rtol} \quad \text{or} \quad \frac{\|\Delta w^{J,k}\|_2}{\|w^{J,k}\|_2} < \id{stol}, \label{eq:stoppingcriterion}
-\end{equation}
-where $\Delta w^{J,k} = w^{J,k} -  w^{J,k-1}$, as the stopping criterion.\footnote{The defaults in our Firedrake implementation match those of PETSc SNES \cite{Balayetal2023}: $\id{atol}=10^{-50}$, $\id{rtol}=10^{-8}$, $\id{stol}=10^{-8}$.
-See Section \ref{sec:results} for specific settings in Examples.}
-
-In Section \ref{sec:results} we will show that V-cycle iterations alone, starting with some finest-mesh initial iterate $w^{J,0}$ and repeated until \eqref{eq:stoppingcriterion} holds, forms an effective solver.
-However, an FMG cycle, Algorithm \ref{alg:fascd-fmg} and Figure \ref{fig:fascdcycles} (right), is even better.
+\|\rSS(w^{J,k})\|_2 < \id{atol} \quad \text{or} \quad \frac{\|\rSS(w^{J,k})\|_2}{\|\rSS(w^{J,0})\|_2} < \id{rtol} \quad \text{or} \quad \frac{\|\Delta w^{J,k}\|_{L^2}}{\|w^{J,k}\|_{L^2}} < \id{stol}, \label{eq:stoppingcriterion}
+\end{equation}
+where $\Delta w^{J,k} = w^{J,k} -  w^{J,k-1}$, $\|\cdot\|_2$ is the euclidean norm on $\RR^{m_J}$, and $\|\cdot\|_{L^2}$ is the norm on $L^2(\Omega)$, as the stopping criterion.
+Defaults in our Firedrake implementation match those of PETSc SNES \cite{Balayetal2023}: $\id{atol}=10^{-50}$, $\id{rtol}=\id{stol}=10^{-8}$; see Section \ref{sec:results} for specific settings in Examples.
+
+In Section \ref{sec:results} we will show that V-cycle iterations alone, starting with some finest-mesh initial iterate $w^{J,0}$ and repeated until \eqref{eq:stoppingcriterion} holds, serve as an effective solver.
+However, an FMG cycle, Algorithm \ref{alg:fascd-fmg} below and Figure \ref{fig:fascdcycles} (right), is even better.
 The well-known principle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate $w^0\in\mathcal{K}^0$; one may generate it by injection and projection/truncation of $w^{J,0}$.
 Then an admissible initial iterate on the next-finer mesh is generated via prolongation.
-In the VI context one must also project (truncate) to the bounds after the prolongation; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fmg}.
-During this nested-iteration ``ramp'' one approximately solves on each level using a fixed number of V-cycles, each down to the coarsest mesh.
+However, in the VI context one must also project (truncate) to the bounds after the prolongation; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fmg}.
+During this ``ramp'' stage one approximately solves on each level using a fixed number (\id{rampv}) of V-cycles, each down to the coarsest mesh.
 When the ramp is complete, one does full-depth V-cycles to convergence.
@@
 
-In contrast to the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, standard injection is used to generate the constraints $\underline{\gamma}^j,\overline{\gamma}^j$ on each level because, during the ramp, a full solution iterate is immediately needed, not just a correction to an initial iterate.
-Essentially, our choice is between \eqref{eq:fe:badstandard} (where the coarse constraint set is always nonempty, but the solution may not be admissible on the next level) or \eqref{eq:fe:badmonotone} (where the coarse solution is always admissible if it exists, but the constraint set may be empty).
-We choose \eqref{eq:fe:badstandard} so that the coarse problems always have nonempty constraint sets, and project (truncate) pointwise to the bounds as we refine.
-
-A large space of smoothers could be explored, but in the next Section we will only apply an active-set (reduced-space) Newton method \cite{Balayetal2023,BensonMunson2006}, denoted as \emph{RS Newton}.
+In contrast to the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, during the ramp standard injection is used to generate the constraints $\underline{\gamma}^j,\overline{\gamma}^j$ on each level because a full solution iterate is initially needed, not just a correction.
+Essentially, our choice is between \eqref{eq:fe:badstandard}, where the coarse constraint set is nonempty but then the prolonged solution may not be admissible, or \eqref{eq:fe:badmonotone}, where the constraint set may be empty.
+We choose \eqref{eq:fe:badstandard}, with nonempty sets, but we must project (truncate) pointwise to the bounds, potentially adding high-frequency noise to the initial iterate.
+
+A large space of smoothers could be explored, but in the next Section we will only apply an reduced-space Newton method \cite{Balayetal2023,BensonMunson2006}, denoted as \emph{RS Newton}.
 A fixed number of preconditioned Krylov iterations approximately solve the arising linear systems, thus $O(m_j)$ work is done per smoother application.
@@
 
-For $P_1$ and $Q_1$ elements, Algorithms \ref{alg:fascd} and \ref{alg:fascd-fmg} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}.
-The numerical results in this Section, which follow the Section \ref{sec:vi} examples, focus on the scaling of FASCD iterations with mesh resolution.
+For $P_1$ and $Q_1$ elements, FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fmg} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}.
+The numerical results in this Section, which follow the Section \ref{sec:vi} examples, focus on the scaling of iterations with mesh resolution.
 We see that each V-cycle greatly reduces the residual norm, and that a few FMG iterations solve most of these problems to within discretization error.
@@
 \begin{example} \label{ex:results:classical}
-Consider the classical unilateral 2D Laplacian obstacle problem, Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\}$.
+Consider the classical unilateral 2D Laplacian obstacle problem, namely Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\}$.
 We consider two cases, over square domains $\Omega$, with different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$ (Figure \ref{fig:results:classical}).
 The standard ``ball'' problem has an upper hemisphere obstacle \cite[Chapter 12]{Bueler2021}, but the ``spiral'' problem's obstacle is constructed to produce the coincidence set shown \cite[problem 7.1.1]{GraeserKornhuber2009}.
-For this experiment we used a regular coarsest mesh with $m_0=41$ nodes (64 triangles), and the finest mesh had $m_{10} \approx 3.4 \times 10^7$ nodes.
-The smoother applied a single RS Newton iteration with three IC-preconditioned CG iterations.
-The initial iterate was $w=\max\{0,\psi\}$.
+For both cases we used a regular coarsest mesh with $m_0=41$ nodes (64 triangles), and the finest mesh had $m_{10} \approx 3.4 \times 10^7$ nodes.
+A single RS Newton iteration with three IC-preconditioned CG iterations was used as the smoother, and the initial iterate was $w=\max\{0,\psi\}$.
 
@@
 By contrast, V-cycle iterations increase slowly, perhaps logarithmically with the resolution.
-For the spiral problem, \cite[Figure 7.11]{GraeserKornhuber2009} reports that a $m_J \approx 5.2 \times 10^5$ problem needs approximately 25 V-cycle iterations for the best (truncated, TNMG/TMMG) monotone multilevel algorithms.\footnote{The stopping criterion and asymptotic convergence rates in \cite{GraeserKornhuber2009} are based on norm differences from a last numerical iterate with unknown accuracy.
-Our method directly measures satisfaction of the VI by a residual norm, namely $\|\rSS(w)\|_2$.}  The 12 V-cycle iterations needed here, at comparable resolution (8 levels), represent somewhat more arithmetic because our smoother is RS Newton using CG+IC, rather than projected Gauss-Seidel.
+For the spiral problem, \cite[Figure 7.11]{GraeserKornhuber2009} reports that a $m_J \approx 5.2 \times 10^5$ problem needs approximately 25 V-cycle iterations for the best (truncated;  ``TNMG'' and ``TMMG'') monotone multilevel algorithms.\footnote{The stopping criterion and asymptotic convergence rates in \cite{GraeserKornhuber2009} are based on norm differences from a last numerical iterate with unknown accuracy.
+Our method directly measures satisfaction of the VI, via the residual norm $\|\rSS(w)\|_2$.}  The 12 V-cycle iterations needed here, at comparable resolution (8 levels), represent somewhat more arithmetic because our smoother is RS Newton using CG+IC, rather than the simpler projected Gauss-Seidel iteration.
 Resolution-dependent asymptotic convergence rates for V-cycles, defined as $\rho_J=(\|\rSS(w^{J,k})\|_2/\|\rSS(w^{J,0})\|_2)^{1/k}$ where $w^{J,k}$ is the first iterate to satisfy \eqref{eq:stoppingcriterion}, are shown in Figure \ref{fig:results:asymp}.
-Both iterations and rates are similar for the ball and spiral problems suggesting that the geometric complexity of the coincidence set and free boundary minimally influences FASCD V-cycle performance.
-(The highest-resolution rate difference has no explanation for now.)  Contrast the rate of about $0.1$ for 8 levels here with about $0.4$ reported by \cite[Figure 7.11]{GraeserKornhuber2009} for TNMG/TMMG at comparable resolution.
+Both iterations and rates are similar for the ball and spiral problems; the geometric complexity of the coincidence set and free boundary minimally influences FASCD V-cycle performance.
+Contrast the rate of about $0.1$ for 8 levels here with about $0.4$ reported by \cite[Figure 7.11]{GraeserKornhuber2009} for TNMG/TMMG at comparable resolution.
 These rates are all excellent; our better rate mostly reflects a stronger smoother.
@@
 Since each V-cycle does the work of about eight finest-level Krylov iterations, the ball results show nearly textbook multigrid efficiency at the highest resolutions despite the low solution regularity.
-More V-cycles are needed to clean-up error remaining after the ramp in the spiral problem, presumably from the FMG initial-iterate injection scheme (Section \ref{sec:implementation}).
+More V-cycles are needed to clean-up error remaining after the ramp in the spiral problem.
 \end{example}
@@
 \begin{example}  \label{ex:results:plap}
-We again consider Example \ref{ex:plaplacian}, with a unilateral constraint, but now for the nonlinear $p$-Laplacian \cite{ChoeLewis1991}.
+We again consider Example \ref{ex:plaplacian}, with a unilateral constraint, but now for a nonlinear $p$-Laplacian \cite{ChoeLewis1991} in one dimension.
 Setting $p=1.5$ yields a fast diffusion ($D=|\grad u|^{p-2}\to \infty$ as $|\grad u|\to 0$).
-Note that Example \ref{ex:results:sia} below considers a doubly-nonlinear operator with $p=4$ degenerate diffusion.
+Note that Example \ref{ex:results:sia} below, by constrast, has a doubly-nonlinear operator with $p=4$ degenerate diffusion.
 
@@
 For this experiment the coarsest mesh has $m_0=7$ nodes and the finest $m_9=3073$.
-The default smoother uses three (fixed) iterations of RS Newton, with LU.
+The default smoother uses three iterations of RS Newton, with LU soluton of the linear systems.
 We set \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $=$ \texttt{stol} $= 10^{-12}$.
@@
 FMG is even more efficient; a single iteration suffices at all higher resolutions.
-Note that convergence is at rate $O(h^{2.005})$; compare \cite{ChoeLewis1991}.
+Note that convergence is at rate $O(h^{2.0})$; compare \cite{ChoeLewis1991}.
 \end{example}
@@
 However, the smoother was the same in all cases: two (fixed) RS Newton iterations, with three preconditioned (additive Schwarz method (ASM) with ILU on each process) GMRES iterations.
-The coarse solver was RS Newton and LU, iterated to convergence.
 Tolerances were set to \id{rtol} $=10^{-5}$ and \id{atol} $=$ \id{stol} $= 10^{-9}$.
@@
 
-For a weak-scaling study we fixed the number of degrees of freedom per process at $m_J/P=641^2 \approx 4.1 \times 10^5$ and solved with $6, \dots, 11$ refinements.\footnote{The experiments were conducted on ARCHER2, the UK national supercomputer.
-Each node has 16 memory channels, so a maximum of 16 cores per node were employed.}  The results in Table \ref{tab:results:siaweak} show some degradation in time per cycle with increasing core counts, perhaps due to the use of one-level additive Schwarz.
+For a weak-scaling study we fixed the number of degrees of freedom per process at $m_J/P=641^2 \approx 4.1 \times 10^5$ and solved with $6, \dots, 11$ levels.\footnote{The experiments were conducted on ARCHER2, the UK national supercomputer.
+Each node has 16 memory channels, so a maximum of 16 cores per node were employed.}  The results in Table \ref{tab:results:siaweak} show some degradation in time per cycle with increasing core counts, perhaps due to the use of one-level additive Schwarz in the Krylov iteration.
 Nevertheless, the finest mesh result, with $m_J=20481^2=4.1 \times 10^8$ degrees of freedom, required less than three minutes to solve for the steady surface elevation of an ice sheet larger than the current Greenland ice sheet ($1.7$ million $\text{km}^2$) at uniform $87.5$ m resolution.
@@
 For smoothers of Newton-Krylov type one could also add linear geometric or algebraic multigrid preconditioning \cite{Trottenbergetal2001} for the linear step problems, instead of our simpler choice of incomplete factorization preconditioners.
-A restructured multilevel CD approach which applies the additive/parallel CD iteration \cite{Tai2003} might be considered, in pursuit of a VI algorithm analogous to the BPX multigrid method for PDEs \cite{BramblePasciakXu1990}.
-However, the performance observed in Section \ref{sec:results} is already excellent on the tested examples.
+In a different direction, a restructured multilevel CD approach which applies the additive/parallel CD iteration \cite{Tai2003} might be considered, in pursuit of a VI algorithm analogous to the BPX multigrid method for PDEs \cite{BramblePasciakXu1990}.
+In any case, the performance observed in Section \ref{sec:results} is already excellent on the tested examples.
 

### 2023-08-14 paper/fascd.tex
cite Hintermuller & Kopacka (2011), and massage the prior-literature paragraph

@@
-\documentclass[review,hidelinks,onefignum,onetabnum]{siamart220329}
+\documentclass[review,hidelinks,onefignum,onetabnum,final]{siamart220329}
 
@@
 PEF was supported by the Engineering and Physical Sciences Research Council [EPSRC grants EP/R029423/1 and EP/W026163/1].
-This work used the ARCHER2 UK National Supercomputing Service (https://www.archer2.ac.uk).
-PEF thanks Lawrence Mitchell for advice on implementation, and Jack D.~Betteridge for assistance with Example \ref{ex:results:sia}.}}}
+This work used the ARCHER2 UK National Supercomputing Service \href{https://www.archer2.ac.uk}{www.archer2.ac.uk}).
+PEF thanks Lawrence Mitchell for advice on implementation, and Jack D.~Betteridge for assistance with running Example \ref{ex:results:sia}.}}}
 
@@
 
-\elb{address H\&K 2011 in following paragraph}
-
 Prior multilevel algorithms for VIs have limitations compared to the FASCD framework.
-The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} applies to linear complementarity problems and projected (truncated) Gauss--Seidel smoothers.
+The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} applies to linear complementarity problems and it uses projected (truncated) Gauss--Seidel smoothers.
 While FASCD reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed, PFAS reduces to linear multigrid.
+Another work which combines FAS and VIs is \cite{HintermullerKopacka2011}, but the problems solved, and the approach, are completely different.
+For an optimal control problem with an elliptic VI constraint, \cite{HintermullerKopacka2011} applies FAS to a system of PDEs derived from penalizing the VI and the control constraints; their V- and W-cycles are not altered in any manner to maintain admissibility.
 On the other hand, the multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so as stated it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
-The latter property is improved by reformulation using defect constraints and monotone injection operators (see \cite[Algorithm 4.2]{GraeserKornhuber2009}, or Section \ref{sec:cdmultilevel} herein).
+The latter limitation is removed by Gr\"aser \& Kornhuber's \cite[Algorithm 4.2]{GraeserKornhuber2009} reformulation using defect constraints and monotone injection operators; see Section \ref{sec:cdmultilevel} below.
 The monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be formulated only with Gauss--Seidel type smoothers, and may have been applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.
-Unlike the fastest variants of monotone multigrid, no zeroing of the basis functions associated with the active set (truncation) is employed in FASCD.
-This simplifies the implementation and allows the coarse grid corrections to update the estimate of the active set on the fine grid.
-Numerical evidence indicates that FASCD coarse corrections are effective even when the coarse meshes are too coarse to resolve the active set.
-On a problem for which the truncated Newton/monotone multigrid methods were designed \cite[problem 7.1.1]{GraeserKornhuber2009}, the performance of FASCD is at least as good (Example \ref{ex:results:classical}).
+Furthermore, unlike the fastest variants of monotone multigrid, no zeroing of the basis functions associated with the active set (truncation) is employed in FASCD, and this simplifies the implementation.
+Nonetheless FASCD coarse grid corrections can update the estimate of the active set on the fine grid, and indeed the numerical evidence indicates that FASCD coarse corrections are effective even when the coarse meshes are far too coarse to resolve the active set.
+On a problem with a complicated coincidence set, for which the truncated Newton/monotone multigrid methods were apparently designed \cite[problem 7.1.1]{GraeserKornhuber2009}, the performance of FASCD is at least as good (Example \ref{ex:results:classical}).
 A limitation of FASCD is that, as formulated, it only solves bound-constrained VI problems.

### 2023-08-14 paper/fascd.tex
make target publish.zip was used in submission to arxiv, which is now processing

@@
   (\email{patrick.farrell@maths.ox.ac.uk}).}}
+
+\graphicspath{{figs/}{fixfigs/}{fixfigs/vcycle/}{genfigs/}}
 
@@
 \centering
-\includegraphics[width=0.45\textwidth]{genfigs/cartoon.pdf}
+\includegraphics[width=0.45\textwidth]{cartoon.pdf}
 \caption{A constraint decomposition (CD) for a unilateral obstacle problem on a two-point space $\Omega=\{x_1,x_2\}$, with $\mathcal{V}=\left\{v \,:\, \Omega \to \RR\right\}$ and $\mathcal{K}=\{v\ge \psi\}$.}
@@
 \centering
-\includegraphics[width=0.35\textwidth]{fixfigs/ball-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/spiral-set.png}
+\includegraphics[width=0.35\textwidth]{ball-set.png} \qquad \includegraphics[width=0.35\textwidth]{spiral-set.png}
 \caption{Coincidence sets for the ball and spiral solutions in Example \ref{ex:results:classical}.}
@@
 \centering
-\includegraphics[width=0.65\textwidth]{genfigs/asymprates.pdf}
+\includegraphics[width=0.65\textwidth]{asymprates.pdf}
 \caption{V-cycle asymptotic convergence rates in Example \ref{ex:results:classical}.}
@@
 \centering
-\includegraphics[width=0.35\textwidth]{fixfigs/poll2d-zero-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/poll2d-one-set.png}
+\includegraphics[width=0.35\textwidth]{poll2d-zero-set.png} \qquad \includegraphics[width=0.35\textwidth]{poll2d-one-set.png}
 \caption{Coincidence sets for the 2D results in Example \ref{ex:results:advdiff}: $u=0$ (left) and $u=1$ (right).}
@@
 \centering
-\includegraphics[width=1.0\textwidth]{fixfigs/sialev8scene.png}
+\includegraphics[width=1.0\textwidth]{sialev8scene.png}
 \caption{Example \ref{ex:results:sia} solution over a bumpy bedrock topography (grey), with 100-times vertical exaggeration and surface coloring by flow speed.}

### 2023-08-14 paper/tikz/imagesvcycle.tex
implement figure reductions/removals, and some text edits; do more after finalizing results
make target publish.zip was used in submission to arxiv, which is now processing

@@
-\begin{tikzpicture}[scale=1.0]
-    \pgfmathsetmacro{\HEIGHT}{2}
+\begin{tikzpicture}[scale=0.75]
+    \pgfmathsetmacro{\HEIGHT}{1.5}
     \pgfmathsetmacro{\XSTEP}{3.7}
-    \pgfmathsetmacro{\YSTEP}{3.7}
+    \pgfmathsetmacro{\YSTEP}{4.3}
     \pgfmathsetmacro{\YLABEL}{-0.4}
     \pgfmathsetmacro{\NUDGE}{0.5}
-    \node (start2) [] at (0,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/start2.png}};
+    % regarding /includegraphics commands, note /graphicspath in main file
+    \node (start2) [] at (0,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{start2.png}};
     \node (start2label) [below of = start2, yshift=\YLABEL cm] {$w^2 \ge \underline{\gamma}^2$};
-    \node (down2) [] at (1.0*\XSTEP,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/down2.png}};
+    \node (down2) [] at (1.0*\XSTEP,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{down2.png}};
     \node (down2label) [below of = down2, yshift=\YLABEL cm] {$y^2 \ge \underline{\phi}^2$};
-    \node (down1) [] at (1.4*\XSTEP,1*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/down1.png}};
+    \node (down1) [] at (1.4*\XSTEP,1*\YSTEP) {\includegraphics[height=\HEIGHT cm]{down1.png}};
     \node (down1label) [below of = down1, yshift=\YLABEL cm] {$y^1 \ge \underline{\phi}^1$};
-    \node (up0) [] at (1.9*\XSTEP,0) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/up0.png}};
+    \node (up0) [] at (1.9*\XSTEP,0) {\includegraphics[height=\HEIGHT cm]{up0.png}};
     \node (up0label) [below of = up0, yshift=\YLABEL cm] {$z^0 \ge \underline{\chi}^0$};
-    \node (up1) [] at (2.4*\XSTEP,1*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/up1.png}};
+    \node (up1) [] at (2.4*\XSTEP,1*\YSTEP) {\includegraphics[height=\HEIGHT cm]{up1.png}};
     \node (up1label) [below of = up1, yshift=\YLABEL cm] {$z^1 \ge \underline{\chi}^1$};
-    \node (up2) [] at (2.8*\XSTEP,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/up2.png}};
+    \node (up2) [] at (2.8*\XSTEP,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{up2.png}};
     \node (up2label) [below of = up2, yshift=\YLABEL cm] {$z^2 \ge \underline{\chi}^2$};
-    \node (next2) [] at (3.8*\XSTEP,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{fixfigs/vcycle/next2.png}};
+    \node (next2) [] at (3.8*\XSTEP,2*\YSTEP) {\includegraphics[height=\HEIGHT cm]{next2.png}};
     \node (next2label) [below of = next2, yshift=\YLABEL cm] {$\tilde{w}^2 \ge \underline{\gamma}^2$};

### 2023-08-18 paper/fascd.tex
after submission arxiv checks for line numbers .. which they do not allow
Fix bug in Example 8.2.
The error is seen in the unregularized full cycle results
in the old plap1d.txt file.  Some such runs converged
because of du=0 steps, so _stol was satisfied.
The regularization is needed, at least for full cycles.
In these results, which are better than before, all runs
use -epsreg 1.0-8 (new default) and -fascd_stol 0.

@@
-\documentclass[review,hidelinks,onefignum,onetabnum,final]{siamart220329}
+%\documentclass[hidelinks,onefignum,onetabnum,final]{siamart220329}  % for arxiv
+\documentclass[review,hidelinks,onefignum,onetabnum,final]{siamart220329}  % for submission
 
@@
 \end{abstract}
+
+% REQUIRED NOVELTY STATEMENT FOR SIAM SUBMISSION:
+% Our new multilevel algorithm for general variational inequality problems requires neither an objective function nor linearity of the residual.
+It is carefully constructed to reduce to general, nonlinear FAS for PDEs when the inequality constraints are removed (this is new) and the method is smoother-agnostic (this is also new relative to any prior work with demonstrated implementations).
+Box constraints, i.e.~both upper and lower obstacles, are allowed.
+The method is strictly admissible, and so can be applied even if the residual operator is defined only on those functions which satisfy the constraints; this is new relative to a variety of semi-smooth and/or penalized solvers.
+Besides the V-cycle, a full multigrid cycle is proposed.
+In a new theoretical aspect, we identify a telescoping subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
+High resolution and performance computational results, using a reduced-space Newton method as a smoother, are given, with evidence of numerical accuracy, convergence, (serial scaling) optimality, and weak parallel scaling.
 
@@
 We again consider Example \ref{ex:plaplacian}, with a unilateral constraint, but now for a nonlinear $p$-Laplacian \cite{ChoeLewis1991} in one dimension.
-Setting $p=1.5$ yields a fast diffusion ($D=|\grad u|^{p-2}\to \infty$ as $|\grad u|\to 0$).
-Note that Example \ref{ex:results:sia} below, by constrast, has a doubly-nonlinear operator with $p=4$ degenerate diffusion.
-
-This example reveals the strong contrast between $V(1,0)$ and $V(0,1)$ results (\pr{up}, \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively).
-Set $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, $\mathcal{K} = \{v \ge \psi\} \subset W^{1,p}(\Omega)$, and $\ip\ell v = \int_\Omega g v$, where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$, in VI \eqref{eq:vi}.
+Setting $p=1.5$ yields a fast diffusion with $D=|\grad u|^{p-2}\to \infty$ as $|\grad u|\to 0$.
+(By contrast, Example \ref{ex:results:sia} below has a doubly-nonlinear operator with $p=4$ degenerate diffusion.)  The diffusivity is regularized, with $|\grad v|^p$ in \eqref{eq:plaplacianobjective} replaced by $\left(\eps + |\grad v|^2\right)^{p/2}$, and $\eps=10^{-8}$.
+We set $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, $\mathcal{K} = \{v \ge \psi\} \subset W^{1,p}(\Omega)$, and $\ip\ell v = \int_\Omega g v$, where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$, in VI \eqref{eq:vi}.
 The exact solution is easily computed for any $p$; see the codes accompanying this paper.
 For this experiment the coarsest mesh has $m_0=7$ nodes and the finest $m_9=3073$.
-The default smoother uses three iterations of RS Newton, with LU soluton of the linear systems.
-We set \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $=$ \texttt{stol} $= 10^{-12}$.
+The default smoother uses three iterations of RS Newton, with LU solution of the linear systems.
+We set \texttt{rtol} $= 10^{-6}$, \texttt{atol} $= 10^{-12}$, and \texttt{stol} $=0$.
 Figure \ref{fig:imagesvcycle} illustrates the major steps of a V-cycle, showing each of the coarse-correction VI subproblems and solutions; the result $\tilde w^2$ is visually close to the exact solution.
@@
 
-Iteration results are shown in Table \ref{tab:results:fastplap1d}.
-We see that $V(1,0)$ cycles, Algorithm 4.2 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.
-By contrast, $V(0,1)$ cycles, up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are as good as the default $V(1,1)$ cycles (Algorithm \ref{alg:fascd}).
-FMG is even more efficient; a single iteration suffices at all higher resolutions.
+Iteration results in Table \ref{tab:results:fastplap1d} show a strong contrast between $V(1,0)$ and $V(0,1)$ results (\pr{up} $=0$ and \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively).
+The $V(1,0)$ cycles, Algorithm 4.2 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.
+By contrast, $V(0,1)$ cycles, up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are as good as the default $V(1,1)$ cycles.
+FMG is even more efficient; a single iteration suffices at all resolutions.
 Note that convergence is at rate $O(h^{2.0})$; compare \cite{ChoeLewis1991}.
@@
 \midrule
- 1 &    $7$ &  1 &  1 &  1 &  1 & $2.3\times 10^{-1}$ \\
- 2 &   $13$ &  3 &  3 &  2 &  2 & $3.3 \times 10^{-2}$ \\
+ 1 &    $7$ &  1 &  1 &  1 &  1 & $2.3 \times 10^{-1}$ \\
+ 2 &   $13$ &  3 &  2 &  2 &  1 & $3.3 \times 10^{-2}$ \\
  3 &   $25$ & 12 &  4 &  2 &  1 & $9.1 \times 10^{-3}$ \\
@@
  7 &  $385$ & NC &  3 &  3 &  1 & $4.7 \times 10^{-5}$ \\
- 8 &  $769$ & NC &  4 &  7 &  1 & $9.5 \times 10^{-6}$ \\
- 9 & $1537$ & NC &  5 &  6 &  1 & $3.6 \times 10^{-6}$ \\
-10 & $3073$ & NC & 11 & 19 &  1 & $4.9 \times 10^{-7}$ \\
+ 8 &  $769$ & NC &  3 &  6 &  1 & $9.2 \times 10^{-6}$ \\
+ 9 & $1537$ & NC &  3 &  3 &  1 & $3.4 \times 10^{-6}$ \\
+10 & $3073$ & NC &  2 &  4 &  1 & $4.1 \times 10^{-7}$ \\
 \bottomrule

### 2023-08-18 paper/results/plap1d.m

@@ -1,17 +1,11 @@
 % generate convergence rates from results in plap1d.txt
+% errors are from default V(1,1) runs
 
 h = 2.^(-(0:9));
-fasterr = [2.263e-01 3.255e-02 9.108e-03 3.248e-03 5.505e-04 ...
-           1.690e-04 4.717e-05 9.522e-06 3.591e-06 4.948e-07];
-%degerr = [8.700e-02 3.812e-02 1.589e-02 6.318e-03 2.330e-03 ...
-%          6.992e-04 2.047e-04 1.078e-04 3.997e-05 1.506e-05];
-%loglog(h,fasterr,'*',h,degerr,'o')
-loglog(h,fasterr,'*')
+err = [2.263e-01 3.255e-02 9.108e-03 3.248e-03 5.501e-04 ...
+       1.684e-04 4.656e-05 9.203e-06 3.430e-06 4.139e-07];
+loglog(h,err,'*')
 grid on,  xlabel h,  ylabel('|error|_{inf}')
-%legend('p=1.5','p=4.0')
 legend('p=1.5')
-fastp = polyfit(log(h),log(fasterr),1)
-%degp = polyfit(log(h),log(degerr),1)
-
-%   fastp = 2.0047  -1.7579
-%   degp  = 1.4203  -2.2802
+p = polyfit(log(h),log(err),1)
+%   p = 2.0047  -1.7579

### 2023-08-18 paper/results/plap1d.sh

@@ -6,10 +6,11 @@
 
 LLIST="1 2 3 4 5 6 7 8 9 10"
 
+# do: V(1,0), V(0,1), V(1,1)
 for VOPT in "-fascd_skip_upsmooth" "-fascd_skip_downsmooth" ""; do
     echo "******** p = 1.5 V-cycle VOPT=${VOPT} *******"
     for LEVS in $LLIST; do
-        python3 plap1d.py -p 1.5 $VOPT -levs $LEVS
+        python3 plap1d.py $VOPT -p 1.5 -fascd_stol 0.0 -levs $LEVS
     done
     echo ""
 done
@@ -17,21 +18,7 @@ done
 for EPSREG in 0.0 1.0e-8; do
     echo "******** p = 1.5 F-cycle EPSREG=${EPSREG} *******"
     for LEVS in $LLIST; do
-        python3 plap1d.py -fascd_cycle_type full -p 1.5 -epsreg $EPSREG -levs $LEVS
+        python3 plap1d.py -fascd_cycle_type full -epsreg $EPSREG -p 1.5 -fascd_stol 0.0 -levs $LEVS
     done
     echo ""
 done
-
-# for ITS in 3 4; do
-#     echo "******** p = 4.0 V-cycle ITS=${ITS} *******"
-#     for LEVS in $LLIST; do
-#         python3 plap1d.py -p 4.0 -fascd_levels_snes_max_it $ITS -levs $LEVS
-#     done
-#     echo ""
-# done
-
-# echo "******** p = 4.0 F-cycle *******"
-# for LEVS in $LLIST; do
-#     python3 plap1d.py -fascd_cycle_type full -p 4.0 -levs $LEVS
-# done
-# echo ""

### 2023-11-04 talk-dms/references.tex
basic slides by copy of oxford talk ... much to do

--- 2023-02-09 talk/references.tex
@@
 \begin{itemize}
-\item A.~Brandt \& C.~Cryer (1983). \emph{Multigrid algorithms for the solution of linear complementarity problems \dots}, SIAM J.~Sci.~Stat.~Comput.~4 (4), 655--684 \sdoi{10.1137/0904046}
-%FTITLE Multigrid algorithms for the solution of linear complementarity problems arising from free boundary problems
-\item P.~Brune, M.~Knepley, B.~Smith, \& X.~Tu (2015). \emph{Composing scalable nonlinear algebraic solvers}, SIAM Review~57 (4), 535--565 \sdoi{10.1137/130936725}
+\item A.~Brandt (1977). \emph{Multi-level adaptive solutions to boundary-value problems}, Mathematics of Computation 31 (138), 333--390
+%\item A.~Brandt \& C.~Cryer (1983). \emph{Multigrid algorithms for the solution of linear complementarity problems \dots}, SIAM J.~Sci.~Stat.~Comput.~4 (4), 655--684 \sdoi{10.1137/0904046}
+%%FTITLE Multigrid algorithms for the solution of linear complementarity problems arising from free boundary problems
+%\item P.~Brune, M.~Knepley, B.~Smith, \& X.~Tu (2015). \emph{Composing scalable nonlinear algebraic solvers}, SIAM Review~57 (4), 535--565 \sdoi{10.1137/130936725}
 \item E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032 \sdoi{10.1137/20M135217X}
 %\item E.~Bueler (2022). \emph{Performance analysis of high-resolution ice-sheet simulations}, J.~Glaciol., \sdoi{10.1017/jog.2022.113}
-\item G.~de Diego, P.~Farrell, \& I.~Hewitt (2022). \emph{Numerical approximation of viscous contact problems applied to glacial sliding}, J.~Fluid Mech.~938 (A21). \sdoi{10.1017/jfm.2022.178}
-\item P.~Farrell, M.~Croci, \& T.~Surowiec (2020). \emph{Deflation for semismooth equations}, Optimization Methods \& Software 35 (6), 1248--1271 \sdoi{10.1080/10556788.2019.1613655}
+\item E.~Bueler \& P.~Farrell (2023).
+\emph{A full approximation storage multilevel method for nonlinear variational inequalities}, submitted \, {\tiny \href{https://arxiv.org/abs/2308.06888}{arxiv:2308.06888}}
+%\item G.~Gonzalez de Diego, P.~Farrell, \& I.~Hewitt (2022).
+\emph{Numerical approximation of viscous contact problems applied to glacial sliding}, J.~Fluid Mech.~938 (A21).
+\sdoi{10.1017/jfm.2022.178}
+%\item P.~Farrell, M.~Croci, \& T.~Surowiec (2020). \emph{Deflation for semismooth equations}, Optimization Methods \& Software 35 (6), 1248--1271 \sdoi{10.1080/10556788.2019.1613655}
 \item C.~Gr{\"a}ser \& R.~Kornhuber (2009). \emph{Multigrid methods for obstacle problems}, J.~Comput.~Math., 1--44
@@
 %FTITLE Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics
-\item G.~Jouvet \& E.~Bueler (2012). \emph{Steady, shallow ice sheets as obstacle problems \dots}, SIAM J.~Appl.~Math.~72 (4), 1292--1314 \sdoi{10.1137/110856654}
-%FTITLE Steady, shallow ice sheets as obstacle problems: well-posedness and finite element approximation
+\item G.~Jouvet \& E.~Bueler (2012).
+\emph{Steady, shallow ice sheets as obstacle problems: well-posedness and finite element approximation}, SIAM J.~Appl.~Math.~72 (4), 1292--1314 \sdoi{10.1137/110856654}
 %\item G.~Jouvet \& J.~Rappaz (2011). \emph{Analysis and finite element approximation of a nonlinear stationary {S}tokes problem \dots}, Adv.~Numer.~Analysis 2011 (164581) \sdoi{10.1155/2011/164581}
 %FTITLE Analysis and finite element approximation of a nonlinear stationary {S}tokes problem arising in glaciology
-\item R.~Kornhuber (1994). \emph{Monotone multigrid methods for elliptic variational inequalities I}, Numer.~Math.~69, 167--184 \sdoi{10.1007/BF03325426}
-\item A.~Reusken (1987). \emph{Convergence of the multigrid full approximation scheme \dots}, Numer.~Math.~52, 251--277 \sdoi{10.1007/BF01398879}
-%FTITLE Convergence of the multigrid full approximation scheme for a class of elliptic mildly nonlinear boundary value problems
+%\item R.~Kornhuber (1994). \emph{Monotone multigrid methods for elliptic variational inequalities I}, Numer.~Math.~69, 167--184 \sdoi{10.1007/BF03325426}
+\item A.~Reusken (1987).
+\emph{Convergence of the multigrid full approximation scheme for a class of elliptic mildly nonlinear boundary value problems}, Numer.~Math.~52, 251--277 \sdoi{10.1007/BF01398879}
 %\item J.~Seguinot \& 5 others (2018).  \emph{Modelling last glacial cycle ice dynamics in the Alps}, The Cryosphere, 12 (10), 3265--3285 \sdoi{10.5194/tc-12-3265-2018}

### 2023-11-04 talk-dms/slides.tex
basic slides by copy of oxford talk ... much to do

--- 2023-02-09 talk/slides.tex
@@
 
-\title[Multigrid for nonlinear and nonlocal VIs]{Toward nonlinear multigrid \\ for nonlinear and nonlocal variational inequalities}
+\title[Fast solvers for PDEs subject to inequalities]{Fast solvers for partial differential equations \\ subject to inequalities}
 
@@
 
-\date[]{February 2023}
+\date[]{November 2023}
 
@@
 
-\titlegraphic{\hfill \includegraphics[width=0.15\textwidth]{images/uafbw.png} \quad \includegraphics[width=0.10\textwidth]{images/oxford.png}}
+\titlegraphic{\vspace{-8mm} \includegraphics[width=0.25\textwidth]{../talk-oxford/images/uafbw.png} \hfill \begin{minipage}{0.16\textwidth}
+\includegraphics[width=\textwidth]{../talk-oxford/images/oxford.png}
+
+\vspace{18mm}
+\end{minipage} \vspace{-10mm}}
 
@@
 {
-  %\usebackgroundtemplate{\includegraphics[width=\paperwidth]{images/gray-british-clark2022.png}}
+  %\usebackgroundtemplate{\includegraphics[width=\paperwidth]{../talk-oxford/images/gray-british-clark2022.png}}
   \begin{frame}
@@
 \begin{frame}{Outline}
-  \tableofcontents[hideallsubsections]
+  %\tableofcontents[hideallsubsections]
+  \tableofcontents
 \end{frame}
@@
 
-\includegraphics[width=0.55\textwidth]{images/obstacle65.pdf} \qquad \includegraphics[width=0.35\textwidth]{images/obstacle-sets.png}
+\includegraphics[width=0.55\textwidth]{../talk-oxford/images/obstacle65.pdf} \qquad \includegraphics[width=0.35\textwidth]{../talk-oxford/images/obstacle-sets.png}
 
@@
 \item elastic contact, Signorini problems (e.g.~Kikuchi \& Oden 1988)
-\item viscous contact problems (de Diego et al.~2022)
+\item viscous contact problems (Gonzalez de Diego et al.~2022)
 \item pricing of American options in the Black-Scholes model
@@
 \begin{center}
-\includegraphics[width=0.9\textwidth]{images/calcone.png}
+\includegraphics[width=0.9\textwidth]{../talk-oxford/images/calcone.png}
 \end{center}
@@
 
-\section{full approximation scheme (FAS) multigrid for PDEs}
+\section{multigrid for PDEs}
+
+\subsection{nonlinear multigrid = full approximation scheme (FAS)}
 
@@
 \begin{center}
-$\Omega^h$\, \includegraphics[height=0.16\textheight]{images/fine-grid.png} \hspace{25mm} \includegraphics[height=0.16\textheight]{images/coarse-grid.png} \,$\Omega^H$
+$\Omega^h$\, \includegraphics[height=0.16\textheight]{../talk-oxford/images/fine-grid.png} \hspace{25mm} \includegraphics[height=0.16\textheight]{../talk-oxford/images/coarse-grid.png} \,$\Omega^H$
 \end{center}
@@
 
-\includegraphics[height=0.15\textheight]{images/mg-grids.png}
+\includegraphics[height=0.15\textheight]{../talk-oxford/images/mg-grids.png}
 
@@
 \only<1>{
-\hfill \includegraphics[height=0.15\textheight]{images/mg-vcycle.png}
+\hfill \includegraphics[height=0.15\textheight]{../talk-oxford/images/mg-vcycle.png}
 }
-\only<2>{\hfill \includegraphics[height=0.15\textheight]{images/mg-fcycle.png} \quad \begin{tabular}{l} F-cycle = \\ nested iteration \end{tabular}
+\only<2>{\hfill \includegraphics[height=0.15\textheight]{../talk-oxford/images/mg-fcycle.png} \quad \begin{tabular}{l} F-cycle = \\ nested iteration \end{tabular}
 }
@@
 
-\begin{frame}[fragile]
-\frametitle{multigrid solver composition in PETSc}
-
-\begin{itemize}
-\item implemented here using an FD discretization and PETSc:\footnote{Portable Extensible Toolkit for Scientific computing \quad \href{https://petsc.org/release/}{\texttt{petsc.org}} \quad \includegraphics[height=3mm]{images/petsc.png}}
-    \begin{itemize}
-    \item[$\circ$] multigrid solvers in PETSc are \emph{composed} from smoothers on each level, and a coarse-level solver \dots here these are nonlinear GS
-    \item[$\circ$] FAS multigrid is a nonlinear solver (SNES) type
-    \item[$\circ$] PETSc $=$ C/Fortran/python
-    \end{itemize}
-\item FAS multigrid F-cycle:
-\end{itemize}
-\begin{Verbatim}[xleftmargin=15mm,fontsize=\scriptsize]
-./bratu -da_grid_x 5 -da_grid_y 5 -da_refine J \
-    -snes_rtol 1.0e-12 \
-    -snes_type fas \
-    -snes_fas_type full \
-    -fas_levels_snes_type ngs \
-    -fas_levels_snes_ngs_sweeps 2 \
-    -fas_levels_snes_ngs_max_it 1 \
-    -fas_levels_snes_norm_schedule none \
-    -fas_coarse_snes_type ngs \
-    -fas_coarse_snes_max_it 1 \
-    -fas_coarse_snes_ngs_sweeps 4 \
-    -fas_coarse_snes_ngs_max_it 1 
-\end{Verbatim}
-
-%NOT IN MY BOOK \hfill \includegraphics[width=8mm]{images/frontcover.jpg}
-\end{frame}
-
 \begin{frame}{Bratu model problem: optimality}
@@
 \begin{column}{0.55\textwidth}
-\includegraphics<1>[width=\textwidth]{images/bratu-exps.png}
+\includegraphics<1>[width=\textwidth]{../talk-oxford/images/bratu-exps.png}
 
-\includegraphics<2>[width=\textwidth]{images/bratu-time.png}
+\includegraphics<2>[width=\textwidth]{../talk-oxford/images/bratu-time.png}
 \end{column}
@@
 
-\section{the nonlinear and nonlocal VI for a fluid layer in a climate}
-
 \begin{frame}{problem: fluid layer in a climate}
@@
 \bigskip
-\hfill \mbox{\includegraphics[height=0.25\textheight]{images/domain-data.png} \hspace{7mm} \includegraphics[height=0.25\textheight]{images/domain-velocity.png}}
+\hfill \mbox{\includegraphics[height=0.25\textheight]{../talk-oxford/images/domain-data.png} \hspace{7mm} \includegraphics[height=0.25\textheight]{../talk-oxford/images/domain-velocity.png}}
 \end{frame}
@@
 
-\includegraphics[width=1.02\textwidth]{images/alps-seguinot2018.png}
+\includegraphics[width=1.02\textwidth]{../talk-oxford/images/alps-seguinot2018.png}
 
@@
 \bigskip
-\hfill \mbox{\includegraphics[height=0.2\textheight]{images/domain-data.png} \hspace{7mm} \includegraphics[height=0.2\textheight]{images/domain-velocity.png}}
+\hfill \mbox{\includegraphics[height=0.2\textheight]{../talk-oxford/images/domain-data.png} \hspace{7mm} \includegraphics[height=0.2\textheight]{../talk-oxford/images/domain-velocity.png}}
 \end{frame}
@@
 \begin{itemize}
-\only<1>{
-\item \alert{Stokes model}
-
-solve the Stokes problem, then evaluate velocity at surface:
-    $$\int_{\Lambda(s) = \{b < z <s\}} 2 \nu(D\bu) D\bu : D\bv - p \Div\bv - (\Div\bu) q - \rhoi \bg \cdot \bv = 0 \quad \forall \bv,q$$
-    $$\Phi(s) = - \bu|_s \cdot \bn_s$$
-
-    \begin{itemize}
-    \item[$\circ$] assuming incompressibility and non-Newtonian viscosity: $\nu(D\bu) = \frac{1}{2} \Gamma |D\bu|^{\pp-2}$ with $\pp = \frac{4}{3}$
-    \item[$\circ$] given $s$, this is a well-posed problem for velocity $\bu\in \mathbf{W}^{1,\pp}$ and pressure $p \in L^\qq$ on domain $\Lambda(s)$
-    \item[$\circ$] near-optimal solvers available (Isaac et al 2015)
-    \end{itemize}
-}
-
-\only<2>{
 \item \alert{lubrication approximation}\footnote{also known as the \emph{shallow ice approximation}} \alert{model}
@@
     \end{itemize}
-}
 \end{itemize}
@@
 
-\begin{frame}{CP form of viscous fluid layer in a climate}
+\section{multigrid for VIs}
 
-\begin{itemize}
-\item VI form on previous slide is too abstracted for clarity
-\item the strong form of the same problem is a \alert{complementarity problem (CP) coupled to a Stokes problem}:
-\begin{align*}
-s - b &\ge 0 && \text{ in $\Omega \subset \RR^2$} \\
-- \bu|_s \cdot \bn_s - a &\ge 0 && \text{''} \\
-(s-b)(- \bu|_s \cdot \bn_s - a) &= 0 && \text{''} \\
-- \nabla \cdot \left(2 \nu(D\bu)\, D\bu\right) + \nabla p - \rhoi \mathbf{g} &= \bzero && \text{in $\Lambda(s) \subset \RR^3$} \\
-\nabla \cdot \bu &= 0 && \text{''} \\
-\left(2 \nu(D\bu) D\bu - pI\right) \bn &= \bzero && \{z=s\} \subset \partial \Lambda(s) \\
-\bu &= \bzero && \{z=b\} \subset \partial \Lambda(s)
-\end{align*}
-\item solve this for $s$ on $\Omega$, and simultaneously for $\bu,p$ on $\Lambda(s) = \{b < z < s\}$
-\end{itemize}
-\end{frame}
-
-\begin{frame}{a non-local VI problem}
-
-\begin{itemize}
-\item in the Stokes case, the residual $r(s) = a - \Phi(s)$ $= a + \bu|_s \cdot \bn_s$ \alert{depends non-locally on $s$}
-\item for example, consider $\bu_{(s+\psi)} - \bu_{(s)}$ from surface perturbation
-
-(hat function) $\psi$
-\end{itemize}
-
-\medskip
-\begin{center}
-\includegraphics[width=0.85\textwidth]{images/stokes-greens-arndt.png}
-\end{center}
-
-\vspace{-5mm}
-\hfill {\tiny P.~Arndt figure}
-\end{frame}
-
-\begin{frame}{what's needed for multigrid to work here?}
-
-\begin{block}{viscous fluid layer geometry problem}
-	$$\ip{\Phi(s)}{r-s} \ge \ip{a}{r-s} \quad \text{ for all } r \in \mathcal{K}$$
-where
-\begin{itemize}
-\item $\mathcal{K} = \left\{r \in \mathcal{V} \,:\, r \ge b\right\}$
-\item $\Phi(s)=- \bu|_s \cdot \bn_s$
-\item $s$ is solution surface elevation
-\item $\bu$ is Stokes solution on $\Lambda(s) = \{b < z < s\}$
-\item $a$ (climate) and $b$ (bed elevation) are the input data
-\end{itemize} 
-\end{block}
-
-\begin{itemize}
-\item[]
-    \begin{itemize}
-    \item[$\circ$] for more on this problem class see (Bueler, 2021)
-    \end{itemize}
-\end{itemize}
-
-\begin{itemize}
-\item {\color{FireBrick} what is needed for scalable multilevel solutions?}
-    \begin{enumerate}
-    \item[1.] iterates must be admissible
-    \item[2.] global linearization of $\Phi(s)$ must be avoided
-    \item[3.] smoother cost must be comparable to one residual    
-    \end{enumerate}
-\end{itemize}
-\end{frame}
-
-\section{multigrid approaches for VIs}
+\subsection{earlier approaches}
 
@@
 \medskip
-\hspace{10mm} \includegraphics[height=0.2\textheight]{images/obstacle65.pdf}
+\hspace{10mm} \includegraphics[height=0.2\textheight]{../talk-oxford/images/obstacle65.pdf}
 
 \vspace{-20mm}
-\hfill \includegraphics[height=0.25\textheight]{images/vi-newton-gmg-bad.png}
+\hfill \includegraphics[height=0.25\textheight]{../talk-oxford/images/vi-newton-gmg-bad.png}
 \end{frame}
@@
 \vspace{-12mm}
-\hfill \includegraphics[width=0.25\textwidth]{images/mg-fcycle.png}
+\hfill \includegraphics[width=0.25\textwidth]{../talk-oxford/images/mg-fcycle.png}
 \end{itemize}
@@
 \bigskip
-\includegraphics[height=0.25\textheight]{images/vi-newton-gmg-good.png}
+\includegraphics[height=0.25\textheight]{../talk-oxford/images/vi-newton-gmg-good.png}
 
 \vspace{-22mm}
-\hspace{40mm} \includegraphics[height=0.45\textheight]{images/obstacle-flops-per-n.png} \qquad \includegraphics<2>[width=20mm]{images/frontcover.jpg}
-\end{frame}
-
-\begin{frame}{multigrid strategies for VIs}
-
-\begin{itemize}
-\item semi-smooth Newton also yields mesh-independent iterations
-    \begin{itemize}
-    \item[$\circ$] penalty scaling argument (Farrell et al.~2020)
-    \end{itemize}
-\item other VI multilevel strategies:
-    \begin{itemize}
-    \item[$\circ$] projected FAS multigrid for linear CPs (Brandt \& Cryer, 1983)
-    \item[$\circ$] monotone multigrid (Kornhuber, 1994)
-    \item[$\circ$] multilevel constraint decomposition (Tai, 2003)  \hfill $\gets$ \emph{more below}
-    \end{itemize}
-\end{itemize}
+\hspace{40mm} \includegraphics[height=0.45\textheight]{../talk-oxford/images/obstacle-flops-per-n.png} \qquad \includegraphics<2>[width=20mm]{../talk-oxford/images/frontcover.jpg}
 \end{frame}
@@
 
-\section{FAS multigrid for VIs?}
+\subsection{full approximation scheme constraint decomposition (FASCD)}
 
@@
 \vspace{-25mm}
-\hfill \includegraphics[width=0.4\textwidth]{images/fascd-vcycle.png}
-\end{frame}
-
-\begin{frame}{as decomposition of the fluid layer}
-
-\begin{itemize}
-\item again this is too abstract
-\item what does it look like for the fluid layer?
-    \begin{itemize}
-    \item[$\circ$] coarse grids have \emph{admissible} pieces of the fine-grid iterate
-    \end{itemize}
-\end{itemize}
-
-\bigskip
-\begin{center}
-\includegraphics[width=\textwidth]{images/icedecomp.png}
-\end{center}
+\hfill \includegraphics[width=0.4\textwidth]{../talk-oxford/images/fascd-vcycle.png}
 \end{frame}
@@
 
+\section{results}
+
+\subsection{classical obstacle problem}
+
+\subsection{advection-diffusion of a concentration}
+
+\subsection{glacier surface elevations}
+
 \begin{frame}{2D shallow ice approximation results (\emph{very fresh})}
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{images/fascd-firedrake-dome.png}
+\includegraphics[width=0.8\textwidth]{../talk-oxford/images/fascd-firedrake-dome.png}
 \end{center}
@@
 
-\begin{frame}{evidence of mesh independence}
+\begin{frame}{mesh independence}
 
 \begin{itemize}
-\item same lubrication approximation, but in 1D
-\item FASCD V-cycles with NGS and NJacobi smoothers
-\item up-smoothing preferred: V(0,2) beats V(1,1)
-\item evidence of mesh independence of factors $\|r^{(k+1)}\| / \|r^{(k)}\|$
+\item FIXME
 \end{itemize}
-
-\begin{center}
-\includegraphics[width=0.5\textwidth]{images/sia-asymprates.pdf}
-\end{center}
 \end{frame}
@@
 \begin{itemize}
-\item the \alert{variational inequality (VI)} problem class is good to know
-\item likewise \alert{full approximation storage (FAS) multigrid}
-    \begin{itemize}
-    \item[$\circ$] need for better support and documentation in PETSc/Firedrake
-    \end{itemize}
-\item multigrid treatment of nonlinear and \alert{nonlocal VIs}?
-    \begin{itemize}
-    \item[$\circ$] smoothers not obvious in nonlocal cases
-    \item[$\circ$] seeking practical evidence of mesh-independent convergence
-    \end{itemize}
-\item glacier evolution, as fluid-layer-in-climate problems, \alert{needs attention} from applied mathematicians and numerical analysts
-    \begin{itemize}
-    \item[$\circ$] VI form not widely recognized
-    \item[$\circ$] current state of the art $=$ explicit time stepping of surface
-        \begin{itemize}
-        \item[$\vartriangleright$] slow for science, intrinsically \emph{not} scalable
-        \end{itemize}
-    \item[$\circ$] to do: \alert{steady-state and implicit step VI problems}
-    \item[$\circ$] more on this view in my Math.~Geosci.~seminar tomorrow 2pm L5
-    \end{itemize}
+\item FIXME
 \end{itemize}
@@
 
-\begin{frame}{additional background references}
+\begin{frame}{background references}
 

### 2023-11-04 talk-dms/slides.tex
a bit of progress

@@
 \hypersetup{
-    pdftitle = {Toward nonlinear multigrid for nonlinear and nonlocal variational inequalities},
+    pdftitle = {Fast solvers for PDEs subject to inequalities},
     pdfauthor = {Ed Bueler and Patrick Farrell},
@@
 
-%\begin{frame}
-%  \maketitle
-%\end{frame}
-
 {
@@
 \begin{frame}{Outline}
-  %\tableofcontents[hideallsubsections]
-  \tableofcontents
+  \tableofcontents[hideallsubsections]
+  %\tableofcontents
 \end{frame}
@@
 \section{variational inequalities (VIs)}
+
+\begin{frame}{example: a classical obstacle problem}
+
+\begin{center}
+\includegraphics[width=0.65\textwidth]{../talk-oxford/images/obstacle65.pdf}
+\end{center}
+
+\only<1>{
+\begin{itemize}
+\item \emph{problem.} on a domain $\Omega \subset \RR^2$, find the displacement $u(x)$ of a membrane, with fixed value $u = g$ on $\partial \Omega$, above an \emph{obstacle} $\psi(x)$, which minimizes the elastic energy
+    $$J(v) = \int_\Omega \frac{1}{2} |\grad v|^2 - f\, v$$
+\item shown above: \quad $\Omega$ a square, $\psi(x)$ a hemisphere
+\end{itemize}
+
+\phantom{x}
+}
+\only<2>{
+\begin{itemize}
+\item i.e.~constrained optimization over an \emph{admissible set}
+	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \text{ and } v \ge \psi\right\}$$
+\item $\mathcal{K}$ is a closed and convex subset of a Sobolev space:
+	$$H^1(\Omega) = \left\{v \,:\, \int_\Omega |v|^2 + |\grad v|^2 < \infty\right\}$$
+\end{itemize}
+}
+\end{frame}
 
@@
 \begin{itemize}
-\item \emph{problem.} on a domain $\Omega \subset \RR^2$, find the displacement $u(x)$ of a membrane, with fixed value $u = g$ on $\partial \Omega$, above an \emph{obstacle} $\psi(x)$, which minimizes the elastic energy
-    $$J(v) = \int_\Omega \frac{1}{2} |\grad v|^2 - f\, v$$
-\item shown above: \quad $\Omega=(-2,2)^2$, $\psi(x)$ a hemisphere, $f(x)=0$
-\end{itemize}
-
-\phantom{x}
-}
-\only<2>{
-\begin{itemize}
-\item i.e.~constrained optimization over a convex \emph{admissible set}
-	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \text{ and } v \ge \psi\right\}$$
-\item $J'(u)$ points directly into $\mathcal{K}$, the {\color{FireBrick} \emph{variational inequality} (VI)}:
-    $$\ip{J'(u)}{v-u} = \int_\Omega \grad u\cdot \grad (v-u) - f (v-u) \ge 0 \quad \text{for all } v \in \mathcal{K}$$
-\end{itemize}
-}
-\only<3>{
-\begin{itemize}
 \item the solution defines \emph{active} $A_u = \{u = \psi\}$ and \emph{inactive} $R_u = \{u> \psi\}$ subsets of $\Omega$, and a \emph{free boundary} $\Gamma_u=\partial R_u \cap \Omega$
@@
 
-\phantom{x}
-
-\phantom{x}
+   \begin{itemize}
+   \item[$\circ$] this is not acceptable \dots
+   \end{itemize}
 \end{itemize}
 }
-\only<4>{
+\only<2>{
 \begin{itemize}
@@
 \end{align*}
+
    \begin{itemize}
-   \item[$\circ$] for optimization problems: \quad CP $=$ KKT conditions
+   \item[$\circ$] CP $=$ (KKT conditions in $\infty$-dimensions)
    \end{itemize}
@@
 \phantom{x}
+\end{itemize}
+}
+\only<3>{
+\begin{itemize}
+\item FIXME i.e.~constrained optimization over a convex \emph{admissible set}
+	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \text{ and } v \ge \psi\right\}$$
+\item $J'(u)$ points directly into $\mathcal{K}$, the {\color{FireBrick} \emph{variational inequality} (VI)}:
+    $$\ip{J'(u)}{v-u} = \int_\Omega \grad u\cdot \grad (v-u) - f (v-u) \ge 0 \quad \text{for all } v \in \mathcal{K}$$
 \end{itemize}

### 2023-11-05 talk-dms/references.tex
progress slides

@@
 \item E.~Bueler \& P.~Farrell (2023).
-\emph{A full approximation storage multilevel method for nonlinear variational inequalities}, submitted \, {\tiny \href{https://arxiv.org/abs/2308.06888}{arxiv:2308.06888}}
+\emph{A full approximation scheme multilevel method for nonlinear variational inequalities}, submitted \, {\tiny \href{https://arxiv.org/abs/2308.06888}{arxiv:2308.06888}}
 %\item G.~Gonzalez de Diego, P.~Farrell, \& I.~Hewitt (2022).

### 2023-11-05 talk-dms/slides.tex
progress slides

@@
 
+\newcommand{\ds}{\displaystyle}
+
 \newcommand{\bus}{\bu|_s}
@@
 
-%\subtitle{\emph{x}}
+%\subtitle{\emph{}}
 
@@
   %\tableofcontents
+
+\vfill
+{\footnotesize \emph{PDEs = partial differential equations}}
 \end{frame}
@@
 
-\begin{frame}{example: a classical obstacle problem}
+\begin{frame}{example: classical obstacle problem}
 
@@
 \begin{itemize}
-\item \emph{problem.} on a domain $\Omega \subset \RR^2$, find the displacement $u(x)$ of a membrane, with fixed value $u = g$ on $\partial \Omega$, above an \emph{obstacle} $\psi(x)$, which minimizes the elastic energy
+\item \emph{problem.} on a domain $\Omega \subset \RR^2$, find the displacement $u(x)$ of a membrane, with fixed value $u = g$ on $\partial \Omega$, above an \emph{obstacle} $\psi(x)$, which minimizes elastic (plus some potential) energy
     $$J(v) = \int_\Omega \frac{1}{2} |\grad v|^2 - f\, v$$
@@
 \begin{itemize}
-\item i.e.~constrained optimization over an \emph{admissible set}
-	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \text{ and } v \ge \psi\right\}$$
-\item $\mathcal{K}$ is a closed and convex subset of a Sobolev space:
-	$$H^1(\Omega) = \left\{v \,:\, \int_\Omega |v|^2 + |\grad v|^2 < \infty\right\}$$
+\item this is constrained optimization over an \emph{admissible set}
+	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \,\text{ and }\, v \ge \psi\right\}$$
+
+    \begin{itemize}
+    \item[$\circ$] $\mathcal{K}$ is a closed and convex subset of the Sobolev space $H^1(\Omega) = \left\{v \,:\, \int_\Omega |v|^2 + |\grad v|^2 < \infty\right\}$
     \end{itemize}
+\end{itemize}
+
+\vspace{10.5mm}
 }
@@
 \begin{itemize}
-\item the solution defines \emph{active} $A_u = \{u = \psi\}$ and \emph{inactive} $R_u = \{u> \psi\}$ subsets of $\Omega$, and a \emph{free boundary} $\Gamma_u=\partial R_u \cap \Omega$
-\item naive strong form would pose the problem in terms of its solution:
+\item the solution defines subsets of $\Omega$:
+   \begin{itemize}
+   \item[$\circ$] \emph{active set} $A_u = \{u = \psi\}$
+   \item[$\circ$] \emph{inactive set} $R_u = \{u> \psi\}$
+   \item[$\circ$] \emph{free boundary} $\Gamma_u=\partial R_u \cap \Omega$
+   \end{itemize}
+\end{itemize}
+
+\vspace{25mm}
+}
+\only<2-3>{
+\begin{itemize}
+\item a naive strong form would pose the problem in terms of its solution:
 \begin{align*}
@@
    \begin{itemize}
-   \item[$\circ$] this is not acceptable \dots
+   \item[$\circ$] Poisson equation $-\grad^2 u = f$ is ``$J'(u)=0$'' on $R_u$
+   \item<3>[$\circ$] using the solution $u$ to define the set $R_u$ on which to solve the PDE $-\grad^2 u = f$ \alert{is not acceptable}
    \end{itemize}
@@
 }
-\only<2>{
+\only<4>{
 \begin{itemize}
-\item the \emph{complementarity problem} (CP) is meaningful as a strong form:
+\item the \emph{complementarity problem} (CP) is a meaningful strong form:
 \begin{align*}
@@
 }
-\only<3>{
+\only<5>{
 \begin{itemize}
-\item FIXME i.e.~constrained optimization over a convex \emph{admissible set}
-	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \text{ and } v \ge \psi\right\}$$
-\item $J'(u)$ points directly into $\mathcal{K}$, the {\color{FireBrick} \emph{variational inequality} (VI)}:
-    $$\ip{J'(u)}{v-u} = \int_\Omega \grad u\cdot \grad (v-u) - f (v-u) \ge 0 \quad \text{for all } v \in \mathcal{K}$$
+\item the correct weak form is a {\color{FireBrick} \emph{variational inequality} (VI)}
+\item namely, $J'(u)$ points directly into $\mathcal{K}$:
+    $$\ip{J'(u)}{v-u} = \int_\Omega \grad u\cdot \grad (v-u) - f (v-u) \ge 0$$
+for all $v \in \mathcal{K}$
+
+\phantom{x}
+
+\phantom{x}
 \end{itemize}
 }
+\end{frame}
+
+\begin{frame}{VI $=$ weak form with $v-u$ as the test function}
+
+{\small
+   $$J(u) \le J(v) \quad \forall v \in \mathcal{K} \qquad \iff \qquad \ip{J'(u)}{v-u} \ge 0 \quad \forall v \in \mathcal{K}$$
+}
+
+\bigskip
+\begin{center}
+\includegraphics[width=0.5\textwidth]{figs/convexuv.png}
+\end{center}
 \end{frame}
@@
 \item let $\mathcal{K}$ be a closed and convex subset of a Banach space $\mathcal{V}$
-\item suppose $F:\mathcal{K} \to \mathcal{V}'$ is a continuous, generally nonlinear operator
+\item suppose $F:\mathcal{K} \to \mathcal{V}'$ is a continuous operator
     \begin{itemize}
-    \item[$\circ$] $F$ may be defined only on $\mathcal{K}$
+    \item[$\circ$] $F$ is generally nonlinear
+    \item[$\circ$] $F$ may be defined \emph{only} on $\mathcal{K}$
     \item[$\circ$] $F$ may \emph{not}\, be the derivative of an objective function $J$
+    \item[$\circ$] $F=J'$, a linear operator, in classical obstacle problem
     \end{itemize}
@@
 	$${\color{FireBrick} \ip{F(u)}{v-u} \ge 0 \quad \text{ for all } v \in \mathcal{K}}$$
-\item when $\mathcal{K}$ is nontrivial the problem {\color{FireBrick} VI($F$,$\mathcal{K}$)} is nonlinear \emph{even when $F$ is a linear operator}
+\item<2> when $\mathcal{K}$ is nontrivial the problem {\color{FireBrick} VI($F$,$\mathcal{K}$)} is nonlinear \emph{even when $F$ is a linear operator}
 \end{itemize}
@@
 
-\begin{frame}{VI $=$ ``constrained equation''}
+\begin{frame}{VI $=$ constrained ``system of equations''}
 
 \begin{center}
-\begin{tabular}{l|l}
-\begin{minipage}[t][16mm][t]{0.4\textwidth}
-unconstrained optimization:
+\begin{tabular}{r|l|l}
+& \qquad unconstrained & \qquad constrained \\ \hline
+optimization &
+\begin{minipage}[t][16mm][t]{0.32\textwidth}
 $$\min_{u\in\mathcal{V}} J(u)$$
@@
 &
-\begin{minipage}[t][16mm][t]{0.4\textwidth}
-constrained optimization:
-$$\min_{u\in\mathcal{K}} J(u)$$
+\begin{minipage}[t][16mm][t]{0.35\textwidth}
+$$\min_{u\in\mathcal{K} \subset \mathcal{V}} J(u)$$
 \end{minipage}
 \\ \hline
-\begin{minipage}[t][16mm][t]{0.4\textwidth}
-equation for $u \in \mathcal{V}$: {\LARGE \strut}
+\only<1>{equations}\only<2>{\begin{minipage}[t][16mm][t]{0.15\textwidth} weak form \par equations \end{minipage}} &
+\begin{minipage}[t][16mm][t]{0.32\textwidth}
 
-$$F(u)=0$$
+\vspace{-2mm}
+find $u \in \mathcal{V}$:
+\only<1>{$$F(u)=0$$}
+\only<2>{$$\ip{F(u)}{v} = 0 \quad \forall v \in \mathcal{V}$$}
 \end{minipage}
 &
-\begin{minipage}[t][16mm][t]{0.4\textwidth}
+\begin{minipage}[t][16mm][t]{0.35\textwidth}
 
 \vspace{-2mm}
-{\color{FireBrick} VI} for $u \in \mathcal{K}$:
+find $u \in \mathcal{K} \subset \mathcal{V}$:
 $${\color{FireBrick} \ip{F(u)}{v-u} \ge 0 \quad \forall v \in \mathcal{K}}$$
@@
 \begin{itemize}
-\item elastic contact, Signorini problems (e.g.~Kikuchi \& Oden 1988)
-\item viscous contact problems (Gonzalez de Diego et al.~2022)
-\item pricing of American options in the Black-Scholes model
-\item the geometry of glaciers \hfill $\longleftarrow$ \emph{more soon}
-
-\bigskip
-\item first-semester calculus
+\item elastic contact
+    \begin{itemize}
+    \item[$\circ$] car tires, for example
     \end{itemize}
 
+\vspace{-10mm}
+\hfill \includegraphics[width=0.2\textwidth]{figs/tirecontact.png}
+
+\vspace{-20mm}
+\item pricing of American options
+    \begin{itemize}
+    \item[$\circ$] inequality-constrained Black-Scholes model
+    \end{itemize}
+
+\vspace{1.5mm}
+\item the geometry of glaciers %\hfill $\longleftarrow$ \emph{more soon}
+
+\vspace{3mm}
+\item first-semester calculus: \quad minimize $f(x)$ over $\mathcal{K} = [a,b]$
+\end{itemize}
+
+%\vspace{20mm}
 \begin{center}
@@
 
-\subsection{nonlinear multigrid = full approximation scheme (FAS)}
+\subsection{\emph{only} nonlinear multigrid!}
+\subsection{full approximation scheme (FAS)}
 

### 2023-11-05 talk-dms/slides.tex
progress talk

@@
 
-\title[Fast solvers for PDEs subject to inequalities]{Fast solvers for partial differential equations \\ subject to inequalities}
+\title[Fast solvers for PDE+inequalities]{Fast solvers for partial differential equations \\ subject to inequalities}
 
@@
 
-\titlegraphic{\vspace{-8mm} \includegraphics[width=0.25\textwidth]{../talk-oxford/images/uafbw.png} \hfill \begin{minipage}{0.16\textwidth}
+\titlegraphic{\vspace{-8mm} \includegraphics[width=0.26\textwidth]{../talk-oxford/images/uafbw.png} \hfill \begin{minipage}{0.14\textwidth}
 \includegraphics[width=\textwidth]{../talk-oxford/images/oxford.png}
@@
 
-\vfill
-{\footnotesize \emph{PDEs = partial differential equations}}
+%\vfill
+%{\footnotesize \emph{PDEs = partial differential equations}}
 \end{frame}
@@
 
-\begin{frame}{nonlinear 2-grid scheme}
+\begin{frame}{nonlinear 2-mesh scheme}
 
@@
 	\begin{itemize}
-	\item for example, $F : \mathcal{V} \to \mathcal{V}'$ for $\mathcal{V}=H^1(\Omega)$, with $\ell\in \mathcal{V}'$
+	\item[$\circ$] $\mathcal{V}=H^1(\Omega)$, $\ell\in \mathcal{V}'$
+	\item[$\circ$] $F : \mathcal{V} \to \mathcal{V}'$ continuous and injective
+	\item[$\circ$] for example, the Liouville-Bratu problem: $-\grad^2 u - e^u = f(x)$
 	\end{itemize}
-\item discretization gives algebraic system on fine grid $\Omega^h$:
+\item discretization gives algebraic system on fine mesh $\Omega^h$:
     $$F^h(u^h) = \ell^h$$
-\item suppose $w^h$ yields residual norm $\|\ell^h - F^h(w^h)\| > \text{TOL}$
+
+	\begin{itemize}
+	\item[$\circ$] $u^h$ denotes exact (algebraic) solution
+	\end{itemize}
 \end{itemize}
 
-\phantom{x}
+\vspace{7mm}
 }
@@
 \begin{itemize}
+\item \emph{goal}: to solve $F^h(u^h) = \ell^h$ on $\Omega^h$
+\item suppose $w^h$ is a not-yet-converged iterate:
+    $$r^h=\ell^h - F^h(w^h), \qquad \|r^h\| > \text{TOL}$$
 \item how can we improve $w^h$ \emph{without} globally linearizing $F^h$?
 
-(are there alternatives to Newton's method?)
-\item note the \emph{residual} $r^h(w^h) = \ell^h - F^h(w^h)$ is computable,
-
-while the \emph{error} $e^h = w^h-u^h$ is unknown
-\item the residual definition can be rewritten
-    $$F^h(u^h) - F^h(w^h) \stackrel{*}{=} r^h(w^h)$$
-\only<2>{\item for $F^h$ linear, try to solve this \emph{error equation} $F^h(e^h) = -r^h(w^h)$ for $\tilde e^h$, and correct $w^h \leftarrow w^h-\tilde e^h$ to improve $w^h$?}
-\only<3>{\item \st{for $F^h$ linear, try to solve this \emph{error equation} $F^h(e^h) = -r^h(w^h)$ for $\tilde e^h$, and correct $w^h \leftarrow w^h-\tilde e^h$ to improve $w^h$?}}
+	\begin{itemize}
+	\item are there alternatives to Newton's method?
+	\end{itemize}
+\item<3> notes:
+    \setbeamertemplate{enumerate item}{\emph{\roman{enumi})}}
+    \begin{enumerate}
+    \item the \emph{residual} \quad $r^h = \ell^h - F^h(w^h)$ \quad is computable
+    \item the \emph{error} \quad $e^h = u^h-w^h$ \quad is unknown
+    \item our equation can be rewritten
+    $$F^h(u^h) - F^h(w^h) = r^h$$
+    \end{enumerate}
 \end{itemize}
 }
-\only<4>{
+\only<4-5>{
 \begin{itemize}
-\item \emph{goal}: use a coarser mesh to estimate the error in $\ast$
-\item \emph{nodewise problem}: for $\psi_i^h$ a hat function or dof, solve for $c\in\RR$:
+\item \emph{goal}: from iterate $w^h$, to solve
+    $$F^h(u^h) - F^h(w^h) = r^h$$
+\item \alert{for $F^h$ linear}, convert this to the \emph{error equation}
+    $$F^h(e^h) = r^h$$
+\item an approximation solution $\tilde e^h$ would improve our iterate:
+    $$w^h \leftarrow w^h+\tilde e^h$$
+\item<5> but $F^h$ is not linear!
+\end{itemize}
+
+\vspace{5mm}
+}
+\only<6>{
+\begin{itemize}
+\item \emph{goal}: use a coarser mesh $\Omega^H$ to somehow estimate the solution $u^h$ in the \emph{correction equation}
+    $${\color{FireBrick} F^h(u^h) - F^h(w^h) = r^h}$$
+\item basic multigrid idea: there are algorithms (\alert{smoothers}) which inefficiently improve $w^h$ \dots use them a little first \dots then correct from the coarser mesh
+\end{itemize}
+
+\vspace{24mm}
+}
+\only<7-8>{
+\begin{itemize}
+\item \emph{nodewise problem}: for $\psi_i^h$ a hat function or dof, solve for $c\in\RR$ to make the residual at that location zero:
 	$${\color{FireBrick} \phi_i(c) = r^h(w^h + c \psi_i^h)[\psi_i^h] = 0}$$
-\item sweeping through and solving nodewise problems is a \emph{smoother}
+\item sweeping through and solving nodewise problems is a smoother
     \begin{itemize}
     \item[$\circ$] Fourier analysis on linear PDEs shows smoothing property
-    \item[$\circ$] post-smoothing, $e^h$ and $r^h(w^h)$ have smaller high-frequencies
+    \item[$\circ$] post-smoothing, $e^h$ and $r^h$ have smaller high-frequencies
     \end{itemize}
-\item Brandt (1977): after smoothing, $F^h(u^h) - F^h(w^h) = r^h(w^h)$ should be accurately approximate-able on a coarser grid
+\item<8> after smoothing, the correction equation should be accurately approximate-able on the coarser mesh $\Omega^H$
 \end{itemize}
+
+\vspace{14mm}
 }
-\only<5>{
+\only<9-11>{
 \begin{itemize}
-\item \emph{goal}: use a coarser mesh to estimate the error in $\ast$
-\item \emph{full approximation storage} (FAS) equation:
-	$${\color{FireBrick} F^H(w^H) - F^H(\iR w^h) = R \, r^h(w^h)}$$
+\item \emph{goal}: use a coarser mesh $\Omega^H$ to somehow estimate the solution $u^h$ in $F^h(u^h) - F^h(w^h) = r^h(w^h)$
+\item Brandt's (1977) \emph{full approximation scheme} (FAS) equation:
+	$${\color{FireBrick} F^H(u^H) - F^H(\iR w^h) = R \, r^h(w^h)}$$
 
@@
     \item[$\circ$] $R:(\mathcal{V}^h)' \to (\mathcal{V}^H)'$ is \emph{canonical restriction}
-    \item[$\circ$] if $w^h=u^h$ exactly then $w^H = \iR w^h$ by well-posedness
+    \item<10-11>[$\circ$] note: if $w^h=u^h$ exactly then $u^H = \iR w^h$ since $F^H$ injective
     \end{itemize}
 
-\item rewritten: \quad ${\color{FireBrick} F^H(w^H) = \ell^H}$ where ${\color{FireBrick} \ell^H = F^H(\iR w^h) + R\, r^h(w^h)}$
+\item<11> rewritten FAS equation: let ${\color{FireBrick} \ell^H = F^H(\iR w^h) + R\, r^h(w^h)}$ then
+    $${\color{FireBrick} F^H(u^H) = \ell^H}$$
+\end{itemize}
 
-\phantom{x}
-
-\phantom{x}
-\end{itemize}
+\vspace{-3mm}
 }
@@
 
-\begin{frame}{FAS 2-grid solver}
+\begin{frame}{full approximation scheme (FAS) 2-mesh solver}
 
 \begin{align*}
-&\text{smooth by sweeps over grid:} & &w^h \leftarrow \left[\phi_i(c) = 0 \,\forall i\right] \\
+&\text{pre-smooth over fine:} & & [\text{smoother updates } w^h] \\
 &\text{restrict:}                   & &\ell^H = F^H(\iR w^h) + R\, r^h(w^h) \\
@@
 &\text{correct}:                    & &w^h \leftarrow w^h + P(w^H - \iR w^h) \\
-&\text{smooth by sweeps over grid:} & &w^h \leftarrow \left[\phi_i(c) = 0 \,\forall i\right]
+&\text{post-smooth over fine:} & & [\text{smoother updates } w^h]
 \end{align*}
@@
 \item $P: \mathcal{V}^H \to \mathcal{V}^h$ is \emph{prolongation}
-\item recall: \quad $\phi_i(c) = r^h(w^h + c \psi_i^h)[\psi_i^h]$
-\item restrict$+$(solve coarse)$+$correct \, $=$ \, \emph{coarse grid correction}
+\item restrict$+$(solve coarse)$+$correct \, $=$ \, \emph{FAS coarse grid correction}
 \end{itemize}
@@
 \hfill {\tiny Sequinot et al.~(2018)}
-
-%\vspace{-5mm}
-{\footnotesize
-\begin{itemize}
-\item<2> more ice sheet modeling at my Math.~Geosci.~Seminar tomorrow 2pm L5
-\end{itemize}
-}
 \end{frame}
@@
 
-{\color{FireBrick} FASCD = full approximation storage constraint decomposition}
+{\color{FireBrick} FASCD = full approximation scheme constraint decomposition}
 
@@
 
-\begin{frame}{full approximation storage constraint decomposition}
+\begin{frame}{full approximation scheme constraint decomposition}
 

### 2023-11-06 talk-dms/slides.tex
progress slides

@@
 
-\begin{frame}{nonlinear multigrid by FAS V-cycle or F-cycle}
+\begin{frame}{nonlinear multigrid by FAS: \only<1>{V-cycle}\only<2>{F-cycle}}
+
+\includegraphics[height=0.2\textheight]{../talk-oxford/images/mg-grids.png}
+
+\hspace{4mm} $j=3$ \hspace{13mm} $j=2$ \hspace{13mm} $j=1$ \hspace{12mm} $j=0$
 
 \bigskip
-
-\includegraphics[height=0.15\textheight]{../talk-oxford/images/mg-grids.png}
-
-\bigskip
-
 \only<1>{
-\hfill \includegraphics[height=0.15\textheight]{../talk-oxford/images/mg-vcycle.png}
-}
-\only<2>{\hfill \includegraphics[height=0.15\textheight]{../talk-oxford/images/mg-fcycle.png} \quad \begin{tabular}{l} F-cycle = \\ nested iteration \end{tabular}
-}
-
-\only<2>{\vspace{-3mm}}
-
+\begin{columns}
+\begin{column}{0.75\textwidth}
 {\small
@@
       $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j; w^j)$ \\
-      $w^{j-1} \gets \iR w^j$ \\
+      $w^{j-1} = \iR w^j$ \\
       $\ell^{j-1} = F^{j-1}(w^{j-1}) + R \left(\ell^j - F^j(w^j)\right)$ \\-
@@
 }
+\end{column}
+\begin{column}{0.25\textwidth}
+\includegraphics[width=0.7\textwidth]{../talk-oxford/images/mg-vcycle.png}
+\end{column}
+\end{columns}
+}
+\only<2>{
+\vspace{12mm}
+
+\hfill \input{tikz/fcycle.tex}
+
+\vspace{6mm}
+}
 \end{frame}
@@
 \begin{itemize}
-\item FAS multigrid works well on the right nonlinear PDE problem
-\item example: Liouville-Bratu equation\footnote{exact solution by Liouville (1853) makes a nice test case}
+\item FAS multigrid works \alert{very well} on the right nonlinear PDE problem
+\item example: Liouville-Bratu equation
     $$-\nabla^2 u - e^u = 0$$
 with Dirichlet boundary conditions on $\Omega=(0,1)^2$
-\item implement with minimal problem-specific code:
+\item discretize by straightforward finite differences
+\item minimal problem-specific code:
     \begin{enumerate}
@@
         \begin{itemize}
-        \item[$\circ$] e.g.~nonlinear Jacobi or Gauss-Seidel iteration
+        \item[$\circ$] nonlinear Gauss-Seidel iteration
         \end{itemize}
-    \item[3.] coarse solve can be same as smoother, or use Newton etc.
+    \item[3.] coarse solve can be same as smoother, or more sophisticated-but-not-optimal (e.g.~Newton iteration)
     \end{enumerate}
@@
 
-\begin{frame}{Bratu model problem: optimality}
+\begin{frame}{the meaning of ``fast solver''}
+
+\begin{itemize}
+\item what does ``very well'' on the previous slide mean?
+\end{itemize}
+
+\begin{block}{definition} a solver is \emph{optimal} if work in flops, and/or run-time, is $O(N)$ for $N$ unknowns
+\end{block}
+
+\begin{itemize}
+\item multigrid: optimality is a frequently-achieved goal for PDE problems with reasonably-smooth solutions
+\item<2> in fact, multigrid people get greedy
+\end{itemize}
+
+\begin{block}<2>{definition} a solver exhibits \emph{textbook multigrid efficiency} if it is optimal and it does work less than 10 times that of a single smoother sweep
+\end{block}
+\end{frame}
+
+\begin{frame}{Bratu model problem: optimality and TME}
 
@@
 \end{align*}
-\item<1-> up to $N\approx 10^8$ dofs
-    \begin{itemize}
-    \item[$\circ$] $J=11$ refinements
-    \item[$\circ$] laptop is memory-limited
-    \end{itemize}
-\item<2> compare $\approx 20\,\mu\,\text{s}/\text{N}$ for Poisson equation using Firedrake $P_1$ elements and geometric multigrid
+\item $N\approx 10^8$ unknowns at $J=11$ levels V-cycle
+\item compare $\approx 20\,\mu\,\text{s}/\text{N}$ for Poisson using Firedrake ($P_1$, geometric multigrid)
 \end{itemize}
@@
 \begin{column}{0.55\textwidth}
-\includegraphics<1>[width=\textwidth]{../talk-oxford/images/bratu-exps.png}
-
-\includegraphics<2>[width=\textwidth]{../talk-oxford/images/bratu-time.png}
+\includegraphics[width=\textwidth]{../talk-oxford/images/bratu-time.png}
 \end{column}
@@
     \end{enumerate}
-\end{itemize}
-\end{frame}
-
-\begin{frame}{problem: fluid layer in a climate}
-
-\begin{itemize}
-\item let's not get stuck on textbook example problems!
-\item multigrid for a real-world VI problem?
-\item consider an incompressible, viscous layer with surface elevation $s(x,y)$, flowing with velocity $\bu(x,y,z)$, driven by gravity, over fixed bed topography with elevation $b(x,y)$, in a \emph{climate} which adds or removes fluid at a signed rate $a(x,y)$ [$\text{m}\,\text{s}^{-1}$]
-    \begin{itemize}
-    \item[$\circ$] data $a,b$ defined on domain $\Omega \subset \RR^2$
-    \end{itemize}
-\item geophysical examples: \alert{glaciers and ice sheets}, sea ice, lakes
-\end{itemize}
-
-\bigskip
-\hfill \mbox{\includegraphics[height=0.25\textheight]{../talk-oxford/images/domain-data.png} \hspace{7mm} \includegraphics[height=0.25\textheight]{../talk-oxford/images/domain-velocity.png}}
-\end{frame}
-
-\begin{frame}{example: glacier ice coverage of the Alps in prior climates}
-
-\includegraphics[width=1.02\textwidth]{../talk-oxford/images/alps-seguinot2018.png}
-
-\vspace{-2mm}
-\hfill {\tiny Sequinot et al.~(2018)}
-\end{frame}
-
-\begin{frame}{naive strong form}
-
-\begin{itemize}
-\item naive strong form of the steady model: % ($\bn_s = \left<-s_x,-s_y,1\right>$ is surface normal):
-\begin{align*}
-s &\ge b                    & &\text{everywhere in } \Omega \\
--\bu|_s \cdot \bn_s &= a    & &\text{where } s(x,y) > b(x,y)
-\end{align*}
-
-    \begin{itemize}
-    \item[$\circ$] surface velocity $\bu|_s$ is determined by fluid domain geometry $s$
-    \item[$\circ$] $\bn_s=\left<-\grad s,1\right>$ is upward surface normal
-    \item[$\circ$] generally: $-\bu|_s \cdot \bn_s$ is a \emph{non-local} function of $s$
-    \end{itemize}
-\item the inequality constraint $s \ge b$ \alert{generates a free boundary}
-
-if an ablative climate $a < 0$ forces surface down to bed
-\end{itemize}
-
-\bigskip
-\hfill \mbox{\includegraphics[height=0.2\textheight]{../talk-oxford/images/domain-data.png} \hspace{7mm} \includegraphics[height=0.2\textheight]{../talk-oxford/images/domain-velocity.png}}
-\end{frame}
-
-\begin{frame}{how to evaluate $\Phi(s) = - \bu|_s \cdot \bn_s$ for glacier ice?}
-
-\begin{itemize}
-\item \alert{lubrication approximation}\footnote{also known as the \emph{shallow ice approximation}} \alert{model}
-
-apply a nonlinear elliptic differential operator to $s$:
-    $$\Phi(s) = - \frac{\gamma}{\qq} (s-b)^{\qq} |\grad s|^{\qq} - \grad \cdot\left(\frac{\gamma}{\qq+1} (s-b)^{\qq+1} |\grad s|^{\qq-2} \grad s\right)$$
-    \begin{itemize}
-    \item $\qq = 4$
-    \item $\grad$ is in $x,y$ only
-    \item $\Phi(s)$ is a nonlinear \alert{differential operator} in this model because membrane stresses are \emph{not} balanced
-    \item $\Phi(s)$ is doubly-degenerate
-    \end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}{VI for fluid layer in a climate}
-
-\begin{itemize}
-\item admissible surface elevations:
-    $$\mathcal{K} = \left\{r \in \mathcal{V} \,:\, r \ge b\right\}$$
-
-    \begin{itemize}
-    \item[$\circ$] $\mathcal{V}$ to be determined by viscous fluid model\footnote{in shallow ice approximation, $(s-b)^{8/3} \in W^{1,4}(\Omega)$ (Jouvet \& Bueler, 2012)}
-    \end{itemize}
-\item VI problem for surface elevation $s\in\mathcal{K}$:
-	$$\ip{\Phi(s)}{r-s} \ge \ip{a}{r-s} \quad \text{ for all } r \in \mathcal{K}$$
-where
-    $$\Phi(s)=- \bu|_s \cdot \bn_s,$$
-with extension by 0 to all of $\Omega$, and $\bu$ is the velocity solution on
-    $$\Lambda(s) = \{(x,y,z) : b(x,y) < z < s(x,y)\}$$
 \end{itemize}
@@
 
-\begin{frame}{2D shallow ice approximation results (\emph{very fresh})}
+\begin{frame}{mesh independence}
+
+\begin{itemize}
+\item FIXME
+\end{itemize}
+\end{frame}
+
+\begin{frame}{problem: fluid layer in a climate}
+
+\begin{itemize}
+\item consider an incompressible, viscous layer with surface elevation $s(x,y)$, flowing with velocity $\bu(x,y,z)$, driven by gravity, over fixed bed topography with elevation $b(x,y)$, in a \emph{climate} which adds or removes fluid at a signed rate $a(x,y)$ [$\text{m}\,\text{s}^{-1}$]
+    \begin{itemize}
+    \item[$\circ$] data $a,b$ defined on domain $\Omega \subset \RR^2$
+    \end{itemize}
+\item geophysical examples: \alert{glaciers and ice sheets}, sea ice, lakes
+\end{itemize}
+
+\bigskip
+\hfill \mbox{\includegraphics[height=0.25\textheight]{../talk-oxford/images/domain-data.png} \hspace{7mm} \includegraphics[height=0.25\textheight]{../talk-oxford/images/domain-velocity.png}}
+\end{frame}
+
+\begin{frame}{example: glacier ice coverage of the Alps in prior climates}
+
+\includegraphics[width=1.02\textwidth]{../talk-oxford/images/alps-seguinot2018.png}
+
+\vspace{-2mm}
+\hfill {\tiny Sequinot et al.~(2018)}
+\end{frame}
+
+\begin{frame}{naive strong form}
+
+\begin{itemize}
+\item naive strong form of the steady model: % ($\bn_s = \left<-s_x,-s_y,1\right>$ is surface normal):
+\begin{align*}
+s &\ge b                    & &\text{everywhere in } \Omega \\
+-\bu|_s \cdot \bn_s &= a    & &\text{where } s(x,y) > b(x,y)
+\end{align*}
+
+    \begin{itemize}
+    \item[$\circ$] surface velocity $\bu|_s$ is determined by fluid domain geometry $s$
+    \item[$\circ$] $\bn_s=\left<-\grad s,1\right>$ is upward surface normal
+    \item[$\circ$] generally: $-\bu|_s \cdot \bn_s$ is a \emph{non-local} function of $s$
+    \end{itemize}
+\item the inequality constraint $s \ge b$ \alert{generates a free boundary}
+
+if an ablative climate $a < 0$ forces surface down to bed
+\end{itemize}
+
+\bigskip
+\hfill \mbox{\includegraphics[height=0.2\textheight]{../talk-oxford/images/domain-data.png} \hspace{7mm} \includegraphics[height=0.2\textheight]{../talk-oxford/images/domain-velocity.png}}
+\end{frame}
+
+\begin{frame}{shallow ice approximation}
+
+\begin{itemize}
+\item with enough simplifications we may compute the surface motion term by applying a nonlinear elliptic differential operator to $s$:
+    $$\Phi(s) = - \frac{\gamma}{\qq} (s-b)^{\qq} |\grad s|^{\qq} - \grad \cdot\left(\frac{\gamma}{\qq+1} (s-b)^{\qq+1} |\grad s|^{\qq-2} \grad s\right)$$
+    \begin{itemize}
+    \item $\qq = 4$
+    \item $\grad$ is in $x,y$ only
+    \item $\Phi(s)$ is a nonlinear \alert{differential operator} in this model because membrane stresses are \emph{not} balanced
+    \item $\Phi(s)$ is doubly-degenerate
+    \end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{VI for fluid layer in a climate}
+
+\begin{itemize}
+\item admissible surface elevations:
+    $$\mathcal{K} = \left\{r \in \mathcal{V} \,:\, r \ge b\right\}$$
+
+    \begin{itemize}
+    \item[$\circ$] $\mathcal{V}$ to be determined by viscous fluid model\footnote{in shallow ice approximation, $(s-b)^{8/3} \in W^{1,4}(\Omega)$ (Jouvet \& Bueler, 2012)}
+    \end{itemize}
+\item VI problem for surface elevation $s\in\mathcal{K}$:
+	$$\ip{\Phi(s)}{r-s} \ge \ip{a}{r-s} \quad \text{ for all } r \in \mathcal{K}$$
+where
+    $$\Phi(s)=- \bu|_s \cdot \bn_s,$$
+with extension by 0 to all of $\Omega$, and $\bu$ is the velocity solution on
+    $$\Lambda(s) = \{(x,y,z) : b(x,y) < z < s(x,y)\}$$
+\end{itemize}
+\end{frame}
+
+\begin{frame}{2D shallow ice approximation}
 
@@
     \end{itemize}
-\end{itemize}
-
-\medskip
-\begin{center}
-\includegraphics[width=0.8\textwidth]{../talk-oxford/images/fascd-firedrake-dome.png}
-\end{center}
-\end{frame}
-
-\begin{frame}{mesh independence}
-
-\begin{itemize}
-\item FIXME
 \end{itemize}

### 2023-11-06 talk-dms/slides.tex
sketch out remainder of talk

@@
 
-\begin{frame}{nonlinear multigrid by FAS: \only<1>{V-cycle}\only<2>{F-cycle}}
+\begin{frame}{nonlinear multigrid by FAS: \only<1>{V-cycle}\only<2>{``FMG'' cycle}}
 
@@
 \end{align*}
-\item $N\approx 10^8$ unknowns at $J=11$ levels V-cycle
+\item $N\approx 10^8$ unknowns for $12$-level V-cycle
 \item compare $\approx 20\,\mu\,\text{s}/\text{N}$ for Poisson using Firedrake ($P_1$, geometric multigrid)
@@
 \item VIs are nonlinear problems, even for linear operators like $-\nabla^2$
-\item Newton-multigrid is straightforward in PETSc:
-\begin{Verbatim}[xleftmargin=13mm,fontsize=\scriptsize]
-./obstacle -da_grid_x 3 -da_grid_y 3 \
-    -snes_type vinewtonrsls -ksp_type cg -pc_type mg \
-    -da_refine J
-\end{Verbatim}
+\item most obvious method (?): Newton-multigrid
     \begin{itemize}
@@
         \begin{itemize}
-        \item[] \texttt{rsls} = reduced space line search
+        \item[] reduced space line search
         \end{itemize}
-    \item[$\circ$] Newton step equations solved by CG with GMG V-cycles
+    \item[$\circ$] Newton step equations solved by CG with linear multigrid V-cycles
     \end{itemize}
@@
     \begin{itemize}
-    \item[$\circ$] grid-dependent (growing) Newton iterations
+    \item[$\circ$] consequence: grid-dependent (growing) Newton iterations
     \end{itemize}
@@
 \begin{itemize}
-\item applying nested iteration (nonlinear F-cycle) resolves this:
-
-\vspace{2mm}
-\begin{Verbatim}[xleftmargin=13mm,fontsize=\scriptsize]
-./obstacle -da_grid_x 3 -da_grid_y 3 \
-    -snes_type vinewtonrsls -ksp_type cg -pc_type mg \
-\end{Verbatim}
-
-\vspace{-1.5mm}
-\hspace{18.7mm} {\scriptsize \color{FireBrick} \texttt{-snes\_grid\_sequence J}}
-
-\vspace{2mm}
+\item applying nested iteration helps this:
     \begin{itemize}
+    \item[$\circ$] this is effectively an FMG-cycle, an alternative to FASCD
     \item[$\circ$] grid-independent Newton iterations
-    \item[$\circ$] optimal $O(N^1)$ flops and time
-    \item<2>[$\circ$] Chapter 12 example in my new book
+    \item[$\circ$] optimal $O(N)$ flops and time; Chapter 12 example in my new book
     \end{itemize}
-
-\vspace{-12mm}
-\hfill \includegraphics[width=0.25\textwidth]{../talk-oxford/images/mg-fcycle.png}
 \end{itemize}
@@
 \vspace{-22mm}
-\hspace{40mm} \includegraphics[height=0.45\textheight]{../talk-oxford/images/obstacle-flops-per-n.png} \qquad \includegraphics<2>[width=20mm]{../talk-oxford/images/frontcover.jpg}
+\hspace{40mm} \includegraphics[width=20mm]{../talk-oxford/images/frontcover.jpg}
 \end{frame}
@@
 {\small
-\begin{tabular}{l|ccccc}
+\begin{tabular}{r|cccc}
    & \stacktwo{admissible}{iterates} & \stacktwo{mesh-indep.}{rates}
-       & \stacktwo{no global}{linearization}
-         & \stacktwo{PETSc or}{Firedrake}  \\ \hline
-RS NM                      & $\checkmark$ &  & & $\checkmark$ \\
-\quad $+$ NI & $\checkmark$ & $\checkmark$ & & $\checkmark$ \\
-SS NM             &              & $\checkmark$ & & $\checkmark$ \\
-{\color{FireBrick} FASCD}  & $\checkmark$ & {\color{FireBrick} \textbf{?}} & $\checkmark$ & \only<2>{{\Large {\color{FireBrick} $\checkmark$}}} {\Large \strut}
+       & \stacktwo{no need for}{global linearization}  \\ \hline
+NM                         & $\checkmark$ &              & \\
+\quad $+$ NI               & $\checkmark$ & $\checkmark$ & \\
+{\color{FireBrick} FASCD}  & $\checkmark$ & $\checkmark$ & $\checkmark$ {\Large \strut}
 \end{tabular}
@@
 \begin{center}
-RS = reduced space, SS = semi-smooth, NM = Newton-multigrid, NI = nested iteration
+NM = Newton-multigrid, NI = nested iteration
 \end{center}
@@
 \begin{itemize}
-\item for the non-local fluid layer VI problem we need all 4 checked
-\item we are trying-out a new algorithm,
+\item new algorithm,
 
 {\color{FireBrick} FASCD = full approximation scheme constraint decomposition}
-
-    \begin{itemize}
-    \item[$\circ$] Firedrake implementation \only<2>{\dots as of yesterday}
-    \end{itemize}
 \end{itemize}
@@
 where $\mathcal{K}^i \subset \mathcal{V}^i$, with projections $\Pi_i : \mathcal{K} \to \mathcal{K}_i$
-\item CD additive and multiplicative iterations exist for $VI(F,\ell,\mathcal{K})$
+\item CD multiplicative iterations for $VI(F,\ell,\mathcal{K})$
 
 \begin{pseudo}
-\pr{cd-add}(u)\text{:} \\+
-    for $i \in \{0,\dots,m-1\}$: \\+
-        \rm{find} $\hat w_i\in \cK_i$ \rm{so that for all} $v_i\in \cK_i$, \\+
-            $\boxed{\ip{F(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i}}$ \\--
-    $\hat w = \sum_i \hat w_i\in\cK$ \\
-    return $w=(1-\alpha) u + \alpha \hat w$
+\pr{cd-mult}
 \end{pseudo}
@@
 \begin{pseudo}
-\pr{fascd-vcycle}(\ell^J,\psi^J; w^J)\text{:} \\+
-    $\chi^J = \psi^J - w^J {\large \strut}$ \\
-    for $j=J$ downto $j=1$ \\+
-      $\chi^{j-1} = \maxR \chi^j {\large \strut}$ \\
-      $\phi^j = \chi^j - P\chi^{j-1} {\large \strut}$ \\
-      $y^j = 0$ \\
-      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\phi^j,w^j;y^j)$ \hspace{8mm} \ct{smoothing in $\mathcal{D}^j$}\\
-      $w^{j-1} = \iR(w^j + y^j)$ \\
-      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j+y^j)\right)$ \\-
-    $z^0 = 0$ \\
-    $\text{\pr{solve}}(\ell^0,\chi^0,w^0;z^0)$ \hspace{10mm} \ct{coarse solve in $\mathcal{U}^0$} \\
-    for $j=1$ to $j=J$ \\+
-      $z^j = y^{j} + P z^{j-1}$ \\
-      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\chi^j,w^j;z^j)$ \hspace{8mm} \ct{smoothing in $\mathcal{U}^j$} \\-
-    $w^J \gets w^J+z^J$
+FIXME
 \end{pseudo}
@@
 
-\begin{frame}{mesh independence}
+\begin{frame}{classical obstacle problem}
+
+\only<1-2>{\includegraphics[width=0.4\textwidth]{../talk-oxford/images/obstacle65.pdf}}
+\hfill
+\only<1-2>{\includegraphics[width=0.25\textwidth]{../paper/fixfigs/ball-set.png} \hspace{1mm}}
+\only<3>{\includegraphics[width=0.25\textwidth]{../paper/fixfigs/spiral-set.png} \hspace{2mm}}
 
 \begin{itemize}
-\item FIXME
+\item<1> plot $m_J$ versus V-cycle iterations for ball from Table 2 and fit $\log(m_J)$
+\item<2> add $m_J$ versus F-cycle iterations for ball from Table 2
+\item<3> plot $m_J$ versus V-cycle, F-cycle iterations for spiral from Table 2 and fit $\log(m_J)$
+\end{itemize}
+\end{frame}
+
+\begin{frame}{advection-diffusion of a concentration}
+
+\mbox{
+\includegraphics[width=0.3\textwidth]{../paper/fixfigs/poll2d-zero-set.png}
+\includegraphics[width=0.3\textwidth]{../paper/fixfigs/poll2d-one-set.png}
+}
+
+\begin{itemize}
+\item 2D problem: $-\eps \grad^2 u + \bm{X}\cdot \grad u = \phi$ with $0\le u\le 1$
+\item plot $m_J$ versus FMG-cycle iterations from Table 3
 \end{itemize}
@@
 
-\begin{frame}{2D shallow ice approximation}
+\begin{frame}{weak-scaling results: shallow ice approximation}
+
+\hfill \includegraphics[width=0.8\textwidth]{../paper/fixfigs/sialev8scene.png}
 
 \begin{itemize}
-\item preliminary results
-\item dome test case in lubrication approximation
-    \begin{itemize}
-    \item[$\circ$] here $\Phi(s)$ \emph{is} a differential operator
-    \item[$\circ$] note $s^{8/3} \in W^{1,4}(\Omega)$ but not in $C^2$
-    \end{itemize}
-\item FASCD algorithm result
-    \begin{itemize}
-    \item[$\circ$] Firedrake $P_1$ elements
-    \item[$\circ$] strong smoother (\texttt{vinewtonrsls})
-    \end{itemize}
+\item bar plot $P$ on $x$-axis and time on $y$-axis; label bars with $m_J$ and FMG iterations
 \end{itemize}

### 2023-11-09 talk-dms/slides.tex
basically the figure sequence ...

@@
 \hfill
-\only<1-2>{\includegraphics[width=0.25\textwidth]{../paper/fixfigs/ball-set.png} \hspace{1mm}}
-\only<3>{\includegraphics[width=0.25\textwidth]{../paper/fixfigs/spiral-set.png} \hspace{2mm}}
+\only<1-2>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/ball-set.png} \hspace{1mm}}
+\only<3>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/spiral-set.png} \hspace{2mm}}
 
-\begin{itemize}
-\item<1> plot $m_J$ versus V-cycle iterations for ball from Table 2 and fit $\log(m_J)$
-\item<2> add $m_J$ versus F-cycle iterations for ball from Table 2
-\item<3> plot $m_J$ versus V-cycle, F-cycle iterations for spiral from Table 2 and fit $\log(m_J)$
-\end{itemize}
+\bigskip
+
+\begin{center}
+%\only<1>{\includegraphics[width=0.25\textwidth]{figs/balliterslog.png}}
+\only<1>{\includegraphics[width=0.5\textwidth]{figs/balliters.png}}
+\only<2>{\includegraphics[width=0.5\textwidth]{figs/balliters.png}}
+\only<3>{\includegraphics[width=0.5\textwidth]{figs/bothiters.png}}
+\end{center}
+
+%\item<1> plot $m_J$ versus V-cycle iterations for ball from Table 2 and fit $\log(m_J)$
+%\item<2> add $m_J$ versus F-cycle iterations for ball from Table 2
+%\item<3> plot $m_J$ versus V-cycle, F-cycle iterations for spiral from Table 2 and fit $\log(m_J)$
 \end{frame}

### 2023-11-09 talk-dms/py/classical.py

@@ -1,0 +1,51 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+_SHOW = False
+def writeout(outname):
+    if _SHOW:
+        plt.show()
+    else:
+        print('writing file ' + outname)
+        plt.savefig(outname, bbox_inches='tight')
+
+# data from Table 2 in paper
+mJ = [41, 145, 545, 2.1e3, 8.3e3, 3.3e4, 1.3e5, 5.3e5, 2.1e6, 8.4e6, 3.4e7]
+ballV = [1,2,3,7,9,11,11,12,13,13,14]
+ballF = [1,2,3,3,4,5,4,4,3,2,2]
+spirV = [1,4,5,7,9,10,11,12,13,14,18]
+spirF = [1,2,3,4,4,5,5,5,6,6,6]
+
+# ball iterations: V only, with fit
+plt.semilogx(mJ,ballV,'o',label='V-cycles (ball problem)')
+p = np.polyfit(np.log(mJ),ballV,1)
+print(p)
+fitstr = r'$O(log(m_J))$'
+line, = plt.semilogx(mJ,p[0]*np.log(mJ)+p[1],'b--',label=fitstr)
+plt.xlabel(r'$m_J$  (number of unknowns)')
+plt.legend(loc='upper left', fontsize=12.0)
+plt.xlim([10.0,1.0e8])
+plt.ylim([0.0,19.0])
+plt.gca().set_yticks(range(2,20,2))
+writeout('balliterslog.png')
+line.remove()
+
+# ball iterations: both V and F
+#plt.semilogx(mJ,ballV,'o',label='V-cycles (ball problem)')
+plt.semilogx(mJ,ballF,'+',ms=14.0,label='F-cycles (ball problem)')
+plt.xlabel(r'$m_J$  (number of unknowns)')
+plt.legend(loc='upper left', fontsize=12.0)
+plt.xlim([10.0,1.0e8])
+plt.ylim([0.0,19.0])
+plt.gca().set_yticks(range(2,20,2))
+writeout('balliters.png')
+
+# add spiral iterations
+plt.semilogx(mJ,spirV,'o',label='V-cycles (spiral problem)')
+plt.semilogx(mJ,spirF,'+',ms=14.0,label='F-cycles (spiral problem)')
+plt.xlabel(r'$m_J$  (number of unknowns)')
+plt.legend(loc='upper left', fontsize=12.0)
[…]

### 2023-11-09 talk-dms/slides.tex
plot a bunch of results

--- 2023-11-06 talk-dms/slides.tex
@@
 
-\only<1-2>{\includegraphics[width=0.4\textwidth]{../talk-oxford/images/obstacle65.pdf}}
-\hfill
-\only<1-2>{\includegraphics[width=0.25\textwidth]{../paper/fixfigs/ball-set.png} \hspace{1mm}}
-\only<3>{\includegraphics[width=0.25\textwidth]{../paper/fixfigs/spiral-set.png} \hspace{2mm}}
-
-\begin{itemize}
-\item<1> plot $m_J$ versus V-cycle iterations for ball from Table 2 and fit $\log(m_J)$
-\item<2> add $m_J$ versus F-cycle iterations for ball from Table 2
-\item<3> plot $m_J$ versus V-cycle, F-cycle iterations for spiral from Table 2 and fit $\log(m_J)$
-\end{itemize}
+\only<1>{\includegraphics[width=0.75\textwidth]{figs/ballitersV.png}}\only<2>{\includegraphics[width=0.75\textwidth]{figs/ballitersVlog.png}}\only<3>{\includegraphics[width=0.75\textwidth]{figs/ballitersVF.png}}\only<4>{\includegraphics[width=0.75\textwidth]{figs/bothitersVF.png}} \hfill \only<1-3>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/ball-set.png}}\only<4>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/spiral-set.png}}
 \end{frame}
@@
 \item 2D problem: $-\eps \grad^2 u + \bm{X}\cdot \grad u = \phi$ with $0\le u\le 1$
-\item plot $m_J$ versus FMG-cycle iterations from Table 3
 \end{itemize}
+\end{frame}
+
+\begin{frame}{advection-diffusion of a concentration}
+
+\includegraphics[width=0.75\textwidth]{figs/advdiff.png} \hfill \begin{minipage}[t]{0.2\textwidth}
+\vspace{-50mm}
+
+\includegraphics[width=\textwidth]{../paper/fixfigs/poll2d-zero-set.png}
+
+\includegraphics[width=\textwidth]{../paper/fixfigs/poll2d-one-set.png}
+\end{minipage}
 \end{frame}
@@
 
-\begin{frame}{weak-scaling results: shallow ice approximation}
+\begin{frame}{computed ice sheet}
 
-\hfill \includegraphics[width=0.8\textwidth]{../paper/fixfigs/sialev8scene.png}
+\centering
+\includegraphics[width=\textwidth]{../paper/fixfigs/sialev8scene.png}
 
-\begin{itemize}
-\item bar plot $P$ on $x$-axis and time on $y$-axis; label bars with $m_J$ and FMG iterations
-\end{itemize}
+FIXME ice sheet the size of Greenland
+\end{frame}
+
+\begin{frame}{weak-scaling}
+
+\centering
+\includegraphics[width=0.7\textwidth]{figs/siaweaktime.png}
+
+FIXME final run had $m_J=20481^2=4.1\times 10^8$ unknowns and a 88 meter grid resolution
 \end{frame}

### 2023-11-09 talk-dms/py/advdiff.py

@@ -1,0 +1,27 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+_SHOW = False
+def writeout(outname):
+    if _SHOW:
+        plt.show()
+    else:
+        print('writing file ' + outname)
+        plt.savefig(outname, bbox_inches='tight')
+
+# data from Table 4 in paper
+n = 15 * 2**np.arange(0,9) + 1
+mJ2D = n**2
+mJ3D = n[:5]**3
+F2D = [1,2,2,2,2,2,3,3,3]
+F3D = [1,2,2,2,3]
+
+plt.clf()
+plt.semilogx(mJ2D,F2D,'+',ms=12.0,mew=2.0,label='FMG cycles (2D)')
+plt.semilogx(mJ3D,F3D,'+',ms=14.0,mew=2.0,label='FMG cycles (3D)')
+plt.xlabel(r'$m_J$  (number of unknowns)')
+plt.legend(loc='upper left', fontsize=12.0)
+plt.xlim([10.0,1.0e8])
+plt.ylim([0.0,4.0])
+plt.gca().set_yticks(range(1,5))
+writeout('advdiff.png')

### 2023-11-09 talk-dms/py/classical.py

@@ -16,36 +16,22 @@ ballF = [1,2,3,3,4,5,4,4,3,2,2]
 spirV = [1,4,5,7,9,10,11,12,13,14,18]
 spirF = [1,2,3,4,4,5,5,5,6,6,6]
 
-# ball iterations: V only, with fit
-plt.semilogx(mJ,ballV,'o',label='V-cycles (ball problem)')
-p = np.polyfit(np.log(mJ),ballV,1)
-print(p)
-fitstr = r'$O(log(m_J))$'
-line, = plt.semilogx(mJ,p[0]*np.log(mJ)+p[1],'b--',label=fitstr)
-plt.xlabel(r'$m_J$  (number of unknowns)')
-plt.legend(loc='upper left', fontsize=12.0)
-plt.xlim([10.0,1.0e8])
-plt.ylim([0.0,19.0])
-plt.gca().set_yticks(range(2,20,2))
-writeout('balliterslog.png')
-line.remove()
-
-# ball iterations: both V and F
-#plt.semilogx(mJ,ballV,'o',label='V-cycles (ball problem)')
-plt.semilogx(mJ,ballF,'+',ms=14.0,label='F-cycles (ball problem)')
-plt.xlabel(r'$m_J$  (number of unknowns)')
-plt.legend(loc='upper left', fontsize=12.0)
-plt.xlim([10.0,1.0e8])
-plt.ylim([0.0,19.0])
-plt.gca().set_yticks(range(2,20,2))
-writeout('balliters.png')
-
-# add spiral iterations
-plt.semilogx(mJ,spirV,'o',label='V-cycles (spiral problem)')
-plt.semilogx(mJ,spirF,'+',ms=14.0,label='F-cycles (spiral problem)')
-plt.xlabel(r'$m_J$  (number of unknowns)')
-plt.legend(loc='upper left', fontsize=12.0)
-plt.xlim([10.0,1.0e8])
-plt.ylim([0.0,19.0])
-plt.gca().set_yticks(range(2,20,2))
-writeout('bothiters.png')
+names = ['ballitersV', 'ballitersVlog', 'ballitersVF', 'bothitersVF']
+for j in range(4):
+    plt.clf()
+    plt.semilogx(mJ,ballV,'o',label='V-cycles (ball problem)')
+    if j == 1:
+        p = np.polyfit(np.log(mJ),ballV,1)
+        fitstr = r'$O(log(m_J))$'
[…]

### 2023-11-09 talk-dms/py/siaweak.py

@@ -1,0 +1,27 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+_SHOW = False
+def writeout(outname):
+    if _SHOW:
+        plt.show()
+    else:
+        print('writing file ' + outname)
+        plt.savefig(outname, bbox_inches='tight')
+
+# data from Table 5 in paper
+P = ['1','4','16','64','256','1024']
+#iters = [3,3,3,3,2,2]
+time = [98.0,98.0,124.0,136.0,95.0,177.0]
+plt.clf()
+#plt.bar(P,iters)
+#plt.ylabel('FMG cycles')
+#plt.ylim([0.0,4.0])
+#plt.gca().set_yticks(range(1,5))
+#writeout('siaweakiters.png')
+plt.bar(P,time)
+plt.xlabel('processors')
+plt.ylabel('run time (s)')
+plt.ylim([0.0,200.0])
+plt.gca().set_yticks([50.0,100.0,150.0,200.0])
+writeout('siaweaktime.png')

### 2023-11-09 talk-dms/slides.tex
progress

@@
 
-\only<1>{
+\only<1-2>{
 \begin{itemize}
@@
 \item shown above: \quad $\Omega$ a square, $\psi(x)$ a hemisphere
+\item<2>[\alert{Q.}] how to solve this as a PDE with boundary conditions?
 \end{itemize}
 
-\phantom{x}
+\vspace{2mm}
 }
-\only<2>{
+\only<3>{
 \begin{itemize}
-\item this is constrained optimization over an \emph{admissible set}
+\item this is constrained optimization over an infinite-dimensional \emph{admissible set}
 	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \,\text{ and }\, v \ge \psi\right\}$$
@@
     \begin{itemize}
-    \item[$\circ$] $\mathcal{K}$ is a closed and convex subset of the Sobolev space $H^1(\Omega) = \left\{v \,:\, \int_\Omega |v|^2 + |\grad v|^2 < \infty\right\}$
+    \item[$\circ$] $\mathcal{K}$ is a closed and convex subset of the Sobolev space
+   $$H^1(\Omega) = \left\{v \,:\, \int_\Omega |v|^2 + |\grad v|^2 < \infty\right\}$$
     \end{itemize}
 \end{itemize}
-
-\vspace{10.5mm}
 }
@@
    \item[$\circ$] Poisson equation $-\grad^2 u = f$ is ``$J'(u)=0$'' on $R_u$
-   \item<3>[$\circ$] using the solution $u$ to define the set $R_u$ on which to solve the PDE $-\grad^2 u = f$ \alert{is not acceptable}
+   \item<3>[$\circ$] using the solution $u$ to define the set $R_u$ on which to solve the PDE $-\grad^2 u = f$ \alert{does not lead to solution algorithms}
    \end{itemize}
@@
 \begin{itemize}
-\item the correct weak form is a {\color{FireBrick} \emph{variational inequality} (VI)}
-\item namely, $J'(u)$ points directly into $\mathcal{K}$:
+\item the weak form is a {\color{FireBrick} \emph{variational inequality} (VI)}, which says that $J'(u)$ points directly into $\mathcal{K}$:
     $$\ip{J'(u)}{v-u} = \int_\Omega \grad u\cdot \grad (v-u) - f (v-u) \ge 0$$
@@
 
-\phantom{x}
-
-\phantom{x}
+\vspace{11mm}
 \end{itemize}
@@
     \end{itemize}
-\item the general problem {\color{FireBrick} VI($F$,$\mathcal{K}$)} is
+\item the general variational inequality {\color{FireBrick} VI($F$,$\mathcal{K}$)} is
 	$${\color{FireBrick} \ip{F(u)}{v-u} \ge 0 \quad \text{ for all } v \in \mathcal{K}}$$
-\item<2> when $\mathcal{K}$ is nontrivial the problem {\color{FireBrick} VI($F$,$\mathcal{K}$)} is nonlinear \emph{even when $F$ is a linear operator}
+\item when $\mathcal{K}$ is nontrivial the problem {\color{FireBrick} VI($F$,$\mathcal{K}$)} is nonlinear \emph{even when $F$ is a linear operator}
 \end{itemize}
@@
 \vspace{3mm}
-\item first-semester calculus: \quad minimize $f(x)$ over $\mathcal{K} = [a,b]$
+\item first-semester calculus:
+    $$u \gets \min_{x\in[a,b]} f(x) \quad \iff \quad f'(u)(v-u) \ge 0 \quad \forall v \in[a,b]$$
 \end{itemize}
@@
 
-\section{multigrid for PDEs}
+\section{nonlinear multigrid for PDEs}
 
-\subsection{\emph{only} nonlinear multigrid!}
 \subsection{full approximation scheme (FAS)}
@@
 	\begin{itemize}
-	\item[$\circ$] $\mathcal{V}=H^1(\Omega)$, $\ell\in \mathcal{V}'$
+	\item[$\circ$] $u\in \mathcal{V}=H^1(\Omega)$
+	\item[$\circ$] $\ell\in \mathcal{V}'$
 	\item[$\circ$] $F : \mathcal{V} \to \mathcal{V}'$ continuous and injective
-	\item[$\circ$] for example, the Liouville-Bratu problem: $-\grad^2 u - e^u = f(x)$
+	\item[$\circ$] for example, the Liouville-Bratu problem: $-\grad^2 u - e^u = f$
 	\end{itemize}
@@
 
-\vspace{7mm}
+\vspace{4mm}
 }
@@
 \begin{itemize}
-\item \emph{goal}: from iterate $w^h$, to solve
+\item \emph{updated goal}: from iterate $w^h$, to solve
     $$F^h(u^h) - F^h(w^h) = r^h$$
@@
 \begin{itemize}
-\item \emph{goal}: use a coarser mesh $\Omega^H$ to somehow estimate the solution $u^h$ in the \emph{correction equation}
+\item \emph{updated goal}: use a coarser mesh $\Omega^H$ to somehow estimate the solution $u^h$ in the nonlinear \emph{correction equation}
     $${\color{FireBrick} F^h(u^h) - F^h(w^h) = r^h}$$
-\item basic multigrid idea: there are algorithms (\alert{smoothers}) which inefficiently improve $w^h$ \dots use them a little first \dots then correct from the coarser mesh
+\item basic multigrid idea: there are algorithms (\alert{smoothers}) which ``improve'' $w^h$ \dots use them a little first \dots then correct from the coarser mesh
+    \begin{itemize}
+    \item[$\circ$] ``improve'' means they remove high-frequency error components efficiently
+    \end{itemize}
 \end{itemize}
 
-\vspace{24mm}
+\vspace{15mm}
 }
-\only<7-8>{
+\only<7>{
 \begin{itemize}
@@
 	$${\color{FireBrick} \phi_i(c) = r^h(w^h + c \psi_i^h)[\psi_i^h] = 0}$$
-\item sweeping through and solving nodewise problems is a smoother
+\item sweeping through and solving nodewise problems is a \alert{smoother}
     \begin{itemize}
-    \item[$\circ$] Fourier analysis on linear PDEs shows smoothing property
-    \item[$\circ$] post-smoothing, $e^h$ and $r^h$ have smaller high-frequencies
+    \item[$\circ$] Fourier analysis shows smoothing property
+    \item[$\circ$] after smoothing, $e^h$ and $r^h$ have smaller high-frequencies
     \end{itemize}
-\item<8> after smoothing, the correction equation should be accurately approximate-able on the coarser mesh $\Omega^H$
+\item after smoothing, the correction equation on $\Omega^h$ should be accurately approximate-able on the coarser mesh $\Omega^H$
 \end{itemize}
@@
 }
-\only<9-11>{
+\only<8-9>{
 \begin{itemize}
-\item \emph{goal}: use a coarser mesh $\Omega^H$ to somehow estimate the solution $u^h$ in $F^h(u^h) - F^h(w^h) = r^h(w^h)$
+\item \emph{updated goal}: use a coarser mesh $\Omega^H$ to somehow estimate the solution $u^h$ in $F^h(u^h) - F^h(w^h) = r^h(w^h)$
 \item Brandt's (1977) \emph{full approximation scheme} (FAS) equation:
@@
     \begin{itemize}
-    \item[$\circ$] $\iR:\mathcal{V}^h \to \mathcal{V}^H$ is \emph{injection}
+    \item[$\circ$] $\iR:\mathcal{V}^h \to \mathcal{V}^H$ is node-wise \emph{injection}
     \item[$\circ$] $R:(\mathcal{V}^h)' \to (\mathcal{V}^H)'$ is \emph{canonical restriction}
-    \item<10-11>[$\circ$] note: if $w^h=u^h$ exactly then $u^H = \iR w^h$ since $F^H$ injective
+    \item[$\circ$] note: if $w^h=u^h$ exactly then $u^H = \iR w^h$ since $F^H$ injective
     \end{itemize}
 
-\item<11> rewritten FAS equation: let ${\color{FireBrick} \ell^H = F^H(\iR w^h) + R\, r^h(w^h)}$ then
+\item<9> rewritten FAS equation: let ${\color{FireBrick} \ell^H = F^H(\iR w^h) + R\, r^h(w^h)}$ then
     $${\color{FireBrick} F^H(u^H) = \ell^H}$$
@@
 
+\begin{center}
+fine mesh $=\Omega^h$\, \includegraphics[height=0.14\textheight]{../talk-oxford/images/fine-grid.png} \hspace{15mm} \includegraphics[height=0.14\textheight]{../talk-oxford/images/coarse-grid.png} \,$\Omega^H=$ coarse mesh
+\end{center}
+
 \begin{align*}
 &\text{pre-smooth over fine:} & & [\text{smoother updates } w^h] \\
-&\text{restrict:}                   & &\ell^H = F^H(\iR w^h) + R\, r^h(w^h) \\
-&\text{solve coarse:}                      & &F^H(w^H) = \ell^H \\
-&\text{correct}:                    & &w^h \leftarrow w^h + P(w^H - \iR w^h) \\
+&\text{\textbf{restrict}:}                   & &\ell^H = F^H(\iR w^h) + R\, r^h(w^h) \\
+&\text{\textbf{solve coarse}:}                      & &F^H(w^H) = \ell^H \\
+&\text{\textbf{correct}:}                    & &w^h \leftarrow w^h + P(w^H - \iR w^h) \\
 &\text{post-smooth over fine:} & & [\text{smoother updates } w^h]
@@
 \begin{itemize}
-\item $P: \mathcal{V}^H \to \mathcal{V}^h$ is \emph{prolongation}
-\item restrict$+$(solve coarse)$+$correct \, $=$ \, \emph{FAS coarse grid correction}
+\item $P: \mathcal{V}^H \to \mathcal{V}^h$ is \emph{canonical prolongation}
+\item \textbf{restrict}$+$(\textbf{solve coarse})$+$\textbf{correct} \, $=$ \, \emph{FAS coarse grid correction}
 \end{itemize}
@@
 
-\begin{frame}{nonlinear multigrid by FAS: \only<1>{V-cycle}\only<2>{``FMG'' cycle}}
+\begin{frame}{nonlinear multigrid by FAS: \only<1>{V-cycle}\only<2>{FMG cycle}}
 
@@
 
-\hspace{4mm} $j=3$ \hspace{13mm} $j=2$ \hspace{13mm} $j=1$ \hspace{12mm} $j=0$
+\hspace{4mm} $J=3$ \hspace{13mm} $j=2$ \hspace{13mm} $j=1$ \hspace{12mm} $j=0$
 
@@
 \only<2>{
-\vspace{12mm}
+\vspace{6mm}
 
-\hfill \input{tikz/fcycle.tex}
+\centering
+\input{tikz/fcycle.tex}
+
+FMG $=$ full multigrid
 
@@
 \begin{itemize}
-\item FAS multigrid works \alert{very well} on the right nonlinear PDE problem
+\item FAS multigrid works \alert{very well} on nice nonlinear PDE problems
 \item example: Liouville-Bratu equation
@@
 with Dirichlet boundary conditions on $\Omega=(0,1)^2$
-\item discretize by straightforward finite differences
+\item discretize by (straightforward) finite differences
 \item minimal problem-specific code:
@@
         \end{itemize}
-    \item[3.] coarse solve can be same as smoother, or more sophisticated-but-not-optimal (e.g.~Newton iteration)
+    \item[3.] coarsest-level solve can be same as smoother, or more sophisticated (e.g.~Newton iteration)
     \end{enumerate}
@@
 \begin{itemize}
-\item multigrid: optimality is a frequently-achieved goal for PDE problems with reasonably-smooth solutions
+\item since $\sim$1980: optimality can be achieved by multigrid for PDE problems with reasonably-smooth solutions
 \item<2> in fact, multigrid people get greedy
@@
 
-\begin{block}<2>{definition} a solver exhibits \emph{textbook multigrid efficiency} if it is optimal and it does work less than 10 times that of a single smoother sweep
+\begin{block}<2>{definition} a solver exhibits \emph{textbook multigrid efficiency} if it is optimal and it does total work less than 10 times that of a single smoother sweep
 \end{block}
@@
 \begin{itemize}
+\item \texttt{bratu.c}
 \item observed optimality:
@@
 \end{align*}
-\item $N\approx 10^8$ unknowns for $12$-level V-cycle
+\item {\color{FireBrick} $12$-level V-cycle} has $N\approx 10^8$ unknowns
 \item compare $\approx 20\,\mu\,\text{s}/\text{N}$ for Poisson using Firedrake ($P_1$, geometric multigrid)
@@
 \begin{column}{0.55\textwidth}
-\includegraphics[width=\textwidth]{../talk-oxford/images/bratu-time.png}
+\includegraphics[width=\textwidth]{figs/bratu-time.png}
 \end{column}
 \end{columns}
-\end{frame}
-
-\begin{frame}{benefits of FAS multigrid for nonlinear PDEs?}
-
-\begin{itemize}
-\item \alert{benefits} of FAS multigrid?
-    \begin{enumerate}
-    \item[1.] minimal code, esp.~in from-scratch implementations
-        \begin{itemize}
-        \item[$\circ$] just write residual plus pointwise smoother!
-        \end{itemize}
-    \item[2.] composition with nonlinear preconditioners (Brune et al.~2015)
-    \end{enumerate}
-
-\bigskip
-\item \alert{disadvantages?}
-    \begin{enumerate}
-    \item[1.] Firedrake/FENiCs \emph{do} automatically provide linearizations from UFL statements of weak forms
-    \item[2.] small literature of convergence or descriptive performance for FAS (Trottenberg et al.~(2001), Reusken (1987))
-    \item[3.] not enough tutorial literature?
-    \end{enumerate}
-\end{itemize}
 \end{frame}
@@
 \begin{itemize}
-\item VIs are nonlinear problems, even for linear operators like $-\nabla^2$
-\item most obvious method (?): Newton-multigrid
+\item best-known multigrid method for VIs: Newton-multigrid
     \begin{itemize}
-    \item[$\circ$] linear solver applies to inactive variables
+    \item[$\circ$] linear solver applies to currently-inactive variables
         \begin{itemize}
-        \item[] reduced space line search
+        \item ``reduced space line search''
         \end{itemize}
@@
     \end{itemize}
-\item issue: the outer Newton iteration must converge on the active set \alert{before} multigrid can provide effective preconditioning
-    \begin{itemize}
-    \item[$\circ$] consequence: grid-dependent (growing) Newton iterations
-    \end{itemize}
+\item observed in actual use: grid-dependent Newton iterations
+\item the (outer) Newton iteration must converge on the active set \alert{before} linear multigrid can provide effective preconditioning
 \end{itemize}
@@
 \medskip
-\hspace{10mm} \includegraphics[height=0.2\textheight]{../talk-oxford/images/obstacle65.pdf}
-
-\vspace{-20mm}
-\hfill \includegraphics[height=0.25\textheight]{../talk-oxford/images/vi-newton-gmg-bad.png}
+\includegraphics[height=0.15\textheight]{../talk-oxford/images/obstacle65.pdf} \qquad \includegraphics[height=0.25\textheight]{../talk-oxford/images/vi-newton-gmg-bad.png}
 \end{frame}
@@
 \begin{itemize}
-\item applying nested iteration helps this:
+\item on simpler VI problems, \alert{nested iteration} helps
     \begin{itemize}
-    \item[$\circ$] this is effectively an FMG-cycle, an alternative to FASCD
-    \item[$\circ$] grid-independent Newton iterations
-    \item[$\circ$] optimal $O(N)$ flops and time; Chapter 12 example in my new book
+    \item[$\circ$] nested-iteration = grid-sequencing
+    \item[$\circ$] on classical obstacle problem: grid-independent Newton iterations, optimal $O(N)$ flops and time; Chapter 12 in my book
     \end{itemize}
+\item not as effective on harder (\emph{lower-regularity, complicated free boundary}) problems
 \end{itemize}
 
-\bigskip
-\includegraphics[height=0.25\textheight]{../talk-oxford/images/vi-newton-gmg-good.png}
-
-\vspace{-22mm}
-\hspace{40mm} \includegraphics[width=20mm]{../talk-oxford/images/frontcover.jpg}
+\medskip
+\includegraphics[height=0.15\textheight]{../talk-oxford/images/obstacle65.pdf} \qquad \includegraphics[height=0.25\textheight]{../talk-oxford/images/vi-newton-gmg-good.png} \qquad \includegraphics[width=15mm]{../talk-oxford/images/frontcover.jpg}
 \end{frame}
@@
 
+\begin{center}
+\begin{tabular}{r|ccc}
+   & \stacktwo{mesh-indep.}{rates}
+       & \stacktwo{no need for}{global linearization}  \\ \hline
+NM                         &              & \\
+NM $+$ NI                  & $\checkmark$?& \\
+{\color{FireBrick} FASCD}  & $\checkmark$ & $\checkmark$ {\Large \strut}
+\end{tabular}
+\end{center}
+
+\medskip
 {\small
-\begin{tabular}{r|cccc}
-   & \stacktwo{admissible}{iterates} & \stacktwo{mesh-indep.}{rates}
-       & \stacktwo{no need for}{global linearization}  \\ \hline
-NM                         & $\checkmark$ &              & \\
-\quad $+$ NI               & $\checkmark$ & $\checkmark$ & \\
-{\color{FireBrick} FASCD}  & $\checkmark$ & $\checkmark$ & $\checkmark$ {\Large \strut}
-\end{tabular}
-}
-
-\bigskip
-{\scriptsize
 \begin{center}
@@
 \begin{itemize}
-\item new algorithm,
+\item new algorithm (2023):
 
@@
 
+\begin{frame}{subspace decomposition}
+
+\hfill \includegraphics[height=0.12\textheight]{../talk-oxford/images/mg-grids.png}
+
+{\footnotesize
+\hfill $\Omega^3$ \hspace{8.5mm} $\Omega^2$ \hspace{8.5mm} $\Omega^1$ \hspace{8.5mm} $\Omega^0$ \hspace{1mm}
+}
+
+\begin{itemize}
+\item to explain ``constraint decomposition'' in FAS\underline{CD}, start with subspace decomposition in the multilevel case
+\item assume mesh levels are nested: $\Omega^j \subset \Omega^{j+1}$ nodewise
+\item let $\mathcal{V}^j$ be the space of FE functions on $\Omega^j$
+    \begin{itemize}
+    \item[$\circ$] standard FE spaces: \, $\mathcal{V}^j \subset \mathcal{V}^{j+1}$
+    \end{itemize}
+\begin{block}{definition}
+$\ds \mathcal{V}^J = \sum_{i=0}^J \mathcal{V}^i$ \quad is called a \emph{subspace decomposition} (Xu 1992)
+\end{block}
+    \begin{itemize}
+    \item[$\circ$] \emph{non}-unique vector space sum
+    \end{itemize}
+\end{itemize}
+\end{frame}
+
 \begin{frame}{constraint decomposition}
@@
 \begin{itemize}
-\item Tai's (2003) \emph{constraint decomposition} (CD) for VIs follows the subspace decomposition idea (Xu 1992)
+\item Tai's (2003) \emph{constraint decomposition} (CD) for VIs extends subspace decomposition to convex subsets
+    \begin{itemize}
+    \item[$\circ$] non-trivial
+    \end{itemize}
 \item suppose $\mathcal{K} \subset \mathcal{V}$ is a closed and convex admissible subset
@@
 where $\mathcal{K}^i \subset \mathcal{V}^i$, with projections $\Pi_i : \mathcal{K} \to \mathcal{K}_i$
-\item CD multiplicative iterations for $VI(F,\ell,\mathcal{K})$
+\end{itemize}
+\end{frame}
 
-\begin{pseudo}
-\pr{cd-mult}
+\begin{frame}{constraint decomposition: multiplicative iterations}
+
+\begin{itemize}
+\item multiplicative iterations for $VI(F,\ell,\mathcal{K})$ over the CD
+
+{\small
+\begin{pseudo}[left-margin=-5mm]
+\pr{cd-mult}(u)\text{:} \\+
+    for $i = 0,\dots,m-1$: \\+
+        find $w_i\in \cK_i$ s.t. \\+
+            $\displaystyle \Big<f\Big(\sum_{j<i} w_j + w_i + \sum_{j>i} \Pi_j u\Big),\, v_i - w_i\Big> \ge \ip{\ell}{v_i - w_i} \,\forall v_i \in \cK_i$ \\--
+    return $w=\sum_i w_i\in\cK$
 \end{pseudo}
+}
 \end{itemize}

### 2023-11-09 talk-dms/py/advdiff.py
tidy python figure gens

@@ -1,13 +1,8 @@
+#!/usr/bin/env python3
+
 import numpy as np
 import matplotlib.pyplot as plt
-
-_SHOW = False
-def writeout(outname):
-    if _SHOW:
-        plt.show()
-    else:
-        print('writing file ' + outname)
-        plt.savefig(outname, bbox_inches='tight')
+from writeout import writeout
 
 # data from Table 4 in paper
 n = 15 * 2**np.arange(0,9) + 1

### 2023-11-09 talk-dms/py/classical.py

@@ -1,13 +1,8 @@
+#!/usr/bin/env python3
+
 import numpy as np
 import matplotlib.pyplot as plt
-
-_SHOW = False
-def writeout(outname):
-    if _SHOW:
-        plt.show()
-    else:
-        print('writing file ' + outname)
-        plt.savefig(outname, bbox_inches='tight')
+from writeout import writeout
 
 # data from Table 2 in paper
 mJ = [41, 145, 545, 2.1e3, 8.3e3, 3.3e4, 1.3e5, 5.3e5, 2.1e6, 8.4e6, 3.4e7]

### 2023-11-09 talk-dms/py/siaweak.py

@@ -1,13 +1,8 @@
+#!/usr/bin/env python3
+
 import numpy as np
 import matplotlib.pyplot as plt
-
-_SHOW = False
-def writeout(outname):
-    if _SHOW:
-        plt.show()
-    else:
-        print('writing file ' + outname)
-        plt.savefig(outname, bbox_inches='tight')
+from writeout import writeout
 
 # data from Table 5 in paper
 P = ['1','4','16','64','256','1024']

### 2023-11-09 talk-dms/py/innerconeapprox.py
inner cone approx figure

@@ -1,0 +1,47 @@
+#!/usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.patches import Arc
+from writeout import writeout
+
+plt.rcParams["font.family"] = "serif"
+plt.rcParams["mathtext.fontset"] = "cm"
+
+fsize=15.0
+
+def makeaxes(xmin,xmax,ymin,ymax):
+    x = np.linspace(xmin,xmax,2)
+    y = np.linspace(ymin,ymax,2)
+    zero = np.zeros(np.shape(x))
+    plt.plot(x,zero,'k',lw=1.0)
+    plt.plot(zero,y,'k',lw=1.0)
+
+plt.figure(figsize=(8,6))
+xm, ym = 0.6, 0.4
+makeaxes(-0.4, xm, -0.25, ym)
+cornerx = [-0.3, -0.29, -0.2, -0.1]
+cornery = [-0.2, -0.14, -0.09, 0.0]
+sh = 0.02
+for n, xc, yc in zip(range(4), cornerx, cornery):
+    plt.fill([xc,xm,xm,xc,xc],[yc,yc,ym,ym,yc],'k',alpha=0.13)
+    plt.plot(cornerx[n],cornery[n],'k.',ms=10.0)
+    plt.text(cornerx[n]+(n-3)*sh,cornery[n]-2*sh,r'$\chi^%d$' % (3-n),
+             fontsize=fsize)
+plt.plot(0.0,0.0,'k.',ms=10.0)
+plt.text(0.0+sh,0.0+sh,r'$0$',fontsize=fsize)
+plt.text(-0.4,0.35,r'$\mathcal{V}^3$',
+         fontsize=fsize)
+plt.text(0.4,0.03,r'$\mathcal{U}^0=\mathcal{D}^0$',
+         fontsize=fsize)
+plt.text(0.33,cornery[2]+0.02,r'$\mathcal{U}^1=\mathcal{D}^0+\mathcal{D}^1$',
+         fontsize=fsize)
+plt.text(0.27,cornery[1]+0.01,r'$\mathcal{U}^2=\mathcal{D}^0+\mathcal{D}^1+\mathcal{D}^2$',
+         fontsize=fsize)
+plt.text(0.19,cornery[0]+0.01,
+         r'$\mathcal{U}^3=\mathcal{D}^0+\mathcal{D}^1+\mathcal{D}^2+\mathcal{D}^3$',
[…]

### 2023-11-09 talk-dms/slides.tex
inner cone approx figure
what I practice on

@@
 
-\subsection{earlier approaches}
+\subsection{an earlier approach}
 
@@
     \item[$\circ$] nested-iteration = grid-sequencing
-    \item[$\circ$] on classical obstacle problem: grid-independent Newton iterations, optimal $O(N)$ flops and time; Chapter 12 in my book
+    \item[$\circ$] on classical obstacle problem: mesh-independent Newton iterations, optimal $O(N)$ flops and time; Chapter 12 in my book
     \end{itemize}
@@
 \begin{itemize}
-\item to explain ``constraint decomposition'' in FAS\underline{CD}, start with subspace decomposition in the multilevel case
+\item to explain the ``CD'' (constraint decomposition) in FAS\underline{CD}, start with subspace decomposition in the multilevel case
 \item assume mesh levels are nested: $\Omega^j \subset \Omega^{j+1}$ nodewise
@@
 \begin{itemize}
-\item Tai's (2003) \emph{constraint decomposition} (CD) for VIs extends subspace decomposition to convex subsets
+\item Tai's (2003) constraint decomposition for VIs extends subspace decomposition to convex subsets
     \begin{itemize}
-    \item[$\circ$] non-trivial
+    \item[$\circ$] extends non-trivially
     \end{itemize}
-\item suppose $\mathcal{K} \subset \mathcal{V}$ is a closed and convex admissible subset
-\item for a subspace decomposition $\mathcal{V} = \sum_i \mathcal{V}_i$, write the admissible subset as a sum
-    $$\mathcal{K} = \sum_i \mathcal{K}_i$$
-where $\mathcal{K}^i \subset \mathcal{V}^i$, with projections $\Pi_i : \mathcal{K} \to \mathcal{K}_i$
+\item suppose $\mathcal{K}^J \subset \mathcal{V}^J$ is a closed and convex subset
+\item assume $\mathcal{V}^J = \sum_i\mathcal{V}^i$ is a subspace decomposition
+\begin{block}{definition}
+$\ds \mathcal{K}^J = \sum_{i=0}^J \mathcal{K}^i$ \quad is a \emph{constraint decomposition} (CD) if there are closed and convex subsets $\mathcal{K}^i\subset \mathcal{V}^i$, and (nonlinear) projections $\Pi_i : \mathcal{K}^J \to \mathcal{K}^i$, so that $\ds v = \sum_{i=0}^J \Pi_i v$ and a stability condition applies (not shown)
+\end{block}
 \end{itemize}
+\end{frame}
+
+\begin{frame}{constraint decomposition}
+
+\begin{itemize}
+\item observation: generally $\mathcal{K}^i \not\subset \mathcal{K}^J$
+\item picture below: obstacle problem on a two-point mesh ($\mathcal{V} \cong \RR^2$)
+\end{itemize}
+
+\bigskip
+\begin{center}
+\includegraphics[width=0.6\textwidth]{../paper/genfigs/cartoon.pdf}
+\end{center}
 \end{frame}
@@
 
-\begin{frame}{multilevel constraint decomposition}
+\begin{frame}{defect obstacles}
 
@@
 \item recall $\mathcal{K} = \{v \ge \psi\}$ in classical obstacle problem
-\item define \alert{defect obstacle} for a fine-level iterate $w^J$:
+\begin{block}{definition}
+the \emph{defect obstacle} for a fine-level iterate $w^J \in \mathcal{K}^J = \{v\,:\,v\ge \psi^J\}$ is an FE function in $\mathcal{V}^J$:
     $$\chi^J = \psi^J - w^J$$
-\item \alert{monotone restriction} generates obstacles on each level:
-	$$\chi^j = R^{\oplus} \chi^{j+1}$$
-\item let $\mathcal{U}^j = \{z \ge \chi^j\}$, $\mathcal{D}^j = \{y \ge \chi^j - \chi^{j-1}\}$
+\end{block}
+\item we generate the CD through defect obstacles $\chi^j$ on each level
+\item via \emph{monotone restriction}: \quad $\ds \chi^j = R^{\oplus} \chi^{j+1}$
+\end{itemize}
+
+\begin{center}
+\input{tikz/Rplusminus.tex}
+\end{center}
+\end{frame}
+
+\begin{frame}{down and up sets}
+
+\begin{itemize}
+\item let $\mathcal{U}^j = \{z \ge \chi^j\}$, $\mathcal{D}^j = \{y \ge \phi^j=\chi^j - \chi^{j-1}\}$
 \item get CD of fine-level constraint set:
 	$$\mathcal{U}^J = \sum_{i=0}^J \mathcal{D}^i \phantom{KLDFJSKJDS SDF}$$
-\item multiplicative CD iteration $\to$ V-cycle
 \end{itemize}
 
-\vspace{-25mm}
-\hfill \includegraphics[width=0.4\textwidth]{../talk-oxford/images/fascd-vcycle.png}
+\begin{center}
+\includegraphics[width=0.4\textwidth]{../talk-oxford/images/fascd-vcycle.png}
+\end{center}
+\end{frame}
+
+\begin{frame}{multilevel constraint decomposition}
+
+\centering
+\includegraphics[width=0.7\textwidth]{figs/innerconeapprox.png}
 \end{frame}
@@
 \begin{pseudo}
-FIXME
+\pr{fascd-vcycle}(J,\ell^J,\psi^J;w^J)\text{:} \\+
+    $\chi^J = \psi^J - w^J$ \\
+    for $j=J$ downto $j=1$ \\+
+      $\chi^{j-1} = \maxR \chi^j$ \\
+      $\phi^j = \chi^j - P\chi^{j-1}$ \\
+      $y^j = 0$ \\
+      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\phi^j,w^j;y^j)$  \hspace{1.0cm} \ct{in $\mathcal{D}^j$}\\
+      $w^{j-1} = \iR(w^j + y^j)$ \\
+      $\ell^{j-1} = f^{j-1}(w^{j-1}) + R \left(\ell^j - f^j(w^j+y^j)\right)$ \\-
+    $z^0 = 0$ \\
+    $\text{\pr{solve}}(\ell^0,\chi^0,w^0;z^0)$ \hspace{1.0cm} \ct{in $\mathcal{U}^0$} \\
+    for $j=1$ to $j=J$ \\+
+      $z^j = y^{j} + P z^{j-1}$ \\
+      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\chi^j,w^j;z^j)$  \hspace{1.0cm} \ct{in $\mathcal{U}^j$} \\-
+    return $w^J+z^J$
 \end{pseudo}

### 2023-11-12 talk-dms/references.tex
slides in good shape through start of glaciers example

@@
 %\item R.~Kornhuber (1994). \emph{Monotone multigrid methods for elliptic variational inequalities I}, Numer.~Math.~69, 167--184 \sdoi{10.1007/BF03325426}
-\item A.~Reusken (1987).
+%\item A.~Reusken (1987).
 \emph{Convergence of the multigrid full approximation scheme for a class of elliptic mildly nonlinear boundary value problems}, Numer.~Math.~52, 251--277 \sdoi{10.1007/BF01398879}
@@
 \item X.~Tai (2003). \emph{Rate of convergence for some constraint decomposition methods for nonlinear variational inequalities}, Numer.~Math.~93 (4), 755--786 \sdoi{10.1007/s002110200404}
+\item J.~Xu (1992). \emph{Iterative methods by space decomposition and subspace correction}, SIAM Rev.~34 (4), 581--613 \sdoi{10.1137/1034116}
 \end{itemize}

### 2023-11-12 talk-dms/slides.tex
slides in good shape through start of glaciers example

@@
 
-\title[Fast solvers for PDE+inequalities]{Fast solvers for partial differential equations \\ subject to inequalities}
+\title[Fast solvers for PDE subject to inequalities]{Fast solvers for partial differential equations \\ subject to inequalities}
 
@@
 
-\begin{frame}{Outline}
+\begin{frame}{Outline: \emph{Fast solvers for PDEs subject to inequalities}}
   \tableofcontents[hideallsubsections]
-  %\tableofcontents
-
-%\vfill
-%{\footnotesize \emph{PDEs = partial differential equations}}
 \end{frame}
@@
 }
-\only<2-3>{
+\only<2>{
 \begin{itemize}
@@
    \item[$\circ$] Poisson equation $-\grad^2 u = f$ is ``$J'(u)=0$'' on $R_u$
-   \item<3>[$\circ$] using the solution $u$ to define the set $R_u$ on which to solve the PDE $-\grad^2 u = f$ \alert{does not lead to solution algorithms}
+   \item[$\circ$] using the solution $u$ to define the set $R_u$ on which to solve the PDE $-\grad^2 u = f$ \alert{does not lead to solution algorithms}
    \end{itemize}
@@
 }
-\only<4>{
+\only<3>{
 \begin{itemize}
@@
    \begin{itemize}
-   \item[$\circ$] CP $=$ (KKT conditions in $\infty$-dimensions)
+   \item[$\circ$] CP $=$ KKT conditions
+       \begin{itemize}
+       \item but in $\infty$-dimensions
        \end{itemize}        
-
-\phantom{x}
+   \end{itemize}
 
@@
 }
-\only<5>{
+\only<4>{
 \begin{itemize}
@@
 
-\begin{frame}{VI $=$ weak form with $v-u$ as the test function}
+\begin{frame}{VI $=$ weak form}
 
+\begin{itemize}
+\item for problems of optimization type, the VI is the weak form, with $v-u$ as the test function:
 {\small
@@
 }
+\end{itemize}
 
@@
 	\item[$\circ$] $\ell\in \mathcal{V}'$
-	\item[$\circ$] $F : \mathcal{V} \to \mathcal{V}'$ continuous and injective
+	\item[$\circ$] $F : \mathcal{V} \to \mathcal{V}'$ continuous and one-to-one
 	\item[$\circ$] for example, the Liouville-Bratu problem: $-\grad^2 u - e^u = f$
@@
 	\end{itemize}
-\item<3> notes:
+\item notes:
     \setbeamertemplate{enumerate item}{\emph{\roman{enumi})}}
@@
 }
-\only<4-5>{
+\only<3-4>{
 \begin{itemize}
@@
     $$w^h \leftarrow w^h+\tilde e^h$$
-\item<5> but $F^h$ is not linear!
+\item<4> but $F^h$ is not linear!
 \end{itemize}
@@
 }
-\only<6>{
+\only<5>{
 \begin{itemize}
@@
 }
-\only<7>{
+\only<6>{
 \begin{itemize}
@@
 }
-\only<8-9>{
+\only<7>{
 \begin{itemize}
@@
 
-\item<9> rewritten FAS equation: let ${\color{FireBrick} \ell^H = F^H(\iR w^h) + R\, r^h(w^h)}$ then
+\item rewritten FAS equation: let ${\color{FireBrick} \ell^H = F^H(\iR w^h) + R\, r^h(w^h)}$ then
     $${\color{FireBrick} F^H(u^H) = \ell^H}$$
 \end{itemize}
-
-\vspace{-3mm}
 }
@@
 &\text{pre-smooth over fine:} & & [\text{smoother updates } w^h] \\
-&\text{\textbf{restrict}:}                   & &\ell^H = F^H(\iR w^h) + R\, r^h(w^h) \\
-&\text{\textbf{solve coarse}:}                      & &F^H(w^H) = \ell^H \\
-&\text{\textbf{correct}:}                    & &w^h \leftarrow w^h + P(w^H - \iR w^h) \\
+&\text{restrict:}                   & &\ell^H = F^H(\iR w^h) + R\, r^h(w^h) \\
+&\text{solve coarse:}                      & &F^H(w^H) = \ell^H \\
+&\text{correct:}                    & &w^h \leftarrow w^h + P(w^H - \iR w^h) \\
 &\text{post-smooth over fine:} & & [\text{smoother updates } w^h]
@@
 \item since $\sim$1980: optimality can be achieved by multigrid for PDE problems with reasonably-smooth solutions
-\item<2> in fact, multigrid people get greedy
+\item in fact, multigrid people get greedy
+\begin{block}{definition} a solver shows \emph{textbook multigrid efficiency} if it does total work less than 10 times that of a single smoother sweep
+\end{block}
+    \begin{itemize}
+    \item[$\circ$] TME $\implies$ optimal
     \end{itemize}
-
-\begin{block}<2>{definition} a solver exhibits \emph{textbook multigrid efficiency} if it is optimal and it does total work less than 10 times that of a single smoother sweep
-\end{block}
+\end{itemize}
 \end{frame}
 
-\begin{frame}{Bratu model problem: optimality and TME}
+\begin{frame}{Bratu model problem: TME}
 
@@
 \text{flops} &= O(N^1) \\
-\text{exp evaluations} &= O(N^1) \\
 \text{processor time} &= O(N^1)
 \end{align*}
-\item {\color{FireBrick} $12$-level V-cycle} has $N\approx 10^8$ unknowns
+\item highest-resolution {\color{FireBrick} $12$-level V-cycle} has $N\approx 10^8$ unknowns
 \item compare $\approx 20\,\mu\,\text{s}/\text{N}$ for Poisson using Firedrake ($P_1$, geometric multigrid)
@@
 
-\subsection{an earlier approach}
+\subsection{full approximation scheme constraint decomposition (FASCD)}
 
-\begin{frame}[fragile]
-\frametitle{Newton-multigrid for the classical obstacle problem}
+\begin{frame}{an FAS multigrid strategy for VIs}
 
 \begin{itemize}
-\item best-known multigrid method for VIs: Newton-multigrid
-    \begin{itemize}
-    \item[$\circ$] linear solver applies to currently-inactive variables
-        \begin{itemize}
-        \item ``reduced space line search''
-        \end{itemize}
-    \item[$\circ$] Newton step equations solved by CG with linear multigrid V-cycles
-    \end{itemize}
-\item observed in actual use: grid-dependent Newton iterations
-\item the (outer) Newton iteration must converge on the active set \alert{before} linear multigrid can provide effective preconditioning
-\end{itemize}
-
-\medskip
-\includegraphics[height=0.15\textheight]{../talk-oxford/images/obstacle65.pdf} \qquad \includegraphics[height=0.25\textheight]{../talk-oxford/images/vi-newton-gmg-bad.png}
-\end{frame}
-
-\begin{frame}[fragile]
-\frametitle{nested iteration}
-
-\begin{itemize}
-\item on simpler VI problems, \alert{nested iteration} helps
-    \begin{itemize}
-    \item[$\circ$] nested-iteration = grid-sequencing
-    \item[$\circ$] on classical obstacle problem: mesh-independent Newton iterations, optimal $O(N)$ flops and time; Chapter 12 in my book
-    \end{itemize}
-\item not as effective on harder (\emph{lower-regularity, complicated free boundary}) problems
-\end{itemize}
-
-\medskip
-\includegraphics[height=0.15\textheight]{../talk-oxford/images/obstacle65.pdf} \qquad \includegraphics[height=0.25\textheight]{../talk-oxford/images/vi-newton-gmg-good.png} \qquad \includegraphics[width=15mm]{../talk-oxford/images/frontcover.jpg}
-\end{frame}
-
-\newcommand{\stacktwo}[2]{\begin{tabular}{c} #1 \\ #2 \end{tabular}}
-
-\begin{frame}{multigrid strategies for VIs: feature table}
-
-\begin{center}
-\begin{tabular}{r|ccc}
-   & \stacktwo{mesh-indep.}{rates}
-       & \stacktwo{no need for}{global linearization}  \\ \hline
-NM                         &              & \\
-NM $+$ NI                  & $\checkmark$?& \\
-{\color{FireBrick} FASCD}  & $\checkmark$ & $\checkmark$ {\Large \strut}
-\end{tabular}
-\end{center}
-
-\medskip
-{\small
-\begin{center}
-NM = Newton-multigrid, NI = nested iteration
-\end{center}
-}
-
-\vspace{5mm}
-\begin{itemize}
-\item new algorithm (2023):
+\item new algorithm (Bueler \& Farrell 2023):
 
 {\color{FireBrick} FASCD = full approximation scheme constraint decomposition}
+
+\bigskip
+\item what is ``constraint decomposition'' in FAS\underline{CD}?
 \end{itemize}
 \end{frame}
-
-\subsection{full approximation scheme constraint decomposition (FASCD)}
 
@@
 \begin{itemize}
-\item to explain the ``CD'' (constraint decomposition) in FAS\underline{CD}, start with subspace decomposition in the multilevel case
-\item assume mesh levels are nested: $\Omega^j \subset \Omega^{j+1}$ nodewise
-\item let $\mathcal{V}^j$ be the space of FE functions on $\Omega^j$
-    \begin{itemize}
-    \item[$\circ$] standard FE spaces: \, $\mathcal{V}^j \subset \mathcal{V}^{j+1}$
-    \end{itemize}
+\item start with subspace decomposition over nested meshes:
+    $$\Omega^j \subset \Omega^{j+1}$$
+\item the FE function spaces $\mathcal{V}^j$ over $\Omega^j$ are also nested:
+    $$\mathcal{V}^j \subset \mathcal{V}^{j+1}$$
+
 \begin{block}{definition}
@@
     \item[$\circ$] \emph{non}-unique vector space sum
+    \item[$\circ$] Xu's paper explains how to analyze linear multigrid for PDEs via subspace decomposition
     \end{itemize}
@@
 \begin{itemize}
-\item Tai's (2003) constraint decomposition for VIs extends subspace decomposition to convex subsets
-    \begin{itemize}
-    \item[$\circ$] extends non-trivially
-    \end{itemize}
+\item Tai's (2003) constraint decomposition \emph{non-trivially} extends a subspace decomposition $\mathcal{V}^J = \sum_i\mathcal{V}^i$ to convex subsets
 \item suppose $\mathcal{K}^J \subset \mathcal{V}^J$ is a closed and convex subset
-\item assume $\mathcal{V}^J = \sum_i\mathcal{V}^i$ is a subspace decomposition
 \begin{block}{definition}
@@
 \item observation: generally $\mathcal{K}^i \not\subset \mathcal{K}^J$
-\item picture below: obstacle problem on a two-point mesh ($\mathcal{V} \cong \RR^2$)
 \end{itemize}
@@
 \includegraphics[width=0.6\textwidth]{../paper/genfigs/cartoon.pdf}
+
+\medskip
+{\small obstacle problem on a two-point mesh with $\mathcal{V} \cong \RR^2$}
 \end{center}
@@
 
-\begin{frame}{constraint decomposition: multiplicative iterations}
+\begin{frame}{iterations over constraint decomposition}
 
 \begin{itemize}
-\item multiplicative iterations for $VI(F,\ell,\mathcal{K})$ over the CD
+\item Tai proposed abstract iterations for solving $VI(F,\ell,\mathcal{K})$ over a CD $\mathcal{K}^J = \sum_{i=0}^J \mathcal{K}^i$
 
@@
         find $w_i\in \cK_i$ s.t. \\+
-            $\displaystyle \Big<f\Big(\sum_{j<i} w_j + w_i + \sum_{j>i} \Pi_j u\Big),\, v_i - w_i\Big> \ge \ip{\ell}{v_i - w_i} \,\forall v_i \in \cK_i$ \\--
+            $\displaystyle \Big<F\Big(\sum_{j<i} w_j + w_i + \sum_{j>i} \Pi_j u\Big),\, v_i - w_i\Big> \ge \ip{\ell}{v_i - w_i} \,\forall v_i \in \cK_i$ \\--
     return $w=\sum_i w_i\in\cK$
@@
 }
+
+\medskip
+\item Tai's iterations are not practical because you compute on the finest level in fact
+\item we added two techniques: \emph{defect obstacles} on each level, and \emph{FAS coarse corrections}
 \end{itemize}
@@
 \begin{itemize}
-\item recall $\mathcal{K} = \{v \ge \psi\}$ in classical obstacle problem
+\item recall $\mathcal{K} = \{v \ge \psi\}$ in an obstacle problem
 \begin{block}{definition}
-the \emph{defect obstacle} for a fine-level iterate $w^J \in \mathcal{K}^J = \{v\,:\,v\ge \psi^J\}$ is an FE function in $\mathcal{V}^J$:
-    $$\chi^J = \psi^J - w^J$$
+for finest-level admissible set $\mathcal{K}^J = \{v^J\ge \psi^J\} \subset \mathcal{V}^J$ and an iterate $w^J \in \mathcal{K}^J$, the \emph{defect obstacle} (Gr\"aser \& Kornhuber 2009) is
+    $$\chi^J = \psi^J - w^J \in \mathcal{V}^J$$
 \end{block}
-\item we generate the CD through defect obstacles $\chi^j$ on each level
-\item via \emph{monotone restriction}: \quad $\ds \chi^j = R^{\oplus} \chi^{j+1}$
+
+    \begin{itemize}
+    \item[$\circ$] note $\chi^J \le 0$
+    \end{itemize}
+\item we generate the CD through
+
+defect obstacles $\chi^j$ on each
+
+level via \emph{monotone restriction}:
+
+$$\chi^j = R^{\oplus} \chi^{j+1} \phantom{smdlfkaj asdfklj asdf sdfaa asddfas dsa}$$
+
+    \begin{itemize}
+    \item[$\circ$] a \emph{nonlinear} operator
+    \item[$\circ$] {\footnotesize also due to (Gr\"aser \& Kornhuber 2009)}
+    \end{itemize}
 \end{itemize}
 
-\begin{center}
-\input{tikz/Rplusminus.tex}
-\end{center}
+\vspace{-25mm}
+\hfill \mbox{\input{tikz/Rplusminus.tex}}
 \end{frame}
 
-\begin{frame}{down and up sets}
+\begin{frame}{up and down sets}
 
 \begin{itemize}
-\item let $\mathcal{U}^j = \{z \ge \chi^j\}$, $\mathcal{D}^j = \{y \ge \phi^j=\chi^j - \chi^{j-1}\}$
-\item get CD of fine-level constraint set:
-	$$\mathcal{U}^J = \sum_{i=0}^J \mathcal{D}^i \phantom{KLDFJSKJDS SDF}$$
+\item upward part in the FASCD V-cycle uses large admissible sets:
+    $$\mathcal{U}^j = \{z^j \ge \chi^j\}$$
+\item downward sets are smaller to guarantee admissibility of the upcoming coarse correction:
+    $$\mathcal{D}^j = \{y^j \ge \phi^j=\chi^j - \chi^{j-1}\}$$
+\item $\ds \mathcal{U}^j = \sum_{i=0}^j \mathcal{D}^i$ is a CD of the $j$th-level admissible set
 \end{itemize}
@@
 
-\begin{frame}{multilevel constraint decomposition}
+\begin{frame}{multilevel constraint decomposition in FASCD}
 
@@
 % case of lower obstacle only
-\begin{pseudo}
+\begin{pseudo}[font=\small]
 \pr{fascd-vcycle}(J,\ell^J,\psi^J;w^J)\text{:} \\+
@@
       $y^j = 0$ \\
-      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\phi^j,w^j;y^j)$  \hspace{1.0cm} \ct{in $\mathcal{D}^j$}\\
+      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\phi^j,w^j;y^j)$ \\
       $w^{j-1} = \iR(w^j + y^j)$ \\
@@
     $z^0 = 0$ \\
-    $\text{\pr{solve}}(\ell^0,\chi^0,w^0;z^0)$ \hspace{1.0cm} \ct{in $\mathcal{U}^0$} \\
+    $\text{\pr{solve}}(\ell^0,\chi^0,w^0;z^0)$  \\
     for $j=1$ to $j=J$ \\+
       $z^j = y^{j} + P z^{j-1}$ \\
-      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\chi^j,w^j;z^j)$  \hspace{1.0cm} \ct{in $\mathcal{U}^j$} \\-
+      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\chi^j,w^j;z^j)$  \\-
     return $w^J+z^J$
 \end{pseudo}
+\end{frame}
+
+\begin{frame}{FASCD specifics}
+
+\begin{itemize}
+\item see paper (Bueler \& Farrell 2023) for generalization to upper and lower obstacles:
+    $$\mathcal{K}^J = \{\underline{\psi}^J \le v^J \le \overline{\psi}^J\}$$
+\item see paper for stopping criteria
+    \begin{itemize}
+    \item[$\circ$] need to evaluate whether CP/KKT conditions are satisfied
+    \end{itemize}
+\item see paper for FASCD FMG cycle
+\end{itemize}
 \end{frame}
@@
 
-\begin{frame}{classical obstacle problem}
+\begin{frame}{classical obstacle problem by FASCD}
 
-\only<1>{\includegraphics[width=0.75\textwidth]{figs/ballitersV.png}}\only<2>{\includegraphics[width=0.75\textwidth]{figs/ballitersVlog.png}}\only<3>{\includegraphics[width=0.75\textwidth]{figs/ballitersVF.png}}\only<4>{\includegraphics[width=0.75\textwidth]{figs/bothitersVF.png}} \hfill \only<1-3>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/ball-set.png}}\only<4>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/spiral-set.png}}
+\only<1>{\includegraphics[width=0.75\textwidth]{figs/ballitersV.png}}\only<2>{\includegraphics[width=0.75\textwidth]{figs/ballitersVlog.png}}\only<3>{\includegraphics[width=0.75\textwidth]{figs/ballitersVF.png}}\only<4>{\includegraphics[width=0.75\textwidth]{figs/bothitersVF.png}} \hfill
+\only<1-3>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/ball-set.png}}%
+\only<4>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/spiral-set.png}}
 \end{frame}
@@
 
-\mbox{
-\includegraphics[width=0.3\textwidth]{../paper/fixfigs/poll2d-zero-set.png}
-\includegraphics[width=0.3\textwidth]{../paper/fixfigs/poll2d-one-set.png}
-}
+\begin{itemize}
+\item suppose $u(x)$ is a concentration in $\Omega \subset \RR^d$: \qquad $0\le u\le 1$
+\item suppose it moves by combination of diffusion, advection (wind $\bm{X}(x)$), and source ($\phi(x)$):
+    $$-\eps \grad^2 u + \bm{X}\cdot \grad u = \phi$$
+\item two active sets (\emph{pictured in $d=2$ case}):
+    $$\underline{A}_u = \{u(x) = 0\} \hspace{22mm} \overline{A}_u = \{u(x) = 1\} \hspace{8mm}$$
+\end{itemize}
 
-\begin{itemize}
-\item 2D problem: $-\eps \grad^2 u + \bm{X}\cdot \grad u = \phi$ with $0\le u\le 1$
-\end{itemize}
+\centering
+\includegraphics[width=0.27\textwidth]{../paper/fixfigs/poll2d-zero-set.png} \hspace{18mm}
+\includegraphics[width=0.27\textwidth]{../paper/fixfigs/poll2d-one-set.png}
 \end{frame}
@@
 
-\includegraphics[width=0.75\textwidth]{figs/advdiff.png} \hfill \begin{minipage}[t]{0.2\textwidth}
-\vspace{-50mm}
+\begin{center}
+\includegraphics[width=0.75\textwidth]{figs/advdiff.png}
+\end{center}
 
-\includegraphics[width=\textwidth]{../paper/fixfigs/poll2d-zero-set.png}
-
-\includegraphics[width=\textwidth]{../paper/fixfigs/poll2d-one-set.png}
-\end{minipage}
+\scriptsize
+\begin{itemize}
+\item<2> compare Klee-Minty cube?
+\end{itemize}
 \end{frame}

### 2023-11-12 talk-dms/references.tex
I think this is adequate

@@
 \begin{itemize}
-\item A.~Brandt (1977). \emph{Multi-level adaptive solutions to boundary-value problems}, Mathematics of Computation 31 (138), 333--390
+\item A.~Brandt (1977). \emph{Multi-level adaptive solutions to boundary-value problems}, Mathematics of Computation 31 (138), 333--390 \sdoi{10.1090/S0025-5718-1977-0431719-X}
 %\item A.~Brandt \& C.~Cryer (1983). \emph{Multigrid algorithms for the solution of linear complementarity problems \dots}, SIAM J.~Sci.~Stat.~Comput.~4 (4), 655--684 \sdoi{10.1137/0904046}
@@
 %\item P.~Brune, M.~Knepley, B.~Smith, \& X.~Tu (2015). \emph{Composing scalable nonlinear algebraic solvers}, SIAM Review~57 (4), 535--565 \sdoi{10.1137/130936725}
-\item E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032 \sdoi{10.1137/20M135217X}
+%\item E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032 \sdoi{10.1137/20M135217X}
 %\item E.~Bueler (2022). \emph{Performance analysis of high-resolution ice-sheet simulations}, J.~Glaciol., \sdoi{10.1017/jog.2022.113}

### 2023-11-12 talk-dms/slides.tex
I think this is adequate

@@
 
+\begin{frame}{FASCD V-cycle: visualization on a 1D problem}
+
+\centering
+\includegraphics[width=0.85\textwidth]{figs/vcycle-visualized.png}
+\end{frame}
+
 \begin{frame}{FASCD specifics}
 
+see paper (Bueler \& Farrell 2023) for:
 \begin{itemize}
-\item see paper (Bueler \& Farrell 2023) for generalization to upper and lower obstacles:
+\item generalization to upper and lower obstacles:
     $$\mathcal{K}^J = \{\underline{\psi}^J \le v^J \le \overline{\psi}^J\}$$
-\item see paper for stopping criteria
+\item stopping criteria
     \begin{itemize}
-    \item[$\circ$] need to evaluate whether CP/KKT conditions are satisfied
+    \item[$\circ$] evaluate whether CP/KKT conditions are satisfied
     \end{itemize}
-\item see paper for FASCD FMG cycle
+\item FMG cycle
+\item details of $O(m_J)$ smoother
 \end{itemize}
@@
 \item suppose $u(x)$ is a concentration in $\Omega \subset \RR^d$: \qquad $0\le u\le 1$
-\item suppose it moves by combination of diffusion, advection (wind $\bm{X}(x)$), and source ($\phi(x)$):
+\item suppose it moves by combination of diffusion, advection by wind $\bm{X}(x)$, and source function $\phi(x)$:
     $$-\eps \grad^2 u + \bm{X}\cdot \grad u = \phi$$
-\item two active sets (\emph{pictured in $d=2$ case}):
+\item two active sets (\emph{$d=2$ case}):
     $$\underline{A}_u = \{u(x) = 0\} \hspace{22mm} \overline{A}_u = \{u(x) = 1\} \hspace{8mm}$$
@@
 \begin{itemize}
-\item<2> compare Klee-Minty cube?
+\item[] \emph{compare: linear programming (Klee-Minty cube?), spatial correlations}
 \end{itemize}
@@
 
-\begin{frame}{problem: fluid layer in a climate}
+\begin{frame}{problem: geometry of flowing glacier ice in a climate}
 
 \begin{itemize}
-\item consider an incompressible, viscous layer with surface elevation $s(x,y)$, flowing with velocity $\bu(x,y,z)$, driven by gravity, over fixed bed topography with elevation $b(x,y)$, in a \emph{climate} which adds or removes fluid at a signed rate $a(x,y)$ [$\text{m}\,\text{s}^{-1}$]
-    \begin{itemize}
-    \item[$\circ$] data $a,b$ defined on domain $\Omega \subset \RR^2$
+\item ``where are there glaciers?'' is a free-boundary problem
 \end{itemize}
-\item geophysical examples: \alert{glaciers and ice sheets}, sea ice, lakes
-\end{itemize}
-
-\bigskip
-\hfill \mbox{\includegraphics[height=0.25\textheight]{../talk-oxford/images/domain-data.png} \hspace{7mm} \includegraphics[height=0.25\textheight]{../talk-oxford/images/domain-velocity.png}}
-\end{frame}
-
-\begin{frame}{example: glacier ice coverage of the Alps in prior climates}
 
@@
 
-\begin{frame}{naive strong form}
+\begin{frame}{free-boundary problem: flowing glacier ice in a climate}
 
 \begin{itemize}
-\item naive strong form of the steady model: % ($\bn_s = \left<-s_x,-s_y,1\right>$ is surface normal):
-\begin{align*}
-s &\ge b                    & &\text{everywhere in } \Omega \\
--\bu|_s \cdot \bn_s &= a    & &\text{where } s(x,y) > b(x,y)
-\end{align*}
-
+\item glacier = incompressible, viscous fluid driven by gravity
+\item to find: ice surface elevation $s(t,x,y)$ and velocity $\bu(t,x,y,z)$
+\item over fixed bed topography with elevation $b(x,y)$
     \begin{itemize}
-    \item[$\circ$] surface velocity $\bu|_s$ is determined by fluid domain geometry $s$
-    \item[$\circ$] $\bn_s=\left<-\grad s,1\right>$ is upward surface normal
-    \item[$\circ$] generally: $-\bu|_s \cdot \bn_s$ is a \emph{non-local} function of $s$
+    \item[$\circ$] $s(t,x,y) \ge b(x,y)$
     \end{itemize}
-\item the inequality constraint $s \ge b$ \alert{generates a free boundary}
-
-if an ablative climate $a < 0$ forces surface down to bed
+\item in a \emph{climate} which adds or removes ice at a signed rate $a(t,x,y)$
+    \begin{itemize}
+    \item[$\circ$] data $a,b$ is defined on domain $\Omega \subset \RR^2$
+    \end{itemize}
 \end{itemize}
@@
 \bigskip
-\hfill \mbox{\includegraphics[height=0.2\textheight]{../talk-oxford/images/domain-data.png} \hspace{7mm} \includegraphics[height=0.2\textheight]{../talk-oxford/images/domain-velocity.png}}
+\hfill \mbox{\includegraphics[height=0.24\textheight]{../talk-oxford/images/domain-data.png} \quad $\stackrel{?}{\to}$ \quad \includegraphics[height=0.24\textheight]{../talk-oxford/images/domain-velocity.png}}
+\end{frame}
+
+\begin{frame}{glacier free-boundary problem: naive strong form}
+
+\begin{itemize}
+\item is this an adequate description?:
+\begin{align*}
+s &\ge b                    & &\text{everywhere in } \Omega \\
+\frac{\partial s}{\partial t} &= a + \bu|_s \cdot \bn_s & &\text{where } s(t,x,y) > b(x,y)
+\end{align*}
+\item notes:
+    \begin{itemize}
+    \item[$\circ$] surface velocity $\bu|_s$ is, \alert{in some manner}, determined by $s$
+        \begin{itemize}
+        \item $\bu|_s$ is generally a \emph{non-local} function of $s$
+        \end{itemize}
+    \item[$\circ$] $\bn_s=\left<-\grad s,1\right>$ is upward surface normal
+    \end{itemize}
+\end{itemize}
+
+\bigskip
+\hfill \mbox{\includegraphics[height=0.2\textheight]{../talk-oxford/images/domain-data.png} \quad $\stackrel{?}{\to}$ \quad \includegraphics[height=0.2\textheight]{../talk-oxford/images/domain-velocity.png}}
+\end{frame}
+
+\begin{frame}{glacier free-boundary problem: steady VI form}
+
+\begin{itemize}
+\item admissible surface elevations:\footnote{$(s-b)^{8/3} \stackrel{?}{\in} W^{1,4}(\Omega)$ so $\mathcal{V} \stackrel{?}{=} (W^{1,4})^{3/8}$ \dots  see (Jouvet \& Bueler, 2012)}
+    $$\mathcal{K} = \left\{r \in \mathcal{V} \,:\, r \ge b\right\}$$
+\item steady ($\frac{\partial s}{\partial t}=0$) VI problem for surface elevation $s\in\mathcal{K}$:
+	$$\ip{\Phi(s) - a}{r-s} \ge 0 \quad \text{ for all } r \in \mathcal{K}$$
+where
+    $$\Phi(s)=- \bu|_s \cdot \bn_s$$
+with extension by 0 to all of $\Omega$
+%\item $\bu$ is the velocity solution on
+%    $$\Lambda(s) = \{(x,y,z) : b(x,y) < z < s(x,y)\} \subset \RR^3$$
+\end{itemize}
 \end{frame}
@@
 \begin{itemize}
-\item with enough simplifications we may compute the surface motion term by applying a nonlinear elliptic differential operator to $s$:
-    $$\Phi(s) = - \frac{\gamma}{\qq} (s-b)^{\qq} |\grad s|^{\qq} - \grad \cdot\left(\frac{\gamma}{\qq+1} (s-b)^{\qq+1} |\grad s|^{\qq-2} \grad s\right)$$
-    \begin{itemize}
-    \item $\qq = 4$
-    \item $\grad$ is in $x,y$ only
-    \item $\Phi(s)$ is a nonlinear \alert{differential operator} in this model because membrane stresses are \emph{not} balanced
-    \item $\Phi(s)$ is doubly-degenerate
-    \end{itemize}
+\item the \emph{shallow ice approximation} is a highly-simplified view of conservation of momentum
+\item isothermal, nonsliding case:
+\begin{align*}
+\Phi(s) &= - \bu|_s \cdot \bn_s \\
+        &= - \frac{\gamma}{4} (s-b)^{4} |\grad s|^{4} - \grad \cdot\left(\frac{\gamma}{5} (s-b)^{5} |\grad s|^{2} \grad s\right)
+\end{align*}
 \end{itemize}
@@
 
-\begin{frame}{VI for fluid layer in a climate}
+\begin{frame}{FASCD test case: simplified ice sheet the size of Greenland}
 
 \begin{itemize}
-\item admissible surface elevations:
-    $$\mathcal{K} = \left\{r \in \mathcal{V} \,:\, r \ge b\right\}$$
+\item \emph{ice sheet} = big glacier
+\end{itemize}
+
+\bigskip\bigskip
+\centering
+\includegraphics[width=\textwidth]{../paper/fixfigs/sialev8scene.png}
+\end{frame}
+
+\begin{frame}{FASCD: parallel weak scaling}
 
 \begin{itemize}
-    \item[$\circ$] $\mathcal{V}$ to be determined by viscous fluid model\footnote{in shallow ice approximation, $(s-b)^{8/3} \in W^{1,4}(\Omega)$ (Jouvet \& Bueler, 2012)}
+\item observed optimality of FMG solver
+\item good parallel \emph{weak scaling} as well
+    \begin{itemize}
+    \item[$\circ$] each processor owns $641\times 641$ (sub) mesh
+    \item[$\circ$] $P=1024$ run had $20481^2=4.1\times 10^8$ unknowns
+    \item[] \dots and 88 meter resolution
     \end{itemize}
-\item VI problem for surface elevation $s\in\mathcal{K}$:
-	$$\ip{\Phi(s)}{r-s} \ge \ip{a}{r-s} \quad \text{ for all } r \in \mathcal{K}$$
-where
-    $$\Phi(s)=- \bu|_s \cdot \bn_s,$$
-with extension by 0 to all of $\Omega$, and $\bu$ is the velocity solution on
-    $$\Lambda(s) = \{(x,y,z) : b(x,y) < z < s(x,y)\}$$
 \end{itemize}
-\end{frame}
 
-\begin{frame}{computed ice sheet}
-
+\bigskip
 \centering
-\includegraphics[width=\textwidth]{../paper/fixfigs/sialev8scene.png}
-
-FIXME ice sheet the size of Greenland
-\end{frame}
-
-\begin{frame}{weak-scaling}
-
-\centering
-\includegraphics[width=0.7\textwidth]{figs/siaweaktime.png}
-
-FIXME final run had $m_J=20481^2=4.1\times 10^8$ unknowns and a 88 meter grid resolution
+\includegraphics[width=0.55\textwidth]{figs/siaweaktime.png}
 \end{frame}
@@
 \begin{itemize}
-\item FIXME
+\item FASCD = new multilevel solver for VI (free-boundary) problems
+    \begin{itemize}
+    \item[$\circ$] implemented in Python Firedrake (over PETSc)
+    \end{itemize}
+\item observed optimality, even TME, in many cases
+    \begin{itemize}
+    \item[$\circ$] actually fast
+    \end{itemize}
+
+\bigskip\bigskip
+\item[] \textbf{{\color{FireBrick} to do:}}
+    \begin{itemize}
+    \item add mesh adaptivity to free boundary (Stefano)
+    \item implement in C inside PETSc
+    \item apply to space-time (\emph{parabolic}) VI problems
+    \item prove convergence
+    \item identify smoothers for problems like elastic contact
+    \item include membrane stresses in glacier case
+    \end{itemize}
 \end{itemize}
@@
 %{\scriptsize
-{\notsotiny
+%{\notsotiny
+{\footnotesize
 \input{references.tex}
@@
 %{\scriptsize
-{\notsotiny
+%{\notsotiny
+{\footnotesize
 \input{background.tex}

### 2023-12-18 paper/reviews/response.tex
start a responses document

\documentclass[letterpaper,final,12pt,reqno]{amsart}

\usepackage[total={6.3in,9.2in},top=1.1in,left=1.1in]{geometry}

\usepackage{times,bm,bbm,empheq,fancyvrb,graphicx,amsthm,amssymb}
\usepackage[dvipsnames]{xcolor}
\usepackage{longtable}
\usepackage{booktabs}

\usepackage{tabto}
\TabPositions{1.5cm}

\usepackage{float}

% hyperref should be the last package we load
\usepackage[pdftex,
colorlinks=true,
plainpages=false, % only if colorlinks=true
linkcolor=blue,   % ...
citecolor=Red,    % ...
urlcolor=black    % ...
]{hyperref}

\renewcommand{\baselinestretch}{1.05}

\allowdisplaybreaks[1]  % allow display breaks in align environments, if they avoid major underfull

\newcommand{\eps}{\epsilon}

\newcommand{\RR}{\mathbb{R}}
\newcommand{\ZZ}{\mathbb{Z}}

\newcommand{\grad}{\nabla}
\newcommand{\Div}{\nabla\cdot}
\newcommand{\trace}{\operatorname{tr}}

\newcommand{\hbn}{\hat{\mathbf{n}}}

\newcommand{\bb}{\mathbf{b}}
\newcommand{\be}{\mathbf{e}}
\newcommand{\bbf}{\mathbf{f}}
\newcommand{\bg}{\mathbf{g}}
\newcommand{\bn}{\mathbf{n}}
\newcommand{\br}{\mathbf{r}}
\newcommand{\bu}{\mathbf{u}}
\newcommand{\bv}{\mathbf{v}}
\newcommand{\bw}{\mathbf{w}}
\newcommand{\bx}{\mathbf{x}}
\newcommand{\by}{\mathbf{y}}
\newcommand{\bz}{\mathbf{z}}

\newcommand{\bF}{\mathbf{F}}
\newcommand{\bV}{\mathbf{V}}
\newcommand{\bX}{\mathbf{X}}

\newcommand{\bxi}{\bm{\xi}}
\newcommand{\bzero}{\bm{0}}

\newcommand{\cK}{\mathcal{K}}
\newcommand{\cV}{\mathcal{V}}

\newcommand{\rhoi}{\rho_{\text{i}}}

\newcommand{\ip}[2]{\left<#1,#2\right>}

\newcommand{\maxR}{R^{\bm{\oplus}}}
\newcommand{\minR}{R^{\bm{\ominus}}}
\newcommand{\iR}{R^{\bullet}}

\newcommand{\nn}{{\text{n}}}
\newcommand{\pp}{{\text{p}}}
\newcommand{\qq}{{\text{q}}}
\newcommand{\rr}{{\text{r}}}

\newcommand{\supp}{\operatorname{supp}}
\newcommand{\Span}{\operatorname{span}}

\newenvironment{review}%
{\bigskip \par \begin{quote} \selectfont \sl}%
{\end{quote}}

\begin{document}
\title{Response to reviews of \emph{A full approximation scheme multilevel method for nonlinear variational inequalities}}

\author{Ed Bueler}

\author{Patrick Farrell}

\date{\today}

\maketitle

%\tableofcontents

\thispagestyle{empty}
%\bigskip

FIXME The thoughtful comments of the two reviewers are much appreciated.  Below we respond to their concerns---in block italics---and, in the important cases, quote the new or rewritten content of the paper which addresses these concerns.  The aggregate responses to these comments have made the paper about one page longer.

\section{Responses to Referee 1}

%%% ELB CONFIDENTIAL  I am guessing this is Tai himself.  I will look at Tai and Espedal 1998 to see if a citation is appropriate.

\begin{review}
This paper is about using multilevel algorithm to solve nonlinear problems with constraints. Many real-world applications need to solve problem of this kind. Thus, the fast algorithm proposed here is of great importance.

This work proposed to improve the CD (constraint decomposition) method of [32] in several aspects. The new techniques are innovative (see detailed comments below) and show very good performances in numerical experiments report in this paper. I strongly suggest accepting this work for SISC after some minor revisions suggested below.

The reasons that I suggest accepting this paper are based on the following observations:

1. This paper proposed a new way to decompose the constraint set as shown in section 5. This is a good and innovative idea. In the original algorithm of [32], all the decompositions need to go to the finest level. With this new technique, only the nearest level of grid needs to be involved. This is a great advantage.

2. Another advantage of the method proposed in this paper is to use FAS (full approximation Scheme) to solve the subproblems over the different levels, while the original scheme of [32] needs to solve a nonlinear problems over the different grid levels that is projected from the finest mesh and also exactly (even though, approximation solver can be considered as in Tai-Espedal (SINUM 1998). This is making the coarse grid solving much fast and cheaper.

3. This paper proposed to use the algorithm for general monotone nonlinear problems with constraints, while the algorithm proposed in [32] is only for minimization problems. The operator from convex minimization problems is monotone, but there are many monotone operators that are not from minimization problems as shown in Example 2.5 and 2.6 of this work.
\end{review}

\noindent This summary is accurate and appreciated.

\begin{review}
The weak point of this work is that there is no convergence proof for the proposed algorithms. Intuitively, the constraint decomposition given in this paper is much tighter. It would give a better correction during the iterations, but this is not shown theoretically nor numerically.
\end{review}

\noindent xx

\begin{review}
The proposed FMG solver is an optimal solver as claimed by the authors. It also improves CD method of [32] in several aspects. The authors need to supply comparisons between the two methods in the numerical section.
\end{review}

\noindent xx

\begin{review}
Good if the authors can plot the error of the numerical solutions, i.e. $u^J - u^*$ and its convergence order with respect to mesh size $h_J$ with different mesh size $h_J$. This observed convergence order could be beneficial for people working with numerical analysis.
\end{review}

\noindent xx

\section{Responses to Referee 2}

\begin{review}
This paper merges the FAS multigrid method, which is a highly optimal solver, with a particular optimization method (CD).  They are able to get close to optimal work complexity this class of optimization solvers, which is a significant accomplishment.  The paper is clearly written and demonstrates good performance.
\end{review}

\noindent This summary is accurate and appreciated.

\begin{review}
Abstract: ``FASCD is a common extension of \dots'' this sounds like FASCD is, well, common, yet this seems to be the first paper that develops this method.  I think you are saying others have done the same, but it reads like FASCD is common.  Maybe something like ``while many multilevel VI solvers have been developed, FASCD has unique desirable properties''.
\end{review}

\noindent xx

\begin{review}
The prior work section starting at line 40, is hard to follow.  There are three projects that you comparing to.  There are many qualities discussed.  A table or some graphic might be useful and some signposts like ``three MG VI methods have been developed in prior works [6,24,22]'' and a paragraph for each.
\end{review}

\noindent xx

\begin{review}
You might move all discussion of desirable properties to the beginning at line 25. and refer to numbered properties in the prior work section as far as what other do well and don't do well.
\end{review}

\noindent xx

\begin{review}
The FAS treatment, while good, is a little confusing.  Perhaps only to an expert. FMG is not an iterative method, yet you start with ``FASCD \dots by a small, mesh-independent number of FMG iterations \dots''.
\end{review}

\noindent xx

\begin{review}
The FAS algorithm with V-cycles attached to make an iterative method to arbitrary solver tolerance is fine, but reporting multiple iteration is not really correct.  Right?  It is FMG + k V-cycle iterates.  Maybe report ``extra V-cycles in FASCD'', which should be 0, but you ``over solve'' to compare with others.  This is subtle and maybe could be put in the caption of Table 2 to reduce the text for the column heading from what is now ``FMG'' (``FMG V-cycles'' maybe).
\end{review}

\noindent xx

\begin{review}
This issue becomes a real problem in Table 5 where you devide the solver time by what you call the FASCD iterations.  This does not make sense. I would suggest, since you are just looking at time per ``iteration'', that you just do full MG and forget about iterating.
\end{review}

\noindent xx

\begin{review}
FASCD is different in that you have to do strange things with the constraints in the down smoothing, so it is not surprising that V(1,0) fails (table 3) and V(1,1) is WORSE that V(0,1) for on test, which indicates a certain fragility with the down smoothing process. It seems that should be discussed, but I don't see this.
\end{review}

\noindent xx

\end{document}

### 2023-12-19 paper/reviews/response.tex
little bits of responses as I go through our text with a green pen

@@
 
+\newcommand\short[1]{\medskip\noindent #1}   % short form of response
+
+\newenvironment{response}% long form of response
+{\medskip\noindent}%
+{}
+
 \begin{document}
-\title{Response to reviews of \emph{A full approximation scheme multilevel method for nonlinear variational inequalities}}
+\title[Response to reviews]{Response to reviews of \\ \emph{A full approximation scheme multilevel method \\ for nonlinear variational inequalities}}
 
@@
 
-\noindent This summary is accurate and appreciated.
+\short{FIXME This summary is accurate and appreciated.}
 
@@
 
-\noindent xx
+\begin{response}
+FIXME good point
+\end{response}
 
@@
 
-\noindent xx
+\short{xx}
 
@@
 
-\noindent xx
+\short{FIXME ``common'' $\to$ ``joint''}
 
@@
 
-\noindent xx
+\short{FIXME yes, indeed four projects (which makes the reviewers point). cut into paragraphs}
 
@@
 
-\noindent xx
+\short{FIXME appreciated.  mostly kept the same, but we think the greater organization of prior work paragraphs makes referring to numbered properties unneeded}
 
 \begin{review}
-The FAS treatment, while good, is a little confusing.
-Perhaps only to an expert.
+The FAS treatment, while good, is a little confusing.  Perhaps only to an expert.
+\end{review}
+
+\short{xx}
+
+\begin{review}
 FMG is not an iterative method, yet you start with ``FASCD \dots by a small, mesh-independent number of FMG iterations \dots''.

### 2023-12-23 paper/genfigs/asymprates.py
Figure 6 now generated from fascd/reproduce/table2.py

@@ -1,57 +1,0 @@
-#!/usr/bin/env python3
-
-import numpy as np
-import matplotlib.pyplot as plt
-from writeout import writeout
-
-# from V-cycle results, mcd-extended/paper/results/drape_fascd.txt
-# line J:  [k, |r_ss(w^0)|, |r_ss(w^k)|]
-resnormball = [[ 1, 2.977687712552e+00, 5.592593521511e-16],  # J=0, single level
-               [ 3, 4.597248896467e+00, 9.299029163684e-15],  # J=1, two-level
-               [ 6, 4.019378820472e+00, 3.548401366476e-14],  # ...
-               [ 7, 3.518652347563e+00, 4.898369539597e-13],
-               [ 9, 3.989265817522e+00, 5.346032356444e-13],
-               [11, 5.149110069199e+00, 1.683050131619e-12],
-               [11, 7.072193804686e+00, 6.219113662073e-12],
-               [12, 9.929467647243e+00, 6.630087376663e-12],
-               [13, 1.401524016881e+01, 5.806872030903e-12],
-               [13, 1.981030784949e+01, 1.876069712594e-11],
-               [14, 2.801197654214e+01, 1.004260356580e-11]]
-resnormspir = [[ 1, 1.250364911422e+01, 1.441420557646e-15],
-               [ 4, 3.581030295829e+01, 1.356576646565e-14],
-               [ 5, 5.195274377477e+01, 2.683842615342e-11],
-               [ 7, 7.324445644144e+01, 8.422103502963e-12],
-               [ 9, 1.069563682455e+02, 1.703081752902e-11],
-               [10, 1.517759488345e+02, 6.241169671924e-11],
-               [11, 2.119960943342e+02, 6.428954633272e-11],
-               [12, 3.016400256848e+02, 1.293809842620e-10],
-               [13, 4.275614401535e+02, 1.177739303183e-10],
-               [14, 6.031903763444e+02, 1.310547032348e-10],
-               [18, 8.529374614123e+02, 5.472658274103e-10]]
-
-plt.rcParams["font.family"] = "serif"
-plt.rcParams["mathtext.fontset"] = "cm"
-
-fsize = 20.0
-
-plt.figure(figsize=(8, 6))
[…]

### 2023-12-23 paper/results/drape_fascd.sh
all of the results are now reported or generated in reproduce/ of git@bitbucket.org:pefarrell/fascd.git

@@ -1,19 +1,0 @@
-# generate FASCD results for Example 8.1 in paper
-# warning: the level 10,11 runs assume decent memory on you machine
-# note: these runs are serial so as to fix the smoother
-
-for PROB in ball spiral; do
-    echo ""
-    echo "******** PROB=${PROB}, V-cycles, tight tolerance (over-solving) *******"
-    for LEVS in 1 2 3 4 5 6 7 8 9 10 11; do
-        python3 drape.py -prob $PROB -levs $LEVS -fascd_cycle_type multiplicative -fascd_monitor -fascd_atol 1.0e-12 -fascd_rtol 1.0e-12 -fascd_stol 1.0e-12
-    done
-done
-
-for PROB in ball spiral; do
-    echo ""
-    echo "******** PROB=${PROB}, F-cycles, to discretization error, with defaults rtol=stol=1.0e-8 and atol=1.0e-50 *******"
-    for LEVS in 1 2 3 4 5 6 7 8 9 10 11; do
-        python3 drape.py -prob $PROB -levs $LEVS -fascd_cycle_type full -fascd_monitor
-    done
-done

### 2023-12-23 paper/results/drape_nofas.sh

@@ -1,9 +1,0 @@
-# generate -nofas results for Example 8.1 in paper
-
-for PROB in ball spiral; do
-    echo ""
-    echo "******** -nofas PROB=${PROB} *******"
-    for LEVS in 1 2 3 4 5 6 7 8; do
-        python3 drape.py -prob $PROB -levs $LEVS -nofas -snes_rtol 1.0e-12
-    done
-done

### 2023-12-23 paper/results/plap1d.m

@@ -1,11 +1,0 @@
-% generate convergence rates from results in plap1d.txt
-% errors are from default V(1,1) runs
-
-h = 2.^(-(0:9));
-err = [2.263e-01 3.255e-02 9.108e-03 3.248e-03 5.501e-04 ...
-       1.684e-04 4.656e-05 9.203e-06 3.430e-06 4.139e-07];
-loglog(h,err,'*')
-grid on,  xlabel h,  ylabel('|error|_{inf}')
-legend('p=1.5')
-p = polyfit(log(h),log(err),1)
-%   p = 2.0047  -1.7579

### 2023-12-23 paper/results/plap1d.sh

@@ -1,24 +1,0 @@
-# generate results for Example 8.2 in paper
-# run as
-#   $ bash PATHTOHERE/plap1d.sh &> plap1d.txt
-# from git@bitbucket.org:pefarrell/fascd.gitexamples/
-# for convergence rates, run plap1d.m in octave
-
-LLIST="1 2 3 4 5 6 7 8 9 10"
-
-# do: V(1,0), V(0,1), V(1,1)
-for VOPT in "-fascd_skip_upsmooth" "-fascd_skip_downsmooth" ""; do
-    echo "******** p = 1.5 V-cycle VOPT=${VOPT} *******"
-    for LEVS in $LLIST; do
-        python3 plap1d.py $VOPT -p 1.5 -fascd_stol 0.0 -levs $LEVS
-    done
-    echo ""
-done
-
-for EPSREG in 0.0 1.0e-8; do
-    echo "******** p = 1.5 F-cycle EPSREG=${EPSREG} *******"
-    for LEVS in $LLIST; do
-        python3 plap1d.py -fascd_cycle_type full -epsreg $EPSREG -p 1.5 -fascd_stol 0.0 -levs $LEVS
-    done
-    echo ""
-done

### 2023-12-23 paper/results/pollutant.sh

@@ -1,21 +1,0 @@
-# generate results for Example 8.3 in paper
-# run as
-#   $ bash PATHTOHERE/pollutant.sh &> pollutant.txt
-# from git@bitbucket.org:pefarrell/fascd.git in examples/
-
-LIST2D="1 2 3 4 5 6 7 8 9"
-LIST3D="1 2 3 4 5"  # level 5 almost fills 128GB memory machine
-
-echo "******** -dim 2: serial, P1 *******"
-for LEVS in $LIST2D; do
-    python3 pollutant.py -dim 2 -levs $LEVS
-done
-echo ""
-
-# which mpiexec
-echo "******** -dim 3: P = 16 parallel, Q1 *******"
-for LEVS in $LIST3D; do
-    mpiexec -n 16 -map-by numa -bind-to core python3 pollutant.py -dim 3 -q1 -levs $LEVS
-done
-echo ""
-

### 2023-12-23 paper/results/siaweak.sh

@@ -1,12 +1,0 @@
-# generate weak-scaling results for SIA Example 8.4 in paper
-# uncomment later lines if you are on a big machine
-# siaweak.txt is result from this script; see archer2/ for large machine logs
-
-MPI="mpiexec -map-by numa -bind-to core"
-OPTS="-bumps -mcoarse 20 -fascd_cycle_type full -fascd_monitor -fascd_rtol 1.0e-4 -fascd_atol 1.0e-8 -quad"
-time $MPI -n  1 python3 sia.py $OPTS -levs 6
-time $MPI -n  4 python3 sia.py $OPTS -levs 7
-time $MPI -n 16 python3 sia.py $OPTS -levs 8
-#time $MPI -n 64 python3 sia.py $OPTS -levs 9
-#time $MPI -n 256 python3 sia.py $OPTS -levs 10
-#time $MPI -n 1024 python3 sia.py $OPTS -levs 11

### 2023-12-23 paper/fascd.tex
Figure 6 now generated from fascd/reproduce/table2.py
edits through end Section 1

@@
 We present the \emph{full approximation scheme constraint decomposition} (FASCD) multilevel method for solving variational inequalities (VIs).
-FASCD is a common extension of both the full approximation scheme (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and the constraint decomposition (CD) method introduced by X.-C.~Tai for VIs arising in optimization.
-We extend the CD idea by exploiting the telescoping nature of certain function space subset decompositions arising from multilevel mesh hierarchies.
-When a reduced-space (active set) Newton method is applied as a smoother, with work proportional to the number of unknowns on a given mesh level, FASCD V-cycles exhibit nearly mesh-independent convergence rates, and full multigrid cycles are optimal solvers.
+FASCD is a joint extension of both the full approximation scheme (FAS) multigrid technique for nonlinear partial differential equations, due to A.~Brandt, and the constraint decomposition (CD) method introduced by X.-C.~Tai for VIs arising in optimization.
+We extend the CD idea by exploiting the telescoping nature of certain subset decompositions arising from multilevel mesh hierarchies.
+When a reduced-space (active set) Newton method is applied as a smoother, with work proportional to the number of unknowns on a given mesh level, FASCD V-cycles exhibit nearly mesh-independent convergence rates.
+The full multigrid cycle version is an optimal solver.
 The example problems include differential operators which are symmetric linear, nonsymmetric linear, and nonlinear, in unilateral and bilateral VI problems.
@@
 In a finite element (FE) context, these methods converge for multilevel and domain decompositions.
-The multilevel CD method has almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}, \cite[Theorem 4.6]{GraeserKornhuber2009}.
+The multilevel CD method has almost grid-independent error bounds, and near-optimal complexity, for elliptic, linear obstacle problems \cite[Subsection 5.4]{Tai2003}, \cite[Theorem 4.6]{GraeserKornhuber2009}, but returns to the finest mesh are required when evaluating functions on coarse levels.
 
@@
 \item The FAS coarse corrections are adapted for general VI problems. The method requires neither an objective function nor linearity of the residual.
-\item Box constraints, i.e.~both upper and lower obstacles, are allowed.
+\item Box constraints, i.e.~both upper and lower obstacles, are allowed.  (However, other convex admissible sets are not supported.)
 \item The method is strictly admissible, and so can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
-\item V-cycles (Algorithm \ref{alg:fascd}) and full multigrid (FMG; \cite{Trottenbergetal2001}) cycles (Algorithm \ref{alg:fascd-fmg}) are proposed.
-\item We identify a ``telescoping'' subset property which can be exploited on the up-smoothing side of a V-cycle for greater performance.
-\item The method is (nonlinear) smoother-agnostic.
-When a (mesh-dependent) reduced-space Newton method \cite{BensonMunson2006} is used as relaxation, we demonstrate that application of FMG is an optimal solver for the examples in Section \ref{sec:results}.
+\item The work at each level is proportional to the number of nodes on that level.  Coarser-level evaluations do not require finest-level information.
+\item V-cycles (Algorithm \ref{alg:fascd}) and full multigrid (FMG; \cite{Trottenbergetal2001}) cycles (Algorithm \ref{alg:fascd-fmg}) are implemented.
+\item We identify a ``telescoping'' subset property, which is exploited on the up-smoothing side of a V-cycle, yielding better performance.
+\item The method is smoother-agnostic.
+\item Using a (mesh-dependent) reduced-space Newton method \cite{BensonMunson2006} as the smoother, in Section \ref{sec:results} we demonstrate that application of FMG is an optimal solver.
 \end{enumerate}
 
-Prior multilevel algorithms for VIs have limitations compared to the FASCD framework.
-The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} applies to linear complementarity problems and it uses projected (truncated) Gauss--Seidel smoothers.
-While FASCD reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed, PFAS reduces to linear multigrid.
-Another work which combines FAS and VIs is \cite{HintermullerKopacka2011}, but the problems solved, and the approach, are completely different.
-For an optimal control problem with an elliptic VI constraint, \cite{HintermullerKopacka2011} applies FAS to a system of PDEs derived from penalizing the VI and the control constraints; their V- and W-cycles are not altered in any manner to maintain admissibility.
-On the other hand, the multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function, so as stated it only applies to optimization-type problems, and furthermore each of its coarser-level problems refers to the finest-level discretization.
-The latter limitation is removed by Gr\"aser \& Kornhuber's \cite[Algorithm 4.2]{GraeserKornhuber2009} reformulation using defect constraints and monotone injection operators; see Section \ref{sec:cdmultilevel} below.
-The monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be formulated only with Gauss--Seidel type smoothers, and may have been applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.
+Prior multilevel algorithms for VIs have significant limitations compared to the FASCD framework.  Listed chronologically, here are some comparisons:
+\begin{itemize}
+\item The projected full approximation scheme (PFAS) algorithm of Brandt \& Cryer \cite{BrandtCryer1983} applies to linear complementarity problems and it uses projected (truncated) Gauss--Seidel smoothers.
+While our method reduces to general, nonlinear FAS for PDEs when the inequality constraints are removed, PFAS reduces to linear multigrid.
+\item The multilevel CD iteration of Tai \cite{Tai2003} is formulated in terms of an objective function; as stated it only applies to optimization-type problems.
+Furthermore, each of its coarser-level problems refers to the finest-level discretization.
+\item The last-mentioned limitation of the \cite{Tai2003} method is removed by Gr\"aser \& Kornhuber's \cite[Algorithm 4.2]{GraeserKornhuber2009} reformulation using defect constraints and monotone injection operators.
+We adopt these techniques; see Section \ref{sec:cdmultilevel} below.
+However, the monotone multigrid methods of Kornhuber and co-authors \cite{GraeserKornhuber2009,Kornhuber1994} appear to be formulated only with Gauss--Seidel type smoothers, and may have been applied only to optimization-type VI problems \cite[for example]{JouvetGraeser2013}.
 Furthermore, unlike the fastest variants of monotone multigrid, no zeroing of the basis functions associated with the active set (truncation) is employed in FASCD, and this simplifies the implementation.
-Nonetheless FASCD coarse grid corrections can update the estimate of the active set on the fine grid, and indeed the numerical evidence indicates that FASCD coarse corrections are effective even when the coarse meshes are far too coarse to resolve the active set.
-On a problem with a complicated coincidence set, for which the truncated Newton/monotone multigrid methods were apparently designed \cite[problem 7.1.1]{GraeserKornhuber2009}, the performance of FASCD is at least as good (Example \ref{ex:results:classical}).
-A limitation of FASCD is that, as formulated, it only solves bound-constrained VI problems.
+FASCD coarse grid corrections can indeed update the finest-level active set, and the numerical evidence indicates that FASCD coarse corrections are effective even when the coarse meshes are far too coarse to resolve the active set.
+On a problem with a complicated coincidence set, for which truncated Newton/monotone multigrid methods were apparently designed \cite[problem 7.1.1]{GraeserKornhuber2009}, the performance of FASCD is at least as good (Example \ref{ex:results:classical}).
+\item Another work which combines FAS and VIs is \cite{HintermullerKopacka2011}, but the problems solved, and the approach, are completely different.
+For an optimal control problem with an elliptic VI constraint, \cite{HintermullerKopacka2011} applies FAS to a system of PDEs derived from penalizing the VI and the control constraints; their FAS algorithm is not applied to the VI itself while ours is.
+\end{itemize}
 
 In our experiments, FASCD V-cycles exhibit essentially the same mesh-independent convergence rates for VI problems that FAS multigrid V-cycles exhibit for the corresponding unconstrained PDE problems.
-The FMG results in Section \ref{sec:results} show near-textbook multigrid efficiency, yielding solutions within discretization error in work comparable to a few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}.
-Specifically, we see this performance for classical (unilateral, Laplacian) obstacle problems, $p$-Laplacian obstacle problems, and box-constrained advection-diffusion equations.
+In fact, the FMG results in Section \ref{sec:results} show near-textbook multigrid efficiency, yielding solutions within discretization error in work comparable to a relatively few smoother sweeps on the finest-level mesh \cite{BrandtLivne2011}.
+We see this performance for classical (unilateral, Laplacian) obstacle problems, $p$-Laplacian obstacle problems, and box-constrained advection-diffusion equations.
 Furthermore, V-cycle results are largely insensitive to the geometric complexity of the coincidence set and free boundary.
@@
 All iterates in FASCD, on all mesh levels, are admissible, and thus the nonlinear operator in the VI needs only be defined for admissible states.
-For example, we solve an ice sheet geometry problem (Examples \ref{ex:sia} and \ref{ex:results:sia}) in which the ice flow formulas are only meaningful for surface elevation iterates that do not penetrate the bedrock.
-(Non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural modifications of the operator formula.)  FASCD solves this low-regularity ice sheet problem, wherein the solution has unbounded gradient at the free boundary, by a small, mesh-independent number of FMG iterations (Example \ref{ex:results:sia}).
-The same Example investigates the parallel weak scaling \cite{Bueler2021} of our implementation.
+For example, we solve an ice sheet geometry problem (Example \ref{ex:sia}) in which the flow formulas are only meaningful for surface elevation iterates that do not penetrate the bedrock.
+(Non-admissible methods, including semi-smooth methods \cite{BensonMunson2006}, would require unnatural modifications of the operator formula.)  The FASCD FMG cycle solves this low-regularity problem, in which the solution gradient is unbounded at the free boundary, optimally; see Example \ref{ex:results:sia}.
+Our results on this Example also demonstrate the parallel weak scaling \cite{Bueler2021} of our implementation.
 
@@
 In Section \ref{sec:vi} we recall the theory of coercive and monotone VIs, and describe several motivating examples.
-Section \ref{sec:cd} reviews CDs and the iterations they induce.
-Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, and then Section \ref{sec:cdmultilevel} extends the CD method via defect constraints generated by monotone injection operators \cite{GraeserKornhuber2009}.
-Here we propose an apparently new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD and an associated iteration.
-Section \ref{sec:vcycle} then states the V-cycle, with Section \ref{sec:implementation} addressing convergence criteria, FMG cycles, and our smoother choice.
+Section \ref{sec:cd} reviews CDs and their induced iterations.
+Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, and then Section \ref{sec:cdmultilevel} implements the CD method using defect constraints which are generated by monotone injection \cite{GraeserKornhuber2009}.
+In Section \ref{sec:cdmultilevel} we also propose a new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD and an associated iteration.
+Section \ref{sec:vcycle} then states the FASCD V-cycle, and Section \ref{sec:implementation} addresses practical implementation including convergence criteria and the FMG cycle.
 Section \ref{sec:results} gives numerical results from our open source Python/Firedrake \cite{Rathgeberetal2016} implementation.
@@
 
-%REGENERATE using results/asymprates.py
+%REGENERATE using reproduce/table2.py at git@bitbucket.org:pefarrell/fascd.git
 \begin{figure}[ht]

### 2023-12-24 paper/fascd.tex
edits and clean up through most of section 2

@@
 Section \ref{sec:cd} reviews CDs and their induced iterations.
-Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, and then Section \ref{sec:cdmultilevel} implements the CD method using defect constraints which are generated by monotone injection \cite{GraeserKornhuber2009}.
+Section \ref{sec:femultilevel} sets up multilevel FE hierarchies, and then Section \ref{sec:cdmultilevel} implements the CD method using defect constraints which are generated by monotone injection.
 In Section \ref{sec:cdmultilevel} we also propose a new understanding of telescoping sets for up-smoothing, based on an ``incomplete'' CD and an associated iteration.
@@
 \end{equation}
-Because $f$ is (generally) nonlinear, the source term $\ell$ is not strictly needed in stating this class of problems---by redefining $f$ we may take $\ell=0$---but introducing $\ell$ clarifies the algorithms which follow.
-
-VI \eqref{eq:vi} generalizes nonlinear systems of equations $f(u^\star)=\ell$ to problems where $u^\star$ is also constrained to be in $\cK$.
-Pretending the dual pairing is an inner product, by \eqref{eq:vi} the angle between $f(u^\star)-\ell$ and any arbitrary vector $v-u^\star$ pointing from $u^\star$ into $\cK$ is at most $90^\circ$; the dual vector $f(u^\star)-\ell$ must point into $\cK$.
-If $u^\star$ is in the interior of the constraint set then \eqref{eq:vi} implies $f(u^\star)=\ell$.
-Variational inequalities may generally permit multiple solutions~\cite{Farrell2019}.
-
-The following definitions are standard \cite{KinderlehrerStampacchia1980}.
+Because $f$ is (generally) nonlinear, the source term $\ell$ is not strictly needed in stating this class of problems---by redefining $f$ we may take $\ell=0$---but introducing $\ell$ will help to clarify later algorithms.
+
+VI \eqref{eq:vi} generalizes nonlinear systems of equations (i.e.~$f(u^\star)=\ell$) to cases where $u^\star$ is also constrained to be in $\cK$.
+Pretending the dual pairing is an inner product, \eqref{eq:vi} says that the angle between $f(u^\star)-\ell$ and any arbitrary vector $v-u^\star$ pointing from $u^\star$ into $\cK$ is at most $90^\circ$; if $u^\star$ solves \eqref{eq:vi} then the dual vector $f(u^\star)-\ell$ points straight into $\cK$.
+If $u^\star$ is in the interior of $\cK$ then \eqref{eq:vi} implies $f(u^\star)=\ell$.
+Note that variational inequalities generally permit multiple solutions~\cite{Farrell2019}.
+
+The following definitions are standard \cite[for example]{KinderlehrerStampacchia1980}.
 
@@
 It is well-known that if $f:\cK \to \cV'$ is continuous, monotone, and coercive then VI \eqref{eq:vi} has a solution \cite[Corollary III.1.8]{KinderlehrerStampacchia1980}, and that the solution is unique when $f$ is strictly monotone.
-As is standard in the calculus of variations \cite{Evans2010}, coercivity permits a compactness argument for unbounded sets $\cK$; recall that the closed and bounded subsets of a reflexive Banach space are weakly compact.
-Some VIs solved in this paper satisfy a stronger inequality than \eqref{eq:coercive}.
-
-\begin{definition}  Let $p>1$.  The map $f:\cK \to \cV'$ is \emph{$p$-coercive} if there exists $\kappa>0$ such that
+As in the calculus of variations, coercivity permits a compactness argument for unbounded sets $\cK$ using the fact that closed and bounded subsets of a reflexive Banach space are weakly compact.
+
+Some operators considered below satisfy a stronger inequality than \eqref{eq:coercive}.
+
+\begin{definition}  Let $p>1$.  The operator $f:\cK \to \cV'$ is \emph{$p$-coercive} if there exists $\kappa>0$ such that
 \begin{equation}
@@
 It is easy to see that if $f$ is $p$-coercive then it is monotone, strictly monotone, and coercive.
-Thus if $f:\cK \to \cV'$ is continuous and $p$-coercive then there exists a unique $u^\star\in \cK$ solving VI \eqref{eq:vi}.
+Thus if $f:\cK \to \cV'$ is continuous and $p$-coercive ($p>1$) then there exists a unique $u^\star\in \cK$ solving VI \eqref{eq:vi}.
 
 Let $\Omega \subset \RR^d$ denote a bounded, open set with piecewise-smooth boundary.
-Sobolev spaces \cite{Evans2010} will be denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$, with $W^{1,p}_0(\Omega)$ denoting the subspace with zero trace.
+Sobolev spaces \cite{Evans2010} will be denoted by $W^{k,p}(\Omega)$, for integer $k$ and $1\le p \le \infty$, with $W^{1,p}_0(\Omega)$ denoting the subspace with zero trace along $\partial\Omega$.
 The following example includes the classical and $p$-Laplacian obstacle problems \cite{ChoeLewis1991}.
@@
 \begin{equation}
-\ip{f(u)}{v} = \int_\Omega |\grad u|^{p-2} \grad u \cdot \grad v\dx, \label{eq:plaplacian}
-\end{equation}
-a continuous map \cite[Theorem A.0.6]{Peral1997}.
-It follows from inequalities for $|\cdot|^p$ in $\RR^d$, and the Poincar\'e inequality, that $f$ is $p$-coercive \cite{ChoeLewis1991}; \cite[Appendix A]{Bueler2021conservation}.
+\ip{f(u)}{v} = \int_\Omega |\grad u|^{p-2} \grad u \cdot \grad v\dx. \label{eq:plaplacian}
+\end{equation}
+It follows from inequalities for $|\cdot|^p$ in $\RR^d$, and the Poincar\'e inequality, that $f$ is $p$-coercive \cite{ChoeLewis1991}; see also \cite[Appendix A]{Bueler2021conservation}.
 \end{example}
 
-For monotone operators $f$, VI \eqref{eq:vi} generalizes the problem of minimizing a convex function over $\cK$.
+Note that for monotone operators VI \eqref{eq:vi} also generalizes the problem of minimizing a convex function over $\cK$.
 Suppose $F:\cK \to \RR$ is lower semi-continuous and (G\^ateaux) differentiable with continuous derivative $F':\cK \to \cV'$.
-Then $F$ is convex if and only if $F'$ is monotone \cite[Proposition I.5.5]{EkelandTemam1976}.
-Furthermore, Proposition II.2.1 in \cite{EkelandTemam1976} shows that if $F$ is convex then \eqref{eq:vi} holds for $f=F'$ if and only if
+Then $F$ is convex if and only if $f=F'$ is monotone, in which case \eqref{eq:vi} holds if and only if
 \begin{equation}
@@
 \end{equation}
-The CD methods of Tai \cite{Tai2003} address problem \eqref{eq:minimization}; the analysis in \cite{Tai2003} supposes that an objective $F$ exists with $F'$ 2-coercive.
-For Example \ref{ex:plaplacian} we may define
-\begin{equation}
-F(v) = \frac{1}{p} \int_\Omega |\grad v|^p\dx. \label{eq:plaplacianobjective}
-\end{equation}
-Then $F$ is a convex functional and $F'(v) = f(v)$ for $f$ given in \eqref{eq:plaplacian}.
-For any closed and convex $\cK\subset \cV$, VI problem \eqref{eq:vi} is then equivalent to optimization problem \eqref{eq:minimization}.
+(See e.g.~Propositions I.5.5 and II.2.1 in \cite{EkelandTemam1976} for these facts.)  The CD methods of Tai \cite{Tai2003} address problem \eqref{eq:minimization}; his analysis supposes that an objective $F$ exists with $F'$ 2-coercive.
+In Example \ref{ex:plaplacian}
+\begin{equation}
+F(v) = \frac{1}{p} \int_\Omega |\grad v|^p\dx \label{eq:plaplacianobjective}
+\end{equation}
+is a convex functional and $F' = f$ is given by \eqref{eq:plaplacian}.
 
@@
 Let $\cV \subset W^{1,2}(\Omega)$ be the space of functions which have zero trace on $\partial_D\Omega$.
-Consider a velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, and assume that the flow is outward on the Neumann boundary: $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.
+Consider a velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, and also assume that the flow is outward on the Neumann boundary: $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.
 For $u,v \in \cV$, $\eps>0$, and $\phi \in \cV'$, define
@@
 \end{equation}
-With this operator, consider VI \eqref{eq:vi} with $\ell = 0$ for any closed and convex $\cK \subset \cV$. The interior condition in strong form is the following linear advection-diffusion equation
+With this operator, consider VI \eqref{eq:vi} with $\ell = 0$ for any closed and convex $\cK \subset \cV$.
+The interior condition \cite{KinderlehrerStampacchia1980} of \eqref{eq:vi} is the following linear advection-diffusion equation:
 \begin{equation}
@@
 \end{equation}
-On the other hand, it is easy to see that $|\!\ip{f(u)}{v}\!| \le \left((\eps + \|\bX\|_\infty) \|u\| + \|\phi\|'\right) \|v\|$, so that $f:\cK \to \cV'$ is continuous.
+On the other hand, it is easy to see that $|\!\ip{f(u)}{v}\!| \le \left((\eps + \|\bX\|_\infty) \|u\| + \|\phi\|'\right) \|v\|$, so $f:\cK \to \cV'$ is continuous.
 Lemma \ref{lem:advectionskew} says that the bilinear form $b(u,v)$ is skew-symmetric up to a term which is nonnegative by the outward flow assumption.
@@
 \end{align*}
-Thus $f$ is 2-coercive, and so the associated VI problem is well-posed.
-If $\bX \ne 0$ then $f\ne F'$ for any objective $F$, because $\ip{f(u)}{v}$ does not possess the symmetry of a gradient: for $u,v$ which are zero on $\partial \Omega$, note that $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.
-References \cite{Bueler2021conservation,ChangNakshatrala2017} consider similar advection-diffusion VI problems over $\cK = \{v\ge 0\}$.
-\end{example}
+Thus $f$ is 2-coercive and the associated VI problem is well-posed.
+If $\bX \ne 0$ then $f\ne F'$ for any objective $F$, because $\ip{f(u)}{v}$ does not possess the symmetry of a gradient: for $u,v$ which are zero on $\partial \Omega$, note that $\ip{f(u)}{v} - \ip{f(v)}{u} = -2 b(u,v)$.\end{example}
+
+References \cite{Bueler2021conservation,ChangNakshatrala2017} consider similar advection-diffusion VI problems to Example \ref{ex:advectiondiffusion}, over $\cK = \{v\ge 0\}$.
 
@@
 It is fully nonlinear and not of optimization type.
-Related VI problems arise for any fluid layer subject to processes which can remove fluid mass \cite{Bueler2021conservation}.
+Related VI problems arise for any fluid layer subject to local surface processes which can remove fluid mass \cite{Bueler2021conservation}.
 

### 2023-12-25 paper/fascd.tex
editing progress

@@
 
-\begin{example}  \label{ex:sia}  Let $\Omega \subset \RR^2$ be a fixed region of land, and suppose $b \in C^1(\Omega)$ is a given bedrock elevation.
+\begin{example}  \label{ex:sia}  Let $\Omega \subset \RR^2$ be a fixed region of land, and suppose $b \in C^1(\Omega)$ is the bedrock elevation on $\Omega$.
 Let $a \in L^2(\Omega)$ denote a given ``surface mass balance'' function, the annually-averaged rate of ice accumulation (snow) minus melt and runoff.
@@
 \begin{equation}
-\int_\Omega \Gamma (s-b)^{n+2} |\grad s|^{n-1} \grad s \cdot \grad (v-s) \dx \ge \int_\Omega a (v-s)\dx \quad \text{ for all } v \in \mathcal{K}, \label{eq:siavi}
-\end{equation}
-where $n>1$ and $\Gamma>0$ are physical constants for ice; $n=3$ is a typical value \cite{GreveBlatter2009}.  The doubly-nonlinear and doubly-degenerate operator here is
+\int_\Omega \Gamma (s-b)^{n+2} |\grad s|^{n-1} \grad s \cdot \grad (v-s) \dx \ge \int_\Omega a (v-s)\dx \quad \text{ for all } v \in \mathcal{K}. \label{eq:siavi}
+\end{equation}
+Here $n>1$ and $\Gamma>0$ are physical constants for ice; $n=3$ is a typical exponent \cite{GreveBlatter2009}.  The doubly-nonlinear and doubly-degenerate operator here is
 \begin{equation}
@@
 If the bed is flat ($b=0$) then a power transformation converts \eqref{eq:siaoperator} into form \eqref{eq:plaplacian} with $p=n+1$.
-Less is understood for general beds $b$, but existence holds for the VI \cite{JouvetBueler2012}, and a power of the solution lives in a Sobolev space: $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.
-This theory, and also observations of ice sheets, shows that $|\grad s|$ is usually unbounded as one approaches the free boundary from the icy ($s>b$) side.
+Less is understood for general smooth beds $b$, but existence holds for \eqref{eq:siavi} \cite{JouvetBueler2012}, and a power of a solution will live in a Sobolev space: $(s-b)^{2p/(p-1)} \in W^{1,p}(\Omega)$.
+This theory, and also observations of ice sheets, shows that $|\grad s|$ is often unbounded as one approaches the free boundary from the icy ($s>b$) side.
 \end{example}
 
-The numerical results in Section \ref{sec:results} will apply FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fmg} to the above Examples.
+The numerical results in Section \ref{sec:results} come from applying FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fmg} to the above Examples.
 
@@
 We say \eqref{eq:subspacedecomp} is a \emph{space decomposition} of $\cV$ \cite{Xu1992}.
-For a closed, convex subset $\cK \subset \cV$, suppose further that $\cK_i \subset \cV_i$ are nonempty, closed, and convex subsets such that
+For a closed, convex subset $\cK \subset \cV$, suppose that $\cK_i \subset \cV_i$ are nonempty, closed, and convex subsets such that
 \begin{equation}
@@
 A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:decompositionstability} \cite{Tai2003}.
-Figure \ref{fig:cartoon} suggests how a low-dimensional example might look; note $\cK_i \not\subset \cK$ in this and many other cases.
+Figure \ref{fig:cartoon} suggests how a low-dimensional example might look; note $\cK_i \not\subset \cK$ in this and most other cases.
 
@@
 
-We will construct a multilevel FE CD discretization in Sections \ref{sec:femultilevel}--\ref{sec:implementation}.
-However, the CD concept can be applied at the level of the continuum problem.
+We will construct a finite-dimensional multilevel CD in Sections \ref{sec:femultilevel}--\ref{sec:implementation}.
+However, the CD concept can also be applied at the level of the continuum problem.
 For example, in a unilateral obstacle problem ($\cK = \{v \ge \psi\}$) one can construct an overlapping domain decomposition of $\cV$ by using a smooth, finite partition of unity $\{\varphi_i\}$, subspaces $\cV_i = \{w \in \cV : w = 0 \text{ if } \varphi_i = 0\}$, constraint sets $\cK_i = \{v \in \cV_i : v \ge \varphi_i \psi\}$, and decomposition operators $\Pi_i(v) = \varphi_i v$.
-One may also construct a CD which is a disjoint frequency decomposition of a Hilbert space of periodic functions on a cube $\Omega=(0,1)^d$.
-The multilevel CD, over a finite-dimensional FE discretization, will, in a vague sense, approximate such a frequency decomposition.
+More relevant to the current effort, one may also construct a CD via a disjoint frequency decomposition of a Hilbert space of periodic functions on a cube $\Omega=(0,1)^d$.
+Reflecting the underlying multigrid idea \cite{Trottenbergetal2001}, the finite-dimensional FE CD in Section \ref{sec:cdmultilevel}, will, in a vague sense, approximate such a frequency decomposition.
 
@@
 The following multiplicative/successive \pr{cd-mult} algorithm, which generalizes the Gauss--Seidel iteration, replaces $u \in \cK$ with a new iterate $w\in\cK$ which should be closer to $u^\star \in \cK$ solving \eqref{eq:vi}.
-The idea is to solve smaller VI problems over each subset $\cK_i$.
-Each subset solution corrects the global iterate immediately, so the ordering of the $\cK_i$ is important.
+Each solution of the ``smaller'' VI problem over $\cK_i$ corrects the global iterate immediately, and thus the ordering of $\{\cK_i\}$ affects the rate of convergence.
 

### 2023-12-26 paper/fascd.tex
more edits, through middlle section 4

@@
 
-\noindent The sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``those parts of $u$ which need improvement''.
+\noindent The sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``portions of $u$ needing improvement''.
 Tai \cite{Tai2003} also proposes an additive/parallel version of the algorithm, but it is not used here.
 
-The VI problems in \pr{cd-mult} reduce to constrained line searches when the $\cV_i$ are one-dimensional.
-If $e_i = w_i - \Pi_i u \in \cV_i$ denotes the correction from $\cK_i$, then the new iterate $w = u + \sum_i e_i$ is computed by adding a correction from each sub\emph{space} $\cV_i$, but so that $w \in \cK$; the corrections preserve admissibility.
-Furthermore, if VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization} then \pr{cd-mult} generates monotonically-decreasing objective values.
+The VI problems inside \pr{cd-mult} reduce to constrained line searches when the $\cV_i$ are one-dimensional.
+If $e_i = w_i - \Pi_i u \in \cV_i$ denotes the correction from $\cK_i$, then the new iterate $w = u + \sum_i e_i$ is computed by adding a correction from each subspace $\cV_i$, but so that $w \in \cK$; the corrections preserve admissibility.
+
+If VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization} then \pr{cd-mult} generates monotonically-decreasing objective values.
 
@@
 
-\pr{cd-mult} is meaningful whether or not we are in the optimization case $f=F'$, and even if $f$ is nonlinear, non-local, or defined only on $\cK$, but its practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.
+However, \pr{cd-mult} is meaningful whether or not we are in the optimization case $f=F'$, and even if $f$ is nonlinear, non-local, or defined only on $\cK$.
+Practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.
 In particular, references \cite{GraeserKornhuber2009,Tai2003} only apply the multilevel CD method to the classical obstacle problem, wherein $f=F'$ is linear, local, and defined on all of $\mathcal{V}$.
-In Section \ref{sec:vcycle} we will extend multilevel CD iterations to general nonlinear operators by applying the full approximation scheme (FAS) idea of Brandt \cite{Brandt1977}, and then illustrate its application to general operators in Examples \ref{ex:results:plap}--\ref{ex:results:sia}.
-
-From now on we restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bilateral obstacle problems}, over $\mathcal{V}=W^{1,p}(\Omega)$.
+In Section \ref{sec:vcycle} we will extend multilevel CD iterations to general nonlinear operators by applying the full approximation scheme (FAS) idea of Brandt \cite{Brandt1977}, and then illustrate its application to general operators in Section \ref{sec:results}.
+
+From now on we restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, also called bilateral obstacle problems, over $\mathcal{V}=W^{1,p}(\Omega)$.
 The \emph{lower obstacle} $\underline{\gamma}$ and \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions, defined on the closure $\overline{\Omega}$, with values in the extended real line $\tilde\RR = [-\infty,+\infty]$, satisfying the following properties: $\underline{\gamma} < +\infty$, $\overline{\gamma}>-\infty$, and $\underline{\gamma} \le \overline{\gamma}$.
-Also suppose $\partial\Omega$ is split into disjoint sets $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive Hausdorff measure, and $\partial_N \Omega$ possibly empty.
+We also suppose that $\partial\Omega$ is split into disjoint sets $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive Hausdorff measure, and $\partial_N \Omega$ possibly empty.
 Dirichlet data is given by $g_D:\partial_D \Omega \to \RR$, and we assume $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.
@@
 \end{equation}
-the closed and convex \emph{constraint set}; this treats inequality constraints and Dirichlet conditions in a unified manner.
-Dirichlet values are in the trace sense \cite{Evans2010}, and $g_D$ is assumed to be as regular as needed.
+the closed and convex \emph{constraint set}.
+(Observe that this treats inequality constraints and Dirichlet conditions in a unified manner.)  Dirichlet values are in the trace sense \cite{Evans2010}, and $g_D$ is assumed to be as regular as needed.
 Any Neumann conditions over $\partial_N \Omega$ are imposed weakly by the operator $f$, via an integral over $\partial_N\Omega$.
-We will numerically solve VI problems \eqref{eq:vi} over constraint sets \eqref{eq:originalconstraintset}.
 
@@
 
-We seek to rapidly compute the approximate solution of box-constrained VI problems \eqref{eq:vi} and \eqref{eq:originalconstraintset} using finite elements (FE).
-In this Section a nested multilevel $P_1$ or $Q_1$ element \cite{Elmanetal2014} approximation to \eqref{eq:vi} will be constructed by refinement of a coarse mesh.
-The mesh-refinement hierarchy permits the construction of a multilevel CD of the finest-level constraint set.
-
-The mesh hierarchy and FE spaces can be constructed by standard nested refinement.
+We will approximate solutions of box-constrained VI problems \eqref{eq:vi}, \eqref{eq:originalconstraintset} using finite elements (FE).
+In this Section a nested, multilevel, and $P_1$ or $Q_1$ element \cite{Elmanetal2014} mesh hierarchy is constructed by refinement of a coarse mesh.
+This hierarchy permits the construction of a multilevel CD of the finest-level constraint set in Section \ref{sec:cdmultilevel}.
+
+The construction is by standard nested refinement as follows.
 Let $\Omega \subset \RR^d$ be an open bounded polygon and suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate cells with union equal to $\overline{\Omega}$ and maximum cell diameter (mesh size) $h_0>0$.
@@
 We call $\mathcal{T}^J$ the \emph{finest-level mesh} and $J+1$ the \emph{number of levels}.
-Let $m_j$ be the number of vertices in $\mathcal{T}^j$, the \emph{degrees of freedom} at the $j$th level.
-
-Let $\mathcal{V}^j$ be nested FE spaces over each mesh $\mathcal{T}^j$, with arbitrary boundary values:
+Let $m_j$ be the number of vertices in $\mathcal{T}^j$, the number of degrees of freedom at the $j$th level.
+
+As is common in the discretization of variational inequalities, we employ lowest-order continuous Lagrange elements $P_1$ or $Q_1$ on simplices and tensor-product cells respectively.\footnote{Keith \& Surowiec have recently proposed a scheme for solving VIs using high-order finite elements~\cite{Keith2023}.}  Let $\mathcal{V}^j$ be such nested FE spaces over each mesh $\mathcal{T}^j$, with arbitrary boundary values:
 \begin{equation}
@@
 \end{equation}
-As is common in the discretization of variational inequalities, we employ lowest-order continuous Lagrange elements $P_1$ or $Q_1$ on simplices and tensor-product cells respectively\footnote{Keith \& Surowiec have recently proposed a scheme for solving VIs using high-order finite elements~\cite{Keith2023}.}.
-This choice is convenient because it enjoys \emph{nodal monotonicity}:
+The choice of $P_1$ or $Q_1$ is convenient because it enjoys \emph{nodal monotonicity}:
 \begin{equation}
@@
 for all $w,z \in \mathcal{V}^j$, where $\bw,\bz \in \RR^{m_j}$ denote vectors of nodal values.
-In other words, to enforce an inequality constraint, with $P_1$ or $Q_1$ we only enforce it pointwise on the degrees of freedom.
-For $k\ge 2$ there are functions in the standard bases of $P_k$ and $Q_k$ which take on negative values \cite[Figure 1.7]{Elmanetal2014}, violating \eqref{eq:nodalmonotonicity}.
-
-To construct box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the vertices of $\mathcal{T}^j$ with values in $\tilde{\RR} = [-\infty,+\infty]$.
+In other words, to enforce an inequality constraint over $\cV^j$ we need only enforce it pointwise on the degrees of freedom.
+By contrast, for $k\ge 2$ there are functions in the standard bases of $P_k$ and $Q_k$ which take on negative values \cite[Figure 1.7]{Elmanetal2014}, violating \eqref{eq:nodalmonotonicity}.
+
+To construct box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of $\tilde{\RR} = [-\infty,+\infty]$ valued functions on the vertices of $\mathcal{T}^j$.
 The finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ are constructed by suitable interpolation of $\underline{\gamma}$, $\overline{\gamma}$, e.g.~by applying the Cl\'ement interpolation operator~\cite{Carstensen2006}.
-Again the obstacles must satisfy $\underline{\gamma}^J < +\infty$, $-\infty < \overline{\gamma}^J$, and $\underline{\gamma}^J \le \overline{\gamma}^J$.
-On the finest-level mesh the discrete boundary data $g_D^J \in \mathcal{V}^J$ is assumed to satisfy $g_D^J = g_D$ on $\partial_D \Omega$---this restricts $g_D$ in \eqref{eq:originalconstraintset} to be piecewise-linear---and $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those vertices of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.
+These obstacles must satisfy $\underline{\gamma}^J < +\infty$, $-\infty < \overline{\gamma}^J$, and $\underline{\gamma}^J \le \overline{\gamma}^J$.
+On the finest-level mesh the discrete boundary data $g_D^J \in \mathcal{V}^J$ is assumed to satisfy $g_D^J = g_D$ on $\partial_D \Omega$---this restricts $g_D$ in \eqref{eq:originalconstraintset} to be piecewise-linear---and $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ at those vertices of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.
 Then we define
@@
 Let $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ be a discretization of $f$ in \eqref{eq:vi}.
-Note that the functional $\ell^J=\ell$ is defined and continuous over $\mathcal{V}^J$.
-We seek the FE solution $u^J \in \mathcal{K}^J$ of the finest-level, finite-dimensional, nonlinear VI
+Note that, by inclusion \eqref{eq:fe:nestedspaces}, $\ell^J=\ell$ is defined and continuous over $\mathcal{V}^J$.
+We seek the FE solution $u^J \in \mathcal{K}^J$ of the finest-level, finite-dimensional, nonlinear VI:
 \begin{equation}
@@
 \end{equation}
-If $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, then \eqref{eq:fe:vi} is well-posed.
+If $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$ then \eqref{eq:fe:vi} is well-posed.
 

### 2023-12-26 paper/fascd.tex
edits through end section 5

@@
 
-While \eqref{eq:fe:vi} approximates \eqref{eq:vi}, we warn the reader that it is usually not conforming.
+While \eqref{eq:fe:vi} approximates \eqref{eq:vi}, we warn the reader that it is usually not a conforming FE method.
 That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, from definitions \eqref{eq:originalconstraintset} and \eqref{eq:fe:fineconstraintset} we do not expect $\mathcal{K}^J \subset \mathcal{K}$ in general.
-Even for continuous obstacles $\underline{\gamma}, \overline{\gamma}$, admissibility with respect to the FE obstacles $\underline{\gamma}^J, \overline{\gamma}^J$ will generally not imply $\mathcal{K}$-admissibility.
-For a practical example, nodal elevation data in a glacier application misses the topographic detail in-between the vertices \cite{Bueler2016}.
-
-The multilevel algorithm defined in Section \ref{sec:vcycle} requires five transfer operators between meshes.
-Of these, three are employed in standard FAS \cite{Trottenbergetal2001}; Table \ref{tab:transfers} will help the reader keep track.
-Canonical prolongation $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ implements the vector space inclusion.
-Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $\ip{R\sigma}{z} = \ip{\sigma}{z}$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
-These canonical transfers generally involve a change of FE representation, and in particular $R$ is ``full-weighting'' \cite{Trottenbergetal2001}.
-To define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, note that the vertices of $\mathcal{T}^{j-1}$ are vertices of $\mathcal{T}^j$ and so for $z\in\mathcal{V}^j$ define $\iR z$ as the element of $\mathcal{V}^{j-1}$ with the same (coarse) nodal values.
+Even for continuous obstacles $\underline{\gamma}, \overline{\gamma}$, FE admissibility \eqref{eq:fe:fineconstraintset} will generally not imply $\mathcal{K}$-admissibility \eqref{eq:originalconstraintset}.
+For a practical consequence of this situation, regarding Example \ref{ex:sia}, nodal bed elevation data in a glacier application will miss the topographic detail between the nodes \cite{Bueler2016}.
+
+The multilevel algorithm of Section \ref{sec:vcycle} requires five transfer operators between meshes, and Table \ref{tab:transfers} will help the reader keep track.
+Three of these are already employed in standard FAS \cite{Trottenbergetal2001}, and they are linear maps defined as follows.
+First, canonical prolongation $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ implements the vector space inclusion \eqref{eq:fe:nestedspaces}.
+Second, canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ is defined for $\sigma \in (\mathcal{V}^j)'$ as $\ip{R\sigma}{z} = \ip{\sigma}{z}$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.
+(These canonical transfers imply a change of FE representation; $R$ is ``full-weighting'' \cite{Trottenbergetal2001}.)  Finally, to define injection $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$, note that the vertices of $\mathcal{T}^{j-1}$ are vertices of $\mathcal{T}^j$.
+Thus for $z\in\mathcal{V}^j$ we define $\iR z$ as the element of $\mathcal{V}^{j-1}$ with the same (coarse) nodal values.
 Injection is monotonic in the sense that if $w \le z$ then $\iR w \le \iR z$.
@@
 
-The monotone injections $\maxR,\minR$ \cite{Tai2003,GraeserKornhuber2009}, which will only be applied to constraints in $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the stars of the coarse vertices).
-To make the construction precise, let $x_q^j$ denote the vertices of mesh $\mathcal{T}^j$, with corresponding basis functions $\psi_q^j$.
-For an $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$:
+The monotone injections $\maxR,\minR$ \cite{Tai2003,GraeserKornhuber2009} in Table \ref{tab:transfers} will only be applied to constraints in the $\tilde\RR$-valued nodal spaces $\tilde{\mathcal{V}}^j$.
+They are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the stars of the coarse vertices).
+To make the construction precise, let $x_q^j$ denote a vertex of mesh $\mathcal{T}^j$, with corresponding basis function $\psi_q^j$.
+For $z\in\tilde{\mathcal{V}}^j$ we define new functions in $\tilde{\mathcal{V}}^{j-1}$:
 \begin{align}
@@
 \end{align}
-These operators are constructed to enforce the property
-\begin{equation}
-\minR z \le z \le \maxR z,  \label{eq:fe:monotoneproperty}
-\end{equation}
-which does not hold for the ordinary injection operator $\iR$.
-An illustration of the construction on nested meshes is given in Figure \ref{fig:Rplusminus}; implementing these operators on non-nested meshes could be efficiently done using a supermesh~\cite{Farrell2011}.
-In the next Section we will use $\maxR,\minR$ to construct ``level defect constraints'' such that the zero function is always admissible.
-This construction depends on additional ordering properties which follow from definition \eqref{eq:fe:monotoneinjections}:
+These operators are constructed to imply
+\begin{equation}
+\minR z \le z \le \maxR z.  \label{eq:fe:monotoneproperty}
+\end{equation}
+An illustration of the construction is given in Figure \ref{fig:Rplusminus}.
+Implementing these operators on non-nested meshes could be done efficiently using a supermesh~\cite{Farrell2011}.
+In the next Section we will use $\maxR,\minR$ to construct ``level defect constraints'' such that the zero function is always admissible, a construction which depends on these additional ordering properties which follow from \eqref{eq:fe:monotoneinjections}:
 \begin{equation}
@@
 
-\section{Multilevel CDs from level defect constraints} \label{sec:cdmultilevel}
-
-A key choice in devising a multilevel scheme for VIs is how to construct bounds for the coarser levels.
+\section{Multilevel CDs via level defect constraints} \label{sec:cdmultilevel}
+
+A key choice in devising a multilevel scheme for VIs is in the construction of bounds for the coarser levels.
 Our multilevel CD scheme uses \emph{defect constraints} \cite{GraeserKornhuber2009} derived from an admissible finest-level iterate.
-The next Example illustrates why we do \emph{not}\footnote{An exception occurs in FMG; see Section \ref{sec:implementation}.} apply injection operators directly to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.
+The next Example illustrates why we do \emph{not}, with the exception of an interpolation needed for the FMG method of Section \ref{sec:implementation}, apply injection operators directly to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.
 
@@
 Let $\underline{\gamma}^1$ be a sawtooth function with minimum value zero, and suppose $\overline{\gamma}^1=C+\underline{\gamma}^1$ with $C>0$ strictly less than the maximum of $\underline{\gamma}^1$.
-Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ vertices, respectively.
+Figure \ref{fig:directRbad} shows such functions on $\mathcal{T}^1$ with five vertices.
 For appropriate boundary conditions the fine-level constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.
 However, various schemes to coarsen the constraints by direct transfers of $\underline{\gamma}^1,\overline{\gamma}^1$ are doomed to failure.
-Suppose we propose
+Suppose
 \begin{equation}
@@
 Then $\underline{\gamma}^0=0$ and $\overline{\gamma}^0=C$ identically.
-In that case the corresponding constraint set $\mathcal{K}^0$ is nonempty, but prolongation $Pw^0$ of an admissible solution of the coarse problem ($w^0\in\mathcal{K}^0$; $\underline{\gamma}^0 \le w^0 \le \overline{\gamma}^0$) is not admissible on the finer level ($Pw^0 \notin \mathcal{K}^1$).
+In that case the corresponding constraint set $\mathcal{K}^0$ is nonempty, but prolongation $Pw^0$ of an admissible solution of the coarse problem ($w^0\in\mathcal{K}^0$; $\underline{\gamma}^0 \le w^0 \le \overline{\gamma}^0$) is never admissible on the finer level ($Pw^0 \notin \mathcal{K}^1$).
 Attempting to fix this by projecting (truncating) into $\mathcal{K}^1$ would introduce high frequencies not present in $w^0$.
@@
 Possibility \eqref{eq:fe:badmonotone}, which generates an empty coarser-level set $\mathcal{K}^0$, is at right.
-Defect constraints \eqref{eq:fe:defectconstraints} avoid this difficulty.}
+Defect constraints \eqref{eq:fe:defectconstraints} avoid such difficulties.}
 \label{fig:directRbad}
@@
 
-Iterate-dependent defect constraints will satisfy ordering \eqref{eq:fe:chiordering} below and bypass the difficulties in the above Example.
-Suppose $w^J \in \cK^J$ is an admissible iterate for problem \eqref{eq:fe:vi}.
+However, we can construct iterate-dependent constraints which satisfy ordering \eqref{eq:fe:chiordering} below, and which bypass the difficulties in the above Example.
+Suppose $w^J \in \cK^J$ is an admissible iterate; see \eqref{eq:fe:fineconstraintset}.
 We define the \emph{finest-level defect constraints} \cite{GraeserKornhuber2009} as functions in $\tilde{\mathcal{V}}^J$:
@@
 \end{equation}
-(If $\underline{\gamma}^J=-\infty$ then $\underline{\chi}^J=-\infty$ by definition, and similarly if $\overline{\gamma}^J=+\infty$ then $\overline{\chi}^J=+\infty$.)  Definition \eqref{eq:fe:fineconstraintset} implies that $\underline{\chi}^J \le 0 \le \overline{\chi}^J$ everywhere in $\overline{\Omega}$, including on the Dirichlet boundary $\partial_D\Omega$.
-Observe that a corrected iterate $w^J + z^J$ is admissible, $w^J + z^J \in \cK^J$, if and only if the perturbation $z^J \in \mathcal{V}^J$ is zero on $\partial_D\Omega$ and bounded by the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$.
-
-Now descending to coarser levels, $j=J-1,\dots,1,0$, we apply the monotone injections to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:\footnote{The definition of $\underline{\chi}^j$ matches that in Section 4 of \cite{GraeserKornhuber2009}; see (4.22) and Figure 4.1 therein.}
+(If $\underline{\gamma}^J=-\infty$ then $\underline{\chi}^J=-\infty$ by definition, and similarly if $\overline{\gamma}^J=+\infty$ then $\overline{\chi}^J=+\infty$.)  Definition \eqref{eq:fe:fineconstraintset} implies that $\underline{\chi}^J \le 0 \le \overline{\chi}^J$ on $\overline{\Omega}$.
+Observe that a corrected iterate $w^J + z^J$ is admissible, i.e.~$w^J + z^J \in \cK^J$, if and only if $z^J \in \mathcal{V}^J$ is zero on $\partial_D\Omega$ and bounded by the defect constraints, $\underline{\chi}^J \le z^J \le \overline{\chi}^J$.
+
+To descend to coarser levels, $j=J-1,\dots,1,0$, we apply the monotone injections to generate \emph{level defect constraints} (LDCs) in $\tilde{\mathcal{V}}^j$:\footnote{The definition of $\underline{\chi}^j$ matches that in Section 4 of \cite{GraeserKornhuber2009}; see (4.22) and Figure 4.1 therein.}
 \begin{equation}
@@
 \end{equation}
-The LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if another element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$; similar comments apply to $\overline{\chi}^{j}$.
-Note that for Algorithm \ref{alg:fascd} below, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not required.
-
-This construction of multilevel box constraints, likewise the unilateral construction in \cite{GraeserKornhuber2009}, permits large, yet admissible, coarse corrections, which promotes multilevel solver efficiency.
-When coarsening the problem downward in a V-cycle, one cannot predict the upcoming coarser corrections and so one must construct constraint sets from which any sum will be admissible.
+The LDC $\underline{\chi}^j$ is the most-negative function from $\tilde{\mathcal{V}}^j$ such that if an element of $\tilde{\mathcal{V}}^j$ is above $\underline{\chi}^j$ then it is also above $\underline{\chi}^{j+1}$; corresponding comments apply to $\overline{\chi}^{j}$.
+Note that for Algorithm \ref{alg:fascd} below, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not actually required.
+
+This construction of multilevel box constraints, likewise of the unilateral construction in \cite{GraeserKornhuber2009}, should permit large, yet admissible, coarse corrections, which promotes multilevel solver efficiency.
+However, when coarsening the problem downward in a V-cycle, we observe that one cannot predict the upcoming coarser corrections and so one must construct constraint sets from which any sum will be admissible.
 With this in mind, and looking forward to telescoping sums \eqref{eq:fe:telescoping}, we compute the LDC differences \cite{GraeserKornhuber2009}:
@@
 (Take $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ so that $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.
-By definition, if $\underline{\chi}^j(x_p)=-\infty$ then we set $\underline{\phi}^j(x_p)=-\infty$ regardless of the value of $\underline{\chi}^{j-1}(x_p)$, and etc.)  Note that while $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ again bracket zero, they are not ordered as in \eqref{eq:fe:chiordering}.
-
-The corresponding \emph{downward} and \emph{upward constraint sets}, respectively, are closed and convex subsets of the FE spaces $\mathcal{V}^j$:
+By definition, if $\underline{\chi}^j(x_p)=-\infty$ then we set $\underline{\phi}^j(x_p)=-\infty$ regardless of the value of $\underline{\chi}^{j-1}(x_p)$, and etc.)  Note that while $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ do bracket zero, they are not ordered as in \eqref{eq:fe:chiordering}.
+
+Now we define \emph{downward} and \emph{upward constraint sets}, respectively, as certain closed and convex subsets of the FE spaces $\mathcal{V}^j$:
 \begin{align}
@@
 Because $\underline{\chi}^j \le \underline{\phi}^j \le 0 \le \overline{\phi}^j \le \overline{\chi}^j$, it follows that $\mathcal{D}^j \subseteq \mathcal{U}^j$.
-Note that the finest-level set $\mathcal{U}^J$ is equivalent to the original constraint set $\mathcal{K}^J$: for all $z^J \in \mathcal{V}^J$ , $z^J \in \mathcal{U}^J$ if and only if $w^J+z^J \in \mathcal{K}^J$.
+Note that the finest-level upward set $\mathcal{U}^J$ is equivalent to the original constraint set $\mathcal{K}^J$: for all $z^J \in \mathcal{V}^J$ , $z^J \in \mathcal{U}^J$ if and only if $w^J+z^J \in \mathcal{K}^J$.
 
@@
 for $j=0,1,\dots,J$.
-In unilateral cases, where either $\underline{\gamma}^J=-\infty$ identically or $\overline{\gamma}^J=+\infty$ identically, inclusion \eqref{eq:fe:downwardsuminclusion} becomes a full CD with $\mathcal{U}^j=\sum_{i=0}^j \mathcal{D}^i$.
+In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$ identically, inclusion \eqref{eq:fe:downwardsuminclusion} becomes a full CD with $\mathcal{U}^j=\sum_{i=0}^j \mathcal{D}^i$.
 \end{lemma}
@@
 
-One might attempt to convert \eqref{eq:fe:downwardsuminclusion} into a full CD for arbitrary box constraints by constructing different decomposition maps $\Pi_i$, for example by splitting $z^j = \min\{z^j,0\} + \max\{z^j,0\}$ and applying unilateral formulas \eqref{eq:fe:unilateraldecompositionoperator} to $\min\{z^j,0\}$ and $\max\{z^j,0\}$ separately.
-The above proof shows that the lower bound $\Pi_i (\min\{z^j,0\}) \ge \underline{\phi}^i$ indeed holds.
+One might try to convert \eqref{eq:fe:downwardsuminclusion} into a full CD for arbitrary box constraints by attempting to construct decomposition maps $\Pi_i$ as follows.
+One can split $z^j = \min\{z^j,0\} + \max\{z^j,0\}$ and apply unilateral formulas \eqref{eq:fe:unilateraldecompositionoperator} to $\min\{z^j,0\}$ and $\max\{z^j,0\}$ separately.
+The above proof shows that the lower bound $\Pi_i (\min\{z^j,0\}) \ge \underline{\phi}^i$ does hold.
 However, the following Example shows why $\Pi_i(\min\{z^j,0\})$ may have an arbitrarily large maximum, and thus exceed any finite upper LDC $\overline{\phi}^i$.
-In fact, $\Pi_i$ defined above does not generally map all non-positive elements of $\mathcal{U}^J$ into $\mathcal{D}^i$ if a finite upper LDC is present.
-The Example does not exclude other strategies for showing full CDs in the general box-constrained case.
+In fact, $\Pi_i$ defined above does not generally map all non-positive elements of $\mathcal{U}^J$ into $\mathcal{D}^i$ in cases where a finite upper LDC is present.
 
@@
 
-Each upward set $\mathcal{U}^j$ can also be decomposed, incompletely in the sense of \eqref{eq:fe:downwardsuminclusion}, down to an arbitrary coarser level.
-The next Lemma, which does not follow from Lemma \ref{lem:downwardadmissibility}, justifies upward admissibility in Algorithm \ref{alg:fascd}.
+This Example does not exclude other strategies for showing full CDs in the general box-constrained case.
+
+In any case, each upward set $\mathcal{U}^j$ can be decomposed, incompletely in the sense of \eqref{eq:fe:downwardsuminclusion}, down to an arbitrary coarser level.
+The next Lemma does not follow from Lemma \ref{lem:downwardadmissibility}.
 
@@
 \end{equation}
-In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, actually $\mathcal{U}^j = \mathcal{U}^k + \sum_{i=k+1}^j \mathcal{D}^i$ is a full CD. \end{lemma}
+In unilateral cases, where either $\underline{\gamma}^J=-\infty$ or $\overline{\gamma}^J=+\infty$, it follows that $\mathcal{U}^j = \mathcal{U}^k + \sum_{i=k+1}^j \mathcal{D}^i$ is a full CD.
+\end{lemma}
 
@@
 
-Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} show the admissibility of all iterates and corrections in Algorithm \ref{alg:fascd} below.
-These inclusions are special to multilevel schemes based on defect constraints, and we have seen that they hold regardless of whether decomposition operators $\Pi_i$ can be constructed.
-In fact we can propose a new, incomplete variant of the \pr{cd-mult} iteration.
-For $w^J\in\mathcal{K}^J$, and existing downward perturbations $y^i\in\mathcal{D}^i$, the iteration applies in the special case where inclusion \eqref{eq:fe:upwardsuminclusion} holds:
+Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} imply the admissibility of all iterates and corrections in Algorithm \ref{alg:fascd} below.
+These inclusions are special to multilevel schemes based on defect constraints, and they hold regardless of whether decomposition operators $\Pi_i$ can be constructed.
+We propose a new, incomplete variant of the \pr{cd-mult} iteration which requires no operators $\Pi_i$.
+For $w^J\in\mathcal{K}^J$, and (input) downward perturbations $y^i\in\mathcal{D}^i$, it applies in the special case where inclusion \eqref{eq:fe:upwardsuminclusion} holds:
 
@@
 \noindent
-Note that the downward perturbations $y^i$, which apparently disappear as \pr{icd-tele} proceeds, can actually play a practical role in solving each VI problem, namely when constructing an initial iterate for $z^j$; see the next Section.
+The downward perturbation $y^j$, which apparently disappears as \pr{icd-tele} proceeds, actually helps in constructing an initial iterate for $z^j$; see Algorithm \ref{alg:fascd}.
 
@@
 For comparison, $V(1,1)$ cycles in \cite{GraeserKornhuber2009} use half-size versions of the smaller sets $\mathcal{D}^j$ both on descent and ascent; see also \cite[section 5.4]{Tai2003}.
-However, any convergence analysis extending the arguments of \cite{GraeserKornhuber2009,Tai2003} might require full CDs.
+On the other hand, any convergence analysis extending the arguments of \cite{GraeserKornhuber2009,Tai2003} might require full CDs.
 

### 2023-12-26 paper/fascd.tex
edits through middle Section 7

@@
 
-The V-cycle in this Section extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.2]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}; it reduces to FAS for PDEs when inequality constraints are removed.
+The V-cycle in this Section extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.2]{GraeserKornhuber2009} by following a nonlinear full approximation scheme (FAS) approach \cite{BrandtLivne2011}.
+Our V-cycle reduces to FAS for PDEs when inequality constraints are removed.
 Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$, via information from each coarser FE subspace $\mathcal{V}^j$, by approximately solving VI problems in constraint sets $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.
-The corrections are inductively constructed so that iterates remain admissible at every level.
-A key idea is that up-smoothing solves less-restricted VI problems than down-smoothing.
-Going downward one does not know what coarse corrections are coming, thus one must choose smaller constraint sets ($\mathcal{D}^j$) such that the correction sum is admissible once we return to a given level.
-Going upward, however, no further coarse corrections can violate admissibility, so we relax in the larger sets $\mathcal{U}^j$.
-
-Suppose $y^i \in \mathcal{D}^i$, for $i=J,\dots,j+1$, are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdcycles}, left).
+The corrections are constructed so that iterates remain admissible at every level.
+A key idea is that up-smoothing solves less-restricted VI problems, in larger subsets, than down-smoothing.
+Going downward one does not know what coarse corrections are coming, thus one must choose the smaller constraint sets ($\mathcal{D}^j$) so the correction sum is admissible once we return to a given level.
+Going upward, however, no further coarse corrections can violate admissibility, so we may relax in the larger sets $\mathcal{U}^j$.
+
+Suppose that $y^i \in \mathcal{D}^i$, for $i=J,\dots,j+1$, are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdcycles}, left).
 By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.
@@
 Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for relaxation on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using inclusion \eqref{eq:fe:upwardsuminclusion}.
-Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.
+Continuing in this ``telescoping'' way by up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.
 The new finest-level iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.
@@
 \end{tabular}
-\caption{Left: The FASCD V-cycle Algorithm \ref{alg:fascd} computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.
-Right: FMG Algorithm \ref{alg:fascd-fmg} generates the initial iterate on finer levels by injection and truncation (doubled edges).}
+\caption{Left: FASCD V-cycle Algorithm \ref{alg:fascd} computes downward corrections $y^j \in \mathcal{D}^j$, but the upward corrections $z^j\in\mathcal{U}^j$ are in larger sets.
+Right: FMG Algorithm \ref{alg:fascd-fmg} generates initial iterates on finer levels via injection and truncation (doubled edges).}
 \label{fig:fascdcycles}
@@
 
-We must specify the problems solved at each mesh level.
+We must specify the problems which are to be solved at each mesh level.
 As in FAS multigrid for nonlinear PDEs \cite{BrandtLivne2011,Bruneetal2015,Trottenbergetal2001}, both the solution approximation and the residual must be coarsened, and a source functional created.
-On mesh level $j$ we assume a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:vi}.
-At the start of each V-cycle we assume that an admissible iterate $w^J \in \mathcal{K}^J$ has been used to generate the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}); they do not depend on the coarse-level corrections.
+On any mesh level $j$ we assume there is a re-discretized nonlinear operator $f^j$, with output in $(\cV^j)'$, which is an FE discretization of $f$ in \eqref{eq:vi}.
+At the start of each V-cycle we assume that an admissible iterate $w^J \in \mathcal{K}^J$ has been used to generate the LDCs and constraint sets $\mathcal{D}^j$, $\mathcal{U}^j$ at each level (Section \ref{sec:cdmultilevel}).
 
@@
 \end{equation}
-As is typical of FAS \cite[section 5.3]{Trottenbergetal2001}, injection $\iR$ coarsens the (primal) iterates, but canonical restriction $R$ coarsens (dual) linear functionals.
+As in FAS \cite[section 5.3]{Trottenbergetal2001}, injection $\iR$ coarsens the (primal) iterates, but canonical restriction $R$ coarsens the (dual) linear functionals.
 Descending through the levels $j=J,J-1,\dots,1$ we solve the following VI problems for $y^j \in \mathcal{D}^j$:
@@
 
-On ascent, $j=0,1,\dots,J$, the same VI is solved but for $z^j \in \mathcal{U}^j$:
+On ascent, $j=0,1,\dots,J$, nearly the same VI is solved, but now for $z^j \in \mathcal{U}^j$:
 \begin{equation}
@@
 
-However, because the down-smoother has made progress, $y$ should be approximate-able by solving a coarser, $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps: \emph{(i)} subtract the computable quantity $f^{j+1}(w^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides, \emph{(ii)} replace the relaxed residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction, \emph{(iii)} replace $w^{j+1}+y^{j+1}$, where it appears on the left, by its injection, and \emph{(iv)} where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.
+However, because the down-smoother has made progress, $y$ should be approximate-able via a coarser $j$th-level problem generated by modifying \eqref{eq:fe:downvinotional} in 4 steps: \emph{(i)} subtract the computable quantity $f^{j+1}(w^{j+1}+y^{j+1}) \in (\mathcal{V}^{j+1})'$ from both sides, \emph{(ii)} replace the relaxed residual $\ell^{j+1}-f^{j+1}(w^{j+1}+y^{j+1})$ on the right by its restriction, \emph{(iii)} replace $w^{j+1}+y^{j+1}$ where it appears on the left by its injection, and \emph{(iv)} where it appears on the left, replace $f^{j+1}$ by the coarser rediscretization $f^j$.
 These steps yield a problem for the $j$th-level correction $y=y^j \in \mathcal{D}^j$:
@@
 \end{align}
-for all $v\in \mathcal{D}^{j}$.  Though in cluttered form, this is VI problem \eqref{eq:fe:downvi}, once definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource} are applied.
+for all $v\in \mathcal{D}^{j}$.  Though it is cluttered, this becomes VI problem \eqref{eq:fe:downvi} once definitions \eqref{eq:fe:definew} and \eqref{eq:fe:levelsource} are applied.
 
 The explanation for \eqref{eq:fe:upvi} and \eqref{eq:fe:upwardinitial} is simpler.
-Whether or not problem \eqref{eq:fe:downvi} has been exactly solved by the computed $y^j\in\mathcal{D}^j$, when going upward we can expand problem \eqref{eq:fe:downvi} into the larger set $\mathcal{U}^j \supset \mathcal{D}^j$, and \eqref{eq:fe:upwardsuminclusion} assures admissibly upward from any approximate solution to \eqref{eq:fe:upvi}.
-Generally $y^j$ is not the solution to \eqref{eq:fe:upvi}; e.g.~in the optimization case $f=F'$ it would not be optimal in the enlarged constraint set.
+Whether or not problem \eqref{eq:fe:downvi} has been exactly solved by the computed $y^j\in\mathcal{D}^j$, when going upward we can expand the VI problem into the larger set $\mathcal{U}^j \supset \mathcal{D}^j$.
+Inclusion \eqref{eq:fe:upwardsuminclusion} assures admissibility upward from any approximate solution to \eqref{eq:fe:upvi}.
 Our current best estimate of the solution to \eqref{eq:fe:upvi} is $z^j \approx y^j + Pz^{j-1}\in \mathcal{U}^j$, so this is the initial iterate for relaxation.
@@
 These ideas come together in Algorithm \ref{alg:fascd}, \pr{fascd-vcycle}, which updates $w^J$.
-The \pr{smooth} and \pr{solve} procedures are assumed to do in-place modifications of their final arguments, without modifying the others.
+The \pr{smooth} and \pr{solve} procedures are assumed to do in-place modifications of their final arguments.
+With \id{down} $=1$ and \id{up} $=1$ smoother iterations on each level, Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle.
+The $V(1,0)$ cycle with \id{up} $=0$ generalizes Algorithm 4.2 in \cite{GraeserKornhuber2009} by applying FAS-type corrections and allowing any smoother.
+A single $V(1,1)$ iteration of \pr{fascd-vcycle} approximates an application of \pr{cd-mult} (Section \ref{sec:cd}) over CD \eqref{eq:fe:downwardsuminclusion}, followed by \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over inclusion \eqref{eq:fe:upwardsuminclusion}, and in this sense $V(1,1)=V(1,0)+V(0,1)$.
 
@@
       $y^j = 0$ \\
-      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\underline{\phi}^j,\overline{\phi}^j,w^j;y^j)$  \ct{problem \eqref{eq:fe:downvi} in $\mathcal{D}^j$}\\
+      $\text{\pr{smooth}}^{\text{\id{down}}}(\ell^j,\underline{\phi}^j,\overline{\phi}^j,w^j;y^j)$  \ct{VI problem \eqref{eq:fe:downvi} in $\mathcal{D}^j$}\\
       $w^{j-1} = \iR(w^j + y^j)$ \label{line:vcyclerestrictsolution} \\
@@
     $z^0 = 0$ \\
-    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,\overline{\chi}^0,w^0;z^0)$ \hspace{1.0cm} \ct{problem \eqref{eq:fe:upvi} in $\mathcal{U}^0$} \\
+    $\text{\pr{solve}}(\ell^0,\underline{\chi}^0,\overline{\chi}^0,w^0;z^0)$ \hspace{1.0cm} \ct{VI problem \eqref{eq:fe:upvi} in $\mathcal{U}^0$} \\
     for $j=1$ to $j=J$ \\+
       $z^j = y^{j} + P z^{j-1}$ \label{line:vcycleupsmoothinitial} \\
-      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,\overline{\chi}^j,w^j;z^j)$  \ct{problem \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
+      $\text{\pr{smooth}}^{\text{\id{up}}}(\ell^j,\underline{\chi}^j,\overline{\chi}^j,w^j;z^j)$  \ct{VI problem \eqref{eq:fe:upvi} in $\mathcal{U}^j$} \\-
     return $w^J+z^J$
 \end{pseudo}
-\caption{The FASCD V-cycle, an iteration for solving VI problem \eqref{eq:fe:vi}.}
+\caption{The FASCD V-cycle for solving VI problem \eqref{eq:fe:vi}.}
 \label{alg:fascd}
 \end{pseudofloat}
-
-With \id{down} $=1$ and \id{up} $=1$ smoother iterations on each level, Algorithm \ref{alg:fascd} is a $V(1,1)$ cycle.
-The $V(1,0)$ cycle with \id{up} $=0$ is similar to Algorithm 4.2 in \cite{GraeserKornhuber2009}, but here it is generalized by applying FAS-type corrections and allowing any smoother.
-In unilateral cases one $V(1,1)$ iteration of \pr{fascd-vcycle} approximates applications of \pr{cd-mult} (Section \ref{sec:cd}) over CD \eqref{eq:fe:downwardsuminclusion}, followed by \pr{icd-tele} (Section \ref{sec:cdmultilevel}) over \eqref{eq:fe:upwardsuminclusion}, and in this sense $V(1,1)=V(1,0)+V(0,1)$.
 
@@
 \end{equation}
-namely $16m_J$, $12m_J$, and $10m_J$ in dimensions $d=1,2,3$, respectively.
+at most $16m_J$, $12m_J$, and $10m_J$ in dimensions $d=1,2,3$, respectively.
 For comparison, a single-level method needs $4 m_J$ storage (i.e.~$\underline{\gamma}^J,\overline{\gamma}^J,\ell^J,w^J$).
@@
 
-In this section we propose a convergence criterion, construct the full multigrid (FMG) extension of Algorithm \ref{alg:fascd}, and summarize the reduce-space Newton method used as the smoother and coarse solver.
-
-Repeated application of Algorithm \ref{alg:fascd} to solve finite-dimensional VI problem \eqref{eq:fe:vi} will not generally reduce the residual $f^J(w^J) - \ell^J$ to zero everywhere.
-However, the VI is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003}, which says, by definition, that certain conditions hold for $u^J$ at nodes $x_p \in \mathcal{T}^J$.
-Let $r(u^J)_p = \ip{f^J(u^J)-\ell^J}{\psi_p}$ be the ordinary (pointwise) residual.
-If $x_p \notin \partial_D\Omega$ then exactly one of these four conditions must hold:
+In this Section we propose a convergence criterion, construct the full multigrid (FMG) extension of Algorithm \ref{alg:fascd}, and summarize the reduce-space Newton method used as the smoother and coarse solver.
+
+A solution of finite-dimensional VI problem \eqref{eq:fe:vi} will not generally reduce the residual $f^J(u^J) - \ell^J$ to zero everywhere.
+Instead the VI is equivalent to a \emph{mixed complementarity problem} (MCP) \cite{FacchineiPang2003} which says, by definition, that certain conditions hold for $u^J$ at nodes $x_p \in \mathcal{T}^J$.
+Let $r(u^J)_p = \ip{f^J(u^J)-\ell^J}{\psi_p}$ be the ordinary residual at $x_p$.
+If $x_p \notin \partial_D\Omega$ then from \eqref{eq:fe:vi} exactly one of these four conditions must hold:
 \begin{itemize}
@@
 \end{itemize}
-These 5 cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, \emph{pinched} (\emph{both active}), and \emph{boundary}, respectively.
-Where \emph{pinched} or \emph{boundary} apply, $r(u^J)_p$ can have any value.
+These 5 cases might be labeled \emph{inactive}, \emph{lower active}, \emph{upper active}, \emph{pinched} (\emph{both active}), and \emph{Dirichlet boundary}, respectively.
+Where \emph{pinched} or \emph{Dirichlet boundary} apply, $r(u^J)_p$ can have any value.
 
@@
 Observe that $\phi$ is semi-smooth, thus continuous, and that $\phi(a,b)=0$ if and only if $a,b$ are complementary, i.e.~$a\ge 0$, $b\ge 0$, and $ab=0$.
-For any $w^J\in \mathcal{K}^J$ denote $\underline{g}_p = w^J(x_p) - \underline{\gamma}^J(x_p)$ and $\overline{g}_p = \overline{\gamma}^J(x_p) - w^J(x_p)$, the nonnegative \emph{gaps} at $x_p$.
+For any $w^J\in \mathcal{K}^J$ denote $\underline{g}_p = w^J(x_p) - \underline{\gamma}^J(x_p)$ and $\overline{g}_p = \overline{\gamma}^J(x_p) - w^J(x_p)$ as the nonnegative \emph{gaps} at $x_p$.
 Now define the \emph{semi-smooth residual}
@@
 
-Suppose that Algorithm \ref{alg:fascd} is started with $w^{J,0} \in \mathcal{K}^J$ and generates (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.
+Now suppose that Algorithm \ref{alg:fascd} is iterated from $w^{J,0} \in \mathcal{K}^J$, generating (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.
 Let $\id{atol}$, $\id{rtol}$, $\id{stol} \ge 0$ be given absolute, relative, and step tolerances, respectively.
-We employ
-\begin{equation}
-\|\rSS(w^{J,k})\|_2 < \id{atol} \quad \text{or} \quad \frac{\|\rSS(w^{J,k})\|_2}{\|\rSS(w^{J,0})\|_2} < \id{rtol} \quad \text{or} \quad \frac{\|\Delta w^{J,k}\|_{L^2}}{\|w^{J,k}\|_{L^2}} < \id{stol}, \label{eq:stoppingcriterion}
-\end{equation}
-where $\Delta w^{J,k} = w^{J,k} -  w^{J,k-1}$, $\|\cdot\|_2$ is the euclidean norm on $\RR^{m_J}$, and $\|\cdot\|_{L^2}$ is the norm on $L^2(\Omega)$, as the stopping criterion.
-Defaults in our Firedrake implementation match those of PETSc SNES \cite{Balayetal2023}: $\id{atol}=10^{-50}$, $\id{rtol}=\id{stol}=10^{-8}$; see Section \ref{sec:results} for specific settings in Examples.
+We employ the following stopping criterion:
+\begin{equation}
+\|\rSS(w^{J,k})\|_2 < \id{atol} \quad \text{or} \quad \frac{\|\rSS(w^{J,k})\|_2}{\|\rSS(w^{J,0})\|_2} < \id{rtol} \quad \text{or} \quad \frac{\|\Delta w^{J,k}\|_{L^2}}{\|w^{J,k}\|_{L^2}} < \id{stol}.
+\label{eq:stoppingcriterion}
+\end{equation}
+Here $\Delta w^{J,k} = w^{J,k} -  w^{J,k-1}$, $\|\cdot\|_2$ is the euclidean norm on $\RR^{m_J}$, and $\|\cdot\|_{L^2}$ is the norm on $L^2(\Omega)$.
+Defaults in our implementation (Section \ref{sec:results}) match those of PETSc SNES \cite{Balayetal2023}: $\id{atol}=10^{-50}$, $\id{rtol}=\id{stol}=10^{-8}$; see Examples below for specific settings.
 

### 2023-12-26 paper/fascd.tex
edits through Example 8.3

@@
 
-Now suppose that Algorithm \ref{alg:fascd} is iterated from $w^{J,0} \in \mathcal{K}^J$, generating (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.
+Now suppose that Algorithm \ref{alg:fascd} is iterated from $w^{J,0} \in \mathcal{K}^J$, generating fine-level iterates $w^{J,k} \in \mathcal{K}^J$.
 Let $\id{atol}$, $\id{rtol}$, $\id{stol} \ge 0$ be given absolute, relative, and step tolerances, respectively.
@@
 
-In Section \ref{sec:results} we will show that V-cycle iterations alone, starting with some finest-mesh initial iterate $w^{J,0}$ and repeated until \eqref{eq:stoppingcriterion} holds, serve as an effective solver.
+In Section \ref{sec:results} we will show that V-cycle iterations, starting with some finest-mesh initial iterate $w^{J,0}$ and repeated until \eqref{eq:stoppingcriterion} holds, serve as an effective solver.
 However, an FMG cycle, Algorithm \ref{alg:fascd-fmg} below and Figure \ref{fig:fascdcycles} (right), is even better.
-The well-known principle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate $w^0\in\mathcal{K}^0$; one may generate it by injection and projection/truncation of $w^{J,0}$.
-Then an admissible initial iterate on the next-finer mesh is generated via prolongation.
-However, in the VI context one must also project (truncate) to the bounds after the prolongation; see lines \ref{line:fcycleprolongtruncone} and \ref{line:fcycleprolongtrunctwo} in Algorithm \ref{alg:fascd-fmg}.
+The well-known principle \cite[section 2.6]{Trottenbergetal2001} is to start on the coarsest mesh and solve the problem from any initial iterate $w^{0,0}\in\mathcal{K}^0$.
+(One may generate $w^{0,0}$ by nodal injection and projection/truncation of $w^{J,0}$.)  Each initial iterate on a next-finer level is generated via prolongation, but in our VI context one must also project (truncate) to the bounds to gain admissibility; see line \ref{line:fcycleprolongtruncone} in Algorithm \ref{alg:fascd-fmg} below.
 During this ``ramp'' stage one approximately solves on each level using a fixed number (\id{rampv}) of V-cycles, each down to the coarsest mesh.
 When the ramp is complete, one does full-depth V-cycles to convergence.
+When reporting iterations for the FMG cycle, ``$+n$'' means $n$ V-cycles were required to satisfy \eqref{eq:stoppingcriterion} after the FMG ramp, and thus ``$+0$'' means that the FMG cycle alone sufficed.
 
@@
     $\text{\pr{solve}}(\ell^0,\underline{\gamma}^0,\overline{\gamma}^0;w^0)$ \\
-    for $j=1$ to $j=J-1$ \\+
+    for $j=1$ to $j=J$ \\+
         $w^j = \max\left\{\underline{\gamma}^{j},\min\{\overline{\gamma}^{j}, Pw^{j-1}\}\right\} \in \mathcal{K}^j$ \label{line:fcycleprolongtruncone} \\
         $\pr{fascd-vcycle}^{\text{\id{rampv}}}(j,\ell^j,\underline{\gamma}^j,\overline{\gamma}^j;w^j)$ \\-
-    $w^J = \max\left\{\underline{\gamma}^{J},\min\{\overline{\gamma}^{J}, Pw^{J-1}\}\right\} \in \mathcal{K}^J$ \label{line:fcycleprolongtrunctwo} \\
     while not \eqref{eq:stoppingcriterion} \\+
@@
 
-In contrast to the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, during the ramp standard injection is used to generate the constraints $\underline{\gamma}^j,\overline{\gamma}^j$ on each level because a full solution iterate is initially needed, not just a correction.
+In contrast to the construction of LDCs (Section \ref{sec:cdmultilevel}) for V-cycles, during the ramp standard injection is used to generate the constraints $\underline{\gamma}^j,\overline{\gamma}^j$ on each level.
+This is because a full solution iterate is initially needed, not just a correction.
 Essentially, our choice is between \eqref{eq:fe:badstandard}, where the coarse constraint set is nonempty but then the prolonged solution may not be admissible, or \eqref{eq:fe:badmonotone}, where the constraint set may be empty.
-We choose \eqref{eq:fe:badstandard}, with nonempty sets, but we must project (truncate) pointwise to the bounds, potentially adding high-frequency noise to the initial iterate.
-
-A large space of smoothers could be explored, but in the next Section we will only apply an reduced-space Newton method \cite{Balayetal2023,BensonMunson2006}, denoted as \emph{RS Newton}.
-A fixed number of preconditioned Krylov iterations approximately solve the arising linear systems, thus $O(m_j)$ work is done per smoother application.
+We choose \eqref{eq:fe:badstandard}, with nonempty sets, but we project (truncate) pointwise to the bounds, potentially adding high-frequency noise to the initial iterate.
+
+A large space of smoothers could be explored, but in the next Section we will only apply a fixed number of iterations of a reduced-space Newton method \cite{Balayetal2023,BensonMunson2006}, denoted as \emph{RS Newton}.
+A fixed number of preconditioned Krylov iterations \cite{Bueler2021} approximately solve the arising linear systems, thus $O(m_j)$ work is done per smoother application on level $j$.
 For the symmetric Laplacian (Example \ref{ex:results:classical}) we apply conjugate gradients (CG) with incomplete Cholesky (IC) preconditioning, but otherwise generalized minimum residual (GMRES) with incomplete LU (ILU) preconditioning is used.
@@
 
-For $P_1$ and $Q_1$ elements, FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fmg} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}.
+For $P_1$ and $Q_1$ elements, FASCD Algorithms \ref{alg:fascd} and \ref{alg:fascd-fmg} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}, which applies solvers from the PETSc \cite{Balayetal2023} library.
 The numerical results in this Section, which follow the Section \ref{sec:vi} examples, focus on the scaling of iterations with mesh resolution.
-We see that each V-cycle greatly reduces the residual norm, and that a few FMG iterations solve most of these problems to within discretization error.
+We will see that each V-cycle greatly reduces the residual norm, and that the application of an FMG cycle solves most of these problems to within discretization error, or nearly so.
 
@@
 We consider two cases, over square domains $\Omega$, with different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$ (Figure \ref{fig:results:classical}).
-The standard ``ball'' problem has an upper hemisphere obstacle \cite[Chapter 12]{Bueler2021}, but the ``spiral'' problem's obstacle is constructed to produce the coincidence set shown \cite[problem 7.1.1]{GraeserKornhuber2009}.
-For both cases we used a regular coarsest mesh with $m_0=41$ nodes (64 triangles), and the finest mesh had $m_{10} \approx 3.4 \times 10^7$ nodes.
+The standard ``ball'' problem has the unit upper hemisphere as the obstacle \cite[Chapter 12]{Bueler2021}.
+The ``spiral'' problem's obstacle is constructed to produce the coincidence set shown \cite[problem 7.1.1]{GraeserKornhuber2009}.
+For both cases we used a regular coarsest mesh with $m_0=41$ nodes (64 triangles), and then the finest mesh had $m_{10} \approx 3.4 \times 10^7$ nodes after 10 refinements.
 A single RS Newton iteration with three IC-preconditioned CG iterations was used as the smoother, and the initial iterate was $w=\max\{0,\psi\}$.
@@
 Table \ref{tab:results:classical} shows the resulting iteration counts.
-For a first experiment we over-solve, with \texttt{atol} $=$ \texttt{rtol} $=$ \texttt{stol} $= 10^{-12}$ in stopping criterion \eqref{eq:stoppingcriterion}, to facilitate comparison to results in \cite[section 7]{GraeserKornhuber2009}.
+For a first experiment we over-solve, with \texttt{atol} $=$ \texttt{rtol} $=$ \texttt{stol} $= 10^{-12}$ in \eqref{eq:stoppingcriterion}, to facilitate comparison to results in \cite[section 7]{GraeserKornhuber2009}.
 As expected, RS Newton as a solver shows rapidly-increasing iterations, roughly proportional to resolution.
-By contrast, V-cycle iterations increase slowly, perhaps logarithmically with the resolution.
+By contrast, $V(1,1)$-cycle iterations increase slowly, perhaps logarithmically with the resolution.
 For the spiral problem, \cite[Figure 7.11]{GraeserKornhuber2009} reports that a $m_J \approx 5.2 \times 10^5$ problem needs approximately 25 V-cycle iterations for the best (truncated;  ``TNMG'' and ``TMMG'') monotone multilevel algorithms.\footnote{The stopping criterion and asymptotic convergence rates in \cite{GraeserKornhuber2009} are based on norm differences from a last numerical iterate with unknown accuracy.
-Our method directly measures satisfaction of the VI, via the residual norm $\|\rSS(w)\|_2$.}  The 12 V-cycle iterations needed here, at comparable resolution (8 levels), represent somewhat more arithmetic because our smoother is RS Newton using CG+IC, rather than the simpler projected Gauss-Seidel iteration.
+Our method directly measures satisfaction of the VI, via the residual norm $\|\rSS(w)\|_2$ (Section \ref{sec:implementation}).}  The 12 V-cycle iterations needed here for comparable resolution (8 levels) represent somewhat more arithmetic because our smoother is RS Newton using CG+IC, rather than the simpler projected Gauss-Seidel iteration.
+
 Resolution-dependent asymptotic convergence rates for V-cycles, defined as $\rho_J=(\|\rSS(w^{J,k})\|_2/\|\rSS(w^{J,0})\|_2)^{1/k}$ where $w^{J,k}$ is the first iterate to satisfy \eqref{eq:stoppingcriterion}, are shown in Figure \ref{fig:results:asymp}.
@@
 
-For FMG we instead used the defaults \texttt{rtol} $=$ \texttt{stol} $= 10^{-8}$, sufficient to reach discretization error, and in Table \ref{tab:results:classical} we report the number of V-cycles after the ramp; a small, constant number solves the problem.
+For the FMG cycle we instead used the defaults \texttt{rtol} $=$ \texttt{stol} $= 10^{-8}$, sufficient to reach discretization error.
+In Table \ref{tab:results:classical} we report the number of V-cycles after the ramp as ``$+n$''; a small, constant number solves the ball problem.
 Since each V-cycle does the work of about eight finest-level Krylov iterations, the ball results show nearly textbook multigrid efficiency at the highest resolutions despite the low solution regularity.
-More V-cycles are needed to clean-up error remaining after the ramp in the spiral problem.
+In the spiral problem, more V-cycles are needed to clean-up the remaining (free boundary) error.
 \end{example}
 
-%REGENERATE using git@bitbucket.org:pefarrell/fascd.git, in examples/
-%$ python3 drape.py -prob ball -levs 9 -o ball -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
-%$ mpiexec -n 16 python3 drape.py -prob spiral -levs 10 -o spiral.pvd -fascd_levels_snes_converged_reason -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
-%$ paraview ball|spiral.pvd
+%REGENERATE using git@bitbucket.org:pefarrell/fascd.git in examples/:
+%  $ python3 drape.py -prob ball -levs 9 -o ball -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
+%  $ mpiexec -n 16 python3 drape.py -prob spiral -levs 10 -o spiral.pvd -fascd_levels_snes_converged_reason -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
+%  $ paraview ball|spiral.pvd
 %use preset "xray" and invert colors; for spiral.pvd, show lgap = u - lb as black if above 0.02; for ball i used 0.001
@@
 % see git@bitbucket.org:pefarrell/fascd.git: examples/drape.py
-% and https://github.com/bueler/mcd-extended: paper/results/drape_{nofas,fascd}.sh|txt
+% see reproduce/table2.py
 \begin{table}[ht]
@@
 \midrule
- 1 &             $41$ &   1 &  1 &  1 &   2 &  1 &  1 \\
- 2 &            $145$ &   1 &  3 &  2 &   4 &  4 &  2 \\
- 3 &            $545$ &   3 &  6 &  3 &   6 &  5 &  3 \\
- 4 & $2.1\times 10^3$ &   6 &  7 &  3 &   7 &  7 &  4 \\
- 5 & $8.3\times 10^3$ &  13 &  9 &  4 &  11 &  9 &  4 \\
- 6 & $3.3\times 10^4$ &  23 & 11 &  5 &  20 & 10 &  5 \\
- 7 & $1.3\times 10^5$ &  46 & 11 &  4 &  35 & 11 &  5 \\
- 8 & $5.3\times 10^5$ &  90 & 12 &  4 &  67 & 12 &  5 \\
- 9 & $2.1\times 10^6$ & \XX & 13 &  3 & \XX & 13 &  6 \\
-10 & $8.4\times 10^6$ & \XX & 13 &  2 & \XX & 14 &  6 \\
-11 & $3.4\times 10^7$ & \XX & 14 &  2 & \XX & 18 &  6 \\
+% FIXME GENERATE WHOLE TABLE ON THELIO
+ 2 &            $145$ &   1 &  3 &  +2 &   4 &  4 &  +2 \\
+ 3 &            $545$ &   3 &  6 &  +3 &   6 &  5 &  +3 \\
+ 4 & $2.1\times 10^3$ &   6 &  7 &  +3 &   7 &  7 &  +4 \\
+ 5 & $8.3\times 10^3$ &  13 &  9 &  +4 &  11 &  9 &  +4 \\
+ 6 & $3.3\times 10^4$ &  23 & 11 &  +5 &  20 & 10 &  +5 \\
+ 7 & $1.3\times 10^5$ &  46 & 11 &  +4 &  35 & 11 &  +5 \\
+ 8 & $5.3\times 10^5$ &  90 & 12 &  +4 &  67 & 12 &  +5 \\
+ 9 & $2.1\times 10^6$ & \XX & 13 &  +3 & \XX & 13 &  +6 \\
+10 & $8.4\times 10^6$ & \XX & 13 &  +2 & \XX & 14 &  +6 \\
+11 & $3.4\times 10^7$ & \XX & 14 &  +2 & \XX & 18 &  +6 \\
 \bottomrule
@@
 \bigskip
-\caption{Iterations for FASCD V-cycle and FMG algorithms in Example \ref{ex:results:classical}; see text regarding tolerances.  Runs marked \XX\xspace were not attempted.}
+\caption{Iterations for Example \ref{ex:results:classical}; see text regarding cycle types and tolerances.
+Runs \XX\xspace were not attempted.
+``$+n$'' for FMG denotes number of V-cycles, after the FMG cycle, to satisfy \eqref{eq:stoppingcriterion}.}
 \label{tab:results:classical}
@@
 
-Iteration results in Table \ref{tab:results:fastplap1d} show a strong contrast between $V(1,0)$ and $V(0,1)$ results (\pr{up} $=0$ and \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively).
-The $V(1,0)$ cycles, Algorithm 4.2 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.
-By contrast, $V(0,1)$ cycles, up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are as good as the default $V(1,1)$ cycles.
+Iteration results in Table \ref{tab:results:fastplap1d} show a strong contrast between $V(1,0)$ and $V(0,1)$ results, chosen by setting \pr{up} $=0$ and \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively.
+The $V(1,0)$ cycles, essentially Algorithm 4.2 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.
+By contrast, $V(0,1)$ cycles, which up-smooth in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are as good as the default $V(1,1)$ cycles.
 FMG is even more efficient; a single iteration suffices at all resolutions.
@@
 % see git@bitbucket.org:pefarrell/fascd.git: examples/plap1d.py
-% and https://github.com/bueler/mcd-extended: paper/results/plap1d.sh|txt
+% and reproduce/table3.py
 \begin{table}[ht]
@@
 \midrule
- 1 &    $7$ &  1 &  1 &  1 &  1 & $2.3 \times 10^{-1}$ \\
- 2 &   $13$ &  3 &  2 &  2 &  1 & $3.3 \times 10^{-2}$ \\
- 3 &   $25$ & 12 &  4 &  2 &  1 & $9.1 \times 10^{-3}$ \\
- 4 &   $49$ & NC &  4 &  2 &  1 & $3.2 \times 10^{-3}$ \\
- 5 &   $97$ & NC &  3 &  3 &  1 & $5.5 \times 10^{-4}$ \\
- 6 &  $193$ & NC &  3 &  3 &  1 & $1.7 \times 10^{-4}$ \\
- 7 &  $385$ & NC &  3 &  3 &  1 & $4.7 \times 10^{-5}$ \\
- 8 &  $769$ & NC &  3 &  6 &  1 & $9.2 \times 10^{-6}$ \\
- 9 & $1537$ & NC &  3 &  3 &  1 & $3.4 \times 10^{-6}$ \\
-10 & $3073$ & NC &  2 &  4 &  1 & $4.1 \times 10^{-7}$ \\
+% FIXME GENERATE WHOLE TABLE ON THELIO
+ 2 &   $13$ &  3 &  2 &  2 &  +0 & $3.3 \times 10^{-2}$ \\
+ 3 &   $25$ & 12 &  4 &  2 &  +0 & $9.1 \times 10^{-3}$ \\
+ 4 &   $49$ & NC &  4 &  2 &  +0 & $3.2 \times 10^{-3}$ \\
+ 5 &   $97$ & NC &  3 &  3 &  +0 & $5.5 \times 10^{-4}$ \\
+ 6 &  $193$ & NC &  3 &  3 &  +0 & $1.7 \times 10^{-4}$ \\
+ 7 &  $385$ & NC &  3 &  3 &  +0 & $4.7 \times 10^{-5}$ \\
+ 8 &  $769$ & NC &  3 &  6 &  +0 & $9.2 \times 10^{-6}$ \\
+ 9 & $1537$ & NC &  3 &  3 &  +0 & $3.4 \times 10^{-6}$ \\
+10 & $3073$ & NC &  2 &  4 &  +0 & $4.1 \times 10^{-7}$ \\
 \bottomrule
@@
 \bigskip
-\caption{Iterations for V-cycles and FMG in Example \ref{ex:results:plap}.  NC $=$ did not converge after 50 iterations.}
+\caption{Iterations for Example \ref{ex:results:plap}.  Runs ``NC'' did not converge after 50 iterations.}
 \label{tab:results:fastplap1d}
@@
 Firedrake applications can exploit the PETSc \cite{Balayetal2023} library's parallel solvers.
-The above Examples, which were run in serial to fix the smoother, also run efficiently in parallel.
-The next Examples apply parallel solvers and investigate the weak scaling of FASCD FMG.
+The above Examples, which were run in serial, also run efficiently in parallel.
+The next Examples apply parallel solvers and investigate the weak scaling of FMG.
 
@@
 Consider the advection-diffusion VI problem in Example \ref{ex:advectiondiffusion} with $\Omega=(-1,1)^d$, $d=2,3$, diffusivity $\eps=0.1$ in \eqref{eq:advectiondiffusionstrong}, and zero Dirichlet boundary conditions.
-The 2D problem satisfies the coercivity hypothesis, with a divergence-free velocity field $\bX = (7+5y,-5x)$, but in 3D we set $\bX = (7+5y,-5x,2z)$ so $\grad\cdot\bX \ne 0$.
+The 2D problem satisfies the coercivity hypothesis, with a divergence-free velocity field $\bX = (7+5y,-5x)$.
+In 3D we set $\bX = (7+5y,-5x,2z)$ so $\grad\cdot\bX \ne 0$.
 The discontinuous source term $\phi \in L^\infty(\Omega)$ satisfies $\phi\ge 0$ in half of the domain, with positive circular patches, but values are negative in the other half; details are in the associated code.
 While the other Examples in this Section are unilateral, here the solution is box-constrained: $0 \le u \le 1$.
-Figure \ref{fig:results:advdiff} shows coincidence sets for the 2D problem; the 3D sets are comparable in regularity and general pattern.
+Figure \ref{fig:results:advdiff} shows coincidence sets for the 2D problem.
+The 3D sets are comparable in regularity and general pattern.
 
@@
 A coarse mesh with $m_0=16^d$ vertices minimally-resolved the coincidence set.
-The local Péclet number $\text{Pe}=\|\bX(x,y,z)\|_2 L \eps^{-1}$, with $L=2$ the width of $\Omega$, varied from 40 to 260, and the mesh Péclet number $m_J^{-1/d}\,\text{Pe}$ varied further with resolution, so the $m_J\le 121^d$ problems were advection-dominated, and those with $m_J\ge 481^d$ diffusion-dominated.
-However, the smoother was the same in all cases: two (fixed) RS Newton iterations, with three preconditioned (additive Schwarz method (ASM) with ILU on each process) GMRES iterations.
+The local Péclet number $\text{Pe}=\|\bX(x,y,z)\|_2 L \eps^{-1}$, with $L=2$ the width of $\Omega$, varied from 40 to 260, and the mesh Péclet number $m_J^{-1/d}\,\text{Pe}$ varied further with resolution.
+The $m_J\le 121^d$ problems were numerically advection-dominated and those with $m_J\ge 481^d$ diffusion-dominated.
+However, the smoother was the same in all cases: two (fixed) RS Newton iterations, with three preconditioned GMRES iterations.
+The preconditioner was the additive Schwarz method (ASM) with ILU on each process.
 Tolerances were set to \id{rtol} $=10^{-5}$ and \id{atol} $=$ \id{stol} $= 10^{-9}$.
-FMG iteration results are shown in Table \ref{tab:results:advdiff}; a few iterations sufficed up to these memory-limited resolutions.
+FMG cycle results are shown in Table \ref{tab:results:advdiff}; up to memory-limited resolution a couple of V-cycle iterations sufficed to satisfy \eqref{eq:stoppingcriterion}.
 \end{example}
@@
 % see git@bitbucket.org:pefarrell/fascd.git: examples/pollutant.py
-% and https://github.com/bueler/mcd-extended: paper/results/pollutant.sh|txt
+% and reproduce/table4.py
 \begin{table}[ht]
@@
 \midrule
- 1 &    $16^d$ & 1 & 1 \\
- 2 &    $31^d$ & 2 & 2 \\
- 3 &    $61^d$ & 2 & 2 \\
- 4 &   $121^d$ & 2 & 2 \\
- 5 &   $241^d$ & 2 & 3 \\
- 6 &   $481^d$ & 2 \\
- 7 &   $961^d$ & 3 \\
- 8 &  $1921^d$ & 3 \\
- 9 &  $3841^d$ & 3 \\
+ 2 &    $31^d$ & +1 & +1 \\
+ 3 &    $61^d$ & +1 & +1 \\
+ 4 &   $121^d$ & +1 & +1 \\
+ 5 &   $241^d$ & +1 & +2 \\
+ 6 &   $481^d$ & +1 \\
+ 7 &   $961^d$ & +2 \\
+ 8 &  $1921^d$ & +2 \\
+ 9 &  $3841^d$ & +2 \\
 \bottomrule
@@
 \bigskip
-\caption{FMG iterations for 2D ($P_1$, serial) and 3D ($Q_1$, 16 processes) solutions of the advection-diffusion bi-obstacle VI problem in Example \ref{ex:results:advdiff}.}
+\caption{FMG cycle results for 2D ($P_1$, serial) and 3D ($Q_1$, 16 processes) solutions of Example \ref{ex:results:advdiff}.}
 \label{tab:results:advdiff}

### 2023-12-26 paper/fascd.tex
done with wave of edits

@@
 \begin{example}   \label{ex:results:sia}
-Consider the steady shallow ice approximation (SIA) obstacle problem (Example \ref{ex:sia}), in which the surface elevation is constrained to exceed the bedrock elevation, $s\ge b$, on $\Omega=(0,L)^2$ with $L=1800$ km.
+Recall the steady shallow ice approximation (SIA) obstacle problem in Example \ref{ex:sia}, in which the surface elevation is constrained to exceed the bedrock elevation, $s\ge b$.
+Let $\Omega=(0,L)^2$ with $L=1800$ km.
 The data consist of a radial surface mass balance function $a$ \cite[equation (5.122)]{GreveBlatter2009} and a bumpy, synthetic map $b\in C^1(\Omega)$, with an elevation range of $1800$ m; details are in the associated code.
-A numerical solution is shown in Figure \ref{fig:results:siascene}, colored by the solution surface flow speed, the magnitude of $\mathbf{u} = (n+1)(n+2)^{-1} \Gamma (s-b)^{n+1} |\grad s|^{n-1} \grad s$.
+A numerical solution is shown in Figure \ref{fig:results:siascene}, colored by surface flow speed, namely the magnitude of $\mathbf{u} = (n+1)(n+2)^{-1} \Gamma (s-b)^{n+1} |\grad s|^{n-1} \grad s$ \cite{GreveBlatter2009}.
 Note $|\grad s|$ is unbounded at the free boundary; the solution $s$, while continuous, has very low regularity.
@@
 
-We applied FASCD FMG using $Q_1$ elements and a smoother of four (fixed) iterations of RS Newton (Section \ref{sec:implementation}), with quadratic backtracking line search.
-The arising linear systems were approximately solved by three (fixed) iterations of preconditioned GMRES and an ASM+ILU preconditioner.
-The coarse mesh was relatively fine, with $m_0=21^2$ vertices and 400 cells.
+We applied FMG using $Q_1$ elements and a smoother of four (fixed) iterations of RS Newton (Section \ref{sec:implementation}), with quadratic backtracking line search, and two (\texttt{rampv} $=2$ in Algorithm \ref{alg:fascd-fmg}) V-cycles per level.
+Three iterations of preconditioned GMRES, with an ASM+ILU preconditioner, were applied to the Newton linear systems.
+The coarse mesh was relatively fine, with $m_0=21^2$ nodes.
 Because of the low regularity of the solution we used \id{rtol} $= 2 \times 10^{-4}$ and \id{atol} $= 10^{-8}$.
@@
 
-For a weak-scaling study we fixed the number of degrees of freedom per process at $m_J/P=641^2 \approx 4.1 \times 10^5$ and solved with $6, \dots, 11$ levels.\footnote{The experiments were conducted on ARCHER2, the UK national supercomputer.
-Each node has 16 memory channels, so a maximum of 16 cores per node were employed.}  The results in Table \ref{tab:results:siaweak} show some degradation in time per cycle with increasing core counts, perhaps due to the use of one-level additive Schwarz in the Krylov iteration.
-Nevertheless, the finest mesh result, with $m_J=20481^2=4.1 \times 10^8$ degrees of freedom, required less than three minutes to solve for the steady surface elevation of an ice sheet larger than the current Greenland ice sheet ($1.7$ million $\text{km}^2$) at uniform $87.5$ m resolution.
+For a weak-scaling study, over $P$ processes we fixed the number of degrees of freedom per process at $m_J/P \approx 4.1 \times 10^5$ and solved with $J+1=6, \dots, 11$ levels.\footnote{The experiments were conducted on ARCHER2, the UK national supercomputer.
+Each node has 16 memory channels, so a maximum of 16 cores per node were employed.}  The results in Table \ref{tab:results:siaweak} show some degradation in run time with increasing core counts, perhaps due to the use of single-level additive Schwarz in the Krylov iteration.
+Nevertheless, the finest mesh result, with $m_J=20481^2=4.1 \times 10^8$ degrees of freedom, required less than three minutes to solve for the steady surface elevation of an ice sheet larger in area than the current Greenland ice sheet ($1.7$ million $\text{km}^2$), at uniform $87.5$ m resolution.
 
 % see git@bitbucket.org:pefarrell/fascd.git: examples/sia.py
-% and https://github.com/bueler/mcd-extended: paper/results/siaweak.sh|txt
+% and reproduce/table5.sh
 \begin{table}[ht]
@@
 \toprule
-$P$ & \emph{levels} & $m_J$ & \emph{iterations} & \emph{time} (s) & \emph{time$/$iteration} (s) \\
+$P$ & \emph{levels} & $m_J$ & \emph{iterations} & \emph{time} (s) \\
 \midrule
- 1 & 6 & $641^2$ & 3 & 98 & 32.7 \\
- 4 & 7 & $1281^2$ & 3 & 98 & 32.7 \\
- 16 & 8 & $2561^2$ & 3 & 124 & 41.3 \\
- 64 & 9 & $5121^2$ & 3 & 136 & 45.3 \\
- 256 & 10 & $10241^2$ & 2 & 95 & 47.5 \\
- 1024 & 11 & $20481^2$ & 2 & 177 & 88.5 \\
+ 1 & 6 & $641^2$ & +2 & 98 \\
+ 4 & 7 & $1281^2$ & +2 & 98 \\
+ 16 & 8 & $2561^2$ & +2 & 124 \\
+ 64 & 9 & $5121^2$ & +2 & 136 \\
+ 256 & 10 & $10241^2$ & +1 & 95 \\
+ 1024 & 11 & $20481^2$ & +1 & 177 \\
  \bottomrule
@@
 \bigskip
-\caption{Iterations and wall clock time for parallel FASCD FMG solutions of Example \ref{ex:results:sia} on $P$ processes.}
+\caption{Parallel weak scaling FMG cycle results (V-cycle iterations to satisfy \eqref{eq:stoppingcriterion}, and wall clock time) for Example \ref{ex:results:sia}.}
 \label{tab:results:siaweak}
@@
 
-As a multilevel solver for VI problems, the FASCD method is a strategy for generating coarser-level problems which, once solved, make helpful admissible additions to a V-cycle update.
+As a multilevel solver for VI problems, the FASCD method is a strategy for generating coarser-level problems which, once solved, make helpful admissible corrections within a V-cycle iteration.
 The method is smoother-agnostic, and our implementation within the extensible Firedrake \cite{Rathgeberetal2016} and PETSc \cite{Balayetal2023} library framework will allow easy experimentation with other smoothers (projected Gauss--Seidel, semi-smooth Newton, interior-point, etc.), beyond the active-set Newton method chosen for demonstration here.
-For smoothers of Newton-Krylov type one could also add linear geometric or algebraic multigrid preconditioning \cite{Trottenbergetal2001} for the linear step problems, instead of our simpler choice of incomplete factorization preconditioners.
+For smoothers of Newton-Krylov type one could also add linear geometric or algebraic multigrid preconditioning \cite{Trottenbergetal2001} for the linear Newton step problems, instead of our simpler choice of incomplete factorization.
 In a different direction, a restructured multilevel CD approach which applies the additive/parallel CD iteration \cite{Tai2003} might be considered, in pursuit of a VI algorithm analogous to the BPX multigrid method for PDEs \cite{BramblePasciakXu1990}.
@@
 
-\subsection*{Code availability} \label{sec:code}  The software used to produce the results is archived at tag \texttt{v1.0} in the repository \texttt{\url{https://bitbucket.org/pefarrell/fascd/}}, using the Firedrake version \texttt{3fb16ad47}.
+\subsection*{Code availability} \label{sec:code}  The software used to produce the results is archived at tag \texttt{v1.0} in the repository \texttt{\url{https://bitbucket.org/pefarrell/fascd/}}, using Firedrake version \texttt{3fb16ad47}.
 

### 2023-12-27 paper/reviews/response.tex
actually write some responses

@@
 
-FIXME The thoughtful comments of the two reviewers are much appreciated.
-Below we respond to their concerns---in block italics---and, in the important cases, quote the new or rewritten content of the paper which addresses these concerns.
-The aggregate responses to these comments have made the paper about one page longer.
+The thoughtful comments from the two reviewers are much appreciated.
+Below we respond to their concerns.
+The paper was revised based on these comments and concerns, and some additional edits were made to improve readability.
+(A colored \LaTeX difference PDF is included to show the extent of the revisions.)  Our revisions have not significantly lengthened the paper.
 
@@
 
-%%% ELB CONFIDENTIAL  I am guessing this is Tai himself.  I will look at Tai and Espedal 1998 to see if a citation is appropriate.
+%%% ELB CONFIDENTIAL  I am guessing this is Tai himself.
 
@@
 I strongly suggest accepting this work for SISC after some minor revisions suggested below.
+\end{review}
 
+\short{This summary is accurate, and the expression of support is appreciated!}
+
+\begin{review}
 The reasons that I suggest accepting this paper are based on the following observations:
@@
 
-\short{FIXME This summary is accurate and appreciated.}
+\short{The reviewer points out our new algorithm's avoidance of the ``need to go to the finest level.''  This idea is now a sentence in our revised Introduction.
+We also appreciate the pointer to Tai \& Espedal (1998), which shows convergence of a space decomposition method even each subspace solver is only approximate.
+If we could supply a convergence proof this would be relevant, but we do not see how to do this currently.}
 
@@
 \begin{response}
-FIXME good point
+We agree that this is the weakest point of the work.
+Sadly, a convergence proof is not yet available.
+We believe that the algorithm is sufficiently powerful as to be worthwhile publishing, and that our numerical demonstrations adequately support this claim.
 \end{response}
@@
 
-\short{xx}
+\short{Our numerical demonstrations support the claim of optimality, and the reviewer seems to agree.
+Regarding the need to supply comparisons, we think that our already-included comparison to results from [22] (Graeser \& Kornhuber, 2009), who implemented the [32] (Tai, 2003) method using a constraint decomposition like ours, are more relevant.
+(This comparison is discussed near the end of Example 8.1.)  Our convergence rates are much better than those in [32], and comparable to those in [22], with our rates about $0.10$ compared to $>0.70$ for [32] on the same ``ball'' problem at comparable resolution; smaller numbers are better.
+However, the improvement on this simplest Laplacian obstacle problem is largely our better smoother.}
 

### 2023-12-28 paper/fascd.tex
put in a convergence figure, in response to reviewer 1

@@
 Our method directly measures satisfaction of the VI, via the residual norm $\|\rSS(w)\|_2$ (Section \ref{sec:implementation}).}  The 12 V-cycle iterations needed here for comparable resolution (8 levels) represent somewhat more arithmetic because our smoother is RS Newton using CG+IC, rather than the simpler projected Gauss-Seidel iteration.
+
+The ball problem has an accurately-known solution $u\in W^{2,\infty}(\Omega)$\cite{Bueler2021,Tai2003}.
+Figure \ref{fig:results:convergence} shows that the $L^2$ convergence rate is $O(h^2)$ as expected for $P_1$ elements FIXME: CITE CIARLET?
+GLOWINSKI?
 
@@
 \includegraphics[width=0.65\textwidth]{asymprates.pdf}
-\caption{V-cycle asymptotic convergence rates in Example \ref{ex:results:classical}.}
+\caption{V-cycle asymptotic rates in Example \ref{ex:results:classical}.}
 \label{fig:results:asymp}
+\end{figure}
+
+%REGENERATE using reproduce/table2.py at git@bitbucket.org:pefarrell/fascd.git
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.65\textwidth]{convergence.pdf}
+\caption{Convergence rate for the ball problem in Example \ref{ex:results:classical}.}
+\label{fig:results:convergence}
 \end{figure}

### 2023-12-29 paper/fascd.tex
remove the Figure I just put in, and a few more small edits

--- 2023-12-26 paper/fascd.tex
@@
 Furthermore, each of its coarser-level problems refers to the finest-level discretization.
+FASCD is both more general, as it applies to non-variational monotone operators (Section \ref{sec:vi}), and more efficient on each coarser level.
 \item The last-mentioned limitation of the \cite{Tai2003} method is removed by Gr\"aser \& Kornhuber's \cite[Algorithm 4.2]{GraeserKornhuber2009} reformulation using defect constraints and monotone injection operators.
@@
 The ``spiral'' problem's obstacle is constructed to produce the coincidence set shown \cite[problem 7.1.1]{GraeserKornhuber2009}.
-For both cases we used a regular coarsest mesh with $m_0=41$ nodes (64 triangles), and then the finest mesh had $m_{10} \approx 3.4 \times 10^7$ nodes after 10 refinements.
+For both cases we used $P_1$ elements on a regular coarsest mesh with $m_0=41$ nodes (64 triangles), and then the finest mesh had $m_{10} \approx 3.4 \times 10^7$ nodes after 10 refinements.
 A single RS Newton iteration with three IC-preconditioned CG iterations was used as the smoother, and the initial iterate was $w=\max\{0,\psi\}$.
@@
 Our method directly measures satisfaction of the VI, via the residual norm $\|\rSS(w)\|_2$ (Section \ref{sec:implementation}).}  The 12 V-cycle iterations needed here for comparable resolution (8 levels) represent somewhat more arithmetic because our smoother is RS Newton using CG+IC, rather than the simpler projected Gauss-Seidel iteration.
+
+The ball problem has a known exact solution $u\in W^{2,\infty}(\Omega)$ \cite{Bueler2021,Tai2003}.
+Our measured convergence rate is $\|u^\star-u^J\|_{L^2(\Omega)}=O(h^{1.98})$, as expected for $P_1$ elements.
+FIXME: CITE CIARLET OR GLOWINSKI
 
@@
 \includegraphics[width=0.65\textwidth]{asymprates.pdf}
-\caption{V-cycle asymptotic convergence rates in Example \ref{ex:results:classical}.}
+\caption{V-cycle asymptotic rates in Example \ref{ex:results:classical}.}
 \label{fig:results:asymp}
@@
 FMG is even more efficient; a single iteration suffices at all resolutions.
-Note that convergence is at rate $O(h^{2.0})$; compare \cite{ChoeLewis1991}.
+Note that measured $L^\infty(\Omega)$ convergence is $O(h^{2.00})$; compare \cite{ChoeLewis1991}.
 \end{example}
@@
 \bigskip
-\caption{Parallel weak scaling FMG cycle results (V-cycle iterations to satisfy \eqref{eq:stoppingcriterion}, and wall clock time) for Example \ref{ex:results:sia}.}
+\caption{Parallel weak scaling FMG cycle results (additional V-cycle iterations to satisfy \eqref{eq:stoppingcriterion}, and wall clock time) for Example \ref{ex:results:sia}.}
 \label{tab:results:siaweak}

### 2023-12-29 paper/reviews/response.tex
complete draft of responses to reviews

@@
 
-\short{The reviewer points out our new algorithm's avoidance of the ``need to go to the finest level.''  This idea is now a sentence in our revised Introduction.
-We also appreciate the pointer to Tai \& Espedal (1998), which shows convergence of a space decomposition method even each subspace solver is only approximate.
-If we could supply a convergence proof this would be relevant, but we do not see how to do this currently.}
+\begin{response}
+Appreciated.  The idea that our new algorithm avoids the ``need to go to the finest level''  is now a sentence in the Introduction.
+
+We also appreciate the pointer to Tai \& Espedal (1998), which shows convergence of a space decomposition method when each subspace solver is only approximate (and satisfying certain bounds).
+If we could supply a convergence proof this would be relevant, but we do not see how to do it currently.
+\end{response}
 
@@
 \short{Our numerical demonstrations support the claim of optimality, and the reviewer seems to agree.
-Regarding the need to supply comparisons, we think that our already-included comparison to results from [22] (Graeser \& Kornhuber, 2009), who implemented the [32] (Tai, 2003) method using a constraint decomposition like ours, are more relevant.
-(This comparison is discussed near the end of Example 8.1.)  Our convergence rates are much better than those in [32], and comparable to those in [22], with our rates about $0.10$ compared to $>0.70$ for [32] on the same ``ball'' problem at comparable resolution; smaller numbers are better.
-However, the improvement on this simplest Laplacian obstacle problem is largely our better smoother.}
+Regarding the need to supply comparisons, we think that our already-included comparison is adequate.
+In Example 8.1 we compare to results from [22] (Graeser \& Kornhuber, 2009), who have  implemented the [32] (Tai, 2003) method using a constraint decomposition like ours, whereas the [32] results use less efficient constraint decompositions.
+Our convergence rates are much better than those in [32], and comparable to those in [22], with our rates about $0.10$ compared to $>0.70$ for [32] on the same ``ball'' problem at comparable resolution.
+(Smaller numbers are better here.)  On the other hand, the improvement seen in this simplest Laplacian obstacle problem is largely from our better smoother.}
 
@@
 
-\short{xx}
+\short{In two cases the exact solution is known (``ball'' problem in Example 8.1; Example 8.2).
+The $P_1$ convergence plots themselves are too boring to display, but convergence order of $\|u^\star - u^J\|$ is now reported for both cases.}
 
@@
 
-\short{FIXME This summary is accurate and appreciated.}
+\short{This summary is accurate and appreciated.}
 
@@
 
-\short{FIXME ``common'' $\to$ ``joint''}
+\short{The word ``common'' has been replaced by ``joint'' for clarity.
+All that is intended is that FASCD extends two previously-separate algorithmic ideas, namely FAS and CD.
+Note that the desirable properties are listed clearly in the Introduction.}
 
@@
 
-\short{FIXME yes, indeed four projects (which makes the reviewers point). cut into paragraphs}
+\short{This is quite helpful.
+Indeed we seek to contrast with four projects, i.e.~[6,32,22,24] (old numbering), which makes the reviewer's point!
+This part of the text is now an itemized list, thus easier for the reader to follow.}
 
@@
 
-\short{FIXME appreciated.  mostly kept the same, but we think the greater organization of prior work paragraphs makes referring to numbered properties unneeded}
+\short{While keeping the overall structure mostly the same, we have made an effort to clean and clarify the presentation of desirable properties (an enumerated list) and of prior work (an itemized list).}
 
@@
 
-\short{xx}
+\short{An effort has been made to clarify the presentation of the FAS idea.
+Regarding existing literature, it remains a relatively awkward idea to explain; apparently our presentation is not exceptional.}
 
@@
 FMG is not an iterative method, yet you start with ``FASCD \dots by a small, mesh-independent number of FMG iterations \dots''.
-\end{review}
 
-\short{xx}
-
-\begin{review}
 The FAS algorithm with V-cycles attached to make an iterative method to arbitrary solver tolerance is fine, but reporting multiple iteration is not really correct.
@@
 
-\short{xx}
+\short{The reviewer is correct, and we have settled on a presentation close to that suggested.
+As explained now in the text---see Section 7 before Algorithm 7.1, and at first use in Example 8.1---notation ``$+n$'' is used when $n$ additional V-cycles were required to satisfy the stopping criterion (7.3).
+Thus ``$+0$'' appears when the FMG cycle itself sufficed, e.g.~(new) Table 3.
+}
 
@@
 
-\short{xx}
+\short{Indeed, in Table 5 the time-divided-by-iterations column is not helpful, and we have removed it.
+The ``iterations'' column is now ``$+n$'' as described above.
+We believe this addresses the reviewer's (totally reasonable!) critique.}
 
@@
 
-\short{xx}
+\begin{response}
+The description of ``strange things with the constraints in the down smoothing'' is understandable, but this is inherited from reference [22].
+As we try (hard?) to explain in Section 5, in down-smoothing the constraint sets must be small; this holds already in [22].
+We have improved upon [22] by recognizing that such small sets can be avoided in up-smoothing.
+
+One point of the Example in question (i.e.~Example 8.2) is, indeed, that small constraint sets are bad for the kind of highly nonlinear VI problems we are interested in solving.
+(Note that [22] does not address $p$-Laplacian cases for other than the classical $p=2$ case.)  Exactly why convergence is damaged by too-small sets, in a case like this, might be revealed by a convergence proof, i.e.~the one we don't have.
+
+The subtle and very grid-dependent difference between the better V(0,1) and V(1,1) cycles is something we do not understand.
+It maybe connected to the location of the free boundary relative to the nearest grid point, but this is both too useless and too speculative to include in the paper.
+\end{response}
 

### 2023-12-30 paper/reviews/response.tex
Edits to the response

@@
 The thoughtful comments from the two reviewers are much appreciated.
-Below we respond to their concerns.
-The paper was revised based on these comments and concerns, and some additional edits were made to improve readability.
-(A colored \LaTeX difference PDF is included to show the extent of the revisions.)  Our revisions have not significantly lengthened the paper.
+We have revised the manuscript in response.
+For the convenience of the Referees we include a latexdiff to show the specific changes.
 
@@
 
-\short{This summary is accurate, and the expression of support is appreciated!}
+\short{Thank you for this supportive and accurate summary.}
 
@@
 \begin{response}
-Appreciated.  The idea that our new algorithm avoids the ``need to go to the finest level''  is now a sentence in the Introduction.
+Thank you for raising these points. We now emphasize that [32] requires visiting the finest level in the introduction.
 
 We also appreciate the pointer to Tai \& Espedal (1998), which shows convergence of a space decomposition method when each subspace solver is only approximate (and satisfying certain bounds).
-If we could supply a convergence proof this would be relevant, but we do not see how to do it currently.
+This could form an important part of an eventual proof of convergence.
 \end{response}
@@
 We believe that the algorithm is sufficiently powerful as to be worthwhile publishing, and that our numerical demonstrations adequately support this claim.
+We also hope that by publishing this algorithm we might stimulate theoretical investigations by others.
 \end{response}
@@
 \short{Our numerical demonstrations support the claim of optimality, and the reviewer seems to agree.
-Regarding the need to supply comparisons, we think that our already-included comparison is adequate.
-In Example 8.1 we compare to results from [22] (Graeser \& Kornhuber, 2009), who have  implemented the [32] (Tai, 2003) method using a constraint decomposition like ours, whereas the [32] results use less efficient constraint decompositions.
-Our convergence rates are much better than those in [32], and comparable to those in [22], with our rates about $0.10$ compared to $>0.70$ for [32] on the same ``ball'' problem at comparable resolution.
+We already supply comparisons in Section 8; in Example 8.1 we compare to results from [22] (Graeser \& Kornhuber, 2009), who have implemented the [32] (Tai, 2003) method using a constraint decomposition like ours, whereas the results in [32] use less efficient [FIXME: less efficient in what sense?] constraint decompositions.
+Our achieved convergence rate is about $0.10$, comparable to those in [22], whereas [32] achieves $>0.70$ on the same ``ball'' problem at comparable resolution.
 (Smaller numbers are better here.)  On the other hand, the improvement seen in this simplest Laplacian obstacle problem is largely from our better smoother.}
@@
 
-\short{In two cases the exact solution is known (``ball'' problem in Example 8.1; Example 8.2).
-The $P_1$ convergence plots themselves are too boring to display, but convergence order of $\|u^\star - u^J\|$ is now reported for both cases.}
+\short{We now report the convergence order of $\|u^\star - u^J\|$ for the two cases where the exact solution is known (``ball'' problem in Example 8.1; Example 8.2).}
 
@@
 
-\short{The word ``common'' has been replaced by ``joint'' for clarity.
+\short{Ah, we meant ``common'' as in ``joint'' rather than ``frequent''.
+The word ``common'' has been replaced by ``joint'' for clarity.
 All that is intended is that FASCD extends two previously-separate algorithmic ideas, namely FAS and CD.
@@
 
-\short{This is quite helpful.
+\short{Thank you for this helpful suggestion.
 Indeed we seek to contrast with four projects, i.e.~[6,32,22,24] (old numbering), which makes the reviewer's point!
@@
 
-\short{An effort has been made to clarify the presentation of the FAS idea.
+\short{We have made an effort to clarify the presentation of the FAS idea.
 Regarding existing literature, it remains a relatively awkward idea to explain; apparently our presentation is not exceptional.}
@@
 The ``iterations'' column is now ``$+n$'' as described above.
-We believe this addresses the reviewer's (totally reasonable!) critique.}
+We believe this addresses the reviewer's very reasonable criticism.}
 
@@
 The description of ``strange things with the constraints in the down smoothing'' is understandable, but this is inherited from reference [22].
-As we try (hard?) to explain in Section 5, in down-smoothing the constraint sets must be small; this holds already in [22].
+As we try to explain in Section 5, in down-smoothing the constraint sets must be small; this holds already in [22].
 We have improved upon [22] by recognizing that such small sets can be avoided in up-smoothing.
@@
 The subtle and very grid-dependent difference between the better V(0,1) and V(1,1) cycles is something we do not understand.
-It maybe connected to the location of the free boundary relative to the nearest grid point, but this is both too useless and too speculative to include in the paper.
+It may be connected to the location of the free boundary relative to the nearest grid point, but this is too speculative for us to feel comfortable including a remark the manuscript.
 \end{response}

### 2024-01-08 paper/fascd.tex
Edits to the response
table 3 reproduced

@@
 \item The method is strictly admissible, and so can be applied even if the residual operator is defined only on those functions which satisfy the constraints.
-\item The work at each level is proportional to the number of nodes on that level.  Coarser-level evaluations do not require finest-level information.
+\item The work at each level is proportional to the number of degrees of freedom on that level. Coarser-level evaluations do not require finest-level information.
 \item V-cycles (Algorithm \ref{alg:fascd}) and full multigrid (FMG; \cite{Trottenbergetal2001}) cycles (Algorithm \ref{alg:fascd-fmg}) are implemented.
@@
 FMG is even more efficient; a single iteration suffices at all resolutions.
-Note that measured $L^\infty(\Omega)$ convergence is $O(h^{2.00})$; compare \cite{ChoeLewis1991}.
+Note that measured $L^\infty(\Omega)$ convergence is $O(h^{1.99})$; compare \cite{ChoeLewis1991}.
 \end{example}
@@
 \midrule
-% FIXME GENERATE WHOLE TABLE ON THELIO
  2 &   $13$ &  3 &  2 &  2 &  +0 & $3.3 \times 10^{-2}$ \\
@@
  4 &   $49$ & NC &  4 &  2 &  +0 & $3.2 \times 10^{-3}$ \\
- 5 &   $97$ & NC &  3 &  3 &  +0 & $5.5 \times 10^{-4}$ \\
+ 5 &   $97$ & 14 &  3 &  3 &  +0 & $5.5 \times 10^{-4}$ \\
  6 &  $193$ & NC &  3 &  3 &  +0 & $1.7 \times 10^{-4}$ \\
@@
  9 & $1537$ & NC &  3 &  3 &  +0 & $3.4 \times 10^{-6}$ \\
-10 & $3073$ & NC &  2 &  4 &  +0 & $4.1 \times 10^{-7}$ \\
+10 & $3073$ & NC &  2 &  5 &  +0 & $4.1 \times 10^{-7}$ \\
 \bottomrule

### 2024-01-09 paper/fascd.tex
had not noticed before now that Ciarlet states K_h (notsubset) K too

@@
 While \eqref{eq:fe:vi} approximates \eqref{eq:vi}, we warn the reader that it is usually not a conforming FE method.
-That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, from definitions \eqref{eq:originalconstraintset} and \eqref{eq:fe:fineconstraintset} we do not expect $\mathcal{K}^J \subset \mathcal{K}$ in general.
+That is, even if quadrature and other implementation details for $f^J$ are ignored, and despite the vector space inclusion $\mathcal{V}^J \subset \mathcal{V}$, from definitions \eqref{eq:originalconstraintset} and \eqref{eq:fe:fineconstraintset} we do not expect $\mathcal{K}^J \subset \mathcal{K}$ in general \cite[section 5.1]{Ciarlet2002}.
 Even for continuous obstacles $\underline{\gamma}, \overline{\gamma}$, FE admissibility \eqref{eq:fe:fineconstraintset} will generally not imply $\mathcal{K}$-admissibility \eqref{eq:originalconstraintset}.

### 2024-01-09 paper/fascd.tex
table 2 reproduced
replace fixme (citation) with fixme (actual rate I will now compute)

@@
 
-The ball problem has a known exact solution $u\in W^{2,\infty}(\Omega)$ \cite{Bueler2021,Tai2003}.
-Our measured convergence rate is $\|u^\star-u^J\|_{L^2(\Omega)}=O(h^{1.98})$, as expected for $P_1$ elements.
-FIXME: CITE CIARLET OR GLOWINSKI
+FIXME RATE  The ball problem has a known exact solution $u\in W^{2,\infty}(\Omega)$ \cite{Bueler2021,Tai2003}.
+Our measured convergence rate is $\|u^\star-u^J\|_{H^1(\Omega)}=O(h^{1.xx})$; compare $O(h^1)$ from theory \cite[section 5.1]{Ciarlet2002}.
 
@@
 \midrule
-% FIXME GENERATE WHOLE TABLE ON THELIO
- 2 &            $145$ &   1 &  3 &  +2 &   4 &  4 &  +2 \\
- 3 &            $545$ &   3 &  6 &  +3 &   6 &  5 &  +3 \\
- 4 & $2.1\times 10^3$ &   6 &  7 &  +3 &   7 &  7 &  +4 \\
- 5 & $8.3\times 10^3$ &  13 &  9 &  +4 &  11 &  9 &  +4 \\
- 6 & $3.3\times 10^4$ &  23 & 11 &  +5 &  20 & 10 &  +5 \\
- 7 & $1.3\times 10^5$ &  46 & 11 &  +4 &  35 & 11 &  +5 \\
- 8 & $5.3\times 10^5$ &  90 & 12 &  +4 &  67 & 12 &  +5 \\
- 9 & $2.1\times 10^6$ & \XX & 13 &  +3 & \XX & 13 &  +6 \\
-10 & $8.4\times 10^6$ & \XX & 13 &  +2 & \XX & 14 &  +6 \\
-11 & $3.4\times 10^7$ & \XX & 14 &  +2 & \XX & 18 &  +6 \\
+2 & 145 & 1 & 3 & +1 & 4 & 4 & +1 \\ 
+3 & 545 & 3 & 6 & +2 & 6 & 5 & +2 \\ 
+4 & $2.1\times 10^3$ & 6 & 7 & +2 & 7 & 7 & +3 \\ 
+5 & $8.3\times 10^3$ & 13 & 9 & +3 & 11 & 9 & +3 \\ 
+6 & $3.3\times 10^4$ & 23 & 11 & +4 & 20 & 10 & +4 \\ 
+7 & $1.3\times 10^5$ & 46 & 11 & +3 & 35 & 11 & +4 \\ 
+8 & $5.3\times 10^5$ & 90 & 12 & +3 & 67 & 12 & +4 \\ 
+9 & $2.1\times 10^6$ & \XX & 13 & +2 & \XX & 13 & +5 \\ 
+10 & $8.4\times 10^6$ & \XX & 13 & +1 & \XX & 14 & +5 \\ 
+11 & $3.4\times 10^7$ & \XX & 14 & +1 & \XX & 18 & +5 \\ 
 \bottomrule

### 2024-01-10 paper/fascd.tex
measured convergence rate for Example 8.1

@@
 
-FIXME RATE  The ball problem has a known exact solution $u\in W^{2,\infty}(\Omega)$ \cite{Bueler2021,Tai2003}.
-Our measured convergence rate is $\|u^\star-u^J\|_{H^1(\Omega)}=O(h^{1.xx})$; compare $O(h^1)$ from theory \cite[section 5.1]{Ciarlet2002}.
+The ball problem has a known exact solution $u\in W^{2,\infty}(\Omega)$ \cite{Bueler2021,Tai2003}.
+The measured convergence rate is $\|u^\star-u^J\|_{H^1(\Omega)}=O(h^{1.51})$, somewhat better than $O(h^1)$ provable from theory \cite[section 5.1]{Ciarlet2002}.
 

### 2024-01-10 paper/fascd.tex
table 4 preproduced
add zenodo to code availability; we should be set for reproducibility badge

@@
  5 &   $241^d$ & +1 & +2 \\
- 6 &   $481^d$ & +1 \\
- 7 &   $961^d$ & +2 \\
- 8 &  $1921^d$ & +2 \\
- 9 &  $3841^d$ & +2 \\
+ 6 &   $481^d$ & +1 &  \\
+ 7 &   $961^d$ & +2 &  \\
+ 8 &  $1921^d$ & +2 &  \\
+ 9 &  $3841^d$ & +2 &  \\
 \bottomrule
@@
 
-\subsection*{Code availability} \label{sec:code}  The software used to produce the results is archived at tag \texttt{v1.0} in the repository \texttt{\url{https://bitbucket.org/pefarrell/fascd/}}, using Firedrake version \texttt{3fb16ad47}.
+\subsection*{Code availability} \label{sec:code}  The software used to produce the results in Section \ref{sec:results} is archived at tag \texttt{v1.0} in the repository \texttt{\url{https://bitbucket.org/pefarrell/fascd/}} and at \href{https://doi.org/10.5281/zenodo.10476845}{\texttt{doi:10.5281/zenodo.10476845}}.
+We used Firedrake master revision \texttt{c5e939dde}.
 

### 2024-01-10 paper/reviews/response.tex
tiny changes to language; this is being re-submitted

@@
 We have revised the manuscript in response.
-For the convenience of the Referees we include a latexdiff to show the specific changes.
+For the convenience of the Editor and Referees we include a latexdiff to show the specific changes.
 
@@
 \short{Our numerical demonstrations support the claim of optimality, and the reviewer seems to agree.
-We already supply comparisons in Section 8; in Example 8.1 we compare to results from [22] (Graeser \& Kornhuber, 2009), who have implemented the [32] (Tai, 2003) method using a constraint decomposition like ours, whereas the results in [32] use less efficient [FIXME: less efficient in what sense?] constraint decompositions.
+We already supply comparisons in Section 8; in Example 8.1 we compare to results from [22] (Graeser \& Kornhuber, 2009), who have implemented the [32] (Tai, 2003) method using a constraint decomposition like ours, whereas the results in [32] use rather different constraint decompositions.
 Our achieved convergence rate is about $0.10$, comparable to those in [22], whereas [32] achieves $>0.70$ on the same ``ball'' problem at comparable resolution.
@@
 
-\short{We now report the convergence order of $\|u^\star - u^J\|$ for the two cases where the exact solution is known (``ball'' problem in Example 8.1; Example 8.2).}
+\short{We now report the convergence order of $\|u^\star - u^J\|$ for the two cases where the exact solution is known, namely the ``ball'' problem in Example 8.1 and the $p$-Laplacian problem in Example 8.2.}
 
@@
 \short{The reviewer is correct, and we have settled on a presentation close to that suggested.
-As explained now in the text---see Section 7 before Algorithm 7.1, and at first use in Example 8.1---notation ``$+n$'' is used when $n$ additional V-cycles were required to satisfy the stopping criterion (7.3).
-Thus ``$+0$'' appears when the FMG cycle itself sufficed, e.g.~(new) Table 3.
+As explained now in the text, notation ``$+n$'' is now used when $n$ additional V-cycles were required to satisfy the stopping criterion (7.3).
+(This is documented in Section 7 before Algorithm 7.1, and at first use in Example 8.1.)  Thus ``$+0$'' appears when the FMG cycle itself sufficed, e.g.~(new) Table 3.
 }
@@
 
-\short{Indeed, in Table 5 the time-divided-by-iterations column is not helpful, and we have removed it.
+\short{The reviewer is correct that in Table 5 the time-divided-by-iterations column was not helpful, and we have removed it.
 The ``iterations'' column is now ``$+n$'' as described above.
@@
 The description of ``strange things with the constraints in the down smoothing'' is understandable, but this is inherited from reference [22].
-As we try to explain in Section 5, in down-smoothing the constraint sets must be small; this holds already in [22].
+As we try to explain in Section 5, in down-smoothing the constraint sets must be small; this characterizes the cycles in [22].
 We have improved upon [22] by recognizing that such small sets can be avoided in up-smoothing.
@@
 One point of the Example in question (i.e.~Example 8.2) is, indeed, that small constraint sets are bad for the kind of highly nonlinear VI problems we are interested in solving.
-(Note that [22] does not address $p$-Laplacian cases for other than the classical $p=2$ case.)  Exactly why convergence is damaged by too-small sets, in a case like this, might be revealed by a convergence proof, i.e.~the one we don't have.
+(Note that [22] does not address $p$-Laplacian cases, for other than the classical $p=2$ case, while we address the $1.5$-Laplacian and a doubly-degenerate version of the $4$-Laplacian.)  Exactly why convergence is damaged by too-small sets, in a case like this, might be revealed by a convergence proof, i.e.~the proof we don't have.
 
-The subtle and very grid-dependent difference between the better V(0,1) and V(1,1) cycles is something we do not understand.
-It may be connected to the location of the free boundary relative to the nearest grid point, but this is too speculative for us to feel comfortable including a remark the manuscript.
+The subtle and very grid-dependent difference between the better V(0,1) and V(1,1) cycles is also something we do not understand.
+It may be connected to the location of the free boundary relative to the nearest grid point, but this is too speculative for us to feel comfortable including as a remark in the manuscript.
 \end{response}

### 2024-01-10 paper/fascd.tex
tiny changes to language; this is being re-submitted

@@
 By contrast, $V(0,1)$ cycles, which up-smooth in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are as good as the default $V(1,1)$ cycles.
-FMG is even more efficient; a single iteration suffices at all resolutions.
+FMG is even more efficient; a single cycle suffices at all resolutions.
 Note that measured $L^\infty(\Omega)$ convergence is $O(h^{1.99})$; compare \cite{ChoeLewis1991}.
@@
 The above Examples, which were run in serial, also run efficiently in parallel.
-The next Examples apply parallel solvers and investigate the weak scaling of FMG.
+The next two Examples apply parallel solvers and investigate the weak scaling of FMG.
 
@@
 
-For a weak-scaling study, over $P$ processes we fixed the number of degrees of freedom per process at $m_J/P \approx 4.1 \times 10^5$ and solved with $J+1=6, \dots, 11$ levels.\footnote{The experiments were conducted on ARCHER2, the UK national supercomputer.
+For a weak-scaling study, on $P$ processes we fixed the number of degrees of freedom per process at $m_J/P \approx 4.1 \times 10^5$ and solved with $J+1=6, \dots, 11$ levels.\footnote{The experiments were conducted on ARCHER2, the UK national supercomputer.
 Each node has 16 memory channels, so a maximum of 16 cores per node were employed.}  The results in Table \ref{tab:results:siaweak} show some degradation in run time with increasing core counts, perhaps due to the use of single-level additive Schwarz in the Krylov iteration.

### 2024-07-09 talk-ismp/references.tex
start on slides for Montreal

--- 2023-11-05 talk-dms/references.tex
@@
 \begin{itemize}
-\item A.~Brandt (1977). \emph{Multi-level adaptive solutions to boundary-value problems}, Mathematics of Computation 31 (138), 333--390
-%\item A.~Brandt \& C.~Cryer (1983). \emph{Multigrid algorithms for the solution of linear complementarity problems \dots}, SIAM J.~Sci.~Stat.~Comput.~4 (4), 655--684 \sdoi{10.1137/0904046}
-%%FTITLE Multigrid algorithms for the solution of linear complementarity problems arising from free boundary problems
-%\item P.~Brune, M.~Knepley, B.~Smith, \& X.~Tu (2015). \emph{Composing scalable nonlinear algebraic solvers}, SIAM Review~57 (4), 535--565 \sdoi{10.1137/130936725}
-\item E.~Bueler (2021). \emph{Conservation laws for free-boundary fluid layers}, SIAM J.~Appl.~Math.~81 (5), 2007--2032 \sdoi{10.1137/20M135217X}
-%\item E.~Bueler (2022). \emph{Performance analysis of high-resolution ice-sheet simulations}, J.~Glaciol., \sdoi{10.1017/jog.2022.113}
-\item E.~Bueler \& P.~Farrell (2023).
-\emph{A full approximation scheme multilevel method for nonlinear variational inequalities}, submitted \, {\tiny \href{https://arxiv.org/abs/2308.06888}{arxiv:2308.06888}}
-%\item G.~Gonzalez de Diego, P.~Farrell, \& I.~Hewitt (2022).
-\emph{Numerical approximation of viscous contact problems applied to glacial sliding}, J.~Fluid Mech.~938 (A21).
-\sdoi{10.1017/jfm.2022.178}
-%\item P.~Farrell, M.~Croci, \& T.~Surowiec (2020). \emph{Deflation for semismooth equations}, Optimization Methods \& Software 35 (6), 1248--1271 \sdoi{10.1080/10556788.2019.1613655}
+\item E.~Bueler \& P.~Farrell (2024).
+\emph{A full approximation scheme multilevel method for nonlinear variational inequalities}, to appear SIAM J.~Sci.~Comput.
+\, {\tiny \href{https://arxiv.org/abs/2308.06888}{arxiv:2308.06888}}
 \item C.~Gr{\"a}ser \& R.~Kornhuber (2009). \emph{Multigrid methods for obstacle problems}, J.~Comput.~Math., 1--44
-%\item T.~Isaac, G.~Stadler, \& O.~Ghattas (2015). \emph{Solution of nonlinear Stokes equations \dots ice sheet dynamics}, SIAM J.~Sci.~Comput., 37 (6), B804--B833 \sdoi{10.1137/140974407}
-%FTITLE Solution of nonlinear Stokes equations discretized by high-order finite elements on nonconforming and anisotropic meshes, with application to ice sheet dynamics
 \item G.~Jouvet \& E.~Bueler (2012).
 \emph{Steady, shallow ice sheets as obstacle problems: well-posedness and finite element approximation}, SIAM J.~Appl.~Math.~72 (4), 1292--1314 \sdoi{10.1137/110856654}
-%\item G.~Jouvet \& J.~Rappaz (2011). \emph{Analysis and finite element approximation of a nonlinear stationary {S}tokes problem \dots}, Adv.~Numer.~Analysis 2011 (164581) \sdoi{10.1155/2011/164581}
-%FTITLE Analysis and finite element approximation of a nonlinear stationary {S}tokes problem arising in glaciology
-%\item R.~Kornhuber (1994). \emph{Monotone multigrid methods for elliptic variational inequalities I}, Numer.~Math.~69, 167--184 \sdoi{10.1007/BF03325426}
-\item A.~Reusken (1987).
-\emph{Convergence of the multigrid full approximation scheme for a class of elliptic mildly nonlinear boundary value problems}, Numer.~Math.~52, 251--277 \sdoi{10.1007/BF01398879}
-%\item J.~Seguinot \& 5 others (2018).  \emph{Modelling last glacial cycle ice dynamics in the Alps}, The Cryosphere, 12 (10), 3265--3285 \sdoi{10.5194/tc-12-3265-2018}
-%FAUTHORS Seguinot, J., Ivy-Ochs, S., Jouvet, G., Huss, M., Funk, M., & Preusser, F.
 \item X.~Tai (2003). \emph{Rate of convergence for some constraint decomposition methods for nonlinear variational inequalities}, Numer.~Math.~93 (4), 755--786 \sdoi{10.1007/s002110200404}

### 2024-07-09 talk-ismp/slides.tex
start on slides for Montreal

--- 2023-11-12 talk-dms/slides.tex
@@
 \hypersetup{
-    pdftitle = {Fast solvers for PDEs subject to inequalities},
+    pdftitle = {A full approximation scheme multilevel method for solving nonlinear variational inequalities},
     pdfauthor = {Ed Bueler and Patrick Farrell},
@@
 
-\title[Fast solvers for PDE subject to inequalities]{Fast solvers for partial differential equations \\ subject to inequalities}
+\title[FAS multigrid for VIs]{A full approximation scheme multilevel method \\ for solving nonlinear variational inequalities}
 
@@
 
-\date[]{November 2023}
+\date[]{ISMP 2024}
 
@@
 
-\begin{frame}{Outline: \emph{Fast solvers for PDEs subject to inequalities}}
+\begin{frame}{Outline}
   \tableofcontents[hideallsubsections]
@@
 \section{variational inequalities (VIs)}
+
+\begin{frame}{{\color{FireBrick} variational inequality} $=$ constrained weak form equations}
+
+\begin{center}
+\begin{tabular}{r|l|l}
+& \qquad unconstrained & \qquad constrained \\ \hline
+optimization &
+\begin{minipage}[t][16mm][t]{0.32\textwidth}
+$$\min_{u\in\mathcal{V}} J(u)$$
+\end{minipage}
+&
+\begin{minipage}[t][16mm][t]{0.35\textwidth}
+$$\min_{u\in\mathcal{K}} J(u)$$
+\end{minipage}
+\\ \hline
+\only<1>{equations}\only<2>{\begin{minipage}[t][16mm][t]{0.15\textwidth} weak form \par equations \end{minipage}} &
+\begin{minipage}[t][16mm][t]{0.32\textwidth}
+
+\vspace{-2mm}
+find $u \in \mathcal{V}$:
+\only<1>{$$F(u)=0$$}
+\only<2>{$$\ip{F(u)}{v} = 0 \quad \forall v \in \mathcal{V}$$}
+\end{minipage}
+&
+\begin{minipage}[t][16mm][t]{0.35\textwidth}
+
+\vspace{-2mm}
+\only<2>{
+{\color{FireBrick} find $u \in \mathcal{K}$:}
+$${\color{FireBrick} \ip{F(u)}{v-u} \ge 0 \quad \forall v \in \mathcal{K}}$$
+}
+\end{minipage}
+\end{tabular}
+\end{center}
+
+\bigskip
+
+\begin{itemize}
+\item where $\mathcal{K} \subset \mathcal{V}$ is closed and convex
+\end{itemize}
+\end{frame}
 
@@
 
-\only<1-2>{
 \begin{itemize}
-\item \emph{problem.} on a domain $\Omega \subset \RR^2$, find the displacement $u(x)$ of a membrane, with fixed value $u = g$ on $\partial \Omega$, above an \emph{obstacle} $\psi(x)$, which minimizes elastic (plus some potential) energy
+\item \emph{problem.} on $\Omega \subset \RR^2$, find $u(x)$ which minimizes
     $$J(v) = \int_\Omega \frac{1}{2} |\grad v|^2 - f\, v$$
-\item shown above: \quad $\Omega$ a square, $\psi(x)$ a hemisphere
-\item<2>[\alert{Q.}] how to solve this as a PDE with boundary conditions?
+over $\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \,\text{ and }\, v \ge \psi\right\}$
 \end{itemize}
-
-\vspace{2mm}
-}
-\only<3>{
-\begin{itemize}
-\item this is constrained optimization over an infinite-dimensional \emph{admissible set}
-	$$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \,\text{ and }\, v \ge \psi\right\}$$
-
-    \begin{itemize}
-    \item[$\circ$] $\mathcal{K}$ is a closed and convex subset of the Sobolev space
-   $$H^1(\Omega) = \left\{v \,:\, \int_\Omega |v|^2 + |\grad v|^2 < \infty\right\}$$
-    \end{itemize}
-\end{itemize}
-}
 \end{frame}
@@
 \bigskip
-\only<1>{
 \begin{itemize}
@@
    \end{itemize}
-\end{itemize}
-
-\vspace{25mm}
-}
-\only<2>{
-\begin{itemize}
-\item a naive strong form would pose the problem in terms of its solution:
-\begin{align*}
--\grad^2 u &= f \quad \text{ on $R_u$} \\
-u &= \psi \quad \text{ on $A_u$}
-\end{align*}
+\item the \emph{complementarity problem} (CP) is a PDE-like strong form:
+\begin{equation*}
+u - \psi \ge 0, \qquad -\grad^2 u - f \ge 0, \qquad (u - \psi)(-\grad^2 u - f) = 0
+\end{equation*}
 
    \begin{itemize}
-   \item[$\circ$] Poisson equation $-\grad^2 u = f$ is ``$J'(u)=0$'' on $R_u$
-   \item[$\circ$] using the solution $u$ to define the set $R_u$ on which to solve the PDE $-\grad^2 u = f$ \alert{does not lead to solution algorithms}
+   \item[$\circ$] CP $=$ KKT conditions essentially, but $\infty$-dimensional
    \end{itemize}
-\end{itemize}
-}
-\only<3>{
-\begin{itemize}
-\item the \emph{complementarity problem} (CP) is a meaningful strong form:
-\begin{align*}
-u - \psi &\ge 0 \\
--\grad^2 u - f &\ge 0 \\
-(u - \psi)(-\grad^2 u - f) &= 0
-\end{align*}
-
-   \begin{itemize}
-   \item[$\circ$] CP $=$ KKT conditions
-       \begin{itemize}
-       \item but in $\infty$-dimensions
-       \end{itemize}        
-   \end{itemize}
-
-\phantom{x}
-\end{itemize}
-}
-\only<4>{
-\begin{itemize}
 \item the weak form is a {\color{FireBrick} \emph{variational inequality} (VI)}, which says that $J'(u)$ points directly into $\mathcal{K}$:
@@
 for all $v \in \mathcal{K}$
-
-\vspace{11mm}
 \end{itemize}
-}
 \end{frame}
@@
 \begin{center}
-\includegraphics[width=0.5\textwidth]{figs/convexuv.png}
+\includegraphics[width=0.5\textwidth]{../talk-dms/figs/convexuv.png}
 \end{center}
@@
 
-\begin{frame}{VI $=$ constrained ``system of equations''}
-
-\begin{center}
-\begin{tabular}{r|l|l}
-& \qquad unconstrained & \qquad constrained \\ \hline
-optimization &
-\begin{minipage}[t][16mm][t]{0.32\textwidth}
-$$\min_{u\in\mathcal{V}} J(u)$$
-\end{minipage}
-&
-\begin{minipage}[t][16mm][t]{0.35\textwidth}
-$$\min_{u\in\mathcal{K} \subset \mathcal{V}} J(u)$$
-\end{minipage}
-\\ \hline
-\only<1>{equations}\only<2>{\begin{minipage}[t][16mm][t]{0.15\textwidth} weak form \par equations \end{minipage}} &
-\begin{minipage}[t][16mm][t]{0.32\textwidth}
-
-\vspace{-2mm}
-find $u \in \mathcal{V}$:
-\only<1>{$$F(u)=0$$}
-\only<2>{$$\ip{F(u)}{v} = 0 \quad \forall v \in \mathcal{V}$$}
-\end{minipage}
-&
-\begin{minipage}[t][16mm][t]{0.35\textwidth}
-
-\vspace{-2mm}
-find $u \in \mathcal{K} \subset \mathcal{V}$:
-$${\color{FireBrick} \ip{F(u)}{v-u} \ge 0 \quad \forall v \in \mathcal{K}}$$
-\end{minipage}
-\end{tabular}
-\end{center}
-\end{frame}
-
 \begin{frame}{applications of VIs}
@@
 \vspace{-10mm}
-\hfill \includegraphics[width=0.2\textwidth]{figs/tirecontact.png}
+\hfill \includegraphics[width=0.2\textwidth]{../talk-dms/figs/tirecontact.png}
 
@@
 
-\section{nonlinear multigrid for PDEs}
-
-\subsection{full approximation scheme (FAS)}
+\section{nonlinear (FAS) multigrid for PDEs}
 
@@
 \centering
-\input{tikz/fcycle.tex}
+\input{../talk-dms/tikz/fcycle.tex}
 
@@
 \begin{column}{0.55\textwidth}
-\includegraphics[width=\textwidth]{figs/bratu-time.png}
+\includegraphics[width=\textwidth]{../talk-dms/figs/bratu-time.png}
 \end{column}
@@
 
-\section{multigrid for VIs}
-
-\subsection{full approximation scheme constraint decomposition (FASCD)}
+\section{FASCD = FAS multigrid for VIs}
 
@@
 \vspace{-25mm}
-\hfill \mbox{\input{tikz/Rplusminus.tex}}
+\hfill \mbox{\input{../talk-dms/tikz/Rplusminus.tex}}
 \end{frame}
@@
 \centering
-\includegraphics[width=0.7\textwidth]{figs/innerconeapprox.png}
+\includegraphics[width=0.7\textwidth]{../talk-dms/figs/innerconeapprox.png}
 \end{frame}
@@
 \centering
-\includegraphics[width=0.85\textwidth]{figs/vcycle-visualized.png}
+\includegraphics[width=0.85\textwidth]{../talk-dms/figs/vcycle-visualized.png}
 \end{frame}
@@
 
-\subsection{classical obstacle problem}
-
-\subsection{advection-diffusion of a concentration}
-
-\subsection{glacier surface elevations}
-
 \begin{frame}{classical obstacle problem by FASCD}
 
-\only<1>{\includegraphics[width=0.75\textwidth]{figs/ballitersV.png}}\only<2>{\includegraphics[width=0.75\textwidth]{figs/ballitersVlog.png}}\only<3>{\includegraphics[width=0.75\textwidth]{figs/ballitersVF.png}}\only<4>{\includegraphics[width=0.75\textwidth]{figs/bothitersVF.png}} \hfill
+\only<1>{\includegraphics[width=0.75\textwidth]{../talk-dms/figs/ballitersV.png}}\only<2>{\includegraphics[width=0.75\textwidth]{../talk-dms/figs/ballitersVlog.png}}\only<3>{\includegraphics[width=0.75\textwidth]{../talk-dms/figs/ballitersVF.png}}\only<4>{\includegraphics[width=0.75\textwidth]{../talk-dms/figs/bothitersVF.png}} \hfill
 \only<1-3>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/ball-set.png}}%
@@
 \begin{center}
-\includegraphics[width=0.75\textwidth]{figs/advdiff.png}
+\includegraphics[width=0.75\textwidth]{../talk-dms/figs/advdiff.png}
 \end{center}
@@
 \centering
-\includegraphics[width=0.55\textwidth]{figs/siaweaktime.png}
+\includegraphics[width=0.55\textwidth]{../talk-dms/figs/siaweaktime.png}
 \end{frame}
@@
 
-%{\scriptsize
-%{\notsotiny
 {\footnotesize
@@
 
-\begin{frame}{background references}
-
-%{\scriptsize
-%{\notsotiny
-{\footnotesize
-\input{background.tex}
-}
-\end{frame}
-
 \end{document}

### 2024-07-21 talk-ismp/slides.tex
work some on the talk

@@
 
-\titlegraphic{\vspace{-8mm} \includegraphics[width=0.26\textwidth]{../talk-oxford/images/uafbw.png} \hfill \begin{minipage}{0.14\textwidth}
+\titlegraphic{\vspace{-18mm} \includegraphics[width=0.22\textwidth]{../talk-oxford/images/uafbw.png} \hfill \begin{minipage}{0.12\textwidth}
 \includegraphics[width=\textwidth]{../talk-oxford/images/oxford.png}
@@
   \begin{frame}
+
+\vspace{10mm}
+
     \titlepage
@@
 \begin{frame}{Outline}
-  \tableofcontents[hideallsubsections]
+  \tableofcontents
 \end{frame}
@@
 \\ \hline
-\only<1>{equations}\only<2>{\begin{minipage}[t][16mm][t]{0.15\textwidth} weak form \par equations \end{minipage}} &
+\only<1>{equations}\only<2->{\begin{minipage}[t][16mm][t]{0.15\textwidth} weak form \par equations \end{minipage}} &
 \begin{minipage}[t][16mm][t]{0.32\textwidth}
@@
 \only<1>{$$F(u)=0$$}
-\only<2>{$$\ip{F(u)}{v} = 0 \quad \forall v \in \mathcal{V}$$}
+\only<2->{$$\ip{F(u)}{v} = 0 \quad \forall v \in \mathcal{V}$$}
 \end{minipage}
@@
 \vspace{-2mm}
-\only<2>{
+\only<3>{
 {\color{FireBrick} find $u \in \mathcal{K}$:}
@@
 \item where $\mathcal{K} \subset \mathcal{V}$ is closed and convex
+\item<3> sometimes the {\color{FireBrick} variational inequality (VI)} is the KKT condition of optimization ($F=J'$), and sometimes not
 \end{itemize}
@@
     $$J(v) = \int_\Omega \frac{1}{2} |\grad v|^2 - f\, v$$
-over $\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \,\text{ and }\, v \ge \psi\right\}$
+over
+    $$\mathcal{K} = \left\{v \in H^1(\Omega) \,:\, v\big|_{\partial \Omega} = g \,\text{ and }\, v \ge \psi\right\}$$
 \end{itemize}
@@
 
-\begin{frame}{example: classical obstacle problem}
+\begin{frame}{example, cont.}
 
-\includegraphics[width=0.55\textwidth]{../talk-oxford/images/obstacle65.pdf} \qquad \includegraphics[width=0.35\textwidth]{../talk-oxford/images/obstacle-sets.png}
+\begin{minipage}[t]{0.55\textwidth}
+\vspace{0pt}
+\includegraphics[width=\textwidth]{../talk-oxford/images/obstacle65.pdf}
+\end{minipage}
+\hfill
+\begin{minipage}[t]{0.3\textwidth}
+\vspace{2mm}
+\includegraphics[width=\textwidth]{../talk-oxford/images/obstacle-sets.png}
+\end{minipage}
 
-\bigskip
+\vspace{-1mm}
 \begin{itemize}
@@
 \end{equation*}
-
-   \begin{itemize}
-   \item[$\circ$] CP $=$ KKT conditions essentially, but $\infty$-dimensional
-   \end{itemize}
-\item the weak form is a {\color{FireBrick} \emph{variational inequality} (VI)}, which says that $J'(u)$ points directly into $\mathcal{K}$:
-    $$\ip{J'(u)}{v-u} = \int_\Omega \grad u\cdot \grad (v-u) - f (v-u) \ge 0$$
-for all $v \in \mathcal{K}$
+\item the weak form is the VI:
+    $$\int_\Omega \grad u\cdot \grad (v-u) - f (v-u) \ge 0 \quad \forall v \in \mathcal{K}$$
 \end{itemize}
@@
 
-\begin{frame}{VI $=$ weak form}
+\begin{frame}{geometrical meaning of VI}
 
 \begin{itemize}
-\item for problems of optimization type, the VI is the weak form, with $v-u$ as the test function:
-{\small
-   $$J(u) \le J(v) \quad \forall v \in \mathcal{K} \qquad \iff \qquad \ip{J'(u)}{v-u} \ge 0 \quad \forall v \in \mathcal{K}$$
-}
+\item the VI gives an ``angle condition'' for the test function $v-u$:
+\item in optimization problems the condition is a consequence:
+\begin{align*}
+J(u) &\le J(v) \quad \forall v \in \mathcal{K} \\
+\implies \qquad \ip{J'(u)}{v-u} &\ge 0 \quad \forall v \in \mathcal{K}
+\end{align*}
+\item for general VIs the condition \emph{is} the problem:
+   $$\ip{F(u)}{v-u} \ge 0 \quad \forall v \in \mathcal{K}$$    
 \end{itemize}
@@
 \begin{center}
-\includegraphics[width=0.5\textwidth]{../talk-dms/figs/convexuv.png}
+\includegraphics[width=0.4\textwidth]{../talk-dms/figs/convexuv.png}
 \end{center}
@@
     \item[$\circ$] $F$ may \emph{not}\, be the derivative of an objective function $J$
-    \item[$\circ$] $F=J'$, a linear operator, in classical obstacle problem
     \end{itemize}
@@
 	$${\color{FireBrick} \ip{F(u)}{v-u} \ge 0 \quad \text{ for all } v \in \mathcal{K}}$$
-\item when $\mathcal{K}$ is nontrivial the problem {\color{FireBrick} VI($F$,$\mathcal{K}$)} is nonlinear \emph{even when $F$ is a linear operator}
+\item when $\mathcal{K}$ is nontrivial, VI($F$,$\mathcal{K}$) is nonlinear even when $F$ is a linear operator
 \end{itemize}
@@
 
-\vspace{3mm}
-\item first-semester calculus:
-    $$u \gets \min_{x\in[a,b]} f(x) \quad \iff \quad f'(u)(v-u) \ge 0 \quad \forall v \in[a,b]$$
+\vspace{1.5mm}
+\item<2> first-semester calculus:
+    $$u = \mathop{\textnormal{argmin}}_{x\in[a,b]} f(x) \quad \iff \quad f'(u)(v-u) \ge 0 \quad \forall v \in[a,b]$$
 \end{itemize}
 
-%\vspace{20mm}
+\vspace{-5mm}
 \begin{center}
-\includegraphics[width=0.9\textwidth]{../talk-oxford/images/calcone.png}
+\only<1>{\vspace{20.5mm}}
+
+\includegraphics<2>[height=25mm]{../talk-oxford/images/calcone.png}
 \end{center}
@@
 
-\section{nonlinear (FAS) multigrid for PDEs}
+\section{nonlinear multigrid for PDEs}
 
-\begin{frame}{nonlinear 2-mesh scheme}
+\subsection{full approximation scheme (FAS)}
+
+\begin{frame}{nonlinear 2-mesh scheme for PDEs}
 
@@
 	\begin{itemize}
-	\item[$\circ$] $u\in \mathcal{V}=H^1(\Omega)$
-	\item[$\circ$] $\ell\in \mathcal{V}'$
-	\item[$\circ$] $F : \mathcal{V} \to \mathcal{V}'$ continuous and one-to-one
-	\item[$\circ$] for example, the Liouville-Bratu problem: $-\grad^2 u - e^u = f$
+	\item[$\circ$] $\mathcal{V}=H^1(\Omega)$, $F : \mathcal{V} \to \mathcal{V}'$ continuous, $\ell\in \mathcal{V}'$
+	\item[$\circ$] example is the Liouville-Bratu problem: $-\grad^2 u - e^u = f$
 	\end{itemize}
 \item discretization gives algebraic system on fine mesh $\Omega^h$:
-    $$F^h(u^h) = \ell^h$$
-
-	\begin{itemize}
-	\item[$\circ$] $u^h$ denotes exact (algebraic) solution
-	\end{itemize}
+    $${\color{FireBrick} F^h(u^h) = \ell^h}$$
 \end{itemize}
@@
     \begin{enumerate}
-    \item the \emph{residual} \quad $r^h = \ell^h - F^h(w^h)$ \quad is computable
-    \item the \emph{error} \quad $e^h = u^h-w^h$ \quad is unknown
-    \item our equation can be rewritten
-    $$F^h(u^h) - F^h(w^h) = r^h$$
+    \item \emph{residual} $r^h = \ell^h - F^h(w^h)$ is computable but \emph{error} $e^h = u^h-w^h$ is unknown
+    \item equation can be rewritten as \qquad $F^h(u^h) - F^h(w^h) = r^h$
     \end{enumerate}
@@
 }
-\only<3-4>{
+\only<3>{
 \begin{itemize}
-\item \emph{updated goal}: from iterate $w^h$, to solve
-    $$F^h(u^h) - F^h(w^h) = r^h$$
-\item \alert{for $F^h$ linear}, convert this to the \emph{error equation}
+\item \emph{goal}: from iterate $w^h$, to solve {\color{FireBrick} $F^h(u^h) - F^h(w^h) = r^h$}
+\item for $F^h$ linear, one may convert to the error equation
     $$F^h(e^h) = r^h$$
-\item an approximation solution $\tilde e^h$ would improve our iterate:
+with approximate solution $\tilde e^h$, to improve the iterate
     $$w^h \leftarrow w^h+\tilde e^h$$
-\item<4> but $F^h$ is not linear!
+\end{itemize}
+
+\vspace{5mm}
+}
+\only<4>{
+\begin{itemize}
+\item \emph{updated goal}: use a coarser mesh $\Omega^H$ to somehow estimate the solution $u^h$ in the nonlinear \emph{correction equation}
+    $${\color{FireBrick} F^h(u^h) - F^h(w^h) = r^h}$$
+\item \emph{smoothers}: there are inexpensive algorithms which improve $w^h$ by efficiently removing high-frequency error components
+	\begin{itemize}
+	\item sweeping through and solving nodewise problems (nonlinear Gauss-Seidel) is a smoother
+	\end{itemize}
+\item how about low-frequency error components?
 \end{itemize}
@@
 \begin{itemize}
-\item \emph{updated goal}: use a coarser mesh $\Omega^H$ to somehow estimate the solution $u^h$ in the nonlinear \emph{correction equation}
-    $${\color{FireBrick} F^h(u^h) - F^h(w^h) = r^h}$$
-\item basic multigrid idea: there are algorithms (\alert{smoothers}) which ``improve'' $w^h$ \dots use them a little first \dots then correct from the coarser mesh
-    \begin{itemize}
-    \item[$\circ$] ``improve'' means they remove high-frequency error components efficiently
-    \end{itemize}
-\end{itemize}
-
-\vspace{15mm}
-}
-\only<6>{
-\begin{itemize}
-\item \emph{nodewise problem}: for $\psi_i^h$ a hat function or dof, solve for $c\in\RR$ to make the residual at that location zero:
-	$${\color{FireBrick} \phi_i(c) = r^h(w^h + c \psi_i^h)[\psi_i^h] = 0}$$
-\item sweeping through and solving nodewise problems is a \alert{smoother}
-    \begin{itemize}
-    \item[$\circ$] Fourier analysis shows smoothing property
-    \item[$\circ$] after smoothing, $e^h$ and $r^h$ have smaller high-frequencies
-    \end{itemize}
-\item after smoothing, the correction equation on $\Omega^h$ should be accurately approximate-able on the coarser mesh $\Omega^H$
-\end{itemize}
-
-\vspace{14mm}
-}
-\only<7>{
-\begin{itemize}
-\item \emph{updated goal}: use a coarser mesh $\Omega^H$ to somehow estimate the solution $u^h$ in $F^h(u^h) - F^h(w^h) = r^h(w^h)$
 \item Brandt's (1977) \emph{full approximation scheme} (FAS) equation:
@@
 
-\item rewritten FAS equation: let ${\color{FireBrick} \ell^H = F^H(\iR w^h) + R\, r^h(w^h)}$ then
+\item rewritten FAS equation: let $\ell^H = F^H(\iR w^h) + R\, r^h(w^h)$ then
     $${\color{FireBrick} F^H(u^H) = \ell^H}$$
@@
 
-\begin{frame}{full approximation scheme (FAS) 2-mesh solver}
+\begin{frame}{full approximation scheme (FAS): 2-mesh cycle}
 
@@
 \begin{align*}
-&\text{pre-smooth over fine:} & & [\text{smoother updates } w^h] \\
+&\text{pre-smooth over fine:} & & \text{smoother on } w^h \\
 &\text{restrict:}                   & &\ell^H = F^H(\iR w^h) + R\, r^h(w^h) \\
@@
 &\text{correct:}                    & &w^h \leftarrow w^h + P(w^H - \iR w^h) \\
-&\text{post-smooth over fine:} & & [\text{smoother updates } w^h]
+&\text{post-smooth over fine:} & & \text{smoother on } w^h
 \end{align*}
@@
 \begin{itemize}
+\item $R: (\mathcal{V}^h)' \to (\mathcal{V}^H)'$ is \emph{canonical restriction}
 \item $P: \mathcal{V}^H \to \mathcal{V}^h$ is \emph{canonical prolongation}
+\item $\iR: \mathcal{V}^h \to \mathcal{V}^H$ is \emph{injection}
 \item \textbf{restrict}$+$(\textbf{solve coarse})$+$\textbf{correct} \, $=$ \, \emph{FAS coarse grid correction}
@@
 
-\begin{frame}{does it work?}
+\begin{frame}{FAS reputation}
 
@@
 with Dirichlet boundary conditions on $\Omega=(0,1)^2$
-\item discretize by (straightforward) finite differences
 \item minimal problem-specific code:
@@
         \end{itemize}
-    \item[3.] coarsest-level solve can be same as smoother, or more sophisticated (e.g.~Newton iteration)
+    \item[3.] coarsest-level solve can be same as smoother
     \end{enumerate}
@@
 
-\begin{frame}{the meaning of ``fast solver''}
+\begin{frame}{the meanings of ``fast solver''}
 
@@
 \item what does ``very well'' on the previous slide mean?
+\begin{block}{definition} a solver is \emph{optimal} if work is $O(N)$ for $N$ unknowns
+\end{block}
+    \begin{itemize}
+    \item[$\circ$] measure work in flops or run-time
+    \item[$\circ$] constant iterations of $O(N)$ method $\implies$ optimal
+    \end{itemize}
 \end{itemize}
 
-\begin{block}{definition} a solver is \emph{optimal} if work in flops, and/or run-time, is $O(N)$ for $N$ unknowns
-\end{block}
-
 \begin{itemize}
-\item since $\sim$1980: optimality can be achieved by multigrid for PDE problems with reasonably-smooth solutions
-\item in fact, multigrid people get greedy
+\item multigrid people get greedy:
 \begin{block}{definition} a solver shows \emph{textbook multigrid efficiency} if it does total work less than 10 times that of a single smoother sweep
@@
 
-\begin{frame}{Bratu model problem: TME}
+\section{FASCD = FAS multigrid for VIs using constraint decomposition}
 
-\begin{columns}
-\begin{column}{0.45\textwidth}
-\begin{itemize}
-\item \texttt{bratu.c}
-\item observed optimality:
-\begin{align*}
-\text{flops} &= O(N^1) \\
-\text{processor time} &= O(N^1)
-\end{align*}
-\item highest-resolution {\color{FireBrick} $12$-level V-cycle} has $N\approx 10^8$ unknowns
-\item compare $\approx 20\,\mu\,\text{s}/\text{N}$ for Poisson using Firedrake ($P_1$, geometric multigrid)
-\end{itemize}
-\end{column}
-\begin{column}{0.55\textwidth}
-\includegraphics[width=\textwidth]{../talk-dms/figs/bratu-time.png}
-\end{column}
-\end{columns}
-\end{frame}
+\subsection{results}  % FIXME lead with results
 
-\section{FASCD = FAS multigrid for VIs}
+\subsection{algorithm details}
 
@@
 
-\section{results}
+% FIXME move earlier \section{results}
 

### 2024-07-21 talk-ismp/py/advdiff.py
slight improvements to figs from DMS talk

@@ -1,0 +1,23 @@
+#!/usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+from writeout import writeout
+
+# data from Table 4 in paper
+n = 15 * 2**np.arange(0,9) + 1
+mJ2D = n**2
+mJ3D = n[:5]**3
+F2D = [1,2,2,2,2,2,3,3,3]
+F3D = [1,2,2,2,3]
+
+plt.clf()
+plt.semilogx(mJ2D,F2D,'+',ms=12.0,mew=2.0,label='FMG cycles (2D)')
+plt.semilogx(mJ3D,F3D,'+',ms=14.0,mew=2.0,label='FMG cycles (3D)')
+plt.xlabel(r'$m_J$  (number of unknowns)')
+plt.ylabel('FASCD iterations')
+plt.legend(loc='upper left', fontsize=12.0)
+plt.xlim([10.0,1.0e8])
+plt.ylim([0.0,4.0])
+plt.gca().set_yticks(range(1,5))
+writeout('advdiff.png')

### 2024-07-21 talk-ismp/py/classical.py

@@ -1,0 +1,33 @@
+#!/usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+from writeout import writeout
+
+# data from Table 2 in paper
+mJ = [41, 145, 545, 2.1e3, 8.3e3, 3.3e4, 1.3e5, 5.3e5, 2.1e6, 8.4e6, 3.4e7]
+ballV = [1,2,3,7,9,11,11,12,13,13,14]
+ballF = [1,2,3,3,4,5,4,4,3,2,2]
+spirV = [1,4,5,7,9,10,11,12,13,14,18]
+spirF = [1,2,3,4,4,5,5,5,6,6,6]
+
+names = ['ballitersV', 'ballitersVlog', 'ballitersVF', 'bothitersVF']
+for j in range(4):
+    plt.clf()
+    plt.semilogx(mJ,ballV,'o',ms=12.0,color='C0',label='V-cycles (ball problem)')
+    if j == 1:
+        p = np.polyfit(np.log(mJ),ballV,1)
+        fitstr = r'$O(log(m_J))$'
+        plt.semilogx(mJ,p[0]*np.log(mJ)+p[1],'--',color='C0',label=fitstr)
+    if j >= 2:
+        plt.semilogx(mJ,ballF,'+',ms=16.0,mew=2.0,color='C0',label='FMG cycles (ball problem)')
+    if j == 3:
+        plt.semilogx(mJ,spirV,'o',ms=8.0,color='C1',label='V-cycles (spiral problem)')
+        plt.semilogx(mJ,spirF,'+',ms=12.0,mew=2.0,color='C1',label='FMG cycles (spiral problem)')
+    plt.xlabel(r'$m_J$  (number of unknowns)')
+    plt.ylabel('FASCD iterations')
+    plt.legend(loc='upper left', fontsize=12.0)
+    plt.xlim([10.0,1.0e8])
+    plt.ylim([0.0,19.0])
+    plt.gca().set_yticks(range(2,20,2))
+    writeout(names[j] + '.png')

### 2024-07-21 talk-ismp/slides.tex
practice this version?; need to get below 25 min

@@
 	\end{itemize}
-\item discretization gives algebraic system on fine mesh $\Omega^h$:
+\item discretization\footnote{discretization = finite element (FE) method, throughout this talk} gives algebraic system on fine mesh $\Omega^h$:
     $${\color{FireBrick} F^h(u^h) = \ell^h}$$
@@
 \begin{itemize}
-\item FAS multigrid works \alert{very well} on nice nonlinear PDE problems
+\item FAS multigrid is a fast solver on nice nonlinear PDE problems
 \item example: Liouville-Bratu equation
@@
 
-\begin{frame}{the meanings of ``fast solver''}
+\begin{frame}{a meaning for ``fast solver''}
 
-\begin{itemize}
-\item what does ``very well'' on the previous slide mean?
 \begin{block}{definition} a solver is \emph{optimal} if work is $O(N)$ for $N$ unknowns
 \end{block}
+
+\bigskip
 \begin{itemize}
-    \item[$\circ$] measure work in flops or run-time
-    \item[$\circ$] constant iterations of $O(N)$ method $\implies$ optimal
-    \end{itemize}
-\end{itemize}
-
-\begin{itemize}
-\item multigrid people get greedy:
-\begin{block}{definition} a solver shows \emph{textbook multigrid efficiency} if it does total work less than 10 times that of a single smoother sweep
-\end{block}
-    \begin{itemize}
-    \item[$\circ$] TME $\implies$ optimal
-    \end{itemize}
+\item one can measure work in flops or run-time
+\item constant iterations of $O(N)$ method $\implies$ optimal
 \end{itemize}
@@
 
-\subsection{results}  % FIXME lead with results
+\subsection{results, and an application}
+
+\begin{frame}{classical obstacle problem by FASCD}
+
+\only<1>{recall:
+
+\centerline{\includegraphics[width=0.9\textwidth]{../talk-oxford/images/obstacle65.pdf}}}
+\only<2>{\includegraphics[width=0.75\textwidth]{figs/ballitersV.png}}\only<3>{\includegraphics[width=0.75\textwidth]{figs/ballitersVlog.png}}\only<4>{\includegraphics[width=0.75\textwidth]{figs/ballitersVF.png}}\only<5>{\includegraphics[width=0.75\textwidth]{figs/bothitersVF.png}} \hfill
+\only<2-4>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/ball-set.png}}%
+\only<5>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/spiral-set.png}}
+
+\only<2-5>{\noindent \scriptsize
+\emph{all results from Bueler \& Farrell (2024)}}
+\end{frame}
+
+\begin{frame}{advection-diffusion of a concentration}
+
+\begin{itemize}
+\item $u(x)$ is a concentration in $\Omega \subset \RR^d$: \qquad $\boxed{0\le u\le 1}$
+\item substance moves by diffusion, advection (wind $\bm{X}(x)$), and source ($\sigma(x)$):
+    $$-\eps \grad^2 u + \bm{X}\cdot \grad u = \sigma$$
+\item two active sets:
+    $$\underline{A}_u = \{u(x) = 0\} \hspace{22mm} \overline{A}_u = \{u(x) = 1\} \hspace{8mm}$$
+\end{itemize}
+
+\centering
+\includegraphics[width=0.27\textwidth]{../paper/fixfigs/poll2d-zero-set.png} \hspace{18mm}
+\includegraphics[width=0.27\textwidth]{../paper/fixfigs/poll2d-one-set.png}
+\end{frame}
+
+\begin{frame}{advection-diffusion of a concentration}
+
+\begin{center}
+\includegraphics[width=0.75\textwidth]{figs/advdiff.png}
+\end{center}
+
+\vspace{-3mm}
+\hfill \tiny \emph{Bueler \& Farrell (2024)}
+
+\medskip
+\scriptsize
+\begin{itemize}
+\item<2> compare linear programming (Klee-Minty cube?) \dots with strong spatial correlations
+\end{itemize}
+\end{frame}
+
+\begin{frame}{problem: geometry of flowing glacier ice in a climate}
+
+\begin{itemize}
+\item ``where are the glaciers?'' is a free-boundary problem
+\end{itemize}
+
+\includegraphics[width=1.02\textwidth]{../talk-oxford/images/alps-seguinot2018.png}
+
+\vspace{-2mm}
+\hfill {\tiny Seguinot et al.~(2018)}
+\end{frame}
+
+\begin{frame}{free-boundary problem: flowing glacier ice in a climate}
+
+\begin{itemize}
+\item glacier = incompressible, viscous fluid driven by gravity
+\item in a \emph{climate} which adds or removes ice at a signed rate $a(t,x,y)$
+\item over fixed bed topography with elevation $b(x,y)$
+    \begin{itemize}
+    \item[$\circ$] data $a,b$ is defined on domain $\Omega \subset \RR^2$
+    \end{itemize}
+\item to find: ice surface elevation $s(t,x,y)$ and velocity $\bu(t,x,y,z)$
+\item subject to constraint $\boxed{s\ge b}$
+\end{itemize}
+
+\bigskip
+\hfill \mbox{\includegraphics[height=0.24\textheight]{../talk-oxford/images/domain-data.png} \quad $\stackrel{?}{\to}$ \quad \includegraphics[height=0.24\textheight]{../talk-oxford/images/domain-velocity.png}}
+\end{frame}
+
+\begin{frame}{glacier free-boundary problem: steady VI form}
+
+\begin{itemize}
+\item admissible surface elevations:
+    $$\mathcal{K} = \left\{r \in \mathcal{V} \,:\, r \ge b\right\}$$
+\item steady-state VI problem for surface elevation $s\in\mathcal{K}$:
+	$$\ip{\Phi(s) - a}{r-s} \ge 0 \quad \text{ for all } r \in \mathcal{K}$$
+where $\Phi(s)=- \bu|_s \cdot \bn_s$
+\item in the isothermal, nonsliding, \emph{shallow ice approximation}\footnote{$=$ \emph{a highly-simplified view of conservation of momentum, in which} $(s-b)^{8/3} \in W^{1,4}(\Omega)$ \emph{so} $\mathcal{V} \stackrel{?}{=} (W^{1,4})^{3/8}$ \hfill \emph{(Jouvet \& Bueler, 2012)}} case:
+\begin{align*}
+\Phi(s) &= - \bu|_s \cdot \bn_s \\
+        &= - \frac{\gamma}{4} (s-b)^{4} |\grad s|^{4} - \grad \cdot\left(\frac{\gamma}{5} (s-b)^{5} |\grad s|^{2} \grad s\right)
+\end{align*}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{test case: ice sheet ($=$ big glacier) the size of Greenland}
+
+\centering
+\includegraphics[width=\textwidth]{../paper/fixfigs/sialev8scene.png}
+\end{frame}
+
+\begin{frame}{FASCD: parallel weak scaling}
+
+\begin{itemize}
+\item optimality of FMG solver (not shown)
+\item good parallel \emph{weak scaling}
+    \begin{itemize}
+    \item[$\circ$] each processor owns $641\times 641$ (sub) mesh
+    \item[$\circ$] $P=1024$ run had $20481^2=4.1\times 10^8$ unknowns
+    \item[] \dots and 88 meter resolution
+    \end{itemize}
+\end{itemize}
+
+\bigskip
+\centering
+\includegraphics[width=0.55\textwidth]{../talk-dms/figs/siaweaktime.png}
+
+\hfill \scriptsize
+\emph{Bueler \& Farrell (2024)}
+\end{frame}
 
@@
 \begin{itemize}
-\item new algorithm (Bueler \& Farrell 2023):
+\item new algorithm (Bueler \& Farrell 2024):
 
@@
     $$\Omega^j \subset \Omega^{j+1}$$
-\item the FE function spaces $\mathcal{V}^j$ over $\Omega^j$ are also nested:
+\item the FE vector spaces $\mathcal{V}^j$ over $\Omega^j$ are also nested:
     $$\mathcal{V}^j \subset \mathcal{V}^{j+1}$$
+\item $\mathcal{V}^J = \sum_{i=0}^J \mathcal{V}^i$ is called a \emph{subspace decomposition} (Xu 1992)
 
-\begin{block}{definition}
-$\ds \mathcal{V}^J = \sum_{i=0}^J \mathcal{V}^i$ \quad is called a \emph{subspace decomposition} (Xu 1992)
-\end{block}
     \begin{itemize}
@@
 \begin{itemize}
-\item Tai's (2003) constraint decomposition \emph{non-trivially} extends a subspace decomposition $\mathcal{V}^J = \sum_i\mathcal{V}^i$ to convex subsets
+\item Tai's (2003) constraint decomposition \emph{non-trivially} extends subspace decomposition to convex subsets
 \item suppose $\mathcal{K}^J \subset \mathcal{V}^J$ is a closed and convex subset
@@
 \begin{itemize}
-\item Tai proposed abstract iterations for solving $VI(F,\ell,\mathcal{K})$ over a CD $\mathcal{K}^J = \sum_{i=0}^J \mathcal{K}^i$
+\item Tai proposed abstract iterations for solving $VI(F,\ell,\mathcal{K})$ over a CD $\mathcal{K}^J = \sum_{i=0}^J \mathcal{K}^i$, specifically:
 
@@
 
-\medskip
-\item Tai's iterations are not practical because you compute on the finest level in fact
-\item we added two techniques: \emph{defect obstacles} on each level, and \emph{FAS coarse corrections}
+\item Tai's iterations are not practical because you must evaluate all residuals on the finest level
+\item two techniques make a practical solver:
+
+    \begin{itemize}
+    \item[$\circ$] \emph{defect obstacles} on each level  \hfill {\scriptsize (Gr\"aser \& Kornhuber (2009)}
+    \item[$\circ$] \emph{FAS coarse corrections}  \hfill {\scriptsize (Bueler \& Farrell (2024)}
+    \end{itemize}
 \end{itemize}
@@
 \begin{itemize}
-\item recall $\mathcal{K} = \{v \ge \psi\}$ in an obstacle problem
+\item suppose $\mathcal{K} = \{v \ge \psi\}$ in an obstacle problem
 \begin{block}{definition}
-for finest-level admissible set $\mathcal{K}^J = \{v^J\ge \psi^J\} \subset \mathcal{V}^J$ and an iterate $w^J \in \mathcal{K}^J$, the \emph{defect obstacle} (Gr\"aser \& Kornhuber 2009) is
+for finest-level admissible set $\mathcal{K}^J = \{v^J\ge \psi^J\} \subset \mathcal{V}^J$ and an iterate $w^J \in \mathcal{K}^J$, the \emph{defect obstacle} is
     $$\chi^J = \psi^J - w^J \in \mathcal{V}^J$$
@@
     \end{itemize}
-\item we generate the CD through
+\item generate the CD through
 
@@
     \item[$\circ$] a \emph{nonlinear} operator
-    \item[$\circ$] {\footnotesize also due to (Gr\"aser \& Kornhuber 2009)}
     \end{itemize}
@@
 
-\begin{frame}{up and down sets}
+\begin{frame}{up and down CDs in FASCD}
+
+\begin{center}
+\includegraphics[width=0.4\textwidth]{../talk-oxford/images/fascd-vcycle.png}
+\end{center}
 
 \begin{itemize}
-\item upward part in the FASCD V-cycle uses large admissible sets:
+\item upward direction in V-cycle uses larger admissible sets:
     $$\mathcal{U}^j = \{z^j \ge \chi^j\}$$
@@
 \end{itemize}
+\end{frame}
+
+\begin{frame}{up and down CDs in FASCD}
 
 \begin{center}
-\includegraphics[width=0.4\textwidth]{../talk-oxford/images/fascd-vcycle.png}
+\includegraphics[width=0.7\textwidth]{../talk-dms/figs/innerconeapprox.png}
 \end{center}
+
+\begin{itemize}
+\item all sets $\mathcal{U}^j$ and $\mathcal{D}^j$ are cones which include the origin
+\end{itemize}
 \end{frame}
 
-\begin{frame}{multilevel constraint decomposition in FASCD}
+\begin{frame}{FASCD = full approximation scheme constraint decomposition}
 
-\centering
-\includegraphics[width=0.7\textwidth]{../talk-dms/figs/innerconeapprox.png}
-\end{frame}
-
-\begin{frame}{full approximation scheme constraint decomposition}
-
-% case of lower obstacle only
+\vspace{-2mm}
 \begin{pseudo}[font=\small]
@@
 \end{pseudo}
+
+\vspace{-2mm}
+\tiny
+\begin{itemize}
+\item[] \emph{(unilateral (lower obstacle only) version shown for simplicity)}
+\end{itemize}
 \end{frame}
 
-\begin{frame}{FASCD V-cycle: visualization on a 1D problem}
+\begin{frame}{V-cycle: visualization on a 1D problem}
 
@@
 
-\begin{frame}{FASCD specifics}
+\begin{frame}{specifics}
 
-see paper (Bueler \& Farrell 2023) for:
+see (Bueler \& Farrell 2024) for:
 \begin{itemize}
@@
     \begin{itemize}
-    \item[$\circ$] evaluate whether CP/KKT conditions are satisfied
+    \item[$\circ$] evaluates CP/KKT conditions
     \end{itemize}
@@
 \end{itemize}
-\end{frame}
-
-% FIXME move earlier \section{results}
-
-\begin{frame}{classical obstacle problem by FASCD}
-
-\only<1>{\includegraphics[width=0.75\textwidth]{../talk-dms/figs/ballitersV.png}}\only<2>{\includegraphics[width=0.75\textwidth]{../talk-dms/figs/ballitersVlog.png}}\only<3>{\includegraphics[width=0.75\textwidth]{../talk-dms/figs/ballitersVF.png}}\only<4>{\includegraphics[width=0.75\textwidth]{../talk-dms/figs/bothitersVF.png}} \hfill
-\only<1-3>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/ball-set.png}}%
-\only<4>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/spiral-set.png}}
-\end{frame}
-
-\begin{frame}{advection-diffusion of a concentration}
-
-\begin{itemize}
-\item suppose $u(x)$ is a concentration in $\Omega \subset \RR^d$: \qquad $0\le u\le 1$
-\item suppose it moves by combination of diffusion, advection by wind $\bm{X}(x)$, and source function $\phi(x)$:
-    $$-\eps \grad^2 u + \bm{X}\cdot \grad u = \phi$$
-\item two active sets (\emph{$d=2$ case}):
-    $$\underline{A}_u = \{u(x) = 0\} \hspace{22mm} \overline{A}_u = \{u(x) = 1\} \hspace{8mm}$$
-\end{itemize}
-
-\centering
-\includegraphics[width=0.27\textwidth]{../paper/fixfigs/poll2d-zero-set.png} \hspace{18mm}
-\includegraphics[width=0.27\textwidth]{../paper/fixfigs/poll2d-one-set.png}
-\end{frame}
-
-\begin{frame}{advection-diffusion of a concentration}
-
-\begin{center}
-\includegraphics[width=0.75\textwidth]{../talk-dms/figs/advdiff.png}
-\end{center}
-
-\scriptsize
-\begin{itemize}
-\item[] \emph{compare: linear programming (Klee-Minty cube?), spatial correlations}
-\end{itemize}
-\end{frame}
-
-\begin{frame}{problem: geometry of flowing glacier ice in a climate}
-
-\begin{itemize}
-\item ``where are there glaciers?'' is a free-boundary problem
-\end{itemize}
-
-\includegraphics[width=1.02\textwidth]{../talk-oxford/images/alps-seguinot2018.png}
-
-\vspace{-2mm}
-\hfill {\tiny Sequinot et al.~(2018)}
-\end{frame}
-
-\begin{frame}{free-boundary problem: flowing glacier ice in a climate}
-
-\begin{itemize}
-\item glacier = incompressible, viscous fluid driven by gravity
-\item to find: ice surface elevation $s(t,x,y)$ and velocity $\bu(t,x,y,z)$
-\item over fixed bed topography with elevation $b(x,y)$
-    \begin{itemize}
-    \item[$\circ$] $s(t,x,y) \ge b(x,y)$
-    \end{itemize}
-\item in a \emph{climate} which adds or removes ice at a signed rate $a(t,x,y)$
-    \begin{itemize}
-    \item[$\circ$] data $a,b$ is defined on domain $\Omega \subset \RR^2$
-    \end{itemize}
-\end{itemize}
-
-\bigskip
-\hfill \mbox{\includegraphics[height=0.24\textheight]{../talk-oxford/images/domain-data.png} \quad $\stackrel{?}{\to}$ \quad \includegraphics[height=0.24\textheight]{../talk-oxford/images/domain-velocity.png}}
-\end{frame}
-
-\begin{frame}{glacier free-boundary problem: naive strong form}
-
-\begin{itemize}
-\item is this an adequate description?:
-\begin{align*}
-s &\ge b                    & &\text{everywhere in } \Omega \\
-\frac{\partial s}{\partial t} &= a + \bu|_s \cdot \bn_s & &\text{where } s(t,x,y) > b(x,y)
-\end{align*}
-\item notes:
-    \begin{itemize}
-    \item[$\circ$] surface velocity $\bu|_s$ is, \alert{in some manner}, determined by $s$
-        \begin{itemize}
-        \item $\bu|_s$ is generally a \emph{non-local} function of $s$
-        \end{itemize}
-    \item[$\circ$] $\bn_s=\left<-\grad s,1\right>$ is upward surface normal
-    \end{itemize}
-\end{itemize}
-
-\bigskip
-\hfill \mbox{\includegraphics[height=0.2\textheight]{../talk-oxford/images/domain-data.png} \quad $\stackrel{?}{\to}$ \quad \includegraphics[height=0.2\textheight]{../talk-oxford/images/domain-velocity.png}}
-\end{frame}
-
-\begin{frame}{glacier free-boundary problem: steady VI form}
-
-\begin{itemize}
-\item admissible surface elevations:\footnote{$(s-b)^{8/3} \stackrel{?}{\in} W^{1,4}(\Omega)$ so $\mathcal{V} \stackrel{?}{=} (W^{1,4})^{3/8}$ \dots  see (Jouvet \& Bueler, 2012)}
-    $$\mathcal{K} = \left\{r \in \mathcal{V} \,:\, r \ge b\right\}$$
-\item steady ($\frac{\partial s}{\partial t}=0$) VI problem for surface elevation $s\in\mathcal{K}$:
-	$$\ip{\Phi(s) - a}{r-s} \ge 0 \quad \text{ for all } r \in \mathcal{K}$$
-where
-    $$\Phi(s)=- \bu|_s \cdot \bn_s$$
-with extension by 0 to all of $\Omega$
-%\item $\bu$ is the velocity solution on
-%    $$\Lambda(s) = \{(x,y,z) : b(x,y) < z < s(x,y)\} \subset \RR^3$$
-\end{itemize}
-\end{frame}
-
-\begin{frame}{shallow ice approximation}
-
-\begin{itemize}
-\item the \emph{shallow ice approximation} is a highly-simplified view of conservation of momentum
-\item isothermal, nonsliding case:
-\begin{align*}
-\Phi(s) &= - \bu|_s \cdot \bn_s \\
-        &= - \frac{\gamma}{4} (s-b)^{4} |\grad s|^{4} - \grad \cdot\left(\frac{\gamma}{5} (s-b)^{5} |\grad s|^{2} \grad s\right)
-\end{align*}
-\end{itemize}
-\end{frame}
-
-\begin{frame}{FASCD test case: simplified ice sheet the size of Greenland}
-
-\begin{itemize}
-\item \emph{ice sheet} = big glacier
-\end{itemize}
-
-\bigskip\bigskip
-\centering
-\includegraphics[width=\textwidth]{../paper/fixfigs/sialev8scene.png}
-\end{frame}
-
-\begin{frame}{FASCD: parallel weak scaling}
-
-\begin{itemize}
-\item observed optimality of FMG solver
-\item good parallel \emph{weak scaling} as well
-    \begin{itemize}
-    \item[$\circ$] each processor owns $641\times 641$ (sub) mesh
-    \item[$\circ$] $P=1024$ run had $20481^2=4.1\times 10^8$ unknowns
-    \item[] \dots and 88 meter resolution
-    \end{itemize}
-\end{itemize}
-
-\bigskip
-\centering
-\includegraphics[width=0.55\textwidth]{../talk-dms/figs/siaweaktime.png}
 \end{frame}
@@
     \end{itemize}
-\item observed optimality, even TME, in many cases
+\item observed optimality in many cases
     \begin{itemize}
-    \item[$\circ$] actually fast
+    \item[$\circ$] actually fast on tested nonlinear PDE problems
     \end{itemize}
@@
 \bigskip\bigskip
-\item[] \textbf{{\color{FireBrick} to do:}}
+\item many things \textbf{{\color{FireBrick} to do}}:
     \begin{itemize}
-    \item add mesh adaptivity to free boundary (Stefano)
-    \item implement in C inside PETSc
-    \item apply to space-time (\emph{parabolic}) VI problems
-    \item prove convergence
-    \item identify smoothers for problems like elastic contact
-    \item include membrane stresses in glacier case
+    \item[$\circ$] implement in C inside PETSc
+    \item[$\circ$] prove convergence; compare (Reusken, 1988)
+    \item[$\circ$] apply to space-time (\emph{parabolic}) VI problems
+    \item[$\circ$] add mesh adaptivity
+    \item[$\circ$] identify smoothers for problems like elastic contact
+    \item[$\circ$] include membrane stresses in glacier case
     \end{itemize}

### 2024-07-23 talk-ismp/slides.tex
small improvements after 1st practice

@@
 \begin{itemize}
-\item the VI gives an ``angle condition'' for the test function $v-u$:
-\item in optimization problems the condition is a consequence:
-\begin{align*}
-J(u) &\le J(v) \quad \forall v \in \mathcal{K} \\
-\implies \qquad \ip{J'(u)}{v-u} &\ge 0 \quad \forall v \in \mathcal{K}
-\end{align*}
-\item for general VIs the condition \emph{is} the problem:
+\item the VI gives an ``angle condition'' for the test function $v-u$
+\item in optimization problems the VI is a consequence:
+$$J(u) \le J(v), \, \forall v \in \mathcal{K} \qquad \implies \qquad \ip{J'(u)}{v-u} \ge 0, \, \forall v \in \mathcal{K}$$
+\item for general VIs the angle condition \emph{is} the problem:
    $$\ip{F(u)}{v-u} \ge 0 \quad \forall v \in \mathcal{K}$$    
@@
 \begin{center}
-\includegraphics[width=0.4\textwidth]{../talk-dms/figs/convexuv.png}
+\includegraphics[width=0.4\textwidth]{figs/convexuv.png}
 \end{center}
@@
 \vspace{1.5mm}
-\item<2> first-semester calculus:
+\item first-semester calculus:
     $$u = \mathop{\textnormal{argmin}}_{x\in[a,b]} f(x) \quad \iff \quad f'(u)(v-u) \ge 0 \quad \forall v \in[a,b]$$
@@
 \begin{center}
-\only<1>{\vspace{20.5mm}}
-
-\includegraphics<2>[height=25mm]{../talk-oxford/images/calcone.png}
+\includegraphics[height=25mm]{../talk-oxford/images/calcone.png}
 \end{center}
@@
 with Dirichlet boundary conditions on $\Omega=(0,1)^2$
-\item minimal problem-specific code:
-    \begin{enumerate}
-    \item[1.] residual evaluation on grid level: $F^j(\cdot)$
-    \item[2.] pointwise smoother: $\phi_i(c) = 0 \,\forall i$
-        \begin{itemize}
-        \item[$\circ$] nonlinear Gauss-Seidel iteration
-        \end{itemize}
-    \item[3.] coarsest-level solve can be same as smoother
-    \end{enumerate}
+%\item minimal problem-specific code:
+%    \begin{enumerate}
+%    \item[1.] residual evaluation on grid level: $F^j(\cdot)$
+%    \item[2.] pointwise smoother: $\phi_i(c) = 0 \,\forall i$
+%        \begin{itemize}
+%        \item[$\circ$] nonlinear Gauss-Seidel iteration
+%        \end{itemize}
+%    \item[3.] coarsest-level solve can be same as smoother
+%    \end{enumerate}
 \end{itemize}

### 2024-07-23 talk-ismp/py/sia.py
iterations for SIA example

@@ -1,0 +1,19 @@
+#!/usr/bin/env python3
+
+import numpy as np
+import matplotlib.pyplot as plt
+from writeout import writeout
+
+# data from Table 5 in paper
+n = np.array([641, 1281, 2561, 5121, 10241, 20481])**2
+FMG = [2, 2, 2, 2, 1, 1]
+
+plt.clf()
+plt.semilogx(n,FMG,'+',ms=14.0,mew=2.0,label='FMG cycles')
+plt.xlabel(r'$m_J$  (number of unknowns)')
+plt.ylabel('FASCD iterations')
+plt.legend(loc='upper left', fontsize=12.0)
+plt.xlim([1.0e5,1.0e9])
+plt.ylim([0.0,3.0])
+plt.gca().set_yticks(range(1,5))
+writeout('sia.png')

### 2024-07-23 talk-ismp/references.tex
practice this version?; need to get below 25 min
paper published

@@
 \begin{itemize}
-\item E.~Bueler \& P.~Farrell (2024).
-\emph{A full approximation scheme multilevel method for nonlinear variational inequalities}, to appear SIAM J.~Sci.~Comput.
-\, {\tiny \href{https://arxiv.org/abs/2308.06888}{arxiv:2308.06888}}
+\item E.~Bueler \& P.~Farrell (2024).  \emph{A full approximation scheme multilevel method for nonlinear variational inequalities}, SIAM J.~Sci.~Comput.~46 (4) \sdoi{10.1137/23M1594200}
 \item C.~Gr{\"a}ser \& R.~Kornhuber (2009). \emph{Multigrid methods for obstacle problems}, J.~Comput.~Math., 1--44
@@
 \emph{Steady, shallow ice sheets as obstacle problems: well-posedness and finite element approximation}, SIAM J.~Appl.~Math.~72 (4), 1292--1314 \sdoi{10.1137/110856654}
+\item A.~Reusken (1988). \emph{Convergence of the multilevel full approximation scheme including the V-cycle}, Numerische Mathematik 53, 663--686.
 \item X.~Tai (2003). \emph{Rate of convergence for some constraint decomposition methods for nonlinear variational inequalities}, Numer.~Math.~93 (4), 755--786 \sdoi{10.1007/s002110200404}

### 2024-07-23 talk-ismp/slides.tex
many changes to make it better

@@
 
+\newcommand{\cK}{\mathcal{K}}
+\newcommand{\cV}{\mathcal{V}}
+
 \newcommand{\rhoi}{\rho_{\text{i}}}
@@
 
-\title[FAS multigrid for VIs]{A full approximation scheme multilevel method \\ for solving nonlinear variational inequalities}
+\newcommand{\sdoi}[1]{\,{\tiny \href{https://doi.org/#1}{doi:#1}}}
+
+\title[FASCD = multigrid for VIs]{A full approximation scheme multilevel method \\ for solving nonlinear variational inequalities}
 
@@
 
-\titlegraphic{\vspace{-18mm} \includegraphics[width=0.22\textwidth]{../talk-oxford/images/uafbw.png} \hfill \begin{minipage}{0.12\textwidth}
+\titlegraphic{\vspace{-18mm} \includegraphics[width=0.2\textwidth]{../talk-oxford/images/uafbw.png} \hfill \begin{minipage}{0.12\textwidth}
 \includegraphics[width=\textwidth]{../talk-oxford/images/oxford.png}
@@
   \tableofcontents
+
+\only<1>{\phantom{\emph{warning:} no convergence theorem}}
+\only<2>{\emph{warning:} no convergence theorem}
 \end{frame}
@@
 
-\begin{frame}{example, cont.}
+\begin{frame}{example \dots and VI language}
 
@@
    \begin{itemize}
-   \item[$\circ$] \emph{active set} $A_u = \{u = \psi\}$
-   \item[$\circ$] \emph{inactive set} $R_u = \{u> \psi\}$
-   \item[$\circ$] \emph{free boundary} $\Gamma_u=\partial R_u \cap \Omega$
+   \item[$\circ$] \emph{active set} \quad $A_u = \{u = \psi\}$
+   \item[$\circ$] \emph{inactive set} \quad $R_u = \{u> \psi\}$
+   \item[$\circ$] \emph{free boundary} \quad $\Gamma_u=\partial R_u \cap \Omega$
    \end{itemize}
@@
 
-\begin{frame}{geometrical meaning of VI}
+\begin{frame}{variational inequalities: general setting}
 
 \begin{itemize}
-\item the VI gives an ``angle condition'' for the test function $v-u$
-\item in optimization problems the VI is a consequence:
-$$J(u) \le J(v), \, \forall v \in \mathcal{K} \qquad \implies \qquad \ip{J'(u)}{v-u} \ge 0, \, \forall v \in \mathcal{K}$$
-\item for general VIs the angle condition \emph{is} the problem:
-   $$\ip{F(u)}{v-u} \ge 0 \quad \forall v \in \mathcal{K}$$    
-\end{itemize}
-
-\bigskip
-\begin{center}
-\includegraphics[width=0.4\textwidth]{figs/convexuv.png}
-\end{center}
-\end{frame}
-
-\begin{frame}{general variational inequalities}
-
-\begin{itemize}
-\item let $\mathcal{K}$ be a closed and convex subset of a Banach space $\mathcal{V}$
-\item suppose $F:\mathcal{K} \to \mathcal{V}'$ is a continuous operator
+\item $\mathcal{K}$ is a closed and convex subset of a Banach space $\mathcal{V}$
+\item $F:\mathcal{K} \to \mathcal{V}'$ is a continuous operator
     \begin{itemize}
@@
     \end{itemize}
-\item the general variational inequality {\color{FireBrick} VI($F$,$\mathcal{K}$)} is
+\item the variational inequality {\color{FireBrick} VI($F$,$\mathcal{K}$)} is to find $u\in\cK$ so that
 	$${\color{FireBrick} \ip{F(u)}{v-u} \ge 0 \quad \text{ for all } v \in \mathcal{K}}$$
-\item when $\mathcal{K}$ is nontrivial, VI($F$,$\mathcal{K}$) is nonlinear even when $F$ is a linear operator
+
+    \begin{itemize}
+    \item[$\circ$] for nontrivial $\mathcal{K}$, VI($F$,$\mathcal{K}$) is nonlinear even if $F$ is linear
+    \end{itemize}
 \end{itemize}
@@
 \item first-semester calculus:
-    $$u = \mathop{\textnormal{argmin}}_{x\in[a,b]} f(x) \quad \iff \quad f'(u)(v-u) \ge 0 \quad \forall v \in[a,b]$$
+    $$u = \mathop{\textnormal{argmin}}_{x\in[a,b]} f(x) \quad \implies \quad f'(u)(v-u) \ge 0 \quad \forall v \in[a,b]$$
 \end{itemize}
@@
 &\text{restrict:}                   & &\ell^H = F^H(\iR w^h) + R\, r^h(w^h) \\
-&\text{solve coarse:}                      & &F^H(w^H) = \ell^H \\
+&\text{coarse solve:}               & &F^H(w^H) = \ell^H \\
 &\text{correct:}                    & &w^h \leftarrow w^h + P(w^H - \iR w^h) \\
@@
 \item $\iR: \mathcal{V}^h \to \mathcal{V}^H$ is \emph{injection}
-\item \textbf{restrict}$+$(\textbf{solve coarse})$+$\textbf{correct} \, $=$ \, \emph{FAS coarse grid correction}
+\item (restrict$+$coarse solve$+$correct) \, $=$ \, \emph{FAS coarse grid correction}
 \end{itemize}
@@
 
-\subsection{results, and an application}
+\subsection{results on test problems}
 
-\begin{frame}{classical obstacle problem by FASCD}
+\begin{frame}{test problem 1: classical obstacle problem}
 
@@
 
-\begin{frame}{advection-diffusion of a concentration}
+\begin{frame}{test problem 2: advection-diffusion}
 
@@
 \item $u(x)$ is a concentration in $\Omega \subset \RR^d$: \qquad $\boxed{0\le u\le 1}$
-\item substance moves by diffusion, advection (wind $\bm{X}(x)$), and source ($\sigma(x)$):
+\item concentration evolves by diffusion, advection, and source:
     $$-\eps \grad^2 u + \bm{X}\cdot \grad u = \sigma$$
@@
 \hfill \tiny \emph{Bueler \& Farrell (2024)}
-
-\medskip
-\scriptsize
-\begin{itemize}
-\item<2> compare linear programming (Klee-Minty cube?) \dots with strong spatial correlations
-\end{itemize}
 \end{frame}
 
-\begin{frame}{problem: geometry of flowing glacier ice in a climate}
+\subsection{an application}
 
-\begin{itemize}
-\item ``where are the glaciers?'' is a free-boundary problem
-\end{itemize}
+\begin{frame}{``where are the glaciers?'' is a free-boundary problem}
 
-\includegraphics[width=1.02\textwidth]{../talk-oxford/images/alps-seguinot2018.png}
+\bigskip
+\includegraphics[width=1.01\textwidth]{../talk-oxford/images/alps-seguinot2018.png}
 
 \vspace{-2mm}
-\hfill {\tiny Seguinot et al.~(2018)}
+\hfill {\tiny \emph{Seguinot et al.~(2018)}}
 \end{frame}
 
-\begin{frame}{free-boundary problem: flowing glacier ice in a climate}
+\begin{frame}{inequality-constrained problem: flowing glacier ice in a climate}
 
@@
 where $\Phi(s)=- \bu|_s \cdot \bn_s$
-\item in the isothermal, nonsliding, \emph{shallow ice approximation}\footnote{$=$ \emph{a highly-simplified view of conservation of momentum, in which} $(s-b)^{8/3} \in W^{1,4}(\Omega)$ \emph{so} $\mathcal{V} \stackrel{?}{=} (W^{1,4})^{3/8}$ \hfill \emph{(Jouvet \& Bueler, 2012)}} case:
+\item in the isothermal, nonsliding, \emph{shallow ice approximation}\footnote{$=$ \emph{a highly-simplified view of conservation of momentum, in which} $(s-b)^{8/3} \in W^{1,4}(\Omega)$ \emph{so} $\mathcal{V} \stackrel{?}{=} (W^{1,4})^{3/8}$ \hfill \emph{Jouvet \& Bueler (2012)}} case:
 \begin{align*}
@@
 
-\begin{frame}{test case: ice sheet ($=$ big glacier) the size of Greenland}
+\begin{frame}{test case: ice sheet the size of Greenland}
 
@@
 \begin{itemize}
-\item optimality of FMG solver (not shown)
-\item good parallel \emph{weak scaling}
+\item optimality of FMG solver (left)
+\item good parallel \emph{weak scaling} (right)
     \begin{itemize}
@@
 \bigskip
-\centering
-\includegraphics[width=0.55\textwidth]{../talk-dms/figs/siaweaktime.png}
+\mbox{\includegraphics[width=0.5\textwidth]{figs/sia.png} \includegraphics[width=0.5\textwidth]{../talk-dms/figs/siaweaktime.png}}
 
@@
 
-\subsection{algorithm details}
+\subsection{the algorithm}
 
@@
 \begin{itemize}
-\item new algorithm (Bueler \& Farrell 2024):
+\item new algorithm:\footnote{E.~Bueler \& P.~Farrell (2024).
+\emph{A full approximation scheme multilevel method for nonlinear variational inequalities}, SIAM J.~Sci.~Comput.~46 (4) \sdoi{10.1137/23M1594200}}
 
+\bigskip
 {\color{FireBrick} FASCD = full approximation scheme constraint decomposition}
@@
 \bigskip
-\item what is ``constraint decomposition'' in FAS\underline{CD}?
+    \begin{itemize}
+    \item[$\circ$] what is ``constraint decomposition''?
+    \item[$\circ$] unilateral (lower obstacle) version shown, but bilateral or ``box constraints'' are implemented
+    \end{itemize}
 \end{itemize}
 \end{frame}
-
-\newcommand{\cK}{\mathcal{K}}
 
@@
 \begin{itemize}
-\item start with subspace decomposition over nested meshes:
+\item start with nested meshes:
     $$\Omega^j \subset \Omega^{j+1}$$
@@
     $$\mathcal{V}^j \subset \mathcal{V}^{j+1}$$
-\item $\mathcal{V}^J = \sum_{i=0}^J \mathcal{V}^i$ is called a \emph{subspace decomposition} (Xu 1992)
+\item $\mathcal{V}^J = \sum_{i=0}^J \mathcal{V}^i$ is called a \emph{subspace decomposition}
 
@@
     \item[$\circ$] \emph{non}-unique vector space sum
-    \item[$\circ$] Xu's paper explains how to analyze linear multigrid for PDEs via subspace decomposition
+    \item[$\circ$] one may analyze linear multigrid for PDEs via subspace decomposition
     \end{itemize}
@@
 \begin{itemize}
-\item Tai's (2003) constraint decomposition \emph{non-trivially} extends subspace decomposition to convex subsets
+\item Tai (2003): constraint decomposition \emph{non-trivially} extends subspace decomposition to convex subsets
 \item suppose $\mathcal{K}^J \subset \mathcal{V}^J$ is a closed and convex subset
+\item suppose $\mathcal{V}^J = \sum_{i=0}^J \mathcal{V}^i$ is a subspace decomposition
 \begin{block}{definition}
@@
 
-\begin{frame}{iterations over constraint decomposition}
+\begin{frame}{iterations over a constraint decomposition}
 
 \begin{itemize}
-\item Tai proposed abstract iterations for solving $VI(F,\ell,\mathcal{K})$ over a CD $\mathcal{K}^J = \sum_{i=0}^J \mathcal{K}^i$, specifically:
+\item Tai proposed abstract iterations for solving $VI(F,\ell,\mathcal{K})$ over a CD
+\item specifically the multiplicative/serial iteration:
 
@@
 
-\item Tai's iterations are not practical because you must evaluate all residuals on the finest level
+\item this iteration is not practical because you must evaluate all residuals on the finest level
 \item two techniques make a practical solver:
@@
     \begin{itemize}
-    \item[$\circ$] \emph{defect obstacles} on each level  \hfill {\scriptsize (Gr\"aser \& Kornhuber (2009)}
-    \item[$\circ$] \emph{FAS coarse corrections}  \hfill {\scriptsize (Bueler \& Farrell (2024)}
+    \item[$\circ$] \emph{defect obstacles} on each level  \hfill {\scriptsize Gr\"aser \& Kornhuber (2009)}
+    \item[$\circ$] \emph{FAS coarse corrections}  \hfill {\scriptsize Bueler \& Farrell (2024)}
     \end{itemize}
@@
 \begin{itemize}
-\item all sets $\mathcal{U}^j$ and $\mathcal{D}^j$ are cones which include the origin
+\item all sets $\mathcal{U}^j$ and $\mathcal{D}^j$ include the origin
 \end{itemize}
@@
 
-see (Bueler \& Farrell 2024) for:
+see Bueler \& Farrell (2024) for:
 \begin{itemize}
@@
 \item FMG cycle
-\item details of $O(m_J)$ smoother
+\item details of $O(m_J)$ smoother: a few iterations of projected, reduced-space Newton with a few iterations of CG or GMRES, plus ILU preconditioning, on the linear step equations
 \end{itemize}
@@
     \begin{itemize}
-    \item[$\circ$] implemented in Python Firedrake (over PETSc)
+    \item[$\circ$] implemented in Python using the Firedrake library
+    \item[$\circ$] observed optimality and actually fast on tested problems
     \end{itemize}
-\item observed optimality in many cases
+\item many things \textbf{{\color{FireBrick} to do}}:
     \begin{itemize}
-    \item[$\circ$] actually fast on tested nonlinear PDE problems
+    \item[$\circ$] implement in C within PETSc
+    \item[$\circ$] prove convergence; compare (Reusken, 1988)
+    \item[$\circ$] mesh adaptivity, better smoothers, parabolic examples, \dots
+    \end{itemize}
 \end{itemize}
@@
 \bigskip\bigskip
-\item many things \textbf{{\color{FireBrick} to do}}:
-    \begin{itemize}
-    \item[$\circ$] implement in C inside PETSc
-    \item[$\circ$] prove convergence; compare (Reusken, 1988)
-    \item[$\circ$] apply to space-time (\emph{parabolic}) VI problems
-    \item[$\circ$] add mesh adaptivity
-    \item[$\circ$] identify smoothers for problems like elastic contact
-    \item[$\circ$] include membrane stresses in glacier case
-    \end{itemize}
-\end{itemize}
+\begin{minipage}[T]{0.25\textwidth}
+\vspace{0pt}
+
+\includegraphics[width=0.8\textwidth]{figs/QRbuelerfarrell.png}
+
+\scriptsize paper in
+
+SIAM J.~Sci.~Comput.
+\end{minipage}
+\hfill
+\begin{minipage}[T]{0.25\textwidth}
+\vspace{0pt}
+
+\hfill \includegraphics[width=0.8\textwidth]{figs/QRfascd.png}
+
+\scriptsize \hfill code at bitbucket
+
+\phantom{foo}
+\end{minipage}
 \end{frame}

### 2024-07-24 talk-ismp/slides.tex
a bit shorter and better

@@
 \item consider a nonlinear elliptic PDE problem:
-	$$F(u) = \ell$$
+	$${\color{FireBrick} F(u) = \ell}$$
 
@@
 	\item[$\circ$] $\mathcal{V}=H^1(\Omega)$, $F : \mathcal{V} \to \mathcal{V}'$ continuous, $\ell\in \mathcal{V}'$
-	\item[$\circ$] example is the Liouville-Bratu problem: $-\grad^2 u - e^u = f$
 	\end{itemize}
@@
     $${\color{FireBrick} F^h(u^h) = \ell^h}$$
+\item suppose $w^h$ is a not-yet-converged iterate:
+    $$r^h=\ell^h - F^h(w^h), \qquad \|r^h\| > \text{TOL}$$
 \end{itemize}
-
-\vspace{4mm}
 }
@@
 \begin{itemize}
-\item \emph{goal}: to solve $F^h(u^h) = \ell^h$ on $\Omega^h$
-\item suppose $w^h$ is a not-yet-converged iterate:
-    $$r^h=\ell^h - F^h(w^h), \qquad \|r^h\| > \text{TOL}$$
 \item how can we improve $w^h$ \emph{without} globally linearizing $F^h$?
@@
 	\end{itemize}
-\item notes:
-    \setbeamertemplate{enumerate item}{\emph{\roman{enumi})}}
-    \begin{enumerate}
-    \item \emph{residual} $r^h = \ell^h - F^h(w^h)$ is computable but \emph{error} $e^h = u^h-w^h$ is unknown
-    \item equation can be rewritten as \qquad $F^h(u^h) - F^h(w^h) = r^h$
-    \end{enumerate}
+\item rewrite the equation as the nonlinear \emph{correction equation}
+    $${\color{FireBrick} F^h(u^h) - F^h(w^h) = r^h}$$
+\item \underline{for $F^h$ linear}, convert the correction eqn to the \emph{error equation}
+    $$F^h(e^h) = r^h,$$
+with approximate solution $\tilde e^h$, and update $w^h \leftarrow w^h+\tilde e^h$
 \end{itemize}
@@
 \begin{itemize}
-\item \emph{goal}: from iterate $w^h$, to solve {\color{FireBrick} $F^h(u^h) - F^h(w^h) = r^h$}
-\item for $F^h$ linear, one may convert to the error equation
-    $$F^h(e^h) = r^h$$
-with approximate solution $\tilde e^h$, to improve the iterate
-    $$w^h \leftarrow w^h+\tilde e^h$$
+\item there are fast \emph{smoothers}: inexpensive algorithms which improve $w^h$ by efficiently removing high-frequency error components
+	\begin{itemize}
+	\item sweeping through and solving nodewise problems (nonlinear Gauss-Seidel) is a smoother
+	\end{itemize}
+\item how about low-frequency error components?
+\item \emph{idea}: use a coarser mesh $\Omega^H$ to estimate $u^h$ in the correction eqn
 \end{itemize}
@@
 \only<4>{
-\begin{itemize}
-\item \emph{updated goal}: use a coarser mesh $\Omega^H$ to somehow estimate the solution $u^h$ in the nonlinear \emph{correction equation}
-    $${\color{FireBrick} F^h(u^h) - F^h(w^h) = r^h}$$
-\item \emph{smoothers}: there are inexpensive algorithms which improve $w^h$ by efficiently removing high-frequency error components
-	\begin{itemize}
-	\item sweeping through and solving nodewise problems (nonlinear Gauss-Seidel) is a smoother
-	\end{itemize}
-\item how about low-frequency error components?
-\end{itemize}
-
-\vspace{5mm}
-}
-\only<5>{
 \begin{itemize}
@@
 
-\item rewritten FAS equation: let $\ell^H = F^H(\iR w^h) + R\, r^h(w^h)$ then
+\item let $\ell^H = F^H(\iR w^h) + R\, r^h(w^h)$
+\item rewritten FAS equation:
     $${\color{FireBrick} F^H(u^H) = \ell^H}$$
@@
 \begin{itemize}
-\item FAS multigrid is a fast solver on nice nonlinear PDE problems
-\item example: Liouville-Bratu equation
-    $$-\nabla^2 u - e^u = 0$$
-with Dirichlet boundary conditions on $\Omega=(0,1)^2$
-%\item minimal problem-specific code:
-%    \begin{enumerate}
-%    \item[1.] residual evaluation on grid level: $F^j(\cdot)$
-%    \item[2.] pointwise smoother: $\phi_i(c) = 0 \,\forall i$
-%        \begin{itemize}
-%        \item[$\circ$] nonlinear Gauss-Seidel iteration
-%        \end{itemize}
-%    \item[3.] coarsest-level solve can be same as smoother
-%    \end{enumerate}
+\item FAS multigrid is a \emph{fast solver} on nice nonlinear PDE problems
+    \begin{itemize}
+    \item[$\circ$] example: Liouville-Bratu equation $-\nabla^2 u - e^u = 0$
     \end{itemize}
-\end{frame}
 
-\begin{frame}{a meaning for ``fast solver''}
+\bigskip
+\item<2> what does ``fast solver'' mean?
+\end{itemize}
 
-\begin{block}{definition} a solver is \emph{optimal} if work is $O(N)$ for $N$ unknowns
+\begin{block}<2>{definition} a solver is \emph{optimal} if work is $O(N)$ for $N$ unknowns
 \end{block}
 
-\bigskip
 \begin{itemize}
-\item one can measure work in flops or run-time
-\item constant iterations of $O(N)$ method $\implies$ optimal
+\item<2> one can measure work in flops or run-time
+\item<2> constant iterations of $O(N)$ method $\implies$ optimal
 \end{itemize}
@@
 
-\begin{frame}{inequality-constrained problem: flowing glacier ice in a climate}
+\begin{frame}{the glacier free-boundary problem}
 
 \begin{itemize}
-\item glacier = incompressible, viscous fluid driven by gravity
-\item in a \emph{climate} which adds or removes ice at a signed rate $a(t,x,y)$
-\item over fixed bed topography with elevation $b(x,y)$
-    \begin{itemize}
-    \item[$\circ$] data $a,b$ is defined on domain $\Omega \subset \RR^2$
-    \end{itemize}
-\item to find: ice surface elevation $s(t,x,y)$ and velocity $\bu(t,x,y,z)$
-\item subject to constraint $\boxed{s\ge b}$
-\end{itemize}
+\item a \emph{glacier} is an incompressible, viscous fluid driven by gravity, with mass from a climate which adds or removes ice at a signed rate, over a fixed bed topography with elevation $b$
+\item to find: ice surface elevation $s$ (and velocity $\bu$) subject to constraint $\boxed{s\ge b}$
+\item test case: ice sheet the size of Greenland
 
 \bigskip
-\hfill \mbox{\includegraphics[height=0.24\textheight]{../talk-oxford/images/domain-data.png} \quad $\stackrel{?}{\to}$ \quad \includegraphics[height=0.24\textheight]{../talk-oxford/images/domain-velocity.png}}
+\begin{center}
+\includegraphics[width=0.65\textwidth]{../paper/fixfigs/sialev8scene.png}
+\end{center}
+\end{itemize}
 \end{frame}
@@
 
-\begin{frame}{test case: ice sheet the size of Greenland}
-
-\centering
-\includegraphics[width=\textwidth]{../paper/fixfigs/sialev8scene.png}
-\end{frame}
-
-\begin{frame}{FASCD: parallel weak scaling}
+\begin{frame}{optimality and weak scaling on a glacier problem}
 
@@
     \item[$\circ$] each processor owns $641\times 641$ (sub) mesh
-    \item[$\circ$] $P=1024$ run had $20481^2=4.1\times 10^8$ unknowns
-    \item[] \dots and 88 meter resolution
+    \item[$\circ$] $P=1024$ run had $20481^2=4.1\times 10^8$ unknowns and 88 meter resolution
     \end{itemize}
@@
 \bigskip
-    \begin{itemize}
-    \item[$\circ$] what is ``constraint decomposition''?
-    \item[$\circ$] unilateral (lower obstacle) version shown, but bilateral or ``box constraints'' are implemented
-    \end{itemize}
+\item<2> but what is ``constraint decomposition''?
 \end{itemize}
@@
     \item[$\circ$] \emph{non}-unique vector space sum
-    \item[$\circ$] one may analyze linear multigrid for PDEs via subspace decomposition
+    %\item[$\circ$] one may analyze linear multigrid for PDEs via subspace decomposition
     \end{itemize}
@@
 
-\item this iteration is not practical because you must evaluate all residuals on the finest level
-\item two techniques make a practical solver:
+\item \dots but not practical because you evaluate residuals on finest level
+\item two techniques make a practical solver (Bueler \& Farrell, 2024):
 
@@
     \item[$\circ$] \emph{defect obstacles} on each level  \hfill {\scriptsize Gr\"aser \& Kornhuber (2009)}
-    \item[$\circ$] \emph{FAS coarse corrections}  \hfill {\scriptsize Bueler \& Farrell (2024)}
+    \item[$\circ$] \emph{FAS coarse corrections}  \hfill {\scriptsize Brandt (1977)}
     \end{itemize}
@@
 \item FMG cycle
-\item details of $O(m_J)$ smoother: a few iterations of projected, reduced-space Newton with a few iterations of CG or GMRES, plus ILU preconditioning, on the linear step equations
+\item details of the $O(m_j)$ smoother used in the results: a few iterations of projected, reduced-space Newton with a few iterations of CG+ICC or GMRES+ILU on the linear step equations
 \end{itemize}
@@
     \item[$\circ$] implemented in Python using the Firedrake library
-    \item[$\circ$] observed optimality and actually fast on tested problems
+    \item[$\circ$] observed optimality, and actually fast on tested problems
     \end{itemize}

### 2024-07-25 talk-ismp/slides.tex
fiddles

@@
 %% https://tex.stackexchange.com/questions/170222/change-the-numbering-in-beamers-table-of-content
+
+\setbeamertemplate{page number in head/foot}[appendixframenumber]
 
@@
 	\begin{itemize}
-	\item[$\circ$] $\mathcal{V}=H^1(\Omega)$, $F : \mathcal{V} \to \mathcal{V}'$ continuous, $\ell\in \mathcal{V}'$
+	\item[$\circ$] $\mathcal{V}=H^1(\Omega)$, $F : \mathcal{V} \to \mathcal{V}'$ continuous bijection, $\ell\in \mathcal{V}'$
 	\end{itemize}
-\item discretization\footnote{discretization = finite element (FE) method, throughout this talk} gives algebraic system on fine mesh $\Omega^h$:
+\item finite element (FE) discretization gives algebraic system on fine mesh $\Omega^h$:
     $${\color{FireBrick} F^h(u^h) = \ell^h}$$
@@
 	\begin{itemize}
-	\item are there alternatives to Newton's method?
+	\item alternatives to Newton's method?
 	\end{itemize}
@@
 \end{itemize}
+
+\vspace{6mm}
 }
@@
 	\begin{itemize}
-	\item sweeping through and solving nodewise problems (nonlinear Gauss-Seidel) is a smoother
+	\item nonlinear Gauss-Seidel: sweep through mesh and solving nodewise problems with e.g.~1-variable Newton
 	\end{itemize}
-\item how about low-frequency error components?
+\item but how about low-frequency error components?
 \item \emph{idea}: use a coarser mesh $\Omega^H$ to estimate $u^h$ in the correction eqn
@@
 
-\vspace{5mm}
+\vspace{26.5mm}
 }
@@
     \item[$\circ$] $R:(\mathcal{V}^h)' \to (\mathcal{V}^H)'$ is \emph{canonical restriction}
-    \item[$\circ$] note: if $w^h=u^h$ exactly then $u^H = \iR w^h$ since $F^H$ injective
+    \item[$\circ$] if $w^h=u^h$ exactly then $u^H = \iR w^h$ since $F^H$ injective
     \end{itemize}
@@
 \end{itemize}
+
+\vspace{2mm}
 }
@@
 \begin{center}
-fine mesh $=\Omega^h$\, \includegraphics[height=0.14\textheight]{../talk-oxford/images/fine-grid.png} \hspace{15mm} \includegraphics[height=0.14\textheight]{../talk-oxford/images/coarse-grid.png} \,$\Omega^H=$ coarse mesh
+fine mesh $\Omega^h$ \, \includegraphics[height=0.14\textheight]{../talk-oxford/images/fine-grid.png} \hspace{15mm} \includegraphics[height=0.14\textheight]{../talk-oxford/images/coarse-grid.png} \,coarse mesh $\Omega^H$
 \end{center}
@@
 \begin{align*}
-&\text{pre-smooth over fine:} & & \text{smoother on } w^h \\
-&\text{restrict:}                   & &\ell^H = F^H(\iR w^h) + R\, r^h(w^h) \\
-&\text{coarse solve:}               & &F^H(w^H) = \ell^H \\
-&\text{correct:}                    & &w^h \leftarrow w^h + P(w^H - \iR w^h) \\
-&\text{post-smooth over fine:} & & \text{smoother on } w^h
+&\text{1. pre-smooth:}   & & \text{smoother on } w^h \\
+&\text{2. restrict:}     & &\ell^H = F^H(\iR w^h) + R\, r^h(w^h) \\
+&\text{3. coarse solve:} & & \text{solve } F^H(w^H) = \ell^H \\
+&\text{4. correct:}      & &w^h \leftarrow w^h + P(w^H - \iR w^h) \\
+&\text{5. post-smooth:}  & & \text{smoother on } w^h
 \end{align*}
@@
 \item $\iR: \mathcal{V}^h \to \mathcal{V}^H$ is \emph{injection}
-\item (restrict$+$coarse solve$+$correct) \, $=$ \, \emph{FAS coarse grid correction}
+\item 2,3,4 \, $=$ \, \emph{FAS coarse grid correction}
 \end{itemize}
@@
 \item<2> one can measure work in flops or run-time
-\item<2> constant iterations of $O(N)$ method $\implies$ optimal
+\item<2> constant $O(1)$ iterations of $O(N)$ method $\implies$ optimal
 \end{itemize}
@@
 
-\only<1>{recall:
+\only<1>{\includegraphics[width=0.75\textwidth]{figs/ballitersV.png}}\only<2>{\includegraphics[width=0.75\textwidth]{figs/ballitersVlog.png}}\only<3>{\includegraphics[width=0.75\textwidth]{figs/ballitersVF.png}}\only<4>{\includegraphics[width=0.75\textwidth]{figs/bothitersVF.png}} \hfill
+\only<1-3>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/ball-set.png}}%
+\only<4>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/spiral-set.png}}
 
-\centerline{\includegraphics[width=0.9\textwidth]{../talk-oxford/images/obstacle65.pdf}}}
-\only<2>{\includegraphics[width=0.75\textwidth]{figs/ballitersV.png}}\only<3>{\includegraphics[width=0.75\textwidth]{figs/ballitersVlog.png}}\only<4>{\includegraphics[width=0.75\textwidth]{figs/ballitersVF.png}}\only<5>{\includegraphics[width=0.75\textwidth]{figs/bothitersVF.png}} \hfill
-\only<2-4>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/ball-set.png}}%
-\only<5>{\includegraphics[width=0.2\textwidth]{../paper/fixfigs/spiral-set.png}}
-
-\only<2-5>{\noindent \scriptsize
-\emph{all results from Bueler \& Farrell (2024)}}
+\only<1-4>{\noindent \scriptsize \emph{all results from Bueler \& Farrell (2024)}}
 \end{frame}
@@
 
-\begin{frame}{iterations over a constraint decomposition}
+\begin{frame}{iterations over a multilevel constraint decomposition}
 
 \begin{itemize}
-\item Tai proposed abstract iterations for solving $VI(F,\ell,\mathcal{K})$ over a CD
+\item Tai (2003) proposed abstract iterations for solving $VI(F,\ell,\mathcal{K})$ over a CD
 \item specifically the multiplicative/serial iteration:
-
-{\small
-\begin{pseudo}[left-margin=-5mm]
+\begin{pseudo}[left-margin=-3mm,font=\small]
 \pr{cd-mult}(u)\text{:} \\+
     for $i = 0,\dots,m-1$: \\+
-        find $w_i\in \cK_i$ s.t. \\+
-            $\displaystyle \Big<F\Big(\sum_{j<i} w_j + w_i + \sum_{j>i} \Pi_j u\Big),\, v_i - w_i\Big> \ge \ip{\ell}{v_i - w_i} \,\forall v_i \in \cK_i$ \\--
+        find $w_i\in \cK_i$ s.t.~$\forall v_i \in \cK_i$ \\+
+            $\displaystyle \Big<F\Big(\sum_{j<i} w_j + w_i + \sum_{j>i} \Pi_j u\Big),\, v_i - w_i\Big> \ge \ip{\ell}{v_i - w_i}$ \\--
     return $w=\sum_i w_i\in\cK$
 \end{pseudo}
-}
 
-\item \dots but not practical because you evaluate residuals on finest level
-\item two techniques make a practical solver (Bueler \& Farrell, 2024):
+\item \dots but not practical for a multilevel CD because you evaluate residuals on finest level
+\item two innovations make a practical solver (Bueler \& Farrell, 2024):
 
     \begin{itemize}
-    \item[$\circ$] \emph{defect obstacles} on each level  \hfill {\scriptsize Gr\"aser \& Kornhuber (2009)}
-    \item[$\circ$] \emph{FAS coarse corrections}  \hfill {\scriptsize Brandt (1977)}
+    \item[$\circ$] defect obstacles on each level  \hfill {\scriptsize Gr\"aser \& Kornhuber (2009)}
+    \item[$\circ$] FAS coarse corrections  \hfill {\scriptsize Brandt (1977)}
     \end{itemize}
@@
 \begin{itemize}
-\item suppose $\mathcal{K} = \{v \ge \psi\}$ in an obstacle problem
+\item suppose $\mathcal{K} = \{v \ge \psi\}$, a unilateral obstacle problem
 \begin{block}{definition}
-for finest-level admissible set $\mathcal{K}^J = \{v^J\ge \psi^J\} \subset \mathcal{V}^J$ and an iterate $w^J \in \mathcal{K}^J$, the \emph{defect obstacle} is
+given a finest-level admissible set $\mathcal{K}^J = \{v^J\ge \psi^J\} \subset \mathcal{V}^J$, and an admissible iterate $w^J \in \mathcal{K}^J$, the \emph{defect obstacle} is
     $$\chi^J = \psi^J - w^J \in \mathcal{V}^J$$
@@
     \end{itemize}
-\item generate the CD through
+\item multilevel: generate defect
 
-defect obstacles $\chi^j$ on each
+obstacles $\chi^j$ on each level
 
-level via \emph{monotone restriction}:
+ via \emph{monotone restriction}:
 
@@
     \begin{itemize}
-    \item[$\circ$] a \emph{nonlinear} operator
+    \item[$\circ$] $R^{\oplus}$ is nonlinear
     \end{itemize}
@@
     $\chi^J = \psi^J - w^J$ \\
-    for $j=J$ downto $j=1$ \\+
+    for $j=J$ downto $j=1$: \\+
       $\chi^{j-1} = \maxR \chi^j$ \\
@@
     $\text{\pr{solve}}(\ell^0,\chi^0,w^0;z^0)$  \\
-    for $j=1$ to $j=J$ \\+
+    for $j=1$ to $j=J$: \\+
       $z^j = y^{j} + P z^{j-1}$ \\
@@
 \vspace{-2mm}
-\tiny
-\begin{itemize}
-\item[] \emph{(unilateral (lower obstacle only) version shown for simplicity)}
-\end{itemize}
+\tiny \hfill \emph{unilateral (lower obstacle only) version shown for simplicity}
 \end{frame}
 
-\begin{frame}{V-cycle: visualization on a 1D problem}
+\begin{frame}{V-cycle visualization (1D unilateral obstacle problem)}
 
@@
 \item FMG cycle
-\item details of the $O(m_j)$ smoother used in the results: a few iterations of projected, reduced-space Newton with a few iterations of CG+ICC or GMRES+ILU on the linear step equations
+\item details of the $O(m_j)$ smoother used in the results: a few iterations of a reduced-space Newton method, with a few iterations of CG+ICC or GMRES+ILU on the linear step equations
 \end{itemize}
@@
 \begin{itemize}
-\item FASCD = new multilevel solver for VI (free-boundary) problems
+\item FASCD = new multilevel solver for VI problems
     \begin{itemize}
-    \item[$\circ$] implemented in Python using the Firedrake library
+    \item[$\circ$] implemented in Python using the \href{https://www.firedrakeproject.org/}{Firedrake FE library (\texttt{firedrakeproject.org})}
     \item[$\circ$] observed optimality, and actually fast on tested problems
     \end{itemize}
-\item many things \textbf{{\color{FireBrick} to do}}:
+{\small
+\item much work to do:
     \begin{itemize}
-    \item[$\circ$] implement in C within PETSc
+    \item[$\circ$] implement in C within \href{https://petsc.org/release/}{PETSc (\texttt{petsc.org})}
     \item[$\circ$] prove convergence; compare (Reusken, 1988)
@@
     \end{itemize}
+}
 \end{itemize}
@@
 \bigskip\bigskip
-\begin{minipage}[T]{0.25\textwidth}
+\footnotesize
+
+\noindent \href{https://epubs.siam.org/doi/10.1137/23M1594200}{\texttt{doi/10.1137/23M1594200}} \hfill \href{https://bitbucket.org/pefarrell/fascd/}{\texttt{bitbucket.org/pefarrell/fascd/}}
+
+\smallskip
+\begin{minipage}[T]{0.35\textwidth}
 \vspace{0pt}
 
-\includegraphics[width=0.8\textwidth]{figs/QRbuelerfarrell.png}
+\includegraphics[width=0.57\textwidth]{figs/QRbuelerfarrell.png}
 
-\scriptsize paper in
-
-SIAM J.~Sci.~Comput.
+\scriptsize the paper (SIAM J.~Sci.~Comput.)
 \end{minipage}
@@
 
-\scriptsize \hfill code at bitbucket
-
-\phantom{foo}
+\scriptsize \hfill the code
 \end{minipage}
 \end{frame}
+
+\appendix
 
